% ══════════════════════════════════════════════════════════
\section{Introduction}\label{sec:intro}
% ══════════════════════════════════════════════════════════

The idea that a single primitive can generate a rich family of
objects is a recurring structural motif in mathematics and logic.
In Boolean algebra the Sheffer stroke (NAND) suffices to express
every truth function~\cite{Sheffer1913}.
In the $\lambda$-calculus a single variable-binding construct
generates all computable functions~\cite{Church1936}.
In combinatory logic two combinators $S$ and~$K$ are enough;
with Sch\"onfinkel's encoding even a single combinator
suffices~\cite{Schonfinkel1924}.

A recent and striking addition to this lineage is
Odrzywołek's demonstration that \emph{every} elementary function
available on a standard scientific calculator---the 36-element
basis comprising arithmetic, algebraic, trigonometric, hyperbolic,
and logarithmic operations---can be reconstructed from a single
binary operator
\begin{equation}\label{eq:eml-def-intro}
  \eml(x,y) \;=\; e^{x} - \ln y
\end{equation}
together with the constant~$1$, using the minimal context-free
grammar $S \to 1 \mid \eml(S,S)$~\cite{Odrzywołek2026}.
This collapses the \emph{expression library} to a uniform binary
tree syntax---we call this \textbf{expression universality}.

The present note identifies expression universality as the bottom
rung of a three-level hierarchy whose upper levels are qualitatively
stronger.

\begin{enumerate}[label=\textbf{L\arabic*},leftmargin=2.5em]
\item \textbf{Expression universality} (Odrzywołek):
  a diverse function library is collapsed to a single operator and
  a tree grammar.  The operations of composition and evaluation
  remain external to the formalism.
\item \textbf{Protocol universality} (this paper, building on
  \cite{OmegaSPG2026}):
  integer arithmetic operations---addition and multiplication
  themselves---are not primitive but \emph{emerge} from a single
  scan--projection protocol.  The unique primitive is function
  composition~$\circ$, and $+$~and~$\times$ are its level-0 and
  level-1 readouts.  Canonical forms are computable, and the word
  problem for projection-word equivalence is decidable.
\item \textbf{Certificate universality} (this paper, building on
  \cite{OmegaSPG2026}):
  the fold fibers carry collision-moment sequences
  $\{S_q(m)\}_{q \ge 0}$ whose Hankel matrices are positive
  semi-definite.  A finite Prony--Hankel reconstruction recovers
  the full multiplicity spectrum from $2r$ moments, yielding an
  auditable, finite-dimensional spectral certificate at each
  resolution~$m$.
\end{enumerate}

\begin{table}[ht]
\centering\small
\caption{Three levels of single-primitive universality.}
\label{tab:hierarchy}
\begin{tabular}{@{}llll@{}}
\toprule
Level & What is unified & Primitive & Canonical form \\
\midrule
L0: Expression  & function library  & $\eml + 1$  & none intrinsic \\
L1: Protocol    & arithmetic ops    & $\circ$ (composition) & Zeckendorf NF \\
L2: Certificate & audit structure   & finite-state kernel & spectral fingerprint \\
\bottomrule
\end{tabular}
\end{table}

The relationships between levels are directional rather than
strict class containments (\Cref{sec:strictness}):
\begin{itemize}[nosep]
\item \textbf{L0 vs L1.}  EML and the SPG protocol are
  \emph{incomparable} as classes---EML is expression-universal
  but not protocol-universal (the word problem is undecidable;
  \Cref{thm:separation}), and protocol universality is not
  entailed by expression universality.  When a
  semantically-compatible evaluator exists
  (\Cref{prop:L1-compiles-L0}), L1 strictly strengthens L0
  by providing decidable canonical forms.
\item \textbf{L1 vs L2.}  Clauses~(a) and~(b) of~L2
  (Hankel PSD, finite Prony reconstruction) are generic for
  protocol-universal systems with finite fibers
  (\Cref{rem:hankel-psd-generic}).  Clause~(c) (polynomial-
  dimension transfer-operator computability) is the
  non-trivial structural content of L2, witnessed in the
  SPG setting by \Cref{thm:star-kernel}.  Whether this
  clause fails for some protocol-universal system---hence
  witnessing a strict class containment---is left open
  (\Cref{rem:L1-L2-strictness-open}).
\end{itemize}

\paragraph{Relation to the main work.}
The protocol and certificate layers are developed in full
generality in~\cite{OmegaSPG2026}.  The purpose of this note is
narrower: to make the hierarchy explicit, to state the
level-by-level structural relations cleanly, and to anchor
the comparison with Odrzywołek's EML operator as a concrete
external reference point.

\subsection{Theorem-by-theorem comparison with prior work}

\Cref{def:eml} is Odrzywołek's operator
\eqref{eq:eml-def-intro} on the principal logarithm branch, so it
is a classical input from~\cite{Odrzywołek2026}.  \Cref{thm:eml-universality}
is exactly Odrzywołek's constructive universality theorem for the
36-element calculator basis, restated here only as a supporting
proposition.  \Cref{def:expr-univ} is our formal abstraction of
Odrzywołek's grammar $S\to 1\mid\eml(S,S)$ from one concrete
operator to an arbitrary single binary operator with a constant.

\Cref{def:spg} uses the classical symbolic-dynamics passage from
a dynamical system and a window to binary itineraries, going back
to Morse--Hedlund~\cite{MorseHedlund1940}.  \Cref{def:fold} adds
to Zeckendorf's unique Fibonacci representation~\cite{Zeckendorf1972}
the finite-resolution fold projection modulo~$F_{m+2}$.  \Cref{prop:fold-confluence}
applies Newman's terminating-plus-local-confluence criterion~\cite{Newman1942}
to the four explicit Fibonacci rewrite rules, thereby separating
finite modular normalization from exact inverse-limit normalization.
\Cref{def:completions} specializes the standard inverse-limit
viewpoint on profinite objects~\cite{RibesZalesskii2010} to legal
Zeckendorf words by distinguishing eventually-zero addresses from
the compact golden-mean inverse limit.  \Cref{def:stable-arith}
packages Zeckendorf arithmetic, whose efficient addition algorithms
are classical~\cite{AhlbachEtAl2013,HieronymiTerry2018}, as
operations on stable fold addresses.  \Cref{thm:semiring} is the
supporting transported semiring structure induced by the
Zeckendorf valuation.  \Cref{thm:two-level} is the supporting
Church-numeral iteration identity~\cite{Church1936}.  \Cref{cor:minimal}
is the fold-specific increment: Church iteration is embedded in a
Zeckendorf normal-form system with decidable address equality.
\Cref{def:prot-univ} isolates that increment as a definition by
combining a single primitive, canonical rewriting, and decidable
word problem in the sense of standard term-rewriting theory~\cite{BaaderNipkow1998}.

\Cref{def:collision-moment} turns the fiber-size power sums into
named fold-collision moments, a fold-specific use of the classical
moment viewpoint.  \Cref{thm:hankel-psd} is the supporting
Hamburger moment-matrix positivity criterion~\cite{Hamburger1920}
applied to the finite fiber-size measure.  \Cref{thm:prony} is the
supporting finite Prony reconstruction principle for an atomic
moment sequence~\cite{BatenkovYomdin2013}.  \Cref{thm:star-kernel}
is the supporting Perron--Frobenius and symmetric-tensor compression
for non-negative transfer matrices~\cite{BermanPlemmons1994}.  \Cref{def:cert-univ}
is our certificate-level definition: it requires the generic
moment facts and additionally singles out polynomial-dimensional
finite-state kernel computability as the non-generic clause.

\Cref{prop:L1-compiles-L0} is a conditional normal-form compilation
statement: when a semantic evaluator is compatible with a canonical
rewrite system, equivalence reduces to the decidable word problem
as in standard rewriting theory~\cite{BaaderNipkow1998}.  \Cref{thm:separation}
is the main novel positioning increment, combining Odrzywołek's
EML representation with Richardson's undecidability theorem for
elementary-function identity~\cite{Richardson1968} to show that
expression universality does not supply protocol universality.
\Cref{cor:strict-L0-L1} is the corresponding conditional corollary:
L0 and L1 are incomparable as classes, while compatible instances
gain decidable canonical forms.  \Cref{prop:L2-generic-vs-structural}
credits Hamburger positivity and Prony reconstruction as generic
finite-fiber inputs, and isolates the polynomial transfer-kernel
clause as the actual structural increment of L2.

\Cref{thm:mixture} is the supporting finite disintegration formula
for a uniform law over fibers~\cite{Kallenberg2002}.  \Cref{thm:jacobi}
is the supporting finite spectral realization of an atomic moment
measure by multiplication in an orthogonal-polynomial basis, and
\Cref{cor:stieltjes} is the associated Stieltjes continued fraction
from the same classical theory~\cite{Stieltjes1894}.  \Cref{thm:escort-ldp}
is the supporting finite-alphabet Laplace principle
estimate~\cite{DemboZeitouni2010}.  \Cref{thm:cross-kernel} is the
supporting feature-map construction of a positive-definite kernel
in the Moore--Aronszajn framework~\cite{Aronszajn1950}.  \Cref{thm:zeta}
is the supporting finite Euler product and Witt--Möbius necklace
expansion~\cite{Moree2005}.  \Cref{thm:gibbs} is the supporting
finite Gibbs variational principle~\cite{Ellis1985}.  \Cref{prop:nno}
is the fold-address realization of the classical natural-numbers
object and Church-numeral iteration picture~\cite{Lawvere1964,Church1936}.
