\section{A Zeckendorf protocol}\label{sec:L1}

We now add the normal-form data absent from expression universality.
Let \(\Omega_m=\{0,1\}^m\).  The elements of \(\Omega_m\) may be
regarded as finite itineraries of a one-generator symbolic source,
although no dynamical property will be used.  The fold map sends a
word to its Zeckendorf address.  This elementary construction gives
the two pieces needed later: a computable normal form and readout
maps from normal forms to the intended semantic objects.

\subsection{The fold}
\label{ssec:fold}

Let \(F_1=F_2=1\) and \(F_{n+2}=F_{n+1}+F_n\).  For \(m\ge1\), let
\[
  X_m=\{c_1\cdots c_m\in\{0,1\}^m:
        c_kc_{k+1}=0 \text{ for } 1\le k<m\}
\]
be the set of legal Fibonacci words, and define
\[
  \Val_m(c_1\cdots c_m)=\sum_{k=1}^m c_kF_{k+1}.
\]

\begin{definition}[Fold map]\label{def:fold}
For \(\omega=\omega_1\cdots\omega_m\in\Omega_m\), set
\[
  N(\omega)=\sum_{k=1}^m \omega_kF_{k+1}.
\]
The finite fold map at resolution \(m\) is the map
\(\Fold_m\colon\Omega_m\to X_m\) sending \(\omega\) to the unique
legal word \(c\in X_m\) satisfying
\begin{equation}\label{eq:fold-residue}
  \Val_m(c)\equiv N(\omega)\pmod {F_{m+2}}.
\end{equation}
The existence and uniqueness of this residue representative follow
from \Cref{prop:fold-confluence}.
\end{definition}

\begin{proposition}[Finite fold normal forms]
\label{prop:fold-confluence}
For every \(m\ge1\), the map
\[
  \Val_m\colon X_m\longrightarrow \{0,1,\ldots,F_{m+2}-1\}
\]
is a bijection.  Hence \(\Fold_m\) is well defined and computable.
Moreover, if
\[
  X_{\infty}^{\mathrm{fin}}
  =\{(c_1,c_2,\ldots)\in\{0,1\}^{\NN}:
    c_kc_{k+1}=0 \text{ for all }k,\ c_k=0\text{ for }k\gg0\},
\]
then
\[
  \Val(c)=\sum_{k\ge1}c_kF_{k+1}
\]
is a bijection \(X_{\infty}^{\mathrm{fin}}\to\NN\).
\end{proposition}

\begin{proof}
By Zeckendorf's theorem, every \(n\ge0\) has a unique expression as
a sum of nonconsecutive Fibonacci numbers \(F_{k+1}\)
\cite{Zeckendorf1972}.  If \(0\le n<F_{m+2}\), then no summand
\(F_{m+2}\) or larger appears, so the digits lie in the positions
\(1,\ldots,m\).  This gives a legal word in \(X_m\).  Conversely,
if \(M_m\) is the maximum of \(\Val_m\) on \(X_m\), then
\[
  M_m=\max(M_{m-1},F_{m+1}+M_{m-2}),\qquad M_0=0,\quad M_1=1,
\]
and induction gives \(M_m=F_{m+2}-1\).  Thus every legal word has
value in the displayed range, and uniqueness in Zeckendorf's theorem
proves the finite bijection.  The infinite statement is the same
theorem without the upper bound; eventual vanishing makes the sum
finite.  Computability follows either from the greedy Zeckendorf
algorithm or from standard finite-automaton methods for
Zeckendorf--Ostrowski arithmetic
\cite{AhlbachEtAl2013,HieronymiTerry2018}.
\end{proof}

The finite map \(\Fold_m\) is a residue-level normalizer.  The
stable normalizer of a finite binary word \(\omega\) is the
unreduced Zeckendorf address
\[
  \mathrm{NF}_{\infty}(\omega)=\Val^{-1}(N(\omega))
  \in X_{\infty}^{\mathrm{fin}},
\]
where \(\omega\) is extended by zeros after its last digit.  Thus the
finite maps \(\Fold_m\) provide the certificate fibers, whereas
arithmetic is read on the countable set of eventually-zero normal
forms.

The finite legal-word spaces also form an inverse system under
coordinate truncation.  The arithmetic used below lives on its
eventually-zero part, not on the whole compact limit.

\begin{definition}[Finite and compact completions]
\label{def:completions}
The eventually-zero space \(X_{\infty}^{\mathrm{fin}}\) is the
countable set in \Cref{prop:fold-confluence}, identified with
\(\NN\) by \(\Val\).  The compact inverse limit is
\begin{equation}\label{eq:X-infty}
  \widehat X
  =\varprojlim_m X_m
  \cong
  \{(c_1,c_2,\ldots)\in\{0,1\}^{\NN}:
    c_kc_{k+1}=0\text{ for all }k\}.
\end{equation}
It is a compact zero-dimensional space, analogous to a profinite
completion \cite{RibesZalesskii2010}.  The arithmetic operations
below are carried out on \(X_{\infty}^{\mathrm{fin}}\).
\end{definition}

\begin{remark}\label{rem:finite-vs-profinite}
The valuation is not a bijection \(\widehat X\to\NN\), since
\(\widehat X\) is uncountable.  Stable arithmetic is therefore an
operation on eventually-zero addresses, not on the full compact
inverse limit.
\end{remark}

\subsection{Arithmetic by readout}
\label{ssec:arith}

\begin{definition}[Stable addition and multiplication]
\label{def:stable-arith}
For \(c,d\in X_{\infty}^{\mathrm{fin}}\), define
\begin{align}
  c\oplus d
    &=\Val^{-1}\bigl(\Val(c)+\Val(d)\bigr),\label{eq:stable-add}\\
  c\otimes d
    &=\Val^{-1}\bigl(\Val(c)\Val(d)\bigr).
      \label{eq:stable-mul}
\end{align}
Equivalently, \(c\oplus d\) is obtained by adding the two values and
then normalizing; this is the same as coordinatewise digit addition
followed by Zeckendorf carrying.
\end{definition}

\begin{proposition}[Transported semiring structure]
\label{thm:semiring}
\((X_{\infty}^{\mathrm{fin}},\oplus,\otimes)\) is a commutative
semiring with additive identity \(0_X=(0,0,\ldots)\) and
multiplicative identity \(\Val^{-1}(1)\).  The valuation map
\[
  \Val\colon
  (X_{\infty}^{\mathrm{fin}},\oplus,\otimes)
  \xrightarrow{\ \sim\ }(\NN,+,\times)
\]
is a semiring isomorphism.  At finite resolution, transporting the
ring operations on \(\ZZ/F_{m+2}\ZZ\) through the bijection
\(\Val_m\colon X_m\to\ZZ/F_{m+2}\ZZ\) gives a ring structure on
\(X_m\) isomorphic to \(\ZZ/F_{m+2}\ZZ\).  This finite ring
structure records only the residue-level structure; the stable
semiring operation is the one on
\(X_{\infty}^{\mathrm{fin}}\).
\end{proposition}

\begin{proof}
The operations \(\oplus\) and \(\otimes\) are the pullbacks of
\(+\) and \(\times\) along the bijection \(\Val\).  The semiring
axioms and the asserted isomorphism are immediate.  The finite
statement is the same transport of structure through \(\Val_m\),
with values taken modulo \(F_{m+2}\).
\end{proof}

Thus arithmetic has not been added as a primitive operation on
words.  It is read through valuation and stable Zeckendorf
normalization; residue reduction is only the finite-resolution
version used for certificates.

\subsection{Composition as the primitive}
\label{ssec:sole}

\begin{proposition}[Two-level readout of composition]
\label{thm:two-level}
Let \(A\) be a nonempty set and let \(g\colon A\to A\) be an
endomorphism of infinite order.  Define
\begin{align}
  S_0&=\langle g\rangle
      =\{g^{\circ n}:n\in\NN\}\subseteq\End(A),\label{eq:level0}\\
  S_1&=\{I_n:n\in\NN\}\subseteq\End(\End(A)),
      \qquad I_n(f)=f^{\circ n}.\label{eq:level1}
\end{align}
With \(\iota_0(n)=g^{\circ n}\) and \(\iota_1(n)=I_n\), the maps
\(\iota_0\) and \(\iota_1\) are injective and
\begin{align}
  \iota_0(m+n)&=\iota_0(m)\circ\iota_0(n),
    \label{eq:add-from-comp}\\
  \iota_1(mn)&=\iota_1(m)\circ\iota_1(n).
    \label{eq:mul-from-comp}
\end{align}
\end{proposition}

\begin{proof}
The first identity is
\(g^{\circ m}\circ g^{\circ n}=g^{\circ(m+n)}\), and infinite order
gives injectivity of \(\iota_0\).  The same hypothesis gives
injectivity of \(\iota_1\), since \(I_m(g)=g^{\circ m}\) and
\(I_n(g)=g^{\circ n}\).  Finally,
\[
  (I_m\circ I_n)(f)=I_m(f^{\circ n})
    =(f^{\circ n})^{\circ m}=f^{\circ mn}=I_{mn}(f).
\]
\end{proof}

\begin{corollary}[Zeckendorf readout of Church iteration]
\label{cor:minimal}
Under the identification \(X_{\infty}^{\mathrm{fin}}\cong\NN\),
the address \(c_n=\Val^{-1}(n)\) represents the \(n\)-fold iterate.
Addition of addresses is the level-zero composition law, and
multiplication is the level-one composition law.  This is the usual
Church-numeral iteration identity \cite{Church1936}, expressed in
Zeckendorf normal form.
\end{corollary}

\begin{proof}
Apply \Cref{thm:two-level} after replacing \(n\) by its address
\(c_n\).  The semiring identification in \Cref{thm:semiring}
transports the addition and multiplication laws to stable
addresses.
\end{proof}

No new identity about Church iteration is involved.  The point is
that the usual iteration laws remain decidable after the indices are
placed in Zeckendorf normal form.

\begin{theorem}[Level-zero obstruction for multiplicative readout]
\label{thm:level-zero-obstruction}
Let \(M\) be a monoid and let \(g\in M\) have infinite order.  Put
\[
  C=\{g^n:n\in\NN\}\subset M .
\]
There is no injective map
\(\eta\colon\NN_{\ge1}\to C\) satisfying
\[
  \eta(ab)=\eta(a)\eta(b)
  \qquad(a,b\ge1).
\]
Consequently multiplication on the positive integers cannot be
represented by the ambient monoid product on a single free
monogenic orbit, even after arbitrary recoding of the numerals into
that orbit.  In the composition model of \Cref{thm:two-level},
therefore, multiplicative readout requires a change of level from
the orbit \(\{g^{\circ n}\}\) to the power operators
\(f\mapsto f^{\circ n}\), under this level-zero monogenic-orbit
restriction.
\end{theorem}

\begin{proof}
Since \(g\) has infinite order, each element of \(C\) has a unique
form \(g^e\), \(e\in\NN\).  Write
\[
  \eta(n)=g^{e(n)} .
\]
The displayed multiplicativity condition is equivalent to
\[
  e(ab)=e(a)+e(b)\qquad(a,b\ge1).
\]
Taking \(a=b=1\) gives \(e(1)=0\).  If \(p\) is prime and
\(e(p)=0\), then \(\eta(p)=\eta(1)\), contrary to injectivity.
Thus \(e(p)>0\) for every prime \(p\).

Choose two distinct primes \(p\) and \(r\), and set
\[
  A=e(p)>0,\qquad B=e(r)>0 .
\]
Then
\[
  e(p^B)=B\,e(p)=AB=A\,e(r)=e(r^A).
\]
Hence \(\eta(p^B)=\eta(r^A)\), while unique factorization gives
\(p^B\ne r^A\).  This contradiction proves the theorem.
\end{proof}

\begin{definition}[Protocol universality]\label{def:prot-univ}
Let \(\mathcal O\) be a class of operations on a set \(D\).  A
\emph{single-primitive protocol} for \(\mathcal O\) consists of
\begin{enumerate}
\item a decidable set \(C\) of raw codes and a computable
  normalization map \(\mathrm{NF}\colon C\to N\) onto a set \(N\) of
  unique normal forms, with decidable equality on \(N\);
\item a decoding map \(\nu\colon N\to D\);
\item a finite family of semantic sorts \(Y_\ell\), each equipped
  with an interpretation \(\sigma_\ell\) of one binary operation
  symbol \(\sigma\);
\item computable readout maps
  \(\Phi_\ell\colon N\to Y_\ell\).
\end{enumerate}
An operation \(\theta\in\mathcal O\) of arity \(r\) is realized if
there are a computable operation
\(\widehat\theta\colon N^r\to N\), a level \(\ell\), and a
\(\sigma\)-term \(R_\theta\) in \(r\) variables such that
\[
  \nu(\widehat\theta(n_1,\ldots,n_r))
  =\theta(\nu(n_1),\ldots,\nu(n_r))
\]
and
\[
  \Phi_\ell(\widehat\theta(n_1,\ldots,n_r))
  =
  R_\theta(\Phi_\ell(n_1),\ldots,\Phi_\ell(n_r))
\]
for all \(n_i\in N\).  The protocol is
\emph{protocol-universal for \(\mathcal O\)} if every operation in
\(\mathcal O\) is realized in this sense.  The normalizer is
algorithmic presentation data; the only algebraic operation used in
the readout terms is \(\sigma\), as in a canonical rewriting
system with a decidable word problem \cite{BaaderNipkow1998}.
\end{definition}

\begin{theorem}[Zeckendorf protocol universality]
\label{thm:zeckendorf-protocol-univ}
Let \(A\) be a nonempty set and let \(g\colon A\to A\) be an
endomorphism of infinite order.  The Zeckendorf address system with
raw codes the finite binary words, normal forms
\(N=X_{\infty}^{\mathrm{fin}}\), normalizer
\(\mathrm{NF}_{\infty}\), decoding map \(\nu=\Val\), and the two
composition readouts
\[
  \Phi_0(c_n)=g^{\circ n},\qquad
  \Phi_1(c_n)=I_n,\qquad c_n=\Val^{-1}(n),
\]
is a single-primitive protocol universal for
\(\mathcal O=\{+,\times\}\) on \(\NN\).  Addition is realized by
\(\widehat{+}=\oplus\) at level \(0\), and multiplication is
realized by \(\widehat{\times}=\otimes\) at level \(1\).  In both
cases the readout term is the one binary operation symbol,
interpreted as composition on the relevant semantic sort.
\end{theorem}

\begin{proof}
Let \(C\) be the decidable set of finite binary words and let
\(N=X_{\infty}^{\mathrm{fin}}\).  The normal form of a raw word is
\(\mathrm{NF}_{\infty}\), obtained by computing the integer
\(\sum_k\omega_kF_{k+1}\) and applying the greedy Zeckendorf
algorithm.  Equality of normal forms is literal equality of finite
legal words, hence is decidable by \Cref{prop:fold-confluence}.  The
decoding map is \(\nu=\Val\).

Interpret the single operation symbol as composition on both
semantic sorts
\[
  Y_0=\End(A),\qquad Y_1=\End(\End(A)).
\]
The readout maps are the displayed maps
\(\Phi_0(c_n)=g^{\circ n}\) and \(\Phi_1(c_n)=I_n\).  They are
well defined because \(\Val\) is a bijection
\(X_{\infty}^{\mathrm{fin}}\to\NN\).

For addition take \(\widehat{+}=\oplus\).  This operation is
computable: compute the two values, add them in \(\NN\), and return
the Zeckendorf normal form.  The semantic clause is
\[
  \nu(c_m\oplus c_n)=m+n,
\]
by \Cref{thm:semiring}.  The readout clause at level \(0\) is
\[
  \Phi_0(c_m\oplus c_n)
  =g^{\circ(m+n)}
  =g^{\circ m}\circ g^{\circ n}
  =\Phi_0(c_m)\circ\Phi_0(c_n),
\]
which is \Cref{thm:two-level}.

For multiplication take \(\widehat{\times}=\otimes\).  It is
computable by the same valuation-normalization procedure, using
integer multiplication.  Again \Cref{thm:semiring} gives
\[
  \nu(c_m\otimes c_n)=mn.
\]
At level \(1\), \Cref{thm:two-level} gives
\[
  \Phi_1(c_m\otimes c_n)
  =I_{mn}
  =I_m\circ I_n
  =\Phi_1(c_m)\circ\Phi_1(c_n).
\]
Thus both operations in \(\{+,\times\}\) are realized using only the
single binary operation symbol in the readout.  The finite maps
\(\Fold_m\) are the resolution quotients whose fibers are studied in
\Cref{sec:L2}; they do not alter the stable readout just verified.
\end{proof}
