% ══════════════════════════════════════════════════════════
\section{Expression Universality: The EML Paradigm}\label{sec:L0}
% ══════════════════════════════════════════════════════════

We recall the main result of~\cite{Odrzywołek2026} and
extract from it a precise definition of expression universality.

% ── 2.1  The EML operator ─────────────────────────────────
\subsection{The EML operator and the 36-element basis}

\begin{definition}[EML operator]\label{def:eml}
The \emph{Exp-Minus-Log} binary operator is
\[
  \eml \colon \CC \times \CC^{*} \;\longrightarrow\; \CC,
  \qquad
  \eml(x,y) \;=\; e^{x} - \operatorname{Log} y,
\]
where $\operatorname{Log}$ denotes the principal-branch
logarithm on~$\CC^{*}$.
(The complex domain is essential: constants such as~$i$
and trigonometric functions are generated via Euler's
formula; see~\cite{Odrzywołek2026}.)
\end{definition}

Odrzywołek identifies a 36-element \emph{scientific calculator
basis}~$\mathcal{B}_{36}$ comprising 8~constants (including
$\pi,e,i,-1,1,2$ and input variables), 20~unary functions
(exp, ln, reciprocal, square root, square, trigonometric,
inverse trigonometric, hyperbolic, inverse hyperbolic, and the
sigmoid), and 8~binary operations ($+,-,\times,\div$, logarithm,
power, average, hypotenuse).

\begin{theorem}[Odrzywołek {\cite[Thm.~1]{Odrzywołek2026}}]
\label{thm:eml-universality}
Every function in~$\mathcal{B}_{36}$ is representable, on a
suitable domain of definition for the principal-branch
logarithm, as a finite composition of\/ $\eml$ and the
constant~$1$.  Explicitly, for $x$ in the right half-plane
$\{\Re(x) > 0\}$:
\begin{align}
  e^{x}  &= \eml(x,\, 1), \label{eq:eml-exp}\\
  \operatorname{Log} x
         &= \eml\!\bigl(1,\;
            \eml(\eml(1,x),\,1)\bigr), \label{eq:eml-ln}\\
  e      &= \eml(1,1). \label{eq:eml-e}
\end{align}
Outside the principal domain, the identities hold up to the
branch structure of\/ $\operatorname{Log}$; for example,
at~$x=-1$ the right-hand side of~\eqref{eq:eml-ln} returns
$-i\pi$, whereas $\operatorname{Log}(-1) = i\pi$, reflecting
the standard branch cut along the negative real axis.
From these three identities, the remaining 33 elements
of~$\mathcal{B}_{36}$ follow by further nesting,
domain-by-domain.
\end{theorem}

The representation space is governed by the context-free grammar
\begin{equation}\label{eq:eml-grammar}
  S \;\longrightarrow\; 1 \;\mid\; \eml(S,\,S),
\end{equation}
so that every representable function corresponds to a finite binary
tree with $\eml$ at each internal node and~$1$ (or an input
variable) at each leaf.

% ── 2.2  Definition of expression universality ────────────
\subsection{Formal definition}

\begin{definition}[Expression universality]\label{def:expr-univ}
Let $\mathcal{F}$ be a family of functions and
$\sigma \colon D^{2} \to D$ a binary operator on a domain~$D$
containing a distinguished element~$c_0 \in D$.
We say the pair $(\sigma, c_0)$ is
\emph{expression-universal for~$\mathcal{F}$} if every $f \in
\mathcal{F}$ can be written as a term in the grammar
\[
  S \;\longrightarrow\; c_0 \;\mid\; x_i
    \;\mid\; \sigma(S,\,S)
\]
where $x_i$ range over input variables.
\end{definition}

\begin{remark}\label{rem:no-canon}
Expression universality is purely \emph{existential}:
it asserts that a representation exists, but provides no intrinsic
canonical form.  Two distinct trees may compute the same function,
and deciding their equivalence requires external reasoning
(simplification rules, numerical evaluation, etc.) that is not part
of the grammar~\eqref{eq:eml-grammar}.
\end{remark}

\begin{remark}\label{rem:eml-variants}
Odrzywołek also identifies the variant
$\mathrm{edl}(x,y) = e^{x}/\ln y$ paired with~$e$ as another
expression-universal operator, and notes the negated form
$-\!\eml(y,x) = \ln x - e^{y}$ paired with~$-\infty$.
The universality phenomenon is thus not an isolated accident but a
structural feature of $\exp$--$\ln$ duality.
\end{remark}

% ── 2.3  Structural limits of L0 ─────────────────────────
\subsection{Structural boundaries}

Expression universality collapses the function \emph{library} but
leaves two layers untouched:

\begin{enumerate}[label=(\roman*)]
\item \textbf{Operations are external.}
  The act of composing two EML trees, evaluating a tree at a point,
  or deciding whether two trees represent the same function---none
  of these is internal to the grammar.
  Composition is meta-level tree grafting; evaluation is an external
  interpreter.

\item \textbf{No canonical forms.}
  The function $x \mapsto x$ can be represented as
  $\eml(\ln x, \eml(1,x))$ or via many other trees.
  There is no finite, computable normalization procedure
  that maps every EML tree to a unique representative of its
  equivalence class.  (This will be the key separation from~L1
  in~\Cref{sec:strictness}.)
\end{enumerate}

These boundaries are not defects of the EML construction; they are
intrinsic to expression universality as defined.  Overcoming them
requires moving to a qualitatively different kind of universality.
