\documentclass[reqno]{amsart}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{amscd}
\usepackage{enumitem}
\usepackage[numbers]{natbib}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}
\usepackage{geometry}
\geometry{margin=1.15in}
\raggedbottom
\ExplSyntaxOn
\cs_new_eq:NN \paper_original_Cref \Cref
\bool_new:N \l_paper_cref_labels_known_bool
\RenewDocumentCommand{\Cref}{m}
  {
    \bool_set_true:N \l_paper_cref_labels_known_bool
    \clist_map_inline:nn {#1}
      {
        \cs_if_exist:cF { r@##1 }
          { \bool_set_false:N \l_paper_cref_labels_known_bool }
      }
    \bool_if:NTF \l_paper_cref_labels_known_bool
      { \paper_original_Cref{#1} }
      { \textbf{??} }
  }
\ExplSyntaxOff
\hypersetup{
  pdftitle={Gluing Failure, Visible Quotients, and Pure-Ext Blind Spots},
  pdfkeywords={information states, team semantics, sheafification, gerbes, visible quotient, contextuality}
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\St}{\mathrm{St}}
\newcommand{\Mod}{\mathrm{Mod}}
\newcommand{\Val}{\mathrm{Val}}
\newcommand{\Th}{\mathrm{Th}}
\newcommand{\Frm}{\mathrm{Form}}
\newcommand{\Adm}{\mathsf{Adm}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\LocSec}{\mathsf{LocSec}}
\newcommand{\CompSec}{\mathsf{CompSec}}
\newcommand{\Sec}{\mathsf{Sec}}
\newcommand{\Undef}{\mathsf{Undef}}
\newcommand{\Tot}{\mathsf{Tot}}
\newcommand{\Null}{\mathsf{Null}}
\newcommand{\Upd}{\mathrm{Upd}}
\newcommand{\Com}{\mathrm{Com}}
\newcommand{\OST}{\mathrm{OST}}
\newcommand{\Act}{\mathsf{Act}}
\newcommand{\Ch}{\mathsf{Ch}}
\newcommand{\res}{\mathrm{res}}
\newcommand{\Read}{\mathrm{Read}}
\newcommand{\Vis}{\mathrm{Vis}}
\newcommand{\Ob}{\mathrm{Ob}}
\newcommand{\Band}{\mathcal{A}}
\newcommand{\Cech}{\check{H}}
\newcommand{\Sh}{\mathbf{Sh}}
\newcommand{\Psh}{\mathbf{PSh}}
\newcommand{\KRed}{\mathsf{K}}
\newcommand{\leanverified}[1]{}

\title[Gluing Failure, Visible Quotients, and Pure-Ext Blind Spots]{Gluing Failure, Visible Quotients, and Pure-Ext Blind Spots}
\author{Anonymous}

\begin{document}

\begin{abstract}
We study gluing failure for references evaluated at information states in a focused note that starts from a chosen local-object layer and an explicit gluing-sensitive realization-prestack hypothesis, rather than from a canonical obstruction theory on a bare set-valued presheaf. Naming an abelian band is not an existence theorem for such a prestack. The split lift is only a separatedness criterion: for $\sigma,\tau\in F_{p,s}(U)$ its isomorphism presheaf on the slice $\mathcal C_a/U$ is
\[
V\longmapsto
\begin{cases}
\Band(V), & \res_V^U(\sigma)=\res_V^U(\tau),\\
\varnothing, & \res_V^U(\sigma)\neq\res_V^U(\tau),
\end{cases}
\]
so it may have mixed support. Hence the split lift is a prestack for the original presheaf exactly when equality in $F_{p,s}|_{\mathcal C_a}$ is separated. After sheafifying $F_{p,s}$ the same split construction realizes the sheafified visible classes, not the original presheaf unless the relevant comparison map is an isomorphism. Non-separated presheaves require extra off-diagonal torsor data, including nonzero torsor classes for locally equal but globally distinct pairs, and may have no realization with a fixed band when those torsors are all trivial. Consequently every gerbe or obstruction statement below is conditional on a chosen stack-banded realization input, not on a band alone. The three points proved in this narrowed setting are these. First, on lower-layer formulas state forcing coincides with a flat team semantics on nonempty teams of assignments, and we prove an automorphism-invariant undefinability theorem: compatible local sectionability and gluing-level absence are not definable by lower formulas that do not name the local-object structure. Second, after a stack-banded realization prestack with its off-diagonal isomorphism sheaves and a conjugation-compatible band has been chosen, the induced stack banding identifies $\Band|_U$ with the automorphism sheaf of every object of the stackification, compatible local presentations of a reference define visible branches, and these are exactly the global connected components of the realization stack; under a global conservativity hypothesis, gluing failure occurs precisely when visible branches exist but all component gerbes of those branches are non-neutral. Third, for a finite abelian branch obstruction $\omega$ arising from the chosen realization data on a finite \v{C}ech nerve presentation, the intrinsic visible quotient is $\operatorname{coker}(\mathrm{ev}_{\omega})$, where $\mathrm{ev}_{\omega}:H_2(N(\mathcal U),\mathbb Z)\to A$ is the homological evaluation map arising from the universal coefficient theorem. The paper proves that this quotient is well defined on the branch obstruction's allowed finite-presentation transport class: fixed-coefficient homological refinements preserve it on the nose, and coefficient retivializations by finite-group isomorphisms transport it canonically. Equivalently, this quotient is detected exactly by the class-admissible character channels. In the unique-branch support-presheaf setting, the theory recovers the Abramsky--Brandenburger no-global-section picture and identifies the cohomologically blind residue as the pure $\operatorname{Ext}$-part. Two verified finite-nerve branch-gerbe constructions isolate the visible and blind extremes.
\end{abstract}

\keywords{information states, team semantics, gluing failure, gerbes, visible quotient, contextuality}
\subjclass[2020]{03B45, 03F55, 03G30, 18F20}

\maketitle

\input{sec_introduction}
\input{sec_preliminaries}
\input{sec_information_states}
\input{sec_null_decomposition}
\input{sec_gerbe_obstruction}
\input{sec_homological_visibility}
\input{sec_conclusion}
\input{sec_appendix}
\input{sec_multiaxis_refinement}
\input{sec_observer_spacetime}
\input{sec_conservativity}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
