\subsection{Realization inputs: split lifts and off-diagonal data}\label{subsec:realization-inputs}

The presheaf $F_{p,s}$ records only isomorphism classes of possible local
realizations. Gerbes and obstruction classes require more: one must have actual
groupoids of local realizations and sheaves of isomorphisms between their
objects. This is additional realization data unless it is constructed by a
separate criterion. The banding required later is also a compatible banding
that descends to the stackification, not merely a fibrewise identification of
automorphism groups in the original prestack.

The distinction used below is the following data hierarchy.
\[
\text{set-valued }F_{p,s}
\quad\longrightarrow\quad
\text{fibred groupoid over }\mathcal C_a
\quad\longrightarrow\quad
\text{prestack}.
\]
The first arrow is extra structure: it chooses objects, restrictions, and
isomorphism sets above the set-valued presheaf. It is not supplied by a band. A
band specifies only the intended diagonal automorphism sheaves; it does not
specify the off-diagonal isomorphism sheaves, their restrictions, or their
compositions. The second arrow is the prestack condition: every isomorphism
presheaf on every slice must be a sheaf. Thus, throughout the rest of the paper,
a phrase such as ``a stack-banded realization prestack with band $\Band$'' has a fixed
data-level meaning. It asserts:
\begin{enumerate}[label=(\alph*),nosep]
\item a fibred groupoid over the slice $\mathcal C_a$;
\item an identification of its presheaf of isomorphism classes with
$F_{p,s}|_{\mathcal C_a}$;
\item sheaf-valued off-diagonal isomorphism presheaves, with restriction and
composition;
\item diagonal automorphism sheaves in the prestack identified with $\Band$
compatibly with restriction and conjugation;
\item a stackification banding on every object of the stack
$\mathfrak L_r=a(\mathfrak P_r)$.
\end{enumerate}
The banding items alone are still not a realization theory. The full assertion
means that the whole prestack has already been given, or has been constructed
under an explicit criterion. It never means that the pair $(F_{p,s},\Band)$
determines the missing off-diagonal data, and it never licenses gerbes or
obstruction classes before those data have been supplied. The last item is a
hard hypothesis for the gerbe layer. It may be supplied directly as part of the
realization input, or obtained from a conjugation-compatible prestack-level
banding by \Cref{prop:stackification-band-descends}. The later component-gerbe
arguments use only this stack-level banding.

The split construction below is included as a test case because it exposes the
point at which the tempting band-only argument fails. In a single fibre over
$U$, the split construction puts morphisms only between equal elements of
$F_{p,s}(U)$. The prestack condition, however, is not a fibrewise condition. It
tests the isomorphism presheaf after every refinement $V\to U$. Thus the
calculation that matters is not
\[
\mathrm{Hom}_{\mathfrak P^{\mathrm{split}}(U)}(\sigma,\tau)
\]
alone, but the whole presheaf
\[
V\longmapsto
\mathrm{Hom}_{\mathfrak P^{\mathrm{split}}(V)}
(\res_V^U\sigma,\res_V^U\tau)
\]
on the slice $\mathcal C_a/U$.

\begin{definition}[Equality sieve of a pair]\label{def:equality-sieve-pair}
For an object $U$ of $\mathcal C_a$ and sections
\[
\sigma,\tau\in F_{p,s}(U),
\]
define the equality sieve of the pair by
\[
\mathrm{Eq}_U(\sigma,\tau)
:=\{\,V\to U\mid \res_V^U(\sigma)=\res_V^U(\tau)\,\}
\subseteq \mathcal C_a/U.
\]
This sieve is computed after refinement. It is not determined by the truth value
of $\sigma=\tau$ in the single fibre over $U$ unless the restricted presheaf is
already separated enough to make equality local.
\end{definition}

The split fibred groupoid is defined formally in
\Cref{def:split-fibred-groupoid}. Its decisive calculation is the following:
the isomorphism presheaf between
$\sigma,\tau\in F_{p,s}(U)$ is the restriction-dependent presheaf
\[
V\longmapsto
\begin{cases}
\Band(V), & V\to U \text{ lies in } \mathrm{Eq}_U(\sigma,\tau),\\
\varnothing, & V\to U \text{ does not lie in } \mathrm{Eq}_U(\sigma,\tau).
\end{cases}
\]
Thus the two constant-support extremes occur only in the two extreme equality
sieves:
\[
\mathrm{Eq}_U(\sigma,\tau)=\mathcal C_a/U
\quad\Longrightarrow\quad
I_{\sigma,\tau}\cong \Band|_U,
\]
and
\[
\mathrm{Eq}_U(\sigma,\tau)=\varnothing
\quad\Longrightarrow\quad
I_{\sigma,\tau}\cong\varnothing.
\]
In the general case the equality sieve can be proper and nonempty. Then
$I_{\sigma,\tau}$ has mixed support: it is nonempty on exactly those refinements
where the two restrictions agree and empty on the others. This mixed-support
case is precisely the case missed by the invalid fibrewise dichotomy. In
particular, a constant-support description of the split isomorphism presheaf is
valid only after adding the corresponding constant-support hypothesis on the
equality sieve. Inspecting only
$\mathrm{Hom}_{\mathfrak P^{\mathrm{split}}(U)}(\sigma,\tau)$ is therefore
insufficient for the prestack test: the sheaf axiom sees local sections on all
objects of the slice $\mathcal C_a/U$.

Consequently the split construction is a separatedness criterion, not a general
existence theorem. If a non-separated pair is distinct over $U$ but becomes
equal on a cover of $U$, then the identity sections of $\Band$ on the cover are
local split isomorphisms with no global split isomorphism. No choice of the
abelian band changes this matching-family obstruction. A nonsplit realization,
when one exists, must add genuine off-diagonal torsors and compatible
composition data outside the split construction. This necessary-data statement
is not itself a sufficiency theorem: pairwise torsors must still extend to a
fibred groupoid with restriction functors and associative composition, and its
isomorphism presheaves must still be sheaves. The section therefore proves a
ledger, not a universal construction: separated presheaves have the split lift;
non-separated presheaves do not have the split lift; and any nonsplit lift of a
non-separated presheaf is additional realization data whose coherence has to be
supplied and checked.

\begin{definition}[Separated local-object presheaf on the slice]\label{def:slice-separated-local-object}
Fix an admitted reference $r:\mathsf{Ref}_s$ at $p$, write $a=[r]_p$, and work on
\[
\mathcal C_a:=\mathcal C_{p,s}/a.
\]
The restricted local-object presheaf $F_{p,s}|_{\mathcal C_a}$ is \emph{separated} if, for every nonempty cover $\{U_i\to U\}_{i\in I}$ in $\mathcal C_a$ and every pair of sections
\[
\sigma,\tau\in F_{p,s}(U),
\]
the implication
\[
\res_{U_i}^{U}(\sigma)=\res_{U_i}^{U}(\tau)\ \text{for all }i\in I
\quad\Longrightarrow\quad
\sigma=\tau
\]
holds.
\end{definition}

\begin{lemma}[Equality loci and separatedness]\label{lem:equality-loci-separated}
In the setting of \Cref{def:slice-separated-local-object}, fix an object $U$ of $\mathcal C_a$ and sections
\[
\sigma,\tau\in F_{p,s}(U).
\]
Define the equality-locus presheaf $E_{\sigma,\tau}$ on $\mathcal C_a/U$ by
\[
E_{\sigma,\tau}(V)=
\begin{cases}
\{*\}, & \res_V^U(\sigma)=\res_V^U(\tau),\\
\varnothing, & \res_V^U(\sigma)\neq\res_V^U(\tau).
\end{cases}
\]
Then $F_{p,s}|_{\mathcal C_a}$ is separated if and only if $E_{\sigma,\tau}$ is a sheaf for every such pair $(\sigma,\tau)$.
\end{lemma}

\begin{proof}
For a nonempty cover $\{V_\ell\to V\}$ in $\mathcal C_a/U$, a matching family in $E_{\sigma,\tau}(V_\ell)$ exists exactly when
\[
\res_{V_\ell}^U(\sigma)=\res_{V_\ell}^U(\tau)
\]
for every $\ell$; in that case the matching family is unique because the nonempty values of $E_{\sigma,\tau}$ are singletons.

If $F_{p,s}|_{\mathcal C_a}$ is separated, local equality on the cover forces
\[
\res_V^U(\sigma)=\res_V^U(\tau),
\]
so the unique local matching family has the unique amalgam in $E_{\sigma,\tau}(V)$. If equality fails on some member of the cover, there is no local matching family and the sheaf equalizer condition is vacuous. Hence each $E_{\sigma,\tau}$ is a sheaf.

Conversely, suppose all equality-locus presheaves are sheaves. If two sections of $F_{p,s}(V)$ become equal on a nonempty cover $\{V_\ell\to V\}$, apply the sheaf condition to the equality-locus presheaf for that pair over $V$. The unique local matching family then has an amalgam over $V$, which says precisely that the two original sections were already equal over $V$. This is separatedness.
\end{proof}

\begin{definition}[Prestack-level banding]\label{def:prestack-level-banding}
Let $\mathfrak P\to\mathcal C_a$ be a prestack in groupoids, and let $\Band$
be an abelian sheaf on $\mathcal C_a$. A \emph{prestack-level $\Band$-banding}
of $\mathfrak P$ is a family of isomorphisms of sheaves on the slices
\[
\beta^{\mathfrak P}_y:\Band|_U\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak P}(y)
\qquad (U\in\mathcal C_a,\ y\in\mathfrak P(U))
\]
compatible with restriction and conjugation by morphisms in $\mathfrak P$.
This is a datum only on objects of the prestack. It is not yet a banding of
the stackification, because stackification has additional objects obtained by
descent. Consequently this auxiliary notion is insufficient for the later gerbe
arguments unless it is followed by the descent construction of
\Cref{prop:stackification-band-descends}, or unless a stack-level banding has
been supplied separately.
\end{definition}

\begin{definition}[Stack-level banding]\label{def:stack-level-banding}
Let $\mathfrak X$ be a stack in groupoids on $\mathcal C_a$, and let $\Band$ be
an abelian sheaf on $\mathcal C_a$. A \emph{stack-level $\Band$-banding} of
$\mathfrak X$ is a family of isomorphisms of sheaves on the slices
\[
\beta_x:\Band|_U\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak X}(x)
\qquad (U\in\mathcal C_a,\ x\in\mathfrak X(U))
\]
with the following two compatibilities.
\begin{enumerate}[label=(\roman*),nosep]
\item If $f:V\to U$ is a morphism in $\mathcal C_a$, then the restriction of
$\beta_x$ to $\mathcal C_a/V$ is $\beta_{f^*x}$ under the canonical
identification
\[
\underline{\mathrm{Aut}}_{\mathfrak X}(x)|_V
\cong
\underline{\mathrm{Aut}}_{\mathfrak X}(f^*x).
\]
\item If $\varphi:x\to y$ is a morphism in $\mathfrak X(U)$, then conjugation by
$\varphi$ carries $\beta_x(a)$ to $\beta_y(a)$ for every local section $a$ of
$\Band|_U$.
\end{enumerate}
Thus a stack-level banding is a datum on all objects of the stack, not merely on
objects lying in a chosen prestack presentation.
\end{definition}

\begin{definition}[Stack-banded realization prestack]\label{def:realization-prestack}
Fix an admitted reference $r:\mathsf{Ref}_s$ at a state $p$, and write
\[
a=[r]_p,
\qquad
\mathcal C_a:=\mathcal C_{p,s}/a.
\]
A \emph{stack-banded realization prestack} for $r$ at $p$ is a presentation of
an already $\Band$-banded stack. Concretely, it is the tuple
\[
\bigl(
\mathfrak P_r,\mathfrak L_r,\iota_r,\Band,
\{\beta^{\mathfrak L}_x\}_{x\in\mathfrak L_r}
\bigr)
\]
with the following meaning. The presentation datum is a prestack in groupoids
\[
\mathfrak P_r\to \mathcal C_a
\]
whose isomorphism presheaves are sheaves, together with an identification of the
displayed stack and morphism with its stackification,
\[
\mathfrak L_r:=a(\mathfrak P_r),
\qquad
\iota_r:\mathfrak P_r\longrightarrow \mathfrak L_r
\]
\cite[Tag 02ZP]{StacksProject}. The fourth entry is a fixed abelian sheaf
$\Band$ on $\mathcal C_a$. The fifth entry is a stack-level banding on
\emph{all} objects of $\mathfrak L_r$. These data satisfy:
\begin{enumerate}[label=(\roman*),nosep]
\item The presheaf of isomorphism classes of the presentation is identified with
the restricted local-object functor,
\[
\pi_0^{\mathrm{pre}}(\mathfrak P_r):U\longmapsto
\Ob(\mathfrak P_r(U))/\!\cong,
\qquad
\pi_0^{\mathrm{pre}}(\mathfrak P_r)\cong F_{p,s}|_{\mathcal C_a};
\]
\item The stackification $\mathfrak L_r$ itself, not merely the presenting
prestack $\mathfrak P_r$, is equipped with the listed stack-level
$\Band$-banding in the sense of \Cref{def:stack-level-banding}. Explicitly, for
every object $U$ of $\mathcal C_a$ and every
\[
x\in\mathfrak L_r(U),
\]
including objects obtained only by descent, there is an isomorphism of sheaves
on $\mathcal C_a/U$
\[
\beta^{\mathfrak L}_x:\Band|_U\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x)
\]
compatible with restriction and conjugation by morphisms in $\mathfrak L_r$;
\item The stack banding in \textup{(ii)} restricts to the original prestack:
for every object $U$ of $\mathcal C_a$ and every $y\in\mathfrak P_r(U)$, the
canonical map
\[
\underline{\mathrm{Aut}}_{\mathfrak P_r}(y)\longrightarrow
\underline{\mathrm{Aut}}_{\mathfrak L_r}(\iota_r y)
\]
is an isomorphism of sheaves, and composing $\beta^{\mathfrak L}_{\iota_r y}$
with its inverse gives a prestack-level banding
\[
\beta^{\mathfrak P}_y:\Band|_U\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak P_r}(y)
\]
compatible with restriction and conjugation in $\mathfrak P_r$.
\end{enumerate}
Thus the band used later is explicitly a band of the stackification
$\mathfrak L_r$ on arbitrary stackified objects, not only a diagonal automorphism
identification for objects of $\mathfrak P_r$. In particular, later statements
may use $\beta^{\mathfrak L}_x$ for an arbitrary
$x\in\mathfrak L_r(U)$ only because item \textup{(ii)} is part of the
hypothesis, or because it has first been produced by
\Cref{prop:stackification-band-descends}. A prestack with only the restricted
maps $\beta^{\mathfrak P}_y$ in \textup{(iii)} is not a stack-banded realization
prestack in this sense unless the stack-level family in \textup{(ii)} has also
been supplied or constructed. The definition is an
existence-and-choice requirement on the whole fibred groupoid together with this
stack-level banding. Item \textup{(iii)} recovers the prestack-level banding of
\Cref{def:prestack-level-banding}; item \textup{(ii)} is the stronger datum used
by the component-gerbe theorems. Therefore a later statement that assumes
\Cref{def:realization-prestack} assumes the displayed tuple, not only the
prestack $\mathfrak P_r$ and not only the prestack-level maps
$\beta^{\mathfrak P}_y$. The definition is not a construction from the
pair $(F_{p,s}|_{\mathcal C_a},\Band)$. The next proposition gives a sufficient
construction of the stack-level datum in the common case where the
prestack-level banding has first been supplied.

Equivalently, the object used by the later gerbe theory is the
$\Band$-banded stack
\[
(\mathfrak L_r,\{\beta^{\mathfrak L}_x\}_{x\in\mathfrak L_r})
\]
together with a prestack presentation whose isomorphism classes recover
$F_{p,s}|_{\mathcal C_a}$. The prestack presentation is needed to connect back
to the local-object functor; the band used on component gerbes is already the
stack-level family on $\mathfrak L_r$.
\end{definition}

\begin{proposition}[The realization stack is already banded]\label{prop:realization-stack-already-banded}
Every stack-banded realization prestack in the sense of
\Cref{def:realization-prestack} determines a $\Band$-banded stack
\[
(\mathfrak L_r,\{\beta^{\mathfrak L}_x\}_{x\in\mathfrak L_r}).
\]
Concretely, for every object $U$ of $\mathcal C_a$ and every object
$x\in\mathfrak L_r(U)$ there is a specified isomorphism of sheaves on the slice
\[
\beta^{\mathfrak L}_x:\Band|_U\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x),
\]
compatible with restriction and conjugation. This assertion includes objects
which exist only after stackification. The diagonal maps on $\mathfrak P_r$ are
merely the restrictions of this stack-level family along
$\iota_r:\mathfrak P_r\to\mathfrak L_r$.
\end{proposition}

\begin{proof}
This is exactly item \textup{(ii)} of
\Cref{def:realization-prestack}. Item \textup{(iii)} identifies the restriction
of the same stack-level automorphism sheaves with the original prestack
automorphism sheaves, so the prestack-level banding is recovered by restriction
and is not a substitute for the stack band in later gerbe arguments.
\end{proof}

\begin{proposition}[Descent criterion for stackification bands]\label{prop:stackification-band-descends}
Let $\mathfrak P_r\to\mathcal C_a$ be a prestack in groupoids whose presheaf of
isomorphism classes is identified with $F_{p,s}|_{\mathcal C_a}$, and write
$\mathfrak L_r:=a(\mathfrak P_r)$. Suppose that $\mathfrak P_r$ is equipped with
a prestack-level $\Band$-banding in the sense of
\Cref{def:prestack-level-banding}. Then the fixed band descends through
stackification: there is a unique stackification banding
\[
\beta^{\mathfrak L}_x:\Band|_U\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x)
\qquad (x\in\mathfrak L_r(U))
\]
which is a stack-level $\Band$-banding of $\mathfrak L_r$ in the sense of
\Cref{def:stack-level-banding} and satisfies
\Cref{def:realization-prestack}, items~\textup{(ii)--(iii)}. It is obtained by local
presentation of stackified objects and is independent of all choices. Hence any
later full substack cut out by a connected-component condition inherits the same
band by restriction of this stack-level datum.

This proposition is the descent step used when the primitive input was only a
compatible prestack-level banding. If the realization input already supplies the
stack-level family of \Cref{def:realization-prestack}, item~\textup{(ii)}, the later
arguments use that supplied family directly.
\end{proposition}

\begin{proof}
We spell out the construction because this is the bridge used later for
component gerbes. Since $\mathfrak P_r$ is a prestack, the stackification
morphism $\iota_r:\mathfrak P_r\to\mathfrak L_r$ is fully faithful on
isomorphism sheaves: for any $y,z\in\mathfrak P_r(U)$ the natural map
\[
\underline{\mathrm{Isom}}_{\mathfrak P_r}(y,z)
\longrightarrow
\underline{\mathrm{Isom}}_{\mathfrak L_r}(\iota_r y,\iota_r z)
\]
is an isomorphism of sheaves on $\mathcal C_a/U$. Also every object of
$\mathfrak L_r$ is locally in the essential image of $\iota_r$
\cite[Tag 02ZP]{StacksProject}.

Let $x\in\mathfrak L_r(U)$. Choose a cover $\{U_i\to U\}$, objects
$y_i\in\mathfrak P_r(U_i)$, and isomorphisms
\[
\theta_i:\iota_r(y_i)\xrightarrow{\sim} x|_{U_i}.
\]
On each $U_i$, define
\[
\Band|_{U_i}\longrightarrow
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x|_{U_i})
\]
by first applying $\beta^{\mathfrak P}_{y_i}$ and then conjugating along
$\theta_i$.

These local maps agree on overlaps. Indeed, after restricting to
$U_{ij}$, the comparison
\[
\theta_j^{-1}\circ\theta_i:\iota_r(y_i)|_{U_{ij}}
\xrightarrow{\sim}
\iota_r(y_j)|_{U_{ij}}
\]
corresponds, by the displayed sheaf isomorphism of isomorphism sheaves, to a
unique section of
\[
\underline{\mathrm{Isom}}_{\mathfrak P_r}(y_i|_{U_{ij}},y_j|_{U_{ij}}).
\]
The conjugation-compatibility of the prestack-level banding says precisely that conjugation by
this prestack isomorphism identifies
$\beta^{\mathfrak P}_{y_i}$ with $\beta^{\mathfrak P}_{y_j}$ over the same
section of $\Band|_{U_{ij}}$. Conjugating further by $\theta_i$ and $\theta_j$
therefore makes the two local maps from $\Band|_{U_{ij}}$ to
\[
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x)|_{U_{ij}}
\]
coincide.

Since both $\Band$ and
$\underline{\mathrm{Aut}}_{\mathfrak L_r}(x)$ are sheaves on the slice, the
local identifications glue uniquely to the displayed
$\beta^{\mathfrak L}_x$. This glued map is an isomorphism because being an
isomorphism of sheaves is local on the target slice, and it restricts to the
isomorphisms already constructed on the cover $\{U_i\to U\}$. The construction
is independent of the chosen presentation: two presentations have a common
refinement, and on that common refinement the preceding overlap argument shows
that the resulting maps agree; the sheaf property then forces equality over
$U$.

Restriction compatibility is checked by pulling the chosen presentation along a
morphism $V\to U$; the construction just described becomes the construction for
$x|_V$. For conjugation, let $f:x\to z$ be a morphism in
$\mathfrak L_r(U)$. After refining $U$, the two objects and the arrow $f$ are
represented by objects and an arrow in $\mathfrak P_r$. The
conjugation-compatibility of the prestack-level banding therefore shows locally
that conjugation by $f$ carries $\beta^{\mathfrak L}_x(a)$ to
$\beta^{\mathfrak L}_z(a)$ for every local section $a$ of $\Band$. Since the two
sides are sections of a sheaf of automorphisms, the equality descends to $U$.

For an object $y\in\mathfrak P_r(U)$, choose the identity cover and the
identity presentation $\iota_r(y)\cong\iota_r(y)$. The construction then gives
exactly $\beta^{\mathfrak P}_y$ under the fully faithful map on automorphism
sheaves. Finally, uniqueness follows because every $x\in\mathfrak L_r(U)$ is
locally of the form $\iota_r(y_i)$, where any stackification banding compatible
with the prestack-level banding must restrict to the maps just constructed; sheaf
gluing then determines the global map uniquely. If a later full substack is
defined by a condition on connected components, its automorphism sheaf at an
object is the same automorphism sheaf as in $\mathfrak L_r$, so the same
stack-level banding restricts to it.
\end{proof}

\begin{corollary}[From a prestack band to a stack-banded input]\label{cor:prestack-band-gives-stack-banded-input}
Let $\mathfrak P_r\to\mathcal C_a$ be a prestack in groupoids whose presheaf of
isomorphism classes is identified with $F_{p,s}|_{\mathcal C_a}$. If
$\mathfrak P_r$ is equipped with a prestack-level $\Band$-banding, then
$\mathfrak P_r$ together with the stackification banding of
\Cref{prop:stackification-band-descends} is a stack-banded realization prestack
in the sense of \Cref{def:realization-prestack}. In particular, every object
$x\in\mathfrak L_r(U)$, not only every object coming from
$\mathfrak P_r(U)$, has a specified isomorphism
\[
\beta^{\mathfrak L}_x:\Band|_U\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x).
\]
\end{corollary}

\begin{proof}
Condition \textup{(i)} of \Cref{def:realization-prestack} is the assumed
identification of isomorphism classes. The descent proposition supplies the
stack-level banding required in \textup{(ii)} and proves that it restricts to
the original prestack-level banding as required in \textup{(iii)}.
\end{proof}

\begin{proposition}[Where the band lives after stackification]\label{prop:band-lives-on-stackification}
In the setting of \Cref{def:realization-prestack}, the band used by all later
gerbe and obstruction statements is the stack-level family
\[
\beta^{\mathfrak L}_x:\Band|_U\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x)
\qquad (U\in\mathcal C_a,\ x\in\mathfrak L_r(U)).
\]
This includes objects of $\mathfrak L_r(U)$ obtained by descent and not lying in
the essential image of $\mathfrak P_r(U)$. For an original prestack object
$y\in\mathfrak P_r(U)$, the diagonal prestack banding is only the restriction
of this stack-level banding along
\[
\iota_r:\mathfrak P_r\to\mathfrak L_r.
\]
Conversely, if the input starts with only a conjugation-compatible
prestack-level banding, then \Cref{prop:stackification-band-descends} is exactly
the descent step that produces the displayed stack-level family.
\end{proposition}

\begin{proof}
The displayed family is part of the data in
\Cref{def:realization-prestack}, item~\textup{(ii)}. Its restriction to an object
$y\in\mathfrak P_r(U)$ gives the prestack-level banding because
\Cref{def:realization-prestack}, item~\textup{(iii)} identifies
\[
\underline{\mathrm{Aut}}_{\mathfrak P_r}(y)
\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak L_r}(\iota_r y).
\]
Thus the prestack diagonal identifications are recovered from the stack band,
not the other way around inside later arguments. If a compatible
prestack-level banding is the primitive input, the construction and uniqueness
of the required stack-level banding are precisely
\Cref{prop:stackification-band-descends}. Hence every later occurrence of
$\Band$ as an automorphism sheaf of a stackified object is justified by the
stack-level family above.
\end{proof}

\begin{proposition}[Banded-stack handoff to components]\label{prop:banded-stack-handoff}
Let $\mathfrak P_r$ be a stack-banded realization prestack in the sense of
\Cref{def:realization-prestack}, and let
\[
\xi\in\pi_0(\mathfrak L_r)(U_0)
\]
be a section over an object $U_0$ of $\mathcal C_a$. On the slice
$\mathcal C_a/U_0$, define the full fibred subcategory
\[
\mathfrak L_{r,\xi}(V)
:=
\{\,x\in\mathfrak L_r(V)\mid [x]=\xi|_V\,\}.
\]
Then for every $V\to U_0$ and every $x\in\mathfrak L_{r,\xi}(V)$ one has an
equality of automorphism sheaves on $\mathcal C_a/V$,
\[
\underline{\mathrm{Aut}}_{\mathfrak L_{r,\xi}}(x)
=
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x),
\]
and therefore the stackification banding gives a canonical isomorphism
\[
\Band|_V
\xrightarrow{\beta^{\mathfrak L}_x}
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x)
=
\underline{\mathrm{Aut}}_{\mathfrak L_{r,\xi}}(x).
\]
The displayed map is defined for every object of the component subcategory,
including objects obtained only by descent in $\mathfrak L_r$. If the primitive
input was a compatible prestack-level banding, this is the unique descended map
of \Cref{prop:stackification-band-descends}. If the stack-level banding was
supplied directly, it is exactly the supplied map of
\Cref{def:realization-prestack}, item~\textup{(ii)}. Thus the passage from
$\mathfrak L_r$ to any connected-component subcategory does not introduce a new
band and does not use the prestack-level maps on objects of $\mathfrak P_r$.
\end{proposition}

\begin{proof}
The subcategory $\mathfrak L_{r,\xi}(V)$ is full in
$\mathfrak L_r(V)$ for every $V\to U_0$, and its defining condition is imposed
only on objects. Hence an automorphism of $x$ in the component subcategory is
exactly an automorphism of the same object in $\mathfrak L_r(V)$; the same
identification holds after every further restriction $W\to V$, giving equality
of automorphism sheaves on the slice. The displayed banding is then the
stack-level map $\beta^{\mathfrak L}_x$ from
\Cref{def:realization-prestack}, item~\textup{(ii)}, transported across this equality.
Restriction and conjugation compatibilities are inherited from the stack-level
banding of $\mathfrak L_r$. The final assertions distinguish the two allowed
ways of obtaining that stack-level map: it is either part of the realization
input, or it is the unique descent constructed in
\Cref{prop:stackification-band-descends}.
\end{proof}

\begin{corollary}[No prestack-only banding at component level]\label{cor:no-prestack-only-component-band}
Let $\mathfrak P_r$ be a stack-banded realization prestack and let
$\mathfrak L_{r,\xi}$ be a connected-component subcategory as in
\Cref{prop:banded-stack-handoff}. For an arbitrary object
$x\in\mathfrak L_{r,\xi}(V)$, the assertion
\[
\Band|_V\cong
\underline{\mathrm{Aut}}_{\mathfrak L_{r,\xi}}(x)
\]
is justified only by the stack-level map
\[
\beta^{\mathfrak L}_x:\Band|_V\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x)
\]
from \Cref{def:realization-prestack}, item~\textup{(ii)}, or by the descended map
constructed in \Cref{prop:stackification-band-descends}. The prestack-level
maps
\[
\beta^{\mathfrak P}_y:\Band|_V\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak P_r}(y)
\qquad (y\in\mathfrak P_r(V))
\]
do not apply to such an $x$ unless $x$ has first been identified through this
stack-level banding. In particular, every component gerbe and every branch
obstruction class below uses the band of $\mathfrak L_r$, not a
prestack-only diagonal band.
\end{corollary}

\begin{proof}
\Cref{prop:banded-stack-handoff} identifies the component automorphism sheaf of
$x$ with the ambient automorphism sheaf of the same stackified object in
$\mathfrak L_r$, and then applies $\beta^{\mathfrak L}_x$. That map is part of
\Cref{def:realization-prestack}, item~\textup{(ii)}. If the primitive input was only a
compatible prestack-level banding, \Cref{prop:stackification-band-descends} is
the separate descent argument that constructs this $\beta^{\mathfrak L}_x$ for
all stackified objects. A map $\beta^{\mathfrak P}_y$ is indexed by an object of
$\mathfrak P_r(V)$ and therefore has no domain of application to a general
object of $\mathfrak L_r(V)$ before that descent or supplied stack-level datum
has been invoked.
\end{proof}

\begin{remark}[Realization data are part of the hypotheses]\label{rem:realization-data-hypothesis}
The definition requires an actual fibred groupoid with sheaf-valued isomorphism presheaves, an identification of its presheaf of isomorphism classes with $F_{p,s}|_{\mathcal C_a}$, and a banding of the stackification on arbitrary objects as in \Cref{def:realization-prestack}, item~\textup{(ii)}. If the banding is not supplied directly, \Cref{prop:stackification-band-descends} constructs it as the unique descent of a compatible prestack-level banding in the sense of \Cref{def:prestack-level-banding}; in neither case is it inferred from a bare sheaf of groups without the prestack and its compatibility data. For two objects $x,y\in\mathfrak P_r(U)$, the off-diagonal sheaf
\[
\underline{\operatorname{Isom}}(x,y)(V)
:=
\operatorname{Isom}_{\mathfrak P_r(V)}
(x|_V,y|_V)
\]
on $\mathcal C_a/U$ is part of the supplied prestack. It is not determined by the diagonal band. Where it is locally nonempty, the band makes it a $\Band|_U$-torsor; it becomes a copy of $\Band|_U$ only after choosing a local isomorphism as an origin, and such choices have their own descent obstruction.

A sheaf of abelian groups $\Band$ therefore supplies only the intended automorphism groups. It does not supply the objects, the restriction functors, the off-diagonal isomorphism presheaves, or the coherence maps that make composition associative after restriction.

Thus branch gerbes and branch obstruction classes are attached only after such realization data have been chosen, or after they have been constructed under a genuine criterion such as the separated split-lift criterion of \Cref{prop:split-realization-prestack}. The phrase ``with band $\Band$'' is never used below as a substitute for the existence of the realization prestack itself. If the realization prestack has not been supplied or constructed, the component gerbes and branch-obstruction classes are not defined.
\end{remark}

\begin{lemma}[Non-separated pairs require off-diagonal torsors]\label{lem:nonseparated-pairs-need-torsors}
Assume that a realization prestack $\mathfrak P_r$ as in \Cref{def:realization-prestack} has been chosen. Let $U$ be an object of $\mathcal C_a$, let
\[
\sigma,\tau\in F_{p,s}(U),
\]
and choose representatives $x,y\in \mathfrak P_r(U)$ of these two isomorphism classes. Suppose that $\sigma\neq\tau$ in $F_{p,s}(U)$, but that for some cover $\{U_i\to U\}_{i\in I}$ one has
\[
\res_{U_i}^{U}(\sigma)=\res_{U_i}^{U}(\tau)
\qquad\text{for every }i.
\]
Then the sheaf
\[
\underline{\mathrm{Isom}}(x,y)
\]
on $\mathcal C_a/U$ is a locally nonempty $\Band|_U$-torsor with no global section over $U$. Equivalently, any realization prestack of such a non-separated pair must choose nontrivial off-diagonal torsor data; that data is not determined by the band $\Band$ alone.
\end{lemma}

\begin{proof}
The equality of the restricted classes over each $U_i$ means that
\[
x|_{U_i}\cong y|_{U_i}
\]
in the fibre over $U_i$. Hence $\underline{\mathrm{Isom}}(x,y)$ has local sections on the cover. Since $\mathfrak P_r$ is a prestack, this isomorphism presheaf is a sheaf. Composition with automorphisms of $x$, identified by the banding with $\Band|_U$, acts simply transitively on local isomorphisms wherever they exist, so $\underline{\mathrm{Isom}}(x,y)$ is a $\Band|_U$-torsor \cite[Chap.~III]{MacLaneMoerdijk1994}, \cite[Sec.~11]{StacksProject}. If it had a section over $U$, then $x\cong y$ in $\mathfrak P_r(U)$, contradicting $\sigma\neq\tau$ under the identification $\pi_0^{\mathrm{pre}}(\mathfrak P_r)\cong F_{p,s}|_{\mathcal C_a}$.
\end{proof}

\begin{corollary}[Non-separated pairs force torsors]\label{cor:nonseparated-forces-h1-class}
Under the hypotheses of \Cref{lem:nonseparated-pairs-need-torsors}, the sheaf
\[
\underline{\mathrm{Isom}}(x,y)
\]
defines a nonzero class in the pointed torsor set
\[
H^1(U,\Band).
\]
Equivalently, a realization prestack for a non-separated pair can exist only if
the slice admits the required nontrivial locally nonempty $\Band|_U$-torsor. The
class is part of the nonsplit realization data; it is not determined by the
diagonal band.
\end{corollary}

\begin{proof}
By \Cref{lem:nonseparated-pairs-need-torsors}, the sheaf
$\underline{\mathrm{Isom}}(x,y)$ is a locally nonempty $\Band|_U$-torsor with no
global section. Torsors under the abelian sheaf $\Band|_U$ are classified by
$H^1(U,\Band)$, and the zero class is represented by the trivial torsor,
equivalently by a torsor with a global section
\cite[Chap.~III]{MacLaneMoerdijk1994}, \cite[Sec.~11]{StacksProject}. Hence the
associated class is nonzero.
\end{proof}

\begin{corollary}[A band does not force a realization]\label{cor:band-does-not-force-realization}
Let $\Band$ be an abelian sheaf on $\mathcal C_a$. Suppose that every locally nonempty $\Band|_U$-torsor on $\mathcal C_a/U$ has a global section, equivalently that
\[
H^1(U,\Band)=0
\]
for every object $U$ of $\mathcal C_a$. If $F_{p,s}|_{\mathcal C_a}$ admits a
stack-banded realization prestack with band $\Band$ in the sense of
\Cref{def:realization-prestack}, then $F_{p,s}|_{\mathcal C_a}$ is separated.
Consequently, under this torsor-triviality hypothesis, any non-separated
local-object presheaf admits no such realization prestack with band $\Band$.
\end{corollary}

\begin{proof}
Assume that a stack-banded realization prestack with band $\Band$ exists in the
sense of \Cref{def:realization-prestack}. If $F_{p,s}|_{\mathcal C_a}$ were not separated, then there would be an object $U$, distinct sections
\[
\sigma,\tau\in F_{p,s}(U),
\]
and a cover $\{U_i\to U\}$ such that
\[
\res_{U_i}^{U}(\sigma)=\res_{U_i}^{U}(\tau)
\]
for every $i$. By \Cref{cor:nonseparated-forces-h1-class}, representatives of
$\sigma$ and $\tau$ in the realization prestack would determine a nonzero class
in $H^1(U,\Band)$. This contradicts the assumed vanishing. Hence
$F_{p,s}|_{\mathcal C_a}$ is separated. The final sentence is the contrapositive.
\end{proof}

\begin{proposition}[Realization is additional off-diagonal data]\label{prop:realization-additional-data}
Fix the set-valued local-object presheaf $F_{p,s}|_{\mathcal C_a}$ and an
abelian sheaf $\Band$ on $\mathcal C_a$. These data determine only the desired
isomorphism classes and diagonal automorphism sheaves of a possible
realization. A stack-banded realization prestack with this band exists only after one has
also supplied, compatibly with restriction and composition, sheaves of
off-diagonal isomorphisms between representatives of distinct local classes.

More precisely, if $\sigma,\tau\in F_{p,s}(U)$ are distinct but become equal on
a cover of $U$, then every stack-banded realization prestack with band $\Band$ determines a
locally nonempty $\Band|_U$-torsor with no global section over $U$. This is a
necessary condition on any nonsplit realization, not a construction of one:
the torsors for all pairs must also carry restriction and composition maps
satisfying the fibred-groupoid axioms. Hence, on sites where all such torsors
are trivial, no stack-banded realization prestack with band $\Band$ can realize a
non-separated $F_{p,s}|_{\mathcal C_a}$.
\end{proposition}

\begin{proof}
The diagonal part is exactly the restriction required in \Cref{def:realization-prestack}, item~\textup{(iii)}: for an
object representing a class $\sigma$, its automorphism sheaf is identified
with $\Band$. Nothing in the pair $(F_{p,s},\Band)$ specifies the sheaf of
isomorphisms between representatives of two different classes, nor the
restriction and composition maps involving those isomorphisms. Those are part
of the fibred groupoid required in \Cref{def:realization-prestack}.

For such a non-separated pair, \Cref{lem:nonseparated-pairs-need-torsors}
shows that any realization prestack produces a torsor under $\Band|_U$ that is
locally nonempty and has no global section. If all locally nonempty torsors
under $\Band|_U$ are trivial for every $U$, the previous corollary gives the
claimed non-existence statement.
\end{proof}

\begin{definition}[The split fibred groupoid]\label{def:split-fibred-groupoid}
Fix an admitted reference $r:\mathsf{Ref}_s$ at a state $p$, write
\[
a=[r]_p,
\qquad
\mathcal C_a:=\mathcal C_{p,s}/a,
\]
and let $\Band$ be any abelian sheaf on $\mathcal C_a$. Define the split fibred groupoid
\[
\mathfrak P_r^{\mathrm{split}}\to \mathcal C_a
\]
as follows. For each object $U$ of $\mathcal C_a$, the fibre
\[
\mathfrak P_r^{\mathrm{split}}(U)
\]
has object set $F_{p,s}(U)$ and morphism sets
\[
\mathrm{Hom}_{\mathfrak P_r^{\mathrm{split}}(U)}(\sigma,\tau)
:=
\begin{cases}
\Band(U), & \sigma=\tau,\\
\varnothing, & \sigma\neq\tau,
\end{cases}
\]
with composition given by the group law on $\Band(U)$, and with pullback induced by the restriction maps of $F_{p,s}$ and $\Band$.
These are only the Hom sets in the single fibre over $U$. For a fixed pair
$\sigma,\tau\in F_{p,s}(U)$, the associated isomorphism presheaf on
$\mathcal C_a/U$ is obtained by first restricting both objects to each
$V\to U$ and then applying the displayed Hom rule in the fibre over $V$.
\end{definition}

This fibrewise definition is therefore not the prestack test. The prestack test
is performed on each presheaf of isomorphisms over the slice $\mathcal C_a/U$.
The following lemma is the decisive calculation.

\begin{lemma}[Isomorphism presheaves in the split lift]\label{lem:split-isomorphism-presheaf}
In the setting of \Cref{def:split-fibred-groupoid}, for fixed $\sigma,\tau\in F_{p,s}(U)$, the associated isomorphism presheaf on $\mathcal C_a/U$ is
\[
I_{\sigma,\tau}(V)=
\begin{cases}
\Band(V), & \res_V^U(\sigma)=\res_V^U(\tau),\\
\varnothing, & \res_V^U(\sigma)\neq\res_V^U(\tau).
\end{cases}
\]
Equivalently,
\[
I_{\sigma,\tau}\cong E_{\sigma,\tau}\times \Band|_U
\]
as presheaves on $\mathcal C_a/U$, where $E_{\sigma,\tau}$ is the equality-locus presheaf of \Cref{lem:equality-loci-separated} and the product is interpreted so that the factor $E_{\sigma,\tau}(V)=\varnothing$ gives the empty set. Thus the support of $I_{\sigma,\tau}$ is the equality sieve of $(\sigma,\tau)$. If $\sigma\neq\tau$ over $U$ but the two sections become equal on some refinement $V\to U$, then $I_{\sigma,\tau}$ has empty value at $U$ and nonempty value at $V$; this is the basic non-separated phenomenon. The identifications $I_{\sigma,\tau}\cong \Band|_U$ and $I_{\sigma,\tau}\cong\varnothing$ are valid only in the two special constant-support cases, namely equality on every refinement or inequality on every refinement.
\end{lemma}

\begin{proof}
For a morphism $f:V\to U$ in $\mathcal C_a$, the pullback functor in the split fibred groupoid sends objects by
\[
f^*(\sigma):=\res_V^U(\sigma)\in F_{p,s}(V)
\]
and automorphisms by the sheaf restriction
\[
f^*:\Band(U)\to \Band(V).
\]
Thus, over a refinement $V\to U$, the induced objects are $\res_V^U(\sigma)$ and $\res_V^U(\tau)$. By the defining Hom rule in the fibre over $V$, their isomorphism set is $\Band(V)$ exactly when these two restrictions are equal, and is empty otherwise. This proves the displayed formula. The final description is the same formula written as the product of the equality-locus presheaf with the sheaf of groups $\Band|_U$.
\end{proof}

\begin{proposition}[No fibrewise dichotomy for split isomorphisms]\label{prop:no-fibrewise-split-dichotomy}\label{cor:no-fibrewise-split-dichotomy}
In the setting of \Cref{lem:split-isomorphism-presheaf}, the split isomorphism presheaf between
\[
\sigma,\tau\in F_{p,s}(U)
\]
is controlled by the equality sieve
\[
\mathrm{Eq}_U(\sigma,\tau)=\{\,V\to U\mid \res_V^U(\sigma)=\res_V^U(\tau)\,\}.
\]
It is the full band sheaf $\Band|_U$ exactly when this sieve is all of
$\mathcal C_a/U$. It is empty on every refinement exactly when this sieve is
empty. When the equality sieve is proper and nonempty, the presheaf has mixed
support and is neither constant-support extreme.

In particular, if $\sigma\neq\tau$ in $F_{p,s}(U)$ but there is a refinement $V\to U$ with
\[
\res_V^U(\sigma)=\res_V^U(\tau),
\]
then
\[
I_{\sigma,\tau}(U)=\varnothing,
\qquad
I_{\sigma,\tau}(V)=\Band(V),
\]
so the isomorphism presheaf has mixed support on the slice and cannot be
identified with either constant-support extreme.
\end{proposition}

\begin{proof}
This is the formula of \Cref{lem:split-isomorphism-presheaf} read on every object of the slice $\mathcal C_a/U$. The full band sheaf occurs precisely when the equality sieve is the whole slice, and the everywhere-empty case occurs precisely when the equality sieve is empty. If the equality sieve is proper and nonempty, the presheaf has both nonempty and empty values, so it is neither of the two constant alternatives.
\end{proof}

\begin{remark}[The base fibre is not enough]\label{rem:base-fibre-not-enough}
The Hom set in the single fibre over $U$ only records whether $\sigma=\tau$ in $F_{p,s}(U)$. The prestack condition asks instead for the sheaf condition on the whole slice $\mathcal C_a/U$. Thus a pair with $\sigma\neq\tau$ over $U$ can still produce local split isomorphisms after pullback to a cover, and those local isomorphisms are exactly what obstruct the split construction in the non-separated case.
\end{remark}

\begin{proposition}[Non-separated pairs obstruct the split lift]\label{prop:nonseparated-pairs-obstruct-split}
Assume there are an object $U$ of $\mathcal C_a$, a nonempty cover
\[
\{U_i\to U\}_{i\in I},
\]
and distinct sections
\[
\sigma,\tau\in F_{p,s}(U)
\]
such that
\[
\res_{U_i}^{U}(\sigma)=\res_{U_i}^{U}(\tau)
\qquad\text{for every }i\in I.
\]
Then, for every abelian sheaf $\Band$ on $\mathcal C_a$, the split fibred groupoid
$\mathfrak P_r^{\mathrm{split}}$ is not a prestack.

More explicitly, the split isomorphism presheaf between $\sigma$ and $\tau$ has
\[
I_{\sigma,\tau}(U)=\varnothing,
\qquad
I_{\sigma,\tau}(U_i)=\Band(U_i)
\]
for all $i$. The identity sections of $\Band(U_i)$ form a matching family on the
cover, but there is no section over $U$ to which they can glue. Thus the failure is
an actual failure of the sheaf axiom for one isomorphism presheaf, not a
choice-dependent defect of the band.
\end{proposition}

\begin{proof}
The formula of \Cref{lem:split-isomorphism-presheaf} gives
\[
I_{\sigma,\tau}(U)=\varnothing
\]
because $\sigma\neq\tau$ in $F_{p,s}(U)$. On each member of the cover, the same
formula gives
\[
I_{\sigma,\tau}(U_i)=\Band(U_i)
\]
because the restrictions of $\sigma$ and $\tau$ agree there. On every common
refinement of two members of the cover, the restrictions of the identity
sections $1\in \Band(U_i)$ and $1\in\Band(U_j)$ are both the identity section.
Thus these local identities are a matching family in the pretopology sense.
If $I_{\sigma,\tau}$ were a sheaf, this family would have a unique amalgam in
$I_{\sigma,\tau}(U)$, but that set is empty. Hence an isomorphism presheaf is not
a sheaf, and the split fibred groupoid is not a prestack.
\end{proof}

\begin{proposition}[Exact split-lift criterion]\label{prop:split-realization-prestack}
In the setting of \Cref{def:split-fibred-groupoid}, the following conditions are equivalent:
\begin{enumerate}[label=(\roman*),nosep]
\item the split fibred groupoid $\mathfrak P_r^{\mathrm{split}}$ is a prestack;
\item for every object $U$ of $\mathcal C_a$ and every pair $\sigma,\tau\in F_{p,s}(U)$, the restriction-dependent split isomorphism presheaf
\[
I_{\sigma,\tau}(V)=
\begin{cases}
\Band(V), & \res_V^U(\sigma)=\res_V^U(\tau),\\
\varnothing, & \res_V^U(\sigma)\neq\res_V^U(\tau)
\end{cases}
\]
is a sheaf on $\mathcal C_a/U$;
\item for every such pair, the equality-locus presheaf $E_{\sigma,\tau}$ is a sheaf on $\mathcal C_a/U$;
\item $F_{p,s}|_{\mathcal C_a}$ is separated in the sense of \Cref{def:slice-separated-local-object}.
\end{enumerate}
Consequently, under separatedness, $\mathfrak P_r^{\mathrm{split}}$ is a
stack-banded realization prestack with band $\Band$ and
\[
\pi_0^{\mathrm{pre}}(\mathfrak P_r^{\mathrm{split}})
\cong
F_{p,s}|_{\mathcal C_a}.
\]
Without separatedness the split construction is not a prestack. This failure
already occurs for the explicit isomorphism presheaf in \textup{(ii)}: local
identity sections over a cover need not have a global split amalgam. In that
case the band alone still supplies only the diagonal automorphism sheaves; it
does not supply a replacement realization prestack. Any nonsplit realization
prestack, if one exists, must add the off-diagonal torsor data described in
\Cref{prop:realization-additional-data}.

Equivalently, the proposition proves exactly the following replacement for the
fibrewise dichotomy one might write down before passing to the slice: the split
lift is a realization prestack for the original local-object presheaf if and only
if equality in that presheaf is local for the chosen covers. The conclusion is a
conditional construction in the separated case and a failure criterion in the
non-separated case; it is not a band-only realization principle. The theorem has
no additional positive content in the mixed-support case: if the equality sieve
of a pair is proper but covers the base, the displayed presheaf is exactly the
one whose local identity sections fail to glue.

This is the only positive split-lift statement used later. Whenever later
sections invoke a realization prestack in a non-separated situation, the
prestack is not being supplied by this proposition; it is part of the separate
realization input and includes the off-diagonal torsors and their coherent
composition laws.
\end{proposition}

\begin{proof}
For a fibred groupoid, being a prestack is exactly the sheaf condition for all isomorphism presheaves. In the split fibred groupoid, \Cref{lem:split-isomorphism-presheaf,cor:no-fibrewise-split-dichotomy} identify the relevant presheaf on $\mathcal C_a/U$ with the restriction-dependent presheaf
\[
I_{\sigma,\tau}(V)=
\begin{cases}
\Band(V), & \res_V^U(\sigma)=\res_V^U(\tau),\\
\varnothing, & \res_V^U(\sigma)\neq\res_V^U(\tau).
\end{cases}
\]
This proves the equivalence of \textup{(i)} and \textup{(ii)}. The value of $I_{\sigma,\tau}$ is not decided in the fibre over $U$; it is decided separately after every refinement $V\to U$. Thus the proof never replaces $I_{\sigma,\tau}$ by a constant-support presheaf unless the equality sieve is respectively all of $\mathcal C_a/U$ or empty. The mixed-support presheaf is the generic case, and it is the case that detects non-separatedness.

We compare \textup{(ii)} and \textup{(iii)} directly. Assume first that $E_{\sigma,\tau}$ is a sheaf. Let $\{V_\ell\to V\}$ be a nonempty cover in $\mathcal C_a/U$, and let $(\alpha_\ell)$ be a matching family in $I_{\sigma,\tau}(V_\ell)$. The existence of $\alpha_\ell$ implies
\[
\res_{V_\ell}^U(\sigma)=\res_{V_\ell}^U(\tau)
\]
for every $\ell$, hence the corresponding singleton sections of $E_{\sigma,\tau}(V_\ell)$ form a matching family. Since $E_{\sigma,\tau}$ is a sheaf, this family amalgamates over $V$, so
\[
\res_V^U(\sigma)=\res_V^U(\tau).
\]
Thus all $\alpha_\ell$ are sections of the sheaf $\Band$ over the cover of $V$, and they glue uniquely in $\Band(V)$. The displayed formula then gives a unique amalgam in $I_{\sigma,\tau}(V)$. Hence $I_{\sigma,\tau}$ is a sheaf.

Conversely assume that $I_{\sigma,\tau}$ is a sheaf. Let $\{V_\ell\to V\}$ be a nonempty cover and suppose a matching family exists in $E_{\sigma,\tau}(V_\ell)$. This is exactly the assertion that
\[
\res_{V_\ell}^U(\sigma)=\res_{V_\ell}^U(\tau)
\]
for every $\ell$. The identity elements of $\Band(V_\ell)$ form a matching family in $I_{\sigma,\tau}(V_\ell)$, so the sheaf property for $I_{\sigma,\tau}$ gives an amalgam in $I_{\sigma,\tau}(V)$. The displayed formula for $I_{\sigma,\tau}(V)$ can be nonempty only when
\[
\res_V^U(\sigma)=\res_V^U(\tau).
\]
Therefore the matching family in $E_{\sigma,\tau}$ has its unique amalgam over $V$, so $E_{\sigma,\tau}$ is a sheaf. This proves \textup{(ii)}$\Leftrightarrow$\textup{(iii)} without collapsing $I_{\sigma,\tau}$ to a constant-support presheaf determined in the base fibre.

The equivalence of \textup{(iii)} and \textup{(iv)} is \Cref{lem:equality-loci-separated}.

For clarity, the two directions that matter most can also be read without the auxiliary presheaf $E_{\sigma,\tau}$. If $F_{p,s}|_{\mathcal C_a}$ is separated and local split isomorphisms exist on a cover, then the underlying local equalities of restrictions glue to equality over the base; the sheaf property of $\Band$ then glues the actual automorphism sections. Conversely, if $F_{p,s}|_{\mathcal C_a}$ is not separated, choose a cover $\{V_\ell\to V\}$ and distinct $\sigma,\tau\in F_{p,s}(V)$ whose restrictions agree on every $V_\ell$. The equality sieve of this pair contains the cover but not the identity map $V\to V$. Therefore the identity elements $1\in\Band(V_\ell)$ form a matching family in $I_{\sigma,\tau}(V_\ell)$, while
\[
I_{\sigma,\tau}(V)=\varnothing.
\]
Hence an isomorphism presheaf fails the sheaf condition and the split fibred groupoid is not a prestack.

By construction, two objects of $\mathfrak P_r^{\mathrm{split}}(U)$ are isomorphic if and only if they are equal as elements of $F_{p,s}(U)$. Therefore
\[
\pi_0^{\mathrm{pre}}(\mathfrak P_r^{\mathrm{split}})(U)\cong F_{p,s}(U)
\]
for every $U$, giving condition \textup{(i)} of \Cref{def:realization-prestack}. The split automorphism sheaf of every object $\sigma$ is exactly $\Band|_U$, restrictions are the restrictions of $\Band$, and conjugation is trivial since $\Band$ is abelian. Hence \Cref{prop:stackification-band-descends} supplies the stack banding of arbitrary stackified objects required in \Cref{def:realization-prestack}, item~\textup{(ii)}, and the split diagonal banding agrees with it as required in \textup{(iii)}.
\end{proof}

\begin{remark}[How to read the split criterion]\label{rem:split-criterion-reading}
\Cref{prop:split-realization-prestack} has only the following positive content:
if the restricted local-object presheaf is separated, the split fibred groupoid
is a realization prestack for that presheaf. Its negative content is equally
important: if a pair of sections is distinct over $U$ but equal after a cover,
then the split isomorphism presheaf is supported precisely on the refinements
where equality holds. The local identity sections over that cover are a
matching family with no global split amalgam. Thus the obstruction is visible
inside a single isomorphism presheaf and cannot be removed by changing the
abelian band. A nonsplit realization, when it exists, is therefore a separate
choice of off-diagonal torsors and compatible composition, not an automatic
output of the data $(F_{p,s},\Band)$.
\end{remark}

\begin{example}[A non-separated split lift]\label{ex:nonseparated-split-lift}
Let the slice site contain objects
\[
c\to u\to a,
\qquad
c\to v\to a,
\]
and let $\{u\to a,v\to a\}$ be a cover, with $c$ representing the common overlap. Define a local-object presheaf by
\[
F_{p,s}(a)=\{\sigma,\tau\},
\qquad
F_{p,s}(u)=F_{p,s}(v)=F_{p,s}(c)=\{*\},
\]
with both $\sigma$ and $\tau$ restricting to $*$ on $u$ and on $v$. Let $\Band$ be any abelian sheaf, for instance the constant sheaf $\mathbb Z/2\mathbb Z$.

For the split fibred groupoid, the isomorphism presheaf between the two objects $\sigma,\tau\in F_{p,s}(a)$ satisfies
\[
I_{\sigma,\tau}(a)=\varnothing,
\qquad
I_{\sigma,\tau}(u)=\Band(u),
\qquad
I_{\sigma,\tau}(v)=\Band(v).
\]
The identity sections on $u$ and $v$ agree after restriction to the overlap $c$, so they form a matching family. There is no amalgam over $a$, because $I_{\sigma,\tau}(a)$ is empty. Thus the split fibred groupoid is not a prestack. The failure is not a property of the chosen band; it is caused by the non-separated pair $\sigma,\tau$.
\end{example}

\begin{corollary}[No band-only split prestack theorem]\label{cor:nonseparated-split-fails}
Suppose there is a cover $\{V_\ell\to V\}_{\ell\in L}$ in $\mathcal C_a$ and distinct sections
\[
\sigma,\tau\in F_{p,s}(V)
\]
whose restrictions to every $V_\ell$ are equal. Then, for every abelian sheaf $\Band$ on $\mathcal C_a$, the split fibred groupoid $\mathfrak P_r^{\mathrm{split}}$ is not a prestack. In particular, fixing a band does not, by itself, produce a banded realization prestack for an arbitrary set-valued local-object presheaf; in the non-separated case a nonsplit realization, if it is supplied, must add off-diagonal torsors as in \Cref{lem:nonseparated-pairs-need-torsors}.
\end{corollary}

\begin{proof}
For the pair $\sigma,\tau$, the presheaf of isomorphisms satisfies
\[
I_{\sigma,\tau}(V)=\varnothing
\]
because $\sigma\neq\tau$. On each member of the cover, however,
\[
I_{\sigma,\tau}(V_\ell)=\Band(V_\ell),
\]
and the identity sections form a matching family on common refinements. If $I_{\sigma,\tau}$ were a sheaf, that matching family would have an amalgam in $I_{\sigma,\tau}(V)$, contradicting $I_{\sigma,\tau}(V)=\varnothing$. Thus an isomorphism presheaf fails the sheaf condition, so the split fibred groupoid is not a prestack.
\end{proof}

\begin{corollary}[Split lift after sheafification]\label{cor:sheafified-split-lift}
Let
\[
F^\sharp:=a\bigl(F_{p,s}|_{\mathcal C_a}\bigr)
\]
be the sheafification of the restricted local-object presheaf on $\mathcal C_a$. For any abelian sheaf $\Band$ on $\mathcal C_a$, the split fibred groupoid with object set $F^\sharp(U)$ over $U$ and morphism sets
\[
\mathrm{Hom}(\xi,\zeta)=
\begin{cases}
\Band(U), & \xi=\zeta,\\
\varnothing, & \xi\neq \zeta
\end{cases}
\]
is a stack-banded realization prestack with band $\Band$ and presheaf of
isomorphism classes $F^\sharp$. It is a realization prestack for the original
local-object presheaf $F_{p,s}|_{\mathcal C_a}$ only when the sheafification unit
\[
F_{p,s}|_{\mathcal C_a}\longrightarrow F^\sharp
\]
is an isomorphism.
\end{corollary}

\begin{proof}
The sheaf $F^\sharp$ is separated, so the proof of \Cref{prop:split-realization-prestack} applies verbatim with $F_{p,s}|_{\mathcal C_a}$ replaced by $F^\sharp$. Hence all split isomorphism presheaves are sheaves and the resulting split fibred groupoid is a realization prestack with diagonal automorphism sheaf $\Band$, with restriction and conjugation compatibility. Its stackification banding, required in \Cref{def:realization-prestack}, item~\textup{(ii)}, is supplied by \Cref{prop:stackification-band-descends}, and it restricts back to the split diagonal banding as required in \textup{(iii)}. Its isomorphism classes are exactly $F^\sharp$ by construction. Therefore it realizes the original presheaf precisely in the case where the displayed unit identifies $F_{p,s}|_{\mathcal C_a}$ with $F^\sharp$.
\end{proof}

\begin{proposition}[No band-only realization principle]\label{prop:no-band-only-realization-principle}
Fix the restricted local-object presheaf $F_{p,s}|_{\mathcal C_a}$ and an
abelian sheaf $\Band$ on $\mathcal C_a$. The following are the only
realization conclusions established by the split analysis:
\begin{enumerate}[label=(\roman*),nosep]
\item if $F_{p,s}|_{\mathcal C_a}$ is separated, the split fibred groupoid is a
realization prestack for the original presheaf;
\item for arbitrary $F_{p,s}|_{\mathcal C_a}$, the split construction after
sheafification is a realization prestack for
$a(F_{p,s}|_{\mathcal C_a})$, not for the original presheaf unless the
sheafification unit is an isomorphism;
\item if $F_{p,s}|_{\mathcal C_a}$ is non-separated and all locally nonempty
$\Band|_U$-torsors are trivial on every slice $\mathcal C_a/U$, then no
stack-banded realization prestack with band $\Band$ realizes the original
presheaf;
\item in the remaining non-separated cases, a realization prestack can only be
used after the whole nonsplit fibred groupoid has been supplied as additional
input: off-diagonal isomorphism sheaves, restriction functors, composition, and
the prestack sheaf condition on all isomorphism presheaves. In particular, each
locally equal but globally distinct pair requires a specified nonzero torsor
class as in \Cref{cor:nonseparated-forces-h1-class}.
\end{enumerate}
Consequently, fixing $\Band$ alone never licenses the later gerbe, component,
or obstruction constructions. Every later invocation of those constructions is
conditional on an actual chosen stack-banded realization prestack, or on one
produced by the separated split-lift criterion together with the stackification
banding of \Cref{prop:stackification-band-descends}. These alternatives are exhaustive for the
arguments in this paper:
\begin{enumerate}[label=(\alph*),nosep]
\item if $F_{p,s}|_{\mathcal C_a}$ is separated, the split prestack realizes
$F_{p,s}|_{\mathcal C_a}$;
\item for arbitrary $F_{p,s}|_{\mathcal C_a}$, the split prestack after
sheafification realizes $a(F_{p,s}|_{\mathcal C_a})$;
\item if $F_{p,s}|_{\mathcal C_a}$ is non-separated and all relevant
$\Band$-torsors are trivial, no stack-banded realization prestack with this
band realizes $F_{p,s}|_{\mathcal C_a}$;
\item in the remaining non-separated cases, a nonsplit prestack must be supplied
and checked as additional data.
\end{enumerate}
In particular, there is no implication of the form
\[
(F_{p,s}|_{\mathcal C_a},\Band)
\quad\Longrightarrow\quad
\text{a stack-banded realization prestack with band }\Band
\]
without an additional separatedness hypothesis or an explicit nonsplit
construction of the off-diagonal torsors and their composition laws.
\end{proposition}

\begin{proof}
Statement \textup{(i)} is exactly the positive direction of
\Cref{prop:split-realization-prestack}. Statement \textup{(ii)} is
\Cref{cor:sheafified-split-lift}. Statement \textup{(iii)} is
\Cref{cor:band-does-not-force-realization}. Finally, statement \textup{(iv)}
is \Cref{prop:realization-additional-data}: when distinct local classes become
equal only after refinement, any nonsplit realization must provide locally
nonempty off-diagonal $\Band$-torsors and compatible composition data that are
not present in the pair $(F_{p,s}|_{\mathcal C_a},\Band)$; the required torsor
class is nonzero by \Cref{cor:nonseparated-forces-h1-class}. These alternatives
exhaust the conclusions proved here, so the later gerbe and obstruction
classes are attached to chosen realization input, not to a bare band.
\end{proof}

\begin{proposition}[Dependency ledger for later realization statements]\label{prop:realization-dependency-ledger}
Every later statement in this paper that uses component gerbes, branch
obstruction classes, finite-nerve presentations, or visible quotients is to be
read with the following explicit dependencies.
\begin{enumerate}[label=(\roman*),nosep]
\item The set-valued presheaf $F_{p,s}|_{\mathcal C_a}$ supplies only local
isomorphism classes. It does not supply a fibred groupoid.
\item The bare band $\Band$ supplies only the coefficient sheaf intended for
diagonal automorphisms. Once it is attached to an actual prestack compatibly
with restriction and conjugation, the stack-banded realization input also
includes the sheaf isomorphisms
\[
\beta^{\mathfrak L}_x:\Band|_U\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x)
\]
for every $x\in\mathfrak L_r(U)$. These maps are either supplied directly in
\Cref{def:realization-prestack}, item~\textup{(ii)} or, when the primitive datum is a
compatible prestack-level banding, uniquely obtained by
\Cref{prop:stackification-band-descends}. This still does not supply
off-diagonal isomorphism sheaves, their restriction maps, or their composition
law.
\item A realization prestack for the original presheaf is available from the
split construction exactly under the separatedness condition of
\Cref{prop:split-realization-prestack}.
\item Without separatedness, the split construction fails on the mixed-support
isomorphism presheaves of \Cref{lem:split-isomorphism-presheaf}; in torsor-trivial
slices no nonsplit replacement with the same band exists by
\Cref{cor:band-does-not-force-realization}; in non-torsor-trivial slices a
nonsplit replacement still has to choose the nonzero off-diagonal torsor classes
of \Cref{cor:nonseparated-forces-h1-class}.
\item The unconditional split construction after sheafification realizes
$a(F_{p,s}|_{\mathcal C_a})$, not the original presheaf unless the
sheafification unit is an isomorphism.
\end{enumerate}
Therefore a later phrase such as ``choose a stack-banded realization prestack
with band $\Band$'' is an existence-and-choice hypothesis on the whole
prestack. It is not
an inference from the pair $(F_{p,s}|_{\mathcal C_a},\Band)$, and no gerbe or
obstruction class is attached to that pair before such realization data have
been supplied.
\end{proposition}

\begin{proof}
Items \textup{(i)} and \textup{(ii)} restate the two pieces of data separated in
\Cref{def:realization-prestack,prop:stackification-band-descends}. Item \textup{(iii)} is the exact criterion of
\Cref{prop:split-realization-prestack}. Item \textup{(iv)} combines the
restriction-dependent computation of \Cref{lem:split-isomorphism-presheaf} with
\Cref{cor:nonseparated-split-fails,cor:nonseparated-forces-h1-class,cor:band-does-not-force-realization}. Item
\textup{(v)} is \Cref{cor:sheafified-split-lift}. These are the only realization
conclusions established in \Cref{prop:no-band-only-realization-principle}. The
component-gerbe and obstruction constructions in the following sections are
therefore conditional on the chosen prestack input appearing in
\Cref{def:gluing-sensitive-realization-input,def:compatible-realization-stack}.
\end{proof}

\begin{remark}[Hypothesis ledger for later gerbe statements]\label{rem:realization-hypothesis-ledger}
Every later statement that mentions a stack-banded realization prestack with band $\Band$
uses the prestack itself as part of the hypotheses. Concretely, the input is a
fibred groupoid whose diagonal automorphism sheaves are banded by $\Band$
compatibly with restriction and conjugation, together with the induced
stack-level banding
$\beta^{\mathfrak L}_x:\Band|_U\cong
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x)$ for arbitrary stackified objects.
\Cref{def:realization-prestack}, item~\textup{(ii)} supplies this stack-level banding;
when the input begins with only a compatible prestack-level banding,
\Cref{prop:stackification-band-descends} proves that the same family is its
unique descent. The off-diagonal isomorphism sheaves, restrictions, and
compositions have already been supplied. The split
construction supplies such a prestack for the original presheaf only under the
separatedness condition of
\Cref{prop:split-realization-prestack}. The unconditional split construction of
\Cref{cor:sheafified-split-lift} realizes the sheafified presheaf instead, and
therefore changes the object functor unless the sheafification unit is already
an isomorphism.

In the non-separated case, \Cref{cor:nonseparated-split-fails} rules out the
split lift for every band, while
\Cref{cor:band-does-not-force-realization,prop:no-band-only-realization-principle}
show that on torsor-trivial slices a nonsplit replacement does not exist either.
Thus the later gerbe and obstruction classes are attached to a chosen
realization input, not to the bare pair $(F_{p,s},\Band)$.
\end{remark}

\begin{proposition}[Components of the stackification]\label{prop:stackification-components}
In the setting of \Cref{def:realization-prestack}, there is a canonical isomorphism of sheaves on $\mathcal C_a$:
\[
\pi_0(\mathfrak L_r)\cong a_{p,s}F_{p,s}|_{\mathcal C_a}.
\]
\end{proposition}

\begin{proof}
By stackification, the morphism presheaves of $\mathfrak P_r$ are sheafified and every object of $\mathfrak L_r$ is locally in the essential image of $\mathfrak P_r\to \mathfrak L_r$ \cite[Tag 02ZP]{StacksProject}. Passing to isomorphism classes therefore shows that $\pi_0(\mathfrak L_r)$ is the sheafification of $\pi_0^{\mathrm{pre}}(\mathfrak P_r)$. Since stacks in setoids are precisely sheaves of sets \cite[Tag 042Y]{StacksProject}, we obtain
\[
\pi_0(\mathfrak L_r)\cong a\bigl(\pi_0^{\mathrm{pre}}(\mathfrak P_r)\bigr).
\]
Using \Cref{def:realization-prestack}, item~\textup{(i)}, this becomes
\[
\pi_0(\mathfrak L_r)\cong a_{p,s}F_{p,s}|_{\mathcal C_a}.
\qedhere
\]
\end{proof}

\begin{definition}[Global conservativity at the terminal fibre]\label{def:global-conservativity}
In the setting of \Cref{def:realization-prestack}, let
\[
\iota_r:\mathfrak P_r\to \mathfrak L_r:=a(\mathfrak P_r)
\]
be the stackification morphism. We say that $\mathfrak P_r$ is \emph{globally conservative at $a$} if the induced functor on the terminal fibre
\[
(\iota_r)_a:(\mathfrak P_r)_a\to (\mathfrak L_r)_a
\]
is essentially surjective.
\end{definition}

\begin{definition}[Gluing-sensitive realization input]\label{def:gluing-sensitive-realization-input}
For an admitted reference $r$ at $p$, a \emph{gluing-sensitive realization
input} consists of a stack-banded realization prestack $\mathfrak P_r$ with
band $\Band$ in the sense of \Cref{def:realization-prestack}, including the
stackification banding
\[
\beta^{\mathfrak L}_x:\Band|_U\xrightarrow{\sim}
\underline{\mathrm{Aut}}_{\mathfrak L_r}(x)
\qquad (x\in\mathfrak L_r(U)),
\]
and global conservativity at $a=[r]_p$ in the sense of
\Cref{def:global-conservativity}. The branch-obstruction theory of
\Cref{sec:gerbe-obstruction,sec:homological-visibility} uses this input and
then adds the stated gerbe-adapted cover and finite-nerve hypotheses. This
terminology is only a shorthand for explicit hypotheses; it is not an assertion
that the input is canonically determined by $F_{p,s}$ or by $\Band$. In
particular, the off-diagonal data isolated in
\Cref{prop:realization-additional-data,prop:no-band-only-realization-principle}
are included in the prestack part of the input, while the automorphism sheaves
of arbitrary stackified objects are included through the displayed
stackification banding.
\end{definition}

\begin{proposition}[No new global objects at $a$]\label{prop:no-new-global-objects}
Assume global conservativity at $a$ in the sense of \Cref{def:global-conservativity}. Then
\[
M,p\Vdash \Sec_s(r)
\quad\Longleftrightarrow\quad
\mathfrak L_r(a)\neq\varnothing.
\]
\end{proposition}

\begin{proof}
By \Cref{def:existence-levels} and \Cref{def:realization-prestack}, item~\textup{(i)},
\[
M,p\Vdash \Sec_s(r)
\quad\Longleftrightarrow\quad
F_{p,s}(a)\neq\varnothing
\quad\Longleftrightarrow\quad
\pi_0^{\mathrm{pre}}(\mathfrak P_r)(a)\neq\varnothing
\quad\Longleftrightarrow\quad
(\mathfrak P_r)_a\neq\varnothing.
\]
If $(\mathfrak P_r)_a\neq\varnothing$, choose $y\in (\mathfrak P_r)_a$. Then $\iota_r(y)\in \mathfrak L_r(a)$, so $\mathfrak L_r(a)\neq\varnothing$. Conversely, if $x\in \mathfrak L_r(a)$, essential surjectivity of $(\iota_r)_a$ yields $y\in (\mathfrak P_r)_a$ with $\iota_r(y)\cong x$, so $(\mathfrak P_r)_a\neq\varnothing$.
\end{proof}

\begin{remark}[What remains conditional]\label{rem:gluing-sensitive-lifts}
\Cref{prop:split-realization-prestack} supplies a formal split lift for the original local-object presheaf only in the separated case, and \Cref{ex:nonseparated-split-lift,cor:nonseparated-split-fails} show explicitly why the same construction fails without separatedness. \Cref{cor:sheafified-split-lift} records the only unconditional split alternative: replace $F_{p,s}|_{\mathcal C_a}$ by its sheafification and realize that sheafified object functor instead. \Cref{lem:nonseparated-pairs-need-torsors} explains one necessary feature of a nonsplit realization of a non-separated presheaf: locally nonempty off-diagonal $\Band$-torsors with specified descent behavior. That necessary feature is not sufficient by itself; the torsors must fit into a coherent fibred groupoid whose isomorphism presheaves are sheaves. \Cref{cor:band-does-not-force-realization,prop:no-band-only-realization-principle} make the point sharper: on sites where those torsors are all trivial, a non-separated local-object presheaf admits no stack-banded realization prestack with the chosen band at all. For an arbitrary set-valued local object functor, admitting a stack-banded realization prestack is therefore an explicit existence-and-choice hypothesis, not a consequence of merely fixing the band. What remains further conditional in the later branch-obstruction theory is the choice of a \emph{gluing-sensitive} lift, namely a stack-banded realization prestack whose stackification banding identifies $\Band|_U$ with $\underline{\mathrm{Aut}}_{\mathfrak L_r}(x)$ for every stackified object $x$, whose stackification is globally conservative at $a$, and whose component gerbes satisfy the vanishing hypotheses needed to produce branch obstruction classes. The branch obstruction classes of \Cref{sec:gerbe-obstruction} are therefore attached to a chosen gluing-sensitive lift and its stack band, not to the bare set-valued functor $F_{p,s}$ alone.
\end{remark}
