\section{Conclusion}\label{sec:conclusion}

The present version is a focused note built around a narrower claim than the longer manuscript. It starts from a chosen local-object layer and a chosen gluing-sensitive stack-banded realization prestack as explicit data, rather than from a claim that an arbitrary set-valued local object functor carries a canonical obstruction theory. The split construction is a separatedness criterion, not a general existence theorem: its isomorphism presheaves are restriction-dependent, it is a prestack exactly in the separated case, and in torsor-trivial situations a non-separated local-object presheaf has no realization prestack with the chosen band at all. In non-torsor-trivial situations, the missing nonsplit realization still has to choose the nonzero off-diagonal torsor classes forced by locally equal but globally distinct pairs. The unconditional split construction is available only after replacing the local-object presheaf by its sheafification, which changes the object functor to the visible sheafified one. Thus a realization prestack is not shorthand for a band; it includes the off-diagonal isomorphism sheaves, their descent behavior, and the stackification banding that identifies $\Band|_U$ with the automorphism sheaf of every object of $\mathfrak L_r(U)$. Its first result is logical: compatible local sectionability and gluing-level absence are not definable by automorphism-invariant lower information-state formulas that omit the local-object predicates, so the local-object layer is genuinely necessary. Its second result is semantic: under the stated global-conservativity and branch-gerbe hypotheses, visible branches are exactly the connected components of the realization stack, and gluing failure is exactly the situation in which compatible local data exist but every component gerbe is non-neutral. Its third result is homological: once a branch obstruction class has been fixed, the intrinsic visible quotient is the cokernel of the homological evaluation map, so the visible part of an obstruction is precisely its $H_2$-detected part, while the blind remainder is exactly the pure $\operatorname{Ext}$-residual.

This is also the paper's main external point of contact with established work. In the unique-branch support-presheaf setting, the theory recovers the Abramsky--Brandenburger no-global-section picture and identifies the Car\`u-type cohomological blind cases with the pure $\operatorname{Ext}$-part of the obstruction class \cite{AbramskyBrandenburger2011,AbramskyMansfieldBarbosa2012,AbramskyBarbosaKishidaLalMansfield2015,Caru2017}. The worked $S^2$-type and $\mathbb{R}P^2$-type finite-nerve obstruction examples are realized by the good-cover construction of \Cref{cor:good-cover-semantic-models}; their realization prestacks, bands, branch classes, and global-conservativity hypotheses are checked explicitly before the visible and blind extremes are computed.

The standard ingredients should therefore be read as infrastructure rather than
as novelty claims. The matching-family description of sheafification, the
passage from a prestack to the sheaf of connected components, the classification
of banded gerbes by degree-$2$ classes, and the universal coefficient sequence
are used in their usual forms. The split-lift criterion is also a control
statement rather than a new construction principle: it marks the separated case
where the split prestack exists and explains why non-separated presheaves need
additional coherent realization data. The new content is the way these
ingredients are assembled for chosen gluing-sensitive realization inputs: the
lower-language undefinability of gluing predicates, the branch-wise gerbe
semantics of gluing-level absence, and the intrinsic quotient that separates
$H_2$-visible obstruction from pure-$\operatorname{Ext}$ blindness.

What has been omitted is deliberate. The measurable recursive-addressing, symbolic-dynamical cylinder, inverse-limit, multi-branch budget, refinement-dynamical, and complexity layers belong to a different narrative and would require their own separate development. The present paper is intended to stand or fall on the branch-gerbe semantics of gluing failure and on the identification of the intrinsic visible quotient and its pure-Ext blind residue.
