\section{Introduction}\label{sec:introduction}

Formal semantics is often presented as truth at a fully specified point of evaluation. That picture is exact when the relevant assignment or world has already been fixed. It is too coarse, however, when information is only partial. One may have ruled out some realizations without having selected a unique one; one may have an admissible reference before a global object has formed; or one may have compatible local data that fail to glue. In such situations the natural semantic notion is not pointwise truth but forcing at an information state.

\medskip
\noindent\textbf{Related work.}
Kripke and Beth semantics replace absolute truth by forcing at stages of information and impose persistence under refinement \cite{Beth1956,Kripke1965,TroelstraVanDalen1988,Fitting1969}. The topos-theoretic perspective on forcing, in which geometric morphisms and classifying topoi internalize the forcing relation, is developed in \cite{Tierney1976,Goldblatt1984,MacLaneMoerdijk1994}. Team and inquisitive semantics evaluate formulas on information states or teams; compositional semantics for imperfect information, dependence logic, and inquisitive semantics provide the foundational comparison points for the flat fragment used here \cite{Hodges1997,Vaananen2007,CiardelliRoelofsen2011,CiardelliGroenendijkRoelofsen2015,CiardelliIemhoffYang2020,AlbertGradel2022,CiardelliGrilletti2022,Galliani2021}. Sheaf-theoretic semantics separates local from global existence and treats gluing failure as a structural defect rather than as a mere lack of value \cite{MacLaneMoerdijk1994}; for semantic and linguistic uses of sheaf-theoretic gluing, reference, and ambiguity, see also \cite{AbramskySadrzadeh2014,LoSadrzadehMansfield2022}. On the geometric side, banded gerbes and non-abelian cohomology provide the standard language for obstruction classes \cite{Giraud1971}. Related local-to-global obstruction viewpoints also appear in the sheaf-theoretic analysis of contextuality and its cohomological and homotopical refinements \cite{AbramskyBrandenburger2011,AbramskyBarbosaKishidaLalMansfield2015,AbramskyMansfieldBarbosa2015,Caru2017,Caru2018,OkayRaussendorf2020,OkaySheinbaum2021,AbramskyBarbosa2021,Aasnaess2022}. The present paper combines these themes in order to address a single question: once an admitted reference carries a finite abelian gluing obstruction, what part of that obstruction is semantically visible without enlarging the state space, and what part survives only as a blind residue?

Logical contextuality work such as \cite{AbramskyBarbosa2021} analyzes the deductive strength of contextuality predicates. Our forcing-necessity results show that the gluing predicates form a strictly higher expressive layer even before one passes to contextuality-specific axioms.

Homotopical and classifying-space treatments \cite{OkaySheinbaum2021,OkayRaussendorf2020} study contextuality via classifying spaces and higher invariants. The present paper remains at the level of meet-semilattice sites but keeps track of the same topological data through the branch-dependent component gerbes and the $H_2$-visible quotient.

Work on cohomological completeness and blind spots \cite{Caru2018,Aasnaess2022} catalogues when obstructions detect contextuality and when they miss it. Our $\ker(\eta_A)$ description of the blind residue identifies precisely which part of a branch obstruction can be killed by character channels and which part remains invisible even though the obstruction is nontrivial.

On the flat first-order side, \cite{Galliani2021} and related team-semantics papers isolate the fragment of dependence logic that remains definable on flat teams. Our setting treats that fragment as the $\mathbb L_0$ layer and shows that adding local-object predicates strictly increases expressive power.

The paper is deliberately narrow. We retain only the lower-layer pointwise and information-state semantics, the local-object layer needed to express compatible local presence and gluing failure, and the sheaf/gerbe machinery used to interpret branch obstructions. The dynamic refinement, support, and complexity material of the longer manuscript is omitted here.

The relation to existing APAL-adjacent work is two-sided. On lower-layer formulas, the information-state semantics is just flat team-style semantics on nonempty sets of assignments. The new phenomena begin only when local-object predicates are added. At that point, the logic can express compatible local presence, global absence, and gluing failure, and these predicates are invisible to the lower flat fragment. This is why the paper is not merely a team-semantical definability note and also not merely a sheaf-theoretic repackaging of contextuality.

\subsection{Theorem-by-theorem comparison with prior work}

\Cref{def:logic-layer} packages Beth- and Kripke-style forcing semantics into a single layer interface, so its role is supportive rather than competitive with the classical sources \cite{Beth1956,Kripke1965,Fitting1969}. \Cref{def:conservative-extension} has no exact textbook antecedent in this state-projection form, and its increment is the explicit tuple of language embedding, forgetful model map, and state projection that makes later layer comparisons auditable. \Cref{prop:conservative-composition} is a routine closure property for that bespoke notion, included only to let the semantic layers compose cleanly. \Cref{cor:whole-chain-invariance} is the finite-iteration version of the same support lemma, specialized to the paper's three-layer architecture. \Cref{def:information-state} adapts Hodges-style teams and inquisitive information states to typed contexts and nonempty realization sets \cite{Hodges1997,CiardelliRoelofsen2011}. \Cref{def:refinement} combines Kripke/Beth information growth with context extension and team restriction \cite{Beth1956,Kripke1965}. \Cref{def:state-forcing} is the standard flat team clause rewritten as forcing at a state \cite{Hodges1997,Vaananen2007}. \Cref{prop:flat-team-reading} is therefore a classical input from flat team semantics, used only as a calibration identity \cite{Hodges1997}. \Cref{def:state-theory} imports the familiar notion of a theory satisfied at a stage or world into the state-based setting \cite{Fitting1969}. \Cref{def:decidedness} is likewise standard intuitionistic-style bookkeeping about whether a formula or its negation is forced \cite{TroelstraVanDalen1988}. \Cref{prop:forcing-persistence} is the usual monotonicity of forcing under refinement, here expressed with explicit context growth \cite{Kripke1965}. \Cref{cor:state-theory-monotonicity} is the immediate monotone-theory corollary of that classical persistence principle. \Cref{prop:singleton-conservativity} is the singleton-team flatness fact from team semantics \cite{Hodges1997}. \Cref{def:state-semantic-consequence} is ordinary semantic consequence lifted verbatim to the state semantics. \Cref{prop:lifted-pointwise-soundness} is the standard transfer of pointwise soundness to a flat team semantics, included here as a supporting bridge \cite{Hodges1997,Galliani2021}.

\Cref{def:refinement-stability} introduces the paper's non-splitting hypothesis for visible classes under refinement, and this is genuinely new terminology rather than inherited doctrine. \Cref{def:reference-classes} turns admitted references into state-indexed forcing-equivalence classes, which is a semantic quotient not isolated in the cited team-semantics or sheaf literature. \Cref{def:local-reference-frame} specializes the standard notion of a site to admitted reference classes equipped with meets so that semantic overlaps are computed by semilattice intersections \cite{MacLaneMoerdijk1994}. \Cref{def:local-object-functor} is just a presheaf under paper-specific notation \cite{MacLaneMoerdijk1994}. \Cref{def:existence-levels} contributes the paper's three-way separation between local, compatible-local, and global existence for one admitted reference. \Cref{prop:global-implies-compatible-local} is the identity-cover observation familiar from any site-theoretic local-to-global discussion. \Cref{thm:sheafification-characterization} is the first genuinely semantic increment in this section because it identifies the paper's predicate $\CompSec_s(r)$ exactly with nonemptiness of the sheafified fibre \cite{MacLaneMoerdijk1994,Johnstone2002}. \Cref{cor:sheaf-local-to-global} is the sheaf axiom specialized to a single reference fibre. \Cref{cor:sheafification-removes-glue} is the corresponding associated-sheaf corollary. \Cref{def:realization-prestack} is new terminology for a banded prestack whose presheaf of components is the chosen local object functor, even though prestacks and stackification themselves are classical \cite{StacksProject}. \Cref{prop:split-realization-prestack} adds an explicit split construction with prescribed $\pi_0^{\mathrm{pre}}=F_{p,s}|_{\mathcal C_a}$, rather than merely invoking abstract existence. \Cref{prop:stackification-components} is classical stack theory: it is the Stacks-project fact that stackification sheafifies connected components \cite[Tags 02ZP, 042Y]{StacksProject}. \Cref{def:global-conservativity} isolates terminal-fibre essential surjectivity as the exact extra condition needed to compare semantic global objects with stackified ones. \Cref{prop:no-new-global-objects} then identifies the paper's global-section predicate with nonemptiness of the terminal stack fibre under that new hypothesis. \Cref{def:gluing-sensitive} is a paper-specific package of the two hypotheses actually used later. \Cref{lem:component-local-nonempty} is a derived support lemma showing that each chosen branch substack is locally nonempty once gluing-sensitivity is assumed. \Cref{def:visible-value-classes} recasts associated-sheaf sections as semantic value classes of a reference. \Cref{thm:visible-value-components} is the payoff of that recasting: visible values coincide both with sheafified matching families and with global connected-component sections of the realization stack.

\Cref{def:typed-readout} is a new partial-readout construction that distinguishes undefinedness, nullity, uniqueness, and ambiguity by the cardinality of visible classes. \Cref{prop:address-before-value} extracts the precise semantic obligations of a non-null readout, namely admittedness plus compatible local sectionability. \Cref{cor:nonnull-readout-criterion} gives the converse singleton-branch criterion for when the readout returns a genuine value. \Cref{prop:typed-readout-persistence} is new and depends essentially on the refinement-stability assumption introduced earlier. \Cref{def:undef-total-null} packages the three coarse global statuses into uniform predicate names. \Cref{def:null-failure-modes} is likewise a new three-way taxonomy of local, compatibility, and gluing failure. \Cref{prop:null-modes-imply-null} records that each refined failure mode still lies inside the coarse null predicate. \Cref{prop:null-modes-exclusive} proves that the three refined modes really separate the null cases. \Cref{def:decomposition-decidability} isolates the exact decidability assumptions needed for an exhaustive trichotomy. \Cref{prop:null-trichotomy} is then a genuine paper-specific classification theorem rather than a standard textbook lemma. \Cref{thm:null-classification-under-refinement} propagates that classification to finer states when the requisite decidability is available there. \Cref{prop:finite-three-mode-model} adds a concrete finite meet-semilattice example exhibiting all three absence modes in one structure. \Cref{lem:state-automorphism-invariance} is the usual automorphism-invariance of first-order truth lifted to the flat state semantics \cite{Marker2002}. \Cref{thm:pointwise-irreducibility} is the main logical increment: it constructs four lower-layer-indistinguishable references realizing all positions in $\Sec_s(r)\Rightarrow\CompSec_s(r)\Rightarrow\LocSec_s(r)$ and thereby proves genuine undefinability in the flat fragment, in the spirit of the flatness boundary studied in team semantics \cite{Galliani2021}. \Cref{def:conditional-datum} finally packages, in a single auditable tuple, every external choice needed before the gerbe and homological layers begin.

\Cref{def:compatible-realization-stack} introduces branch-indexed component gerbes and cover-relative branch obstruction classes, which is more specific than the classical single-gerbe presentation of Giraud \cite{Giraud1971}. \Cref{thm:component-gerbe-decomposition} is the corresponding branchwise increment: each visible branch cuts out a full banded gerbe inside the realization stack, using the standard gerbe criterion only after the paper-specific visible-branch selection has been made \cite[Tag 06NY]{StacksProject}. \Cref{def:branch-constancy} is a new restriction-bijection condition ensuring that those branchwise gerbes exhaust the whole stack without overlap. \Cref{cor:branch-decomposition} is the derived coproduct decomposition under that hypothesis. \Cref{thm:cech-bridge-compatible-realizations} adapts the classical \v{C}ech $2$-cocycle construction for banded gerbes to the paper's branchwise, cover-relative setting and adds explicit refinement functoriality \cite{Giraud1971,StacksProject}. \Cref{thm:gerbe-null-semantics} is the semantic payoff: gluing-level absence is characterized by the coexistence of visible branches with non-neutrality of every component gerbe, which is strictly finer than the single-branch no-global-section picture of contextuality \cite{AbramskyBrandenburger2011,AbramskyMansfieldBarbosa2015}. \Cref{def:finite-nerve-presentation} is a new packaging of the finite cover, band trivialization, and obstruction class needed for homological evaluation. \Cref{def:intrinsic-visible-quotient} introduces the visible quotient $A_{\mathrm{vis}}(\mathcal U,A,\omega)=A/\mathrm{Im}(\mathrm{ev}_{\omega})$, and we do not know a prior source using this quotient as an invariant of contextual or gerbe obstruction data. \Cref{lem:finite-abelian-duality} is classical finite abelian duality \cite{DummitFoote2004}. \Cref{prop:visible-quotient-invariance} is new and shows that the visible quotient survives changes of band trivialization and all refinements that are surjective on $H_2$. \Cref{cor:subdivision-visible-quotient} is the subdivision-type specialization of that invariance theorem. \Cref{def:class-admissible-character} is a new character-selection device tied to the chosen obstruction class. \Cref{thm:intrinsic-visible-quotient} then identifies the visible quotient with the quotient detected by exactly those class-admissible characters. \Cref{cor:intrinsic-character-detection} is the dual separation corollary. \Cref{thm:intrinsic-pure-ext-initiality} adds a second new feature, namely an initiality theorem showing that the same quotient is the smallest finite coefficient quotient through which the $H_2$-visible part of the obstruction disappears.

\Cref{cor:h1-vanishing-class-trivializing} is the direct $H_1=0$ specialization of that initiality theorem. \Cref{thm:character-blind-obstructions} is the paper's sharpest comparison with the contextuality literature because it turns Car\`u-style blind examples into the structural criterion $\omega\in\ker(\eta_A)=\operatorname{Ext}^1(H_1,A)$ \cite{Caru2017,Caru2018}. \Cref{cor:h2-vanishing-blindness} is the corresponding automatic-blindness test when $H_2$ vanishes. \Cref{cor:h1-free-no-blindness} gives the opposite test when $H_1$ is free. \Cref{thm:unique-branch-contextuality-comparison} is a genuine bridge theorem: in the unique-branch support-presheaf specialization, the paper's predicates $\CompSec_s$, $\Sec_s$, and $\Null^{\mathrm{glue}}_s$ recover exactly the Abramsky--Brandenburger compatible-family and no-global-section picture without any lift hypothesis \cite{AbramskyBrandenburger2011}. \Cref{thm:unique-branch-banded-gerbe} is then a conditional specialization of the paper's own branch-gerbe and visible-quotient results to that Abramsky--Brandenburger setting, with the concrete gain that the remaining blind residue is identified exactly as the pure $\operatorname{Ext}$-part. The audit therefore supports the paper's overall positioning provided the hypothesis profile is stated accurately: \Cref{sec:null-decomposition} already assumes refinement stability, while \Cref{sec:gerbe-obstruction,sec:homological-visibility} correctly remain conditional on the external data gathered in \Cref{def:conditional-datum}.

\medskip
\noindent\textbf{Main result A: forcing necessity.}
After separating admissible reference, local sectionability, compatible local sectionability, and global sectionability, we show that compatible local existence is characterized by sheafification (\Cref{thm:sheafification-characterization}). The key logical separation theorem is that compatible local sectionability and gluing-level absence are not definable by any formula of the lower information-state forcing language without the local-object layer (\Cref{thm:pointwise-irreducibility}). This is the result that makes the enrichment logically necessary rather than merely convenient.

\medskip
\noindent\textbf{Main result B: conditional branch-gerbe semantics.}
Once an abelian band $\Band$ has been fixed on the slice site, every local object functor admits a canonical split realization prestack (\Cref{prop:split-realization-prestack}). However, \Cref{rem:split-not-conservative} shows that the split prestack is not globally conservative in general. The substantive hypothesis, encoded in the conditional data of \Cref{def:conditional-datum}, is that a gluing-sensitive realization prestack $\mathfrak P_r$ has been chosen together with a visible branch $v$ and a gerbe-adapted cover. Under that hypothesis, visible value classes correspond to the global sections of the sheaf of connected components of the realization stack (\Cref{thm:visible-value-components}); each visible branch determines a banded component gerbe (\Cref{thm:component-gerbe-decomposition}); and gluing failure occurs when visible branches exist but every component gerbe is non-neutral (\Cref{thm:gerbe-null-semantics}). Under branch constancy, the realization stack decomposes into those component gerbes (\Cref{cor:branch-decomposition}). No general existence theorem for gluing-sensitive lifts is proved here, so every statement in Sections \ref{sec:gerbe-obstruction} and \ref{sec:homological-visibility} is to be read relative to a chosen conditional datum.

\medskip
\noindent\textbf{Main result C: visible quotient and $\ker(\eta_A)$ blindness.}
For a chosen adapted cover with finite \v{C}ech nerve and finite abelian band, the homological evaluation map
\[
\mathrm{ev}_{\omega}:H_2(N(\mathcal U),\mathbb Z)\to A
\]
defines the visible quotient
\[
A_{\mathrm{vis}}(\mathcal U,A,\omega):=\operatorname{coker}(\mathrm{ev}_{\omega})=A/K_{\mathcal U,\omega},
\]
and \Cref{thm:intrinsic-visible-quotient} shows that this quotient is detected by the class-admissible character channels. \Cref{prop:visible-quotient-invariance} shows that it is independent of the chosen band trivialization, and invariant under refinements whose induced map on $H_2$ is surjective; subdivision-type refinements are therefore harmless. The blind part is characterized separately: \Cref{thm:character-blind-obstructions} identifies complete character-blindness with the case where the surviving obstruction class lies in $\ker(\eta_A)$, i.e.\ in the pure $\operatorname{Ext}$-subgroup of the universal coefficient sequence. In the support-presheaf specialization, \Cref{thm:unique-branch-contextuality-comparison} recovers the Abramsky--Brandenburger no-global-section picture at the presheaf level without any lift hypothesis, while \Cref{thm:unique-branch-banded-gerbe} turns Car\`u-type blind cases into a conditional theorem about $\ker(\eta_A)$-residuals once a gluing-sensitive lift and a finite obstruction presentation have been fixed.

\medskip
\noindent\textbf{What is new.}
Sheafification, stackification, banded gerbes, \v{C}ech $2$-cocycles, and the universal coefficient sequence are standard inputs and are cited directly. The new contributions are: (i) a forcing-necessity theorem showing that the gluing predicates are not definable in the flat fragment; (ii) a conditional branch-sensitive gerbe semantics that is explicitly formulated relative to the data of \Cref{def:conditional-datum}; and (iii) the identification of the visible quotient $A_{\mathrm{vis}}(\mathcal U,A,\omega)$ as the $H_2$-detected part of the obstruction while the blind residue is exactly the kernel of the universal coefficient map $\eta_A$. These points isolate the genuinely new semantics and keep them separate from the standard geometric machinery.

The main notation is as follows. The lower semantic layers are denoted by $\mathbb L_0$ and $\mathbb L_1$. For an admitted reference $r$, the realization prestack is $\mathfrak P_r$, its stackification is $\mathfrak L_r$, the component gerbe of a visible branch $v$ is $\mathfrak G_v$, and the cover-relative branch obstruction class on an adapted cover $\mathcal U$ is $\omega_v(\mathcal U)$. Given a conditional datum $(r,p,\mathfrak P_r,\Band,v,\mathcal U,A,\omega)$, the hidden subgroup is $K_{\mathcal U,\omega}:=\mathrm{Im}(\mathrm{ev}_\omega)\leq A$ and the corresponding visible quotient is $A_{\mathrm{vis}}(\mathcal U,A,\omega):=A/K_{\mathcal U,\omega}$.

The assumption profile is as follows. \Cref{sec:preliminaries,sec:information-states} are unconditional. \Cref{sec:null-decomposition} works under the standing refinement-stability assumption of \Cref{def:refinement-stability}. \Cref{sec:gerbe-obstruction} begins only after a conditional datum in the sense of \Cref{def:conditional-datum} has been fixed, and defines branch obstructions relative to the chosen gerbe-adapted cover. \Cref{sec:homological-visibility} adds finite abelian band trivializations and finite nerve presentations, and asserts cover-independence only for refinements whose induced maps on $H_2$ are explicitly controlled.

The paper is organized accordingly. \Cref{sec:preliminaries} and \Cref{sec:information-states} isolate the lower-layer semantics and its conservative-extension behavior. \Cref{sec:null-decomposition} develops local objects, sheafification, visible value classes, the forcing-necessity theorem, and the conditional data that drive the later sections. \Cref{sec:gerbe-obstruction} gives the branch-gerbe semantics of gluing failure relative to those data. \Cref{sec:homological-visibility} develops the visible quotient and its invariance properties, the $\ker(\eta_A)$ blindness theorem, three worked examples (including one end-to-end example verifying the conditional hypotheses from the local reference frame through the visible quotient), and the unique-branch comparison with contextuality, now separated into an unconditional support-presheaf part and a conditional banded-gerbe part together with a dictionary of what extra data each layer requires.
