\documentclass[11pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\signature{Wenlin Zhang \\ National University of Singapore}
\address{Wenlin Zhang \\ National University of Singapore \\ Singapore \\ \texttt{e1327962@u.nus.edu}}

\begin{document}

\begin{letter}{Editorial Board \\ SIAM Journal on Applied Dynamical Systems}

\opening{Dear Editors,}

We submit the manuscript ``Folded Histograms of Irrational Rotations:
Sampling Certificates and Structural Mismatch to the Parry Measure''
for consideration by SIAM Journal on Applied Dynamical Systems.

The paper addresses a concrete question in symbolic dynamics: when an
orbit of an irrational rotation is coarse-grained into a shift of
finite type by Fibonacci weighting and Zeckendorf normalization, how
much of the resulting histogram mismatch with the Parry equilibrium
is due to finite sampling, and how much is structural? We give a
complete answer for the golden-mean benchmark.

The main contributions are:

\begin{enumerate}
\item A two-stage audit that cleanly separates finite-sample error
(controlled by star discrepancy) from the structural divergence
that persists in the infinite-sample limit.

\item Closed-form Renyi divergence envelopes and an explicit optimal
injective relabeling of the realized Sturmian law into the
golden-mean shift, valid for all sufficiently large block lengths.

\item A universal lower bound showing that the injective placement
barrier extends to arbitrary primitive shifts of finite type,
not just the golden-mean shift.

\item A proof that the Fibonacci weighting is the unique choice
satisfying a natural contiguous numeration axiom, placing the
fold on a canonical footing.

\item Transfer of the full audit to nonlinear circle diffeomorphisms
via Denjoy's theorem, verified numerically on the standard
circle map.
\end{enumerate}

The paper sits at the intersection of symbolic dynamics, ergodic
theory, and information-theoretic diagnostics. We believe SIADS is a
natural venue because the results link classical dynamical systems
theory (Sturmian sequences, Parry measures, topological conjugacy)
with quantitative statistical tools, and the benchmark is designed to
be directly usable in applications where coarse-grained orbit data
must be compared against equilibrium baselines.

All numerical results are reproducible from the Python scripts
included in the supplementary materials. The generated tables are
checked into the supplement as-is, and the audit-grid conventions
are documented in detail.

Neither the manuscript nor any substantial part of it is under
consideration elsewhere. Both authors have approved the submission.

\closing{Sincerely,}

\end{letter}
\end{document}
