\begin{theorem}\label{thm:fold-groupoid-chi-rigidity}
With the notation of Proposition \ref{prop:fold-groupoid-z2x2-sector-by-orientation}, consider the Wedderburn decomposition
\[
\CC[\mathcal R_m]\cong \bigoplus_{x\in X_m} M_{d_m(x)}(\CC).
\]
For each \(x\in X_m\), let \(P(c_x),P(\iota_x)\in M_{d_m(x)}(\CC)\) be the permutation matrices representing the scan element and the reversal involution on the \(x\)-block. Assume that \(\Phi\in\mathrm{Aut}^{\ast}(\CC[\mathcal R_m])\) is compatible with these symmetries in the following sense: there are a permutation \(\sigma:X_m\to X_m\) and unitaries
\[
U_x\in M_{d_m(\sigma(x))}(\CC)
\]
such that, for every \(x\in X_m\),
\[
\Phi\!\left(M_{d_m(x)}(\CC)\right)=M_{d_m(\sigma(x))}(\CC),
\]
\[
\Phi\!\bigl(P(c_x)\bigr)=U_x\,P(c_{\sigma(x)})\,U_x^{-1},
\qquad
\Phi\!\bigl(P(\iota_x)\bigr)=U_x\,P(\iota_{\sigma(x)})\,U_x^{-1}.
\]
Then
\[
\chi(\sigma(x))=\chi(x)
\qquad\text{for all }x\in X_m,
\]
where
\[
\chi(x)=\bigl(\operatorname{sgn}(c_x),\operatorname{sgn}(\iota_x)\bigr)\in\{\pm1\}^2.
\]
\end{theorem}

\begin{proof}
Since \(\Phi\) restricts to a \(\ast\)-isomorphism
\[
M_{d_m(x)}(\CC)\longrightarrow M_{d_m(\sigma(x))}(\CC),
\]
the two matrix algebras must have the same size, so \(d_m(\sigma(x))=d_m(x)\). By the unitary form of the Skolem--Noether theorem, there exists a unitary \(V_x\in M_{d_m(\sigma(x))}(\CC)\) such that
\[
\Phi(A)=V_xAV_x^{-1}
\qquad
\text{for all }A\in M_{d_m(x)}(\CC).
\]
Hence
\[
V_xP(c_x)V_x^{-1}=U_xP(c_{\sigma(x)})U_x^{-1},
\]
so the matrices \(P(c_x)\) and \(P(c_{\sigma(x)})\) are conjugate in \(M_{d_m(\sigma(x))}(\CC)\). The same argument applies to \(P(\iota_x)\) and \(P(\iota_{\sigma(x)})\).

For any permutation \(\pi\), the determinant of its permutation matrix is \(\operatorname{sgn}(\pi)\). Taking determinants in the conjugacy relations therefore gives
\[
\operatorname{sgn}(c_x)=\operatorname{sgn}(c_{\sigma(x)}),
\qquad
\operatorname{sgn}(\iota_x)=\operatorname{sgn}(\iota_{\sigma(x)}).
\]
This is precisely the claim that \(\chi(\sigma(x))=\chi(x)\).
\end{proof}

\begin{proposition}\label{prop:fold-groupoid-aut-chi-semidirect-classification}
For \((\alpha,\beta)\in\{\pm1\}^2\), let
\[
\CC[\mathcal{R}_m]^{\alpha,\beta}
\]
be the sector ideal of Proposition \ref{prop:fold-groupoid-z2x2-sector-by-orientation}. Define
\[
\mathrm{Aut}^{\ast}_{\chi}(\CC[\mathcal R_m])
:=
\left\{\Phi\in\mathrm{Aut}^{\ast}(\CC[\mathcal R_m]):
\Phi\bigl(\CC[\mathcal{R}_m]^{\alpha,\beta}\bigr)=\CC[\mathcal{R}_m]^{\alpha,\beta}
\text{ for all }(\alpha,\beta)\right\},
\]
and let
\[
\Pi_{\chi}
:=
\left\{\sigma\in\mathrm{Sym}(X_m):d_m(\sigma(x))=d_m(x),\ \chi(\sigma(x))=\chi(x)\text{ for all }x\in X_m\right\}.
\]
Then there is a natural group isomorphism
\[
\mathrm{Aut}^{\ast}_{\chi}(\CC[\mathcal R_m])
\cong
\Bigl(\prod_{x\in X_m}\mathrm{PU}\bigl(d_m(x)\bigr)\Bigr)\rtimes \Pi_{\chi}.
\]
\end{proposition}

\begin{proof}
By Theorem \ref{thm:fold-groupoid-aut-semidirect-pi0}, every \(\ast\)-automorphism of \(\CC[\mathcal R_m]\) is given by blockwise projective-unitary conjugation together with a permutation of blocks of the same size.

If \(\Phi\in\mathrm{Aut}^{\ast}_{\chi}(\CC[\mathcal R_m])\), then the induced block permutation \(\sigma\) must preserve matrix size and keep each sector ideal invariant. Since the sector ideals are exactly the sums of blocks with a fixed fingerprint \(\chi(x)\), it follows that \(\chi(\sigma(x))=\chi(x)\) for every \(x\). Thus \(\sigma\in\Pi_\chi\).

Conversely, any \(\sigma\in\Pi_\chi\), together with arbitrary blockwise inner automorphisms, preserves each sector ideal because both the matrix size and the fingerprint are preserved on every block. This yields a homomorphism
\[
\Bigl(\prod_{x\in X_m}\mathrm{PU}\bigl(d_m(x)\bigr)\Bigr)\rtimes \Pi_{\chi}
\longrightarrow
\mathrm{Aut}^{\ast}_{\chi}(\CC[\mathcal R_m]),
\]
and the previous paragraph shows that it is surjective. Injectivity is immediate from the uniqueness of the induced block permutation and of each inner automorphism modulo scalars. This proves the stated semidirect-product description.
\end{proof}

\begin{remark}
The proposition is the standard structure theorem for automorphisms of a finite direct sum of matrix algebras, with the extra requirement that the scan-reversal sector decomposition be fixed.
\end{remark}

\endinput
