\section{Introduction}\label{sec:introduction}

Finite-state symbolic models impose rigid constraints on counting
functions. In a deterministic automaton, the number of accepted words
of length \(m\) is governed by powers of a finite matrix. In a sofic
shift, the number of admissible words of a given length is likewise
controlled by a finite-state presentation. These mechanisms force
periodic density laws or exponential-polynomial growth, and that is
often enough to rule out arithmetic languages whose size is governed by
the prime number theorem.

The purpose of this note is modest: we place three elementary
obstructions of this kind side by side. The first concerns approximation
of the binary prime language by a fixed complete deterministic
automaton. The second concerns exact regular realizations and eventual
sofic slice realizations inside the canonical Zeckendorf syntax. The
third concerns the analytic behaviour of Euler products and finite
determinant models attached to finite-state data.

Throughout, \(f(m)\asymp g(m)\) means that \(c_1g(m)\le f(m)\le c_2g(m)\)
for some positive constants \(c_1,c_2\) and all sufficiently large
\(m\). We write \(F_0=0\), \(F_1=1\) for the Fibonacci numbers,
\(\varphi=(1+\sqrt5)/2\), and for a Zeckendorf word
\(w=x_1\cdots x_m\) we set
\[
  \val(w):=\sum_{k=1}^m x_kF_{k+1}.
\]

\subsection{Main results}

Our first theorem is a density dichotomy for deterministic automata.

\begin{theorem}\label{thm:intro-density}
Let \(\mathcal A\) be a complete deterministic finite automaton over
\(\Sigma=\{0,1\}\), and let
\[
  a_m(\mathcal A)
  :=\#\{w\in\Sigma^m:\mathcal A\text{ accepts }w\}.
\]
Then there exist an integer \(p\ge 1\), non-negative constants
\(c_0,\dots,c_{p-1}\), and \(\theta\in(0,1)\) such that
\[
  \frac{a_m(\mathcal A)}{2^m}=c_{m\bmod p}+O(\theta^m)
  \qquad (m\to\infty).
\]
Consequently exactly one of the following holds:
\begin{enumerate}[label=\textup{(\roman*)}]
  \item \(a_m(\mathcal A)/2^m\) decays exponentially.
  \item Along some residue class modulo \(p\), the ratio
  \(a_m(\mathcal A)/2^m\) is bounded below by a positive constant.
\end{enumerate}
In particular, no deterministic finite automaton can satisfy
\[
  a_m(\mathcal A)/2^m\asymp m^{-\alpha}
\]
for any \(\alpha>0\).
\leanverified{paper_prime_languages_intro_density}
\end{theorem}

When this is applied to the binary prime language
\(\mathcal P_m^{(2)}\), whose size is \(\asymp 2^m/m\) by the prime
number theorem \cite{HardyWright2008}, one obtains a sharp
incompatibility with finite-state recognition.

\begin{corollary}[Symmetric-difference lower bound for binary primes]
  \label{cor:intro-binary-primes}
Let \(L(\mathcal A)\subseteq\{0,1\}^\ast\) be recognised by a fixed
deterministic finite automaton. Then there exists \(c>0\) such that
\[
  \abs{L(\mathcal A)\cap\{0,1\}^m\triangle\mathcal P_m^{(2)}}
  \ge c\,\frac{2^m}{m}
\]
for all sufficiently large \(m\). If \(L(\mathcal A)\) falls in the
positive-density alternative of \Cref{thm:intro-density}, then along an
infinite arithmetic progression of lengths one has the stronger bound
\[
  \abs{L(\mathcal A)\cap\{0,1\}^m\triangle\mathcal P_m^{(2)}}
  \gg 2^m.
\]
\leanverified{paper_prime_languages_intro_binary_primes}
\end{corollary}

\begin{corollary}[No simultaneous recall and precision for fixed DFA]
  \label{cor:intro-binary-precision}
  Let
  \[
    \mathrm{Rec}_m
    :=\frac{\abs{L(\mathcal A)\cap\mathcal P_m^{(2)}}}
            {\abs{\mathcal P_m^{(2)}}},
    \qquad
    \mathrm{Prec}_m
    :=\frac{\abs{L(\mathcal A)\cap\mathcal P_m^{(2)}}}
            {\abs{L(\mathcal A)\cap\{0,1\}^m}}
  \]
  whenever \(\abs{L(\mathcal A)\cap\{0,1\}^m}>0\). Then every fixed DFA
  satisfies one of the following:
  \begin{enumerate}[label=\textup{(\roman*)}]
    \item \(\mathrm{Rec}_m\to 0\) exponentially fast.
    \item \(\mathrm{Prec}_m\to 0\) along an infinite arithmetic
    progression of lengths.
  \end{enumerate}
  In particular, no fixed DFA has both recall and precision bounded
  away from zero for infinitely many lengths.
\leanverified{paper_prime_languages_intro_binary_precision}
\end{corollary}

Our second theorem concerns the Zeckendorf language
\[
  \mathcal Z
  :=\{x_1\cdots x_m\in\{0,1\}^\ast:
      x_ix_{i+1}\neq 11,\ x_m=1\}.
\]

\begin{theorem}\label{thm:intro-zeckendorf}
Let \(L\subseteq\mathcal Z\) be regular. After deleting states that are
not reachable from the initial state or cannot reach an accepting
state, let \(\rho\) be the maximal spectral radius of a strongly
connected component in a DFA recognizing \(L\). Define
\[
  N_L(T):=\#\{w\in L:\val(w)\le T\},
\]
where \(\val(w)\) is the Zeckendorf value of \(w\). Then
\[
  N_L(T)=T^{\alpha+o(1)},
  \qquad
  \alpha=
  \begin{cases}
    0, & \rho\le 1,\\[0.3em]
    \dfrac{\log \rho}{\log \varphi}, & \rho>1.
  \end{cases}
  \qquad (T\to\infty).
\]
In particular, the language of Zeckendorf representations of prime
numbers is not regular. Combined with the residue-class asymptotics for
sofic word counts proved later
(\Cref{prop:sofic-counts-exponential-polynomial,lem:residue-class-asymptotics}),
the same comparison also shows that there is no sofic shift \(X\) over
\(\{0,1\}\) with \(\mathcal B_m(X)=\mathcal P_m^{(\mathrm Z)}\) for
all sufficiently large \(m\), where \(\mathcal B_m(X)\) denotes the
admissible words of length \(m\) in \(X\).
\leanverified{paper_prime_languages_intro_zeckendorf_package}
\end{theorem}

Our final theorem records two analytic obstructions.

\begin{theorem}\label{thm:intro-analytic}
The following statements hold.
\begin{enumerate}[label=\textup{(\roman*)}]
  \item If \((b_n)_{n\ge 1}\subseteq\ZZ_{\ge 0}\), if \(b_p>0\) for
  infinitely many primes \(p\), and if the Euler product
  \[
    Z_b(z):=\prod_{n\ge 1}(1-z^n)^{-b_n}
  \]
  converges to a holomorphic function on \(\abs{z}<1\), then it has the
  unit circle as a natural boundary.

  \item For every finite matrix \(M\), the function
  \[
    F_M(s):=\det(I-e^{-s}M)^{-1}
  \]
  is \(2\pi i\)-periodic. Consequently, if \(F(s)=\sum_{n\ge 1}a_n n^{-s}\)
  is a Dirichlet series with at least one coefficient \(a_n\neq 0\) for
  some \(n\ge 2\), then \(F\) cannot agree with \(F_M\) on any non-empty
  open subset of a right half-plane. In particular, \(F_M\) cannot agree
  with the Riemann \(\zeta\)-function on any non-empty open subset of
  \(\Re(s)>1\).
\end{enumerate}
\leanverified{paper_prime_languages_intro_analytic}
\end{theorem}

\subsection{Previous work and what is new}

The impossibility of recognizing prime numbers by finite automata in
classical settings goes back to Hartmanis and Shank
\cite{HartmanisShank1968}. In the broader framework of abstract
numeration systems, Rigo proved that for any numeration system \(S\),
the set of primes is never \(S\)-recognizable
\cite{Rigo2000,RigoMaes2002}. Accordingly,
\Cref{cor:zeckendorf-prime-language-not-regular} is a
Zeckendorf-specialized derivation of a known non-recognizability fact,
included here because the transfer-matrix argument directly supports the
sofic obstruction that follows it.

The general background on automatic and recognizable sequences is
classical; see Cobham's foundational papers
\cite{Cobham1969,Cobham1972} and the monograph
\cite{AlloucheShallit2003}. The growth of word-counting functions for
regular and more general \(S\)-recognizable languages is also classical;
see Salomaa and Soittola \cite{SalomaaSoittola1978}, Berstel and
Reutenauer \cite{BerstelReutenauer2011}, and Charlier and Rampersad
\cite{CharlierRampersad2011} for Zeckendorf-adjacent treatments.
\Cref{thm:zeckendorf-regular-powerlaw} is a Zeckendorf-specialized
instance of those standard growth results, included as a self-contained
derivation via the Perron--Frobenius theorem.

The interaction between primes and digital constraints has a much deeper
literature. In particular, the work of Mauduit and Rivat
\cite{MauduitRivat2010} and of Drmota, Mauduit and Rivat
\cite{DrmotaMauduitRivat2009} shows that prime inputs already display
highly nontrivial equidistribution phenomena inside automatic and
digit-restricted settings. The approximation of non-regular languages by
fixed finite automata has also been studied directly; see Sin'ya
\cite{Sinya2015} for asymptotic approximation results in that spirit.

For Zeckendorf and more general linear or Parry numeration systems, the
regularity of admissible expansions is standard; see
\cite{Frougny1992,BruyereHansel1997}. Our treatment uses only this
regular syntax together with the transfer-matrix viewpoint on word
counts, familiar from symbolic dynamics and analytic combinatorics
\cite{LindMarcus1995,Kitchens1998,FlajoletSedgewick2009}. On the
analytic side, the natural-boundary statement belongs to the classical
circle of ideas initiated by P\'olya, Carlson, and Estermann
\cite{Polya1923,Carlson1921,Estermann1928}, while the determinant
formula for shifts of finite type goes back to Artin and Mazur
\cite{ArtinMazur1965}; see also \cite{LindMarcus1995}.

The genuinely new contributions of this note are two. First, the
fixed-DFA density dichotomy (\Cref{thm:dfa-density-dichotomy}) and its
precision/recall corollary (\Cref{sec:automata}) provide an explicit
quantitative obstruction for approximating the binary prime language by
a \emph{single fixed} DFA, in a form that does not appear in the
general Cobham or Rigo framework. Second, the sofic-slice obstruction
(\Cref{thm:zeckendorf-primes-not-sofic}) excludes even eventual sofic
realization of the prime Zeckendorf length slices, by comparing the
prime-number-theorem growth rate \(\varphi^m/m\) with the
exponential-polynomial form forced by any sofic shift; this combination
appears to be new in this exact form. The analytic observations in
\Cref{sec:analytic} are included as a complementary perspective, with
the understanding that they belong to a classical tradition.

\subsection{Organisation}

\Cref{sec:automata} proves the density dichotomy for deterministic
automata and applies it to binary primes. \Cref{sec:zeckendorf}
establishes the growth law for regular Zeckendorf sublanguages and the
resulting regular and sofic obstructions. \Cref{sec:analytic} contains
the two analytic obstructions at the level of Euler products and
determinants.
