\section{Introduction}\label{sec:introduction}

Formal semantics is often presented as truth at a fully specified point of evaluation. That picture is exact when the relevant assignment or world has already been fixed. It is too coarse, however, when information is only partial. One may have ruled out some realizations without having selected a unique one; one may have an admissible reference before a global object has formed; or one may have compatible local data that fail to glue. In such situations the natural semantic notion is not pointwise truth but forcing at an information state. The present note restricts attention to the narrow setting in which this semantic layer is studied \emph{relative} to a chosen gluing-sensitive realization prestack with fixed abelian band. No attempt is made to reconstruct such data from arbitrary empirical models; all results are conditional on the explicit hypotheses described below.

\medskip
\noindent\textbf{Related work.}
Several familiar traditions isolate parts of this picture.

\emph{Team semantics and dependence logic.}
Hodges introduced compositional semantics for a language of imperfect information, evaluating formulas on sets of assignments rather than on single valuations \cite{Hodges1997}. V\"a\"an\"anen developed this into dependence logic \cite{Vaananen2007}, and the resulting family of team-based and inquisitive logics now forms an active area; see \cite{CiardelliGroenendijsenRoelofsen2019} for the inquisitive-semantics perspective and \cite{AlbertGradel2022,CiardelliGrilletti2022} for recent APAL contributions. Galliani isolated upwards closed dependencies and highlighted how forcing across information states interacts with closure properties of teams \cite{Galliani2013}. Abramsky, Puljuj\"arvi, and V\"a\"an\"anen related team-semantical dependence notions directly to quantum-mechanical independences \cite{AbramskyPuljujarviVaananen2021}, underscoring why comparisons with contextuality arise even in low expressive fragments. Kripke and Beth semantics replace absolute truth by forcing at stages of information and impose persistence under refinement \cite{Beth1956,Kripke1965,TroelstraVanDalen1988}. The lower-layer forcing in the present paper coincides with flat team semantics on nonempty teams; what is new begins only after the local-object predicates are added, and every separation result stated below is valid only in the constant-free equality-only fragment on singleton parameter states spelled out in \Cref{def:lower-layer-syntax,def:information-state-definability}.

\emph{Sheaf-theoretic contextuality and cohomological obstructions.}
Sheaf-theoretic semantics separates local from global existence and treats gluing failure as a structural defect rather than as a mere lack of value \cite{MacLaneMoerdijk1994}. Abramsky and Brandenburger gave the foundational sheaf-theoretic formulation of non-locality and contextuality \cite{AbramskyBrandenburger2011}. The cohomological refinement of that picture, using \v{C}ech cohomology of the support presheaf to detect obstructions to global sections, was developed by Abramsky, Mansfield, and Barbosa \cite{AbramskyMansfield2012} and extended to a logical/topological setting by Abramsky et al.\ \cite{AbramskyEtAl2015}. The compatible-family/global-section distinction encoded below by $\CompSec$ and $\Sec$ is also central in the extendability literature: Mansfield and Barbosa show how that distinction is transported from Kochen--Specker models to Bell models \cite{MansfieldBarbosa2013}. Geometric renditions such as bundle diagrams \cite{BeerOsborne2018} and the modern review of Kochen--Specker contextuality by Budroni, Cabello, G\"uhne, Kleinmann, and Larsson \cite{BudroniCabelloGuhneKleinmannLarsson2022} supply the standard local-to-global language against which our semantics should be read.
Car\`u \cite{Caru2017} showed that the cohomological obstruction can be incomplete, a phenomenon we identify in \S\ref{sec:homological-visibility} as the pure $\operatorname{Ext}$-case. Car\`u's later work \cite{Caru2018} pursues complete invariants for contextuality, and additional topological or homotopical approaches include the bundle and surface proofs of Okay, Roberts, Bartlett, and Raussendorf \cite{OkayRobertsBartlettRaussendorf2017}, the cohomological quantum-computation framework of Raussendorf \cite{Raussendorf2019}, the homotopical comparison of Okay and Raussendorf \cite{OkayRaussendorf2020}, and the algebraic/comparison results of Aasn{\ae}ss \cite{Aasnaess2020,Aasnaess2022}. The present paper does not claim to improve on the Abramsky--Brandenburger obstruction theory itself; our contribution is a conditional comparison (\Cref{thm:unique-branch-contextuality-comparison}) showing that, when a compatible realization prestack exists, the formalism developed here recovers the no-global-section picture and refines the blind residue. GHZ and related empirical models enter here only as presheaf-level benchmarks for the compatible-family/global-section distinction; no canonical banded gerbe or intrinsic \v{C}ech $H^2$-class is extracted from them in the present paper.

\emph{Gerbes and obstruction language.}
The gerbe-theoretic and cohomological machinery used in \S\ref{sec:gerbe-obstruction} is standard \cite{Giraud1971,Johnstone2002,Bredon1997,StacksProject}. Our contribution is not in the categorical formalism but in its semantic use: the branch-gerbe decomposition of gluing failure relative to a chosen realization prestack and the identification of the presentation-relative visible quotient.

The present paper combines these themes in order to address a single question: once an admitted reference carries a finite abelian gluing obstruction relative to a chosen realization prestack, what part of that obstruction is semantically visible without enlarging the state space, and what part survives only as a blind residue?

The paper is deliberately narrow and is best read as a focused note. Every gerbe- and presentation-relative-visible-quotient statement is conditioned on the prior choice of a gluing-sensitive realization prestack with fixed abelian band; we do not claim that arbitrary set-valued functors canonically determine such data. We retain only the lower-layer pointwise and information-state semantics, the local-object layer needed to express compatible local presence and gluing failure, and the sheaf/gerbe machinery used to interpret branch obstructions. We do not attempt here to derive the site-theoretic layer from an underlying measurable recursive-addressing hierarchy, and we do not claim that an arbitrary set-valued local object functor canonically determines a branch obstruction class.

Against that background, the novelty of the present paper is two-sided. On lower-layer formulas, the information-state semantics is just flat team semantics on nonempty sets of assignments. The new phenomena begin only when local-object predicates are added. At that point, the logic can express compatible local presence, global absence, and gluing failure, and these predicates are invisible to the lower flat fragment. This is why the paper is not merely a team-semantical definability note and also not merely a sheaf-theoretic repackaging of contextuality.

\medskip
\noindent\textbf{Main result A: forcing necessity.}
After separating admissible reference, local sectionability, compatible local sectionability, and global sectionability, we show that compatible local existence is characterized by sheafification (\Cref{thm:sheafification-characterization}). The key logical separation theorem is that, in the constant-free equality-only lower reduct evaluated on singleton parameter states of parameter-indexed local-object expansions, compatible local sectionability and gluing-level absence are not definable by any formula of the lower information-state forcing language without the local-object layer (\Cref{thm:pointwise-irreducibility}). This is the only genuinely new theorem: it shows that the enrichment is logically necessary \emph{within that narrow comparison class}. No wider forcing-necessity claim is made for richer signatures, multi-parameter states, or higher expressive fragments.

\medskip
\noindent\textbf{Main result B: visible branches and gluing failure.}
After fixing an admitted reference and working with a chosen realization prestack with abelian band on the slice site, one can compare that lift with the canonical split realization prestack of \Cref{prop:split-realization-prestack} (which requires the local object functor to be separated). The substantive hypotheses are semantic: we work with gluing-sensitive lifts whose stackifications recover the sheafification of isomorphism classes (\Cref{prop:stackification-components}), identify visible value classes with the global connected components of the realization stack (\Cref{thm:visible-value-components}), and are globally conservative at the terminal fibre when stack-global objects are compared back to the original local object functor. In that setting, each visible branch determines a banded component gerbe (\Cref{thm:component-gerbe-decomposition}), and gluing failure occurs exactly when visible branches exist but every component gerbe is non-neutral (\Cref{thm:gerbe-null-semantics}). Under branch constancy, the realization stack decomposes into those component gerbes (\Cref{cor:branch-decomposition}). These statements are semantic syntheses of well-known gerbe facts and are presented accordingly as propositions and corollaries rather than as new gerbe theory.

\medskip
\noindent\textbf{Main result C: presentation-relative visible quotient and pure-Ext blindness.}
For a finite abelian branch obstruction attached to such a chosen visible branch on a finite \v{C}ech nerve, the homological evaluation map
\[
\mathrm{ev}_{\omega}:H_2(N(\mathcal U),\mathbb Z)\to A
\]
defines the presentation-relative visible quotient as
\[
A_{\mathrm{vis}}^{\omega}:=\operatorname{coker}(\mathrm{ev}_{\omega}),
\]
and \Cref{thm:intrinsic-visible-quotient} proves that this quotient is exactly the one detected by the class-admissible character channels. All invariance claims are made with respect to the explicitly described presentation changes: equivalence of chosen lifts, change of band trivialization, and common refinements satisfying an explicit $H_2$-surjectivity hypothesis (\Cref{prop:visible-quotient-trivialization-invariance,prop:visible-quotient-lift-invariance,prop:visible-quotient-refinement-invariance}). The blind part is equally sharp: \Cref{thm:character-blind-obstructions} identifies complete character-blindness with the case where the surviving obstruction class lies entirely in the pure $\operatorname{Ext}$-summand of the universal coefficient sequence. In the unique-branch support-presheaf setting, \Cref{thm:unique-branch-contextuality-comparison} is a conditional comparison statement: when a compatible realization prestack exists and is globally conservative, it recovers the Abramsky--Brandenburger no-global-section picture and turns Car\`u-type blind cases into a theorem about pure $\operatorname{Ext}$-residuals. No construction from an arbitrary empirical model to a realization prestack is claimed here.

Standard inputs and new contributions should be kept separate. Sheafification, stackification, banded gerbes, and the universal coefficient sequence are not claimed as new. Several of the results below (the sheafification characterization, the identification of visible presentations with sections of the sheafification, the banded gerbe decomposition, the \v{C}ech cocycle construction, the universal coefficient algebra) are routine consequences of standard machinery; they are included as propositions and lemmas to keep the paper self-contained. The genuinely new contribution at theorem level is the forcing-necessity theorem (\Cref{thm:pointwise-irreducibility}), showing that gluing predicates are not definable in the lower flat layer and only in the explicitly stated fragment. The semantic proposition \Cref{thm:gerbe-null-semantics} and the homological propositions of \S\ref{sec:homological-visibility} organize standard machinery relative to the chosen lift, while \Cref{thm:unique-branch-contextuality-comparison} is a conditional translation statement. The fully verified torus-type construction of \Cref{ex:end-to-end} checks every standing assumption A1--A4 in proposition-proof form.

The main notation is summarized in \Cref{tbl:notation}. All later references to standard background cite the propositions whose proofs have been moved to Appendix~\ref{app:deferred-proofs}.

\begin{table}[h]
\centering
\caption{Frequently used symbols.}\label{tbl:notation}
\begin{tabular}{ll}
\toprule
Symbol & Meaning (section) \\
\midrule
$\mathbb L_0,\mathbb L_1,\mathbb L_2$ & Lower, information-state, and enriched layers (\S\ref{sec:preliminaries}) \\
$\St(M)$ & Information states of a model $M$ (\S\ref{sec:information-states}) \\
$\mathfrak P_r,\mathfrak L_r$ & Chosen realization prestack and its stackification (\S\ref{sec:null-decomposition}) \\
$\Vis_{p,s}(r)$ & Visible value classes of a reference at $p$ (\S\ref{sec:null-decomposition}) \\
$\mathfrak G_v$ & Component gerbe of a visible branch $v$ (\S\ref{sec:gerbe-obstruction}) \\
$\omega_v$ & Branch obstruction class attached to $v$ (\S\ref{sec:gerbe-obstruction}) \\
$\Band$ & Fixed abelian band on the slice site (\S\ref{sec:null-decomposition}) \\
$K_{\omega},\,A_{\mathrm{vis}}^{\omega}$ & Hidden subgroup and presentation-relative visible quotient (\S\ref{sec:homological-visibility}) \\
\bottomrule
\end{tabular}
\end{table}

The standing assumptions and their use are as follows.
\begin{description}[labelwidth=2.6em]
\item[A1] (\emph{Chosen lift}) Every section that invokes visible branches assumes a gluing-sensitive realization prestack with fixed band (\S\ref{sec:null-decomposition}).
\item[A2] (\emph{Global conservativity}) Results connecting stack-global objects back to the presheaf (for instance \Cref{thm:gerbe-null-semantics}) additionally assume global conservativity at the terminal fibre.
\item[A3] (\emph{Finite presentation}) Homological visibility statements (\S\ref{sec:homological-visibility}) assume a finite nerve presentation with compatible band trivialization when \Cref{def:finite-nerve-presentation} is invoked.
\item[A4] (\emph{Branch constancy}) Decompositions such as \Cref{cor:branch-decomposition} assume the explicit branch-constancy hypothesis of \Cref{def:branch-constancy}.
\end{description}
Each major result cites exactly the assumptions it uses; \Cref{fig:assumption-map} in \S\ref{subsec:scope-chosen-lift} spells out a dependency diagram that readers can consult before entering the technical sections.

The paper is organized accordingly. \Cref{sec:preliminaries} and \Cref{sec:information-states} isolate the lower-layer semantics and its conservative-extension behavior. \Cref{sec:null-decomposition} develops local objects, sheafification, visible value classes, and the forcing-necessity theorem. \Cref{sec:gerbe-obstruction} gives the branch-gerbe semantics of gluing failure. \Cref{sec:homological-visibility} develops the presentation-relative visible quotient, the pure-Ext blindness proposition, invariance under change of presentation, and three worked constructions, including the fully verified torus-type case.

\subsection{Theorem-by-theorem comparison with prior work}

The lower-layer material in \Cref{sec:preliminaries,sec:information-states} is classical. \Cref{prop:conservative-composition,cor:whole-chain-invariance} are formal consequences of the definition of conservative extension; \Cref{def:lower-layer-syntax,def:information-state,def:refinement,def:state-forcing,prop:flat-team-reading,prop:singleton-conservativity,prop:lifted-pointwise-soundness} sit squarely in the Hodges--V\"a\"an\"anen team-semantics tradition and in the Beth--Kripke forcing tradition \cite{Hodges1997,Vaananen2007,Beth1956,Kripke1965,TroelstraVanDalen1988}. The increment in this part of the paper is organizational only: \Cref{def:logic-layer,def:conservative-extension,def:state-semantic-consequence} package standard ingredients so that the later local-object layer can be compared to the lower flat fragment without ambiguity.

The sheaf-theoretic material in \Cref{sec:null-decomposition} is likewise mostly standard once a site of admitted references has been fixed. \Cref{prop:local-reference-pretopology,thm:sheafification-characterization,cor:sheaf-local-to-global,cor:sheafification-removes-glue,prop:stackification-components} are applications of Grothendieck-pretopology, sheafification, and stackification results from Mac Lane--Moerdijk and the Stacks Project \cite{MacLaneMoerdijk1994,StacksProject}, while \Cref{prop:global-implies-compatible-local} is the usual local-to-global implication. The paper-specific contribution of this section is semantic rather than categorical: \Cref{def:reference-classes,def:local-reference-frame,def:existence-levels,def:visible-value-classes,def:typed-readout,def:null-failure-modes,def:decomposition-decidability} isolate the admissibility/compatibility/gluing distinctions the later results talk about, and \Cref{prop:split-realization-prestack} gives an explicit split lift available under separatedness.

The only unconditional theorem-level novelty is \Cref{thm:pointwise-irreducibility}. Its proof uses classical automorphism-invariance of first-order formulas and the collapse of singleton states to pointwise semantics \cite[Theorem~1.3.9]{HodgesShorter1997}, but the conclusion is not classical: in the constant-free equality-only comparison class on singleton parameter states, the predicates $\CompSec$, $\Sec$, and $\Null^{\mathrm{glue}}$ are not definable in the lower forcing language. The supporting propositions \Cref{prop:null-modes-imply-null,prop:null-modes-exclusive,prop:null-trichotomy,thm:null-classification-under-refinement,prop:finite-three-mode-model} serve only to sharpen that separation by isolating the three structurally distinct failure modes.

The gerbe section packages standard obstruction theory rather than new gerbe mathematics. \Cref{thm:component-gerbe-decomposition,thm:cech-bridge-compatible-realizations} are standard consequences of the theory of gerbes and their \v{C}ech classes in Giraud and the Stacks Project \cite{Giraud1971,StacksProject}, and \Cref{cor:branch-decomposition} is a straightforward consequence once the manuscript's branch-constancy hypothesis is imposed. The increment here is semantic packaging: \Cref{thm:visible-value-components,thm:gerbe-null-semantics} translate compatible families, visible branches, and non-neutral component gerbes into a dictionary adapted to information-state forcing relative to a chosen gluing-sensitive lift.

The algebra in \Cref{sec:homological-visibility} starts from classical universal-coefficient and character-duality facts \cite[Theorem~3.2]{Hatcher2002}, \cite[Sec.~2.3]{Weibel1994}, and \cite[Chaps.~5, 12]{DummitFoote2004}. What is new is the packaging of those facts into the presentation-relative visible quotient of \Cref{def:intrinsic-visible-quotient}, the explicit invariance statements \Cref{prop:visible-quotient-trivialization-invariance,prop:visible-quotient-lift-invariance,prop:visible-quotient-refinement-invariance,prop:presentation-groupoid}, and the pure-$\operatorname{Ext}$ interpretation of complete character-blindness in \Cref{thm:character-blind-obstructions}. Finally, \Cref{thm:unique-branch-contextuality-comparison} should be read as a conditional translation of the Abramsky--Brandenburger no-global-section picture \cite{AbramskyBrandenburger2011,AbramskyMansfield2012,AbramskyEtAl2015} together with Car\`u's incompleteness phenomenon \cite{Caru2017}: once a compatible realization prestack is supplied, the remaining blind residue is exactly the pure-$\operatorname{Ext}$ part.
