\section{Lower-Layer Semantics and Conservative Extension}\label{sec:preliminaries}

The later sections use only one abstract idea: a higher semantic layer may add coordinates or structure without changing the behavior of formulas already present below. The definitions in this section isolate that invariance principle and nothing more.

\begin{definition}[Logic layer]\label{def:logic-layer}
A \emph{logic layer} is a tuple
\[
\mathbb{L}=(\Sigma,\mathrm{Form},\Mod,\St,\Vdash,\models),
\]
where $\Sigma$ is a signature, $\mathrm{Form}$ is the class of well-formed $\Sigma$-formulas, $\Mod$ is a class of admissible models, $\St$ assigns to every $M\in\Mod$ a nonempty set $\St(M)$ of states, $\Vdash$ is a forcing relation on model-state pairs, and $\models$ is the induced consequence relation defined by
\[
\Gamma\models\varphi
\quad:\Longleftrightarrow\quad
\forall M\in\Mod\;\forall p\in\St(M)\,
\bigl(M,p\Vdash\Gamma \Rightarrow M,p\Vdash\varphi\bigr).
\]
\end{definition}

The forcing relation is the primitive semantic notion. Consequence is derived from uniform preservation of forcing across all admissible models and states.

\begin{definition}[Conservative extension]\label{def:conservative-extension}
Let
\[
\mathbb{L}_m=(\Sigma_m,\mathrm{Form}_m,\Mod_m,\St_m,\Vdash_m,\models_m)
\quad\text{and}\quad
\mathbb{L}_n=(\Sigma_n,\mathrm{Form}_n,\Mod_n,\St_n,\Vdash_n,\models_n)
\]
with $m<n$. We say that $\mathbb{L}_n$ is a \emph{conservative extension} of $\mathbb{L}_m$, written $\mathbb{L}_m\preceq\mathbb{L}_n$, if there exist
\begin{enumerate}[label=(\roman*),nosep]
\item a language embedding $e_{m,n}:\Sigma_m\hookrightarrow\Sigma_n$,
\item a forgetful functor $U_{n,m}:\Mod_n\to\Mod_m$,
\item for every $M\in\Mod_n$, a state projection
\[
\pi^M_{n,m}:\St_n(M)\to \St_m\bigl(U_{n,m}(M)\bigr),
\]
\end{enumerate}
such that for every $\varphi\in \mathrm{Form}_m$, every $M\in\Mod_n$, and every $p\in\St_n(M)$,
\begin{equation}\label{eq:conservative-extension}
M,p\Vdash_n \varphi
\quad\Longleftrightarrow\quad
U_{n,m}(M),\pi^M_{n,m}(p)\Vdash_m \varphi.
\end{equation}
\end{definition}

Thus a higher layer may add expressive power, state coordinates, or geometric constraints, but it may not alter the forcing behavior of formulas that already belonged to a lower layer.

\begin{proposition}[Composition of conservative extensions]\label{prop:conservative-composition}
If $\mathbb{L}_\ell\preceq\mathbb{L}_m$ and $\mathbb{L}_m\preceq\mathbb{L}_n$, then $\mathbb{L}_\ell\preceq\mathbb{L}_n$.
\end{proposition}

\begin{proof}
See Appendix~\ref{app:deferred-proofs}.
\end{proof}

\begin{corollary}[Finite-chain invariance]\label{cor:whole-chain-invariance}
If
\[
\mathbb{L}_0\preceq\mathbb{L}_1\preceq\cdots\preceq\mathbb{L}_n,
\]
then every formula belonging to any lower layer keeps the same forcing behavior when interpreted in any later layer and projected back down.
\end{corollary}

\begin{proof}
Apply \Cref{prop:conservative-composition} repeatedly along the chain.
\end{proof}

The focused note uses three layers. The link between $\mathbb{L}_0$ and $\mathbb{L}_1$ is the singleton embedding $i_M:\Val_M(\Gamma)\to \St(M)$ from \Cref{def:singleton-embedding}, not a conservative extension: \Cref{prop:singleton-conservativity} shows that $\mathbb{L}_1$ agrees with $\mathbb{L}_0$ precisely on those singleton states, and no truth-preserving projection exists from arbitrary information states back to pointwise assignments. The genuine conservative extension asserted in the paper is
\[
\mathbb{L}_1
\preceq
\mathbb{L}_2,
\]
where $\mathbb{L}_1$ is the information-state forcing layer on the lower syntax and $\mathbb{L}_2$ extends $\mathbb{L}_1$ by the reference and local-object predicates used later in the gluing discussion. By \Cref{cor:whole-chain-invariance}, passing upward from $\mathbb{L}_1$ to $\mathbb{L}_2$ is cumulative rather than revisionary.

Gerbes, branch obstructions, and homological evaluation are then built on top of the $\mathbb{L}_2$ data relative to a chosen gluing-sensitive realization prestack rather than as additional formal languages. The higher dynamic and complexity layers of the longer manuscript are intentionally omitted from the present note.
