\section{Branch gerbes and obstruction classes}\label{sec:gerbe-obstruction}

Fix a chosen realization prestack. Since its stackification $\mathfrak L_r$ may have several connected components over $a$, the obstruction theory is naturally branchwise: each visible branch gives rise to a component gerbe, and the relevant cohomology class belongs to that gerbe rather than to the reference as a whole. Throughout this section the results remain relative to the chosen lift and, when derived $H^2$ enters, to a chosen cofinal family of gerbe-adapted covers.

\begin{definition}[Component gerbes and branch obstruction classes]\label{def:compatible-realization-stack}\label{def:cech-gluing-obstruction}
Fix an admitted reference $r:\mathsf{Ref}_s$ at a state $p$, write
\[
a=[r]_p,
\]
and assume that a realization prestack $\mathfrak P_r$ with stackification $\mathfrak L_r$ and band $\Band$ has been chosen as in \Cref{def:realization-prestack}. For a visible value class
\[
v\in \Vis_{p,s}(r)\cong \pi_0(\mathfrak L_r)(a),
\]
define the \emph{component substack}
\[
\mathfrak G_v(U):=\{x\in \mathfrak L_r(U)\mid [x]=v|_U\in \pi_0(\mathfrak L_r)(U)\}
\]
for each object $U$ of $\mathcal C_{p,s}/a$. When \Cref{thm:component-gerbe-decomposition} applies, $\mathfrak G_v$ is called the \emph{component gerbe} of the visible branch $v$.

Assume moreover that, for every visible branch $v$, gerbe-adapted covers of $a$ satisfying the vanishing hypotheses of \Cref{thm:cech-bridge-compatible-realizations} form a cofinal family in $\Cov_{p,s}(a)$. Then the resulting filtered \v{C}ech class of the branch gerbe $\mathfrak G_v$ is denoted
\[
\omega_v\in \varinjlim_{\mathcal U}\Cech^2(\mathcal U,\Band)=:\Cech^2(\mathcal C_{p,s}/a,\Band)
\]
and called the \emph{branch obstruction class} of $v$.
The notation depends on the chosen realization prestack $\mathfrak P_r$ and the chosen band $\Band$.
\end{definition}

\begin{remark}
The cover-relative version appears in \cite[Definition~5.1]{MaZhang2026ConditionalGluingFailure}. The increment here is that, once gerbe-adapted covers form a cofinal family, the obstruction is promoted from a class on one chosen adapted cover to a filtered \v{C}ech branch class intrinsic to the preorder-site branch.
\end{remark}

\begin{remark}[\v{C}ech and derived level]\label{rem:cech-derived-level}
Even without the derived-comparison hypothesis in \Cref{def:cech-gluing-obstruction}, the cover-relative \v{C}ech classes constructed in \Cref{thm:cech-bridge-compatible-realizations} are enough for the neutrality criteria proved below. The class $\omega_v$ is the further datum obtained by identifying the filtered \v{C}ech system with degree-$2$ sheaf cohomology.
\end{remark}

\begin{proposition}[Component gerbes of visible branches]\label{thm:component-gerbe-decomposition}
In the setting of \Cref{def:compatible-realization-stack}, for every
\[
v\in \Vis_{p,s}(r)\cong \pi_0(\mathfrak L_r)(a),
\]
the full substack $\mathfrak G_v$ is a banded $\Band$-gerbe over $\mathcal C_{p,s}/a$.
\end{proposition}

\begin{proof}
Since $\mathfrak L_r$ is a stack in groupoids and $\mathfrak G_v$ is a full substack, $\mathfrak G_v$ is again a stack in groupoids. Let $U$ be an object of $\mathcal C_{p,s}/a$. Because $v|_U\in \pi_0(\mathfrak L_r)(U)$ is a section of the sheaf of connected components, by the definition of $\pi_0(\mathfrak L_r)$ there exists a cover $\{U_i\to U\}$ and objects
\[
x_i\in \mathfrak L_r(U_i)
\]
such that $[x_i]=v|_{U_i}$. Hence $(\mathfrak G_v)_{U_i}\neq\varnothing$, so $\mathfrak G_v$ is locally nonempty.

Now let $x,y\in \mathfrak G_v(U)$. Then $[x]=[y]=v|_U\in \pi_0(\mathfrak L_r)(U)$. By the definition of the sheaf of connected components, there exists a cover $\{U_i\to U\}$ such that
\[
x|_{U_i}\cong y|_{U_i}
\]
for every $i$. Thus any two objects of $\mathfrak G_v(U)$ are locally isomorphic, so $\mathfrak G_v$ is a gerbe \cite[Tag 06NY]{StacksProject}. The banding is inherited from $\mathfrak L_r$: since $\mathfrak G_v$ is a full substack, the automorphism sheaf of any object $x\in \mathfrak G_v(U)$ coincides with its automorphism sheaf in $\mathfrak L_r(U)$, hence is identified with $\Band|_U$.
\end{proof}

\begin{definition}[Branch constancy]\label{def:branch-constancy}
In the setting of \Cref{def:compatible-realization-stack}, we say that $\pi_0(\mathfrak L_r)$ is \emph{branch-constant over $a$} if for every object $U$ of $\mathcal C_{p,s}/a$, the restriction map
\[
\rho_U:\pi_0(\mathfrak L_r)(a)\longrightarrow \pi_0(\mathfrak L_r)(U)
\]
is bijective.
\end{definition}

\begin{corollary}[Global branch decomposition under branch constancy]\label{cor:branch-decomposition}
Assume branch constancy over $a$ in the sense of \Cref{def:branch-constancy}. Then the canonical functor
\[
\coprod_{v\in \Vis_{p,s}(r)} \mathfrak G_v\longrightarrow \mathfrak L_r
\]
is an equivalence of stacks in groupoids.
\end{corollary}

\begin{proof}
Let $U\in \mathcal C_{p,s}/a$ and $x\in \mathfrak L_r(U)$. Put $\xi:=[x]\in \pi_0(\mathfrak L_r)(U)$. By surjectivity of $\rho_U$, there exists $v\in \pi_0(\mathfrak L_r)(a)$ such that $\rho_U(v)=\xi$. Hence $x\in \mathfrak G_v(U)$, so the displayed functor is essentially surjective on every fibre.

For full faithfulness, let $x\in \mathfrak G_v(U)$ and $y\in \mathfrak G_w(U)$, and suppose there is a morphism $f:x\to y$ in $\mathfrak L_r(U)$. Any morphism in a groupoid is an isomorphism, hence $[x]=[y]$ in $\pi_0(\mathfrak L_r)(U)$. Therefore $\rho_U(v)=\rho_U(w)$. By injectivity of $\rho_U$, one gets $v=w$. Thus there are no morphisms in $\mathfrak L_r(U)$ between objects of distinct summands, and within each summand the morphism sets coincide because $\mathfrak G_v$ is a full substack.
\end{proof}

\begin{definition}[Discrete component stack and branch selector]\label{def:branch-selector}
In the setting of \Cref{def:compatible-realization-stack}, write
\[
\pi_0(\mathfrak L_r)_{\mathrm{disc}}
\]
for the stack in groupoids on $\mathcal C_{p,s}/a$ whose fibre over an object $U$ is the discrete groupoid on the set $\pi_0(\mathfrak L_r)(U)$. The canonical component projection is the morphism of stacks
\[
q_r:\mathfrak L_r\longrightarrow \pi_0(\mathfrak L_r)_{\mathrm{disc}}
\]
that sends an object to its connected-component class. A \emph{branch selector} for $r$ is a section of $q_r$, that is, a morphism of stacks
\[
s_r:\pi_0(\mathfrak L_r)_{\mathrm{disc}}\longrightarrow \mathfrak L_r
\]
satisfying
\[
q_r\circ s_r=\mathrm{id}_{\pi_0(\mathfrak L_r)_{\mathrm{disc}}}.
\]
\end{definition}

\begin{theorem}[Branch selectors and neutral component gerbes]\label{thm:branch-selector-neutrality}
In the setting of \Cref{def:branch-selector}, if the component projection
\[
q_r:\mathfrak L_r\to \pi_0(\mathfrak L_r)_{\mathrm{disc}}
\]
admits a branch selector, then every component gerbe $\mathfrak G_v$ is neutral.

If branch constancy over $a$ holds, the converse is also valid. If, in addition, global conservativity at $a$ holds in the sense of \Cref{def:global-conservativity}, then these equivalent conditions are equivalent to surjectivity of the terminal sheafification unit
\[
u_a:F_{p,s}(a)\longrightarrow \bigl(a_{p,s}F_{p,s}\bigr)(a)\cong \Vis_{p,s}(r).
\]
If, moreover, the chosen cofinal family of gerbe-adapted covers computes derived cohomology in degree $2$, they are further equivalent to
\[
\omega_v=0
\qquad
(\forall v\in \Vis_{p,s}(r)).
\]
\end{theorem}

\begin{proof}
Assume first that $q_r$ admits a branch selector
\[
s_r:\pi_0(\mathfrak L_r)_{\mathrm{disc}}\to \mathfrak L_r.
\]
Let
\[
v\in \Vis_{p,s}(r)\cong \pi_0(\mathfrak L_r)(a).
\]
Evaluating at the terminal object $a$ gives an object
\[
s_r(a)(v)\in \mathfrak L_r(a)
\]
whose component class is
\[
q_r(a)\bigl(s_r(a)(v)\bigr)=v.
\]
Hence
\[
s_r(a)(v)\in \mathfrak G_v(a),
\]
so $\mathfrak G_v$ is neutral. This proves the first implication.

Assume next that branch constancy holds and that every component gerbe $\mathfrak G_v$ is neutral. For each
\[
v\in \pi_0(\mathfrak L_r)(a),
\]
choose an object
\[
x_v\in \mathfrak G_v(a).
\]
Let $U$ be any object of $\mathcal C_{p,s}/a$ and let
\[
\xi\in \pi_0(\mathfrak L_r)(U).
\]
By branch constancy, there is a unique
\[
v_\xi\in \pi_0(\mathfrak L_r)(a)
\]
such that
\[
\rho_U(v_\xi)=\xi.
\]
Define
\[
s_r(U)(\xi):=x_{v_\xi}|_U\in \mathfrak L_r(U).
\]
Because the codomain of $q_r$ is discrete, there are no non-identity morphisms to specify. For a morphism
\[
f:V\to U
\]
in $\mathcal C_{p,s}/a$,
\[
f^*s_r(U)(\xi)=x_{v_\xi}|_V.
\]
Since restriction in $\pi_0(\mathfrak L_r)$ sends $\xi$ to $\xi|_V=\rho_V(v_\xi)$, uniqueness in the branch-constancy condition gives
\[
v_{\xi|_V}=v_\xi,
\]
hence
\[
f^*s_r(U)(\xi)=s_r(V)(\xi|_V).
\]
Thus $s_r$ is a morphism of stacks. By construction,
\[
q_r(U)\bigl(s_r(U)(\xi)\bigr)=\xi
\]
for every $U$ and every $\xi\in \pi_0(\mathfrak L_r)(U)$, so
\[
q_r\circ s_r=\mathrm{id}_{\pi_0(\mathfrak L_r)_{\mathrm{disc}}}.
\]
Therefore $q_r$ admits a branch selector.

Assume now global conservativity at $a$, and let
\[
u:F_{p,s}|_{\mathcal C_{p,s}/a}\to a_{p,s}F_{p,s}|_{\mathcal C_{p,s}/a}
\]
be the sheafification unit. If every component gerbe is neutral and
\[
v\in \Vis_{p,s}(r)\cong \pi_0(\mathfrak L_r)(a),
\]
choose
\[
x_v\in \mathfrak G_v(a)\subseteq \mathfrak L_r(a).
\]
Global conservativity yields
\[
y_v\in (\mathfrak P_r)_a
\]
with
\[
\iota_r(y_v)\cong x_v.
\]
Write
\[
\sigma_v:=[y_v]\in \pi_0^{\mathrm{pre}}(\mathfrak P_r)(a)\cong F_{p,s}(a).
\]
Since $u_a$ is the map on connected components induced by $\iota_r$,
\[
u_a(\sigma_v)=v.
\]
Thus $u_a$ is surjective.

Conversely, assume that $u_a$ is surjective and let
\[
v\in \Vis_{p,s}(r)\cong \bigl(a_{p,s}F_{p,s}\bigr)(a).
\]
Choose
\[
\sigma\in F_{p,s}(a)
\]
with
\[
u_a(\sigma)=v.
\]
By the identification
\[
F_{p,s}(a)\cong \pi_0^{\mathrm{pre}}(\mathfrak P_r)(a),
\]
there exists an object
\[
y\in (\mathfrak P_r)_a
\]
whose prestack component class is $\sigma$. Then
\[
\iota_r(y)\in \mathfrak L_r(a)
\]
has connected-component class
\[
[\iota_r(y)]=u_a(\sigma)=v.
\]
Therefore
\[
\iota_r(y)\in \mathfrak G_v(a),
\]
so $\mathfrak G_v$ is neutral. This proves the equivalence with surjectivity of $u_a$.

Under the additional derived-comparison hypothesis, \Cref{thm:cech-bridge-compatible-realizations}(iv) identifies vanishing of the \v{C}ech class on a gerbe-adapted cover with neutrality of $\mathfrak G_v$, and the assumed comparison with derived $H^2$ transports this equivalence to the filtered class $\omega_v$. Hence the displayed conditions are also equivalent to
\[
\omega_v=0
\qquad
(\forall v\in \Vis_{p,s}(r)).
\qedhere
\]
\end{proof}

\begin{proposition}[Constructing the \v{C}ech obstruction of a branch gerbe]\label{thm:cech-bridge-compatible-realizations}
Assume that $\mathfrak G_v$ is a banded component gerbe with band $\Band$ in the sense of \Cref{def:compatible-realization-stack}. Let
\[
\mathcal U=\{a_i\to a\}_{i\in I}\in \Cov_{p,s}(a)
\]
be a cover whose \v{C}ech nerve exists in $\mathcal C_{p,s}/a$, and write
\[
U_{i_0\cdots i_n}
\]
for the corresponding simplices. Assume
\[
H^1(U_i,\Band)=0
\qquad\text{and}\qquad
H^1(U_{ij},\Band)=0
\]
for all $i,j\in I$. Choose objects
\[
x_i\in \mathfrak G_v(U_i).
\]
Then:
\begin{enumerate}[label=(\roman*),nosep]
\item for every pair $(i,j)$ there exists an isomorphism
\[
\phi_{ij}:x_i|_{U_{ij}}\xrightarrow{\sim} x_j|_{U_{ij}};
\]
\item on each triple overlap $U_{ijk}$, the element
\[
g_{ijk}:=\phi_{ik}^{-1}|_{U_{ijk}}\circ \phi_{jk}|_{U_{ijk}}\circ \phi_{ij}|_{U_{ijk}}
\in \Band(U_{ijk})
\]
defines a \v{C}ech $2$-cocycle $g=(g_{ijk})\in \check Z^2(\mathcal U,\Band)$;
\item the class $[g]\in \Cech^2(\mathcal U,\Band)$ depends only on the branch gerbe $\mathfrak G_v$, the chosen realization prestack, and the chosen cover: changing the local objects or the isomorphisms changes $g$ by a \v{C}ech coboundary;
\item the class $[g]$ vanishes if and only if the branch gerbe $\mathfrak G_v$ is neutral over $a$.
\end{enumerate}
\end{proposition}

\begin{proof}
Because $\mathfrak G_v$ is a banded gerbe, the sheaf
\[
\mathrm{Isom}(x_i|_{U_{ij}},x_j|_{U_{ij}})
\]
is a $\Band|_{U_{ij}}$-torsor for every pair $(i,j)$. By the standard classification of torsors by $H^1$ \cite[Chap.~III]{MacLaneMoerdijk1994}, \cite[Sec.~11]{StacksProject}, the vanishing
\[
H^1(U_{ij},\Band)=0
\]
forces this torsor to be trivial. Hence global sections exist, which gives the isomorphisms $\phi_{ij}$ and proves \textup{(i)}.

For \textup{(ii)}, the composite
\[
\phi_{ik}^{-1}|_{U_{ijk}}\circ \phi_{jk}|_{U_{ijk}}\circ \phi_{ij}|_{U_{ijk}}
\]
is an automorphism of $x_i|_{U_{ijk}}$, hence a section of $\Band(U_{ijk})$. The usual associativity calculation on quadruple overlaps shows that $(g_{ijk})$ satisfies the \v{C}ech cocycle identity \cite[Chap.~IV]{Giraud1971}, \cite[Sec.~11]{StacksProject}.

For \textup{(iii)}, first change the $\phi_{ij}$ while keeping the objects $x_i$ fixed. Any other choice has the form
\[
\phi'_{ij}=\phi_{ij}\circ h_{ij}^{-1}
\]
for a $1$-cochain $h=(h_{ij})\in \check C^1(\mathcal U,\Band)$, and the corresponding cocycle is
\[
g'=(\delta h)\, g.
\]
Thus the \v{C}ech class is unchanged.

Next change the local objects. If $x'_i\in \mathfrak G_v(U_i)$ is another choice, then
\[
\mathrm{Isom}(x_i,x'_i)
\]
is a $\Band|_{U_i}$-torsor. Since $H^1(U_i,\Band)=0$, this torsor is trivial, so we may choose an isomorphism
\[
\psi_i:x_i\xrightarrow{\sim} x'_i.
\]
Transporting the $\phi_{ij}$ across the $\psi_i$ shows that the cocycle attached to $(x'_i)$ differs from $g$ by a coboundary. Finally, passage to a refinement pulls the cocycle back along the refinement map, so the induced class is stable in the filtered \v{C}ech system.

If the gerbe is neutral, choose a global object $x\in \mathfrak G_v(a)$. Then each
\[
\mathrm{Isom}(x|_{U_i},x_i)
\]
is a $\Band|_{U_i}$-torsor, hence trivial by $H^1(U_i,\Band)=0$. Choose isomorphisms
\[
\psi_i:x|_{U_i}\xrightarrow{\sim} x_i.
\]
With
\[
\phi_{ij}:=\psi_j|_{U_{ij}}\circ \psi_i^{-1}|_{U_{ij}},
\]
the resulting cocycle is identically zero. Thus neutrality implies $[g]=0$.

Conversely, if $[g]=0$, choose a $1$-cochain $h=(h_{ij})$ with
\[
g=\delta h.
\]
Replacing $\phi_{ij}$ by
\[
\phi'_{ij}:=\phi_{ij}\circ h_{ij}^{-1},
\]
we obtain isomorphisms satisfying the exact cocycle condition on triple overlaps. Since $\mathfrak G_v$ is a stack, the descent datum $(x_i,\phi'_{ij})$ is effective and glues to a global object over $a$. Hence the gerbe is neutral.
\end{proof}

This yields the promised branchwise description of gluing failure.

\begin{proposition}[Branchwise gerbe semantics of gluing-level absence]\label{thm:gerbe-null-semantics}
Assume the hypotheses of \Cref{def:cech-gluing-obstruction} for the chosen realization prestack $\mathfrak P_r$, together with global conservativity at $a$ in the sense of \Cref{def:global-conservativity}. Then, for the admitted reference $r:\mathsf{Ref}_s$ under consideration,
\[
M,p\Vdash \CompSec_s(r)
\quad\Longleftrightarrow\quad
\Vis_{p,s}(r)\neq\varnothing.
\]
Moreover,
\[
M,p\Vdash \Sec_s(r)
\quad\Longleftrightarrow\quad
\exists v\in \Vis_{p,s}(r)\ \text{such that}\ \mathfrak G_v\ \text{is neutral}.
\]
Consequently,
\[
M,p\Vdash \Null^{\mathrm{glue}}_s(r)
\quad\Longleftrightarrow\quad
\Vis_{p,s}(r)\neq\varnothing
\ \text{and}\
\forall v\in \Vis_{p,s}(r),\ \mathfrak G_v\ \text{is non-neutral}.
\]
If, in addition, the chosen cofinal family of gerbe-adapted covers computes derived cohomology in degree $2$, then the last condition is equivalent to
\[
\Vis_{p,s}(r)\neq\varnothing
\ \text{and}\
\omega_v\neq 0\ \text{for every}\ v\in \Vis_{p,s}(r).
\]
In particular, the single obstruction class picture is recovered precisely when $|\Vis_{p,s}(r)|=1$.
\end{proposition}

\begin{proof}
The first equivalence is the combination of \Cref{thm:sheafification-characterization,thm:visible-value-components}. Indeed, compatible local sectionability is equivalent to the nonemptiness of $(a_{p,s}F_{p,s})(a)$, and this set is exactly $\Vis_{p,s}(r)$.

For the second equivalence, \Cref{prop:no-new-global-objects} gives
\[
M,p\Vdash \Sec_s(r)
\quad\Longleftrightarrow\quad
\mathfrak L_r(a)\neq\varnothing.
\]
If $x\in \mathfrak L_r(a)$, then $x$ determines a visible value class
\[
v=[x]\in \pi_0(\mathfrak L_r)(a)\cong \Vis_{p,s}(r),
\]
and $x\in \mathfrak G_v(a)$, so $\mathfrak G_v$ is neutral. Conversely, if $\mathfrak G_v$ is neutral for some visible branch $v$, then any global object of $\mathfrak G_v(a)$ lies in $\mathfrak L_r(a)$, so $\mathfrak L_r(a)\neq\varnothing$.

Since $r$ is admitted, the definition
\[
\Null^{\mathrm{glue}}_s(r):=\Adm_s(r)\land \CompSec_s(r)\land \neg\Sec_s(r)
\]
now yields the third equivalence immediately.

For the final claim, if the chosen cofinal family computes derived cohomology in degree $2$, then \Cref{thm:cech-bridge-compatible-realizations}(iv) identifies the vanishing of each $\omega_v$ with neutrality of the corresponding component gerbe, while the comparison with derived $H^2$ is the standard banded-gerbe classification \cite[Chap.~IV]{Giraud1971}, \cite[Tag 06NY]{StacksProject}.
\end{proof}

\begin{corollary}[No branch selector in the gluing-null mode]\label{cor:no-branch-selector}
Assume the hypotheses of \Cref{thm:gerbe-null-semantics} for the chosen realization prestack. If
\[
M,p\Vdash \Null^{\mathrm{glue}}_s(r),
\]
then the component projection
\[
q_r:\mathfrak L_r\to \pi_0(\mathfrak L_r)_{\mathrm{disc}}
\]
admits no branch selector.
\end{corollary}

\begin{proof}
By \Cref{thm:gerbe-null-semantics},
\[
\Vis_{p,s}(r)\neq\varnothing
\]
and every component gerbe $\mathfrak G_v$ is non-neutral. If a branch selector existed, \Cref{thm:branch-selector-neutrality} would force every component gerbe to be neutral, a contradiction.
\end{proof}

\begin{example}[A torus-type branch gerbe]\label{ex:torus-gerbe}
The torus-type construction of \Cref{ex:end-to-end} yields a unique visible branch $v$ and a $\mathbb Z/4\mathbb Z$-banded component gerbe $\mathfrak G_v$ whose obstruction class evaluates a fundamental $2$-cycle to $2$. Hence the class is nonzero, the gerbe is non-neutral by \Cref{thm:cech-bridge-compatible-realizations}, and \Cref{thm:gerbe-null-semantics} places the admitted reference in the gluing-null mode.
\end{example}
