\section{Conclusion}\label{sec:conclusion}

The present version is a focused note built around a narrower claim than the longer manuscript. It starts from a chosen local-object layer and a chosen gluing-sensitive realization prestack, rather than from a claim that an arbitrary set-valued local object functor carries a canonical obstruction theory. Its first result is logical: compatible local sectionability and gluing-level absence are not definable in the lower information-state forcing language, so the local-object layer is genuinely necessary. Its second result is semantic: visible branches are exactly the connected components of the realization stack, and gluing failure is exactly the situation in which compatible local data exist but every component gerbe is non-neutral. Its third result is homological: once a branch obstruction class has been fixed, the intrinsic visible quotient is the cokernel of the homological evaluation map, so the visible part of an obstruction is precisely its $H_2$-detected part, while the blind remainder is exactly the pure $\operatorname{Ext}$-residual.

This is also the paper's main external point of contact with established work. In the unique-branch support-presheaf setting, the theory recovers the Abramsky--Brandenburger no-global-section picture and identifies the Car\`u-type cohomological blind cases with the pure $\operatorname{Ext}$-part of the obstruction class. The worked $S^2$-type and $\mathbb{R}P^2$-type examples then isolate the visible and blind extremes of the theory.

What has been omitted is deliberate. The measurable recursive-addressing, symbolic-dynamical cylinder, inverse-limit, multi-branch budget, refinement-dynamic, and complexity layers belong to a different narrative and would require their own separate development. The present paper is intended to stand or fall on the branch-gerbe semantics of gluing failure and on the identification of the intrinsic visible quotient and its pure-Ext blind residue.
