\documentclass[11pt,a4paper]{amsart}
% target: Integers: Electronic Journal of Combinatorial Number Theory

\usepackage{amsmath,amssymb,amsthm,mathrsfs}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{enumitem}
\usepackage[protrusion=true,expansion=false]{microtype}
\usepackage{mathtools}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\DeclareMathOperator{\val}{val}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\abs}[1]{\lvert #1 \rvert}

\title[Fixed-DFA Approximation of Prime Languages]{Fixed-DFA approximation obstructions for prime languages in fixed bases}

\author{Haobo Ma}
\address[Haobo Ma]{AELF PTE LTD., \#14-02, Marina Bay Financial Centre Tower 1, 8 Marina Blvd, Singapore 018981}
\email[Haobo Ma]{auric@aelf.io}

\author{Wenlin Zhang}
\address[Wenlin Zhang]{National University of Singapore, Singapore}
\email[Wenlin Zhang]{e1327962@u.nus.edu}

\date{}

\begin{document}

\begin{abstract}
We revisit finite-state approximations of prime languages from the
transfer-matrix viewpoint. First, we record the classical eventual
periodicity of normalized word counts for any fixed complete DFA on the
full base-\(b\) shift, but we retain the quantitative constants in a
form suitable for approximation lower bounds. Second, for every trim
deterministic presentation of a regular numeration system whose
underlying graph is strongly connected with Perron root \(\lambda>1\)
(including the Zeckendorf system), we show that intersections with any
fixed DFA inherit the same dichotomy: either the length-\(m\) slice has
exponentially small density, or it stays bounded away from zero along
an arithmetic progression. These two inputs yield uniform quantitative
obstructions for prime slices: no fixed DFA can keep both recall and
precision bounded below on infinitely many lengths, and the symmetric
difference from the base-\(b\) (respectively Zeckendorf) prime slice is
at least \(c_{\mathcal A,b}\,b^m/m\) (respectively
\(c_{\mathcal A,\varphi}\,\varphi^m/m\)) for all large \(m\). We also
tie the Zeckendorf prime obstruction to Rigo's theorem on
\(S\)-recognizable primes, giving an uniform quantitative restatement
for the canonical Fibonacci numeration system.
\end{abstract}

\subjclass[2020]{68Q45, 11A41, 11A63, 37B10}
\keywords{Deterministic finite automaton, base-\(b\) prime language, regular approximation,
Zeckendorf representation, density dichotomy}

\maketitle

\input{sec_introduction}
\input{sec_automata}
\input{sec_zeckendorf}
\input{sec_analytic}
\input{sec_conclusion}

\begin{thebibliography}{99}

\bibitem{AlloucheShallit2003}
J.-P. Allouche and J. Shallit,
\textit{Automatic Sequences: Theory, Applications, Generalizations},
Cambridge University Press, Cambridge, 2003.

\bibitem{BerstelReutenauer2011}
J. Berstel and C. Reutenauer,
\textit{Noncommutative Rational Series with Applications},
Cambridge University Press, Cambridge, 2011.

\bibitem{Berstel1973}
J. Berstel,
Sur la densit\'e asymptotique de langages formels,
\textit{Automata, Languages and Programming},
North-Holland, Amsterdam, 1973, pp.~345--358.

\bibitem{BodirskyGaertnerVonOertzenSchwinghammer2004}
M. Bodirsky, T. G\"artner, T. von Oertzen, and J. Schwinghammer,
Efficiently computing the density of regular languages,
\textit{LATIN 2004: Theoretical Informatics}, Lecture Notes in Comput.
Sci., vol.~2976, Springer, Berlin, 2004, pp.~262--270.

\bibitem{SalomaaSoittola1978}
A. Salomaa and M. Soittola,
\textit{Automata-Theoretic Aspects of Formal Power Series},
Springer, New York, 1978.

\bibitem{CharlierRampersad2011}
\'E. Charlier and N. Rampersad,
The growth function of \(S\)-recognizable sets,
\textit{Theoret. Comput. Sci.} \textbf{412} (2011), no.~39, 5400--5408.

\bibitem{CharlierRampersadRigoWaxweiler2010}
\'E. Charlier, N. Rampersad, M. Rigo, and L. Waxweiler,
The minimal automaton recognizing \(m\mathbb{N}\) in a linear numeration system,
\textit{Integers} \textbf{11B} (2011), Paper A4, 24~pp.

\bibitem{CordySalomaa2007}
B. Cordy and K. Salomaa,
On the existence of regular approximations,
\textit{Theoret. Comput. Sci.} \textbf{387} (2007), no.~2, 125--135.

\bibitem{KappesKintala2006}
M. Kappes and C. M. R. Kintala,
Tradeoffs between reliability and conciseness of deterministic
finite automata,
\textit{J. Autom. Lang. Comb.} \textbf{11} (2006), no.~3, 281--296.

\bibitem{EismanRavikumar2003}
G. Eisman and B. Ravikumar,
Approximate recognition of non-regular languages by finite automata,
\textit{arXiv:cs/0305029} (2003).

\bibitem{SzilardYuZhangShallit1992}
A. Szilard, S. Yu, K. Zhang, and J. Shallit,
Characterizing regular languages with polynomial densities,
\textit{Mathematical Foundations of Computer Science 1992}, Lecture Notes
in Computer Science, vol.~629, Springer, 1992, pp.~494--503.

\bibitem{Shallit1994}
J. Shallit,
Numeration systems, linear recurrences, and regular sets,
\textit{Inform. and Comput.} \textbf{113} (1994), no.~2, 331--347.

\bibitem{Sakarovitch2009}
J. Sakarovitch,
\textit{Elements of Automata Theory},
Cambridge University Press, Cambridge, 2009.

\bibitem{PerrinPin2004}
D. Perrin and J.-\'E. Pin,
\textit{Infinite Words: Automata, Semigroups, Logic and Games},
Academic Press, London, 2004.

\bibitem{BruyereHansel1997}
V. Bruy{\`e}re and G. Hansel,
Bertrand numeration systems and recognizability,
\textit{Theoret. Comput. Sci.} \textbf{181} (1997), no.~1, 17--43.

\bibitem{Cobham1969}
A. Cobham,
On the base-dependence of sets of numbers recognizable by finite automata,
\textit{Math. Systems Theory} \textbf{3} (1969), 186--192.

\bibitem{Cobham1972}
A. Cobham,
Uniform tag sequences,
\textit{Math. Systems Theory} \textbf{6} (1972), 164--192.

\bibitem{DrmotaMauduitRivat2009}
M. Drmota, C. Mauduit, and J. Rivat,
Primes with an average sum of digits,
\textit{Compos. Math.} \textbf{145} (2009), no.~2, 271--292.

\bibitem{FlajoletSedgewick2009}
P. Flajolet and R. Sedgewick,
\textit{Analytic Combinatorics},
Cambridge University Press, Cambridge, 2009.

\bibitem{Frougny1992}
C. Frougny,
Representations of numbers and finite automata,
\textit{Math. Systems Theory} \textbf{25} (1992), no.~1, 37--60.

\bibitem{HardyWright2008}
G. H. Hardy and E. M. Wright,
\textit{An Introduction to the Theory of Numbers},
6th ed., Oxford University Press, Oxford, 2008.
Revised by D. R. Heath-Brown and J. H. Silverman.

\bibitem{HartmanisShank1968}
J. Hartmanis and H. Shank,
On the recognition of primes by automata,
\textit{J. ACM} \textbf{15} (1968), no.~3, 382--389.

\bibitem{LecomteRigo2001}
P. B. A. Lecomte and M. Rigo,
Numeration systems on a regular language,
\textit{Theory Comput. Syst.} \textbf{34} (2001), no.~1, 27--44.

\bibitem{Rigo2000}
M. Rigo,
Generalization of automatic sequences for numeration systems on a regular language,
\textit{Theoret. Comput. Sci.} \textbf{244} (2000), no.~1--2, 271--281.

\bibitem{RigoMaes2002}
M. Rigo and A. Maes,
More on generalized automatic sequences,
\textit{J. Autom. Lang. Comb.} \textbf{7} (2002), no.~3, 351--376.

\bibitem{Kitchens1998}
B. P. Kitchens,
\textit{Symbolic Dynamics: One-Sided, Two-Sided and Countable State Markov Shifts},
Springer, Berlin, 1998.

\bibitem{LindMarcus1995}
D. Lind and B. Marcus,
\textit{An Introduction to Symbolic Dynamics and Coding},
Cambridge University Press, Cambridge, 1995.

\bibitem{MauduitRivat2010}
C. Mauduit and J. Rivat,
Sur un probl{\`e}me de Gelfond: la somme des chiffres des nombres premiers,
\textit{Ann. of Math.} (2) \textbf{171} (2010), no.~3, 1591--1646.

\bibitem{Sinya2021}
R. Sin'ya,
Asymptotic approximation by regular languages,
\textit{SOFSEM 2021: Theory and Practice of Computer Science},
Lecture Notes in Comput. Sci., vol.~12607, Springer, Cham, 2021,
pp.~74--88; arXiv:2008.01413.

\bibitem{Zeckendorf1972}
E. Zeckendorf,
Repr{\'e}sentation des nombres naturels par une somme de nombres de
Fibonacci ou de nombres de Lucas,
\textit{Bull. Soc. Roy. Sci. Li{\`e}ge} \textbf{41} (1972), 179--182.

\bibitem{Bell2020UpperDensity}
J. P. Bell,
The upper density of an automatic set is rational,
\textit{J. Th{\'e}or. Nombres Bordeaux} \textbf{32} (2020), no.~2, 585--604.

\bibitem{BellLidbetterShallit2020}
J. P. Bell, T. F. Lidbetter, and J. Shallit,
Additive number theory via approximation by regular languages,
\textit{Internat. J. Found. Comput. Sci.} \textbf{31} (2020), no.~6, 667--687.

\bibitem{Mullner2017Sarnak}
C. M{\"u}llner,
Automatic sequences fulfill the Sarnak conjecture,
\textit{Duke Math. J.} \textbf{166} (2017), no.~17, 3219--3290.

\bibitem{AdamczewskiDrmotaMullner2022}
B. Adamczewski, M. Drmota, and C. M{\"u}llner,
(Logarithmic) densities for automatic sequences along primes and squares,
\textit{Trans. Amer. Math. Soc.} \textbf{375} (2022), no.~8, 6657--6697.

\bibitem{ArtinMazur1965}
M. Artin and B. Mazur,
On periodic points,
\textit{Ann. of Math.} (2) \textbf{81} (1965), no.~1, 82--99.

\bibitem{Polya1923}
G. P{\'o}lya,
Sur les s{\'e}ries enti{\`e}res {\`a} coefficients entiers,
\textit{Proc. London Math. Soc.} (2) \textbf{21} (1923), no.~1, 22--38.

\bibitem{Carlson1921}
F. Carlson,
{\"U}ber Potenzreihen mit ganzzahligen Koeffizienten,
\textit{Math. Z.} \textbf{9} (1921), 1--13.

\bibitem{Estermann1928}
T. Estermann,
On certain functions represented by {D}irichlet series,
\textit{Proc. London Math. Soc.} (2) \textbf{27} (1928), no.~1, 435--448.

\bibitem{HanselPerrin1989}
G. Hansel and D. Perrin,
Rational probability measures,
\textit{Inform. and Comput.} \textbf{82} (1989), no.~2, 95--110.

\bibitem{Kozik2005}
J. Kozik,
Conditional densities of regular languages,
Manuscript, 2005.

\bibitem{Kozik2005Conditional}
J. Kozik,
Conditional densities of regular languages,
\textit{Mathematical Foundations of Computer Science 2005}, 2005.

\bibitem{Koga2019}
T. Koga,
On the density of regular languages,
\textit{Internat. J. Algebra Comput.} \textbf{29} (2019), no.~8, 1491--1508.

\bibitem{Koga2019Density}
T. Koga,
On the density of regular languages,
\textit{Inform. and Comput.}, 2019.

\bibitem{Montoya2025}
J.~A. Montoya,
Relative densities of formal languages,
\textit{Descriptional Complexity of Formal Systems}, Lecture Notes in Comput.
Sci., Springer, 2025, to appear.

\bibitem{Montoya2025Relative}
J.~A. Montoya,
Relative densities of formal languages,
\textit{Developments in Language Theory / DCFS}, 2025.

\bibitem{BertheGouletOuelletPerrin2025}
V. Berth{\'e}, H. Goulet-Ouellet, and D. Perrin,
Density of rational languages under shift invariant measures,
\textit{Proceedings of the 52nd International Colloquium on Automata, Languages,
and Programming (ICALP)}, LIPIcs, 2025, to appear.

\bibitem{BertheGouletPerrin2025}
V. Berth{\'e}, H. Goulet-Ouellet, and D. Perrin,
Density of rational languages under shift invariant measures,
\textit{Proceedings of ICALP}, 2025.

\bibitem{Sinya2015}
R. Sin'ya,
Asymptotic approximation by regular languages,
Preprint, 2015; see also \textit{SOFSEM 2021}, Lecture Notes in Comput. Sci.,
vol.~12607, Springer, Cham, 2021, pp.~74--88.

\end{thebibliography}

\end{document}


