\section{Introduction}\label{sec:introduction}

Finite-state symbolic models impose rigid constraints on counting
functions. In a deterministic automaton, the number of accepted words
of length \(m\) is governed by powers of a finite matrix. In a sofic
shift, the number of admissible words of a given length is likewise
controlled by a finite-state presentation. These mechanisms force
periodic density laws or exponential-polynomial growth, and that is
often enough to rule out arithmetic languages whose size is governed by
the prime number theorem.

This note is centered on a quantitative consequence rather than on a
new density theorem. Once the classical density dichotomy for regular
languages is written with explicit residue classes and constants, it
gives fixed-DFA approximation obstructions for prime slices. The
genuinely new outputs are the symmetric-difference and recall/precision
lower bounds in the fixed-base and Zeckendorf settings. The density
dichotomies themselves, the Zeckendorf growth law, and the analytic
remarks are included only as background in the specialized form needed
for those corollaries. The eventual-sofic exclusion for prime
Zeckendorf slices is recorded as a short corollary of regularity plus
non-regularity, not as a separate asymptotic theorem.

Throughout, we work with trim deterministic presentations whose
underlying directed graph is strongly connected and has Perron root
\(>1\); numeration automata with multiple strongly connected components
do occur \cite{CharlierRampersadRigoWaxweiler2010}, so this hypothesis
is part of the setup rather than a claim of full generality.

Throughout, \(f(m)\asymp g(m)\) means that \(c_1g(m)\le f(m)\le c_2g(m)\)
for some positive constants \(c_1,c_2\) and all sufficiently large
\(m\). We write \(F_0=0\), \(F_1=1\) for the Fibonacci numbers,
\(\varphi=(1+\sqrt5)/2\), and for a Zeckendorf word
\(w=x_1\cdots x_m\) we set
\[
  \val(w):=\sum_{k=1}^m x_kF_{k+1}.
\]

\subsection{Main results}

We summarize the background inputs and the quantitative corollaries
needed later; the precise residue-class refinements and proofs appear
in Sections~\ref{sec:automata}--\ref{sec:zeckendorf}.

We begin with the classical density dichotomy for deterministic
automata, recorded here because its constants feed directly into the
prime approximation bounds.

\begin{theorem}\label{thm:intro-density}
Let \(\mathcal A\) be a complete deterministic finite automaton over
\(\Sigma=\{0,1\}\), and let
\[
  a_m(\mathcal A)
  :=\#\{w\in\Sigma^m:\mathcal A\text{ accepts }w\}.
\]
Then there exist an integer \(p\ge 1\), non-negative constants
\(c_0,\dots,c_{p-1}\), and \(\theta\in(0,1)\) such that
\[
  \frac{a_m(\mathcal A)}{2^m}=c_{m\bmod p}+O(\theta^m)
  \qquad (m\to\infty).
\]
Consequently exactly one of the following holds:
\begin{enumerate}[label=\textup{(\roman*)}]
  \item \(a_m(\mathcal A)/2^m\) decays exponentially.
  \item Along some residue class modulo \(p\), the ratio
  \(a_m(\mathcal A)/2^m\) is bounded below by a positive constant.
\end{enumerate}
Section~\ref{sec:automata} proves this dichotomy. In particular, it
rules out the intermediate scale \(a_m(\mathcal A)/2^m\asymp m^{-\alpha}\).
\end{theorem}

When this is applied to the binary prime language
\(\mathcal P_m^{(2)}\), whose size is \(\asymp 2^m/m\) by the prime
number theorem \cite{HardyWright2008}, one obtains a sharp
incompatibility with finite-state recognition.

\begin{corollary}[Symmetric-difference lower bound for binary primes]
  \label{cor:intro-binary-primes}
Let \(L(\mathcal A)\subseteq\{0,1\}^\ast\) be recognised by a fixed
deterministic finite automaton. Then there exists \(c>0\) such that
\[
  \abs{(L(\mathcal A)\cap\{0,1\}^m)\triangle\mathcal P_m^{(2)}}
  \ge c\,\frac{2^m}{m}
\]
for all sufficiently large \(m\). The proof in
\Cref{sec:automata} further shows that positive-density languages have
\(\gg 2^m\) symmetric difference along an arithmetic progression.
\end{corollary}

\begin{corollary}[No simultaneous recall and precision for fixed DFA]
  \label{cor:intro-binary-precision}
  Let
  \[
    \mathrm{Rec}_m
    :=\frac{\abs{L(\mathcal A)\cap\mathcal P_m^{(2)}}}
            {\abs{\mathcal P_m^{(2)}}},
    \qquad
    \mathrm{Prec}_m
    :=\frac{\abs{L(\mathcal A)\cap\mathcal P_m^{(2)}}}
            {\abs{L(\mathcal A)\cap\{0,1\}^m}}
  \]
  whenever \(\abs{L(\mathcal A)\cap\{0,1\}^m}>0\). Then every fixed DFA
  satisfies one of the following:
  \begin{enumerate}[label=\textup{(\roman*)}]
    \item \(\mathrm{Rec}_m\to 0\) exponentially fast.
    \item \(\mathrm{Prec}_m\to 0\) along an infinite arithmetic
    progression of lengths.
  \end{enumerate}
\end{corollary}

\begin{example}[MSD-first base-\(5\) convention]\label{ex:msd-base}
  With digits read from the most significant position first, the word
  \(w=2031\in\{0,1,2,3,4\}^4\) encodes the integer
  \(2\cdot 5^3+0\cdot 5^2+3\cdot 5+1=266\). The leading digit must be
  non-zero when counting prime slices of fixed length: although the
  word \(0031\) has the same numerical value, it is excluded from
  \(\mathcal P^{(5)}_4\) because it lies outside
  \(\{1,\dots,4\}\Sigma_5^{3}\). This is the convention used throughout
  \Cref{sec:automata}.
\end{example}

For the Zeckendorf setting we use the following background growth law
for regular sublanguages of
\[
  \mathcal Z
  :=\{x_1\cdots x_m\in\{0,1\}^\ast:
      x_ix_{i+1}\neq 11,\ x_m=1\}.
\]

\begin{theorem}[Background: Zeckendorf growth law]
  \label{thm:intro-zeckendorf}
Let \(L\subseteq\mathcal Z\) be regular. After deleting states that are
not reachable from the initial state or cannot reach an accepting
state, let \(\rho\) be the maximal spectral radius of a strongly
connected component in a DFA recognizing \(L\). Define
\[
  N_L(T):=\#\{w\in L:\val(w)\le T\},
\]
where \(\val(w)\) is the Zeckendorf value of \(w\). Then
\[
  N_L(T)=T^{\alpha+o(1)},
  \qquad
  \alpha=
  \begin{cases}
    0, & \rho\le 1,\\[0.3em]
    \dfrac{\log \rho}{\log \varphi}, & \rho>1.
  \end{cases}
  \qquad (T\to\infty).
\]
\Cref{sec:zeckendorf} combines this background statement with
prime-counting estimates to deduce the quantitative Zeckendorf
corollaries. The non-regularity and eventual-sofic exclusions then
follow immediately.
\end{theorem}

\begin{example}[LSD-first Zeckendorf convention]\label{ex:lsd-zeck}
  Consider the integer \(23=F_8+F_3\). Its Zeckendorf expansion uses
  digits \(x_k\in\{0,1\}\) with \(x_k\) multiplying \(F_{k+1}\) and the
  least significant digit listed first. The admissible word
  \(w=0100001\) has length \(7\), satisfies the no-\(11\) rule, and
  evaluates to
  \(x_1F_2+x_2F_3+\cdots+x_7F_8=0+1\cdot 2+\cdots+1\cdot 21=23\).
  The terminal \(1\) is the most significant digit, consistent with the
  trim deterministic presentation used in \Cref{sec:zeckendorf}.
\end{example}

For completeness, we record two classical analytic obstructions.

\begin{theorem}[Background]\label{thm:intro-analytic}
The following statements hold.
\begin{enumerate}[label=\textup{(\roman*)}]
  \item If \((b_n)_{n\ge 1}\subseteq\ZZ_{\ge 0}\), if \(b_p>0\) for
  infinitely many primes \(p\), and if the Euler product
  \[
    Z_b(z):=\prod_{n\ge 1}(1-z^n)^{-b_n}
  \]
  converges to a holomorphic function on \(\abs{z}<1\), then it has the
  unit circle as a natural boundary.

  \item For every finite matrix \(M\), the function
  \[
    F_M(s):=\det(I-e^{-s}M)^{-1}
  \]
  is \(2\pi i\)-periodic. Consequently, if \(F(s)=\sum_{n\ge 1}a_n n^{-s}\)
  is a Dirichlet series with at least one coefficient \(a_n\neq 0\) for
  some \(n\ge 2\), then \(F\) cannot agree with \(F_M\) on any non-empty
  open subset of a right half-plane. In particular, \(F_M\) cannot agree
  with the Riemann \(\zeta\)-function on any non-empty open subset of
  \(\Re(s)>1\).
\end{enumerate}
\end{theorem}

\subsection{Relation to prior work}\label{subsec:prior-work}

We give a theorem-by-theorem account of what is known, what is a
specialized reproof, and what is new.

\begin{itemize}
  \item \textbf{\Cref{thm:intro-density} (Theorem~2.2; density dichotomy):
  classical.} The eventual periodicity of normalized word counts for a
  deterministic automaton is a direct Perron--Frobenius consequence and
  is standard in the density literature for regular languages; see
  \cite{Berstel1973,SalomaaSoittola1978,BodirskyGaertnerVonOertzenSchwinghammer2004,Bell2020UpperDensity,Koga2019Density,BerstelReutenauer2011}.
  We record it only in the residue-class form needed for the later
  prime estimates.

  \item \textbf{\Cref{lem:trim-density} and
  \Cref{prop:zeckendorf-regular-powerlaw}: classical.} These results are standard residue-class regroupings of
  the periodic Perron--Frobenius theorem for trim strongly connected
  presentations and of the known Zeckendorf growth law; see, e.g.,
  \cite{SalomaaSoittola1978,LecomteRigo2001,CharlierRampersad2011}.

  \item \textbf{\Cref{thm:parry-density} (Theorem~2.9; trim strongly connected
  presentations): background packaging.} This is a specialized
  relative-density formulation of the conditional-density literature
  \cite{HanselPerrin1989,Kozik2005Conditional,Koga2019Density,Montoya2025Relative,BertheGouletPerrin2025}.
  Its role is to isolate the trim, strongly connected case needed for
  the later prime obstructions.

  \item \textbf{\Cref{cor:intro-binary-primes} and
  \Cref{cor:intro-binary-precision} (Corollaries~2.5 and~2.7): new.}
  These corollaries give a uniform quantitative obstruction for
  approximating the base-\(b\) prime slices by a single fixed DFA, in
  terms of both symmetric difference and simultaneous recall/precision.
  The Zeckendorf analogues
  \Cref{cor:zeck-symmetric-diff,cor:zeck-recall-precision} follow from
  the same mechanism.

  \item \textbf{Sofic-slice exclusion
  (\Cref{rem:zeckendorf-primes-not-sofic}): easy corollary.} Once
  \Cref{cor:zeckendorf-prime-language-not-regular} is known, the
  regularity of sofic block languages immediately rules out eventual
  sofic realization of the prime Zeckendorf slices. We record this only
  because it fits the overall narrative.

  \item \textbf{\Cref{thm:intro-analytic} (analytic obstructions):
  classical background.} The natural-boundary and periodicity
  statements are standard consequences of the work of
  P\'olya--Carlson--Estermann and of the Artin--Mazur determinant
  formula \cite{Polya1923,Carlson1921,Estermann1928,ArtinMazur1965,BowenLanford1970}.
\end{itemize}
\smallskip
\noindent\textbf{Relation to density literature.} For ordinary
automatic sets, asymptotic density questions go back at least to
Berstel \cite{Berstel1973}; matrix and algorithmic treatments appear in
\cite{BodirskyGaertnerVonOertzenSchwinghammer2004,Bell2020UpperDensity},
and Koga's paper \cite{Koga2019Density} revisits Sin'ya's density-zero
criterion \cite{Sinya2021}. The rational-measure viewpoint of Hansel
and Perrin \cite{HanselPerrin1989} interprets regular languages via
probability measures on shifts of finite type. Kozik's MFCS 2005 paper
\cite{Kozik2005Conditional} formulates conditional densities for
deterministic presentations in the trim strongly connected setting,
Montoya \cite{Montoya2025Relative} recasts these questions in terms of
relative densities, and Berth\'e--Goulet-Ouellet--Perrin
\cite{BertheGouletPerrin2025} establish a broad density theorem under
shift-invariant measures. Accordingly,
\Cref{thm:intro-density,thm:intro-zeckendorf,thm:parry-density}
should be read here as background statements in a form suited to the
prime-slice applications.


\medskip
\noindent\textbf{Summary.} \Cref{thm:intro-density} together with
\Cref{lem:trim-density} and \Cref{prop:zeckendorf-regular-powerlaw}
record the classical density statements needed for deterministic
presentations. \Cref{thm:parry-density} packages the conditional-density
framework for precisely those trim, strongly connected presentations
with Perron root \(\lambda>1\) that drive our applications; we do not
claim any results for presentations with multiple strongly connected
components, consistent with the counterexamples in
\cite{CharlierRampersadRigoWaxweiler2010}. The genuinely new
contributions are the quantitative recall/precision and symmetric-difference
obstructions (\Cref{cor:intro-binary-primes,cor:intro-binary-precision}),
their Zeckendorf companions
(\Cref{cor:zeck-symmetric-diff,cor:zeck-recall-precision}). The
eventual-sofic exclusion is recorded only as the easy corollary
\Cref{rem:zeckendorf-primes-not-sofic}.

\medskip
The impossibility of recognizing prime numbers by finite automata in
classical settings goes back to Hartmanis and Shank
\cite{HartmanisShank1968}. The interaction between primes and digital
constraints has a much deeper literature; in particular, the work of
Mauduit and Rivat \cite{MauduitRivat2010} and of Drmota, Mauduit and
Rivat \cite{DrmotaMauduitRivat2009} shows that prime inputs display
highly nontrivial equidistribution phenomena inside automatic and
digit-restricted settings. For Zeckendorf and more general Parry
numeration systems, the regularity of admissible expansions is standard;
see \cite{Frougny1992,BruyereHansel1997}. Our treatment uses only this
regular syntax together with the transfer-matrix viewpoint on word
counts, familiar from symbolic dynamics and analytic combinatorics
\cite{LindMarcus1995,Kitchens1998,FlajoletSedgewick2009}.

\subsection{Organisation}

\Cref{sec:automata} proves the density dichotomy for deterministic
automata and derives the fixed-base prime corollaries.
\Cref{sec:zeckendorf} records the background growth law for regular
Zeckendorf sublanguages, deduces the corresponding quantitative prime
corollaries, and then notes the easy sofic exclusion.
\Cref{sec:analytic} collects background analytic observations
concerning Euler products and determinant models. Later sections cite
the statements recorded above directly, so their proofs and corollaries
need not restate the results in full.

