\section{Refinement Dynamics and Support}\label{sec:multiaxis-refinement}

The dynamics/support material is supplementary to the obstruction-theoretic core. All formal definitions, propositions, and proofs have been moved to Appendix~\ref{sec:appendix-dynamics}. This section records only the informal narrative needed for context.

\begin{itemize}
\item \textbf{Multi-axis updates.} Appendix~\ref{sec:appendix-dynamics} defines multi-axis dynamic structures and shows that axis-wise refinements preserve forcing and can witness delayed decidability without altering the semantic clauses. The explicit lifting principle (Appendix~\ref{sec:appendix-dynamics}, Corollary~\ref{cor:explicit-lifting}) explains how orthogonal axes can be recombined without changing typed readouts.
\item \textbf{Axis support.} The notions of support, indispensability, and minimal support, together with their complexity bounds, now reside in Appendix~\ref{sec:appendix-dynamics}. No later argument depends on these notions, but they provide a template for tracking how much information a refinement direction carries.
\item \textbf{Branch monotonicity and rewrites.} Refinement can only decrease the intrinsic hidden subgroup of a visible branch (Appendix~\ref{sec:appendix-dynamics}, Proposition~\ref{thm:branch-visibility-monotonicity}), and value-preserving rewrites cannot create new readout facts (Appendix~\ref{sec:appendix-dynamics}, Corollary~\ref{cor:normalization-no-fact-creation}). These results explain how the higher-layer semantics behave under additional bookkeeping without affecting the main theorems.
\end{itemize}

Readers interested in multi-axis dynamics can therefore consult Appendix~\ref{sec:appendix-dynamics} without disrupting the main narrative, and all references to the earlier Section~\ref{sec:multiaxis-refinement} now point to that appendix.
