\section{Preliminaries: Conservative Extensions}\label{sec:preliminaries}

The later sections use only one abstract idea: a higher semantic layer may add coordinates or structure without changing the behavior of formulas already present below. The definitions in this section isolate that invariance principle and nothing more.

\subsection*{Standing notation}
States are denoted by $p,q,u$, and admitted references by $r:\mathsf{Ref}_s$ with associated class $a=[r]_p$. The local reference site at a state and sort is $\mathcal C_{p,s}$, the local object functor is $F_{p,s}$, the visible value classes of a reference are $\Vis_{p,s}(r)$, and branch obstruction classes are denoted by $\omega$ or $\omega_v$. Subscripts such as $\mathrm{vis}$, $\mathrm{br}$, and $\mathrm{unif}$ consistently indicate visible, branch-sensitive, and branch-uniform quotients.

\begin{definition}[Logic layer]\label{def:logic-layer}
A \emph{logic layer} is a tuple
\[
\mathbb{L}=(\Sigma,\mathrm{Form},\Mod,\St,\Vdash,\models),
\]
where $\Sigma$ is a signature, $\mathrm{Form}$ is the class of well-formed $\Sigma$-formulas, $\Mod$ is a class of admissible models, $\St$ assigns to every $M\in\Mod$ a nonempty set $\St(M)$ of states, $\Vdash$ is a forcing relation on model-state pairs, and $\models$ is the induced consequence relation defined by
\[
\Gamma\models\varphi
\quad:\Longleftrightarrow\quad
\forall M\in\Mod\;\forall p\in\St(M)\,
\bigl(M,p\Vdash\Gamma \Rightarrow M,p\Vdash\varphi\bigr).
\]
\end{definition}

The forcing relation is the primitive semantic notion. Consequence is derived from uniform preservation of forcing across all admissible models and states.

\begin{definition}[Conservative extension]\label{def:conservative-extension}
Let
\[
\mathbb{L}_m=(\Sigma_m,\mathrm{Form}_m,\Mod_m,\St_m,\Vdash_m,\models_m)
\quad\text{and}\quad
\mathbb{L}_n=(\Sigma_n,\mathrm{Form}_n,\Mod_n,\St_n,\Vdash_n,\models_n)
\]
with $m<n$. We say that $\mathbb{L}_n$ is a \emph{conservative extension} of $\mathbb{L}_m$, written $\mathbb{L}_m\preceq\mathbb{L}_n$, if there exist
\begin{enumerate}[label=(\roman*),nosep]
\item a language embedding $e_{m,n}:\Sigma_m\hookrightarrow\Sigma_n$,
\item a forgetful functor $U_{n,m}:\Mod_n\to\Mod_m$,
\item for every $M\in\Mod_n$, a state projection
\[
\pi^M_{n,m}:\St_n(M)\to \St_m\bigl(U_{n,m}(M)\bigr),
\]
\end{enumerate}
such that for every $\varphi\in \mathrm{Form}_m$, every $M\in\Mod_n$, and every $p\in\St_n(M)$,
\begin{equation}\label{eq:conservative-extension}
M,p\Vdash_n \varphi
\quad\Longleftrightarrow\quad
U_{n,m}(M),\pi^M_{n,m}(p)\Vdash_m \varphi.
\end{equation}
\end{definition}

Thus a higher layer may add expressive power, state coordinates, or geometric constraints, but it may not alter the forcing behavior of formulas that already belonged to a lower layer.

\begin{definition}[Singleton semantic embedding]\label{def:singleton-semantic-embedding}
Let $\mathbb{L}_0=(\Sigma_0,\mathrm{Form}_0,\Mod_0,\St_0,\Vdash_0,\models_0)$ and $\mathbb{L}_1=(\Sigma_1,\mathrm{Form}_1,\Mod_1,\St_1,\Vdash_1,\models_1)$. We say that $\mathbb{L}_0$ \emph{embeds into $\mathbb{L}_1$ by a singleton semantic embedding}, written $\mathbb{L}_0\hookrightarrow\mathbb{L}_1$, if there exist
\begin{enumerate}[label=(\roman*),nosep]
\item a language embedding $e_{0,1}:\Sigma_0\hookrightarrow\Sigma_1$,
\item a forgetful functor $U_{1,0}:\Mod_1\to\Mod_0$,
\item for every $M\in\Mod_1$ and every pointwise state $\rho\in\St_0(U_{1,0}(M))$, a canonical singleton state
\[
\iota_{0,1}^M(\rho)\in\St_1(M)
\]
\end{enumerate}
such that for every $\varphi\in\mathrm{Form}_0$,
\begin{equation}\label{eq:singleton-semantic-embedding}
M,\iota_{0,1}^M(\rho)\Vdash_1\varphi
\quad\Longleftrightarrow\quad
U_{1,0}(M),\rho\Vdash_0\varphi.
\end{equation}
\end{definition}

A singleton semantic embedding is weaker than a conservative extension: it provides a state-preserving injection from pointwise states into information states, rather than a state projection defined on all higher-layer states. The inverse map from an arbitrary information state $p=(\Gamma,R)$ with $|R|>1$ to a single pointwise assignment does not exist in general while preserving all $\mathbb{L}_0$-formulas.

\begin{proposition}[Composition of conservative extensions]\label{prop:conservative-composition}
If $\mathbb{L}_\ell\preceq\mathbb{L}_m$ and $\mathbb{L}_m\preceq\mathbb{L}_n$, then $\mathbb{L}_\ell\preceq\mathbb{L}_n$.
\end{proposition}

\begin{proof}
Compose the language embeddings and forgetful functors:
\[
e_{\ell,n}:=e_{m,n}\circ e_{\ell,m},
\qquad
U_{n,\ell}:=U_{m,\ell}\circ U_{n,m}.
\]
For every $M\in\Mod_n$, define
\[
\pi^M_{n,\ell}:=
\pi^{\,U_{n,m}(M)}_{m,\ell}\circ \pi^M_{n,m}.
\]
If $\varphi\in\mathrm{Form}_\ell$ and $p\in\St_n(M)$, then
\[
M,p\Vdash_n\varphi
\Longleftrightarrow
U_{n,m}(M),\pi^M_{n,m}(p)\Vdash_m\varphi
\Longleftrightarrow
U_{n,\ell}(M),\pi^M_{n,\ell}(p)\Vdash_\ell\varphi,
\]
by two applications of \Cref{def:conservative-extension}. This is exactly the required conservativity condition.
\end{proof}

\begin{corollary}[Finite-chain invariance]\label{cor:whole-chain-invariance}
If
\[
\mathbb{L}_0\preceq\mathbb{L}_1\preceq\cdots\preceq\mathbb{L}_n,
\]
then every formula belonging to any lower layer keeps the same forcing behavior when interpreted in any later layer and projected back down.
\end{corollary}

\begin{proof}
Apply \Cref{prop:conservative-composition} repeatedly along the chain.
\end{proof}

The constructions used later fit into the chain
\[
\mathbb{L}_0
\xhookrightarrow{\iota}
\mathbb{L}_1
\preceq
\mathbb{L}_2
\preceq
\mathbb{L}_3
\preceq
\mathbb{L}_4,
\]
where the hooked arrow $\xhookrightarrow{\iota}$ denotes a singleton semantic embedding (\Cref{def:singleton-semantic-embedding}) and each $\preceq$ denotes a conservative extension in the sense of \Cref{def:conservative-extension}. Here $\mathbb{L}_0$ is the underlying pointwise semantics, $\mathbb{L}_1$ is the information-state forcing layer with reference sorts, $\mathbb{L}_2$ adds local-object predicates, admissibility, site structure, and the geometric machinery needed to formulate local-to-global questions, $\mathbb{L}_3$ adds structural absence and gluing-obstruction data, and $\mathbb{L}_4$ adds refinement dynamics. By \Cref{cor:whole-chain-invariance} applied to the conservative-extension segment $\mathbb{L}_1\preceq\cdots\preceq\mathbb{L}_4$, the passage up the chain from $\mathbb{L}_1$ onward is cumulative rather than revisionary. The base boundary $\mathbb{L}_0\hookrightarrow\mathbb{L}_1$ is not a conservative extension in the sense of \Cref{def:conservative-extension}: pointwise states embed into information states as singletons (\Cref{prop:singleton-conservativity}), but the inverse projection from an arbitrary multi-realization information state to a single pointwise assignment does not exist in general. This embedding nonetheless preserves forcing of $\mathbb{L}_0$-formulas on the canonical singleton image, which is the property actually used later.

\begin{proposition}[Concrete instantiation of the finite chain]\label{prop:chain-instantiation}
The chain
\[
\mathbb L_0\xhookrightarrow{\iota} \mathbb L_1\preceq \mathbb L_2\preceq \mathbb L_3\preceq \mathbb L_4
\]
is instantiated as follows.
\begin{enumerate}[label=(\roman*),leftmargin=3em]
\item \textbf{$\mathbb L_0\hookrightarrow\mathbb L_1$ (pointwise states to information states: singleton semantic embedding).}
The language embedding
\[
e_{0,1}:\Sigma_0\hookrightarrow \Sigma_1
\]
is the inclusion of the underlying many-sorted symbols. In addition, $\Sigma_1$ contains the reference sorts $\mathsf{Ref}_s$ (one for each referenceable sort $s\in S_{\mathrm{ref}}$) together with their equality; no geometric predicates or site structure is imposed at this step. The forgetful functor
\[
U_{1,0}:\Mod_1\to\Mod_0
\]
forgets the information-state presentation and the reference carrier sets, remembering only the underlying many-sorted structure of $\Sigma_0$. A pointwise state $\rho$ in $\mathbb L_0$ is represented inside $\mathbb L_1$ by the singleton information state
\[
\iota_{0,1}^M(\rho):=(\Gamma,\{\rho\}).
\]
By \Cref{prop:singleton-conservativity},
\[
M,\iota_{0,1}^M(\rho)\Vdash_1\varphi
\quad\Longleftrightarrow\quad
U_{1,0}(M),\rho\Vdash_0\varphi
\]
for every $\mathbb L_0$-formula $\varphi$. This is a singleton semantic embedding in the sense of \Cref{def:singleton-semantic-embedding}: it preserves forcing of $\mathbb L_0$-formulas on the canonical singleton image but does not provide a state projection from arbitrary $\mathbb L_1$-states back to $\mathbb L_0$-states.

\item \textbf{$\mathbb L_1\leadsto\mathbb L_2$ (information states to local objects).}
The embedding
\[
e_{1,2}:\Sigma_1\hookrightarrow \Sigma_2
\]
adjoins the admissibility predicates, the local-object predicates $\LocSec_s$, $\CompSec_s$, $\Sec_s$, the site-structure symbols (covers, refinement preorder, pullbacks), and the associated geometric notation. The reference sorts $\mathsf{Ref}_s$ and their equality are already present in $\Sigma_1$; the step from $\mathbb L_1$ to $\mathbb L_2$ equips references with the geometric structure needed to state local-to-global questions. The forgetful functor
\[
U_{2,1}:\Mod_2\to\Mod_1
\]
forgets the local reference sites, covers, and local object functors but retains the reference sorts and their carrier sets. The state projection
\[
\pi^M_{2,1}:\St_2(M)\to \St_1(U_{2,1}(M))
\]
is the identity on the underlying information state $(\Gamma_p,R_p)$. Since the forcing clauses for formulas already in $\mathrm{Form}_1$ never inspect the added local-object data, one has
\[
M,p\Vdash_2\varphi
\quad\Longleftrightarrow\quad
U_{2,1}(M),\pi^M_{2,1}(p)\Vdash_1\varphi
\]
for every $\varphi\in\mathrm{Form}_1$.

\item \textbf{$\mathbb L_2\leadsto\mathbb L_3$ (local objects to obstruction data).}
The embedding
\[
e_{2,3}:\Sigma_2\hookrightarrow \Sigma_3
\]
adjoins the predicates and symbols used to talk about structural absence, visible branches, component gerbes, and obstruction classes. The forgetful functor
\[
U_{3,2}:\Mod_3\to\Mod_2
\]
drops the realization prestack, band, component-gerbe, and obstruction data. The state projection
\[
\pi^M_{3,2}:\St_3(M)\to \St_2(U_{3,2}(M))
\]
again leaves the underlying state unchanged. Formulas from $\mathrm{Form}_2$ are evaluated exactly as before, so
\[
M,p\Vdash_3\varphi
\quad\Longleftrightarrow\quad
U_{3,2}(M),\pi^M_{3,2}(p)\Vdash_2\varphi
\]
for every $\varphi\in\mathrm{Form}_2$.

\item \textbf{$\mathbb L_3\leadsto\mathbb L_4$ (obstruction data to refinement dynamics).}
The embedding
\[
e_{3,4}:\Sigma_3\hookrightarrow \Sigma_4
\]
adjoins the update relations, axis indices, support notions, and rewrite language of the refinement layer. The forgetful functor
\[
U_{4,3}:\Mod_4\to\Mod_3
\]
forgets the dynamic axes, update structure, and rewrite operations. The state projection
\[
\pi^M_{4,3}:\St_4(M)\to \St_3(U_{4,3}(M))
\]
forgets only the new dynamic bookkeeping. Because the forcing clauses for $\mathrm{Form}_3$ do not depend on that added bookkeeping, one has
\[
M,p\Vdash_4\varphi
\quad\Longleftrightarrow\quad
U_{4,3}(M),\pi^M_{4,3}(p)\Vdash_3\varphi
\]
for every $\varphi\in\mathrm{Form}_3$.
\end{enumerate}
The base boundary $\mathbb L_0\hookrightarrow\mathbb L_1$ is a singleton semantic embedding: pointwise states embed faithfully into information states, preserving the forcing of all $\mathbb L_0$-formulas on that canonical image. The three higher boundaries $\mathbb L_1\preceq\mathbb L_2\preceq\mathbb L_3\preceq\mathbb L_4$ are conservative extensions in the sense of \Cref{def:conservative-extension}, each with an explicit forgetful functor and state projection.
\end{proposition}

This chain is the complete setting of the present paper. It is self-contained and does not presuppose any external framework. The actual conservative-extension segment starts at $\mathbb L_1$; the lower boundary $\mathbb L_0\hookrightarrow\mathbb L_1$ is a weaker but still useful embedding that faithfully represents pointwise truth inside the information-state layer.
