\section{Introduction}\label{sec:introduction}

Standard sheaf-theoretic semantics captures local-to-global passage: compatible local sections glue when the underlying presheaf is a sheaf, and fail to glue when it is not. The obstruction to gluing is recorded by cohomology classes, and the sheaf-theoretic analysis of contextuality \cite{AbramskyBrandenburger2011}, together with its cohomological refinements \cite{AbramskyMansfieldSoaresBarbosa2012,AbramskyBarbosaKishidaLalMansfield2015,Caru2017,Caru2018,Montanhano2021} and topological perspectives \cite{OkayRobertsBartlettRaussendorf2017,OkayTyhurstRaussendorf2018}, treats the absence of a global section as the central semantic defect. Mansfield's extendability programme and its Abramsky--Barbosa follow-ups \cite{MansfieldBarbosa2012,MansfieldThesis,AbramskyBarbosaLogicContextuality,CaruAbramskyValuation2021} contrast logical, valuation-algebraic, and topological packaging of disagreement, while the bundle perspective of Barbosa--Kharoof--Okay and the cohomological refinements for measurement-based quantum computation \cite{BarbosaKharoofOkay2024,Aasnaess2022} emphasise the geometric meaning of obstruction classes. In that broader literature, Car\`u's completeness programme and Montanhano's semimodule-valued refinements are designed precisely to control false positives beyond the basic obstruction. The pointwise comparison considered in this paper is narrower: within the automorphism-invariant constant-free pointwise fragment of the lower-layer reduct, the distinction between compatible local sectionability and global sectionability is invisible. A formula evaluated at a single realization can register only pointwise value data; under the syntactic restrictions made precise later, it does not detect that compatible local data exist but fail to glue.

The present work sits in the forcing and sheaf-theoretic tradition in logic, building on the Kripke semantics for intuitionistic logic \cite{Kripke1965}, Beth's semantic tableaux constructions \cite{Beth1956}, Fitting's model-theoretic forcing \cite{Fitting1969}, Tierney's early topos-theoretic forcing perspective \cite{Tierney1972}, and the later expositions of Goldblatt and Johnstone \cite{Goldblatt2006,Johnstone2002}. What distinguishes our setting is the explicit stratification of the semantics into conservative extension layers, so that each enrichment---from pointwise truth to information-state forcing, from forcing to local objects, from local objects to gluing obstructions---is accompanied by a precise invariance guarantee.

This gap motivates the present paper, written for Logical Methods in Computer Science. We construct a four-layer chain
\[
\mathbb{L}_0 \xhookrightarrow{\iota} \mathbb{L}_1 \preceq \mathbb{L}_2 \preceq \mathbb{L}_3 \preceq \mathbb{L}_4,
\]
where $\mathbb{L}_0$ is pointwise semantics, $\mathbb{L}_1$ adds information-state forcing together with reference sorts, $\mathbb{L}_2$ adds admissibility predicates, local-object predicates, and the explicit site structure needed to formulate local-to-global questions, $\mathbb{L}_3$ adds structural absence and gluing-obstruction data, and $\mathbb{L}_4$ adds refinement dynamics. The base boundary $\mathbb{L}_0\hookrightarrow\mathbb{L}_1$ is a singleton semantic embedding: each pointwise state maps faithfully into the information-state layer, but the inverse projection from arbitrary multi-realization information states to single pointwise assignments does not exist in general. The three higher boundaries are conservative extensions in the sense of \Cref{def:conservative-extension}, with explicit forgetful functors and state projections spelled out in \Cref{prop:chain-instantiation}. Repeated composition along the conservative-extension segment gives the layer-preservation statement recorded in \Cref{cor:whole-chain-invariance}. The chain is the minimal framework in which the restricted logical separation theorem, the conditional gerbe semantics, and the homological visibility quotient can all be stated together.

\medskip
\noindent\textbf{Research question.}
Once an admitted reference carries a finite abelian gluing obstruction, what part of that obstruction is semantically visible without enlarging the state space, and how much additional information is required to recover the hidden part?

\medskip
\noindent\textbf{Main results.}
The answer is delivered by three groups of results, each located at a specific boundary of the extension chain. The homological result is the main mathematical payoff; the forcing layer isolates the logical comparison class; and the gerbe layer is retained only as a conditional semantic framework making the later hypotheses explicit. The standard sheafification, gerbe, and homological tools are used transparently as background ingredients; the nonstandard content lies in the way the comparison class, visible branches, and homological quotient are tied together. Table~\ref{tab:roadmap} summarises which statements are new, which inputs are classical, and which hypotheses are required in each case.

\begin{table}[t]
\centering
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{|p{0.22\textwidth}|p{0.32\textwidth}|p{0.36\textwidth}|}
\hline
Statement & Standard input & New contribution / hypotheses \\
\hline
Theorem~A (Theorem~\ref{thm:pointwise-irreducibility}) & Sheafification characterization of $\CompSec_s$ and automorphism actions on lower layers & Automorphism-invariant separation for the one-variable constant-free $\Sigma_1$-fragment of the $\mathbb L_1$-reduct. \\
\hline
Theorem~B (Theorems~\ref{thm:component-gerbe-decomposition}, \ref{thm:gerbe-null-semantics}) & Stackification of realization prestacks, Giraud classification of gerbes & Conditional branched semantics together with a gluing-sensitive realization datum for GHZ/Mermin support presheaves, including the flag-nerve computation that feeds the obstruction. \\
\hline
Theorem~C (Theorem~\ref{thm:intrinsic-visible-quotient} and Corollaries~\ref{cor:strict-to-intrinsic-visible}, \ref{cor:uniform-hidden-budget-factorization}) & Universal coefficient theorem, finite Pontryagin duality & Intrinsic visible quotient independent of the chosen presentation under homology-equivalent refinements, branchwise aggregation exact sequence, and intrinsic hidden-state budget bounds. \\
\hline
\end{tabular}
\caption{Roadmap of statements, indicating which inputs are reused and where new mathematics or hypotheses enter.}
\label{tab:roadmap}
\end{table}

\medskip
\noindent\textbf{Theorem~A} (restricted logical separation; $\mathbb{L}_1$--$\mathbb{L}_2$ boundary).
Compatible local sectionability is characterized by sheafification (\Cref{thm:sheafification-characterization}), and the predicates $\CompSec_s$, $\Sec_s$, and $\Null^{\mathrm{glue}}_s$ are not definable by automorphism-invariant constant-free pointwise formulas of the $\mathbb{L}_1$-reduct with one free reference variable, within the comparison class fixed in \Cref{thm:pointwise-irreducibility}. The restriction to one variable is essential (Section~\ref{sec:null-decomposition}), but it isolates precisely the fragment that is actually needed when the lower layer ranges over references without parameters.

\medskip
\noindent\textbf{Theorem~B} (conditional branched gerbe semantics; $\mathbb{L}_2$--$\mathbb{L}_3$ boundary).
The gerbe layer remains conditional on a gluing-sensitive realization prestack whose stackification recovers the sheafification of isomorphism classes, is globally conservative at the terminal fibre, and is equipped with a cofinal \v{C}ech-to-derived comparison family of covers. Under those hypotheses, visible value classes are the global sections of the sheaf of connected components of the realization stack (\Cref{thm:visible-value-components}), each visible branch $v$ determines a banded component gerbe $\mathfrak{G}_v$ (\Cref{thm:component-gerbe-decomposition}), and gluing-level absence holds exactly when visible branches exist but every component gerbe is non-neutral (\Cref{thm:gerbe-null-semantics}). The GHZ/Mermin verification (\Cref{thm:ghz-realization}) spells out this data explicitly, including the flag-nerve computation required to compare Čech and derived cohomology for that empirical model.

\medskip
\noindent\textbf{Theorem~C} (homological visibility; at $\mathbb{L}_3$).
Under explicit finite-nerve and realizability hypotheses, and after restricting to covers on which the finite abelian band is trivialized, the universal coefficient theorem yields a homological evaluation map $\mathrm{ev}_\omega : H_2(N(\mathcal{U}),\mathbb{Z}) \to A$ whose image is the intrinsic hidden subgroup $K_\omega$ and whose cokernel is the intrinsic visible quotient $A_{\mathrm{vis}}^\omega$. The preparatory algebra is standard; the substantive point is the semantic interpretation: the class-admissible characters are exactly the annihilator of that image, and the blind residual---the part of the obstruction invisible to every admissible character---is precisely the pure $\operatorname{Ext}$-type regime singled out in \Cref{thm:character-blind-obstructions}. The invariance statement (\Cref{prop:invariance-visible-quotients}) now makes the homology-equivalence requirement explicit: homology-preserving refinements, changes of cocycle, and band re-trivializations do not affect $K_\omega$ or $A_{\mathrm{vis}}^\omega$. Under assumptions (i)--(iv) of \Cref{thm:unique-branch-contextuality-comparison}, including the unique-branch condition, the theory yields a formal comparison with the no-global-section criterion in the support-presheaf setting of \cite{AbramskyBrandenburger2011}.

\medskip
\noindent\textbf{Secondary consequences.}
The multi-branch theory yields an exact comparison between branch-sensitive and branch-uniform visibility. The canonical surjection
\[
A_{\mathrm{vis}}^{\mathrm{br}}(r) \twoheadrightarrow A_{\mathrm{vis}}^{\mathrm{unif}}(r)
\]
has kernel $(\sum_v K_v)/(\bigcap_v K_v)$, so the hidden budget behind uniform visibility factors into a branch-resolution term and a within-branch residual (\Cref{thm:branch-comparison-exact-sequence,thm:branch-resolution-budget,cor:uniform-hidden-budget-factorization}). Worked finite examples (\Cref{ex:one-bit-hidden-obstruction,ex:rp2-character-blind,ex:s2-single-branch-maximal-collapse,ex:two-branch-sphere-separation}) start from explicit local-object functors together with chosen local realization data and separate the visible, blind, and multi-branch regimes.

\medskip
\noindent\textbf{Related work.}
The logical separation theorem isolates the precise limitation of automorphism-invariant formulas and therefore complements the logic-of-contextuality programme of \cite{AbramskyBarbosaLogicContextuality}. The realization layer sits next to the extendability analyses of \cite{MansfieldBarbosa2012,MansfieldThesis} and the valuation/algebraic comparison of \cite{CaruAbramskyValuation2021}, while the bundle perspective of \cite{BarbosaKharoofOkay2024} and the MBQC-focused refinements of \cite{Aasnaess2022} provide geometric context for the gerbe packaging used here. Finally, the unique-branch comparison (\Cref{thm:unique-branch-contextuality-comparison}) spells out how the present machinery interacts with the cohomological incompleteness issues studied by \cite{Caru2017,Caru2018,Montanhano2021}: the intrinsic visibility quotient pinpoints the exact obstruction classes that remain blind even after the standard fixes.

\medskip
\noindent\textbf{Hypothesis stratification.}
The hypotheses used in these results are stratified rather than cumulative. The site-theoretic assumptions enter already at the local-object layer: the reference preorder is promoted to a thin category equipped with a Grothendieck pretopology and the pullbacks needed for matching families. Branch constancy enters only for the decomposition corollary (\Cref{cor:branch-decomposition}), not for the component-gerbe theorem itself. Global conservativity is used only when stack-global objects are compared back to the original local object functor. Finite nerve presentations appear only in the finite abelian visibility theorems, and those theorems are explicitly restricted to covers on which the band is trivialized; nontrivial monodromy would require local coefficient systems. The logical separation result is an automorphism-invariant pointwise non-definability theorem in a constant-free lower comparison class, the gerbe semantics is conditional on gluing-sensitive realization data together with a \v{C}ech-to-derived comparison hypothesis, and the cohomological statements are conditional on finite-nerve realizability together with a fixed cohomological model.

\medskip
\noindent\textbf{Plan of the paper.}
\Cref{sec:preliminaries} defines logic layers and conservative extensions and introduces the four-layer chain, together with its concrete adjacent interfaces. \Cref{sec:information-states} defines information states, forcing, and lifted pointwise soundness. \Cref{sec:null-decomposition} upgrades the local-reference geometry to an explicit site, proves the sheafification characterization, states the restricted forcing-necessity theorem (Theorem~A), and isolates the additional gluing-sensitive realization hypotheses needed later. \Cref{sec:gerbe-obstruction} first develops the conditional branched gerbe semantics (Theorem~B) and then proves the GHZ/Mermin realization theorem and the end-to-end worked example demanded by the referee. \Cref{sec:homological-visibility} fixes the cohomological conventions for finite nerves, proves the invariance of the intrinsic visible quotient, and derives the exact branch-aggregation statements (Theorem~C). The refinement-dynamics material previously in \Cref{sec:multiaxis-refinement} has been moved to Appendix~\ref{sec:appendix-dynamics}; the appendix now records both those supplementary constructions and the finite-state complexity bounds.
