\section{Conclusion}\label{sec:conclusion}

\subsection*{Summary of results}

Under the stated hypotheses, this paper addresses the following question: given an admitted reference carrying a finite abelian gluing obstruction, what part of that obstruction is semantically visible without enlarging the state space?

Package~A shows that, under its explicit syntactic and invariance hypotheses, compatible local sectionability and the gluing predicates are not definable by automorphism-invariant constant-free pointwise formulas of the lower-layer reduct with one free reference variable. The predicates $\CompSec_s$, $\Sec_s$, and $\Null^{\mathrm{glue}}_s$ are therefore not reducible to the automorphism-invariant pointwise fragment of the $\mathbb L_1$-reduct considered in \Cref{thm:pointwise-irreducibility}.

Package~B provides a conditional gerbe-theoretic framework rather than an unconditional realization theorem. Under gluing-sensitive realization, global conservativity, and \v{C}ech-to-derived comparison hypotheses, visible value classes are the global sections of the sheaf of connected components of the realization stack; each visible branch determines a banded component gerbe; and gluing-level absence is characterized by the existence of visible branches together with the non-neutrality of all component gerbes. The GHZ/Mermin discussion in \Cref{thm:ghz-realization} makes explicit the extra amplitude-lift data required to place a support presheaf inside that framework; no intrinsic GHZ value of $K_\omega$ is claimed from the support presheaf alone.

Package~C is the main quantitative result under finite-nerve hypotheses with trivialized finite abelian band. The image of the homological evaluation map is the intrinsic hidden subgroup, the quotient by that image is the intrinsic visible quotient, and the blind residual---the part invisible to every admissible character---is exactly the pure $\operatorname{Ext}$ regime. Under assumptions \textup{(i)}--\textup{(iv)} of \Cref{thm:unique-branch-contextuality-comparison}, including the unique-branch condition and the realization hypotheses, this yields a formal comparison with the no-global-section criterion in the support-presheaf setting and identifies the precise mechanism behind the cohomological blind cases discussed by Car\`u and later completeness refinements \cite{Caru2017,Caru2018,Montanhano2021}. The invariance statement emphasizes that homology-equivalent refinements, rather than arbitrary refinements, preserve the intrinsic quantities.

\subsection*{External significance}

The connection to contextuality gives the results an anchor outside pure sheaf theory. The restricted pointwise undefinability theorem shows that, within the automorphism-invariant constant-free pointwise fragment fixed in Package~A, the gluing predicates are not definable; the passage from pointwise semantics to local objects is therefore an enrichment relative to that specific comparison class. The branched gerbe semantics replaces the binary local-to-global defect of standard contextuality analysis with a graded, branch-indexed account, but only after the additional realization hypotheses are verified. The homological visibility theorem then turns the universal coefficient sequence into a tool for describing the boundary between what is semantically accessible and what remains hidden, without claiming a complete cohomological invariant in the sense of \cite{Caru2018,Montanhano2021}.

\subsection*{Relation to topos-theoretic forcing}

The four-layer chain developed here shares its conceptual ancestry with topos-theoretic forcing as developed by Tierney, Johnstone, and Mac Lane--Moerdijk \cite{Tierney1972,Johnstone2002,MacLaneMoerdijk1994}. In the topos-theoretic tradition, forcing is modeled by the internal logic of a sheaf topos, and conservativity results take the form of geometric morphism adjunctions that preserve logical structure. The present framework is narrower in one respect---it works with a fixed site rather than varying the ambient topos---but broader in another: the explicit stratification into numbered layers with explicit forgetful functors and state projections (\Cref{def:conservative-extension}) makes the conservativity claims syntactically verifiable at each boundary, rather than relying on the general preservation properties of geometric morphisms. The gerbe-theoretic layer, in particular, uses the Giraud classification in a site-specific rather than topos-internal way, which permits the finite combinatorial analysis of visibility that drives Package~C. Whether the full four-layer chain admits a natural topos-internal reformulation---with each layer corresponding to a subterminal object or a Lawvere--Tierney topology on a single ambient topos---remains an open question that would connect the present results to the broader categorical semantics programme.

\subsection*{Open problems}

Two directions remain open. First, the finite-state complexity upper bounds in the appendix have no matching lower bounds; determining the exact complexity of the support and minimal-support problems remains a supplementary combinatorial question rather than part of the main proof chain. Second, the present paper treats a single admitted reference at one state and restricts the homological analysis to trivialized bands on finite covers. Extending the theory to local coefficient systems, and verifying realization hypotheses together with amplitude lifts in broader semantic model classes such as empirical support presheaves, remains future work. The multi-observer extension---indexed families of references with coupled states and observer symmetry---is likewise deferred to a sequel.
