\section{Homological visibility and intrinsic quotients}
\label{sec:homological-visibility}

We now fix one visible branch together with a finite nerve presentation. The quotient attached to a chosen cocycle still depends on that presentation, so in order to isolate the part determined by the cohomology class itself we pass to the homological evaluation map. Throughout the section we work with one fixed finite cover and the comparison maps coming from the identification of \v{C}ech cochains with simplicial cochains on its finite nerve. The resulting hidden subgroup and visible quotient are intrinsic only relative to the homology-equivalent changes of presentation singled out below.

We also assume that the chosen cover $\mathcal U$ trivializes the band, so that the coefficients may be viewed as a constant finite abelian group $A$ on the nerve. The case of nontrivial monodromy, which would require local coefficients, is not considered here.

\subsection{Finite nerve presentations}

\begin{definition}[Finite nerve presentation of the obstruction class]\label{def:finite-nerve-presentation}
Fix an admitted reference $r:\mathsf{Ref}_s$ at a state $p$, and assume conditions \textup{(i)}--\textup{(iv)} of \Cref{def:finite-obstruction-presentation}. A \emph{finite nerve presentation} of the branch obstruction class consists of:
\begin{enumerate}[label=(\roman*),nosep]
\item a finite covering family
\[
\mathcal U=\{a_i\to a\}_{i\in I},
\qquad
a=[r]_p,
\]
whose \v{C}ech nerve $N(\mathcal U)$ is finite;
\item a trivialization of the band on $\mathcal U$, so that the band may be viewed as a constant coefficient finite abelian group $A$ on $N(\mathcal U)$;
\item the induced obstruction class
\[
\omega\in H^2(N(\mathcal U),A),
\]
obtained by identifying \v{C}ech cochains on $\mathcal U$ with simplicial cochains on the finite nerve, transporting the branch obstruction cocycle through that identification, and writing
\[
\omega:=\omega_v
\]
for the chosen visible branch. When the class is also viewed as a derived $H^2$-class, this identification is part of the chosen presentation.
\end{enumerate}
Write
\[
\eta_A:H^2(N(\mathcal U),A)\longrightarrow \mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),A)
\]
for the map in the universal coefficient exact sequence
\[
0\to \operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),A)\to H^2(N(\mathcal U),A)\xrightarrow{\eta_A}\mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),A)\to 0
\]
\cite[Sec.~3.1]{Hatcher2002}, and define the \emph{homological evaluation map}
\[
\mathrm{ev}_{\omega}:=\eta_A(\omega).
\]
\end{definition}

\begin{proposition}[Invariance of $K_\omega$ and $A_{\mathrm{vis}}^\omega$]\label{prop:invariance-visible-quotients}
Let $(\mathcal U,A,\omega)$ and $(\mathcal U',A',\omega')$ be finite nerve presentations of the same branch obstruction class. Suppose the two presentations differ by one of the following operations:
\begin{enumerate}[label=(\roman*),nosep]
\item refinement of the cover by a simplicial map $f:N(\mathcal U')\to N(\mathcal U)$ that induces an isomorphism $f_*:H_2(N(\mathcal U'),\mathbb Z)\xrightarrow{\sim} H_2(N(\mathcal U),\mathbb Z)$, with the coefficient group fixed;
\item replacement of the \v{C}ech cocycle by a cohomologous cocycle in $H^2(N(\mathcal U),A)$;
\item re-trivialization of the band by an isomorphism of coefficient groups $\theta:A\xrightarrow{\sim} A'$.
\end{enumerate}
Then there exists a canonical group isomorphism $\phi:A\to A'$ such that
\[
\phi\bigl(K_{\omega}\bigr)=K_{\omega'}
\qquad\text{and}\qquad
\phi\ \text{induces}\ A_{\mathrm{vis}}^{\omega}\cong A_{\mathrm{vis}}^{\omega'}.
\]
In particular, both $K_{\omega}$ and $A_{\mathrm{vis}}^{\omega}$ are invariant under the three comparison operations listed above.
\end{proposition}

\begin{remark}[Typical refinements]\label{rem:refinement-iso}
The refinements used later are simplicial homotopy equivalences on the chosen nerves---for example, barycentric subdivisions and the standard refinements attached to good covers---and therefore satisfy the hypothesis of case \textup{(i)} of \Cref{prop:invariance-visible-quotients}.
\end{remark}

\begin{remark}[Necessity of homology equivalences]\label{rem:homology-equality}
The equalities in \Cref{prop:invariance-visible-quotients} rely on the assumption that the refinement map induces an isomorphism on $H_2$. Without that hypothesis one only obtains inclusions $\mathrm{Im}(\mathrm{ev}_{\omega'})\subseteq \mathrm{Im}(\mathrm{ev}_{\omega})$ and hence a comparison map $A_{\mathrm{vis}}^{\omega}\twoheadrightarrow A_{\mathrm{vis}}^{\omega'}$, but not equality. The term ``intrinsic visible quotient'' is therefore reserved for refinements and presentations that are homology-equivalent in this precise sense.
\end{remark}

\begin{proof}
Case \textup{(ii)} is immediate because cohomologous cocycles have the same image under $\eta_A$.

In case \textup{(i)}, the assumption on $f$ gives an isomorphism on $H_2$. Since $\omega'=f^*(\omega)$ and the universal coefficient map is natural, one has
\[
\mathrm{ev}_{\omega'}=\eta_A\bigl(f^*(\omega)\bigr)=\eta_A(\omega)\circ f_*,
\]
where $f_*:H_2(N(\mathcal U'),\mathbb Z)\to H_2(N(\mathcal U),\mathbb Z)$ is induced on homology and is an isomorphism by hypothesis. Hence $\mathrm{Im}(\mathrm{ev}_{\omega'})=\mathrm{Im}(\mathrm{ev}_\omega)$, so $K_{\omega'}=K_{\omega}$ and $A_{\mathrm{vis}}^{\omega'}\cong A_{\mathrm{vis}}^{\omega}$.

In case \textup{(iii)}, the coefficient isomorphism $\theta$ identifies the two cohomology groups, and $\omega'=\theta_*(\omega)$. Naturality gives
\[
\mathrm{ev}_{\omega'}=\eta_{A'}(\omega')=\eta_{A'}\bigl(\theta_*(\omega)\bigr)=\theta\circ \eta_A(\omega)=\theta\circ \mathrm{ev}_\omega.
\]
Therefore $K_{\omega'}=\theta(K_{\omega})$, and $\theta$ induces an isomorphism between $A_{\mathrm{vis}}^\omega$ and $A_{\mathrm{vis}}^{\omega'}$.

The required $\phi$ is the identity in case \textup{(i)}, the identity after identifying cohomologous cocycles in case \textup{(ii)}, and $\theta$ in case \textup{(iii)}.
\end{proof}

\begin{proposition}[Finite nerve comparison isomorphism]\label{prop:finite-nerve-comparison}
Let $\mathcal U=\{a_i\to a\}_{i\in I}$ and $N(\mathcal U)$ be as above. For each $n\ge 0$, each nonempty $(n{+}1)$-fold overlap $a_{i_0\cdots i_n}$ determines a unique $n$-simplex $(i_0,\dots,i_n)$ of $N(\mathcal U)$, and the assignment
\[
\Theta_{\mathcal U}^n:\check C^n(\mathcal U,A)\to C^n(N(\mathcal U),A),\qquad
(\alpha_{i_0\cdots i_n})\longmapsto\bigl((i_0,\dots,i_n)\mapsto \alpha_{i_0\cdots i_n}\bigr)
\]
is an isomorphism of cochain groups. Under this correspondence,
\[
\delta\leftrightarrow d,\qquad
B^n\leftrightarrow B^n,\qquad
Z^n\leftrightarrow Z^n.
\]
Hence cohomology is identified canonically.
\end{proposition}
\begin{proof}
The $n$-cochains on $\mathcal U$ are families of coefficients indexed by nonempty $(n{+}1)$-fold overlaps $a_{i_0\cdots i_n}$, while the simplicial $n$-cochains on $N(\mathcal U)$ are coefficient functions on the $n$-simplices $(i_0,\dots,i_n)$. By definition of the nerve, these two index sets agree, so $\Theta_{\mathcal U}^n$ is a bijection.

The \v{C}ech coboundary of $(\alpha_{i_0\cdots i_n})$ is the alternating sum of restrictions to the faces of each overlap, and the simplicial coboundary of $\Theta_{\mathcal U}^n(\alpha)$ is the alternating sum over the faces of the simplex $(i_0,\dots,i_{n+1})$. Because the face maps of $N(\mathcal U)$ are induced by omitting one index, the two alternating sums coincide term by term. Hence
\[
\Theta_{\mathcal U}^{n+1}(\delta\alpha)=d(\Theta_{\mathcal U}^n(\alpha)).
\]
Therefore $\Theta_{\mathcal U}^\bullet$ is an isomorphism of cochain complexes, and it induces the stated identifications on cocycles, coboundaries, and cohomology.
\end{proof}

\begin{proposition}[Strict versus intrinsic hidden subgroups]\label{prop:strict-vs-intrinsic-hidden-subgroups}
In the setting of \Cref{def:finite-nerve-presentation}, let $\alpha\in Z^2(N(\mathcal U),A)$ be a simplicial cocycle representing $\omega$, and let
\[
\widetilde\alpha:C_2(N(\mathcal U),\mathbb Z)\to A
\]
be the chain-level evaluation map obtained by linear extension from the oriented $2$-simplices. Then
\[
H_\alpha=\operatorname{Im}(\widetilde\alpha)
\]
and
\[
K_{\omega}:=\mathrm{Im}(\mathrm{ev}_{\omega}).
\]
Thus $H_\alpha$ records presentation-level chain visibility: it is generated by the values of the chosen cocycle on all $2$-chains. By contrast, $K_{\omega}$ records cycle-level class visibility: it is the image that survives after passage to $2$-homology. In particular,
\[
K_{\omega}\subseteq H_\alpha.
\]
\end{proposition}
\begin{proof}
The image of $\widetilde\alpha$ is the subgroup generated by the values of $\alpha$ on the oriented $2$-simplices, equivalently by the values of the original \v{C}ech cocycle on the nonempty triple overlaps of $\mathcal U$. This is exactly the strict subgroup attached to the chosen presentation. The second equality is the definition of $K_\omega$. Since $H_2(N(\mathcal U),\mathbb Z)$ is represented by $2$-cycles inside $C_2(N(\mathcal U),\mathbb Z)$, every homological value counted by $\mathrm{ev}_\omega$ is, a fortiori, a chain-level value of $\widetilde\alpha$, so $K_\omega\subseteq H_\alpha$.
\end{proof}

\subsection{Intrinsic visibility and admissible characters}

\begin{definition}[Intrinsic visible quotient]\label{def:intrinsic-visible-quotient}
In the setting of \Cref{def:finite-nerve-presentation}, define the \emph{intrinsic hidden subgroup}
\[
K_{\omega}:=\mathrm{Im}(\mathrm{ev}_{\omega})
\]
and the associated \emph{intrinsic visible quotient}
\[
A_{\mathrm{vis}}^{\omega}:=\operatorname{coker}(\mathrm{ev}_{\omega})=A/K_{\omega}.
\]
\end{definition}

\begin{definition}[Class-admissible characters]\label{def:class-admissible-character}
In the setting of \Cref{def:finite-nerve-presentation}, let
\[
X_{\omega}:=\{\chi\in\widehat A\mid \chi_*(\omega)=0\},
\qquad
\widehat A:=\mathrm{Hom}(A,\mathbb T).
\]
Elements of $X_{\omega}$ are called \emph{class-admissible} for $\omega$.
\end{definition}

\begin{theorem}[Character-theoretic detection of the intrinsic visible quotient]\label{thm:intrinsic-visible-quotient}
In the setting of \Cref{def:class-admissible-character},
\[
X_{\omega}
=
\{\chi\in\widehat A\mid \chi\circ \mathrm{ev}_{\omega}=0\}
=
\mathrm{Ann}(K_{\omega}).
\]
Consequently,
\[
K_{\omega}
=
\bigcap_{\chi\in X_{\omega}}\ker\chi,
\qquad
A_{\mathrm{vis}}^{\omega}\cong A\Big/\!\bigcap_{\chi\in X_{\omega}}\ker\chi.
\]
\end{theorem}

\begin{proof}
For every character $\chi:A\to\mathbb T$, naturality of the universal coefficient sequence gives a commutative square
\[
\begin{CD}
H^2(N(\mathcal U),A) @>\eta_A>> \mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),A) \\
@V\chi_*VV @VV(\chi\circ -)V \\
H^2(N(\mathcal U),\mathbb T) @>\eta_{\mathbb T}>> \mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),\mathbb T).
\end{CD}
\]
Since $\mathbb T$ is divisible and hence injective in $\mathbf{Ab}$, one has
\[
\operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),\mathbb T)=0
\]
\cite[Sec.~2.3]{Weibel1994}. Thus $\eta_{\mathbb T}$ is an isomorphism by \Cref{def:finite-nerve-presentation}. Therefore
\[
\chi_*(\omega)=0
\Longleftrightarrow
\eta_{\mathbb T}(\chi_*(\omega))=0
\Longleftrightarrow
\chi\circ \eta_A(\omega)=0
\Longleftrightarrow
\chi\circ \mathrm{ev}_{\omega}=0.
\]
This proves the first displayed identity, and the second is simply the definition of the annihilator of a subgroup of $A$.

Let $H:=K_{\omega}=\mathrm{Im}(\mathrm{ev}_{\omega})\le A$. By the first part of the proof, $X_{\omega}=\mathrm{Ann}(H)$. Finite abelian duality implies
\[
\bigcap_{\chi\in \mathrm{Ann}(H)}\ker\chi=H
\]
\cite[Part~I, Chaps.~2, 12]{Terras1999}. Hence
\[
K_{\omega}
=
\bigcap_{\chi\in X_{\omega}}\ker\chi.
\]
The final isomorphism follows immediately:
\[
A_{\mathrm{vis}}^{\omega}
=
A/K_{\omega}
\cong
A\Big/\!\bigcap_{\chi\in X_{\omega}}\ker\chi.
\qedhere
\]
\end{proof}

\begin{corollary}[Character detection of the intrinsic visible quotient]\label{cor:intrinsic-character-detection}
The quotient map
\[
\pi_{\omega}:A\twoheadrightarrow A_{\mathrm{vis}}^{\omega}
\]
induces a natural isomorphism
\[
\widehat{A_{\mathrm{vis}}^{\omega}}\cong X_{\omega}.
\]
For elements $x,y\in A$, one has
\[
\pi_{\omega}(x)=\pi_{\omega}(y)
\quad\Longleftrightarrow\quad
\chi(x)=\chi(y)\ \text{for every class-admissible }\chi.
\]
\end{corollary}

\begin{proof}
By \Cref{thm:intrinsic-visible-quotient},
\[
X_{\omega}=\mathrm{Ann}(K_{\omega})=K_{\omega}^{\perp}.
\]
Characters of $A_{\mathrm{vis}}^{\omega}=A/K_{\omega}$ are exactly the characters of $A$ that kill $K_{\omega}$, so precomposition with $\pi_{\omega}$ identifies $\widehat{A_{\mathrm{vis}}^{\omega}}$ with $X_{\omega}$. The separation statement follows from finite abelian duality applied to the finite group $A_{\mathrm{vis}}^{\omega}$ \cite[Part~I, Chap.~2]{Terras1999}.
\end{proof}

\begin{proposition}[Initiality among pure-extension quotients]\label{thm:intrinsic-pure-ext-initiality}
Let
\[
q:A\twoheadrightarrow B
\]
be a surjective homomorphism onto a finite abelian group $B$. In the setting of \Cref{def:finite-nerve-presentation}, the following are equivalent:
\begin{enumerate}[label=(\roman*),nosep]
\item the pushed-forward class $q_*(\omega)\in H^2(N(\mathcal U),B)$ lies in the subgroup
\[
\operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),B);
\]
\item $\eta_B(q_*(\omega))=0$;
\item $q\circ \mathrm{ev}_{\omega}=0$;
\item $\mathrm{Im}(\mathrm{ev}_{\omega})\subseteq \ker q$;
\item $q$ factors uniquely through
\[
\pi_{\omega}:A\twoheadrightarrow A_{\mathrm{vis}}^{\omega}.
\]
\end{enumerate}
Hence $A_{\mathrm{vis}}^{\omega}$ is the initial coefficient quotient through which the $H_2$-visible part of the obstruction disappears, leaving only a pure $\operatorname{Ext}$-type remnant.
\end{proposition}

\begin{proof}
The equivalence of \textup{(i)} and \textup{(ii)} is exactly the universal coefficient sequence for $B$ \cite[Sec.~3.1]{Hatcher2002}. Naturality gives
\[
\eta_B(q_*(\omega))=q\circ \eta_A(\omega)=q\circ \mathrm{ev}_{\omega},
\]
so \textup{(ii)} and \textup{(iii)} are equivalent. Conditions \textup{(iii)} and \textup{(iv)} are equivalent by definition of the image of a homomorphism. By \Cref{def:intrinsic-visible-quotient}, $\mathrm{Im}(\mathrm{ev}_{\omega})=K_{\omega}$, so \textup{(iv)} is equivalent to
\[
K_{\omega}\subseteq \ker q.
\]
By the universal property of the quotient by $K_{\omega}$, this is equivalent to unique factorization of $q$ through $\pi_{\omega}$, proving \textup{(v)}.
\end{proof}

\begin{corollary}[Class-trivializing initiality when $H_1$ vanishes]\label{cor:h1-vanishing-class-trivializing}
Assume
\[
H_1(N(\mathcal U),\mathbb Z)=0.
\]
Then, for every surjective homomorphism $q:A\twoheadrightarrow B$, the following are equivalent:
\begin{enumerate}[label=(\roman*),nosep]
\item $q_*(\omega)=0$ in $H^2(N(\mathcal U),B)$;
\item $q$ factors uniquely through
\[
\pi_{\omega}:A\twoheadrightarrow A_{\mathrm{vis}}^{\omega}.
\]
\end{enumerate}
Hence, when $H_1(N(\mathcal U),\mathbb Z)=0$, the intrinsic visible quotient is initial among coefficient quotients that annihilate the obstruction class itself.
\end{corollary}

\begin{proof}
If $H_1(N(\mathcal U),\mathbb Z)=0$, then
\[
\operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),B)=0
\]
for every abelian group $B$. Thus the pure-extension condition in \Cref{thm:intrinsic-pure-ext-initiality}(i) is equivalent to $q_*(\omega)=0$, and the conclusion follows.
\end{proof}

\begin{remark}[Canonical visibility without a canonical splitting]\label{rem:no-canonical-splitting}
The quotient $A_{\mathrm{vis}}^{\omega}$ and the subgroup $K_{\omega}=\mathrm{Im}(\mathrm{ev}_{\omega})$ are canonical. By contrast, the universal coefficient sequence need not split canonically. Thus the theory isolates a canonical visible quotient and a canonical blind subgroup without requiring any noncanonical decomposition of $\omega$ into visible and invisible summands.
\end{remark}

\begin{corollary}[Comparison with strict visible quotients]\label{cor:strict-to-intrinsic-visible}
Let $\alpha$ be any finite obstruction presentation of the class $\omega$. Then the intrinsic visible quotient
\[
A_{\mathrm{vis}}^{\omega}
\]
admits a canonical surjective homomorphism onto the strict visible quotient
\[
A/H_\alpha.
\]
Equivalently, every strict visible quotient is a quotient of the intrinsic visible quotient.
\end{corollary}

\begin{proof}
Let $q:A\twoheadrightarrow A/H_\alpha$ be the canonical quotient. By \Cref{thm:initial-visible-quotient}, the quotient kills every cocycle value $\alpha_{ijk}$ pointwise, so the pushed-forward cocycle representative is zero. Hence $q_*(\omega)=0$, and in particular it lies in the $\operatorname{Ext}$-subgroup of $H^2(N(\mathcal U),A/H_\alpha)$. Apply \Cref{thm:intrinsic-pure-ext-initiality}.
\end{proof}

\begin{definition}[Chain-level evaluation of a presentation]\label{def:chain-level-evaluation}
Let $\alpha\in Z^2(N(\mathcal U),A)$ be a finite obstruction presentation of the class $\omega$. Write
\[
C_2(N(\mathcal U),\mathbb Z)
\]
for the simplicial group of integral $2$-chains on the finite nerve. The \emph{chain-level evaluation map} of $\alpha$ is the homomorphism
\[
\widetilde\alpha:C_2(N(\mathcal U),\mathbb Z)\to A
\]
defined by linear extension from the oriented $2$-simplices:
\[
\widetilde\alpha\!\left(\sum_{\sigma} n_{\sigma}\sigma\right)
:=
\sum_{\sigma} n_{\sigma}\alpha(\sigma).
\]
\end{definition}

\begin{proposition}[Strict visibility versus homological visibility]\label{prop:chain-vs-homological-visibility}
In the setting of \Cref{def:chain-level-evaluation}, the following hold.
\begin{enumerate}[label=(\roman*),nosep]
\item $\widetilde\alpha$ vanishes on $2$-boundaries and therefore induces a homomorphism
\[
\overline\alpha:H_2(N(\mathcal U),\mathbb Z)\to A.
\]
\item The induced homomorphism $\overline\alpha$ is exactly the homological evaluation map $\mathrm{ev}_{\omega}$.
\item The strict obstruction subgroup and the intrinsic hidden subgroup satisfy
\[
H_\alpha=\mathrm{Im}(\widetilde\alpha),
\qquad
K_{\omega}=\mathrm{Im}(\overline\alpha)=\widetilde\alpha(Z_2(N(\mathcal U),\mathbb Z)).
\]
\end{enumerate}
Thus the gap between strict visibility and intrinsic visibility is exactly the gap between the image of the cocycle on all $2$-chains and its image on $2$-cycles.
\end{proposition}

\begin{proof}
If $c\in C_3(N(\mathcal U),\mathbb Z)$, then
\[
\widetilde\alpha(\partial c)=(\delta\alpha)(c)=0
\]
because $\alpha$ is a cocycle. Hence $\widetilde\alpha$ vanishes on $B_2(N(\mathcal U),\mathbb Z)$ and induces a homomorphism
\[
\overline\alpha:H_2(N(\mathcal U),\mathbb Z)=Z_2/B_2\to A.
\]
This proves \textup{(i)}.

For any $2$-cycle $z=\sum_{\sigma} n_{\sigma}\sigma\in Z_2(N(\mathcal U),\mathbb Z)$, set
\[
\alpha^{\sharp}(z):=\sum_{\sigma} n_{\sigma}\alpha(\sigma)=\widetilde\alpha(z).
\]
If $z'$ is another $2$-cycle homologous to $z$, then $z-z'=\partial c$ for some $c\in C_3(N(\mathcal U),\mathbb Z)$. By \textup{(i)},
\[
\alpha^{\sharp}(z)-\alpha^{\sharp}(z')
=
\widetilde\alpha(z-z')
=
\widetilde\alpha(\partial c)
=
0.
\]
Therefore $\alpha^{\sharp}$ depends only on the homology class of $z$ and defines a homomorphism
\[
H_2(N(\mathcal U),\mathbb Z)\to A,\qquad
[z]\longmapsto \alpha^{\sharp}(z).
\]
By the definition of the universal coefficient map $\eta_A$, this homomorphism is exactly the image of the cohomology class $\omega=[\alpha]$ under
\[
\eta_A:H^2(N(\mathcal U),A)\to \operatorname{Hom}(H_2(N(\mathcal U),\mathbb Z),A).
\]
Thus
\[
\overline\alpha([z])=\alpha^{\sharp}(z)=\eta_A(\omega)([z])=\mathrm{ev}_{\omega}([z])
\]
for every $[z]\in H_2(N(\mathcal U),\mathbb Z)$, proving \textup{(ii)}.

For \textup{(iii)}, the image of $\widetilde\alpha$ is exactly the subgroup generated by the simplex values $\alpha_{ijk}$, hence it is $H_\alpha$ by \Cref{def:finite-obstruction-presentation}. The equality
\[
\mathrm{Im}(\overline\alpha)=\widetilde\alpha(Z_2(N(\mathcal U),\mathbb Z))
\]
is immediate from \textup{(i)}, and \Cref{def:intrinsic-visible-quotient} identifies this image with $K_{\omega}$.
\end{proof}

\begin{corollary}[Criterion for coincidence of strict and intrinsic visibility]\label{cor:strict-intrinsic-coincidence-criterion}
Let $\alpha$ be a finite obstruction presentation of $\omega$. The following are equivalent:
\begin{enumerate}[label=(\roman*),nosep]
\item $A/H_\alpha\cong A_{\mathrm{vis}}^{\omega}$;
\item $H_\alpha=K_{\omega}$;
\item $\widetilde\alpha(C_2(N(\mathcal U),\mathbb Z))=\widetilde\alpha(Z_2(N(\mathcal U),\mathbb Z))$.
\end{enumerate}
Equivalently, strict visibility agrees with intrinsic visibility exactly when every presentation-level visible value already arises from the evaluation of the cocycle on a genuine $2$-cycle.
\end{corollary}

\begin{proof}
The equivalence of \textup{(i)} and \textup{(ii)} is immediate from the quotient descriptions
\[
A/H_\alpha
\qquad\text{and}\qquad
A_{\mathrm{vis}}^{\omega}=A/K_{\omega}.
\]
The equivalence of \textup{(ii)} and \textup{(iii)} is exactly \Cref{prop:chain-vs-homological-visibility}(iii).
\end{proof}

\begin{theorem}[Character-blind pure-extension obstructions]\label{thm:character-blind-obstructions}
In the setting of \Cref{def:finite-nerve-presentation}, the following are equivalent:
\begin{enumerate}[label=(\roman*),nosep]
\item $K_{\omega}=0$;
\item $A_{\mathrm{vis}}^{\omega}=A$;
\item $\mathrm{ev}_{\omega}=0$;
\item $\omega$ lies in the subgroup
\[
\operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),A)\subseteq H^2(N(\mathcal U),A).
\]
\end{enumerate}
If these conditions hold and $\omega\neq 0$, then the chosen branch gerbe $\mathfrak G_v$ is non-neutral, while every character of $A$ is class-admissible for that branch. In particular, if $|\Vis_{p,s}(r)|=1$, then
\[
M,p\Vdash \Null^{\mathrm{glue}}_s(r).
\]
\end{theorem}

\begin{proof}
By \Cref{def:intrinsic-visible-quotient},
\[
K_{\omega}=\mathrm{Im}(\mathrm{ev}_{\omega}),
\]
so \textup{(i)} and \textup{(iii)} are equivalent. Since $A_{\mathrm{vis}}^{\omega}=A/K_{\omega}$, conditions \textup{(i)} and \textup{(ii)} are equivalent. By the universal coefficient sequence, \textup{(iii)} holds exactly when $\omega\in\ker(\eta_A)$, and this kernel is the stated $\operatorname{Ext}$-subgroup. Hence all four conditions are equivalent.

If they hold and $\omega\neq 0$, then \Cref{prop:derived-branch-classification} shows that the branch gerbe $\mathfrak G_v$ is non-neutral.
On the other hand, $\mathrm{ev}_{\omega}=0$ implies
\[
\chi\circ \mathrm{ev}_{\omega}=0
\]
for every $\chi\in\widehat A$, so \Cref{thm:intrinsic-visible-quotient} shows that every character is class-admissible.

Assume now that $|\Vis_{p,s}(r)|=1$. Since the chosen branch $v$ exists, \Cref{prop:compsecs-vs-visible-branches} gives
\[
M,p\Vdash \CompSec_s(r).
\]
Suppose towards a contradiction that
\[
M,p\Vdash \Sec_s(r).
\]
By \Cref{def:realization-prestack}(i), this yields an object of the terminal fibre $(\mathfrak P_r)_a$, and its image under the stackification morphism
\[
\iota_r:\mathfrak P_r\to \mathfrak L_r
\]
defines an object of $\mathfrak L_r(a)$. Because there is only one visible branch, every object of $\mathfrak L_r(a)$ lies in the unique component gerbe $\mathfrak G_v(a)$. Hence
\[
\mathfrak G_v(a)\neq\varnothing.
\]
By \Cref{lem:terminal-fibre-neutrality}, this implies that $\mathfrak G_v$ is neutral, contradicting the non-neutrality already proved. Therefore
\[
M,p\nVdash \Sec_s(r).
\]
Combining this with $M,p\Vdash \CompSec_s(r)$ and admittedness of $r$, we obtain
\[
M,p\Vdash \Null^{\mathrm{glue}}_s(r).
\]
\end{proof}

\begin{corollary}[Automatic blindness when $H_2$ vanishes]\label{cor:h2-vanishing-blindness}
Assume
\[
H_2(N(\mathcal U),\mathbb Z)=0.
\]
Then, for every obstruction class $\omega\in H^2(N(\mathcal U),A)$,
\[
\mathrm{ev}_{\omega}=0,
\qquad
K_{\omega}=0,
\qquad
A_{\mathrm{vis}}^{\omega}=A.
\]
Every nonzero branch obstruction class on such a nerve is therefore character-blind.
\end{corollary}

\begin{proof}
If $H_2(N(\mathcal U),\mathbb Z)=0$, then
\[
\mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),A)=0,
\]
so $\mathrm{ev}_{\omega}=0$ for every $\omega$. Apply \Cref{thm:character-blind-obstructions}.
\end{proof}

\begin{corollary}[No character-blind obstructions when $H_1$ is free]\label{cor:h1-free-no-blindness}
Assume that $H_1(N(\mathcal U),\mathbb Z)$ is free abelian. Then
\[
K_{\omega}=0
\quad\Longleftrightarrow\quad
\omega=0.
\]
Equivalently, every nonzero branch obstruction class has a nontrivial $H_2$-visible part. In particular, if $|\Vis_{p,s}(r)|=1$, every nontrivial gluing obstruction on such a nerve produces a proper intrinsic visible quotient.
\end{corollary}

\begin{proof}
If $H_1(N(\mathcal U),\mathbb Z)$ is free abelian, then
\[
\operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),A)=0.
\]
Hence the universal coefficient map $\eta_A$ is injective. By \Cref{thm:character-blind-obstructions},
\[
K_{\omega}=0
\quad\Longleftrightarrow\quad
\mathrm{ev}_{\omega}=0
\quad\Longleftrightarrow\quad
\eta_A(\omega)=0.
\]
Injectivity of $\eta_A$ therefore gives $K_{\omega}=0$ if and only if $\omega=0$. If $\omega\neq 0$, then $K_{\omega}\neq 0$, so $A_{\mathrm{vis}}^{\omega}=A/K_{\omega}$ is a proper quotient of the finite group $A$.
\end{proof}

\begin{example}[A character-blind obstruction on an $\mathbb{R}P^2$-type nerve]\label{ex:rp2-character-blind}
Let $\mathcal U$ be a finite cover whose \v{C}ech nerve is a finite triangulation of $\mathbb{R}P^2$, and take
\[
A=\mathbb Z/2\mathbb Z.
\]
Then
\[
H_2(\mathbb{R}P^2,\mathbb Z)=0,
\qquad
H_1(\mathbb{R}P^2,\mathbb Z)\cong \mathbb Z/2\mathbb Z
\]
\cite[Sec.~2.2]{Hatcher2002}. The universal coefficient sequence therefore yields
\[
H^2(\mathbb{R}P^2,\mathbb Z/2\mathbb Z)
\cong
\operatorname{Ext}^1(\mathbb Z/2\mathbb Z,\mathbb Z/2\mathbb Z)
\cong
\mathbb Z/2\mathbb Z
\]
\cite[Sec.~3.1]{Hatcher2002}. Let $\omega$ be the nonzero class. Then
\[
\mathrm{ev}_{\omega}=0,
\qquad
K_{\omega}=0,
\qquad
A_{\mathrm{vis}}^{\omega}=A,
\]
by \Cref{thm:character-blind-obstructions}. Thus no class-level quotient collapse occurs at all, yet the obstruction class is nonzero. Hence this branch is completely character-blind. If it is the unique visible branch of the admitted reference under consideration, then \Cref{thm:gerbe-null-semantics} yields a genuine gluing-level absence that is invisible to every character channel.
\end{example}

\begin{example}[A single-branch maximal-collapse obstruction on $S^2$]\label{ex:s2-single-branch-maximal-collapse}
Let $\mathcal U=\{U_i\to S^2\}_{i\in I}$ be a finite good cover of $S^2$, let $N:=N(\mathcal U)$ be its nerve, and let $\mathcal C_a$ be the corresponding slice site for one admitted reference $r:\mathsf{Ref}_s$ with $a=[r]_p$. Take the constant band
\[
A=\mathbb Z/2\mathbb Z
\]
on $\mathcal C_a$. Define the local object functor
\[
F_{p,s}|_{\mathcal C_a}:\mathcal C_a^{op}\to\mathsf{Set}
\]
by
\[
F_{p,s}(U)=\{v\}
\]
for every nonempty proper finite intersection $U$ of members of $\mathcal U$, by
\[
F_{p,s}(a)=\varnothing,
\]
and by the identity map on every restriction between proper intersections. Then the cover
\[
\mathcal U=\{U_i\to a\}_{i\in I}
\]
supports the compatible family constantly equal to $v$, while there is no global section at $a$. Hence
\[
M,p\Vdash \CompSec_s(r),
\qquad
M,p\nVdash \Sec_s(r).
\]
Because
\[
H^2(N,A)\cong \mathbb Z/2\mathbb Z,
\]
choose a simplicial cocycle
\[
\alpha\in Z^2(N,A)
\]
representing the nonzero class. For each $U_i$ choose a local realization $x_i$ with automorphism group $A$, and choose isomorphisms
\[
\phi_{ij}:x_i|_{U_{ij}}\xrightarrow{\sim} x_j|_{U_{ij}}
\]
on the pairwise overlaps such that
\[
\phi_{ik}^{-1}\phi_{jk}\phi_{ij}=\alpha_{ijk}\in A
\]
on each triple overlap. The standard cocycle-to-gerbe construction for local trivializations \cite[Chap.~IV]{Giraud1971}, \cite[Sec.~5]{Breen2006}, \cite{Murray1996} then yields a prestack presentation $\mathfrak P_r$ on $\mathcal C_a$ with the following properties:
\begin{enumerate}[label=(\roman*),nosep]
\item over each member $U_i$ of the cover there is the chosen local object $x_i$ with automorphism group $A$;
\item the chosen transition maps have triple-overlap defect exactly $\alpha$;
\item the presheaf of isomorphism classes satisfies
\[
 \pi_0^{\mathrm{pre}}(\mathfrak P_r)\cong F_{p,s}|_{\mathcal C_a}.
\]
\end{enumerate}
Thus the example begins with an explicit local object functor and then chooses local realization and transition data realizing that functor; the branch gerbe is recovered from those data by stackification rather than being taken as primitive input.

Let
\[
\mathfrak L_r:=a(\mathfrak P_r).
\]
Because $\pi_0^{\mathrm{pre}}(\mathfrak P_r)\cong F_{p,s}|_{\mathcal C_a}$, the sheaf of connected components of $\mathfrak L_r$ has exactly one global section at $a$, represented by the constant compatible family above. Therefore
\[
\Vis_{p,s}(r)=\{v\}.
\]
But $\alpha$ is nonzero, so the descent datum is ineffective and no object exists over $a$ in the resulting stack; equivalently, the unique component gerbe is non-neutral. Hence
\[
\mathfrak G_v=\mathfrak L_r,
\qquad
\omega_v=[\alpha]\neq 0.
\]
Combining this with the already computed presheaf-level facts gives
\[
M,p\Vdash \Null^{\mathrm{glue}}_s(r),
\qquad
M,p\nVdash \Sec_s(r),
\qquad
\Vis_{p,s}(r)=\{v\},
\qquad
\mathfrak G_v=\mathfrak L_r,
\qquad
\omega_v=[\alpha]\neq 0.
\]
The good cover $\mathcal U$ trivializes the constant band, all finite intersections are contractible, and therefore the $H^1$-vanishing and \v{C}ech-to-derived comparison hypotheses of \Cref{def:cech-gluing-obstruction,def:finite-nerve-presentation} are satisfied.

Since
\[
H_1(N,\mathbb Z) = 0,
\qquad
H_2(N,\mathbb Z) \cong \mathbb Z
\]
\cite[Secs.~2.1, 2.2]{Hatcher2002}, the universal coefficient theorem gives
\[
H^2(N, \mathbb{Z}/2\mathbb{Z})
\cong
\mathrm{Hom}(\mathbb{Z}, \mathbb{Z}/2\mathbb{Z})
\cong
\mathbb{Z}/2\mathbb{Z}.
\]
Let $\omega$ be the nonzero class, and let $[S^2]$ denote the fundamental class in $H_2(N,\mathbb Z)$. Then
\[
\mathrm{ev}_\omega([S^2]) = 1 \in \mathbb{Z}/2\mathbb{Z},
\]
so the homological evaluation map is surjective. By \Cref{cor:maximal-collapse},
\[
K_\omega = A = \mathbb{Z}/2\mathbb{Z},
\qquad
A_{\mathrm{vis}}^\omega = 0.
\]
Hence
\[
M,p\Vdash \Null^{\mathrm{glue}}_s(r),
\qquad
\omega=\omega_v\neq 0,
\qquad
\mathrm{ev}_{\omega}:H_2(N,\mathbb Z)\twoheadrightarrow A,
\qquad
K_\omega=A,
\qquad
A_{\mathrm{vis}}^\omega=0.
\]
Since $K_\omega = A$, \Cref{thm:intrinsic-visible-quotient} gives
\[
X_\omega = \mathrm{Ann}(K_\omega) = \mathrm{Ann}(A) = \{1\},
\]
so only the trivial character is class-admissible. Equivalently, $A_{\mathrm{vis}}^\omega = A/K_\omega = 0$.

This example therefore verifies the full downstream hypothesis chain
\[
(\mathcal C_a,\mathfrak P_r,A,\omega,\mathrm{ev}_\omega,K_\omega,A_{\mathrm{vis}}^\omega)
\]
inside a concrete site-theoretic example, once the realization stack has been fixed by the chosen cocycle data. It is the opposite extreme from the $\mathbb{R}P^2$ blind case of \Cref{ex:rp2-character-blind}: for an $\mathbb{R}P^2$-type nerve one has $K_\omega=0$, every character is class-admissible, and the obstruction is purely $\operatorname{Ext}$-type with $A_{\mathrm{vis}}^\omega=A$; for the present $S^2$-type nerve one has $K_\omega=A$, only the trivial character is class-admissible, and the obstruction is purely $H_2$-type with $A_{\mathrm{vis}}^\omega=0$.
\end{example}

\begin{definition}[Intrinsic strictification code]\label{def:intrinsic-strictification-code}
An \emph{intrinsic strictification code} for the obstruction class $\omega$ consists of a finite set $R$ and an injective map
\[
\iota_{\omega}:A\hookrightarrow A_{\mathrm{vis}}^{\omega}\times R
\]
such that
\[
\mathrm{pr}_1\circ\iota_{\omega}=\pi_{\omega}.
\]
\end{definition}

\begin{corollary}[Intrinsic hidden-state lower bound]\label{cor:intrinsic-hidden-state-lower-bound}
Every intrinsic strictification code satisfies
\[
|R|\ge |K_{\omega}|.
\]
Equivalently,
\[
\log_2|R|\ge \log_2|K_{\omega}|.
\]
This lower bound is sharp.
\end{corollary}

\begin{proof}
By \Cref{def:intrinsic-visible-quotient}, the fibers of $\pi_{\omega}$ are exactly the cosets of $K_{\omega}$, so every fiber has cardinality $|K_{\omega}|$. The same fiber-counting argument as in \Cref{thm:minimal-information-cost-of-strictification} therefore shows that any injective code preserving the first projection must satisfy $|R|\ge |K_{\omega}|$.

For sharpness, take $R:=K_{\omega}$ and choose a set-theoretic section $s_{\omega}:A_{\mathrm{vis}}^{\omega}\to A$ of $\pi_{\omega}$. Then
\[
\iota_{\omega}(x):=\bigl(\pi_{\omega}(x),x-s_{\omega}(\pi_{\omega}(x))\bigr)
\]
is injective.
\end{proof}

\begin{proposition}[Prime decomposition of intrinsic visibility and budget]\label{thm:intrinsic-prime-decomposition}
Write the finite abelian group $A$ as a direct sum of its $p$-primary components:
\[
A=\bigoplus_p A_{(p)}.
\]
Let $\omega_p\in H^2(N(\mathcal U),A_{(p)})$ be the image of $\omega$ under the projection onto the $p$-primary summand. Then
\[
\mathrm{ev}_{\omega}=\bigoplus_p \mathrm{ev}_{\omega_p},
\qquad
K_{\omega}=\bigoplus_p K_{\omega,p},
\qquad
A_{\mathrm{vis}}^{\omega}\cong \bigoplus_p A_{(p)}/K_{\omega,p},
\]
where
\[
K_{\omega,p}:=\mathrm{Im}(\mathrm{ev}_{\omega_p}).
\]
Consequently,
\[
|K_{\omega}|=\prod_p |K_{\omega,p}|,
\qquad
\log_2|K_{\omega}|=\sum_p \log_2|K_{\omega,p}|.
\]
\end{proposition}

\begin{proof}
Both $H^2(N(\mathcal U),-)$ and $\mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),-)$ commute with finite direct sums, so the universal coefficient map is compatible with the primary decomposition of $A$. Hence
\[
\eta_A(\omega)=\bigoplus_p \eta_{A_{(p)}}(\omega_p),
\]
which is exactly the asserted decomposition of $\mathrm{ev}_{\omega}$.

Taking images gives
\[
K_{\omega}
=
\mathrm{Im}(\mathrm{ev}_{\omega})
=
\bigoplus_p \mathrm{Im}(\mathrm{ev}_{\omega_p})
=
\bigoplus_p K_{\omega,p},
\]
by \Cref{def:intrinsic-visible-quotient}. The quotient decomposition follows immediately:
\[
A_{\mathrm{vis}}^{\omega}
=
A/K_{\omega}
\cong
\bigoplus_p A_{(p)}/K_{\omega,p}.
\]
Since finite direct sums correspond to products of cardinalities, the budget formula is immediate.
\end{proof}

\begin{corollary}[Intrinsic budget versus strict budget]\label{cor:strict-intrinsic-budget-comparison}
For every finite obstruction presentation $\alpha$ of the class $\omega$,
\[
|K_{\omega}|\le |H_\alpha|,
\qquad
\log_2|K_{\omega}|\le \log_2|H_\alpha|.
\]
Hence the intrinsic hidden-state budget never exceeds the strict presentation-level budget. Equality holds if and only if the canonical surjection
\[
A_{\mathrm{vis}}^{\omega}\twoheadrightarrow A/H_\alpha
\]
is an isomorphism.
\end{corollary}

\begin{proof}
By \Cref{cor:strict-to-intrinsic-visible}, there is a surjective homomorphism
\[
A_{\mathrm{vis}}^{\omega}=A/K_{\omega}\twoheadrightarrow A/H_\alpha.
\]
For finite groups this implies
\[
|A|/|K_{\omega}|\ge |A|/|H_\alpha|,
\]
hence $|K_{\omega}|\le |H_\alpha|$. The logarithmic inequality is immediate. Equality of cardinalities holds exactly when the displayed surjection is bijective.
\end{proof}

\begin{corollary}[Maximal collapse criterion]\label{cor:maximal-collapse}
In the setting of \Cref{def:finite-nerve-presentation}, the following are equivalent:
\begin{enumerate}[label=(\roman*),nosep]
\item $\mathrm{ev}_{\omega}:H_2(N(\mathcal U),\mathbb Z)\to A$ is surjective;
\item $K_{\omega}=A$;
\item $A_{\mathrm{vis}}^{\omega}=0$;
\item the intrinsic strictification budget is maximal:
\[
\log_2|K_{\omega}|=\log_2|A|.
\]
\end{enumerate}
\end{corollary}

\begin{proof}
By \Cref{def:intrinsic-visible-quotient}, $K_{\omega}=\mathrm{Im}(\mathrm{ev}_{\omega})$, so \textup{(i)} and \textup{(ii)} are equivalent. Condition \textup{(ii)} is equivalent to \textup{(iii)} by the definition of $A_{\mathrm{vis}}^{\omega}=A/K_{\omega}$. For finite groups, \textup{(ii)} is also equivalent to $|K_{\omega}|=|A|$, which is exactly \textup{(iv)} after taking $\log_2$.
\end{proof}
