\documentclass{amsart}

\usepackage{amsmath,amssymb,amsthm,mathtools}
\IfFileExists{stmaryrd.sty}{%
  \usepackage{stmaryrd}%
}{%
  \newcommand{\llbracket}{\mathopen{[\![}}%
  \newcommand{\rrbracket}{\mathclose{]\!]}}%
}
\usepackage{hyperref}
\usepackage{cleveref}
\newcommand{\printreferences}{\bibliographystyle{amsplain}\bibliography{references}}
\emergencystretch=2em

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\DeclareMathOperator{\eml}{eml}
\DeclareMathOperator{\Fold}{Fold}
\DeclareMathOperator{\Val}{Val}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\rank}{rank}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\one}{\mathbf{1}}
\newcommand{\dom}{\operatorname{dom}}

\title[A Single-Primitive Universality Hierarchy]
{A Three-Level Hierarchy for Single-Primitive Universality}

\author[The Omega Institute]{The Omega Institute}
\date{}
\subjclass[2020]{Primary 03D35; Secondary 11B39, 68Q42, 47A57}
\keywords{single primitive, Zeckendorf representation, canonical forms,
moment problem, Richardson theorem}

\begin{document}
\begin{abstract}
We distinguish three kinds of data that are often conflated under
the phrase single-primitive universality.  Expression universality
requires only that a prescribed family of functions have terms over
one operation; Odrzywo{\l}ek's operator
\(\eml(x,y)=e^x-\operatorname{Log}y\) is the model example.
Protocol universality adds computable normal forms and readouts.
For Fibonacci words, Zeckendorf normalization and two levels of
composition recover addition and multiplication on \(\NN\), and a
level-zero rigidity theorem shows that the second composition level
is forced for multiplicative readout on a free monogenic orbit.
Finite certificate data are then attached to normalization fibers:
collision moments form positive Hankel matrices, and finite initial
segments recover the fiber-size spectrum by Prony reconstruction.
These bare finite-fiber certificates are formal; the substantive
strengthening is finite-state transfer computation.  We prove that
a computable finite-fiber enlargement with the same Zeckendorf
normal forms and readouts need not admit any fixed finite-state
transfer realization.
At the expression level, Richardson's theorem rules out a
computable canonical form for analytic equivalence in the full
elementary EML language.  Thus a one-node expression grammar does
not by itself supply a decidable protocol.
\end{abstract}

\maketitle

\input{sections/sec01_introduction}
\input{sections/sec02_expression_universality}
\input{sections/sec03_protocol_universality}
\input{sections/sec04_certificate_universality}
\input{sections/sec05_hierarchy_strictness}
\input{sections/sec06_discussion}

\printreferences

\end{document}
