\section{A bounded-delay transducer model for the finite-window fold}
\label{app:transducer}

To make the finite-state kernel construction of
\Cref{thm:collision-kernel,thm:symmetric-compression} self-contained, we spell
out the transducer-level mechanism behind it.  The only external input is the
classical existence of a deterministic subsequential transducer for Fibonacci
normalization of finite words read from the most significant digit to the least;
see \cite{Frougny1992NumbersAutomata,Frougny1999OnlineAddition}.  For a broader
automata-theoretic framework for certifying Fibonacci representation systems,
see Shallit and Shan~\cite{ShallitShan2023}.

For \(\omega=(\omega_1,\dots,\omega_m)\in\Omega_m\), write
\[
  \omega^{\mathrm{rev}}:=\omega_m\cdots \omega_1
\]
for the corresponding most-significant-digit-first input word.  Fix once and
for all a deterministic subsequential transducer
\[
  \mathcal T=(Q,q_0,\delta,\lambda,\tau)
\]
with input alphabet \(\{0,1\}\), output alphabet \(\{0,1\}\), transition map
\(\delta\), transition-output map \(\lambda\colon Q\times\{0,1\}\to\{0,1\}^*\),
and terminal-output map \(\tau\colon Q\to\{0,1\}^*\), such that the total
output
\[
  \Lambda(a_1\cdots a_m)
  :=
  \lambda^*(q_0,a_1\cdots a_m)\,
  \tau(\delta^*(q_0,a_1\cdots a_m))
\]
is the \((m+1)\)-digit left-padded Zeckendorf normalization of the input word
\(a_1\cdots a_m\), written in the same most-significant-digit-first
orientation.  Thus leading zeroes are part of the output convention.  Set
\[
  L:=\max_{s\in Q}|\tau(s)|.
\]
For the certificate identities in Appendix~\ref{app:certification},
\(\mathcal T\) is treated as a finite transition table.  No minimality of this
table is required: any deterministic subsequential Fibonacci normalizer with the
stated input-output property gives the same collision numbers \(S_q(m)\).  Once
the table is fixed, the histogram kernel is the explicitly defined integer
matrix of \Cref{def:exact-histogram-kernel}; the later recurrence certificates
are identities in that matrix.  Thus the computational object used in
Appendix~\ref{app:certification} is not the scalar list
\((S_q(m))_m\), but the finite transition table of \(\mathcal T\) and the
integer matrix reconstructed from it.

\begin{proposition}[Pinned finite table for the certificate layer]
\label{prop:pinned-normalizer-certificate-table}
The arithmetic-window certificates use one fixed finite transition table, not a
family of fitted scalar data.  The table and the remaining exact arithmetic
records are archived in the paper-local file
\[
  \texttt{certificates/arithmetic\_window\_certificates.json}
\]
with SHA-256 hash
\[
\begin{gathered}
\texttt{712704ff15acbdbb1cb05299c6b32802}\\
\texttt{6818edea4b364878e839eaec39884068},
\\
\text{payload hash }
\texttt{531d8a9f1e8dbb5ae8045d463b619ba5}\\
\texttt{fe650ee5c070936641eb8d3b77020b61}.
\end{gathered}
\]
Under Python with SymPy installed, it is regenerated by
\[
  \texttt{python3 certificates/verify\_arithmetic\_window.py},
\]
whose SHA-256 hash is
\[
\begin{gathered}
  \texttt{87cf1ac098d651f1f46b0749e8497633}\\
  \texttt{7dfc0aa27879fd119806108e1ce79b98}.
\end{gathered}
\]
The field \(\texttt{fixed\_normalizer\_transition\_table}\) is the finite
Frougny carry table from which the verifier reconstructs the binary padded
normalizer of \Cref{thm:pinned-padded-normalizer-soundness}; the one-copy state
updates \(T_a\), committed outputs \(C_a\), and terminal-flush convention used
in the histogram kernels of \Cref{def:exact-histogram-kernel} are the binary
restriction \(a\in\{0,1\}\) of that reconstructed normalizer.  The same file
stores the exact supplementary records
\(\texttt{normalizer\_local\_verification}\) and
\(\texttt{exact\_moment\_window\_records}\) used later for the Hankel and Smith
arguments.
\end{proposition}

\begin{proof}
The assertion is a reproducibility statement about a finite object.  The JSON
file lists the state names, input and output alphabets, initial state, and every
transition of the fixed table.  The verifier script contains the same table in
source form, rebuilds the canonical JSON payload, checks the local carry
identities of the pinned table, reconstructs the exact finite moment-window and
Hankel records used later in the appendix, and verifies the Smith,
finite-field, and discriminant records before writing the payload hash
displayed above.  Since SHA-256 hashes are computed from the byte strings of
the script and the generated JSON file, the displayed hashes pin the exact
finite table and supplementary records used by the rest of the appendix.
\end{proof}

\begin{theorem}[Soundness of the hash-pinned padded normalizer]
\label{thm:pinned-padded-normalizer-soundness}
Let \(\mathcal C\) be the finite carry table stored in the pinned certificate
field \(\texttt{fixed\_normalizer\_transition\_table}\).  Its states are the ten
signed triples
\[
\begin{gathered}
000,\ 001,\ 002,\ 010,\ 100,\ 101,\\
0{-}12,\ 1{-}12,\ 01{-}1,\ 11{-}1,
\end{gathered}
\]
with initial state \(000\), input alphabet \(\{0,1,2\}\), and output alphabet
\(\{0,1\}\).  The deterministic binary transducer \(\mathcal T\) used in this
appendix is the binary-input padded normalizer obtained from this table by
Frougny's finite carry construction with the terminal zero flush and leading
delay convention recorded in the certificate schema.  For every binary word
\(w=a_1\cdots a_m\), \(\Lambda(w)\) has length \(m+1\), contains no adjacent
ones, and satisfies
\[
  \sum_{i=1}^m a_iF_{m+2-i}
  =
  \sum_{i=0}^m \Lambda(w)_{i+1}F_{m+2-i}.
\]
Consequently \(\Lambda(w)\) is the unique \((m+1)\)-digit padded Zeckendorf
normal form of \(w\).
\end{theorem}

\begin{proof}
The theorem is a convention translation from Frougny's standard Fibonacci
normalizer to the pinned table used in this paper.  The local part that depends
on our notation is finite.  Write a state
\(s\) as a signed triple \((\alpha,\beta,\gamma)\).  Each recorded edge
\[
  s\xrightarrow[d/e]{}s',
  \qquad
  s'=(\alpha',\beta',\gamma'),
\]
satisfies the Fibonacci carry identity
\[
  \alpha F_{n+3}+\beta F_{n+2}+\gamma F_{n+1}+dF_n
  =
  eF_{n+3}+\alpha'F_{n+2}+\beta'F_{n+1}+\gamma'F_n
  \qquad(n\ge2).
\]
For example, the edge
\[
  002\xrightarrow[0/0]{}11{-}1
\]
is the identity
\[
  2F_{n+1}=F_{n+2}+F_{n+1}-F_n.
\]
The other recorded edges are checked in exactly the same finite way.  The file
\texttt{certificates/arithmetic\_window\_certificates.json} records this check
for all \(30\) edges in the field
\(\texttt{normalizer\_local\_verification}\): after reducing both sides to the
basis \((F_{n+1},F_n)\), the verifier checks the resulting coefficient pairs
exactly.  Hence every carry pass preserves the represented integer after
accounting for the three pending registers and the terminal zero flush.

The remaining assertion is the standard output-safety conclusion of Frougny's
normalization theorem
\cite{Frougny1992NumbersAutomata,Frougny1999OnlineAddition}: for this
deterministic carry table, after restricting to binary input, appending the
terminal zero flush, and keeping the one-symbol leading delay convention, the
resulting output is the unique Zeckendorf normalization of the same integer.
In the present paper the only extra convention is the fixed left padding by one
digit.  That padding is precisely the leading-delay symbol produced by the
terminal flush, so the output has length \(m+1\).  Thus the reconstructed
binary transducer emits a binary word of length \(m+1\) with no adjacent ones
and with the same represented value as the input.  By the uniqueness part of
Zeckendorf's theorem, this word is the padded Zeckendorf normal form.
\end{proof}

\begin{theorem}[Fixed-width overflow deletion]
\label{thm:appendix-fixed-width-overflow}
For every \(\omega\in\Omega_m\), let \(x=\Fold_m(\omega)\).  Then there exists
\(\varepsilon_m(\omega)\in\{0,1\}\) such that
\[
  \Lambda(\omega^{\mathrm{rev}})
  =
  \varepsilon_m(\omega)\,x_m\cdots x_1.
\]
The word on the right has exactly \(m+1\) symbols.  Hence deleting its first
symbol recovers the length-\(m\) visible fold in reversed order, with leading
zeroes retained.  For \(m=0\), this says
\(\Lambda(\varepsilon)=0\) and deletion gives the empty word.  For \(m=1\),
\[
  \Lambda(0)=00,
  \qquad
  \Lambda(1)=01.
\]
\end{theorem}

\begin{proof}
The cases \(m=0\) and \(m=1\) follow directly from the padded convention and
from \(F_2=1\).  Assume now that \(m\ge2\).  By
\Cref{thm:pinned-padded-normalizer-soundness}, \(\Lambda\) is the padded
Zeckendorf normalizer.  Since
\[
  N_m(\omega)
  =
  \sum_{j=1}^m \omega_jF_{j+1}
  =
  \sum_{i=1}^m \omega_{m+1-i}F_{m+2-i},
\]
the reversed word \(\omega^{\mathrm{rev}}\) represents \(N_m(\omega)\) in
most-significant-digit-first order.  By \Cref{prop:single-overflow}, there are
unique \(\varepsilon_m(\omega)\in\{0,1\}\) and \(0\le r<F_{m+2}\) such that
\[
  N_m(\omega)
  =
  \varepsilon_m(\omega)F_{m+2}+r
  .
\]
By definition of the finite-window fold,
\[
  r=V_m(\Fold_m(\omega))=V_m(x)=\sum_{j=1}^m x_jF_{j+1}.
\]
Hence
\[
  N_m(\omega)
  =
  \varepsilon_m(\omega)F_{m+2}+\sum_{j=1}^m x_jF_{j+1}.
\]
If \(\varepsilon_m(\omega)=1\), then
\[
  r=N_m(\omega)-F_{m+2}<F_{m+1},
\]
because \(N_m(\omega)\le F_{m+3}-2\).  Therefore the top visible digit of
\(x\) vanishes: \(x_m=0\).  Hence the word
\(\varepsilon_m(\omega)x_m\cdots x_1\) is admissible.
Therefore it is the padded Zeckendorf normalization of \(N_m(\omega)\), which is
exactly \(\Lambda(\omega^{\mathrm{rev}})\).
\end{proof}

\begin{lemma}[One-digit overflow decomposition]
\label{lem:appendix-overflow-digit}
For every \(\omega\in\Omega_m\), let \(x=\Fold_m(\omega)\).  Then there exists
\(\varepsilon_m(\omega)\in\{0,1\}\) such that
\[
  \Lambda(\omega^{\mathrm{rev}})
  =
  \varepsilon_m(\omega)\,x_m\cdots x_1.
\]
In particular, after deleting its first symbol one recovers the visible word of
the finite-window fold in reversed order.
\end{lemma}

\begin{proof}
This is \Cref{thm:appendix-fixed-width-overflow}.
\end{proof}

\begin{definition}[Visible truncation of an output word]
\label{def:appendix-visible-truncation}
For \(y\in\{0,1\}^*\), let \(\operatorname{vis}(y)\) be the word obtained from
\(y\) by deleting its first symbol when \(|y|\ge 1\), and let
\(\operatorname{vis}(y):=\varepsilon\) when \(y\) is empty.
\end{definition}

\begin{lemma}[Prefix stability for visible outputs]
\label{lem:appendix-prefix-stability}
If \(u\) is a prefix of \(v\) in \(\{0,1\}^*\), then
\(\operatorname{vis}(\Lambda(u))\) and \(\operatorname{vis}(\Lambda(v))\) have
a common prefix of length at least
\[
  |\operatorname{vis}(\Lambda(u))|-(L+1).
\]
Equivalently, after any processed prefix, at most the final \(L+1\) visible
symbols can still be altered by extending the input.
\end{lemma}

\begin{proof}
Write \(v=uw\), let \(s=\delta^*(q_0,u)\), and decompose the run as
\[
  \Lambda(u)=\alpha\,\tau(s),
  \qquad
  \Lambda(v)=\alpha\,\beta\,\tau(t)
\]
for some word \(\beta\) and some state \(t\).  Thus \(\Lambda(u)\) and
\(\Lambda(v)\) have a common prefix of length
\[
  |\alpha|=|\Lambda(u)|-|\tau(s)|\ge |\Lambda(u)|-L.
\]
Deleting the first symbol from each word shortens the common prefix by at most
one, so the visible truncations share a prefix of length at least
\[
  |\Lambda(u)|-L-1=|\operatorname{vis}(\Lambda(u))|-(L+1).
\]
\end{proof}

\begin{definition}[Committed and residual visible output]
\label{def:appendix-cut}
For a finite word \(y\in\{0,1\}^*\), write
\[
  y=\operatorname{com}(y)\,\operatorname{res}(y),
\]
where \(\operatorname{res}(y)\) is the suffix of \(y\) of length at most
\(L+1\), and \(\operatorname{com}(y)\) is the complementary prefix.
\end{definition}

\begin{proposition}[Finite collision automaton]
\label{prop:appendix-collision-automaton}
Fix \(q\ge 2\).  Set
\[
  \widehat Q:=Q\times\{0,1\}^{\le L+1}\times\{0,1\}.
\]
There is a finite automaton on \(\widehat Q^q\) with one transition per input
digit such that, for every \(m\ge 0\), accepted length-\(m\) paths are in
canonical bijection with \(q\)-tuples
\[
  (\omega^{(1)},\dots,\omega^{(q)})\in\Omega_m^q
\]
satisfying
\[
  \Fold_m(\omega^{(1)})=\cdots=\Fold_m(\omega^{(q)}).
\]
Consequently there exist a nonnegative integer matrix \(A_q\) and nonnegative
integer vectors \(u_q,v_q\) such that
\[
  S_q(m)=u_q^{\mathsf T}A_q^m v_q
  \qquad (m\ge 0).
\]
\end{proposition}

\begin{proof}
For a one-copy augmented state \((s,r,e)\in\widehat Q\), the meaning is:
\(s\) is the current transducer state, \(r\) is the residual visible suffix
that may still change under future extensions, and \(e\in\{0,1\}\) records
whether the initial overflow digit of \Cref{lem:appendix-overflow-digit} has
already been deleted.

Given input digit \(a\in\{0,1\}\), form the temporary word
\[
  w:=r\,\lambda(s,a).
\]
If \(e=0\), delete the first symbol of \(w\) whenever \(|w|\ge 1\), and let
\(e'=1\) precisely when that deletion has occurred.  If \(e=1\), keep \(w\)
unchanged and set \(e'=1\).  Denote the resulting word by \(w_{\mathrm{vis}}\).
Now apply \Cref{def:appendix-cut}:
\[
  w_{\mathrm{vis}}
  =
  c\,r',
\]
with \(|r'|\le L+1\).  The update rule is
\[
  (s,r,e)\xrightarrow{\,a\,}(\delta(s,a),r',e'),
\]
and the newly committed visible block is \(c\).

By \Cref{lem:appendix-prefix-stability}, every continuation of the current input
must begin its visible output with the already committed block \(c\); only the
stored suffix \(r'\) can still change.  Hence, in a synchronized \(q\)-copy
run, a transition labelled by \((a_1,\dots,a_q)\in\{0,1\}^q\) is admissible if
and only if the corresponding committed blocks \(c_1,\dots,c_q\) are all equal.
Whenever they disagree, the visible outputs can never be repaired by a later
continuation, so the run enters a rejecting sink.

After reading \(m\) digits, each copy sits in a state \((s_i,r_i,e_i)\).  The
remaining visible suffix is obtained by appending the terminal word \(\tau(s_i)\)
to \(r_i\), performing the possible initial deletion if \(e_i=0\), and keeping
the resulting word.  Because \(|r_i|\le L+1\) and \(|\tau(s_i)|\le L\), this
terminal suffix has length at most \(2L+1\).  A terminal \(q\)-tuple is
accepting exactly when these final suffixes are identical in all coordinates.

By construction, a synchronized run avoids rejection and ends in an accepting
terminal state if and only if the committed visible blocks agree at every stage
and the remaining unresolved suffixes agree at the end.  This is equivalent to
equality of the full visible words, which are precisely the reversed
finite-window folds by \Cref{lem:appendix-overflow-digit}.  Therefore accepted
paths of length \(m\) are exactly the \(q\)-tuples counted by \(S_q(m)\).

Let \(A_q\) be the adjacency matrix of the non-sink part of this finite
automaton, let \(u_q\) select the unique initial synchronized state
\(((q_0,\varepsilon,0),\dots,(q_0,\varepsilon,0))\), and let \(v_q\) mark the
accepting terminal states.  Then the accepted path count is
\[
  u_q^{\mathsf T}A_q^m v_q=S_q(m).
\]
\end{proof}

The proof of \Cref{thm:symmetric-compression} is now immediate from
\Cref{prop:appendix-collision-automaton}: one simply replaces the one-copy state
set \(Q\) there by the finite augmented state set \(\widehat Q\).  The histogram
quotient therefore acts on weak compositions of \(q\) into \(|\widehat Q|\)
parts, so its dimension remains polynomial in \(q\).
