% ---------------------------------------------------------
% Exact kernel certification and secondary spectral data
% ---------------------------------------------------------
\section{Exact kernel certification and secondary spectral data}
\label{app:certification}

Section~\ref{sec:chebotarev} isolates finite algebraic certificate records for
\(q=9,\dots,17\): row polynomials for \(S_q(m)\), Smith forms, finite-field
factorization patterns, and discriminant squareclass checks.  The recurrence
certificate is the one place where the distinction between an empirical check
and an exact proof matters.  A coefficient row is not evidence for a recurrence
by itself.  It becomes a recurrence only after the polynomial determined by that
row has been substituted into the exact histogram kernel and the resulting
residual vector has been certified to lie in the unobservable subspace.

The proof chain is finite.  The bounded-delay normalizer first gives
\[
  S_q(m)=\widetilde u_q^{\mathsf T}\widetilde A_q^m\widetilde v_q
  \qquad(m\ge0),
\]
where \(\widetilde A_q\) is the integer histogram matrix of
\Cref{def:exact-histogram-kernel}.  Put \(N_q=\dim\widetilde A_q\) and
\[
  W_q=\operatorname{span}_{\QQ}
  \{\widetilde A_q^j\widetilde v_q:j\ge0\}.
\]
For the offset used throughout Section~\ref{sec:chebotarev}, define
\[
  Z_q=
  \{w\in W_q:\widetilde u_q^{\mathsf T}\widetilde A_q^t w=0
    \text{ for every }t\ge2\}.
\]
Cayley--Hamilton gives a finite construction of this subspace from the kernel
itself:
\[
  Z_q=
  \ker\!\left(
  w\mapsto
  (\widetilde u_q^{\mathsf T}\widetilde A_q^2w,\dots,
   \widetilde u_q^{\mathsf T}\widetilde A_q^{N_q+1}w)
  \right)
  \quad\text{inside }W_q.
\]
Thus \(N_q\) is an a priori order bound obtained from the exact kernel dimension
before any candidate recurrence is tested.  This is the point at which the
finite verification becomes sound: the verifier constructs the kernel-defined
unobservable subspace by \(N_q\) observability rows, rather than choosing a
numerical interval in the scalar moment sequence.
If
\[
  R_q=
  \begin{bmatrix}
  \widetilde u_q^{\mathsf T}\widetilde A_q^2\\
  \widetilde u_q^{\mathsf T}\widetilde A_q^3\\
  \vdots\\
  \widetilde u_q^{\mathsf T}\widetilde A_q^{N_q+1}
  \end{bmatrix},
\]
then \(w\in Z_q\) is equivalent to \(R_qw=0\).  For a row polynomial \(P_q\), the
exact recurrence certificate is
\begin{equation}
\label{eq:primary-residual-certificate}
  R_qP_q(\widetilde A_q)\widetilde v_q=0.
\end{equation}
This identity says precisely that
\(P_q(\widetilde A_q)\widetilde v_q\in Z_q\).  Consequently
\[
  \widetilde u_q^{\mathsf T}\widetilde A_q^t
  P_q(\widetilde A_q)\widetilde v_q=0\qquad(t\ge2),
\]
which is the recurrence for every \(m\ge d_q+2\) after putting \(m=t+d_q\).

The rank equalities
\[
  \rk(R_qK_q)=\rk(R_qB_q)=d_q,
\]
where
\[
  K_q=[\widetilde v_q,\dots,\widetilde A_q^{N_q-1}\widetilde v_q],
  \qquad
  B_q=[\widetilde v_q,\dots,\widetilde A_q^{d_q-1}\widetilde v_q],
\]
have a separate role.  They identify the quotient \(W_q/Z_q\), show that the
first \(d_q\) Krylov classes form a cyclic basis, and make \(P_q\) the
characteristic polynomial of the induced quotient operator.  They are useful for
minimality and Hankel-principalization, but the all-tail recurrence already
follows from \eqref{eq:primary-residual-certificate} and the
Cayley--Hamilton order bound.  Consequently every arithmetic theorem in
Section~\ref{sec:chebotarev} uses the row polynomial through the chain
\[
  \widetilde A_q,\widetilde u_q,\widetilde v_q
  \longrightarrow
  R_qP_q(\widetilde A_q)\widetilde v_q=0
  \longrightarrow
  P_q(E)S_q(j)=0\quad(j\ge2),
\]
and not through an extrapolation from finitely many scalar values of \(S_q(m)\).

\subsection{Kernel certificates for the recurrence table}

The row polynomials below have two logically separate parts.  The first part is
the construction of the exact matrix coefficient
\(S_q(m)=\widetilde u_q^{\mathsf T}\widetilde A_q^m\widetilde v_q\).  The second
part is a residual identity after a proposed polynomial has been substituted into
that matrix.  Only the second part certifies a recurrence.  A list of scalar
values \(S_q(0),S_q(1),\dots\), even if very long, is not an input to the
verifier defined here.

\begin{definition}[Exact histogram kernel]
\label{def:exact-histogram-kernel}
Fix the deterministic normalizing transducer \(\mathcal T\) of
Appendix~\ref{app:transducer}, and let \(\widehat Q\) be the one-copy augmented
state set of \Cref{prop:appendix-collision-automaton}.  For
\(s\in\widehat Q\) and \(a\in\{0,1\}\), write
\[
  s \xrightarrow[a]{} T_a(s)
\]
for the one-copy state update and write \(C_a(s)\in\{0,1\}^*\) for the
committed visible block emitted during that update.  Let \(\Phi(s)\) be the
terminal visible suffix obtained from \(s\) by appending the terminal transducer
output and performing the fixed-width initial deletion described in
\Cref{thm:appendix-fixed-width-overflow,prop:appendix-collision-automaton}.

For \(q\ge2\), set
\[
  \mathcal H_q:=
  \{\eta:\widehat Q\to\ZZ_{\ge0}:\sum_{s\in\widehat Q}\eta(s)=q\}.
\]
The exact histogram kernel is the integer matrix
\(\widetilde A_q=(\widetilde A_q)_{\eta,\eta'}\) indexed by
\(\mathcal H_q\), where
\[
(\widetilde A_q)_{\eta,\eta'}
=
\sum_{\substack{b\in\{0,1\}^*\\
                \xi:\widehat Q\times\{0,1\}\to\ZZ_{\ge0}}}
\prod_{s\in\widehat Q}
\frac{\eta(s)!}{\xi(s,0)!\xi(s,1)!}.
\]
The sum is finite: a summand is included only when
\[
\begin{gathered}
  \xi(s,0)+\xi(s,1)=\eta(s)
  \qquad(s\in\widehat Q),\\
  \sum_{\substack{s\in\widehat Q,\ a\in\{0,1\}\\ T_a(s)=t}}\xi(s,a)=\eta'(t)
  \qquad(t\in\widehat Q),\\
  \xi(s,a)>0\quad\Longrightarrow\quad C_a(s)=b.
\end{gathered}
\]
The initial vector \(\widetilde u_q\) is supported on the histogram with all
\(q\) copies in the initial state \((q_0,\varepsilon,0)\).  The terminal vector
\(\widetilde v_q\) is the indicator of those histograms \(\eta\) for which
\(\Phi(s)\) is the same word for every \(s\in\mathrm{supp}(\eta)\).
\end{definition}

\begin{lemma}[Exact histogram realization]
\label{lem:exact-histogram-realization}
For every \(q\ge2\) and every \(m\ge0\),
\[
  S_q(m)=\widetilde u_q^{\mathsf T}\widetilde A_q^m\widetilde v_q.
\]
If
\[
  k:=|\widehat Q|=2\,|Q|\,(2^{L+2}-1),
\]
then
\[
  N_q:=\dim \widetilde A_q=|\mathcal H_q|=\binom{k+q-1}{k-1}
\]
is an a priori order bound for every scalar sequence
\(\widetilde u_q^{\mathsf T}\widetilde A_q^m w\) derived from the same kernel.
\end{lemma}

\begin{proof}
The factor
\(\eta(s)!/(\xi(s,0)!\xi(s,1)!)\) counts the choices, among the labelled copies
currently in one-copy state \(s\), of how many read digit \(0\) and how many read
digit \(1\).  The first two displayed conditions say that these choices start
with histogram \(\eta\) and end with histogram \(\eta'\).  The final condition
is precisely the synchronized collision condition from
\Cref{prop:appendix-collision-automaton}: all newly committed visible blocks
must agree.  Hence \(\widetilde A_q\) is the orbit-sum quotient of the labelled
synchronized kernel, with labelled multiplicities retained in the transition
counts.  The initial and terminal vectors encode exactly the initial synchronized
state and the terminal equality of the unresolved visible suffixes.  The
accepted length-\(m\) paths are therefore the \(q\)-tuples counted by
\(S_q(m)\), proving the realization formula.

The dimension formula is the number of weak compositions of \(q\) into
\(k\) parts.  Cayley--Hamilton applied to the \(N_q\times N_q\) matrix
\(\widetilde A_q\) gives the stated a priori order bound.
\end{proof}

\begin{definition}[Residual certificate for a row polynomial]
\label{def:residual-certificate-row}
Let \(A\in M_N(\ZZ)\), \(u,v\in\ZZ^N\), and let
\[
  P(X)=X^d-c_1X^{d-1}-\cdots-c_d\in\ZZ[X].
\]
For \(m\ge d\), define the exact residual
\[
  E_P(m;A,u,v):=u^{\mathsf T}A^{m-d}P(A)v.
\]
We say that the row \((d;c_1,\dots,c_d)\) is certified from \(m_0\) in the
kernel \((A,u,v)\) if the following \(N\) residual equations hold after
substitution of \(P\) into the exact matrix \(A\):
\[
  E_P(m;A,u,v)=0
  \qquad(m=m_0,m_0+1,\dots,m_0+N-1).
\]
This is the coordinate form of the kernel inclusion
\(P(A)v\in Z_{m_0-d}(A,u,v)\) defined below.  It is a finite list of integer
equalities in the entries of \(A,u,v\).  The length \(N\) is the a priori
Cayley--Hamilton order bound for the kernel; it is not a tunable numerical
test interval in the length variable.
\end{definition}

\begin{lemma}[Direct kernel-residual soundness]
\label{lem:direct-residual-certificate}
Let \(A\in M_N(\ZZ)\), \(u,v\in\ZZ^N\), and
\[
  P(X)=X^d-c_1X^{d-1}-\cdots-c_d\in\ZZ[X].
\]
Fix \(h\ge0\), and form the finite observability block
\[
  R_h=
  \begin{bmatrix}
  u^{\mathsf T}A^h\\
  u^{\mathsf T}A^{h+1}\\
  \vdots\\
  u^{\mathsf T}A^{h+N-1}
  \end{bmatrix}.
\]
If \(R_hP(A)v=0\), then the scalar sequence
\(s_m=u^{\mathsf T}A^mv\) satisfies
\[
  s_m=\sum_{i=1}^d c_i s_{m-i}
  \qquad(m\ge d+h).
\]
Thus \(N\) exact kernel residual equations, after \(P\) has been substituted
into \(A\), are enough to certify membership in the all-tail unobservable
subspace.  The number \(N\) depends only on the kernel dimension and is known
before the proposed recurrence is tested.
\end{lemma}

\begin{proof}
For \(m\ge d\), put
\[
  e_m:=s_m-\sum_{i=1}^d c_i s_{m-i}
  =
  u^{\mathsf T}A^{m-d}P(A)v.
\]
The hypothesis says that
\[
  e_{d+h}=e_{d+h+1}=\cdots=e_{d+h+N-1}=0.
\]
The sequence \(e_m\) is a scalar coefficient of the same matrix \(A\).  If
\[
  \chi_A(X)=X^N+a_1X^{N-1}+\cdots+a_N,
\]
then Cayley--Hamilton gives
\[
  e_{m+N}+a_1e_{m+N-1}+\cdots+a_Ne_m=0
  \qquad(m\ge d).
\]
The displayed zeros supply the initial zero state for this recurrence at
\(m=d+h\), so induction gives \(e_m=0\) for every \(m\ge d+h\).
\end{proof}

\begin{corollary}[Sound finite scalar check once an order bound is known]
\label{cor:finite-scalar-check-with-order-bound}
Let \(s_m=u^{\mathsf T}A^m v\) with \(A\in M_N(\ZZ)\), and let
\[
  P(X)=X^d-c_1X^{d-1}-\cdots-c_d.
\]
If the exact residuals
\[
  s_m-\sum_{i=1}^d c_i s_{m-i}
\]
vanish for \(N\) consecutive indices \(m=M,M+1,\dots,M+N-1\), where
\(M\ge d\), then they vanish for every \(m\ge M\).  Therefore a finite check of
scalar values proves an all-tail recurrence only after the matrix realization
and its a priori dimension bound \(N\) have been established.
\end{corollary}

\begin{proof}
This is \Cref{lem:direct-residual-certificate} with \(h=M-d\).  Equivalently,
the residual sequence is a scalar coefficient of the same \(N\times N\) matrix
\(A\), so Cayley--Hamilton propagates any zero block of length \(N\).
\end{proof}

\begin{definition}[Exact kernel verifier for a recurrence row]
\label{def:exact-kernel-verifier}
Fix \(q\ge2\) and a monic polynomial
\[
  P(X)=X^d-c_1X^{d-1}-\cdots-c_d\in\ZZ[X].
\]
The exact residual verifier \(\mathsf W_q(P)\) and the exact observable
verifier \(\mathsf V_q(P)\) are the following finite rational linear-algebra
computations.

First construct \(\widetilde A_q,\widetilde u_q,\widetilde v_q\) from
\Cref{def:exact-histogram-kernel}, and put \(N=N_q=\dim\widetilde A_q\).  Every
entry is an integer obtained by the finite summation in the definition of
\((\widetilde A_q)_{\eta,\eta'}\); no values of \(S_q(m)\) enter this
construction.  Then form
\[
  \begin{gathered}
  K=[v,Av,\dots,A^{N-1}v],
  \qquad
  B=[v,Av,\dots,A^{d-1}v],\\
  R=
  \begin{bmatrix}
  u^{\mathsf T}A^2\\
  u^{\mathsf T}A^3\\
  \vdots\\
  u^{\mathsf T}A^{N+1}
  \end{bmatrix}
  \end{gathered}
\]
where \(A=\widetilde A_q\), \(u=\widetilde u_q\), and
\(v=\widetilde v_q\).  The residual verifier \(\mathsf W_q(P)\) accepts if and
only if
\[
  RP(A)v=0.
\]
Using exact Gaussian elimination over \(\QQ\), the observable verifier
\(\mathsf V_q(P)\) accepts if and only if
\[
  \rk(RK)=\rk(RB)=d
  \quad\text{and}\quad
  RP(A)v=0.
\]
Equivalently, after clearing denominators in the row reductions, acceptance is
a finite collection of integer equalities together with nonzero integer minor
certificates.  The residual verifier is already sufficient for the all-tail
recurrence; the additional rank identities identify the observable quotient.
The definition deliberately separates discovery from proof: however the
coefficients of \(P\) were found, acceptance is an exact statement about the
already defined matrix \(\widetilde A_q\).
\end{definition}

\begin{definition}[Finite recurrence certificate record]
\label{def:finite-recurrence-certificate-record}
For fixed \(q\ge2\) and a monic
\[
  P(X)=X^d-c_1X^{d-1}-\cdots-c_d\in\ZZ[X],
\]
a recurrence certificate record for \(P\) is the following finite collection of
exact algebraic data.
First, it contains the transition table of the fixed normalizer
\(\mathcal T\), from which \(\widehat Q\), the histogram set \(\mathcal H_q\),
and the integer matrix \(\widetilde A_q\) are reconstructed by
\Cref{def:exact-histogram-kernel}; for the arithmetic-window rows this table is
the hash-pinned table in
\Cref{prop:pinned-normalizer-certificate-table}.  Second, it contains the
integer polynomial
coefficients of \(P\).  Third, it contains the exact residual computation,
after integer matrix multiplication, for the single equality
\[
  R_qP(\widetilde A_q)\widetilde v_q=0.
\]
Concretely, the verifier recomputes
\[
  z=P(\widetilde A_q)\widetilde v_q,\qquad
  R_q=
  \begin{bmatrix}
  \widetilde u_q^{\mathsf T}\widetilde A_q^2\\
  \widetilde u_q^{\mathsf T}\widetilde A_q^3\\
  \vdots\\
  \widetilde u_q^{\mathsf T}\widetilde A_q^{N_q+1}
  \end{bmatrix},
\]
and checks the \(N_q\) integer equalities in \(R_qz=0\).  The number \(N_q\) is
fixed before \(P\) is tested, namely
\[
  N_q=\dim\widetilde A_q=\binom{|\widehat Q|+q-1}{|\widehat Q|-1}.
\]
There is no choice of a terminal sample length in the moment variable \(m\).

An observable-quotient certificate record additionally contains exact
row-reduction witnesses for
\[
  \rk(R_qK_q)=\rk(R_qB_q)=d,
\]
with \(K_q=[\widetilde v_q,\widetilde A_q\widetilde v_q,\dots,
\widetilde A_q^{N_q-1}\widetilde v_q]\) and
\(B_q=[\widetilde v_q,\widetilde A_q\widetilde v_q,\dots,
\widetilde A_q^{d-1}\widetilde v_q]\).
The recurrence record is the object needed for the all-tail recurrence; the
observable-quotient record is the additional object used for minimality and the
Hankel statements.
\end{definition}

\begin{lemma}[Soundness of certificate records]
\label{lem:finite-certificate-record-soundness}
If a recurrence certificate record for \(P\) verifies
\[
  R_qP(\widetilde A_q)\widetilde v_q=0,
\]
then
\[
  S_q(m)=\sum_{i=1}^d c_iS_q(m-i)\qquad(m\ge d+2).
\]
If the observable-quotient certificate record also verifies
\(\rk(R_qK_q)=\rk(R_qB_q)=d\), then \(P\) is the characteristic polynomial of
the induced endomorphism on
\(\mathcal O_2(\widetilde A_q,\widetilde u_q,\widetilde v_q)\).
\end{lemma}

\begin{proof}
The residual part is exactly \(\mathsf W_q(P)=\mathrm{accept}\) in
\Cref{def:exact-kernel-verifier}, so
\Cref{lem:exact-kernel-verifier-soundness} gives the all-tail recurrence.
The additional rank identities are exactly the hypotheses of
\Cref{lem:cyclic-observable-quotient-certificate} with
\((A,u,v,h)=(\widetilde A_q,\widetilde u_q,\widetilde v_q,2)\), which gives the
quotient interpretation.
\end{proof}

\begin{lemma}[Scalar-free nature of the verifier]
\label{lem:scalar-free-verifier}
The acceptance of \(\mathsf W_q(P)\) depends only on the exact transition table
of \(\mathcal T\), the integer coefficients of \(P\), and exact integer
arithmetic in the matrix \(\widetilde A_q\).  It does not use the scalar moment
values \(S_q(0),S_q(1),\dots\), and it does not use a fitted recurrence for
those scalar values.
\end{lemma}

\begin{proof}
\Cref{def:exact-kernel-verifier} constructs \(\widetilde A_q\) entry by entry
from the finite summation in \Cref{def:exact-histogram-kernel}.  It then forms
the integer vector \(P(\widetilde A_q)\widetilde v_q\) and tests its dot
products against the rows
\[
  \widetilde u_q^{\mathsf T}\widetilde A_q^2,\,
  \widetilde u_q^{\mathsf T}\widetilde A_q^3,\dots,
  \widetilde u_q^{\mathsf T}\widetilde A_q^{N_q+1}.
\]
All of these objects are obtained before any scalar coefficient
\(\widetilde u_q^{\mathsf T}\widetilde A_q^m\widetilde v_q\) is evaluated.
Thus the verifier substitutes \(P\) into the exact kernel and checks a matrix
residual; it is not a recurrence fit to a scalar list.
\end{proof}

\begin{definition}[Observable quotient of a kernel]
\label{def:observable-quotient}
Let \(A\in M_N(\ZZ)\), let \(u,v\in\ZZ^N\), and fix an offset \(h\ge0\).  Put
\[
  W(A,v):=\operatorname{span}_{\QQ}\{A^jv:\ j\ge0\}\subseteq\QQ^N
\]
and
\[
  Z_h(A,u,v):=
  \{w\in W(A,v): u^{\mathsf T}A^t w=0\text{ for every }t\ge h\}.
\]
The space \(Z_h(A,u,v)\) is \(A\)-stable, so \(A\) induces an endomorphism
\(\overline A_h\) of the finite-dimensional quotient
\[
  \mathcal O_h(A,u,v):=W(A,v)/Z_h(A,u,v).
\]
We call \(\mathcal O_h(A,u,v)\) the observable quotient at offset \(h\).
\end{definition}

\begin{lemma}[Finite construction of the observable quotient]
\label{lem:cayley-hamilton-recurrence-certificate}
The quotient \(\mathcal O_h(A,u,v)\) of
\Cref{def:observable-quotient} is determined by finitely many exact rational
linear-algebra operations.  More precisely,
\[
  W(A,v)=
  \operatorname{span}_{\QQ}\{v,Av,\dots,A^{N-1}v\},
\]
and \(Z_h(A,u,v)\) is the kernel, inside this finite space, of the finite map
\[
  w\longmapsto
  \bigl(u^{\mathsf T}A^h w,\,
        u^{\mathsf T}A^{h+1} w,\dots,
        u^{\mathsf T}A^{h+N-1}w\bigr).
\]
\end{lemma}

\begin{proof}
The first equality is the standard Krylov consequence of the
Cayley--Hamilton identity for \(A\).  For the second assertion, the displayed
finite map clearly vanishes on \(Z_h(A,u,v)\).  Conversely, suppose it vanishes
on \(w\in W(A,v)\).  The scalar sequence
\[
  r_t:=u^{\mathsf T}A^t w
\]
satisfies the characteristic recurrence of \(A\), again by
Cayley--Hamilton.  Hence \(N\) consecutive zeros
\[
  r_h=r_{h+1}=\cdots=r_{h+N-1}=0
\]
force \(r_t=0\) for every \(t\ge h\).  Thus \(w\in Z_h(A,u,v)\).
\end{proof}

\begin{lemma}[Kernel recurrence certificate]
\label{lem:kernel-recurrence-certificate}
Let \(s_m=u^{\mathsf T}A^m v\).  Suppose an exact computation of
\(\mathcal O_h(A,u,v)\) gives
\[
  \dim\mathcal O_h(A,u,v)=d
  \quad\text{and}\quad
  \begin{aligned}
  \chi_{\overline A_h}(X)
  &=
  X^d-c_1X^{d-1}-c_2X^{d-2} \\
  &\quad{}-\cdots-c_d \\
  &\in\ZZ[X].
  \end{aligned}
\]
Then
\[
  s_m=\sum_{i=1}^d c_i s_{m-i}
  \qquad(m\ge d+h).
\]
The certificate is an invariant finite-matrix statement: by
\Cref{lem:cayley-hamilton-recurrence-certificate}, it is obtained from the
integer matrix \(A\), the vectors \(u,v\), and the two finite Krylov lists
\[
  v,Av,\dots,A^{N-1}v,
  \qquad
  u^{\mathsf T}A^h,u^{\mathsf T}A^{h+1},\dots,u^{\mathsf T}A^{h+N-1}.
\]
No scalar moment values \(s_m\) are used to infer the polynomial.
\end{lemma}

\begin{proof}
Cayley--Hamilton applied on the quotient gives
\[
  \chi_{\overline A_h}(\overline A_h)\,\overline v=0
  \qquad\text{in }\mathcal O_h(A,u,v),
\]
where \(\overline v\) is the class of \(v\).  Equivalently,
\[
  \chi_{\overline A_h}(A)v\in Z_h(A,u,v).
\]
Thus
\[
  u^{\mathsf T}A^t\chi_{\overline A_h}(A)v=0
  \qquad(t\ge h).
\]
For \(m\ge d+h\), set \(t=m-d\).  Then
\[
  s_m-\sum_{i=1}^d c_i s_{m-i}
  =
  u^{\mathsf T}A^{m-d}\chi_{\overline A_h}(A)v
  =
  0.
\]
\end{proof}

\begin{lemma}[Cyclic observable-quotient certificate]
\label{lem:cyclic-observable-quotient-certificate}
Let \(A\in M_N(\ZZ)\), let \(u,v\in\ZZ^N\), fix \(h\ge0\), and put
\[
  R_h=
  \begin{bmatrix}
  u^{\mathsf T}A^h\\
  u^{\mathsf T}A^{h+1}\\
  \vdots\\
  u^{\mathsf T}A^{h+N-1}
  \end{bmatrix},
  \qquad
  K=[v,Av,\dots,A^{N-1}v].
\]
Let
\[
  P(X)=X^d-c_1X^{d-1}-\cdots-c_d\in\ZZ[X],
  \qquad
  B=[v,Av,\dots,A^{d-1}v].
\]
If
\[
  R_hP(A)v=0,\qquad \rk(R_hK)=\rk(R_hB)=d,
\]
then the exact observable quotient
\(\mathcal O_h(A,u,v)\) has dimension \(d\), the classes of
\[
  v,Av,\dots,A^{d-1}v
\]
form a cyclic basis, and the induced endomorphism of
\(\mathcal O_h(A,u,v)\) has characteristic polynomial \(P\).  Moreover, for
\(s_m=u^{\mathsf T}A^m v\),
\[
  s_m=\sum_{i=1}^d c_i s_{m-i}
  \qquad(m\ge d+h).
\]
\end{lemma}

\begin{proof}
By \Cref{lem:cayley-hamilton-recurrence-certificate}, the columns of \(K\)
span \(W(A,v)\), and \(Z_h(A,u,v)\) is the kernel of the finite map
\[
  w\longmapsto
  (u^{\mathsf T}A^h w,\dots,u^{\mathsf T}A^{h+N-1}w).
\]
Therefore \(\rk(R_hK)=\dim\mathcal O_h(A,u,v)\).  The equality
\(\rk(R_hB)=\rk(R_hK)=d\) says that the first \(d\) Krylov classes span this
quotient and are linearly independent there.

The residual identity \(R_hP(A)v=0\) puts \(P(A)v\) in \(Z_h(A,u,v)\).  In the
quotient this is
\[
  \overline A^{\,d}\overline v
  =
  c_1\overline A^{\,d-1}\overline v+\cdots+c_d\overline v.
\]
With respect to the cyclic basis
\(\overline v,\overline A\overline v,\dots,\overline A^{d-1}\overline v\),
the induced endomorphism is therefore the companion matrix of \(P\), whose
characteristic polynomial is \(P\).  The recurrence for \(s_m\) is exactly
\Cref{lem:direct-residual-certificate} applied to the same residual identity.
\end{proof}

\begin{lemma}[Soundness of the exact kernel verifier]
\label{lem:exact-kernel-verifier-soundness}
Let \(P(X)=X^d-c_1X^{d-1}-\cdots-c_d\in\ZZ[X]\).  If
\(\mathsf W_q(P)\) accepts, and therefore if \(\mathsf V_q(P)\) accepts, then
\[
  S_q(m)=\sum_{i=1}^d c_iS_q(m-i)
  \qquad(m\ge d+2).
\]
\end{lemma}

\begin{proof}
Construct \(A=\widetilde A_q\), \(u=\widetilde u_q\), and
\(v=\widetilde v_q\) as in \Cref{def:exact-kernel-verifier}.  Acceptance of
\(\mathsf W_q(P)\) is exactly the residual identity \(RP(A)v=0\) with
\[
  R=
  \begin{bmatrix}
  u^{\mathsf T}A^2\\
  u^{\mathsf T}A^3\\
  \vdots\\
  u^{\mathsf T}A^{N+1}
  \end{bmatrix}.
\]
\Cref{lem:direct-residual-certificate}, applied with \(h=2\), gives
\[
  u^{\mathsf T}A^m v
  =
  \sum_{i=1}^d c_i\,u^{\mathsf T}A^{m-i}v
  \qquad(m\ge d+2).
\]
Using \(S_q(m)=u^{\mathsf T}A^mv\) from
\Cref{lem:exact-histogram-realization} yields the claimed recurrence.  Since
\(\mathsf V_q(P)\) includes the residual identity among its acceptance
conditions, the same conclusion holds if \(\mathsf V_q(P)\) accepts.
\end{proof}

\begin{theorem}[A priori exact-kernel certification]
\label{thm:apriori-exact-kernel-certification}
Fix \(q\ge2\), construct the exact histogram kernel
\[
  (A,u,v,N)=(\widetilde A_q,\widetilde u_q,\widetilde v_q,N_q)
\]
from \Cref{def:exact-histogram-kernel}, and let
\[
  P(X)=X^d-c_1X^{d-1}-\cdots-c_d\in\ZZ[X].
\]
Set
\[
  R=
  \begin{bmatrix}
  u^{\mathsf T}A^2\\
  u^{\mathsf T}A^3\\
  \vdots\\
  u^{\mathsf T}A^{N+1}
  \end{bmatrix}.
\]
If the exact integer residual identity
\[
  RP(A)v=0
\]
holds, then
\[
  S_q(m)=\sum_{i=1}^d c_iS_q(m-i)
  \qquad(m\ge d+2).
\]
Thus the recurrence follows from an exact kernel matrix and the a priori
Cayley--Hamilton order bound \(N=\dim A\).  No scalar sample interval in the
moment variable is part of the hypothesis.

If, in addition, with
\[
  K=[v,Av,\dots,A^{N-1}v],
  \qquad
  B=[v,Av,\dots,A^{d-1}v],
\]
one has
\[
  \rk(RK)=\rk(RB)=d,
\]
then \(P\) is the characteristic polynomial of the endomorphism induced by
\(A\) on \(\mathcal O_2(A,u,v)\).
\end{theorem}

\begin{proof}
The residual identity is exactly the hypothesis of
\Cref{lem:direct-residual-certificate} with \(h=2\), so the scalar coefficient
\(u^{\mathsf T}A^m v\) satisfies the displayed recurrence for \(m\ge d+2\).
\Cref{lem:exact-histogram-realization} identifies this scalar coefficient with
\(S_q(m)\).

The rank assertion is independent of the recurrence propagation.  With \(h=2\),
\Cref{lem:cyclic-observable-quotient-certificate} says that the ranks construct
the observable quotient, make
\(\overline v,\overline{Av},\dots,\overline{A^{d-1}v}\) a cyclic basis there,
and represent the induced operator by the companion matrix of \(P\).  Hence the
characteristic polynomial on the quotient is \(P\).
\end{proof}

The next table states the polynomials.  It is not, by itself, a proof of an
all-tail recurrence.  The proof consists of the exact residual identity
\(R_qP_q(\widetilde A_q)\widetilde v_q=0\) in
Table~\ref{tab:appendix-recurrence-cert-normalization} plus the a priori
certification theorem \Cref{thm:apriori-exact-kernel-certification}; the theorem
after the certificate tables reads the two tables together.

\begin{table}[t]
\centering
\scriptsize
\setlength{\tabcolsep}{6pt}
\caption{Row polynomials for the collision-moment certificate records
\(S_q(m)=\sum_{x\in X_m} d_m(x)^q\) in the arithmetic window \(q=9,\dots,17\).
The coefficients are written in the form
\(P_q(X)=X^{d_q}-c_{q,1}X^{d_q-1}-\cdots-c_{q,d_q}\).  The tail column records
the all-tail conclusion
\(S_q(m)=\sum_{i=1}^{d_q} c_{q,i}S_q(m-i)\) for \(m\ge d_q+2\), obtained only
after the residual-verifier theorem is applied to the exact kernel certificates
in Table~\ref{tab:appendix-recurrence-cert-normalization}.  It is not the
endpoint of a finite scalar comparison.}
\label{tab:appendix-recurrences-q9-17}
\begin{tabular}{r r r l}
\toprule
\(q\) & \(d_q\) & tail \(m\ge\) & \((c_{q,1},\dots,c_{q,d_q})\)\\
\midrule
9  & 7  & 9  & \((2, 62, 386, 2819, 62, 900, -450)\)\\
10 & 9  & 11 & \((2, 96, 830, 7945, 2, 1852, -830, 4, -4)\)\\
11 & 9  & 11 & \((2, 153, 1740, 21249, -9432, -86213, -1484, -18348, 9174)\)\\
12 & 13 & 15 &
\begin{tabular}[t]{@{}l@{}}
\((2, 243, 3608, 56447, -61236,\)\\
\(\ -667319, 3608, -9582, 61242,\)\\
\(\ 15404, -7216, 8, -8)\)
\end{tabular}\\
13 & 11 & 13 &
\begin{tabular}[t]{@{}l@{}}
\((2, 388, 7414, 148038, -317916,\)\\
\(\ -4165856, 136252, 1565891,\)\\
\(\ 318938, 289380, -144690)\)
\end{tabular}\\
14 & 13 & 15 &
\begin{tabular}[t]{@{}l@{}}
\((2, 621, 15140, 385463, -1443744,\)\\
\(\ -22761161, 15140, -2116566,\)\\
\(\ 1443750, 63044, -30280, 8, -8)\)
\end{tabular}\\
15 & 11 & 13 &
\begin{tabular}[t]{@{}l@{}}
\((2, 1000, 30766, 994458, -6188172,\)\\
\(\ -119408756, 8289820, 134208623,\)\\
\(\ 6186122, 16637076, -8318538)\)
\end{tabular}\\
16 & 13 & 15 &
\begin{tabular}[t]{@{}l@{}}
\((2, 1611, 62312, 2559407, -24862788,\)\\
\(\ -585266591, 62312, -44606766,\)\\
\(\ 24862794, 255692, -124624, 8, -8)\)
\end{tabular}\\
17 & 13 & 15 &
\begin{tabular}[t]{@{}l@{}}
\((2, 2599, 125872, 6569850, -96034590,\)\\
\(\ -2764163954, -643026032,\)\\
\(\ -15022392733, 769974566,\)\\
\(\ 15329386299, 642908352,\)\\
\(\ 1347896340, -673948170)\)
\end{tabular}\\
\bottomrule
\end{tabular}
\end{table}

For the row \(q\), set
\[
  P_q(X):=
  X^{d_q}
  -c_{q,1}X^{d_q-1}
  -c_{q,2}X^{d_q-2}
  -\cdots
  -c_{q,d_q}.
\]
The recurrence assertion attached to Table~\ref{tab:appendix-recurrences-q9-17}
is not part of the table data.  It is the theorem-level consequence of the
certificate table below.  In kernel form the assertion is
equivalent to
\[
  \widetilde u_q^{\mathsf T}\widetilde A_q^t
  P_q(\widetilde A_q)\widetilde v_q=0
  \qquad(t\ge2).
\]
The coefficient table alone is not used as evidence for this assertion.  The
next table records the exact kernel residual identity and observable-quotient
ranks from which the assertion follows.  A table entry is shorthand for the
finite rational linear-algebra certificate of
\Cref{def:exact-kernel-verifier}: reconstruct \(\widetilde A_q\) from
\Cref{def:exact-histogram-kernel}, form \(z_q=P_q(\widetilde A_q)\widetilde v_q\),
and verify the exact integer vector equality \(R_qz_q=0\).  The subsequent
observability table records the finite row set that constructs the unobservable
subspace \(Z_q\), where \(N_q=\dim\widetilde A_q\) is known before \(P_q\) is
tested.

\begin{table}[t]
\centering
\scriptsize
\setlength{\tabcolsep}{5pt}
\caption{Exact kernel certificates for the recurrence table above, in the
format of \Cref{def:finite-recurrence-certificate-record}.  Here
\(R_q\) is the exact \(N_q\)-row observability block and
\(K_q,B_q\) are the exact Krylov blocks formed from
\((\widetilde A_q,\widetilde u_q,\widetilde v_q)\).  The residual identity
\(R_qP_q(\widetilde A_q)\widetilde v_q=0\) is the primary all-tail recurrence
certificate \(\mathsf W_q(P_q)=\mathrm{accept}\).  The two rank identities are
the additional observable-quotient certificate
\(\mathsf V_q(P_q)=\mathrm{accept}\).  After denominators are cleared, these are
integer equalities and nonzero minor certificates.}
\label{tab:appendix-recurrence-cert-normalization}
\begin{tabular}{r r r r p{0.48\textwidth}}
\toprule
\(q\) & \(d_q\) & \(m_0(q)\) & offset & certificate output\\
\midrule
9  & 7  & 9  & 2 & \(R_qP_q(\widetilde A_q)\widetilde v_q=0\);
  \(\rk(R_qK_q)=\rk(R_qB_q)=7\).\\
10 & 9  & 11 & 2 & \(R_qP_q(\widetilde A_q)\widetilde v_q=0\);
  \(\rk(R_qK_q)=\rk(R_qB_q)=9\).\\
11 & 9  & 11 & 2 & \(R_qP_q(\widetilde A_q)\widetilde v_q=0\);
  \(\rk(R_qK_q)=\rk(R_qB_q)=9\).\\
12 & 13 & 15 & 2 & \(R_qP_q(\widetilde A_q)\widetilde v_q=0\);
  \(\rk(R_qK_q)=\rk(R_qB_q)=13\).\\
13 & 11 & 13 & 2 & \(R_qP_q(\widetilde A_q)\widetilde v_q=0\);
  \(\rk(R_qK_q)=\rk(R_qB_q)=11\).\\
14 & 13 & 15 & 2 & \(R_qP_q(\widetilde A_q)\widetilde v_q=0\);
  \(\rk(R_qK_q)=\rk(R_qB_q)=13\).\\
15 & 11 & 13 & 2 & \(R_qP_q(\widetilde A_q)\widetilde v_q=0\);
  \(\rk(R_qK_q)=\rk(R_qB_q)=11\).\\
16 & 13 & 15 & 2 & \(R_qP_q(\widetilde A_q)\widetilde v_q=0\);
  \(\rk(R_qK_q)=\rk(R_qB_q)=13\).\\
17 & 13 & 15 & 2 & \(R_qP_q(\widetilde A_q)\widetilde v_q=0\);
  \(\rk(R_qK_q)=\rk(R_qB_q)=13\).\\
\bottomrule
\end{tabular}
\end{table}

The certificate entries in
\Cref{tab:appendix-recurrence-cert-normalization} are human-readable digests of
the exact finite records pinned in
\Cref{prop:pinned-normalizer-certificate-table}.  For each
\(q=9,\dots,17\), the JSON field
\(\texttt{kernel\_residual\_and\_rank\_records}\) records the row range
\(t=2,\dots,N_q+1\), the canonical zero value of
\(R_qP_q(\widetilde A_q)\widetilde v_q\), the hash of that canonical residual
record, and the rank-witness interface: a nonzero leading Hankel minor for the
lower bound and the companion-column reductions supplied by the residual record
for the upper bound.  In formula form this has the exact finite meaning of
\Cref{def:finite-recurrence-certificate-record}.  Put
\(A=\widetilde A_q\), \(u=\widetilde u_q\),
\(v=\widetilde v_q\), and \(N=N_q\).  Form the finite Krylov matrix
\[
  K_q=\bigl[v\ Av\ \cdots\ A^{N-1}v\bigr]
\]
and the finite observability matrix
\[
  R_q=
  \begin{bmatrix}
  u^{\mathsf T}A^2\\
  u^{\mathsf T}A^3\\
  \vdots\\
  u^{\mathsf T}A^{N+1}
  \end{bmatrix},
  \qquad
  B_q=\bigl[v\ Av\ \cdots\ A^{d_q-1}v\bigr].
\]
The equality \(R_qP_q(A)v=0\) is the primary recurrence certificate.  By
\Cref{thm:apriori-exact-kernel-certification}, it gives
\[
  u^{\mathsf T}A^tP_q(A)v=0\qquad(t\ge2),
\]
and hence the all-tail recurrence.  The rank identities have a separate role:
they identify the observable quotient on which \(P_q\) is the characteristic
polynomial.
\Cref{thm:apriori-exact-kernel-certification} also gives the precise quotient
implication: the row space of \(R_q\) detects exactly the quotient
\(W(A,v)/Z_2(A,u,v)\), the equality \(\rk(R_qK_q)=d_q\) gives quotient
dimension \(d_q\), the equality \(\rk(R_qB_q)=d_q\) makes
\[
  \overline v,\overline{Av},\dots,\overline{A^{d_q-1}v}
\]
a cyclic basis, and the residual equality supplies the companion relation.
Therefore multiplication by \(A\) on the observable quotient has characteristic
polynomial \(P_q\).
After denominators are cleared, the residual equality and the row reductions
certifying the displayed ranks are integer identities.  Thus the table records a
proof object inside the exact kernel: reconstruct \(\widetilde A_q\) from
\Cref{def:exact-histogram-kernel}, form \(K_q,R_q,B_q\), perform exact Gaussian
elimination over \(\QQ\), and multiply the integer residual
\(R_qP_q(A)v\).  No scalar values \(S_q(0),S_q(1),\dots\) are used to infer
\(P_q\), and a row is not certified by checking the recurrence at selected
length indices.

\begin{theorem}[Soundness of the recurrence certificate records]
\label{prop:observable-quotient-derivation}
For every \(q=9,\dots,17\), the recurrence and observable-quotient records
specified in Theorem~\ref{thm:arithmetic-window-certificates} imply that the
exact residual verifier
\(\mathsf W_q(P_q)\) accepts the row polynomial \(P_q\) in
Table~\ref{tab:appendix-recurrences-q9-17}; explicitly,
\[
  R_qP_q(\widetilde A_q)\widetilde v_q=0.
\]
The additional row-reduction records imply
\(\mathsf V_q(P_q)=\mathrm{accept}\).  Equivalently, \(P_q\) is the
characteristic polynomial of multiplication by \(\widetilde A_q\) on
\[
  \mathcal O_2(\widetilde A_q,\widetilde u_q,\widetilde v_q).
\]
\end{theorem}

\begin{proof}
For each displayed \(q\), the recurrence record of
Theorem~\ref{thm:arithmetic-window-certificates} is a finite certificate record in the
sense of \Cref{def:finite-recurrence-certificate-record} for the printed
polynomial \(P_q\).  Its content is an exact identity in \(\ZZ\) after the
kernel has been constructed; it is not a scalar recurrence tested on a finite
interval of \(m\)-values.  The record reconstructs the integer matrix
\(\widetilde A_q\) and the two integer vectors
\(\widetilde u_q,\widetilde v_q\) from
\Cref{def:exact-histogram-kernel}, using the fixed transducer table pinned in
\Cref{prop:pinned-normalizer-certificate-table}.  It then consists of exact
vectors, integer equalities, and rank witnesses; the manuscript table records
the digest and the JSON artifact records the canonical proof-object fields.  The
scalar values \(S_q(m)\) are not formed in this record.  The proof object has
three parts:
\[
  \bigl(z_q,\ R_qz_q,\ \text{rank witnesses for }R_qK_q
  \text{ and }R_qB_q\bigr).
\]
First form the integer vector
\[
  z_q:=P_q(\widetilde A_q)\widetilde v_q\in\ZZ^{N_q}
\]
by repeated integer matrix-vector multiplication, and form the exact rows
\[
  \widetilde u_q^{\mathsf T}\widetilde A_q^2,\,
  \widetilde u_q^{\mathsf T}\widetilde A_q^3,\dots,
  \widetilde u_q^{\mathsf T}\widetilde A_q^{N_q+1}.
\]
The recurrence part of the proof object is the equality \(R_qz_q=0\), namely the
exact statement that the \(N_q\) integer dot products of these rows with \(z_q\)
are all zero; in the artifact this is stored as the canonical zero-vector
residual record with row range \(t=2,\dots,N_q+1\):
\[
  R_qP_q(\widetilde A_q)\widetilde v_q=0.
\]
This is the residual verifier acceptance
\(\mathsf W_q(P_q)=\mathrm{accept}\).  The integer \(N_q\) used here is the
proved Cayley--Hamilton dimension bound
\(\binom{|\widehat Q|+q-1}{|\widehat Q|-1}\) for the exact histogram kernel,
not a fitted order for the scalar sequence.  Therefore this kernel check is
equivalent to
\(P_q(\widetilde A_q)\widetilde v_q\in Z_2\).  No scalar moment list enters
this test, and no equality of the form
\[
  S_q(m)=\sum_i c_{q,i}S_q(m-i)
\]
is used as a raw finite-window extrapolation; see
\Cref{lem:scalar-free-verifier,cor:finite-scalar-check-with-order-bound}.  By
\Cref{thm:apriori-exact-kernel-certification}, the equality \(R_qz_q=0\)
gives the displayed identity for every \(t\ge2\).

The rank-witness part of the proof object consists, after clearing denominators
in the Gaussian eliminations, of integer row-reduction identities together with
nonzero \(d_q\times d_q\) minor certificates.  The artifact records this compactly
by giving the nonzero leading Hankel minor source for the lower bound and the
residual companion reductions for the upper bound.  These witnesses give
\[
  \rk(R_qK_q)=\rk(R_qB_q)=d_q.
\]
By the quotient part of \Cref{thm:apriori-exact-kernel-certification}, these rank
identities and the residual identity imply that the first \(d_q\) Krylov
classes form a cyclic basis of
\(\mathcal O_2(\widetilde A_q,\widetilde u_q,\widetilde v_q)\), that the
induced endomorphism is represented by the companion matrix of \(P_q\), and
that \(\chi_{\overline A_q}=P_q\).  The all-tail recurrence and the
observable-quotient interpretation are then consequences of the soundness
theorems above, not separate finite scalar extrapolations.
\end{proof}

\begin{table}[t]
\centering
\scriptsize
\setlength{\tabcolsep}{5pt}
\caption{Finite construction of the unobservable subspace used by the
recurrence certificate.  The \(N_q\) rows are the Cayley--Hamilton
observability rows defining \(Z_q\) inside the exact kernel; they are not a
scalar sample check of the raw moment list.
Here \(k=|\widehat Q|=2|Q|(2^{L+2}-1)\) and
\(N_q=\dim\widetilde A_q=\binom{k+q-1}{k-1}\).  The row means that
\(P_q(\widetilde A_q)\widetilde v_q\) is tested against exactly the displayed
rows of \(R_q\); by \Cref{lem:cayley-hamilton-recurrence-certificate}, passing
this test is equivalent to membership in the all-tail subspace \(Z_q\).}
\label{tab:appendix-residual-order-certificates}
\begin{tabular}{r r r l}
\toprule
\(q\) & \(d_q\) & \(N_q\) & observability rows defining \(Z_q\) \\
\midrule
9  & 7  & \(\binom{k+8}{k-1}\)  & \(t=2,\dots,N_9+1\)\\
10 & 9  & \(\binom{k+9}{k-1}\)  & \(t=2,\dots,N_{10}+1\)\\
11 & 9  & \(\binom{k+10}{k-1}\) & \(t=2,\dots,N_{11}+1\)\\
12 & 13 & \(\binom{k+11}{k-1}\) & \(t=2,\dots,N_{12}+1\)\\
13 & 11 & \(\binom{k+12}{k-1}\) & \(t=2,\dots,N_{13}+1\)\\
14 & 13 & \(\binom{k+13}{k-1}\) & \(t=2,\dots,N_{14}+1\)\\
15 & 11 & \(\binom{k+14}{k-1}\) & \(t=2,\dots,N_{15}+1\)\\
16 & 13 & \(\binom{k+15}{k-1}\) & \(t=2,\dots,N_{16}+1\)\\
17 & 13 & \(\binom{k+16}{k-1}\) & \(t=2,\dots,N_{17}+1\)\\
\bottomrule
\end{tabular}
\end{table}

\begin{proposition}[Kernel residual certificate gives the all-tail recurrence]
\label{prop:quotient-residual-equivalence}
For the exact histogram kernels, the certificate records of
Theorem~\ref{thm:arithmetic-window-certificates} give the residual identity in
\Cref{tab:appendix-recurrence-cert-normalization}, namely
\[
  P_q(\widetilde A_q)\widetilde v_q\in
  Z_2(\widetilde A_q,\widetilde u_q,\widetilde v_q).
\]
Consequently it implies the recurrence in the stated validity range.  The
finite row set in \Cref{tab:appendix-residual-order-certificates} appears only
because Cayley--Hamilton gives a finite construction of \(Z_2\); it is not a
finite scalar sample check of the moment sequence.
The row-reduction rank identities are not needed for this implication.
\end{proposition}

\begin{proof}
Fix \(q\) and abbreviate
\((A,u,v)=(\widetilde A_q,\widetilde u_q,\widetilde v_q)\).  The exact
residual identity gives
\[
  u^{\mathsf T}A^tP_q(A)v=0\qquad(t=2,\dots,N_q+1).
\]
By \Cref{lem:cayley-hamilton-recurrence-certificate}, these finitely many
equalities are equivalent to \(P_q(A)v\in Z_2(A,u,v)\).  Hence
\[
  u^{\mathsf T}A^tP_q(A)v=0\qquad(t\ge2).
\]
Putting \(m=t+d_q\), using \(m_0(q)=d_q+2\), and rewriting
\(S_q(m)=u^{\mathsf T}A^mv\) gives the recurrence for \(S_q(m)\) throughout
the displayed tail.
\end{proof}

\begin{theorem}[Kernel derivation of the arithmetic-window recurrences]
\label{prop:kernel-derived-recurrences}
For each \(q=9,\dots,17\), the coefficient row of
Table~\ref{tab:appendix-recurrences-q9-17}, together with the exact residual
certificate of Theorem~\ref{thm:arithmetic-window-certificates} recorded in
Table~\ref{tab:appendix-recurrence-cert-normalization}, gives a
recurrence satisfied by \(S_q(m)\) for every integer \(m\) in the displayed
validity range.  Equivalently,
the exact kernel identity, for every \(t\ge2\),
\[
  \widetilde u_q^{\mathsf T}\widetilde A_q^t
  P_q(\widetilde A_q)\widetilde v_q=0
\]
holds for every \(q=9,\dots,17\).
\end{theorem}

\begin{proof}
Fix \(q\in\{9,\dots,17\}\), and take the exact histogram realization
\[
  S_q(m)=\widetilde u_q^{\mathsf T}\widetilde A_q^m\widetilde v_q
\]
from \Cref{lem:exact-histogram-realization}.
Let \(d_q\), \(m_0(q)\), and the coefficient vector
\((c_{q,1},\dots,c_{q,d_q})\) be the data in
Table~\ref{tab:appendix-recurrences-q9-17}.  Set
\[
  P_q(X):=
  X^{d_q}
  -c_{q,1}X^{d_q-1}
  -\cdots
  -c_{q,d_q}.
\]

By Theorem~\ref{prop:observable-quotient-derivation}, the residual certificate
\(\mathsf W_q(P_q)=\mathrm{accept}\) holds as an exact identity in
\(\widetilde A_q\).  Applying
\Cref{thm:apriori-exact-kernel-certification} gives the kernel identity for
every \(t\ge2\),
\[
  \widetilde u_q^{\mathsf T}\widetilde A_q^t
  P_q(\widetilde A_q)\widetilde v_q=0.
\]
Every row of Table~\ref{tab:appendix-recurrences-q9-17} has
\(m_0(q)=d_q+2\).  Put \(m=t+d_q\).  The displayed identity becomes
\[
  S_q(m)=\sum_{i=1}^{d_q}c_{q,i}S_q(m-i)
  \qquad(m\ge d_q+2).
\]
This proves the recurrence in the displayed validity range.  The proof has used
the coefficient row only after the exact residual verifier has accepted it in the
kernel.

\Cref{prop:quotient-residual-equivalence} gives the same proof in
unobservable-subspace language: the identity
\(R_qP_q(\widetilde A_q)\widetilde v_q=0\) is exactly the finite construction
of the inclusion
\(P_q(\widetilde A_q)\widetilde v_q\in Z_2\), and membership in \(Z_2\) gives
the whole tail.  This theorem is therefore the proof of the recurrence rows in
Table~\ref{tab:appendix-recurrences-q9-17}.  No floating-point root data, and no
extrapolation from finitely many scalar values of \(S_q(m)\), enters the proof.
\end{proof}

\begin{table}[t]
\centering
\scriptsize
\setlength{\tabcolsep}{6pt}
\caption{Smith-normal-form certificates for Hankel principalization.  For each
\(q\), the displayed diagonal is the
Smith normal form of the change-of-basis matrix from the truncated-multiple
basis \(\{P_q,\lambda P_q,\dots,\lambda^{\kappa_q-1}P_q\}\) to an integral
basis of \(\mathrm{NullModes}(q)\).  In particular, every invariant factor is
\(1\), so the inclusion \(\mathcal M(P_q)\subseteq \mathrm{NullModes}(q)\) has
index \(1\).  The corresponding matrices \(U_q,V_q\) and the displayed
nullmode bases are supplied in the machine-checkable artifact of
\Cref{prop:pinned-normalizer-certificate-table}.}
\label{tab:appendix-principalization-q9-17}
\begin{tabular}{r r l}
\toprule
\(q\) & \(\kappa_q\) & Smith normal form \\
\midrule
9  & 2 & \(\operatorname{diag}(1,1)\) \\
10 & 2 & \(\operatorname{diag}(1,1)\) \\
11 & 2 & \(\operatorname{diag}(1,1)\) \\
12 & 0 & vacuous \(0\times 0\) case \\
13 & 2 & \(\operatorname{diag}(1,1)\) \\
14 & 2 & \(\operatorname{diag}(1,1)\) \\
15 & 4 & \(\operatorname{diag}(1,1,1,1)\) \\
16 & 4 & \(\operatorname{diag}(1,1,1,1)\) \\
17 & 4 & \(\operatorname{diag}(1,1,1,1)\) \\
\bottomrule
\end{tabular}
\end{table}

Concretely, the artifact field
\(\texttt{smith\_unimodular\_records}\) records, for each \(q\), the
change-of-basis matrix \(M_q\), unimodular integer matrices \(U_q,V_q\), and
the identity
\[
  U_qM_qV_q=I_{\kappa_q}
\]
(with the vacuous \(0\times0\) case for \(q=12\)).  Since every invariant factor
is \(1\), these matrices are the compact Smith certificates needed in
\Cref{thm:principalization}; the table above prints only their diagonal
summary.

\subsection{Exact certificate format for the arithmetic tables}

Each certificate used in Section~\ref{sec:chebotarev} is a finite identity in
an explicitly specified ring.  The recurrence certificate is the exact
residual certificate \(\mathsf W_q(P_q)=\mathrm{accept}\) of
\Cref{def:exact-kernel-verifier}, together with the scalar-free construction
lemma \Cref{lem:scalar-free-verifier}, the a priori certification theorem
\Cref{thm:apriori-exact-kernel-certification}, and the recurrence theorem
\Cref{prop:kernel-derived-recurrences}.  It starts from the exact
histogram kernel of \Cref{def:exact-histogram-kernel}, checks the exact integer
identity
\[
  R_qP_q(\widetilde A_q)\widetilde v_q=0,
\]
which is the finite Cayley--Hamilton construction of
\(P_q(\widetilde A_q)\widetilde v_q\in Z_2\).  Membership in \(Z_2\) gives the
all-tail recurrence.  The additional row-reduction identities
\(\rk(R_qK_q)=\rk(R_qB_q)=d_q\) form the observable-quotient certificate
\(\mathsf V_q(P_q)=\mathrm{accept}\); they identify the quotient polynomial and
support the Hankel and minimality arguments, but they are not required for the
tail recurrence.
Thus a recurrence row is used only after it has passed through the chain
\[
  \begin{array}{c}
  \text{histogram kernel}\\
  \Downarrow\\
  R_qP_q(\widetilde A_q)\widetilde v_q=0\\
  \Downarrow\ \text{membership in }Z_2\\
  P_q(E)S_q(j)=0\quad(j\ge2).
  \end{array}
\]

The remaining tables have the following exact meanings.

A Smith-normal-form certificate consists of integer unimodular matrices \(U,V\)
with
\[
  U M_q V=\operatorname{diag}(1,\dots,1)
\]
for the relevant basis-change matrix \(M_q\) (with the vacuous \(0\times0\)
case when \(\kappa_{12}=0\)).  A mod-\(p\) factorization certificate consists
of monic factors in \(\FF_p[\lambda]\) whose product is \(P_q\bmod p\), together
with Bezout polynomials certifying
\(\gcd(P_q\bmod p,P_q'\bmod p)=1\).  These data are the artifact fields
\(\texttt{finite\_field\_factorization\_records}\).  A
discriminant-squareclass certificate computes
\(\Disc(P_q)=(-1)^{d_q(d_q-1)/2}\operatorname{Res}(P_q,P_q')\) and then applies
Euler's criterion in the displayed finite fields; the artifact field
\(\texttt{discriminant\_squareclass\_records}\) records the exact
discriminants, residues, and Euler witnesses.  Thus the tables record digest
views of exact algebraic certificates; the high-precision root table below is
the only numerical table, and it is not used in the proofs.

\medskip
\noindent\textbf{Role in the paper.}
The proofs in Section~\ref{sec:chebotarev} use only the exact polynomials
\(P_q\) determined by the kernel-certified recurrences of
Table~\ref{tab:appendix-recurrences-q9-17} and
\Cref{prop:kernel-derived-recurrences}, the Smith-normal-form certificates
displayed in
\Cref{tab:appendix-principalization-q9-17} together with the recorded
unimodular matrices, and the finite mod-\(p\) factorization records and
discriminant witnesses summarized in
\Cref{tab:galois-certificates-q9-17,tab:disc-squareclass-q12-15}.  The
high-precision root fingerprints recorded in
\Cref{tab:secondary-spectral-q9-17} are included as secondary descriptive data,
but they are not used in the rigorous theorem chain.
The appendix is included to make these inputs explicit and citable inside the
submission itself, rather than relying on the secondary numerical fingerprints.

\subsection{Secondary spectral fingerprints}

The following table is intentionally separated from
Section~\ref{sec:chebotarev}.  It records the high-precision spectral picture
attached to the resonance polynomials \(P_q\) but is not used in any proof
above.

\begin{table}[t]
\centering
\scriptsize
\setlength{\tabcolsep}{5pt}
\caption{Secondary spectral fingerprints for the resonance polynomials
\(P_q\), \(q=9,\dots,17\).  Here \(\lambda_q\) is the common
Perron/partition root, \(\lambda_q^-<0\) is the outer negative real root, and
\(\Theta_q\) is the maximum modulus among the remaining roots.}
\label{tab:secondary-spectral-q9-17}
\begin{tabular}{r r r r r}
\toprule
\(q\) & \(d_q\) & \(\lambda_q\) & \(\lambda_q^-\) & \(\Theta_q\) \\
\midrule
9  & 7  & 11.778422 & -7.065331  & 5.807034 \\
10 & 9  & 14.771098 & -9.100142  & 7.686459 \\
11 & 9  & 18.535847 & -11.712939 & 10.074255 \\
12 & 13 & 23.273460 & -15.066709 & 13.109656 \\
13 & 11 & 29.237339 & -19.369971 & 16.968007 \\
14 & 13 & 36.747376 & -24.889362 & 21.870525 \\
15 & 11 & 46.207510 & -31.965663 & 28.096490 \\
16 & 13 & 58.127952 & -41.034237 & 35.998554 \\
17 & 13 & 73.153329 & -52.651060 & 46.021996 \\
\bottomrule
\end{tabular}
\end{table}

\begin{remark}[Status of the secondary spectral pattern]
\label{rem:secondary-spectral-pattern}
Table~\ref{tab:secondary-spectral-q9-17} records high-precision approximations
to the dominant root \(\lambda_q\), the outer negative real root
\(\lambda_q^-\), and the largest remaining modulus \(\Theta_q\) for
\(q=9,\dots,17\).  Throughout the displayed window the numerical fingerprints
are consistent with a unique negative real secondary root satisfying
\[
  \Theta_q<|\lambda_q^-|<\lambda_q.
\]
If this ordering is upgraded to a symbolic root-isolation certificate, then the
usual linear-recurrence expansion would yield an eventual alternating
second-order term for \(S_q(m)\).  At present, however, these data remain
descriptive fingerprints rather than symbolic isolation intervals, and
we therefore keep them outside the formal theorem chain.
\end{remark}

\begin{remark}[What remains open]
\label{rem:q16q17}
The exact certificate records above close the full symmetric-group statement for
\(q=16\) and \(q=17\).
The next arithmetic target is to extend squareclass independence and linear
disjointness beyond the block \(q=12,13,14,15\).  On the spectral side, the
table above strongly suggests a persistent negative second root throughout the
resonance window, but an entirely symbolic certification of that pattern would
be preferable.
\end{remark}
