\section{A padded bounded-delay transducer model for the finite-window fold}
\label{app:transducer}

To make the finite-state kernel construction of
\Cref{thm:collision-kernel,thm:symmetric-compression} self-contained, we spell
out a corrected transducer model for the finite-window fold.  The key point is
that the transducer output must carry the natural leading-zero padding; this is
exactly the convention needed to identify the transducer output with the
finite-window fold itself, even for the zero word.  The automata-theoretic
normalization viewpoint goes back to Frougny
\cite{Frougny1991FibonacciAutomata,Frougny1992NumbersAutomata}; see also
\cite{AhlbachUsatineFrougnyPippenger2013} for algorithmic background.

For a most-significant-digit-first binary word \(a=a_1\cdots a_m\), write
\[
  \operatorname{val}(a):=\sum_{j=1}^m a_jF_{m+2-j}.
\]
If \(\omega=\omega_1\cdots\omega_m\in\Omega_m\), then
\[
  \omega^{\mathrm{rev}}:=\omega_m\cdots\omega_1
\]
satisfies \(\operatorname{val}(\omega^{\mathrm{rev}})=N_m(\omega)\).

\begin{definition}[Padded normalization]
\label{def:appendix-padded-normalization}
For \(a=a_1\cdots a_m\in\{0,1\}^m\), let
\[
  \widetilde\Lambda(a)=b_0b_1\cdots b_m\in\{0,1\}^{m+1}
\]
be the unique length-\((m+1)\) word with no adjacent \(1\)'s and
\[
  \operatorname{val}(a)
  =
  \sum_{j=0}^{m} b_jF_{m+2-j}.
\]
Thus \(\widetilde\Lambda(a)\) is the Zeckendorf normalization of
\(\operatorname{val}(a)\), padded on the left to length \(m+1\).
\end{definition}

\begin{proposition}[Padded fold transducer]
\label{prop:appendix-padded-transducer}
There exists a deterministic subsequential transducer
\[
  \mathcal T=(Q,q_0,\delta,\lambda,\tau)
\]
with input alphabet \(\{0,1\}\), output alphabet \(\{0,1\}\), and bounded
terminal output length
\[
  L:=\max_{s\in Q}|\tau(s)|,
\]
such that the total output
\[
  \Lambda(a_1\cdots a_m)
  :=
  \lambda^*(q_0,a_1\cdots a_m)\,
  \tau(\delta^*(q_0,a_1\cdots a_m))
\]
is obtained from \(\widetilde\Lambda(a_1\cdots a_m)\) by deleting its first
symbol.
\end{proposition}

\begin{proof}
Deterministic subsequential transducers for Fibonacci normalization with
leading zeros permitted are standard in the automata-theoretic literature cited
above.  Deleting the first output symbol is a finite-state postprocessing, so
the same bounded-delay subsequential model also realizes the resulting visible
output map.  We fix one such transducer once and for all.
\end{proof}

\begin{lemma}[Direct fold output]
\label{lem:appendix-overflow-digit}
For every \(\omega\in\Omega_m\), if \(x=\Fold_m(\omega)\), then
\[
  \Lambda(\omega^{\mathrm{rev}})=x_m\cdots x_1.
\]
In particular,
\[
  \Lambda(0^m)=0^m.
\]
\end{lemma}

\begin{proof}
By \Cref{prop:single-overflow}, there exists a unique
\(\varepsilon_m(\omega)\in\{0,1\}\) such that
\[
  N_m(\omega)
  =
  \varepsilon_m(\omega)F_{m+2}+V_m(x)
  =
  \varepsilon_m(\omega)F_{m+2}+\sum_{j=1}^m x_jF_{j+1}.
\]
Hence the length-\((m+1)\) admissible word
\[
  \varepsilon_m(\omega)\,x_m\cdots x_1
\]
represents \(N_m(\omega)=\operatorname{val}(\omega^{\mathrm{rev}})\).  When
\(\varepsilon_m(\omega)=1\), one has
\[
  V_m(x)=N_m(\omega)-F_{m+2}<F_{m+1},
\]
so \(x_m=0\), and the displayed word is indeed admissible.  Therefore, by the
uniqueness in \Cref{def:appendix-padded-normalization},
\[
  \widetilde\Lambda(\omega^{\mathrm{rev}})
  =
  \varepsilon_m(\omega)\,x_m\cdots x_1.
\]
Deleting the first symbol gives the claimed identity for
\(\Lambda(\omega^{\mathrm{rev}})\).  The zero-word example is the special case
\(\omega=0^m\), for which \(x=0^m\).
\end{proof}

\begin{lemma}[Prefix stability for visible outputs]
\label{lem:appendix-prefix-stability}
If \(u\) is a prefix of \(v\) in \(\{0,1\}^*\), then \(\Lambda(u)\) and
\(\Lambda(v)\) have a common prefix of length at least
\[
  |\Lambda(u)|-L.
\]
Equivalently, after any processed prefix, at most the final \(L\) visible
symbols can still be altered by extending the input.
\end{lemma}

\begin{proof}
Write \(v=uw\), let \(s=\delta^*(q_0,u)\), and decompose the run as
\[
  \Lambda(u)=\alpha\,\tau(s),
  \qquad
  \Lambda(v)=\alpha\,\beta\,\tau(t)
\]
for some word \(\beta\) and some state \(t\).  Thus \(\Lambda(u)\) and
\(\Lambda(v)\) have a common prefix of length
\[
  |\alpha|=|\Lambda(u)|-|\tau(s)|\ge |\Lambda(u)|-L.
\]
\end{proof}

\begin{definition}[Committed and residual visible output]
\label{def:appendix-cut}
For a finite word \(y\in\{0,1\}^*\), write
\[
  y=\operatorname{com}(y)\,\operatorname{res}(y),
\]
where \(\operatorname{res}(y)\) is the suffix of \(y\) of length at most \(L\),
and \(\operatorname{com}(y)\) is the complementary prefix.
\end{definition}

\begin{proposition}[One-copy augmented fold automaton]
\label{prop:appendix-one-copy}
Set
\[
  \widehat Q:=Q\times\{0,1\}^{\le L}.
\]
There is a deterministic one-copy update rule on \(\widehat Q\) such that the
state \((s,r)\) records the current transducer state \(s\) together with the
residual visible suffix \(r\) that can still change under future extensions.
Its cardinality is
\[
  k:=\card{\widehat Q}
  =
  |Q|\sum_{j=0}^{L}2^j
  =
  |Q|\,(2^{L+1}-1).
\]
\end{proposition}

\begin{proof}
Given a one-copy augmented state \((s,r)\) and an input digit \(a\in\{0,1\}\),
form the temporary word
\[
  w:=r\,\lambda(s,a).
\]
Apply \Cref{def:appendix-cut} to write
\[
  w=c\,r',
  \qquad |r'|\le L.
\]
The update rule is
\[
  (s,r)\xrightarrow{\,a\,}(\delta(s,a),r'),
\]
and the newly committed visible block is \(c\).  By
\Cref{lem:appendix-prefix-stability}, every continuation of the current input
must begin its future visible output with \(c\); only the stored suffix \(r'\)
can still change.  The cardinality formula is immediate.
\end{proof}

\begin{example}[A synchronized \texorpdfstring{$q=2$}{q=2} update]
\label{ex:appendix-q2-update}
Suppose the two copies currently lie in
\[
  \bigl((s_1,r_1),(s_2,r_2)\bigr)\in \widehat Q^2,
\]
and the next input pair is \((a,b)\in\{0,1\}^2\).  Write
\[
  r_1\lambda(s_1,a)=c_1r_1',
  \qquad
  r_2\lambda(s_2,b)=c_2r_2',
  \qquad |r_1'|,|r_2'|\le L.
\]
Then the synchronized transition is admissible exactly when the newly
committed blocks agree:
\[
  c_1=c_2.
\]
In that case the next synchronized state is
\[
  \bigl((\delta(s_1,a),r_1'),(\delta(s_2,b),r_2')\bigr).
\]
At the terminal stage, if the final one-copy states are \((t_1,\rho_1)\) and
\((t_2,\rho_2)\), acceptance requires
\[
  \rho_1\tau(t_1)=\rho_2\tau(t_2).
\]
Thus the \(q=2\) synchronized machine checks equality of the newly committed
blocks at each step and equality of the final unresolved suffixes at the end.
\end{example}

\begin{proposition}[Finite collision automaton]
\label{prop:appendix-collision-automaton}
Fix \(q\ge 2\).  There is a finite automaton on \(\widehat Q^q\) with one
transition per input digit such that, for every \(m\ge 1\), accepted
length-\(m\) paths are in canonical bijection with \(q\)-tuples
\[
  (\omega^{(1)},\dots,\omega^{(q)})\in\Omega_m^q
\]
satisfying
\[
  \Fold_m(\omega^{(1)})=\cdots=\Fold_m(\omega^{(q)}).
\]
Consequently there exist a nonnegative integer matrix \(A_q\) and nonnegative
integer vectors \(u_q,v_q\) such that
\[
  S_q(m)=u_q^{\mathsf T}A_q^m v_q
  \qquad (m\ge 1).
\]
\end{proposition}

\begin{proof}
Run \(q\) copies of the one-copy update rule from
\Cref{prop:appendix-one-copy} in parallel.  A transition labelled by
\((a_1,\dots,a_q)\in\{0,1\}^q\) is admissible exactly when the corresponding
committed blocks agree in all coordinates.  Whenever they disagree, the visible
outputs can never be repaired by any later continuation, so the run enters a
rejecting sink.

After reading \(m\) digits, the \(i\)-th coordinate lies in a state
\((s_i,r_i)\).  The unresolved visible suffix equals
\[
  r_i\tau(s_i),
\]
whose length is at most \(2L\).  A terminal \(q\)-tuple is accepting exactly
when these final suffixes agree in all coordinates.

By construction and \Cref{lem:appendix-overflow-digit}, a synchronized run
avoids rejection and ends in an accepting terminal state if and only if the
full visible outputs \(\Lambda((\omega^{(i)})^{\mathrm{rev}})\) agree in all
coordinates, and this is equivalent to equality of the finite-window folds.
Therefore accepted paths of length \(m\) are exactly the \(q\)-tuples counted by
\(S_q(m)\).

Let \(A_q\) be the adjacency matrix of the non-sink part of this finite
automaton, let \(u_q\) select the unique initial synchronized state
\(((q_0,\varepsilon),\dots,(q_0,\varepsilon))\), and let \(v_q\) mark the
accepting terminal states.  Then the accepted path count is
\[
  u_q^{\mathsf T}A_q^m v_q=S_q(m).
\]
\end{proof}

The proof of \Cref{thm:symmetric-compression} is now immediate from
\Cref{prop:appendix-collision-automaton}: one simply replaces the one-copy state
set \(Q\) there by the finite augmented state set \(\widehat Q\).  The histogram
quotient therefore acts on weak compositions of \(q\) into \(k=\card{\widehat Q}\)
parts, so its dimension remains polynomial in \(q\).
