% main.tex -- Cayley transform, Poisson entropy, and reconstruction
\documentclass[12pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{microtype}
\usepackage{geometry}
\geometry{margin=2.8cm}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}

\numberwithin{equation}{section}

\DeclareMathOperator{\im}{im}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\DD}{\mathbb{D}}
\newcommand{\UU}{\mathcal{U}}
\newcommand{\dd}{\,\mathrm{d}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\norm}[1]{\lVert #1\rVert}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\mathscr}[1]{\mathcal{#1}}

\title{Poisson Entropy Asymptotics via Cayley--Chebyshev Modes}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We study the large-time relative entropy of a Poisson average
$P_t*\nu$ relative to the translated Poisson kernel
$P_t(\cdot-\bar\gamma)$ with matching mean.  The Cayley transform
normalizes the problem by pulling normalized Haar measure on the circle
back to the Cauchy law on~$\RR$, and this normalization is rigid among
orientation-preserving $C^1$ charts.  In the Cayley variable the ratio
\[
R_\varepsilon(y)=\frac{1+y^2}{1+(y-\varepsilon)^2}
\]
has a finite Chebyshev--Fourier mode calculus.  The mode parity eliminates
odd powers through the finite expansion order, and the uniform remainder
estimates give an expansion of
$D_{\mathrm{KL}}(P_t*\nu\,\|\,P_t(\cdot-\bar\gamma))$
through order~$t^{-8}$ under a finite centred seventh absolute moment.
The sixth- and eighth-order coefficients admit explicit non-negative
defect coordinates.  The eighth-order defect strictly amplifies the
sixth-order one, with equality only for the symmetric two-point law, and
the sixth-order asymptotic defect yields a coefficient-level quantitative
$W_2$ stability theorem together with a finite-time corollary after the
explicit $C_8t^{-2}$ correction.  The finite phase section supplies the
classical root-of-unity sampling interface needed to recover the Cayley
Fourier coefficients from exact finite grids.  The same secant mode family
then yields the Poisson strip space $\sqrt{\pi}\,e^{-|D|}L^2(\RR)$; for
the integer orbit, the standard positive-symbol criterion becomes a closed
hyperbolic lattice symbol with explicit cardinal and interpolation-norm
formulas.
\end{abstract}

\smallskip
\noindent\textbf{Keywords:}
Cayley transform, Haar measure, Poisson semigroup, relative entropy,
Chebyshev polynomials, reproducing kernels, Hardy spaces, sampling
theorem, cardinal interpolation.

\smallskip
\noindent\textbf{MSC 2020:}
94A17, 47D07, 46E22, 42C15, 30H10.

\input{sec_introduction}
\input{sec_preliminaries}
\input{sec_cayley_gate}
\input{sec_haar_pullback}
\input{sec_finite_phase_sampling}
\input{sec_entropy_asymptotics}
\input{sec_doob_phi_entropy}
\input{sec_gram_space}
\input{sec_strip_lattice}
\appendix
\input{sec_appendix}

\begin{thebibliography}{99}

\bibitem{Aronszajn1950}
N.~Aronszajn.
\newblock Theory of reproducing kernels.
\newblock \emph{Trans. Amer. Math. Soc.} 68 (1950), no.~3, 337--404.

\bibitem{ArtsteinBallBartheNaor2004}
S.~Artstein, K.~M. Ball, F.~Barthe, and A.~Naor.
\newblock On the rate of convergence in the entropic central limit theorem.
\newblock \emph{Probab. Theory Related Fields} 129 (2004), no.~3, 381--390.

\bibitem{AldroubiGrochenig2001}
A.~Aldroubi and K.~Gr{\"o}chenig.
\newblock Nonuniform sampling and reconstruction in shift-invariant spaces.
\newblock \emph{SIAM Rev.} 43 (2001), no.~4, 585--620.

\bibitem{Barron1986}
A.~R. Barron.
\newblock Entropy and the central limit theorem.
\newblock \emph{Ann. Probab.} 14 (1986), no.~1, 336--342.

\bibitem{Buhmann2003}
M.~D. Buhmann.
\newblock \emph{Radial Basis Functions: Theory and Implementations}.
\newblock Cambridge University Press, Cambridge, 2003.

\bibitem{BobkovChistyakovGotze2013}
S.~G. Bobkov, G.~P. Chistyakov, and F.~G\"otze.
\newblock Convergence to stable laws in relative entropy.
\newblock \emph{J. Theoret. Probab.} 26 (2013), no.~3, 803--818.

\bibitem{BobkovChistyakovGotzeBerryEsseen2014}
S.~G. Bobkov, G.~P. Chistyakov, and F.~G\"otze.
\newblock Berry--Esseen bounds in the entropic central limit theorem.
\newblock \emph{Probab. Theory Related Fields} 159 (2014), nos.~3--4,
435--478.

\bibitem{BobkovChistyakovGotzeEdgeworth2013}
S.~G. Bobkov, G.~P. Chistyakov, and F.~G\"otze.
\newblock Rate of convergence and Edgeworth-type expansion in the entropic
central limit theorem.
\newblock \emph{Ann. Probab.} 41 (2013), no.~4, 2479--2512.

\bibitem{BrigatiDolbeaultSimonov2024LogSobolev}
G.~Brigati, J.~Dolbeault, and N.~Simonov.
\newblock Stability for the logarithmic Sobolev inequality.
\newblock \emph{J. Funct. Anal.} 287 (2024), no.~8, 110562.
\newblock doi:10.1016/j.jfa.2024.110562.

\bibitem{CarlenCarvalho1992}
E.~A. Carlen and M.~C. Carvalho.
\newblock Strict entropy production bounds and stability of the rate of
convergence to equilibrium for the Boltzmann equation.
\newblock \emph{J. Stat. Phys.} 67 (1992), nos.~3--4, 575--608.

\bibitem{Carlen2025LogHLS}
E.~A. Carlen.
\newblock Stability for the logarithmic Hardy--Littlewood--Sobolev inequality
with application to the Keller--Segel equation.
\newblock \emph{J. Funct. Anal.} 288 (2025), no.~6, 110818.
\newblock doi:10.1016/j.jfa.2024.110818.

\bibitem{ChengJiangSun2016}
C.~Cheng, Y.~Jiang, and Q.~Sun.
\newblock Sampling and Galerkin reconstruction in reproducing kernel spaces.
\newblock \emph{Appl. Comput. Harmon. Anal.} 41 (2016), no.~2, 638--659.

\bibitem{Bownik2000}
M.~Bownik.
\newblock The structure of shift-invariant subspaces of
{$L^2(\mathbb{R}^n)$}.
\newblock \emph{J. Funct. Anal.} 177 (2000), no.~2, 282--309.

\bibitem{Folland1995}
G.~B. Folland.
\newblock \emph{A Course in Abstract Harmonic Analysis}.
\newblock CRC Press, Boca Raton, FL, 1995.

\bibitem{Feller1971}
W.~Feller.
\newblock \emph{An Introduction to Probability Theory and Its Applications},
Vol.~II.
\newblock 2nd ed., John Wiley \& Sons, New York, 1971.

\bibitem{Duren1970}
P.~L. Duren.
\newblock \emph{Theory of {$H^p$} Spaces}.
\newblock Academic Press, New York, 1970.

\bibitem{deBoorDeVoreRon1994}
C.~de~Boor, R.~DeVore, and A.~Ron.
\newblock The structure of finitely generated shift-invariant spaces in
{$L_2(\mathbb{R}^d)$}.
\newblock \emph{J. Funct. Anal.} 119 (1994), no.~1, 37--78.

\bibitem{Grafakos2014}
L.~Grafakos.
\newblock \emph{Classical Fourier Analysis}.
\newblock 3rd ed., Springer, New York, 2014.

\bibitem{JohnsonBarron2004}
O.~T. Johnson and A.~R. Barron.
\newblock Fisher information inequalities and the central limit theorem.
\newblock \emph{Probab. Theory Related Fields} 129 (2004), no.~3, 391--409.

\bibitem{JohnsonStableDeBruijn2013}
O.~Johnson.
\newblock A de Bruijn identity for symmetric stable laws.
\newblock \emph{arXiv:1310.2045} (2013).

\bibitem{FuhrGrochenigHaimiKlotzRomero2018}
H.~F{\"u}hr, K.~Gr{\"o}chenig, A.~Haimi, A.~Klotz, and J.~L. Romero.
\newblock Density of sampling and interpolation in reproducing kernel Hilbert
spaces.
\newblock \emph{J. Lond. Math. Soc.} (2) 96 (2017), no.~3, 663--686.

\bibitem{GrochenigRomeroStockler2020}
K.~Gr{\"o}chenig, J.~L. Romero, and J.~St{\"o}ckler.
\newblock Sharp results on sampling with derivatives in shift-invariant spaces
and multi-window Gabor frames.
\newblock \emph{Constr. Approx.} 51 (2020), no.~1, 1--25.

\bibitem{Lukacs1970}
E.~Lukacs.
\newblock \emph{Characteristic Functions}.
\newblock 2nd ed., Charles Griffin, London, 1970.

\bibitem{NashedSun2010}
M.~Z. Nashed and Q.~Sun.
\newblock Sampling and reconstruction of signals in a reproducing kernel
subspace of {$L^p(\mathbb{R}^d)$}.
\newblock \emph{J. Funct. Anal.} 258 (2010), no.~7, 2422--2452.

\bibitem{NashedWalter1991}
M.~Z. Nashed and G.~G. Walter.
\newblock General sampling theorems for functions in reproducing kernel
Hilbert spaces.
\newblock \emph{Math. Control Signals Systems} 4 (1991), no.~4, 363--390.

\bibitem{NielsenOkamura2023Cauchy}
F.~Nielsen and K.~Okamura.
\newblock On {$f$}-divergences between Cauchy distributions.
\newblock \emph{IEEE Trans. Inform. Theory} 69 (2023), no.~5, 3150--3171.

\bibitem{Rudin1987}
W.~Rudin.
\newblock \emph{Real and Complex Analysis}.
\newblock 3rd ed., McGraw--Hill, New York, 1987.

\bibitem{RonShen1995}
A.~Ron and Z.~Shen.
\newblock Frames and stable bases for shift-invariant subspaces of
{$L^2(\mathbb{R}^d)$}.
\newblock \emph{Canad. J. Math.} 47 (1995), no.~5, 1051--1094.

\bibitem{KiselevMininNovikov2016}
E.~A. Kiselev, L.~A. Minin, and I.~Ya. Novikov.
\newblock Calculation of the Riesz constants and orthogonalization for
incomplete systems of coherent states by means of theta functions.
\newblock \emph{Sb. Math.} 207 (2016), no.~8, 1127--1141.

\bibitem{KiselevMininNovikov2019}
E.~A. Kiselev, L.~A. Minin, and I.~Ya. Novikov.
\newblock Limit properties of systems of integer translates and functions
generating tight Gabor frames.
\newblock \emph{Math. Notes} 106 (2019), no.~1, 71--80.

\bibitem{Kwasnicki2017}
M.~Kwa{\'s}nicki.
\newblock Ten equivalent definitions of the fractional Laplace operator.
\newblock \emph{Fract. Calc. Appl. Anal.} 20 (2017), no.~1, 7--51.
\newblock doi:10.1515/fca-2017-0002.

\bibitem{RomeroUlanovskiiZlotnikov2024}
J.~L. Romero, A.~Ulanovskii, and I.~Zlotnikov.
\newblock Sampling in the shift-invariant space generated by the bivariate
Gaussian function.
\newblock \emph{J. Funct. Anal.} 287 (2024), no.~9, 110600.

\bibitem{Stein1971}
E.~M. Stein and G.~Weiss.
\newblock \emph{Introduction to Fourier Analysis on Euclidean Spaces}.
\newblock Princeton University Press, Princeton, NJ, 1971.

\bibitem{Toscani2016}
G.~Toscani.
\newblock Entropy inequalities for stable densities and strengthened central
limit theorems.
\newblock \emph{J. Stat. Phys.} 164 (2016), no.~6, 371--389.

\bibitem{BaranovBelovGrochenig2022}
A.~Baranov, Y.~Belov, and K.~Gr{\"o}chenig.
\newblock Complete interpolating sequences for the Gaussian shift-invariant
space.
\newblock \emph{Appl. Comput. Harmon. Anal.} 61 (2022), 191--201.

\bibitem{Verdu2023Cauchy}
S.~Verd{\'u}.
\newblock The Cauchy distribution in information theory.
\newblock \emph{Entropy} 25 (2023), no.~2, 346.

\bibitem{Widder1941}
D.~V. Widder.
\newblock \emph{The Laplace Transform}.
\newblock Princeton University Press, Princeton, NJ, 1941.

\bibitem{Toscani2015}
G.~Toscani.
\newblock The fractional Fisher information and the central limit theorem for
stable laws.
\newblock \emph{Ricerche Mat.} 65 (2016), no.~1, 71--91.

\end{thebibliography}

\end{document}
