% sec_gram_space.tex -- Mode Gram kernel, RKHS completion, and centered reconstruction
\section{Mode Gram Space and Centered Reconstruction}%
\label{sec:gram-space}

\subsection{The mode Gram kernel and Hilbert completion}

Let
\[
\omega(\dd y):=\frac{\dd y}{\pi(1+y^2)}.
\]
For $\varepsilon\in\RR$ define the secant family
\begin{equation}\label{eq:mode-secant-family}
\Psi_\varepsilon(y):=\frac{R_\varepsilon(y)-1}{\varepsilon}
=\frac{2y-\varepsilon}{1+(y-\varepsilon)^2},
\qquad y\in\RR,
\end{equation}
where the formula remains valid at $\varepsilon=0$ and gives
\[
\Psi_0(y)=u_1(y)=\frac{2y}{1+y^2}.
\]
Since
\[
R_\varepsilon(y)\,\omega(\dd y)=P_1(y-\varepsilon)\,\dd y,
\]
one has
\[
\int_{\RR}R_\varepsilon(y)\,\omega(\dd y)=1
\qquad(\varepsilon\in\RR),
\]
and therefore $\Psi_\varepsilon\in L^2_0(\omega)$ for every
$\varepsilon\in\RR$.

\begin{theorem}[The mode Gram kernel]%
\label{thm:mode-gram-kernel}
For every $\varepsilon,\eta\in\RR$,
\begin{equation}\label{eq:mode-gram-kernel}
\langle \Psi_\varepsilon,\Psi_\eta\rangle_{L^2(\omega)}
=\frac{2}{4+(\varepsilon-\eta)^2}.
\end{equation}
Equivalently,
\begin{equation}\label{eq:mode-gram-RR}
\int_{\RR}R_\varepsilon(y)R_\eta(y)\,\omega(\dd y)
=\frac{\varepsilon^2+\eta^2+4}{(\varepsilon-\eta)^2+4}.
\end{equation}
\end{theorem}

\begin{proof}
For $\varepsilon\neq\eta$ define
\[
I(\varepsilon,\eta)
:=\frac{1}{\pi}\int_{\RR}
\frac{1+y^2}{(1+(y-\varepsilon)^2)(1+(y-\eta)^2)}\,\dd y.
\]
The integrand is $O(|z|^{-2})$ on the upper half-plane and has simple
poles at $z=\varepsilon+\ii$ and $z=\eta+\ii$.  A direct residue
computation gives
\[
\operatorname{Res}_{z=\varepsilon+\ii}
=\frac{\varepsilon(\varepsilon+2\ii)}
{2\ii(\varepsilon-\eta)(\varepsilon-\eta+2\ii)},
\]
\[
\operatorname{Res}_{z=\eta+\ii}
=\frac{\eta(\eta+2\ii)}
{2\ii(\eta-\varepsilon)(\eta-\varepsilon+2\ii)}.
\]
Hence, by the residue theorem,
\[
I(\varepsilon,\eta)
=2\ii\Bigl(
\operatorname{Res}_{z=\varepsilon+\ii}
+\operatorname{Res}_{z=\eta+\ii}
\Bigr)
=\frac{\varepsilon^2+\eta^2+4}{(\varepsilon-\eta)^2+4}.
\]
This proves \eqref{eq:mode-gram-RR} for $\varepsilon\neq\eta$, and the
diagonal case follows by continuity.

Since
\[
\int_{\RR}R_\varepsilon(y)\,\omega(\dd y)=1,
\]
one obtains
\[
\int_{\RR}(R_\varepsilon(y)-1)(R_\eta(y)-1)\,\omega(\dd y)
=\frac{2\varepsilon\eta}{4+(\varepsilon-\eta)^2}.
\]
For $\varepsilon\eta\neq 0$, division by $\varepsilon\eta$ yields
\eqref{eq:mode-gram-kernel}; the remaining cases follow by continuity in
$(\varepsilon,\eta)$.
\end{proof}

\begin{corollary}[Closed formulas for quadratic mode integrals]%
\label{cor:mode-quadratic-integrals}
For $|\varepsilon|<1$ one has the $L^2(\omega)$ expansion
\[
\Psi_\varepsilon
=\sum_{n=1}^{\infty}u_n\,\varepsilon^{n-1}.
\]
Consequently, for every $p,q\ge 1$,
\[
\langle u_{2p},u_{2q}\rangle_{L^2(\omega)}
=(-1)^{p+q}2^{-(2p+2q-1)}
\binom{2p+2q-2}{2p-1},
\]
for every $p,q\ge 0$,
\[
\langle u_{2p+1},u_{2q+1}\rangle_{L^2(\omega)}
=(-1)^{p+q}2^{-(2p+2q+1)}
\binom{2p+2q}{2p},
\]
and
\[
\langle u_{2p},u_{2q+1}\rangle_{L^2(\omega)}=0.
\]
\end{corollary}

\begin{proof}
Use $y=\tan\theta$ to identify $L^2(\omega)$ with
$L^2((-\pi/2,\pi/2),\dd\theta/\pi)$.  From
Theorem~\ref{thm:cayley-chebyshev-mode},
\[
u_n(\tan\theta)=\cos^n\theta\,U_n(\sin\theta).
\]
For $s\in[-1,1]$ the Chebyshev polynomials of the second kind satisfy
$|U_n(s)|\le n+1$.  Hence, for every $0<r<1$ and
$|\varepsilon|\le r$,
\[
\sum_{n=1}^{\infty}
\sup_{\theta}\bigl|u_n(\tan\theta)\varepsilon^{n-1}\bigr|
\le
\sum_{n=1}^{\infty}(n+1)r^{n-1}<\infty.
\]
The Weierstrass test gives uniform, and therefore $L^2(\omega)$,
convergence.  The Chebyshev generating function used in
Theorem~\ref{thm:cayley-chebyshev-mode} is valid because
$|\varepsilon\cos\theta|<1$, so the uniform limit is
$(R_\varepsilon-1)/\varepsilon=\Psi_\varepsilon$.

Moreover $\|u_n\|_{L^2(\omega)}\le n+1$, and therefore, for
$|\varepsilon|,|\eta|\le r<1$,
\[
\sum_{m,n\ge1}
\|u_m\|_{L^2(\omega)}\|u_n\|_{L^2(\omega)}
|\varepsilon|^{m-1}|\eta|^{n-1}<\infty.
\]
Thus the inner product of the two $L^2$ series may be expanded
absolutely:
\[
\frac{2}{4+(\varepsilon-\eta)^2}
=\sum_{m,n\ge 1}
\langle u_m,u_n\rangle_{L^2(\omega)}\,
\varepsilon^{m-1}\eta^{n-1}.
\]
On the other hand,
\[
\frac{2}{4+(\varepsilon-\eta)^2}
=\frac12\sum_{\ell=0}^{\infty}
(-1)^\ell 4^{-\ell}(\varepsilon-\eta)^{2\ell}.
\]
Comparing coefficients gives the stated formulas.  Mixed parity terms
vanish because the kernel only contains even total degree.
\end{proof}

\begin{theorem}[The mode space and its RKHS completion]%
\label{thm:mode-space-rkhs}
Let
\[
L^2_0(\omega):=
\left\{
f\in L^2(\omega):\ \int_{\RR}f(y)\,\omega(\dd y)=0
\right\},
\]
and let $\mathcal H_K$ be the reproducing-kernel Hilbert space generated
by the kernel
\[
K(a,b):=\frac{2}{4+(a-b)^2}.
\]
Set
\[
\mathcal S_0:=\operatorname{span}\{\Psi_\varepsilon:\varepsilon\in\RR\},
\qquad
\mathcal S:=\overline{\mathcal S_0}^{\,L^2(\omega)}
\subset L^2_0(\omega),
\]
Then
\[
\mathcal S=L^2_0(\omega).
\]
Equivalently, the algebraic span
$\mathcal S_0$ is dense in~$L^2_0(\omega)$.
Moreover, the linear assignment
\[
\mathcal U_0:\mathcal S_0\longrightarrow \mathcal H_K,
\qquad
\mathcal U_0(\Psi_\varepsilon):=K(\cdot,\varepsilon)
\qquad(\varepsilon\in\RR)
\]
is an isometry and extends uniquely to a unitary operator
\[
\mathcal U:L^2_0(\omega)\longrightarrow \mathcal H_K.
\]
\end{theorem}

\begin{proof}
We already noted that $\Psi_\varepsilon\in L^2_0(\omega)$, so
$\mathcal S_0\subset L^2_0(\omega)$.  Let
$f\in L^2_0(\omega)$ be orthogonal to~$\mathcal S_0$.  Then for every
$\varepsilon\in\RR$,
\[
0=\langle f,\Psi_\varepsilon\rangle_{L^2(\omega)}.
\]
If $\varepsilon\neq 0$, multiplying by~$\varepsilon$ gives
\[
0=\int_{\RR}f(y)\bigl(R_\varepsilon(y)-1\bigr)\,\omega(\dd y)
=\int_{\RR}f(y)R_\varepsilon(y)\,\omega(\dd y),
\]
because $\int f\,\dd\omega=0$.  Since
$R_\varepsilon(y)\,\omega(\dd y)=P_1(y-\varepsilon)\,\dd y$, this means
\[
0=\int_{\RR}f(y)P_1(\varepsilon-y)\,\dd y
\qquad(\varepsilon\in\RR).
\]
At $\varepsilon=0$ this is exactly the zero-mean condition.
In other words, the Poisson regularization $F:=P_1*f$ vanishes
identically as a function on~$\RR$.  We now justify the injectivity in the
present weighted space.  Since $f\in L^2(\omega)$,
\[
\left|\int_{\RR} f(y)\varphi(y)\dd y\right|
\le
\left(\int_{\RR}\frac{|f(y)|^2}{1+y^2}\dd y\right)^{1/2}
\left(\int_{\RR}(1+y^2)|\varphi(y)|^2\dd y\right)^{1/2}
\]
for every $\varphi\in\mathscr S(\RR)$.  Thus $f$ defines a tempered
distribution.  The same bound applies to any test factor with finite
weighted norm $\int(1+y^2)|q(y)|^2\dd y$.  In particular,
$P_1*\varphi$ has this property for $\varphi\in\mathscr S(\RR)$, because
$P_1*\varphi(y)=O((1+|y|)^{-2})$ at infinity.  Hence the convolution
identity above agrees with the distributional convolution by $P_1$: for
$\varphi\in\mathscr S(\RR)$, Fubini is justified by this weighted estimate,
and
\[
\langle F,\varphi\rangle
=\langle f,\widetilde P_1*\varphi\rangle,
\qquad \widetilde P_1(x):=P_1(-x)=P_1(x).
\]
Using $\widehat{P_1}(\xi)=e^{-|\xi|}$, we obtain the Fourier-side identity
away from the nonsmooth point of the multiplier:
\[
\langle \widehat F,\psi\rangle
=\langle \widehat f,e^{-|\xi|}\psi\rangle,
\qquad \psi\in C_c^\infty(\RR\setminus\{0\}).
\]
Since $F=0$, the left-hand side vanishes.  If
$\phi\in C_c^\infty(\RR\setminus\{0\})$, then
$\psi=e^{|\xi|}\phi$ is again a compactly supported smooth test function
with support away from~$0$, and hence
\[
\langle\widehat f,\phi\rangle
=\langle\widehat f,e^{-|\xi|}\psi\rangle
=0.
\]
Therefore $\operatorname{supp}\widehat f\subset\{0\}$.  By the standard
structure theorem for distributions supported at one point
\cite{Grafakos2014}, there are constants $a_0,\dots,a_M$ such that
\[
\widehat f=\sum_{j=0}^{M}a_j\delta_0^{(j)}.
\]
After inverse Fourier transform, the distribution $f$ is represented by a
polynomial.  The weighted condition $f\in L^2(\omega)$ excludes every
nonconstant polynomial, since a nonzero polynomial of degree $d\ge1$ gives
an integrand comparable to $|y|^{2d-2}$ at infinity.  Thus $f$ is constant
almost everywhere.  The zero-mean condition
$\int_{\RR}f\,\omega(\dd y)=0$ forces that constant to be zero.  Therefore
$\mathcal S_0^\perp=\{0\}$ in $L^2_0(\omega)$, so
\[
\overline{\mathcal S_0}^{\,L^2(\omega)}=L^2_0(\omega).
\]
By definition this closure is~$\mathcal S$, hence
$\mathcal S=L^2_0(\omega)$.

Theorem~\ref{thm:mode-gram-kernel} shows that
\[
\langle \Psi_\varepsilon,\Psi_\eta\rangle_{L^2(\omega)}
=K(\varepsilon,\eta)
=\langle K(\cdot,\varepsilon),K(\cdot,\eta)\rangle_{\mathcal H_K},
\]
so $\mathcal U_0$ is an isometry on~$\mathcal S_0$.  Since the
kernel sections span a dense subspace of~$\mathcal H_K$, the isometry
extends uniquely to a surjective isometry
$\mathcal U:L^2_0(\omega)\to\mathcal H_K$, that is, to a unitary
operator.
\end{proof}

\subsection{Higher secant kernels and coefficient representation}

For the kernel under consideration, the multilinear integrals appearing
in the entropy expansion from Section~\ref{sec:precision-potential} can
be organized as derivatives of explicit multilinear kernels built from
the same secant family.

\begin{theorem}[Higher secant kernels and coefficient representation]
\label{thm:higher-secant-kernels}
For each integer $m\ge 2$ define
\[
\mathcal J_m(\varepsilon_1,\dots,\varepsilon_m)
:=
\int_{\RR}\prod_{j=1}^m R_{\varepsilon_j}(y)\,\omega(\dd y),
\]
\[
\Gamma_m(\varepsilon_1,\dots,\varepsilon_m)
:=
\int_{\RR}\prod_{j=1}^m \Psi_{\varepsilon_j}(y)\,\omega(\dd y).
\]
Then the following hold.
\begin{enumerate}[label=(\roman*)]
\item If $\varepsilon_1,\dots,\varepsilon_m$ are pairwise distinct, then
      \begin{equation}\label{eq:higher-secant-residue}
      \mathcal J_m(\varepsilon_1,\dots,\varepsilon_m)
      =
      \sum_{j=1}^m
      \frac{\varepsilon_j(\varepsilon_j+2\ii)}
      {\displaystyle\prod_{\ell\ne j}
      (\varepsilon_j-\varepsilon_\ell)
      (\varepsilon_j-\varepsilon_\ell+2\ii)}.
      \end{equation}
\item The function $\Gamma_m$ is continuous on $\RR^m$, is
      real-analytic in a neighborhood of $(0,\dots,0)$, and for
      nonzero pairwise distinct parameters satisfies
      \begin{equation}\label{eq:higher-secant-inclusion-exclusion}
      \Gamma_m(\varepsilon_1,\dots,\varepsilon_m)
      =
      \frac{1}{\varepsilon_1\cdots\varepsilon_m}
      \sum_{S\subseteq\{1,\dots,m\}}(-1)^{m-|S|}
      \mathcal J_{|S|}\bigl((\varepsilon_j)_{j\in S}\bigr),
      \end{equation}
      with the conventions $\mathcal J_0:=1$ and
      $\mathcal J_1(\varepsilon):=1$.  For $m=2$ one has
      \[
      \Gamma_2(\varepsilon_1,\varepsilon_2)=K(\varepsilon_1,\varepsilon_2).
      \]
\item Let $N\ge 2$, and assume that $\nu$ has finite centered
      $(2N+1)$st absolute moment.  If
      \[
      D_{\mathrm{KL}}(h_t\|g_t)
      =\sum_{r=2}^{N}c_{2r}\,t^{-2r}+o(t^{-2N})
      \qquad(t\to\infty)
      \]
      is the expansion from
      Theorem~\ref{thm:poisson-kl-even-orders}, then
      \begin{equation}\label{eq:higher-secant-coefficients}
      c_{2N}
      =
      \sum_{k=2}^{N}\frac{(-1)^k}{k(k-1)}
      \sum_{\substack{n_1,\dots,n_k\ge 2\\n_1+\cdots+n_k=2N}}
      \mu_{n_1}\cdots\mu_{n_k}\,
      \frac{
      \partial_1^{n_1-1}\cdots \partial_k^{n_k-1}
      \Gamma_k(0,\dots,0)}
      {(n_1-1)!\cdots(n_k-1)!}.
      \end{equation}
      Consequently, within this model every coefficient in the Poisson
      entropy expansion is recovered as a Taylor coefficient of the
      higher secant kernels $\Gamma_k$.
\end{enumerate}
\end{theorem}

\begin{proof}
Fix $m\ge 2$.  For pairwise distinct
$\varepsilon_1,\dots,\varepsilon_m\in\RR$ set
\[
F_m(z)
:=
\frac{1+z^2}{\prod_{j=1}^m(1+(z-\varepsilon_j)^2)}.
\]
Since $F_m(z)=O(|z|^{2-2m})$ in the upper half-plane, the residue
theorem applies on semicircles of radius tending to infinity.  The poles
in the upper half-plane are the simple poles
$z=\varepsilon_j+\ii$ $(1\le j\le m)$.  For each such $j$,
\[
\operatorname{Res}_{z=\varepsilon_j+\ii}F_m
=
\frac{1+(\varepsilon_j+\ii)^2}
{2\ii\displaystyle\prod_{\ell\ne j}
\bigl(1+(\varepsilon_j+\ii-\varepsilon_\ell)^2\bigr)}.
\]
Because
\[
1+(\varepsilon_j+\ii)^2=\varepsilon_j(\varepsilon_j+2\ii)
\]
and
\[
1+(\varepsilon_j+\ii-\varepsilon_\ell)^2
=(\varepsilon_j-\varepsilon_\ell)
(\varepsilon_j-\varepsilon_\ell+2\ii),
\]
the residue identity becomes
\[
\operatorname{Res}_{z=\varepsilon_j+\ii}F_m
=
\frac{\varepsilon_j(\varepsilon_j+2\ii)}
{2\ii\displaystyle\prod_{\ell\ne j}
(\varepsilon_j-\varepsilon_\ell)
(\varepsilon_j-\varepsilon_\ell+2\ii)}.
\]
Hence
\[
\mathcal J_m(\varepsilon_1,\dots,\varepsilon_m)
=\frac{1}{\pi}\int_{\RR}F_m(y)\,\dd y
=2\ii\sum_{j=1}^m \operatorname{Res}_{z=\varepsilon_j+\ii}F_m,
\]
which is exactly \eqref{eq:higher-secant-residue}.  This proves~(i).

For~(ii), continuity of $\Gamma_m$ on $\RR^m$ follows from dominated
convergence on compact parameter sets.  Indeed, if
$|\varepsilon_j|\le M$ for every~$j$, then
\[
\Psi_{\varepsilon_j}(y)
=\frac{2(y-\varepsilon_j)+\varepsilon_j}{1+(y-\varepsilon_j)^2},
\]
so
\[
\abs{\Psi_{\varepsilon_j}(y)}
\le \frac{2\abs{y-\varepsilon_j}}{1+(y-\varepsilon_j)^2}
+\frac{|\varepsilon_j|}{1+(y-\varepsilon_j)^2}
\le 1+M
\qquad(y\in\RR).
\]
Since $\omega$ is a probability measure,
$\prod_{j=1}^m\Psi_{\varepsilon_j}\in L^1(\omega)$ uniformly on such
compact sets.

If every $\varepsilon_j$ is nonzero, then
\[
\Psi_{\varepsilon_j}=\frac{R_{\varepsilon_j}-1}{\varepsilon_j}.
\]
Multiplying these identities and integrating against $\omega$ gives
\eqref{eq:higher-secant-inclusion-exclusion} by expanding the product
\[
\prod_{j=1}^m(R_{\varepsilon_j}-1)
\]
and using the conventions $\mathcal J_0:=1$ and
$\mathcal J_1(\varepsilon):=1$.

To prove the analyticity statement near the origin, fix $0<\rho<1$.
Proposition~\ref{prop:cayley-mode-remainder} yields, for every integer
$N\ge 1$,
\[
\Psi_\varepsilon(y)
=\sum_{n=1}^{N}u_n(y)\varepsilon^{n-1}
+\varepsilon^N v_N(y,\varepsilon),
\qquad |\varepsilon|\le \rho,
\]
where $v_N$ is bounded on $\RR\times[-\rho,\rho]$.  Since
$|u_n(y)|\le 1$ for every $n\ge 1$ and $y\in\RR$, multiplying the
preceding expansions for $\Psi_{\varepsilon_1},\dots,\Psi_{\varepsilon_m}$
shows that
\[
\Gamma_m(\varepsilon_1,\dots,\varepsilon_m)
=
\sum_{n_1,\dots,n_m=1}^{N}
I_{n_1,\dots,n_m}\,
\varepsilon_1^{n_1-1}\cdots\varepsilon_m^{n_m-1}
+\sum_{j=1}^m\varepsilon_j^N G_{j,N}(\varepsilon_1,\dots,\varepsilon_m),
\]
where each $G_{j,N}$ is bounded on $[-\rho,\rho]^m$ and
\[
I_{n_1,\dots,n_m}
:=
\int_{\RR}
u_{n_1}(y)\cdots u_{n_m}(y)\,\omega(\dd y).
\]
This is exactly the Taylor expansion of $\Gamma_m$ at the origin, so
$\Gamma_m$ is real-analytic there.  The identity $\Gamma_2=K$ is
Theorem~\ref{thm:mode-gram-kernel}.

For~(iii), the Taylor expansion from~(ii) gives
\[
I_{n_1,\dots,n_k}
=
\frac{
\partial_1^{n_1-1}\cdots\partial_k^{n_k-1}\Gamma_k(0,\dots,0)}
{(n_1-1)!\cdots(n_k-1)!}
\qquad(n_1,\dots,n_k\ge 1).
\]
The proof of Theorem~\ref{thm:poisson-kl-even-orders} yields
\[
c_{2N}
=
\sum_{k=2}^{N}\frac{(-1)^k}{k(k-1)}
\sum_{\substack{n_1,\dots,n_k\ge 2\\n_1+\cdots+n_k=2N}}
\mu_{n_1}\cdots\mu_{n_k}\,I_{n_1,\dots,n_k}.
\]
Substituting the derivative formula for $I_{n_1,\dots,n_k}$ proves
\eqref{eq:higher-secant-coefficients}.
\end{proof}

\subsection{Centered reconstruction from Poisson data}

\begin{proposition}[Centered Poisson reconstruction by Laplace uniqueness]%
\label{prop:poisson-central-two-channel}
Let $\nu$ be a probability measure on~$\RR$ with finite mean~$\bar\gamma$.
Let $\nu_c$ denote the law of the centred variable
$\Delta:=\gamma-\bar\gamma$.  Define
\[
h_t(x):=\int_{\RR}P_t(x-\gamma)\dd\nu(\gamma),
\]
the central observation functions
\[
A(t):=\pi\,h_t(\bar\gamma),
\qquad
B(t):=\pi\,\partial_x h_t(x)\big|_{x=\bar\gamma}
\qquad(t>0).
\]
\[
H(t):=\int_t^\infty B(s)\dd s
\qquad(t>0),
\]
and let
\[
\phi_{\nu_c}(r):=\int_{\RR}e^{ir\Delta}\dd\nu(\gamma),
\qquad r\in\RR,
\]
be the centred characteristic function.  Then for every $t>0$ one has
\begin{equation}\label{eq:poisson-laplace-reconstruction}
A(t)+\ii H(t)
=\int_0^\infty e^{-tr}\phi_{\nu_c}(r)\dd r.
\end{equation}
Consequently, the pair $(A,B)$ determines the centred law~$\nu_c$
uniquely.  Equivalently, the triple $(\bar\gamma,A,B)$ determines the
original measure~$\nu$ uniquely.
\end{proposition}

\begin{proof}
By the elementary Laplace identities
\[
\int_0^\infty e^{-tr}\cos(r\Delta)\dd r=\frac{t}{t^2+\Delta^2},
\qquad
\int_0^\infty e^{-tr}\sin(r\Delta)\dd r=\frac{\Delta}{t^2+\Delta^2},
\]
one obtains
\[
A(t)
=\int_{\RR}\frac{t}{t^2+\Delta^2}\dd\nu(\gamma)
=\int_0^\infty e^{-tr}\Re\phi_{\nu_c}(r)\dd r.
\]
Differentiation under the integral is justified by dominated convergence,
since for each fixed $t>0$,
\[
\left|\frac{2t(x-\gamma)}{((x-\gamma)^2+t^2)^2}\right|
\le \frac{1}{2t}
\qquad(x,\gamma\in\RR),
\]
and
\[
\partial_x h_t(x)\big|_{x=\bar\gamma}
=\frac{2}{\pi}\int_{\RR}\frac{t\,\Delta}{(\Delta^2+t^2)^2}\dd\nu(\gamma),
\]
in particular
\[
B(t)=2t\int_{\RR}\frac{\Delta}{(\Delta^2+t^2)^2}\dd\nu(\gamma).
\]
Since
\[
\int_t^\infty \frac{2s|\Delta|}{(\Delta^2+s^2)^2}\dd s
=\frac{|\Delta|}{\Delta^2+t^2}
\le \frac{1}{2t},
\]
Fubini's theorem applies and gives
\[
H(t)
=\int_{\RR}\frac{\Delta}{t^2+\Delta^2}\dd\nu(\gamma)
=\int_0^\infty e^{-tr}\Im\phi_{\nu_c}(r)\dd r.
\]
Combining the two formulas yields
\eqref{eq:poisson-laplace-reconstruction}.  By uniqueness of the Laplace
transform on bounded continuous functions \cite{Widder1941}, $(A,H)$, and
therefore $(A,B)$, determine $\Re\phi_{\nu_c}(r)$ and
$\Im\phi_{\nu_c}(r)$ for every $r\ge 0$.  Hence $\phi_{\nu_c}(r)$ is known
on $[0,\infty)$.  For $r>0$ set
\[
\phi_{\nu_c}(-r):=\overline{\phi_{\nu_c}(r)}.
\]
This is the standard characteristic-function symmetry, so the entire
characteristic function is determined.  Levy's inversion theorem for
characteristic functions then recovers the centred law~$\nu_c$
\cite{Lukacs1970,Feller1971}, and adding the mean recovers~$\nu$.
\end{proof}

\begin{corollary}[The symmetric channel]%
\label{cor:poisson-symmetrization}
Under the hypotheses of Proposition~\ref{prop:poisson-central-two-channel}, the
function $A$ determines the symmetrization of the centred law~$\nu_c$.
\end{corollary}

\begin{proof}
The real part of $\phi_{\nu_c}$ is exactly the characteristic function of
the centred symmetrization
$\frac12(\nu_c+\check\nu_c)$, where $\check\nu_c(E):=\nu_c(-E)$.  Since
$A$ is the Laplace transform of $\Re\phi_{\nu_c}$, the claim follows from
Laplace inversion.
\end{proof}

\begin{corollary}[The symmetric case]%
\label{cor:poisson-symmetric-case}
Under the hypotheses of Proposition~\ref{prop:poisson-central-two-channel}, the
centred law~$\nu_c$ is symmetric if and only if $B\equiv 0$.  In that
case $A$ alone determines~$\nu_c$, and the pair $(\bar\gamma,A)$
determines~$\nu$.
\end{corollary}

\begin{proof}
The centred law is symmetric if and only if
$\Im\phi_{\nu_c}\equiv 0$.  By
\eqref{eq:poisson-laplace-reconstruction}, this is equivalent to
$H\equiv 0$, hence to $B\equiv 0$.
\end{proof}

\begin{theorem}[Observation channels as evaluation functionals]%
\label{thm:poisson-channel-rkhs}
Let $\nu$ be a probability measure on~$\RR$ with finite mean~$\bar\gamma$.
For $t>0$ define the centred fluctuation profile
\[
\widetilde\delta_t(y):=
\frac{h_t(\bar\gamma+ty)}{g_t(\bar\gamma+ty)}-1,
\qquad y\in\RR.
\]
Then $\widetilde\delta_t\in L^2_0(\omega)$ and
\begin{equation}\label{eq:delta-t-secant}
\widetilde\delta_t(y)
=\frac{1}{t}\int_{\RR}\Delta\,\Psi_{\Delta/t}(y)\,\dd\nu(\gamma).
\end{equation}
Consequently,
\begin{equation}\label{eq:delta-t-rkhs}
(\mathcal U\widetilde\delta_t)(a)
=\frac{1}{t}\int_{\RR}\Delta\,K(a,\Delta/t)\,\dd\nu(\gamma),
\qquad a\in\RR.
\end{equation}
Moreover,
\begin{equation}\label{eq:delta-t-origin}
\widetilde\delta_t(0)=tA(t)-1,
\end{equation}
and
\begin{equation}\label{eq:delta-t-rkhs-origin}
(\mathcal U\widetilde\delta_t)(0)=2t\,H(2t)
=2t\int_{2t}^{\infty}B(s)\,\dd s.
\end{equation}
\end{theorem}

\begin{proof}
By definition,
\[
\widetilde\delta_t(y)
=\int_{\RR}\bigl(R_{\Delta/t}(y)-1\bigr)\,\dd\nu(\gamma)
=\frac{1}{t}\int_{\RR}\Delta\,\Psi_{\Delta/t}(y)\,\dd\nu(\gamma),
\]
which is \eqref{eq:delta-t-secant}.  By Theorem~\ref{thm:mode-gram-kernel},
\[
\|\Psi_\varepsilon\|_{L^2(\omega)}^2
=\frac{2}{4+(\varepsilon-\varepsilon)^2}
=\frac12.
\]
Hence
\[
\left\|\widetilde\delta_t\right\|_{L^2(\omega)}
\le\frac{1}{t}\int_{\RR}|\Delta|\,\left\|\Psi_{\Delta/t}\right\|_{L^2(\omega)}\,\dd\nu(\gamma)
=\frac{1}{\sqrt2\,t}\int_{\RR}|\Delta|\,\dd\nu(\gamma)
<\infty,
\]
so \eqref{eq:delta-t-secant} is a Bochner integral in $L^2_0(\omega)$.
Therefore,
\[
\int_{\RR}\widetilde\delta_t(y)\,\omega(\dd y)
=\int_{\RR}(h_t(x)-g_t(x))\,\dd x=0,
\]
and thus $\widetilde\delta_t\in L^2_0(\omega)$.  Applying the unitary map
$\mathcal U$ from Theorem~\ref{thm:mode-space-rkhs} to
\eqref{eq:delta-t-secant} yields \eqref{eq:delta-t-rkhs}.

At $y=0$ one has
\[
g_t(\bar\gamma)=\frac{1}{\pi t},
\]
so
\[
\widetilde\delta_t(0)
=\frac{h_t(\bar\gamma)}{g_t(\bar\gamma)}-1
=tA(t)-1,
\]
which proves \eqref{eq:delta-t-origin}.  Finally, evaluating
\eqref{eq:delta-t-rkhs} at $a=0$ and using
$K(0,b)=2/(4+b^2)$ gives
\[
(\mathcal U\widetilde\delta_t)(0)
=\frac{1}{t}\int_{\RR}\Delta\frac{2}{4+(\Delta/t)^2}\,\dd\nu(\gamma)
=2t\int_{\RR}\frac{\Delta}{4t^2+\Delta^2}\,\dd\nu(\gamma).
\]
The definition of $H$ in
Proposition~\ref{prop:poisson-central-two-channel} gives
\[
H(2t)=\int_{\RR}\frac{\Delta}{4t^2+\Delta^2}\,\dd\nu(\gamma),
\]
and \eqref{eq:delta-t-rkhs-origin} follows.
\end{proof}

\begin{remark}
The previous theorem recovers the centred Cauchy transform on the
imaginary axis:
\[
G_{\nu_c}(\ii t)
=\int_{\RR}\frac{1}{\ii t-\Delta}\dd\nu(\gamma)
=-H(t)-\ii A(t)
\qquad(t>0).
\]
Thus the Laplace-transform formulation strengthens the earlier
Cauchy-transform identity while also isolating the loss of the translation
parameter.  In particular, the proof of
Proposition~\ref{prop:poisson-central-two-channel} is better viewed as a
Laplace-uniqueness plus characteristic-function-inversion argument than
as a Stieltjes inversion argument.
\end{remark}
