% sec_entropy_asymptotics.tex -- Poisson smoothing and entropy asymptotics
\section{Poisson Smoothing and Entropy Asymptotics}%
\label{sec:precision-potential}

\paragraph{Assumption and dependency map.}
Throughout this section we write
$\Delta:=\gamma-\bar\gamma$ and
$\mu_n:=\int_{\RR}\Delta^n\,\dd\nu(\gamma)$ whenever the corresponding
moment exists.  For ease of reference, the moment hypotheses and logical
dependencies are as follows.
\begin{itemize}[leftmargin=*]
\item Proposition~\ref{thm:poisson-second-order} uses the mean, variance,
and finite centred fourth moment, through a direct Taylor expansion of
$R_\varepsilon$.
\item Theorem~\ref{thm:poisson-kl-even-orders} uses a finite centred
$(2N+1)$st absolute moment,
Lemma~\ref{lem:tail-uniform-cayley-mode-remainder}, and
Lemma~\ref{lem:mode-parity-vanish}.
\item Proposition~\ref{prop:cdim-poisson-cauchy-moment-inversion-by-phase-modes}
uses the same finite mode calculus at order four and a finite centred
fifth absolute moment to recover $(\sigma^2,\mu_3,\mu_4)$ from phase
Fourier coefficients.
\item Theorem~\ref{thm:poisson-kl-eighth} uses the mean, variance, finite
centred seventh absolute moment,
Lemma~\ref{lem:tail-uniform-cayley-mode-remainder},
Lemma~\ref{lem:mode-parity-vanish}, and
Appendix~\ref{app:entropy-integrals}.
\item Theorems~\ref{thm:poisson-defect-ladder}
and~\ref{thm:poisson-defect-amplification} use the same finite centred
seventh absolute moment hypothesis and the explicit coefficients from
Theorem~\ref{thm:poisson-kl-eighth}.
\item The fourth-moment identity \eqref{eq:defect-sixth-fourth-moment} and
the coupling estimate \eqref{eq:fourth-moment-algebraic-coupling} use only
the mean, variance, and finite centred fourth moment, and define the
standalone algebraic defect $\Delta_6^{(4)}(\nu)$.
\item Theorem~\ref{thm:poisson-defect-stability} is a seventh-moment
entropy theorem.  It again uses the mean, variance, and finite centred
seventh absolute moment; only under this hypothesis does the bridge lemma in
Subsection~\ref{subsec:fourth-moment-stability} identify
$\Delta_6^{(4)}(\nu)$ with the genuine $t^{-6}$ entropy defect.
\end{itemize}
In particular, the appendix enters only in the explicit evaluation of the
$t^{-6}$ and $t^{-8}$ coefficients in
Theorem~\ref{thm:poisson-kl-eighth}.  The defect ladder and amplification
results are algebraic consequences of those coefficients, while the
fourth-moment identity below isolates the polynomial which later becomes
the sixth-order entropy defect.  The stability theorem is stated only after
imposing the stronger seventh-moment hypothesis, so its defect variable is
the genuine $t^{-6}$ entropy-asymptotic defect.  Without that stronger
hypothesis this section uses only the superscripted algebraic notation
$C_6^{(4)}$ and $\Delta_6^{(4)}$.

\subsection{Entropy in Cayley coordinates}

The logarithmic Jacobian of the Cayley coordinate is
\begin{equation}\label{eq:precision-potential}
\mathcal P(x):=-\log J(x),\qquad
J(x):=\upsilon'(x)=\frac{1}{\pi(1+x^{2})},
\end{equation}
equivalently,
\begin{equation}\label{eq:precision-potential-expanded}
\mathcal P(x)=\log\pi+\log(1+x^{2}).
\end{equation}

\begin{theorem}[Entropy identity in Cayley coordinates]%
\label{thm:precision-ledger}
Let $X$ be a real-valued random variable with density $f_X$ such that
$h_{\mathrm{diff}}(X)>-\infty$ and $\EE[\mathcal P(X)]<\infty$.  Set
$U:=\upsilon(X)\in\UU$.  Then
\begin{equation}\label{eq:entropy-ledger}
h_{\mathrm{diff}}(X)
=h_{\mathrm{diff}}(U)+\EE[\mathcal P(X)].
\end{equation}
\end{theorem}

\begin{proof}
The density of $U$ is
\[
f_U(u)=f_X(\tan(\pi u))\,\pi(1+\tan^2(\pi u)).
\]
The standard change-of-variables formula for differential entropy gives
\[
h_{\mathrm{diff}}(X)
=h_{\mathrm{diff}}(U)+\EE\!\left[\log\frac{1}{\abs{\upsilon'(X)}}\right]
=h_{\mathrm{diff}}(U)+\EE[\mathcal P(X)].
\]
\end{proof}
\leanverified{paper_circle_dimension_precision_ledger}

\subsection{Poisson smoothing}

For $t>0$ and $x\in\RR$ define
\begin{equation}\label{eq:poisson-kernel}
P_t(x):=\frac{1}{\pi}\frac{t}{x^{2}+t^{2}}.
\end{equation}

\begin{proposition}[Basic identities for the Poisson kernels]%
\label{prop:poisson-identities}
For $s,t>0$ one has
\begin{equation}\label{eq:poisson-basic}
P_t(x)=\frac{1}{t}P_1\!\left(\frac{x}{t}\right),
\qquad
\widehat{P_t}(\xi)=e^{-t\abs{\xi}},
\qquad
P_t*P_s=P_{t+s}.
\end{equation}
In particular,
\[
\mu=P_1(x)\dd x,
\qquad
P_t*\mu=P_{t+1}.
\]
\end{proposition}

\begin{proof}
The scaling identity is immediate from the definition.  The Fourier
transform of~$P_1$ is $e^{-\abs{\xi}}$
\cite[Chapter~III]{Stein1971}; scaling yields
$\widehat{P_t}(\xi)=e^{-t\abs{\xi}}$.  The convolution law follows by
multiplying Fourier transforms.
\end{proof}

\begin{proposition}[Standard facts on the Poisson semigroup]%
\label{thm:poisson-semigroup-law}
For $1\le p\le\infty$ and $t>0$ define
$T_t f:=P_t*f$.  Then:
\begin{enumerate}[label=(\roman*)]
\item $\|T_t f\|_{L^p(\RR)}\le \|f\|_{L^p(\RR)}$ for every $f\in L^p$;
\item $T_{t+s}=T_tT_s$ for all $s,t>0$;
\item for $1\le p<\infty$, $T_t f\to f$ in $L^p(\RR)$ as $t\downarrow 0$;
\item on $L^2(\RR)$, $\widehat{T_t f}(\xi)=e^{-t\abs{\xi}}\widehat f(\xi)$
      and the generator is $-|D|$.
\end{enumerate}
\end{proposition}

\begin{proof}
Since $\int_{\RR}P_t(x)\dd x=1$ and $P_t\ge 0$, Young's inequality gives
(i).  The semigroup law is Proposition~\ref{prop:poisson-identities}.
Strong continuity for $1\le p<\infty$ follows because $\{P_t\}_{t>0}$ is
an approximate identity.  The $L^2$-multiplier statement is exactly
\eqref{eq:poisson-basic}; the generator follows from
\[
\frac{e^{-t\abs{\xi}}-1}{t}\longrightarrow -\abs{\xi}
\qquad(t\downarrow 0)
\]
in the Fourier representation.
\end{proof}

\begin{remark}
The preceding proposition is standard and is included only to fix the
normalization of the Poisson kernel and of the generator $-|D|$ used in
the sequel.
\end{remark}

\subsection{Large-time comparison with a translated Poisson profile}

\begin{proposition}[Second-order approximation]%
\label{thm:poisson-second-order}
\leanverified{paper_circle_dimension_poisson_second_order}
Let $\nu$ be a probability measure on~$\RR$ with mean
\[
\bar\gamma:=\int_{\RR}\gamma\dd\nu(\gamma),
\]
variance
\[
\sigma^{2}:=\int_{\RR}(\gamma-\bar\gamma)^{2}\dd\nu(\gamma),
\]
and finite centred fourth moment.  Define
\[
h_t:=P_t*\nu,
\qquad
g_t(x):=P_t(x-\bar\gamma).
\]
Let
\[
U(y):=\frac{3y^{2}-1}{(1+y^{2})^{2}}.
\]
Then there exists a constant $C_\nu>0$ such that for every $t\ge 1$,
\begin{equation}\label{eq:poisson-profile}
\sup_{x\in\RR}\left|
\frac{h_t(x)}{g_t(x)}
-1
-\frac{\sigma^{2}}{t^{2}}U\!\left(\frac{x-\bar\gamma}{t}\right)
\right|
\le \frac{C_\nu}{t^{3}}.
\end{equation}
Consequently,
\begin{equation}\label{eq:poisson-l1-rate}
\|h_t-g_t\|_{L^{1}(\RR)}=O(t^{-2}),
\qquad
\|h_t-g_t\|_{L^\infty(\RR)}=O(t^{-3}).
\end{equation}
Moreover,
\begin{equation}\label{eq:poisson-kl-rate}
D_{\mathrm{KL}}(h_t\|g_t)=O(t^{-4}).
\end{equation}
\end{proposition}

\begin{proof}
Write $\Delta:=\gamma-\bar\gamma$ and $x=\bar\gamma+ty$.  Then
\[
\frac{P_t(x-\gamma)}{g_t(x)}
=\frac{1+y^{2}}{1+(y-\Delta/t)^{2}}.
\]
Set $\varepsilon:=\Delta/t$.  A direct algebraic computation gives the
exact identity
\begin{align*}
\frac{1+y^{2}}{1+(y-\varepsilon)^{2}}
&=1+\frac{2y}{1+y^{2}}\varepsilon
+\frac{3y^{2}-1}{(1+y^{2})^{2}}\varepsilon^{2} \\
&\quad
+\frac{4y(y^{2}-1)\varepsilon^{3}-(3y^{2}-1)\varepsilon^{4}}
{(1+y^{2})^{2}(1+(y-\varepsilon)^{2})}.
\end{align*}
Since
\[
\frac{4\abs{y}\abs{y^{2}-1}}{(1+y^{2})^{2}}\le 2,
\qquad
\frac{\abs{3y^{2}-1}}{(1+y^{2})^{2}}\le 1,
\]
the remainder is bounded by $2\abs{\varepsilon}^{3}+\abs{\varepsilon}^{4}$
uniformly in~$y$.  Taking expectation and using $\EE[\Delta]=0$ yields
\[
\frac{h_t(\bar\gamma+ty)}{g_t(\bar\gamma+ty)}
=1+\frac{\sigma^{2}}{t^{2}}U(y)+R_t(y),
\]
with
\[
\sup_{y\in\RR}\abs{R_t(y)}
\le \frac{2\EE[\abs{\Delta}^{3}]+\EE[\abs{\Delta}^{4}]}{t^{3}}
\]
for $t\ge 1$.  This proves \eqref{eq:poisson-profile}.

The function~$U$ is bounded, so
$\sup_x\abs{h_t(x)/g_t(x)-1}=O(t^{-2})$.  Since $\int_{\RR}g_t=1$,
\[
\|h_t-g_t\|_{L^1}
\le \sup_{x}\abs{\frac{h_t(x)}{g_t(x)}-1}\int_{\RR}g_t(x)\dd x
=O(t^{-2}).
\]
Also $\|g_t\|_{L^\infty}=1/(\pi t)$, so
\[
\|h_t-g_t\|_{L^\infty}
\le \|g_t\|_{L^\infty}\sup_{x}\abs{\frac{h_t(x)}{g_t(x)}-1}
=O(t^{-3}).
\]
Let
\[
\delta_t(x):=\frac{h_t(x)}{g_t(x)}-1.
\]
The bound \eqref{eq:poisson-profile} implies that
$\|\delta_t\|_{L^\infty}=O(t^{-2})$, so for large $t$ one has
$\|\delta_t\|_{L^\infty}\le 1/2$.  Since
$\int g_t\delta_t\dd x=\int(h_t-g_t)\dd x=0$ and
\[
(1+s)\log(1+s)-s\le 2s^{2}
\qquad(\abs{s}\le 1/2),
\]
it follows that
\[
D_{\mathrm{KL}}(h_t\|g_t)
=\int_{\RR}g_t(x)\bigl((1+\delta_t(x))\log(1+\delta_t(x))-\delta_t(x)\bigr)\dd x
\le 2\|\delta_t\|_{L^\infty}^{2}.
\]
Hence \eqref{eq:poisson-kl-rate} holds.
\end{proof}

\subsection{Cayley--Chebyshev mode calculus}

For $\varepsilon\in\RR$ and $y\in\RR$ set
\begin{equation}\label{eq:cayley-comparison-kernel}
R_\varepsilon(y):=\frac{1+y^2}{1+(y-\varepsilon)^2}.
\end{equation}
For fixed $y$, the Taylor expansion of $R_\varepsilon(y)$ at
$\varepsilon=0$ defines coefficients $u_n(y)$ through
\begin{equation}\label{eq:cayley-mode-series}
R_\varepsilon(y)=\sum_{n=0}^{\infty}u_n(y)\varepsilon^n,
\qquad
|\varepsilon|<\sqrt{1+y^2}.
\end{equation}

\begin{theorem}[Cayley--Chebyshev mode formula]%
\label{thm:cayley-chebyshev-mode}
Let $U_n$ denote the Chebyshev polynomials of the second kind.  If
$y=\tan\theta$ with $\theta\in(-\pi/2,\pi/2)$, then for every $n\ge 0$,
\begin{equation}\label{eq:cayley-chebyshev-mode}
u_n(\tan\theta)=\cos^n\theta\,U_n(\sin\theta).
\end{equation}
For $n\ge 1$ this is equivalent to
\begin{equation}\label{eq:cayley-chebyshev-trig}
u_n(\tan\theta)
=\cos^{n-1}\theta\,
\sin\!\Bigl((n+1)\Bigl(\frac{\pi}{2}-\theta\Bigr)\Bigr).
\end{equation}
Consequently
\begin{equation}\label{eq:cayley-mode-parity}
u_n(-y)=(-1)^n u_n(y)
\qquad(n\ge 0).
\end{equation}
Moreover, for every integer $m\ge 1$,
\begin{equation}\label{eq:cayley-mode-even-fourier}
u_{2m}(\tan\theta)
=(-1)^m 2^{-(2m-1)}
\sum_{j=0}^{2m-1}\binom{2m-1}{j}\cos(2(j+1)\theta),
\end{equation}
and for every integer $m\ge 0$,
\begin{equation}\label{eq:cayley-mode-odd-fourier}
u_{2m+1}(\tan\theta)
=(-1)^m 2^{-2m}
\sum_{j=0}^{2m}\binom{2m}{j}\sin(2(j+1)\theta).
\end{equation}
\leanverified{paper_circle_dimension_cayley_chebyshev_mode}
\end{theorem}

\begin{proof}
With $y=\tan\theta$ one has
\[
R_\varepsilon(\tan\theta)
=\frac{1}{1-2(\varepsilon\cos\theta)\sin\theta+(\varepsilon\cos\theta)^2}.
\]
The generating function
\[
\sum_{n=0}^{\infty}U_n(x)r^n=\frac{1}{1-2xr+r^2}
\]
therefore yields \eqref{eq:cayley-chebyshev-mode} with
$x=\sin\theta$ and $r=\varepsilon\cos\theta$.  Since
\[
U_n(\sin\theta)
=\frac{\sin((n+1)(\pi/2-\theta))}{\cos\theta},
\]
formula \eqref{eq:cayley-chebyshev-trig} follows for $n\ge 1$.  The
parity relation \eqref{eq:cayley-mode-parity} is immediate from
\eqref{eq:cayley-chebyshev-trig}.

For the finite Fourier expansions, write
\[
u_{2m}(\tan\theta)=(-1)^m\cos^{2m-1}\theta\,\cos((2m+1)\theta),
\]
\[
u_{2m+1}(\tan\theta)=(-1)^m\cos^{2m}\theta\,\sin((2m+2)\theta),
\]
and use
\[
1+e^{2\ii\theta}=2e^{\ii\theta}\cos\theta.
\]
Taking real and imaginary parts after the binomial expansion gives
\eqref{eq:cayley-mode-even-fourier} and
\eqref{eq:cayley-mode-odd-fourier}.
\end{proof}

\begin{proposition}[Finite mode truncations]%
\label{prop:cayley-mode-remainder}
For every integer $N\ge 1$ there exists a rational function
$v_N:\RR^2\to\RR$, bounded on~$\RR^2$, such that
\begin{equation}\label{eq:cayley-mode-remainder}
R_\varepsilon(y)
=\sum_{n=0}^{N}u_n(y)\varepsilon^n+\varepsilon^{N+1}v_N(y,\varepsilon)
\qquad(y,\varepsilon\in\RR).
\end{equation}
\end{proposition}

\begin{proof}
For every $x$ and $r$ one has the finite-sum identity
\[
\sum_{n=0}^{N}U_n(x)r^n
=\frac{1-r^{N+1}U_{N+1}(x)+r^{N+2}U_N(x)}{1-2xr+r^2}.
\]
Substituting $x=\sin\theta$ and $r=\varepsilon\cos\theta$ gives
\eqref{eq:cayley-mode-remainder} with
\[
v_N(\tan\theta,\varepsilon)
=\frac{\cos^{N+1}\theta\,U_{N+1}(\sin\theta)
-\varepsilon\cos^{N+2}\theta\,U_N(\sin\theta)}
{1-2(\varepsilon\cos\theta)\sin\theta+(\varepsilon\cos\theta)^2}.
\]
Equivalently, since
\[
1-2(\varepsilon\cos\theta)\sin\theta
   +(\varepsilon\cos\theta)^2
=\frac{1+(\tan\theta-\varepsilon)^2}{1+\tan^2\theta},
\]
we have
\[
v_N(y,\varepsilon)
=\frac{(1+y^2)u_{N+1}(y)-\varepsilon u_N(y)}
       {1+(y-\varepsilon)^2},
\]
so $v_N$ is rational in $(y,\varepsilon)$.

It remains to prove boundedness.  Put
$c=\cos\theta$, $s=\sin\theta$, and $d=\tan\theta-\varepsilon$.  Then
$\varepsilon c=s-dc$, and the preceding trigonometric formula becomes
\[
v_N(\tan\theta,\varepsilon)
=\frac{c^{N-1}\bigl(U_{N+1}(s)-sU_N(s)+dc\,U_N(s)\bigr)}
       {1+d^2}.
\]
For $|s|\le1$, the polynomials $U_N$ and $U_{N+1}$ are bounded by
constants depending only on~$N$.  Since $N\ge1$, $|c|^{N-1}\le1$, and
\[
\frac{|d|\,|c|^N}{1+d^2}\le \frac{|d|}{1+d^2}\le \frac12,
\]
the right-hand side is bounded uniformly in $\theta$ and~$\varepsilon$.
This proves that $v_N$ is bounded on~$\RR^2$.
\end{proof}

\begin{lemma}[Tail-uniform Cayley mode remainder]%
\label{lem:tail-uniform-cayley-mode-remainder}
Let $N\ge1$, let $X$ be a real random variable with
$\EE|X|^{N+1}<\infty$, and let
\[
\mathcal R_{N,t}(y):=
\EE\!\left[
R_{X/t}(y)-\sum_{n=0}^{N}u_n(y)\frac{X^n}{t^n}
\right]
\qquad(t>0).
\]
Then there is a constant $M_N<\infty$, depending only on~$N$, such that
\begin{equation}\label{eq:tail-uniform-cayley-mode-remainder}
\sup_{y\in\RR}|\mathcal R_{N,t}(y)|
\;+\;\|\mathcal R_{N,t}\|_{L^1(\omega)}
\le
M_N\,\EE|X|^{N+1}\,t^{-(N+1)},
\end{equation}
where $\omega(\dd y)=\pi^{-1}(1+y^2)^{-1}\dd y$.  Thus the remainder is
uniform in the phase variable after integration over the full, unbounded
tail of~$X$.
\end{lemma}

\begin{proof}
By Proposition~\ref{prop:cayley-mode-remainder},
\[
R_{X/t}(y)-\sum_{n=0}^{N}u_n(y)\frac{X^n}{t^n}
=\frac{X^{N+1}}{t^{N+1}}v_N(y,X/t)
\]
for every $y$ and every value of~$X$.  Let
$M_N:=\|v_N\|_{L^\infty(\RR^2)}$.  Taking expectations gives
\[
|\mathcal R_{N,t}(y)|
\le M_N\,\EE|X|^{N+1}\,t^{-(N+1)}
\qquad(y\in\RR).
\]
This is the asserted uniform estimate.  Since $\omega$ is a probability
measure, the same bound also gives the displayed $L^1(\omega)$ estimate.
\end{proof}

For later use we record the first modes explicitly:
\begin{equation}\label{eq:cayley-first-modes}
u_2(y)=\frac{3y^2-1}{(1+y^2)^2},
\qquad
u_3(y)=\frac{4y(y^2-1)}{(1+y^2)^3},
\qquad
u_4(y)=\frac{5y^4-10y^2+1}{(1+y^2)^4},
\end{equation}
\begin{equation}\label{eq:cayley-next-modes}
u_5(y)=\frac{6y^5-20y^3+6y}{(1+y^2)^5},
\qquad
u_6(y)=\frac{7y^6-35y^4+21y^2-1}{(1+y^2)^6}.
\end{equation}

Proposition~\ref{prop:phase-coordinate-interface} is the formal link
with the finite phase-sampling block: the trigonometric coefficients
below are Haar-Fourier coefficients of the same circle characters whose
finite grids are counted by~$\Sigma_G(N)$.

\begin{proposition}[Poisson--Cauchy moment inversion by phase modes]%
\label{prop:cdim-poisson-cauchy-moment-inversion-by-phase-modes}
Let $\nu$ be a probability measure on~$\RR$ with mean~$\bar\gamma$ and
finite centred fifth absolute moment.  Put
\[
\Delta:=\gamma-\bar\gamma,\qquad
\mu_n:=\int_{\RR}\Delta^n\,\dd\nu(\gamma),\qquad
\sigma^2:=\mu_2,
\]
and let $h_t:=P_t*\nu$ and $g_t(x):=P_t(x-\bar\gamma)$.  For
$\theta\in(-\pi/2,\pi/2)$ define the phase profile
\[
F_t(\theta):=
\frac{h_t(\bar\gamma+t\tan\theta)}
     {g_t(\bar\gamma+t\tan\theta)}-1.
\]
For $k\ge1$ set
\[
\mathcal C_k(t):=
2\int_{-\pi/2}^{\pi/2}F_t(\theta)\cos(2k\theta)\,\frac{\dd\theta}{\pi},
\qquad
\mathcal S_k(t):=
2\int_{-\pi/2}^{\pi/2}F_t(\theta)\sin(2k\theta)\,\frac{\dd\theta}{\pi}.
\]
Then
\[
\sigma^2=\lim_{t\to\infty}-2t^2\bigl(\mathcal C_1(t)-\mathcal C_4(t)\bigr),
\qquad
\mu_3=\lim_{t\to\infty}-4t^3\mathcal S_3(t),
\qquad
\mu_4=\lim_{t\to\infty}8t^4\mathcal C_4(t).
\]
More precisely,
\[
\mathcal C_1(t)-\mathcal C_4(t)
=-\frac{\sigma^2}{2t^2}+O(t^{-5}),\qquad
\mathcal S_3(t)=-\frac{\mu_3}{4t^3}+O(t^{-5}),
\qquad
\mathcal C_4(t)=\frac{\mu_4}{8t^4}+O(t^{-5}).
\]
\end{proposition}

\begin{proof}
Write $y=\tan\theta$.  For $x=\bar\gamma+ty$ one has
\[
\frac{P_t(x-\gamma)}{P_t(x-\bar\gamma)}
=\frac{1+y^2}{1+(y-\Delta/t)^2}
=R_{\Delta/t}(y).
\]
Therefore
\[
F_t(\theta)=\int_{\RR}\bigl(R_{\Delta/t}(\tan\theta)-1\bigr)
\,\dd\nu(\gamma).
\]
Proposition~\ref{prop:cayley-mode-remainder} gives the pointwise
order-four identity
\[
R_{\Delta/t}(y)-1
=\sum_{n=1}^{4}u_n(y)\frac{\Delta^n}{t^n}
+\frac{\Delta^5}{t^5}v_4(y,\Delta/t).
\]
Taking $\nu$-expectations and using $\mu_1=0$ gives, uniformly in
$\theta$,
\[
F_t(\theta)
=\frac{\sigma^2}{t^2}u_2(\tan\theta)
+\frac{\mu_3}{t^3}u_3(\tan\theta)
+\frac{\mu_4}{t^4}u_4(\tan\theta)
+O(t^{-5}).
\]
The finite fifth absolute moment is used only through
Lemma~\ref{lem:tail-uniform-cayley-mode-remainder}, with $X=\Delta$ and
truncation order~$4$, to bound the last expectation uniformly.

The first three nonconstant modes have the finite phase-Fourier
expansions
\[
u_2(\tan\theta)=-\frac12\bigl(\cos2\theta+\cos4\theta\bigr),
\]
\[
u_3(\tan\theta)
=-\frac14\bigl(\sin2\theta+2\sin4\theta+\sin6\theta\bigr),
\]
\[
u_4(\tan\theta)
=\frac18\bigl(\cos2\theta+3\cos4\theta+3\cos6\theta+\cos8\theta\bigr),
\]
which are the cases $n=2,3,4$ of
Theorem~\ref{thm:cayley-chebyshev-mode}.  The functions
$\cos(2k\theta)$ and $\sin(2k\theta)$ are orthogonal on
$(-\pi/2,\pi/2)$, and the prefactor $2/\pi$ in the definitions of
$\mathcal C_k$ and $\mathcal S_k$ extracts the corresponding coefficient.
Thus the preceding uniform expansion gives
\[
\mathcal C_1(t)
=-\frac{\sigma^2}{2t^2}+\frac{\mu_4}{8t^4}+O(t^{-5}),
\qquad
\mathcal C_4(t)=\frac{\mu_4}{8t^4}+O(t^{-5}),
\]
and
\[
\mathcal S_3(t)=-\frac{\mu_3}{4t^3}+O(t^{-5}).
\]
Subtracting the two cosine coefficients and multiplying by the displayed
powers of~$t$ proves the three inversion formulas.
\end{proof}

\begin{lemma}[Mode parity cancellation]\label{lem:mode-parity-vanish}
For integers $n_1,\dots,n_k\ge 1$ let
\[
I_{n_1,\dots,n_k}:=
\int_{\RR}
\frac{u_{n_1}(y)\cdots u_{n_k}(y)}{\pi(1+y^2)}\,\dd y.
\]
Then $I_{n_1,\dots,n_k}=0$ whenever $n_1+\cdots+n_k$ is odd.
\end{lemma}

\begin{proof}
From \eqref{eq:cayley-mode-parity}, $u_n(-y)=(-1)^n u_n(y)$ for every
$n\ge1$, while $(1+y^2)^{-1}$ is even.  Hence the integrand is odd if
$n_1+\cdots+n_k$ is odd, so the integral vanishes.
\end{proof}
\leanverified{paper_circle_dimension_mode_parity_vanish}

\subsection{Entropy asymptotics and rigidity}

The next lemma is the bookkeeping device used by the coefficient
computations.  It fixes the order of operations for powers of the density
ratio.  First one truncates the one-body mode expansion additively, then
one expands the resulting finite polynomial exactly, and only after that
one discards monomials whose total $t$-degree is larger than the target
degree.  The retained polynomial is denoted $P_{k,N,t}$, and the power
remainder is defined by
\[
\varepsilon_{k,N,t}
:=\bigl(\delta_t(\bar\gamma+t\,\cdot)\bigr)^k-P_{k,N,t}.
\]
Thus the power formula below is a pointwise identity with an additive
remainder in $L^\infty(\RR)\cap L^1(\omega)$.

\begin{lemma}[Uniform additive entropy-expansion remainders]%
\label{lem:entropy-uniform-remainder}
Let $N\ge2$ and let $\nu$ have mean~$\bar\gamma$ and finite centred
$(2N+1)$st absolute moment.  Put
\[
\delta_t(\bar\gamma+ty):=
\frac{h_t(\bar\gamma+ty)}{g_t(\bar\gamma+ty)}-1.
\]
Let $\omega(\dd y):=\pi^{-1}(1+y^2)^{-1}\dd y$.
Then there are constants $C_{\nu,N},T_{\nu,N}>0$ and, for each
$2\le k\le N$, a constant $C_{\nu,N,k}>0$ with the following properties.
\begin{enumerate}[label=\textup{(\alph*)},leftmargin=*]
\item For every $t\ge1$,
\begin{equation}\label{eq:uniform-delta-expansion}
\delta_t(\bar\gamma+ty)
=\sum_{n=2}^{2N}\frac{\mu_n}{t^n}u_n(y)+\rho_{N,t}(y),
\qquad
\sup_{y\in\RR}|\rho_{N,t}(y)|
\;+\;\|\rho_{N,t}\|_{L^1(\omega)}
\le C_{\nu,N} t^{-(2N+1)},
\end{equation}
and
\begin{equation}\label{eq:uniform-delta-bound}
\sup_{y\in\RR}|\delta_t(\bar\gamma+ty)|\le C_{\nu,N} t^{-2}.
\end{equation}

\item For $t\ge T_{\nu,N}$ the function
\[
y\longmapsto
\Phi(\delta_t(\bar\gamma+ty))
:=(1+\delta_t(\bar\gamma+ty))
\log(1+\delta_t(\bar\gamma+ty))
-\delta_t(\bar\gamma+ty)
\]
is integrable against $\omega$, and
\begin{equation}\label{eq:uniform-Phi-expansion}
\Phi(\delta_t(\bar\gamma+ty))
=\sum_{j=2}^{N}\frac{(-1)^j}{j(j-1)}
\bigl(\delta_t(\bar\gamma+ty)\bigr)^j+\eta_{N,t}(y),
\end{equation}
where
\[
\sup_{y\in\RR}|\eta_{N,t}(y)|
\;+\;\|\eta_{N,t}\|_{L^1(\omega)}
\le C_{\nu,N} t^{-(2N+2)}.
\]

\item Put
\[
A_{N,t}(y):=\sum_{n=2}^{2N}\frac{\mu_n}{t^n}u_n(y),
\qquad
\delta_t(\bar\gamma+ty)=A_{N,t}(y)+\rho_{N,t}(y).
\]
For every $2\le k\le N$, define the retained polynomial
\[
P_{k,N,t}(y):=
\sum_{m=2k}^{2N}\frac{1}{t^m}
\sum_{\substack{n_1,\dots,n_k\ge2\\ n_1+\cdots+n_k=m}}
\mu_{n_1}\cdots\mu_{n_k}
u_{n_1}(y)\cdots u_{n_k}(y).
\]
The power remainder is the exact additive difference
\[
\varepsilon_{k,N,t}(y):=
\bigl(\delta_t(\bar\gamma+ty)\bigr)^k-P_{k,N,t}(y).
\]
Thus the power expansion is the pointwise additive identity
\begin{equation}\label{eq:uniform-power-expansion}
\begin{split}
\bigl(\delta_t(\bar\gamma+ty)\bigr)^k
\;=&
\sum_{m=2k}^{2N}\frac{1}{t^m}
\sum_{\substack{n_1,\dots,n_k\ge2\\ n_1+\cdots+n_k=m}}
\mu_{n_1}\cdots\mu_{n_k}
u_{n_1}(y)\cdots u_{n_k}(y)
\;+\;\varepsilon_{k,N,t}(y).
\end{split}
\end{equation}
Equivalently, in the retained-polynomial notation this is
\[
\bigl(\delta_t(\bar\gamma+ty)\bigr)^k
=P_{k,N,t}(y)+\varepsilon_{k,N,t}(y).
\]
The last term is the whole difference between the exact power and the
retained polynomial $P_{k,N,t}$.  In terms of the one-body truncation,
\[
\varepsilon_{k,N,t}
=\bigl((A_{N,t}+\rho_{N,t})^k-A_{N,t}^k\bigr)
+\bigl(A_{N,t}^k-P_{k,N,t}\bigr).
\]
For every $t\ge1$ the same additive remainder satisfies the uniform and
integrated bounds
\begin{equation}\label{eq:uniform-power-error-bound}
\sup_y |\varepsilon_{k,N,t}(y)|
+\|\varepsilon_{k,N,t}\|_{L^1(\omega)}
\le C_{\nu,N,k}t^{-(2N+1)}.
\end{equation}
In particular,
\[
\left|\int_{\RR}\varepsilon_{k,N,t}\,\dd\omega\right|
\le C_{\nu,N,k}t^{-(2N+1)}.
\]
\end{enumerate}
Consequently all remainders in \eqref{eq:uniform-Phi-expansion} and
\eqref{eq:uniform-power-expansion} are $o(t^{-2N})$ in $L^1(\omega)$.
\end{lemma}

\begin{proof}
\emph{Mode remainder.}
All constants below may depend on $\nu$ and $N$, and in the power estimate
also on $k$, but not on $y$ or $t$.
Apply Lemma~\ref{lem:tail-uniform-cayley-mode-remainder} with
$X=\Delta$ and truncation order~$2N$.  Taking $\nu$-expectations in the
mode expansion, using $\mu_0=1$ and $\mu_1=0$, and subtracting the
constant term gives \eqref{eq:uniform-delta-expansion}.  The lemma gives
the displayed uniform and $L^1(\omega)$ bound for $\rho_{N,t}$ directly
from the finite centred $(2N+1)$st absolute moment.
Each fixed mode $u_n$ is bounded by
Theorem~\ref{thm:cayley-chebyshev-mode}, so
\eqref{eq:uniform-delta-expansion} implies \eqref{eq:uniform-delta-bound}.

\emph{Entropy Taylor remainder.}
Choose $T_{\nu,N}$ so large that the right-hand side of
\eqref{eq:uniform-delta-bound} is at most $1/2$ for $t\ge T_{\nu,N}$.  On
$|z|\le1/2$ the Taylor series
\[
\Phi(z)=\sum_{k=2}^{\infty}\frac{(-1)^k}{k(k-1)}z^k
\]
has a remainder after degree~$N$ bounded by a constant times
$|z|^{N+1}$.  Since $\delta_t=O(t^{-2})$ uniformly, this gives
\eqref{eq:uniform-Phi-expansion}.  The measure~$\omega$ has total mass
one, so the same supremum estimate gives the displayed $L^1(\omega)$
bound and the asserted integrability.

\emph{Power remainders.}
It remains to justify \eqref{eq:uniform-power-expansion} and
\eqref{eq:uniform-power-error-bound}.  From
\eqref{eq:uniform-delta-expansion},
\[
\delta_t(\bar\gamma+ty)=A_{N,t}(y)+\rho_{N,t}(y),
\qquad
\sup_y|\rho_{N,t}(y)|
\;+\;\|\rho_{N,t}\|_{L^1(\omega)}
=O(t^{-(2N+1)}).
\]
The modes are bounded, so $A_{N,t}=O(t^{-2})$ uniformly.  First expand the
finite product $A_{N,t}(y)^k$ exactly:
\[
A_{N,t}(y)^k
=
\sum_{m=2k}^{2Nk}\frac{1}{t^m}
\sum_{\substack{2\le n_1,\dots,n_k\le 2N\\
n_1+\cdots+n_k=m}}
\mu_{n_1}\cdots\mu_{n_k}
u_{n_1}(y)\cdots u_{n_k}(y).
\]
No estimate has entered this step: it is a finite algebraic identity.
Splitting this exact finite sum at total degree $2N$ gives
\[
A_{N,t}(y)^k=P_{k,N,t}(y)+r_{k,N,t}(y),
\]
where
\[
r_{k,N,t}(y)=
\sum_{m=2N+1}^{2Nk}\frac{1}{t^m}
\sum_{\substack{2\le n_1,\dots,n_k\le 2N\\
n_1+\cdots+n_k=m}}
\mu_{n_1}\cdots\mu_{n_k}
u_{n_1}(y)\cdots u_{n_k}(y).
\]
The sum is finite.  Since every mode in it is bounded and $t\ge1$,
\[
\sup_y|r_{k,N,t}(y)|=O(t^{-(2N+1)}).
\]
Because $\omega(\RR)=1$, this also gives
\[
\|r_{k,N,t}\|_{L^1(\omega)}=O(t^{-(2N+1)}).
\]
Second, compare the exact power of $\delta_t=A_{N,t}+\rho_{N,t}$ with the
finite product just expanded.  The binomial formula gives
\[
\bigl(A_{N,t}(y)+\rho_{N,t}(y)\bigr)^k-A_{N,t}(y)^k
=
\sum_{\ell=1}^{k}\binom{k}{\ell}
A_{N,t}(y)^{k-\ell}\rho_{N,t}(y)^\ell.
\]
For the term with index~$\ell$ the bound $A_{N,t}=O(t^{-2})$ and
$\rho_{N,t}=O(t^{-(2N+1)})$ gives
$O(t^{-2(k-\ell)})O(t^{-\ell(2N+1)})$, which is
$O(t^{-(2N+1)})$ for $t\ge1$.  Hence
\[
\bigl(A_{N,t}(y)+\rho_{N,t}(y)\bigr)^k-A_{N,t}(y)^k
=O(t^{-(2N+1)}).
\]
For the corresponding $L^1(\omega)$ estimate, use
\[
\|A_{N,t}^{\,k-\ell}\rho_{N,t}^{\,\ell}\|_{L^1(\omega)}
\le
\|A_{N,t}\|_\infty^{k-\ell}
\|\rho_{N,t}\|_\infty^{\ell-1}
\|\rho_{N,t}\|_{L^1(\omega)}
=O(t^{-(2N+1)})
\]
for every $\ell\ge1$.
By the definition of $\varepsilon_{k,N,t}$, this proves the explicit
additive decomposition
\[
\varepsilon_{k,N,t}(y)
=r_{k,N,t}(y)
+\bigl(A_{N,t}(y)+\rho_{N,t}(y)\bigr)^k-A_{N,t}(y)^k
\]
for the additive remainder in \eqref{eq:uniform-power-expansion}.  The
displayed uniform and $L^1(\omega)$ estimates for $r_{k,N,t}$ and the
binomial difference give \eqref{eq:uniform-power-error-bound}.  The bounds
$t^{-(2N+2)}=o(t^{-2N})$ and
$t^{-(2N+1)}=o(t^{-2N})$ give the final claim.
\end{proof}

\begin{theorem}[Odd orders vanish in the entropy expansion]%
\label{thm:poisson-kl-even-orders}
Let $N\ge 2$ and let $\nu$ be a probability measure on~$\RR$ with mean
$\bar\gamma$ and finite centred $(2N+1)$st absolute moment.  Set
\[
\Delta:=\gamma-\bar\gamma,
\qquad
\mu_n:=\int_{\RR}\Delta^n\dd\nu(\gamma),
\]
and keep the notation
\[
h_t:=P_t*\nu,
\qquad
g_t(x):=P_t(x-\bar\gamma).
\]
Then there exist real coefficients $c_4,c_6,\dots,c_{2N}$ such that
\begin{equation}\label{eq:poisson-kl-even-orders}
D_{\mathrm{KL}}(h_t\|g_t)
=\sum_{m=2}^{N}c_{2m}\,t^{-2m}+o(t^{-2N})
\qquad(t\to\infty).
\end{equation}
In particular, every odd power $t^{-(2m+1)}$ has zero coefficient.
\end{theorem}

\begin{proof}
With $x=\bar\gamma+ty$ one has
\[
\delta_t(x):=\frac{h_t(x)}{g_t(x)}-1
=\int_{\RR}\bigl(R_{\Delta/t}(y)-1\bigr)\dd\nu(\gamma).
\]
Lemma~\ref{lem:entropy-uniform-remainder} gives the uniform mode
expansion, the Taylor remainder for
$\Phi(z)=(1+z)\log(1+z)-z$, and the $L^1(\omega)$ control needed to
integrate term by term.  More explicitly, for each $2\le k\le N$ it gives
\[
\bigl(\delta_t(\bar\gamma+ty)\bigr)^k
=P_{k,N,t}(y)+\varepsilon_{k,N,t}(y),
\qquad
\|\varepsilon_{k,N,t}\|_{L^1(\omega)}
=O(t^{-(2N+1)}),
\]
so integrating a power of $\delta_t$ contributes the integral of the
complete polynomial $P_{k,N,t}$ plus only an additive
$o(t^{-2N})$ error.  Therefore
\[
D_{\mathrm{KL}}(h_t\|g_t)
=\int_{\RR}\frac{\Phi(\delta_t(\bar\gamma+ty))}{\pi(1+y^2)}\dd y
\]
\[
=\sum_{k=2}^{N}\frac{(-1)^k}{k(k-1)}
\sum_{m=2k}^{2N}\frac{1}{t^m}
\sum_{\substack{n_1,\dots,n_k\ge 2\\n_1+\cdots+n_k=m}}
\mu_{n_1}\cdots\mu_{n_k}\,I_{n_1,\dots,n_k}
+o(t^{-2N}),
\]
where $I_{n_1,\dots,n_k}$ is given in
Lemma~\ref{lem:mode-parity-vanish}.  By the lemma all odd~$m$-terms
vanish, so only even powers remain.  Reindexing $m=2r$ gives
\[
=\sum_{r=2}^{N} c_{2r}\,t^{-2r}+o(t^{-2N}),
\]
which is \eqref{eq:poisson-kl-even-orders}.
\end{proof}
\leanverified{paper_circle_dimension_poisson_kl_even_orders}

\begin{theorem}[Eighth-order entropy expansion]%
\label{thm:poisson-kl-eighth}
Let $\nu$ be a probability measure on~$\RR$ with mean~$\bar\gamma$,
variance $\sigma^2>0$, and finite centred seventh absolute moment.  Then,
as $t\to\infty$,
\begin{equation}\label{eq:poisson-kl-eighth}
\begin{split}
D_{\mathrm{KL}}(h_t\|g_t)
=\frac{\sigma^4}{8t^4}
+\frac{\sigma^6+6\mu_3^2-8\sigma^2\mu_4}{64\,t^6}
\\
\quad
+\frac{3\sigma^8-12\sigma^4\mu_4+9\sigma^2\mu_3^2}{256\,t^8}
+\frac{12\sigma^2\mu_6-30\mu_3\mu_5+20\mu_4^2}{256\,t^8}
+o(t^{-8}).
\end{split}
\end{equation}
\end{theorem}

\begin{proof}
Let $\omega(\dd y)=\pi^{-1}(1+y^2)^{-1}\dd y$.  With
$x=\bar\gamma+ty$, Lemma~\ref{lem:tail-uniform-cayley-mode-remainder}
with $X=\Delta$ and truncation order~$6$ gives
\[
\delta_t(\bar\gamma+ty)
=\frac{\sigma^2}{t^2}u_2(y)
+\frac{\mu_3}{t^3}u_3(y)
+\frac{\mu_4}{t^4}u_4(y)
+\frac{\mu_5}{t^5}u_5(y)
+\frac{\mu_6}{t^6}u_6(y)
+O(t^{-7})
\]
uniformly in $y\in\RR$.  Also
$g_t(\bar\gamma+ty)\dd(\bar\gamma+ty)=\omega(\dd y)$, hence
\[
D_{\mathrm{KL}}(h_t\|g_t)
=\int_{\RR}\Phi(\delta_t(\bar\gamma+ty))\,\omega(\dd y).
\]
Set
\[
a_2:=\sigma^2u_2,\qquad a_3:=\mu_3u_3,\qquad a_4:=\mu_4u_4,
\qquad a_5:=\mu_5u_5,\qquad a_6:=\mu_6u_6.
\]
Then
\[
\delta_t=t^{-2}a_2+t^{-3}a_3+t^{-4}a_4+t^{-5}a_5+t^{-6}a_6+O(t^{-7})
\]
uniformly.  Since $\delta_t=O(t^{-2})$ and
\[
\Phi(z)=\frac12z^2-\frac16z^3+\frac1{12}z^4+O(z^5)
\qquad(z\to0),
\]
we have, uniformly in $y$,
\[
\Phi(\delta_t)
=\frac12\delta_t^2-\frac16\delta_t^3+\frac1{12}\delta_t^4
+O(t^{-10}).
\]
Write
\[
d_t:=t^{-2}a_2+t^{-3}a_3+t^{-4}a_4+t^{-5}a_5+t^{-6}a_6,
\qquad
\delta_t=d_t+\theta_t,
\]
where $\sup_y|\theta_t(y)|=O(t^{-7})$.  Then
$d_t=O(t^{-2})$ uniformly.  The finite seventh absolute moment is used
exactly here: it makes the omitted mode $\theta_t$ small enough that any
occurrence of it lies past order~$t^{-8}$.  Indeed
\[
\delta_t^2=d_t^2+2d_t\theta_t+\theta_t^2=d_t^2+O(t^{-9}),
\]
while
\[
\delta_t^3=d_t^3+O(t^{-11}),\qquad
\delta_t^4=d_t^4+O(t^{-13}).
\]
Combining these estimates with the Taylor remainder for~$\Phi$ gives the
following uniform rule for coefficients:
\[
\begin{aligned}
[t^{-m}]\Phi(\delta_t)
&=[t^{-m}]\bigl(\tfrac12d_t^2-\tfrac16d_t^3+\tfrac1{12}d_t^4\bigr),
\\
&\hspace{5em}4\le m\le8.
\end{aligned}
\]
Here coefficient extraction is understood modulo a uniform $O(t^{-9})$
remainder.  This is the only point in the proof where the seventh absolute
moment is needed beyond the lower moments appearing in the displayed
coefficients.  Equivalently,
\[
\Phi(\delta_t)
=\frac12d_t^2-\frac16d_t^3+\frac1{12}d_t^4+O(t^{-9})
=\sum_{m=4}^{8}t^{-m}B_m(y)+O(t^{-9})
\]
uniformly in $y$.  The coefficient functions are therefore obtained by
expanding only the following finite products in powers of~$t^{-1}$.  No
term containing $\theta_t$ can contribute to any retained coefficient, and
all terms of total order larger than eight are absorbed in the uniform
$O(t^{-9})$ remainder:
\begin{align*}
d_t^2
&=t^{-4}a_2^2+2t^{-5}a_2a_3
+t^{-6}(2a_2a_4+a_3^2) \\
&\quad
+t^{-7}(2a_2a_5+2a_3a_4)
+t^{-8}(2a_2a_6+2a_3a_5+a_4^2)+O(t^{-9}),\\
d_t^3
&=t^{-6}a_2^3+3t^{-7}a_2^2a_3
+t^{-8}(3a_2^2a_4+3a_2a_3^2)+O(t^{-9}),\\
d_t^4
&=t^{-8}a_2^4+O(t^{-9}).
\end{align*}
Thus
\begin{align*}
B_4&=\frac12a_2^2,\\
B_5&=a_2a_3,\\
B_6&=a_2a_4+\frac12a_3^2-\frac16a_2^3,\\
B_7&=a_2a_5+a_3a_4-\frac12a_2^2a_3,\\
B_8&=a_2a_6+a_3a_5+\frac12a_4^2
-\frac12a_2^2a_4-\frac12a_2a_3^2+\frac1{12}a_2^4.
\end{align*}
Thus the extraction rule gives $[t^{-m}]\Phi(\delta_t)=B_m$ for
$4\le m\le8$.  In particular, the $t^{-8}$ coefficient is exactly the sum
of three contributions:
\[
\frac12[t^{-8}]d_t^2
=a_2a_6+a_3a_5+\frac12a_4^2,
\]
\[
-\frac16[t^{-8}]d_t^3
=-\frac12a_2^2a_4-\frac12a_2a_3^2,
\qquad
\frac1{12}[t^{-8}]d_t^4=\frac1{12}a_2^4.
\]
The functions $B_5$ and $B_7$ are sums of products
$u_{n_1}\cdots u_{n_k}$ with odd total index, so
Lemma~\ref{lem:mode-parity-vanish} gives
\[
\int_{\RR}B_5\,\dd\omega=\int_{\RR}B_7\,\dd\omega=0.
\]
The remaining coefficients are obtained from the identities in
Appendix~\ref{app:entropy-integrals}:
\[
\int_{\RR}B_4\,\dd\omega
=\frac{\sigma^4}{2}\int_{\RR}u_2^2\,\dd\omega
=\frac{\sigma^4}{8},
\]
\[
\int_{\RR}B_6\,\dd\omega
=\sigma^2\mu_4\!\int_{\RR}u_2u_4\,\dd\omega
+\frac{\mu_3^2}{2}\!\int_{\RR}u_3^2\,\dd\omega
-\frac{\sigma^6}{6}\!\int_{\RR}u_2^3\,\dd\omega
=\frac{\sigma^6+6\mu_3^2-8\sigma^2\mu_4}{64},
\]
and
\begin{align*}
\int_{\RR}B_8\,\dd\omega
&=\sigma^2\mu_6\!\int_{\RR}u_2u_6\,\dd\omega
+\mu_3\mu_5\!\int_{\RR}u_3u_5\,\dd\omega
+\frac{\mu_4^2}{2}\!\int_{\RR}u_4^2\,\dd\omega\\
&\quad
-\frac{\sigma^4\mu_4}{2}\!\int_{\RR}u_2^2u_4\,\dd\omega
-\frac{\sigma^2\mu_3^2}{2}\!\int_{\RR}u_2u_3^2\,\dd\omega
+\frac{\sigma^8}{12}\!\int_{\RR}u_2^4\,\dd\omega\\
&=\frac{3\sigma^8-12\sigma^4\mu_4+9\sigma^2\mu_3^2}{256}
+\frac{12\sigma^2\mu_6-30\mu_3\mu_5+20\mu_4^2}{256}.
\end{align*}
Since $\omega(\RR)=1$, the integrated $O(t^{-9})$ error is
$o(t^{-8})$.  Substituting these three coefficients proves
\eqref{eq:poisson-kl-eighth}.
\end{proof}
\leanverified{paper_circle_dimension_poisson_kl_eighth}

The next two defect theorems use $C_6$ and $C_8$ only in the sense just
established: they are entropy-asymptotic coefficients under the finite
centred seventh absolute moment hypothesis of
Theorem~\ref{thm:poisson-kl-eighth}.  The fourth-moment algebra in
Subsection~\ref{subsec:fourth-moment-stability} deliberately uses separate
superscripted notation until that same seventh-moment hypothesis is imposed.

\begin{theorem}[A two-level defect ladder]%
\label{thm:poisson-defect-ladder}
Let $\nu$ be a probability measure on~$\RR$ with mean~$\bar\gamma$,
variance $\sigma^2>0$, and finite centred seventh absolute moment.  Define
\begin{align*}
C_6&:=\frac{\sigma^6+6\mu_3^2-8\sigma^2\mu_4}{64}, \\
C_8&:=
\frac{3\sigma^8-12\sigma^4\mu_4+9\sigma^2\mu_3^2+12\sigma^2\mu_6}{256}
+\frac{-30\mu_3\mu_5+20\mu_4^2}{256}.
\end{align*}
Let $X:=(\gamma-\bar\gamma)/\sigma$ on the probability space
$(\RR,\nu)$, set
\[
\beta_n:=\EE[X^n],\qquad n\in\NN,\ n\ge1,\qquad
\beta_3:=\EE[X^3],
\qquad
Q_2:=X^2-1-\beta_3X,
\]
and define
\[
Q_3:=XQ_2-\|Q_2\|_{L^2(\nu)}^2X-\frac{\beta_3}{4}Q_2.
\]
Then
\begin{equation}\label{eq:defect-c6}
\frac{64}{\sigma^6}C_6
=-7-2\beta_3^2-8\|Q_2\|_{L^2(\nu)}^2,
\end{equation}
and
\begin{equation}\label{eq:defect-c8}
\frac{256}{\sigma^8}C_8
=23
+12\|Q_3\|_{L^2(\nu)}^2
+32\|Q_2\|_{L^2(\nu)}^4
+\frac{61}{4}\beta_3^2\|Q_2\|_{L^2(\nu)}^2
+52\|Q_2\|_{L^2(\nu)}^2
+2\beta_3^4
+13\beta_3^2.
\end{equation}
In particular
\begin{equation}\label{eq:defect-c8-lower}
C_8\ge \frac{23}{256}\sigma^8,
\end{equation}
and equality holds if and only if
\[
\nu=\frac12\delta_{\bar\gamma+\sigma}
+\frac12\delta_{\bar\gamma-\sigma}.
\]
The identity \eqref{eq:defect-c6} recovers the classical Pearson
moment relation, while \eqref{eq:defect-c8} gives the next higher-order
component in the same defect coordinates.
\leanverified{paper_circle_dimension_poisson_defect_ladder}
\end{theorem}

\begin{proof}
Write
\[
\kappa:=\|Q_2\|_{L^2(\nu)}^2=\EE[Q_2^2],
\qquad
L:=\EE[XQ_2^2],
\qquad
M:=\EE[Q_2^3].
\]
Since $Q_2=X^2-1-\beta_3X$, one has
\[
\kappa=\beta_4-1-\beta_3^2,
\]
and therefore
\[
\frac{64}{\sigma^6}C_6
=1+6\beta_3^2-8\beta_4
=-7-2\beta_3^2-8\kappa,
\]
which is exactly \eqref{eq:defect-c6}.

The identity $X^2=1+\beta_3X+Q_2$ yields
\[
\beta_5=\beta_3(\beta_3^2+2+2\kappa)+L,
\]
\[
\beta_6=1+3\beta_3^2+3\kappa+\beta_3^4+3\beta_3^2\kappa+3\beta_3L+M.
\]
Hence, if
\[
P_8:=3-12\beta_4+9\beta_3^2+12\beta_6-30\beta_3\beta_5+20\beta_4^2,
\]
then a direct substitution gives
\[
P_8
=23+20\kappa^2+16\beta_3^2\kappa+64\kappa+2\beta_3^4+13\beta_3^2+6\beta_3L+12M.
\]
On the other hand,
\[
\|Q_3\|_{L^2(\nu)}^2
=M+\frac{\beta_3}{2}L+\kappa-\kappa^2+\frac{\beta_3^2}{16}\kappa.
\]
Substituting this identity into the preceding expression for $P_8$
yields \eqref{eq:defect-c8}.  Since
$P_8=256C_8/\sigma^8$, inequality \eqref{eq:defect-c8-lower} follows.

If equality holds in \eqref{eq:defect-c8-lower}, then every non-negative
term on the right-hand side of \eqref{eq:defect-c8} vanishes.  In
particular $\beta_3=0$ and $\|Q_2\|_{L^2(\nu)}=0$, so $X^2=1$
$\nu$-almost everywhere.  Since $\EE[X]=0$ and $\EE[X^2]=1$, it follows
that $X=\pm 1$ with equal weights.  Thus
\[
\nu=\frac12\delta_{\bar\gamma+\sigma}
+\frac12\delta_{\bar\gamma-\sigma}.
\]
The converse is immediate.
\end{proof}

\begin{theorem}[Defect amplification]\label{thm:poisson-defect-amplification}
Let $\nu$ be a probability measure on~$\RR$ with mean~$\bar\gamma$,
variance $\sigma^2>0$, and finite centred seventh absolute moment.  Define
\[
\Delta_6(\nu):=-\frac{7}{64}\sigma^6-C_6,
\qquad
\Delta_8(\nu):=C_8-\frac{23}{256}\sigma^8,
\]
where $C_6$ and $C_8$ are the entropy coefficients from
Theorem~\ref{thm:poisson-kl-eighth}.  Set
\[
X:=\frac{\gamma-\bar\gamma}{\sigma},
\qquad
\beta_3:=\EE[X^3],
\qquad
Q_2:=X^2-1-\beta_3X,
\]
\[
\kappa:=\|Q_2\|_{L^2(\nu)}^2,
\qquad
Q_3:=XQ_2-\kappa X-\frac{\beta_3}{4}Q_2.
\]
Then
\[
\Delta_6(\nu)
=\frac{\sigma^6}{32}\beta_3^2+\frac{\sigma^6}{8}\kappa,
\]
and
\begin{align}
\Delta_8(\nu)
&=
\frac{13\sigma^2}{8}\Delta_6(\nu)
\notag\\
&\quad
+\frac{\sigma^8}{256}\left(
12\|Q_3\|_{L^2(\nu)}^2
+32\kappa^2
+\frac{61}{4}\beta_3^2\kappa
+2\beta_3^4
\right). \label{eq:defect-amplification}
\end{align}
In particular,
\[
\Delta_8(\nu)\ge \frac{13\sigma^2}{8}\Delta_6(\nu),
\]
and equality holds if and only if
\[
\nu=\frac12\delta_{\bar\gamma+\sigma}
+\frac12\delta_{\bar\gamma-\sigma}.
\]
\leanverified{paper_circle_dimension_poisson_defect_amplification}
\end{theorem}

\begin{proof}
Equation~\eqref{eq:defect-c6} gives
\[
\Delta_6(\nu)
=-\frac{\sigma^6}{64}\left(7+\frac{64}{\sigma^6}C_6\right)
=\frac{\sigma^6}{32}\beta_3^2+\frac{\sigma^6}{8}\kappa.
\]
Likewise, subtracting $23$ from \eqref{eq:defect-c8} yields
\[
\Delta_8(\nu)
=\frac{\sigma^8}{256}\Bigl(
12\|Q_3\|_{L^2(\nu)}^2
+32\kappa^2
+\frac{61}{4}\beta_3^2\kappa
+52\kappa
+2\beta_3^4
+13\beta_3^2
\Bigr).
\]
By the preceding formula for $\Delta_6(\nu)$,
\[
\frac{13\sigma^2}{8}\Delta_6(\nu)
=\frac{\sigma^8}{256}\bigl(52\kappa+13\beta_3^2\bigr),
\]
and substituting this identity proves \eqref{eq:defect-amplification}.

If equality holds in the lower bound, then every non-negative term in
\eqref{eq:defect-amplification} vanishes.  In particular,
$\beta_3=0$ and $\kappa=0$, hence $Q_2=X^2-1$ vanishes
$\nu$-almost everywhere.  Since $\EE[X]=0$ and $\EE[X^2]=1$, it follows
that $X=\pm 1$ with equal weights, which is exactly the stated
two-point law.  The converse is immediate.
\end{proof}

\begin{example}[Gaussian input]\label{ex:gaussian-defects}
Let $\nu$ be the centred Gaussian law with variance~$\sigma^2$.  For
$X=(\gamma-\bar\gamma)/\sigma$ one has
\[
\beta_3=\beta_5=0,\qquad \beta_4=3,\qquad \beta_6=15,
\qquad \kappa=\beta_4-1=2.
\]
Hence the defect formulas give
\[
\Delta_6(\nu)=\frac{\sigma^6}{4},
\qquad
\Delta_8(\nu)=\frac{19}{16}\sigma^8.
\]
Equivalently, inserting the Gaussian moments into the coefficient
formula for $C_8$ gives
\[
\frac{256}{\sigma^8}C_8
=3-12\cdot3+12\cdot15+20\cdot9=327,
\]
and subtracting the extremal value~$23$ yields the same
$\Delta_8=(327-23)\sigma^8/256$.
\end{example}

\begin{example}[Zero-inflated symmetric two-point family]%
\label{ex:zero-inflated-defects}
Fix $0<p\le1$ and let $X_p$ be the centred variance-one random variable
defined by
\[
\mathbb P(X_p=0)=1-p,
\qquad
\mathbb P(X_p=p^{-1/2})=\mathbb P(X_p=-p^{-1/2})=\frac p2.
\]
For the law of $\bar\gamma+\sigma X_p$ one has
\[
\beta_3=\beta_5=0,\qquad
\beta_4=\frac1p,\qquad
\beta_6=\frac1{p^2},\qquad
\kappa=\frac1p-1.
\]
Therefore
\[
\Delta_6(\nu)
=\frac{\sigma^6}{8}\left(\frac1p-1\right),
\]
and
\[
\Delta_8(\nu)
=\frac{\sigma^8}{256}
\left(\frac{32}{p^2}-\frac{12}{p}-20\right)
=\frac{\sigma^8}{64}\,
\frac{(1-p)(5p+8)}{p^2}.
\]
At $p=1$ the law is the symmetric two-point extremizer and both defects
vanish; for $0<p<1$ the extra atom at the mean produces a positive
sixth- and eighth-order defect.
\end{example}

\subsection{Seventh-moment entropy stability and fourth-moment algebra}
\label{subsec:fourth-moment-stability}

The symbols $C_6$ and $C_8$ are reserved for coefficients in an entropy
asymptotic expansion.  In this paper that expansion is supplied by
Theorem~\ref{thm:poisson-kl-eighth}, whose proof uses a finite centred
seventh absolute moment to control the mode remainder after $u_6$ by
$O(t^{-7})$.  A different object is available under fourth moments alone:
the polynomial obtained by writing down the displayed formula for the
$t^{-6}$ coefficient.  The notation below separates these two objects.

\begin{itemize}[leftmargin=*]
\item With only
\[
\int_{\RR}\abs{\gamma-\bar\gamma}^{4}\,\dd\nu(\gamma)<\infty,
\]
the quantities $C_6^{(4)}(\nu)$ and $\Delta_6^{(4)}(\nu)$ are algebraic
moment polynomials.  They are not entropy-asymptotic coefficients.
\item With the stronger hypothesis
\[
\int_{\RR}\abs{\gamma-\bar\gamma}^{7}\,\dd\nu(\gamma)<\infty,
\]
Theorem~\ref{thm:poisson-kl-eighth} defines the genuine coefficient $C_6$,
and the bridge lemma below identifies
\[
C_6=C_6^{(4)}(\nu),
\qquad
\Delta_6(\nu)=\Delta_6^{(4)}(\nu).
\]
\end{itemize}
Consequently finite fourth moment alone never defines the unsuperscripted
coefficient $C_6$ or the unsuperscripted entropy defect $\Delta_6$ in this
subsection.  The order of the arguments reflects this distinction: pure
fourth-moment algebra first, then a seventh-moment bridge, and finally the
entropy-stability theorem for the genuine coefficient defect.

\paragraph{Fourth-moment algebra.}

\begin{algebralemma}[Fourth-moment algebraic defect]
Let $\nu$ have mean~$\bar\gamma$, variance $\sigma^2>0$, and finite
centred fourth moment.  Define
\[
C_6^{(4)}(\nu):=
\frac{\sigma^6+6\mu_3^2-8\sigma^2\mu_4}{64},
\qquad
\Delta_6^{(4)}(\nu):=
-\frac{7}{64}\sigma^6-C_6^{(4)}(\nu).
\]
The superscript $(4)$ is part of the notation: it records that the
quantity is a fourth-moment polynomial and carries no asymptotic assertion
by itself.  On $(\RR,\nu)$ set
\begin{align*}
X&:=\frac{\gamma-\bar\gamma}{\sigma},
&
\beta_3&:=\EE[X^3],
&
\beta_4&:=\EE[X^4],
\\
Q_2&:=X^2-1-\beta_3X,
&
\kappa&:=\|Q_2\|_{L^2(\nu)}^2.
\end{align*}
Then the pure fourth-moment identity is
\begin{equation}\label{eq:defect-sixth-fourth-moment}
\Delta_6^{(4)}(\nu)
=\frac{\sigma^6}{32}\beta_3^2+\frac{\sigma^6}{8}\kappa,
\qquad
\kappa=\beta_4-1-\beta_3^2\ge0.
\end{equation}
Moreover, $\Delta_6^{(4)}(\nu)=0$ if and only if the centred law of
$\gamma-\bar\gamma$ is $\rho_\sigma=\frac12(\delta_{-\sigma}
+\delta_\sigma)$.
\end{algebralemma}

\begin{proof}
The identity for $\kappa$ follows by expanding $Q_2^2$ and using
$\EE[X]=0$, $\EE[X^2]=1$, and $\EE[X^3]=\beta_3$:
\[
\kappa
=\EE[(X^2-1-\beta_3X)^2]
=\beta_4-1-\beta_3^2.
\]
Since $\mu_3=\sigma^3\beta_3$ and $\mu_4=\sigma^4\beta_4$,
\[
-\frac{7}{64}\sigma^6-C_6^{(4)}(\nu)
=\frac{\sigma^6}{64}\bigl(-7-1-6\beta_3^2+8\beta_4\bigr)
=\frac{\sigma^6}{32}\beta_3^2+\frac{\sigma^6}{8}\kappa.
\]
This proves \eqref{eq:defect-sixth-fourth-moment} and non-negativity.  If
$\Delta_6^{(4)}(\nu)=0$, then $\beta_3=0$ and $\kappa=0$, so
$X^2=1$ almost surely.  The conditions $\EE[X]=0$ and $\EE[X^2]=1$ then
force $X=\pm1$ with equal probabilities, i.e. the centred law is
$\rho_\sigma$.  The converse is immediate by substitution.
\end{proof}

\begin{algebralemma}[Fourth-moment algebraic coupling]
Let $\nu$ have mean~$\bar\gamma$, variance $\sigma^2>0$, and finite
centred fourth moment.  Let $\nu_c$ be the centred law of
$\Delta=\gamma-\bar\gamma$ and set
\[
\rho_\sigma:=\frac12(\delta_{-\sigma}+\delta_\sigma).
\]
For the fourth-moment polynomial $\Delta_6^{(4)}(\nu)$ defined above,
\[
\Delta_6^{(4)}(\nu)\ge0
\]
and
\begin{equation}\label{eq:fourth-moment-algebraic-coupling}
W_2(\nu_c,\rho_\sigma)
\le
4\,\sigma^{-1/2}\Delta_6^{(4)}(\nu)^{1/4}
+8\,\sigma^{-2}\Delta_6^{(4)}(\nu)^{1/2}.
\end{equation}
Moreover, $\Delta_6^{(4)}(\nu)=0$ if and only if
$\nu_c=\rho_\sigma$.  This is an algebraic fourth-moment statement; it does
not introduce an entropy coefficient.
\end{algebralemma}

\begin{proof}
All random variables below live on $(\RR,\nu)$.  Set
\[
X:=\frac{\gamma-\bar\gamma}{\sigma},
\qquad
\beta_3:=\EE[X^3],
\qquad
Q_2:=X^2-1-\beta_3X,
\qquad
\kappa:=\|Q_2\|_{L^2(\nu)}^2 .
\]
By \eqref{eq:defect-sixth-fourth-moment},
\begin{equation}\label{eq:algebraic-coupling-delta-six}
\Delta_6^{(4)}(\nu)
=\frac{\sigma^6}{32}\beta_3^2+\frac{\sigma^6}{8}\kappa .
\end{equation}
Define
\[
Y:=\mathbf{1}_{\{X\ge 0\}}-\mathbf{1}_{\{X<0\}}\in\{-1,1\}.
\]
For every real $x$, with $\operatorname{sgn}(0)=1$,
\[
|x-\operatorname{sgn}(x)|\le |x^2-1|.
\]
Therefore
\[
\EE[(X-Y)^2]\le \EE[(X^2-1)^2].
\]
The decomposition $X^2-1=\beta_3X+Q_2$ is orthogonal in $L^2(\nu)$ because
\[
\EE[XQ_2]=\EE[X^3]-\EE[X]-\beta_3\EE[X^2]=0.
\]
Thus
\[
\EE[(X^2-1)^2]
=\beta_3^2+\kappa
\le 2\beta_3^2+8\kappa
=\frac{64}{\sigma^6}\Delta_6^{(4)}(\nu),
\]
where the last equality is \eqref{eq:algebraic-coupling-delta-six}.  Since
$\EE[X]=0$,
\[
|\EE[Y]|=|\EE[Y-X]|
\le \EE[|Y-X|]
\le \EE[(Y-X)^2]^{1/2}
\le \frac{8}{\sigma^3}\Delta_6^{(4)}(\nu)^{1/2}.
\]
Let $Z$ be a symmetric Rademacher random variable.  If
$p=\mathbb P(Y=1)$, then $\EE[Y]=2p-1$ and the optimal transport between
$\mathcal L(Y)$ and the symmetric law on $\{-1,1\}$ moves
$|p-1/2|$ mass across distance~$2$.  Hence
\[
W_2(\mathcal L(Y),\mathcal L(Z))^2
=4|p-1/2|=2|\EE[Y]|
\le \frac{16}{\sigma^3}\Delta_6^{(4)}(\nu)^{1/2}.
\]
Also
\[
W_2(\mathcal L(X),\mathcal L(Y))
\le \EE[(X-Y)^2]^{1/2}
\le \frac{8}{\sigma^3}\Delta_6^{(4)}(\nu)^{1/2}.
\]
Since $\nu_c=\mathcal L(\sigma X)$ and
$\rho_\sigma=\mathcal L(\sigma Z)$, taking square roots in the preceding
bound for $W_2(\mathcal L(Y),\mathcal L(Z))^2$ and then using scaling and
the triangle inequality gives \eqref{eq:fourth-moment-algebraic-coupling}.
The zero statement follows from the preceding fourth-moment algebraic
defect lemma.
\end{proof}

\paragraph{Seventh-moment bridge.}

\begin{algebralemma}[Seventh-moment bridge from algebra to entropy]
Assume that $\nu$ has mean~$\bar\gamma$, variance $\sigma^2>0$, and
\[
\int_{\RR}\abs{\gamma-\bar\gamma}^{7}\,\dd\nu(\gamma)<\infty.
\]
Let $C_6$ be the $t^{-6}$ entropy coefficient supplied by
Theorem~\ref{thm:poisson-kl-eighth}, and let $C_6^{(4)}(\nu)$ and
$\Delta_6^{(4)}(\nu)$ be the fourth-moment polynomials defined in the
fourth-moment algebraic defect lemma.  Then
\begin{equation}\label{eq:delta-six-bridge}
C_6=C_6^{(4)}(\nu),
\qquad
\Delta_6(\nu)
:=-\frac{7}{64}\sigma^6-C_6
=\Delta_6^{(4)}(\nu).
\end{equation}
Thus the unsuperscripted entropy defect $\Delta_6$ is introduced only under
the finite seventh absolute moment hypothesis.
\end{algebralemma}

\begin{proof}
The finite seventh absolute moment implies the fourth moment needed to
define $C_6^{(4)}(\nu)$ and $\Delta_6^{(4)}(\nu)$.  It also gives the
uniform remainder in the proof of Theorem~\ref{thm:poisson-kl-eighth}.
Comparing the $t^{-6}$ coefficient in
\eqref{eq:poisson-kl-eighth} with the definition of $C_6^{(4)}(\nu)$ gives
$C_6=C_6^{(4)}(\nu)$.  Substituting this identity into
$\Delta_6(\nu):=-7\sigma^6/64-C_6$ and using
\eqref{eq:defect-sixth-fourth-moment} gives the displayed bridge identity.
\end{proof}

\paragraph{Entropy layer.}
With only a finite centred fourth moment, the algebraic lemmas prove
\eqref{eq:defect-sixth-fourth-moment} and
\eqref{eq:fourth-moment-algebraic-coupling} for the superscripted
polynomial $\Delta_6^{(4)}(\nu)$.  Without the additional assumption
\[
\int_{\RR}\abs{\gamma-\bar\gamma}^{7}\,\dd\nu(\gamma)<\infty,
\]
Theorem~\ref{thm:poisson-kl-eighth} is not available, so the
unsuperscripted coefficient $C_6$ and defect $\Delta_6$ are not defined.

\begin{theorem}[Finite-seventh-moment stability for the sixth-order entropy defect]%
\label{thm:poisson-defect-stability}
Let $\nu$ be a probability measure on~$\RR$ with mean~$\bar\gamma$,
variance $\sigma^2>0$, and finite centred seventh absolute moment, i.e.
\[
\int_{\RR}\abs{\gamma-\bar\gamma}^{7}\,\dd\nu(\gamma)<\infty.
\]
Let
$C_6$ be the $t^{-6}$ coefficient in
Theorem~\ref{thm:poisson-kl-eighth}, which exists under the displayed
seventh-moment hypothesis, and define the sixth-order entropy defect by
\[
\Delta_6(\nu):=-\frac{7}{64}\sigma^6-C_6.
\]
The finite seventh absolute moment hypothesis is part of the definition of
the object used in this theorem: $\Delta_6$ is the entropy coefficient
defect supplied by Theorem~\ref{thm:poisson-kl-eighth}.  The fourth-moment
polynomial is the superscripted quantity $\Delta_6^{(4)}$ from the
fourth-moment algebraic defect lemma, and the bridge identity
\eqref{eq:delta-six-bridge} identifies it with $\Delta_6$ only under the
seventh-moment hypothesis.  Under finite fourth moment alone this theorem
is not asserted; only the separate algebraic coupling estimate
\eqref{eq:fourth-moment-algebraic-coupling} for $\Delta_6^{(4)}$ remains.
Let $\nu_c$ be the centred law of $\Delta=\gamma-\bar\gamma$ and set
\[
\rho_\sigma:=\frac12(\delta_{-\sigma}+\delta_\sigma).
\]
Then $\Delta_6(\nu)\ge0$ and
\[
W_2(\nu_c,\rho_\sigma)
\le
4\,\sigma^{-1/2}\Delta_6(\nu)^{1/4}
+8\,\sigma^{-2}\Delta_6(\nu)^{1/2}.
\]
In particular, for fixed $\sigma$,
\[
W_2(\nu_c,\rho_\sigma)=O\!\bigl(\Delta_6(\nu)^{1/4}\bigr)
\qquad\text{as }\Delta_6(\nu)\downarrow 0.
\]
Moreover $\Delta_6(\nu)=0$ if and only if $\nu_c=\rho_\sigma$.
\leanverified{paper_circle_dimension_poisson_defect_stability}
\end{theorem}

\begin{proof}
The displayed seventh absolute moment condition is exactly the hypothesis
of the bridge lemma above.  Hence
\[
\Delta_6(\nu)=\Delta_6^{(4)}(\nu).
\]
The finite fourth moment needed in the preceding algebraic coupling lemma
is automatic.  Substituting the bridge identity into
\eqref{eq:fourth-moment-algebraic-coupling} gives the stated $W_2$ estimate,
and the non-negativity and zero characterization are those of the
fourth-moment algebraic defect after the same substitution.  The asymptotic
$O(\Delta_6^{1/4})$ statement for fixed $\sigma$ follows immediately from
the displayed estimate.
\end{proof}

\begin{remark}[Fourth-moment coupling for the algebraic defect]\label{rem:fourth-moment-stability-algebraic}
The finite-fourth-moment-only conclusion is exactly the algebraic estimate
\eqref{eq:fourth-moment-algebraic-coupling}.  It is deliberately stated
before the seventh-moment bridge because it concerns only the polynomial
$\Delta_6^{(4)}$.  It does not assert the existence of a $t^{-6}$ entropy
coefficient under only fourth moments, and it intentionally contains
neither $C_6$ nor the unsuperscripted entropy defect $\Delta_6$.
\end{remark}

\begin{remark}
The estimate is a coefficient-level analogue of stability-to-extremizers
results for entropy deficits, such as the logarithmic
Hardy--Littlewood--Sobolev stability theorem of Carlen
\cite{Carlen2025LogHLS} and the logarithmic Sobolev stability theorem of
Brigati--Dolbeault--Simonov
\cite{BrigatiDolbeaultSimonov2024LogSobolev}.  Here the defect is not a
functional-inequality deficit: the asymptotic defect $\Delta_6$ is the
coefficient-level quantity obtained from the Poisson expansion, and the
proof reduces it to the fourth-moment polynomial
\eqref{eq:defect-sixth-fourth-moment} before applying an explicit coupling
to the symmetric two-point law.
\end{remark}
