\section{Introduction}\label{sec:introduction}

For a primitive finite matrix \(A\), the dynamical \(\zeta\)-function
\[
  \zeta_A(z)=\det(I-zA)^{-1}
\]
packages periodic-orbit data, transfer-operator traces, and non-Perron
spectral data into one holomorphic object. In symbolic dynamics this
identity is classical and can be viewed as the finite-state shadow of a
broader Fredholm determinant identity valid for general trace-class
operators.

This paper separates three layers that are often used together. The
first layer is connective and classical: the Fredholm determinant
recovers both the trace sequence and the associated Witt coordinates.
The second layer is the organizing result of the paper: using the
standard spectral description of compact normal trace-class operators, we
identify exactly when an abstract Euler product admits a trace-class
normal realisation by cyclic blocks, and we show that in the normal
class this cyclic-block geometry is forced and minimal for the closed
span of the non-zero spectral subspaces. The
third layer returns to symbolic dynamics as an application: once the
determinant is taken as the primary object, finite-part constants, Abel
renormalisations, orbit-weight moments, and limit-law data can all be
organized within one operator-theoretic package.

The novelty claimed here is deliberately narrow. We do not propose a new
Fredholm determinant identity, a new symbolic model, or a new spectral-gap
limit theorem. The cyclic-block theorem is a precise trace-class normal
packaging of standard compact-normal spectral facts for prescribed
countable Euler products: it gives an if-and-only-if realisation
criterion, identifies the forced closed non-zero spectral support, and
records the exact trace-norm threshold. That threshold is not merely
constructional: in the normal category it is the exact trace budget forced
by the cyclic zero divisor of the determinant, and in infinite rank it
forces a genuine support/range dichotomy because the algebraic range is
then dense and nonclosed inside the closed non-zero spectral support. The
Fredholm--Witt bridge and the
spectral rigidity statement are included because they are the two
classical links that make this characterisation useful in the symbolic
applications: the former turns periodic data into determinants, and the
latter turns determinant identities back into spectral information. The
symbolic section is application-level throughout: it repackages standard
perturbative and spectral-gap consequences at the determinant level, with
the separate expansion/nonresonance assumption stated whenever secondary
determinants enter; it is not presented as a second source of structural
novelty.

\subsection{Main results}

We begin by recording the classical bridge in the form used throughout
the paper.
\begin{proposition}[Connecting Fredholm--Witt identity]
  \label{prop:intro-fredholm-witt}
  Let \(T\in \sone(H)\) be a trace-class operator on a complex Hilbert
  space. Define
  \[
    a_n(T):=\Tr(T^n),
    \qquad
    p_n(T):=\frac{1}{n}\sum_{d\mid n}\mu(d)\,a_{n/d}(T)
    \qquad (n\ge 1).
  \]
  Then for every \(\abs{z}<\min(1,\norm{T}^{-1})\),
  \[
    \det(I-zT)^{-1}
    =\exp\!\left(\sum_{n\ge 1}\frac{a_n(T)}{n}z^n\right)
    =\prod_{n\ge 1}(1-z^n)^{-p_n(T)}.
  \]
\end{proposition}

The main structural theorem records the realization statement with
necessity and closed-support optimality in the normal trace-class
category. Its determinant convention is part of the statement: finite
truncations give positive-exponent determinants, and reciprocal
negative-exponent products are used only after a zero-free domain has
been fixed.
\begin{theorem}[Positive-exponent cyclic-block determinant and closed-support optimality]
  \label{thm:intro-cyclic-characterization}
  \leanverified{paper_fredholm_witt_intro_cyclic_characterization}
  Let \(\{(n_j,\beta_j,m_j)\}_{j\in J}\) be a countable family with
  \(n_j,m_j\in\NN\), \(\beta_j\in\CC\setminus\{0\}\), and
  \[
    \sum_{j\in J}m_j n_j\abs{\beta_j}^{1/n_j}<\infty.
  \]
  Then the positive-exponent product
  \[
    \prod_{j\in J}(1-\beta_j z^{n_j})^{m_j}
  \]
  is the Fredholm determinant of the cyclic-block trace-class normal
  operator
  \[
    \bigoplus_{j\in J}\bigl(C(n_j,\alpha_j)\otimes I_{m_j}\bigr),
    \qquad \alpha_j^{n_j}=\beta_j.
  \]
  Conversely, if a trace-class normal operator has this Fredholm
  determinant, then its non-zero spectral support is unitarily equivalent
  to the non-zero part of this cyclic-block direct sum; equivalently, after
  adjoining an arbitrary zero summand it is unitarily equivalent to the
  displayed model. This is a closed-support classification, not an
  assertion that the algebraic range is the whole support. Here, in the
  infinite case, the negative-exponent product
  \[
    \prod_{j\in J}(1-\beta_j z^{n_j})^{-m_j}
  \]
  means the reciprocal of the locally uniform positive-exponent Fredholm
  determinant on a zero-free neighbourhood of the origin; it is not the
  Fredholm determinant computed on finite truncations. Equivalently,
  realising the reciprocal Euler product is always shorthand for
  realising this positive-exponent Fredholm determinant and then taking
  its reciprocal only where it is non-zero.
  Consequently every trace-class normal realization has closed span of
  non-zero spectral subspaces of dimension
  \(
    \sum_{j\in J}m_j n_j
  \)
  and trace norm
  \(
    \sum_{j\in J}m_j n_j\abs{\beta_j}^{1/n_j}
  \).
  In particular, for a single factor \((1-\beta z^n)^{-m}\), the closed
  span of the non-zero spectral subspaces has dimension exactly \(mn\).
  In infinite direct sums this is a statement about the closed span, not
  about equality of the algebraic range with that span; the algebraic
  range may be a proper dense nonclosed subspace.
  These counts are attached to the chosen Euler data
  \(\mathcal D=\{(n_j,\beta_j,m_j)\}\); distinct but equivalent
  factorizations of the same Euler product produce unitarily equivalent
  normal models. The source theorem chain makes this sharp through
  \Cref{prop:sharp-trace-budget-support-obstruction}: the same cyclic
  zero divisor already forces this trace norm and this closed support for
  every normal trace-class realization.
\end{theorem}

Once the determinant is fixed, the non-zero spectrum is fixed as well.
The next statement is classical but important because it is the
mechanism that turns determinant identities into rigidity.
\begin{proposition}[Spectral rigidity]
  \label{prop:intro-spectral-rigidity}
  Let \(T_1,T_2\in \sone\). If
  \[
    \det(I-zT_1)=\det(I-zT_2)
  \]
  in a neighbourhood of \(z=0\), then \(T_1\) and \(T_2\) have the same
  non-zero eigenvalues with algebraic multiplicity.
\end{proposition}

Section~4 records determinant-level applications in two directions. These
results are kept explicitly at the application level: they show how the
determinant package interfaces with standard perturbative spectral-gap
theory, but they are not presented as a second structural novelty beside
\Cref{thm:cyclic-fredholm-realisation}. The
first concerns finite-part constants and Abel renormalisations. The
reduced determinant uses only relative branch isolation. Every statement
that involves \(I-\lambda^{-k}A\) uses explicit secondary admissibility:
the expanded relative branch. Concretely, this includes the uniform
complementary relative gap and the expansion condition
\[
  \inf_{\theta\in U}\abs{\lambda(\theta)}>1,
\]
which is equivalent, under the relative gap, to the exact subunit
secondary-radius condition
\[
  \sup_{\theta\in U}\sup_{k\ge2}
  \rho(\lambda(\theta)^{-k}A_\theta)<1 .
\]
The radius condition is the condition used to define the principal
trace-log branches; it is not a consequence of the relative gap alone. The
second concerns
twisted determinants attached to locally constant
observables, which
simultaneously encode periodic-orbit moments and limit-law data. These
are applications of standard finite-dimensional perturbation theory and
spectral-gap methods, rather than additional claims of structural
novelty. In particular,
\Cref{prop:secondary-fredholm-normal-convergence},
\Cref{thm:logM-holomorphic-variation}, and
\Cref{prop:logM-truncation} separate the normal-convergence,
differentiation, and truncation statements needed for
\(\log\MM(A_\theta)\).
\Cref{prop:twisted-determinant-orbit-moments},
\Cref{thm:twisted-determinant-rigidity}, and
\Cref{cor:twisted-determinant-cumulants} separate periodic-orbit moment
recovery from cumulant rigidity for twisted determinants. Finally,
\Cref{prop:off-origin-decay} and \Cref{thm:clt-spectral} state the off-origin
decay and central limit theorem in the exact operator form actually
used: decay is controlled by the spectral radius of the centered
twisted matrix family on compact sets away from the arithmetic lattice
obstruction.

\subsection{Hypothesis map}

The logical dependencies are as follows.
\begin{enumerate}[label=\textup{(\roman*)}]
  \item The Fredholm--Witt identity uses only \(T\in\sone(H)\).
  \item The cyclic-block theorem uses
  \[
    \sum_j m_jn_j\abs{\beta_j}^{1/n_j}<\infty
  \]
  for existence of the trace-class model, and normality for the converse
  unitary classification. Its optimality statement concerns the closed
  support \(\mathsf S(L)\), not the algebraic range.
  \item The spectral-rigidity proposition uses only equality of trace-class
  Fredholm determinant germs.
  \item In the perturbative section, branch isolation alone is used only for
  the Riesz branch, reduced resolvent, and reduced determinant. Every full
  secondary factor \(I-\lambda^{-k}A\), every \(\Phi_k\), and the full Abel
  finite part \(F(A,\lambda)\) require secondary admissibility: expansion
  of the selected eigenvalue, \(\abs{\lambda}>1\) pointwise or
  \(\inf_U\abs{\lambda}>1\) in families, together with the complementary
  relative gap. Under that relative gap, these expansion conditions are
  equivalent to \(s(A,\lambda)<1\) pointwise and \(s_U<1\) in families.
  \item The twisted determinant and limit-law statements are finite-state
  applications under primitivity and local constancy of the observable.
\end{enumerate}

\subsection{Previous work}

The dynamical \(\zeta\)-function viewpoint goes back at least to
Artin--Mazur \cite{ArtinMazur1965}. In the symbolic finite-state setting,
the rationality of these \(\zeta\)-functions goes back to
Bowen--Lanford and Manning
\cite{BowenLanford1970Zeta,Manning1971Axiom}; see also
\cite{Ruelle1976ZetaExpanding,Ruelle1978Thermodynamic,ParryPollicott1990Zeta,LindMarcus1995,Kitchens1998}.
The necklace and Witt-vector algebra behind the M\"obius inversion formulas
used here is classical; see Metropolis--Rota \cite{MetropolisRota1983}.
The theory of Fredholm determinants for trace-class operators, and more
generally for operators in trace ideals, originates with Grothendieck
\cite{Grothendieck1956}, is extended to the nuclear setting by Ruelle
\cite{Ruelle1990Extension}, and is developed systematically in Simon
\cite{Simon2005TraceIdeals}. The product representation for trace-class
Fredholm determinants that underlies the spectral rigidity discussion is
\cite[Theorem~3.7]{Simon2005TraceIdeals}.

The operator-theoretic approach to dynamical zeta-functions via
Fredholm determinants of transfer operators was developed by Pollicott
\cite{Pollicott1986Meromorphic} in the Axiom~A setting, and the
Fredholm theory of dynamical determinants for piecewise invertible
systems in the trace-class sense was carried out by Keller
\cite{Keller1989Fredholm}. The perturbative arguments in the symbolic
applications draw on transfer-operator spectral theory in the sense of
Ruelle and Baladi
\cite{Ruelle1969,Ruelle1978Thermodynamic,Baladi2000PositiveTransfer}.
For finite-state spectral-gap CLT, cumulants, and perturbative
regularity of eigenvalues, we also follow the standard finite-dimensional
framework described in
\cite[Chapter~5]{Ruelle1978Thermodynamic},
\cite[Chapter~3]{Baladi2000PositiveTransfer},
\cite[Chapters~I and~IV]{HennionHerve2001},
and \cite[Chapter~II]{Kato1995Perturbation}.
The spectral-gap approach to limit theorems for Markov chains was
initiated by Nagaev \cite{Nagaev1957} and extended to the dynamical
setting by Guivarc'h and Hardy \cite{GuivarchHardy1988}; for broader
stability results for transfer-operator spectra, see also
\cite[Theorem~1 and Corollary~1]{KellerLiverani1999}. For group inverses
and reduced resolvents we use standard matrix-analytic material from
\cite{CampbellMeyer1979}, while the analytic perturbation theory for
isolated eigenvalues is taken from Kato \cite{Kato1995Perturbation}.
The finite-state Liv\v{s}ic-type coboundary criterion used in the
off-origin Fourier estimate is cited in the form of Parry's cocycle
theorem \cite[Theorem~1]{Parry1992Cocycle}. For the parameter
regularity of holomorphic families of dynamical determinants beyond the
finite-dimensional setting, see J\'ezéquel \cite{Jezequel2017Parameter}.

The present note does not claim a new symbolic model, a new Fredholm
identity, or a new spectral-gap limit theorem. The structural point is
the cyclic-block theorem of
\Cref{thm:cyclic-fredholm-realisation}: sufficiency, necessity in the
normal trace-class class, and optimality of the resulting non-zero block
size. The remaining operator statements are classical connective input
or determinant-level applications, included because they place the
cyclic-block theorem in the right framework and show how the same
determinant package interfaces with standard symbolic perturbation
theory.

\subsection{Organisation}

\Cref{sec:preliminaries} recalls the finite-state \(\zeta\)-function,
primitive orbit counts, and the finite-part constants used later.
\Cref{sec:fredholm} contains the Fredholm--Witt bridge, the
cyclic-block characterization theorem, its immediate closed
spectral-subspace optimality corollary, and the classical determinant
rigidity statement.
\Cref{sec:perturbation} develops the symbolic applications in separate
blocks: normal convergence and truncation for the Abel finite part,
twisted-determinant moment and cumulant rigidity, the spectral-radius
form of off-origin decay together with the standard CLT, explicit
computational examples, and a brief comparison with classical
finite-dimensional perturbation theory. \Cref{sec:conclusion} closes
with the logical status of the main ingredients and the scope of the
operator-theoretic package.
