\section{Fredholm Determinants and Cyclic-Block Realisations}
\label{sec:fredholm}

\subsection{The Fredholm--Witt bridge}

Let \(H\) be a complex Hilbert space and let \(T\in \sone(H)\) be a
trace-class operator. Define
\[
  a_n(T):=\Tr(T^n),
  \qquad
  p_n(T):=\frac{1}{n}\sum_{d\mid n}\mu(d)\,a_{n/d}(T)
  \qquad (n\ge 1).
\]

\begin{proposition}[Connecting Fredholm--Witt identity]
  \label{thm:fredholm-witt-bridge}
  The following is a direct consequence of the classical Fredholm
  determinant identity; we record it here to fix notation and to make
  the Euler-product structure explicit. For every trace-class operator
  \(T\) and every \(\abs{z}<\min(1,\norm{T}^{-1})\),
  \[
    \det(I-zT)^{-1}
    =\exp\!\left(\sum_{n\ge 1}\frac{a_n(T)}{n}z^n\right)
    =\prod_{n\ge 1}(1-z^n)^{-p_n(T)},
  \]
  where \(\log(1-z^m)\) denotes the principal branch on \(\abs{z}<1\).
  This identity holds as an equality of holomorphic germs at \(z=0\);
  equivalently, it is an identity of formal power series. The
  exponents \(p_n(T)\) are in general not integers, so the Euler
  product on the right is not a finite collection of zero-free factors,
  and the expression should not be extended beyond the common disc of
  convergence without an explicit branch choice.
\end{proposition}

\begin{proof}
  The Fredholm determinant identity
  \[
    \det(I-zT)^{-1}
    =\exp\!\left(\sum_{n\ge 1}\frac{\Tr(T^n)}{n}z^n\right)
  \]
  is classical for trace-class operators; see Grothendieck
  \cite{Grothendieck1956} and \cite[Chapter~3]{Simon2005TraceIdeals}.
  By definition of \(p_n(T)\), M\"obius inversion gives
  \[
    a_n(T)=\sum_{d\mid n}d\,p_d(T).
  \]
  Hence
  \begin{align*}
    \sum_{n\ge 1}\frac{a_n(T)}{n}z^n
    &=\sum_{n\ge 1}\sum_{d\mid n}\frac{d\,p_d(T)}{n}z^n \\
    &=\sum_{m\ge 1}\sum_{r\ge 1}\frac{p_m(T)}{r}z^{mr}
      =-\sum_{m\ge 1}p_m(T)\log(1-z^m).
  \end{align*}
  Here \(\log(1-z^m)\) is the principal branch, which is single-valued
  for \(\abs{z^m}<1\), i.e.\ for \(\abs{z}<1\). Together with the
  convergence of the trace series for \(\abs{z}<\norm{T}^{-1}\), the
  calculation is valid for \(\abs{z}<\min(1,\norm{T}^{-1})\).
  Exponentiating gives the Euler product.
\end{proof}

\begin{remark}[Loss of arithmetic constraints]
  \label{rem:witt-audit-constraints}
  \Cref{thm:fredholm-witt-bridge} is purely operator-theoretic. In
  contrast with \Cref{rem:finite-state-arithmetic}, the coordinates
  \(p_n(T)\) need not be integers and need not be non-negative. Thus
  trace-class realisability is strictly broader than finite-state
  realisability.
\end{remark}

\subsection{A converse realisation by cyclic blocks}

For \(n\ge 1\), let \(\Pi_n\) denote the cyclic permutation matrix on
\(\CC^n\). For \(\alpha\in\CC\) define the cyclic block
\[
  C(n,\alpha):=\alpha\Pi_n.
\]
Since the eigenvalues of \(\Pi_n\) are the \(n\)th roots of unity, one
has
\begin{equation}\label{eq:cycle-block-det}
  \det(I-zC(n,\alpha))=1-\alpha^n z^n.
\end{equation}

\begin{lemma}[Single cyclic-block spectrum]
  \label{lem:single-cyclic-block-spectrum}
  Let \(n,m\in\NN\), \(\beta\in\CC\setminus\{0\}\), and
  \(\alpha^n=\beta\). Then the non-zero eigenvalues of
  \(C(n,\alpha)\otimes I_m\) are
  \[
    \alpha\omega,\qquad \omega^n=1,
  \]
  each with multiplicity \(m\), and
  \[
    \det\!\bigl(I-z(C(n,\alpha)\otimes I_m)\bigr)
    =(1-\beta z^n)^m .
  \]
\end{lemma}

\begin{proof}
  The Fourier basis diagonalizes \(\Pi_n\) with eigenvalues
  \(\omega\), \(\omega^n=1\). Hence \(C(n,\alpha)\otimes I_m\) has
  eigenvalues \(\alpha\omega\), each repeated \(m\) times. Therefore
  \[
    \det\!\bigl(I-z(C(n,\alpha)\otimes I_m)\bigr)
    =\prod_{\omega^n=1}(1-z\alpha\omega)^m
    =(1-\alpha^nz^n)^m
    =(1-\beta z^n)^m .
  \]
\end{proof}

\begin{lemma}[Finite cyclic truncation determinant with positive exponents]
  \label{lem:finite-cyclic-truncation-determinant}
  Let \(F\) be a finite set of indices and put
  \[
    L_F:=\bigoplus_{j\in F}\bigl(C(n_j,\alpha_j)\otimes I_{m_j}\bigr),
    \qquad \alpha_j^{n_j}=\beta_j.
  \]
  Then, as a finite-dimensional determinant identity,
  \[
    \det(I-zL_F)
    =\prod_{j\in F}(1-\beta_j z^{n_j})^{m_j},
    \qquad
    \det(I-zL_F)^{-1}
    =\prod_{j\in F}(1-\beta_j z^{n_j})^{-m_j}
  \]
  wherever the reciprocal is defined.
\end{lemma}

\begin{proof}
  The eigenvalues of the \(j\)-th block are
  \(\alpha_j\omega\), \(\omega^{n_j}=1\), each repeated \(m_j\) times.
  Hence
  \begin{align*}
    \det(I-zL_F)
    &=\prod_{j\in F}\prod_{\omega^{n_j}=1}
      (1-z\alpha_j\omega)^{m_j} \\
    &=\prod_{j\in F}(1-\alpha_j^{n_j}z^{n_j})^{m_j}
      =\prod_{j\in F}(1-\beta_j z^{n_j})^{m_j}.
  \end{align*}
  Taking reciprocals gives the second identity on the complement of the
  finite zero set of the determinant.
\end{proof}

\begin{lemma}[Trace-norm limit convention for cyclic products]
  \label{lem:cyclic-product-trace-norm-limit}
  Assume
  \[
    \sum_{j\in J}m_j n_j\abs{\beta_j}^{1/n_j}<\infty
  \]
  and choose \(\alpha_j^{n_j}=\beta_j\). Let
  \[
    H_{\mathcal D}
    :=\bigoplus_{j\in J}\bigl(\CC^{n_j}\otimes\CC^{m_j}\bigr),
    \qquad
    L_{\mathcal D}
    :=\bigoplus_{j\in J}\bigl(C(n_j,\alpha_j)\otimes I_{m_j}\bigr)
  \]
  on \(H_{\mathcal D}\), and let \(L_{\mathcal D}^{(N)}\) be any finite
  truncation along an enumeration of \(J\). Then
  \(L_{\mathcal D}^{(N)}\to L_{\mathcal D}\) in trace norm and
  \[
    \det(I-zL_{\mathcal D})
    =\lim_{N\to\infty}
      \prod_{j\le N}(1-\beta_j z^{n_j})^{m_j}
  \]
  locally uniformly for \(z\in\CC\). On any connected zero-free
  neighbourhood \(\Omega\ni0\) for this determinant,
  \[
    \det(I-zL_{\mathcal D})^{-1}
    =\lim_{N\to\infty}
      \prod_{j\le N}(1-\beta_j z^{n_j})^{-m_j}
  \]
  locally uniformly on \(\Omega\). Thus the negative-exponent product is
  used only as the reciprocal of the non-vanishing Fredholm determinant on
  such a domain.
\end{lemma}

\begin{proof}
  The singular values of the \(j\)-th block are all
  \(\abs{\alpha_j}=\abs{\beta_j}^{1/n_j}\), with multiplicity
  \(m_jn_j\). Hence
  \[
    \norm{L_{\mathcal D}-L_{\mathcal D}^{(N)}}_1
    =\sum_{j>N}m_j n_j\abs{\beta_j}^{1/n_j}\to0 .
  \]
  Equivalently, the net of finite partial direct sums over finite
  subsets of \(J\) converges unconditionally in trace norm; no ordering of
  the factors is part of the determinant definition.
  For each fixed \(N\), \Cref{lem:finite-cyclic-truncation-determinant}
  gives
  \[
    \det(I-zL_{\mathcal D}^{(N)})
    =\prod_{j\le N}(1-\beta_j z^{n_j})^{m_j}.
  \]
  Trace-norm continuity of the Fredholm determinant
  \cite[Theorem~3.4]{Simon2005TraceIdeals},
  \[
    \abs{\det(I+K)-\det(I+L)}
    \le\norm{K-L}_1\exp\!\bigl(1+\norm{K}_1+\norm{L}_1\bigr),
  \]
  applied to \(K=-zL_{\mathcal D}^{(N)}\) and
  \(L=-zL_{\mathcal D}\), gives locally uniform convergence in \(z\).
  If the limiting determinant is non-zero on \(\Omega\), then its
  reciprocal is holomorphic there; local uniform convergence of the
  reciprocals follows as follows. On a compact set
  \(K\Subset\Omega\), the limiting determinant has a positive minimum
  modulus. The locally uniform convergence of the finite determinants
  therefore makes all sufficiently large finite determinants non-zero on
  \(K\), and the reciprocal difference is controlled by the usual
  identity
  \[
    f_N^{-1}-f^{-1}=\frac{f-f_N}{f_Nf}.
  \]
  Hence the reciprocal finite products converge locally uniformly to
  \(D_{\mathcal D}^{-1}\) on \(\Omega\).
\end{proof}

\noindent\emph{Determinant convention.}
In the infinite cyclic-block products below, the positive-exponent product
is the Fredholm determinant and is obtained as an entire locally uniform
limit of finite truncation determinants. The product with negative
exponents is not an independently defined infinite product. It always
means the reciprocal of that non-vanishing Fredholm determinant on a
specified connected zero-free neighbourhood of the origin. Thus finite
truncations are computed first, the Fredholm determinant is then obtained
by trace-norm continuity, and reciprocals are taken only after a zero-free
domain has been fixed.

\noindent\emph{Reciprocal-product realisation.}
For data \(\mathcal D=\{(n_j,\beta_j,m_j)\}_{j\in J}\) satisfying the
summability hypothesis, write
\[
  D_{\mathcal D}(z):=\prod_{j\in J}(1-\beta_j z^{n_j})^{m_j}
\]
for the locally uniform limit of the positive-exponent finite products.
A trace-class operator \(L\) will be said to realise the reciprocal Euler
product
\(\prod_{j\in J}(1-\beta_j z^{n_j})^{-m_j}\) only in the following
determinant-side sense: on some connected zero-free neighbourhood
\(\Omega\ni0\) for \(D_{\mathcal D}\),
\[
  \begin{aligned}
    \det(I-zL)&=D_{\mathcal D}(z),\\
    \det(I-zL)^{-1}
      &=D_{\mathcal D}(z)^{-1}
        =\prod_{j\in J}(1-\beta_j z^{n_j})^{-m_j}
        \quad (z\in\Omega).
  \end{aligned}
\]
This terminology never reverses the finite-dimensional calculation:
for every finite truncation \(F\), the Fredholm determinant is
\(\prod_{j\in F}(1-\beta_j z^{n_j})^{m_j}\), and the negative-exponent
expression is its reciprocal where the determinant is non-zero.

\begin{definition}[Closed non-zero spectral support]
  \label{def:closed-nonzero-support}
For a normal trace-class operator \(L\), write
\[
  \mathsf S(L):=
  \overline{\mathrm{span}}
  \{\,\ker(L-\lambda I):\lambda\in\spec(L)\setminus\{0\}\,\}.
\]
\end{definition}

\begin{lemma}[Closed support is the closure of the range]
  \label{lem:normal-support-range-closure}
  If \(L\) is normal and trace-class, then
  \[
    \mathsf S(L)=(\ker L)^\perp=\overline{\mathrm{ran}(L)}.
  \]
  This identifies the closed support with the closure of the algebraic
  range. It does not assert that the algebraic range is closed.
\end{lemma}

\begin{proof}
  A trace-class normal operator is compact, so the spectral theorem gives
  an orthogonal decomposition into \(\ker L\) and the finite-dimensional
  eigenspaces corresponding to non-zero eigenvalues, with possible
  accumulation only at \(0\). The closed span of the non-zero eigenspaces is
  therefore \((\ker L)^\perp\). The range contains the finite linear span of
  the non-zero eigenspaces, because \(L\) is invertible on each such
  eigenspace, and it is contained in \((\ker L)^\perp\). Taking closures
  gives the displayed identity.
\end{proof}

\noindent\emph{Range convention.}
For an infinite direct sum of the cyclic blocks satisfying the trace-class
summability hypothesis below, the block singular values tend to \(0\), and
\(\mathrm{ran}(L_{\mathcal D})\) is a proper dense nonclosed subspace. Thus
identifying the closed support of \(L_{\mathcal D}\) with
\(H_{\mathcal D}\) is never shorthand for saying that the algebraic range is
all of \(H_{\mathcal D}\). The exact algebraic range is computed separately
in \Cref{lem:cyclic-sum-range-closure}, and every
optimality statement in this subsection is a statement about
\(\mathsf S(L)\).

\begin{proposition}[Sharp trace-norm budget for cyclic zero divisors]
  \label{prop:sharp-trace-budget-support-obstruction}
  Let \(\mathcal D=\{(n_j,\beta_j,m_j)\}_{j\in J}\) be a countable
  family with \(n_j,m_j\in\NN\) and
  \(\beta_j\in\CC\setminus\{0\}\). Choose roots
  \(\alpha_j^{n_j}=\beta_j\), and let
  \[
    \Lambda_{\mathcal D}
    :=
    \bigl\{\!\bigl\{\alpha_j\omega:
      j\in J,\ \omega^{n_j}=1,\text{ repeated }m_j\text{ times}
    \bigr\}\!\bigr\}
  \]
  be the prescribed non-zero spectral multiset, with multiplicities added
  when the same complex number occurs more than once.

  Suppose \(L\in\sone(H)\) is normal and its non-zero spectrum, counted
  with algebraic multiplicity, is exactly \(\Lambda_{\mathcal D}\). Then
  the trace-norm budget is forced:
  \[
    \|L\|_1
    =
    \sum_{\lambda\in\Lambda_{\mathcal D}}\abs{\lambda}
    =
    \sum_{j\in J}m_j n_j\abs{\beta_j}^{1/n_j}
    <\infty .
  \]
  Moreover
  \[
    \dim \mathsf S(L)=\sum_{j\in J}m_jn_j,
  \]
  as a Hilbert-space dimension. If \(J\) is infinite, then the algebraic
  range is a proper dense subspace of its closed support:
  \[
    \mathrm{ran}(L)\subsetneq \mathsf S(L)
    =\overline{\mathrm{ran}(L)} .
  \]
  Conversely, when the displayed budget is finite, the cyclic direct sum
  \(L_{\mathcal D}\) constructed in
  \Cref{thm:cyclic-fredholm-realisation} attains exactly this trace norm
  and this closed support.
\end{proposition}

\begin{proof}
  Since \(L\) is compact and normal, the spectral theorem diagonalizes it
  on the orthogonal sum of its non-zero eigenspaces and its kernel. For a
  normal trace-class operator, the trace norm is the sum of the moduli of
  the non-zero eigenvalues, counted with multiplicity. Therefore
  \[
    \|L\|_1
    =\sum_{\lambda\in\spec(L)\setminus\{0\}}
      \abs{\lambda}\,\dim\ker(L-\lambda I).
  \]
  By the assumed cyclic zero divisor, this is the sum over the multiset
  \(\Lambda_{\mathcal D}\). Each \(j\)-th cyclic factor contributes
  \(n_j\) eigenvalues of modulus
  \(\abs{\alpha_j}=\abs{\beta_j}^{1/n_j}\), each repeated \(m_j\) times.
  Hence
  \[
    \sum_{\lambda\in\Lambda_{\mathcal D}}\abs{\lambda}
    =\sum_{j\in J}m_j n_j\abs{\beta_j}^{1/n_j}.
  \]
  The left side is \(\|L\|_1<\infty\), so the budget is necessary.

  The closed support \(\mathsf S(L)\) is the Hilbert direct sum of the
  non-zero eigenspaces. Its dimension is therefore the total multiplicity
  of \(\Lambda_{\mathcal D}\), namely
  \[
    \sum_{\lambda\in\Lambda_{\mathcal D}}1
    =\sum_{j\in J}m_jn_j,
  \]
  with the same convention that multiplicities are counted. This remains
  true when different cyclic factors contribute the same eigenvalue,
  because the eigenspace dimension is the sum of the corresponding
  algebraic multiplicities.

  By \Cref{lem:normal-support-range-closure},
  \(\overline{\mathrm{ran}(L)}=\mathsf S(L)\). If \(J\) is infinite, then
  the multiset \(\Lambda_{\mathcal D}\) is infinite and the trace-class
  condition just proved forces its elements to have moduli tending to
  \(0\). Choose an orthonormal sequence of eigenvectors \(e_r\) with
  eigenvalues \(\lambda_r\ne0\) and \(\lambda_r\to0\). The closed range
  theorem would give a constant \(c>0\) with
  \(\norm{Lx}\ge c\norm{x}\) for all
  \(x\in(\ker L)^\perp=\mathsf S(L)\) if \(\mathrm{ran}(L)\) were closed.
  Applying this to \(e_r\) gives \(\abs{\lambda_r}\ge c\), a
  contradiction. Thus the algebraic range is not closed. Since its closure
  is \(\mathsf S(L)\), it is a proper dense subspace of that support.

  Finally assume
  \(\sum_jm_jn_j\abs{\beta_j}^{1/n_j}<\infty\). The block sum
  \(L_{\mathcal D}\) is trace-class by
  \Cref{lem:cyclic-product-trace-norm-limit}. Its blocks have exactly the
  multiset \(\Lambda_{\mathcal D}\), so the equalities above give both the
  asserted trace norm and the asserted closed support.
\end{proof}

\begin{theorem}[Positive-exponent cyclic-block determinant and closed-support optimality]
  \label{thm:cyclic-fredholm-realisation}
  \leanverified{paper_fredholm_witt_cyclic_block_characterization}
  Let \(\mathcal D=\{(n_j,\beta_j,m_j)\}_{j\in J}\) be a countable
  family with
  \[
    n_j,m_j\in\NN,\qquad
    \beta_j\in\CC\setminus\{0\},
  \]
  and assume that
  \[
    \sum_{j\in J}m_j n_j\abs{\beta_j}^{1/n_j}<\infty.
  \]
  Choose \(\alpha_j\in\CC\) with \(\alpha_j^{n_j}=\beta_j\), and set
  \[
    H_{\mathcal D}
    :=\bigoplus_{j\in J}\bigl(\CC^{n_j}\otimes\CC^{m_j}\bigr),
    \qquad
    L_{\mathcal D}
    :=\bigoplus_{j\in J}\bigl(C(n_j,\alpha_j)\otimes I_{m_j}\bigr).
  \]
  Then \(L_{\mathcal D}\in \sone(H_{\mathcal D})\). The entire function
  \[
    D_{\mathcal D}(z):=\det(I-zL_{\mathcal D})
  \]
  is the positive-exponent cyclic product in the convention above. More
  explicitly, for every finite \(F\subset J\), if
  \(L_F:=\bigoplus_{j\in F}(C(n_j,\alpha_j)\otimes I_{m_j})\), then
  \[
    \det(I-zL_F)
    =\prod_{j\in F}\prod_{\omega^{n_j}=1}
      (1-z\alpha_j\omega)^{m_j}
    =\prod_{j\in F}(1-\beta_j z^{n_j})^{m_j},
  \]
  with positive exponent \(m_j\). This is the determinant identity; its
  reciprocal is a different expression and is used only where the displayed
  determinant is non-zero. The infinite determinant \(D_{\mathcal D}\) is
  the locally uniform finite-truncation limit
  \[
    D_{\mathcal D}(z)
    =\prod_{j\in J}(1-\beta_j z^{n_j})^{m_j}
  \]
  on compact subsets of \(\CC\), where the right-hand side denotes the
  limit of the positive-exponent products over finite subsets of \(J\).
  On any connected zero-free neighbourhood \(\Omega\ni0\) for
  \(D_{\mathcal D}\), the reciprocal Euler product means
  \[
    D_{\mathcal D}(z)^{-1}
    =\prod_{j\in J}(1-\beta_j z^{n_j})^{-m_j}.
  \]
  Thus the determinant itself is never the negative-exponent product; the
  latter is only the reciprocal of this already constructed Fredholm
  determinant on a zero-free domain. Therefore realising the reciprocal
  Euler product is exactly the same assertion as realising the
  determinant \(D_{\mathcal D}\) on the determinant side.
  Since \(\sum_{j\in J}m_jn_j\abs{\beta_j}^{1/n_j}<\infty\) and
  \(m_jn_j\ge 1\), the sequence
  \(\bigl(\abs{\beta_j}^{1/n_j}\bigr)_{j\in J}\) is bounded. Writing
  \[
    r_{\mathcal D}:=
    \begin{cases}
      \sup_{j\in J}\abs{\beta_j}^{1/n_j}, & J\ne\varnothing,\\
      0, & J=\varnothing,
    \end{cases}
  \]
  this zero-free reciprocal identity holds in particular for
  \(\abs{z}<r_{\mathcal D}^{-1}\), with the convention
  \(r_{\mathcal D}^{-1}=+\infty\) if \(r_{\mathcal D}=0\).

  Moreover, for a normal trace-class operator \(L\in \sone(H)\), the
  following are equivalent:
  \begin{enumerate}[label=\textup{(\roman*)}]
    \item The Fredholm determinants agree as holomorphic germs at the
      origin:
      \[
        \det(I-zL)=D_{\mathcal D}(z).
      \]
      Equivalently, on some common zero-free neighbourhood of \(0\), their
      reciprocals agree and the right-hand side is the reciprocal Euler
      product just defined.
    \item There exist a Hilbert space \(H_0\) and a unitary operator
      \(U:H\to H_0\oplus H_{\mathcal D}\) such that
      \[
        ULU^{-1}=0_{H_0}\oplus L_{\mathcal D}.
      \]
  \end{enumerate}
  The unitary normal form in \textup{(ii)} is a statement about the
  reducing non-zero spectral subspace \((\ker L)^\perp\). It is not an
  identification of the algebraic range with \(H_{\mathcal D}\). Under
  this unitary,
  \[
    U\,\mathsf S(L)=U\,\overline{\mathrm{ran}(L)}
    =\{0\}\oplus H_{\mathcal D},
  \]
  while the algebraic range is a different subspace. If \(J\)
  is infinite, then the algebraic range of
  \(0_{H_0}\oplus L_{\mathcal D}\) is
  \(\{0\}\oplus\mathrm{ran}(L_{\mathcal D})\), which is a proper dense
  subspace of \(\{0\}\oplus H_{\mathcal D}\); the exact blockwise formula
  is given in \Cref{lem:cyclic-sum-range-closure}.
  In particular the determinant fixes the non-zero spectral multiset
  \[
    \bigl\{\!\bigl\{\alpha_j\omega:
      j\in J,\ \omega^{n_j}=1,\text{ repeated }m_j\text{ times}
    \bigr\}\!\bigr\},
  \]
  with multiplicities added when different indices give the same
  eigenvalue. Hence every trace-class normal realization of these Euler
  data has closed non-zero spectral support, in the sense of
  \Cref{def:closed-nonzero-support},
  satisfying
  \[
    \dim\mathsf S(L)=\sum_{j\in J}m_j n_j,
  \]
  counted as a Hilbert-space dimension. For a single factor
  \((1-\beta z^n)^{-m}\), this says that
  \(\dim\mathsf S(L)=mn\) for every normal realization of that factor. In
  the infinite direct-sum case this closed support is not the algebraic
  range; the algebraic range is a proper dense nonclosed
  subspace and is addressed separately in
  \Cref{lem:cyclic-sum-range-closure}. These counts are
  attached to the chosen Euler data \(\mathcal D\); distinct but
  equivalent factorizations of the same Euler product produce the same
  collapsed spectral multiset and therefore unitarily equivalent normal
  models. Equivalently, the determinant prescribes a sharp trace-class
  budget: in the normal category, any realization with this cyclic zero
  divisor must satisfy the budget and support conclusions of
  \Cref{prop:sharp-trace-budget-support-obstruction}.
\end{theorem}

\begin{proof}
  We keep the finite determinant calculation, the trace-norm limit, the
  reciprocal convention, and the support statement as separate steps. This
  prevents the reciprocal Euler product from being confused with the
  Fredholm determinant itself.

  \emph{Finite blocks and trace class.}
  Each block \(C(n_j,\alpha_j)\otimes I_{m_j}\) has singular value
  \(\abs{\alpha_j}\) with multiplicity \(m_j n_j\). Therefore
  \[
    \norm{L_{\mathcal D}}_1
    =\sum_{j\in J}m_j n_j\abs{\alpha_j}
    =\sum_{j\in J}m_j n_j\abs{\beta_j}^{1/n_j}
    <\infty,
  \]
  so \(L_{\mathcal D}\) is trace-class. The determinant is first computed
  on finite truncations, before any limiting argument or reciprocal is
  used. For a finite subset \(F\subset J\), put
  \(L_F:=\bigoplus_{j\in F}(C(n_j,\alpha_j)\otimes I_{m_j})\). The
  eigenvalues of the \(j\)-th block are \(\alpha_j\omega\),
  \(\omega^{n_j}=1\), each repeated \(m_j\) times. Hence
  \begin{align*}
    \det(I-zL_F)
    &=\prod_{j\in F}\prod_{\omega^{n_j}=1}
      (1-z\alpha_j\omega)^{m_j} \\
    &=\prod_{j\in F}(1-\alpha_j^{n_j}z^{n_j})^{m_j}
      =\prod_{j\in F}(1-\beta_j z^{n_j})^{m_j}.
  \end{align*}
  Thus finite truncations have positive exponents in the determinant. The
  negative-exponent finite product is only the reciprocal of this determinant
  away from its zero set.

  \emph{Trace-norm limit and reciprocal domain.}
  Passing to an increasing finite exhaustion of \(J\),
  \Cref{lem:cyclic-product-trace-norm-limit} gives the locally uniform
  Fredholm-determinant limit
  \[
    D_{\mathcal D}(z)=\det(I-zL_{\mathcal D})
    =\prod_{j\in J}(1-\beta_j z^{n_j})^{m_j}
  \]
  as the trace-norm limit of the finite products with positive exponents.
  On a connected zero-free neighbourhood for \(D_{\mathcal D}\), its
  reciprocal is holomorphic and the finite reciprocal products converge to
  \(D_{\mathcal D}^{-1}\), again by
  \Cref{lem:cyclic-product-trace-norm-limit}. The block diagonal norm
  satisfies \(\norm{L_{\mathcal D}}=r_{\mathcal D}\) by the definition of
  \(r_{\mathcal D}\) and the fact that each block has operator norm
  \(\abs{\alpha_j}=\abs{\beta_j}^{1/n_j}\). If \(r_{\mathcal D}=0\), then
  \(J=\varnothing\), \(L_{\mathcal D}=0\), and \(D_{\mathcal D}\equiv1\).
  Otherwise, on the disc \(\abs{z}<r_{\mathcal D}^{-1}\), one has
  \(\norm{zL_{\mathcal D}}<1\), so \(I-zL_{\mathcal D}\) is invertible and
  the determinant is non-zero. Taking reciprocals on this disc gives
  \[
    D_{\mathcal D}(z)^{-1}
    =\prod_{j\in J}(1-\beta_j z^{n_j})^{-m_j}.
  \]

  \emph{Converse normal form.}
  The implication \textup{(ii)}\(\Rightarrow\)\textup{(i)} is immediate
  from the positive-exponent direct-sum formula above and the fact that
  zero summands do not change the Fredholm determinant. Conversely,
  assume \textup{(i)}. Then
  \[
    \det(I-zL)=\det(I-zL_{\mathcal D})
  \]
  as holomorphic germs at the origin. Both sides are entire, hence
  identical on \(\CC\). By \cite[Theorem~3.7]{Simon2005TraceIdeals},
  the zeros of the Fredholm determinant are precisely the reciprocals
  of the non-zero eigenvalues, counted with algebraic multiplicity.
  Therefore \(L\) and \(L_{\mathcal D}\) have the same non-zero
  spectrum with multiplicity. Explicitly, the non-zero eigenvalues of
  \(L_{\mathcal D}\) are
  \[
    \alpha_j\omega,\qquad \omega^{n_j}=1,
  \]
  each repeated \(m_j\) times, with multiplicities added if the same
  complex number occurs for more than one pair \((j,\omega)\). The
  trace-class summability condition rules out infinite multiplicity at
  any non-zero eigenvalue. Thus the determinant determines this
  collapsed spectral multiset; it does not carry any additional data
  about how equal eigenvalues were grouped in a particular Euler
  factorization.

  The operator \(L_{\mathcal D}\) is normal. Since \(L\) is normal, the
  spectral theorem gives an orthogonal decomposition
  \[
    H=\ker L\oplus\bigoplus_{\lambda\in\spec(L)\setminus\{0\}}E_\lambda,
  \]
  where \(L|_{E_\lambda}=\lambda I\). If the same eigenvalue \(\lambda\)
  is contributed by several indices \(j\), its eigenspace \(E_\lambda\)
  has dimension equal to the sum of the corresponding multiplicities.
  Choose an orthogonal decomposition of each \(E_\lambda\) into pieces
  with these multiplicities and then regroup those pieces by the index
  \(j\). This produces reducing subspaces \(H_j\subset H\) such that
  \(L|_{H_j}\) is unitarily equivalent to
  \[
    \bigoplus_{\omega^{n_j}=1}\alpha_j\omega\,I_{m_j}.
  \]
  The discrete Fourier transform on \(\CC^{n_j}\) diagonalizes
  \(\Pi_{n_j}\), so
  \[
    \bigoplus_{\omega^{n_j}=1}\alpha_j\omega\,I_{m_j}
    \cong C(n_j,\alpha_j)\otimes I_{m_j}.
  \]
  The remaining part of \(H\) is \(\ker L\). Hence \(L\) is unitarily
  equivalent to \(0_{\ker L}\oplus L_{\mathcal D}\), proving
  \textup{(ii)}.

  \emph{Closed support.}
  The support and minimality assertions are then immediate from this
  unitary normal form, but they concern closed spectral support rather than
  algebraic range. The closed non-zero spectral support of
  \(L_{\mathcal D}\) is all of
  \[
    H_{\mathcal D}
    =\bigoplus_{j\in J}\bigl(\CC^{n_j}\otimes\CC^{m_j}\bigr),
  \]
  because every summand is a non-zero spectral subspace after Fourier
  diagonalisation of the cyclic block. Therefore any normal realization
  with determinant \(D_{\mathcal D}\) has
  \[
    \dim\mathsf S(L)=\dim H_{\mathcal D}
    =\sum_{j\in J}m_jn_j.
  \]
  By \Cref{lem:normal-support-range-closure}, this support is
  \((\ker L)^\perp=\overline{\mathrm{ran}(L)}\). No surjectivity of the
  algebraic range is used here; when \(J\) is infinite, the exact
  algebraic range is the dense proper subspace described in
  \Cref{lem:cyclic-sum-range-closure}.
\end{proof}

\begin{definition}[Block-cyclic normal realization]
  \label{def:block-cyclic-normal-realisation}
  A normal operator \(L\in \sone(H)\) satisfying
  \Cref{thm:cyclic-fredholm-realisation}\textup{(ii)} is called a
  block-cyclic normal realization of the Euler data
  \(\mathcal D=\{(n_j,\beta_j,m_j)\}_{j\in J}\). This definition records
  only the unitary normal form of the non-zero spectral part. It includes no
  assertion that the algebraic range is the whole cyclic-block space; that
  range is the separate object computed in
  \Cref{lem:cyclic-sum-range-closure}.
\end{definition}

\begin{lemma}[Algebraic range and closed support of a cyclic direct sum]
  \label{lem:cyclic-sum-range-closure}
  In the notation of \Cref{thm:cyclic-fredholm-realisation}, write
  \[
    H_j:=\CC^{n_j}\otimes\CC^{m_j},
    \qquad
    L_j:=C(n_j,\alpha_j)\otimes I_{m_j}.
  \]
  Then
  \[
    \mathrm{ran}(L_{\mathcal D})
    =
    \left\{
      y=\bigoplus_{j\in J}y_j\in H_{\mathcal D}:
      \sum_{j\in J}\norm{L_j^{-1}y_j}^{2}<\infty
    \right\}.
  \]
  In particular the finite-support algebraic direct sum is contained in
  \(\mathrm{ran}(L_{\mathcal D})\), and
  \[
    \overline{\mathrm{ran}(L_{\mathcal D})}=H_{\mathcal D}.
  \]
  If \(J\) is infinite under the trace-class summability hypothesis, then
  \[
    \mathrm{ran}(L_{\mathcal D})
    \subsetneq H_{\mathcal D}
    =\overline{\mathrm{ran}(L_{\mathcal D})}
    =\mathsf S(L_{\mathcal D}).
  \]
  Thus the canonical support invariant is the closure of the range,
  equivalently the closed span of the non-zero spectral subspaces, and not
  the algebraic range itself.
\end{lemma}

\begin{proof}
  Each \(L_j\) is invertible because \(\alpha_j^{n_j}=\beta_j\ne0\), and
  \(L_j=\alpha_j(\Pi_{n_j}\otimes I_{m_j})\) is a scalar multiple of a
  unitary. Hence a vector \(y=\bigoplus_j y_j\) belongs to
  \(\mathrm{ran}(L_{\mathcal D})\) exactly when
  \(x_j:=L_j^{-1}y_j\) defines a square-summable vector
  \(x=\bigoplus_jx_j\in H_{\mathcal D}\). This proves the displayed
  characterization of the algebraic range.

  If \(y\) has finite support, then \(L_j^{-1}y_j\) has finite support as
  well, so the finite-support algebraic direct sum lies in the range. That
  algebraic direct sum is dense in \(H_{\mathcal D}\), giving
  \(\overline{\mathrm{ran}(L_{\mathcal D})}=H_{\mathcal D}\).

  Finally assume \(J\) is infinite and the trace-class summability condition
  holds. Since \(m_jn_j\ge1\), the sequence \(\abs{\alpha_j}\) tends to
  \(0\), and
  \[
    \sum_j\abs{\alpha_j}
    \le\sum_jm_jn_j\abs{\alpha_j}<\infty .
  \]
  Choose unit vectors \(e_j\in H_j\) and put
  \(y=\bigoplus_j\alpha_j e_j\). Since
  \(\sum_j\abs{\alpha_j}<\infty\), also
  \(\sum_j\abs{\alpha_j}^2<\infty\); hence this vector belongs to
  \(H_{\mathcal D}\). However, the
  unique blockwise preimage has \(j\)-th component
  \[
    (\Pi_{n_j}^{-1}\otimes I_{m_j})e_j,
  \]
  of norm \(1\), for every \(j\).
  This preimage is not square-summable. Thus \(y\notin
  \mathrm{ran}(L_{\mathcal D})\), even though \(y\in
  \overline{\mathrm{ran}(L_{\mathcal D})}\).
\end{proof}

\noindent\emph{Range caveat.}
The preceding lemma is the precise range statement used in the support
claims. In an infinite cyclic direct sum the algebraic range is the dense
proper subspace described in \Cref{lem:cyclic-sum-range-closure}; only its
closure is the support invariant \(\mathsf S(L_{\mathcal D})\).

\begin{corollary}[Closed support, exact range, and trace norm]
  \label{thm:block-cyclic-rigidity}
  \leanverified{paper_fredholm_witt_block_cyclic_rigidity}
  Let \(L\) be a block-cyclic normal realization of \(\mathcal D\) in the
  sense of \Cref{def:block-cyclic-normal-realisation}. Let
  \(U:H\to H_0\oplus H_{\mathcal D}\) be any unitary representation with
  \(ULU^{-1}=0_{H_0}\oplus L_{\mathcal D}\), and write
  \(L_j=C(n_j,\alpha_j)\otimes I_{m_j}\). This is the exact range
  statement for the model. The closed support
  \[
    \mathsf S(L)=
    \overline{\mathrm{span}}
    \{\,\ker(L-\lambda I):\lambda\in\spec(L)\setminus\{0\}\,\},
  \]
  and the algebraic range are different objects. They are given by
  \[
    U\,\mathsf S(L)
    =
    U\,\overline{\mathrm{ran}(L)}
    =
    \overline{U\,\mathrm{ran}(L)}
    =
    \{0\}\oplus H_{\mathcal D},
  \]
  while
  \[
    U\,\mathrm{ran}(L)
    =
    \{0\}\oplus
    \left\{
      y=\bigoplus_{j\in J}y_j\in H_{\mathcal D}:
      \sum_{j\in J}\norm{L_j^{-1}y_j}^{2}<\infty
    \right\}.
  \]
  Therefore the determinant fixes the closed non-zero spectral support,
  equivalently the closure of the range, not the algebraic range itself.
  The block-cyclic normal form is thus a classification of the reducing
  non-zero spectral subspace. It is not a claim that
  \(\mathrm{ran}(L)\) is unitarily equivalent to
  \(H_{\mathcal D}\) in infinite rank.
  In particular,
  \[
    \dim\mathsf S(L)=\sum_{j\in J}m_j n_j,
    \qquad
    \|L\|_1
    =\sum_{j\in J}m_j n_j\abs{\beta_j}^{1/n_j}.
  \]
  If \(J\) is finite, then only finitely many block inverses occur and
  \[
    U\,\mathrm{ran}(L)=\{0\}\oplus H_{\mathcal D}.
  \]
  If \(J\) is infinite, then the trace-class summability condition forces
  \(\abs{\alpha_j}\to0\), and
  \[
    U\,\mathrm{ran}(L)
    \subsetneq \{0\}\oplus H_{\mathcal D}
    =U\,\overline{\mathrm{ran}(L)}
    =U\,\mathsf S(L).
  \]
  Thus the equality
  \(U\,\mathrm{ran}(L)=\{0\}\oplus H_{\mathcal D}\) holds precisely in the
  finite-block case; in the infinite trace-class case the algebraic range
  is proper dense and nonclosed.
  Consequently the summability condition in
  \Cref{thm:cyclic-fredholm-realisation} is both necessary and
  sufficient for trace-class normal realizability of the cyclic-block
  data, and for a single factor \((1-\beta z^n)^{-m}\) no normal
  realization can have closed span of non-zero spectral subspaces of
  dimension less than \(mn\).
\end{corollary}

\begin{proof}
  Choose a unitary \(U:H\to H_0\oplus H_{\mathcal D}\) with
  \(ULU^{-1}=0_{H_0}\oplus L_{\mathcal D}\). The zero summand contributes
  only the zero vector to the algebraic range, so
  \[
    U\,\mathrm{ran}(L)
    =\{0\}\oplus\mathrm{ran}(L_{\mathcal D}).
  \]
  Applying \Cref{lem:cyclic-sum-range-closure} gives exactly the blockwise
  range displayed in the statement. If \(J\) is infinite, the same lemma
  shows that this algebraic range is dense and proper in
  \(H_{\mathcal D}\), hence nonclosed. If \(J\) is finite, each block is
  invertible and the displayed range equals \(H_{\mathcal D}\).

  The non-zero spectral subspaces of \(L_{\mathcal D}\) span
  \(H_{\mathcal D}\). Since \(L_{\mathcal D}\) is normal compact,
  \(\mathsf S(L_{\mathcal D})
  =(\ker L_{\mathcal D})^\perp
  =\overline{\mathrm{ran}(L_{\mathcal D})}\).
  By \Cref{lem:cyclic-sum-range-closure},
  \[
    \overline{\mathrm{ran}(L_{\mathcal D})}=H_{\mathcal D},
    \qquad
    U\,\mathsf S(L)=\{0\}\oplus H_{\mathcal D}.
  \]
  Thus the invariant used here is explicitly the closed support of the
  non-zero spectrum. In the infinite case, the algebraic range is dense in
  \(H_{\mathcal D}\) but not equal to \(H_{\mathcal D}\). Equivalently,
  \[
    U\,\mathrm{ran}(L)
    \subsetneq U\,\overline{\mathrm{ran}(L)}
    =U\,\mathsf S(L)
  \]
  whenever \(J\) is infinite. If \(J\) is finite, the condition defining
  \(\mathrm{ran}(L_{\mathcal D})\) contains only finitely many inverse
  norms and is automatic for every vector of \(H_{\mathcal D}\). Hence
  only in that finite-block case does the algebraic range equal
  \(\{0\}\oplus H_{\mathcal D}\) after applying \(U\).
  Hence equality of the algebraic range with the cyclic-block support space
  occurs precisely in the finite-block case. No closed-range assertion is
  being used in the infinite case.
  Consequently
  \[
    \dim\mathsf S(L)
    =\sum_{j\in J}m_j n_j.
  \]
  The singular values of \(C(n_j,\alpha_j)\otimes I_{m_j}\) are all
  equal to \(\abs{\alpha_j}=\abs{\beta_j}^{1/n_j}\), each with
  multiplicity \(m_j n_j\). Therefore
  \[
    \|L\|_1
    =\|L_{\mathcal D}\|_1
    =\sum_{j\in J}m_j n_j\abs{\beta_j}^{1/n_j}.
  \]
  The final statement is the single-factor specialization.
\end{proof}

\begin{remark}[Why the range is not the support]
  \label{rem:range-not-support}
  The unitary normal form
  \(ULU^{-1}=0_{H_0}\oplus L_{\mathcal D}\) identifies the closed
  non-zero spectral support, not the algebraic range. In the infinite case
  the vector constructed in \Cref{lem:cyclic-sum-range-closure},
  \(y=\bigoplus_j\alpha_j e_j\), lies in \(H_{\mathcal D}\) but has no
  square-summable blockwise preimage under \(L_{\mathcal D}\). Thus
  \(y\in\mathsf S(L_{\mathcal D})\setminus\mathrm{ran}(L_{\mathcal D})\).
  This is the precise obstruction to replacing
  \(\overline{\mathrm{ran}(L)}\) by \(\mathrm{ran}(L)\) in infinite rank.
\end{remark}

\begin{remark}[Sharpness of the summability condition]
  \label{rem:summability-sharpness}
  \Cref{prop:sharp-trace-budget-support-obstruction},
  \Cref{thm:cyclic-fredholm-realisation}, and
  \Cref{thm:block-cyclic-rigidity} show that
  \[
    \sum_{j\in J}m_j n_j\abs{\beta_j}^{1/n_j}<\infty
  \]
  is exactly the criterion for trace-class normal realizability of the
  prescribed Euler data, and that the resulting closed support is forced.
  Thus, within the normal category, the summability hypothesis is not
  merely sufficient but necessary.
\end{remark}

\begin{remark}[Choice of roots]\label{rem:choice-of-roots}
  Replacing \(\alpha_j\) by another \(n_j\)th root of \(\beta_j\)
  merely rotates the same regular \(n_j\)-gon of eigenvalues. Hence the
  corresponding blocks \(C(n_j,\alpha_j)\otimes I_{m_j}\) are unitarily
  equivalent. The determinant data determine the block only up to this
  harmless relabelling and the addition of zero summands.
\end{remark}

\begin{corollary}[Finite Euler products give finite-rank determinant models]
  \label{cor:finite-rank-realisation}
  \leanverified{paper_fredholm_witt_finite_rank_realisation}
  If \(J\) is finite in \Cref{thm:cyclic-fredholm-realisation}, then
  \(L_{\mathcal D}\) is finite-rank. As an immediate specialization,
  every finite reciprocal Euler product
  \[
    \prod_{j=1}^N(1-\beta_j z^{n_j})^{-m_j}
  \]
  is realized in the determinant-side sense: there is a finite-rank normal
  operator \(L\) with
  \[
    \det(I-zL)=\prod_{j=1}^N(1-\beta_j z^{n_j})^{m_j},
  \]
  and the displayed reciprocal product is \(\det(I-zL)^{-1}\) wherever
  this determinant is non-zero. The non-zero rank of \(L\) is exactly
  \(\sum_{j=1}^N m_j n_j\).
\end{corollary}

\subsection{Spectral rigidity}

\begin{proposition}[Spectral rigidity of Fredholm determinants]
  \label{thm:fredholm-spectral-rigidity}
  Let \(T_1,T_2\in \sone\). If
  \[
    \det(I-zT_1)=\det(I-zT_2)
  \]
  in a neighbourhood of \(z=0\), then the non-zero spectra of \(T_1\)
  and \(T_2\) coincide with algebraic multiplicity. Consequently,
  \[
    \Tr(T_1^n)=\Tr(T_2^n),
    \qquad
    p_n(T_1)=p_n(T_2)
    \qquad (n\ge 1).
  \]
\end{proposition}

\begin{proof}
  For any \(T\in\sone(H)\), the Fredholm determinant
  \(z\mapsto\det(I-zT)\) is an entire function of order at most~\(1\)
  \cite[Theorem~3.7]{Simon2005TraceIdeals}. Since the two Fredholm
  determinants agree as holomorphic germs at the origin, the identity
  theorem extends the equality to all of~\(\CC\). By the canonical
  product representation \cite[Theorem~3.7]{Simon2005TraceIdeals}, the
  zeros of \(\det(I-zT)\) are precisely the reciprocals
  \(\lambda_k^{-1}\) of the non-zero eigenvalues of~\(T\), each
  occurring with its algebraic multiplicity. Hence the non-zero spectra
  coincide.

  Equality of determinants near \(0\) also gives equality of the Taylor
  expansions
  \[
    \log\det(I-zT_i)^{-1}
    =\sum_{n\ge 1}\frac{\Tr(T_i^n)}{n}z^n
    \qquad (i=1,2),
  \]
  so the trace sequences agree coefficientwise. The primitive
  coordinates \(p_n(T_i)\) are obtained from the trace sequences by
  M\"obius inversion.
\end{proof}

\begin{corollary}[Realisability versus rigidity]
  \label{cor:realisability-vs-rigidity}
  \leanverified{paper_fredholm_witt_realisability_vs_rigidity}
  Let \(A\) be a finite matrix and let \(T\in \sone\) satisfy
  \[
    \det(I-zT)=\det(I-zA).
  \]
  Then \(T\) and \(A\) have the same non-zero eigenvalues with
  algebraic multiplicity. Thus cyclic-block constructions may enlarge
  the ambient operator category, but they do not create spectral
  freedom once the determinant is prescribed.
\end{corollary}
