\section{Folding projections and multiscale commuting structure}
\label{sec:multiscale}

\subsection{A rewriting realization and categorical properties}
The map \(\Fold_m\) can be computed by a finite-time local rewriting system realizing Zeckendorf normalization. Two indexing conventions appear in the literature and both are used in this paper; we fix them precisely here.

\begin{lemma}[Convention translation]\label{lem:convention}
Let \(\omega=(\omega_1,\dots,\omega_m)\in\Om_m\) be a microstate word written in the \emph{low-to-high} (LH) convention of \eqref{eq:Nomega}, so that position \(k\) carries weight \(\Fib_{k+1}\). Let \(\omega^\sharp=(\omega_m,\dots,\omega_1)\) be the same word written in the \emph{most-significant-digit-first} (MSD-first) convention. Then:
\begin{enumerate}
\item[(i)] The local MSD-first rewrite rule \(011\to 100\) applied at positions \((k,k+1,k+2)\) (counting from the left of \(\omega^\sharp\)) corresponds, in the LH convention, to the rule \(110\to 001\) applied at positions \((m-k-1,m-k,m-k+1)\).
\item[(ii)] After a leading zero is prepended in MSD-first convention (equivalently, \(\omega_m=0\) is available as a carry slot), scanning the MSD-first word left-to-right and applying \(011\to 100\) until no occurrence remains is equivalent to scanning the LH word right-to-left and applying \(110\to 001\) until no occurrence remains.
\item[(iii)] In either convention, each application of the rule preserves the Fibonacci value \(N(\omega)=\sum_k\omega_k\Fib_{k+1}\) and strictly decreases the total number of \(1\)s. The unique fixed point is the Zeckendorf representation of \(N(\omega)\).
\end{enumerate}
\end{lemma}

\begin{proof}
Part (i) is immediate: reversing indices maps \((a,b,c)\mapsto (c,b,a)\), so \(011\to 100\) becomes \(110\to 001\). Part (ii) follows because left-to-right scanning in MSD-first order corresponds to scanning from the highest-weight position to the lowest in LH order, i.e., right-to-left. Part (iii): the identity \(\Fib_{k}+\Fib_{k+1}=\Fib_{k+2}\) shows that replacing \(011\) at MSD positions \((k,k+1,k+2)\) by \(100\) replaces \(\Fib_{m-k+1}+\Fib_{m-k}\) with \(\Fib_{m-k+2}\) and does not alter the total value; the count of \(1\)s decreases by \(1\). Termination and uniqueness of the normal form are proved in Appendix~\ref{app:rewriting}.
\end{proof}

One convenient formulation uses MSD-first convention with the rule
\[
011 \;\to\; 100,
\]
scanning left-to-right after prefixing a leading zero; by Lemma~\ref{lem:convention}(i)--(ii), in the window convention of \eqref{eq:Nomega} this is the equivalent rule \(110\to 001\) acting from high to low Fibonacci weights. See \cite{Frougny1992,Frougny1999,ShallitShan2023} for automata-theoretic formulations. Termination is immediate since each rewrite strictly decreases the total number of \(1\)s. Confluence holds and yields a unique normal form, which is precisely the Zeckendorf representation.

In our window model, this implies:
\begin{proposition}[Well-definedness, idempotence, surjectivity]\label{prop:idempotent}
For each \(m\ge 1\), the fold \(\Fold_m:\Om_m\to X_m\) is well-defined, idempotent, and surjective:
\[
\Fold_m\circ \Fold_m=\Fold_m,
\qquad
\Fold_m(\Om_m)=X_m.
\]
\end{proposition}

\begin{proof}
Appendix~\ref{app:rewriting} proves that the rewriting system terminates and is confluent, so every \(\omega\in\Om_m\) has a unique Zeckendorf normal form with the same Fibonacci value. This is exactly \(\Fold_m(\omega)\), hence \(\Fold_m\) is well-defined.

If \(x\in X_m\), then \(x\) already has no adjacent \(1\)s and therefore is itself the Zeckendorf expansion of \(N(x)\). Hence \(\Fold_m(x)=x\). Applying this to \(x=\Fold_m(\omega)\in X_m\) gives
\[
\Fold_m(\Fold_m(\omega))=\Fold_m(\omega),
\]
so \(\Fold_m\) is idempotent. The same fixed-point property implies surjectivity: every \(x\in X_m\) lies in the image because \(\Fold_m(x)=x\).
\end{proof}

\subsection{Stable restriction and the inverse limit}
For \(m_2\ge m_1\), define the stable truncation
\[
\proj_{m_2\to m_1}: X_{m_2}\to X_{m_1},
\qquad
\proj_{m_2\to m_1}(x_1\cdots x_{m_2}) := x_1\cdots x_{m_1}.
\]
Then \(\bigl(X_m,\proj_{m_2\to m_1}\bigr)\) is an inverse system, and the inverse limit
\[
X_\infty := \varprojlim_m X_m
\]
is naturally identified with the one-sided golden-mean shift (equivalently, the one-sided subshift forbidding \(11\)); see \cite[Chapter~1]{LindMarcus1995} or \cite[Chapter~1]{Kitchens1998}.

\subsection{The fold-aware restriction operator}
Let \(\naive_{m+1\to m}:\Om_{m+1}\to \Om_m\) be the naive truncation (drop the last coordinate),
\[
\naive_{m+1\to m}(\omega_1\cdots \omega_{m+1}) := \omega_1\cdots \omega_m.
\]
Multiscale consistency suggests enforcing the commuting relation
\begin{equation}\label{eq:commuting}
\Fold_m\circ \aware_{m+1\to m}
\;=\;
\proj_{m+1\to m}\circ \Fold_{m+1},
\end{equation}
so that ``restrict-then-fold'' agrees with ``fold-then-restrict'' at the stable level.

\begin{definition}[Fold-aware restriction]\label{def:fold-aware}
Define the fold-aware restriction \(\aware_{m+1\to m}:\Om_{m+1}\to \Om_m\) by the explicit formula
\begin{equation}\label{eq:re-def}
\aware_{m+1\to m}(\omega)
:= \proj_{m+1\to m}\!\bigl(\Fold_{m+1}(\omega)\bigr)\in X_m\subset \Om_m.
\end{equation}
\end{definition}

\begin{lemma}[Strict commutation]\label{lem:strict-commute}
The operator \(\aware_{m+1\to m}\) satisfies \eqref{eq:commuting}. Moreover, it is the unique map \(\Om_{m+1}\to X_m\) satisfying \eqref{eq:commuting}.
\end{lemma}

\begin{proof}
Since \(\Fold_{m+1}(\omega)\in X_{m+1}\), its truncation \(\proj_{m+1\to m}(\Fold_{m+1}(\omega))\) lies in \(X_m\), hence is fixed by \(\Fold_m\). Therefore
\[
\Fold_m(\aware_{m+1\to m}(\omega))
=
\Fold_m\!\left(\proj_{m+1\to m}(\Fold_{m+1}(\omega))\right)
=
\proj_{m+1\to m}(\Fold_{m+1}(\omega)),
\]
which is \eqref{eq:commuting}. For uniqueness among maps into \(X_m\), note that \(\Fold_m\) restricts to the identity on \(X_m\), so \eqref{eq:commuting} forces \(\aware_{m+1\to m}(\omega)=\proj_{m+1\to m}(\Fold_{m+1}(\omega))\).
\end{proof}

