\section{Restriction discrepancy: worst cases and bulk statistics}
\label{sec:gauge}

\subsection{Definition}
We now compare the naive restriction with the compatible restriction from Section~\ref{sec:multiscale}. For \(\omega\in \Om_{m+1}\), define the \emph{restriction discrepancy}, or \emph{gauge anomaly}, by
\begin{equation}\label{eq:gauge-anomaly}
G_m(\omega)
:=
\Hamming{\naive_{m+1\to m}(\omega)\oplus \aware_{m+1\to m}(\omega)}.
\end{equation}
Equivalently, if
\[
\Delta^{(m)}(\omega)
:=
\naive_{m+1\to m}(\omega)\oplus \aware_{m+1\to m}(\omega)\in \set{0,1}^m,
\]
then \(G_m(\omega)=\sum_{k=1}^m \Delta^{(m)}_k(\omega)\).
Thus \(G_m\) measures exactly how far naive truncation fails to agree with the restriction dictated by normalization.

The logic of this section is as follows. Propositions~\ref{prop:worst} and \ref{prop:tail-patch} show that the obstruction is a multiscale effect and not a bounded boundary effect. The structural step is Theorem~\ref{thm:pair-language}, which identifies the bulk pair process with an explicit four-state synchronized presentation. The exact laws, limit theorems, and spectral formulas that follow are consequences of this identification.

\subsection{Worst-case discrepancy and the absence of bounded repairs}
\begin{proposition}[Extensive worst case]\label{prop:worst}
For every \(m\ge 1\) there exists an explicit microstate \(\omega^\star\in \Om_{m+1}\) such that
\[
G_m(\omega^\star)=m.
\]
\end{proposition}

\begin{remark}[An explicit worst-case family]\label{rem:explicit-worst}
Let \(L=m+1\) and write \(L=3n+r\) with \(r\in\{0,1,2\}\). Define a length-\(L\) word \(\omega^\star\) by
\[
\omega^\star=
\begin{cases}
(110)^n, & r=0,\\
(110)^n0, & r=1,\\
(110)^n11, & r=2,
\end{cases}
\]
where concatenation is in the window order \((\omega_1,\dots,\omega_{m+1})\).
\end{remark}

\begin{proof}
The identity \(\Fib_{k+1}+\Fib_{k+2}=\Fib_{k+3}\) shows that each block \(110\) may be replaced by \(001\) without changing Fibonacci value. Repeating this blockwise gives
\[
\Fold_L\bigl((110)^n\bigr)=(001)^n,
\qquad
\Fold_L\bigl((110)^n0\bigr)=(001)^n0.
\]
Indeed, the words on the right are already in \(X_L\), so uniqueness of the Zeckendorf normal form identifies them as the folds.
In the remaining case \(L=3n+2\), so the terminal block \(11\) occupies the LH positions \(L-1\) and \(L\) and contributes
\[
\Fib_L+\Fib_{L+1}=\Fib_{L+2},
\]
which is exactly the weight of a single \(1\) at the new position \(L+1\). Hence
\[
N\bigl((110)^n11\bigr)=N\bigl((001)^n00\bigr)+\Fib_{L+2},
\]
and the word \((001)^n00\in X_L\) gives the first \(L\) digits of the Zeckendorf normalization. Therefore the first \(L\) digits of \(\Fold_L((110)^n11)\) are \((001)^n00\).

Now delete the last coordinate. When \(r=0\), the naive truncation is \((110)^{n-1}11\) and the fold-aware truncation is \((001)^{n-1}00\); when \(r=1\), they are \((110)^n\) and \((001)^n\); when \(r=2\), they are \((110)^n1\) and \((001)^n0\). In each case the two length-\(m\) words are coordinatewise complementary. Therefore every coordinate contributes \(1\) to the Hamming distance, so \(G_m(\omega^\star)=m\).
\end{proof}

\begin{proposition}[No bounded tail patch]\label{prop:tail-patch}
Fix \(\ell\ge 1\). There is no family of maps
\[
P_{m,\ell}:\set{0,1}^{\ell+1}\to \set{0,1}^{\ell}
\]
such that, for all sufficiently large \(m\) and all \(\omega\in \Om_{m+1}\),
\begin{equation}\label{eq:tail-patch-form}
\aware_{m+1\to m}(\omega)
=
\omega_1\cdots \omega_{m-\ell}\,
P_{m,\ell}(\omega_{m-\ell+1}\cdots \omega_{m+1}).
\end{equation}
In particular, the anomaly cannot be uniformly removed by modifying only a bounded tail of the naive truncation.
\end{proposition}

\begin{proof}
Assume that \eqref{eq:tail-patch-form} holds for some fixed \(\ell\) and all sufficiently large \(m\). Choose \(m>\ell\) in that range and let \(\omega^\star\) be the word from Proposition~\ref{prop:worst}. Since \(G_m(\omega^\star)=m\), the words \(\naive_{m+1\to m}(\omega^\star)\) and \(\aware_{m+1\to m}(\omega^\star)\) differ at every coordinate. In particular, their first \(m-\ell\) coordinates differ.

But the right-hand side of \eqref{eq:tail-patch-form} leaves those first \(m-\ell\) coordinates unchanged: it agrees with \(\omega^\star_1\cdots \omega^\star_{m-\ell}\) there and only modifies the last \(\ell\) positions. This contradicts the previous paragraph. Hence no such bounded tail patch exists.
\end{proof}

\begin{example}[A worst-case word at \(m=4\)]\label{ex:worst4}
For \(m=4\), the explicit family of Remark~\ref{rem:explicit-worst} gives \(\omega^\star=11011\in\Om_5\). Its Fibonacci value is
\[
N(\omega^\star)=\Fib_2+\Fib_3+\Fib_5+\Fib_6=1+2+5+8=16,
\]
so \(Z(16)=\Fib_7+\Fib_4\) with \(\Fib_7=13\) at position \(6\) (outside the window) and \(\Fib_4=3\) at position \(3\), giving
\[
\Fold_5(\omega^\star)=00100,
\qquad
\aware_{5\to4}(\omega^\star)=0010,
\qquad
\naive_{5\to4}(\omega^\star)=1101.
\]
Thus \(G_4(\omega^\star)=4\), with disagreement at every site.
\end{example}

\subsection{The bulk pair process and its one-site law}
\label{subsec:joint-law}
To formulate the limit laws, we pass to the stationary bulk model. Let \(\eta=(\eta_t)_{t\in\ZZ}\) be a bi-infinite Bernoulli-\(\frac12\) sequence, and let \((X_t,Y_t)_{t\in\ZZ}\) be the sitewise microstate/normalized-digit process encoded by the reversed pair shift of Appendix~\ref{app:pressure}. Thus \(X_t=\eta_t\), while \(Y_t\) is the normalized digit seen at site \(t\) in the bulk limit. By construction the pair process is stationary; we refer to its law as the \emph{bulk law}.

The next theorem identifies the finite-state structure behind this bulk process. Once it is established, the later statements of this section are obtained from standard finite-state arguments applied to a completely explicit presentation.

\begin{theorem}[Interior language theorem for reversed pair factors]\label{thm:pair-language}
Let \(G_{\mathrm{pair}}\) be the labeled graph with state set \(\mathcal{S}=\{00,01,10,11\}\) and edges
\[
00 \xrightarrow{00} 00,\quad
00 \xrightarrow{01} 01,\quad
00 \xrightarrow{11} 11,\quad
01 \xrightarrow{10} 10,
\]
\[
10 \xrightarrow{10} 00,\quad
10 \xrightarrow{00} 01,\quad
10 \xrightarrow{11} 01,\quad
11 \xrightarrow{00} 00.
\]
For a finite word \(w\in(\{00,01,10,11\})^\ell\), the following are equivalent.
\begin{enumerate}
\item[(i)] \(w\) occurs as an interior factor of some reversed pair word \(\mathcal{R}_M(\omega)\).
\item[(ii)] \(w\) is the label of a path in \(G_{\mathrm{pair}}\).
\end{enumerate}
Moreover, every path label in \(G_{\mathrm{pair}}\) is realized at arbitrarily large distance from the boundary by a reversed pair word \(\mathcal{R}_M(\omega)\).
\end{theorem}

\begin{proof}
Suppose first that \(w\) occurs as an interior factor of \(\mathcal{R}_M(\omega)\). By Lemma~\ref{lem:zero-padding}, we may prepend arbitrarily many high-weight zeros and therefore assume that the factor is preceded by a synchronized block \(00|00\). Appendix~\ref{app:pressure} classifies all synchronized length-\(3\) and length-\(4\) factors. The length-\(3\) classification shows that once the synchronized regime has been entered, the future depends only on the current suffix state \(s\in\mathcal{S}\), namely on the last pair symbol in
\[
00|00,\qquad 00|01,\qquad 00|10,\qquad 00|11.
\]
The length-\(4\) classification then shows that the admissible one-step extensions are exactly the edges displayed in the statement. Reading the factor from left to right therefore determines a path in \(G_{\mathrm{pair}}\) labeled by \(w\).

Conversely, every edge of \(G_{\mathrm{pair}}\) is realized by one of the synchronized length-\(4\) factors listed in Appendix~\ref{app:pressure}. Start with a long block \(00^L\), which keeps the presentation in state \(00\). We claim by induction on the path length that any labeled path in \(G_{\mathrm{pair}}\) can be realized after this synchronized prefix. The initial step is exactly the preceding sentence. For the induction step, suppose a synchronized factor ending in state \(s\) has been realized. By construction of the states as synchronized follower classes, the admissible continuations from that suffix are precisely the outgoing edges from \(s\), and Appendix~\ref{app:pressure} lists an explicit synchronized length-\(4\) witness for each such continuation. Appending that witness realizes the next edge and moves the suffix state to the prescribed target state. Iterating along the whole path realizes the complete path label.

Finally, Lemma~\ref{lem:zero-padding} allows us to increase the initial high-weight padding arbitrarily, so the realized path label can be placed at arbitrary distance from the boundary. This proves both directions.
\end{proof}

Since the first coordinate of each pair symbol is the microstate digit itself, the map \(\omega\mapsto \mathcal{R}_M(\omega)\) is injective, so the uniform measure on \(\Om_M\) is exactly the uniform measure on admissible reversed pair words of length \(M\). Proposition~\ref{prop:bulk-limit} therefore implies that the bulk law is the local weak limit of the uniform measures on \(\Om_M\): if \(\omega^{(M)}\) is uniform on \(\Om_M\) and \(k_M\to\infty\) with \(M-k_M\to\infty\), then
\[
\bigl(\omega^{(M)}_{k_M},(\Fold_M(\omega^{(M)}))_{k_M}\bigr)
\Rightarrow
(X_0,Y_0);
\]
see \cite[Section~8.4]{LindMarcus1995} and \cite[Chapter~2]{Kitchens1998}.

\begin{theorem}[Exact bulk joint law]\label{thm:joint}
Under the bulk law, the one-site joint distribution of \((X_0,Y_0)\) is
\[
\begin{array}{c|cc}
& Y_0=0 & Y_0=1\\\hline
X_0=0 & \frac{7}{18} & \frac{1}{9}\\
X_0=1 & \frac{1}{3} & \frac{1}{6}.
\end{array}
\]
Consequently,
\[
\PP(X_0\neq Y_0)=\frac{4}{9},
\qquad
\PP(Y_0=1)=\frac{5}{18}.
\]
\end{theorem}

\begin{proof}
By Theorem~\ref{thm:pair-language}, the graph \(G_{\mathrm{pair}}\) is a primitive right-resolving presentation of the interior reversed-pair language. In the state order \((00,01,10,11)\), its adjacency matrix is
\[
A=\begin{pmatrix}
1&1&0&1\\
0&0&1&0\\
1&2&0&0\\
1&0&0&0
\end{pmatrix},
\]
and the edges are labeled by \((X_t,Y_t)\in\{0,1\}^2\) as follows:
\[
00 \xrightarrow{00} 00,\quad
00 \xrightarrow{01} 01,\quad
00 \xrightarrow{11} 11,\quad
01 \xrightarrow{10} 10,
\]
\[
10 \xrightarrow{10} 00,\quad
10 \xrightarrow{00} 01,\quad
10 \xrightarrow{11} 01,\quad
11 \xrightarrow{00} 00.
\]
The graph is primitive, so its bulk measure is the Parry measure \cite{Parry1964,ParryPollicott1990,Ruelle2004}. Appendix~\ref{app:pressure} computes the Perron eigenvalue \(\lambda=2\), right/left eigenvectors \(r=(2,1,2,1)^{\mathsf T}\) and \(\ell=(2,2,1,1)^{\mathsf T}\), and the stationary state distribution
\[
\pi=\frac{1}{9}(4,2,2,1).
\]
The probability of a labeled edge \(i\to j\) is \(\pi_iP_{ij}/A_{ij}\), where \(P\) is the Parry transition matrix and the multiplicity-\(2\) edge \(10\to 01\) is split uniformly between the labels \(00\) and \(11\). Summing the corresponding edge probabilities over the four labels gives
\[
\PP(00)=\frac{7}{18},\qquad
\PP(01)=\frac{1}{9},\qquad
\PP(10)=\frac{1}{3},\qquad
\PP(11)=\frac{1}{6},
\]
which is exactly the stated table. The two displayed marginals follow by summation.
\end{proof}

\begin{remark}[Comparison with the normalized-digit literature]
The two-sided normalization transducers studied by Alexander--Zagier, Sidorov--Vershik, Kempton, and Bezhaeva--Oseledets \cite{AlexanderZagier1991,SidorovVershik1998,KemptonDigits2014,BezhaevaOseledets2007,BezhaevaOseledets2010} also produce the marginal density \(\PP(Y_0=1)=5/18\). Our four-state graph is not a replacement for those classical models of the normalized-digit process. It is a synchronized presentation for the \emph{pair} observable \((X_t,Y_t)\), and therefore for the restriction discrepancy \(X_t\oplus Y_t\). The agreement of the \(Y\)-marginal confirms compatibility with the earlier Erd\H{o}s-measure literature.
\end{remark}

\subsection{Limit theorems for the discrepancy process}
\label{subsec:lln-clt}
Let \(\Delta_t:=X_t\oplus Y_t\) be the stationary discrepancy indicator process in the bulk. For \(n\ge 1\), set
\[
S_n:=\sum_{t=1}^n \Delta_t.
\]

\begin{corollary}[LLN and CLT for anomaly density]\label{thm:llnclt}
Under the bulk law,
\[
\frac{S_n}{n}\xrightarrow{\ \mathrm{a.s.}\ }\frac{4}{9},
\]
and
\[
\frac{S_n-\frac{4}{9}n}{\sqrt n}\ \Rightarrow\ \mathcal{N}\!\left(0,\frac{118}{243}\right).
\]
\end{corollary}

\begin{proof}
Appendix~\ref{app:pressure} realizes \(\Delta_t\) as a bounded observable of an irreducible, aperiodic, finite-state Markov chain, namely the edge chain of the four-state synchronized presentation. The almost-sure law of large numbers is therefore the ergodic theorem for finite-state Markov chains \cite[Section~4]{Billingsley1961}. The central limit theorem follows from the standard Markov-chain CLT, for instance in Gordin's martingale-coboundary form \cite{Gordin1969}; see also \cite[Chapter~4]{ParryPollicott1990}. The asymptotic variance is the second derivative at \(0\) of the logarithmic pressure of the tilted matrix \(A_\theta\) computed in Appendix~\ref{app:pressure}, namely
\[
\Lambda''(0)=\frac{118}{243}.
\]
\end{proof}
\begin{remark}
The ergodic and central limit theorems are standard once the four-state Markov structure is available; what is specific to the present model is the explicit density \(4/9\) and the variance constant \(118/243\).
\end{remark}

\begin{remark}[Finite-window interpretation]
The process \((X_t,Y_t)\) is the bulk limit of the uniform measures on finite windows. Consequently, the constants in Theorem~\ref{thm:joint} and Corollary~\ref{thm:llnclt} govern the interior statistics of the finite-window anomaly \(G_m\); boundary effects contribute only \(O(1)\) sites and do not alter the limiting density or variance.
\end{remark}

\subsection{Autocovariances and power spectrum}
\label{subsec:spectrum}
Let \((\Delta_t)_{t\in\ZZ}\) be the stationary bulk discrepancy process.

\begin{theorem}[Autocovariance and spectrum]\label{thm:spectrum}
The autocovariances \(c_j:=\Cov(\Delta_0,\Delta_j)\) satisfy
\[
c_0=\frac{20}{81},
\]
and, for \(j\ge1\),
\[
c_j=2^{-j}\!\left(\frac{1}{8}+(-1)^j\!\left(\frac{7}{648}+\frac{j}{108}\right)\right).
\]
The two-sided power spectral density
\[
S(\omega):=\sum_{j\in \ZZ} c_j\,e^{-i\omega j}
\]
is the rational function
\[
S(\omega)
=
\frac{2\bigl(32\cos^3\omega+12\cos^2\omega-116\cos\omega-105\bigr)}
{9\,(4\cos\omega-5)\,(4\cos\omega+5)^2}.
\]
In particular,
\[
S(0)=\sum_{j\in\ZZ}c_j=\frac{118}{243},
\]
in agreement with Corollary~\ref{thm:llnclt}.
\end{theorem}

\begin{proof}
Appendix~\ref{app:pressure} expresses the covariances as
\[
c_j=qP^{j-1}h-\left(\frac49\right)^2\qquad (j\ge1),
\]
for explicit vectors \(q\) and \(h\) attached to the Parry transition matrix \(P\). Because \(P\) is similar to \(A/2\), its nontrivial spectral modes are \(1/2\) and \(-1/2\), with a double root at \(-1/2\). Hence \(c_j\) is a linear combination of \(2^{-j}\), \((-2)^{-j}\), and \(j(-2)^{-j}\), and matching the first three values yields the displayed formula. The series is absolutely summable, so the power spectral density is obtained by summing the resulting cosine series. Appendix~\ref{app:pressure} performs that sum explicitly and obtains the stated rational function. Evaluating at \(\omega=0\) gives \(S(0)=118/243\).
\end{proof}

\begin{remark}[Internal consistency]
The identity
\[
S(0)=c_0+2\sum_{j\ge 1} c_j
\]
is the Green--Kubo/Wiener--Khinchin consistency relation linking the CLT variance, the covariance sum, and the power spectrum. The formulas above therefore provide three equivalent descriptions of the same second-order statistic.
\end{remark}
