\section{Sliding block codes and a conjugacy threshold}
\label{sec:sequence}

The image of a sliding block code on a shift of finite type is automatically sofic, so Proposition~\ref{prop:sofic} is only the starting point. The substantive issue is injectivity: one must show that every ambiguity in the labeled graph is local and then verify that one more label resolves it.

\subsection{The folding code and its image}
Let \(\eta=(\eta_t)_{t\in \ZZ}\in \set{0,1}^{\ZZ}\) be a two-sided microstate sequence. Fix \(m\ge 1\) and define the length-\(m\) sliding block folding code
\begin{equation}\label{eq:Phi}
(\Phi_m(\eta))_t
:=
\Fold_m(\eta_{t-m+1}\cdots \eta_t)\in X_m,
\qquad t\in \ZZ.
\end{equation}
The image shift is
\[
Y_m := \Phi_m(\set{0,1}^{\ZZ})\subset X_m^{\ZZ},
\]
equipped with the left shift on the \(X_m\)-alphabet.

\subsection{A graph presentation}
\begin{proposition}[Sofic model]\label{prop:sofic}
For each \(m\ge 1\), the image shift \(Y_m\) is sofic and admits an explicit labeled-graph presentation.
\end{proposition}

\begin{proof}[Construction]
Let \(V_m:=\set{0,1}^{m-1}\). For each \(v\in V_m\) and \(b\in \set{0,1}\), define the successor vertex \(v'\) by shifting and appending \(b\),
\[
v'= \text{suffix}_{m-1}(vb),
\]
and label the directed edge \(v\to v'\) by
\[
\lambda_m(vb):=\Fold_m(vb)\in X_m.
\]
Denote the resulting labeled graph by \(G_m\).
Let \(\Sigma_m\) be the edge shift of this finite directed graph and let \(\Lambda_m:\Sigma_m\to X_m^{\ZZ}\) be the label map. By construction, \(\Lambda_m(\Sigma_m)=Y_m\), hence \(Y_m\) is sofic.
\end{proof}

\subsection{Ambiguous cores and a three-label decoder}
The conjugacy question is best studied in the fiber product of the graph from Proposition~\ref{prop:sofic}. We say that a two-label block \((y_{t-1},y_t)\in X_m^2\) is \emph{ambiguous} if it admits two distinct lifts in \(G_m\), or equivalently if it labels an off-diagonal path of length \(2\) in the fiber product \(\widehat G_m\).

Some coordinates of an ambiguous block may be passive, in the sense that deleting them leaves the ambiguity unchanged. Removing all such coordinates yields a minimal witness, called an \emph{ambiguous core}. Appendix~\ref{app:sofic} contains the corresponding fiber-product bookkeeping. For the argument in the main text, the relevant point is that the local relation \(110\leftrightarrow 001\) has span \(3\), so the active part of a minimal witness can occupy at most four consecutive coordinates.

\begin{theorem}[Ambiguous cores and three-label resolution]\label{thm:core-resolution}
Fix \(m\ge 3\). Every ambiguous two-label block in \(Y_m\) strips to one of the following minimal cores.
\begin{enumerate}
\item[(a)] Two length-\(3\) cores:
\[
(000,001),\qquad (100,001).
\]
\item[(b)] Eight length-\(4\) cores:
\[
(0000,0001),\quad (1000,0001),\quad (0010,1001),\quad (0100,1001),
\]
\[
(0001,0000),\quad (0001,0010),\quad (1001,0000),\quad (1001,0010).
\]
\end{enumerate}
No minimal ambiguous core has length at least \(5\). In every listed case, the third label \(y_t\) determines the current appended microstate bit uniquely. Consequently there is a well-defined decoder
\[
\psi_m:\mathcal{B}_3(Y_m)\to\{0,1\},
\qquad
\eta_t=\psi_m(y_{t-2},y_{t-1},y_t),
\]
where \(\mathcal{B}_3(Y_m)\) is the set of admissible three-label blocks in \(Y_m\).
\end{theorem}

\begin{proof}
Appendix~\ref{app:sofic} formulates the stripping operation on off-diagonal fiber-product paths: removing passive coordinates preserves ambiguity and terminates at a minimal core. In such a core, both the left boundary and the right boundary of the active interval must lie inside a local rewrite \(110\leftrightarrow 001\). Since each rewrite triple has length \(3\), the two boundary triples must overlap in a minimal witness; otherwise a coordinate strictly between them would be passive and could be stripped. Therefore the active interval has length at most \(4\), which proves that no minimal ambiguous core has length \(\ge 5\).

The remaining possibilities are then the complete \(m=3\) and \(m=4\) core lists recorded in Appendix~\ref{app:sofic}. For length \(3\), the only ambiguous two-label blocks are
\[
(000,001)\quad\text{and}\quad (100,001).
\]
After the second label, the two competing lifts terminate at the vertices \(01\) and \(10\). Their successor label sets are
\[
\mathcal{L}^+(01)=\{000,010\},
\qquad
\mathcal{L}^+(10)=\{100,101\},
\]
which are disjoint. Hence the third label resolves the ambiguity, and therefore determines the appended bit on the third edge.

For length \(4\), Appendix~\ref{app:sofic} shows that the only ambiguous cores are the eight listed in part (b). They split into three finite patterns. The four cores
\[
(0000,0001),\quad (1000,0001),\quad (0010,1001),\quad (0100,1001)
\]
leave terminal vertices \(001\) and \(110\) after the second label. Both vertices have successor label set \(\{0000,0010\}\), and in both states the label \(0000\) is produced by appended bit \(1\) while the label \(0010\) is produced by appended bit \(0\). The two cores
\[
(0001,0000),\qquad (1001,0000)
\]
leave terminal vertices \(011\) and \(101\), whose successor label sets \(\{0001,0100\}\) and \(\{1000,1010\}\) are disjoint. Finally, the two cores
\[
(0001,0010),\qquad (1001,0010)
\]
leave terminal vertices \(010\) and \(100\), whose successor label sets \(\{0100,0101\}\) and \(\{1000,1001\}\) are again disjoint. Hence in every length-\(4\) core the third label determines the correct lift, and therefore the current appended bit.

Finally, every ambiguous two-label block strips to one of the cores above, and reattaching passive coordinates does not alter the third label. Therefore the same three-label rule resolves every ambiguity in \(Y_m\), so the decoder \(\psi_m\) is well defined on all admissible three-label blocks.
\end{proof}

\begin{example}[Continuation of Example~\ref{ex:running}]
For the block \(11011\) from Example~\ref{ex:running}, the three consecutive \(3\)-window labels are
\[
(001,101,000).
\]
They form the path
\[
11 \xrightarrow{001} 10 \xrightarrow{101} 01 \xrightarrow{000} 11.
\]
Hence the decoder reads the appended bit on the last edge as \(1\), exactly as claimed in Example~\ref{ex:running}.
\end{example}

\subsection{The threshold \texorpdfstring{\(m\ge 3\)}{m >= 3}}
\begin{theorem}[Conjugacy for \(m \ge 3\)]\label{thm:conjugacy}
For \(m\ge 3\), the sliding block code \(\Phi_m:\set{0,1}^{\ZZ}\to Y_m\) is a topological conjugacy: there exists a sliding block inverse \(\Psi_m:Y_m\to \set{0,1}^{\ZZ}\) such that
\[
\Psi_m\circ \Phi_m=\mathrm{id}_{\set{0,1}^{\ZZ}},
\qquad
\Phi_m\circ \Psi_m=\mathrm{id}_{Y_m}.
\]
\end{theorem}

\begin{proof}
Define
\[
(\Psi_m(y))_t:=\psi_m(y_{t-2},y_{t-1},y_t),\qquad y\in Y_m,
\]
where \(\psi_m\) is the three-label decoder from Theorem~\ref{thm:core-resolution}. If \(y=\Phi_m(\eta)\), then each triple \((y_{t-2},y_{t-1},y_t)\) arises from three consecutive edges of \(G_m\), so Theorem~\ref{thm:core-resolution} gives
\[
(\Psi_m(\Phi_m(\eta)))_t=\eta_t
\qquad\text{for every }t\in\ZZ.
\]
Hence \(\Psi_m\circ\Phi_m=\mathrm{id}_{\set{0,1}^{\ZZ}}\).

Conversely, fix \(y\in Y_m\). By definition of \(Y_m\), there exists at least one lift path in \(G_m\) with label sequence \(y\). Read from that path the appended microstate bits \(b_t\). Applying Theorem~\ref{thm:core-resolution} to each three-label block shows that
\[
b_t=\psi_m(y_{t-2},y_{t-1},y_t)=(\Psi_m(y))_t
\qquad\text{for every }t\in\ZZ.
\]
Hence every lift path of \(y\) has the same appended-bit sequence \(\Psi_m(y)\). The corresponding vertex sequence is then forced, because the vertex at time \(t\) is the \((m-1)\)-block
\[
v_t=\bigl((\Psi_m(y))_{t-m+1},\dots,(\Psi_m(y))_{t-1}\bigr).
\]
The edge from \(v_t\) to \(v_{t+1}\) is labeled by
\[
\Fold_m\bigl((\Psi_m(y))_{t-m+1}\cdots (\Psi_m(y))_t\bigr),
\]
which is exactly \(y_t\) because the chosen lift path already has those appended bits and label \(y_t\). Therefore
\[
\Phi_m(\Psi_m(y))=y.
\]
Thus \(\Phi_m\) is bijective onto \(Y_m\) with sliding block inverse \(\Psi_m\).
\end{proof}

\begin{proposition}[The degenerate case \(m=2\)]\label{thm:m2}
When \(m=2\), the only failure of injectivity is that the two fixed microstate sequences \(0^\ZZ\) and \(1^\ZZ\) have the same image \((00)^\ZZ\). Away from this pair, \(\Phi_2\) is injective.
\end{proposition}

\begin{proof}
Appendix~\ref{app:sofic} identifies the only off-diagonal bi-infinite cycle in the fiber product for \(m=2\): it is the constant label sequence \((00)^\ZZ\), lifted once by \(0^\ZZ\) and once by \(1^\ZZ\). Every other label sequence contains \(01\) or \(10\), and those labels determine the adjacent microstate bits uniquely. Hence \((00)^\ZZ\) is the unique non-injective fiber.
\end{proof}

\begin{remark}
Theorem~\ref{thm:core-resolution} gives the local inverse mechanism behind Theorem~\ref{thm:conjugacy}: two labels may leave a finite ambiguity, but after passive coordinates are removed only finitely many cores remain, and in each case the third label resolves the ambiguity.
\end{remark}
