% ------------------------------------------------------------------
%  Recursive address factors and finite observation
% ------------------------------------------------------------------
\documentclass[12pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{enumitem}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{microtype}
\geometry{margin=2.7cm}
\microtypesetup{expansion=false}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\DeclareMathOperator{\card}{card}
\DeclareMathOperator{\diam}{diam}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\Col}{Col}
\DeclareMathOperator{\spec}{spec}

\newcommand{\NN}{\mathbb{N}}
\newcommand{\NNzero}{\mathbb{N}_0}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\norm}[1]{\lVert #1\rVert}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\ind}{\mathbf{1}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\Leb}{\mathrm{Leb}}

\title{Finite Observation of Open Symbolic Dynamics:\\
Escape Rates, Cyclotomic Resonances, and Survivor R\'enyi Spectra}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Let $(X,\mathcal{B},\mu,T)$ be a measure-preserving system and let
$W\in\mathcal{B}$ be a binary readout set.  The coding map
$x\mapsto (\mathbf{1}_W(T^t x))_{t\geq 0}$ defines the visible symbolic
factor.  We study what finite observation of that factor detects for
first-entry residue events of symbolic holes.

For visible events, the depth-$m$ Bayes error admits an exact cylinder
decomposition and a Tanaka-type martingale identity.  In the Markov
regime the boundary mass is carried by a finite weighted automaton, so
the finite-observation exponent is a pressure gap.  For first-entry
residue events of Markov holes, this pressure gap is exactly the escape
rate of the open system.  More precisely, the Bayes error is a
residue-twisted killed resolvent, its poles lie in the cyclotomic lift
of the killed spectrum, and its ratio to survivor mass converges to a
quasistationary ambiguity amplitude determined by the Perron left
eigenvector of the killed chain.

We then introduce the survivor R\'enyi pressure spectrum
\[
  \mathcal{Z}_{H,m}(s)
  := \sum_{a\in\mathcal{L}_m(Y_H)} \nu_\phi([a])^s.
\]
For one-step Gibbs-Markov sources we prove
\[
  -\frac1m\log \mathcal{Z}_{H,m}(s)
  \longrightarrow
  sP_Y(\phi)-P_{Y_H}(s\phi),
\]
while conditioning on survival to depth $m$ yields
$sP_{Y_H}(\phi)-P_{Y_H}(s\phi)$.  At $s=2$ this yields the collision
entropy governing birthday thresholds for surviving samples, and the
ambient two-sample survival-collision exponent splits into twice the
escape rate plus the survivor R\'enyi-$2$ entropy rate.

Finally, we lift the finite-state cyclotomic resolvent picture to open
Ruelle spectra for H\"older Gibbs systems under the standard spectral-gap
hypothesis.  A short structural section records that recursive visible
refinements remain inside ordinary factor theory and do not create
entropy.
\end{abstract}

{\small
\noindent\textbf{Keywords:}
open symbolic dynamics, escape rates, Ruelle resonances, R\'enyi
pressures, Poisson approximation, hidden conditional entropy.

\medskip
\noindent\textbf{MSC 2020:}
37B10, 37A50, 37D35, 94A17.
}

\bigskip

\input{sec_introduction}
\input{sec_preliminaries}
\input{sec_spg}
\input{sec_recursive_addressing}
\input{sec_clarity}
\input{sec_open_system}
\input{sec_double_budget}
\input{sec_conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
