\section{Introduction}\label{sec:introduction}

Open dynamical systems ask what remains visible after trajectories are
allowed to escape through a hole.  In symbolic models one can measure
that loss in several ways: by the decay of survivor mass, by the poles
of the killed resolvent, by quasistationary laws on the survivor set,
and by pressure gaps between the ambient system and the survivor
subshift.  The present paper studies a different but closely related
observable: the Bayes error seen at finite observation depth when one
tries to decide an open-system event from the first \(m\) visible
symbols.

Let \((X,\mathcal{B},\mu,T)\) be a measure-preserving system, let
\(W\in\mathcal{B}\) be a binary readout set, and let
\(\omega_\infty(x)=(\mathbf{1}_W(T^t x))_{t\geq 0}\) be the visible
symbolic factor.  For a visible event \(\mathsf P\), the quantity
\[
  \varepsilon_m(\mathsf P;\mu)
\]
denotes the smallest possible classification error among all decisions
based only on the first \(m\) visible symbols.  Our main question is:
what dynamical invariants are encoded in the sequence
\(m\mapsto \varepsilon_m(\mathsf P;\mu)\) when \(\mathsf P\) is defined
by first entry into a symbolic hole?

The paper has a short structural component and a longer open-system
component.  The structural part records that recursive visible
refinements remain inside ordinary factor theory: every recursive step
is a sliding-block factor, visible \(\sigma\)-algebras decrease, and no
new entropy is created.  This bookkeeping plays only a preparatory role
here.  The real content begins once one passes from general visible
events to first-entry residue events for open symbolic systems.

Fix therefore a topologically mixing one-step subshift of finite type
\(Y\), a one-step potential \(\varphi\), and the corresponding
equilibrium Markov measure \(\nu_\varphi\).  Let \(H\) be a Markov hole,
let \(Y_H\) be the survivor subshift, and write
\[
  \gamma_H(\varphi):=P_Y(\varphi)-P_{Y_H}(\varphi)
\]
for the escape rate.  For a non-empty proper subset
\(R\subsetneq \ZZ/q\ZZ\), let \(\mathsf P_R\) be the event that the
first entry time into \(H\) is finite and belongs to \(R\) modulo
\(q\).  The finite-observation problem is then to understand the depth
profile
\[
  m\longmapsto \varepsilon_m(\mathsf P_R;\nu_\varphi).
\]

The first theorem identifies not only the exponential decay rate of this
profile but also its quasistationary amplitude.

\begin{theorem}[Quasistationary ambiguity amplitudes]
\label{thm:main-amplitude}
Assume the setting above and the residue non-degeneracy hypothesis from
Section~\ref{sec:open-system}.  Let \(B_H\) be the killed transition
matrix on the safe alphabet, and let \(\ell_H\) be a non-negative left
eigenvector of \(B_H\) for \(\rho_H:=\rho(B_H)\), normalized by
\[
  \xi_H:=\frac{\ell_H}{\ell_H^{\mathsf T}\mathbf 1}.
\]
For \(r\in \ZZ/q\ZZ\), let \(b_r\) be the residue ambiguity vector from
Theorem~\ref{thm:first-entry-escape-rate}.  Then for every residue class
\(r\),
\[
  \lim_{n\to\infty}
  \frac{\varepsilon_{nq+r+1}(\mathsf P_R;\nu_\varphi)}
       {\nu_\varphi(\tau_H\ge nq+r+1)}
  = \xi_H^{\mathsf T}b_r.
\]
Equivalently,
\[
  \varepsilon_m(\mathsf P_R;\nu_\varphi)
  = \bigl(\xi_H^{\mathsf T}b_{m-1\bmod q}+o(1)\bigr)\,
    \nu_\varphi(\tau_H\ge m).
\]
In particular, finite observation records not only the escape exponent
\(\gamma_H(\varphi)\) but also a residue-dependent quasistationary
ambiguity amplitude.
\end{theorem}

Thus the Bayes error contains a second dynamical datum beyond the escape
rate itself.  In the doubling-hole example, the explicit constant
\(2/5\) arises precisely from this quasistationary amplitude.

The second theorem is the central organizing statement of the paper.  It
unifies escape, collision, and R\'enyi entropy into a single survivor
pressure family.

\begin{theorem}[Survivor R\'enyi pressure spectrum]
\label{thm:main-renyi}
Assume again that \(Y\) is a mixing one-step subshift of finite type,
\(\nu_\varphi\) is the equilibrium Markov measure of a one-step
potential \(\varphi\), and \(H\) is a Markov hole with mixing survivor
subshift \(Y_H\).  For \(m\geq 1\), let \(\mathcal L_m^H\) be the set of
admissible length-\(m\) words that avoid \(H\) for the first \(m\)
steps, and for \(s>0\) define
\[
  \mathcal Z_{H,m}(s)
  := \sum_{a\in \mathcal L_m^H} \nu_\varphi([a])^s.
\]
Then
\[
  \lim_{m\to\infty}
  -\frac1m\log \mathcal Z_{H,m}(s)
  = sP_Y(\varphi)-P_{Y_H}(s\varphi).
\]

If \(\lambda_m^H\) denotes the law of the survivor word conditioned on
survival to depth \(m\),
\[
  \lambda_m^H(a)
  := \frac{\nu_\varphi([a])}{\nu_\varphi(\tau_H\ge m)},
  \qquad a\in\mathcal L_m^H,
\]
then
\[
  \lim_{m\to\infty}
  -\frac1m\log
  \sum_{a\in\mathcal L_m^H} \lambda_m^H(a)^s
  = sP_{Y_H}(\varphi)-P_{Y_H}(s\varphi).
\]

In particular, at \(s=2\) the conditioned survivor law has R\'enyi-\(2\)
entropy rate
\[
  h_{2,H}:=2P_{Y_H}(\varphi)-P_{Y_H}(2\varphi),
\]
while the ambient two-sample exponent for simultaneous survival and
visible collision is
\[
  \chi_H:=2P_Y(\varphi)-P_{Y_H}(2\varphi)
        = 2\gamma_H(\varphi)+h_{2,H}.
\]
\end{theorem}

The identity
\[
  \chi_H = 2\gamma_H(\varphi)+h_{2,H}
\]
is the point at which the collision story becomes part of open-system
thermodynamic formalism rather than an unrelated probabilistic add-on.
Combined with the Chen--Stein argument in
Section~\ref{sec:double-budget}, it yields Poisson birthday thresholds
for surviving samples on the scale
\[
  \binom{N_m}{2}e^{-m h_{2,H}}\asymp 1.
\]
The information-theoretic register estimates from the final subsection
then become a supplementary statement about what it costs to restore a
collapsed R\'enyi spectrum injectively.

The third theorem explains why the finite-state resolvent picture should
be regarded as the first case of a spectral statement about open Ruelle
operators.

\begin{theorem}[Cyclotomic resonance lift]
\label{thm:main-resonance}
For first-entry residue events of Markov holes, the Bayes error
generating function is a residue-twisted killed resolvent,
\[
  \mathcal E_{\mathsf P_R}(z)
  = z\,\widehat u^{\mathsf T}(I-z\widehat B_H)^{-1}\widehat b,
\]
and every pole belongs to the \(q\)-th cyclotomic lift of the killed
spectrum.  The positive real pole of smallest modulus is
\[
  z_H=e^{P_{Y_H}(\varphi)-P_Y(\varphi)},
\]
which is the reciprocal escape factor.

More generally, let \(Y\) be a mixing subshift of finite type,
\(\varphi\) a H\"older potential, and \(H\) a finite union of cylinders.
After passing to a sufficiently high block presentation in which \(H\)
becomes one-step, the residue-error sequence admits an exact backward
transfer-operator representation on the past-history shift of the block
system.  Its generating function is therefore a scalar resolvent
coefficient of a \(q\)-cyclic lift of the open backward Gibbs
specification operator.  Under the standard past--future identification
coming from the natural extension, the isolated spectrum of that
backward operator agrees with the isolated spectrum of the usual forward
open Ruelle operator \(\mathcal L_{\varphi,H}\).  Consequently the error
generating function extends meromorphically throughout every spectral
annulus on which the open resolvent is meromorphic, and its poles belong
to the \(q\)-th cyclotomic lift of the open Ruelle spectrum.  If the
leading open resonance \(e^{-\gamma_H(\varphi)}\) is simple and
isolated, then there exist \(\eta>0\), an integer
\(\ell\geq 1\) determined by the hole length, and positive constants
\(c_r\) such that
\[
  \varepsilon_m(\mathsf P_R;\nu_\varphi)
  = c_{m-\ell\bmod q}e^{-m\gamma_H(\varphi)}
    + O\bigl(e^{-m(\gamma_H(\varphi)+\eta)}\bigr).
\]
\end{theorem}

The first part of Theorem~\ref{thm:main-resonance} is exact and
finite-state.  The second part is the open-system spectral lift that
places the finite-observation problem inside the usual Ruelle
resonance framework for H\"older Gibbs systems.

Thermodynamic formalism for symbolic systems and equilibrium states is
classical; we use standard material from
\citet{Bowen2008,Ruelle1978,Walters1975GMeasures,WaltersErgodicTheory}.
The open-system viewpoint---escape through holes, killed transfer
operators, and quasistationary behaviour---goes back to
\citet{PianigianiYorke1979} and was developed for symbolic systems by
\citet{ColletMartinezSchmitt1997}, \citet{DemersYoung2006},
\citet{KellerLiverani2009}, \citet{DemersWrightYoung2012}, and
\citet{FergusonPollicott2012}.  The Poisson approximation input is the
Chen--Stein method of \citet{ArratiaGoldsteinGordon1989}.  What is
specific here is the claim that finite observation detects these open
invariants directly: escape rate, residue-dependent quasistationary
amplitudes, cyclotomic resonance lifts, and survivor R\'enyi pressure
spectra.

The paper is organized as follows.  Section~\ref{sec:preliminaries}
introduces the visible symbolic factor and finite observation
\(\sigma\)-algebras.  Section~\ref{sec:spg} records visible words and a
brief Sturmian illustration.  Section~\ref{sec:recursive-addressing}
contains only the structural factor-theoretic background on recursive
visible refinements.  Section~\ref{sec:clarity} develops the exact Bayes
error formulas and the weighted boundary transfer theorem.  Section
~\ref{sec:open-system} proves
Theorems~\ref{thm:main-amplitude} and~\ref{thm:main-resonance}.  Section
~\ref{sec:double-budget} proves
Theorem~\ref{thm:main-renyi}, derives the collision consequences, and
then records the hidden-entropy cost of restoring a collapsed collision
spectrum by injective refinement.  The paper ends with brief final
remarks.
