\input{preamble}

\title{Golden-Mean Folding in Binary Scan--Projection Systems:\\Stable Types, Normal Forms, and Auditable Addressing}
\author{
  Haobo Ma\thanks{AELF PTE LTD., \#14-02, Marina Bay Financial Centre Tower 1, 8 Marina Blvd, Singapore 018981. Email: \texttt{auric@aelf.io}}
  \and
  Wenlin Zhang\thanks{Corresponding author. National University of Singapore, Singapore. Email: \texttt{e1327962@u.nus.edu}}
}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Finite-window observations in symbolic dynamics, coding theory, automata, and auditable computation naturally generate exponentially large binary microstate spaces \(\Om_m=\{0,1\}^m\).  If microstates are used directly as types, increasing the resolution \(m\) produces sparsity, representational instability, and poor cross-scale comparability.  This paper studies scan--projection systems induced by a singly generated action \(T^t:X\to X\) and a binary readout \(s_t(x)=\one_W(T^t x)\), and constructs a golden-mean folding mechanism that maps microstates \(\Om_m\) to stable type spaces \(X_m\).

Using Fibonacci/Zeckendorf syntax, we define a folding map \(\Fold_m:\Om_m\to X_m\) and prove termination, confluence, idempotence, and order independence.  Thus every finite-window microstate admits a unique stable normal form.  We then introduce fold-aware restriction maps and prove cross-resolution compatibility, while showing that naive truncation and folding need not commute.  Recursive addressing is formulated as an internal reorganization of previous readout layers, and we prove that derived addresses do not enlarge the existing visible sigma-algebra.

To quantify the intuition that deeper scans improve clarity, we define the optimal prefix decision error \(\varepsilon_m(P;\mu)\), derive its cylinder decomposition, characterize the Bayes-optimal prefix classifier, and obtain exponential clarity bounds controlled by boundary-cylinder dimension.  Together, these results define a computable, cross-scale compatible, and statistically auditable generator of stable types.  The claims of this paper are restricted to the mathematical and operational construction of stable types, addressing, and clarity certificates.  Spectral, arithmetic, \v{C}ech-residual, Hilbert-space, quantum, and spacetime interpretations are treated only as possible downstream bridges and are not used as assumptions or conclusions of the main proof chain.
\end{abstract}

\paragraph{Keywords.}
scan--projection systems; golden mean; Zeckendorf normal forms; symbolic dynamics; stable types; recursive addressing; statistical clarity; auditable computation.

\paragraph{2020 Mathematics Subject Classification.}
Primary 37B10, 68Q45; Secondary 11B39, 37A50, 94A17.

\tableofcontents

\input{sections/01-introduction}
\input{sections/02-scope-claim-hierarchy}
\input{sections/03-scan-projection-systems}
\input{sections/04-golden-syntax}
\input{sections/05-folding-map}
\input{sections/06-cross-resolution}
\input{sections/07-recursive-addressing}
\input{sections/08-clarity}
\input{sections/09-certificates}
\input{sections/10-reproducible-experiments}
\input{sections/11-discussion}
\input{sections/12-conclusion}

\appendix
\input{appendices/appA-rewriting}
\input{appendices/appB-inverse-systems}
\input{appendices/appC-clarity-proofs}
\input{appendices/appD-finite-graph-certificates}
\input{appendices/appE-reproducibility}
\input{appendices/appF-extended-examples}

\bibliographystyle{plain}
\bibliography{references}

\end{document}
