\section{Recursive visible refinements: the key reduction}
\label{sec:recursive-addressing}

This section records, for later use, that recursive visible refinements
remain inside ordinary factor theory and reduce to the base-level
visible process.  The single result needed in
Sections~\ref{sec:open-system}--\ref{sec:double-budget} is
Theorem~\ref{thm:finite-depth-collapse} below.

\begin{definition}[Recursive addressing scheme]
\label{def:recursive-scheme}
A recursive addressing scheme consists of finite families
\(\mathcal{C}^{(L)}=\{C^{(L)}_1,\ldots,C^{(L)}_{r_L}\}\), \(L\geq 0\),
where \(\mathcal{C}^{(0)}\) is given and each
\(C^{(L+1)}_j:=(\pi_{\mathcal{C}^{(L)}})^{-1}(B^{(L+1)}_j)\) for
cylinder sets \(B^{(L+1)}_j\) in the level-\(L\) symbolic shift.  We
write \(\pi^{(L)}:=\pi_{\mathcal{C}^{(L)}}\) for the level-\(L\) coded
process and \(\mathcal{F}_m^{(L)}\) for its depth-\(m\) observation
\(\sigma\)-algebra.
\end{definition}

For each \(q\geq 1\), let \(s_q\) denote the maximum cylinder length
among the defining cylinders used to construct level \(q\) from level
\(q-1\), and set
\(R_0:=1\), \(R_L:=1+\sum_{q=1}^{L}(s_q-1)\) for \(L\geq 1\).

\begin{theorem}[Finite-depth collapse]
\label{thm:finite-depth-collapse}
For every \(L\geq 0\) there exists a sliding-block code
\(\Psi_L\) of window length \(R_L\) such that
\(\pi^{(L)}=\Psi_L\circ\pi^{(0)}\) \(\mu\)-almost everywhere.
Consequently, for every \(m\geq 1\),
\[
  \mathcal{F}_m^{(L)}\subseteq \mathcal{F}_{m+R_L-1}^{(0)}.
\]
In particular, if
\(\varepsilon_m^{(L)}(P;\mu)\) denotes the optimal depth-\(m\) decision
error at level \(L\), then
\[
  \varepsilon_{m+R_L-1}^{(0)}(P;\mu)
  \leq \varepsilon_m^{(L)}(P;\mu)
  \qquad\text{for every } m\geq 1.
\]
\end{theorem}

\begin{proof}
Compose the sliding-block codes level by level:
\(\Psi_L:=\Phi_{L-1}\circ\cdots\circ\Phi_0\), where each \(\Phi_q\)
maps the level-\(q\) symbolic shift to level \(q+1\) by evaluating the
defining cylinders.  The window length satisfies the induction
\(R_{L+1}=R_L+s_{L+1}-1\).  The \(\sigma\)-algebra inclusion and error
inequality follow by pulling back \(m\)-cylinders through a code of
window \(R_L\).
\end{proof}

\begin{remark}
Recursive visible refinement is an internal factor construction:
the visible \(\sigma\)-algebra does not expand and no new entropy is
created.  The full factor-theoretic development---sliding-block
realization, inverse-limit address models, and complete address
reconstruction---is available from the authors on request and is not
needed for the main open-system theorems.
\end{remark}
