\section{Final remarks}\label{sec:conclusion}

The central finding is that finite observation of first-entry residue
events for a Markov hole in a mixing shift of finite type recovers
three layers of open-system information while remaining in a
finite-state framework: the escape rate
\(\gamma_H(\varphi)=P_Y(\varphi)-P_{Y_H}(\varphi)\), the
residue-dependent quasistationary ambiguity amplitude
\(\xi_H^{\mathsf T}b_r\), and a scalar resolvent coefficient whose poles
lie in the cyclotomic lift of the killed spectrum.  In the one-step
Markov setting, these objects are explicit: the killed-matrix formula
(Theorem~\ref{thm:first-entry-escape-rate}) gives exact Bayes error
asymptotics, Example~\ref{ex:amplitude-contrast} separates identical
escape exponents from different ambiguity amplitudes, and
Example~\ref{ex:cyclotomic-example} displays the predicted cyclotomic
pole structure.

The survivor R\'enyi pressure spectrum
\(s\mapsto sP_{Y_H}(\varphi)-P_{Y_H}(s\varphi)\) packages the
collision behaviour of surviving samples into open-system thermodynamic
formalism.  In the present paper this spectrum is a secondary
consequence of the same killed-matrix calculation.  At \(s=2\) it
determines the birthday scale of the conditioned survivor law, with the
ambient two-sample exponent splitting as
\(\chi_H = 2\gamma_H(\varphi)+h_{2,H}\).  The Poisson threshold is
therefore a consequence of the same calculation rather than a separate
phenomenon.

Two directions remain natural.  First, the H\"older extension
mentioned in Remark~\ref{rem:holder-open-resonance} would require a
complete operator-theoretic verification of the spectral-gap and
regularity hypotheses; establishing this would extend the Bayes-error
observable beyond the finite-state setting.  That extension is not
proved here, and the present paper should be read as a focused
finite-state contribution.
Second, it would be interesting to understand whether finer residue
observables or multidimensional holes detect larger pieces of the open
resonance set than the cyclotomic lifts accessed by the scalar error
resolvent considered here.
