\documentclass[11pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\signature{Wenlin Zhang \\ National University of Singapore}
\address{Wenlin Zhang \\ National University of Singapore \\ Singapore \\ \texttt{e1327962@u.nus.edu}}

\begin{document}

\begin{letter}{Editorial Board \\ Ergodic Theory and Dynamical Systems}

\opening{Dear Editors,}

We submit the manuscript ``Finite-Block Folds of Irrational Rotations and
Parry Block Laws'' for consideration by Ergodic Theory and Dynamical Systems.

The paper studies a finite symbolic factor of irrational rotation
words into the golden-mean language, defined through Fibonacci
valuation, Zeckendorf normalization and prefix truncation. The main
point is dynamical: in the Sturmian slice the fold is lossless on
the realized language and gives a relabeling of the higher-block
presentation, yet its block laws remain separated from the Parry
law of the ambient shift by an explicit finite-resolution barrier.

The main contributions are:

\begin{enumerate}
\item A no-collision and higher-block conjugacy theorem for the
canonical Sturmian slice of the fold.

\item Closed-form R\'enyi divergence envelopes and an explicit optimal
injective relabeling of the realized Sturmian law into the
golden-mean shift, valid for all sufficiently large block lengths.

\item A finite-block factor rigidity theorem showing that the
Parry divergence barrier is intrinsic for aperiodic finite-block
factors of bounded-type Sturmian systems inside primitive
positive-entropy shifts.

\item A proof that the Fibonacci weighting is the unique choice
satisfying a natural contiguous numeration axiom, placing the
fold on a canonical footing.

\item Explicit discrepancy and relative-entropy estimates that
separate finite sampling error from the limiting Parry comparison.
\end{enumerate}

The paper is centered on symbolic dynamics and ergodic theory:
Sturmian systems, higher-block presentations, Parry measures,
primitive shifts of finite type, and finite-block factors. The
numerical section is included only to check the explicit estimates
and is not the source of the main conclusions.

All numerical tables are reproducible from the Python scripts
included in the supplementary materials.

Neither the manuscript nor any substantial part of it is under
consideration elsewhere. Both authors have approved the submission.

\closing{Sincerely,}

\end{letter}
\end{document}
