\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{microtype}

\numberwithin{equation}{section}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\TT}{\mathbb{T}}
\newcommand{\Fold}{\mathrm{Fold}}
\newcommand{\leb}{\mathrm{Leb}}
\newcommand{\pref}{\mathrm{pref}}
\DeclareMathOperator{\TV}{TV}
\DeclareMathOperator{\KL}{KL}
\DeclareMathOperator{\supp}{supp}

\title{Finite-Block Folds of Irrational Rotations and Parry Block Laws}
\author{Haobo Ma\thanks{AELF PTE LTD., \#14-02, Marina Bay Financial Centre Tower 1, 8 Marina Blvd, Singapore 018981, Singapore. \texttt{auric@aelf.io}.}
\and
Wenlin Zhang\thanks{National University of Singapore, Singapore. Corresponding author. \texttt{e1327962@u.nus.edu}.}}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We consider a finite symbolic map from binary rotation words to the
golden-mean language, obtained by taking a Fibonacci valuation, passing to
Zeckendorf normal form and truncating the resulting prefix.  Star-discrepancy
estimates for irrational rotations give explicit total-variation and relative
entropy bounds for the empirical laws of these folded words.  In the Sturmian
case \(\beta=\alpha\), the map has a stronger property: on the realized
length-\(m\) language it is injective, and the induced process is conjugate to
the usual \(m\)-block presentation of the Sturmian subshift.

The comparison with the golden-mean Parry block law can then be made at fixed
resolution.  We give an exact identity for \(\KL(\nu\|\pi_m)\), determine the
support-constrained minima for all R\'enyi orders, and, for \(m\ge 6\), identify
the optimal injective placement of the Sturmian factor law in the golden-mean
language.  For bounded-type slopes the canonical folded law stays within a
bounded additive error of this optimum, uniformly in the R\'enyi order, and
\[
D_q(\pi_m^{\mathrm{fold}}\|\pi_m)=m\log\varphi-\log m+O_\alpha(1).
\]
The same leading term holds for aperiodic finite-block factors of
bounded-type Sturmian systems inside any primitive positive-entropy shift,
with \(m\log\varphi\) replaced by \(mh\); periodic factors give \(mh+O(1)\).
Thus the Parry discrepancy is a finite-block factor phenomenon, not merely a
feature of the particular Zeckendorf normalization.  The construction is also
carried over, by conjugacy, to \(C^2\) circle diffeomorphisms with irrational
rotation number.  Numerical tables are included only to illustrate the
separation between sampling error and the limiting Parry comparison.
\end{abstract}

\medskip
\noindent\textbf{Key words.} irrational rotations, circle maps, Sturmian
words, Parry measure, R\'enyi divergence, discrepancy, shifts of finite type

\medskip
\noindent\textbf{AMS subject classifications.} 37B10, 37A55, 94A17, 11B39

\section{Introduction}

Let an irrational rotation be coded by a finite binary window.  If the
resulting words are then passed through a deterministic map, the empirical law
may differ from a natural reference law for two unrelated reasons: the orbit
has been sampled only for a finite time, and the limiting symbolic law may
occupy a small part of the reference state space.  This paper treats a case in
which both effects can be separated at a fixed block length.

The map considered below is defined as follows.  Given
\(\omega\in\{0,1\}^m\), take its Fibonacci value, write this integer in
Zeckendorf normal form, and retain the first \(m\) digits.  This produces a
map
\[
\Fold_m:\{0,1\}^m\longrightarrow X_m,
\]
where \(X_m\) is the length-\(m\) language of the golden-mean shift.  Although
\(\Fold_m\) has many fibers on the full cube, it is injective on the
length-\(m\) factors that occur in the Sturmian coding with \(\beta=\alpha\).
Equivalently, the folded sequence is just a relabelled \(m\)-block
presentation of the Sturmian subshift.  No entropy, recurrence or complexity
is lost by the fold on the realized language.

This observation is the point at which the sampling question and the Parry
comparison separate.  Star discrepancy controls the empirical mass of each
rotation cylinder, giving total-variation bounds for the microstate law and
for its folded pushforward.  The constants are \(2m\) in the general rotation
window and \(m+1\) in the Sturmian case.  With a lower bound on the smallest
atom of the associated finite rotation partition, the same estimates give
relative-entropy bounds; in the Sturmian slice this lower bound has the usual
continued-fraction and Diophantine forms.  The remaining discrepancy from the
golden-mean Parry law is therefore a property of the limiting folded law, not
of the orbit length.

For the golden-mean shift this limiting comparison is unusually explicit.
Parry cylinder weights collapse to three endpoint classes, and this gives an
exact finite-\(m\) identity for \(\KL(\nu\|\pi_m)\), as well as two-sided
R\'enyi bounds uniform in the order.  The associated support problem is also
finite-dimensional: the largest Parry mass of a \(k\)-point set is obtained by
filling endpoint classes in decreasing weight.  If the Sturmian weights are
kept fixed and only injective placements of the realized words in \(X_m\) are
allowed, then for \(m\ge 6\) the optimum is obtained by placing all words in
the heaviest endpoint class \(X_m^{00}\), and
\[
\inf_{\iota} D_q(\iota_*\mu_m\|\pi_m)
=(m-1)\log\varphi-\log\pi(0)-H_q(\mu_m).
\]
For bounded-type slopes this is
\[
m\log\varphi-\log m+O_\alpha(1),
\]
uniformly in \(q\).  The canonical folded law is never more than
\(2\log\varphi\) above this optimum.

The same mechanism persists beyond the golden-mean example.  For a primitive
shift of finite type with entropy \(h\), an injective coding of a
low-complexity source has Parry divergence \(mh-H_q(\mu_m)+O(1)\).  Moreover,
for bounded-type Sturmian systems, every finite-block factor is either
periodic or remains linearly recurrent with linear complexity.  Thus an
aperiodic finite-block factor, regarded inside a primitive positive-entropy
shift, has divergence \(mh-\log m+O(1)\), uniformly in the R\'enyi order.  The
canonical fold is one concrete instance of this more general finite-block
obstruction.

All of the separate ingredients are classical.  Sturmian codings, factor
complexity and factor frequencies go back to
\cite{MorseHedlund1940,Slater1950DistributionIntegers,Slater1967GapsSteps} and
are naturally part of symbolic dynamics and substitutions
\cite{LindMarcus1995,QueffelecSubstitution,Fogg2002}. The Parry measure and
the thermodynamic formalism of shifts of finite type are classical
\cite{Parry1964IntrinsicMarkovChains,LindMarcus1995,WaltersErgodicTheory,
ParryPollicott1990Zeta}. Fibonacci and Ostrowski numeration provide the
arithmetic background for the normalization
\cite{Zeckendorf1972,Frougny1991FibonacciFiniteAutomata,
Frougny1992NumbersAutomata,Berthe2001Ostrowski,
BaranwalaSchaefferShallit2021OstrowskiAutomatic}. The discrepancy and
empirical-divergence estimates use
\cite{KuipersNiederreiter1974,Goda2024QuasiUniformKronecker,
WeissmanEtAl2003L1Empirical,SasonVerdu2016fDivergence,
DemboZeitouni2010LargeDeviations}.

The contribution is the finite-resolution comparison obtained when these
standard structures are coupled by the canonical fold: the no-collision
theorem on the realized Sturmian language, the exact Parry block optimization,
and the finite-block factor obstruction for bounded-type Sturmian sources
inside primitive positive-entropy shifts.

Section~2 sets up the rotation model, the fold and the Parry block law.
Sections~3 and~4 prove the sampling estimates and the Sturmian rigidity
statements.  Section~5 contains the Parry comparison, including the support
minima, the injective placement problem, the bounded-type consequences and the
extension to primitive shifts.  Section~6 records the numerical checks and the
transfer to nonlinear circle maps.  The appendices contain the fiber-uniform
lift identities, a local normalization description of the fold, and the
uniqueness of the Fibonacci weights under a contiguous numeration axiom.

\section{Model and target measures}

\subsection{Rotations, windows and the folded observable}

Let $\alpha\in(0,1)$ be irrational, let $\beta\in(0,1)$, and let
$x_0\in\TT:=\mathbb{R}/\mathbb{Z}$.  Define the Kronecker orbit
\[
x_t:=x_0+t\alpha \pmod 1,\qquad t\ge 0,
\]
and the binary readout
\[
s_t:=\mathbf{1}_{[0,\beta)}(x_t)\in\{0,1\}.
\]
For a fixed resolution $m\ge 1$, let
\[
\Omega_m:=\{0,1\}^m,\qquad
X_m:=\{u=(u_1,\dots,u_m)\in\{0,1\}^m:\ u_i u_{i+1}=0\text{ for }1\le i<m\}.
\]
Throughout \(F_0:=0\), \(F_1:=1\), and
\(F_{n+2}=F_{n+1}+F_n\).

\begin{definition}[Canonical fold]\label{def:canonical-fold}
For \(\omega=\omega_1\cdots\omega_m\in\Omega_m\), define its Fibonacci value by
\[
N_m(\omega):=\sum_{j=1}^m \omega_j F_{j+1}.
\]
Let \(Z(n)=(z_1,z_2,\dots)\) be the Zeckendorf expansion of \(n\)
\cite{Zeckendorf1972}, written in the convention
\[
n=\sum_{j\ge 1} z_jF_{j+1},\qquad z_j\in\{0,1\},\qquad z_jz_{j+1}=0.
\]
The canonical fold is
\[
\Fold_m(\omega):=z_1z_2\cdots z_m\in X_m,
\qquad Z(N_m(\omega))=(z_1,z_2,\dots).
\]
\end{definition}

\begin{proposition}[Basic properties of the canonical fold]\label{prop:canonical-fold-basic}
For every \(m\ge 1\), the map \(\Fold_m:\Omega_m\to X_m\) is surjective and
fixes \(X_m\) pointwise. Moreover, \(N_m\) restricts to a bijection
\[
N_m:X_m\longrightarrow \{0,1,\dots,F_{m+2}-1\}.
\]
For every \(\omega\in\Omega_m\),
\[
N_m(\Fold_m(\omega))\equiv N_m(\omega)\pmod {F_{m+2}}.
\]
\end{proposition}

\begin{proof}
Zeckendorf uniqueness \cite{Zeckendorf1972} gives a one-to-one correspondence
between legal \(\{0,1\}\)-digit strings supported in positions \(1,\dots,m\)
and the integers \(0,\dots,F_{m+2}-1\). Hence \(N_m|_{X_m}\) is bijective. If
\(x\in X_m\), then \(x\) is already the Zeckendorf expansion of \(N_m(x)\), so
\(\Fold_m(x)=x\); surjectivity follows.

Finally, for \(\omega\in\Omega_m\),
\[
N_m(\omega)\le \sum_{j=1}^m F_{j+1}=F_{m+3}-2<F_{m+3}.
\]
Thus \(Z(N_m(\omega))\) has no nonzero digit beyond position \(m+1\). Write
\[
Z(N_m(\omega))=(z_1,\dots,z_m,z_{m+1},0,\dots).
\]
Then
\[
N_m(\omega)=N_m(\Fold_m(\omega))+z_{m+1}F_{m+2},
\]
which proves the congruence.
\end{proof}

\begin{proposition}[Residue characterization of the canonical fold]\label{prop:fold-residue-characterization}
For \(\omega\in\Omega_m\) and \(x\in X_m\),
\[
\Fold_m(\omega)=x
\quad\Longleftrightarrow\quad
N_m(\omega)\equiv N_m(x)\pmod {F_{m+2}}.
\]
\end{proposition}

\begin{proof}
The forward implication is the congruence in
Proposition~\ref{prop:canonical-fold-basic}. Conversely, let
\(y:=\Fold_m(\omega)\). Then \(y\in X_m\), and
\[
N_m(y)\equiv N_m(\omega)\equiv N_m(x)\pmod {F_{m+2}}.
\]
Since both \(N_m(x)\) and \(N_m(y)\) lie in
\(\{0,\dots,F_{m+2}-1\}\), they are equal. The bijectivity of \(N_m|_{X_m}\)
then gives \(x=y\).
\end{proof}

\begin{definition}[Microstates and folded histograms]\label{def:microstates-folded-histograms}
For \(x\in\TT\), set
\[
\omega_m(x):=\bigl(\mathbf{1}_{[0,\beta)}(x+j\alpha)\bigr)_{j=0}^{m-1}
\in\Omega_m.
\]
Along the orbit \(x_t=x_0+t\alpha\), write
\[
\omega_m(t):=\omega_m(x_t)=s_t\cdots s_{t+m-1},
\qquad
y_m(t):=\Fold_m(\omega_m(t)).
\]
The empirical microstate and folded laws are
\[
\hat\mu_{m,N}(a):=\frac1N\#\{0\le t<N:\omega_m(t)=a\},
\qquad a\in\Omega_m,
\]
and
\[
\hat\pi_{m,N}(x):=\frac1N\#\{0\le t<N:y_m(t)=x\},
\qquad x\in X_m.
\]
Equivalently, \(\hat\pi_{m,N}=(\Fold_m)_*\hat\mu_{m,N}\).
\end{definition}

\begin{remark}[Use of the fold]\label{rem:scope-fold-assumptions}
The map \(\Fold_m\) is fixed throughout the paper.  The discrepancy estimates
below use only that it is deterministic.  The residue formula, the
cross-resolution boundary law and the no-collision theorem, however, use the
particular Zeckendorf normalization in Definition~\ref{def:canonical-fold}.
The Parry identities in Section~5 are statements about laws on \(X_m\) and are
independent of how those laws arise.  We do not compare different
Zeckendorf-type normalizations here; the rigidity result is specific to the
canonical residue fold.
\end{remark}

Because irrational rotations are uniquely ergodic, the empirical microstate
law converges to a deterministic target law induced by Lebesgue measure.

\begin{definition}[Target laws]\label{def:target-laws}
For \(a\in\Omega_m\), let
\[
C_m(a):=\{x\in\TT:\omega_m(x)=a\}
\]
and define
\[
\mu_m(a):=\leb(C_m(a)).
\]
The folded target law is
\[
\pi_m^{\mathrm{fold}}:=(\Fold_m)_*\mu_m.
\]
\end{definition}

\begin{remark}[Spaces and laws]\label{rem:model-spaces-maps}
The notation is summarized by
\[
\begin{array}{ccc}
\Omega_m & \xrightarrow{\Fold_m} & X_m\\[0.3em]
\mu_m \downarrow\phantom{(\Fold_m)_*} &  & \phantom{(\Fold_m)_*}\downarrow \pi_m^{\mathrm{fold}}\\[0.3em]
\hat\mu_{m,N} & \xrightarrow{(\Fold_m)_*} & \hat\pi_{m,N}
\end{array}
\qquad
\begin{array}{ccc}
\Omega_{m+1} & \xrightarrow{\Fold_{m+1}} & X_{m+1}\\[0.3em]
\pref_m \downarrow &  & \downarrow \rho_{m+1\to m}\\[0.3em]
\Omega_m & \xrightarrow{\Fold_m} & X_m.
\end{array}
\]
The left diagram distinguishes the microstate space \(\Omega_m\) from the
folded space \(X_m\).  In the right diagram, \(\pref_m\) and
\(\rho_{m+1\to m}\) denote prefix truncations on \(\Omega_{m+1}\) and
\(X_{m+1}\), respectively.  Their failure to commute with the fold is measured
in Section~3.2.
\end{remark}

\subsection{The Parry measure}

The reference law on \(X_m\) is the \(m\)-block marginal of the Parry measure
of the golden-mean shift.  Let
\[
A=\begin{pmatrix}1&1\\1&0\end{pmatrix},\qquad
\varphi=\frac{1+\sqrt5}{2}.
\]
The associated two-state Parry chain has transition matrix
\[
P=\begin{pmatrix}\varphi^{-1} & \varphi^{-2}\\ 1&0\end{pmatrix},
\]
and stationary distribution
\[
\pi(1)=\frac{1}{\varphi^2+1},\qquad
\pi(0)=1-\pi(1).
\]
For $u=u_1\cdots u_m\in X_m$, define the length-$m$ Parry cylinder law
\[
\pi_m(u):=\pi(u_1)\prod_{j=1}^{m-1}P_{u_j u_{j+1}}.
\]
\(\pi_m\) is used only as a reference measure; in general
\(\pi_m^{\mathrm{fold}}\ne\pi_m\).  Since \(|X_m|=F_{m+2}\), the ambient
shift has topological entropy
\[
h_{\mathrm{top}}(X)=\lim_{m\to\infty}\frac1m\log |X_m|=\log\varphi,
\]
and the Parry measure is its unique measure of maximal entropy
\cite{LindMarcus1995,WaltersErgodicTheory}.

\begin{proposition}[Endpoint collapse of Parry cylinders]\label{prop:parry-endpoint-collapse}
For every \(m\ge1\) and every \(u=u_1\cdots u_m\in X_m\),
\[
\pi_m(u)=
\pi(0)\,\varphi^{-m+\mathbf{1}_{\{u_1=0\}}-\mathbf{1}_{\{u_m=1\}}}.
\]
Equivalently, for \(m\ge2\), the endpoint classes have weights
\[
\pi_m(u)=
\begin{cases}
\pi(0)\varphi^{-(m-1)}, & u_1u_m=00,\\
\pi(0)\varphi^{-m}, & u_1u_m\in\{01,10\},\\
\pi(0)\varphi^{-(m+1)}, & u_1u_m=11.
\end{cases}
\]
\end{proposition}

\begin{proof}
For an admissible word, write \(n_{ij}\) for the number of transitions
\(i\to j\). Since \(11\) is forbidden,
\[
\pi_m(u)=\pi(u_1)\varphi^{-n_{00}-2n_{01}}.
\]
The identities
\[
n_{01}-n_{10}=u_m-u_1,
\qquad
n_{00}+n_{01}+n_{10}=m-1
\]
give \(n_{00}+2n_{01}=m-1-u_1+u_m\). Also
\(\pi(1)=\pi(0)\varphi^{-2}\). Substituting these two relations into the
Parry cylinder formula yields the displayed expression. The endpoint table is
immediate.
\end{proof}

\subsection{Discrepancy of the rotation orbit}

Our only quantitative input is the one-dimensional star discrepancy of the
rotation orbit:
\[
D_N^*(\alpha;x_0)
:=
\sup_{0\le a\le 1}
\left|
\frac{1}{N}\sum_{t=0}^{N-1}\mathbf{1}_{[0,a)}(x_t)-a
\right|.
\]
This quantity is effective because length-$m$ rotation words are determined by a
finite partition of the circle into interval atoms.

\begin{proposition}[Continued-fraction discrepancy rate]\label{prop:rotation-discrepancy-cf-rate}
Let \(\alpha=[0;a_1,a_2,\dots]\), let \(q_{-1}=0,q_0=1\), and let
\(q_n\) be the continued-fraction denominators.  Choose \(s=s(N)\) so that
\(q_s\le N<q_{s+1}\), and write the standard Ostrowski expansion
\cite{Cassels1957}
\[
N=\sum_{j=0}^s b_jq_j,\qquad 0\le b_j\le a_{j+1}.
\]
Then, uniformly in \(x_0\),
\[
N D_N^*(\alpha;x_0)\le 2\sum_{j=0}^s b_j
\le 2\sum_{j=1}^{s+1}a_j .
\]
In particular, if the partial quotients of \(\alpha\) are bounded by \(A\),
then
\[
D_N^*(\alpha;x_0)
\le
\frac{2A}{N}
\left(2+\frac{\log(N+1)}{\log\varphi}\right).
\]
For the golden slope \(\alpha=\varphi^{-1}=[0;1,1,\dots]\),
\[
D_N^*(\varphi^{-1};x_0)
\le
\frac{2}{N}
\left(2+\frac{\log(N+1)}{\log\varphi}\right).
\]
\end{proposition}

\begin{proof}
For \(0\le a\le1\), put \(f_a(x)=\mathbf{1}_{[0,a)}(x)-a\).  The
Denjoy--Koksma inequality for rotations at continued-fraction denominators
\cite[Chapter~2]{KuipersNiederreiter1974} gives, for every starting point
\(y\),
\[
\left|\sum_{t=0}^{q_j-1} f_a(y+t\alpha)\right|
\le \operatorname{Var}(f_a)\le 2 .
\]
Decompose the first \(N\) iterates into the consecutive Ostrowski blocks of
lengths \(q_j\).  Since the preceding estimate is uniform in the starting
point of each block,
\[
\left|\sum_{t=0}^{N-1}f_a(x_0+t\alpha)\right|
\le 2\sum_{j=0}^s b_j .
\]
After division by \(N\) and taking the supremum over \(a\), this proves the
continued-fraction bound.  The digit bound \(b_j\le a_{j+1}\) gives the second
inequality.

If \(a_j\le A\), then \(\sum_{j=1}^{s+1}a_j\le A(s+1)\).  The universal
growth lower bound \(q_s\ge F_{s+1}\ge \varphi^{s-1}\) gives
\[
s+1\le 2+\frac{\log(N+1)}{\log\varphi},
\]
and the bounded-type estimate follows.  The golden case is the special case
\(A=1\).
\end{proof}

\begin{proposition}[Sturmian cylinder lengths]\label{thm:spg-sturmian-cylinder-measure-continued-fraction-sandwich}
Assume \(\beta=\alpha\), and let \(p_n/q_n\) be the continued-fraction
convergents of \(\alpha\), with the usual denominator convention
\(q_0=1\). If \(q_n\le m<q_{n+1}\), then every
\(u\in S_m(\alpha):=\supp(\mu_m)\) satisfies
\[
\frac{1}{q_{n+1}+q_n}<\mu_m(u)<\frac{2}{q_n}.
\]
\end{proposition}

\begin{proof}
For the Sturmian partition, the nonempty length-\(m\) cylinders are exactly the
gaps between the ordered points
\[
\{(1-j)\alpha\bmod 1:0\le j\le m\}.
\]
The displayed lower and upper estimates are the standard continued-fraction
gap bounds obtained from the three-distance theorem for rotations; see
Slater's gap theorem \cite{Slater1967GapsSteps} and the classical treatment in
\cite[Chapter~2]{KuipersNiederreiter1974}. These gap lengths are precisely the
Lebesgue cylinder masses \(\mu_m(u)\).
\end{proof}

\section{Sampling estimates}

\subsection{Discrepancy bounds}

We first record the elementary estimate which passes from star discrepancy to
histogram error.  The only point to note is the number of interval atoms in the
rotation partition.

\begin{lemma}[Interval partitions from star discrepancy]\label{lem:interval-partition-star-discrepancy}
Let \(\nu_N:=N^{-1}\sum_{t=0}^{N-1}\delta_{x_0+t\alpha}\) and let
\(\lambda\) be Lebesgue measure on \(\TT\).  If
\(\mathcal P=\{I_1,\dots,I_r\}\) is a finite partition of \(\TT\) into
half-open intervals, then
\[
\TV\bigl((\nu_N(I))_{I\in\mathcal P},(\lambda(I))_{I\in\mathcal P}\bigr)
\le r\,D_N^*(\alpha;x_0).
\]
\end{lemma}

\begin{proof}
For every half-open interval \(I\subset\TT\), possibly wrapping around the
origin,
\[
|\nu_N(I)-\lambda(I)|\le 2D_N^*(\alpha;x_0),
\]
because \(I\) is the difference, modulo complements if necessary, of two
anchored intervals \([0,a)\).  Therefore
\[
\TV\bigl((\nu_N(I))_{I\in\mathcal P},(\lambda(I))_{I\in\mathcal P}\bigr)
=\frac12\sum_{I\in\mathcal P}|\nu_N(I)-\lambda(I)|
\le rD_N^*(\alpha;x_0).
\]
\end{proof}

\begin{theorem}[Discrepancy control of microstate and folded laws]\label{thm:tv-certificate-hist}
For every irrational \(\alpha\), every \(\beta\in(0,1)\), every
\(x_0\in\TT\), and every \(m,N\ge1\),
\[
\TV(\hat\mu_{m,N},\mu_m)
\le 2m\,D_N^*(\alpha;x_0),
\qquad
\TV(\hat\pi_{m,N},\pi_m^{\mathrm{fold}})
\le 2m\,D_N^*(\alpha;x_0).
\]
In the Sturmian slice \(\beta=\alpha\), these improve to
\[
\TV(\hat\mu_{m,N},\mu_m)
\le (m+1)D_N^*(\alpha;x_0),
\qquad
\TV(\hat\pi_{m,N},\pi_m^{\mathrm{fold}})
\le (m+1)D_N^*(\alpha;x_0).
\]
\end{theorem}

\begin{proof}
The length-\(m\) coding map
\[
x\longmapsto \omega_m(x)
=\bigl(\mathbf 1_{[0,\beta)}(x+j\alpha)\bigr)_{j=0}^{m-1}
\]
can change only when \(x\) crosses one of the boundary points
\[
\Gamma_m(\alpha,\beta)
:=\{-j\alpha,\ \beta-j\alpha \pmod 1:\ 0\le j<m\}.
\]
After deleting repetitions and arranging these points cyclically, they cut
\(\TT\) into a half-open interval partition \(\mathcal P_m\) with
\(|\mathcal P_m|\le 2m\).  The convention \([0,\beta)\) makes
\(\omega_m\) constant on each atom of this partition, so the microstate laws
\(\hat\mu_{m,N}\) and \(\mu_m\) are pushforwards of the two partition laws
\((\nu_N(I))_{I\in\mathcal P_m}\) and \((\lambda(I))_{I\in\mathcal P_m}\).
Total variation cannot increase under a deterministic pushforward.  Hence
Lemma~\ref{lem:interval-partition-star-discrepancy} gives
\[
\TV(\hat\mu_{m,N},\mu_m)
\le |\mathcal P_m|D_N^*(\alpha;x_0)
\le 2mD_N^*(\alpha;x_0).
\]
Since \(\hat\pi_{m,N}=(\Fold_m)_*\hat\mu_{m,N}\) and
\(\pi_m^{\mathrm{fold}}=(\Fold_m)_*\mu_m\), another application of
monotonicity under deterministic pushforward gives the folded estimate.

If \(\beta=\alpha\), then
\[
\Gamma_m(\alpha,\alpha)
=\{-j\alpha,\ \alpha-j\alpha \pmod 1:\ 0\le j<m\}
=\{-(j-1)\alpha\pmod 1:\ 0\le j\le m\}.
\]
These \(m+1\) points are distinct because \(\alpha\) is irrational.  Thus the
same argument applies with \(|\mathcal P_m|=m+1\), proving the sharpened
Sturmian bounds.
\end{proof}

\begin{remark}
Thus any arithmetic or numerical bound on \(D_N^*(\alpha;x_0)\) gives the same
order of control for the folded histogram.  No independence approximation is
being used.
\end{remark}

\subsection{Cross-resolution residuals}

Let $r_{m+1\to m}:\Omega_{m+1}\to\Omega_m$ denote prefix truncation,
$r_{m+1\to m}(u_1\cdots u_{m+1})=u_1\cdots u_m$, and let
$\rho_{m+1\to m}:X_{m+1}\to X_m$ be the analogous truncation on folded words.
The following residual measures the failure of folding and truncation to
commute between adjacent window lengths:
Define the projective residual
\[
E_{m,N}
:=
\TV\!\left(\hat\pi_{m,N},(\rho_{m+1\to m})_*\hat\pi_{m+1,N}\right).
\]

\begin{theorem}[Boundary criterion for the canonical fold]\label{prop:multiscale-residual-by-boundary}
For \(u=u_1\cdots u_{m+1}\in\Omega_{m+1}\), let
\(a:=u_1\cdots u_m\). Then
\[
\rho_{m+1\to m}\Fold_{m+1}(u)\ne \Fold_m(a)
\]
if and only if
\[
u_{m+1}=1
\quad\text{and}\quad
N_m(a)\ge F_{m+1}.
\]
Consequently, with
\[
B_m:=\{u\in\Omega_{m+1}:u_{m+1}=1,\ N_m(u_1\cdots u_m)\ge F_{m+1}\},
\]
one has
\[
E_{m,N}\le \hat\mu_{m+1,N}(B_m).
\]
\end{theorem}

\begin{proof}
Write \(n:=N_m(a)\). If \(u_{m+1}=0\), then
\(N_{m+1}(u)=n\), so the first \(m\) Zeckendorf digits are the same at levels
\(m\) and \(m+1\).

Now suppose \(u_{m+1}=1\). If \(n<F_{m+1}\), the Zeckendorf expansion of \(n\)
uses no digit in positions \(m\) or \(m+1\). Adding \(F_{m+2}\) therefore
places an isolated \(1\) in position \(m+1\) and leaves the first \(m\) digits
unchanged.

It remains to show that the first \(m\) digits cannot remain unchanged when
\(n\ge F_{m+1}\). Write \(p:=N_m(\Fold_m(a))\). By
Proposition~\ref{prop:canonical-fold-basic}, \(n=p+\varepsilon F_{m+2}\) with
\(\varepsilon\in\{0,1\}\). Also \(N_{m+1}(u)<F_{m+4}\), so no Zeckendorf digit
beyond position \(m+2\) can occur. If \(F_{m+1}\le n<F_{m+2}\), then
\(\varepsilon=0\), \(p=n\), and the Zeckendorf expansion of \(p\) has digit
\(1\) in position \(m\). Keeping the first \(m\) digits after adding
\(F_{m+2}\) would force the digit in position \(m+1\) also to be \(1\), giving
an illegal adjacent pair. No alternative choice of higher digits has the same
value, because the only representation of
\[
p+F_{m+2}=p+bF_{m+2}+cF_{m+3},
\qquad b,c\in\{0,1\},
\]
with the first \(m\) digits fixed is \(b=1,c=0\).

If \(n\ge F_{m+2}\), then \(\varepsilon=1\). Were the first \(m\) digits
unchanged after adding \(F_{m+2}\), the higher digits would have to satisfy
\[
2F_{m+2}=bF_{m+2}+cF_{m+3},
\qquad b,c\in\{0,1\},
\]
which is impossible. Hence the first \(m\) digits change exactly in the stated
case.

For deterministic maps \(A(u)=\Fold_m(r_{m+1\to m}u)\) and
\(B(u)=\rho_{m+1\to m}\Fold_{m+1}(u)\), the total variation distance between
the empirical laws of \(A(u)\) and \(B(u)\) is bounded by the empirical mass of
\(\{A\ne B\}\). The criterion identifies this set with \(B_m\).
\end{proof}

\begin{proposition}[Residuals relative to the Parry measure]\label{prop:multiscale-residual-parry-bound}
Let \(\pi_m\) be the \(m\)-block Parry law on \(X_m\). Since the Parry family
is projectively consistent,
\[
\pi_m=(\rho_{m+1\to m})_*\pi_{m+1},
\]
one has
\[
E_{m,N}
\le
\TV(\hat\pi_{m,N},\pi_m)+\TV(\hat\pi_{m+1,N},\pi_{m+1}).
\]
\end{proposition}

\begin{proof}
Using the projective consistency of the Parry family and the triangle
inequality,
\[
\begin{aligned}
E_{m,N}
&=
\TV\!\left(\hat\pi_{m,N},(\rho_{m+1\to m})_*\hat\pi_{m+1,N}\right)\\
&\le
\TV(\hat\pi_{m,N},\pi_m)
+
\TV\!\left(\pi_m,(\rho_{m+1\to m})_*\hat\pi_{m+1,N}\right)\\
&=
\TV(\hat\pi_{m,N},\pi_m)
+
\TV\!\left((\rho_{m+1\to m})_*\pi_{m+1},
(\rho_{m+1\to m})_*\hat\pi_{m+1,N}\right).
\end{aligned}
\]
Total variation does not increase under deterministic pushforward, so the last
term is at most \(\TV(\hat\pi_{m+1,N},\pi_{m+1})\).
\end{proof}

\begin{corollary}[Residual comparison bound]
\label{cor:multiscale-residual-audit-bound}
For every irrational \(\alpha\), every \(\beta\in(0,1)\), and every
\(m,N\ge1\),
\[
E_{m,N}
\le
(4m+2)D_N^*(\alpha;x_0)
+
\TV(\pi_m^{\mathrm{fold}},\pi_m)
+
\TV(\pi_{m+1}^{\mathrm{fold}},\pi_{m+1}).
\]
In the Sturmian slice \(\beta=\alpha\), one may replace \(4m+2\) by
\(2m+3\).
\end{corollary}

\begin{proof}
By Proposition~\ref{prop:multiscale-residual-parry-bound},
\[
E_{m,N}
\le
\TV(\hat\pi_{m,N},\pi_m)+\TV(\hat\pi_{m+1,N},\pi_{m+1}).
\]
Apply the triangle inequality at levels \(m\) and \(m+1\):
\[
\TV(\hat\pi_{\ell,N},\pi_\ell)
\le
\TV(\hat\pi_{\ell,N},\pi_\ell^{\mathrm{fold}})
+
\TV(\pi_\ell^{\mathrm{fold}},\pi_\ell),
\qquad \ell\in\{m,m+1\}.
\]
Now use Theorem~\ref{thm:tv-certificate-hist}. In the general window this gives
\[
\TV(\hat\pi_{m,N},\pi_m^{\mathrm{fold}})
+
\TV(\hat\pi_{m+1,N},\pi_{m+1}^{\mathrm{fold}})
\le
(2m+2m+2)D_N^*(\alpha;x_0),
\]
while in the Sturmian slice it gives
\[
\TV(\hat\pi_{m,N},\pi_m^{\mathrm{fold}})
+
\TV(\hat\pi_{m+1,N},\pi_{m+1}^{\mathrm{fold}})
\le
((m+1)+(m+2))D_N^*(\alpha;x_0).
\]
Substituting these two estimates proves the corollary.
\end{proof}
\section{Relative-entropy bounds}

\subsection{From total variation to relative entropy}

To pass from total variation to relative entropy one needs a lower bound on the
smallest atom of the target distribution.

\begin{lemma}[Relative entropy from total variation under strict positivity]\label{lem:kl-from-tv-qmin}
Let \(p\) and \(q\) be probability laws on a finite set \(E\), with
\(\supp(p)\subset\supp(q)\), and put
\[
q_{\min}:=\min_{x\in\supp(q)}q(x)>0.
\]
Then
\[
\KL(p\|q)\le \frac{4}{q_{\min}}\TV(p,q)^2.
\]
\end{lemma}

\begin{proof}
The standard inequality \(\KL(p\|q)\le \chi^2(p,q)\) gives
\[
\KL(p\|q)
\le \sum_{x\in\supp(q)}\frac{(p(x)-q(x))^2}{q(x)}
\le \frac{1}{q_{\min}}\sum_{x\in E}|p(x)-q(x)|^2.
\]
Since \(\sum_x |p(x)-q(x)|=2\TV(p,q)\), the last expression is at most
\(4q_{\min}^{-1}\TV(p,q)^2\).
\end{proof}

\subsection{Lower bounds on the smallest rotation atom}

\begin{definition}[Diophantine lower bound]\label{def:diophantine-lower-bound}
For an irrational \(\alpha\), set
\[
c(\alpha):=\inf_{k\ge1} k\|k\alpha\|,
\]
where \(\|x\|\) denotes distance to the nearest integer. Thus
\(\alpha\) is badly approximable if and only if \(c(\alpha)>0\).
\end{definition}

\begin{definition}[General window partition scale]\label{def:general-window-partition-scale}
For \(m\ge1\), let
\[
\Delta_m(\alpha,\beta)
:=\{-j\alpha,\ \beta-j\alpha \pmod 1:\ 0\le j<m\}\subset\TT .
\]
Let \(\mathcal P_m(\alpha,\beta)\) be the half-open circle partition whose
endpoints are the distinct points of \(\Delta_m(\alpha,\beta)\), taken in
cyclic order and oriented counterclockwise.  Define
\[
\lambda_m(\alpha,\beta)
:=
\min_{I\in\mathcal P_m(\alpha,\beta)}\leb(I).
\]
\end{definition}

\begin{proposition}[General window atom separation]\label{prop:general-window-atom-separation}
For every irrational \(\alpha\), every \(\beta\in(0,1)\), and every
\(m\ge1\), each positive atom of \(\mu_m\) and of
\(\pi_m^{\mathrm{fold}}\) has mass at least
\(\lambda_m(\alpha,\beta)\).  Moreover
\[
\lambda_m(\alpha,\beta)
=
\min\Bigl(
\{\|k\alpha\|:1\le k\le m-1\}
\cup
\{\|\beta+r\alpha\|:\ |r|\le m-1,\ \beta+r\alpha\notin\mathbb Z\}
\Bigr),
\]
with the first set omitted when \(m=1\).  Consequently, if
\(c(\alpha)>0\) and
\[
s_m(\alpha,\beta)
:=
\min_{\substack{|r|\le m-1\\ \beta+r\alpha\notin\mathbb Z}}
\|\beta+r\alpha\|,
\]
then
\[
\lambda_m(\alpha,\beta)\ge
\min\left\{\frac{c(\alpha)}{m},\,s_m(\alpha,\beta)\right\}.
\]
If, in addition,
\[
c_\beta(\alpha):=
\inf_{\substack{r\in\mathbb Z\\ \beta+r\alpha\notin\mathbb Z}}
(|r|+1)\|\beta+r\alpha\|>0,
\]
then
\[
\lambda_m(\alpha,\beta)
\ge
\frac{\min\{c(\alpha),c_\beta(\alpha)\}}{m}.
\]
\end{proposition}

\begin{proof}
The discontinuity set of the word map \(x\mapsto\omega_m(x)\) is contained in
\(\Delta_m(\alpha,\beta)\).  On each half-open atom of
\(\mathcal P_m(\alpha,\beta)\) the word is constant.  The half-open convention
is compatible with the coding interval \([0,\beta)\): at a point
\(-j\alpha\) the \(j\)-th coordinate takes the right-hand value \(1\), while at
\(\beta-j\alpha\) it takes the right-hand value \(0\).  Thus every observed
boundary word also occurs on the adjacent half-open atom of positive length.
Therefore each cylinder \(C_m(a)\) is a union of atoms of
\(\mathcal P_m(\alpha,\beta)\), and every positive cylinder mass
\(\mu_m(a)\) is at least \(\lambda_m(\alpha,\beta)\).  A positive folded atom
is a sum of positive cylinder masses over one fiber of \(\Fold_m\), so it is
also at least \(\lambda_m(\alpha,\beta)\).

The minimum atom length of a finite circle partition is the minimum positive
circular distance between two distinct endpoints.  Two endpoints from the same
family \(\{-j\alpha\}\) or \(\{\beta-j\alpha\}\) differ by \(k\alpha\), with
\(1\le |k|\le m-1\).  Endpoints from different families differ by
\(\beta+r\alpha\), with \(|r|\le m-1\).  Removing the zero cross-differences
coming from coincident endpoints gives the displayed formula for
\(\lambda_m(\alpha,\beta)\).

For \(1\le k\le m-1\),
\[
\|k\alpha\|\ge \frac{c(\alpha)}{k}\ge \frac{c(\alpha)}{m}.
\]
The cross-family distances are bounded below by \(s_m(\alpha,\beta)\), which
proves the finite-scale bound.  If \(c_\beta(\alpha)>0\), then for
\(|r|\le m-1\),
\[
\|\beta+r\alpha\|
\ge \frac{c_\beta(\alpha)}{|r|+1}
\ge \frac{c_\beta(\alpha)}{m}.
\]
Combining this with the same-family estimate gives the final inequality.
\end{proof}

\begin{proposition}[Finite-scale lower atom bound]\label{prop:qmin-lowerbound-markov}
In the Sturmian slice \(\beta=\alpha\), write \(q_{\min}(m)\) for the least
positive atom of \(\mu_m\). If \(\alpha\) has Diophantine constant
\(c(\alpha)\), then
\[
q_{\min}(m)\ge \frac{c(\alpha)}{m}.
\]
For \(\alpha=\varphi^{-1}\), the all-\(k\) constant \(c(\alpha)=\varphi^{-2}\)
is valid, and therefore
\[
q_{\min}(m)\ge \frac{\varphi^{-2}}{m}.
\]
\end{proposition}

\begin{proof}
In the Sturmian slice, the positive atoms of \(\mu_m\) are the gaps between the
points \(\{(1-j)\alpha:0\le j\le m\}\) on the circle. Every such gap has length
at least
\[
\min_{1\le k\le m}\|k\alpha\|
\ge \frac{c(\alpha)}{m},
\]
which proves the first claim.

For \(\alpha=\varphi^{-1}\), choose \(n\ge2\) with \(F_n\le k<F_{n+1}\). The
best-approximation property of continued-fraction denominators gives
\(\|k\alpha\|\ge \|F_n\alpha\|=\varphi^{-n}\). Hence
\[
k\|k\alpha\|\ge F_n\varphi^{-n}\ge \varphi^{-2},
\]
where the last inequality uses \(F_n\ge\varphi^{n-2}\). This proves the stated
uniform choice of \(c(\alpha)\).
\end{proof}

\subsection{Microstate and folded relative-entropy bounds}

\begin{theorem}[General-window relative-entropy certificate]\label{thm:general-window-relative-entropy-certificate}
For every irrational \(\alpha\), every \(\beta\in(0,1)\), and every
\(m,N\ge1\),
\[
\KL(\hat\mu_{m,N}\|\mu_m)
\le
\frac{16m^2}{\lambda_m(\alpha,\beta)}
D_N^*(\alpha;x_0)^2
\]
and
\[
\KL(\hat\pi_{m,N}\|\pi_m^{\mathrm{fold}})
\le
\frac{16m^2}{\lambda_m(\alpha,\beta)}
D_N^*(\alpha;x_0)^2 .
\]
Consequently,
\[
\KL(\hat\mu_{m,N}\|\mu_m),\
\KL(\hat\pi_{m,N}\|\pi_m^{\mathrm{fold}})
\le
\frac{16m^2}
{\min\{c(\alpha)/m,s_m(\alpha,\beta)\}}
D_N^*(\alpha;x_0)^2
\]
whenever \(c(\alpha)>0\) and the finite-scale window separation in
Proposition~\ref{prop:general-window-atom-separation} is available.  If
\(c(\alpha)>0\) and \(c_\beta(\alpha)>0\), then both relative entropies are at
most
\[
\frac{16m^3}{\min\{c(\alpha),c_\beta(\alpha)\}}
D_N^*(\alpha;x_0)^2 .
\]
\end{theorem}

\begin{proof}
By Proposition~\ref{prop:general-window-atom-separation}, the least positive
atom of \(\mu_m\) and the least positive atom of
\(\pi_m^{\mathrm{fold}}\) are both at least
\(\lambda_m(\alpha,\beta)\).  The same proposition also shows that the
half-open coding convention puts every empirical word, including a boundary
word, inside a cylinder of positive Lebesgue mass.  Hence
\(\supp(\hat\mu_{m,N})\subset\supp(\mu_m)\), and after applying the
deterministic fold,
\(\supp(\hat\pi_{m,N})\subset\supp(\pi_m^{\mathrm{fold}})\).

Apply Lemma~\ref{lem:kl-from-tv-qmin} to the microstate law and to the folded
law, and then use the general-window part of
Theorem~\ref{thm:tv-certificate-hist}:
\[
\TV(\hat\mu_{m,N},\mu_m),\
\TV(\hat\pi_{m,N},\pi_m^{\mathrm{fold}})
\le 2mD_N^*(\alpha;x_0).
\]
This gives the two displayed \(\lambda_m(\alpha,\beta)\)-bounds.  The
remaining estimates are obtained by substituting the separation lower bounds
from Proposition~\ref{prop:general-window-atom-separation}.
\end{proof}

The next two corollaries keep the sharper \((m+1)\)-atom Sturmian constant and
the explicit bad-approximation lower bound used later in the Parry comparison.

\begin{corollary}[Golden-slope microstate improvement]\label{cor:rotation-microstate-kl-certificate}
Assume \(\beta=\alpha=\varphi^{-1}\).  Then
\[
\KL(\hat\mu_{m,N}\|\mu_m)
\le
4\varphi^2\,m(m+1)^2 D_N^*(\alpha;x_0)^2.
\]
\end{corollary}

\begin{remark}
The general badly-approximable certificate and the \(q_{\min}(m)\)-form appear
in \cite{MaZhang2026FoldedHistogramsSIADS}, Corollary~4.4; the increment here
is the replacement of the golden Markov constant by the all-\(k\) lower bound
\(q_{\min}(m)\ge \varphi^{-2}/m\) from
Proposition~\ref{prop:qmin-lowerbound-markov}.
\end{remark}

\begin{proof}
Lemma~\ref{lem:kl-from-tv-qmin}, the Sturmian total-variation estimate in
Theorem~\ref{thm:tv-certificate-hist}, and
Proposition~\ref{prop:qmin-lowerbound-markov} give
\[
\KL(\hat\mu_{m,N}\|\mu_m)
\le
\frac{4}{q_{\min}(m)}(m+1)^2D_N^*(\alpha;x_0)^2
\le
4\varphi^2\,m(m+1)^2D_N^*(\alpha;x_0)^2.
\]
\end{proof}

\begin{lemma}[KL monotonicity under pushforward]\label{lem:kl-monotone-pushforward}
Let $F:S\to T$ be a map between finite sets and let $p,q$ be probability
measures on $S$ with $p\ll q$.  Then
\[
\KL(F_*p\|F_*q)\le \KL(p\|q).
\]
\end{lemma}

\begin{proof}
This is the data-processing inequality for relative entropy under deterministic
coarse graining \cite{CoverThomas2006}.
\end{proof}

\begin{corollary}[Folded relative-entropy bound]\label{cor:rotation-folded-kl-certificate}
Assume \(\beta=\alpha\) and let \(\alpha\) be badly approximable. Then
\[
\KL(\hat\pi_{m,N}\|\pi_m^{\mathrm{fold}})
\le
\frac{4(m+1)^2}{q_{\min}(m)}D_N^*(\alpha;x_0)^2.
\]
For the golden slope,
\[
\KL(\hat\pi_{m,N}\|\pi_m^{\mathrm{fold}})
\le
4\varphi^2\,m(m+1)^2 D_N^*(\alpha;x_0)^2.
\]
\end{corollary}

\begin{proof}
Since \(\hat\pi_{m,N}=(\Fold_m)_*\hat\mu_{m,N}\) and
\(\pi_m^{\mathrm{fold}}=(\Fold_m)_*\mu_m\), data processing
(Lemma~\ref{lem:kl-monotone-pushforward}) gives
\[
\KL(\hat\pi_{m,N}\|\pi_m^{\mathrm{fold}})
\le
\KL(\hat\mu_{m,N}\|\mu_m).
\]
The \(q_{\min}(m)\)-bound follows from
\cite{MaZhang2026FoldedHistogramsSIADS}, Corollary~4.4, and the golden-slope
bound follows from Corollary~\ref{cor:rotation-microstate-kl-certificate}.
\end{proof}

\subsection{No collisions on the Sturmian support}

In the canonical Sturmian slice the fold does not identify distinct words
which are realized by the rotation.  The proof is arithmetic: as one moves
around the Sturmian cylinder partition, the Fibonacci values of the factors
remain distinct modulo \(F_{m+2}\).

\begin{theorem}[No collisions on the Sturmian language]\label{thm:sturmian-no-collision-rigidity}
Assume \(\beta=\alpha\). For every \(m\ge1\), the restriction
\[
B_m:=\Fold_m|_{S_m(\alpha)}:S_m(\alpha)\longrightarrow X_m
\]
is injective. Equivalently, if \(u,v\in S_m(\alpha)\) and \(u\ne v\), then
\[
\Fold_m(u)\ne \Fold_m(v).
\]
\end{theorem}

\begin{proof}
Fix distinct words \(u,v\in S_m(\alpha)\), and write
\[
d_j:=u_j-v_j\in\{-1,0,1\},
\qquad
\Delta:=N_m(u)-N_m(v)=\sum_{j=1}^m d_jF_{j+1}.
\]
The Sturmian language is balanced
\cite{MorseHedlund1940,LindMarcus1995}: any two factors of the same length
differ in their number of \(1\)'s by at most \(1\). We claim that the nonzero
entries of \(d_1,\dots,d_m\) therefore alternate in sign.

Indeed, let \(i<j\) be consecutive mismatch positions. If \(d_i=d_j=1\), then
the equal-length subwords \(u_i\cdots u_j\) and \(v_i\cdots v_j\) are both
Sturmian factors, but their numbers of \(1\)'s differ by at least \(2\), since
the endpoints contribute \(2\) and there is no opposite mismatch between them.
This contradicts balance. The case \(d_i=d_j=-1\) is identical. Thus, if
\[
1\le i_1<\cdots<i_t\le m
\]
are the mismatch positions, then
\[
\Delta
=
\pm\bigl(F_{i_t+1}-F_{i_{t-1}+1}+\cdots+(-1)^{t-1}F_{i_1+1}\bigr),
\]
with strictly increasing positive terms inside the parentheses.

After exchanging \(u\) and \(v\) if necessary, we may suppose that the sign is
\(+\). For strictly increasing positive numbers \(a_1<\cdots<a_t\), the
truncated alternating sum
\[
a_t-a_{t-1}+\cdots+(-1)^{t-1}a_1
\]
lies in \((0,a_t)\); this follows by a backward induction on \(t\). Applying
this to \(a_s:=F_{i_s+1}\) gives
\[
0<\Delta<F_{i_t+1}\le F_{m+1}<F_{m+2}.
\]
Hence \(\Delta\not\equiv 0\pmod{F_{m+2}}\). By
Proposition~\ref{prop:fold-residue-characterization},
\(\Fold_m(u)\ne \Fold_m(v)\), as required.
\end{proof}

\begin{remark}[Dependence on the canonical slice]
\label{rem:sturmian-rigidity-special}
The injectivity proof uses only two Sturmian facts: the balancedness of the
canonical slice \(\beta=\alpha\), and the residue characterization of the
canonical fold. We do not claim a comparable residue-rigidity statement for
general windows \(\beta\ne\alpha\).
\end{remark}

\begin{corollary}[Exact relabeling on the realized Sturmian language]\label{cor:sturmian-exact-relabeling}
Assume \(\beta=\alpha\), and write
\[
Y_m(\alpha):=\Fold_m(S_m(\alpha))\subset X_m,
\qquad
B_m:=\Fold_m|_{S_m(\alpha)}.
\]
Then \(B_m:S_m(\alpha)\to Y_m(\alpha)\) is a bijection. For every probability
law \(\nu\) supported on \(S_m(\alpha)\), the pushforward \(B_{m*}\nu\) is an
exact relabeling of \(\nu\). In particular,
\[
\pi_m^{\mathrm{fold}}(B_m(u))=\mu_m(u)
\qquad (u\in S_m(\alpha)),
\]
and for every \(q\in[0,\infty]\),
\[
H_q(\pi_m^{\mathrm{fold}})=H_q(\mu_m).
\]
\end{corollary}

\begin{proof}
Theorem~\ref{thm:sturmian-no-collision-rigidity} shows that \(B_m\) is
injective, and surjectivity onto \(Y_m(\alpha)\) is tautological. If
\(\nu\in\Delta(S_m(\alpha))\), then \(B_m\) merely renames the atoms of
\(\nu\), so the masses are preserved one-for-one:
\[
(B_{m*}\nu)(B_m(u))=\nu(u)
\qquad (u\in S_m(\alpha)).
\]
Applying this to \(\nu=\mu_m\) gives the formula for
\(\pi_m^{\mathrm{fold}}=(\Fold_m)_*\mu_m\), and every R\'enyi entropy is
invariant under bijective relabeling.
\end{proof}

\subsection{Higher-block conjugacy on the Sturmian source}

\begin{theorem}[Higher-block conjugacy on the Sturmian source]
\label{thm:sequence-level-transparency}
Assume \(\beta=\alpha\), and let \(X_\alpha\subset\{0,1\}^{\mathbb Z}\) be the
Sturmian subshift of slope \(\alpha\). For \(m\ge1\), let
\[
\Phi_m:X_\alpha\longrightarrow S_m(\alpha)^{\mathbb Z},
\qquad
(\Phi_m(s))_t:=s_t\cdots s_{t+m-1},
\]
be the usual \(m\)-block presentation, and let
\[
\Psi_m:X_\alpha\longrightarrow Y_m(\alpha)^{\mathbb Z},
\qquad
(\Psi_m(s))_t:=\Fold_m(s_t\cdots s_{t+m-1}),
\]
be the folded block process. Then
\[
\Psi_m=B_m^{\mathbb Z}\circ \Phi_m,
\]
where \(B_m^{\mathbb Z}\) is the coordinatewise relabeling induced by the
bijection \(B_m:S_m(\alpha)\to Y_m(\alpha)\). Consequently,
\(B_m^{\mathbb Z}\) is a topological conjugacy from the usual \(m\)-block
presentation \(\Phi_m(X_\alpha)\) onto the folded image \(\Psi_m(X_\alpha)\).
\end{theorem}

\begin{proof}
For every \(s\in X_\alpha\) and every \(t\in\mathbb Z\),
\[
(\Psi_m(s))_t
=
\Fold_m(s_t\cdots s_{t+m-1})
=
B_m\bigl((\Phi_m(s))_t\bigr),
\]
which is exactly the identity \(\Psi_m=B_m^{\mathbb Z}\circ\Phi_m\). By
Corollary~\ref{cor:sturmian-exact-relabeling}, \(B_m\) is a bijection on the
realized alphabet, so \(B_m^{\mathbb Z}\) and \((B_m^{-1})^{\mathbb Z}\) are
inverse one-block codes. They commute with the shift by construction, hence
define a topological conjugacy between the two subshifts.
\end{proof}
\noindent
For the next corollary, write
\[
d_m(x):=|\Fold_m^{-1}(x)|,\qquad x\in X_m,
\]
and define the fiber-uniform lift of a law \(\eta\in\Delta(X_m)\) by
\[
(Q_m\eta)(a):=\frac{\eta(\Fold_m(a))}{d_m(\Fold_m(a))},
\qquad a\in\Omega_m.
\]
These lifts are developed systematically in Appendix~\ref{app:fiber-lift}.

\begin{corollary}[Fiber degeneracy on the Sturmian slice]\label{cor:sturmian-pure-degeneracy-ledger}
Assume \(\beta=\alpha\). Let
\[
\mu_m^e:=Q_m\pi_m^{\mathrm{fold}},
\qquad
\hat\mu_{m,N}^e:=Q_m\hat\pi_{m,N}.
\]
Then
\[
\KL(\mu_m\|\mu_m^e)
=
\sum_{u\in S_m(\alpha)}\mu_m(u)\log d_m(B_m(u))
=
\mathbb E_{\pi_m^{\mathrm{fold}}}\big[\log d_m(Y)\big],
\]
and for every \(N\ge1\),
\[
\KL(\hat\mu_{m,N}\|\hat\mu_{m,N}^e)
=
\sum_{u\in S_m(\alpha)}\hat\mu_{m,N}(u)\log d_m(B_m(u))
=
\mathbb E_{\hat\pi_{m,N}}\big[\log d_m(Y)\big].
\]
\end{corollary}

\begin{proof}
By Corollary~\ref{cor:sturmian-exact-relabeling}, each realized fiber contains
exactly one realized word. Therefore, for \(u\in S_m(\alpha)\),
\[
\mu_m^e(u)
=
\frac{\pi_m^{\mathrm{fold}}(B_m(u))}{d_m(B_m(u))}
=
\frac{\mu_m(u)}{d_m(B_m(u))},
\]
and likewise
\[
\hat\mu_{m,N}^e(u)
=
\frac{\hat\pi_{m,N}(B_m(u))}{d_m(B_m(u))}
=
\frac{\hat\mu_{m,N}(u)}{d_m(B_m(u))},
\]
because \(\supp(\hat\mu_{m,N})\subset S_m(\alpha)\). Hence
\[
\KL(\mu_m\|\mu_m^e)
=
\sum_{u\in S_m(\alpha)}
\mu_m(u)\log\frac{\mu_m(u)}{\mu_m^e(u)}
=
\sum_{u\in S_m(\alpha)}\mu_m(u)\log d_m(B_m(u)).
\]
Pushing this sum forward under the bijection \(B_m\) gives the expectation
under \(\pi_m^{\mathrm{fold}}\). The empirical identity is identical.
\end{proof}
A microstate form of the same decomposition is recorded in
Appendix~\ref{app:fiber-lift}.  If \(\mu\) is a law on \(\Omega_m\),
\(\pi:=(\Fold_m)_*\mu\), and \(Q_m\eta\) is the fiber-uniform lift of a law
\(\eta\) on \(X_m\), then
\[
\KL(\mu\|Q_m\eta)=\KL(\pi\|\eta)+\KL(\mu\|Q_m\pi),
\]
so the observable divergence and the within-fiber term split exactly.  This
identity is independent of the Parry comparison.

\section{Comparison with the Parry measure}

The Parry measure is the measure of maximal entropy for the golden-mean shift.
We compare folded rotation laws with its \(m\)-block marginal.  Since the
Sturmian support has only \(m+1\) words, the comparison is controlled both by
the source weights and by the placement of their support in \(X_m\).  The
golden-mean shift has the useful additional feature that Parry cylinder weights
depend only on the two endpoints of the word.

For \(ij\in\{0,1\}^2\), write
\[
X_m^{ij}:=\{u\in X_m:u_1=i,\ u_m=j\}.
\]
If \(\nu\in\Delta(X_m)\), define the endpoint-class masses
\[
\eta_{ij}(\nu):=\nu(X_m^{ij}).
\]

\begin{corollary}[Decomposition relative to the Parry measure]\label{cor:tv-decomposition-parry}
Let \(\mu\in\Delta(\Omega_m)\), let \(\pi=(\Fold_m)_*\mu\), and let
\(\mu^e:=Q_m\pi\). Then
\[
\KL(\mu\|Q_m\pi_m)
=
\KL(\mu\|\mu^e)+\KL(\pi\|\pi_m).
\]
In particular, for \(\mu=\mu_m\) or \(\mu=\hat\mu_{m,N}\), the divergence from
the Parry lift splits exactly into the within-fiber term and the folded Parry
comparison.
\end{corollary}

\begin{proof}
For \(a\in\Omega_m\), write \(x_a:=\Fold_m(a)\). Then
\[
\mu^e(a)=\frac{\pi(x_a)}{d_m(x_a)},
\qquad
(Q_m\pi_m)(a)=\frac{\pi_m(x_a)}{d_m(x_a)}.
\]
Therefore
\[
\begin{aligned}
\KL(\mu\|Q_m\pi_m)-\KL(\mu\|\mu^e)
&=
\sum_{a\in\Omega_m}\mu(a)\log\frac{\mu^e(a)}{(Q_m\pi_m)(a)}\\
&=
\sum_{a\in\Omega_m}\mu(a)\log\frac{\pi(x_a)}{\pi_m(x_a)}\\
&=
\sum_{x\in X_m}\pi(x)\log\frac{\pi(x)}{\pi_m(x)}
=
\KL(\pi\|\pi_m).
\end{aligned}
\]
\end{proof}

\subsection{Parry identity and support minima}

\begin{definition}[R\'enyi entropy and divergence]\label{def:renyi-entropy-divergence}
Let \(E\) be a finite set, let \(p\in\Delta(E)\), and let \(r\in\Delta(E)\)
with \(p\ll r\).

For \(q\in(0,\infty)\setminus\{1\}\), define the R\'enyi entropy
\[
H_q(p):=\frac{1}{1-q}\log\!\left(\sum_{x\in E}p(x)^q\right).
\]
The endpoint values are
\[
H_1(p):=-\sum_{x\in E}p(x)\log p(x),\qquad
H_0(p):=\log|\supp(p)|,\qquad
H_\infty(p):=-\log\max_{x\in E}p(x).
\]

For \(q\in(0,\infty)\setminus\{1\}\), define the R\'enyi divergence
\[
D_q(p\|r):=\frac{1}{q-1}\log\!\left(\sum_{x\in E}p(x)^q r(x)^{1-q}\right).
\]
The endpoint values are
\[
D_1(p\|r):=\KL(p\|r),\qquad
D_0(p\|r):=-\log r(\supp(p)),
\]
\[
D_\infty(p\|r):=
\log\max_{x\in\supp(p)}\frac{p(x)}{r(x)}.
\]
\end{definition}

\begin{theorem}[Exact divergence identity relative to the Parry measure]\label{thm:parry-kl-identity}
For every probability law \(\nu\in\Delta(X_m)\),
\[
\KL(\nu\|\pi_m)
=
m\log\varphi-\log\pi(0)-H_1(\nu)
+
\bigl(\eta_{11}(\nu)-\eta_{00}(\nu)\bigr)\log\varphi.
\]
\end{theorem}

\begin{proof}
By Proposition~\ref{prop:parry-endpoint-collapse},
\[
\log \pi_m(u)
=
\log\pi(0)-m\log\varphi
+
\mathbf{1}_{\{u_1=0\}}\log\varphi
-
\mathbf{1}_{\{u_m=1\}}\log\varphi.
\]
Hence
\[
\begin{aligned}
\KL(\nu\|\pi_m)
&=
\sum_{u\in X_m}\nu(u)\log\frac{\nu(u)}{\pi_m(u)}\\
&=
-H_1(\nu)-\log\pi(0)+m\log\varphi\\
&\quad
-\left(\sum_{u\in X_m}\nu(u)\mathbf{1}_{\{u_1=0\}}\right)\log\varphi
+
\left(\sum_{u\in X_m}\nu(u)\mathbf{1}_{\{u_m=1\}}\right)\log\varphi.
\end{aligned}
\]
Now
\[
\sum_{u\in X_m}\nu(u)\mathbf{1}_{\{u_1=0\}}
=
\eta_{00}(\nu)+\eta_{01}(\nu),
\]
while
\[
\sum_{u\in X_m}\nu(u)\mathbf{1}_{\{u_m=1\}}
=
\eta_{01}(\nu)+\eta_{11}(\nu).
\]
Subtracting these two expressions gives
\(\eta_{11}(\nu)-\eta_{00}(\nu)\), which proves the formula.
\end{proof}

\begin{remark}[Endpoint collapse]\label{rem:endpoint-collapse-specificity}
Theorem~\ref{thm:parry-kl-identity} depends essentially on
Proposition~\ref{prop:parry-endpoint-collapse}. For a general shift of finite
type, Parry cylinder weights are controlled by boundary states in an edge
presentation, but they need not reduce to the scalar endpoint statistic used
here.  The exact identity is therefore a special feature of the golden-mean
case.
\end{remark}

\begin{corollary}[No fold-entropy loss on the Sturmian slice]\label{cor:no-fold-entropy-loss}
Assume \(\beta=\alpha\). Then for every \(q\in[0,\infty]\),
\[
H_q(\pi_m^{\mathrm{fold}})=H_q(\mu_m).
\]
In particular,
\[
\KL(\pi_m^{\mathrm{fold}}\|\pi_m)
=
m\log\varphi-\log\pi(0)-H_1(\mu_m)
+
\bigl(\eta_{11}(\pi_m^{\mathrm{fold}})
-\eta_{00}(\pi_m^{\mathrm{fold}})\bigr)\log\varphi.
\]
\end{corollary}

\begin{proof}
The entropy identity is Corollary~\ref{cor:sturmian-exact-relabeling}. Apply
Theorem~\ref{thm:parry-kl-identity} with \(\nu=\pi_m^{\mathrm{fold}}\) and
substitute \(H_1(\pi_m^{\mathrm{fold}})=H_1(\mu_m)\).
\end{proof}

\begin{proposition}[Order-uniform R\'enyi window relative to the Parry measure]
\label{prop:parry-renyi-window}
For every \(\nu\in\Delta(X_m)\) and every \(q\in[0,\infty]\),
\[
m\log\varphi-\log\pi(0)-H_q(\nu)-\log\varphi
\le
D_q(\nu\|\pi_m)
\le
m\log\varphi-\log\pi(0)-H_q(\nu)+\log\varphi.
\]
\end{proposition}

\begin{proof}
Put
\[
c_m:=\pi(0)\varphi^{-m}.
\]
By Proposition~\ref{prop:parry-endpoint-collapse}, every admissible word
\(u\in X_m\) has
\[
\pi_m(u)=c_m r(u),
\qquad
r(u)\in\{\varphi,1,\varphi^{-1}\}.
\]

First suppose \(q\in(0,\infty)\setminus\{1\}\). Then
\[
\begin{aligned}
D_q(\nu\|\pi_m)
&=
\frac{1}{q-1}
\log\!\left(
\sum_{u\in X_m}\nu(u)^q(c_mr(u))^{1-q}
\right)\\
&=
-\log c_m
+
\frac{1}{q-1}
\log\!\left(\sum_{u\in X_m}\nu(u)^q r(u)^{1-q}\right).
\end{aligned}
\]
Since
\[
H_q(\nu)=\frac{1}{1-q}\log\!\left(\sum_{u\in X_m}\nu(u)^q\right),
\]
subtracting \(-\log c_m-H_q(\nu)\) leaves
\[
\frac{1}{q-1}
\log
\frac{\sum_{u\in X_m}\nu(u)^q r(u)^{1-q}}
{\sum_{u\in X_m}\nu(u)^q}.
\]
Because \(r(u)\in\{\varphi,1,\varphi^{-1}\}\), the quotient inside the
logarithm lies between \(\varphi^{-(q-1)}\) and \(\varphi^{q-1}\). Hence the
last term lies in \([-\log\varphi,\log\varphi]\), giving the stated bounds.

For \(q=1\), Theorem~\ref{thm:parry-kl-identity} shows that
\[
D_1(\nu\|\pi_m)
=
m\log\varphi-\log\pi(0)-H_1(\nu)
+
\bigl(\eta_{11}(\nu)-\eta_{00}(\nu)\bigr)\log\varphi,
\]
and \(\eta_{11}(\nu)-\eta_{00}(\nu)\in[-1,1]\).

For \(q=0\), let \(S:=\supp(\nu)\). Every word in \(S\) has Parry mass between
\(c_m\varphi^{-1}\) and \(c_m\varphi\), so
\[
|S|c_m\varphi^{-1}
\le
\pi_m(S)
\le
|S|c_m\varphi.
\]
Since \(D_0(\nu\|\pi_m)=-\log\pi_m(S)\) and \(H_0(\nu)=\log|S|\), this is
exactly the required window.

For \(q=\infty\),
\[
D_\infty(\nu\|\pi_m)
=
\log\max_{u\in\supp(\nu)}\frac{\nu(u)}{\pi_m(u)}.
\]
Again \(\pi_m(u)\in[c_m\varphi^{-1},c_m\varphi]\), so
\[
\frac{\max_u \nu(u)}{c_m\varphi}
\le
\max_{u\in\supp(\nu)}\frac{\nu(u)}{\pi_m(u)}
\le
\frac{\max_u \nu(u)}{c_m\varphi^{-1}}.
\]
Taking logarithms and using \(H_\infty(\nu)=-\log\max_u\nu(u)\) proves the
endpoint case.
\end{proof}

\begin{proposition}[Support conditioning for R\'enyi divergence]
\label{prop:renyi-support-conditioning}
Let \(q\in[0,\infty]\), let \(p,\rho\in\Delta(E)\) with \(p\ll\rho\), and let
\(S\subset E\) satisfy \(\supp(p)\subset S\subset\supp(\rho)\). Write
\[
\rho^S(x):=\frac{\rho(x)}{\rho(S)},\qquad x\in S.
\]
Then
\[
D_q(p\|\rho)=D_q(p\|\rho^S)-\log \rho(S).
\]
\end{proposition}

\begin{proof}
For \(q\in(0,\infty)\setminus\{1\}\),
\[
\begin{aligned}
D_q(p\|\rho)
&=
\frac{1}{q-1}\log\!\left(\sum_{x\in S}p(x)^q\rho(x)^{1-q}\right)\\
&=
\frac{1}{q-1}
\log\!\left(\rho(S)^{1-q}\sum_{x\in S}p(x)^q(\rho^S(x))^{1-q}\right)\\
&=
D_q(p\|\rho^S)-\log\rho(S).
\end{aligned}
\]
For \(q=1\),
\[
\KL(p\|\rho)
=
\sum_{x\in S}p(x)\log\frac{p(x)}{\rho^S(x)\rho(S)}
=
\KL(p\|\rho^S)-\log\rho(S).
\]
For \(q=0\), one has
\[
D_0(p\|\rho^S)=-\log \rho^S(\supp p)
=
-\log\frac{\rho(\supp p)}{\rho(S)},
\]
so
\[
D_0(p\|\rho^S)-\log\rho(S)=-\log\rho(\supp p)=D_0(p\|\rho).
\]
For \(q=\infty\),
\[
D_\infty(p\|\rho)
=
\log\max_{x\in\supp p}\frac{p(x)}{\rho^S(x)\rho(S)}
=
D_\infty(p\|\rho^S)-\log\rho(S).
\]
\end{proof}

\begin{lemma}[Endpoint class counts]\label{lem:endpoint-class-counts}
For \(m\ge2\),
\[
|X_m^{00}|=F_m,\qquad
|X_m^{01}|=|X_m^{10}|=F_{m-1},\qquad
|X_m^{11}|=F_{m-2}.
\]
\end{lemma}

\begin{proof}
For \(X_m^{00}\), deleting the two endpoint zeros gives a bijection
\[
X_m^{00}\longleftrightarrow X_{m-2},
\]
so \(|X_m^{00}|=|X_{m-2}|=F_m\).

For \(X_m^{11}\), admissibility forces the shape \(10w01\) with
\(w\in X_{m-4}\), and this gives a bijection
\[
X_m^{11}\longleftrightarrow X_{m-4}.
\]
Hence \(|X_m^{11}|=|X_{m-4}|=F_{m-2}\).

Reversal preserves \(X_m\) and swaps \(X_m^{01}\) with \(X_m^{10}\), so
\(|X_m^{01}|=|X_m^{10}|\). Since
\[
|X_m|=F_{m+2}
=
|X_m^{00}|+|X_m^{01}|+|X_m^{10}|+|X_m^{11}|,
\]
the remaining classes satisfy
\[
2|X_m^{01}|
=
F_{m+2}-F_m-F_{m-2}
=
2F_{m-1},
\]
which proves the formula.
\end{proof}

\begin{definition}[Parry support envelope]\label{def:parry-support-envelope}
For \(m\ge2\) and \(1\le k\le |X_m|=F_{m+2}\), define
\[
M_m(k):=
\max\{\pi_m(S): S\subset X_m,\ |S|\le k\}.
\]
\end{definition}

\begin{theorem}[Support-constrained minima]\label{thm:parry-support-barrier}
Let \(m\ge2\). Put
\[
c_{00}:=\pi(0)\varphi^{-(m-1)},\qquad
c_{01}:=\pi(0)\varphi^{-m},\qquad
c_{11}:=\pi(0)\varphi^{-(m+1)}.
\]
Then \(c_{00}>c_{01}>c_{11}\), and for \(1\le k\le F_{m+2}\),
\[
M_m(k)=
\begin{cases}
kc_{00}, & 1\le k\le F_m,\\[0.3em]
F_m c_{00}+(k-F_m)c_{01},
& F_m<k\le F_m+2F_{m-1},\\[0.3em]
F_m c_{00}+2F_{m-1}c_{01}
+(k-F_m-2F_{m-1})c_{11},
& F_m+2F_{m-1}<k\le F_{m+2}.
\end{cases}
\]
Moreover, if \(\nu\in\Delta(X_m)\) satisfies \(|\supp(\nu)|\le k\), then for
every \(q\in[0,\infty]\),
\[
D_q(\nu\|\pi_m)\ge -\log M_m(k).
\]
Equality holds if and only if \(\nu=\pi_m(\cdot\mid S)\) for some set
\(S\subset X_m\) with \(|S|\le k\) and \(\pi_m(S)=M_m(k)\).
\end{theorem}

\begin{proof}
By Proposition~\ref{prop:parry-endpoint-collapse}, every word in \(X_m^{00}\)
has mass \(c_{00}\), every word in \(X_m^{01}\cup X_m^{10}\) has mass
\(c_{01}\), and every word in \(X_m^{11}\) has mass \(c_{11}\). Since
\(c_{00}>c_{01}>c_{11}\), the maximal Parry mass of a \(k\)-point set is
obtained by filling endpoint classes in this decreasing order. Lemma
~\ref{lem:endpoint-class-counts} gives the available multiplicities in the
three classes, so the displayed piecewise formula for \(M_m(k)\) follows.

Now let \(S:=\supp(\nu)\). Proposition~\ref{prop:renyi-support-conditioning}
gives
\[
D_q(\nu\|\pi_m)=D_q(\nu\|\pi_m(\cdot\mid S))-\log\pi_m(S)\ge -\log\pi_m(S),
\]
because every R\'enyi divergence is nonnegative. Since \(|S|\le k\),
\(\pi_m(S)\le M_m(k)\), hence
\[
D_q(\nu\|\pi_m)\ge -\log M_m(k).
\]
Equality holds precisely when both inequalities are equalities, namely when
\(\pi_m(S)=M_m(k)\) and \(\nu=\pi_m(\cdot\mid S)\).
\end{proof}

\begin{corollary}[Complexity-divergence law]\label{cor:exact-complexity-divergence-law}
Let \(m\ge2\) and \(1\le k\le F_m\). If \(\nu\in\Delta(X_m)\) satisfies
\(|\supp(\nu)|\le k\), then for every \(q\in[0,\infty]\),
\[
D_q(\nu\|\pi_m)
\ge
(m-1)\log\varphi-\log\pi(0)-\log k.
\]
Equality holds if and only if \(|\supp(\nu)|=k\), the support of \(\nu\) is
contained in \(X_m^{00}\), and \(\nu\) is uniform on its support.
\end{corollary}

\begin{proof}
If \(k\le F_m\), then Theorem~\ref{thm:parry-support-barrier} gives
\[
M_m(k)=k\,\pi(0)\varphi^{-(m-1)}.
\]
Therefore
\[
-\log M_m(k)
=(m-1)\log\varphi-\log\pi(0)-\log k,
\]
which is the stated lower bound. Equality in
Theorem~\ref{thm:parry-support-barrier} means that \(\nu\) is the conditioned
Parry law on a maximizing \(k\)-point set. Since every maximizing set lies in
the constant-mass class \(X_m^{00}\), that conditioned law is uniform.
\end{proof}
\subsection{Optimal injective relabeling of the realized Sturmian law}

\begin{theorem}[Optimal injective relabeling of the realized Sturmian law]
\label{thm:optimal-injective-relabeling}
Assume \(\beta=\alpha\) and \(m\ge 6\). Let \(S_m(\alpha)\) be the realized
Sturmian language of length \(m\), let \(\mu_m\) be its factor law, and define
\[
\mathcal J_m
:=
\{\iota:S_m(\alpha)\hookrightarrow X_m:\ \iota\text{ injective}\},
\qquad
L_q(m)
:=
\inf_{\iota\in\mathcal J_m}D_q(\iota_*\mu_m\|\pi_m).
\]
Then for every \(q\in[0,\infty]\),
\[
L_q(m)
=
(m-1)\log\varphi-\log\pi(0)-H_q(\mu_m).
\]
Every injection whose image is contained in \(X_m^{00}\) attains the infimum.
Moreover, the canonical folded law satisfies
\[
0\le D_q(\pi_m^{\mathrm{fold}}\|\pi_m)-L_q(m)\le 2\log\varphi,
\qquad q\in[0,\infty],
\]
and for \(q=1\),
\[
\KL(\pi_m^{\mathrm{fold}}\|\pi_m)-L_1(m)
=
\bigl(1+\eta_{11}(\pi_m^{\mathrm{fold}})
-\eta_{00}(\pi_m^{\mathrm{fold}})\bigr)\log\varphi.
\]
\end{theorem}

\begin{proof}
Let
\[
c_m:=\pi(0)\varphi^{-(m-1)}.
\]
By Proposition~\ref{prop:parry-endpoint-collapse}, every word in \(X_m^{00}\)
has Parry mass \(c_m\), and no word in \(X_m\) has larger Parry mass. By
Lemma~\ref{lem:endpoint-class-counts},
\[
|X_m^{00}|=F_m.
\]
Since \(|S_m(\alpha)|=m+1\) and \(F_m\ge m+1\) for \(m\ge 6\), there exist
injections \(\iota\in\mathcal J_m\) with \(\iota(S_m(\alpha))\subset X_m^{00}\).

Fix such an injection and write \(S:=\iota(S_m(\alpha))\). Then \(|S|=m+1\),
\(\pi_m(S)=(m+1)c_m\), and the conditioned Parry law \(\pi_m^S\) is uniform on
\(S\). Because \(\iota_*\mu_m\) is a relabeling of \(\mu_m\), it has the same
R\'enyi entropies. Therefore
\[
D_q(\iota_*\mu_m\|\pi_m^S)=\log|S|-H_q(\mu_m)
\qquad (q\in[0,\infty]),
\]
and Proposition~\ref{prop:renyi-support-conditioning} gives
\[
D_q(\iota_*\mu_m\|\pi_m)
=
\log(m+1)-H_q(\mu_m)-\log((m+1)c_m)
=
-\log c_m-H_q(\mu_m).
\]
This proves the upper bound
\[
L_q(m)\le -\log c_m-H_q(\mu_m)
=(m-1)\log\varphi-\log\pi(0)-H_q(\mu_m).
\]

For the matching lower bound, fix an arbitrary \(\iota\in\mathcal J_m\), write
\(p_u:=\mu_m(u)\), and set \(\rho_u:=\pi_m(\iota(u))\). Since \(\rho_u\le c_m\)
for every \(u\in S_m(\alpha)\), the Shannon case satisfies
\[
D_1(\iota_*\mu_m\|\pi_m)
=
\sum_{u\in S_m(\alpha)}p_u\log\frac{p_u}{\rho_u}
\ge
\sum_{u\in S_m(\alpha)}p_u\log\frac{p_u}{c_m}
=
-H_1(\mu_m)-\log c_m.
\]
For \(q\in(0,\infty)\setminus\{1\}\), monotonicity of \(x\mapsto x^{1-q}\)
implies
\[
\frac{1}{q-1}\log\!\Bigl(\sum_{u\in S_m(\alpha)}p_u^q\rho_u^{1-q}\Bigr)
\ge
\frac{1}{q-1}\log\!\Bigl(c_m^{1-q}\sum_{u\in S_m(\alpha)}p_u^q\Bigr),
\]
because \(q>1\) gives a direct lower bound inside the logarithm while
\(q<1\) gives an upper bound inside the logarithm and a negative prefactor.
Hence
\[
D_q(\iota_*\mu_m\|\pi_m)\ge -\log c_m-H_q(\mu_m)
\qquad \bigl(q\in(0,\infty)\setminus\{1\}\bigr).
\]
For \(q=\infty\),
\[
D_\infty(\iota_*\mu_m\|\pi_m)
=
\log\max_{u\in S_m(\alpha)}\frac{p_u}{\rho_u}
\ge
\log\max_{u\in S_m(\alpha)}\frac{p_u}{c_m}
=
-\log c_m-H_\infty(\mu_m).
\]
Finally, since \(|S_m(\alpha)|=m+1\),
\[
\pi_m(\iota(S_m(\alpha)))\le (m+1)c_m=c_m e^{H_0(\mu_m)},
\]
and therefore
\[
D_0(\iota_*\mu_m\|\pi_m)
=
-\log \pi_m(\iota(S_m(\alpha)))
\ge
-\log c_m-H_0(\mu_m).
\]
The lower bound is now proved for every \(q\in[0,\infty]\), so the formula for
\(L_q(m)\) follows. Every injection into \(X_m^{00}\) attains equality by the
first part of the argument.

The left inequality
\[
0\le D_q(\pi_m^{\mathrm{fold}}\|\pi_m)-L_q(m)
\]
is immediate because \(\pi_m^{\mathrm{fold}}=(B_m)_*\mu_m\) with
\(B_m\in\mathcal J_m\) by Corollary~\ref{cor:sturmian-exact-relabeling}. The
upper bound follows from the R\'enyi invariance
\[
H_q(\pi_m^{\mathrm{fold}})=H_q(\mu_m)
\]
and Proposition~\ref{prop:parry-renyi-window}, which give
\[
D_q(\pi_m^{\mathrm{fold}}\|\pi_m)
\le
m\log\varphi-H_q(\mu_m)-\log\pi(0)+\log\varphi
=
L_q(m)+2\log\varphi.
\]
For \(q=1\), Corollary~\ref{cor:no-fold-entropy-loss} yields
\[
\KL(\pi_m^{\mathrm{fold}}\|\pi_m)
=
m\log\varphi-H_1(\mu_m)-\log\pi(0)
+\bigl(\eta_{11}(\pi_m^{\mathrm{fold}})
-\eta_{00}(\pi_m^{\mathrm{fold}})\bigr)\log\varphi,
\]
so subtracting the displayed formula for \(L_1(m)\) gives the exact identity.
\end{proof}

\begin{remark}[Injective relabeling and support budgets]
\label{rem:injective-relabeling-stronger}
Theorem~\ref{thm:optimal-injective-relabeling} does more than constrain the
image support size. It fixes the realized Sturmian weight vector \(\mu_m\) and
allows only injective placements of those \(m+1\) probabilities inside the
ambient language \(X_m\). The resulting optimum is therefore a coding
invariance statement for the realized source, not just a best \((m+1)\)-state
approximation problem over arbitrary laws on \(X_m\).  The finite exceptional
cases \(m<6\) play no role in the asymptotic statements and are absorbed into
the \(O_\alpha(1)\) constants.
\end{remark}

\subsection{Support budgets and bounded-type asymptotics}

\begin{proposition}[Bounded-type Sturmian R\'enyi entropy]\label{prop:bounded-type-sturmian-entropy}
Assume \(\beta=\alpha\), and suppose that \(\alpha\) has bounded partial
quotients, equivalently that \(\alpha\) is badly approximable in dimension one
\cite{Cassels1957,Schmidt1969BadlyApproximableSystems}. Then there exist
constants \(c_\alpha,C_\alpha>0\) such that for
every \(m\ge 1\) and every \(u\in S_m(\alpha)\),
\[
\frac{c_\alpha}{m}\le \mu_m(u)\le \frac{C_\alpha}{m}.
\]
Consequently,
\[
H_q(\mu_m)=\log m+O_\alpha(1)
\qquad (q\in[0,\infty]),
\]
with the \(O_\alpha(1)\) term uniform in \(q\).
\end{proposition}

\begin{proof}
Let \(p_n/q_n\) be the convergents of \(\alpha\), and choose \(n\) so that
\(q_n\le m<q_{n+1}\). By
Theorem~\ref{thm:spg-sturmian-cylinder-measure-continued-fraction-sandwich},
every length-\(m\) Sturmian cylinder satisfies
\[
\frac{1}{q_{n+1}+q_n}<\mu_m(u)<\frac{2}{q_n}.
\]
If the partial quotients of \(\alpha\) are bounded by \(A\), then
\(q_{n+1}\le (A+1)q_n\), and since \(q_n\le m<q_{n+1}\), one has
\[
\frac{1}{(A+2)m}\le \mu_m(u)\le \frac{2(A+1)}{m}.
\]
This proves the uniform \(m^{-1}\) bounds.

Because \(|S_m(\alpha)|=m+1\), one has
\[
H_0(\mu_m)=\log(m+1),
\qquad
H_\infty(\mu_m)=-\log\max_{u\in S_m(\alpha)}\mu_m(u)\ge \log m-\log C_\alpha.
\]
Since R\'enyi entropy on a finite alphabet is nonincreasing in \(q\),
\[
\log m-\log C_\alpha
\le
H_q(\mu_m)
\le
\log(m+1)
\qquad (q\in[0,\infty]).
\]
This yields the claimed uniform asymptotic.
\end{proof}

\begin{proposition}[Explicit bounded-type R\'enyi window]\label{prop:explicit-bounded-type-kl-window}
Assume \(\beta=\alpha\), and suppose that the partial quotients of \(\alpha\)
are bounded by \(A\). Then for every \(m\ge 1\) and every \(q\in[0,\infty]\),
\[
\log m-\log(2(A+1))
\le
H_q(\mu_m)
\le
\log(m+1).
\]
Consequently,
\[
m\log\varphi-\log(m+1)-\log\pi(0)-\log\varphi
\le
D_q(\pi_m^{\mathrm{fold}}\|\pi_m)
\]
\[
\le
m\log\varphi-\log m+\log(2(A+1))-\log\pi(0)+\log\varphi.
\]
In particular,
\[
D_q(\pi_m^{\mathrm{fold}}\|\pi_m)
-
\bigl(m\log\varphi-\log m\bigr)
\]
remains bounded by explicit constants depending only on \(A\) and \(\pi(0)\),
uniformly in \(q\).
\end{proposition}

\begin{proof}
The proof of Proposition~\ref{prop:bounded-type-sturmian-entropy} gives
\[
\frac{1}{(A+2)m}\le \mu_m(u)\le \frac{2(A+1)}{m}
\qquad (u\in S_m(\alpha)).
\]
Because \(|S_m(\alpha)|=m+1\), one has
\[
H_0(\mu_m)=\log(m+1),
\]
while the pointwise upper mass bound yields
\[
H_\infty(\mu_m)
=
-\log\max_{u\in S_m(\alpha)}\mu_m(u)
\ge
\log m-\log(2(A+1)).
\]
By monotonicity in \(q\),
\[
\log m-\log(2(A+1))
\le
H_q(\mu_m)
\le
\log(m+1)
\qquad (q\in[0,\infty]).
\]
Corollary~\ref{cor:sturmian-exact-relabeling} gives
\[
H_q(\pi_m^{\mathrm{fold}})=H_q(\mu_m)
\qquad (q\in[0,\infty]),
\]
and Proposition~\ref{prop:parry-renyi-window} then yields
\[
D_q(\pi_m^{\mathrm{fold}}\|\pi_m)
=
m\log\varphi-H_q(\mu_m)-\log\pi(0)+O(1),
\]
where the \(O(1)\) term is bounded in absolute value by \(\log\varphi\),
uniformly in \(q\). Combining these estimates gives the displayed window.
\end{proof}

\begin{corollary}[Bounded-type injective encodings]
\label{cor:injective-relabeling-bounded-type}
Assume \(\beta=\alpha\), and suppose that \(\alpha\) has bounded partial
quotients. Then, for \(m\ge 6\),
\[
\inf_{\iota\in\mathcal J_m}D_q(\iota_*\mu_m\|\pi_m)
=
m\log\varphi-\log m+O_\alpha(1),
\qquad q\in[0,\infty],
\]
where the \(O_\alpha(1)\) term may be chosen uniformly in \(q\). Equivalently,
every injective encoding of the realized Sturmian law into \(X_m\) has the same
linear Parry penalty \(\log\varphi\) and the same logarithmic second-order term
\(-\log m\) up to bounded constants. In particular,
\[
D_q(\pi_m^{\mathrm{fold}}\|\pi_m)
=
\inf_{\iota\in\mathcal J_m}D_q(\iota_*\mu_m\|\pi_m)+O_\alpha(1),
\qquad q\in[0,\infty],
\]
and the additive error is in fact bounded between \(0\) and \(2\log\varphi\).
\end{corollary}

\begin{proof}
Theorem~\ref{thm:optimal-injective-relabeling} gives
\[
\inf_{\iota\in\mathcal J_m}D_q(\iota_*\mu_m\|\pi_m)
=
(m-1)\log\varphi-\log\pi(0)-H_q(\mu_m).
\]
Now apply Proposition~\ref{prop:bounded-type-sturmian-entropy} to substitute
\(H_q(\mu_m)=\log m+O_\alpha(1)\), uniformly in \(q\). The final claim is the
last part of Theorem~\ref{thm:optimal-injective-relabeling}.
\end{proof}

\begin{theorem}[Lossless deterministic maps]
\label{thm:lossless-coarse-graining-envelope}
Assume \(\beta=\alpha\) and \(m\ge 6\). Let \(S_m(\alpha)\) be the realized
Sturmian language and \(\mu_m\) its factor law. Define the class of
deterministic maps that are injective on the realized source:
\[
\mathcal G_m
:=
\{G_m:\Omega_m\to X_m \text{ deterministic}:\
G_m|_{S_m(\alpha)}\text{ injective}\}.
\]
For \(G_m\in\mathcal G_m\), write
\(\nu_m^G:=(G_m)_*\mu_m\). Then for every \(q\in[0,\infty]\),
\[
\inf_{G_m\in\mathcal G_m}D_q(\nu_m^G\|\pi_m)
=
(m-1)\log\varphi-\log\pi(0)-H_q(\mu_m),
\]
and every map \(G_m\) with \(G_m(S_m(\alpha))\subset X_m^{00}\) attains the
infimum. For each fixed order \(0\le q<\infty\), this endpoint-class
condition is also necessary for optimality. Moreover, the canonical fold
satisfies
\[
0\le D_q(\pi_m^{\mathrm{fold}}\|\pi_m)
-\inf_{G_m\in\mathcal G_m}D_q(\nu_m^G\|\pi_m)
\le 2\log\varphi.
\]
\end{theorem}

\begin{proof}
For any \(G_m\in\mathcal G_m\), the restriction
\(\iota:=G_m|_{S_m(\alpha)}:S_m(\alpha)\hookrightarrow X_m\) is injective, and
\(\nu_m^G=\iota_*\mu_m\). Thus the optimization over \(\mathcal G_m\) reduces
to the injective relabeling problem of
Theorem~\ref{thm:optimal-injective-relabeling}. The value of the infimum, the
sufficiency of the condition \(G_m(S_m(\alpha))\subset X_m^{00}\), and the
bounded gap for the canonical fold follow from that theorem.

It remains only to record the necessity statement for finite R\'enyi orders.
Let \(c_m=\pi(0)\varphi^{-(m-1)}\).  In the lower-bound part of the proof of
Theorem~\ref{thm:optimal-injective-relabeling}, equality for \(q=1\) or for
\(q\in(0,\infty)\setminus\{1\}\) requires
\(\pi_m(\iota(u))=c_m\) for every \(u\in S_m(\alpha)\), since all Sturmian
atoms have positive mass. For \(q=0\), equality requires
\(\pi_m(\iota(S_m(\alpha)))=(m+1)c_m\), and this again forces every image word
to have the maximal Parry mass \(c_m\). By
Proposition~\ref{prop:parry-endpoint-collapse}, the words with this maximal
mass are exactly \(X_m^{00}\). Hence every finite-order optimizer has realized
image contained in \(X_m^{00}\). No such necessity is asserted for
\(q=\infty\).
\end{proof}

\subsection{Primitive-SFT asymptotic extension}

\begin{theorem}[Primitive SFT support asymptotics]
\label{thm:primitive-sft-renyi-barrier}
Let \(\Sigma_A\) be a primitive one-step shift of finite type with topological
entropy \(h=\log\lambda\). Write \(X_m^A\) for its admissible words of length
\(m\), and let \(\pi_m^A\) be the \(m\)-block law of its Parry measure. Then
there exist constants \(c_A,C_A>0\) such that
\[
c_A e^{-mh}\le \pi_m^A(w)\le C_A e^{-mh}
\qquad (w\in X_m^A).
\]
Consequently, if \((\nu_m)_{m\ge 1}\) is any family with
\(\nu_m\in\Delta(X_m^A)\) and
\[
\limsup_{m\to\infty}\frac{1}{m}\log |\supp(\nu_m)|\le \gamma<h,
\]
then
\[
\liminf_{m\to\infty}\frac{1}{m}D_q(\nu_m\|\pi_m^A)\ge h-\gamma
\qquad (q\in[0,\infty]).
\]
Conversely, for every \(\gamma\in[0,h]\), there exists a family
\((\nu_m)_{m\ge 1}\) with \(\nu_m\in\Delta(X_m^A)\) such that
\[
\frac{1}{m}\log |\supp(\nu_m)|\to \gamma,
\qquad
\frac{1}{m}D_q(\nu_m\|\pi_m^A)\to h-\gamma
\qquad (q\in[0,\infty]).
\]
\end{theorem}

\begin{proof}
Let \(u\) and \(v\) be positive left and right Perron-Frobenius eigenvectors
for \(A\), normalized so that \(\sum_i u_i v_i=1\). The standard Parry formula
for primitive one-step shifts \cite{LindMarcus1995,ParryPollicott1990Zeta}
gives
\[
\pi_m^A(i_0\cdots i_{m-1})=u_{i_0}v_{i_{m-1}}\lambda^{-(m-1)}
\]
for every admissible word \(i_0\cdots i_{m-1}\in X_m^A\). Since the alphabet is
finite and \(u_i v_j>0\) on admissible endpoint pairs, there are constants
\(c_A,C_A>0\) such that
\[
c_A e^{-mh}\le \pi_m^A(w)\le C_A e^{-mh}
\qquad (w\in X_m^A).
\]

Now let \(S_m:=\supp(\nu_m)\). By
Proposition~\ref{prop:renyi-support-conditioning},
\[
D_q(\nu_m\|\pi_m^A)\ge -\log \pi_m^A(S_m)
\qquad (q\in[0,\infty]).
\]
Using the upper block-mass bound,
\[
\pi_m^A(S_m)\le C_A |S_m| e^{-mh},
\]
so
\[
\frac{1}{m}D_q(\nu_m\|\pi_m^A)
\ge
h-\frac{1}{m}\log |S_m|-\frac{\log C_A}{m}.
\]
Taking \(\liminf\) proves the lower bound.

For sharpness, first fix \(\gamma\in[0,h)\) and set
\(k_m:=\max\{1,\lfloor e^{\gamma m}\rfloor\}\). Choose any subset
\(S_m\subset X_m^A\) with \(|S_m|=k_m\), which is possible for all large
\(m\) because \(|X_m^A|=e^{hm+o(m)}\). Define
\(\nu_m:=\pi_m^A(\cdot\mid S_m)\). Then
\[
c_A k_m e^{-mh}
\le
\pi_m^A(S_m)
\le
C_A k_m e^{-mh},
\]
and Proposition~\ref{prop:renyi-support-conditioning} yields
\[
D_q(\nu_m\|\pi_m^A)=-\log \pi_m^A(S_m)=mh-\log k_m+O_A(1)
\qquad (q\in[0,\infty]).
\]
Since \(m^{-1}\log k_m\to \gamma\), this gives the desired limit. For
\(\gamma=h\), take \(\nu_m:=\pi_m^A\); then every \(D_q(\nu_m\|\pi_m^A)\) is
zero, while the language growth of a primitive shift satisfies
\(|X_m^A|=e^{hm+o(m)}\) \cite{LindMarcus1995}, so
\(m^{-1}\log |\supp(\nu_m)|\to h\).
\end{proof}

\begin{remark}[Asymptotic nature of the primitive-SFT statement]
\label{rem:primitive-sft-no-finite-m-claim}
Theorem~\ref{thm:primitive-sft-renyi-barrier} is an asymptotic statement. It
uses only the comparability \(\pi_m^A(w)\asymp e^{-mh}\) of Parry block masses
on a primitive shift. Unlike the golden-mean result
Theorem~\ref{thm:parry-support-barrier}, it does not give a finite-\(m\) mass
envelope \(M_m(k)\) or support-constrained optimizers.
\end{remark}

\begin{theorem}[Primitive SFT injective placements]
\label{thm:primitive-sft-injective-barrier}
Let \(\Sigma_A\) be a primitive one-step shift of finite type with topological
entropy \(h=\log\lambda\). Let \(u,v\) be positive left and right
Perron-Frobenius eigenvectors for \(A\), normalized so that
\(\sum_i u_i v_i=1\). For an admissible word \(w=i_0\cdots i_{m-1}\), the
Parry formula gives
\(\pi_m^A(w)=u_{i_0}v_{i_{m-1}}\lambda^{-(m-1)}\). Define
\[
c_*:=\max_{i,j}u_i v_j,
\]
and let \((i_*,j_*)\) attain this maximum. Write
\[
\mathcal C_m(i_*,j_*):=\{w\in X_m^A: w_1=i_*,\;w_m=j_*\}.
\]
Then there exist \(\kappa_A>0\) and \(m_0\) such that for all \(m\ge m_0\),
\[
|\mathcal C_m(i_*,j_*)|\ge \kappa_A e^{mh}.
\]
Moreover, if \(\mu_m\) is any probability vector on a finite set \(S_m\), then
for every injection \(\iota:S_m\hookrightarrow X_m^A\) and every
\(q\in[0,\infty]\),
\[
D_q(\iota_*\mu_m\|\pi_m^A)
\ge
(m-1)h-\log c_*-H_q(\mu_m).
\]
If \(|S_m|\le |\mathcal C_m(i_*,j_*)|\), the lower bound is attained by any
injection into \(\mathcal C_m(i_*,j_*)\), and
\[
\inf_{\iota:S_m\hookrightarrow X_m^A}D_q(\iota_*\mu_m\|\pi_m^A)
=
(m-1)h-\log c_*-H_q(\mu_m).
\]
In particular, if \(\log|S_m|=o(m)\), then
\[
\inf_{\iota:S_m\hookrightarrow X_m^A}D_q(\iota_*\mu_m\|\pi_m^A)
=
mh-H_q(\mu_m)+O_A(1),
\]
where the \(O_A(1)\) term is uniform in \(q\).
\end{theorem}

\begin{proof}
By the Perron-Frobenius theorem,
\((A^{m-1})_{i_*j_*}=\lambda^{m-1}(v_{i_*}u_{j_*}+o(1))\), so there exist
\(\kappa_A>0\) and \(m_0\) with
\(|\mathcal C_m(i_*,j_*)|=(A^{m-1})_{i_*j_*}\ge \kappa_A\lambda^{m-1}\) for
\(m\ge m_0\).

Fix an injection \(\iota\) and write \(p_s=\mu_m(s)\),
\(\rho_s=\pi_m^A(\iota(s))\). By definition of \(c_*\),
\(\rho_s\le c_*\lambda^{-(m-1)}\) for all \(s\in S_m\).

For \(q=1\), we have
\[
D_1=\sum_s p_s\log(p_s/\rho_s)
\ge \sum_s p_s\log(p_s/(c_*\lambda^{-(m-1)}))
=(m-1)h-\log c_*-H_1(\mu_m).
\]

For \(q\in(0,\infty)\setminus\{1\}\), the bound
\(\rho_s\le c_*\lambda^{-(m-1)}\) gives
\(\rho_s^{1-q}\ge (c_*\lambda^{-(m-1)})^{1-q}\) when \(q>1\) and
\(\rho_s^{1-q}\le (c_*\lambda^{-(m-1)})^{1-q}\) when \(q<1\). In both cases
the prefactor \((q-1)^{-1}\) produces the same lower bound
\(D_q\ge (m-1)h-\log c_*-H_q(\mu_m)\).

For \(q=\infty\),
\[
D_\infty=\log\max_s(p_s/\rho_s)\ge (m-1)h-\log c_*-H_\infty(\mu_m).
\]

For \(q=0\),
\(\pi_m^A(\iota(S_m))\le |S_m|c_*\lambda^{-(m-1)}\), so
\(D_0=-\log\pi_m^A(\iota(S_m))\ge (m-1)h-\log c_*-H_0(\mu_m)\).

For attainability, if \(|S_m|\le |\mathcal C_m(i_*,j_*)|\), inject
\(S_m\) into \(\mathcal C_m(i_*,j_*)\). All image words then have identical
Parry mass \(c_*\lambda^{-(m-1)}\), so the lower bound is attained for
every \(q\). If \(\log|S_m|=o(m)\), then
\(|S_m|\le \kappa_A\lambda^{m-1}\) for large \(m\), and absorbing
\((m-1)h-\log c_*\) into \(mh+O_A(1)\) yields the last claim.
\end{proof}

\begin{corollary}[Low-complexity source comparison]
\label{cor:low-complexity-source-barrier}
Let \(\Sigma_A\) be a primitive one-step SFT with topological entropy
\(h=\log\lambda\) and Parry \(m\)-block law \(\pi_m^A\). Let
\((S_m,\mu_m)_{m\ge 1}\) be a family of finite probability spaces satisfying
\(|S_m|=e^{o(m)}\) and \(H_q(\mu_m)=\log|S_m|+O(1)\) uniformly in
\(q\in[0,\infty]\). Then for all sufficiently large \(m\), any injection
\(\iota_m:S_m\hookrightarrow X_m^A\) satisfies
\[
D_q(\iota_{m*}\mu_m\|\pi_m^A)
=
mh-\log|S_m|+O(1),
\qquad q\in[0,\infty],
\]
with constants uniform in \(q\). In particular, if \(|S_m|=m+O(1)\), then
\[
D_q(\iota_{m*}\mu_m\|\pi_m^A)
=
mh-\log m+O(1).
\]
\end{corollary}

\begin{proof}
Theorem~\ref{thm:primitive-sft-renyi-barrier} gives constants \(c_A,C_A>0\)
such that
\[
c_A e^{-mh}\le \pi_m^A(w)\le C_A e^{-mh}
\qquad (w\in X_m^A).
\]
Let \(\rho_s:=\pi_m^A(\iota_m(s))\). The same calculation used in
Theorem~\ref{thm:primitive-sft-injective-barrier}, now with both the upper and
lower pointwise bounds on \(\rho_s\), gives
\[
D_q(\iota_{m*}\mu_m\|\pi_m^A)=mh-H_q(\mu_m)+O_A(1),
\qquad q\in[0,\infty],
\]
uniformly over the injection and over the R\'enyi order: for \(q=1\) this follows
from \(-\log\rho_s=mh+O_A(1)\); for
\(q\in(0,\infty)\setminus\{1\}\), the factor \(\rho_s^{1-q}\) differs from
\(e^{mh(q-1)}\) by \(e^{O_A(|1-q|)}\), which becomes \(O_A(1)\) after division
by \(q-1\); and the cases \(q=0,\infty\) follow from the same pointwise mass
comparison. Substituting \(H_q(\mu_m)=\log|S_m|+O(1)\) proves the first claim.
The linear-complexity case follows from \(\log|S_m|=\log m+O(1)\).
\end{proof}

\subsection{Finite-block factors of Sturmian sources}

The preceding primitive-SFT estimates are stated for arbitrary finite
probability spaces.  The next result applies them to finite-block observables
of a bounded-type Sturmian system.  It shows that the Parry discrepancy is
already present at the level of topological factors, before any particular
choice of fold is made.

\begin{theorem}[Finite-block factors of bounded-type Sturmian systems]
\label{thm:finite-block-factor-parry-barrier}
Let \(X_\alpha\) be the Sturmian subshift of a bounded-type irrational
\(\alpha\), and let \(\mu_\alpha\) be its unique invariant measure. Let
\(\Sigma_A\) be a primitive one-step SFT with topological entropy
\(h=\log\lambda>0\), and let \(\pi_n^A\) denote its Parry \(n\)-block law.
Let
\[
\Phi:X_\alpha\longrightarrow \Sigma_A
\]
be a sliding-block code with finite coding radius, assume that
\(Y:=\Phi(X_\alpha)\subset \Sigma_A\), write
\(\nu^\Phi:=\Phi_*\mu_\alpha\), and let \(\nu_n^\Phi\) be its length-\(n\)
block law. If \(Y\) is aperiodic, then
\[
H_q(\nu_n^\Phi)=\log n+O_{\alpha,\Phi}(1)
\qquad (q\in[0,\infty])
\]
and
\[
D_q(\nu_n^\Phi\|\pi_n^A)
=
nh-\log n+O_{\alpha,\Phi,A}(1)
\qquad (q\in[0,\infty]),
\]
where both error terms are uniform in the R\'enyi order. If \(Y\) is periodic,
then
\[
H_q(\nu_n^\Phi)=O_\Phi(1),
\qquad
D_q(\nu_n^\Phi\|\pi_n^A)=nh+O_{\Phi,A}(1)
\qquad (q\in[0,\infty]),
\]
again uniformly in \(q\). Consequently every finite-block factor of a
bounded-type Sturmian source has Parry divergence density \(h\) inside
\(\Sigma_A\), and the logarithmic correction \(-\log n\) is forced precisely
for aperiodic factors.
\end{theorem}

\begin{proof}
Bounded-type Sturmian subshifts are linearly recurrent, and a nonperiodic
factor of a linearly recurrent subshift is again linearly recurrent
\cite{Durand2000LinearlyRecurrentFactors,Fogg2002}. Hence, in the aperiodic
case, \(Y\) is linearly recurrent with some constant \(L\). The image system is
also uniquely ergodic, since it is a factor of the uniquely ergodic system
\((X_\alpha,\mu_\alpha)\).

Let \(\mathcal L_n(Y)\) denote the length-\(n\) language of \(Y\), and let
\(p_Y(n):=|\mathcal L_n(Y)|\). If the coding radius of \(\Phi\) is \(r\), then
each length-\(n\) word of \(Y\) is induced by at least one
length-\((n+2r)\) word of \(X_\alpha\). Since
\(p_{X_\alpha}(k)=k+1\),
\[
p_Y(n)\le n+2r+1.
\]
As \(Y\) is aperiodic, the Morse-Hedlund theorem gives \(p_Y(n)\ge n+1\).
Thus
\[
p_Y(n)=n+O_\Phi(1).
\]

Standard return-word estimates for nonperiodic linearly recurrent subshifts
\cite{DurandHostSkau1999Substitutional,Durand2000LinearlyRecurrentFactors}
give the following uniform frequency bounds; this is the only property of
linear recurrence used below. After increasing \(L\) if necessary, every
return word to every \(v\in\mathcal L_n(Y)\) has length between \(L^{-1}n\)
and \(Ln\). Unique ergodicity then gives
\[
\frac{1}{Ln}\le \nu^\Phi([v])\le \frac{L}{n}
\qquad (v\in\mathcal L_n(Y)).
\]
Together with \(p_Y(n)=n+O_\Phi(1)\), these bounds imply
\[
\log n-\log L
\le H_\infty(\nu_n^\Phi)
\le H_q(\nu_n^\Phi)
\le H_0(\nu_n^\Phi)
=\log p_Y(n)
\le \log n+O_\Phi(1)
\]
for every \(q\in[0,\infty]\). Hence
\[
H_q(\nu_n^\Phi)=\log n+O_{\alpha,\Phi}(1)
\]
uniformly in \(q\).

The Parry block law of the primitive SFT satisfies
\[
c_Ae^{-nh}\le \pi_n^A(w)\le C_Ae^{-nh}
\qquad (w\in X_n^A)
\]
by Theorem~\ref{thm:primitive-sft-renyi-barrier}. Therefore
\[
D_q(\nu_n^\Phi\|\pi_n^A)
=
nh-H_q(\nu_n^\Phi)+O_A(1)
\qquad (q\in[0,\infty]),
\]
uniformly in \(q\). For \(q=1\) this follows by averaging
\(-\log\pi_n^A(w)=nh+O_A(1)\). For
\(q\in(0,\infty)\setminus\{1\}\), writing
\(\pi_n^A(w)=b_w e^{-nh}\) with \(b_w\in[c_A,C_A]\) gives
\[
D_q(\nu_n^\Phi\|\pi_n^A)
=
nh+\frac{1}{q-1}
\log\sum_w \nu_n^\Phi(w)^q b_w^{1-q},
\]
and the factor \(b_w^{1-q}\) changes the last term by at most
\(\max\{|\log c_A|,|\log C_A|\}\). The endpoint cases \(q=0\) and
\(q=\infty\) follow from the same pointwise comparison:
\[
D_0(\nu_n^\Phi\|\pi_n^A)
=nh-H_0(\nu_n^\Phi)+O_A(1),
\qquad
D_\infty(\nu_n^\Phi\|\pi_n^A)
=nh-H_\infty(\nu_n^\Phi)+O_A(1).
\]
Substituting \(H_q(\nu_n^\Phi)=\log n+O_{\alpha,\Phi}(1)\) proves the
aperiodic assertion.

If \(Y\) is periodic, its language has uniformly bounded complexity and the
block frequencies of \(\nu^\Phi\) take only finitely many values determined by
one period. Consequently \(H_q(\nu_n^\Phi)=O_\Phi(1)\) uniformly in \(q\).
The same Parry comparison then gives
\[
D_q(\nu_n^\Phi\|\pi_n^A)=nh+O_{\Phi,A}(1)
\]
for every \(q\in[0,\infty]\). Dividing by \(n\) yields the asserted divergence
density in both cases.
\end{proof}

\begin{remark}[Numerical interpretation]
\label{rem:coarse-graining-audit}
The numerical examples follow the two comparisons used in the proofs.  First
we compare the empirical histogram with the deterministic folded law.  We then
compare the limiting folded law with the Parry block law.  The examples are
included to illustrate this separation in the present model, not to survey
other possible folds.
\end{remark}

\section{Numerical checks}

This section gives numerical checks of the preceding estimates.  We start with
the canonical slice \(\beta=\alpha\), where the finite-\(m\) statements of
Sections~4 and~5 are sharp, and then consider nearby parameter families and an
independent sampling baseline.  The relevant distinction is between
convergence to \(\pi_m^{\mathrm{fold}}\) and comparison with \(\pi_m\).  In
the canonical slice, Theorem~\ref{thm:sequence-level-transparency} rules out
information loss under the fold, while
Theorem~\ref{thm:optimal-injective-relabeling} identifies the best injective
placement of the realized Sturmian law in \(X_m\).

\subsection{Numerical setup}

\begin{table}[t]
\centering
\small
\begin{tabular}{>{\raggedright\arraybackslash}p{0.25\linewidth} >{\raggedright\arraybackslash}p{0.63\linewidth}}
\toprule
Experiment family & Parameter grid \\
\midrule
Rotation sweep &
$m\in\{6,8,10,12,14,16,18\}$,
$N\in\{10^4,3\times 10^4,10^5,3\times 10^5\}$,
$\beta\in\{0.2,0.3,0.5,0.6180339887\}$,
$x_0\in\{0.123,0.271,0.731\}$ \\
Rotation families &
golden, metallic means $A\in\{2,3,5\}$, and large first partial quotient
families $[0;B,1,1,\dots]$ with $B\in\{10,30,100\}$ \\
Independent block baselines &
$m\in\{6,8,10,12,14,16,18\}$,
$N\in\{2000,5000,10^4,3\times 10^4,10^5\}$,
seeds $\{1,2,3,4,5\}$, confidence level $95\%$ \\
\bottomrule
\end{tabular}
\caption{Parameter ranges used in the numerical examples.}
\label{tab:parameter-grid}
\end{table}

All tables are computed from this fixed grid.  In the rotation experiments the
same orbit sample is used for the empirical microstate law, the folded
histogram, the discrepancy bound and the cross-resolution residual.  The
overlines in Table~\ref{tab:rotation_multiscale_residual_summary} denote
averages over the listed orbit phases; Table~\ref{tab:rotation-parry-gap-summary}
also averages over the displayed \(\beta\)-grid.  The residual and Parry tables
use \(N=3\times 10^5\), for which the discrepancy term is already on the order
of \(10^{-5}\).

All slope families in Table~\ref{tab:parameter-grid} are of bounded type.  The
family \([0;B,1,1,\dots]\) is included to show the effect of a large first
partial quotient, not as an unbounded-type example.

\subsection{Canonical slice}

Theorem~\ref{thm:sequence-level-transparency} is exact.  The first numerical
check is therefore the injective placement formula in
Theorem~\ref{thm:optimal-injective-relabeling}.  The table below is computed
from the Lebesgue factor law, with no averaging over \(\beta\) or \(x_0\), and
not from a finite orbit sample.  It begins at \(m=6\), the first resolution at
which \(X_m^{00}\) contains enough words to realize the injective optimum.

\input{generated/tab_canonical_injective_floor_exact_en}

Table~\ref{tab:canonical-injective-floor-exact} shows the finite-\(m\)
content of Theorem~\ref{thm:optimal-injective-relabeling}. The
optimal injective floor \(L_1(m)\) and the exact folded-to-Parry divergence
\(\KL(\pi_m^{\mathrm{fold}}\|\pi_m)\) both grow on the scale
\(m\log\varphi-\log m\). In the displayed range, direct enumeration of the
finite Sturmian partition gives the constant difference
\[
\KL(\pi_m^{\mathrm{fold}}\|\pi_m)-L_1(m)
=
(1-\varphi^{-3})\log\varphi
=
0.3676131227688\ldots
\]
This is the identity of Theorem~\ref{thm:optimal-injective-relabeling}
together with the enumerated endpoint imbalance
\(\eta_{11}(\pi_m^{\mathrm{fold}})-\eta_{00}(\pi_m^{\mathrm{fold}})
=-\varphi^{-3}\) for these canonical golden-slope rows.
At these resolutions the canonical folded law is far inside the universal
\(2\log\varphi\) additive window above the injective optimum.  This table is
the reference point for the empirical comparisons below.

\subsection{Finite-sample validation}

\input{generated/tab_rotation_folded_kl_certificate_en}

Table~\ref{tab:rotation-folded-kl-certificate} is for the golden slope with
the canonical window \(\beta=\alpha\).  In the parameter range shown, the
observed folded relative entropy is below the explicit upper bound.  The
constants are conservative, but the dependence on \((m,N)\) is at the
predicted scale.  The same canonical slice also shows why comparison with the
Parry baseline is different: increasing \(N\) improves agreement with
\(\pi_m^{\mathrm{fold}}\), but it does not remove the deterministic gap to
\(\pi_m\).  Once the empirical histogram is close to
\(\pi_m^{\mathrm{fold}}\), the injective floor in
Table~\ref{tab:canonical-injective-floor-exact} is the relevant reference.

\subsection{Cross-resolution residuals}

\input{generated/tab_rotation_multiscale_residual_summary_en}

Table~\ref{tab:rotation_multiscale_residual_summary} displays the separation
between finite-sample and limiting effects.  At \(N=3\times 10^5\), the mean
star discrepancy in the reported golden slice is of order \(10^{-5}\), whereas
the projective residual remains of order \(10^{-1}\).  For the family
\([0;100,1,1,\dots]\), the residual is larger.  This is the phenomenon
captured by
Theorem~\ref{prop:multiscale-residual-by-boundary}: small orbit discrepancy
does not force small cross-resolution residual, because the residual is driven
by the frequency with which a final \(1\) is appended to prefixes whose
Fibonacci value has already crossed the threshold \(F_{m+1}\).

Corollary~\ref{cor:multiscale-residual-audit-bound} gives the corresponding
quantitative bound: the empirical residual is controlled by an explicit
discrepancy term and by the deterministic total-variation gaps to the Parry
baselines at levels \(m\) and \(m+1\).  Once the discrepancy term is smaller
than these gaps, the source/reference comparison controls the bound.

\subsection{Additional parameter choices}

\input{generated/tab_rotation_parry_gap_summary_en}

Table~\ref{tab:rotation-parry-gap-summary} moves away from the canonical
slice.  At the fixed sample size \(N=3\times 10^5\), the golden discrepancy
upper bound is already \(9.33\times 10^{-5}\), while the total-variation and
relative-entropy gaps to the Parry measure still grow with \(m\) across both
bounded-type and resonant families.  For the golden branch, for example, the
total variation distance rises from \(0.506\) at \(m=6\) to \(0.995\) at
\(m=18\).  This is the same qualitative regime predicted by
Corollary~\ref{cor:tv-decomposition-parry},
Corollary~\ref{cor:injective-relabeling-bounded-type}, and the exact
support-budget formulas of Section~5: once \(m\) grows, the dominant
contribution is the limiting Parry mismatch rather than sampling error.

The comparison across families is also informative.  The metallic and resonant
examples differ in constants, but they exhibit the same obstruction: after
folding into \(X_m\), low effective complexity prevents convergence to the
maximal-entropy baseline.  By the support-minimum results of Section~5, any
alternative fold producing a subexponential-support family in \(X_m\) retains
a positive Parry entropy density.

\subsection{Comparison with independent baselines}

\input{generated/tab_iid_sources_fold_vs_parry_ci_en}

The independent comparison illustrates the decomposition in
Corollary~\ref{cor:tv-decomposition-parry} in two regimes.  When the
data-generating law is matched to the reference baseline, the sampling term
controls the observed error.  For folded rotation data, the deterministic
Parry term remains and eventually dominates.  A different fold may change
constants, or change the effective support exponent, but it cannot remove the
Parry penalty while staying in the same low-complexity regime.

For reference, the standard high-probability envelopes on a finite alphabet
\(E\) with \(|E|=F_{m+2}\) are
\[
\TV(\hat p_N,p)
\lesssim
\sqrt{\frac{\log(2^{F_{m+2}}/\delta)}{2N}},
\qquad
\KL(\hat p_N\|p)
\lesssim
\frac{F_{m+2}\log(N+1)+\log(1/\delta)}{N}
\]
at confidence level \(1-\delta\); see
\cite{WeissmanEtAl2003L1Empirical,DemboZeitouni2010LargeDeviations}. These
are independent sampling envelopes on the finite alphabet \(E\). The rotation
experiments are controlled instead by the deterministic discrepancy
certificates of Section~3. When
\(p=\pi_m\) and the data are Parry-distributed, the observed errors in the
table fall well within this envelope. When \(p=\pi_m\) but the data come from
folded rotation orbits, the deterministic Parry gap dominates.

\subsection{A nonlinear circle-map example}

The construction is not tied to rigid rotations as state equations.  The
following observation records the conjugacy invariance needed for nonlinear
circle maps.

\begin{theorem}[Conjugacy invariance of the comparison]
\label{thm:audit-conjugacy-invariance}
Let \(f:S^1\to S^1\) be an orientation-preserving circle homeomorphism
topologically conjugate to the irrational rotation
\(R_\alpha:x\mapsto x+\alpha\pmod 1\) via an orientation-preserving
homeomorphism \(h\) satisfying \(h\circ f=R_\alpha\circ h\). Fix an anchor
\(a\in S^1\) with \(h(a)=0\), and define the coding window
\(I_a:=[a,f(a))\), the positively oriented half-open arc from \(a\) to
\(f(a)\). For \(x\in S^1\), let
\[
s_t^{(f)}(x):=\mathbf{1}_{I_a}(f^t(x)),
\qquad
s_t^{(\alpha)}(y):=\mathbf{1}_{[0,\alpha)}(y+t\alpha\bmod 1).
\]
Then for every \(x\in S^1\) and \(y:=h(x)\),
\[
s_t^{(f)}(x)=s_t^{(\alpha)}(y)
\qquad(t\ge 0).
\]
Consequently, for every \(m,N\ge 1\), the length-\(m\) microstate laws,
folded laws, empirical histograms, discrepancy certificates, Parry
comparisons, higher-block conjugacy, and all R\'enyi/KL conclusions of
Sections~3--5 are identical for the two systems, with the invariant target law
for \(f\) understood as \(h^{-1}_*\leb\) and the discrepancy certificate taken
in the rotation coordinate \(h(x)\). In particular, if \(f\) is a \(C^2\)
orientation-preserving circle diffeomorphism with irrational rotation number
\(\alpha\), the Denjoy theorem
\cite{Denjoy1932CourbesTore,KatokHasselblatt1995} provides the required
conjugacy, and the preceding comparison applies.
\end{theorem}

\begin{proof}
Since \(h\) is orientation-preserving, the image of \(I_a=[a,f(a))\) is the
positively oriented arc from \(h(a)=0\) to
\(h(f(a))=R_\alpha(h(a))=\alpha\), i.e., \(h(I_a)=[0,\alpha)\). For any
\(x\in S^1\) with \(y=h(x)\), the conjugacy gives
\(h(f^t(x))=R_\alpha^t(y)=y+t\alpha\pmod 1\), so
\[
s_t^{(f)}(x)
=\mathbf{1}_{I_a}(f^t(x))
=\mathbf{1}_{h(I_a)}(h(f^t(x)))
=\mathbf{1}_{[0,\alpha)}(y+t\alpha\bmod 1)
=s_t^{(\alpha)}(y).
\]
Since the symbol sequences agree pointwise for all \(t\), the length-\(m\)
sliding windows, and therefore all empirical counts, folded histograms, and
derived quantities, are identical.
\end{proof}

For example, consider the standard circle map
\[
f_{K,\Omega}(x)=x+\Omega-\frac{K}{2\pi}\sin(2\pi x)\pmod 1
\]
with \(K=0.3\) and \(\Omega\) tuned by bisection on the lift so that the
rotation number equals \(\varphi^{-1}\), giving
\(\Omega\approx 0.6166931270470741\). With anchor \(a=0.2\), coding window
\(I=[a,f(a))\), and initial point \(x_0=a\), the resulting symbolic sequence
is the canonical golden Sturmian coding in conjugacy coordinates.
At \(m=12\) and \(N=10^5\), the empirical folded histogram satisfies
\[
\TV(\hat\pi_{12,N},\pi_{12}^{\mathrm{fold}})
=2.43\times 10^{-5},
\qquad
\KL(\hat\pi_{12,N}\|\pi_{12}^{\mathrm{fold}})
=2.95\times 10^{-9},
\]
while the gaps to the Parry baseline remain
\[
\TV(\hat\pi_{12,N},\pi_{12})=0.9650,
\qquad
\KL(\hat\pi_{12,N}\|\pi_{12})=3.4443.
\]
The distance to the folded limit is already at the \(10^{-5}\) level, while the
Parry discrepancy remains of order \(10^0\).  The same comparison therefore
applies to this nonlinear state-space model.

\section{Concluding remarks}

The construction gives a finite-resolution setting in which sampling error and
Parry comparison can be separated explicitly.  Discrepancy estimates control
the empirical folded histogram, while the support and placement results
describe the limiting folded law.  In the Sturmian slice the fold is injective
on realized blocks and is conjugate to the classical higher-block
presentation, so the Parry gap is not caused by collisions under the fold.

The main obstruction is the low effective complexity of the source inside the
ambient golden-mean language. Even after optimizing over all injective
placements of the realized Sturmian law, one obtains
\[
D_q=m\log\varphi-\log m+O_\alpha(1)
\]
for bounded-type slopes, uniformly in the R\'enyi order. The canonical fold is
within a bounded additive distance of this optimum. The primitive-SFT
statements show that the same leading term \(mh-H_q(\mu_m)\) is the general
injective-placement obstruction, although outside the golden-mean case we no
longer have the finite-\(m\) endpoint envelope.

The exact formulas rely on the endpoint collapse of golden-mean Parry
cylinders.  Thus the paper should be read as an explicit model, rather than as
a classification of possible folds.  The finite-block factor theorem shows,
however, that the leading Parry obstruction is not special to this fold: every
aperiodic finite-block factor of a bounded-type Sturmian source inside a
primitive positive-entropy SFT has the same divergence density and the same
\(-\log m\) correction.  The appendices record the fiber-uniform lift
identities, the local Zeckendorf normalization and the contiguous-numeration
uniqueness of the Fibonacci weights.

\appendix

\section{Fiber-uniform lifts and the Parry lift}\label{app:fiber-lift}

This appendix records the lift identities associated with the canonical fold.
They are used to interpret the degeneracy term in
Corollary~\ref{cor:sturmian-pure-degeneracy-ledger}.

\begin{definition}[Fiber multiplicity and uniform lift]\label{def:fiber-uniform-lift}
For \(x\in X_m\), define the fiber multiplicity
\[
d_m(x):=\bigl|\Fold_m^{-1}(x)\bigr|.
\]
For \(\eta\in\Delta(X_m)\), define its fiber-uniform lift
\[
(Q_m\eta)(a):=\frac{\eta(\Fold_m(a))}{d_m(\Fold_m(a))},
\qquad a\in\Omega_m.
\]
Then \((\Fold_m)_*(Q_m\eta)=\eta\). For \(\mu\in\Delta(\Omega_m)\), let
\[
\pi:=(\Fold_m)_*\mu\in\Delta(X_m),
\qquad
\mu^e:=Q_m\pi.
\]
\end{definition}

\begin{proposition}[Exact entropy decomposition]\label{prop:fold-entropy-ledger}
For every \(\mu\in\Delta(\Omega_m)\), with \(\pi=(\Fold_m)_*\mu\) and
\(\mu^e=Q_m\pi\),
\[
H(\mu)
=
H(\pi)+\sum_{x\in X_m}\pi(x)\log d_m(x)-\KL(\mu\|\mu^e).
\]
Equivalently,
\[
H(\mu)-H(\pi)
=
\mathbb{E}_{\pi}[\log d_m(X)]-\KL(\mu\|\mu^e).
\]
\end{proposition}

\begin{proof}
Write \(x_a:=\Fold_m(a)\). By Definition~\ref{def:fiber-uniform-lift},
\[
\mu^e(a)=\frac{\pi(x_a)}{d_m(x_a)}.
\]
Hence
\[
\begin{aligned}
\KL(\mu\|\mu^e)
&=
\sum_{a\in\Omega_m}\mu(a)\log\frac{\mu(a)}{\mu^e(a)}\\
&=
-H(\mu)-\sum_{a\in\Omega_m}\mu(a)\log\pi(x_a)
+\sum_{a\in\Omega_m}\mu(a)\log d_m(x_a).
\end{aligned}
\]
Since \(\sum_{a:\,x_a=x}\mu(a)=\pi(x)\), the two sums become
\[
\sum_{a\in\Omega_m}\mu(a)\log\pi(x_a)=\sum_{x\in X_m}\pi(x)\log\pi(x)=-H(\pi)
\]
and
\[
\sum_{a\in\Omega_m}\mu(a)\log d_m(x_a)
=
\sum_{x\in X_m}\pi(x)\log d_m(x).
\]
Rearranging gives the claim.
\end{proof}

\begin{proposition}[KL isometry of fiber-uniform lifts]\label{prop:fold-fiber-lift-kl-isometry}
For every \(\pi,\eta\in\Delta(X_m)\),
\[
\KL(Q_m\pi\|Q_m\eta)=\KL(\pi\|\eta).
\]
\end{proposition}

\begin{proof}
Again write \(x_a:=\Fold_m(a)\). Then
\[
\frac{(Q_m\pi)(a)}{(Q_m\eta)(a)}
=
\frac{\pi(x_a)}{\eta(x_a)},
\]
so
\[
\begin{aligned}
\KL(Q_m\pi\|Q_m\eta)
&=
\sum_{a\in\Omega_m}\frac{\pi(x_a)}{d_m(x_a)}
\log\frac{\pi(x_a)}{\eta(x_a)}\\
&=
\sum_{x\in X_m}
\left(\sum_{a:\,x_a=x}\frac{\pi(x)}{d_m(x)}\right)
\log\frac{\pi(x)}{\eta(x)}
=
\KL(\pi\|\eta).
\end{aligned}
\]
\end{proof}

\begin{theorem}[Pythagorean decomposition for the canonical fold]\label{thm:fold-kl-pythagoras}
Let \(\mu\in\Delta(\Omega_m)\), let \(\pi=(\Fold_m)_*\mu\), and let
\(\mu^e=Q_m\pi\). Then for every \(\eta\in\Delta(X_m)\),
\[
\KL(\mu\|Q_m\eta)
=
\KL(\mu\|\mu^e)+\KL(\pi\|\eta).
\]
Consequently, \(\mu^e\) is the unique I-projection of \(\mu\) onto the
fiber-uniform model family
\[
\mathcal E_m:=\{Q_m\eta:\eta\in\Delta(X_m)\}\subset\Delta(\Omega_m),
\]
and
\[
\inf_{\eta\in\Delta(X_m)}\KL(\mu\|Q_m\eta)=\KL(\mu\|\mu^e).
\]
\end{theorem}

\begin{proof}
Let \(x_a:=\Fold_m(a)\). Since
\[
\mu^e(a)=\frac{\pi(x_a)}{d_m(x_a)},
\qquad
(Q_m\eta)(a)=\frac{\eta(x_a)}{d_m(x_a)},
\]
one has
\[
\log\frac{\mu^e(a)}{(Q_m\eta)(a)}=\log\frac{\pi(x_a)}{\eta(x_a)}.
\]
Therefore
\[
\begin{aligned}
\KL(\mu\|Q_m\eta)-\KL(\mu\|\mu^e)
&=
\sum_{a\in\Omega_m}\mu(a)\log\frac{\mu^e(a)}{(Q_m\eta)(a)}\\
&=
\sum_{a\in\Omega_m}\mu(a)\log\frac{\pi(x_a)}{\eta(x_a)}\\
&=
\sum_{x\in X_m}
\left(\sum_{a:\,x_a=x}\mu(a)\right)\log\frac{\pi(x)}{\eta(x)}
=
\KL(\pi\|\eta).
\end{aligned}
\]
This proves the decomposition. Since \(\KL(\pi\|\eta)\ge 0\), with equality if
and only if \(\eta=\pi\), the minimizer is unique and equals \(\mu^e=Q_m\pi\).
\end{proof}

\begin{proposition}[Dual characterization of the fiber-uniform lift]
\label{prop:dual-projection-maxent}
For any \(\pi\in\Delta(X_m)\), define the linear family
\[
\mathcal L_\pi:=\{\nu\in\Delta(\Omega_m):(\Fold_m)_*\nu=\pi\}.
\]
Then \(Q_m\pi\) is the unique maximum-entropy element of
\(\mathcal L_\pi\):
\[
H(\nu)\le H(Q_m\pi)
\qquad(\nu\in\mathcal L_\pi),
\]
with equality if and only if \(\nu=Q_m\pi\). Hence, for any
\(\mu\in\Delta(\Omega_m)\) with \(\pi=(\Fold_m)_*\mu\), the lift
\(Q_m\pi\) is simultaneously:
\begin{enumerate}
\item the maximum-entropy projection onto \(\mathcal L_\pi\), and
\item the I-projection of \(\mu\) onto the fiber-uniform family
  \(\mathcal E_m=\{Q_m\eta:\eta\in\Delta(X_m)\}\).
\end{enumerate}
\end{proposition}

\begin{proof}
For any \(\nu\in\mathcal L_\pi\),
Proposition~\ref{prop:fold-entropy-ledger} gives
\[
H(\nu)
=
H(\pi)+\sum_{x\in X_m}\pi(x)\log d_m(x)-\KL(\nu\|Q_m\pi).
\]
The first two terms are independent of \(\nu\), and
\(\KL(\nu\|Q_m\pi)\ge 0\), so \(H(\nu)\le H(Q_m\pi)\) with equality if and
only if \(\nu=Q_m\pi\). The I-projection statement is
Theorem~\ref{thm:fold-kl-pythagoras}.
\end{proof}

\begin{remark}[Relation to classical information projection]\label{rem:fold-kl-pythagoras-classical}
Theorem~\ref{thm:fold-kl-pythagoras} and
Proposition~\ref{prop:dual-projection-maxent} are the present finite-state
instance of Csisz\'ar's information-projection duality
\cite{Csiszar1975IdivergenceGeometry}.  The sufficient statistic is the folded
observable \(\Fold_m\), and \(\mathcal E_m\) is precisely the family of laws
that are uniform on each fiber.
\end{remark}

\begin{corollary}[Microstate decomposition relative to the Parry lift]\label{cor:microstate-parry-lift-decomposition}
Let \(\mu\in\Delta(\Omega_m)\), let \(\pi=(\Fold_m)_*\mu\), and let
\(\mu^e=Q_m\pi\). Then
\[
\KL(\mu\|Q_m\pi_m)
=
\KL(\mu\|\mu^e)+\KL(\pi\|\pi_m).
\]
If \(|\supp(\pi)|\le k\), then
\[
\KL(\mu\|Q_m\pi_m)
\ge
\KL(\mu\|\mu^e)-\log M_m(k).
\]
In particular, assume \(\beta=\alpha\) and \(m\ge 6\). Then
\[
\KL(\mu_m\|Q_m\pi_m)
=
\mathbb E_{\pi_m^{\mathrm{fold}}}\big[\log d_m(Y)\big]
+\KL(\pi_m^{\mathrm{fold}}\|\pi_m),
\]
\[
\KL(\mu_m\|Q_m\pi_m)
\ge
\mathbb E_{\pi_m^{\mathrm{fold}}}\big[\log d_m(Y)\big]
+(m-1)\log\varphi-\log\pi(0)-\log(m+1),
\]
and for every \(N\ge 1\),
\[
\KL(\hat\mu_{m,N}\|Q_m\pi_m)
=
\mathbb E_{\hat\pi_{m,N}}\big[\log d_m(Y)\big]
+\KL(\hat\pi_{m,N}\|\pi_m),
\]
\[
\KL(\hat\mu_{m,N}\|Q_m\pi_m)
\ge
\mathbb E_{\hat\pi_{m,N}}\big[\log d_m(Y)\big]
+(m-1)\log\varphi-\log\pi(0)-\log(m+1).
\]
\end{corollary}

\begin{proof}
The first identity is Theorem~\ref{thm:fold-kl-pythagoras} with \(\eta=\pi_m\).
If \(|\supp(\pi)|\le k\), then
Theorem~\ref{thm:parry-support-barrier} gives
\(\KL(\pi\|\pi_m)\ge -\log M_m(k)\), which yields the second inequality.

Under \(\beta=\alpha\), Corollary~\ref{cor:sturmian-pure-degeneracy-ledger}
identifies \(\KL(\mu_m\|\mu_m^e)\) and \(\KL(\hat\mu_{m,N}\|\hat\mu_{m,N}^e)\)
with the corresponding degeneracy averages. The folded laws
\(\pi_m^{\mathrm{fold}}\) and \(\hat\pi_{m,N}\) are supported on at most \(m+1\)
words, and for \(m\ge 6\), one has \(m+1\le F_m\). Hence
Corollary~\ref{cor:exact-complexity-divergence-law} applies with \(k=m+1\),
yielding the lower bounds.
\end{proof}

\section{A local normalization description of the fold}\label{app:canonical-fold}

This appendix gives a local rewrite description of the fold in
Definition~\ref{def:canonical-fold}.

Let \(\mathcal{D}\) be the set of finitely supported digit configurations
\[
a=(a_1,a_2,\dots),\qquad a_k\in\mathbb{N},
\]
and define the Fibonacci value
\[
\mathrm{Val}(a):=\sum_{k\ge 1} a_k F_{k+1}.
\]
Embed \(\omega=\omega_1\cdots \omega_m\in\Omega_m\) as
\[
\iota_m(\omega):=(\omega_1,\dots,\omega_m,0,0,\dots)\in\mathcal{D}.
\]
Consider the following value-preserving local rewrites:
\[
e_k+e_{k+1}\to e_{k+2},
\qquad
2e_1\to e_2,
\qquad
2e_2\to e_1+e_3,
\qquad
2e_k\to e_{k-2}+e_{k+1}\ \ (k\ge 3),
\]
where \(e_k\) is the \(k\)-th unit vector in \(\mathcal{D}\).

\begin{proposition}[Order-independent local normalization]\label{prop:canonical-fold-rewrite}
Starting from \(\iota_m(\omega)\), every maximal rewrite sequence terminates at
the same normal form \(\mathrm{NF}(\iota_m(\omega))\in\{0,1\}^{\mathbb{N}}\)
with no adjacent \(1\)'s. This normal form is the Zeckendorf expansion
\(Z(N_m(\omega))\). Consequently,
\[
\Fold_m(\omega)=\pref_m\bigl(\mathrm{NF}(\iota_m(\omega))\bigr)
\]
is independent of rewrite order.
\end{proposition}

\begin{proof}
Each rewrite preserves \(\mathrm{Val}\) because it encodes one of the Fibonacci
identities
\[
F_{k+1}+F_{k+2}=F_{k+3},
\qquad
2F_2=F_3,
\qquad
2F_3=F_2+F_4,
\qquad
2F_{k+1}=F_{k-1}+F_{k+2}\ \ (k\ge 3).
\]
Termination follows from a lexicographic descent argument. Let
\[
A(a):=\sum_{k\ge 1} a_k,
\qquad
B(a):=\sum_{k\ge 1} k\,a_k,
\qquad
C(a):=a_2.
\]
The rules \(e_k+e_{k+1}\to e_{k+2}\) and \(2e_1\to e_2\) decrease \(A\); the
rules \(2e_k\to e_{k-2}+e_{k+1}\) for \(k\ge 3\) preserve \(A\) and decrease
\(B\); and the rule \(2e_2\to e_1+e_3\) preserves \(A\) and \(B\) but decreases
\(C\). Hence \((A,B,C)\) decreases lexicographically at each step, so rewriting
must terminate.

A normal form has no adjacent \(1\)'s and no digit \(\ge 2\), hence is a
Zeckendorf representation. Since every rewrite preserves \(\mathrm{Val}\) and
Zeckendorf representation is unique \cite{Zeckendorf1972}, all terminating
branches have the same normal form. The final formula for \(\Fold_m\) is then
just the definition of the canonical fold through prefix truncation.
\end{proof}

\section{Uniqueness of the Fibonacci weighting}\label{app:fibonacci-uniqueness}

The preceding appendix describes the canonical fold by Zeckendorf
normalization.  The next result records one sense in which the Fibonacci
weights are forced.

\begin{theorem}[Contiguous numeration uniqueness]
\label{thm:fibonacci-weight-uniqueness}
Let \(w_2,w_3,\dots\) be positive integers. Define the valuation
\[
V_m^{(w)}(x_1\cdots x_m):=\sum_{k=1}^m x_k\,w_{k+1}.
\]
If for every \(m\ge 1\) the map \(V_m^{(w)}\) restricts to a bijection from
\(X_m\) onto \(\{0,1,\dots,F_{m+2}-1\}\), then \(w_{k+1}=F_{k+1}\) for all
\(k\ge 1\).
\end{theorem}

\begin{proof}
For \(m=1\), \(X_1=\{0,1\}\) and the bijection condition gives
\(\{0,w_2\}=\{0,1\}\), so \(w_2=1=F_2\).

For \(m=2\), \(X_2=\{00,01,10\}\) and \(\{0,w_2,w_3\}=\{0,1,2\}\); since
\(w_2=1\), one obtains \(w_3=2=F_3\).

Proceed by induction. Suppose \(w_2=F_2,\dots,w_m=F_m\) for some
\(m\ge 3\). Decompose \(X_m\) by its last symbols:
\[
X_m
=
\{u0: u\in X_{m-1}\}
\;\sqcup\;
\{v01: v\in X_{m-2}\}.
\]
On the first part, \(V_m^{(w)}(u0)=V_{m-1}^{(w)}(u)\). On the second,
\(V_m^{(w)}(v01)=V_{m-2}^{(w)}(v)+w_{m+1}\). By the inductive hypothesis,
\[
V_{m-1}^{(w)}(X_{m-1})=\{0,\dots,F_{m+1}-1\},
\quad
V_{m-2}^{(w)}(X_{m-2})=\{0,\dots,F_m-1\}.
\]
Hence
\[
V_m^{(w)}(X_m)
=
\{0,\dots,F_{m+1}-1\}
\;\sqcup\;
\{w_{m+1},\dots,w_{m+1}+F_m-1\}.
\]
For this to equal \(\{0,\dots,F_{m+2}-1\}=\{0,\dots,F_{m+1}+F_m-1\}\), the
second interval must start at \(F_{m+1}\), giving \(w_{m+1}=F_{m+1}\).
\end{proof}

\begin{corollary}[Uniqueness of the canonical residue fold]
\label{cor:canonical-fold-uniqueness}
Let \((G_m)_{m\ge1}\), \(G_m:\Omega_m\to X_m\), be a family of deterministic
maps satisfying the following conditions for a single positive integer weight
sequence \(w=(w_2,w_3,\dots)\):
\begin{enumerate}
\item \(G_m(x)=x\) for all \(m\ge1\) and all \(x\in X_m\).
\item For every \(m\ge1\), the restriction \(V_m^{(w)}|_{X_m}\) is a
  bijection onto \(\{0,\dots,F_{m+2}-1\}\).
\item For every \(m\ge1\) and every \(\omega\in\Omega_m\),
  \(V_m^{(w)}(G_m(\omega))\equiv \sum_{k=1}^m\omega_k w_{k+1}
  \pmod{F_{m+2}}\).
\end{enumerate}
Then \(G_m=\Fold_m\) for every \(m\ge1\).
\end{corollary}

\begin{proof}
Theorem~\ref{thm:fibonacci-weight-uniqueness} forces \(w_{k+1}=F_{k+1}\).
For each fixed \(m\), condition~2 then provides a unique legal representative
in each residue class modulo \(F_{m+2}\), and condition~3 selects that
representative. This is exactly the residue characterization of
Proposition~\ref{prop:fold-residue-characterization}, so \(G_m=\Fold_m\).
\end{proof}

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
