\documentclass[../../../main.tex]{subfiles}

\begin{document}
\input{sec__zeta-finite-part}
\end{document}

% --- Backflow from publication papers (auto-generated) ---
\begin{remark}[Results proven in dedicated publications]
  The following results from this section have been developed into
  standalone publications with full proofs and independent verification:
  \begin{itemize}
    \item \textbf{Reduced-determinant gradient formula} (Theorem \texttt{thm:finite-part-reduced-determinant-group-inverse-gradient}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Holomorphic variation formula} (Theorem \texttt{thm:logM-holomorphic-variation}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Truncation bound} (Theorem \texttt{thm:logM-truncation-bound}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Parseval identity and aliasing formula} (Theorem \texttt{thm:finite-part-nyquist-parseval-aliasing}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Exponential bound for the sampling error} (Corollary \texttt{cor:finite-part-nyquist-error-exp}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Gauss expansion for character sums at roots of unity} (Proposition \texttt{prop:finite-part-dirichlet-torsion-gauss-expansion}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Character inversion for prime moduli} (Theorem \texttt{thm:finite-part-dirichlet-character-inversion-prime}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{A growth criterion for the square-root bound} (Corollary \texttt{cor:finite-part-near-rh-q-axis-envelope}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Boundary multiplicities from Cesaro averages} (Theorem \texttt{thm:finite-part-boundary-multiplicity-qaxis-energy}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Galois covariance of root-of-unity values} (Proposition \texttt{prop:finite-part-torsion-table-galois-covariance}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{A rationality criterion from one root-of-unity layer} (Corollary \texttt{cor:finite-part-torsion-table-galois-certificate}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{A finite criterion for the square-root bound} (Theorem \texttt{thm:finite-part-single-layer-torsion-near-rh}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{A Dirichlet generating series for the cyclic-lift discrepancy} (Theorem \texttt{thm:finite-part-cyclic-lift-polylog-dirichlet-xi}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Bounds for the reduced determinant under the square-root bound} (Proposition \texttt{prop:finite-part-residue-constant-rh-squeeze}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Linear growth of the reduced determinant} (Theorem \texttt{thm:finite-part-residue-drift-trichotomy}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Conjugacy-class expansion} (Proposition \texttt{prop:class-indicator-expansion}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Peter--Weyl determinant factorisation} (Proposition \texttt{prop:kernel-peter-weyl-block-diagonalisation}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Twisted spectral-radius bound} (Proposition \texttt{prop:twisted-spectral-radius-bound}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Class-function determinant linearisation} (Theorem \texttt{thm:class-function-linearisation}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Adams--M\"obius primitive inversion} (Corollary \texttt{cor:class-function-adams-mobius}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Irreducible-channel determinant formula} (Corollary \texttt{cor:primitive-peter-weyl-determinant}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Primitive Peter--Weyl trace formula} (Corollary \texttt{cor:primitive-peter-weyl-trace}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Conjugacy-class primitive decomposition} (Theorem \texttt{thm:class-primitive-decomposition}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Conjugacy-class Artin--M\"obius trace formula} (Corollary \texttt{cor:class-artin-mobius-trace}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Explicit class Mertens theorem and Chebotarev density
  under twisted spectral gap} (Theorem \texttt{thm:class-mertens-explicit}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Class-constant deviations and Perron-point
  determinant evaluation} (Proposition \texttt{prop:class-mertens-deviations}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Non-abelian Fourier reconstruction of class Mertens
  constants} (Theorem \texttt{thm:class-mertens-fourier}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Primitive-orbit closed form for the finite part} (Theorem \texttt{thm:finite-part-primitive-closed-form}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Reduced determinant identity} (Proposition \texttt{prop:finite-part-residue-reduced-determinant}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Orbit-Mertens asymptotic} (Corollary \texttt{cor:finite-part-mertens-asymptotic}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{A formula for the reduced determinants of cyclic lifts} (Theorem \texttt{thm:finite-part-cyclic-lift-reduced-constant-closed}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Multiple-sum expansion for cyclic lifts} (Proposition \texttt{prop:finite-part-cyclic-lift-dirichlet-multiple-sum}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{M\"obius inversion on multiples} (Proposition \texttt{prop:finite-part-cyclic-lift-mobius-inversion}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Determination of the reduced spectrum from cyclic lifts} (Theorem \texttt{thm:finite-part-cyclic-lift-spectrum-identifiability}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Reconstruction from one root-of-unity layer} (Theorem \texttt{thm:finite-part-cyclic-lift-single-q-torsion-reconstruction}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Root-of-unity values of \(\zeta_A\) determine the
  reduced spectrum} (Corollary \texttt{cor:finite-part-cyclic-lift-zeta-torsion-determines-spectrum}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Finite square-root test from one root-of-unity layer} (Corollary \texttt{cor:finite-part-single-layer-square-root-test}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Positivity of the primitive gap} (Corollary \texttt{cor:finite-part-gap-positive}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Theorem A: Finite-part formula at the Perron pole} (Theorem \texttt{thm:intro-finite-part}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Theorem B: Cyclic reconstruction of the reduced spectrum} (Theorem \texttt{thm:intro-cyclic-reconstruction}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Theorem C: Primitive determinant calculus for
  finite-group extensions} (Theorem \texttt{thm:intro-finite-group}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Characteristic polynomial and residue constant} (Proposition \texttt{prop:sync-kernel-spectrum}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Explicit completed determinant} (Proposition \texttt{prop:sync-hatdelta-completion}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{The quotient is a smooth plane quartic} (Proposition \texttt{prop:sync-hatdelta-quotient}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Two-sheeted cover of genus six} (Proposition \texttt{prop:sync-hatdelta-double-cover}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Generic Galois group} (Proposition \texttt{prop:sync-hatdelta-galois-s6-generic}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Cyclotomic resultant degree law} (Theorem \texttt{thm:sync-cyclotomic-degree-law}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{The worst twist $m=2$} (Corollary \texttt{cor:rho-m2-closed-form}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Endpoint jet} (Proposition \texttt{prop:sync-endpoint-jet}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Asymptotic expansion of the twisted Perron roots} (Theorem \texttt{thm:rho-gap-m12}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Cyclic-lift primitive asymptotics} (Proposition \texttt{prop:cyclic-lift-primitive-asymptotics}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Fredholm--Witt bridge} (Theorem \texttt{thm:fredholm-witt-bridge}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Cyclic-block Fredholm realisation} (Theorem \texttt{thm:cyclic-fredholm-realisation}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Spectral-gap central limit theorem} (Theorem \texttt{thm:clt-spectral}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Trace, primitive orbits, and Euler product} (Proposition \texttt{prop:prelim-trace-primitive}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Quotient functoriality} (Theorem \texttt{thm:quotient-functoriality}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Abelian and cyclic shadows} (Corollary \texttt{cor:abelian-cyclic-shadow}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Abelian shadow and genuine non-abelian defect} (Theorem \texttt{thm:abelian-shadow-defect}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Exact boundary of cyclic detection} (Theorem \texttt{thm:cyclic-detection-boundary}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Class indicator for \(C_3\)} (Proposition \texttt{prop:c3-class-indicator-check}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Quadratic Adams obstruction} (Proposition \texttt{thm:quadratic-adams-obstruction}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Twisted channels for the \(S_3\)-example} (Proposition \texttt{prop:s3-example-channels}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Primitive channels for the \(S_3\)-example} (Proposition \texttt{prop:s3-example-primitive-channels}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Primitive profile and class constants for the
  \(S_3\)-example} (Corollary \texttt{cor:s3-example-profile}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Density dichotomy for DFA languages} (Theorem \texttt{thm:dfa-density-dichotomy}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Symmetric-difference lower bound for binary primes} (Corollary \texttt{cor:dfa-prime-symmetric-diff}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{No simultaneous recall and precision for fixed DFA} (Corollary \texttt{cor:dfa-prime-recall-precision}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Regular Zeckendorf languages have power-law growth} (Theorem \texttt{thm:zeckendorf-regular-powerlaw}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Prime Zeckendorf language is not regular} (Corollary \texttt{cor:zeckendorf-prime-language-not-regular}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Prime Zeckendorf shift is not sofic} (Theorem \texttt{thm:zeckendorf-primes-not-sofic}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Prime-supported Euler products have a natural boundary} (Theorem \texttt{thm:euler-product-natural-boundary}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Imaginary-periodicity rigidity of finite determinants} (Proposition \texttt{prop:finite-zeta-imaginary-periodicity}) --- Ergodic Theory and Dynamical Systems paper \texttt{dynamical_zeta_finite}, status: ACCEPT.
    \item \textbf{Connecting Fredholm--Witt identity} (Proposition \texttt{thm:fredholm-witt-bridge}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Cyclic-block characterization and optimality} (Theorem \texttt{thm:cyclic-fredholm-realisation}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Trace-norm rigidity and support optimality} (Corollary \texttt{thm:block-cyclic-rigidity}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Finite Euler products give finite-rank models} (Corollary \texttt{cor:finite-rank-realisation}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Spectral rigidity of Fredholm determinants} (Proposition \texttt{thm:fredholm-spectral-rigidity}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Realisability versus rigidity} (Corollary \texttt{cor:realisability-vs-rigidity}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Connecting Fredholm--Witt identity} (Proposition \texttt{prop:intro-fredholm-witt}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Cyclic-block characterization and optimality} (Theorem \texttt{thm:intro-cyclic-characterization}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Spectral rigidity} (Proposition \texttt{prop:intro-spectral-rigidity}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Reduced-determinant gradient formula} (Proposition \texttt{thm:finite-part-reduced-determinant-group-inverse-gradient}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Normal convergence of the secondary Fredholm factors} (Proposition \texttt{prop:secondary-fredholm-normal-convergence}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Holomorphic variation of the Abel finite part} (Proposition \texttt{thm:logM-holomorphic-variation}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Determinant truncation for the Abel finite part} (Proposition \texttt{prop:logM-truncation}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Twisted determinants encode periodic-orbit moments} (Proposition \texttt{prop:twisted-determinant-orbit-moments}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Twisted determinant rigidity for locally constant observables} (Proposition \texttt{thm:twisted-determinant-rigidity}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Cumulant rigidity} (Corollary \texttt{cor:twisted-determinant-cumulants}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Off-origin decay from a spectral-radius gap} (Proposition \texttt{prop:off-origin-decay}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Spectral-gap central limit theorem} (Proposition \texttt{thm:clt-spectral}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Trace, primitive orbits, and Euler product} (Proposition \texttt{prop:prelim-trace-primitive}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{Primitive-orbit closed form} (Theorem \texttt{thm:finite-part-primitive-closed-form}) --- Journal of Spectral Theory paper \texttt{fredholm_witt_cyclic}, status: ACCEPT.
    \item \textbf{thm:intro-kernel-geometry} (Theorem \texttt{thm:intro-kernel-geometry}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{thm:intro-kernel-cyclotomic} (Theorem \texttt{thm:intro-kernel-cyclotomic}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{Explicit determinant} (Proposition \texttt{prop:kernel-delta-explicit}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{Characteristic polynomial and residue constant} (Proposition \texttt{prop:sync-kernel-spectrum}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{Explicit completed determinant} (Proposition \texttt{prop:sync-hatdelta-completion}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{The quotient is a smooth plane quartic} (Proposition \texttt{prop:sync-hatdelta-quotient}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{Two-sheeted cover of genus six} (Proposition \texttt{prop:sync-hatdelta-double-cover}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{Generic Galois group} (Proposition \texttt{prop:sync-hatdelta-galois-s6-generic}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{Cyclotomic resultant degree law} (Theorem \texttt{thm:sync-cyclotomic-degree-law}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{The twist \(m=2\)} (Corollary \texttt{cor:rho-m2-closed-form}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{Endpoint jet} (Proposition \texttt{prop:sync-endpoint-jet}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{Asymptotic expansion of the twisted Perron roots} (Theorem \texttt{thm:rho-gap-m12}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{Cyclic-lift primitive asymptotics} (Proposition \texttt{prop:cyclic-lift-primitive-asymptotics}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
    \item \textbf{Self-duality} (Proposition \texttt{prop:sync-self-duality}) --- IMRN paper \texttt{self_dual_synchronisation}, status: ACCEPT.
  \end{itemize}
\end{remark}
% --- End backflow ---
