\begin{theorem}[\texorpdfstring{$\log\mathfrak{M}(\theta)$}{logM(theta)} 的解析性与变分公式]\label{thm:finite-part-logM-holomorphic-variation}
在小节 \ref{subsec:logM-analyticity} 的设定下，设 \(\theta\mapsto M_\theta\) 在某复邻域内全纯，并假设 \(\lambda(\theta)=\rho(M_\theta)\) 在该邻域内为简单特征值；记 \(z_\star(\theta)=\lambda(\theta)^{-1}\) 且 \(\zeta_\theta(z)=\det(I-zM_\theta)^{-1}\)。
则 \(\theta\mapsto \log\mathfrak{M}(\theta)\) 在该邻域内全纯，并且满足变分恒等式
\[
\boxed{\;
\frac{d}{d\theta}\log\mathfrak{M}(\theta)
=\frac{d}{d\theta}\log C(\theta)
\;+\;
\sum_{k\ge 2}\frac{\mu(k)}{k}\,\frac{d}{d\theta}\log \zeta_\theta\!\bigl(z_\star(\theta)^k\bigr)\; }.
\]
此外，对每个固定 \(z\)（满足 \(|z|<|z_\star(\theta)|\)）有基本微分恒等式
\[
\boxed{\;
\partial_\theta \log \zeta_\theta(z)
=
\mathrm{Tr}\!\Bigl((I-zM_\theta)^{-1}\,z\,\partial_\theta M_\theta\Bigr)\; }.
\]
\end{theorem}

\begin{proof}
由 \(\log\mathfrak{M}(\theta)=\log C(\theta)+\sum_{k\ge 2}\frac{\mu(k)}{k}\log \zeta_\theta(z_\star(\theta)^k)\) 的接口式（见 \S\ref{app:real-input-40-logm-theta}），对 \(k\ge 2\) 有 \(|z_\star(\theta)^k|<|z_\star(\theta)|\)，从而各项在邻域内全纯且尾部一致收敛，故和函数全纯并可逐项求导，得到第一式。
\par
对第二式，记 \(A_\theta(z)=I-zM_\theta\)，则 \(\log\zeta_\theta(z)=-\log\det A_\theta(z)\)。
对 \(\theta\) 求导并用 \(\partial_\theta\log\det A=\mathrm{Tr}(A^{-1}\partial_\theta A)\) 与 \(\partial_\theta A_\theta(z)=-z\,\partial_\theta M_\theta\) 即得。证毕。
\end{proof}
\leanverified{paper\_finite\_part\_logM\_holomorphic\_variation}

\begin{theorem}[\texorpdfstring{$\log\mathfrak{M}(\theta)$}{logM(theta)} 的可审计截断误差界]\label{thm:finite-part-logM-truncation-explicit-bound}
沿用定理 \ref{thm:finite-part-logM-holomorphic-variation} 的设定，并记 \(d=\dim M_\theta\)、\(\lambda=\lambda(\theta)>1\)。
定义 \(K\) 阶截断
\[
\log\mathfrak{M}_K(\theta):=\log C(\theta)+\sum_{k=2}^{K}\frac{\mu(k)}{k}\log \zeta_\theta(\lambda^{-k}).
\]
则对所有整数 \(K\ge 2\) 有显式误差上界
\[
\boxed{\;
\bigl|\log\mathfrak{M}(\theta)-\log\mathfrak{M}_K(\theta)\bigr|
\le
\frac{d}{K+1}\cdot\frac{\lambda^{-K}}{(1-\lambda^{-1})^2}\; }.
\]
\end{theorem}

\begin{proof}
写出特征值分解（按代数重数计）\(\det(I-zM_\theta)=\prod_{j=1}^{d}(1-\lambda_j z)\)，其中 \(|\lambda_j|\le\lambda\)。
当 \(k\ge 2\) 且 \(z=\lambda^{-k}\) 时，有 \(|\lambda_j z|\le \lambda^{1-k}\le\lambda^{-1}<1\)。
由不等式 \(|\log(1-w)|\le |w|/(1-|w|)\)（\(|w|<1\)）得
\[
|\log \zeta_\theta(\lambda^{-k})|
=\left|\sum_{j=1}^{d}\log(1-\lambda_j\lambda^{-k})^{-1}\right|
\le \sum_{j=1}^{d}\frac{|\lambda_j|\lambda^{-k}}{1-|\lambda_j|\lambda^{-k}}
\le d\cdot \frac{\lambda^{1-k}}{1-\lambda^{-1}}.
\]
因此
\[
\bigl|\log\mathfrak{M}(\theta)-\log\mathfrak{M}_K(\theta)\bigr|
\le
\sum_{k>K}\frac{1}{k}\,|\log \zeta_\theta(\lambda^{-k})|
\le
\frac{d}{1-\lambda^{-1}}\sum_{k>K}\frac{\lambda^{1-k}}{k}.
\]
对 \(k>K\) 有 \(1/k\le 1/(K+1)\)，故
\[
\sum_{k>K}\frac{\lambda^{1-k}}{k}
\le
\frac{1}{K+1}\sum_{k>K}\lambda^{1-k}
=
\frac{1}{K+1}\cdot\frac{\lambda^{-K}}{1-\lambda^{-1}}.
\]
合并即得所示界。证毕。
\end{proof}
\leanverified{paper\_finite\_part\_logM\_truncation\_explicit\_bound}

\begin{corollary}[截断层间的比例传递与穷竭误差]\label{cor:finite-part-logM-relative-scale-transfer}
沿用定理 \ref{thm:finite-part-logM-truncation-explicit-bound} 的记号。则对任意整数 \(L>K\ge 2\)，有
\[
\bigl|\log\mathfrak{M}_L(\theta)-\log\mathfrak{M}_K(\theta)\bigr|
\le
\frac{d}{K+1}\cdot\frac{\lambda^{-K}}{(1-\lambda^{-1})^2}.
\]
若进一步有 \(\log\mathfrak{M}(\theta)\neq 0\)，则
\[
\left|
\frac{\log\mathfrak{M}_K(\theta)}{\log\mathfrak{M}(\theta)}-1
\right|
\le
\frac{d}{(K+1)\,|\log\mathfrak{M}(\theta)|}\cdot\frac{\lambda^{-K}}{(1-\lambda^{-1})^2}.
\]
\end{corollary}

% --- Distillation writeback ---
\begin{theorem}[completed Euler denominator ledger and the commensurable--singular alternative]\label{distill:euclid_elements-proportion_and_denominator_spectrum_families-zeta-completed-denominator-ledger-alternative}
\textbf{Date-time: 2026-04-23T12:21:04+08:00. Family: Euclid--Elements / proportion and denominator-spectrum families. Dependency status: conditional audit.}
Work in the setting of Theorem \ref{thm:finite-part-logM-holomorphic-variation}, and fix one parameter value \(\theta\).  Put \(\lambda=\lambda(\theta)>1\) and write \(M=M_\theta\).  Define the completed Euler denominator ledger
\[
\widehat{\mathcal L}:=\prod_{k\ge2}\prod_{r\ge1}\mathbb C\,\mathbf e_{k,r}
\]
and the coefficient family
\[
A_{k,r}(\theta):=\frac{\mu(k)}{kr}\,\lambda^{-kr}\,\operatorname{Tr}(M^r)
\qquad(k\ge2,\ r\ge1).
\]
Then
\[
\mathfrak D_\theta:=\bigl(A_{k,r}(\theta)\bigr)_{k,r}\in\widehat{\mathcal L}
\]
is a well-defined completed ledger, and its scalar evaluation is absolutely convergent:
\[
\sum_{k\ge2}\sum_{r\ge1}|A_{k,r}(\theta)|<\infty.
\]
Moreover, if
\[
\operatorname{Supp}^{\mu}(\mathfrak D_\theta):=\bigl\{(k,r):\mu(k)\operatorname{Tr}(M^r)\ne0\bigr\},
\]
then the denominator spectrum
\[
\mathsf{Den}(\theta):=\bigl\{k\ge2:\exists r\ge1\text{ with }(k,r)\in\operatorname{Supp}^{\mu}(\mathfrak D_\theta)\bigr\}
\]
satisfies the following exact alternative.
\begin{enumerate}
  \item \textbf{Commensurable periodic case.}  If there exists \(N\in\mathbb N_{>0}\) such that \(k\mid N\) for every \(k\in\mathsf{Den}(\theta)\), then the whole non-zero Euler denominator support factors through the finite divisibility lattice of \(N\).  Equivalently, every active denominator atom is periodic with respect to the single common denominator ring \(\mathbb Z/N\mathbb Z\).
  \item \textbf{Singular finite-prime case.}  If \(\mathsf{Den}(\theta)\) is supported on a finite prime set \(P\) but is not contained in the divisor lattice of any one \(N\), then for every height bound \(H\) there is an active atom \((k,r)\in\operatorname{Supp}^{\mu}(\mathfrak D_\theta)\) such that all prime divisors of \(k\) lie in \(P\) and
  \[
  \sum_{p\in P}v_p(k)>H.
  \]
  Thus the obstruction is not the appearance of a new prime but unbounded \(P\)-adic denominator height.
  \item \textbf{Incommensurable multiprime case.}  If the union of prime divisors of elements of \(\mathsf{Den}(\theta)\) is infinite, then no finite prime-register can contain the active Euler denominator support.
\end{enumerate}
For a finite audit window \(\mathcal W\subset\{(k,r):k\ge2,r\ge1\}\), the same statements hold for the certified subledger obtained by replacing \(\operatorname{Supp}^{\mu}(\mathfrak D_\theta)\) with
\[
\operatorname{Supp}^{\mu}_{\mathcal W}(\mathfrak D_\theta)
:=\bigl\{(k,r)\in\mathcal W:\mu(k)\operatorname{Tr}(M^r)\ne0\text{ has been certified}\bigr\}.
\]
Any conclusion about the full ledger requires, in addition, zero certificates for every atom outside the certified non-zero support.  Atoms with \(\mu(k)=0\) are not denominator obstructions for this Euler ledger.
\end{theorem}

% --- Distillation writeback ---
% --- Distillation writeback: Euclid exhaustion finite-budget closure ---
\begin{theorem}[squarefree Euler-tail exhaustion and stabilized denominator certificates]\label{distill:euclid_elements-exhaustion_and_finite_budget_closure_families-zeta-squarefree-euler-tail-exhaustion}
\textbf{Date-time: 2026-04-23T12:38:38+08:00. Family: Euclid--Elements / exhaustion and finite-budget closure families. Dependency status: rigidity.}
Work in the setting of Theorem \ref{thm:finite-part-logM-holomorphic-variation}, and fix \(\theta\).  Put \(M=M_\theta\), \(d=\dim M\), and \(\lambda=\rho(M)>1\).  For \(k\ge2\) and \(r\ge1\), define the Euler coefficient atom
\[
A_{k,r}(\theta):=\frac{\mu(k)}{kr}\,\lambda^{-kr}\operatorname{Tr}(M^r).
\]
For \(K\ge2\), define the absolute tail budget
\[
\mathcal E_K(\theta):=\sum_{k>K}\sum_{r\ge1}|A_{k,r}(\theta)|.
\]
Then
\[
\boxed{\;
\mathcal E_K(\theta)\le
\frac{d}{K+1}\cdot\frac{\lambda^{-K}}{(1-\lambda^{-1})^2}.
\;}
\]
In particular, the completed Euler ledger is exhausted by finite denominator windows with an explicit residual budget tending to zero.

Let
\[
\mathsf{Den}(\theta):=\{k\ge2:\exists r\ge1\text{ with }\mu(k)\operatorname{Tr}(M^r)\ne0\},
\qquad
\mathsf{Den}_{\le K}(\theta):=\mathsf{Den}(\theta)\cap\{2,\ldots,K\}.
\]
Then every active denominator is squarefree.  Consequently the finite denominator certificates \(\mathsf{Den}_{\le K}(\theta)\) stabilize as \(K\to\infty\) if and only if the active prime support
\[
\mathsf P(\theta):=\bigcup_{k\in\mathsf{Den}(\theta)}\{p:p\mid k\}
\]
is finite.  In that case all active denominators lie in the divisor lattice of the single squarefree integer
\[
N_\theta:=\prod_{p\in\mathsf P(\theta)}p.
\]
If the certificates do not stabilize, then \(\mathsf P(\theta)\) is infinite; equivalently, for every finite prime register \(P_0\) and every height bound \(K_0\), there exists an active squarefree denominator \(k>K_0\) with a prime divisor outside \(P_0\).  Thus, under the present \(\lambda>1\) Euler finite-part hypotheses, the completed denominator-ledger alternative specializes to a squarefree dichotomy: finite commensurable support versus an incommensurable infinite-prime tail.  The singular finite-prime height branch of the broader denominator-ledger alternative cannot occur here, because \(\mu(k)\ne0\) forces \(k\) to be squarefree.
\end{theorem}

% --- Distillation writeback ---
\begin{proposition}[有限 Euler 窗口归一化的 \v Cech 秩界与规范唯一性]\label{distill:cech_cohomology-finite_nerve_obstruction_bounds-zeta-euler-window-normalization-gluing}
沿用定理 \ref{distill:euclid_elements-proportion_and_denominator_spectrum_families-zeta-completed-denominator-ledger-alternative} 中的 completed Euler denominator ledger 记号。固定一个有限 audit window
\[
\mathcal W\subset\{(k,r):k\ge2,\ r\ge1\},
\]
并令 \(\{U_i\}_{i\in I}\) 为 \(\mathcal W\) 的有限覆盖，\(\mathcal N\) 为其 nerve。设 \(\mathcal N\) 的一骨架连通。令 \(A\) 为有限阿贝尔规范群，并假设该有限窗口的局部 Euler 归一化审计具有如下形式：每个顶点 \(i\) 上给出局部归一化选择，任意非空重叠边 \(ij\) 上的两种选择之差由元素
\[
g_{ij}\in A,\qquad g_{ji}=-g_{ij}
\]
记录，且三重重叠相容性正是
\[
g\in Z^1(\mathcal N;A).
\]
顶点规范 \(a=(a_i)\in C^0(\mathcal N;A)\) 的作用约定为把边账本改为
\[
g'_{ij}=g_{ij}-(a_j-a_i).
\]
并规定该有限窗口的归一化可粘接，当且仅当存在某个顶点规范使所有 \(g'_{ij}\) 同时为零。

则该有限窗口中所有可能的归一化粘接障碍都落在
\[
[g]\in\check H^1(\mathcal N;A),
\]
其可登记障碍秩不超过 \(\mathrm d(\check H^1(\mathcal N;A))\)，其中 \(\mathrm d\) 为有限阿贝尔群的最小生成元数。若 \(\check H^1(\mathcal N;A)=0\)，则有限窗口归一化总可粘接；并且在固定一个根顶点 \(o\in I\) 且要求 \(a_o=0\) 后，杀死 \(g\) 的顶点规范唯一。未固定根时，该规范唯一性只成立到全局 \(A\)-平移为止。

因此，在该有限 \((\mathcal W,\mathcal N,A)\) 审计接口内，\(\check H^1(\mathcal N;A)=0\) 排除了 Euler 窗口归一化的有限 \v Cech holonomy 失败；对完整 \(\mathfrak D_\theta\) 的任何剩余认证失败仍可能来自窗口外零证书缺失、Euler 尾预算或解析截断误差，而不是该有限窗口的粘接类。
\end{proposition}
\begin{proof}
由于三重重叠相容性被假设为 \(g\in Z^1(\mathcal N;A)\)，边账本定义了上同调类 \([g]\in\check H^1(\mathcal N;A)\)。若顶点规范 \(a\) 被施加，则
\[
g' = g-\delta a.
\]
因此 \([g']=[g]\)。换言之，\([g]\) 是有限窗口归一化审计在允许规范下的不变量。

按粘接规则，有限窗口可粘接当且仅当存在 \(a\in C^0(\mathcal N;A)\) 使 \(g-\delta a=0\)，也就是当且仅当 \([g]=0\)。所以所有可能的粘接障碍构成 \(\check H^1(\mathcal N;A)\) 的子集；它们生成的有限阿贝尔子群所需生成元数不超过 \(\check H^1(\mathcal N;A)\) 本身的最小生成元数。由于 \(\mathcal N\) 与 \(A\) 均有限，该上同调群可由有限 incidence 矩阵的 Smith 标准形计算。

若 \(\check H^1(\mathcal N;A)=0\)，则每个上循环 \(g\) 都是余边界，故存在 \(a\) 使 \(g=\delta a\)。施加该规范后 \(g'=0\)，从而有限窗口归一化可粘接。

最后证明唯一性陈述。设 \(a,b\in C^0(\mathcal N;A)\) 都满足 \(\delta a=g=\delta b\)。则 \(c:=a-b\) 满足 \(\delta c=0\)，即对每条边 \(ij\) 都有 \(c_j-c_i=0\)。由于 \(\mathcal N\) 的一骨架连通，沿任意路径传递可得所有 \(c_i\) 相等。因此两个规范相差一个全局常量 \(c\in A\)。若再规定根顶点 \(o\) 上 \(a_o=b_o=0\)，则该常量为零，故 \(a=b\)。这说明零上同调给出的是存在性与模全局规范的唯一性，而非未定规范下的绝对唯一性。

完整 Euler denominator ledger 还包含 \(\mathcal W\) 外的原子与无穷尾部。上述论证只处理固定有限窗口的 \v Cech 粘接类；因此即使该类消失，窗口外零证书、尾和预算或解析截断仍可独立失败。证毕。
\end{proof}

% --- End distillation writeback ---

\begin{proof}
Let \(\alpha_1,\ldots,\alpha_d\) be the eigenvalues of \(M\), counted with algebraic multiplicity.  Since \(\lambda=\rho(M)\), one has \(|\alpha_i|\le\lambda\), and hence
\[
|\operatorname{Tr}(M^r)|\le\sum_{i=1}^d|\alpha_i|^r\le d\lambda^r.
\]
Therefore
\[
|A_{k,r}(\theta)|
\le \frac{d}{kr}\lambda^{r-kr}
=\frac{d}{kr}\lambda^{-(k-1)r}.
\]
Summing over \(k>K\) and \(r\ge1\), and using \(1/k\le1/(K+1)\), gives
\[
\mathcal E_K(\theta)\le
\frac{d}{K+1}\sum_{k>K}\sum_{r\ge1}\frac{\lambda^{-(k-1)r}}{r}.
\]
Since \(r^{-1}\le1\) and \(k>K\ge2\),
\[
\sum_{r\ge1}\frac{\lambda^{-(k-1)r}}{r}\le
\sum_{r\ge1}\lambda^{-(k-1)r}\le
\frac{\lambda^{-(k-1)}}{1-\lambda^{-1}}.
\]
Thus
\[
\mathcal E_K(\theta)\le
\frac{d}{K+1}\cdot\frac1{1-\lambda^{-1}}
\sum_{k>K}\lambda^{-(k-1)}
=\frac{d}{K+1}\cdot\frac{\lambda^{-K}}{(1-\lambda^{-1})^2},
\]
which is the asserted residual budget.

It remains to identify which finite certificates stabilize.  If \(k\in\mathsf{Den}(\theta)\), then \(\mu(k)\ne0\).  The Möbius function vanishes exactly on integers divisible by a square; hence every active denominator \(k\) is squarefree.

Suppose first that \(\mathsf P(\theta)\) is finite.  Since every active \(k\) is squarefree and uses only primes in \(\mathsf P(\theta)\), it divides
\[
N_\theta=\prod_{p\in\mathsf P(\theta)}p.
\]
There are only finitely many such divisors, so \(\mathsf{Den}(\theta)\) is finite and the increasing certificates \(\mathsf{Den}_{\le K}(\theta)\) stabilize once \(K\ge\max\mathsf{Den}(\theta)\), with the empty case understood in the evident way.

Conversely, if \(\mathsf{Den}_{\le K}(\theta)\) stabilizes, then \(\mathsf{Den}(\theta)\) is finite, and therefore the union of its prime divisors is finite.  This proves the equivalence between stabilization and finite active prime support.

If stabilization fails, \(\mathsf{Den}(\theta)\) is infinite.  Were \(\mathsf P(\theta)\) finite, squarefreeness would force \(\mathsf{Den}(\theta)\) into the finite divisor lattice of \(\prod_{p\in\mathsf P(\theta)}p\), a contradiction.  Hence \(\mathsf P(\theta)\) is infinite.  Given a finite prime register \(P_0\) and a bound \(K_0\), only finitely many active denominators \(k\le K_0\) exist, and they involve only finitely many primes.  Since \(\mathsf P(\theta)\) is infinite, choose an active denominator involving a prime outside the union of those finitely many primes and outside \(P_0\).  That denominator is squarefree and necessarily exceeds \(K_0\).

The analytic tail estimate and the denominator-certificate statement therefore separate two notions of limiting closure: scalar Euler tails are controlled by the explicit budget \(\mathcal E_K(\theta)\), while finite denominator certificates stabilize only in the finite-prime commensurable case.  This proves the theorem.
\end{proof}
% --- End distillation writeback ---

% --- End distillation writeback ---

\begin{proof}
Let \(\alpha_1,\ldots,\alpha_d\) be the eigenvalues of \(M\), counted with algebraic multiplicity.  Since \(\lambda=\rho(M)\), one has \(|\alpha_j|\le\lambda\), and hence
\[
|\operatorname{Tr}(M^r)|\le\sum_{j=1}^{d}|\alpha_j|^r\le d\lambda^r.
\]
Therefore
\[
|A_{k,r}(\theta)|
\le \frac{d}{kr}\lambda^{r-kr}
=\frac{d}{kr}\lambda^{-(k-1)r}.
\]
Summing first in \(k\) and then in \(r\),
\[
\sum_{k\ge2}\sum_{r\ge1}|A_{k,r}(\theta)|
\le d\sum_{k\ge2}\frac1k\sum_{r\ge1}\frac{\lambda^{-(k-1)r}}{r}
\le d\sum_{k\ge2}\frac1k\sum_{r\ge1}\lambda^{-(k-1)r}.
\]
The inner geometric series is \(\lambda^{-(k-1)}/(1-\lambda^{-(k-1)})\), which is bounded by \(C\lambda^{-(k-1)}\) for a constant \(C=C(\lambda)\).  Thus the last series is dominated by
\[
 dC\sum_{k\ge2}\frac{\lambda^{-(k-1)}}{k}<\infty.
\]
This proves both that \(\mathfrak D_\theta\) is a legitimate element of the completed product ledger and that its scalar evaluation is absolutely convergent.  The use of \(\operatorname{Supp}^{\mu}\) is essential: if \(\mu(k)=0\), then every coefficient with that denominator is zero regardless of \(\operatorname{Tr}(M^r)\), so such an atom cannot contribute an Euler denominator obstruction.

It remains to prove the alternative.  If a common denominator \(N\) exists, then every active \(k\) belongs to the finite set \(\{d:d\mid N\}\).  Hence all active denominator comparisons are performed inside the finite divisibility lattice of \(N\), or equivalently inside the finite common denominator ring \(\mathbb Z/N\mathbb Z\).  This is precisely the commensurable periodic case.

Assume next that the active denominators use only a finite prime set \(P\), but are not contained in the divisor lattice of any fixed \(N\).  If the heights \(\sum_{p\in P}v_p(k)\) were bounded by some \(H\) on all active denominators, then only finitely many exponent vectors \((v_p(k))_{p\in P}\) could occur.  The least common multiple of the corresponding finitely many \(k\)'s would be a single integer \(N\) divisible by every active denominator, contradicting the hypothesis.  Hence for every \(H\) there is an active \(k\) supported on \(P\) with height greater than \(H\), which is exactly the singular finite-prime alternative.  Notice that the statement explicitly restricts to denominators whose prime divisors lie in \(P\); no outside-prime atom is included in this projection.

Finally, if the active prime support is infinite, then any register using only finitely many prime coordinates misses an active denominator whose prime divisor lies outside the register.  Such a support therefore cannot be represented by a finite prime-register.  The three cases are exhaustive: either the active prime support is infinite, or it is finite; in the finite case either the exponent heights are bounded, giving a common denominator by the least common multiple construction, or they are unbounded.  They are also mutually exclusive by construction.

The finite-window assertion is the same argument applied to the certified subset \(\operatorname{Supp}^{\mu}_{\mathcal W}(\mathfrak D_\theta)\).  Since uncertified atoms may later be certified non-zero, a full-ledger conclusion is justified only when the audit also supplies zero certificates outside the certified non-zero support.  This is the Omega denominator-spectrum rule: finite truncations are admissible certificates for what they explicitly contain, while global commensurability requires an all-atoms ledger audit rather than a projection illusion.
\end{proof}

% --- End distillation writeback ---

\begin{proof}
第一式中，\(\log\mathfrak{M}_L(\theta)-\log\mathfrak{M}_K(\theta)\) 正是从 \(k=K+1\) 到 \(k=L\) 的有限尾和，因此其绝对值不超过定理 \ref{thm:finite-part-logM-truncation-explicit-bound} 给出的从 \(k=K+1\) 到无穷的总尾和界。第二式则由
\[
\left|
\frac{\log\mathfrak{M}_K(\theta)}{\log\mathfrak{M}(\theta)}-1
\right|
=
\frac{\bigl|\log\mathfrak{M}_K(\theta)-\log\mathfrak{M}(\theta)\bigr|}{|\log\mathfrak{M}(\theta)|}
\]
与定理 \ref{thm:finite-part-logM-truncation-explicit-bound} 直接得到。证毕。
\end{proof}
\leanverified{paper\_finite\_part\_logM\_relative\_scale\_transfer}

\begin{remark}[穷竭法式比例读法]\label{rem:finite-part-logM-exhaustion-ratio-reading}
推论 \ref{cor:finite-part-logM-relative-scale-transfer} 表明，截断 \(\log\mathfrak{M}_K(\theta)\) 不应被看成与 \(\log\mathfrak{M}(\theta)\) 并列的另一个对象，而应被看成同一解析对象在有限层上的穷竭近似。真正稳定的是“截断层与全量对象的比例误差”而不是某个孤立的 \(K\) 层值；这正是 Euclid 式有限阶段逼近的现代版本。
\end{remark}
