\section{Finite-Field Equational Saturation of Linear Magmas}\label{sec:finite-field-equational-saturation}

This section records the finite-field linear fragment of the equational theory
search.  Throughout, \(p\) is a prime and a linear magma over \(\FF_p\) means a
binary operation
\[
    x\diamond y=ax+by,\qquad a,b\in\FF_p.
\]
For an equation \(E:s=t\), satisfaction in this family is not an opaque
finite-model condition: it is the vanishing of explicitly computable
coefficient polynomials in the two parameters \(a,b\).  The order-four ETP
universe therefore becomes a finite incidence geometry in the affine
coefficient plane.

\begin{theorem}[Linearization Theorem]\label{thm:finite-field-linearization}
Let \(R\) be a commutative ring and let \(t(x_1,\ldots,x_n)\) be a magma term.
Under the interpretation \(x\diamond y=ax+by\), the value of \(t\) has the form
\[
    t=\sum_{i=1}^n c_i(t;a,b)x_i
\]
with \(c_i(t;a,b)\in\ZZ[a,b]\).  If \(|t|\) denotes the number of internal
\(\diamond\)-nodes of \(t\), then \(\deg c_i(t;a,b)\leq |t|\) for every \(i\).
\end{theorem}

\begin{proof}
For the variable term \(x_i\), the coefficient of \(x_i\) is \(1\) and all
other coefficients are \(0\).  If \(t=u\diamond v\), then the recursive rule is
\[
    c_i(t;a,b)=a\,c_i(u;a,b)+b\,c_i(v;a,b).
\]
Thus the interpretation remains linear in the input variables, and each
internal node raises total degree in \(a,b\) by at most one.  Induction on the
binary tree of \(t\) proves both assertions.
\end{proof}
\leanverified{paper\_finite\_field\_linearization}

\begin{corollary}[Coefficient ideal criterion]\label{cor:finite-field-coefficient-ideal}
For an equation \(E:s=t\), define
\[
    I_E=\bigl\langle c_i(s;a,b)-c_i(t;a,b)\ :\ i\geq 1\bigr\rangle
        \subset \ZZ[a,b],
\]
where missing variables have coefficient \(0\).  For any prime \(p\) and any
\((a,b)\in\FF_p^2\), the linear magma \(x\diamond y=ax+by\) satisfies \(E\) for
all input assignments if and only if every generator of \(I_E\) vanishes at
\((a,b)\) modulo \(p\).
\end{corollary}

\begin{proof}
By Theorem \ref{thm:finite-field-linearization}, the difference \(s-t\) is a
linear form in independent variables \(x_i\).  Over \(\FF_p\), this linear form
vanishes on all assignments exactly when each of its coefficients vanishes.
\end{proof}

Consequently each equation determines an integral closed subscheme
\[
    X_E=\operatorname{Spec}\ZZ[a,b]/I_E
\]
of the coefficient plane, and an anti-implication \(E_i\not\Rightarrow E_j\)
over \(\FF_p\) is exactly the existence of a point in the constructible set
\[
    \bigl(X_{E_i}\setminus X_{E_j}\bigr)(\FF_p).
\]
This is the bridge from finite magma search to two-variable algebraic
geometry.

\begin{theorem}[Content Bound]\label{thm:finite-field-content-bound}
\leanverified{paper\_finite\_field\_content\_bound}
Let
\[
    f=\sum_{\alpha} n_\alpha a^{\alpha_1}b^{\alpha_2}\in\ZZ[a,b]
\]
be nonzero, and let
\[
    \operatorname{cont}(f)=\gcd_{\alpha} n_\alpha
\]
be its content.  The reduction of \(f\) in \(\FF_p[a,b]\) is the zero
polynomial if and only if \(p\mid \operatorname{cont}(f)\).
\end{theorem}

\begin{proof}
The monomials \(a^r b^s\) form an \(\FF_p\)-basis of \(\FF_p[a,b]\).  Hence the
reduction of \(f\) is zero precisely when every integer coefficient
\(n_\alpha\) is divisible by \(p\), which is equivalent to
\(p\mid\operatorname{cont}(f)\).
\end{proof}

For the order-four ETP enumeration considered here, the coefficient analysis
has the following numerical envelope: every generator has degree at most \(4\);
every term coefficient and coefficient-difference coefficient has absolute
value at most \(2\); and the only prime at which a nonzero defining polynomial
can collapse to the zero polynomial is \(p=2\).  In particular, all odd primes
see the same nonzero integral defining equations.  Later variation in rational
points is therefore arithmetic variation on fixed low-degree schemes, not a
change in the defining equations themselves.

\begin{lemma}[Geometric degree control]\label{lem:finite-field-geometric-control}
In the order-four linear ETP family, every independent pair of coefficient
constraints cuts out, after removal of common components over an algebraic
closure, a zero-dimensional scheme of degree at most \(16\).
\end{lemma}

\begin{proof}
Each coefficient constraint has total degree at most \(4\).  The assertion is
the plane Bezout bound \(4\cdot 4=16\), applied to the two curves after common
components have been removed.
\end{proof}

The point of Lemma \ref{lem:finite-field-geometric-control} is not that degree
alone forces saturation.  A bounded-degree variety may acquire rational points
over infinitely many primes.  Rather, the lemma identifies the finite
low-complexity algebraic arrangement on which the scan operates.  Implication
over an algebraically closed field \(k\) can also be expressed by the radical
containment
\[
    I_{E_j}\subseteq \sqrt{I_{E_i}}\subset k[a,b],
\]
and failures of containment are constructible differences.  This
Nullstellensatz viewpoint is computationally verified (scripts/equational\_theory/coefficient\_analysis.py, scripts/equational\_theory/nullstellensatz\_check.py)
for the finite certificate described below.

Let \({\mathcal E}_{\leq 2}\) denote the \(810\) order-four ETP laws involving
at most two variables, and let \({\mathcal E}_{=3}\) denote the \(2090\) laws
involving exactly three variables.  For a finite set \({\mathcal S}\) of laws
and a prime \(p\), define
\[
    A_{\mathcal S}(p)=
    \{(i,j)\in{\mathcal S}^2:\exists(a,b)\in\FF_p^2
      \text{ such that }E_i(a,b)\text{ holds and }E_j(a,b)\text{ fails}\}.
\]
The computed prime scan is:
\[
\begin{array}{c|rrr|rrr}
p
& \multicolumn{3}{c|}{{\mathcal E}_{\leq 2}}
& \multicolumn{3}{c}{{\mathcal E}_{=3}}\\
\hline
& \text{new} & \text{cumulative} & \text{patterns}
& \text{new} & \text{cumulative} & \text{patterns}\\
\hline
2  & 439263 & 439263 & 4   & 2519673 & 2519673 & 4\\
3  & 90422  & 529685 & 9   & 106071  & 2625744 & 9\\
5  & 56259  & 585944 & 23  & 18310   & 2644054 & 17\\
7  & 20008  & 605952 & 45  & 48279   & 2692333 & 23\\
11 & 2612   & 608564 & 97  & 0       & 2692333 & 23\\
13 & 384    & 608948 & 129 & 0       & 2692333 & 39\\
17 & 0      & 608948 & 187 & 0       & 2692333 & 37\\
19 & 0      & 608948 & 201 & 0       & 2692333 & 29\\
23 & 0      & 608948 & 193 & 0       & 2692333 & 23\\
29 & 0      & 608948 & 248 & 0       & 2692333 & 33\\
31 & 0      & 608948 & 223 & 0       & 2692333 & 29
\end{array}
\]

\begin{theorem}[Prime-Field Saturation Theorem]\label{thm:prime-field-saturation}
For the scanned primes
\[
    P=\{2,3,5,7,11,13,17,19,23,29,31\},
\]
the anti-implication relation for \({\mathcal E}_{\leq 2}\) is already
saturated by \(p\leq 13\):
\[
    \bigcup_{p\in P} A_{{\mathcal E}_{\leq 2}}(p)
    =
    \bigcup_{p\in\{2,3,5,7,11,13\}} A_{{\mathcal E}_{\leq 2}}(p),
\]
and this common set has cardinality \(608948\).  For
\({\mathcal E}_{=3}\), the anti-implication relation is already saturated by
\(p\leq 7\):
\[
    \bigcup_{p\in P} A_{{\mathcal E}_{=3}}(p)
    =
    \bigcup_{p\in\{2,3,5,7\}} A_{{\mathcal E}_{=3}}(p),
\]
and this common set has cardinality \(2692333\).
\end{theorem}

\begin{proof}
For each equation \(E\), Corollary
\ref{cor:finite-field-coefficient-ideal} converts satisfaction in the linear
magma \(x\diamond y=ax+by\) into a finite list of polynomial vanishing tests in
\(\FF_p^2\).  The verification enumerates all \((a,b)\in\FF_p^2\), records the
truth vector of the selected law family, and inserts every ordered pair
\((i,j)\) for which the vector has \(E_i\) true and \(E_j\) false.  The table
shows the incremental contribution of each prime in \(P\).  After \(p=13\) in
the at-most-two-variable family, and after \(p=7\) in the exactly-three-variable
family, the incremental contribution is zero through the remainder of \(P\).
The stated equalities and cardinalities are exactly these finite unions.  The
certificate is computationally verified (scripts/equational\_theory/coefficient\_analysis.py, scripts/equational\_theory/nullstellensatz\_check.py).
\end{proof}

The theorem should be read as saturation of ordered separations, not as
stabilization of point counts or truth-vector sets.  Indeed, the number of
observed satisfaction patterns continues to change after the saturation prime:
for \({\mathcal E}_{\leq2}\), the scan records \(129\) patterns at \(p=13\),
\(187\) at \(p=17\), and \(248\) at \(p=29\), even though no new ordered
anti-implication appears.  Thus the implication graph is a coarser invariant
than the full finite-field moduli problem.  It records the order-theoretic
shadow of the coefficient geometry.

Several consequences follow.  First, the ETP linear fragment is a finite
scheme-theoretic incidence problem in \(\mathbb A^2\), rather than an
unstructured finite algebra search.  Second, the Catalan--Bell syntactic
complexity of order-four magma terms collapses, under linearization, to
degree-four coefficient geometry in two parameters.  Third, the content bound
isolates \(p=2\) as the only characteristic at which a nonzero defining
polynomial disappears identically.  The remaining primes may vary in rational
point structure, but they vary on a fixed family of low-degree schemes.
