\section{Introduction}\label{sec:introduction}

Zeckendorf's theorem identifies every positive integer with a unique Fibonacci
expansion having no adjacent \(1\)'s \cite{Lekkerkerker1952,Zeckendorf1972}.
The normalization of arbitrary Fibonacci representations is likewise classical.
We therefore take normalization as input data and focus on a different object:
the exact fiber spectrum of finite-window Zeckendorf stabilization.

Fix \(m\ge 1\).  Let
\[
  \Omega_m=\{0,1\}^m,
  \qquad
  X_m=\{x\in\Omega_m:\ x_ix_{i+1}=0\text{ for }1\le i<m\},
\]
and define
\[
  V_m(x)=\sum_{j=1}^m x_jF_{j+1},
  \qquad
  N_m(\omega)=\sum_{j=1}^m \omega_jF_{j+1}.
\]
Since \(V_m\) is a bijection from \(X_m\) onto \(\{0,\dots,F_{m+2}-1\}\), the
finite-window fold map is
\[
  \Fold_m(\omega):=V_m^{-1}\bigl(N_m(\omega)\bmod F_{m+2}\bigr).
\]
Equivalently, \(\Fold_m(\omega)\) is obtained by taking the full Zeckendorf
expansion of \(N_m(\omega)\) and retaining the first \(m\) visible digits.
Write
\[
  d_m(x):=\card{\Fold_m^{-1}(x)},
  \qquad
  d_m^\#(r):=d_m(V_m^{-1}(r)).
\]

Two structural facts, proved in Sections~\ref{sec:preliminaries} and
\ref{sec:residue-affine}, drive the whole paper.  First, the fibers of
\(\Fold_m\) are exact residue classes modulo \(F_{m+2}\).  Second, there is an
explicit affine permutation
\[
  \pi_m(r)\equiv F_{m+1}r+F_{m+1}-1 \pmod{F_{m+2}}
\]
such that \(d_m^\#(\pi_m(r))\) is the coefficient of \(z^r\) in
\(\prod_{j=1}^m(1+z^{F_j})\).  The decisive point is that this coefficient
spectrum is not merely analogous to the classical Fibonacci partition function:
it is exactly a Fibonacci-lag discrete derivative of it.  Once that
single-fiber identity is combined with the power-sum theory of Fibonacci
partitions, finite-window Zeckendorf fibers close to a genuinely discrete
thermodynamic system.

\subsection{Main results}

Following Chow and Slattery \cite{ChowSlattery2021}, let \(R(n)\) denote the
number of partitions of \(n\) into distinct Fibonacci numbers, where the two
unit weights \(F_1=F_2=1\) are not distinguished.  Set \(R(n)=0\) for \(n<0\),
define
\[
  R^\dagger(n):=R(n)+R(n-1),
\]
and write
\[
  T_q^\dagger(N):=\sum_{n=0}^{N-1} R^\dagger(n)^q.
\]
Sanna \cite{Sanna2025} proved that for every integer \(q\ge 1\) there exists a
constant \(\lambda_q>1\) such that
\[
  \sum_{n=0}^{N-1} R(n)^q \asymp_q N^{(\log \lambda_q)/\log\varphi},
  \qquad
  \lim_{q\to\infty}\lambda_q^{1/q}=\varphi^{1/2}.
\]
We show that the same constants govern the finite-window collision moments and
that, after adjoining \(P_0:=\log\varphi\), they form the pressure sequence of a
discrete canonical/microcanonical formalism.

\begin{theoremA}
For every integer \(n\) with \(0\le n<F_{m+2}\),
\[
  d_m^\#(\pi_m(n))
  =
  R^\dagger(n)-R^\dagger(n-F_{m+1}).
\]
Equivalently,
\[
  d_m^\#(\pi_m(n))
  =
  R(n)+R(n-1)-R(n-F_{m+1})-R(n-F_{m+1}-1).
\]
\end{theoremA}

\noindent
Thus the finite-window fiber spectrum is exactly the Fibonacci-lag discrete
derivative of the classical Fibonacci partition function.

\begin{theoremB}
For every integer \(q\ge 1\), the collision moment
\[
  S_q(m):=\sum_{x\in X_m} d_m(x)^q
\]
satisfies the exact Fibonacci-window bounds
\[
  T_q^\dagger(F_{m+1})\le S_q(m)\le T_q^\dagger(F_{m+2}).
\]
Consequently
\[
  S_q(m)\asymp_q \lambda_q^m.
\]
If \(q\ge 2\), then the Perron root \(r_q\) of the finite-state collision
kernel of Section~\ref{sec:moment-kernel} satisfies
\[
  r_q=\lambda_q.
\]
\end{theoremB}

\begin{theoremC}
Set
\[
  P_0:=\log\varphi,
  \qquad
  P_q:=\log\lambda_q
  \qquad (q\ge 1).
\]
For every integer \(q\ge 2\), the constant \(\lambda_q\) is the spectral
radius of a nonnegative integer matrix of dimension polynomial in \(q\).  In
particular, \(\lambda_q\) is an algebraic integer.  The pressure sequence
\((P_q)_{q\ge 0}\) is convex on the nonnegative integers.  Hence the discrete
slopes
\[
  \Delta_q:=P_q-P_{q-1}
  \qquad (q\ge 1)
\]
form an increasing sequence, and
\[
  \lim_{q\to\infty}\Delta_q=\frac12\log\varphi.
\]
\end{theoremC}

For each integer \(q\ge 1\), define the \(q\)-canonical law on \(X_m\) by
\[
  \mu_{m,q}(x):=\frac{d_m(x)^q}{S_q(m)}.
\]

\begin{theoremD}
Fix an integer \(q\ge 1\).  For every \(\varepsilon>0\),
\[
  \mu_{m,q}\!\left\{
    x:\frac1m\log d_m(x)\le \Delta_q-\varepsilon
  \right\}
  \le
  \exp\bigl(-(\varepsilon+o(1))m\bigr),
\]
and
\[
  \mu_{m,q}\!\left\{
    x:\frac1m\log d_m(x)\ge \Delta_{q+1}+\varepsilon
  \right\}
  \le
  \exp\bigl(-(\varepsilon+o(1))m\bigr).
\]
Equivalently, the \(q\)-canonical ensemble is exponentially concentrated on the
adjacent pressure-slope band \([\Delta_q,\Delta_{q+1}]\).
\end{theoremD}

\begin{theoremE}
For a fixed integer \(q\ge 1\) and \(\varepsilon>0\), let
\[
  B_{m,q}(\varepsilon)
  :=
  \left\{
    x\in X_m:
    \Delta_q-\varepsilon\le \frac1m\log d_m(x)\le \Delta_{q+1}+\varepsilon
  \right\}.
\]
Then
\[
  \mu_{m,q}\bigl(B_{m,q}(\varepsilon)\bigr)
  =
  1-\exp\bigl(-(\varepsilon+o(1))m\bigr).
\]
Moreover the cardinality and visible mass of this band satisfy the two-sided
exponential estimates
\[
  P_q-q(\Delta_{q+1}+\varepsilon)
  \le
  \liminf_{m\to\infty}\frac1m\log |B_{m,q}(\varepsilon)|,
\]
\[
  \limsup_{m\to\infty}\frac1m\log |B_{m,q}(\varepsilon)|
  \le
  P_q-q(\Delta_q-\varepsilon),
\]
and
\[
  P_q-\log 2-(q-1)(\Delta_{q+1}+\varepsilon)
  \le
  \liminf_{m\to\infty}\frac1m\log p_m\bigl(B_{m,q}(\varepsilon)\bigr),
\]
\[
  \limsup_{m\to\infty}\frac1m\log p_m\bigl(B_{m,q}(\varepsilon)\bigr)
  \le
  P_q-\log 2-(q-1)(\Delta_q-\varepsilon),
\]
where \(p_m(x):=d_m(x)/2^m\) is the visible law under uniform input.
\end{theoremE}

\begin{theoremF}
Let \(q=q(m)\) be any integer sequence with \(q(m)\to\infty\), and set
\[
  \nu_m(x):=\frac{d_m(x)^{q(m)}}{S_{q(m)}(m)}.
\]
Then for every \(\varepsilon>0\),
\[
  \nu_m\!\left\{
    x:
    \left|\frac1m\log d_m(x)-\frac12\log\varphi\right|>\varepsilon
  \right\}\to 0.
\]
In particular,
\[
  D_m:=\max_{x\in X_m}d_m(x)
  \qquad\text{satisfies}\qquad
  D_m^{1/m}\to\varphi^{1/2},
\]
and
\[
  \log S_{q(m)}(m)
  =
  \Bigl(\frac12\log\varphi+o(1)\Bigr)m\,q(m).
\]
\end{theoremF}

At the visible endpoint \(q=1\), one has \(\mu_{m,1}=p_m\).  Therefore
Theorem~D yields a deterministic typical-thickness band for the visible law:
for every \(\varepsilon>0\),
\[
  \begin{aligned}
    p_m\!\left\{
      x:
      \begin{array}{c}
        \log 2-\log\varphi-\varepsilon
        \le
        \frac1m\log d_m(x) \\[2pt]
        \le
        \log\lambda_2-\log 2+\varepsilon
      \end{array}
    \right\}
    =
    1-\exp\bigl(-(\varepsilon+o(1))m\bigr).
  \end{aligned}
\]
Section~\ref{sec:ledger} shows moreover that the same pressure sequence governs
the full R\'enyi entropy-rate spectrum of the visible law:
for every fixed integer \(q>1\),
\[
  \lim_{m\to\infty}\frac1m H_q(p_m)
  =
  \frac{q\log 2-P_q}{q-1},
\]
and the min-entropy endpoint equals
\[
  \log 2-\frac12\log\varphi.
\]

\medskip
\noindent
The quadratic case remains completely explicit.  Section~\ref{sec:second-collision}
proves that
\[
  S_2(m)=2S_2(m-1)+2S_2(m-2)-2S_2(m-3)
  \qquad (m\ge 4),
\]
so the dominant root \(\lambda_2\) is the largest real zero of
\(\lambda^3-2\lambda^2-2\lambda+2\), in agreement with Chow and Jones
\cite{ChowJones2024}.  In the present version we treat this cubic law as a
fully explicit sample of the general pressure formalism, not as a competing
narrative.

\subsection{Relation to previous work}

Theorem~A gives the paper its decisive bridge to the existing literature.
Chow and Slattery proved exact formulas for the Fibonacci partition function in
terms of the Zeckendorf expansion of \(n\) \cite{ChowSlattery2021}.  Our
Theorem~A shows that finite-window Zeckendorf fibers are not merely related to
that partition theory: after an explicit affine permutation, they are exactly
the Fibonacci-lag discrete derivative of the indexed partition function
\(R^\dagger\).  Consequently their individual values inherit explicit closed
formulas from Chow and Slattery without any partition-theoretic reproof.

Theorem~B is the basic global transfer principle.  The collision moments are
not merely comparable with the Fibonacci partition power sums studied by Sanna
\cite{Sanna2025}; they are squeezed between two adjacent Fibonacci windows of
those sums.  This transfers Sanna's growth constants to the finite-window fold
model and identifies the Perron roots of our finite-state collision kernels
with the same constants \(\lambda_q\).

The new contribution of the present revision is that these constants are no
longer used only as moment-growth exponents.  Once \(P_0=\log\varphi\) is
adjoined, the sequence \((P_q)\) becomes a discrete pressure function, and the
adjacent slope differences \(\Delta_q\) select the canonical thickness bands of
the model.  Theorems~D--F therefore convert the previous collection of upper
envelopes and zero-temperature consequences into a closed canonical and
microcanonical formalism.  In particular, the visible law \(p_m=\mu_{m,1}\) is
not governed by a vague average-fiber heuristic: it occupies a deterministic
band bounded by the first two pressure slopes.

This point of view also clarifies the role of the R\'enyi entropy spectrum.
Section~\ref{sec:ledger} does not supply an unrelated information-theoretic
appendix to the moment theory.  Rather, the R\'enyi rates and visible-tail
bounds are the \(q=1\) endpoint and weighted projections of the same pressure
geometry that governs the canonical ensembles \(\mu_{m,q}\).

\subsection{Organization}

Section~\ref{sec:preliminaries} fixes the finite-window fold map and its basic
overflow structure.  Section~\ref{sec:residue-affine} proves the residue
description, the affine transfer principle, and the partition-difference
formula for individual fibers.  Section~\ref{sec:second-collision} proves the
Fibonacci-window transfer of collision moments and records the explicit
quadratic recurrence.  Section~\ref{sec:moment-kernel} turns the transferred
partition constants into finite-state Perron roots, develops the convex
pressure geometry, proves the canonical selection theorem and the
microcanonical band bounds, and derives the zero-temperature concentration
results.  Section~\ref{sec:ledger} treats the visible law as the \(q=1\)
canonical ensemble and deduces its typical-thickness band, R\'enyi entropy
rates, and visible-tail estimates.  Appendix~\ref{app:transducer} gives a
self-contained bounded-delay transducer model for the finite-state collision
kernels.
