\abstract{%
We study what the full stationary click record of a sampled Unruh--DeWitt detector contributes beyond transition rates and Tolman profiles in static Kubo--Martin--Schwinger backgrounds. For an explicit two-level detector model, ideal one-step dead time yields a Markov law on the golden-mean shift, whereas finite recovery yields a renewal process with infinite symbolic memory and a sharp correlation-sign threshold. These two regimes define two detector-dependent codimension-one hypersurfaces: an exact-clock shell and a memory-transition shell. In the homogeneous spectral-weight regime, the exact-clock shell is a calibrated Tolman isotherm, whereas the finite-recovery shell is characterized by exact $1$-dependence and separates purely negative from sign-alternating covariances. We prove an exact ordering law for the two shells, show that finite recovery never realizes exact symbolic thermality, and derive universal near-horizon asymptotics for their proper distances and area excesses. In static spherical settings, the shell pair fixes the redshift scale, and in the Schwarzschild geometric-optics static-detector regime it determines the mass parameter within the response model. More generally, the shell pair obeys an exact spectral-transport identity whose homogeneous specialization is a mode-by-mode shell-ratio law with exact cancellation of the common response amplitude; two operating modes of the same hardware then yield a local inverse principle for static spherical two-parameter families. These inverse statements are conditional on the fixed weak-coupling and homogeneous-spectral-weight response model rather than claims of detector-independent metric reconstruction. The paper provides a gravity-facing application of full-record detector statistics within that controlled model class.%
}

\keywords{%
  stationary spacetimes; Unruh--DeWitt detectors; quantum field theory in curved spacetime;
  KMS states; local temperature; black hole thermodynamics}
