\section{Introduction}\label{sec:bridge:intro}

In Hawking and Unruh settings, observer dependence is usually discussed through transition rates, detailed balance, and Tolman redshift. We study a sharper operational question: what does the full stationary click record of a sampled detector contribute beyond a rate formula, and can that record single out distinguished stationary observers or codimension-one hypersurfaces in a static curved spacetime?

This question belongs to quantum field theory in curved spacetime, but it is not identical to the standard problems of detector rates, repeated-measurement formalisms, or autonomous quantum clocks. Renewal-like tick statistics occur in autonomous clock models \cite{ErkerEtAl2017}. Detector outputs have been used to infer geometry from measurements \cite{PercheMartinMartinez2022}, while covariant algebraic measurement frameworks stress that operational observables depend on the chosen measurement scheme as well as on the state \cite{Fewster2020}. Information-geometric and metrological approaches instead emphasize state distinguishability and Fisher-information observables rather than the full symbolic record \cite{BraunsteinCaves1994,WangEtAl2014Metrology,HaoWu2016Unruh,FengZhang2022QFI,DuMann2021AdS}. Repeated strings for Unruh--DeWitt detectors and the domain of validity of Markovian master-equation descriptions have recently been analyzed in \cite{PranziniEtAl2024,HanEtAl2025}, and even the interpretation of detailed balance as a complete indicator of detector thermality can fail outside the minimal two-level setting \cite{LimaEtAl2023Qudit}. The issue addressed here is narrower and more geometric: within an explicit detector model, we ask whether full-record detector statistics lift stationary thermal response to detector-defined hypersurfaces in static Kubo--Martin--Schwinger (KMS) backgrounds, and whether the resulting shell data have inverse content.

On the detector-model side, the ideal dead-time rule is the non-paralyzable counting model familiar from counting theory \cite{Muller1973,CantorTeich1975}. The finite-recovery model is still deliberately minimal: recovery is exponential, sampling is discrete, and the hidden detector remains two-state. Paralyzable detectors, nonexponential recovery, and more elaborate quantum-trajectory descriptions lie outside the present scope \cite{Carmichael1993Book,WisemanMilburn2009}. On the thermodynamic side, we use only the stationary proper-time KMS relation along a worldline; we do not identify Tolman scaling by itself with a general notion of local thermal equilibrium \cite{BuchholzSchlemmer2007,Solveen2012,BuchholzSolveen2013,GranseePinamontiVerch2017}.

We study a two-level detector sampled along a stationary worldline. At the algebraic level, a normal binary instrument and a stationary channel state determine a shift-invariant symbolic law on $\{0,1\}^{\NN}$. Two detector regimes are then analyzed. In the ideal one-step dead-time model, a click forbids the next click and the output is a Markov chain on the golden-mean shift $X_\varphi$, which is simply the space of binary records with no consecutive clicks. In the finite-recovery model, excitation and recovery compete within one sampling interval; the hidden detector remains two-state, but the visible process becomes a renewal process with infinite symbolic memory and an explicit correlation threshold. The role of these stochastic laws is to turn detector parameters into geometric selectors once the local excitation rate acquires its Tolman redshift dependence in a static KMS background.

Three terms will recur throughout the paper. The \emph{homogeneous spectral-weight regime} means that the local commutator spectral weight $\mathcal S_x(\Omega)$ is effectively constant on the static region under study, so that Tolman redshift supplies the only spatial dependence in the leading-order excitation rate. \emph{Exact symbolic thermality} means that the sampled output law realizes the distinguished maximum-entropy law compatible with its symbolic grammar; in the ideal dead-time model this law is the Parry measure on the golden-mean shift. The \emph{exact-clock shell} is the codimension-one locus where that ideal sampled law is realized.

A brief notation guide may help fix the operational roles of the main parameters. The coupling amplitude is $\lambda_c$, the sampling step is $\Delta\tau$, the finite-recovery rate is $\kappa_{\mathrm r}$, and $\Gamma_c(\kappa_{\mathrm r},\Delta\tau)$ is the corresponding memory-transition threshold, while $\Gamma_\varphi=\log\varphi/\Delta\tau$ is the exact-clock threshold in the ideal model. The symbols $p$ and $q$ denote one-step click and recovery probabilities in the sampled channel, $(a,b)$ denote the renewal parameters in the finite-recovery symbolic law, $c_m$ denotes mode constants in the shell-inversion formulas, and $\rho$ denotes proper distance to a Killing horizon.

The main results are fourfold.
\begin{enumerate}[label=\textup{(\roman*)},leftmargin=*,itemsep=4pt]
\item \textup{(Finite recovery obstructs exact symbolic thermality.)} Finite recovery never realizes exact symbolic thermality in the full-support renewal class. Zero regenerative jitter and the Bernoulli entropy benchmark are excluded, so exact symbolic thermality is a singular ideal-dead-time phenomenon rather than a generic feature of stationary detector thermality.
\item \textup{(Finite recovery defines a second shell.)} The correlation-sign threshold lifts to a codimension-one memory-transition hypersurface on which the output is exact $1$-dependent; on one side all covariances are negative, while on the other side they alternate in sign.
\item \textup{(The ideal and finite-recovery shells compare exactly.)} The ideal dead-time model defines an exact-clock hypersurface. In the homogeneous spectral-weight regime this shell is a calibrated Tolman isotherm, and the exact-clock shell and the memory-transition shell satisfy a universal ordering law controlled by detector parameters rather than by the background alone.
\item \textup{(The shell pair has inverse and spectral-transport content.)} In static spherical settings the two shell radii fix the redshift scale within the homogeneous-spectral-weight response model, and in the Schwarzschild geometric-optics static-detector regime they determine the mass parameter within that same model class. Outside the homogeneous limit, the same shell pair obeys an exact spectral-transport identity coupling redshift to the local commutator spectral weight. Its homogeneous specialization is the shell-ratio law with exact cancellation of the common response amplitude, and two operating modes of the same hardware then yield a local inverse principle for static spherical two-parameter families.
\end{enumerate}

The gravitational content of the paper is therefore not a derivation of field equations or a detector-independent invariant of the background, in contrast with thermodynamic programs that tie horizon response directly to gravitational dynamics \cite{Jacobson1995}. It is a concrete quantum-field-theoretic statement in curved spacetime: under a fixed weak-coupling detector model, stationary thermality selects two inequivalent codimension-one shells that refine observer selection and near-horizon analysis beyond transition rates alone.

The scope is deliberately limited. We work with a single stationary worldline, explicit detector dynamics, and the weak-coupling KMS bridge used to convert detector response rates into local temperature dependence. The shells are detector-defined threshold hypersurfaces, not intrinsic geometric invariants, and the inverse statements are model-conditional rather than detector-independent metric reconstruction. Rindler space gives an exact realization within the model, whereas the explicit Schwarzschild formulas are presented only in the geometric-optics static-detector regime. The auxiliary finite-window coarse graining is retained only in Appendix~\ref{app:bridge:microhistory}, where it records an information balance compatible with, but not needed for, the main shell-geometry results.

From this perspective, the present shell construction is complementary to other operational frameworks in curved spacetime. Fisher-information programs seek optimal thermometric or geometric estimators \cite{FengZhang2022QFI,DuMann2021AdS}, interferometric proposals track phase or visibility shifts \cite{CostaEtAl2020Ramsey}, recent operational temperature studies emphasize finite-time detector readout and switching effects \cite{ShallueCarroll2025}, and fluctuation-theorem approaches characterize observer-dependent entropy production \cite{BassoMazieroCeleri2025}. Our focus is different: we use the full symbolic detector record itself to define detector-dependent shells and shell-ratio observables.

The paper is organized as follows. Section~\ref{sec:bridge:operational} constructs the stationary symbolic law. Section~\ref{sec:bridge:deadtime} analyzes the ideal dead-time and finite-recovery models. Section~\ref{sec:bridge:kms} studies KMS thermality, the exact-clock shell, the finite-recovery memory shell, their near-horizon asymptotics, and the inverse shell geometry they induce in static curved spacetime. Section~\ref{sec:bridge:discussion} concludes with scope and outlook. Appendix~\ref{app:bridge:microhistory} records the auxiliary finite-window coarse graining, and Appendix~\ref{app:bridge:companion} constructs the stabilization map underlying it.
