
\section{State diagram, timing convention, and carryover rule}\label{app:state-diagram}

The counter has two hidden states, $R$ (ready) and $D$ (refractory). The visible output is produced by the interval click register described in Appendix~\ref{app:gkls}, not by a direct measurement of the hidden state. The state diagram and timing convention for one sampling interval $[0,\Delta\tau]$ are recorded here for reference.

To remove any ambiguity in the sampling protocol, we distinguish three moments in each interval:
\begin{enumerate}[label=(\roman*),itemsep=0.15em]
\item immediately after the interval-end readout at time $n\Delta\tau^+$, which resets the click register and sets the hidden state for interval $n$;
\item during the open interval $(n\Delta\tau,(n+1)\Delta\tau)$, where recovery and excitation can occur;
\item immediately before the next readout at time $(n+1)\Delta\tau^-$, whose hidden state is then carried into interval $n+1$ unless a click occurred.
\end{enumerate}
The carryover rule is the following: a click in interval $n$ forces interval $n+1$ to start in $D$; a no-click does \emph{not} reset the detector to $R$, but instead carries the hidden end-of-interval state into the next interval.

\begin{center}
\begin{tikzpicture}[
  state/.style={draw, circle, minimum size=1.2cm, thick},
  arr/.style={-Latex, thick},
  every node/.style={font=\small}
]
\node[state, fill=green!12] (R) at (0,0) {$R$};
\node[state, fill=red!12] (D) at (4,0) {$D$};

\draw[arr] (R) to[bend left=25] node[above] {excite at rate $\Gamma$} node[below] {output $A_n=1$} (D);
\draw[arr] (R) to[loop left, looseness=8] node[left] {no excitation} node[right,xshift=-0.3cm,yshift=-0.5cm] {$A_n=0$} (R);
\draw[arr] (D) to[bend left=25] node[below] {recover at rate $\kappa_{\mathrm r}$} (R);
\draw[arr] (D) to[loop right, looseness=8] node[right] {no recovery} node[left,xshift=0.3cm,yshift=-0.5cm] {$A_n=0$} (D);

\node[anchor=north, text width=7cm, align=left] at (2,-2.2) {
\textbf{Key rule.} Every click ($A_n=1$) leaves the detector in state $D$ at the start of interval $n+1$, regardless of the hidden state at the start of interval $n$. This is the single assumption that makes click times regeneration points.
};
\end{tikzpicture}
\end{center}

\noindent\textbf{Timing within one interval.}
\begin{enumerate}[label=(\alph*),itemsep=0.15em]
\item \emph{Start in $R$:} The detector may be excited during $[0,\Delta\tau]$ at rate $\Gamma$. If excited (probability $1-p$, with $p=e^{-\Gamma\Delta\tau}$), it ends in $D$ and emits $A_n=1$. If not excited (probability $p$), it ends in $R$ and emits $A_n=0$.
\item \emph{Start in $D$:} Recovery to $R$ occurs at an exponential time $s$ with rate $\kappa_{\mathrm r}$. If $s\le\Delta\tau$, the detector is ready for the remaining subinterval $[s,\Delta\tau]$ and may be excited at rate $\Gamma$. If excited after recovery, it ends in $D$ with $A_n=1$ (probability $a$). If not excited after recovery, it ends in $R$ with $A_n=0$ (probability $b$). If $s>\Delta\tau$ (no recovery during the interval), it remains in $D$ with $A_n=0$ (probability $c=e^{-\kappa_{\mathrm r}\Delta\tau}$). Here $a+b=1-c$, so $a+b+c=1$.
\end{enumerate}

The output-labelled kernels $T_0$ and $T_1$ in~\eqref{eq:t0-t1-main} encode precisely these transition-and-output probabilities. Note that the transition $(D,D)$ can produce either output: $A_n=0$ with probability $c$ (no recovery) or $A_n=1$ with probability $a$ (recovery followed by re-excitation). This is why the output is \emph{not} a deterministic function of the hidden-state pair $(H_n,H_{n+1})$, a point exploited in the factorization proof of $1$-dependence in \cref{sec:records}.

\section{Interval click-register instrument for the phenomenological kernels}\label{app:gkls}

This appendix records one explicit interval instrument realizing the kernels used in the main text. The role of the construction is modest: the UDW detector supplies the stationary excitation input rate $\Gamma_+(x,\Omega)$, while dead time, recovery and interval readout are modelled as counter dynamics. For the record-level analysis of \cref{sec:records}, it is enough to have a consistent interval instrument with these properties. The construction should therefore be read only as an internal realization of the phenomenological kernels, not as a derivation from a microscopic repeated-measurement UDW model.

Let the hidden counter state be $H\in\{R,D\}$ and the click register be $C\in\{0,1\}$. At the beginning of each interval the register is reset to $C=0$. During the interval, the combined hidden/register process has three relevant states,
\[
(R,0),\qquad (D,0),\qquad (D,1),
\]
where $(D,1)$ means that the register has already fired and is then held until the interval-end readout. In the ordered basis $((R,0),(D,0),(D,1))$, the interval generator is the classical Markov generator
\begin{equation}\label{eq:register-generator}
\mathcal L_{\mathrm{ctr}}=
\begin{pmatrix}
-\Gamma & \kappa_{\mathrm r} & 0\\
0 & -\kappa_{\mathrm r} & 0\\
\Gamma & 0 & 0
\end{pmatrix}.
\end{equation}
The transition $(R,0)\to(D,1)$ represents a click. The transition $(D,0)\to(R,0)$ represents recovery. Once the register has fired, the state $(D,1)$ is absorbing until the interval-end measurement.

Let
\[
J=
\begin{pmatrix}
1 & 0\\
0 & 1\\
0 & 0
\end{pmatrix}
\]
embed the hidden-state distribution at the beginning of the interval into the combined hidden/register space, and let
\[
M_0=
\begin{pmatrix}
1 & 0 & 0\\
0 & 1 & 0
\end{pmatrix},
\qquad
M_1=
\begin{pmatrix}
0 & 0 & 0\\
0 & 0 & 1
\end{pmatrix}
\]
denote the interval-end readout maps: outcome $0$ keeps the hidden state reached with register value $0$, while outcome $1$ resets the next interval to the refractory state $D$. The output-labelled instrument is therefore
\begin{equation}\label{eq:instrument-main}
\mathcal I_0=M_0e^{\Delta\tau\mathcal L_{\mathrm{ctr}}}J,
\qquad
\mathcal I_1=M_1e^{\Delta\tau\mathcal L_{\mathrm{ctr}}}J.
\end{equation}

A direct evaluation of~\eqref{eq:instrument-main} gives
\[
e^{\Delta\tau\mathcal L_{\mathrm{ctr}}}J=
\begin{pmatrix}
p & b\\
0 & c\\
1-p & a
\end{pmatrix},
\]
where
\[
p=e^{-\Gamma\Delta\tau},
\qquad
c=e^{-\kappa_{\mathrm r}\Delta\tau},
\]
and
\[
a=\int_0^{\Delta\tau}\kappa_{\mathrm r}e^{-\kappa_{\mathrm r}s}\bigl(1-e^{-\Gamma(\Delta\tau-s)}\bigr)\,ds,
\qquad
b=\int_0^{\Delta\tau}\kappa_{\mathrm r}e^{-\kappa_{\mathrm r}s}e^{-\Gamma(\Delta\tau-s)}\,ds.
\]
Hence
\[
T_0=\mathcal I_0=
\begin{pmatrix}
p & 0\\
b & c
\end{pmatrix},
\qquad
T_1=\mathcal I_1=
\begin{pmatrix}
0 & 1-p\\
0 & a
\end{pmatrix},
\]
which are exactly the kernels used in~\eqref{eq:t0-t1-main}. The construction is consistent for all positive $\Gamma\Delta\tau$ and $\kappa_{\mathrm r}\Delta\tau$; no assumption such as $\kappa_{\mathrm r}\Delta\tau\ll1$ is used. This is the interval instrument implicitly underlying the main text.

\section{Further structural results}\label{app:structural-results}

This appendix records two structural statements that support, but are not needed to read, the main line of the paper. The first places the exponential-recovery counter inside a wider two-state interval class with the same reset architecture. The second shows how exact $1$-dependence fits into the full hierarchy of stationary binary renewal processes.

\subsection{A wider two-state interval class}

The exponential-recovery counter occupies one point in a broader three-parameter family of sampled instruments with the same ready/refractory architecture and the same post-click reset rule. Fix parameters
\[
0<u,w<1,\qquad 0<v<1-w,\qquad u\neq w,
\]
and define
\begin{equation}\label{eq:general-kernels-main}
T_0(u,v,w)=
\begin{pmatrix}
u & 0\\
v & w
\end{pmatrix},
\qquad
T_1(u,v,w)=
\begin{pmatrix}
0 & 1-u\\
0 & 1-v-w
\end{pmatrix}.
\end{equation}
The exponential model corresponds to $(u,v,w)=(p,b,c)$.

\begin{theorem}\label{thm:two-state-class-main}
Let $(A_n)$ be the visible record generated by the kernels~\eqref{eq:general-kernels-main}, and let $G$ denote the zero-run length between successive clicks. Set
\[
a=1-v-w,\qquad \lambda=u-v,\qquad \rho=\frac{(1-u)(1-w)}{1-u+v}.
\]
Then the following statements hold.
\begin{enumerate}[label=\textup{(\roman*)},leftmargin=2.2em,itemsep=0.25em]
\item For every $k\ge0$,
\begin{equation}\label{eq:general-tail-main}
\PP(G\ge k)=\tau_k=\frac{v\,u^k+(\lambda-w)w^k}{u-w},
\qquad
\PP(G=k)=\tau_k-\tau_{k+1}.
\end{equation}
\item The hazard satisfies
\begin{equation}\label{eq:general-hazard-diff-main}
h_{k+1}-h_k=
\frac{v(w-\lambda)u^k w^k}{\tau_k\tau_{k+1}},
\qquad k\ge0.
\end{equation}
Hence the hazard is strictly increasing when $\lambda<w$, constant when $\lambda=w$, and strictly decreasing when $\lambda>w$.
\item The covariance profile is
\begin{equation}\label{eq:general-cov-main}
\operatorname{Cov}(A_0,A_n)=\rho(a-\rho)\lambda^{n-1},
\qquad n\ge1.
\end{equation}
The visible process is Bernoulli if and only if $\lambda=w$. It is exact $1$-dependent and non-Bernoulli if and only if $\lambda=0$.
\item In the exponential-recovery model $(u,v,w)=(p,b,c)$ one has
\begin{equation}\label{eq:general-specialization-main}
w-\lambda=\frac{\Gamma\,(p-c)}{\kappa_{\mathrm r}-\Gamma}>0.
\end{equation}
Accordingly the exponential model lies entirely in the increasing-hazard sector, and its unique non-Bernoulli exact $1$-dependence point is the rank-one condition $p=b$, equivalently $\Gamma=\Gamma_c(\kappa_{\mathrm r},\Delta\tau)$.
\end{enumerate}
\end{theorem}

\subsection{Finite-dependence hierarchy}\label{app:renewal-hierarchy-app}

For the reset architecture, the visible record is determined by its renewal sequence
\[
u_n=\PP(A_n=1\mid A_0=1),\qquad n\ge0,
\]
and by the inter-click interval law. Classical renewal theory relates these two objects through the generating-function identity $U=1/(1-F)$ \cite{Cinlar1975}. The following result identifies exact finite dependence with a rational inter-click law.

\begin{theorem}[Finite-dependence classification]\label{thm:renewal-md-main}
Let $(A_n)_{n\in\mathbb Z}$ be a stationary binary renewal process with click probability $\rho=\PP(A_0=1)\in(0,1)$. Let $Y\in\{1,2,\dots\}$ denote the inter-click interval, set
\[
f_n=\PP(Y=n),
\qquad
F(z)=\sum_{n\ge1}f_n z^n,
\qquad
u_n=\PP(A_n=1\mid A_0=1),
\qquad
u_0=1,
\]
and fix an integer $m\ge1$. Then the following statements are equivalent.
\begin{enumerate}[label=\textup{(\roman*)},leftmargin=2.2em,itemsep=0.25em]
\item The record $(A_n)$ is exact $m$-dependent.
\item One has $u_n=\rho$ for every $n\ge m+1$.
\item The inter-click generating function satisfies
\begin{equation}\label{eq:renewal-q-main}
F(z)=1-\frac{1-z}{Q_m(z)},
\qquad |z|<1,
\end{equation}
where
\begin{equation}\label{eq:renewal-poly-main}
Q_m(z)=1+\sum_{j=1}^{m}(u_j-u_{j-1})z^j+(\rho-u_m)z^{m+1}.
\end{equation}
\end{enumerate}
If the dependence order is exactly $m$, then $\deg Q_m=m+1$. In that case the gap law satisfies the linear recurrence
\begin{equation}\label{eq:renewal-rec-main}
f_n=
(1-u_1)f_{n-1}
+\sum_{j=2}^{m}(u_{j-1}-u_j)f_{n-j}
+(u_m-\rho)f_{n-m-1},
\qquad n\ge m+2.
\end{equation}
\end{theorem}

\section{Key derivations}\label{app:derivations}

This appendix collects the proofs of the main technical results. Further intermediate calculations are recorded in the Supplementary Material.

\subsection{Zero-run law and hazard monotonicity}

\begin{proof}[Proof of \cref{prop:recovery-main}: zero-run law]
Every observed click places the detector in the refractory state $D$ for the next interval, independently of the earlier history. Hence the post-click hidden state is deterministic, and the strong Markov property of the hidden two-state chain makes click times regeneration times. The resulting zero-run lengths are therefore i.i.d., and a gap of length $k$ consists of $k$ successive output-$0$ transitions followed by one output-$1$ transition, giving~\eqref{eq:gap-main}.

To evaluate it explicitly, note that $T_0$ is lower triangular with eigenvalues $p$ and $c$. Direct integration of~\eqref{eq:a-b-main} yields, for $\kappa_{\mathrm r}\neq\Gamma$,
\[
a=\frac{\kappa_{\mathrm r}(1-p)-\Gamma(1-c)}{\kappa_{\mathrm r}-\Gamma},
\qquad
b=\frac{\kappa_{\mathrm r}(p-c)}{\kappa_{\mathrm r}-\Gamma}.
\]
Since $T_0$ is lower triangular,
\[
T_0^k=
\begin{pmatrix}
 p^k & 0 \\
 b\,\dfrac{p^k-c^k}{p-c} & c^k
\end{pmatrix}.
\]
Substituting into~\eqref{eq:gap-main} with $e_D=(0,1)^\top$ and $T_1\mathbf{1}=(1-p,\,a)^\top$ gives
\[
g_k = e_D^\top T_0^k T_1\mathbf{1}
= qb\frac{p^k-c^k}{p-c}+ac^k
= \frac{\kappa_{\mathrm r}(1-p)p^k-\Gamma(1-c)c^k}{\kappa_{\mathrm r}-\Gamma},
\]
where $q=1-p$ and the simplification uses the explicit forms of $a$ and $b$.
\end{proof}

\begin{proof}[Proof of \cref{prop:recovery-main}: hazard monotonicity]
The tail of the gap law is
\[
\sum_{j\ge k}g_j=\frac{\kappa_{\mathrm r}p^k-\Gamma c^k}{\kappa_{\mathrm r}-\Gamma}.
\]
Therefore
\[
h_k = \frac{g_k}{\sum_{j\ge k}g_j}
= \frac{\kappa_{\mathrm r}(1-p)p^k-\Gamma(1-c)c^k}{\kappa_{\mathrm r}p^k-\Gamma c^k}.
\]
Direct subtraction yields
\[
h_{k+1}-h_k = \frac{\kappa_{\mathrm r}\Gamma(p-c)^2p^kc^k}{(\kappa_{\mathrm r}p^k-\Gamma c^k)(\kappa_{\mathrm r}p^{k+1}-\Gamma c^{k+1})} > 0,
\]
since all factors in the numerator and denominator are strictly positive for $p\neq c$ (i.e.\ $\Gamma\neq\kappa_{\mathrm r}$). Strict monotonicity excludes the possibility that the visible process is an $r$-step Markov chain for any finite $r$: if it were, $h_k$ would eventually be constant.
\end{proof}

\subsection{Covariance formula, exact 1-dependence, and edge cases}

\begin{proof}[Proof of \cref{lem:stationary-main}]
The aggregate hidden-state matrix is
\[
P=T_0+T_1=
\begin{pmatrix}
 p & 1-p\\
 b & 1-b
\end{pmatrix}.
\]
Solving $\pi P=\pi$ with $\pi_R+\pi_D=1$ gives
\[
\pi_R=\frac{b}{1-p+b},
\qquad
\pi_D=\frac{1-p}{1-p+b}.
\]
The stationary click rate is therefore
\[
\rho=\pi_R(1-p)+\pi_D a
=
\frac{b(1-p)+(1-p)a}{1-p+b}
=
\frac{(1-p)(a+b)}{1-p+b}
=
\frac{(1-p)(1-c)}{1-p+b}.
\]
Finally,
\[
a=\int_0^{\Delta\tau}\kappa_{\mathrm r}e^{-\kappa_{\mathrm r}s}\bigl(1-e^{-\Gamma(\Delta\tau-s)}\bigr)\,ds
<
(1-p)\int_0^{\Delta\tau}\kappa_{\mathrm r}e^{-\kappa_{\mathrm r}s}\,ds
=(1-p)(1-c)
<
1-p.
\]
Since $\rho=\pi_R(1-p)+\pi_D a$ is a strict convex combination of $a$ and $1-p$, one gets $a<\rho<1-p$.
\end{proof}

\begin{proof}[Proof of \cref{prop:recovery-main}: covariance formula]
The aggregate transition matrix is
\[
P = T_0+T_1 =
\begin{pmatrix}
 p & 1-p\\
 b & 1-b
\end{pmatrix}.
\]
Its eigenvalues are $1$ (with stationary left eigenvector $\pi$) and $\lambda=p-b$. The column vector $u=T_1\mathbf{1}=(1-p,\,a)^\top$ satisfies $\rho=\pi u$. Whenever $A_0=1$, the next hidden state is deterministically $D$, so
\[
\PP(A_0=1,A_n=1)=\rho\,e_D^\top P^{n-1}u,\qquad n\ge1.
\]
The two-state spectral decomposition gives
\[
P^{n-1}u = \rho\mathbf{1}+\lambda^{n-1}(u-\rho\mathbf{1}),
\]
so
\[
e_D^\top P^{n-1}u = \rho+\lambda^{n-1}(a-\rho),
\]
and therefore
\[
\operatorname{Cov}(A_0,A_n) = \PP(A_0=1,A_n=1)-\rho^2 = \rho(a-\rho)\lambda^{n-1}.
\]
The bounds $a<\rho<1-p$ are given by \cref{lem:stationary-main}. Hence $\operatorname{Cov}(A_0,A_1)=\rho(a-\rho)<0$: the record is negatively correlated at lag~$1$.
\end{proof}

\begin{proof}[Proof of exact $1$-dependence]
At the critical line~\eqref{eq:gamma-c-main}, $\lambda=p-b=0$, so $P=\mathbf{1}\pi$ (rank one), and hence $P^\ell=P$ for every $\ell\ge1$. Write $A_i^j=(A_i,\dots,A_j)$ and let $u=u_0\cdots u_r$ and $v=v_0\cdots v_s$ be visible words, with corresponding output-labelled kernel products
\[
T_u=T_{u_0}\cdots T_{u_r},
\qquad
T_v=T_{v_0}\cdots T_{v_s}.
\]
If the two words are separated by at least one sampling step, then for some $\ell\ge1$ their joint cylinder probability is
\[
\pi\,T_u\,P^\ell\,T_v\,\mathbf{1}.
\]
Using $P^\ell=P=\mathbf{1}\pi$ gives
\[
\pi\,T_u\,P^\ell\,T_v\,\mathbf{1}
=
\pi\,T_u\,(\mathbf{1}\pi)\,T_v\,\mathbf{1}
=
(\pi\,T_u\,\mathbf{1})(\pi\,T_v\,\mathbf{1}),
\]
which is the product of the two marginal cylinder probabilities. By linearity, any events $B\in\sigma(A_i:i\le m)$ and $C\in\sigma(A_j:j\ge m+2)$ factorize in the same way, so the process is exact $1$-dependent.

Note that the output is not a deterministic function of the hidden-state pair $(H_n,H_{n+1})$: both $T_0$ and $T_1$ have non-zero $(D,D)$ entries, so the same hidden transition $(D,D)$ can produce either visible outcome. The kernel argument above avoids that issue by working directly with output-labelled cylinders.
\end{proof}

\begin{remark}[Equal-rate limit and Lambert $W$ representation of $\Gamma_c$]\label{rem:edge-cases}
Proposition~\ref{prop:recovery-main} is stated for $\kappa_{\mathrm r}\neq\Gamma$. In the equal-rate limit $\kappa_{\mathrm r}\to\Gamma$, the expression~\eqref{eq:gap-closed-main} is a $0/0$ form. Applying l'H\^opital's rule in $\kappa_{\mathrm r}$ gives
\[
g_k\big|_{\kappa_{\mathrm r}=\Gamma}
=
p^k\bigl[(1-p)+\Gamma\Delta\tau\bigl(k(1-p)-p\bigr)\bigr],
\]
and the formulas for $a$, $b$, and $\lambda$ extend continuously in the same way. The hazard remains strictly increasing by continuity of the positivity argument in \cref{app:derivations}.

The critical rate $\Gamma_c$ is defined implicitly by~\eqref{eq:gamma-c-main}. Setting $s=\kappa_{\mathrm r}\Delta\tau$ and $x=\Gamma_c/\kappa_{\mathrm r}$, one rewrites that equation as $xe^{-sx}=e^{-s}$. With $u=-sx$, this becomes $ue^u=-se^{-s}$, so
\begin{equation}\label{eq:lambert-Gc}
\frac{\Gamma_c}{\kappa_{\mathrm r}}
=
-\frac{W_k(-se^{-s})}{s},
\qquad
k=
\begin{cases}
-1 & \text{if } 0<s<1, \\
 0 & \text{if } s>1,
\end{cases}
\end{equation}
with $\Gamma_c=\kappa_{\mathrm r}$ at $s=1$. For $s<1$ the non-trivial root lies on the $W_{-1}$ branch, whereas for $s>1$ it lies on the $W_0$ branch.
\end{remark}

\subsection{Spectral density and Fano factor}

\begin{proof}[Proof of \cref{prop:spectrum-main}]
From the covariance formula, $\operatorname{Cov}(A_0,A_n)=\rho(a-\rho)\lambda^{n-1}$ for $n\ge1$ with $|\lambda|<1$. Hence
\[
f(\theta) = \rho(1-\rho)+2\rho(a-\rho)\sum_{n\ge1}\lambda^{n-1}\cos(n\theta).
\]
The geometric sum evaluates via
\[
\sum_{n\ge1}\lambda^{n-1}e^{in\theta} = \frac{e^{i\theta}}{1-\lambda e^{i\theta}},
\]
taking real parts gives~\eqref{eq:spectrum-main}. Monotonicity: write $x=\cos\theta$ and $R_\lambda(x)=(x-\lambda)/(1-2\lambda x+\lambda^2)$; then $R_\lambda'(x)=(1-\lambda^2)/(1-2\lambda x+\lambda^2)^2>0$. Since $a-\rho<0$, the spectral density is decreasing in $x$, hence increasing in $\theta$ on $[0,\pi]$.

For the counting variance,
\[
\Var(N_N) = N\gamma_0 + 2\sum_{n=1}^{N-1}(N-n)\gamma_n
= N\rho(1-\rho) + 2\rho(a-\rho)\sum_{n=1}^{N-1}(N-n)\lambda^{n-1}.
\]
The geometric sum evaluates to $(N(1-\lambda)-1+\lambda^N)/(1-\lambda)^2$, giving~\eqref{eq:fano-main} in the limit $N\to\infty$. The strict inequality $F<1$ follows from $a-\rho<0$ and $1-\lambda>0$.
\end{proof}

\subsection{Two-state interval classification and gap inference}

\begin{proof}[Proof of \cref{thm:two-state-class-main}]
Let
\[
T_0=
\begin{pmatrix}
u & 0\\
v & w
\end{pmatrix},
\qquad
T_1=
\begin{pmatrix}
0 & 1-u\\
0 & a
\end{pmatrix},
\qquad
a=1-v-w,
\]
and set $P=T_0+T_1=\bigl(\begin{smallmatrix}u&1-u\\ v&1-v\end{smallmatrix}\bigr)$, $e_D=(0,1)^\top$, and $\lambda=u-v$.

Every click resets the next interval to the hidden state $D$. The visible record is therefore renewal, and the zero-run tail is
\[
\tau_k=\PP(G\ge k)=e_D^\top T_0^k\mathbf 1.
\]
Because $u\neq w$,
\[
T_0^k=
\begin{pmatrix}
u^k & 0\\
v\,\dfrac{u^k-w^k}{u-w} & w^k
\end{pmatrix},
\]
and hence
\[
\tau_k=
v\,\frac{u^k-w^k}{u-w}+w^k
=
\frac{v\,u^k+(\lambda-w)w^k}{u-w},
\]
which is~\eqref{eq:general-tail-main}. Since $\PP(G=k)=\PP(G\ge k)-\PP(G\ge k+1)$, the gap law follows.

For the hazard, write
\[
\tau_k=A u^k + B w^k,
\qquad
A=\frac{v}{u-w},
\qquad
B=\frac{\lambda-w}{u-w}.
\]
Then
\[
h_k
=
\frac{\tau_k-\tau_{k+1}}{\tau_k}
=
1-\frac{\tau_{k+1}}{\tau_k}.
\]
A direct calculation gives
\[
\tau_{k+1}^2-\tau_k\tau_{k+2}
=
-AB(u-w)^2u^kw^k
=
v(w-\lambda)u^kw^k.
\]
Therefore
\[
h_{k+1}-h_k
=
\frac{\tau_{k+1}^2-\tau_k\tau_{k+2}}{\tau_k\tau_{k+1}}
=
\frac{v(w-\lambda)u^kw^k}{\tau_k\tau_{k+1}},
\]
which is~\eqref{eq:general-hazard-diff-main}. Since $v>0$, $u^kw^k>0$, and $\tau_k>0$, the sign of $h_{k+1}-h_k$ is exactly the sign of $w-\lambda$.

The stationary distribution of $P$ is
\[
\pi_R=\frac{v}{1-u+v},
\qquad
\pi_D=\frac{1-u}{1-u+v}.
\]
Hence the stationary click probability is
\[
\rho=\pi_R(1-u)+\pi_Da
=
\frac{(1-u)(1-w)}{1-u+v}.
\]
Let $q=(1-u,a)^\top=T_1\mathbf 1$. As in the proof of \cref{prop:recovery-main},
\[
\PP(A_0=1,A_n=1)=\rho\,e_D^\top P^{n-1}q,
\qquad n\ge1.
\]
The eigenvalues of $P$ are $1$ and $\lambda$, so
\[
P^{n-1}q=\rho\mathbf 1+\lambda^{n-1}(q-\rho\mathbf 1).
\]
Evaluating at $e_D$ yields
\[
e_D^\top P^{n-1}q=\rho+\lambda^{n-1}(a-\rho),
\]
and therefore
\[
\operatorname{Cov}(A_0,A_n)=\rho(a-\rho)\lambda^{n-1},
\]
which is~\eqref{eq:general-cov-main}.

The difference between the post-refractory click probability and the stationary click probability is
\[
\rho-a
=
\frac{(1-u)(1-w)}{1-u+v}-(1-v-w)
=
\frac{v(w-\lambda)}{1-u+v}.
\]
Hence $a=\rho$ if and only if $\lambda=w$. If $\lambda=w$, then $v=u-w$ and the tail formula reduces to $\tau_k=u^k$, so
\[
\PP(G=k)=(1-u)u^k,
\]
which is the geometric zero-run law of a Bernoulli process with click probability $1-u$. Conversely, a Bernoulli process has constant hazard, and the hazard-difference formula forces $\lambda=w$. This proves the Bernoulli characterization.

If $\lambda=0$, then $P=\mathbf 1\pi$ has rank one. The cylinder-factorization argument from the proof of exact $1$-dependence above therefore applies verbatim, so the visible process is exact $1$-dependent. Conversely, if the visible process is exact $1$-dependent and non-Bernoulli, then $\operatorname{Cov}(A_0,A_2)=0$ while $a-\rho\neq0$, and the covariance formula gives $\lambda=0$. This proves the characterization of the non-Bernoulli exact $1$-dependence locus.

In the exponential-recovery model $(u,v,w)=(p,b,c)$, one has
\[
w-\lambda
=
c-(p-b)
=
-(p-c)+\frac{\kappa_{\mathrm r}(p-c)}{\kappa_{\mathrm r}-\Gamma}
=
\frac{\Gamma(p-c)}{\kappa_{\mathrm r}-\Gamma}.
\]
Because $p-c$ has the same sign as $\kappa_{\mathrm r}-\Gamma$, the final expression is strictly positive. Accordingly the model lies in the increasing-hazard sector. The identity $\lambda=0$ is exactly $p=b$, which is equivalent to \eqref{eq:gamma-c-main}.
\end{proof}

\subsection{Finite-dependence hierarchy for renewal records}

\begin{proof}[Proof of \cref{thm:renewal-md-main}]
Let
\[
u_n=\PP(A_n=1\mid A_0=1),
\qquad n\ge0,
\]
with $u_0=1$. If
\[
0=n_0<n_1<\cdots<n_r,
\]
then successive clicks are regeneration points, so repeated conditioning at the renewal times gives
\begin{equation}\label{eq:appendix-renewal-joint}
\PP(A_{n_0}=1,\dots,A_{n_r}=1)
=
\rho\prod_{j=1}^{r}u_{n_j-n_{j-1}}.
\end{equation}

Assume first that $(A_n)$ is exact $m$-dependent. For every $n\ge m+1$, the events $\{A_0=1\}$ and $\{A_n=1\}$ are independent. Hence
\[
u_n
=
\PP(A_n=1\mid A_0=1)
=
\PP(A_n=1)
=
\rho,
\qquad n\ge m+1.
\]
This proves \textup{(i)}$\Rightarrow$\textup{(ii)}.

Assume next that \textup{(ii)} holds. Let $E$ and $F$ be cylinder events determined respectively by finitely many coordinates in $(-\infty,0]$ and $[m+1,\infty)$. By inclusion--exclusion, it is enough to consider cylinders specifying finitely many ones. Thus let
\[
E=\bigcap_{i\in I}\{A_i=1\},
\qquad
F=\bigcap_{j\in J}\{A_j=1\},
\]
with finite sets $I\subset(-\infty,0]$ and $J\subset[m+1,\infty)$. Order $I\cup J$ as
\[
n_0<n_1<\cdots<n_r<n_{r+1}<\cdots<n_s,
\]
where $n_r\le0<n_{r+1}$. Stationarity and \eqref{eq:appendix-renewal-joint} give
\[
\PP(E\cap F)
=
\rho\prod_{j=1}^{s}u_{n_j-n_{j-1}}
=
\left(\rho\prod_{j=1}^{r}u_{n_j-n_{j-1}}\right)
\left(u_{n_{r+1}-n_r}\prod_{j=r+2}^{s}u_{n_j-n_{j-1}}\right)
\]
Since $n_{r+1}-n_r\ge m+1$, assumption \textup{(ii)} gives $u_{n_{r+1}-n_r}=\rho$. The same stationarity argument applied separately to the sets $I$ and $J$ yields
\[
\PP(E)=\rho\prod_{j=1}^{r}u_{n_j-n_{j-1}},
\qquad
\PP(F)=\rho\prod_{j=r+2}^{s}u_{n_j-n_{j-1}}.
\]
Consequently $\PP(E\cap F)=\PP(E)\PP(F)$.
The monotone-class theorem extends this factorization from cylinder events to all
\[
E\in\sigma(A_i:i\le0),
\qquad
F\in\sigma(A_j:j\ge m+1),
\]
so the process is exact $m$-dependent. This proves \textup{(ii)}$\Rightarrow$\textup{(i)}.

Let
\[
U(z)=\sum_{n\ge0}u_n z^n,
\qquad
F(z)=\sum_{n\ge1}f_n z^n,
\qquad |z|<1.
\]
The renewal equation
\[
u_n=\sum_{k=1}^{n}f_k u_{n-k},
\qquad n\ge1,
\]
implies
\[
U(z)=1+F(z)U(z),
\]
and therefore
\begin{equation}\label{eq:appendix-renewal-U}
U(z)=\frac{1}{1-F(z)}.
\end{equation}
If \textup{(ii)} holds, then the coefficient of $z^n$ in $(1-z)U(z)$ is $u_n-u_{n-1}$ for $n\ge1$, and these differences vanish for $n\ge m+2$. Hence
\[
(1-z)U(z)
=
1+\sum_{j=1}^{m}(u_j-u_{j-1})z^j+(\rho-u_m)z^{m+1}
=
Q_m(z).
\]
Combining this identity with \eqref{eq:appendix-renewal-U} yields \eqref{eq:renewal-q-main}. This proves \textup{(ii)}$\Rightarrow$\textup{(iii)}.

Conversely, if \textup{(iii)} holds, then
\[
U(z)=\frac{Q_m(z)}{1-z}.
\]
Since the coefficients of $(1-z)^{-1}$ are all equal to $1$, the coefficients of $U$ are partial sums of the coefficients of $Q_m$. Writing
\[
Q_m(z)=1+\sum_{j=1}^{m}(u_j-u_{j-1})z^j+(\rho-u_m)z^{m+1},
\]
one obtains
\[
u_n=\sum_{j=0}^{\min\{n,m+1\}}[z^j]Q_m(z)=\rho,
\qquad n\ge m+1.
\]
This proves \textup{(iii)}$\Rightarrow$\textup{(ii)} and completes the equivalence.

If the dependence order is exactly $m$, then the process is not exact $(m-1)$-dependent, so $u_m\neq\rho$. The coefficient of $z^{m+1}$ in $Q_m$ is therefore nonzero, and $\deg Q_m=m+1$.

Finally,
\[
Q_m(z)F(z)=Q_m(z)-1+z
\]
is a polynomial identity whose right-hand side has degree at most $m+1$. Writing
\[
Q_m(z)=1+(u_1-1)z+\sum_{j=2}^{m}(u_j-u_{j-1})z^j+(\rho-u_m)z^{m+1}
\]
and comparing the coefficient of $z^n$ for $n\ge m+2$ gives
\[
f_n+(u_1-1)f_{n-1}+\sum_{j=2}^{m}(u_j-u_{j-1})f_{n-j}+(\rho-u_m)f_{n-m-1}=0,
\]
which is exactly \eqref{eq:renewal-rec-main}.
\end{proof}

\begin{proof}[Proof of \cref{thm:gap-inference-main}]
Since $Y_1=G+1$ and \cref{prop:recovery-main} gives
\[
\PP_\Gamma(G=k)=
\frac{\kappa_{\mathrm r}(1-p)p^k-\Gamma(1-c)c^k}{\kappa_{\mathrm r}-\Gamma},
\qquad k\ge0,
\]
the moment generating function of $Y_1$ is the sum of two geometric series:
\[
M_\Gamma(t)
=
\sum_{k\ge0}e^{t(k+1)}\PP_\Gamma(G=k)
=
\frac{e^t}{\kappa_{\mathrm r}-\Gamma}
\left(
\frac{\kappa_{\mathrm r}(1-p)}{1-pe^t}
-
\frac{\Gamma(1-c)}{1-ce^t}
\right),
\]
valid whenever $|pe^t|<1$ and $|ce^t|<1$, that is, for $t<-\log\max\{p,c\}$. In particular, $M_\Gamma$ is finite on a neighbourhood of the origin, and $Y_1$ has moments of all orders.

The identity $\mu(\Gamma)=1/\rho(\Gamma)$ follows from the renewal law. To prove strict monotonicity, write
\[
\rho(\Gamma)=\frac{(1-p)(1-c)}{1-p+b},
\qquad
b(\Gamma)=\int_0^{\Delta\tau}\kappa_{\mathrm r}e^{-\kappa_{\mathrm r}s}e^{-\Gamma(\Delta\tau-s)}\,ds.
\]
Differentiation gives
\[
b'(\Gamma)
=
-\int_0^{\Delta\tau}\kappa_{\mathrm r}(\Delta\tau-s)e^{-\kappa_{\mathrm r}s}e^{-\Gamma(\Delta\tau-s)}\,ds<0,
\]
and therefore
\[
\rho'(\Gamma)
=
\frac{(1-c)\bigl[\Delta\tau\,p\,b-(1-p)b'(\Gamma)\bigr]}{(1-p+b)^2}
>0.
\]
Hence $\mu'(\Gamma)=-\rho'(\Gamma)/\rho(\Gamma)^2<0$. The map $\mu$ is therefore $C^\infty$ and strictly decreasing on $(0,\infty)$, with inverse defined on its image.

The variables $(Y_j)_{j\ge1}$ are i.i.d. The classical Cram\'er theorem and the contraction principle \cite{DemboZeitouni1998} yield the good large-deviation principle for $\bar Y_m$ with rate function~\eqref{eq:gap-rate-main}, and then for $\widehat\Gamma_m=\mu^{-1}(\bar Y_m)$ with rate function~\eqref{eq:gap-rate-gamma-main}. The ordinary i.i.d.\ central limit theorem gives
\[
\sqrt m\,(\bar Y_m-\mu(\Gamma))
\Longrightarrow
\mathcal N\!\bigl(0,\sigma_Y^2(\Gamma)\bigr).
\]
Applying the delta method to the $C^\infty$ inverse map $\mu^{-1}$ \cite{vanDerVaart1998} yields
\[
\sqrt m\,(\widehat\Gamma_m-\Gamma)
\Longrightarrow
\mathcal N\!\left(0,\frac{\sigma_Y^2(\Gamma)}{\mu'(\Gamma)^2}\right),
\]
which is~\eqref{eq:gap-clt-main}.

For the wrong-side error exponent, suppose first that $\Gamma>\Gamma_c$. Since $\mu$ is strictly decreasing, $\mu(\Gamma)<\mu(\Gamma_c)$ and
\[
\{\widehat\Gamma_m\le\Gamma_c\}
=
\{\bar Y_m\ge\mu(\Gamma_c)\}.
\]
The convex rate function $J_\Gamma$ has unique minimum at $\mu(\Gamma)$, so the infimum of $J_\Gamma$ on $[\mu(\Gamma_c),\infty)$ is attained at the boundary point $\mu(\Gamma_c)$. The large-deviation principle therefore gives
\[
\PP_\Gamma(\widehat\Gamma_m\le\Gamma_c)
=
\exp\!\bigl(-mJ_\Gamma(\mu(\Gamma_c))+o(m)\bigr)
=
\exp\!\bigl(-mI_\Gamma(\Gamma_c)+o(m)\bigr).
\]
The case $\Gamma<\Gamma_c$ is identical, with the set $(-\infty,\mu(\Gamma_c)]$ in place of $[\mu(\Gamma_c),\infty)$.
\end{proof}

\subsection{Quantitative shell control under spectral drift}

\begin{proof}[Proof of \cref{thm:spectral-drift-main}]
Write
\[
\sigma(r)=\log\mathcal S(r),
\qquad
g(r)=\log\Gamma_+(r)=2\log\lambda_c+\sigma(r)-L(r).
\]
By hypothesis,
\[
-\eta L'(r)\le \sigma'(r)\le \eta L'(r),
\qquad r\in I.
\]
Hence
\begin{equation}\label{eq:appendix-gprime-bounds}
-(1+\eta)L'(r)\le g'(r)\le -(1-\eta)L'(r)<0,
\qquad r\in I,
\end{equation}
because $L'(r)>0$ and $\eta<1$. Therefore $g$, and hence $\Gamma_+$, is strictly decreasing on $I$. This proves part~(i).

For part~(ii), let $T_1,T_2\in\Gamma_+(I)$. If $T_1=T_2$, the claim is trivial. Assume $T_1>T_2$ and write $r_j=r(T_j)$. Since $\Gamma_+$ is strictly decreasing, $r_1<r_2$. Using $g(r_j)=\log T_j$ gives
\[
\log\frac{T_1}{T_2}
=
g(r_1)-g(r_2)
=
\int_{r_1}^{r_2}\bigl(L'(r)-\sigma'(r)\bigr)\,dr.
\]
The slope bound implies
\[
(1-\eta)L'(r)\le L'(r)-\sigma'(r)\le (1+\eta)L'(r),
\qquad r\in[r_1,r_2].
\]
Integrating over $[r_1,r_2]$ yields
\[
(1-\eta)\bigl(L(r_2)-L(r_1)\bigr)
\le
\log\frac{T_1}{T_2}
\le
(1+\eta)\bigl(L(r_2)-L(r_1)\bigr).
\]
Dividing by the positive constants $1\pm\eta$ proves~\eqref{eq:threshold-comparison-main} in the case $T_1>T_2$; the general form with absolute values follows by symmetry.

For part~(iii), keep the notation $T_1>T_2$ and $r_1<r_2$. The left equality in~\eqref{eq:threshold-comparison-main} is equivalent to
\[
\log\frac{T_1}{T_2}
=(1-\eta)\bigl(L(r_2)-L(r_1)\bigr)
=
\int_{r_1}^{r_2}(1-\eta)L'(r)\,dr.
\]
Comparing with the identity
\[
\log\frac{T_1}{T_2}
=
\int_{r_1}^{r_2}\bigl(L'(r)-\sigma'(r)\bigr)\,dr
\]
shows that equality holds if and only if
\[
L'(r)-\sigma'(r)=(1-\eta)L'(r)
\qquad\text{for every }r\in[r_1,r_2],
\]
because the continuous nonnegative function
\[
r\longmapsto \bigl(L'(r)-\sigma'(r)\bigr)-(1-\eta)L'(r)
=
\eta L'(r)-\sigma'(r)
\]
has zero integral only when it vanishes identically. Thus $\sigma'(r)=\eta L'(r)$ on $[r_1,r_2]$, and integration gives
\[
\sigma(r)=\eta L(r)+\log K
\]
for some $K>0$, which is equivalent to~\eqref{eq:extremal-plus-main}. The converse is immediate by substitution.

The proof of the right equality is identical: equality holds if and only if
\[
L'(r)-\sigma'(r)=(1+\eta)L'(r)
\qquad\text{for every }r\in[r_1,r_2],
\]
that is, $\sigma'(r)=-\eta L'(r)$ on $[r_1,r_2]$. Integration yields~\eqref{eq:extremal-minus-main}, and substitution again proves the converse.
\end{proof}

\begin{proof}[Proof of \cref{cor:greybody-control-main}]
Applying \cref{thm:spectral-drift-main}(ii) with $T_1=\Gamma_\varphi$ and $T_2=\Gamma_{\mathrm{mem}}$ gives
\[
\frac{1}{1+\eta}\left|\log\frac{\Gamma_\varphi}{\Gamma_{\mathrm{mem}}}\right|
\le
\left|
\log
\frac{e^{\beta_\infty\Omega N(r_\varphi)}-1}
     {e^{\beta_\infty\Omega N(r_{\mathrm{mem}})}-1}
\right|
\le
\frac{1}{1-\eta}\left|\log\frac{\Gamma_\varphi}{\Gamma_{\mathrm{mem}}}\right|.
\]
Exponentiating yields~\eqref{eq:greybody-ratio-band-main}, because $C=\Gamma_{\mathrm{mem}}/\Gamma_\varphi$.

For the displacement estimate, fix $T\in\{\Gamma_\varphi,\Gamma_{\mathrm{mem}}\}$. The inhomogeneous and homogeneous shell equations give
\[
T=\lambda_c^2\,
\frac{\mathcal S_0e^{\sigma(r(T))}}{e^{\beta_\infty\Omega N(r(T))}-1}
=
\lambda_c^2\,
\frac{\mathcal S_0}{e^{\beta_\infty\Omega N(r_0(T))}-1}.
\]
Taking logarithms yields
\[
L(r(T))-L(r_0(T))=\sigma(r(T)),
\]
and therefore
\[
\left|L(r(T))-L(r_0(T))\right|
\le
\sup_I|\sigma|
\le
\varepsilon.
\]
This proves~\eqref{eq:greybody-log-displacement-main}. By the mean-value theorem, there exists $\xi_T\in I_T$ such that
\[
L(r(T))-L(r_0(T))=L'(\xi_T)\bigl(r(T)-r_0(T)\bigr).
\]
Since $L'(\xi_T)\ge m_T>0$,
\[
\left|r(T)-r_0(T)\right|
\le
\frac{\left|L(r(T))-L(r_0(T))\right|}{m_T}
\le
\frac{\varepsilon}{m_T},
\]
which is~\eqref{eq:greybody-radius-displacement-main}.
\end{proof}

\subsection{Exact shell rigidity beyond homogeneity}

\begin{proof}[Proof of \cref{thm:tolman-power-main}]
Assume \textup{(i)}. For every threshold $T\in\Gamma_+(I)$, the shell identity $T=\Gamma_+(r(T))$ and \eqref{eq:tolman-power-main} give
\[
T=A\,e^{-(1-\alpha)L(r(T))}.
\]
Hence
\[
L(r(T))
=
\frac{\log A-\log T}{1-\alpha}.
\]
This is \eqref{eq:tolman-affine-main} with
\[
q=\frac{1}{1-\alpha}>0,
\qquad
u=\frac{\log A}{1-\alpha}.
\]
Thus \textup{(i)}$\Rightarrow$\textup{(ii)}.

Assume \textup{(ii)}. For any distinct thresholds $T_1,T_2,T_3\in\Gamma_+(I)$,
\[
L(r(T_2))-L(r(T_1))
=
q(\log T_1-\log T_2),
\]
\[
L(r(T_3))-L(r(T_1))
=
q(\log T_1-\log T_3).
\]
Dividing the two identities gives \eqref{eq:tolman-crossratio-main}. Thus \textup{(ii)}$\Rightarrow$\textup{(iii)}.

Assume \textup{(iii)}. Fix two distinct thresholds $T_a,T_b\in\Gamma_+(I)$. For any threshold $T\in\Gamma_+(I)\setminus\{T_a,T_b\}$, apply \eqref{eq:tolman-crossratio-main} with $(T_1,T_2,T_3)=(T_a,T,T_b)$:
\[
\frac{L(r(T))-L(r(T_a))}{L(r(T_b))-L(r(T_a))}
=
\frac{\log T_a-\log T}{\log T_a-\log T_b}.
\]
Therefore
\[
L(r(T))
=
L(r(T_a))
+
\frac{L(r(T_b))-L(r(T_a))}{\log T_b-\log T_a}
(\log T_a-\log T).
\]
Setting
\[
q=
\frac{L(r(T_b))-L(r(T_a))}{\log T_b-\log T_a},
\qquad
u=L(r(T_a))+q\log T_a,
\]
one obtains
\[
L(r(T))=u-q\log T
\]
for every $T\in\Gamma_+(I)\setminus\{T_a,T_b\}$, and the same identity holds trivially at $T=T_a,T_b$. Since $r\mapsto\Gamma_+(r)$ is strictly decreasing and $r\mapsto L(r)$ is strictly increasing, the maps $T\mapsto r(T)$ and $T\mapsto L(r(T))$ are strictly decreasing. Accordingly $L(r(T_b))-L(r(T_a))$ and $\log T_b-\log T_a$ have the same sign, so $q>0$. This proves \textup{(iii)}$\Rightarrow$\textup{(ii)}.

Assume again \textup{(ii)}. For arbitrary $r\in I$, set $T=\Gamma_+(r)$; then $r(T)=r$, so \eqref{eq:tolman-affine-main} gives
\[
L(r)=u-q\log\Gamma_+(r).
\]
Equivalently,
\[
\Gamma_+(r)=e^{u/q}e^{-L(r)/q}.
\]
Define
\[
A=e^{u/q},
\qquad
\alpha=1-\frac{1}{q}.
\]
Since $q>0$, one has $\alpha<1$, and
\[
\Gamma_+(r)=A\,e^{-(1-\alpha)L(r)},
\]
which is \eqref{eq:tolman-power-main}. Using
\[
\Gamma_+(r)=\lambda_c^2\mathcal S(r)e^{-L(r)}
\]
then gives
\[
\mathcal S(r)=\frac{A}{\lambda_c^2}e^{\alpha L(r)}
=
\frac{A}{\lambda_c^2}\bigl(e^{\beta_\infty\Omega N(r)}-1\bigr)^\alpha.
\]
Thus \textup{(ii)}$\Rightarrow$\textup{(i)} and the equivalence is complete.

Under the equivalent conditions, the identities for $q$ and $u$ were already obtained in the proof of \textup{(i)}$\Rightarrow$\textup{(ii)}. Exponentiating the difference of the affine law at thresholds $T_1$ and $T_2$ gives
\[
\frac{e^{\beta_\infty\Omega N(r(T_1))}-1}
     {e^{\beta_\infty\Omega N(r(T_2))}-1}
=
e^{L(r(T_1))-L(r(T_2))}
=
e^{-q\log T_1+q\log T_2}
=
\left(\frac{T_2}{T_1}\right)^q,
\]
which is \eqref{eq:tolman-ratio-main}.

For the Schwarzschild inversion with known $\alpha$, set
\[
q=\frac{1}{1-\alpha},
\qquad
B=A^q.
\]
At a shell radius $r_i$ corresponding to threshold $T_i$, equation \eqref{eq:tolman-power-main} yields
\[
T_i^q
=
\frac{B}{e^{\beta_\infty\Omega N(r_i)}-1},
\qquad i=1,2,3.
\]
This is exactly the homogeneous three-threshold system of \cref{thm:three-threshold-main}, with common amplitude $B$ and threshold triple $(T_1^q,T_2^q,T_3^q)$. Since $q>0$, the threshold ordering is preserved. Therefore \cref{thm:three-threshold-main} determines $B$, $\beta_\infty$, and $M$ uniquely. Consequently $A=B^{1/q}$ is also uniquely determined.
\end{proof}

\subsection{Near-horizon expansion and local inverse theorem}

\begin{proof}[Proof of \cref{thm:horizon-main}]
In Gaussian normal coordinates from the bifurcation surface, the lapse expands as $N(\rho)=\kappa_{\mathrm H}\rho+O(\rho^3)$ with $\beta_\infty=2\pi/\kappa_{\mathrm H}$. Therefore
\[
e^{\beta_\infty\Omega N(\rho)}-1 = 2\pi\Omega\rho\,(1+O(\rho)).
\]
Applying \cref{thm:ratio-main} and cancelling the common factor $2\pi\Omega$ gives
\[
\frac{\rho_\varphi}{\rho_{\mathrm{mem}}} = \frac{\Delta\tau\,\Gamma_c}{\log\varphi}\,(1+O(\rho_\varphi+\rho_{\mathrm{mem}})).
\]
For the area excess in Schwarzschild, $d\rho=(1-2M/r)^{-1/2}\,dr$ gives $r-r_H=\rho^2/(8M)+O(\rho^4)$ near the horizon. Hence
\[
A(\rho)-A_H = 4\pi(r(\rho)^2-r_H^2) = 4\pi(2r_H)(r-r_H)+O(\rho^4) = 2\pi\rho^2+O(\rho^4),
\]
and the area-excess ratio is the square of the proper-distance ratio.
\end{proof}

\subsection{Amplitude-free three-threshold inversion}

\begin{proof}[Proof of \cref{thm:three-threshold-main}]
For the three operating modes, write
\[
A=\lambda_c^2\mathcal S(\Omega).
\]
The homogeneous shell equations are
\begin{equation}\label{eq:appendix-three-shell}
\frac{A}{e^{\beta_\infty\Omega N(r_i)}-1}=T_i,
\qquad i=1,2,3,
\end{equation}
with $N(r)=\sqrt{1-2M/r}$ in Schwarzschild. Hence
\begin{equation}\label{eq:appendix-three-log}
\beta_\infty\Omega N(r_i)=L_i(A)=\log\!\left(1+\frac{A}{T_i}\right),
\qquad i=1,2,3.
\end{equation}

Set
\[
\lambda_i=T_i^{-1},
\qquad
0<\lambda_1<\lambda_2<\lambda_3,
\]
and define
\[
f_A(\lambda)=\log(1+A\lambda)^2.
\]
Then
\[
\Xi(A)=
\frac{f_A(\lambda_2)-f_A(\lambda_1)}
     {f_A(\lambda_3)-f_A(\lambda_1)}.
\]
Since
\[
\partial_\lambda f_A(\lambda)
=
p_A(\lambda)
:=
\frac{2A\log(1+A\lambda)}{1+A\lambda}
>0,
\]
one has
\[
f_A(\lambda_j)-f_A(\lambda_i)=\int_{\lambda_i}^{\lambda_j}p_A(\lambda)\,d\lambda.
\]
Differentiating $p_A$ with respect to $A$ gives
\[
\partial_A p_A(\lambda)
=
\frac{2\bigl(\log(1+A\lambda)+A\lambda\bigr)}{(1+A\lambda)^2}.
\]
Therefore
\[
q_A(\lambda)
:=
\frac{\partial_A p_A(\lambda)}{p_A(\lambda)}
=
\frac{\log(1+A\lambda)+A\lambda}{A(1+A\lambda)\log(1+A\lambda)}.
\]
Writing $u=A\lambda$ yields
\[
q_A(\lambda)=\frac{1}{A}m(u),
\qquad
m(u)=\frac{\log(1+u)+u}{(1+u)\log(1+u)}.
\]
A direct differentiation gives
\[
m'(u)=
\frac{\log(1+u)-u-\log(1+u)^2}
{(1+u)^2\log(1+u)^2}
<0,
\qquad u>0,
\]
because $\log(1+u)<u$ for every $u>0$. Thus $q_A$ is strictly decreasing in $\lambda$.

Now set
\[
N(A)=f_A(\lambda_2)-f_A(\lambda_1),
\qquad
D(A)=f_A(\lambda_3)-f_A(\lambda_1),
\]
so that $\Xi(A)=N(A)/D(A)$. Then
\[
\frac{d}{dA}\log\Xi(A)
=
\frac{N'(A)}{N(A)}-\frac{D'(A)}{D(A)}
=
\frac{\int_{\lambda_1}^{\lambda_2}q_Ap_A\,d\lambda}
     {\int_{\lambda_1}^{\lambda_2}p_A\,d\lambda}
-
\frac{\int_{\lambda_1}^{\lambda_3}q_Ap_A\,d\lambda}
     {\int_{\lambda_1}^{\lambda_3}p_A\,d\lambda}.
\]
The first term is the $p_A$-weighted average of the decreasing function $q_A$ over $[\lambda_1,\lambda_2]$, while the second term is the corresponding weighted average over the strictly larger interval $[\lambda_1,\lambda_3]$. Writing the latter as the convex combination of the averages on $[\lambda_1,\lambda_2]$ and $[\lambda_2,\lambda_3]$, and using the strict decrease of $q_A$, yields
\[
\frac{\int_{\lambda_1}^{\lambda_2}q_Ap_A\,d\lambda}
     {\int_{\lambda_1}^{\lambda_2}p_A\,d\lambda}
>
\frac{\int_{\lambda_1}^{\lambda_3}q_Ap_A\,d\lambda}
     {\int_{\lambda_1}^{\lambda_3}p_A\,d\lambda}.
\]
Hence $\frac{d}{dA}\log\Xi(A)>0$, and therefore $\Xi'(A)>0$ for every $A>0$. This proves strict monotonicity.

The endpoint limits follow from elementary expansions. As $A\downarrow0$,
\[
L_i(A)=\frac{A}{T_i}+O(A^2),
\]
so
\[
\Xi(A)=
\frac{T_2^{-2}-T_1^{-2}}{T_3^{-2}-T_1^{-2}}
+O(A).
\]
As $A\to\infty$,
\[
L_i(A)=\log A-\log T_i+o(1),
\]
and therefore
\[
L_j(A)^2-L_i(A)^2
=
\log(T_i/T_j)\bigl(2\log A-\log T_i-\log T_j\bigr)+o(1).
\]
Consequently
\[
\lim_{A\to\infty}\Xi(A)=
\frac{\log(T_1/T_2)}{\log(T_1/T_3)}.
\]
This proves part~\textup{(i)}.

For part~\textup{(ii)}, square \eqref{eq:appendix-three-log} to obtain
\begin{equation}\label{eq:appendix-three-linear}
L_i(A)^2=\beta_\infty^2\Omega^2\left(1-\frac{2M}{r_i}\right),
\qquad i=1,2,3.
\end{equation}
Subtracting the identities for $(i,j)=(1,2)$ and $(1,3)$ gives
\[
\frac{L_2(A)^2-L_1(A)^2}{L_3(A)^2-L_1(A)^2}
=
\frac{r_1^{-1}-r_2^{-1}}{r_1^{-1}-r_3^{-1}}
=
\eta.
\]
Thus every compatible shell triple satisfies $\Xi(A)=\eta$. By part~\textup{(i)}, this equation has a unique positive solution if and only if $\eta$ lies in the image interval \eqref{eq:Xi-compatibility-main}; this proves part~\textup{(ii)}.

For part~\textup{(iii)}, multiply \eqref{eq:appendix-three-linear} by $r_i$:
\[
r_iL_i(A_\ast)^2
=
\beta_\infty^2\Omega^2(r_i-2M).
\]
Subtracting the identities for any pair $1\le i<j\le3$ gives
\[
r_iL_i(A_\ast)^2-r_jL_j(A_\ast)^2
=
\beta_\infty^2\Omega^2(r_i-r_j),
\]
which is \eqref{eq:beta-infty-three-main}. Solving the same two equations for $M$ gives \eqref{eq:mass-three-main}. Since $A_\ast$ is unique, these formulas determine $(\beta_\infty,M)$ uniquely as well.
\end{proof}

\begin{proof}[Proof of \cref{cor:three-threshold-stability-main}]
The strict monotonicity of $\Xi$ on $(0,\infty)$ implies that $\Xi^{-1}$ is $C^1$ on $\Xi(I_A)$. For every perturbed shell triple in $\mathcal O$, define
\[
\eta(\rho_1,\rho_2,\rho_3)
=
\frac{\rho_1^{-1}-\rho_2^{-1}}{\rho_1^{-1}-\rho_3^{-1}}.
\]
This map is $C^\infty$ on $\mathcal O$. By the mean-value theorem and the lower bound $\Xi'(A)\ge m_A$ on $I_A$,
\[
|\widehat A-A_\ast|
=
|\Xi^{-1}(\widehat\eta)-\Xi^{-1}(\eta)|
\le
\frac{|\widehat\eta-\eta|}{m_A},
\]
which proves \eqref{eq:amp-stability-main}.

Choose the neighborhood $U\Subset\mathcal O$ small enough that $\eta(U)\subset\Xi(I_A)$ and that the denominators in \eqref{eq:beta-infty-three-main} and \eqref{eq:mass-three-main} stay uniformly away from zero when $(A,\rho_1,\rho_2,\rho_3)\in I_A\times U$. On this compact set the maps
\[
(\rho_1,\rho_2,\rho_3)\longmapsto A=\Xi^{-1}(\eta(\rho_1,\rho_2,\rho_3)),
\]
\[
(\rho_1,\rho_2,\rho_3)\longmapsto \beta_\infty^2,
\qquad
(\rho_1,\rho_2,\rho_3)\longmapsto M
\]
are $C^1$. Their derivatives are therefore bounded on $U$. A second application of the mean-value theorem gives a constant $C_U>0$ such that
\[
|\widehat\beta_\infty^2-\beta_\infty^2|
+
|\widehat M-M|
\le
C_U\max_{1\le i\le3}|\widehat r_i-r_i|,
\]
which is \eqref{eq:geom-stability-main}.
\end{proof}

\section{Simulation methodology for Sections~\ref{subsec:finite-sample} and~\ref{subsec:noisy-inverse}}\label{app:monte}

All Monte Carlo statements in the main text were produced with reproducible scripts that sample directly from the analytic laws of the ideal and finite-recovery models. We use fixed seeds for each experiment; the precise generator is not important for the conclusions because all reported uncertainties are empirical standard deviations across independent replicates.

\paragraph{Finite-recovery diagnostics.} For a chosen $(\Gamma,\kappa_{\mathrm r},\Delta\tau)$ we draw successive zero-run lengths from the discrete distribution $g_k$ in~\eqref{eq:gap-closed-main} by inverse transform sampling: we precompute the cumulative distribution up to $k=200$ (which already captures probability $>0.999$ for the parameter ranges displayed) and sample from the remaining tail by rejection. Each simulated record contains $N_{\mathrm{samp}}$ sampled intervals after discarding an initial burn-in of $5\times10^3$ intervals. Hazards $\widehat h_k$ use the exact run-length counts, covariances $\widehat\gamma_n$ use the unbiased $(N_{\mathrm{samp}}-n)^{-1}$ estimator, and Fano factors $\widehat F$ use block sizes $B=50$. The figures in Sections~\ref{sec:records} and~\ref{subsec:finite-sample} use $\Delta\tau=0.3$, $\kappa_{\mathrm r}=5$, and $\Gamma$ ranging from $0.8\Gamma_c$ to $1.6\Gamma_c$ in steps of $0.05\Gamma_c$. For each grid point we simulate $5\times10^3$ independent records at $N_{\mathrm{samp}}=5\times10^3$, $5\times10^4$, and $10^5$ and report the empirical frequency with which $\widehat\gamma_2$ matches the sign of $\gamma_2$. The error bars quoted in the main text correspond to $\pm$ one empirical standard deviation divided by $\sqrt{\text{replicates}}$.

\paragraph{End-to-end shell pipeline.} The experiment of Section~\ref{subsec:noisy-inverse} uses the mode-$D_1$ design of \cref{tab:worked}. In the ideal-blanking mode we simulate records on the five-point radial grid
\[
r/r_{\mathrm s}\in\{1.006,\,1.007,\,1.008,\,1.009,\,1.010\},
\]
estimate
\[
\widehat p(r)=\frac{\#\{(A_n,A_{n+1})=(0,0)\}}{\#\{A_n=0\}},
\]
and linearly interpolate the zero of $\widehat p(r)-\varphi^{-1}$. In the finite-recovery mode we simulate records on the five-point grid
\[
r/r_{\mathrm s}\in\{1.041,\,1.043,\,1.045,\,1.047,\,1.049\},
\]
estimate the empirical click fraction $\widehat\rho(r)$, solve the monotone equation $\rho(\Gamma)=\widehat\rho(r)$ with $\rho(\Gamma)$ from~\eqref{eq:rho-main}, and then linearly interpolate the zero of $\widehat\Gamma(r)-\Gamma_c$. Each record again has $N_{\mathrm{samp}}=1.2\times10^5$ sampled intervals after a burn-in of $5\times10^3$ intervals. We use $120$ independent Monte Carlo replicates. Whenever both radial grids bracket the corresponding root, the resulting estimates $(\widehat r_\varphi,\widehat r_{\mathrm{mem}})$ are substituted into~\eqref{eq:beta-infty-main}--\eqref{eq:mass-main}; the reported uncertainties are empirical standard deviations across those end-to-end replicates.
