\IfFileExists{etds.cls}{%
  \documentclass{etds}%
}{%
  \documentclass[11pt,letterpaper]{article}%
}

\IfFileExists{etds.cls}{}{\usepackage[margin=1in]{geometry}}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{array}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{enumitem}

\numberwithin{equation}{section}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem*{propstar}{Proposition}

\newcommand{\EE}{\mathbb{E}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ind}{\mathbf{1}}
\newcommand{\leanverified}[1]{}
\DeclareMathOperator{\sgn}{sgn}

% Notation: \mathsf{P} denotes a measurable event (subset of the probability
% space X). For subsets of the symbolic shift space
% \Omega_\infty=\{0,1\}^\mathbb{N} we use symbols such as A\subseteq\Omega_\infty.
% When X=\Omega_\infty both viewpoints coincide and we keep the event notation
% \mathsf{P}.

\title{Scan Error Profiles on Prefix Partitions and Boundary-Controlled Convergence Rates}
\author{Haobo Ma\\{\small Research Department, CHRONOAI PTE.\ LTD., Singapore}\\{\small\texttt{haobo@chronoai.co}}}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We study how refinement of prefix information changes the optimal decision
error for a measurable event.  Given an increasing binary filtration
$(\mathcal{F}_m)_{m\ge 0}$, the scan error profile
$\varepsilon_m(\mathsf{P};\mu)$ is the smallest classification error
achievable using only $\mathcal{F}_m$-measurable rules.  The symbolic-dynamical
setup starts from events whose decisive prefix languages admit a finite-state
posterior presentation; the stronger \emph{observer-regular} hypothesis adds a
bounded resolving window.  In the Parry measure setting,
Theorem~\ref{thm:finite-state-posterior} shows that a finite
follower-resolving observer graph, refined by the current terminal symbol,
determines a higher-block invariant ambiguity exponent directly from
finite-state data and yields a uniform boundary-thickness constant under the
resolving refinement.  In the H\"older Gibbs setting,
Theorem~\ref{thm:weighted-boundary-pressure} shows that the same refined
ambiguous graph carries an event-relative boundary pressure whose reachable
recurrent components govern the weighted boundary partition sums through the
gap $P_X(\varphi)-P_\partial(\varphi)$; under the additional
observer-regular/resolving hypothesis, the same gap then transfers to
scan-error decay.  Theorems~\ref{ex:golden-mean-specialization}
and~\ref{ex:golden-mean-gibbs-weighted} make this explicit on the
golden-mean shift: for an explicit event $\mathsf{E}$ one gets an exact
parity-resolved scan-error formula with
$\delta_\partial(\mathsf{E})=1/2$, and for the locally constant Gibbs
potential $\varphi_t(x)=t\,\mathbf{1}_{\{x_0=0\}}$ one gets an exact weighted
boundary partition formula and, after the same resolving-thickness step, the
resulting scan-error exponent, with pressure gap
$P_X(\varphi_t)-\log\rho(M_{\partial,t})
=\frac12\log\!\left(\frac{1+\sqrt{1+4e^{-t}}}{2}\right)$ on the same refined
observer graph.  The exact atomwise Bayes decomposition, the discrete Tanaka
representation
$\varepsilon_m(\mathsf{P};\mu)=\varepsilon_0(\mathsf{P};\mu)-\EE[L_m^{1/2}]$,
and the transfer inequalities are included as preparatory filtration-level
tools that make these symbolic theorems effective.  Appendix~\ref{sub:nonuniform}
records a separate auxiliary extension to log-regular cylinder measures,
showing via a non-stationary product example that polynomial losses are
genuine in that nonuniform setting.  The observer-regularity hypothesis is an
additional event-specific assumption, not a generic property of arbitrary
events on sofic or hidden Markov systems.
\end{abstract}

\noindent\textbf{AMS 2020 Mathematics Subject Classification.}
Primary 37B10; Secondary 60G42, 28A80, 62C20.

\medskip
\noindent\textbf{Keywords.}
prefix filtrations; scan error; boundary cylinders; symbolic dynamics;
golden-mean shift; convergence rates.

\section{Introduction}
\label{sec:intro}

The purpose of this paper is narrow.  We fix a binary readout process and ask
how refinement of prefix information improves the best possible decision error
for a target event.  Let $(X,\mathcal{B},\mu)$ be a probability space with an
increasing filtration $(\mathcal{F}_m)_{m\ge 0}$ generated by $\{0,1\}$-valued
functions $s_0,s_1,\ldots$  The scan error profile
\[
m\longmapsto\varepsilon_m(\mathsf{P};\mu)
:=\inf\bigl\{\mu(\mathsf{P}\triangle\mathsf{Q}):\mathsf{Q}\in\mathcal{F}_m\bigr\}
\]
records the smallest classification error at each prefix depth.

The atom decomposition of Bayes risk on the finite partitions
$\mathcal{F}_m$ (equivalently, the cylinder decomposition on the symbolic
readout space) and the optimality of the threshold classifier
$p_m\ge 1/2$ are classical
\cite{Tsybakov2009Nonparametric,DevroyeGyorfiLugosi1996}.  We recall them in
Sections~\ref{sec:prefix}--\ref{sec:tanaka}; those sections only fix notation for the exact
boundary decomposition and its discrete Tanaka rewrite; the transfer bounds
later in Section~\ref{sec:rates} are then immediate once one inserts the
relevant cylinder-mass estimates.  The symbolic content begins only when one
imposes a finite-state observer on the ambiguous prefixes of the event.
Throughout the symbolic part, we call $\mathsf{P}$ \emph{observer-regular}
for a measure $\mu$ when the decisive prefix languages of $\mathsf{P}$ admit a
finite-state posterior presentation together with a bounded resolving window
with respect to $\mu$; this is an extra event-specific hypothesis, not an
automatic property of arbitrary events on sofic shifts or hidden Markov
sources.

The ETDS-facing contributions of the paper are:
\begin{enumerate}[label=(\roman*)]
\item A \emph{finite-state boundary observer theorem for $\mu$-finite-presented
  events on Parry SFTs, with thickness under the $\mu$-resolving refinement}
  (Theorem~\ref{thm:finite-state-posterior}): a finite follower-resolving
  observer graph determines both the thickness parameter and, together with
  Proposition~\ref{prop:higher-block-invariance}, a higher-block invariant
  normalized ambiguity exponent through the spectral radius of the refined
  ambiguous product graph on terminal symbols and observer states.
\item A \emph{componentwise boundary pressure theorem for $\mu_\phi$-finite
  presentations with respect to H\"older Gibbs states}
  (Theorem~\ref{thm:weighted-boundary-pressure}): the refined ambiguous
  product graph defines a boundary path shift, and the weighted boundary
  partition sum is controlled by the reachable recurrent boundary components,
  so the exponent for $Z_m(\partial\mathsf{P},\varphi)$ is the pressure gap
  $P_X(\varphi)-P_\partial(\varphi)$ with
  $P_\partial(\varphi)=\max_{C\in\mathfrak{C}_\partial}P_C(\varphi)$; under the
  separate $\mu_\phi$-resolving thickness hypothesis, the same gap transfers to
  scan error.
\item Two \emph{fully explicit golden-mean computations}:
  Theorem~\ref{ex:golden-mean-specialization} gives an exact parity-resolved
  scan-error formula with $\delta_\partial(\mathsf{E})=1/2$, and
  Theorem~\ref{ex:golden-mean-gibbs-weighted} gives an exact weighted boundary
  partition formula and, after verifying the resolving words, the resulting
  scan-error exponent on the same refined boundary graph.
\end{enumerate}

Sections~\ref{sec:prefix}--\ref{sec:tanaka} therefore stay brief: they supply
the exact boundary identity, the discrete Tanaka representation, and the
transfer statements needed later.  The new symbolic mechanism begins in
Subsection~\ref{sub:finite-posterior} with
Theorem~\ref{thm:finite-state-posterior},
Theorem~\ref{thm:weighted-boundary-pressure}, and the explicit golden-mean
computations of
Theorems~\ref{ex:golden-mean-specialization}
and~\ref{ex:golden-mean-gibbs-weighted}.  Appendix~\ref{sub:nonuniform}
isolates the auxiliary log-regular variant.

\subsection{Heuristic comparison with margin assumptions}
\label{sub:margin}

The thickness hypothesis in Corollary~\ref{thm:volume-boundary-scaling} is
best compared heuristically with margin and low-noise conditions in
nonparametric classification.
Mammen and Tsybakov \cite{MammenTsybakov1999} introduce a margin condition
with exponent $\alpha\ge 0$: for the regression function
$p_\infty(x)=\mu(\mathsf{P}\mid\mathcal{F}_\infty)(x)$,
\[
\mu\!\left(\left|p_\infty(x)-\tfrac12\right|\le t\right)\le C_0\,t^\alpha
\qquad\text{for all }t>0.
\]
Tsybakov \cite{Tsybakov2004} extends this to a general noise-exponent
framework, showing that margin conditions control the rate of optimal
plug-in classifiers.

The exact boundary decomposition of Proposition~\ref{prop:scan-error-cylinder}
is
\[
\varepsilon_m(\mathsf{P};\mu)
=\sum_{C\in \partial_m\mathsf{P}}
\mu(C)\min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\}.
\]
Hence the uniform thickness condition
$\min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\}\ge\theta>0$ on
boundary cylinders implies the lower bound
\[
\varepsilon_m(\mathsf{P};\mu)\ge
\theta\sum_{C\in\partial_m\mathsf{P}}\mu(C),
\]
so thickness controls scan error through the \emph{mass} of the depth-$m$
boundary.  What it does \emph{not} control is the global distribution of the
terminal posterior $p_\infty$ near the threshold $1/2$; that global
distribution is exactly what the Mammen--Tsybakov condition measures.  Thus
the two notions live at different scales: thickness is a cylinderwise lower
bound on ambiguous atoms, while the margin condition is an integrated bound on
how much of the space has $p_\infty$ close to $1/2$.  Without additional
geometric information, neither hypothesis implies the other, so the comparison
should be read only as orientation.

The closer multiscale analogue is the dyadic tree analysis of
Scott and Nowak \cite{ScottNowak2006Dyadic}, where partition geometry and
near-boundary noise are coupled in a multiscale classifier.  Tsybakov's noise
framework remains the right broad backdrop for fast-rate phenomena, but it
should not be read as giving a direct substitute for the thickness hypothesis,
and we do not identify either $\theta$ or the boundary-growth exponents with a
classical margin exponent.
The precise relationship between the higher-block invariant ambiguity exponent
in the finite-state symbolic setting, or the raw filtration-level boundary
count exponent in the general setting, and the margin exponent $\alpha$
remains open.

\subsection{Related work}
\label{sub:related}

Bayes risk on partitions and optimality of threshold classifiers are classical;
see \cite{DevroyeGyorfiLugosi1996} and \cite{AudibertTsybakov2007} for refined
plug-in rates.  For dyadic tree methods that explicitly couple multiscale
partition geometry and low-noise behavior near a decision boundary, see
\cite{ScottNowak2006Dyadic}.  The underlying martingale convergence is Doob's
upward theorem \cite{Doob1953,Williams1991,Durrett2019}.  The discrete Tanaka
formula is a lattice analogue of Tanaka's formula
\cite{Tanaka1963,RevuzYor1999}; throughout,
$L_m^{1/2}$ is used only as a discrete threshold local-time functional, not
as an occupation-density local time.

Cylinder partition structure in symbolic dynamics is treated in
\cite{LindMarcus1995,Kitchens1998}; boundary combinatorics connect to
\cite{Fischer1975,Krieger1984,Boyle1983}.  Same-length cylinder bounds for the
Parry (maximal-entropy) measure are in \cite{Parry1964}; the transfer-operator
and thermodynamic-formalism background for H\"older equilibrium states is
classical in
\cite{Ruelle1978,ParryPollicott1990,Baladi2000,Bowen1975,WaltersErgodicTheory}.
Finite-state observable factors of Markov chains go back to Blackwell and
Heller \cite{Blackwell1957,Heller1965}, while deterministic finite automata,
their synchronous product construction, and regular-language recognition are
standard in \cite{HopcroftMotwaniUllman2001}; for sofic shifts this observer
viewpoint is standard in Fischer and Krieger presentations and their later
symbolic-dynamical treatments
\cite{Fischer1975,Krieger1984,Kitchens1998,LindMarcus1995}.  Our finite-state posterior
theorem uses this observer philosophy for a different object: not the output
process itself, but the evolving ambiguity class of a fixed event
$\mathsf{P}$, from which thickness and higher-block invariant boundary growth
become finite-state spectral quantities.
For the literature closest to the symbolic-dynamics side of our setup, Boyle
and Petersen \cite{BoylePetersen2011} survey hidden Markov (sofic) measures
and the factor-map questions around them.  The relative-pressure background
starts with the relativized variational principle of Ledrappier and Walters
\cite{LedrappierWalters1977}, Walters's treatment of relative pressure,
relative equilibrium states, and compensation functions
\cite{Walters1986RelativePressure}, the measures-of-maximal-relative-entropy
analysis of Petersen, Quas, and Shin \cite{PetersenQuasShin2003}, and the
comparison of relative-pressure definitions by Petersen and Shin
\cite{PetersenShin2005}.  More recent developments include Chazottes and
Ugalde \cite{ChazottesUgalde2003,ChazottesUgalde2011}, Pollicott and Kempton
\cite{PollicottKempton2011}, Piraino \cite{Piraino2018}, Feng
\cite{Feng2011FactorMaps}, Yoo \cite{Yoo2018Decomposition}, Allahbakhshi,
Antonioli, and Yoo \cite{AllahbakhshiAntonioliYoo2019}, Antonioli, Hong, and
Quas \cite{AntonioliHongQuas2023}, and Yayama
\cite{Yayama2023RelativePressure,Yayama2026RelativePressureGibbs}, who study
factor codes, relative pressure, relative equilibrium states, and Gibbs
structure for projected or fibered measures.  In particular, the recent ETDS
paper of Yayama \cite{Yayama2026RelativePressureGibbs} analyzes almost
additive sequences arising from relative pressure functions and gives new
criteria for when factor images of Gibbs measures remain Gibbs.  In a
different direction, Garc{\'\i}a-Ramos and Pavlov
\cite{GarciaRamosPavlov2019} use extender-set containment to compare cylinder
weights for measures of maximal entropy.

The thermodynamic language around our quantity $P_\partial(\varphi)$ is also
close in notation, but not in definition, to the \emph{weighted topological
pressure} literature.  Barral and Feng
\cite{BarralFeng2012WeightedThermodynamic}, Feng and Huang
\cite{FengHuang2016WeightedPressure}, Tsukamoto
\cite{Tsukamoto2023WeightedPressure}, and Alibabaei
\cite{Alibabaei2025WeightedPressure} study weighted entropy and pressure for
factor towers or factor maps, with an explicit weight vector and a variational
principle mixing the entropy of the ambient system with the entropies of its
projected systems.  By contrast, the present paper fixes a single event
$\mathsf{P}$ and builds a boundary observer for its ambiguous prefixes; the
quantity $P_\partial(\varphi)$ is then the ordinary topological pressure of
$\varphi\circ\pi_\partial$ on the event-dependent boundary path shift cut out
by that observer.  Thus our pressure is \emph{event-relative} and
\emph{observer-relative}, not a weighted-pressure invariant attached to a
factor map or an entropy-weighted tower.

Recent ETDS work at the symbolic-information interface also includes
MacDonald \cite{MacDonald2024}, who studies when a lower-entropy subshift can
be encoded injectively through a prescribed sliding-block code after
restricting the domain, and Han, Marcus, and Wu \cite{HanMarcusWu2024}, who
characterize when factor codes with an unambiguous symbol admit one-to-one or
finite-to-one SFT restrictions and relate the finite-to-one case to Markov
capacity.  These papers are close in spirit to our finite-state observer
viewpoint because they isolate finite symbolic obstructions to ambiguity under
sliding-block coding; however, the object tracked here is different.  Our
observer records posterior ambiguity for a fixed event $\mathsf{P}$ rather
than fiber multiplicities or injective restrictions of a factor code.  The
extender-set methods of Garc{\'\i}a-Ramos and Pavlov are likewise
conceptually adjacent, because both frameworks control cylinder weights
through finitely describable continuation classes; in our case those classes
are the refined ambiguous boundary states and their induced boundary
cylinders.

These methods are adjacent to, but distinct from,
Theorem~\ref{thm:finite-state-posterior}.  Factor-code and relative-pressure
results begin with a sliding-block factor map $\pi:X\to Y$ and analyze the
thermodynamic structure of the pushed-forward measure or of measures in a
fiber $\pi_*^{-1}\{\nu\}$: induced potentials, compensation functions,
Ledrappier--Walters relative pressure, and the number or uniqueness of
relative equilibrium states.  Extender-set methods compare the admissible
continuation contexts of words in the ambient shift and derive entropy or
measure inequalities from those continuation classes.  Our finite-state
posterior theorem does neither.  It fixes a single event
$\mathsf{P}\in\mathcal{F}_\infty$, builds a finite observer for the
\emph{event-relative} ambiguous prefix classes of $\mathsf{P}$, and then
reads off boundary thickness and the higher-block invariant ambiguity exponent
from the refined ambiguous product graph.  Accordingly, the boundary pressure
$P_\partial(\varphi)$ introduced later is an event-relative weighted boundary
growth rate for this observer graph, not relative pressure for a factor map
and not a thermodynamic invariant attached to fibers of $\pi$.  Thus the
output is a boundary-counting and scan-error theorem for one event, not a
factor-closure theorem for Gibbs classes and not a relative-pressure
characterization of projected measures.

\paragraph{Theorem-by-theorem ETDS comparison against cited adjacent results.}
Table~\ref{tab:etds-novelty-boundary} records only theorem-level symbolic
results, because Proposition~\ref{prop:scan-error-cylinder},
Theorem~\ref{thm:scan-tanaka}, and the transfer inequalities are classical
filtration-level tools rather than new symbolic-dynamical claims.  The table
is a local theorem-by-theorem comparison with the ETDS papers cited in this
manuscript whose statements overlap most directly with the present symbolic
results, namely the cited papers on sliding-block/factor-code ambiguity and on
relative or weighted pressure for factor maps.  It should therefore be read as
an adjacency comparison for the specific cited results below, not as a census
claim about all ETDS papers in any fixed publication window.

\begin{table}[t]
\centering\tiny
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{0.92}
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{1.06in}>{\raggedright\arraybackslash}p{1.63in}>{\raggedright\arraybackslash}p{1.76in}>{\raggedright\arraybackslash}p{1.72in}@{}}
\toprule
present theorem & hypotheses and conclusion here & closest cited ETDS comparison point & exact increment over that comparison \\
\midrule
Theorem~\ref{thm:finite-state-posterior} &
Mixing one-step SFT with Parry measure; event $\mathsf{P}$ has a finite-state
posterior presentation and bounded resolving window.  The refined ambiguous
matrix $\widetilde B$ counts boundary cylinders, gives
$\Lambda_\partial(\mathsf{P})=\rho(\widetilde B)$ and the higher-block
invariant exponent $\delta_\partial(\mathsf{P})$, yields a uniform thickness
constant, and hence gives two-sided scan-error bounds. &
MacDonald \cite{MacDonald2024} (doi: 10.1017/etds.2023.56; arXiv:2210.08150):
for a surjective sliding-block code $\pi:X\to Y$ from a mixing SFT onto a
mixing sofic shift and a lower-entropy subshift $Z$, gives necessary and
sufficient conditions for an embedding $\psi:Z\to X$ with $\pi\circ\psi$
injective.  Han--Marcus--Wu \cite{HanMarcusWu2024} (doi:
10.1017/etds.2023.103; arXiv:2210.12251): for factor codes with an
unambiguous symbol, gives criteria for one-to-one or finite-to-one SFT
restrictions. &
The present theorem fixes no factor map and no restricted subsystem.  Its new
object is an event-relative posterior observer built from the decisive prefix
languages of one event.  From that observer it computes boundary growth,
thickness, higher-block invariant ambiguity exponent, and scan-error rate for
that event, which the factor-code restriction theorems do not address. \\
\addlinespace
Theorem~\ref{thm:weighted-boundary-pressure} &
Mixing one-step SFT, H\"older potential $\phi$, and an event with a
finite-state posterior presentation for $\mu_\phi$.  The weighted boundary
sum equals the exact finite-path sum on the refined observer graph; if no
reachable recurrent boundary component exists then it vanishes eventually,
otherwise the exponential rate is
$P_X(\phi)-P_\partial(\phi)$, with explicit periodic lower bounds and the
polynomial factor $m^{q_\partial-1}$ when several maximal components are
reachable. &
Yayama \cite{Yayama2023RelativePressure} (doi: 10.1017/etds.2022.30;
arXiv:2109.02180): for factor maps from an irreducible SFT and subadditive
sequences associated to relative pressure, studies existence of an additive
representative and equilibrium states.  Tsukamoto
\cite{Tsukamoto2023WeightedPressure} (doi: 10.1017/etds.2021.173;
arXiv:2108.03795) and Alibabaei \cite{Alibabaei2025WeightedPressure} (doi:
10.1017/etds.2024.35; arXiv:2307.16772): develop weighted pressure for factor
maps or towers.  Yayama \cite{Yayama2026RelativePressureGibbs} (doi:
10.1017/etds.2025.10262; arXiv:2402.10199): studies Gibbs images under
one-block factor maps via relative-pressure sequences. &
The theorem here has no factor map, no fiber, and no weighted tower.  It
defines ordinary topological pressure on an event-dependent refined boundary
graph, identifies exactly which reachable recurrent components control the
rate for $Z_m(\partial\mathsf{P},\phi)$, keeps acyclic tails and residue
classes explicit, and then, after the separate resolving-thickness step,
transfers that pressure gap to the scan-error profile of the fixed event. \\
\addlinespace
Theorem~\ref{ex:golden-mean-specialization} &
For the explicit event $\mathsf{E}$ on the golden-mean shift, the theorem
gives the exact refined matrix $B_{\mathsf E}$, the exact counts
$N_{2k}(\partial\mathsf{E})=N_{2k+1}(\partial\mathsf{E})=F_{k+2}$, the exact
value $\delta_\partial(\mathsf{E})=1/2$, a uniform thickness bound, and a
closed parity-resolved scan-error formula. &
Among the cited ETDS papers with overlapping symbolic content, the closest
comparison points are again the ambiguity/coding papers of
MacDonald \cite{MacDonald2024} and Han--Marcus--Wu \cite{HanMarcusWu2024},
because they also isolate finite symbolic obstructions to ambiguity, but their
conclusions are existence and restriction criteria for coding problems rather
than exact formulas for one event on one SFT. &
This theorem is not merely a specialization of the general observer result to
an example.  It upgrades the finite-state mechanism to a closed-form,
citable computation: exact observer matrix, exact ambiguity exponent,
exact scan-error constants, and explicit parity dependence on a fixed
event/SFT pair. \\
\addlinespace
Theorem~\ref{ex:golden-mean-gibbs-weighted} &
For the same event $\mathsf{E}$ and the Gibbs family
$\varphi_t(x)=t\,\mathbf{1}_{\{x_0=0\}}$, the theorem gives the exact weighted
boundary matrix $M_{\partial,t}$, the closed formula
$P_\partial(\varphi_t)=\frac12(t+\log\lambda_t)$, the exact partition-sum
identity
$Z_m(\partial\mathsf{E},\varphi_t)=e^{-mP_X(\varphi_t)}
\alpha_tM_{\partial,t}^{m-2}\mathbf{1}$, and therefore the exact pressure-gap
exponent for the weighted boundary partition sum and, after the explicit
resolving-thickness step, for the scan-error law. &
Among the cited ETDS papers with overlapping thermodynamic content, the
closest comparison points are Tsukamoto
\cite{Tsukamoto2023WeightedPressure}, Alibabaei
\cite{Alibabaei2025WeightedPressure}, and Yayama
\cite{Yayama2026RelativePressureGibbs}, which treat weighted pressure or Gibbs
images for factor maps and almost-additive relative-pressure sequences. &
The increment is a closed-form event-relative pressure computation on a single
finite boundary graph.  The theorem does not give a variational principle for
a factor tower or a Gibbs-image criterion for a factor map; it computes the
actual matrix, the actual pressure gap, and, after the explicit
resolving-thickness step, the resulting scan-error exponent for one explicit
observer graph. \\
\bottomrule
\end{tabular}
\caption{Theorem-by-theorem comparison with the cited adjacent ETDS results
whose statements most directly overlap with the present symbolic theorems.
Each row lists the hypotheses and conclusion of the present theorem, the
closest cited theorem-level comparison point, and the precise increment proved
here.}
\label{tab:etds-novelty-boundary}
\renewcommand{\arraystretch}{1}
\end{table}

Table~\ref{tab:etds-novelty-boundary} is the comparison table used in the
introduction and again in Section~\ref{sec:discussion}.  Its role is a local
theorem-by-theorem separation from the cited adjacent ETDS results above, not
a bibliographic maximality claim.  In that local comparison,
Theorem~\ref{thm:finite-state-posterior} adds an event-relative observer that
computes boundary exponents and scan error, beyond factor-code restriction
criteria.  Theorem~\ref{thm:weighted-boundary-pressure} adds an event-defined
boundary graph whose reachable recurrent components govern the pressure gap,
beyond relative-pressure or weighted-pressure statements for factor maps.
Theorems~\ref{ex:golden-mean-specialization}
and~\ref{ex:golden-mean-gibbs-weighted} then push those two general results to
closed formulas on one explicit SFT/event pair.
The Shannon--McMillan--Breiman theorem \cite{Shannon1948,McMillan1953,Breiman1957}
gives almost-sure asymptotics for the cylinder \emph{containing} a given
sample point; this is an almost-sure pointwise statement, weaker than the
uniform all-cylinder bounds assumed in our transfer principles (see
Section~\ref{sec:discussion}).  Background on dynamics and information theory
is in \cite{WaltersErgodicTheory,LindMarcus1995,CoverThomas2006,MorseHedlund1940}.

\section{Preliminaries: Prefix Filtration Setup}
\label{sec:prefix}

Sections~\ref{sec:prefix}--\ref{sec:tanaka} hold in the following general
setting; no invariance or symbolic structure is used there.

\begin{definition}[Prefix filtration]
\label{def:prefix-filtration}
Let $(X,\mathcal{B},\mu)$ be a probability space and $(s_t)_{t\ge 0}$ a
sequence of $\{0,1\}$-valued measurable functions.  Set $\mathcal{F}_0:=\{\varnothing,X\}$
and $\mathcal{F}_m:=\sigma(s_0,\ldots,s_{m-1})$ for $m\ge 1$.  The family
$(\mathcal{F}_m)_{m\ge 0}$ is the \emph{prefix filtration}; set
$\mathcal{F}_\infty:=\bigvee_m\mathcal{F}_m$.
\end{definition}

Let $\Omega_\infty:=\{0,1\}^{\mathbb{N}}$ and, for
$a=a_0\cdots a_{m-1}\in\{0,1\}^m$, write
\[
[a]:=\{\xi\in\Omega_\infty:\xi_0\cdots\xi_{m-1}=a\},
\qquad
\mathcal{A}_m:=\sigma([a]:a\in\{0,1\}^m).
\]
The readout map
\[
\omega_\infty:X\to\Omega_\infty,
\qquad
\omega_\infty(x):=(s_0(x),s_1(x),\ldots),
\]
is measurable and satisfies
\[
\mathcal{F}_m=\omega_\infty^{-1}(\mathcal{A}_m)
\qquad(m\ge 0).
\]
In the symbolic specialization $X\subseteq\Omega_\infty$ with the coordinate
readout, $\omega_\infty$ is the inclusion map and the positive atoms of
$\mathcal{F}_m$ are exactly the positive length-$m$ cylinders of $X$.  This is
the only symbolic identification needed in Sections~\ref{sec:prefix}--\ref{sec:tanaka}.

\section{Preliminaries: The Scan Error Profile}
\label{sec:scan}

\begin{definition}[Scan error profile and admissible classifiers]
\label{def:scan-error-profile}
Let $\mathsf{P}\in\mathcal{B}$ be a target event.  A depth-$m$
\emph{admissible classifier} is an event $\mathsf{Q}\in\mathcal{F}_m$.
The depth-$m$ scan error is
\[
\varepsilon_m(\mathsf{P};\mu)
:=\inf\{\mu(\mathsf{P}\triangle\mathsf{Q}):\mathsf{Q}\in\mathcal{F}_m\}.
\]
Classifiers and target events are identified up to $\mu$-null symmetric
difference.  Thus statements such as $\mathsf{P}=\mathsf{Q}$, membership in
$\mathcal{F}_m$, and ``$\mathsf{P}$ is decided by depth $m$'' are understood
in the $\mu$-a.s.\ sense whenever null sets are irrelevant; this convention
does not change $\varepsilon_m(\mathsf{P};\mu)$.
\end{definition}

\begin{definition}[Positive-atom convention and conditional probabilities]
\label{def:positive-atoms}
For each $m$, $\mathcal{C}_m$ denotes the finite family of positive atoms of
$\mathcal{F}_m$ on $X$, after identifying atoms that agree modulo
$\mu$-null sets.
Equivalently, one may take the atoms of the finite algebra $\mathcal{F}_m$ and
discard the atoms of $\mu$-measure zero.  All sums and cardinalities over
$\mathcal{C}_m$ below are over this positive-atom family.  If
$C\in\mathcal{C}_m$, then
\[
\mu(\mathsf{A}\mid C):=\frac{\mu(\mathsf{A}\cap C)}{\mu(C)}
\]
for every event $\mathsf{A}$.  In the symbolic specialization
$X\subseteq\Omega_\infty$ with the coordinate readout, the elements of
$\mathcal{C}_m$ are exactly the positive length-$m$ cylinders in $X$.  Thus
all filtration-level statements below are atomwise on $X$; when later phrased
symbolically, they are the corresponding cylinder statements.
\end{definition}

\begin{definition}[Boundary atoms]
\label{def:boundary-atoms}
Let $\mathsf{P}\in\mathcal{B}$.  For each depth $m\ge 0$, define the
\emph{boundary family}
\[
\partial_m\mathsf{P}
:=\{C\in\mathcal{C}_m:\mu(\mathsf{P}\cap C)>0\text{ and }\mu(C\setminus\mathsf{P})>0\},
\]
with cardinality $N_m(\partial\mathsf{P}):=\#(\partial_m\mathsf{P})$.
Thus $\partial_m\mathsf{P}$ is the collection of positive depth-$m$ atoms on
which the target event is still undecided.  In the symbolic specialization,
these are exactly the positive boundary cylinders at depth $m$.
\end{definition}

\begin{proposition}[Exact atom decomposition]
\label{prop:scan-error-cylinder}
Let $\mathcal{C}_m$ be the positive atoms of $\mathcal{F}_m$ and
$p_m:=\mu(\mathsf{P}\mid\mathcal{F}_m)$.
For every $\mathsf{P}\in\mathcal{B}$ and $m\ge 0$,
\[
\varepsilon_m(\mathsf{P};\mu)
=\sum_{C\in\mathcal{C}_m}\min\{\mu(\mathsf{P}\cap C),\mu(C\setminus\mathsf{P})\}
=\EE[\min\{p_m,1-p_m\}].
\]
\end{proposition}

\begin{proof}
Any $\mathcal{F}_m$-measurable event $\mathsf{Q}$ is, modulo a null set, a
union of positive atoms.  On each positive atom $C$, inclusion contributes
local error $\mu(C\setminus\mathsf{P})$ and exclusion contributes
$\mu(\mathsf{P}\cap C)$; the optimum is their minimum.  Since
$\min\{\mu(\mathsf{P}\cap C),\mu(C\setminus\mathsf{P})\}=\mu(C)\min\{p_C,1-p_C\}$
where $p_C:=\mu(\mathsf{P}\mid C)$, summing over atoms gives the expectation form.
\end{proof}

\begin{corollary}[Monotonicity and decidability]
\label{cor:basic-properties}
For every $\mathsf{P}\in\mathcal{B}$, the profile
$m\mapsto\varepsilon_m(\mathsf{P};\mu)$ is nonincreasing, and
$\varepsilon_m(\mathsf{P};\mu)=0$ iff $\mathsf{P}$ coincides $\mu$-a.s.\ with
an event in $\mathcal{F}_m$.
\leanverified{paper_prefix_scan_error_basic_properties}
\end{corollary}

\begin{proof}
Since $\mathcal{F}_m\subseteq\mathcal{F}_{m+1}$, the admissible classifiers at
depth $m+1$ contain those at depth $m$, and the infimum defining
$\varepsilon_{m+1}$ is no larger than the infimum defining $\varepsilon_m$.
For the zero-error criterion, Proposition~\ref{prop:scan-error-cylinder}
shows that $\varepsilon_m=0$ precisely when, on every positive atom
$C\in\mathcal{C}_m$, either $\mu(\mathsf{P}\cap C)=0$ or
$\mu(C\setminus\mathsf{P})=0$.  Let $\mathsf{Q}$ be the union of the positive
atoms $C$ for which $\mu(C\setminus\mathsf{P})=0$.  Then
$\mathsf{Q}\in\mathcal{F}_m$ and $\mu(\mathsf{P}\triangle\mathsf{Q})=0$.
The converse is immediate from the definition of $\varepsilon_m$, since an
$\mathcal{F}_m$-event equal to $\mathsf{P}$ modulo null sets has zero
symmetric-difference error.
\end{proof}

\section{Preliminaries: Discrete Tanaka Representation}
\label{sec:tanaka}

The quantity $L_m^{1/2}$ below is the discrete threshold compensator for the
posterior martingale $p_m=\mu(\mathsf{P}\mid\mathcal{F}_m)$ at level $1/2$.

\begin{theorem}[Discrete Tanaka representation of the scan error]
\label{thm:scan-tanaka}
Let $\mathsf{P}\in\mathcal{B}$ and set
\[
p_m:=\mu(\mathsf{P}\mid\mathcal{F}_m),
\qquad
p_\infty:=\mu(\mathsf{P}\mid\mathcal{F}_\infty).
\]
For $m\ge 1$ define the discrete threshold local-time functional at level
$1/2$ by
\[
L_m^{1/2}
:=\sum_{k=0}^{m-1}
\left(
\left|p_{k+1}-\tfrac12\right|
-\left|p_k-\tfrac12\right|
-\sgn\!\left(p_k-\tfrac12\right)(p_{k+1}-p_k)
\right),
\]
with $\sgn(0):=0$.  Then $L_m^{1/2}\ge 0$ and $m\mapsto L_m^{1/2}$ is
nondecreasing.  Moreover,
\[
\varepsilon_m(\mathsf{P};\mu)=\varepsilon_0(\mathsf{P};\mu)-\EE[L_m^{1/2}],
\qquad
\varepsilon_0(\mathsf{P};\mu)=\min\{\mu(\mathsf{P}),1-\mu(\mathsf{P})\}.
\]
As $m\to\infty$,
\[
\varepsilon_m(\mathsf{P};\mu)\to
\varepsilon_\infty(\mathsf{P};\mu)
:=\inf\{\mu(\mathsf{P}\triangle\mathsf{Q}):\mathsf{Q}\in\mathcal{F}_\infty\}
=\EE[\min\{p_\infty,1-p_\infty\}],
\]
and therefore
\[
\EE[L_\infty^{1/2}]
=\varepsilon_0(\mathsf{P};\mu)-\varepsilon_\infty(\mathsf{P};\mu).
\]
\leanverified{paper_prefix_scan_error_tanaka_representation}
\end{theorem}

\begin{proof}
Proposition~\ref{prop:scan-error-cylinder} gives
\[
\varepsilon_m(\mathsf{P};\mu)
=\EE[\min\{p_m,1-p_m\}]
=\frac12-\EE\left|p_m-\frac12\right|.
\]
Since $p_0=\mu(\mathsf{P})$ is constant,
\[
\varepsilon_0(\mathsf{P};\mu)
=\frac12-\left|p_0-\frac12\right|
=\min\{\mu(\mathsf{P}),1-\mu(\mathsf{P})\}.
\]
Let $f(u):=|u-\tfrac12|$.  For each $k\ge 0$, define
\[
r_k
:=f(p_{k+1})-f(p_k)-\sgn\!\left(p_k-\frac12\right)(p_{k+1}-p_k).
\]
Since $\sgn(x-\tfrac12)\in\partial f(x)$ for every $x\in[0,1]$, the
subgradient inequality for the convex function $f$ yields $r_k\ge 0$
pointwise.  Hence
\[
f(p_{k+1})
=f(p_k)+\sgn\!\left(p_k-\frac12\right)(p_{k+1}-p_k)+r_k,
\]
and summing from $k=0$ to $m-1$ gives
\[
\left|p_m-\frac12\right|
=\left|p_0-\frac12\right|
+\sum_{k=0}^{m-1}\sgn\!\left(p_k-\frac12\right)(p_{k+1}-p_k)
+L_m^{1/2},
\]
where $L_m^{1/2}=\sum_{k=0}^{m-1}r_k\ge 0$.  This is the discrete analogue of
Tanaka's formula \cite{Tanaka1963,RevuzYor1999}; here $L_m^{1/2}$ is exactly
the convexity compensator from that identity, and we do not assert any
occupation-density interpretation for it.  Because
$\sgn(p_k-\tfrac12)$ is bounded and $\mathcal{F}_k$-measurable, while
$p_{k+1}-p_k$ is integrable and has conditional expectation zero given
$\mathcal{F}_k$, we have
\[
\EE\!\left[\sgn\!\left(p_k-\tfrac12\right)(p_{k+1}-p_k)\right]
=\EE\!\left[\sgn\!\left(p_k-\tfrac12\right)\EE[p_{k+1}-p_k\mid\mathcal{F}_k]\right]
=0.
\]
Taking expectations in the summed identity yields
\[
\EE\left|p_m-\frac12\right|
=\left|p_0-\frac12\right|+\EE[L_m^{1/2}],
\]
and therefore
\[
\varepsilon_m(\mathsf{P};\mu)
=\varepsilon_0(\mathsf{P};\mu)-\EE[L_m^{1/2}].
\]
L\'{e}vy's upward theorem \cite{Doob1953,Williams1991,Durrett2019} gives
$p_m\to p_\infty$ a.s.\ and in $L^1$.  Since
$u\mapsto\min\{u,1-u\}$ is $1$-Lipschitz on $[0,1]$,
Proposition~\ref{prop:scan-error-cylinder} yields
\[
\varepsilon_m(\mathsf{P};\mu)=\EE[\min\{p_m,1-p_m\}]
\longrightarrow
\EE[\min\{p_\infty,1-p_\infty\}].
\]
For any $\mathsf{Q}\in\mathcal{F}_\infty$,
\[
\mu(\mathsf{P}\triangle\mathsf{Q})
=\EE\!\left[p_\infty\ind_{\mathsf{Q}^c}+(1-p_\infty)\ind_{\mathsf{Q}}\right]
\ge \EE[\min\{p_\infty,1-p_\infty\}],
\]
with equality for the threshold classifier
$\mathsf{Q}_\infty^\star:=\{p_\infty\ge 1/2\}\in\mathcal{F}_\infty$.  Thus
$\varepsilon_\infty(\mathsf{P};\mu)=\EE[\min\{p_\infty,1-p_\infty\}]$, and the
identity for $\EE[L_\infty^{1/2}]$ follows from monotone convergence applied
to $L_m^{1/2}\uparrow L_\infty^{1/2}$.  No interchange of an infinite sum with
expectation is used at finite $m$; the only terminal passage is the
$L^1$ convergence of $p_m$ together with this final monotone-convergence step.
\end{proof}

\begin{corollary}[Tail-measurable specialization]
\label{cor:scan-tanaka-tail}
If $\mathsf{P}\in\mathcal{F}_\infty$, then
\[
\varepsilon_m(\mathsf{P};\mu)\to 0
\qquad\text{and}\qquad
\EE[L_\infty^{1/2}]=\varepsilon_0(\mathsf{P};\mu).
\]
\end{corollary}

\begin{proof}
If $\mathsf{P}\in\mathcal{F}_\infty$, then
$p_\infty=\mu(\mathsf{P}\mid\mathcal{F}_\infty)=\ind_{\mathsf{P}}$ a.s.  Hence
\[
\varepsilon_\infty(\mathsf{P};\mu)
=\EE[\min\{p_\infty,1-p_\infty\}]=0,
\]
and the conclusion follows from Theorem~\ref{thm:scan-tanaka}.
\end{proof}

\section{Finite-State Boundary Observers and Pressure Laws}
\label{sec:rates}

This section contains both the short transfer preliminaries used throughout
the paper and the main symbolic theorems.  Subsection~\ref{sub:boundary-growth}
records the finite-depth transfer consequences that follow directly from
Proposition~\ref{prop:scan-error-cylinder} once cylinder-mass information is
available.  The genuinely new symbolic content starts in
Subsection~\ref{sub:finite-posterior} with
Theorem~\ref{thm:finite-state-posterior} and culminates in
Theorem~\ref{thm:weighted-boundary-pressure}.

\subsection{Preparatory transfer consequences}
\label{sub:boundary-growth}

\begin{definition}[Boundary growth scale]
\label{def:boundary-cylinder}
For $\mathsf{P}\in\mathcal{B}$, let $\partial_m\mathsf{P}$ and
$N_m(\partial\mathsf{P})$ be the boundary family and boundary count from
Definition~\ref{def:boundary-atoms}.  Let $M_m:=\#\mathcal{C}_m$.  When the limit
$\lambda_{\mathrm{cyl}}:=\lim_{m\to\infty}M_m^{1/m}$ exists and exceeds $1$,
it is the \emph{cylinder growth base}.  For a mixing SFT, $\lambda_{\mathrm{cyl}}$
is the spectral radius of the transition matrix \cite{LindMarcus1995}.  The
\emph{upper boundary dimension} is
\[
\overline{\dim}_B(\partial\mathsf{P})
:=\limsup_{m\to\infty}
\frac{\log \max\{N_m(\partial\mathsf{P}),1\}}{m\log\lambda_{\mathrm{cyl}}}.
\]
This limsup records only an upper exponential growth exponent.  It does not,
by itself, imply the stronger exact asymptotic
$N_m(\partial\mathsf{P})=\lambda_{\mathrm{cyl}}^{\delta m+o(m)}$, which is an
additional hypothesis when invoked later.
\end{definition}

The limsup exponent in Definition~\ref{def:boundary-cylinder} is a property of
the chosen filtration and presentation.  For arbitrary filtrations we make no
representation-invariance claim at this level.  The symbolic finite-state
theorem below isolates the higher-block invariant quantity relevant for
shift-dynamical
applications: the ambiguity growth base of the refined ambiguous product
graph, whose
normalization agrees with
$\overline{\dim}_B(\partial\mathsf{P})$ and is unchanged under higher-block
recoding (Proposition~\ref{prop:higher-block-invariance}).  No broader
invariance under general finite-to-one sliding block conjugacies is claimed
here.  The maximum in Definition~\ref{def:boundary-cylinder} is only a
zero-count convention: if the boundary is eventually empty, then the resulting
exponent is $0$ and the scan error is eventually zero.
In general $\partial_m\mathsf{P}$ is a family of positive atoms of
$\mathcal{F}_m$; in the symbolic specialization those atoms are precisely the
positive cylinders, which is why the later symbolic discussion keeps the
cylinder terminology.

\subsection{Finite-state posterior presentations}
\label{sub:finite-posterior}

The next theorem supplies a finite-state symbolic source of the thickness
hypothesis and of the ambiguity exponent.  The input is a finite observer that
tracks, from a prefix alone, whether the target event is already forced true,
already forced false, or still ambiguous, and that also records the one-step
follower class needed to continue admissibly.  To keep classifier notation
$\mathsf{Q}$ disjoint from observer notation, we write the finite observer
state space as $\mathcal{S}$.

\begin{definition}[Finite-state posterior presentation with respect to a measure]
\label{def:finite-posterior}
Let $\mathcal{A}$ be a finite alphabet, let $X\subseteq\mathcal{A}^{\NN}$ be a
one-sided shift space, let $\mu$ be a Borel probability measure on $X$, and
let $\mathsf{P}\in\mathcal{F}_\infty$.  A
\emph{finite-state posterior presentation for $\mathsf{P}$ with respect to
$\mu$} consists of a
finite set
\[
\mathcal{S}=\mathcal{S}_1\sqcup \mathcal{S}_0\sqcup \mathcal{S}_\partial,
\]
an initial state $q_\emptyset\in \mathcal{S}$, and partial update maps
$T_b:\mathcal{S}\to \mathcal{S}$ for $b\in\mathcal{A}$ such that:
\begin{enumerate}[label=(\arabic*)]
\item for every admissible word $a=a_0\cdots a_{m-1}$ of positive length in
  the language of $X$,
  the iterated state
  \[
  q(a):=T_{a_{m-1}}\circ\cdots\circ T_{a_0}(q_\emptyset)
  \]
  is defined, and $q(\emptyset):=q_\emptyset$;
\item for any admissible words $a,a'$ (possibly empty) with $q(a)=q(a')$ and
  any symbol $b\in\mathcal{A}$,
  \[
  ab\in\mathcal{L}(X)\Longleftrightarrow a'b\in\mathcal{L}(X),
  \]
  and in this case
  \[
  q(ab)=T_b(q(a))=T_b(q(a')).
  \]
  Equivalently, the state $q(a)$ determines the admissible one-symbol
  continuations after $a$, and $T_b$ is defined precisely for those
  continuations;
\item $\mathcal{S}_0$ and $\mathcal{S}_1$ are absorbing under every defined transition;
\item for every admissible word $a$ with $\mu([a])>0$,
  \[
  q(a)\in \mathcal{S}_1 \Rightarrow \mu([a]\setminus\mathsf{P})=0,\qquad
  q(a)\in \mathcal{S}_0 \Rightarrow \mu([a]\cap\mathsf{P})=0,
  \]
  and
  \[
  q(a)\in \mathcal{S}_\partial \Longleftrightarrow [a]\in\partial_{|a|}\mathsf{P}.
  \]
\end{enumerate}
\end{definition}

For $\mathsf{P}\in\mathcal{F}_\infty$, write
\[
L_{1,\mu}(\mathsf{P})
:=\{a\in\mathcal{L}(X):\mu([a]\setminus\mathsf{P})=0\},
\qquad
L_{0,\mu}(\mathsf{P})
:=\{a\in\mathcal{L}(X):\mu([a]\cap\mathsf{P})=0\},
\]
\[
L_{\partial,\mu}(\mathsf{P})
:=\{a\in\mathcal{L}(X):[a]\in\partial_{|a|}\mathsf{P}\}.
\]
These are the $\mu$-decisive prefix languages of $\mathsf{P}$, and they are
genuinely measure-dependent because the classification is modulo $\mu$-null
sets rather than determined by the topological event alone.  Thus a
finite-state posterior presentation is always relative to the measure named in
the theorem that uses it.  In particular,
Theorem~\ref{thm:weighted-boundary-pressure} requires a presentation with
respect to the Gibbs state $\mu_\phi$ itself.  A purely topological
trichotomy by nonempty intersections is not enough to transfer an observer
between full-support measures, because full support gives positive mass to
nonempty open sets, not to arbitrary nonempty Borel subsets.  What suffices is
the continuation-robust criterion below: every ambiguous cylinder must contain
subcylinders on both sides of the event.  Outside that situation, no transfer
of the observer between different measures is claimed here.  When every
admissible cylinder has positive $\mu$-measure, the three languages above
partition $\mathcal{L}(X)$ into prefixes already decided true, already decided
false, and still ambiguous.  This gives a concrete sufficient criterion for
the existence of a finite-state posterior presentation with respect to $\mu$.

\begin{proposition}[Continuation-robust decisive languages for full-support measures]
\label{prop:full-support-invariant-decisive-languages}
Let $\mathcal{A}$ be a finite alphabet, let $X\subseteq\mathcal{A}^{\NN}$ be a
shift space, and let $\mathsf{P}\in\mathcal{F}_\infty$.  Assume that for every
admissible word $a\in\mathcal{L}(X)$ exactly one of the following alternatives
holds:
\begin{enumerate}[label=(\alph*)]
\item $[a]\subseteq \mathsf{P}$;
\item $[a]\subseteq X\setminus \mathsf{P}$;
\item there exist admissible finite words $u,v$ such that
  \[
  [au]\subseteq \mathsf{P},
  \qquad
  [av]\subseteq X\setminus \mathsf{P}.
  \]
\end{enumerate}
Then every full-support Borel probability measure $\nu$ on $X$ induces the
same decisive languages, namely
\[
L_{1,\nu}(\mathsf{P})=\{a\in\mathcal{L}(X):[a]\subseteq \mathsf{P}\},
\]
\[
L_{0,\nu}(\mathsf{P})=\{a\in\mathcal{L}(X):[a]\subseteq X\setminus \mathsf{P}\},
\]
\[
L_{\partial,\nu}(\mathsf{P})
=\{a\in\mathcal{L}(X):\exists\,u,v\in\mathcal{L}(X)\text{ with }
[au]\subseteq \mathsf{P},\ [av]\subseteq X\setminus \mathsf{P}\}.
\]
In particular, any finite-state posterior presentation whose states realize
this trichotomy for one full-support measure on $X$ realizes the same
trichotomy for every full-support measure on $X$.
\end{proposition}

\begin{proof}
Let $\nu$ be a full-support Borel probability measure on $X$.
If $[a]\subseteq \mathsf{P}$, then $\nu([a]\setminus\mathsf{P})=0$, so
$a\in L_{1,\nu}(\mathsf{P})$.  If
$[a]\subseteq X\setminus \mathsf{P}$, then $\nu([a]\cap\mathsf{P})=0$, so
$a\in L_{0,\nu}(\mathsf{P})$.

Now assume that admissible words $u,v$ satisfy
\[
[au]\subseteq \mathsf{P},
\qquad
[av]\subseteq X\setminus \mathsf{P}.
\]
Because $[au]$ and $[av]$ are nonempty cylinders, they are nonempty open sets
in the subspace topology of $X$.  Full support therefore gives
\[
\nu([au])>0,
\qquad
\nu([av])>0.
\]
Since $[au]\subseteq [a]\cap\mathsf{P}$ and
$[av]\subseteq [a]\setminus\mathsf{P}$, it follows that
\[
\nu([a]\cap\mathsf{P})>0,
\qquad
\nu([a]\setminus\mathsf{P})>0,
\]
so $a\in L_{\partial,\nu}(\mathsf{P})$.

By hypothesis, every admissible word falls into exactly one of the three
alternatives~\emph{(a)}--\emph{(c)}.  The previous paragraph shows that each
alternative determines the same $\nu$-decisive class for every full-support
measure $\nu$, which proves the displayed identities and the final claim about
transferring a finite-state observer between full-support measures.
\end{proof}

\begin{theorem}[Constructive observer synthesis from decisive-language automata]
\label{prop:automaton-regular-sufficient}
Let $\mathcal{A}$ be a finite alphabet, let $X\subseteq\mathcal{A}^{\NN}$ be a
one-step shift of finite type with transition matrix
$A=(A_{ij})_{i,j\in\mathcal{A}}$, let $\mu$ be a Borel probability measure on
$X$, and assume that every admissible cylinder in $X$ has positive
$\mu$-measure.  Let $\mathsf{P}\in\mathcal{F}_\infty$.  Assume that the three
decisive languages
$L_{1,\mu}(\mathsf{P})$, $L_{0,\mu}(\mathsf{P})$, and
$L_{\partial,\mu}(\mathsf{P})$ are recognized on admissible words by a single
deterministic finite automaton
\[
\mathcal{M}=(Q,q_{\mathrm{init}},\delta,F_1,F_0,F_\partial),
\]
meaning that $F_1,F_0,F_\partial\subseteq Q$ are pairwise disjoint and
\[
\delta(q_{\mathrm{init}},a)\in F_\sigma
\Longleftrightarrow
a\in L_{\sigma,\mu}(\mathsf{P})
\qquad
(\sigma\in\{1,0,\partial\})
\]
for every admissible word $a\in\mathcal{L}(X)$.  If one starts with separate
DFAs for the three languages, replace them first by their synchronous product
\cite{HopcroftMotwaniUllman2001} and let $F_\sigma$ be the product states whose
$\sigma$-component accepts.

Form the terminal-symbol product observer
\[
\widehat{\mathcal{S}}
:=\{\widehat q_\emptyset\}\cup
\{(i,z)\in\mathcal{A}\times Q:
\exists a\in\mathcal{L}(X),\ |a|\ge 1,\ a_{|a|-1}=i,\
\delta(q_{\mathrm{init}},a)=z\},
\]
and partition it by
\[
\widehat{\mathcal{S}}_\sigma
:=\{(i,z)\in\widehat{\mathcal{S}}:z\in F_\sigma\}
\cup
\begin{cases}
\{\widehat q_\emptyset\},& \sigma=\sigma_\emptyset,\\
\varnothing,& \sigma\neq \sigma_\emptyset,
\end{cases}
\]
where $\sigma_\emptyset\in\{1,0,\partial\}$ is determined by
\[
\sigma_\emptyset=
\begin{cases}
1,& \mu(X\setminus\mathsf{P})=0,\\
0,& \mu(\mathsf{P})=0,\\
\partial,& \mu(\mathsf{P})>0\text{ and }\mu(X\setminus\mathsf{P})>0.
\end{cases}
\]
Define partial updates by
\[
\widehat T_j(\widehat q_\emptyset):=(j,\delta(q_{\mathrm{init}},j))
\qquad\text{when }j\in\mathcal{L}_1(X),
\]
and
\[
\widehat T_j(i,z):=(j,\delta(z,j))
\qquad\text{when }A_{ij}=1.
\]
Then:
\begin{enumerate}[label=(\arabic*)]
\item $\widehat{\mathcal{S}}
  =\widehat{\mathcal{S}}_1\sqcup\widehat{\mathcal{S}}_0\sqcup
  \widehat{\mathcal{S}}_\partial$ with the transition maps
  $(\widehat T_j)_{j\in\mathcal{A}}$ is a finite-state posterior presentation
  for $\mathsf{P}$ in the sense of Definition~\ref{def:finite-posterior}.
\item The presentation is follower-resolving, because the state
  $\widehat q(a)=(a_{|a|-1},\delta(q_{\mathrm{init}},a))$ records the terminal
  symbol $a_{|a|-1}$ and therefore determines exactly the admissible one-step
  followers after $a$ in the one-step SFT.
\item Let $\widehat G$ be the finite directed graph of defined transitions of
  the product observer.  For each ambiguous state
  $r\in\widehat{\mathcal{S}}_\partial$, let $d_\sigma(r)$ be the shortest path
  length in $\widehat G$ from $r$ to $\widehat{\mathcal{S}}_\sigma$, with
  $d_\sigma(r)=+\infty$ if no such path exists.  Then the uniform
  resolving-window hypothesis of
  Theorem~\ref{thm:finite-state-posterior} holds if and only if
  $d_0(r),d_1(r)<\infty$ for every $r\in\widehat{\mathcal{S}}_\partial$.  In
  that case the least possible resolving bound is
  \[
  R_{\min}
  =\max_{r\in\widehat{\mathcal{S}}_\partial}\max\{d_0(r),d_1(r)\},
  \]
  and $R_{\min}$ is computable by reverse breadth-first search from
  $\widehat{\mathcal{S}}_0$ and $\widehat{\mathcal{S}}_1$ in the finite graph
  $\widehat G$.
\end{enumerate}
\end{theorem}

\begin{proof}
For every admissible nonempty word $a=a_0\cdots a_{m-1}$, an induction on $m$
shows that the run in the product observer is
\[
\widehat q(a)
:=\widehat T_{a_{m-1}}\circ\cdots\circ \widehat T_{a_0}(\widehat q_\emptyset)
=\bigl(a_{m-1},\delta(q_{\mathrm{init}},a)\bigr),
\]
while $\widehat q(\emptyset)=\widehat q_\emptyset$ by definition.  Thus the
product observer reads the DFA state together with the terminal symbol of the
current admissible word.

We verify Definition~\ref{def:finite-posterior}.  Property~(1) holds because
every admissible word in $\mathcal{L}(X)$ has a well-defined DFA run, and the
terminal-symbol product was built exactly from those reachable runs.  For
Property~(2), suppose that admissible words $a,a'$ satisfy
$\widehat q(a)=\widehat q(a')=(i,z)$.  Then both words end in the same symbol
$i$.  Because $X$ is one-step, for every $j\in\mathcal{A}$,
\[
aj\in\mathcal{L}(X)\Longleftrightarrow A_{ij}=1
\Longleftrightarrow a'j\in\mathcal{L}(X).
\]
Whenever these equivalent conditions hold, determinism of the DFA gives
\[
\widehat q(aj)
=\bigl(j,\delta(z,j)\bigr)
=\widehat q(a'j).
\]
Thus the product observer is follower-resolving.

For Property~(4), positivity of admissible cylinders implies that
$L_{1,\mu}(\mathsf{P})$, $L_{0,\mu}(\mathsf{P})$, and
$L_{\partial,\mu}(\mathsf{P})$ partition $\mathcal{L}(X)$.  Hence, for every
admissible nonempty word $a$,
\[
\widehat q(a)\in \widehat{\mathcal{S}}_1
\Longleftrightarrow a\in L_{1,\mu}(\mathsf{P}),
\]
\[
\widehat q(a)\in \widehat{\mathcal{S}}_0
\Longleftrightarrow a\in L_{0,\mu}(\mathsf{P}),
\qquad
\widehat q(a)\in \widehat{\mathcal{S}}_\partial
\Longleftrightarrow a\in L_{\partial,\mu}(\mathsf{P}),
\]
which is exactly the state classification required there.  The empty word was
classified separately by the same trichotomy on $X$ itself.

For Property~(3), if $\widehat q(a)\in \widehat{\mathcal{S}}_1$ and $aj$ is
admissible, then $[aj]\subseteq[a]$ and $\mu([a]\setminus\mathsf{P})=0$, hence
$\mu([aj]\setminus\mathsf{P})=0$ and therefore
$aj\in L_{1,\mu}(\mathsf{P})$.  So $\widehat q(aj)\in \widehat{\mathcal{S}}_1$.
The same argument with $\mathsf{P}^{\mathrm c}$ shows that every defined
transition from $\widehat{\mathcal{S}}_0$ stays in
$\widehat{\mathcal{S}}_0$.  Therefore the decided states are absorbing.  This
proves (1) and (2).

For (3), first assume the uniform resolving-window hypothesis holds with some
bound $R$.  Fix $r\in\widehat{\mathcal{S}}_\partial$.  By hypothesis there are
words $w_r^+,w_r^-$ of lengths at most $R$ whose runs from $r$ stay defined at
every intermediate prefix and end in $\widehat{\mathcal{S}}_1$ and
$\widehat{\mathcal{S}}_0$, respectively.  These runs are directed paths in
$\widehat G$, so
$d_1(r)\le |w_r^+|\le R$ and $d_0(r)\le |w_r^-|\le R$.

Conversely, suppose that $d_0(r),d_1(r)<\infty$ for every ambiguous state $r$.
Choose shortest directed paths from $r$ to
$\widehat{\mathcal{S}}_1$ and $\widehat{\mathcal{S}}_0$; reading off their edge
labels gives words $w_r^+,w_r^-$ of lengths
$d_1(r)$ and $d_0(r)$.  Because these words come from actual directed paths in
the product observer, every intermediate prefix remains admissible after every
admissible representative of $r$, and the terminal states lie in the desired
decided classes.  Hence the resolving-window hypothesis holds with
$R=R_{\min}$.  Minimality is immediate from the inequalities
$d_\sigma(r)\le R$ valid for every admissible resolving bound $R$.  Since
$\widehat G$ is finite, the distances $d_0$ and $d_1$ are computed by reverse
breadth-first search from the target sets
$\widehat{\mathcal{S}}_0$ and $\widehat{\mathcal{S}}_1$.
\end{proof}

\begin{definition}[Observer-regularity hierarchy for a fixed measure]
\label{def:observer-regular}
Let $\mathcal{A}$ be a finite alphabet, let $X\subseteq\mathcal{A}^{\NN}$ be a
one-sided shift space, let $\mu$ be a Borel probability measure on $X$, and
let $\mathsf{P}\in\mathcal{F}_\infty$.
\begin{enumerate}[label=(\roman*)]
\item We say that $\mathsf{P}$ is \emph{$\mu$-finite-presented} if it admits a
  finite-state posterior presentation with respect to $\mu$ in the sense of
  Definition~\ref{def:finite-posterior}.
\item Fix a $\mu$-finite presentation
  $(\mathcal{S},q_\emptyset,(T_b)_{b\in\mathcal{A}})$.  We call this
  presentation \emph{$\mu$-resolving} if there exists $R\ge 1$ such that for
  every accessible ambiguous state
  $u\in\mathcal{S}_\partial^{\mathrm{acc}}
  :=\{q(a): |a|\ge 1,\ [a]\in\partial_{|a|}\mathsf{P}\}$ there are admissible
  words $w_u^+,w_u^-$ of lengths at most $R$ satisfying:
  \begin{enumerate}[label=(\alph*)]
  \item for every admissible boundary word $a$ with $q(a)=u$, the
    concatenations $aw_u^\pm$ are admissible in $X$;
  \item every intermediate prefix of each continuation $w_u^\pm$ remains
    admissible after every such boundary word $a$;
  \item $q(aw_u^+)\in\mathcal{S}_1$ and $q(aw_u^-)\in\mathcal{S}_0$ for every
    such boundary word $a$.
  \end{enumerate}
\item We say that $\mathsf{P}$ is \emph{$\mu$-observer-regular} if it is
  $\mu$-finite-presented and admits some $\mu$-resolving finite presentation.
\end{enumerate}
\end{definition}

For a general Borel event $\mathsf{P}\in\mathcal{F}_\infty$, the decisive
languages $L_{1,\mu}(\mathsf{P})$, $L_{0,\mu}(\mathsf{P})$,
$L_{\partial,\mu}(\mathsf{P})$ need not be regular, so no finite observer is
automatic.  But once those languages are given by finite automata, the
one-step SFT graph makes the observer constructive: adjoining the terminal
symbol forces follower-resolvingness, the refined graph
$\widetilde B$ below is a finite product graph, and
Theorem~\ref{prop:automaton-regular-sufficient} computes the least resolving
window by graph distances rather than additional oracle data.  In the weighted
Gibbs setting we keep the later refinements separate: the chosen
$\mu_\phi$-finite presentation defines the refined ambiguous graph and the
boundary pressure, while the names \emph{primitive},
\emph{residue-complete}, and \emph{no-maximal-chain} will be attached below to
the stronger graph-theoretic hypotheses used only for sharper asymptotics.  A
compact dependency table appears immediately before
Theorem~\ref{thm:weighted-boundary-pressure}.

\begin{theorem}[Finite-state posterior theorem from a $\mu$-finite presentation]
\label{thm:finite-state-posterior}
Let $\mathcal{A}$ be a finite alphabet, let $X$ be a mixing one-step SFT on
$\mathcal{A}$ with transition matrix $A=(A_{ij})_{i,j\in\mathcal{A}}$, and let
$\mu$ be its Parry measure.  Assume
\[
h_{\mathrm{top}}(X)=\log\lambda_{\mathrm{cyl}}>0,
\]
equivalently $\lambda_{\mathrm{cyl}}>1$; this excludes only the one-point
zero-entropy mixing SFT, for which the normalization defining
$\delta_\partial(\mathsf{P})$ is not used.  Let
$\mathsf{P}\in\mathcal{F}_\infty$ admit a finite-state posterior presentation
with respect to $\mu$ as in Definition~\ref{def:finite-posterior}.  In
particular, whenever the decisive languages
$L_{0,\mu}(\mathsf{P})$, $L_{1,\mu}(\mathsf{P})$, and
$L_{\partial,\mu}(\mathsf{P})$ are given by DFAs, one may take the explicit
terminal-symbol product observer and the breadth-first-search resolving bound
from Theorem~\ref{prop:automaton-regular-sufficient}.  If
$q_\emptyset\in\mathcal{S}_0\cup\mathcal{S}_1$, then $\mathsf{P}$ is already
decided at depth $0$, so $\partial_m\mathsf{P}=\varnothing$ and
$\varepsilon_m(\mathsf{P};\mu)=0$ for all $m\ge 0$.  We therefore restrict to
the nontrivial case
$q_\emptyset\in \mathcal{S}_\partial$
and let $\mathcal{S}_\partial^{\mathrm{acc}}\subseteq \mathcal{S}_\partial$ be the ambiguous states
attained by at least one boundary word $a$ with $\mu([a])>0$:
\[
\mathcal{S}_\partial^{\mathrm{acc}}
:=\{q(a):|a|\ge 1\text{ and }[a]\in \partial_{|a|}\mathsf{P}\}.
\]
Define the refined ambiguous state space
\[
\mathcal{R}_\partial
:=\{(i,u)\in\mathcal{A}\times \mathcal{S}_\partial^{\mathrm{acc}}:
\exists a\in\mathcal{L}(X)\text{ with }[a]\in\partial_{|a|}\mathsf{P},
\ a_{|a|-1}=i,\ q(a)=u\},
\]
the refined ambiguous transition matrix
$\widetilde{B}
=(\widetilde{B}_{(i,u),(j,v)})_{(i,u),(j,v)\in\mathcal{R}_\partial}$ by
\[
\widetilde{B}_{(i,u),(j,v)}
:=
\begin{cases}
1,& A_{ij}=1\text{ and }T_j(u)=v,\\
0,& \text{otherwise},
\end{cases}
\]
and the initial row vector $r_1\in\mathbb{R}^{\mathcal{R}_\partial}$ by
\[
(r_1)_{(i,u)}
:=\#\{j\in\mathcal{A}:[j]\in\partial_1\mathsf{P}\text{ and }(j,q(j))=(i,u)\}.
\]
If $\mathcal{R}_\partial=\varnothing$, then $\partial_m\mathsf{P}=\varnothing$
for every $m\ge 1$; equivalently, all positive length-$1$ cylinders are already
decided even though $q_\emptyset\in\mathcal{S}_\partial$.  In that degenerate
case the boundary is empty for all $m\ge 1$, and all assertions below are
interpreted with the zero-dimensional matrix $\widetilde{B}$ and the empty row
vector $r_1$, so every displayed matrix expression equals $0$.  Assume from
now on that $\mathcal{R}_\partial\neq\varnothing$ unless stated otherwise.
For each refined ambiguous state $r=(i,u)\in\mathcal{R}_\partial$, let
\[
\mathcal{F}(r):=
\bigl\{w\in\mathcal{L}(X):
\begin{array}{l}
\text{for some, equivalently every, admissible word $a$ represented by $r$,}\\
aw\in\mathcal{L}(X)
\end{array}
\bigr\},
\]
where ``$a$ is represented by $r=(i,u)$'' means that
$[a]\in\partial_{|a|}\mathsf{P}$, $a_{|a|-1}=i$, and $q(a)=u$.  Because
  Definition~\ref{def:finite-posterior}\emph{(2)} makes the presentation
  follower-resolving, the set $\mathcal{F}(r)$ is well-defined.  Then the
  following hold.
\begin{enumerate}[label=(\arabic*)]
\item For every $m\ge 1$,
  \[
  N_m(\partial\mathsf{P})=r_1\widetilde{B}^{\,m-1}\mathbf{1},
  \]
  where $\mathbf{1}$ is the all-ones column vector on
  $\mathcal{R}_\partial$.
\item If $\widetilde{B}$ is nilpotent, then $\partial_m\mathsf{P}=\varnothing$
  for all
  large $m$.  Otherwise
  \[
  \Lambda_\partial(\mathsf{P})
  :=\limsup_{m\to\infty}\max\{N_m(\partial\mathsf{P}),1\}^{1/m}
  =\rho(\widetilde{B}),
  \]
  where $\rho(\widetilde{B})$ is the spectral radius of $\widetilde{B}$.
  In the nilpotent case set $\Lambda_\partial(\mathsf{P})=1$.  Then the
  normalized ambiguity exponent
  \[
  \delta_\partial(\mathsf{P})
  :=\frac{\log\Lambda_\partial(\mathsf{P})}{\log\lambda_{\mathrm{cyl}}}
  \]
  satisfies
  \[
  \delta_\partial(\mathsf{P})
  =\overline{\dim}_B(\partial\mathsf{P}).
  \]
\item Suppose the chosen $\mu$-finite presentation is $\mu$-resolving in the
  sense of Definition~\ref{def:observer-regular}, and let $R\ge 1$ be a
  resolving bound.  Let
\[
  p_{ij}^{(\mu)}:=\mu(x_{n+1}=j\mid x_n=i)\qquad(A_{ij}=1),
  \qquad
  p_*:=\min\{p_{ij}^{(\mu)}:A_{ij}=1\}>0,
  \qquad
  \theta:=p_*^R.
\]
  Then for every $m\ge 1$ and every boundary cylinder
  $C\in\partial_m\mathsf{P}$,
  \[
  \min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\}\ge \theta.
  \]
\item Under the same $\mu$-resolving hypothesis, there exist constants
  $c_-,c_+>0$ such that for every $m\ge 1$,
  \[
  \theta c_-\lambda_{\mathrm{cyl}}^{-m}r_1\widetilde{B}^{\,m-1}\mathbf{1}
  \le \varepsilon_m(\mathsf{P};\mu)
  \le \frac{c_+}{2}\lambda_{\mathrm{cyl}}^{-m}r_1\widetilde{B}^{\,m-1}\mathbf{1}.
  \]
  If the chosen presentation is primitive in the sense of
  Definition~\ref{def:regular-event}, then
  \[
  r_1\widetilde{B}^{\,m-1}\mathbf{1}\asymp \rho(\widetilde{B})^m,
  \]
  and therefore
  \[
  \varepsilon_m(\mathsf{P};\mu)\asymp
  \left(\frac{\rho(\widetilde{B})}{\lambda_{\mathrm{cyl}}}\right)^m
  =\left(\frac{\Lambda_\partial(\mathsf{P})}{\lambda_{\mathrm{cyl}}}\right)^m
  =\lambda_{\mathrm{cyl}}^{-(1-\delta_\partial(\mathsf{P}))m}.
  \]
\end{enumerate}
\end{theorem}

The counting statement in \emph{(1)} is the next lemma: boundary words are in
bijective correspondence with directed paths in the refined ambiguous graph, so
the matrix formula is literal path counting.

\begin{lemma}[Boundary words versus refined ambiguous paths]
\label{lem:boundary-word-refined-path}
Assume the hypotheses of Theorem~\ref{thm:finite-state-posterior}.  For each
$m\ge 1$, the assignment
\[
a=a_0\cdots a_{m-1}
\longmapsto
r(a):=r_0\cdots r_{m-1},
\qquad
r_k:=(a_k,q(a_0\cdots a_k)),
\]
is a bijection from the set of length-$m$ boundary words
\[
\{a\in\mathcal{L}_m(X):[a]\in\partial_m\mathsf{P}\}
\]
onto the directed paths
$r_0\to\cdots\to r_{m-1}$ of length $m-1$ in the refined graph
$G(\widetilde B)$ whose initial vertex lies in $\operatorname{supp}(r_1)$.
Consequently, for every $m\ge 1$,
\[
N_m(\partial\mathsf{P})=r_1\widetilde B^{\,m-1}\mathbf{1}.
\]
\end{lemma}

\begin{proof}
Because $\mu$ is the Parry measure on a mixing one-step SFT, every admissible
cylinder has positive measure.  Thus, by
Definition~\ref{def:finite-posterior}\emph{(4)}, a word is boundary if and only
if its terminal observer state lies in $\mathcal{S}_\partial$.

Let $a=a_0\cdots a_{m-1}$ be a boundary word.  Since
$q(a)\in\mathcal{S}_\partial$ and $\mathcal{S}_0,\mathcal{S}_1$ are absorbing,
no prefix state $q(a_0\cdots a_k)$ can lie in $\mathcal{S}_0\cup\mathcal{S}_1$;
otherwise the remaining suffix would keep the run inside the same absorbing
class and force $q(a)\in\mathcal{S}_0\cup\mathcal{S}_1$, contradiction.  Hence
every prefix is also ambiguous, so every $r_k$ lies in $\mathcal{R}_\partial$.
For each $k<m-1$, the one-step SFT condition gives $A_{a_ka_{k+1}}=1$, and
Definition~\ref{def:finite-posterior}\emph{(2)} gives
\[
T_{a_{k+1}}(q(a_0\cdots a_k))=q(a_0\cdots a_{k+1}),
\]
hence $\widetilde B_{r_k,r_{k+1}}=1$.  Thus $r(a)$ is a directed path in
$G(\widetilde B)$.  Its initial vertex lies in $\operatorname{supp}(r_1)$
because $[a_0]\in\partial_1\mathsf{P}$ and $r_1$ records precisely the
length-$1$ boundary states.

Conversely, let $r_0\to\cdots\to r_{m-1}$ be a directed path in
$G(\widetilde B)$ with initial vertex in $\operatorname{supp}(r_1)$, and write
$r_k=(i_k,u_k)$.  Since $\widetilde B_{r_k,r_{k+1}}=1$, one has
$A_{i_ki_{k+1}}=1$ for every $k<m-1$, so
$a:=i_0\cdots i_{m-1}\in\mathcal{L}_m(X)$.  Because
$r_0\in\operatorname{supp}(r_1)$, the one-letter cylinder $[i_0]$ is boundary
and $q(i_0)=u_0$.  Inductively,
Definition~\ref{def:finite-posterior}\emph{(2)} gives
\[
q(i_0\cdots i_k)=u_k
\qquad(0\le k\le m-1).
\]
In particular, $q(a)=u_{m-1}\in\mathcal{S}_\partial$, so
Definition~\ref{def:finite-posterior}\emph{(4)} implies
$[a]\in\partial_m\mathsf{P}$.  Therefore the path is exactly $r(a)$.

The two constructions are inverse, so the correspondence is bijective.  The
final matrix identity follows because $r_1\widetilde B^{\,m-1}\mathbf{1}$ counts
directed paths of length $m-1$ in $G(\widetilde B)$ whose initial vertex lies
in $\operatorname{supp}(r_1)$.
\end{proof}

\begin{lemma}[Reachable maximal components, including periodic ones, control limsup growth]
\label{lem:reachable-maximal-components}
Let $B$ be a nonnegative $d\times d$ matrix, let $r\in[0,\infty)^d$ be a
nonzero row vector, and let $\mathbf{1}$ be the all-ones column vector.  Write
\[
\operatorname{supp}(r):=\{i\in\{1,\dots,d\}:r_i>0\},
\]
let $G(B)$ be the directed graph with an edge $i\to j$ whenever $B_{ij}>0$,
and for each strongly connected component $C$ of $G(B)$ let $B_C$ denote the
principal submatrix of $B$ indexed by $C$.  Assume that every strongly
connected component $C$ with $\rho(B_C)=\rho(B)$ is reachable in $G(B)$ from
some vertex of $\operatorname{supp}(r)$.  Then
\[
\limsup_{m\to\infty}(rB^m\mathbf{1})^{1/m}=\rho(B).
\]
No primitivity or aperiodicity is assumed for the maximal components.
\end{lemma}

\begin{proof}
Throughout this proof, $(B^m)_{ij}$ counts directed walks of exactly $m$
edges from $i$ to $j$ in $G(B)$.

The upper bound is immediate from any matrix norm:
\[
rB^m\mathbf{1}\le \|r\|\,\|B^m\|\,\|\mathbf{1}\|,
\]
so Gelfand's formula gives
$\limsup_m(rB^m\mathbf{1})^{1/m}\le \rho(B)$.

If $\rho(B)=0$, then $B$ is nilpotent, hence $rB^m\mathbf{1}=0$ for all large
$m$, and the conclusion follows.  Assume now that $\rho(B)>0$.  Choose a
strongly connected component $C$ with $\rho(B_C)=\rho(B)$.  By hypothesis,
there exist $i\in\operatorname{supp}(r)$, a state $a\in C$, and $\ell\ge 0$
such that $(B^\ell)_{ia}>0$, equivalently there is a directed walk of exactly
$\ell$ edges from $i$ to $a$.  Since $r_i>0$, we have
\[
c_0:=r_i(B^\ell)_{ia}>0.
\]
Let $p$ be the period of $C$, and let
$C=C_0\sqcup\cdots\sqcup C_{p-1}$ be the corresponding cyclic decomposition,
indexed so that $a\in C_0$.  Then $(B_C^p)|_{C_0}$ is primitive and has
spectral radius $\rho(B_C)^p=\rho(B)^p$.  By Perron--Frobenius for primitive
matrices, for every $b\in C_0$ there are constants $c_b>0$ and $n_b$ such
that
\[
\bigl((B_C^p)|_{C_0}\bigr)^n_{ab}\ge c_b\,\rho(B)^{pn}
\qquad(n\ge n_b).
\]
Choose $b=a$.  Then for some $c_1>0$ and all large $n$,
\[
(B_C^{pn})_{aa}\ge c_1\,\rho(B)^{pn}.
\]
Since paths counted by $(B_C^{pn})_{aa}$ remain inside $C$, they are also
counted by $(B^{pn})_{aa}$.  Therefore, for all sufficiently large $n$,
\[
rB^{\ell+pn}\mathbf{1}
\ge r_i(B^{\ell+pn})_{ia}
\ge r_i(B^\ell)_{ia}(B^{pn})_{aa}
\ge c_0(B_C^{pn})_{aa}
\ge c_0c_1\,\rho(B)^{pn}.
\]
Taking $(\ell+pn)$th roots and letting $n\to\infty$ gives
\[
\limsup_{m\to\infty}(rB^m\mathbf{1})^{1/m}
\ge
\limsup_{n\to\infty}(rB^{\ell+pn}\mathbf{1})^{1/(\ell+pn)}
\ge \rho(B),
\]
because $(c_0c_1)^{1/(\ell+pn)}\to 1$ and
$\rho(B)^{pn/(\ell+pn)}\to \rho(B)$.  Combined with the upper bound, this
proves the lemma.
\end{proof}

\begin{proof}
If $\mathcal{R}_\partial=\varnothing$, then by definition there is no boundary
word of positive length.  Indeed, if $[a]\in\partial_m\mathsf{P}$ for some
$m\ge 1$, then $(a_{m-1},q(a))\in\mathcal{R}_\partial$, contrary to
$\mathcal{R}_\partial=\varnothing$.  Hence $\partial_m\mathsf{P}=\varnothing$
for every $m\ge 1$, so $N_m(\partial\mathsf{P})=0$ and
$\varepsilon_m(\mathsf{P};\mu)=0$ for all $m\ge 1$.  Every displayed matrix
expression in the theorem is therefore interpreted on the zero-dimensional
matrix $\widetilde{B}$ and equals $0$, so \emph{(1)}--\emph{(4)} are immediate.
Assume for the remainder of the proof that
$\mathcal{R}_\partial\neq\varnothing$.

Lemma~\ref{lem:boundary-word-refined-path} gives the exact boundary-word/path
bijection and therefore the counting identity
\[
N_m(\partial\mathsf{P})=r_1\widetilde{B}^{\,m-1}\mathbf{1}
\qquad(m\ge 1),
\]
which proves~\emph{(1)}.  The point of the refinement is not merely to encode a
recurrence on $\mathcal{S}_\partial^{\mathrm{acc}}$.  Rather, the refined graph
packages the one-step adjacency test and the deterministic observer update into
a single $0$--$1$ matrix on pairs $(i,u)$, so boundary words are counted by
finite directed paths with no hidden multiplicities.

If $\widetilde{B}$ is nilpotent then $\widetilde{B}^{\,m-1}=0$ for all large $m$,
so (1) gives eventual emptiness of the boundary.  Assume now that
$\widetilde{B}$ is not nilpotent.  The upper
bound
\[
N_m(\partial\mathsf{P})\le \|r_1\|\,\|\widetilde{B}^{\,m-1}\|\,\|\mathbf{1}\|
\]
implies
$\limsup_m N_m(\partial\mathsf{P})^{1/m}\le \rho(\widetilde{B})$ by Gelfand's formula.
For the matching lower bound, let $(i,u)\in\mathcal{R}_\partial$ be arbitrary.
Choose a boundary word
$a=a_0\cdots a_{\ell-1}\in\mathcal{L}(X)$ with $\ell\ge 1$, $a_{\ell-1}=i$,
and $q(a)=u$.  For $0\le k\le \ell-1$ set
\[
r_k:=(a_k,q(a_0\cdots a_k)).
\]
By Lemma~\ref{lem:boundary-word-refined-path},
$r_0\to r_1\to\cdots\to r_{\ell-1}$ is a directed walk of exactly $\ell-1$
edges in the refined graph whose initial vertex $r_0=(a_0,q(a_0))$ lies in
$\operatorname{supp}(r_1)$ and whose terminal vertex is $r_{\ell-1}=(i,u)$.
Because $(i,u)$ was arbitrary, every vertex of
the refined ambiguous product graph, and hence every strongly connected
component with spectral radius $\rho(\widetilde{B})$, is reachable from the
support of $r_1$.  This is precisely the situation covered by
Lemma~\ref{lem:reachable-maximal-components}: even if a maximal component is
periodic, it still contributes to the limsup growth once it is reachable from
the initial support.
Lemma~\ref{lem:reachable-maximal-components} applied to
$B=\widetilde{B}$ and $r=r_1$ yields
\[
\limsup_{n\to\infty}(r_1\widetilde{B}^{\,n}\mathbf{1})^{1/n}
=\rho(\widetilde{B}).
\]
Since $N_m(\partial\mathsf{P})=r_1\widetilde{B}^{\,m-1}\mathbf{1}$ for $m\ge 1$,
we have
\[
N_m(\partial\mathsf{P})^{1/m}
=\bigl(r_1\widetilde{B}^{\,m-1}\mathbf{1}\bigr)^{1/m}
=\Bigl(\bigl(r_1\widetilde{B}^{\,m-1}\mathbf{1}\bigr)^{1/(m-1)}\Bigr)^{(m-1)/m},
\]
and $(m-1)/m\to 1$, so replacing $n$ by $m-1$ does not change the
$\frac1m$-limsup.  Using~\emph{(1)} therefore gives
\[
\limsup_{m\to\infty}N_m(\partial\mathsf{P})^{1/m}=\rho(\widetilde{B}),
\]
and the dimension formula in (2) follows from
Definition~\ref{def:boundary-cylinder}.

We next verify that the $\mu$-resolving refinement yields the thickness input
used in~\emph{(4)}.  Assume therefore that the chosen $\mu$-finite
presentation is $\mu$-resolving with resolving bound $R$, and for each
accessible ambiguous state $u\in\mathcal{S}_\partial^{\mathrm{acc}}$ choose
resolving words $w_u^+,w_u^-$ as in
Definition~\ref{def:observer-regular}.

\smallskip
\noindent\emph{Claim 2 (uniform resolving window implies boundary thickness).}
Let $C=[a]\in\partial_m\mathsf{P}$ with $m\ge 1$, and write
$r:=(a_{m-1},q(a))\in\mathcal{R}_\partial$.  Then
\[
\mu(\mathsf{P}\mid C)\ge \theta,
\qquad
1-\mu(\mathsf{P}\mid C)\ge \theta,
\]
where $\theta=p_*^R$.

\smallskip
\noindent\emph{Proof of Claim 2.}
Write $r=(i,u)$ and $w_r^\pm:=w_u^\pm$.  By the $\mu$-resolving hypothesis these words have
lengths at most $R$, belong to $\mathcal{F}(r)$, and satisfy
$T_{w_r^+}(u)\in \mathcal{S}_1$ and $T_{w_r^-}(u)\in \mathcal{S}_0$.  Because
$w_r^\pm\in\mathcal{F}(r)$ and $a$ is represented by $r$, the words
$aw_r^\pm$ are admissible in $X$; the resolving condition also ensures that
every intermediate prefix along these continuations stays admissible.  Hence
$[aw_r^\pm]$ are positive subcylinders of $[a]$.  Moreover
$q(aw_r^+)=T_{w_r^+}(u)\in \mathcal{S}_1$ and
$q(aw_r^-)=T_{w_r^-}(u)\in \mathcal{S}_0$, so
$[aw_r^+]\subseteq \mathsf{P}$ modulo $\mu$-null sets and
$[aw_r^-]\subseteq [a]\setminus \mathsf{P}$ modulo $\mu$-null sets by
Definition~\ref{def:finite-posterior}.  Since the Parry measure is the Markov
measure with transition matrix derived from $A$, every admissible one-step
transition has conditional probability at least $p_*$.  Therefore any
admissible continuation word $w$ of length at most $R$ satisfies
\[
\mu([aw]\mid[a])\ge p_*^{|w|}\ge p_*^R=\theta.
\]
Applying this to $w_r^+$ and $w_r^-$ gives
\[
\mu(\mathsf{P}\mid C)\ge \mu([aw_r^+]\mid[a])\ge \theta
\]
and similarly
\[
1-\mu(\mathsf{P}\mid C)\ge \mu([aw_r^-]\mid[a])\ge \theta.
\]
This proves Claim~2 and hence~\emph{(3)} for every boundary cylinder of depth
at least one.

Finally, Parry's formula \cite{Parry1964} gives constants $c_-,c_+>0$ such
that
$c_-\lambda_{\mathrm{cyl}}^{-m}\le \mu(C)\le c_+\lambda_{\mathrm{cyl}}^{-m}$
for every positive $m$-cylinder $C$.  Combining these cylinder bounds with
Proposition~\ref{prop:scan-error-cylinder},
Lemma~\ref{lem:boundary-word-refined-path}, and Claim~2, we obtain
\[
\varepsilon_m(\mathsf{P};\mu)
=\sum_{C\in\partial_m\mathsf{P}}\mu(C)
\min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\}.
\]
For the lower bound, each boundary cylinder contributes at least
$\theta c_-\lambda_{\mathrm{cyl}}^{-m}$, so
\[
\varepsilon_m(\mathsf{P};\mu)
\ge \theta c_-\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})
= \theta c_-\lambda_{\mathrm{cyl}}^{-m}r_1\widetilde{B}^{\,m-1}\mathbf{1}.
\]
For the upper bound, $\min\{p,1-p\}\le \frac12$ and each positive
depth-$m$ cylinder has mass at most $c_+\lambda_{\mathrm{cyl}}^{-m}$, hence
\[
\varepsilon_m(\mathsf{P};\mu)
\le \frac{c_+}{2}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})
= \frac{c_+}{2}\lambda_{\mathrm{cyl}}^{-m}r_1\widetilde{B}^{\,m-1}\mathbf{1}.
\]
This proves~\emph{(4)}.  If $\widetilde{B}$ is primitive, then
Perron--Frobenius theory gives
$r_1\widetilde{B}^{\,m-1}\mathbf{1}\asymp \rho(\widetilde{B})^{m-1}$, hence
also $r_1\widetilde{B}^{\,m-1}\mathbf{1}\asymp \rho(\widetilde{B})^m$ after
absorbing the constant factor $\rho(\widetilde{B})^{-1}$.  Substituting this
into the two-sided estimate and using~\emph{(2)} yields the displayed
asymptotic law with base $\rho(\widetilde{B})/\lambda_{\mathrm{cyl}}$.
\end{proof}

\begin{proposition}[Higher-block invariance of the ambiguity exponent]
\label{prop:higher-block-invariance}
Assume the hypotheses of Theorem~\ref{thm:finite-state-posterior}.  Then
$\Lambda_\partial(\mathsf{P})$ and
$\delta_\partial(\mathsf{P})$ are independent of the chosen finite-state
posterior presentation.  Let $k\ge 1$, let
$\beta_k:X\to X^{[k]}$ be the standard $k$-block conjugacy to the higher-block
presentation \cite{LindMarcus1995}, let
$\mu^{[k]}:=\mu\circ\beta_k^{-1}$, and let
$\mathsf{P}^{[k]}:=\beta_k(\mathsf{P})$.  Then for every $m\ge 1$,
\[
N_m^{[k]}(\partial\mathsf{P}^{[k]})
=N_{m+k-1}(\partial\mathsf{P}),
\qquad
\varepsilon_m(\mathsf{P}^{[k]};\mu^{[k]})
=\varepsilon_{m+k-1}(\mathsf{P};\mu),
\]
and the cylinder growth bases satisfy
$\lambda_{\mathrm{cyl}}^{[k]}=\lambda_{\mathrm{cyl}}$.  Consequently
\[
\Lambda_\partial(\mathsf{P}^{[k]})=\Lambda_\partial(\mathsf{P}),
\qquad
\delta_\partial(\mathsf{P}^{[k]})=\delta_\partial(\mathsf{P}).
\]
\end{proposition}

\begin{proof}
Theorem~\ref{thm:finite-state-posterior}(2) expresses
$\Lambda_\partial(\mathsf{P})$ as the limsup growth base of the boundary
counts, so any two finite-state posterior presentations yield the same
$\Lambda_\partial(\mathsf{P})$ and hence the same
$\delta_\partial(\mathsf{P})$.

For the higher-block claim, a length-$m$ cylinder in $X^{[k]}$ is determined by
admissible overlapping $k$-blocks
$u_0,\dots,u_{m-1}$ and therefore corresponds bijectively to a unique original
word $a_0\cdots a_{m+k-2}$ of length $m+k-1$.  If
$D=[u_0\cdots u_{m-1}]_{[k]}\subset X^{[k]}$ denotes the $k$-block cylinder and
$C=[a_0\cdots a_{m+k-2}]\subset X$ the associated original cylinder, then
$\beta_k^{-1}(D)=C$.  Therefore
\[
\mu^{[k]}(\mathsf{P}^{[k]}\cap D)=\mu(\mathsf{P}\cap C),
\qquad
\mu^{[k]}(D\setminus \mathsf{P}^{[k]})=\mu(C\setminus\mathsf{P}),
\]
so $D$ is boundary for $\mathsf{P}^{[k]}$ if and only if $C$ is boundary for
$\mathsf{P}$.  This gives
$N_m^{[k]}(\partial\mathsf{P}^{[k]})=N_{m+k-1}(\partial\mathsf{P})$.  The same
cylinder correspondence shows
$M_m^{[k]}=M_{m+k-1}$, hence
$\lambda_{\mathrm{cyl}}^{[k]}=\lambda_{\mathrm{cyl}}$.

Likewise the pullback of the depth-$m$ prefix sigma algebra on $X^{[k]}$ is the
depth-$(m+k-1)$ prefix sigma algebra on $X$:
\[
\beta_k^{-1}\bigl(\mathcal{F}_m^{[k]}\bigr)=\mathcal{F}_{m+k-1}.
\]
Taking the infimum defining the scan error over corresponding measurable sets
gives
$\varepsilon_m(\mathsf{P}^{[k]};\mu^{[k]})
=\varepsilon_{m+k-1}(\mathsf{P};\mu)$.  The displayed invariance of
$\Lambda_\partial$ and $\delta_\partial$ now follows from the count identity
and the equality of cylinder growth bases.
\end{proof}

\begin{remark}[Sofic and hidden-Markov origin]
\label{rem:sofic-hidden-markov}
The state $q(a)$ in Definition~\ref{def:finite-posterior} should be read as an
observer state only after the event $\mathsf{P}$ has been shown to be
\emph{$\mu$-observer-regular} in the sense of
Definition~\ref{def:observer-regular}: the prefix languages
$L_{1,\mu}(\mathsf{P}),L_{0,\mu}(\mathsf{P}),L_{\partial,\mu}(\mathsf{P})$ must
be recognized by finitely many follower-resolving states, and the ambiguous
states must admit a bounded decision window.  In a right-resolving sofic presentation
this is often achieved by adjoining the current presentation vertex, or
equivalently the current follower class, to a finite automaton recognizing the
three decisive prefix languages; for a finite hidden Markov model one can do the
same with a finite observer on hidden-state posterior classes when those
classes already separate decided and boundary prefixes.  After recoding to an
edge shift, such observer-regular events fall under
Theorem~\ref{thm:finite-state-posterior}.  What is \emph{not} claimed is that
every Borel event $\mathsf{P}\in\mathcal{F}_\infty$ on a sofic shift or finite
hidden Markov source admits such a finite observer.  For arbitrary tail events
the decisive prefix languages need not be regular, and existence of a finite
posterior presentation or bounded resolving window remains an additional
assumption here; compare
\cite{Blackwell1957,Heller1965,Fischer1975,Krieger1984,Kitchens1998,LindMarcus1995}.
\end{remark}

\subsection{Transfer consequences of the exact boundary decomposition}

Theorem~\ref{thm:finite-state-posterior} and
Proposition~\ref{prop:higher-block-invariance} identify a higher-block
invariant symbolic exponent $\delta_\partial(\mathsf{P})$ in the finite-state
SFT setting.  Proposition~\ref{prop:scan-error-cylinder} already gives the
exact Bayes-risk identity.  The next results are recorded only as transfer
consequences of that identity under additional cylinder-mass hypotheses: the
first gives an upper bound, the second adds a thickness lower bound, and the
third allows polynomial cylinder distortions.  Their role is to convert
boundary information into scan-error estimates once external measure bounds are
available, not to introduce further symbolic structure.  The new symbolic
mechanism is therefore the finite-state theorem above, which furnishes these
inputs in the Parry-measure setting.  The
hypotheses here are stated in terms of the raw boundary counts
$N_m(\partial\mathsf{P})$, with no representation-invariance claim beyond the
symbolic higher-block setting above.  In particular, the finite-depth
inequalities below require only $\mathsf{P}\in\mathcal{B}$.  The stronger
assumption $\mathsf{P}\in\mathcal{F}_\infty$ enters only when one combines
these bounds with Corollary~\ref{cor:scan-tanaka-tail} to interpret decay of
$\varepsilon_m(\mathsf{P};\mu)$ as a tail-error statement for the limiting
event.

\begin{proposition}[Transfer proposition from the exact boundary decomposition: boundary-dimension upper bound]
\label{thm:boundary-dimension}
Let $\mathsf{P}\in\mathcal{B}$.  Assume that the cylinder growth base
$\lambda_{\mathrm{cyl}}$ exists and $\lambda_{\mathrm{cyl}}>1$.  Suppose
$\mu(C)\le K\lambda_{\mathrm{cyl}}^{-m}$ for all $m\ge 1$ and every positive
depth-$m$ atom $C\in\mathcal{C}_m$.  Then
\[
\varepsilon_m(\mathsf{P};\mu)\le K\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P}).
\]
If $\overline{\dim}_B(\partial\mathsf{P})=\delta<1$, then
$\varepsilon_m(\mathsf{P};\mu)\le\lambda_{\mathrm{cyl}}^{-(1-\delta)m+o(m)}$.
\leanverified{paper_prefix_scan_error_boundary_dimension_upper}
\end{proposition}

\begin{proof}
This is an immediate consequence of Proposition~\ref{prop:scan-error-cylinder}.
Non-boundary cylinders contribute zero to the decomposition
(Proposition~\ref{prop:scan-error-cylinder}), since off $\partial_m\mathsf{P}$
the local term
\[
\min\{\mu(\mathsf{P}\cap C),\mu(C\setminus\mathsf{P})\}
\]
vanishes.  On boundary cylinders,
$\min\{\mu(\mathsf{P}\cap C),\mu(C\setminus\mathsf{P})\}\le\mu(C)\le
K\lambda_{\mathrm{cyl}}^{-m}$.  Summing gives
$\varepsilon_m\le K\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})$.
For $\delta<1$: fix $\eta>0$; for large $m$,
$N_m(\partial\mathsf{P})\le\lambda_{\mathrm{cyl}}^{(\delta+\eta)m}$, giving
$\varepsilon_m\le K\lambda_{\mathrm{cyl}}^{-(1-\delta-\eta)m}$.  Since $\eta$ and
the constant $K$ contribute $o(m)$ after taking logs, the bound follows.
\end{proof}

\begin{corollary}[Transfer corollary from boundary thickness]
\label{thm:volume-boundary-scaling}
Let $\mathsf{P}\in\mathcal{B}$.  Assume that the cylinder growth base
$\lambda_{\mathrm{cyl}}$ exists and $\lambda_{\mathrm{cyl}}>1$.  Suppose
$c_-\lambda_{\mathrm{cyl}}^{-m}\le\mu(C)\le c_+\lambda_{\mathrm{cyl}}^{-m}$ for
all $m\ge 1$ and every positive depth-$m$ atom $C\in\mathcal{C}_m$, and suppose
there exist
$\theta\in(0,\frac12]$ and $m_0$ such that for all $m\ge m_0$ and all
$C\in\partial_m\mathsf{P}$,
\[
\min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\}\ge\theta.
\]
Then for all $m\ge m_0$,
\[
\theta c_-\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})
\le\varepsilon_m(\mathsf{P};\mu)
\le\frac{c_+}{2}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P}).
\]
If, in addition, the boundary count satisfies the exact exponential asymptotic
$N_m(\partial\mathsf{P})=\lambda_{\mathrm{cyl}}^{\delta m+o(m)}$ (a stronger
assumption than merely $\overline{\dim}_B(\partial\mathsf{P})=\delta$), then
\[
\lim_{m\to\infty}\left(-\frac{1}{m}\log_{\lambda_{\mathrm{cyl}}}\varepsilon_m\right)=1-\delta,
\]
so $\varepsilon_m$ has exponential decay exponent $1-\delta$, where $\delta$
is the exponential growth exponent of $N_m(\partial\mathsf{P})$.
\end{corollary}

\begin{proof}
This is an immediate consequence of Proposition~\ref{prop:scan-error-cylinder},
once the cylinder-mass and thickness hypotheses are imposed.
By Proposition~\ref{prop:scan-error-cylinder},
$\varepsilon_m=\sum_{C\in\partial_m\mathsf{P}}\mu(C)\min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\}$.
Lower: thickness and the lower cylinder bound give
$\varepsilon_m\ge\theta c_-\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})$.
Upper: $\min\{p,1-p\}\le\frac12$ and the upper cylinder bound give
$\varepsilon_m\le\frac{c_+}{2}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})$.
Inserting $N_m=\lambda_{\mathrm{cyl}}^{\delta m+o(m)}$ and taking
$-\log_{\lambda_{\mathrm{cyl}}}/m$ yields $1-\delta$ in the limit.
\end{proof}

\begin{remark}[Tail specialization of the transfer principles]
\label{rem:tail-specialization-transfer}
Proposition~\ref{thm:boundary-dimension}
and Corollary~\ref{thm:volume-boundary-scaling}, together with the later
Theorem~\ref{thm:nonuniform-cylinder}, are finite-depth statements for
arbitrary Borel events $\mathsf{P}\in\mathcal{B}$.  If, in addition,
$\mathsf{P}\in\mathcal{F}_\infty$, then Corollary~\ref{cor:scan-tanaka-tail}
identifies $\varepsilon_m(\mathsf{P};\mu)$ as a tail error converging to zero.
Thus any decaying upper bound furnished by these transfer principles becomes a
quantitative tail-error law for $\mathsf{P}$.
\end{remark}

\begin{remark}[Parry versus general Gibbs measures]
\label{rem:parry-vs-gibbs}
The hypotheses of Proposition~\ref{thm:boundary-dimension}
and Corollary~\ref{thm:volume-boundary-scaling} are genuinely same-length
assumptions.
For a mixing SFT with Parry measure $\mu_{\max}$, every admissible word
$a=a_0\cdots a_{m-1}$ satisfies
\[
\mu_{\max}([a])
=\frac{\ell_{a_0}r_{a_{m-1}}}{\lambda_{\mathrm{cyl}}^{m-1}\langle \ell,r\rangle},
\]
where $\ell,r$ are positive left/right Perron--Frobenius eigenvectors of the
transition matrix.  Since only the endpoints appear, one obtains constants
$c_\pm>0$ with $c_-\lambda_{\mathrm{cyl}}^{-m}\le\mu_{\max}([a])\le
c_+\lambda_{\mathrm{cyl}}^{-m}$ uniformly over all admissible words of length
$m$ \cite{Parry1964}.  By contrast, a H\"older Gibbs state for a potential
$\phi$ satisfies
\[
\mu_\phi([a])\asymp \exp\!\bigl(S_m\phi(x)-mP(\phi)\bigr)
\qquad(x\in[a]),
\]
so same-length cylinders generally differ exponentially with the word
\cite{Ruelle1978,ParryPollicott1990,Baladi2000,Bowen1975,WaltersErgodicTheory}.  In that regime a count-only law in
terms of $N_m(\partial\mathsf{P})$ is no longer natural; the corresponding
quantity is a weighted boundary partition sum.  The next subsection turns that
observation into a pressure theorem for finite-state events with possibly
reducible or periodic boundary graphs.  In
particular, the invariant Gibbs replacement for the same-length count laws in
Proposition~\ref{thm:boundary-dimension}
and Corollary~\ref{thm:volume-boundary-scaling} is the boundary-pressure law
$P_X(\phi)-P_\partial(\phi)$ below, not the auxiliary log-regular statement of
Appendix~\ref{sub:nonuniform}.
\end{remark}

\begin{remark}[Assumption ledger for the count and weighted boundary regimes]
\label{rem:boundary-assumption-ledger}
Before passing from the count-based Parry setting to the Gibbs-weighted one,
it is useful to separate the standing hypotheses from the refinements that
improve specific conclusions.
\begin{enumerate}[label=(\alph*)]
\item Proposition~\ref{thm:boundary-dimension} and
  Corollary~\ref{thm:volume-boundary-scaling} are transfer statements with no
  observer graph in the hypotheses.  Proposition~\ref{thm:boundary-dimension}
  uses only an upper bound on cylinder masses.  Corollary~\ref{thm:volume-boundary-scaling}
  adds thickness as a separate hypothesis; $\mu$-observer-regularity,
  residue-complete, primitive, and no-maximal-chain refinements do not enter
  there.
\item In Theorem~\ref{thm:finite-state-posterior}, the finite-state posterior
  presentation alone yields the exact boundary-word recursion in~\emph{(1)} and
  the limsup-growth/dimension statement in~\emph{(2)}.  The $\mu$-resolving
  refinement from Definition~\ref{def:observer-regular}, equivalently the
  $\mu$-observer-regular input for the event, is used only for the thickness
  conclusion~\emph{(3)} and hence for the two-sided scan-error comparison
  in~\emph{(4)}.  The extra primitive refinement is used only in the final
  sentence of \emph{(4)} to replace general limsup growth by a pure
  exponential asymptotic.
\item In the Gibbs-weighted regime, Definition~\ref{def:regular-event} needs
  only a finite-state posterior presentation with respect to $\mu_\phi$; this
  defines the refined graph, the initial state set $\mathcal{I}$, the boundary
  pressure $P_\partial(\phi)$, and the chain-length parameter $q_\partial$.
  In Theorem~\ref{thm:weighted-boundary-pressure}, the finite-path
  decomposition and componentwise pressure analysis yield the upper bound and
  limsup formula in~\emph{(1)}, the residue-class lower bound in~\emph{(2)},
  and the exponential-scale statement in~\emph{(3)}.  The
  $\mu_\phi$-resolving refinement, via the bounded resolving window, is used
  only for the thickness statement \emph{(4)} and therefore for the scan-error
  comparison in \emph{(5)} and
  Corollary~\ref{cor:weighted-boundary-transfer}.
\item The residue-complete refinement is needed exactly for the
  all-large-$m$ lower bounds in
  Theorem~\ref{thm:weighted-boundary-pressure}\emph{(3)},
  Corollary~\ref{cor:weighted-boundary-transfer}\emph{(4)}, and
  Corollary~\ref{cor:weighted-ambiguous-matrix}.  It does not affect the upper
  bounds or the limsup identities.
\item The no-maximal-chain refinement, equivalently $q_\partial=1$, is used
  exactly to remove the polynomial
  factor $m^{q_\partial-1}$ from the upper bounds in
  Theorem~\ref{thm:weighted-boundary-pressure}\emph{(3)},
  Corollary~\ref{cor:weighted-boundary-transfer}\emph{(4)}, and
  Corollary~\ref{cor:weighted-ambiguous-matrix}.  By itself it does not imply
  any lower bound.
\item The primitive refinement of Definition~\ref{def:regular-event} is
  stronger than the previous two refinements: it forces a single aperiodic
  maximal component, so both residue-complete and no-maximal-chain hold
  automatically.  It is a sufficient all-length special case, not a standing
  hypothesis of the general weighted theorem.
\end{enumerate}
\end{remark}

\subsection{Weighted boundary pressure from reachable boundary components}
\label{sub:weighted-pressure}

We isolate one point of terminology before the formal definition.  The
quantity $P_\partial(\phi)$ introduced in this subsection is \emph{not} the
weighted topological pressure of Barral--Feng
\cite{BarralFeng2012WeightedThermodynamic}, Feng--Huang
\cite{FengHuang2016WeightedPressure}, Tsukamoto
\cite{Tsukamoto2023WeightedPressure}, or Alibabaei
\cite{Alibabaei2025WeightedPressure}.  Those theories start from a factor map
or a tower of factor maps together with a weight vector and define a global
thermodynamic invariant through weighted Bowen balls or equivalent weighted
covering formalisms.  Here there is no factor tower and no entropy-weight
vector.  Instead, one fixes the event $\mathsf{P}$, forms the refined
ambiguous observer graph, and then takes the ordinary pressure of
$\phi\circ\pi_\partial$ on the reachable recurrent part of the resulting
boundary path shift.  The adjective ``boundary'' refers to ambiguous prefixes
of $\mathsf{P}$, and ``event-relative'' means that changing $\mathsf{P}$
changes the graph and hence the pressure object itself.

\begin{definition}[Refined ambiguous graph, boundary path shift, and weighted refinements]
\label{def:regular-event}
Let $\mathcal{A}$ be a finite alphabet, let $X\subseteq\mathcal{A}^{\NN}$ be a
one-sided shift space, let $\mu$ be a Borel probability measure on $X$, and
let $\mathsf{P}\in\mathcal{F}_\infty$ admit a finite-state posterior
presentation with respect to $\mu$ in the sense of
Definition~\ref{def:finite-posterior}.  Define the accessible ambiguous states
\[
\mathcal{S}_\partial^{\mathrm{acc}}
:=\{q(a): |a|\ge 1,\ a\in\mathcal{L}(X),\ [a]\in\partial_{|a|}\mathsf{P}\}
\subseteq \mathcal{S}_\partial,
\]
the refined ambiguous state space
\[
\mathcal{R}_\partial
:=\{(i,u)\in\mathcal{A}\times\mathcal{S}_\partial^{\mathrm{acc}}:
\exists a\in\mathcal{L}(X)\text{ with }[a]\in\partial_{|a|}\mathsf{P},\
\ a_{|a|-1}=i,\ q(a)=u\},
\]
the refined ambiguous transition matrix
\[
\widetilde{B}
=(\widetilde{B}_{(i,u),(j,v)})_{(i,u),(j,v)\in\mathcal{R}_\partial},
\qquad
\widetilde{B}_{(i,u),(j,v)}
:=
\begin{cases}
1,& ij\in\mathcal{L}_2(X)\text{ and }T_j(u)=v,\\
0,& \text{otherwise},
\end{cases}
\]
and the initial refined boundary states
\[
\mathcal{I}
:=\{(i,u)\in\mathcal{R}_\partial:[i]\in\partial_1\mathsf{P}\text{ and }q(i)=u\}.
\]
Let
\[
\mathbf{1}_{\mathcal{I}}^{\!\top}\in\mathbb{R}^{\mathcal{R}_\partial},
\qquad
(\mathbf{1}_{\mathcal{I}}^{\!\top})_r:=\mathbf{1}_{\{r\in\mathcal{I}\}},
\]
be the initial indicator row.  It records the length-$1$ refined boundary
states.  When this same refined state space is used in the Parry counting
theorem, this is exactly the initial row vector there denoted by $r_1$.
For $m\ge 1$, write
\[
\mathcal{P}_m^{\mathcal{I}}(\widetilde{B})
:=\bigl\{r_0\cdots r_{m-1}\in\mathcal{R}_\partial^m:
r_0\in\mathcal{I},\ \widetilde{B}_{r_k,r_{k+1}}=1\text{ for }0\le k<m-1\bigr\}
\]
for the finite refined boundary paths of length $m$ starting in
$\mathcal{I}$.  These are the exact finite paths counted by boundary words.
A path in $\mathcal{P}_m^{\mathcal{I}}(\widetilde{B})$ need not extend to an
infinite boundary path: the refined graph may contain reachable dead-end
ambiguous tails, and those terminal paths contribute to the finite partition
sums below even though they are absent from the one-sided shift
$\Sigma_\partial$.
This construction uses only the chosen posterior presentation and the shift
language of $X$; no Parry-measure hypothesis enters at this stage.  We call
the chosen presentation \emph{primitive}, or \emph{regular with respect to this
presentation}, if the refined ambiguous matrix $\widetilde{B}$ is primitive.
The associated \emph{boundary path shift} is the one-sided SFT
\[
\Sigma_\partial
:=\{r_0r_1\cdots\in\mathcal{R}_\partial^{\NN}:
\widetilde{B}_{r_k,r_{k+1}}=1\text{ for all }k\ge 0\}.
\]
Writing each refined state as $r_k=(i_k,u_k)$, define the symbol factor
\[
\pi_\partial:\Sigma_\partial\to X,
\qquad
\pi_\partial(r_0r_1\cdots):=(i_0i_1\cdots).
\]
Because $\widetilde{B}_{(i,u),(j,v)}=1$ forces $ij\in\mathcal{L}_2(X)$,
the image $\pi_\partial(\Sigma_\partial)$ lies in $X$.  For any continuous
potential $\phi:X\to\mathbb{R}$, let $\mathfrak{C}_\partial$ be the collection of strongly connected components
of $G(\widetilde{B})$ that are reachable from $\mathcal{I}$ and contain at
least one directed cycle, and for each $C\in\mathfrak{C}_\partial$ let
$\Sigma_C\subseteq\Sigma_\partial$ denote the irreducible edge shift carried by
$C$.  Let $\Sigma_\partial^{\mathrm{vis}}\subseteq\Sigma_\partial$ be the
subshift obtained by removing the states that are not reachable from
$\mathcal{I}$.  We then write
\[
P_C(\phi):=P_{\Sigma_C}(\phi\circ\pi_\partial|_{\Sigma_C}),
\qquad
P_\partial(\phi):=P_{\Sigma_\partial^{\mathrm{vis}}}(\phi\circ\pi_\partial)
=\max_{C\in\mathfrak{C}_\partial}P_C(\phi)
\]
for the \emph{boundary pressure} induced by the refined ambiguous graph.  If
$\mathfrak{C}_\partial\neq\varnothing$, the first expression is only shorthand
for the recurrent maximum on the right: it ignores reachable acyclic tails,
whereas the finite boundary partition sums below still count those tails
through $\mathcal{P}_m^{\mathcal{I}}(\widetilde{B})$.  If
$\mathfrak{C}_\partial=\varnothing$, we set $P_\partial(\phi):=-\infty$ and
$q_\partial:=0$.  Otherwise let $q_\partial$ be the maximal integer $q\ge 1$
for which there exist distinct components
$C_1,\dots,C_q\in\mathfrak{C}_\partial$ with
$P_{C_j}(\phi)=P_\partial(\phi)$ for all $j$ and with a directed path from
$C_j$ to $C_{j+1}$ in the condensation graph of
$\Sigma_\partial^{\mathrm{vis}}$ for each $1\le j<q$.  We call the chosen
presentation \emph{residue-complete} if some component
$C\in\mathfrak{C}_\partial$ with $P_C(\phi)=P_\partial(\phi)$ admits bridge
paths from $\mathcal{I}$ whose lengths realize every residue class modulo the
period $p_C$ of $C$.  We call it \emph{no-maximal-chain} if $q_\partial=1$.
\end{definition}

\begin{lemma}[Restricted initial-state partition sums on a primitive SFT]
\label{lem:restricted-initial-partition-sum}
Let $Y$ be a primitive one-step SFT over a finite alphabet $\mathcal{A}$, let
$\psi:Y\to\mathbb{R}$ be H\"older, and let $I\subseteq\mathcal{A}$ be nonempty.
For $m\ge 1$, define
\[
\mathcal{Z}_m(\psi)
:=\sum_{w\in\mathcal{L}_m(Y)}
\exp\!\bigl(\sup_{y\in[w]}S_m\psi(y)\bigr)
\]
and
\[
\mathcal{Z}_m^I(\psi)
:=\sum_{\substack{w\in\mathcal{L}_m(Y)\\ w_0\in I}}
\exp\!\bigl(\sup_{y\in[w]}S_m\psi(y)\bigr).
\]
Then there exist constants $A_I^-,A_I^+>0$ such that
\[
A_I^- e^{mP_Y(\psi)}
\le \mathcal{Z}_m^I(\psi)
\le A_I^+ e^{mP_Y(\psi)}
\qquad(m\ge 1).
\]
\end{lemma}

\begin{proof}
The upper bound is immediate once one compares with the unrestricted partition
sum.  Since $Y$ is a mixing SFT and $\psi$ is H\"older, standard
thermodynamic formalism gives constants $A_-,A_+>0$ such that
\[
A_-e^{mP_Y(\psi)}
\le \mathcal{Z}_m(\psi)
\le A_+e^{mP_Y(\psi)}
\qquad(m\ge 1)
\]
\cite[Ch.~1]{Bowen1975},\cite[Ch.~9]{WaltersErgodicTheory}.  Therefore
\[
\mathcal{Z}_m^I(\psi)\le \mathcal{Z}_m(\psi)\le A_+e^{mP_Y(\psi)}.
\]

For the lower bound, fix $i_\ast\in I$.  Let $A$ be the adjacency matrix of
$Y$.  Because $A$ is primitive, there exists $N\ge 1$ such that every entry of
$A^N$ is positive.  For each symbol $b\in\mathcal{A}$ choose one admissible
word
\[
\beta(b)=i_\ast c_1(b)\cdots c_{N-1}(b)
\]
of length $N$ such that $\beta(b)b$ is admissible in $Y$; this is possible
because $(A^N)_{i_\ast b}>0$.

Now let $w=bw_1\cdots w_{m-1}\in\mathcal{L}_m(Y)$.  Define
\[
\Phi(w):=\beta(b)w\in\mathcal{L}_{m+N}(Y).
\]
Then $\Phi(w)$ starts in $I$, and $\Phi$ is injective.  If $x\in[w]$, prefixing
the bridge $\beta(b)$ produces a point $y\in[\Phi(w)]$, so
\[
S_{m+N}\psi(y)=S_N\psi(y)+S_m\psi(x)\ge -N\|\psi\|_\infty+S_m\psi(x).
\]
Taking suprema over $x\in[w]$ gives
\[
\exp\!\bigl(\sup_{y\in[\Phi(w)]}S_{m+N}\psi(y)\bigr)
\ge
e^{-N\|\psi\|_\infty}
\exp\!\bigl(\sup_{x\in[w]}S_m\psi(x)\bigr).
\]
Summing over all $w\in\mathcal{L}_m(Y)$ and using injectivity of $\Phi$,
\[
\mathcal{Z}_{m+N}^I(\psi)
\ge e^{-N\|\psi\|_\infty}\mathcal{Z}_m(\psi)
\ge e^{-N\|\psi\|_\infty}A_-e^{mP_Y(\psi)}.
\]
Thus, for every $n\ge N+1$,
\[
\mathcal{Z}_n^I(\psi)
\ge e^{-N\|\psi\|_\infty}A_-e^{(n-N)P_Y(\psi)}
=\bigl(e^{-N\|\psi\|_\infty-NP_Y(\psi)}A_-\bigr)e^{nP_Y(\psi)}.
\]
For the finitely many values $1\le n\le N$, the sums $\mathcal{Z}_n^I(\psi)$
are strictly positive because $I\neq\varnothing$ and each one-cylinder $[i]$
with $i\in I$ is nonempty in a primitive SFT.  Therefore
\[
A_I^-:=\min\Bigl\{
e^{-N\|\psi\|_\infty-NP_Y(\psi)}A_-,
\min_{1\le n\le N}\mathcal{Z}_n^I(\psi)e^{-nP_Y(\psi)}
\Bigr\}>0
\]
gives the lower bound for all $n\ge 1$.  Taking $A_I^+:=A_+$ completes the
proof.
\end{proof}

\begin{lemma}[Bounded distortion on nonempty subsets of a cylinder]
\label{lem:subset-cylinder-distortion}
Let $X$ be a mixing one-step SFT over a finite alphabet, and let
$\phi:X\to\mathbb{R}$ be H\"older.  Then there exists a constant
$V_\phi<\infty$ such that for every admissible word $a$ of length $m\ge 1$ and
every nonempty subset $U\subseteq[a]$,
\[
\sup_{x\in[a]}S_m\phi(x)-V_\phi
\le \sup_{x\in U}S_m\phi(x)
\le \sup_{x\in[a]}S_m\phi(x).
\]
\end{lemma}

\begin{proof}
Because $\phi$ is H\"older on the mixing one-step SFT $X$, it has a finite
Bowen-variation constant
\[
V_\phi:=\sup_{m\ge 1}\sup_{x,z\in[a],\ a\in\mathcal{L}_m(X)}
|S_m\phi(x)-S_m\phi(z)|<\infty
\]
\cite[Ch.~1]{Bowen1975},\cite[Ch.~9]{WaltersErgodicTheory}.  Fix an admissible
word $a$ of length $m$ and a nonempty subset $U\subseteq[a]$.  The upper bound
is immediate since $U\subseteq[a]$.  For the lower bound, choose $z\in U$.  By
the definition of $V_\phi$, for every $x\in[a]$ one has
\[
S_m\phi(x)\le S_m\phi(z)+V_\phi\le \sup_{y\in U}S_m\phi(y)+V_\phi.
\]
Taking the supremum over $x\in[a]$ gives
\[
\sup_{x\in[a]}S_m\phi(x)\le \sup_{y\in U}S_m\phi(y)+V_\phi,
\]
which is equivalent to the stated lower bound.
\end{proof}

\begin{lemma}[Finite-path pressure lemma]
\label{lem:finite-path-pressure}
Let $G$ be a finite directed graph with vertex set $\mathcal{B}$, let
$I\subseteq\mathcal{B}$ be nonempty, let $X$ be a mixing one-step SFT over a
finite alphabet $\mathcal{A}$, let $\kappa:\mathcal{B}\to\mathcal{A}$ be a
symbol map such that the coordinatewise image of every directed path in $G$ is
an admissible word of $X$, and let $\phi:X\to\mathbb{R}$ be H\"older.  For
$m\ge 1$, define
\[
\mathcal{P}_m^I(G)
:=\{b_0\cdots b_{m-1}\in\mathcal{B}^m:
b_0\in I,\ b_k\to b_{k+1}\text{ in }G\text{ for }0\le k<m-1\}.
\]
Let $\mathfrak{C}_I$ be the cyclic strongly connected components of $G$ that
are reachable from $I$, and assume that $\mathfrak{C}_I\neq\varnothing$.  For
each $C\in\mathfrak{C}_I$, let $\Sigma_C$ be the irreducible edge shift
carried by $C$, let $\pi_C:\Sigma_C\to X$ be the one-block code induced by
$\kappa$, and set
\[
P_C(\phi):=P_{\Sigma_C}(\phi\circ\pi_C).
\]
Write
\[
P_*:=\max_{C\in\mathfrak{C}_I}P_C(\phi),
\]
and let $q$ be the maximal length of a directed chain
$C_1\to\cdots\to C_q$ in the condensation graph of the reachable part of $G$
such that $P_{C_j}(\phi)=P_*$ for every $j$.

For every finite path $\gamma\in\mathcal{P}_m^I(G)$, there is a unique ordered
list $D_1,\dots,D_t$ of cyclic strongly connected components visited by
$\gamma$, and there is a constant $B_G<\infty$, depending only on $G$, such
that the total number of vertices of $\gamma$ outside
$D_1\cup\cdots\cup D_t$ is at most $B_G$.  Equivalently, if $n_j$ denotes the
number of vertices of $\gamma$ spent in $D_j$ and
$L(\gamma_{\mathrm{acyc}})$ denotes the number of vertices of $\gamma$ outside
the visited cyclic components, then
\[
n_1+\cdots+n_t=m-L(\gamma_{\mathrm{acyc}})
\qquad\text{and}\qquad
L(\gamma_{\mathrm{acyc}})\le B_G.
\]
Then there exists $K>0$ such that
\[
\sum_{\gamma\in\mathcal{P}_m^I(G)}
\exp\!\bigl(\sup_{x\in[\kappa(\gamma)]}S_m\phi(x)\bigr)
\le K\,m^{q-1}e^{mP_*}
\qquad(m\ge 1).
\]
\end{lemma}

\begin{proof}
Let $\mathcal{T}$ be the set of vertices of $G$ not contained in any cyclic
strongly connected component.  Every strongly connected component of the
induced graph $G|\mathcal{T}$ is a single vertex with no loop, so
$G|\mathcal{T}$ is acyclic.  Therefore every path in $G|\mathcal{T}$ visits at
most $|\mathcal{T}|$ vertices, and we may take
\[
B_G:=|\mathcal{T}|.
\]
If $\gamma\in\mathcal{P}_m^I(G)$ meets cyclic components
$D_1,\dots,D_t$, then the condensation graph of the reachable part of $G$ is
acyclic, so $\gamma$ can enter each cyclic component at most once.  Hence the
ordered list $D_1,\dots,D_t$ is unique, and all vertices of $\gamma$ not
belonging to these components lie in the acyclic graph $G|\mathcal{T}$.  Thus
their total number is at most $B_G$, which is exactly the asserted bound on
$L(\gamma_{\mathrm{acyc}})$.

Let $V_\phi<\infty$ be a Bowen-variation constant for $\phi$ on $X$.  For each
$C\in\mathfrak{C}_I$, standard thermodynamic formalism on the irreducible SFT
$\Sigma_C$ gives a constant $K_C^+>0$ such that
\[
\sum_{\eta\in\mathcal{L}_n(\Sigma_C)}
\exp\!\bigl(\sup_{y\in[\eta]}S_n(\phi\circ\pi_C)(y)\bigr)
\le K_C^+e^{nP_C(\phi)}
\qquad(n\ge 1)
\]
\cite[Ch.~1]{Bowen1975},\cite[Ch.~9]{WaltersErgodicTheory}.  Fix a cyclic
component $C$ and an initial vertex $e\in C$.  Summing only over the words in
$\Sigma_C$ that start at $e$ can only decrease the left-hand side, so if
$\mathcal{Y}_{C,e}(n)$ denotes the sum of
\[
\exp\!\bigl(\sup_{x\in[\kappa(\eta)]}S_n\phi(x)\bigr)
\]
over all directed paths $\eta$ of length $n$ inside $C$ with first vertex
$\eta_0=e$, then Lemma~\ref{lem:subset-cylinder-distortion} gives
\[
\mathcal{Y}_{C,e}(n)\le e^{V_\phi}K_C^+e^{nP_C(\phi)}
\qquad(n\ge 1).
\]

Because $G$ is finite and the total acyclic length is bounded by $B_G$, there
are only finitely many decorated itineraries
\[
\Gamma=(\gamma_{\mathrm{acyc}};D_1,e_1,f_1;\dots;D_t,e_t,f_t)
\]
consisting of the transient part $\gamma_{\mathrm{acyc}}$, the cyclic
components visited, and the entry and exit vertices $(e_j,f_j)$ inside each
$D_j$.  Fix such an itinerary $\Gamma$.  The corresponding paths contribute at
most
\[
A_\Gamma
\sum_{\substack{n_1,\dots,n_t\ge 1\\
n_1+\cdots+n_t=m-L(\gamma_{\mathrm{acyc}})}}
\prod_{j=1}^t e^{n_jP_{D_j}(\phi)}
\]
for some constant $A_\Gamma>0$ absorbing the finitely many acyclic pieces and
the factors $e^{V_\phi}K_{D_j}^+$.  Let
\[
J_*(\Gamma):=\{j:P_{D_j}(\phi)=P_*\},
\qquad
t_*(\Gamma):=\#J_*(\Gamma).
\]
Because the maximal-pressure components visited by $\Gamma$ form a directed
chain in the condensation graph, one has $t_*(\Gamma)\le q$.

If $t_*(\Gamma)=0$, then every visited cyclic component has pressure strictly
smaller than $P_*$.  Since there are only finitely many cyclic components, the
positive gap
\[
\delta:=\min\{P_*-P_C(\phi):C\in\mathfrak{C}_I,\ P_C(\phi)<P_*\}
\]
is well defined, and the whole $\Gamma$-contribution is
$O(e^{m(P_*-\delta)})$, hence also $O(e^{mP_*})$.

Assume now that $t_*(\Gamma)\ge 1$.  For the components with
$P_D(\phi)<P_*$, write each factor as
\[
e^{n_jP_D(\phi)}=e^{n_jP_*}e^{-n_j(P_*-P_D(\phi))}.
\]
Summing first over those non-maximal lengths gives a finite product of
geometric series of the form
\[
\sum_{n\ge 1}e^{-n(P_*-P_D(\phi))},
\]
independent of $m$.  For the remaining $t_*(\Gamma)$ maximal components, the
positive lengths form a composition of
$m-L(\gamma_{\mathrm{acyc}})-\sum_{j\notin J_*(\Gamma)}n_j$ into
$t_*(\Gamma)$ parts.  The number of such compositions is
\[
\binom{m-L(\gamma_{\mathrm{acyc}})-\sum_{j\notin J_*(\Gamma)}n_j-1}
{t_*(\Gamma)-1}
\le (m+1)^{t_*(\Gamma)-1}
\le (m+1)^{q-1}.
\]
Thus the total $\Gamma$-contribution is bounded by
\[
A_\Gamma'(m+1)^{q-1}e^{mP_*}
\]
for some constant $A_\Gamma'>0$.  Since only finitely many itineraries occur,
summing over them gives
\[
\sum_{\gamma\in\mathcal{P}_m^I(G)}
\exp\!\bigl(\sup_{x\in[\kappa(\gamma)]}S_m\phi(x)\bigr)
\le K\,m^{q-1}e^{mP_*}
\]
for a suitable constant $K>0$ and all $m\ge 1$.
\end{proof}

\begin{lemma}[Finite graph pressure with transient bridges]
\label{lem:reachable-component-partition-sums}
Let $G$ be a finite directed graph with vertex set $\mathcal{B}$, let
$I\subseteq\mathcal{B}$ be nonempty, let $X$ be a mixing one-step SFT over a
finite alphabet $\mathcal{A}$, let $\kappa:\mathcal{B}\to\mathcal{A}$ be a
symbol map such that the coordinatewise image of every directed path in $G$ is
an admissible word of $X$, and let $\phi:X\to\mathbb{R}$ be H\"older.  For a
finite directed path $\gamma=b_0\cdots b_{m-1}$ write
\[
|\gamma|:=m
\]
for its vertex length.  If $\alpha=a_0\cdots a_{r-1}$ and
$\beta=b_0\cdots b_{s-1}$ are directed paths with $a_{r-1}=b_0$, define
\[
\alpha\star\beta:=a_0\cdots a_{r-1}b_1\cdots b_{s-1},
\]
so that
\[
|\alpha\star\beta|=|\alpha|+|\beta|-1.
\]
Thus a bridge and a recurrent segment share their common endpoint.  For
$m\ge 1$ define
\[
\mathcal{P}_m^I(G)
:=\{b_0\cdots b_{m-1}\in\mathcal{B}^m:
b_0\in I,\ b_k\to b_{k+1}\text{ in }G\text{ for }0\le k<m-1\}
\]
and
\[
w_\phi(\gamma)
:=\exp\!\bigl(\sup_{x\in[\kappa(\gamma)]}S_m\phi(x)\bigr),
\qquad
\mathfrak{Z}_m^I(G,\phi,\kappa)
\ :=\sum_{\gamma\in\mathcal{P}_m^I(G)} w_\phi(\gamma),
\]
where $\kappa(\gamma)$ denotes the projected word
$\kappa(b_0)\cdots\kappa(b_{m-1})$.  The sum is over all finite directed paths
starting in $I$, whether or not they extend to infinite paths.  Thus
$\mathfrak{Z}_m^I(G,\phi,\kappa)$ is an exact finite-path partition sum: it
already includes initial transient pieces, acyclic bridges between recurrent
components, and a possible terminal acyclic tail after the last recurrent
component.

Let $\mathfrak{C}_I$ be the strongly connected components of $G$ that are
reachable from $I$ and contain a directed cycle.  For each
$C\in\mathfrak{C}_I$, let $\Sigma_C$ be the irreducible edge shift carried by
$C$, let
\[
\pi_C:\Sigma_C\to X,
\qquad
(\pi_C y)_n:=\kappa(y_n),
\]
let $p_C$ be its period, and let
\[
P_C(\phi):=P_{\Sigma_C}(\phi\circ\pi_C).
\]
If $\mathfrak{C}_I=\varnothing$, then
$\mathfrak{Z}_m^I(G,\phi,\kappa)=0$ for all large $m$.  Assume henceforth that
$\mathfrak{C}_I\neq\varnothing$, and set
\[
P_*:=\max_{C\in\mathfrak{C}_I}P_C(\phi).
\]
Let $q$ be the maximal integer $q\ge 1$ for which there exist distinct
components $C_1,\dots,C_q\in\mathfrak{C}_I$ with
$P_{C_j}(\phi)=P_*$ for all $j$ and with a directed path from $C_j$ to
$C_{j+1}$ in the condensation graph of the reachable part of $G$ for each
$1\le j<q$.
Then the following hold.
\begin{enumerate}[label=(\roman*)]
\item There exists $A_+>0$ such that
  \[
  \mathfrak{Z}_m^I(G,\phi,\kappa)
  \le A_+\,m^{q-1}e^{mP_*}
  \qquad(m\ge 1).
  \]
  In particular,
  \[
  \limsup_{m\to\infty}\frac1m
  \log\mathfrak{Z}_m^I(G,\phi,\kappa)=P_*.
  \]
\item Let $C\in\mathfrak{C}_I$ satisfy $P_C(\phi)=P_*$, and let
  $\beta=b_0\cdots b_{\ell-1}$ be a directed path in $G$ satisfying
  $b_0\in I$ and $b_{\ell-1}\in C$.  Then there exists
  $A_-(\beta)>0$ such that
  \[
  \mathfrak{Z}_{\ell+np_C}^I(G,\phi,\kappa)
  \ge A_-(\beta)e^{(\ell+np_C)P_*}
  \qquad(n\ge 0).
  \]
\item Suppose there is a component $C\in\mathfrak{C}_I$ with
  $P_C(\phi)=P_*$ and directed paths
  $\beta_0,\dots,\beta_{p_C-1}$ from $I$ into $C$ whose lengths realize every
  residue class modulo $p_C$:
  \[
  |\beta_r|\equiv r \pmod{p_C}
  \qquad(0\le r\le p_C-1).
  \]
  Then there exist $A_->0$ and $m_0\ge 1$ such that
  \[
  \mathfrak{Z}_m^I(G,\phi,\kappa)\ge A_-e^{mP_*}
  \qquad(m\ge m_0).
  \]
\end{enumerate}
\end{lemma}

\begin{proof}
Let $V_\phi<\infty$ be a Bowen-variation constant for $\phi$ on $X$, so that
for every admissible word $a$ of length $n$ one has
\[
\sup_{x,z\in[a]}|S_n\phi(x)-S_n\phi(z)|\le V_\phi.
\]

If $\mathfrak{C}_I=\varnothing$, then every directed path starting from $I$
remains in the finite acyclic subgraph of $G$ reachable from $I$.  Hence there
is a maximal path length and
$\mathfrak{Z}_m^I(G,\phi,\kappa)=0$ for all large $m$.
If $\mathfrak{C}_I\neq\varnothing$, then the upper bound in~\emph{(i)} is
exactly Lemma~\ref{lem:finite-path-pressure}, applied to the same graph,
initial set, symbol map, and potential, so~\emph{(i)} already follows.
Assume from now on that $\mathfrak{C}_I\neq\varnothing$.

For the lower bound along a residue class, fix $C$ and $\beta$ as in
~\emph{(ii)}, write $p:=p_C$, and let $a:=b_{\ell-1}$.  If
$C=C_0\sqcup\cdots\sqcup C_{p-1}$ is the cyclic decomposition with
$a\in C_0$, then $\sigma^p$ restricted to
\[
(\Sigma_C)_0:=\{y\in\Sigma_C:y_0\in C_0\}
\]
is topologically mixing.  The potential
\[
\phi_C^{(p)}
:=\sum_{k=0}^{p-1}(\phi\circ\pi_C)\circ\sigma^k
\]
is H\"older on $((\Sigma_C)_0,\sigma^p)$, and its pressure is
$pP_C(\phi)$.  Applying
Lemma~\ref{lem:restricted-initial-partition-sum} to a primitive one-step
presentation of $((\Sigma_C)_0,\sigma^p)$ with initial symbol corresponding to
$a$ gives a constant $B_a^->0$ such that the partition sum over words in
$\Sigma_C$ of length $np+1$ starting at $a$ is bounded below by
$B_a^-e^{npP_C(\phi)}$ for every $n\ge 0$.  Every such word determines a
directed path $\eta$ of length $np+1$ in $C$ starting at $a$.  Concatenating
it to the fixed bridge $\beta$ produces a directed path $\beta\star\eta$ of
length
\[
|\beta\star\eta|
=|\beta|+|\eta|-1
=|\beta|+np
\]
starting in $I$.  Set
\[
U_{\beta,\eta}
:=\sigma^{|\beta|-1}([\kappa(\beta\star\eta)])
\subseteq[\kappa(\eta)].
\]
This subset is nonempty because $[\kappa(\beta\star\eta)]$ is nonempty.
Applying Lemma~\ref{lem:subset-cylinder-distortion} to the word $\kappa(\eta)$
and the nonempty subset $U_{\beta,\eta}$ gives
\[
\sup_{y\in U_{\beta,\eta}}S_{np+1}\phi(y)
\ge
\sup_{z\in[\kappa(\eta)]}S_{np+1}\phi(z)-V_\phi.
\]
Hence
\begin{align*}
\sup_{x\in[\kappa(\beta\star\eta)]}S_{|\beta|+np}\phi(x)
&\ge
-(|\beta|-1)\|\phi\|_\infty
+\sup_{y\in U_{\beta,\eta}}S_{np+1}\phi(y)\\
&\ge
-(|\beta|-1)\|\phi\|_\infty
-V_\phi
+\sup_{z\in[\kappa(\eta)]}S_{np+1}\phi(z)\\
&\ge
-(|\beta|-1)\|\phi\|_\infty
-V_\phi
+\sup_{y\in[\eta]_{\Sigma_C}}S_{np+1}(\phi\circ\pi_C)(y).
\end{align*}
Thus prepending the fixed bridge $\beta$ changes the weights by at most a
uniform multiplicative constant, so there exists $A_-(\beta)>0$ with
\[
\mathfrak{Z}_{|\beta|+np}^I(G,\phi,\kappa)
\ge A_-(\beta)e^{(|\beta|+np)P_C(\phi)}
=A_-(\beta)e^{(|\beta|+np)P_*}
\qquad(n\ge 0).
\]
This proves~\emph{(ii)}.  Combining~\emph{(ii)} with the upper bound already
proved shows that
\[
\limsup_{m\to\infty}\frac1m
\log\mathfrak{Z}_m^I(G,\phi,\kappa)=P_*,
\]
completing~\emph{(i)}.

Finally, assume the residue-complete hypothesis of~\emph{(iii)}.
For each residue class $r$ modulo $p_C$, apply~\emph{(ii)} to $\beta_r$ to
obtain a constant $A_r^->0$ such that
\[
\mathfrak{Z}_{|\beta_r|+np_C}^I(G,\phi,\kappa)
\ge A_r^-e^{(|\beta_r|+np_C)P_*}
\qquad(n\ge 0).
\]
Set
\[
m_0:=\max_{0\le r<p_C}|\beta_r|.
\]
Fix $m\ge m_0$, and let $r\in\{0,\dots,p_C-1\}$ be the residue class of $m$
modulo $p_C$.  Since the residue-complete hypothesis gives
$|\beta_r|\equiv r\pmod{p_C}$, the quantity
\[
n_r(m):=\frac{m-|\beta_r|}{p_C}
\]
is an integer.  It is also nonnegative because $m\ge m_0\ge |\beta_r|$.
Hence every $m\ge m_0$ admits the exact representation
\[
m=|\beta_r|+n_r(m)p_C
\]
for the bridge $\beta_r$ attached to its own residue class.  Applying the
preceding bound for this $r$ gives
\[
\mathfrak{Z}_m^I(G,\phi,\kappa)
\ge A_r^-e^{mP_*}
\ge \Bigl(\min_{0\le s<p_C}A_s^-\Bigr)e^{mP_*}
\qquad(m\ge m_0).
\]
Thus~\emph{(iii)} holds with
$A_-:=\min_{0\le s<p_C}A_s^- >0$.
\end{proof}

\begin{lemma}[Gibbs boundary-counting lemma]
\label{lem:gibbs-counting}
Assume the hypotheses of
Theorem~\ref{thm:weighted-boundary-pressure}.  For each $m\ge 1$, define the
row vector $r_\phi^{(m)}\in\mathbb{R}^{\mathcal{R}_\partial}$ by
\[
r_{\phi,(i,u)}^{(m)}
:=\#\{a\in\mathcal{L}_m(X):[a]\in\partial_m\mathsf{P},\ a_{m-1}=i,\ q(a)=u\}.
\]
Using the initial indicator row
$\mathbf{1}_{\mathcal{I}}^{\!\top}$ from
Definition~\ref{def:regular-event}, one has
$r_\phi^{(1)}=\mathbf{1}_{\mathcal{I}}^{\!\top}$ and
\[
r_\phi^{(m+1)}=r_\phi^{(m)}\widetilde{B}\qquad(m\ge 1).
\]
Consequently
\[
r_\phi^{(m)}=\mathbf{1}_{\mathcal{I}}^{\!\top}\widetilde{B}^{\,m-1}
\]
and
\[
N_m(\partial\mathsf{P})
=\mathbf{1}_{\mathcal{I}}^{\!\top}\widetilde{B}^{\,m-1}\mathbf{1}
\qquad(m\ge 1).
\]
Equivalently, Lemma~\ref{lem:boundary-word-refined-path} remains valid for the
full-support Gibbs state $\mu_\phi$: the recurrence is symbolic and does not
depend on the Parry transition probabilities.

For each boundary word $a=a_0\cdots a_{m-1}\in\mathcal{L}_m(X)$, the refined
state sequence
\[
r(a):=(a_0,q(a_0))(a_1,q(a_0a_1))\cdots(a_{m-1},q(a_0\cdots a_{m-1}))
\]
is an admissible length-$m$ path in the refined ambiguous graph whose first
symbol lies in
\[
\mathcal{I}.
\]
Conversely, every admissible length-$m$ path $r_0\cdots r_{m-1}$ in the
refined ambiguous graph with $r_0\in\mathcal{I}$ is uniquely of the form
$r(a)$ for an admissible boundary word $a\in\mathcal{L}_m(X)$.
In particular, the exact finite-path language of boundary words is
$\mathcal{P}_m^{\mathcal{I}}(\widetilde{B})$, not merely the length-$m$
language of the one-sided shift $\Sigma_\partial$.
\end{lemma}

\begin{proof}
Because $\mu_\phi$ is the equilibrium state of a H\"older potential on the
mixing SFT $X$, it has full support.  Hence every admissible cylinder in $X$
has positive $\mu_\phi$-measure.  Therefore Definition~\ref{def:finite-posterior}\emph{(4)}
classifies admissible words exactly as in the Parry case: an admissible word is
in $L_{1,\mu_\phi}(\mathsf{P})$, $L_{0,\mu_\phi}(\mathsf{P})$, or $L_{\partial,\mu_\phi}(\mathsf{P})$
according to whether its observer state lies in $\mathcal{S}_1$,
$\mathcal{S}_0$, or $\mathcal{S}_\partial$.

The proof of Lemma~\ref{lem:boundary-word-refined-path} now applies
verbatim with ``Parry measure'' replaced by ``full-support Gibbs state.''  To
keep the counting identity explicit, set
\[
\mathcal{B}_m^\phi(i,u)
:=\{a\in\mathcal{L}_m(X):[a]\in\partial_m\mathsf{P},\ a_{m-1}=i,\ q(a)=u\},
\qquad
r_{\phi,(i,u)}^{(m)}=\#\mathcal{B}_m^\phi(i,u).
\]
By Definition~\ref{def:regular-event}, a refined state $r=(i,u)$ lies in
$\mathcal{I}$ exactly when $[i]\in\partial_1\mathsf{P}$ and $q(i)=u$.  Hence
the length-$1$ count vector is precisely the indicator row
$r_\phi^{(1)}=\mathbf{1}_{\mathcal{I}}^{\!\top}$.  For fixed
$(j,v)\in\mathcal{R}_\partial$, every
$c\in\mathcal{B}_{m+1}^\phi(j,v)$ has a unique parent $a\in\mathcal{L}_m(X)$
with $c=aj$.  As above, full support and the absorbing property force
$a\in\mathcal{B}_m^\phi(i,u)$ for some $(i,u)$, and this parent-child relation
is equivalent to the edge condition $A_{ij}=1$ and $T_j(u)=v$.  Conversely,
every $a\in\mathcal{B}_m^\phi(i,u)$ and every edge
$\widetilde{B}_{(i,u),(j,v)}=1$ produce a boundary child $aj$ in
$\mathcal{B}_{m+1}^\phi(j,v)$.  Therefore
\[
r_{\phi,(j,v)}^{(m+1)}
=\sum_{(i,u)\in\mathcal{R}_\partial}
r_{\phi,(i,u)}^{(m)}\,\widetilde{B}_{(i,u),(j,v)}
=\bigl(r_\phi^{(m)}\widetilde{B}\bigr)_{(j,v)}.
\]
Iterating gives
\[
r_\phi^{(m)}=\mathbf{1}_{\mathcal{I}}^{\!\top}\widetilde{B}^{\,m-1},
\]
and summing its coordinates gives
\[
N_m(\partial\mathsf{P})
=\mathbf{1}_{\mathcal{I}}^{\!\top}\widetilde{B}^{\,m-1}\mathbf{1}.
\]

Now let $a=a_0\cdots a_{m-1}$ be an admissible boundary word.  For each
$0\le k\le m-1$, the prefix $a_0\cdots a_k$ is admissible and has positive
measure.  If $q(a_0\cdots a_k)$ lay in $\mathcal{S}_0\cup\mathcal{S}_1$, then
the remaining suffix $a_{k+1}\cdots a_{m-1}$ would keep the run inside the same
absorbing class by Definition~\ref{def:finite-posterior}\emph{(3)}, forcing
$q(a)\in\mathcal{S}_0\cup\mathcal{S}_1$ and hence $[a]\notin\partial_m\mathsf{P}$,
a contradiction.  Therefore $q(a_0\cdots a_k)\in\mathcal{S}_\partial$ for every
$k$, and thus
$r(a)\in\mathcal{R}_\partial^m$.  Moreover,
Definition~\ref{def:finite-posterior}\emph{(2)} gives
$T_{a_{k+1}}(q(a_0\cdots a_k))=q(a_0\cdots a_{k+1})$, so
$\widetilde{B}_{r_k,r_{k+1}}=1$ for each $k<m-1$.  Hence
$r(a)$ is an admissible refined path of length $m$.  Its first symbol lies in
$\mathcal{I}$ because $[a_0]\in\partial_1\mathsf{P}$.

Conversely, let $r_0\cdots r_{m-1}$ be an admissible refined path with
$r_0\in\mathcal{I}$, and write $r_k=(i_k,u_k)$.  Since
$\widetilde{B}_{r_k,r_{k+1}}=1$, we have $A_{i_ki_{k+1}}=1$ for every $k<m-1$,
so $a:=i_0\cdots i_{m-1}$ is admissible in $X$.  Because $r_0\in\mathcal{I}$,
$[i_0]\in\partial_1\mathsf{P}$ and $q(i_0)=u_0$.  Inductively,
Definition~\ref{def:finite-posterior}\emph{(2)} gives
$q(i_0\cdots i_k)=u_k$ for every $k$.  In particular,
$q(a)=u_{m-1}\in\mathcal{S}_\partial$, so Definition~\ref{def:finite-posterior}\emph{(4)}
and positivity of admissible cylinders imply $[a]\in\partial_m\mathsf{P}$.
Thus $r_0\cdots r_{m-1}=r(a)$.  Uniqueness follows because the symbol sequence
$a_0\cdots a_{m-1}$ is read off from the first coordinates of the refined
states, and the observer states are then forced by the deterministic update
rule in Definition~\ref{def:finite-posterior}\emph{(2)}.
\end{proof}

\begin{proposition}[Finite boundary language versus recurrent pressure]
\label{prop:finite-boundary-vs-recurrent-pressure}
Let $\mathcal{A}$ be a finite alphabet, let $X$ be a mixing one-step SFT on
$\mathcal{A}$, let $\phi:X\to\mathbb{R}$ be H\"older, and let $\mu_\phi$ be
its equilibrium state.  Let $\mathsf{P}\in\mathcal{F}_\infty$ admit a
finite-state posterior presentation with respect to $\mu_\phi$, and apply
Definition~\ref{def:regular-event} to this presentation.  For $m\ge 1$, define
\[
Z_m(\partial\mathsf{P},\phi)
:=\sum_{C=[a]\in\partial_m\mathsf{P}}
\exp\!\bigl(\sup_{x\in C}S_m\phi(x)-mP_X(\phi)\bigr)
\]
and
\[
\widehat Z_m(\partial\mathsf{P},\phi)
:=\sum_{\gamma\in\mathcal{P}_m^{\mathcal{I}}(\widetilde{B})}
\exp\!\bigl(\sup_{x\in[a(\gamma)]}S_m\phi(x)-mP_X(\phi)\bigr),
\]
where $a(\gamma)$ is the first-coordinate word of the refined path
$\gamma$.  Then, for every $m\ge 1$,
\[
Z_m(\partial\mathsf{P},\phi)=\widehat Z_m(\partial\mathsf{P},\phi).
\]
Hence the exact finite boundary language is
$\mathcal{P}_m^{\mathcal{I}}(\widetilde{B})$, not merely the length-$m$
language of the one-sided boundary shift $\Sigma_\partial$.

If $\mathfrak{C}_\partial=\varnothing$, then
$Z_m(\partial\mathsf{P},\phi)=\widehat Z_m(\partial\mathsf{P},\phi)=0$ for all
large $m$.  Assume henceforth that $\mathfrak{C}_\partial\neq\varnothing$, so
$q_\partial\ge 1$ and $P_\partial(\phi)\in\mathbb{R}$.
Then reachable acyclic prefixes, bridges between recurrent components, and
terminal dead-end tails affect the partition sums only through multiplicative
constants and the polynomial factor $m^{q_\partial-1}$; the exponential rate
is governed solely by the recurrent boundary pressure
$P_\partial(\phi)=\max_{C\in\mathfrak{C}_\partial}P_C(\phi)$.  More precisely:
\begin{enumerate}[label=(\roman*)]
\item There exists a constant $A_+>0$ such that for every $m\ge 1$,
  \[
  Z_m(\partial\mathsf{P},\phi)
  =\widehat Z_m(\partial\mathsf{P},\phi)
  \le A_+\,m^{q_\partial-1}e^{-m(P_X(\phi)-P_\partial(\phi))}.
  \]
  In particular,
  \[
  \limsup_{m\to\infty}\frac1m\log Z_m(\partial\mathsf{P},\phi)
  =P_\partial(\phi)-P_X(\phi).
  \]
\item Let $C\in\mathfrak{C}_\partial$ satisfy $P_C(\phi)=P_\partial(\phi)$, and
  let $\beta$ be a bridge path in the refined ambiguous graph from a state of
  $\mathcal{I}$ into $C$.  If $p_C$ is the period of $C$ and $|\beta|=\ell$,
  then there exists $A_-(\beta)>0$ such that
  \[
  Z_{\ell+np_C}(\partial\mathsf{P},\phi)
  =\widehat Z_{\ell+np_C}(\partial\mathsf{P},\phi)
  \ge A_-(\beta)e^{-(\ell+np_C)(P_X(\phi)-P_\partial(\phi))}
  \qquad(n\ge 0).
  \]
\item Suppose some component $C\in\mathfrak{C}_\partial$ with
  $P_C(\phi)=P_\partial(\phi)$ makes the chosen presentation
  residue-complete.  Thus there exist bridge paths
  $\beta_0,\dots,\beta_{p_C-1}$ from $\mathcal{I}$ into $C$ whose lengths
  realize every residue class modulo $p_C$.  Then there exist constants
  $A_-,A_+>0$ and $m_0\ge 1$ such that for every $m\ge m_0$,
  \[
  A_-\,e^{-m(P_X(\phi)-P_\partial(\phi))}
  \le Z_m(\partial\mathsf{P},\phi)
  =\widehat Z_m(\partial\mathsf{P},\phi)
  \le A_+\,m^{q_\partial-1}e^{-m(P_X(\phi)-P_\partial(\phi))}.
  \]
  If, in addition, the chosen presentation is no-maximal-chain, equivalently
  $q_\partial=1$, then the upper bound sharpens to
  \[
  Z_m(\partial\mathsf{P},\phi)
  =\widehat Z_m(\partial\mathsf{P},\phi)
  \le A_+\,e^{-m(P_X(\phi)-P_\partial(\phi))}
  \qquad(m\ge m_0).
  \]
\end{enumerate}
\end{proposition}

\begin{proof}
The equality
$Z_m(\partial\mathsf{P},\phi)=\widehat Z_m(\partial\mathsf{P},\phi)$ is exactly
the boundary-word/refined-path bijection from Lemma~\ref{lem:gibbs-counting}.
If $\mathfrak{C}_\partial=\varnothing$, then every refined boundary path lies
in the reachable acyclic part of $G_\partial:=G(\widetilde{B})$, so
$\widehat Z_m(\partial\mathsf{P},\phi)=0$ for all large $m$.  Hence
$Z_m(\partial\mathsf{P},\phi)=0$ for all large $m$ as well.  Assume now that
$\mathfrak{C}_\partial\neq\varnothing$.  Applying
Lemma~\ref{lem:finite-path-pressure} to the refined graph $G_\partial$, with
initial set $\mathcal{I}$, symbol map $\kappa(i,u):=i$, and potential $\phi$,
shows that
\[
e^{mP_X(\phi)}\widehat Z_m(\partial\mathsf{P},\phi)
\le K\,m^{q_\partial-1}e^{mP_\partial(\phi)}
\]
for some $K>0$.  After multiplying by $e^{-mP_X(\phi)}$, this is exactly the
upper bound and limsup statement in~\emph{(i)}.  For the lower bounds, apply
Lemma~\ref{lem:reachable-component-partition-sums}\emph{(ii)--(iii)} to the
same refined graph.  Since its reachable cyclic components are exactly the
elements of $\mathfrak{C}_\partial$, that lemma yields~\emph{(ii)} and the
all-length lower bound together with the no-maximal-chain
($q_\partial=1$) sharpening in~\emph{(iii)} after multiplying by
$e^{-mP_X(\phi)}$.  In particular, the eventual all-$m$ lower bound in
~\emph{(iii)} comes from the explicit residue-by-residue representation
$m=|\beta_r|+np_C$ established in the proof of
Lemma~\ref{lem:reachable-component-partition-sums}\emph{(iii)}.
\end{proof}

\begin{remark}[Hypothesis dependency map for the weighted boundary theorem]
\label{rem:weighted-hypothesis-map}
Before the weighted theorem, the following table summarizes which hypothesis
does what.  It condenses
Remark~\ref{rem:boundary-assumption-ledger} into the form used in the theorem
statement.
\begin{center}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{@{}p{1.45in}p{1.95in}p{2.55in}@{}}
\toprule
Label & Meaning in this section & Used for in Theorem~\ref{thm:weighted-boundary-pressure} \\
\midrule
Finite-state posterior presentation & Definition~\ref{def:finite-posterior}; gives the ambiguous states and refined graph & Base hypothesis.  Needed to define $\widetilde B$, $\mathcal{I}$, $\mathfrak{C}_\partial$, $P_\partial(\phi)$, and the finite partition sums. \\
$\mu_\phi$-observer-regular & Definition~\ref{def:observer-regular}: a $\mu_\phi$-finite presentation together with the $\mu_\phi$-resolving refinement & Used only for the uniform thickness conclusion in \emph{(4)} and hence for the scan-error comparison in \emph{(5)} and the transfer corollary; not needed for \emph{(1)}--\emph{(3)}. \\
Primitive & Definition~\ref{def:regular-event}; the refined ambiguous matrix $\widetilde B$ is primitive & Strong all-length special case.  It forces a single aperiodic maximal component, so residue-complete and no-maximal-chain hold automatically. \\
Residue-complete & Definition~\ref{def:regular-event}; some maximal-pressure component is reachable in every residue class modulo its period & Upgrades the residue-class lower bound in \emph{(2)} to the all-large-$m$ lower bound in \emph{(3)}. \\
No-maximal-chain & Definition~\ref{def:regular-event}; equivalently $q_\partial=1$ & Removes the polynomial factor $m^{q_\partial-1}$ from the upper bound in \emph{(3)}; by itself it gives no lower bound. \\
\bottomrule
\end{tabular}
\end{center}
\end{remark}

\begin{theorem}[Boundary pressure theorem from a $\mu_\phi$-finite presentation via reachable boundary components]
\label{thm:weighted-boundary-pressure}
Let $\mathcal{A}$ be a finite alphabet, let $X$ be a mixing one-step SFT on
$\mathcal{A}$, let
$\phi:X\to\mathbb{R}$ be H\"older, and let $\mu_\phi$ be its equilibrium
state.  Let $\mathsf{P}\in\mathcal{F}_\infty$ admit a finite-state posterior
presentation with respect to $\mu_\phi$ as in
Definition~\ref{def:finite-posterior}.  Apply
Definition~\ref{def:regular-event} to this $\mu_\phi$-posterior presentation,
and let
\[
(\mathcal{R}_\partial,\widetilde{B},\mathcal{I},\mathfrak{C}_\partial)
\]
denote the resulting refined ambiguous graph, initial refined boundary states,
and reachable recurrent boundary components.  Thus the observer and the
boundary graph in this theorem are built for the Gibbs measure $\mu_\phi$, not
merely imported from the Parry measure unless
Proposition~\ref{prop:full-support-invariant-decisive-languages} (or an
equivalent cylinder-robust argument) has been verified for the event at hand.
The role of the $\mu_\phi$-resolving, residue-complete, primitive, and
no-maximal-chain refinements in the statements below is summarized in
Remark~\ref{rem:weighted-hypothesis-map}; see also
Remark~\ref{rem:boundary-assumption-ledger} for the longer version.  Let
$P_X(\phi)$ denote the topological pressure of $\phi$ on $X$, and set
$\phi_\partial:=\phi\circ\pi_\partial$.  For $m\ge 1$, define
the weighted boundary sum
\[
Z_m(\partial\mathsf{P},\phi)
:=\sum_{C=[a]\in\partial_m\mathsf{P}}
\exp\!\bigl(\sup_{x\in C}S_m\phi(x)-mP_X(\phi)\bigr).
\]
For a finite refined boundary path
$\gamma=r_0\cdots r_{m-1}\in\mathcal{P}_m^{\mathcal{I}}(\widetilde{B})$, write
\[
a(\gamma):=\kappa(\gamma):=\kappa(r_0)\cdots\kappa(r_{m-1}),
\qquad
\kappa(i,u):=i.
\]
Define the exact finite-path partition sum
\[
\widehat Z_m(\partial\mathsf{P},\phi)
:=\sum_{\gamma\in\mathcal{P}_m^{\mathcal{I}}(\widetilde{B})}
\exp\!\bigl(\sup_{x\in[a(\gamma)]}S_m\phi(x)-mP_X(\phi)\bigr).
\]
Proposition~\ref{prop:finite-boundary-vs-recurrent-pressure} gives the exact identity
$Z_m(\partial\mathsf{P},\phi)=\widehat Z_m(\partial\mathsf{P},\phi)$ and shows
that reachable acyclic tails contribute only constants and the polynomial
factor $m^{q_\partial-1}$, while the exponential scale is governed solely by
the maximum boundary-component pressure
\[
P_\partial(\phi)=\max_{C\in\mathfrak{C}_\partial}P_C(\phi),
\]
whenever $\mathfrak{C}_\partial\neq\varnothing$, and by eventual vanishing
when $\mathfrak{C}_\partial=\varnothing$.  This is the event-relative boundary pressure attached to the
refined observer graph, not relative pressure for a factor map.  More
precisely, if $\mathfrak{C}_\partial=\varnothing$ then
\[
Z_m(\partial\mathsf{P},\phi)=0
\]
for all sufficiently large $m$.  The exponential statements below require
$\mathfrak{C}_\partial\neq\varnothing$, but the resolving-window thickness
statement~\emph{(4)} is independent of whether recurrent boundary components
exist, because it is proved directly from the finite set
$\mathcal{R}_\partial$ of reachable refined boundary states.  Assume now that
$\mathfrak{C}_\partial\neq\varnothing$, so $q_\partial\ge 1$ and
$P_\partial(\phi)\in\mathbb{R}$.  Then:
\begin{enumerate}[label=(\arabic*)]
\item There exists a constant $C_+>0$ such that for every $m\ge 1$,
  \[
  Z_m(\partial\mathsf{P},\phi)\le
  C_+\,m^{q_\partial-1}e^{-m(P_X(\phi)-P_\partial(\phi))},
  \]
  and
  \[
  \limsup_{m\to\infty}\frac1m\log Z_m(\partial\mathsf{P},\phi)
  =P_\partial(\phi)-P_X(\phi).
  \]
\item Let $C\in\mathfrak{C}_\partial$ satisfy $P_C(\phi)=P_\partial(\phi)$, and
  let $\beta$ be a bridge path in the refined ambiguous graph from a state of
  $\mathcal{I}$ into $C$.  If $p_C$ is the period of $C$ and $|\beta|=\ell$,
  then there exists $C_-(\beta)>0$ such that
  \[
  Z_{\ell+np_C}(\partial\mathsf{P},\phi)
  \ge C_-(\beta)\,
  e^{-(\ell+np_C)(P_X(\phi)-P_\partial(\phi))}
  \qquad(n\ge 0).
  \]
\item Suppose some component $C\in\mathfrak{C}_\partial$ with
  $P_C(\phi)=P_\partial(\phi)$ makes the chosen presentation
  residue-complete in the sense of Definition~\ref{def:regular-event}.  Thus
  there exist bridge paths $\beta_0,\dots,\beta_{p_C-1}$ from
  $\mathcal{I}$ into $C$ whose lengths realize every residue class modulo
  $p_C$, and there are constants
  $C_-,C_+>0$ and $m_0\ge 1$ such that for every $m\ge m_0$,
  \[
  C_-\,e^{-m(P_X(\phi)-P_\partial(\phi))}
  \le Z_m(\partial\mathsf{P},\phi)
  \le
  C_+\,m^{q_\partial-1}e^{-m(P_X(\phi)-P_\partial(\phi))}.
  \]
  In particular,
  \[
  \lim_{m\to\infty}\frac1m\log Z_m(\partial\mathsf{P},\phi)
  =P_\partial(\phi)-P_X(\phi).
  \]
  This residue-complete refinement is automatic when a maximal-pressure
  component is primitive, and hence in particular when the chosen
  presentation is primitive in the sense of
  Definition~\ref{def:regular-event}.  If, in addition, the chosen
  presentation is no-maximal-chain, equivalently $q_\partial=1$, then the
  upper bound sharpens to
  \[
  Z_m(\partial\mathsf{P},\phi)\le
  C_+\,e^{-m(P_X(\phi)-P_\partial(\phi))}
  \qquad(m\ge m_0).
  \]
\end{enumerate}

Independently of whether $\mathfrak{C}_\partial$ is empty, if the chosen
$\mu_\phi$-finite presentation is $\mu_\phi$-resolving in the sense of
Definition~\ref{def:observer-regular}, then:
\begin{enumerate}[label=(\arabic*),resume*]
\item There exists $\theta_\phi>0$ such that every boundary cylinder
  $C\in\partial_m\mathsf{P}$ satisfies
  \[
  \min\{\mu_\phi(\mathsf{P}\mid C),1-\mu_\phi(\mathsf{P}\mid C)\}\ge \theta_\phi.
  \]
\item There exist constants $K_-,K_+>0$ such that for every $m\ge 1$,
  \[
  K_-\,Z_m(\partial\mathsf{P},\phi)
  \le \varepsilon_m(\mathsf{P};\mu_\phi)
  \le K_+\,Z_m(\partial\mathsf{P},\phi).
  \]
  In particular, if $\mathfrak{C}_\partial=\varnothing$, then
  $\varepsilon_m(\mathsf{P};\mu_\phi)=0$ for all large $m$, while if
  $\mathfrak{C}_\partial\neq\varnothing$, then every exponential statement in
  \emph{(1)}--\emph{(3)} transfers to $\varepsilon_m(\mathsf{P};\mu_\phi)$
  with the same rate $P_X(\phi)-P_\partial(\phi)$.
\end{enumerate}
\medskip
\begin{center}
\fbox{\parbox{0.94\linewidth}{\textbf{Dependency statement.}
Conclusions \emph{(1)}--\emph{(3)} are statements about the weighted boundary
partition sums $Z_m(\partial\mathsf{P},\phi)$ on the refined boundary graph.
Conclusion \emph{(4)} is the separate $\mu_\phi$-resolving thickness input.
Only after combining \emph{(4)} with
Proposition~\ref{prop:scan-error-cylinder} does \emph{(5)} transfer the
pressure-gap exponent to the scan-error profile
$\varepsilon_m(\mathsf{P};\mu_\phi)$.}}
\end{center}
\end{theorem}

\begin{lemma}[Quasi-Bernoulli lower bound for exact admissible concatenations]
\label{lem:qb-exact-concat}
Let $X$ be a mixing one-step SFT on a finite alphabet, let
$\phi:X\to\mathbb{R}$ be H\"older, and let $\mu_\phi$ be its equilibrium
state.  Then there exists a constant $C_{\mathrm{qb}}\ge 1$ such that for
every pair of admissible words $u,v\in\mathcal{L}(X)$ with $uv$ admissible,
\[
\mu_\phi([uv])\ge C_{\mathrm{qb}}^{-1}\mu_\phi([u])\mu_\phi([v]).
\]
Equivalently,
\[
\mu_\phi([v]\mid [u])\ge C_{\mathrm{qb}}^{-1}\mu_\phi([v])
\]
whenever $uv$ is admissible.
\end{lemma}

\begin{proof}
By Bowen's Gibbs estimate on the mixing SFT $X$
\cite{Bowen1975,WaltersErgodicTheory}, there exist constants $c_-,c_+>0$ such
that for every admissible word $a$ of length $m\ge 1$,
\[
c_-\exp\!\bigl(\sup_{x\in[a]}S_m\phi(x)-mP_X(\phi)\bigr)
\le \mu_\phi([a])
\le
c_+\exp\!\bigl(\sup_{x\in[a]}S_m\phi(x)-mP_X(\phi)\bigr)
\]
Since $\phi$ is H\"older, its Bowen
variation is uniformly bounded: there exists $V_\phi<\infty$ such that for
every admissible word $a$ of length $m$,
\[
\sup_{x,y\in[a]}|S_m\phi(x)-S_m\phi(y)|\le V_\phi.
\]

Now let $u,v$ be admissible, with lengths $m$ and $n$, and assume that $uv$ is
admissible.  Choose $y\in[v]$ with
$S_n\phi(y)=\sup_{z\in[v]}S_n\phi(z)$; this exists because $[v]$ is compact and
$S_n\phi$ is continuous.  Prefixing $u$ to $y$ gives a point $z:=uy\in[uv]$.
Because the first $m$ coordinates of $z$ are exactly $u$, shifting past them
recovers the tail $y$, so $\sigma^m(z)=\sigma^m(uy)=y$.
Since every point of $[u]$ has the same first $m$ symbols,
\[
S_m\phi(z)\ge \sup_{x\in[u]}S_m\phi(x)-V_\phi.
\]
Therefore
\[
\sup_{x\in[uv]}S_{m+n}\phi(x)
\ge S_{m+n}\phi(z)
=S_m\phi(z)+S_n\phi(y)
\ge \sup_{x\in[u]}S_m\phi(x)+\sup_{x\in[v]}S_n\phi(x)-V_\phi.
\]
Applying the lower Gibbs bound to $[uv]$ and the upper Gibbs bounds to $[u]$
and $[v]$ yields
\[
\mu_\phi([uv])
\ge c_-e^{-V_\phi}
\exp\!\bigl(\sup_{x\in[u]}S_m\phi(x)-mP_X(\phi)\bigr)
\exp\!\bigl(\sup_{x\in[v]}S_n\phi(x)-nP_X(\phi)\bigr),
\]
and therefore
\[
\mu_\phi([uv])\ge c_-c_+^{-2}e^{-V_\phi}\mu_\phi([u])\mu_\phi([v]).
\]
Thus the claim holds with
$C_{\mathrm{qb}}:=c_+^2e^{V_\phi}/c_-$.  The conditional-probability form is
the same inequality divided by $\mu_\phi([u])>0$.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:weighted-boundary-pressure}]
Assertion~\emph{(4)} is exactly
Lemma~\ref{lem:weighted-thickness-without-components} below.  In particular,
the thickness bound is proved directly from the finite resolving family over
$\mathcal{R}_\partial$ and does not use the recurrent boundary-component set
$\mathfrak{C}_\partial$.
Let $\theta_\phi>0$ be the constant supplied by~\emph{(4)}.

Bowen's Gibbs estimate on $X$ \cite{Bowen1975,WaltersErgodicTheory} gives
constants $c_-,c_+>0$ such that for every positive cylinder $C=[a]$ of length
$m$,
\[
c_-\exp\!\bigl(\sup_{x\in C}S_m\phi(x)-mP_X(\phi)\bigr)
\le \mu_\phi(C)
\le
c_+\exp\!\bigl(\sup_{x\in C}S_m\phi(x)-mP_X(\phi)\bigr).
\]
Combining this with Proposition~\ref{prop:scan-error-cylinder} and~\emph{(4)}
yields
\[
\theta_\phi c_- Z_m(\partial\mathsf{P},\phi)
\le \varepsilon_m(\mathsf{P};\mu_\phi)
\le \frac{c_+}{2}Z_m(\partial\mathsf{P},\phi),
\]
so~\emph{(5)} holds with $K_-:=\theta_\phi c_-$ and $K_+:=c_+/2$.

If $\mathfrak{C}_\partial=\varnothing$, then the eventual vanishing of
$Z_m(\partial\mathsf{P},\phi)$ was already proved in
Proposition~\ref{prop:finite-boundary-vs-recurrent-pressure}, so there is
nothing more to prove.  Assume henceforth that
$\mathfrak{C}_\partial\neq\varnothing$.  Then
Proposition~\ref{prop:finite-boundary-vs-recurrent-pressure} proves
~\emph{(1)},~\emph{(2)}, and~\emph{(3)}, and combining those statements with
~\emph{(5)} gives the advertised transfer of the same pressure-gap exponent to
$\varepsilon_m(\mathsf{P};\mu_\phi)$.
\end{proof}

\begin{lemma}[Resolving thickness does not require recurrent boundary components]
\label{lem:weighted-thickness-without-components}
Assume the hypotheses of Theorem~\ref{thm:weighted-boundary-pressure}, and
suppose that the chosen $\mu_\phi$-finite presentation is
$\mu_\phi$-resolving in the sense of
Definition~\ref{def:observer-regular}.  Then there exists $\theta_\phi>0$ such
that for every $m\ge 1$ and every boundary cylinder
$C\in\partial_m\mathsf{P}$,
\[
\min\{\mu_\phi(\mathsf{P}\mid C),1-\mu_\phi(\mathsf{P}\mid C)\}\ge \theta_\phi.
\]
This remains true regardless of whether the recurrent boundary-component set
$\mathfrak{C}_\partial$ is empty.
\end{lemma}

\begin{proof}
If $\mathcal{R}_\partial=\varnothing$, then there is no boundary cylinder of
positive length.  Indeed, if $[a]\in\partial_m\mathsf{P}$ for some $m\ge 1$,
then the final refined state $(a_{m-1},q(a))$ lies in $\mathcal{R}_\partial$
by definition, contradiction.  Hence the assertion is vacuous in this case.

Assume now that $\mathcal{R}_\partial\neq\varnothing$.  For each
$r\in\mathcal{R}_\partial$, choose resolving words $w_r^+,w_r^-$ as in
Definition~\ref{def:observer-regular}, and let
\[
\mathcal{W}:=\{w_r^+,w_r^-:r\in\mathcal{R}_\partial\}.
\]
This is a finite set of admissible words.  Bowen's Gibbs estimate on cylinders
\cite{Bowen1975,WaltersErgodicTheory} gives positive $\mu_\phi$-measure to
every admissible cylinder, so
\[
\min_{w\in\mathcal{W}}\mu_\phi([w])>0.
\]
Let $C_{\mathrm{qb}}$ be the exact-concatenation constant from
Lemma~\ref{lem:qb-exact-concat}, and set
\[
\theta_\phi
:=C_{\mathrm{qb}}^{-1}\min_{w\in\mathcal{W}}\mu_\phi([w])>0.
\]

Now fix $m\ge 1$ and a boundary cylinder $C=[a]\in\partial_m\mathsf{P}$.
Then
\[
r:=(a_{m-1},q(a))\in\mathcal{R}_\partial.
\]
By the resolving property, the concatenations $aw_r^+$ and $aw_r^-$ are
admissible.  Applying Lemma~\ref{lem:qb-exact-concat} with $u=a$ and
$v=w_r^\pm$ gives
\[
\mu_\phi([aw_r^\pm]\mid[a])
\ge C_{\mathrm{qb}}^{-1}\mu_\phi([w_r^\pm])
\ge \theta_\phi.
\]
Again by the resolving property,
$q(aw_r^+)\in\mathcal{S}_1$ and $q(aw_r^-)\in\mathcal{S}_0$.  Therefore
Definition~\ref{def:finite-posterior}\emph{(4)} implies that $[aw_r^+]$ lies
in $\mathsf{P}$ modulo $\mu_\phi$-null sets and $[aw_r^-]$ lies in
$[a]\setminus\mathsf{P}$ modulo $\mu_\phi$-null sets.  Consequently,
\[
\mu_\phi(\mathsf{P}\mid[a])\ge \theta_\phi
\qquad\text{and}\qquad
1-\mu_\phi(\mathsf{P}\mid[a])\ge \theta_\phi,
\]
which is the claimed lower bound.

No part of this proof refers to recurrent components or to
$\mathfrak{C}_\partial$, so the same constant works whether or not
$\mathfrak{C}_\partial$ is empty.
\end{proof}

\begin{corollary}[Weighted transfer corollary from the boundary pressure theorem]
\label{cor:weighted-boundary-transfer}
Assume the hypotheses of Theorem~\ref{thm:weighted-boundary-pressure} and, in
addition, that the chosen $\mu_\phi$-finite presentation is
$\mu_\phi$-resolving, so that
Theorem~\ref{thm:weighted-boundary-pressure}\emph{(4)}--\emph{(5)} apply.  Then
there exist constants $K_-,K_+>0$ such that for every $m\ge 1$,
\[
K_-\,Z_m(\partial\mathsf{P},\phi)
\le \varepsilon_m(\mathsf{P};\mu_\phi)
\le K_+\,Z_m(\partial\mathsf{P},\phi).
\]
If $\mathfrak{C}_\partial=\varnothing$, then
$\varepsilon_m(\mathsf{P};\mu_\phi)=0$ for all large $m$.  Otherwise the
following hold.
\begin{enumerate}[label=(\arabic*)]
\item There exists $L_+>0$ such that for every $m\ge 1$,
  \[
  \varepsilon_m(\mathsf{P};\mu_\phi)
  \le L_+\,m^{q_\partial-1}e^{-m(P_X(\phi)-P_\partial(\phi))}.
  \]
\item Let $C\in\mathfrak{C}_\partial$ satisfy $P_C(\phi)=P_\partial(\phi)$, and
  let $\beta$ be a bridge path from $\mathcal{I}$ into $C$.  If $p_C$ is the
  period of $C$ and $|\beta|=\ell$, then there exists $L_-(\beta)>0$ such that
  \[
  \varepsilon_{\ell+np_C}(\mathsf{P};\mu_\phi)
  \ge L_-(\beta)\,
  e^{-(\ell+np_C)(P_X(\phi)-P_\partial(\phi))}
  \qquad(n\ge 0).
  \]
\item One has
  \[
  P_X(\phi)>P_\partial(\phi)
  \]
  and
  \[
  \limsup_{m\to\infty}\frac1m\log\varepsilon_m(\mathsf{P};\mu_\phi)
  =-(P_X(\phi)-P_\partial(\phi)).
  \]
\item If the residue-complete hypothesis from
  Theorem~\ref{thm:weighted-boundary-pressure}\emph{(3)} holds, then there
  exist $L_-,L_+>0$ and $m_0\ge 1$ such that
  \[
  L_-\,e^{-m(P_X(\phi)-P_\partial(\phi))}
  \le \varepsilon_m(\mathsf{P};\mu_\phi)
  \le
  L_+\,m^{q_\partial-1}e^{-m(P_X(\phi)-P_\partial(\phi))}
  \qquad(m\ge m_0),
  \]
  and consequently
  \[
  -\lim_{m\to\infty}\frac1m\log\varepsilon_m(\mathsf{P};\mu_\phi)
  =P_X(\phi)-P_\partial(\phi).
  \]
  If, in addition, the chosen presentation is no-maximal-chain, equivalently
  $q_\partial=1$, then the upper bound sharpens to
  \[
  \varepsilon_m(\mathsf{P};\mu_\phi)
  \le L_+\,e^{-m(P_X(\phi)-P_\partial(\phi))}
  \qquad(m\ge m_0).
  \]
\end{enumerate}
\end{corollary}

\begin{proof}
The first display is exactly
Theorem~\ref{thm:weighted-boundary-pressure}\emph{(5)}.  The point is that the
pressure estimates from Theorem~\ref{thm:weighted-boundary-pressure}\emph{(1)}--\emph{(3)}
control only $Z_m(\partial\mathsf{P},\phi)$; the extra
$\mu_\phi$-resolving hypothesis supplies the thickness input from
Theorem~\ref{thm:weighted-boundary-pressure}\emph{(4)}, and only then does the
same exponent transfer to $\varepsilon_m(\mathsf{P};\mu_\phi)$.

If $\mathfrak{C}_\partial=\varnothing$, then
Theorem~\ref{thm:weighted-boundary-pressure} gives
$Z_m(\partial\mathsf{P},\phi)=0$ for all large $m$, hence also
$\varepsilon_m(\mathsf{P};\mu_\phi)=0$ for all large $m$.  Assume now that
$\mathfrak{C}_\partial\neq\varnothing$.  Combining the first display with
Theorem~\ref{thm:weighted-boundary-pressure}\emph{(1)} proves~\emph{(1)} with
$L_+:=K_+C_+$.  To obtain~\emph{(2)}, write
\[
G_\partial:=G(\widetilde{B}),
\qquad
\kappa:\mathcal{R}_\partial\to\mathcal{A},
\qquad
\kappa(i,u):=i,
\]
and let
\[
\widehat Z_m(\partial\mathsf{P},\phi)
:=\sum_{\gamma\in\mathcal{P}_m^{\mathcal{I}}(G_\partial)}
\exp\!\bigl(\sup_{x\in[\kappa(\gamma)]}S_m\phi(x)-mP_X(\phi)\bigr).
\]
By Lemma~\ref{lem:gibbs-counting}, boundary words of length $m$ are in
bijection with $\mathcal{P}_m^{\mathcal{I}}(G_\partial)$, so
\[
Z_m(\partial\mathsf{P},\phi)
=\widehat Z_m(\partial\mathsf{P},\phi)
\qquad(m\ge 1).
\]
Applying Lemma~\ref{lem:reachable-component-partition-sums}\emph{(ii)} to the
refined graph $G_\partial$, with initial set $\mathcal{I}$ and symbol map
$\kappa$, gives a constant $A_-(\beta)>0$ such that
\[
\widehat Z_{\ell+np_C}(\partial\mathsf{P},\phi)
\ge A_-(\beta)e^{-(\ell+np_C)P_X(\phi)}e^{(\ell+np_C)P_\partial(\phi)}
\qquad(n\ge 0).
\]
Multiplying by the comparison constant $K_-$ from the first display proves
~\emph{(2)} with $L_-(\beta):=K_-A_-(\beta)$.

Because $\mathsf{P}\in\mathcal{F}_\infty$, Corollary~\ref{cor:scan-tanaka-tail}
gives
\[
\varepsilon_m(\mathsf{P};\mu_\phi)\to 0
\qquad(m\to\infty).
\]
Choose a component $C\in\mathfrak{C}_\partial$ with
$P_C(\phi)=P_\partial(\phi)$ and a bridge path $\beta$ from $\mathcal{I}$ into
$C$, with $|\beta|=\ell$.  By~\emph{(2)}, there exists $L_-(\beta)>0$ such that
\[
\varepsilon_{\ell+np_C}(\mathsf{P};\mu_\phi)
\ge L_-(\beta)\,
e^{-(\ell+np_C)(P_X(\phi)-P_\partial(\phi))}
\qquad(n\ge 0).
\]
If $P_X(\phi)-P_\partial(\phi)\le 0$, then the right-hand side is bounded below
by the positive constant $L_-(\beta)$ for all $n\ge 0$, contradicting
$\varepsilon_m(\mathsf{P};\mu_\phi)\to 0$.  Hence
\[
P_X(\phi)>P_\partial(\phi).
\]
The upper bound from~\emph{(1)} and the residue-class lower bound from
~\emph{(2)} then imply the limsup identity in~\emph{(3)}.  Finally, combining
the first display with Theorem~\ref{thm:weighted-boundary-pressure}\emph{(3)}
proves~\emph{(4)}.
\end{proof}

\begin{lemma}[Standard reducible Perron--Frobenius polynomial bound along the condensation DAG]
\label{lem:reducible-pf-dag-bound}
Let $M$ be a nonnegative $d\times d$ matrix, let
$\alpha\in[0,\infty)^d$ be a nonzero row vector, and let $\mathbf{1}$ be the
all-ones column vector.  Let $\mathfrak{C}_\alpha$ be the strongly connected
components of $G(M)$ that are reachable from $\operatorname{supp}(\alpha)$ and
contain a directed cycle.  If $\mathfrak{C}_\alpha=\varnothing$, then
$\alpha M^n\mathbf{1}=0$ for all large $n$.  Assume
henceforth that $\mathfrak{C}_\alpha\neq\varnothing$, and set
\[
\rho_\alpha(M):=\max_{C\in\mathfrak{C}_\alpha}\rho(M_C).
\]
Let $q_\alpha(M)$ be the maximal integer $q\ge 1$ for which there exist
distinct components $C_1,\dots,C_q\in\mathfrak{C}_\alpha$ with
$\rho(M_{C_j})=\rho_\alpha(M)$ for all $j$ and with a directed path from
$C_j$ to $C_{j+1}$ in the reachable condensation graph of $G(M)$ for each
$1\le j<q$.  Then there exists a constant $K>0$ such that
\[
\alpha M^n\mathbf{1}\le K\,n^{q_\alpha(M)-1}\rho_\alpha(M)^n
\qquad(n\ge 1).
\]
In particular, when $q_\alpha(M)=1$ one has $\alpha M^n\mathbf{1}=O(\rho_\alpha(M)^n)$.
This is the standard reducible Perron--Frobenius polynomial-loss phenomenon
for chains of maximal spectral-radius components in Frobenius normal form;
compare \cite[Ch.~1]{Seneta2006},\cite[Chs.~3--4]{BermanPlemmons1994}.
\end{lemma}

\begin{proof}
If $\mathfrak{C}_\alpha=\varnothing$, then every path in $G(M)$ starting from
$\operatorname{supp}(\alpha)$ lies in the finite acyclic reachable subgraph.
Hence there is a maximal possible path length, so $\alpha M^n\mathbf{1}=0$ for
all large $n$.

Assume now that $\mathfrak{C}_\alpha\neq\varnothing$.  After a simultaneous row
and column permutation, the reachable part of $M$ is in Frobenius normal form,
so each diagonal block $M_C$ indexed by $C\in\mathfrak{C}_\alpha$ is
irreducible.  Standard Perron--Frobenius theory for irreducible nonnegative
matrices gives constants $K_C>0$ such that
\[
\mathbf{1}_C^\top M_C^n\mathbf{1}_C\le K_C\,\rho(M_C)^n
\qquad(n\ge 0,\ C\in\mathfrak{C}_\alpha)
\]
\cite[Ch.~1]{Seneta2006},\cite[Chs.~3--4]{BermanPlemmons1994}.  Let
\[
\tau:=\max\{\rho(M_C):C\in\mathfrak{C}_\alpha,\ \rho(M_C)<\rho_\alpha(M)\}
\]
if such components exist, and set $\tau:=0$ otherwise.  Then
$0\le \tau<\rho_\alpha(M)$.

Every weighted walk counted by $\alpha M^n\mathbf{1}$ visits the strongly
connected components of the reachable graph in a unique order
$D_1,\dots,D_s$.  Because the condensation graph is acyclic, each component
appears at most once, and because the reachable vertices outside
$\bigcup_{C\in\mathfrak{C}_\alpha}C$ form a finite acyclic graph, only
uniformly bounded time can be spent outside the cyclic components.  Therefore
there is a finite family $\Gamma$ of decorated component itineraries with the
following property: for each $\gamma\in\Gamma$, every length-$n$ walk following
$\gamma$ is obtained by choosing nonnegative integers $n_1,\dots,n_s$ with
$n_1+\cdots+n_s=n-L(\gamma)$, where $n_j$ is the time spent inside $D_j$ and
$L(\gamma)$ is the total length of the fixed acyclic bridges, and then choosing
the internal walks inside the components $D_j$.

Fix $\gamma\in\Gamma$, with cyclic components $D_1,\dots,D_s$, and let
$J(\gamma)\subseteq\{1,\dots,s\}$ be the indices of the components satisfying
$\rho(M_{D_j})=\rho_\alpha(M)$.  Write $t(\gamma):=\#J(\gamma)$.  Since the
maximal-spectral-radius components appearing in $\gamma$ form a directed chain
in the reachable condensation graph, one has $t(\gamma)\le q_\alpha(M)$.
Nonnegativity of $M$ gives a constant $B_\gamma>0$ such that the total weight
of all length-$n$ walks following $\gamma$ is at most
\[
B_\gamma
\sum_{\substack{n_1,\dots,n_s\ge 0\\
n_1+\cdots+n_s=n-L(\gamma)}}
\prod_{j=1}^s \mathbf{1}_{D_j}^\top M_{D_j}^{n_j}\mathbf{1}_{D_j}.
\]
Using the block bounds above, this is at most
\[
B_\gamma\Bigl(\prod_{j=1}^s K_{D_j}\Bigr)
\sum_{\substack{n_1,\dots,n_s\ge 0\\
n_1+\cdots+n_s=n-L(\gamma)}}
\rho_\alpha(M)^{\sum_{j\in J(\gamma)}n_j}\tau^{\sum_{j\notin J(\gamma)}n_j}.
\]
Summing first over the lower-spectral-radius components gives a geometric
factor depending only on $\gamma$ and $\tau/\rho_\alpha(M)$, while the
remaining choices of lengths inside the $t(\gamma)$ maximal components are
bounded by the number of weak compositions of an integer into $t(\gamma)$
parts, namely $O(n^{t(\gamma)-1})$.  Hence there exists $A_\gamma>0$ such that
the $\gamma$-contribution is at most
\[
A_\gamma\,n^{t(\gamma)-1}\rho_\alpha(M)^n
\le A_\gamma\,n^{q_\alpha(M)-1}\rho_\alpha(M)^n.
\]
There are only finitely many itineraries in $\Gamma$, so summing over them
gives
\[
\alpha M^n\mathbf{1}\le K\,n^{q_\alpha(M)-1}\rho_\alpha(M)^n
\qquad(n\ge 1)
\]
for some $K>0$, as claimed.
\end{proof}

\begin{corollary}[Weighted ambiguous transfer matrix in the locally constant case]
\label{cor:weighted-ambiguous-matrix}
Under the hypotheses of Theorem~\ref{thm:weighted-boundary-pressure}, assume
in addition that there is a function
\[
F:\mathcal{R}_\partial^{\mathrm{vis}}\to\mathbb{R}
\]
such that for every $m\ge 1$ and every finite refined boundary path
$\gamma=r_0\cdots r_{m-1}\in\mathcal{P}_m^{\mathcal{I}}(\widetilde{B})$ one has
\[
\sup_{x\in[a(\gamma)]}S_m\phi(x)=\sum_{k=0}^{m-1}F(r_k).
\]
On the recurrent part this is exactly the statement that
$\phi_\partial=\phi\circ\pi_\partial$ is locally constant on one-cylinders of
$\Sigma_\partial^{\mathrm{vis}}$; the stronger formulation records the same
one-step weight on every reachable transient state as well, so finite
dead-end boundary tails are counted exactly.
Let $\mathcal{R}_\partial^{\mathrm{vis}}$ be the set of refined boundary states
reachable from $\mathcal{I}$.  Define the visible weighted ambiguous transfer
matrix
\[
M_{\partial,\phi}
 :=\bigl((M_{\partial,\phi})_{rs}\bigr)_{r,s\in\mathcal{R}_\partial^{\mathrm{vis}}},
\qquad
(M_{\partial,\phi})_{rs}:=\widetilde{B}_{rs}e^{F(s)}.
\]
Also define
\[
\alpha_{\partial,\phi}
:=\bigl(\mathbf{1}_{\{r\in\mathcal{I}\}}e^{F(r)}\bigr)_{r\in\mathcal{R}_\partial^{\mathrm{vis}}}.
\]
Then for every $m\ge 1$,
\[
Z_m(\partial\mathsf{P},\phi)
=e^{-mP_X(\phi)}\alpha_{\partial,\phi}M_{\partial,\phi}^{\,m-1}\mathbf{1}.
\]
This is the exact finite refined-path partition sum, including reachable
dead-end tails, not merely the partition sum over the recurrent subshift.
If $\mathfrak{C}_\partial=\varnothing$, then $M_{\partial,\phi}$ is nilpotent
and therefore $Z_m(\partial\mathsf{P},\phi)=0$ for all sufficiently large $m$.
Assume henceforth that $\mathfrak{C}_\partial\neq\varnothing$.  Then
\[
P_\partial(\phi)=\log\rho(M_{\partial,\phi}).
\]
Consequently,
\[
Z_m(\partial\mathsf{P},\phi)\ll
m^{q_\partial-1}
e^{-m(P_X(\phi)-\log\rho(M_{\partial,\phi}))}
\]
for all $m$, and the polynomial factor $m^{q_\partial-1}$ is exactly the
standard reducible Perron--Frobenius correction from
Lemma~\ref{lem:reducible-pf-dag-bound}.  There are matching lower bounds on
every visible residue class of a maximal-weight component.  If the
residue-complete hypothesis from
Theorem~\ref{thm:weighted-boundary-pressure}\emph{(3)} holds, then
\[
Z_m(\partial\mathsf{P},\phi)\gg
e^{-m(P_X(\phi)-\log\rho(M_{\partial,\phi}))}
\]
and
\[
Z_m(\partial\mathsf{P},\phi)\ll
m^{q_\partial-1}
e^{-m(P_X(\phi)-\log\rho(M_{\partial,\phi}))}
\]
and
\[
\varepsilon_m(\mathsf{P};\mu_\phi)\gg
e^{-m(P_X(\phi)-\log\rho(M_{\partial,\phi}))}
\]
and
\[
\varepsilon_m(\mathsf{P};\mu_\phi)\ll
m^{q_\partial-1}
e^{-m(P_X(\phi)-\log\rho(M_{\partial,\phi}))}
\]
for all large $m$.
If, in addition, the chosen presentation is no-maximal-chain, equivalently
$q_\partial=1$, then
\[
Z_m(\partial\mathsf{P},\phi)\asymp
e^{-m(P_X(\phi)-\log\rho(M_{\partial,\phi}))}
\]
and
\[
\varepsilon_m(\mathsf{P};\mu_\phi)\asymp
e^{-m(P_X(\phi)-\log\rho(M_{\partial,\phi}))}
\]
for all large $m$.
\end{corollary}

\begin{proof}
By Proposition~\ref{prop:finite-boundary-vs-recurrent-pressure}, the finite
boundary partition sum is exactly the finite refined-path partition sum, so
every admissible boundary path $\gamma=r_0\cdots r_{m-1}$ of length $m$
contributes exactly the weight
\[
e^{-mP_X(\phi)}e^{F(r_0)}\prod_{k=0}^{m-2}(M_{\partial,\phi})_{r_kr_{k+1}}
=e^{-mP_X(\phi)}e^{\sum_{k=0}^{m-1}F(r_k)},
\]
so
\[
Z_m(\partial\mathsf{P},\phi)
=e^{-mP_X(\phi)}\alpha_{\partial,\phi}M_{\partial,\phi}^{\,m-1}\mathbf{1}
\qquad(m\ge 1).
\]
If $\mathfrak{C}_\partial=\varnothing$, then the visible refined graph is
acyclic, so $M_{\partial,\phi}$ is nilpotent.  The displayed exact identity
therefore gives $Z_m(\partial\mathsf{P},\phi)=0$ for all sufficiently large
$m$, as claimed.

Assume now that $\mathfrak{C}_\partial\neq\varnothing$.  For each
$C\in\mathfrak{C}_\partial$, every state of $C$ lies on an infinite boundary
path, so the restriction of $F$ to $C$ is exactly the one-cylinder potential
for $\phi_\partial$ on $\Sigma_C$.  Hence standard one-step
Perron--Frobenius thermodynamic formalism on $\Sigma_C$ gives
\[
P_C(\phi)=\log\rho\bigl((M_{\partial,\phi})|_C\bigr)
\]
\cite[Ch.~1]{Bowen1975},\cite[Ch.~9]{WaltersErgodicTheory}.  Transient states
do not change the spectral radius, so
\[
P_\partial(\phi)=\max_{C\in\mathfrak{C}_\partial}P_C(\phi)
=\log\rho(M_{\partial,\phi}).
\]
Substituting this identity into
Theorem~\ref{thm:weighted-boundary-pressure} and
Corollary~\ref{cor:weighted-boundary-transfer} yields the displayed upper and
lower bounds, including the residue-class lower bounds and the
no-maximal-chain ($q_\partial=1$) sharpening.
\end{proof}

\begin{remark}[Primitive regularity is the all-length special case]
\label{rem:regularity-excludes-transients}
Theorem~\ref{thm:weighted-boundary-pressure} no longer requires the boundary
graph to be primitive.  Its invariant content is componentwise: the boundary
pressure is the maximum pressure over the reachable recurrent boundary
components, and every visible maximal component contributes a matching lower
bound on its own residue class.  Two separate reducible-graph effects remain.
First, a directed chain of several maximal-pressure components can produce the
polynomial upper factor $m^{q_\partial-1}$.  Second, periodicity can obstruct
all-length lower bounds.  What the no-maximal-chain condition removes is the
polynomial loss on the upper side, while primitivity, or more generally the
residue-complete hypothesis in
Theorem~\ref{thm:weighted-boundary-pressure}\emph{(3)}, adds is an
\emph{all-length} lower bound.  This is exactly the point at which transient
entry states and periodic cores matter.  By
Lemma~\ref{lem:reducible-pf-dag-bound}, the factor $m^{q_\partial-1}$ is the
standard Frobenius-normal-form correction coming from chains of maximal
spectral-radius components in the visible condensation DAG.  Theorem~\ref{ex:golden-mean-gibbs-weighted}
falls under the theorem componentwise, but its exact prefactor is still most
transparent when computed directly from the weighted transfer matrix of the
transient-plus-periodic graph.
\end{remark}

\noindent
Proposition~\ref{prop:weighted-matrix-transient-core} below is the
$q_\alpha(M)=1$ specialization of
Lemma~\ref{lem:reducible-pf-dag-bound} in the case of a unique terminal
maximal component; it is the form needed for the explicit golden-mean
calculation.

\begin{proposition}[Weighted matrix growth with a transient entry and periodic terminal core]
\label{prop:weighted-matrix-transient-core}
Let $M$ be a nonnegative $d\times d$ matrix, let $\alpha\in[0,\infty)^d$ be a
nonzero row vector, let $\mathbf{1}$ be the all-ones column vector, and write
$G(M)$ for the directed graph with an edge $r\to s$ whenever $M_{rs}>0$.
Assume that among the strongly connected components of $G(M)$ reachable from
$\operatorname{supp}(\alpha)$ there is a unique component $C$ with
\[
\rho(M_C)=\rho(M),
\]
and that $C$ is terminal in the reachable subgraph of $G(M)$.  Then
\[
\alpha M^n\mathbf{1}\asymp \rho(M)^n
\qquad(n\ge 0).
\]
No primitivity is assumed for $C$.
\end{proposition}

\begin{proof}
Reorder the reachable states so that the states outside $C$ come first and the
states of $C$ come last.  With this ordering,
\[
M=
\begin{pmatrix}
N&B\\
0&M_C
\end{pmatrix},
\]
where $\rho(N)<\rho(M)$ because $C$ is the unique reachable component with
maximal spectral radius.  Write $\alpha=(\alpha_N,\alpha_C)$ accordingly.
Standard spectral-radius bounds give constants $K_N,K_C>0$ and
$0<\tau<\rho(M)$ such that
\[
\|N^n\|\le K_N\tau^n,
\qquad
\|M_C^n\|\le K_C\rho(M)^n
\qquad(n\ge 0).
\]
Expanding the block power yields
\[
\alpha M^n\mathbf{1}
=\alpha_NN^n\mathbf{1}
+\sum_{k=0}^{n-1}\alpha_NN^kBM_C^{n-1-k}\mathbf{1}
+\alpha_CM_C^n\mathbf{1}.
\]
The first term is $O(\tau^n)$, the last term is $O(\rho(M)^n)$, and the sum
is bounded by
\[
\sum_{k=0}^{n-1}O(\tau^k\rho(M)^{n-1-k})
=O(\rho(M)^n)
\]
because $\tau/\rho(M)<1$.  Hence $\alpha M^n\mathbf{1}\ll \rho(M)^n$.

For the lower bound, choose $a\in C$, $m_0\ge 0$, and $c_0>0$ such that
$(\alpha M^{m_0})_a=c_0>0$; this is possible because $C$ is reachable from
$\operatorname{supp}(\alpha)$.  Let $p$ be the period of the irreducible
matrix $M_C$, and write the cyclic decomposition as
\[
C=C_0\sqcup\cdots\sqcup C_{p-1},
\qquad a\in C_0.
\]
For each residue class $s\in\{0,\dots,p-1\}$ choose a state $b_s\in C_s$ and a
path inside $C$ from $a$ to $b_s$ of length $s$; its weight is some
$d_s>0$.  The matrix $(M_C^p)|_{C_s}$ is primitive with spectral radius
$\rho(M)^p$, so Perron--Frobenius gives constants $e_s>0$ and $n_s$ such that
\[
e_{b_s}^\top M_C^{pn}\mathbf{1}\ge e_s\,\rho(M)^{pn}
\qquad(n\ge n_s).
\]
Therefore, for $n\ge n_s$,
\[
e_a^\top M_C^{pn+s}\mathbf{1}
\ge d_s\,e_{b_s}^\top M_C^{pn}\mathbf{1}
\ge d_se_s\,\rho(M)^{pn}
\asymp \rho(M)^{pn+s}.
\]
Since there are only finitely many residues, this gives a constant $c_1>0$
such that
\[
e_a^\top M_C^n\mathbf{1}\ge c_1\rho(M)^n
\]
for all sufficiently large $n$.  Hence
\[
\alpha M^{m_0+n}\mathbf{1}
\ge c_0\,e_a^\top M_C^n\mathbf{1}
\ge c_0c_1\,\rho(M)^n
\asymp \rho(M)^{m_0+n}.
\]
After absorbing finitely many small $n$ into the constants, the lower bound
holds for every $n\ge 0$.
\end{proof}

For completeness, Appendix~\ref{sub:nonuniform} records an auxiliary
filtration-level transfer principle for log-regular cylinder measures and a
non-stationary sharpness example showing that its polynomial factors are
genuine.  We now turn to the explicit symbolic computations.

\section{Explicit Computations}
\label{sec:examples}

\begin{example}[Finite-horizon events]
\label{ex:finite-horizon-event}
If $\mathsf{P}\in\mathcal{F}_r$ for some finite $r$, then
$\varepsilon_m(\mathsf{P};\mu)=0$ for all $m\ge r$.  Indeed, $\partial_m\mathsf{P}=\varnothing$
for $m\ge r$ since membership in $\mathsf{P}$ is determined by the first $r$
symbols.  Under Definition~\ref{def:boundary-cylinder}, the upper boundary
dimension is $0$ in this eventually empty case.
\end{example}

For the golden-mean example, let
\[
X:=\{x\in\{0,1\}^{\NN}:x_ix_{i+1}\ne 11\text{ for all }i\ge 0\}
\]
and let $\mu$ be its Parry measure.  The transition matrix is
\[
A=\begin{pmatrix}1&1\\1&0\end{pmatrix},
\qquad
\lambda_{\mathrm{cyl}}=\lambda_{\mathrm G}:=\frac{1+\sqrt{5}}{2}.
\]
For a finite word $a$, the cylinder $[a]$ is always understood as a cylinder
inside $X$.

\begin{lemma}[Explicit Parry cylinder bounds]
\label{lem:golden-parry-bounds}
For every admissible word $a=a_0\cdots a_{m-1}$ of length $m\ge 1$,
\[
\mu([a])=\frac{r_{a_0}r_{a_{m-1}}}{\lambda_{\mathrm G}^2+1}\,\lambda_{\mathrm G}^{-(m-1)},
\qquad r_0=\lambda_{\mathrm G},\quad r_1=1.
\]
Consequently
\[
\frac{1}{\sqrt{5}}\,\lambda_{\mathrm G}^{-m}
\le \mu([a])
\le \frac{\lambda_{\mathrm G}^2}{\sqrt{5}}\,\lambda_{\mathrm G}^{-m}.
\]
The positive transition probabilities are
\[
p_{00}^{(\mu)}=\lambda_{\mathrm G}^{-1},\qquad
p_{01}^{(\mu)}=\lambda_{\mathrm G}^{-2},\qquad
p_{10}^{(\mu)}=1.
\]
In particular, if $m\ge 1$ and $w$ is an admissible continuation of $a$ of
length $\ell$, then
\[
\mu([aw]\mid[a])\ge p_*^\ell,\qquad p_*:=\lambda_{\mathrm G}^{-2}.
\]
\end{lemma}

\begin{proof}
The vector $r=(\lambda_{\mathrm G},1)$ is a right and left Perron--Frobenius eigenvector
of $A$.  The Parry transition matrix is
$\bigl(p_{ij}^{(\mu)}\bigr)_{A_{ij}=1}$ with
$p_{ij}^{(\mu)}=A_{ij}r_j/(\lambda_{\mathrm G} r_i)$, and the stationary distribution is
$\pi_i=r_i^2/(r_0^2+r_1^2)$ \cite{Parry1964}.  Hence, for an admissible word,
\[
\mu([a])=\pi_{a_0}\prod_{j=0}^{m-2}p_{a_j a_{j+1}}^{(\mu)}
=\frac{r_{a_0}r_{a_{m-1}}}{\lambda_{\mathrm G}^2+1}\,\lambda_{\mathrm G}^{-(m-1)}.
\]
Since $1\le r_{a_0}r_{a_{m-1}}\le\lambda_{\mathrm G}^2$ and
$\lambda_{\mathrm G}^2+1=\sqrt{5}\,\lambda_{\mathrm G}$, the displayed cylinder bounds follow.  The
transition probabilities are obtained from the same formula for $p_{ij}^{(\mu)}$, and
the continuation estimate follows because every positive transition is at
least $p_*=\lambda_{\mathrm G}^{-2}$.
\end{proof}

Let
\[
Y:=\{x\in X:x_{2j}=0\text{ for all }j\ge 0,\quad
x_{2j+1}x_{2j+3}\ne 11\text{ for all }j\ge 0\}.
\]
A finite word $a$ is called $Y$-admissible if $[a]\cap Y\ne\varnothing$.  For
$x\in X\setminus Y$, define
\[
\rho(x):=\min\{m\ge 1:x_0\cdots x_{m-1}\text{ is not }Y\text{-admissible}\}.
\]
Define $\mathsf{E}$ to be the set of points whose first obstruction at time
$\rho(x)$ is an even coordinate equal to $1$, and include $Y$ in $\mathsf{E}$.

\begin{lemma}[Bounded first-obstruction continuations]
\label{lem:golden-bounded-continuations}
Let $a$ be a $Y$-admissible word of length $m\ge 1$.  There are admissible
continuations $u_{\mathrm e}$ and $u_{\mathrm o}$, each of length at most four,
such that every proper prefix of $au_{\mathrm e}$ and $au_{\mathrm o}$ extending
$a$ is still $Y$-admissible, while $au_{\mathrm e}$ first leaves the
$Y$-language by an even-coordinate obstruction and $au_{\mathrm o}$ first
leaves it by an odd-coordinate obstruction.  Moreover,
\[
\mu([au_{\mathrm e}]\mid[a])\ge p_*^4,\qquad
\mu([au_{\mathrm o}]\mid[a])\ge p_*^4.
\]
\end{lemma}

\begin{proof}
Let $\ell(a)\in\{\ast,0,1\}$ be the last visible odd-coordinate symbol of $a$,
with $\ell(a)=\ast$ when no odd coordinate has yet been seen.  For a
$Y$-admissible word all visible even coordinates are $0$, so the possible
continuations forcing the first obstruction can be written explicitly in the
following case table:
\[
\begin{array}{c|c|c|c}
\text{parity of }m & \ell(a) & u_{\mathrm e} & u_{\mathrm o}\\
\hline
m\text{ even} & 0 & 1 & 0101\\
m\text{ even} & 1 & 001 & 01\\
m\text{ odd} & \ast & 01 & 101\\
m\text{ odd} & 0 & 01 & 101\\
m\text{ odd} & 1 & 01 & 1
\end{array}
\]
Here the omitted pair $(m\text{ even},\ast)$ cannot occur because a word of
positive even length already contains an odd coordinate.

We check the table row by row.  For the even-obstruction words
$u_{\mathrm e}$, every newly inserted odd coordinate is $0$ until the final
even symbol $1$ appears, so every proper prefix remains $Y$-admissible and the
first violation is the displayed even coordinate equal to $1$.  In the row
$(m\text{ even},1)$ one must delay that $1$ by writing $001$, because writing
$1$ immediately at the even coordinate $m$ would create the forbidden adjacent
pair $11$ in $X$.

For the odd-obstruction words $u_{\mathrm o}$, every newly inserted even
coordinate is $0$, so no even obstruction occurs beforehand.  If
$\ell(a)=1$, the continuation $1$ (for $m$ odd) or $01$ (for $m$ even) makes
the next visible odd symbol equal to $1$, creating the first adjacent pair
$11$ in the odd subsequence immediately.  If $\ell(a)\in\{\ast,0\}$, the first
new odd symbol is set to $1$, which is still $Y$-admissible, and the second new
odd symbol is again set to $1$ after an intervening even $0$; this creates the
first odd obstruction and no earlier prefix leaves the $Y$-language.  Every
listed word is admissible in $X$, and the longest one has length four.

Lemma~\ref{lem:golden-parry-bounds}
gives conditional mass at least $p_*^r$ for a continuation of length
$r\le 4$, and $p_*^r\ge p_*^4$.
\end{proof}

\begin{lemma}[The skeleton has zero Parry measure]
\label{lem:golden-y-null}
The set $Y$ satisfies $\mu(Y)=0$.
\end{lemma}

\begin{proof}
Let $A_m$ be the union of all $Y$-admissible cylinders of length $m$.  Then
$Y=\bigcap_{m\ge 1}A_m$.  By Lemma~\ref{lem:golden-bounded-continuations}, for
each $Y$-admissible word $a$ of length $m\ge 1$ there is a subcylinder of
$[a]$ with conditional mass at least $p_*^4$ whose points have left the
$Y$-language by time $m+4$.  Hence
\[
\mu(A_{m+4}\cap[a])\le (1-p_*^4)\mu([a]).
\]
Summing over the $Y$-admissible words of length $m$ gives
$\mu(A_{m+4})\le(1-p_*^4)\mu(A_m)$.  Iterating from $m=1$,
\[
\mu(A_{1+4n})\le (1-p_*^4)^n\mu(A_1)\to 0.
\]
Since $Y\subseteq A_{1+4n}$ for every $n$, $\mu(Y)=0$.
\end{proof}

\begin{proposition}[Golden-mean boundary skeleton]
\label{prop:golden-boundary-skeleton}
Let $a=a_0\cdots a_{m-1}$ be a finite word of length $m$, and let
$\mathrm{odd}(a):=a_1a_3\cdots a_{2\lfloor m/2\rfloor-1}$, interpreted as the
empty word when $\lfloor m/2\rfloor=0$.  Then $a$ is
$Y$-admissible if and only if all even coordinates visible in $a$ are $0$ and
$\mathrm{odd}(a)$ is golden-mean admissible.  For $x\in X\setminus Y$, the
first exit time $\rho(x)$ has exactly one of the following two forms:
\[
\rho(x)=2j+1\text{ and }x_{2j}=1,
\qquad\text{or}\qquad
\rho(x)=2j+4\text{ and }x_{2j+1}=x_{2j+3}=1.
\]
Finally, for every positive length-$m$ cylinder $[a]$,
\[
[a]\in\partial_m\mathsf{E}\quad\Longleftrightarrow\quad a
\text{ is }Y\text{-admissible}.
\]
Consequently, with $F_0=0,F_1=1$,
\[
N_{2k}(\partial\mathsf{E})=N_{2k+1}(\partial\mathsf{E})=F_{k+2}
\qquad(k\ge 0).
\]
\end{proposition}

\begin{proof}
The characterization of $Y$-admissible words is immediate in one direction.
Conversely, if the visible even coordinates of $a$ are $0$ and
$\mathrm{odd}(a)$ has no adjacent $1$s, then extend $a$ by putting all future
even and odd coordinates equal to $0$; the extension lies in $Y$.

If $x\notin Y$, the first failure is either the first even coordinate carrying
$1$, which is seen at prefix length $2j+1$, or the first adjacent pair $11$ in
the odd-coordinate subsequence, whose second symbol is seen at prefix length
$2j+4$.  These prefix lengths have opposite parity, so the newly visible
obstruction is unique.

If $a$ is not $Y$-admissible, let $r\le m$ be the first prefix length at which
$a_0\cdots a_{r-1}$ is not $Y$-admissible.  The obstruction type at $r$ is
already fixed by $a$, and every point of $[a]$ has the same first obstruction;
therefore $[a]$ is contained either in $\mathsf{E}$ or in $\mathsf{E}^c$.
If $a$ is $Y$-admissible and $m\ge 1$, Lemma~\ref{lem:golden-bounded-continuations}
gives admissible continuations $u_{\mathrm e},u_{\mathrm o}$ with
$[au_{\mathrm e}]\subseteq\mathsf{E}$ and
$[au_{\mathrm o}]\subseteq\mathsf{E}^c$; under the Parry measure both
subcylinders have positive mass, so $[a]$ is a boundary cylinder.  For $m=0$, the empty cylinder
$X$ is boundary because the positive cylinders $[1]$ and $[0101]$ have first
even and first odd obstructions respectively.

For $m=2k$ or $m=2k+1$, the visible odd skeleton has length $k$.  The number
of golden-mean admissible words of length $k$ is $F_{k+2}$, by the recurrence
$b_k=b_{k-1}+b_{k-2}$ with $b_0=1$ and $b_1=2$.  The boundary equivalence gives
the displayed formula for $N_m(\partial\mathsf{E})$.
\end{proof}

\begin{proposition}[Explicit follower-resolving observer for $\mathsf{E}$]
\label{prop:golden-finite-posterior}
The event $\mathsf{E}$ admits a finite-state posterior presentation with
respect to the Parry measure, in the sense of
Definition~\ref{def:finite-posterior}.  For a positive
$Y$-admissible word $a$, let $\tau(a)\in\{\mathrm{o},\mathrm{e}\}$ denote the
parity of the next coordinate to be read, let $\ell(a)\in\{\ast,0,1\}$ denote
the last visible odd-coordinate symbol (with $\ell(a)=\ast$ if none has yet
been seen), and let $f(a)\in\{0,1\}$ be the current golden-mean follower
state, equivalently the last symbol of $a$.  The reachable ambiguous triples
are
\[
u=(\mathrm{o},\ast,0),\qquad
e_0=(\mathrm{e},0,0),\qquad
e_1=(\mathrm{e},1,1),\qquad
o_0=(\mathrm{o},0,0),\qquad
o_1=(\mathrm{o},1,0).
\]
Let $q_\emptyset$ be the initial state, and let
\[
\mathcal{S}_1=\{q_{1,0},q_{1,1}\},
\qquad
\mathcal{S}_0=\{q_{0,0},q_{0,1}\}
\]
record whether the first obstruction is even or odd together with the current
follower state in $X$.  For every admissible word $a$, define
\[
q(a):=
\begin{cases}
q_{1,a_{|a|-1}}, & [a]\subseteq \mathsf{E},\\
q_{0,a_{|a|-1}}, & [a]\subseteq \mathsf{E}^c,\\
(\tau(a),\ell(a),f(a)), & a\text{ is }Y\text{-admissible and }|a|\ge 1,
\end{cases}
\]
and set $q(\emptyset):=q_\emptyset$.  The defined one-step updates are
\[
\begin{array}{c|cc}
\text{state} & 0 & 1\\
\hline
q_\emptyset & u & q_{1,1}\\
u & e_0 & e_1\\
e_0 & o_0 & q_{1,1}\\
e_1 & o_1 & \text{undefined}\\
o_0 & e_0 & e_1\\
o_1 & e_0 & q_{0,1}\\
q_{1,0} & q_{1,0} & q_{1,1}\\
q_{1,1} & q_{1,0} & \text{undefined}\\
q_{0,0} & q_{0,0} & q_{0,1}\\
q_{0,1} & q_{0,0} & \text{undefined}
\end{array}
\]
and no other transitions are defined.  These updates are follower-resolving,
the sets $\mathcal{S}_0$ and $\mathcal{S}_1$ are absorbing under every defined transition, and the
uniform resolving-window hypothesis of Theorem~\ref{thm:finite-state-posterior}
holds with $R=4$.

For the reachable refined ambiguous states, ordered as
$(u,e_0,e_1,o_0,o_1)$, the ambiguous transition matrix is
\[
B_{\mathsf{E}}:=\widetilde{B}_{\mathsf{E}}
=\begin{pmatrix}
0&1&1&0&0\\
0&0&0&1&0\\
0&0&0&0&1\\
0&1&1&0&0\\
0&1&0&0&0
\end{pmatrix}.
\]
Its first row is transient.  On the recurrent core $(e_0,e_1,o_0,o_1)$ one has
\[
C_{\mathsf{E}}
=\begin{pmatrix}
0&0&1&0\\
0&0&0&1\\
1&1&0&0\\
1&0&0&0
\end{pmatrix},
\qquad
C_{\mathsf{E}}^2
=\begin{pmatrix}
1&1&0&0\\
1&0&0&0\\
0&0&1&1\\
0&0&1&0
\end{pmatrix}
=\begin{pmatrix}
F&0\\
0&F
\end{pmatrix},
\quad
F=\begin{pmatrix}1&1\\ 1&0\end{pmatrix}.
\]
Hence
\[
\rho(B_{\mathsf{E}})=\rho(C_{\mathsf{E}})
=\sqrt{\rho(F)}
=\lambda_{\mathrm G}^{1/2}.
\]
Consequently Theorem~\ref{thm:finite-state-posterior} gives
\[
\Lambda_\partial(\mathsf{E})=\lambda_{\mathrm G}^{1/2},
\qquad
\delta_\partial(\mathsf{E})=\frac12.
\]
\end{proposition}

\begin{proof}
If a nonempty admissible word $a$ is $Y$-admissible, then
Proposition~\ref{prop:golden-boundary-skeleton} shows that $[a]$ is a boundary
cylinder.  In that case the future behaviour of the first obstruction is
determined exactly by three pieces of prefix data: the parity of the next
coordinate, the last visible odd-coordinate symbol, and the current follower
state in the golden-mean SFT.  These are precisely the recorded coordinates of
$u,e_0,e_1,o_0,o_1$, and the transition table above is just the explicit case
split on whether the next symbol preserves $Y$-admissibility, creates the
first even obstruction, or creates the first odd obstruction.

If $a$ is not $Y$-admissible, then
Proposition~\ref{prop:golden-boundary-skeleton} identifies a unique first
obstruction, which is either even or odd.  Therefore $[a]$ is contained in
either $\mathsf{E}$ or $\mathsf{E}^c$, and $q(a)$ records this decided status
together with the current follower state $a_{|a|-1}$.  Because each state
stores the current follower class, the admissible one-symbol continuations
depend only on the state, so Definition~\ref{def:finite-posterior}(2) holds.
The same definition of $q(a)$ also gives
Definition~\ref{def:finite-posterior}(4): the ambiguous states are exactly the
boundary cylinders, while the decided states lie entirely in $\mathsf{E}$ or
in $\mathsf{E}^c$.

The absorbing transitions on $\mathcal{S}_0$ and $\mathcal{S}_1$ preserve the status of the first
obstruction, so Definition~\ref{def:finite-posterior}(3) holds.  For the
uniform resolving window, Lemma~\ref{lem:golden-bounded-continuations} gives,
from every $Y$-admissible cylinder, admissible continuations of length at most
four that force an even first obstruction and an odd first obstruction.  Those
continuation recipes depend only on the same parity and last-odd-symbol data
recorded in the ambiguous states above, so the hypothesis of
Theorem~\ref{thm:finite-state-posterior} holds with $R=4$.

The displayed matrix $\widetilde{B}_{\mathsf{E}}$ is read directly from the
ambiguous part of the transition table.  In this example each ambiguous
observer state already records the current follower symbol, so the theorem's
refined ambiguity construction adds no extra splitting and the refined matrix
is exactly the displayed one, namely $B_{\mathsf{E}}$.  Its first state $u$
has no incoming ambiguous edge, so it is transient and does not affect the
spectral radius.  The
computation of $C_{\mathsf{E}}^2$ is immediate, and the Fibonacci matrix $F$
has spectral radius $\lambda_{\mathrm G}$.  Therefore
$\rho(B_{\mathsf{E}})=\rho(C_{\mathsf{E}})=\lambda_{\mathrm G}^{1/2}$.  The
final display now follows from Theorem~\ref{thm:finite-state-posterior}.
\end{proof}

\begin{lemma}[Diagonalisation of the golden transfer matrix]
\label{lem:golden-transfer-diagonalization}
Let $a=\lambda_{\mathrm G}^{-1}$ and
\[
B=\begin{pmatrix}a^2&a^3\\ a&0\end{pmatrix},\qquad
v=\frac1{\sqrt{5}}(1,a),\qquad
w=\begin{pmatrix}a^2/2\\[2pt]1/2\end{pmatrix}.
\]
Define
\[
r_+=\begin{pmatrix}1\\ 1\end{pmatrix},
\qquad
r_-=\begin{pmatrix}-a^2\\ 1\end{pmatrix}.
\]
Then $Br_+=ar_+$ and $Br_-=-a^3r_-$, so
\[
B=P\begin{pmatrix}a&0\\ 0&-a^3\end{pmatrix}P^{-1},
\qquad
P=\begin{pmatrix}1&-a^2\\ 1&1\end{pmatrix}.
\]
Moreover,
\[
w=\frac{a^2}{1+a^2}\,r_+ + \frac{a}{2(1+a^2)}\,r_-,
\]
and for every $n\ge 0$,
\[
vB^n w=\frac15\,a^n-\frac1{10}\,(-a^3)^{n+1}.
\]
\end{lemma}

\begin{proof}
Using $a^2+a=1$, one checks
\[
B\begin{pmatrix}1\\ 1\end{pmatrix}
=\begin{pmatrix}a^2+a^3\\ a\end{pmatrix}
=\begin{pmatrix}a\\ a\end{pmatrix},
\qquad
B\begin{pmatrix}-a^2\\ 1\end{pmatrix}
=\begin{pmatrix}0\\ -a^3\end{pmatrix}
=-a^3\begin{pmatrix}-a^2\\ 1\end{pmatrix}.
\]
Thus $r_\pm$ are eigenvectors with eigenvalues $a$ and $-a^3$ respectively,
and the displayed diagonalisation follows because $\det P=1+a^2>0$.

To decompose $w$, write $w=\alpha r_+ + \beta r_-$.  Then
\[
\alpha-a^2\beta=\frac{a^2}{2},
\qquad
\alpha+\beta=\frac12.
\]
Subtracting gives $(1+a^2)\beta=(1-a^2)/2=a/2$, hence
$\beta=\frac{a}{2(1+a^2)}$ and
$\alpha=\frac12-\beta=\frac{a^2}{1+a^2}$.

Now
\[
vr_+=\frac{1+a}{\sqrt{5}}=\frac{\lambda_{\mathrm G}}{\sqrt{5}},
\qquad
vr_-=\frac{a-a^2}{\sqrt{5}}=\frac{a^3}{\sqrt{5}},
\]
since $1-a=a^2$.  Therefore
\[
vB^n w
=\alpha\,vr_+\,a^n+\beta\,vr_-\,(-a^3)^n.
\]
Using $\lambda_{\mathrm G} a=1$ and
$1+a^2=1+\lambda_{\mathrm G}^{-2}=\sqrt{5}\,\lambda_{\mathrm G}^{-1}=\sqrt{5}\,a$, this becomes
\[
vB^n w
=\frac{a}{\sqrt{5}(1+a^2)}\,a^n
+\frac{a^4}{2\sqrt{5}(1+a^2)}\,(-a^3)^n
=\frac15\,a^n+\frac{a^3}{10}\,(-a^3)^n
=\frac15\,a^n-\frac1{10}\,(-a^3)^{n+1}.
\]
\end{proof}

\begin{proposition}[Exact parity-resolved scan error in the golden example]
\label{prop:golden-parity-sharp}
For every $k\ge 1$,
\[
\gamma_k
:=\frac{\lambda_{\mathrm G}}{5}\,\lambda_{\mathrm G}^{-k}
-\frac{1}{10}\,(-\lambda_{\mathrm G}^{-3})^k.
\]
\[
\varepsilon_{2k}(\mathsf{E};\mu)
=\varepsilon_{2k+1}(\mathsf{E};\mu)
=\gamma_k.
\]
Consequently
\[
\lim_{k\to\infty}\lambda_{\mathrm G}^k\varepsilon_{2k}(\mathsf{E};\mu)
=\lim_{k\to\infty}\lambda_{\mathrm G}^k\varepsilon_{2k+1}(\mathsf{E};\mu)
=\frac{\lambda_{\mathrm G}}{5}.
\]
Equivalently, with normalization by $\lambda_{\mathrm G}^{m/2}$,
\[
\varepsilon_{2k}(\mathsf{E};\mu)\sim \frac{\lambda_{\mathrm G}}{5}\lambda_{\mathrm G}^{-k},
\qquad
\varepsilon_{2k+1}(\mathsf{E};\mu)\sim
\frac{\lambda_{\mathrm G}^{3/2}}{5}\lambda_{\mathrm G}^{-(2k+1)/2}.
\]
\end{proposition}

\begin{proof}
Put $a:=\lambda_{\mathrm G}^{-1}$ and $b:=\lambda_{\mathrm G}^{-2}=a^2$.  Suppose the process has not
yet left the $Y$-language, and let the most recent odd-coordinate symbol be
$y\in\{0,1\}$.  Let $e_y$ be the conditional probability that the first future
obstruction is even when the next coordinate to be read is even, and let $o_y$
be the same probability when the next coordinate to be read is odd.  The Parry
transition probabilities in Lemma~\ref{lem:golden-parry-bounds} give
\[
e_0=b+ao_0,\qquad e_1=o_1,\qquad
o_0=ae_0+be_1,\qquad o_1=ae_0.
\]
Solving,
\[
e_0=\frac{\lambda_{\mathrm G}}{2},\qquad e_1=\frac12,\qquad
o_0=\frac{1+b}{2},\qquad o_1=\frac12.
\]
Thus the local classification errors at an even-read state are $b/2$ and
$1/2$ for last odd symbol $0$ and $1$, respectively.  At an odd-read state the
local errors are $a/2$ and $1/2$; the corresponding length-$(2k+1)$ cylinder
masses are obtained from the length-$2k$ masses by multiplying by $a$ when the
last odd symbol is $0$ and by $1$ when it is $1$.  Hence the odd-depth
contribution has the same weights, since $a(a/2)=a^2/2=b/2$.

Let $s_k=(s_k(0),s_k(1))$ be the row vector of total Parry masses of
$Y$-admissible length-$2k$ cylinders whose last visible odd symbol is $0$ or
$1$.  Since $\pi_0=\lambda_{\mathrm G}/\sqrt{5}$,
\[
s_k=vB^{k-1},\qquad
v=\frac1{\sqrt{5}}(1,a),\qquad
B=\begin{pmatrix}a^2&a^3\\ a&0\end{pmatrix}
\quad(k\ge 1).
\]
Indeed, the initial vector is
\[
s_1=\bigl(\mu([00]),\mu([01])\bigr)
=\bigl(\pi_0 p_{00}^{(\mu)},\pi_0 p_{01}^{(\mu)}\bigr)
=\frac1{\sqrt{5}}(1,a)=v.
\]
If a $Y$-admissible cylinder contributing to $s_k(0)$ has last visible odd
symbol $0$, then the next even symbol must be $0$, after which the next odd
symbol may be $0$ or $1$.  By Lemma~\ref{lem:golden-parry-bounds}, the
corresponding two-step cylinder-mass multipliers are
\[
p_{00}^{(\mu)}p_{00}^{(\mu)}=a^2,
\qquad
p_{00}^{(\mu)}p_{01}^{(\mu)}=a^3.
\]
If the last visible odd symbol is $1$, then the next even symbol is again
forced to be $0$, and $Y$-admissibility forbids the next odd symbol from being
$1$; hence the only allowed return is $1\to 0$, with multiplier
\[
p_{10}^{(\mu)}p_{00}^{(\mu)}=a.
\]
Therefore
\[
s_{k+1}(0)=a^2s_k(0)+as_k(1),\qquad
s_{k+1}(1)=a^3s_k(0),
\]
which is exactly the matrix recurrence $s_{k+1}=s_kB$.  With
\[
w=\begin{pmatrix}b/2\\[2pt]1/2\end{pmatrix},
\]
the preceding paragraph gives
$\varepsilon_{2k}(\mathsf{E};\mu)=\varepsilon_{2k+1}(\mathsf{E};\mu)
=vB^{k-1}w$.  Lemma~\ref{lem:golden-transfer-diagonalization} applied with
$n=k-1$ gives
\[
vB^{k-1}w=\frac{\lambda_{\mathrm G}}{5}a^k-\frac{1}{10}(-a^3)^k.
\]
Substituting $a=\lambda_{\mathrm G}^{-1}$ gives the displayed formula and its two
asymptotic normalizations.
\end{proof}

\begin{theorem}[Golden-mean Parry law: refined boundary matrix, ambiguity exponent, and exact scan error]
\label{ex:golden-mean-specialization}
Let
\[
u,\ E_0,\ E_1,\ O_0,\ O_1
\]
be the ambiguous observer states from
Proposition~\ref{prop:golden-finite-posterior}, where $E_y$ means that the
next coordinate to be read is even and the last visible odd symbol is $y$,
while $O_y$ means that the next coordinate to be read is odd and the last
visible odd symbol is $y$.  Then the event $\mathsf{E}$ belongs to
$\mathcal{F}_\infty$, is not decided by any finite prefix, and satisfies
\[
N_{2k}(\partial\mathsf{E})=N_{2k+1}(\partial\mathsf{E})=F_{k+2}
\qquad(k\ge 0).
\]
Its refined ambiguous matrix is
\[
B_{\mathsf{E}}=\widetilde{B}_{\mathsf{E}}
=\begin{pmatrix}
0&1&1&0&0\\
0&0&0&1&0\\
0&0&0&0&1\\
0&1&1&0&0\\
0&1&0&0&0
\end{pmatrix},
\]
with recurrent core
\[
C_{\mathsf{E}}
=\begin{pmatrix}
0&0&1&0\\
0&0&0&1\\
1&1&0&0\\
1&0&0&0
\end{pmatrix},
\qquad
C_{\mathsf{E}}^2
=\begin{pmatrix}
F&0\\
0&F
\end{pmatrix},
\qquad
F=\begin{pmatrix}1&1\\ 1&0\end{pmatrix}.
\]
Consequently
\[
\rho(B_{\mathsf{E}})=\rho(C_{\mathsf{E}})=\lambda_{\mathrm G}^{1/2},
\qquad
\Lambda_\partial(\mathsf{E})=\lambda_{\mathrm G}^{1/2},
\qquad
\delta_\partial(\mathsf{E})=\frac12.
\]
The resolving window is $R=4$, and every boundary cylinder satisfies
\[
\min\{\mu(\mathsf{E}\mid[a]),1-\mu(\mathsf{E}\mid[a])\}
\ge p_*^4=\lambda_{\mathrm G}^{-8}
\qquad([a]\in\partial_m\mathsf{E},\ m\ge 1).
\]
Finally, for every $k\ge 1$,
\[
\varepsilon_{2k}(\mathsf{E};\mu)
=\varepsilon_{2k+1}(\mathsf{E};\mu)
=\frac{\lambda_{\mathrm G}}{5}\,\lambda_{\mathrm G}^{-k}
-\frac{1}{10}\,(-\lambda_{\mathrm G}^{-3})^k,
\]
so
\[
\varepsilon_m(\mathsf{E};\mu)\asymp \lambda_{\mathrm G}^{-m/2},
\qquad
\lim_{k\to\infty}\lambda_{\mathrm G}^k\varepsilon_{2k}(\mathsf{E};\mu)
=\lim_{k\to\infty}\lambda_{\mathrm G}^k\varepsilon_{2k+1}(\mathsf{E};\mu)
=\frac{\lambda_{\mathrm G}}{5}.
\]
\end{theorem}

\begin{proof}
The set $\mathsf{E}$ is the union of the cylinders on which a first even
obstruction occurs together with the closed set $Y$.  Since the one-sided SFT
Borel sigma algebra is generated by cylinders, $\mathsf{E}\in\mathcal{F}_\infty$.
Lemma~\ref{lem:golden-y-null} gives $\mu(Y)=0$.  The event is not decided by
any finite prefix because the all-zero length-$m$ cylinder is $Y$-admissible
for every $m\ge 1$, hence belongs to $\partial_m\mathsf{E}$ by
Proposition~\ref{prop:golden-boundary-skeleton}.

Proposition~\ref{prop:golden-boundary-skeleton} identifies the boundary
cylinders exactly with the $Y$-admissible cylinders and yields the boundary
count formula.  Proposition~\ref{prop:golden-finite-posterior} gives the
finite-state posterior presentation, the matrices $B_{\mathsf{E}}$ and
$C_{\mathsf{E}}$, the resolving window $R=4$, and the identity
\[
\rho(B_{\mathsf{E}})=\rho(C_{\mathsf{E}})=\lambda_{\mathrm G}^{1/2}.
\]
Theorem~\ref{thm:finite-state-posterior} therefore gives
\[
\Lambda_\partial(\mathsf{E})=\lambda_{\mathrm G}^{1/2},
\qquad
\delta_\partial(\mathsf{E})=\frac12.
\]
The same finite-state presentation is higher-block invariant by
Proposition~\ref{prop:higher-block-invariance}.

Lemma~\ref{lem:golden-bounded-continuations} gives, from every boundary
cylinder, admissible even-obstruction and odd-obstruction continuations of
length at most four.  Each has conditional mass at least $p_*^4$, so this
proves the thickness estimate.  Proposition~\ref{prop:golden-parity-sharp} gives the
exact parity-resolved scan-error formula and the limit.  The asymptotic
$\varepsilon_m(\mathsf{E};\mu)\asymp \lambda_{\mathrm G}^{-m/2}$ is immediate
from that formula.
\end{proof}

Theorem~\ref{ex:golden-mean-specialization} packages the whole Parry-side
computation into one citable statement: the same refined observer graph yields
the exact boundary counts, the ambiguity exponent
$\delta_\partial(\mathsf{E})=1/2$, and the exact parity-resolved scan-error
law.

\begin{theorem}[Golden-mean Gibbs law: exact boundary matrix formula and pressure gap]
\label{ex:golden-mean-gibbs-weighted}
Keep the same golden-mean shift $X$, event $\mathsf{E}$, and refined
ambiguous core states $(E_0,E_1,O_0,O_1)$ as in
Theorem~\ref{ex:golden-mean-specialization}.  For $t\in\mathbb{R}$ define the
locally constant potential
\[
\varphi_t(x):=t\,\mathbf{1}_{\{x_0=0\}},
\]
let $\mu_t$ be its equilibrium state, and write
\[
M_X(t)
=\begin{pmatrix}
e^t&1\\
e^t&0
\end{pmatrix},
\qquad
\lambda_t:=\rho(M_X(t))
=\frac{e^t+\sqrt{e^{2t}+4e^t}}{2}.
\]
Then
\[
P_X(\varphi_t)=\log\lambda_t.
\]
In the state order $(E_0,E_1,O_0,O_1)$, the weighted boundary matrix is
\[
M_{\partial,t}
=\begin{pmatrix}
0&0&e^t&0\\[2pt]
0&0&0&e^t\\[2pt]
e^t&1&0&0\\[2pt]
e^t&0&0&0
\end{pmatrix},
\qquad
M_{\partial,t}^2
=\operatorname{diag}\!\bigl(e^tM_X(t),\,e^tM_X(t)\bigr).
\]
Hence
\[
\rho(M_{\partial,t})=e^{t/2}\lambda_t^{1/2},
\qquad
P_\partial(\varphi_t)=\log\rho(M_{\partial,t})
=\frac12\bigl(t+\log\lambda_t\bigr),
\]
and therefore
\[
P_X(\varphi_t)-P_\partial(\varphi_t)
=\frac12\bigl(\log\lambda_t-t\bigr)
=\frac12\log\!\left(\frac{1+\sqrt{1+4e^{-t}}}{2}\right).
\]
Also
\[
Z_1(\partial\mathsf{E},\varphi_t)=e^{t-P_X(\varphi_t)}.
\]
For $m\ge 2$, with
\[
\alpha_t:=(e^{2t},e^t,0,0),
\]
the weighted boundary partition sum is exactly
\[
Z_m(\partial\mathsf{E},\varphi_t)
=e^{-mP_X(\varphi_t)}\alpha_t M_{\partial,t}^{\,m-2}\mathbf{1}.
\]
Consequently
\[
Z_m(\partial\mathsf{E},\varphi_t)\asymp
\exp\!\Bigl(-m\bigl(P_X(\varphi_t)-P_\partial(\varphi_t)\bigr)\Bigr),
\qquad
\varepsilon_m(\mathsf{E};\mu_t)\asymp
\exp\!\Bigl(-m\bigl(P_X(\varphi_t)-P_\partial(\varphi_t)\bigr)\Bigr).
\]
\end{theorem}

\begin{proof}
Because $\varphi_t$ is H\"older on the mixing one-step SFT $X$, Bowen's Gibbs
estimate on cylinders \cite{Bowen1975,WaltersErgodicTheory} gives a constant
$c_t^->0$ such that every admissible cylinder $[a]$ satisfies
\[
\mu_t([a])\ge c_t^-\exp\!\bigl(\sup_{x\in[a]}S_{|a|}\varphi_t(x)-|a|P_X(\varphi_t)\bigr)>0.
\]
Hence $\mu_t$ has full support.  Moreover, $\mathsf{E}$
satisfies the continuation-robust criterion from
Proposition~\ref{prop:full-support-invariant-decisive-languages}: if $a$ is
not $Y$-admissible, then Proposition~\ref{prop:golden-boundary-skeleton}
already gives either $[a]\subseteq\mathsf{E}$ or
$[a]\subseteq\mathsf{E}^c$, while if $a$ is $Y$-admissible, then
Lemma~\ref{lem:golden-bounded-continuations} provides admissible
continuations $u_{\mathrm e},u_{\mathrm o}$ with
$[au_{\mathrm e}]\subseteq\mathsf{E}$ and
$[au_{\mathrm o}]\subseteq\mathsf{E}^c$.  Therefore every full-support measure
on $X$ induces the same decisive languages for $\mathsf{E}$, so the observer
from Proposition~\ref{prop:golden-finite-posterior} is also a finite-state
posterior presentation for $(\mathsf{E},\mu_t)$ with the same resolving
window $R=4$.

The ambient transfer matrix is the displayed matrix $M_X(t)$, so
\[
P_X(\varphi_t)=\log\lambda_t,
\qquad
\lambda_t=\frac{e^t+\sqrt{e^{2t}+4e^t}}{2}.
\]
To compute the boundary pressure explicitly, work on the recurrent boundary
core with states $E_0,E_1,O_0,O_1$ from
Theorem~\ref{ex:golden-mean-specialization}.  For a target state
$s\in\{E_0,E_1,O_0,O_1\}$, let $\sigma(s)$ be the newly read symbol attached
to $s$:
\[
\sigma(E_0)=0,\qquad \sigma(E_1)=1,\qquad \sigma(O_0)=0,\qquad \sigma(O_1)=0.
\]
If $C_{\mathsf{E}}$ is the unweighted adjacency matrix of this core, define
\[
\bigl(M_{\partial,t}\bigr)_{rs}
:=(C_{\mathsf{E}})_{rs}\exp\!\bigl(t\,\mathbf{1}_{\{\sigma(s)=0\}}\bigr).
\]
This gives the displayed matrix $M_{\partial,t}$, and a direct multiplication
shows
\[
M_{\partial,t}^2=\operatorname{diag}\!\bigl(e^tM_X(t),\,e^tM_X(t)\bigr).
\]
Therefore
\[
\rho(M_{\partial,t})=\sqrt{e^t\lambda_t}=e^{t/2}\lambda_t^{1/2},
\qquad
P_\partial(\varphi_t)=\frac12\bigl(t+\log\lambda_t\bigr),
\]
which yields the displayed pressure-gap formula.

For $m\ge 2$, every boundary word begins with $0$ and enters the recurrent
core after one more symbol.  Since $\varphi_t$ is constant on cylinders, the
initial two-symbol weights are exactly the entries of
$\alpha_t=(e^{2t},e^t,0,0)$, so
\[
Z_m(\partial\mathsf{E},\varphi_t)
=e^{-mP_X(\varphi_t)}\alpha_t M_{\partial,t}^{\,m-2}\mathbf{1}.
\]
Set
\[
B_t:=e^tM_X(t)
=\begin{pmatrix}
e^{2t}&e^t\\
e^{2t}&0
\end{pmatrix},
\qquad
w_t:=(e^{2t},e^t).
\]
Then $M_{\partial,t}^2=\operatorname{diag}(B_t,B_t)$, so for every $n\ge 0$,
\[
\alpha_tM_{\partial,t}^{2n}\mathbf{1}=w_tB_t^n\mathbf{1},
\qquad
\alpha_tM_{\partial,t}^{2n+1}\mathbf{1}=e^t\,w_tB_t^n\mathbf{1}.
\]
Since $B_t$ is a positive multiple of the primitive matrix $M_X(t)$,
Perron--Frobenius theory gives constants $c_-,c_+>0$ such that
\[
c_-(e^t\lambda_t)^n
\le
w_tB_t^n\mathbf{1}
\le
c_+(e^t\lambda_t)^n
\qquad(n\ge 0).
\]
Substituting into the exact formula for $Z_m$ gives
\[
Z_m(\partial\mathsf{E},\varphi_t)\asymp
\exp\!\Bigl(-m\bigl(P_X(\varphi_t)-P_\partial(\varphi_t)\bigr)\Bigr).
\]
The same resolving words as in Theorem~\ref{ex:golden-mean-specialization}
still force a decision within four additional symbols from every boundary
state.  Applying
Lemma~\ref{lem:weighted-thickness-without-components} to the finite
$\mu_t$-resolving presentation just identified, and using the cylinder lower
bound established at the start of this proof, gives a
constant $\theta_t>0$ such that
\[
\min\{\mu_t(\mathsf{E}\mid[a]),1-\mu_t(\mathsf{E}\mid[a])\}\ge \theta_t
\qquad([a]\in\partial_m\mathsf{E}).
\]
Bowen's Gibbs estimate on cylinders, together with
Proposition~\ref{prop:scan-error-cylinder}, therefore yields
\[
\varepsilon_m(\mathsf{E};\mu_t)\asymp
\exp\!\Bigl(-m\bigl(P_X(\varphi_t)-P_\partial(\varphi_t)\bigr)\Bigr).
\]
\end{proof}

Theorem~\ref{ex:golden-mean-gibbs-weighted} can also be verified directly from
the explicit matrix: the exact identity for
$Z_m(\partial\mathsf{E},\varphi_t)$ and the explicit resolving words give the
same pressure-gap exponent first for the weighted boundary sum and, after the
resolving-thickness step in the proof, for scan error on the
transient-plus-periodic observer graph.  At $t=0$ this reduces to the Parry value
$\lambda_0=\lambda_{\mathrm G}$, while for $t\neq 0$ the exponent varies with
the potential even though
$N_m(\partial\mathsf{E})\asymp\lambda_{\mathrm G}^{m/2}$.

\section{Discussion}
\label{sec:discussion}

The symbolic contribution is concentrated in
Theorem~\ref{thm:finite-state-posterior},
Theorem~\ref{thm:weighted-boundary-pressure}, and the explicit golden-mean
computations of
Theorems~\ref{ex:golden-mean-specialization}
and~\ref{ex:golden-mean-gibbs-weighted}.  All four statements use the same
event-specific finite observer: the Parry theorem reads boundary growth and
thickness from the refined ambiguous graph, while the Gibbs theorem replaces
raw counts by the event-relative pressure
$P_\partial(\varphi)=\max_{C\in\mathfrak{C}_\partial}P_C(\varphi)$ and its gap
from $P_X(\varphi)$.  Sections~\ref{sec:prefix}--\ref{sec:tanaka} and
Appendix~\ref{sub:nonuniform} serve only to convert those symbolic boundary
estimates into scan-error statements once the relevant cylinder-mass bounds are
available.

\paragraph{Theorem-by-theorem ETDS comparison.}
Table~\ref{tab:etds-novelty-boundary} should be read as a local comparison
table against the cited adjacent ETDS results whose theorem statements
actually overlap with the present symbolic theorems.  It is not intended as a
bibliographic census.  The ETDS theorem-level claims begin only with
Theorem~\ref{thm:finite-state-posterior}, whose increment over
MacDonald~\cite{MacDonald2024} and Han--Marcus--Wu~\cite{HanMarcusWu2024} is
the event-relative observer that computes boundary growth, thickness, and scan
error, rather than a criterion for injective or finite-to-one factor-code
restrictions.  Theorem~\ref{thm:weighted-boundary-pressure} and
Theorem~\ref{ex:golden-mean-gibbs-weighted} then shift the thermodynamic
comparison away from factor-map relative pressure: against
Tsukamoto~\cite{Tsukamoto2023WeightedPressure},
Yayama~\cite{Yayama2023RelativePressure,Yayama2026RelativePressureGibbs}, and
Alibabaei~\cite{Alibabaei2025WeightedPressure}, the increment proved here is
an event-defined refined boundary graph whose reachable recurrent components,
acyclic tails, and residue classes determine the ordinary pressure gap for the
boundary partition sums and, with the separate resolving-thickness input, the
scan-error decay law of one fixed event.  Finally,
Theorem~\ref{ex:golden-mean-specialization} and
Theorem~\ref{ex:golden-mean-gibbs-weighted} supply explicit matrix formulas of
that event-relative type on the golden-mean shift, rather than only an
existence or variational-principle statement.

\paragraph{Shannon--McMillan--Breiman and uniform bounds.}
The SMB theorem \cite{Shannon1948,McMillan1953,Breiman1957} states that for
$\mu$-a.e.\ $x$,
\[
-\frac{1}{m}\log\mu([x]_m)\to h(\mu)
\quad\text{as }m\to\infty,
\]
where $[x]_m$ is the length-$m$ cylinder containing $x$.  This is a pointwise
almost-sure statement about a single cylinder along a single orbit; it does
\emph{not} give a uniform bound over all admissible cylinders simultaneously.
The uniform bounds $\mu(C)\asymp\lambda_{\mathrm{cyl}}^{-m}$ in
Proposition~\ref{thm:boundary-dimension}
and Corollary~\ref{thm:volume-boundary-scaling}
are strictly stronger.  For a mixing SFT with its Parry measure, these
same-length bounds are established by Parry \cite{Parry1964}, so the
count-based theorems apply directly.  For a general H\"older Gibbs measure,
Bowen's estimate \cite{Bowen1975} is instead
\[
\mu([a])\asymp \exp\!\bigl(S_m\phi(x)-mP(\phi)\bigr)\qquad(x\in[a]),
\]
which depends on the word $a$ through the Birkhoff sum $S_m\phi$.
Accordingly, the correct invariant Gibbs analogue is the weighted boundary
pressure law of Theorem~\ref{thm:weighted-boundary-pressure}, where the
boundary exponent is the pressure gap $P_X(\phi)-P_\partial(\phi)$ rather than
a bare count exponent.

\paragraph{Relation to margin conditions.}
The exact boundary identity
\[
\varepsilon_m(\mathsf{P};\mu)
=\sum_{C\in\partial_m\mathsf{P}}
\mu(C)\min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\}
\]
shows the right comparison point.  Uniform thickness gives a lower bound by
boundary mass, but it does not by itself control how much of the ambient space
has terminal posterior $p_\infty$ close to $1/2$.  That latter question is the
domain of the Mammen--Tsybakov condition
\cite{MammenTsybakov1999,Tsybakov2004}, so the comparison is only heuristic.
A closer multiscale analogue to our setup is the dyadic tree analysis of Scott and Nowak
\cite{ScottNowak2006Dyadic}.  In the finite-state symbolic setting the relevant
exponent is the higher-block invariant quantity
$\delta_\partial(\mathsf{P})$; in the general filtration-level theorems it is
only the raw count exponent of $N_m(\partial\mathsf{P})$.  We do not interpret
these quantities, or the thickness parameter $\theta$, as a classical margin
exponent.  Connecting either one to a margin exponent through filtration
geometry remains open.

\paragraph{Future directions.}
We do not claim any extension beyond the discrete finite-depth setting treated
above.  Two natural open problems are whether the filtration-level Tanaka
bookkeeping admits a genuine $q$-ary analogue, with $q-1$ threshold parameters,
and whether any continuous-time scaling limit leads to a Tanaka--Meyer type
formula \cite{RevuzYor1999}.  Other directions include weakening the thickness
hypothesis to an $L^p$ condition on boundary-atom probabilities and finding an
intrinsic pressure or information-theoretic description of ambiguity beyond the
finite-state observer setting.

\paragraph{Limitations.}
The general transfer principles still assume exponential growth of the positive-atom
counts $M_m=\#\mathcal{C}_m$ (equivalently, exponential cylinder complexity in
the symbolic specialization);
subexponential systems (Sturmian codings, substitution shifts) require
separate analysis.  The finite-state posterior theorem covers sofic and hidden
Markov situations whenever the event is $\mu$-finite-presented, and its
thickness/scan-error transfer conclusions require the extra
$\mu$-observer-regular resolving refinement; it does not treat arbitrary Borel
events in $\mathcal{F}_\infty$.  Even on a
mixing sofic shift or finite hidden Markov source, the decisive prefix
languages of a general tail event need not be regular, and the existence of a
finite posterior presentation is open in that level of generality.  For events
whose ambiguity cannot be compressed into finitely many observer states one
would need a weighted thermodynamic formalism for the boundary rather than the
count-based one used here.  Even in the finite-state symbolic setting, the paper proves invariance of
$\delta_\partial(\mathsf{P})$ only under higher-block recoding; invariance
under general finite-to-one sliding block conjugacies remains open.

\appendix

\section{Auxiliary Filtration-Level Log-Regular Transfer and Sharpness}
\label{sub:nonuniform}
\label{app:log-regular}

This appendix records a filtration-only extension of the exact boundary
decomposition.  It only isolates the polynomial loss permitted by log-regular
cylinder bounds; no symbolic observer structure enters here.

\begin{definition}[Log-regular measure]
\label{def:log-regular}
A measure $\mu$ is \emph{log-regular with exponent $\alpha>0$}, relative to an
existing cylinder growth base $\lambda_{\mathrm{cyl}}>1$, if there exist
$0<c_1\le c_2<\infty$ such that for all $m\ge 1$ and every positive depth-$m$
atom $C\in\mathcal{C}_m$,
\[
c_1 m^{-\alpha}\lambda_{\mathrm{cyl}}^{-m}\le\mu(C)\le c_2 m^{\alpha}\lambda_{\mathrm{cyl}}^{-m}.
\]
\end{definition}

\begin{theorem}[Log-regular transfer principle]
\label{thm:nonuniform-cylinder}
Let $\mathsf{P}\in\mathcal{B}$.  Assume that the cylinder growth base
$\lambda_{\mathrm{cyl}}$ exists and $\lambda_{\mathrm{cyl}}>1$.  Let $\mu$ be
log-regular with exponent $\alpha>0$.  Assume the thickness condition of
Corollary~\ref{thm:volume-boundary-scaling} holds with some $\theta>0$ and
$m_0$.  Then for all $m\ge m_0$,
\[
\theta c_1 m^{-\alpha}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})
\le\varepsilon_m(\mathsf{P};\mu)
\le \frac{c_2}{2}m^{\alpha}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P}).
\]
If, in addition, the boundary count satisfies the exact exponential asymptotic
$N_m(\partial\mathsf{P})=\lambda_{\mathrm{cyl}}^{\delta m+o(m)}$ (again stronger
than merely $\overline{\dim}_B(\partial\mathsf{P})=\delta$), then
$\varepsilon_m(\mathsf{P};\mu)=\lambda_{\mathrm{cyl}}^{-(1-\delta)m+o(m)}$ and the
exponential rate $1-\delta$ is unchanged from the uniform case.  More quantitatively,
if for some $\beta\ge 0$ and $A_-,A_+>0$,
\[
A_-m^{-\beta}\lambda_{\mathrm{cyl}}^{\delta m}
\le N_m(\partial\mathsf{P})
\le A_+m^\beta\lambda_{\mathrm{cyl}}^{\delta m}
\]
for all large $m$, then for all large $m$,
\[
\theta c_1A_-m^{-(\alpha+\beta)}\lambda_{\mathrm{cyl}}^{-(1-\delta)m}
\le\varepsilon_m(\mathsf{P};\mu)
\le \frac{c_2A_+}{2}m^{\alpha+\beta}\lambda_{\mathrm{cyl}}^{-(1-\delta)m}.
\]
\leanverified{paper_prefix_scan_error_nonuniform_cylinder}
\end{theorem}

\begin{proof}
Apply Proposition~\ref{prop:scan-error-cylinder} exactly as in
Corollary~\ref{thm:volume-boundary-scaling}, but replace the uniform
same-length bounds by
$c_1m^{-\alpha}\lambda_{\mathrm{cyl}}^{-m}\le\mu(C)\le
c_2m^\alpha\lambda_{\mathrm{cyl}}^{-m}$.  This gives the displayed two-sided
estimate immediately.  If
$N_m(\partial\mathsf{P})=\lambda_{\mathrm{cyl}}^{\delta m+o(m)}$, then
$\log m=o(m)$ shows that the polynomial factors do not alter the exponential
rate, so
$\varepsilon_m(\mathsf{P};\mu)=\lambda_{\mathrm{cyl}}^{-(1-\delta)m+o(m)}$.
The final quantitative estimate is obtained by substituting the two-sided
polynomial bounds for $N_m(\partial\mathsf{P})$ into the first display.
\end{proof}

\begin{example}[Filtration-level sharpness example]
\label{ex:log-regular-polynomial-correction}
This non-stationary product example shows that the polynomial factors in
Theorem~\ref{thm:nonuniform-cylinder} cannot in general be removed.

Let $X=\{0,1\}^{\NN}$ with its prefix filtration and let
$0<\alpha<1$.  Define the independent, non-identically distributed product
measure $\mu_\alpha$ by
\[
\mu_\alpha(x_n=0)=\frac12\left(1+\frac{\alpha}{n+2}\right),
\qquad
\mu_\alpha(x_n=1)=\frac12\left(1-\frac{\alpha}{n+2}\right).
\]
Then $\lambda_{\mathrm{cyl}}=2$ and $\mu_\alpha$ is log-regular with exponent
$\alpha$.  Indeed, for every word
$a=a_0\cdots a_{m-1}$,
\[
\mu_\alpha([a])=2^{-m}\prod_{n=0}^{m-1}
\left(1+\sigma_n(a_n)\frac{\alpha}{n+2}\right),
\qquad \sigma_n(0)=1,\quad \sigma_n(1)=-1,
\]
and the two extremal products satisfy
\[
\prod_{n=0}^{m-1}\left(1+\frac{\alpha}{n+2}\right)\asymp m^\alpha,
\qquad
\prod_{n=0}^{m-1}\left(1-\frac{\alpha}{n+2}\right)\asymp m^{-\alpha},
\]
by the standard gamma-ratio estimate
$\Gamma(m+c)/\Gamma(m)\asymp m^c$.

Let $Y:=\{0^\infty\}$ and, for $x\notin Y$, let
$\rho(x):=\min\{n\ge 0:x_n=1\}$.  Define
\[
\mathsf{P}:=\{x\notin Y:\rho(x)\text{ is even}\}\cup Y.
\]
At depth $m$, the only boundary cylinder is $[0^m]$, so
$N_m(\partial\mathsf{P})=1$ and the raw count exponent is $\delta=0$.  If the
prefix is not $0^m$, the first $1$ has already appeared and the event is
decided.  From $[0^m]$, however, one can force either parity of the first $1$
within at most two further symbols; since all one-coordinate probabilities are
bounded below by $q_*:=(1-\alpha/2)/2>0$,
\[
\min\{\mu_\alpha(\mathsf{P}\mid[0^m]),
1-\mu_\alpha(\mathsf{P}\mid[0^m])\}\ge q_*^2.
\]
Hence the thickness hypothesis holds and
\[
q_*^2\,\mu_\alpha([0^m])
\le \varepsilon_m(\mathsf{P};\mu_\alpha)
\le \frac12\,\mu_\alpha([0^m]).
\]
Finally,
\[
\mu_\alpha([0^m])
=2^{-m}\prod_{n=0}^{m-1}\left(1+\frac{\alpha}{n+2}\right)
\asymp m^\alpha 2^{-m}.
\]
Thus
\[
\varepsilon_m(\mathsf{P};\mu_\alpha)\asymp m^\alpha 2^{-m}.
\]
The exponential rate is still $1-\delta=1$, but the polynomial factor is
real; it cannot be replaced by a uniform constant under log-regularity alone.
Reversing the signs in the one-coordinate probabilities gives the matching
$m^{-\alpha}2^{-m}$ lower-side behaviour.
\end{example}

\section*{Acknowledgements}

The author acknowledges institutional support from CHRONOAI PTE.\ LTD.

\bibliographystyle{plain}
\bibliography{references}

\end{document}
