\section{Scope and claim hierarchy}
\label{sec:scope}

The source manuscript for this article is intentionally broader than a standard paper: it contains stable arithmetic, projection-word calculus, moment kernels, Hankel certificates, spectral fingerprints, \v{C}ech residuals, Hilbert-space interfaces, spacetime skeletons, and physicalization boundaries.  This submission extracts only the dependency-minimal main thread.

\begin{definition}[Claim hierarchy]
\label{def:claim-hierarchy}
We distinguish six levels of claims.
\begin{itemize}
\item \(C0\): the definition layer, including the scan--projection interface, window microstates, cylinder events, and visible sigma-algebras.
\item \(C1\): the finite combinatorial layer, including golden-mean syntax, Zeckendorf normal forms, folding rewrites, termination, confluence, and finite graph interfaces.
\item \(C2\): the cross-scale and statistical layer, including fold-aware restriction, recursive addressing, visible sigma-algebra non-expansion, scan error profiles, and boundary-dimension clarity laws.
\item \(C3\): the certificate layer, including finite graphs, entropy, spectral fingerprints, moment kernels, and reproducible experiments.
\item \(C4\): the bridge layer, including stable arithmetic, projection-word calculus, \v{C}ech residuals, local gluing obstructions, and pre-geometric candidates.
\item \(C5\): the physical interpretation layer, including Hilbert representations, quantum probabilities, spacetime skeletons, and Einstein-type closure.
\end{itemize}
\end{definition}

The main theorems of this paper cover only
\[
C0+C1+C2,
\]
and give a finite audit interface for \(C3\).  Claims in \(C4\) and \(C5\) are not proof assumptions and are not conclusions of the paper.  They are mentioned only as possible downstream admissible bridges.

\begin{remark}[Main-paper discipline]
The distinction in \Cref{def:claim-hierarchy} is structural rather than cosmetic.  Without it, local certificates, physical analogies, and main theorems compete for the same narrative weight.  The proof chain in this paper is
\[
\begin{aligned}
\mathsf{Scan}
&\to \mathsf{GoldenSyntax}
\to \mathsf{Fold}
\to \mathsf{NormalForm}\\
&\to \mathsf{CrossResolution}
\to \mathsf{Address}
\to \mathsf{ClarityAudit}.
\end{aligned}
\]
\end{remark}

\begin{definition}[Auditable stable-type generator]
\label{def:main-output-object}
An auditable stable-type generator is a family
\[
(\Om_m,X_m,\Fold_m,\pi_{m+1\to m},\widetilde\rho_{m+1,m},\mathcal F_m,\mathcal G_{\le L},\varepsilon_m,\Cert_m)_{m\ge1,L\ge0},
\]
where \(\Om_m\) is a finite window microstate space, \(X_m\) is a stable type space, \(\Fold_m\) is a golden folding map, \(\pi_{m+1\to m}\) is stable prefix restriction, \(\widetilde\rho_{m+1,m}\) is fold-aware restriction, \(\mathcal F_m\) is the finite visible sigma-algebra at scan depth \(m\), \(\mathcal G_{\le L}\) is the cumulative visible sigma-algebra up to address layer \(L\), \(\varepsilon_m\) is a scan error profile, and \(\Cert_m\) is a finite certificate package.
\end{definition}
