\section{Golden-mean folding}
\label{sec:folding}

The raw microstate space \(\Om_m=\{0,1\}^m\) is too large to serve as a stable type space.  We fold raw microstates into Zeckendorf-legal representatives by passing through Fibonacci residues.

For \(\omega=\omega_1\cdots\omega_m\in\Om_m\), define the raw Fibonacci value
\[
N_m(\omega):=\sum_{k=1}^{m}\omega_kF_{k+1}.
\]
The finite golden-mean fold is
\[
\boxed{
\Fold_m(\omega):=s_m\bigl(N_m(\omega)\bmod F_{m+2}\bigr).
}
\]
Equivalently,
\[
\Om_m\xrightarrow{N_m}\ZZ
\xrightarrow{\bmod F_{m+2}}\{0,\ldots,F_{m+2}-1\}
\xrightarrow{s_m}X_m.
\]
This is the primary mathematical definition.  Local Fibonacci carry elimination is an implementation of this map, not its definition.

\begin{theorem}[Golden folding normal form]
\label{thm:golden-fold-normal-form}
For each \(m\ge1\), the map
\[
\Fold_m:\Om_m\to X_m
\]
is well-defined, computable, idempotent, and surjective.  Moreover,
\[
\Fold_m(x)=x\qquad(x\in X_m),
\]
and every \(\omega\in\Om_m\) has a unique stable representative
\[
\operatorname{NF}_m(\omega)=\Fold_m(\omega)\in X_m.
\]
\end{theorem}

\begin{proof}
By \Cref{lem:finite-zeckendorf-section}, the finite Zeckendorf section
\[
s_m:\{0,\ldots,F_{m+2}-1\}\to X_m
\]
is well-defined and unique.  Therefore
\[
\Fold_m(\omega)=s_m(N_m(\omega)\bmod F_{m+2})
\]
is well-defined and lies in \(X_m\).

If \(x\in X_m\), then \(N_m(x)=V_m(x)\in\{0,\ldots,F_{m+2}-1\}\).  Hence
\[
N_m(x)\bmod F_{m+2}=N_m(x),
\]
and therefore
\[
\Fold_m(x)=s_m(N_m(x))=x.
\]
Thus \(\Fold_m\) fixes every stable word.

For arbitrary \(\omega\in\Om_m\), the first fold produces an element of \(X_m\).  Since all elements of \(X_m\) are fixed by the fold,
\[
\Fold_m^2(\omega)=\Fold_m(\Fold_m(\omega))=\Fold_m(\omega).
\]
Surjectivity follows because every \(x\in X_m\) satisfies \(\Fold_m(x)=x\).  Finally, uniqueness follows from the bijectivity of \(s_m\): each residue class modulo \(F_{m+2}\) has exactly one legal representative in \(X_m\), and \(\Fold_m\) selects it.
\end{proof}

\begin{proposition}[Fold equivalence equals residue equivalence]
\label{prop:fold-residue-equivalence}
For \(\omega,\omega'\in\Om_m\),
\[
\Fold_m(\omega)=\Fold_m(\omega')
\]
if and only if
\[
N_m(\omega)\equiv N_m(\omega')\pmod{F_{m+2}}.
\]
\end{proposition}

\begin{proof}
By definition,
\[
\Fold_m(\omega)=s_m(N_m(\omega)\bmod F_{m+2}).
\]
Since \(s_m\) is injective, two folds agree if and only if the corresponding residues modulo \(F_{m+2}\) agree.
\end{proof}

\begin{corollary}[Stable state-count reduction]
\label{cor:state-count-reduction}
\[
\frac{|X_m|}{|\Om_m|}
=\frac{F_{m+2}}{2^m}
=\Theta\!\left(\left(\frac{\varphi}{2}\right)^m\right).
\]
\end{corollary}

\begin{remark}[Local implementation]
The map \(\Fold_m\) may be computed by finite Zeckendorf normalization: starting from the Fibonacci digit string representing \(N_m(\omega)\), apply carry-elimination rules until no adjacent ones remain, reduce modulo \(F_{m+2}\), and retain the unique legal length-\(m\) representative.  Zeckendorf uniqueness guarantees path independence.  A local implementation audit belongs in \Cref{app:rewriting}.
\end{remark}
