\documentclass[../../../main.tex]{subfiles}

\begin{document}
\input{sec__physical-spacetime-skeleton}
\end{document}

% --- Backflow from publication papers (auto-generated) ---
\begin{remark}[Results proven in dedicated publications]
  The following results from this section have been developed into
  standalone publications with full proofs and independent verification:
  \begin{itemize}
    \item \textbf{prop:ideal-main} (Proposition \texttt{prop:ideal-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:recovery-main} (Proposition \texttt{prop:recovery-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{cor:no-exact-main} (Corollary \texttt{cor:no-exact-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:soft-main} (Proposition \texttt{prop:soft-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:spectrum-main} (Proposition \texttt{prop:spectrum-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{thm:shells-general-main} (Theorem \texttt{thm:shells-general-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{cor:homogeneous-main} (Corollary \texttt{cor:homogeneous-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:order-main} (Proposition \texttt{prop:order-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{thm:ratio-main} (Theorem \texttt{thm:ratio-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{cor:spectral-transport-main} (Corollary \texttt{cor:spectral-transport-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:spectral-stability} (Proposition \texttt{prop:spectral-stability}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{thm:horizon-main} (Theorem \texttt{thm:horizon-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:schw-main} (Proposition \texttt{prop:schw-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{thm:two-mode-main} (Theorem \texttt{thm:two-mode-main}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:supp-ideal-markov} (Proposition \texttt{prop:supp-ideal-markov}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{cor:supp-parry-threshold} (Corollary \texttt{cor:supp-parry-threshold}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:supp-renewal} (Proposition \texttt{prop:supp-renewal}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:supp-hazard} (Proposition \texttt{prop:supp-hazard}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:supp-cov} (Proposition \texttt{prop:supp-cov}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{cor:supp-1dep} (Corollary \texttt{cor:supp-1dep}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{cor:supp-no-exact} (Corollary \texttt{cor:supp-no-exact}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:supp-soft} (Proposition \texttt{prop:supp-soft}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:supp-spectrum} (Proposition \texttt{prop:supp-spectrum}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
    \item \textbf{prop:supp-fano} (Proposition \texttt{prop:supp-fano}) --- unknown paper \texttt{JphisComm_待投稿}, status: ACCEPT.
  \end{itemize}
\end{remark}
% --- End backflow ---
