\begin{theorem}[Uniqueness of binary jump structures under zero auxiliary registers and permutation naturality]\label{thm:bdry-z2-jump-functor-uniqueness}
Let \(k\ge 2\), and let \(\mathbf{FinSet}_k^{\cong}\) be the groupoid of finite \(k\)-element sets and bijections.
Let \(\mathbf{Tors}(\ZZ_2)\) be the groupoid of \(\ZZ_2\)-torsors, that is, two-point sets endowed with a free transitive \(\ZZ_2\)-action, together with their isomorphisms.
A \emph{binary jump structure} is a functor
\[
\mathcal J:\mathbf{Set}_k^{\cong}\longrightarrow \mathbf{Tors}(\ZZ_2).
\]
Then, for \(k\ge 3\), besides the everywhere-trivial functor there exists exactly one nontrivial such \(\mathcal J\), unique up to natural isomorphism.
This unique nontrivial functor is given by the orientation construction: for any \(k\)-element set \(S\),
\[
\mathrm{Or}(S):=\mathrm{Bij}([k],S)\big/\mathfrak{A}_k,
\qquad [k]:=\{1,\dots,k\},
\]
where \(\mathfrak{A}_k\trianglelefteq\mathfrak{S}_k\) is the alternating group and \(\mathfrak{S}_k\) acts on \(\mathrm{Bij}([k],S)\) by precomposition.
The induced action of \(\mathfrak{S}_k\) on \(\mathrm{Or}(S)\) factors precisely through the sign homomorphism \(\operatorname{sgn}:\mathfrak{S}_k\to \ZZ_2\), so that \(\mathrm{Or}(S)\) is naturally a \(\ZZ_2\)-torsor, and the assignment \(S\mapsto \mathrm{Or}(S)\) realizes the unique nontrivial binary jump structure.
\leanverified{signHom}
\leanverified{signHom\_exists}
\leanverified{trivHom}
\leanverified{sign\_range\_pm\_one}
\leanverified{sign\_sq\_eq\_one}
\leanverified{trivHom\_apply}
\leanverified{sign\_neq\_triv}
\leanverified{paper\_bdry\_z2\_jump\_functor\_skeleton}
\end{theorem}

\begin{proof}
Take the base object \(S_0=[k]\), so that \(\mathrm{Aut}_{\mathbf{FinSet}_k^{\cong}}(S_0)\cong \mathfrak{S}_k\).
The automorphism group of any \(\ZZ_2\)-torsor is canonically isomorphic to \(\ZZ_2\).
Hence any functor \(\mathcal J\) determines a group homomorphism
\[
\chi_{\mathcal J}:\mathfrak{S}_k\to \mathrm{Aut}(\mathcal J(S_0))\cong \ZZ_2.
\]
Conversely, given any homomorphism \(\chi:\mathfrak{S}_k\to \ZZ_2\), one may let \(\mathcal J_\chi(S_0)\) be a two-point set with \(\mathfrak{S}_k\)-action prescribed by \(\chi\).
Since \(\mathbf{FinSet}_k^{\cong}\) is a connected groupoid, this action transports along any bijection to every object \(S\), producing a functor \(\mathcal J_\chi\), and its natural-isomorphism class depends only on \(\chi\).
Thus natural-isomorphism classes of binary jump structures are in bijection with \(\mathrm{Hom}(\mathfrak{S}_k,\ZZ_2)\).
\par
If \(k\ge 3\), then \(\mathfrak{S}_k^{\mathrm{ab}}\cong \ZZ_2\), so \(\mathrm{Hom}(\mathfrak{S}_k,\ZZ_2)\) contains only the trivial homomorphism and the sign homomorphism.
Hence the nontrivial binary jump structure is unique up to natural isomorphism.
Finally, \(\mathrm{Or}(S)\) is obtained by quotienting \(\mathrm{Bij}([k],S)\) by \(\mathfrak{A}_k\), so it is a two-point set on which \(\mathfrak{S}_k\) acts through \(\mathfrak{S}_k/\mathfrak{A}_k\cong \ZZ_2\), namely through the sign representation.
Therefore it realizes the unique nontrivial class. \qedhere
\end{proof}

\begin{corollary}[Orientation parity for boundary uplift: from layer labels to an orientation torsor]\label{cor:bdry-uplift-orientation-parity}
Assume that no auxiliary register is introduced and that the construction must be natural under relabeling of uplift branches.
For each boundary uplift difference set \(\Delta_m^{\mathrm{bdry}}(w)\), define its orientation parity by
\[
\Pi_m(w):=\mathrm{Or}\!\bigl(\Delta_m^{\mathrm{bdry}}(w)\bigr)\in \mathbf{Tors}(\ZZ_2).
\]
Then:
\begin{enumerate}
  \item \(\Pi_m(w)\) is strictly functorial under every relabeling of the uplift branches; in particular, it depends only on the isomorphism class of \(\Delta_m^{\mathrm{bdry}}(w)\) as an unordered finite set.
  \item If \(|\Delta_m^{\mathrm{bdry}}(w)|=2\), then \(\Pi_m(w)\) reduces to the unique nontrivial \(\ZZ_2\)-torsor on a two-point fiber, which is exactly the familiar two-sheet parity of the window-\(6\) case.
  \item If \(|\Delta_m^{\mathrm{bdry}}(w)|=3\), then \(\Pi_m(w)\) produces the unique nontrivial binary jump object without selecting a distinguished sheet: it is not a permutation-invariant bipartition of the layer set itself, but rather the \(\ZZ_2\)-torsor induced by orientation, and therefore it is compatible with the obstruction to choosing a preferred \(1\sqcup 2\) split on a three-sheet fiber.
\end{enumerate}
\end{corollary}

\begin{proof}
Assertion (1) follows directly from the functoriality of \(\mathrm{Or}:\mathbf{Set}_k^{\cong}\to\mathbf{Tors}(\ZZ_2)\).
For (2), when \(|\Delta|=2\), the nontrivial action of \(\mathfrak{S}_2\cong\ZZ_2\) gives the unique nontrivial torsor structure on a two-point set, precisely the same structure carried by the two-sheet parity.
For (3), the case \(k=3\) of Theorem \ref{thm:bdry-z2-jump-functor-uniqueness} shows that the nontrivial binary jump structure is unique up to natural isomorphism and is realized by \(\mathrm{Or}\).
By contrast, a permutation-invariant bipartition of a three-element set would amount to choosing a distinguished \(1\sqcup 2\) split, which is a different type of object. \qedhere
\end{proof}

\begin{proposition}[Sign reduction on uplift fibers: binary jumps move from the center to the abelianization]\label{prop:uplift-fiber-sign-dimension-reduction}
Let \(S\) be a finite set with \(|S|=d\ge 2\), and let
\[
\mathrm{Lab}(S):=\mathrm{Bij}([d],S)
\]
be its set of labelings.
Then \(\mathrm{Lab}(S)\) is a principal \(\mathfrak{S}_d\)-torsor.
Writing \(\operatorname{sgn}:\mathfrak{S}_d\to \ZZ_2\) for the sign homomorphism, there is a natural isomorphism
\[
\mathrm{Or}(S)\ \cong\ \mathrm{Lab}(S)\times_{\mathfrak{S}_d,\operatorname{sgn}}\ZZ_2.
\]
Consequently, when \(d=2\), the group \(\mathfrak{S}_2\cong\ZZ_2\) is abelian, so the binary jump can be realized as a ``central'' flip inside the fiber.
When \(d=3\), however, \(Z(\mathfrak{S}_3)=\{e\}\) while \(\mathfrak{S}_3^{\mathrm{ab}}\cong\ZZ_2\) is still generated by \(\operatorname{sgn}\), so the binary jump can survive only through the sign representation, namely as an orientation torsor rather than a sheet label.
\end{proposition}

\begin{proof}
By definition,
\(
\mathrm{Or}(S)=\mathrm{Bij}([d],S)/\mathfrak{A}_d
\).
On the other hand, the contracted product \(\mathrm{Lab}(S)\times_{\mathfrak{S}_d,\operatorname{sgn}}\ZZ_2\) identifies two labelings precisely when they differ by an even permutation, so it is the same quotient by \(\mathfrak{A}_d\).
The assertions about the center and the abelianization are standard structural facts about symmetric groups: \(Z(\mathfrak{S}_d)=\{e\}\) for \(d\ge 3\), while \(\mathfrak{S}_d^{\mathrm{ab}}\cong\ZZ_2\) for \(d\ge 2\), generated by the sign homomorphism. \qedhere
\end{proof}
