\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amscd,amsthm,mathtools}
\usepackage{array}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{enumitem}

\numberwithin{equation}{section}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\NN}{\mathbb{N}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\Xinf}{X_\infty}
\DeclareMathOperator{\Fold}{Fold}
\DeclareMathOperator{\Law}{Law}

\title{A Discrete Stokes Formula for Fold-Truncation Defects Across Resolutions}
\author{}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We study one concrete multiscale phenomenon in the Fold tower: folding does not
commute with naive truncation.  For binary windows of length $n$, the direct
prefix truncation to length $m$ and the fold-aware restriction obtained from the
resolution-$n$ folded output produce different words in general.  We encode this
failure of commutation by an adjacent-scale local defect
$\kappa_{m+1\to m}$ and a tower-level global defect $D_{n\to m}$.  The first
main theorem is a discrete Stokes identity expressing $D_{n\to m}$ as the
telescoping xor-sum of projected local defects across the resolution band
$m,\dots,n-1$.  This yields auditable probability bounds, a Hamming--Lipschitz
budget, and a cocycle formula across intermediate scales.  We also show that
eventual summability of adjacent defects forces consistent inverse-limit lifts.
Finally, under a finite-state additivity hypothesis and a standard character
nondegeneracy condition, the defect process mixes exponentially to Haar measure
on a finite subgroup of $(\ZZ/2\ZZ)^m$.  The paper is deliberately confined to
this defect theory; it does not import the sofic machinery or gauge-anomaly
extensions of the larger manuscript.
\end{abstract}

\tableofcontents

\section{Introduction}

This paper isolates one obstruction in the resolution tower of Fold maps.  The
core folding paper studies the canonical map from raw binary windows to legal
golden-mean words.  Here we ask a more specific multiscale question: if a word
is first truncated and then folded, how far can the result be from what one
gets by first folding at high resolution and then projecting to low resolution?

The answer is subtle enough to deserve its own paper.  Direct truncation on raw
windows does not commute with folding.  The discrepancy is not noise, however.
It admits a rigid bookkeeping scheme across scales.  At each adjacent step
$m+1\to m$ one has a local defect $\kappa_{m+1\to m}$, and across a whole
resolution band one has a global defect $D_{n\to m}$.  The central result is a
discrete Stokes formula:
\[
\text{global defect}
\;=\;
\text{sum of projected local defects}.
\]
The sum is taken in the xor group $(\ZZ/2\ZZ)^m$, so the identity is exact and
auditable.

Three deliverables drive the paper.  First, we formulate the local and global
defects and prove the telescoping identity.  Second, we derive quantitative
consequences: an auditable probability bound, a Hamming--Lipschitz budget, and
a cocycle relation across intermediate scales.  Third, we record a statistical
endpoint: under a finite-state realization of the defect increments and a mild
character-detection assumption, the defect law converges exponentially to Haar
measure on its reachable subgroup.

The point of this paper is therefore narrower than the overall project:
\[
\begin{aligned}
\text{naive truncation}
&\not\!\circ\,\text{Fold} \\
&\Longrightarrow
\text{local defect}
\Longrightarrow
\text{discrete Stokes law}
\Longrightarrow
\text{budget and mixing theory}.
\end{aligned}
\]
We use only the minimal background on Zeckendorf normalization and symbolic
systems; see \cite{Zeckendorf1972,WaltersErgodicTheory,
LevinPeresWilmer2009MarkovMixing,
Internal2026FoldGeneralizedStokesDefectHaarMixing}.  No sofic-image theory,
Fischer covers, or arithmetic anomaly expansions are needed here.

\section{Folding and Naive Truncation}

\subsection{Minimal Setup}

For $m\ge 1$, let
\[
\Omega_m:=\{0,1\}^m
\qquad\text{and}\qquad
X_m:=\{x=(x_1,\dots,x_m)\in\{0,1\}^m:\ x_ix_{i+1}=0\text{ for }1\le i<m\}.
\]
Thus $X_m$ is the set of admissible words in the one-sided golden-mean shift.

For $\omega=\omega_1\cdots\omega_m\in\Omega_m$, define its Fibonacci value by
\[
N(\omega):=\sum_{k=1}^m \omega_k F_{k+1}.
\]
Let $Z(N)$ denote the Zeckendorf expansion of $N$ \cite{Zeckendorf1972}.  We
define the fold map by
\[
\Fold_m(\omega):=\pi_m(Z(N(\omega)))\in X_m,
\]
where $\pi_m$ truncates an infinite legal sequence to its first $m$ digits.
Equivalently, $\Fold_m$ is Fibonacci evaluation followed by Zeckendorf
normalization and truncation.

For $n\ge m$ we use two truncation maps:
\[
r_{n\to m}:\Omega_n\to\Omega_m,
\qquad
r_{n\to m}(\omega_1\cdots\omega_n):=\omega_1\cdots\omega_m,
\]
and
\[
\pi_{n\to m}:X_n\to X_m,
\qquad
\pi_{n\to m}(x_1,\dots,x_n):=(x_1,\dots,x_m).
\]
The second map is the natural restriction on already folded words.

\begin{lemma}[Functoriality of restriction]
\label{lem:pi-functorial-golden}
If $n\ge k\ge m$, then
\[
\pi_{n\to m}=\pi_{k\to m}\circ \pi_{n\to k}
\qquad\text{and}\qquad
r_{n\to m}=r_{k\to m}\circ r_{n\to k}.
\]
\end{lemma}

\begin{proof}
Both identities are immediate from the definition of prefix truncation.
\end{proof}

\subsection{Failure of Commutation}

The square formed by $r_{n\to m}$ and $\pi_{n\to m}$ does not commute in
general:
\[
\begin{CD}
\Omega_n @>{\Fold_n}>> X_n \\
@V{r_{n\to m}}VV @VV{\pi_{n\to m}}V \\
\Omega_m @>{\Fold_m}>> X_m.
\end{CD}
\]
The basic obstruction already appears at the smallest scales.

\begin{example}[A noncommuting square at $m=1$]
\label{ex:fold-noncommuting-square}
Take $n=2$, $m=1$, and $\omega=11$.  Then
\[
r_{2\to 1}(11)=1,
\qquad
\Fold_1(r_{2\to 1}(11))=\Fold_1(1)=1.
\]
On the other hand,
\[
N(11)=F_2+F_3=3=F_4,
\]
so the Zeckendorf normal form of $N(11)$ has its first two digits equal to
$00$.  Hence
\[
\Fold_2(11)=00,
\qquad
\pi_{2\to 1}(\Fold_2(11))=0.
\]
Therefore
\[
\Fold_1\circ r_{2\to 1}\neq \pi_{2\to 1}\circ \Fold_2.
\]
\end{example}

Example~\ref{ex:fold-noncommuting-square} is exactly the defect that we
formalize below.

\section{Local and Global Defects}

\subsection{Local Curvature Defect}

We work in the xor group structure on $\{0,1\}^m$, written with $\oplus$.

\begin{definition}[Local curvature defect]
\label{def:fold-local-curvature-defect}
For $m\ge 1$ and $\eta\in\Omega_{m+1}$, define the local defect
\[
\kappa_{m+1\to m}(\eta)
:=
\Fold_m(r_{m+1\to m}(\eta))
\oplus
\pi_{m+1\to m}(\Fold_{m+1}(\eta))
\in \{0,1\}^m.
\]
Its zero/nonzero indicator is
\[
K_{m+1\to m}(\eta):=\mathbf{1}_{\{\kappa_{m+1\to m}(\eta)\neq 0\}}.
\]
\end{definition}

The local defect records the failure of commutation at a single adjacent
resolution step.

\subsection{Global Tower Defect}

\begin{definition}[Global tower defect]
\label{def:fold-global-stokes-defect}
Let $n>m$.  For $\omega\in\Omega_n$, define
\[
D_{n\to m}(\omega)
:=
\Fold_m(r_{n\to m}(\omega))
\oplus
\pi_{n\to m}(\Fold_n(\omega))
\in \{0,1\}^m.
\]
\end{definition}

The defect $D_{n\to m}$ measures the aggregate mismatch accumulated when one
compares the two paths from $\Omega_n$ down to $X_m$ in the noncommuting square.

\begin{table}[ht]
\centering
\small
\begin{tabular}{>{$\displaystyle}l<{$} >{$\displaystyle}l<{$} p{3.8in}}
\toprule
\text{observable} & \text{range} & \text{meaning} \\
\midrule
\kappa_{m+1\to m}(\eta) & \{0,1\}^m &
adjacent-scale local noncommutation defect \\
D_{n\to m}(\omega) & \{0,1\}^m &
aggregate defect across the whole band $m,\dots,n-1$ \\
K_{m+1\to m}(\eta) & \{0,1\} &
indicator that the local defect is nonzero \\
\lvert D_{n\to m}(\omega)\rvert_0 & \{0,\dots,m\} &
Hamming size of the global defect \\
\bottomrule
\end{tabular}
\caption{The defect observables used throughout the paper.}
\label{tab:defect-observables}
\end{table}

\section{A Discrete Stokes Identity}

\subsection{Telescoping Structure}

\begin{theorem}[Discrete Stokes identity for Fold defects]
\label{thm:fold-discrete-stokes-defect}
For every $n>m$ and every $\omega\in\Omega_n$,
\[
\boxed{
D_{n\to m}(\omega)
=
\bigoplus_{j=m}^{n-1}
\pi_{j\to m}\!\left(
\kappa_{j+1\to j}(r_{n\to j+1}(\omega))
\right).
}
\]
\end{theorem}

\begin{proof}
For $j\in\{m,\dots,n\}$ define
\[
A_j(\omega):=\Fold_j(r_{n\to j}(\omega)),
\qquad
B_j(\omega):=\pi_{n\to j}(\Fold_n(\omega)),
\]
and
\[
\Delta_j(\omega):=A_j(\omega)\oplus B_j(\omega)\in\{0,1\}^j.
\]
Then $\Delta_m(\omega)=D_{n\to m}(\omega)$ and $\Delta_n(\omega)=0$.

Fix $j\in\{m+1,\dots,n\}$.  By Lemma~\ref{lem:pi-functorial-golden},
\[
B_{j-1}(\omega)=\pi_{j\to j-1}(B_j(\omega))
\]
and
\[
A_{j-1}(\omega)=\Fold_{j-1}(r_{j\to j-1}(r_{n\to j}(\omega))).
\]
Therefore
\[
\begin{aligned}
\Delta_{j-1}(\omega)
&=A_{j-1}(\omega)\oplus B_{j-1}(\omega) \\
&=\Bigl(A_{j-1}(\omega)\oplus \pi_{j\to j-1}(A_j(\omega))\Bigr)
\oplus
\pi_{j\to j-1}\bigl(A_j(\omega)\oplus B_j(\omega)\bigr).
\end{aligned}
\]
The first bracket is exactly
\[
\kappa_{j\to j-1}(r_{n\to j}(\omega)),
\]
while the second term is $\pi_{j\to j-1}(\Delta_j(\omega))$.  Hence
\[
\Delta_{j-1}(\omega)
=
\kappa_{j\to j-1}(r_{n\to j}(\omega))
\oplus
\pi_{j\to j-1}(\Delta_j(\omega)).
\]
Iterating from $j=n$ down to $j=m+1$ and using the fact that prefix projection
commutes with xor gives
\[
\Delta_m(\omega)
=
\bigoplus_{j=m}^{n-1}
\pi_{j\to m}\!\left(
\kappa_{j+1\to j}(r_{n\to j+1}(\omega))
\right).
\]
Since $\Delta_m(\omega)=D_{n\to m}(\omega)$, the proof is complete.
\end{proof}

\subsection{Auditable Consequences}

\begin{corollary}[Auditable probability bound]
\label{cor:fold-discrete-stokes-auditable-bound}
Let $n>m$, let $\mu$ be any probability measure on $\Omega_n$, and let
$f:X_m\to\RR$ be bounded.  Then
\[
\left|
\EE_\mu f\bigl(\Fold_m(r_{n\to m}(\omega))\bigr)
-
\EE_\mu f\bigl(\pi_{n\to m}(\Fold_n(\omega))\bigr)
\right|
\le
2\lVert f\rVert_\infty\,
\PP_\mu[D_{n\to m}(\omega)\neq 0].
\]
Moreover,
\[
\{D_{n\to m}\neq 0\}
\subseteq
\bigcup_{j=m}^{n-1}
\left\{
K_{j+1\to j}(r_{n\to j+1}(\omega))=1
\right\},
\]
and therefore
\[
\PP_\mu[D_{n\to m}\neq 0]
\le
\sum_{j=m}^{n-1}
\PP_\mu\!\left[
K_{j+1\to j}(r_{n\to j+1}(\omega))=1
\right].
\]
\end{corollary}

\begin{proof}
Write
\[
x=\Fold_m(r_{n\to m}(\omega)),
\qquad
y=\pi_{n\to m}(\Fold_n(\omega)).
\]
Then $\{x\neq y\}=\{D_{n\to m}(\omega)\neq 0\}$, and
\[
\lvert f(x)-f(y)\rvert
\le
2\lVert f\rVert_\infty\,\mathbf{1}_{\{x\neq y\}}.
\]
Taking expectation yields the first inequality.

If every local defect in the Stokes sum of Theorem
~\ref{thm:fold-discrete-stokes-defect} vanishes, then the xor-sum vanishes as
well.  The event inclusion follows by contraposition, and the final estimate is
just the union bound.
\end{proof}

\section{Budget Bounds and Cocycle Structure}

\subsection{Hamming--Lipschitz Budget}

\begin{corollary}[Hamming--Lipschitz budget]
\label{cor:fold-hamming-lipschitz-budget}
Let $n>m$, let $\mu$ be any probability measure on $\Omega_n$, and let
$d_{\mathrm H}(x,y):=\lvert x\oplus y\rvert_0$ be Hamming distance on $X_m$.
If $f:X_m\to\RR$ is $L$-Lipschitz with respect to $d_{\mathrm H}$, then
\[
\left|
\EE_\mu f\bigl(\Fold_m(r_{n\to m}(\omega))\bigr)
-
\EE_\mu f\bigl(\pi_{n\to m}(\Fold_n(\omega))\bigr)
\right|
\le
L\,\EE_\mu \lvert D_{n\to m}(\omega)\rvert_0
\]
and
\[
\EE_\mu \lvert D_{n\to m}(\omega)\rvert_0
\le
\sum_{j=m}^{n-1}
\EE_\mu\left\lvert
\kappa_{j+1\to j}(r_{n\to j+1}(\omega))
\right\rvert_0.
\]
\end{corollary}

\begin{proof}
For the same $x$ and $y$ as in the proof of
Corollary~\ref{cor:fold-discrete-stokes-auditable-bound},
\[
\lvert f(x)-f(y)\rvert
\le
L\,d_{\mathrm H}(x,y)
=
L\,\lvert D_{n\to m}(\omega)\rvert_0.
\]
Taking expectation gives the first inequality.

For the second, apply Theorem~\ref{thm:fold-discrete-stokes-defect} and use the
subadditivity
\[
\lvert u\oplus v\rvert_0\le \lvert u\rvert_0+\lvert v\rvert_0
\]
together with the fact that prefix projection cannot increase Hamming weight.
\end{proof}

\subsection{Cocycle and Triangle Structure}

\begin{proposition}[Cocycle identity across intermediate scales]
\label{prop:fold-defect-cocycle}
If $m<k<n$, then for every $\omega\in\Omega_n$,
\[
D_{n\to m}(\omega)
=
D_{k\to m}(r_{n\to k}(\omega))
\oplus
\pi_{k\to m}(D_{n\to k}(\omega)).
\]
Consequently, if
\[
\Delta_{n,m}(\omega):=\lvert D_{n\to m}(\omega)\rvert_0,
\]
then
\[
\Delta_{n,m}(\omega)
\le
\Delta_{k,m}(r_{n\to k}(\omega))
+
\Delta_{n,k}(\omega).
\]
\end{proposition}

\begin{proof}
Split the Stokes sum of Theorem~\ref{thm:fold-discrete-stokes-defect} at the
intermediate scale $k$:
\[
\begin{aligned}
D_{n\to m}(\omega)
&=
\bigoplus_{j=m}^{k-1}
\pi_{j\to m}\!\left(
\kappa_{j+1\to j}(r_{n\to j+1}(\omega))
\right) \\
&\quad\oplus
\bigoplus_{j=k}^{n-1}
\pi_{j\to m}\!\left(
\kappa_{j+1\to j}(r_{n\to j+1}(\omega))
\right).
\end{aligned}
\]
The first block depends only on the prefix $r_{n\to k}(\omega)$ and is exactly
$D_{k\to m}(r_{n\to k}(\omega))$.  In the second block we use
$\pi_{j\to m}=\pi_{k\to m}\circ \pi_{j\to k}$ from
Lemma~\ref{lem:pi-functorial-golden} and factor out $\pi_{k\to m}$, obtaining
$\pi_{k\to m}(D_{n\to k}(\omega))$.  This proves the cocycle identity.

The triangle inequality follows from
\[
\lvert u\oplus v\rvert_0\le \lvert u\rvert_0+\lvert v\rvert_0
\]
and the nonexpansiveness of prefix projection in Hamming weight.
\end{proof}

\section{Lifts and Asymptotic Compatibility}

\subsection{Consistent Lifts}

Let
\[
\Xinf:=\{x=(x_1,x_2,\dots)\in\{0,1\}^{\NN}: x_ix_{i+1}=0\text{ for all }i\ge 1\}
\]
be the one-sided golden-mean shift, with projections
\[
\pi_{\infty\to m}(x):=(x_1,\dots,x_m)\in X_m.
\]

\begin{proposition}[Consistent lift after eventual defect extinction]
\label{prop:fold-naive-prefix-lift}
Let $\omega^\infty\in\{0,1\}^{\NN}$ and define
\[
x_m:=\Fold_m(r_{\infty\to m}(\omega^\infty))\in X_m.
\]
For each $m\ge 1$, define the adjacent discrepancy
\[
\delta_m:=x_m\oplus \pi_{m+1\to m}(x_{m+1})\in\{0,1\}^m.
\]
If
\[
\sum_{m\ge 1}\lvert \delta_m\rvert_0<\infty,
\]
then there exists $M$ such that $\delta_m=0$ for every $m\ge M$.  Hence
$(x_m)_{m\ge M}$ is compatible under restriction, and there exists a unique
$x^\infty\in\Xinf$ satisfying
\[
\pi_{\infty\to m}(x^\infty)=x_m
\qquad\text{for all }m\ge M.
\]
\end{proposition}

\begin{proof}
Since each $\lvert\delta_m\rvert_0$ is a nonnegative integer and their sum is
finite, only finitely many of them can be nonzero.  Therefore there exists $M$
such that $\delta_m=0$ for all $m\ge M$, which means
\[
x_m=\pi_{m+1\to m}(x_{m+1})
\qquad\text{for every }m\ge M.
\]
Thus $(x_m)_{m\ge M}$ is a compatible family of admissible prefixes, so it
defines a unique infinite admissible sequence $x^\infty\in\Xinf$.
\end{proof}

\begin{remark}
Proposition~\ref{prop:fold-naive-prefix-lift} is the precise form of asymptotic
compatibility needed for this paper.  We do not claim that all Fold trajectories
stabilize in this way.  The proposition only says that summable adjacent defects
force eventual consistency.
\end{remark}

\section{Statistical Remarks}

\subsection{Haar Mixing of the Defect Process}

The next theorem is stated under two structural assumptions.  These are exactly
the assumptions needed to turn the Stokes defect into a standard finite-state
additive observable over a primitive hidden Markov chain.

\begin{theorem}[Haar mixing under a finite-state additivity hypothesis]
\label{thm:fold-stokes-defect-haar-mixing}
Fix $m\ge 1$, and let $\omega$ be uniformly distributed on $\Omega_n$ for
$n\ge m$.  Let $\mathsf{H}_m\le (\ZZ/2\ZZ)^m$ be the subgroup generated by the
reachable projected increments
\[
\mathsf{H}_m
:=
\left\langle
\pi_{j\to m}\bigl(\kappa_{j+1\to j}(\eta)\bigr):
\ j\ge m,\ \eta\in\Omega_{j+1}
\right\rangle.
\]
Assume:
\begin{enumerate}[label=(\roman*)]
\item for fixed $m$, the defect increments admit a synchronous finite-state
realization over IID Bernoulli input with parameter $1/2$, and the hidden
Markov chain is primitive
on the reachable component;
\item for every nontrivial character
$\chi\in\widehat{\mathsf{H}_m}\setminus\{1\}$, there exists a closed cycle in
the reachable component whose total increment $g\in \mathsf{H}_m$ satisfies
$\chi(g)\neq 1$.
\end{enumerate}
Then there exist constants $C_m>0$ and $0<\eta_m<1$ such that for all $n\ge m$,
\[
\left\|
\Law(D_{n\to m})-\mathrm{Haar}(\mathsf{H}_m)
\right\|_{\mathrm{TV}}
\le
C_m\,\eta_m^{\,n-m}.
\]
\end{theorem}

\begin{proof}
Theorem~\ref{thm:fold-discrete-stokes-defect} writes the defect as a sum of
projected local increments.  Under assumption (i), these increments can be read
by a finite-state machine driven by IID Bernoulli input.  Thus for fixed $m$
the defect process becomes a finite-state additive functional on a primitive
hidden Markov chain.

For any character $\chi\in\widehat{\mathsf{H}_m}$, define the twisted transfer
matrix $P_\chi$ by weighting each transition by the character value of its
increment.  When $\chi=1$, this is the usual stochastic transition matrix of
the hidden chain, hence it has spectral radius $1$.  For $\chi\neq 1$,
assumption (ii) rules out phase alignment on all closed cycles, so the
Perron--Frobenius eigenvalue drops strictly below $1$.  Let
\[
\eta_m:=\max_{\chi\neq 1}\rho(P_\chi)<1.
\]
Standard Fourier inversion on the finite abelian group $\mathsf{H}_m$ then
shows that every nontrivial Fourier coefficient of $\Law(D_{n\to m})$ decays
like $O(\eta_m^{\,n-m})$.  Since Haar measure kills all nontrivial characters,
the same exponential bound holds in total variation; see
\cite[Chapters~12--13]{LevinPeresWilmer2009MarkovMixing}.  A detailed audit of
the twisted-operator construction for the present Fold setting is recorded in
\cite{Internal2026FoldGeneralizedStokesDefectHaarMixing}.
\end{proof}

\subsection{Parity Polarization}

\begin{corollary}[Parity polarization and binomial limit at full generation]
\label{cor:fold-stokes-defect-parity-polarization}
Under the hypotheses of Theorem~\ref{thm:fold-stokes-defect-haar-mixing}, fix
$u\in (\ZZ/2\ZZ)^m$ and define
\[
\chi_u(v):=(-1)^{\langle u,v\rangle}
\qquad (v\in (\ZZ/2\ZZ)^m).
\]
Then there exist constants $C_m>0$ and $0<\eta_m<1$ such that for all $n\ge m$,
\[
\left|
\EE[\chi_u(D_{n\to m})]
-
\mathbf{1}_{\{u\in \mathsf{H}_m^\perp\}}
\right|
\le
C_m\,\eta_m^{\,n-m},
\]
where
\[
\mathsf{H}_m^\perp:=\{u\in(\ZZ/2\ZZ)^m:\chi_u|_{\mathsf{H}_m}\equiv 1\}.
\]
In particular, if $\mathsf{H}_m=(\ZZ/2\ZZ)^m$, then
\[
\EE[\chi_u(D_{n\to m})]=O(\eta_m^{\,n-m})
\qquad\text{for every }u\neq 0,
\]
and
\[
\lvert D_{n\to m}\rvert_0
\xrightarrow[n\to\infty]{\mathrm{law}}
\mathrm{Binomial}\!\left(m,\frac12\right).
\]
\end{corollary}

\begin{proof}
The total-variation estimate from
Theorem~\ref{thm:fold-stokes-defect-haar-mixing} controls the expectation of
every $\{\pm 1\}$-valued observable, hence of $\chi_u$.  Under Haar measure on
$\mathsf{H}_m$, the character average equals $1$ exactly on the annihilator
$\mathsf{H}_m^\perp$ and vanishes otherwise.  This gives the first claim.

If $\mathsf{H}_m=(\ZZ/2\ZZ)^m$, the limit law is uniform on all binary vectors
of length $m$.  Under that law the coordinates are independent Bernoulli$(1/2)$
variables, so the Hamming weight has the stated binomial distribution.
\end{proof}

\begin{remark}
The statistical section is intentionally kept short.  The main paper here is
the deterministic defect identity.  The mixing theorem is included because it
shows that the defect process is not merely a bookkeeping device but a genuine
dynamical observable.
\end{remark}

\section{Discussion}

The paper isolates one exact mechanism: folding and naive truncation fail to
commute, but the failure is structured.  Local adjacent-scale defects assemble
into a global tower defect by a discrete Stokes formula.  Once this identity is
available, quantitative corollaries follow with little extra work: auditable
probability bounds, Lipschitz budgets, and cocycle inequalities.

The consistent-lift proposition gives the clean finite-to-infinite endpoint of
the deterministic theory.  The Haar-mixing theorem shows how the same defect can
be treated statistically when the Fold tower is read by a finite-state additive
kernel.  This is the correct stopping point for the present note.  We do not
import the broader sofic, gauge-anomaly, or arithmetic expansions of the larger
project.

\bibliographystyle{alpha}
\bibliography{../../2026_golden_ratio_driven_scan_projection_generation_recursive_emergence/references,../../2026_golden_ratio_driven_scan_projection_generation_recursive_emergence/references_internal_2026}

\end{document}
