\subsection{Homological evaluation and intrinsic visibility}

The strict quotient $A_{\mathrm{vis}}=A/H_\alpha$ depends on a chosen cocycle representative. To isolate the part of the obstruction that belongs to the class itself, the following development takes the homological evaluation map as primary. The intrinsic quotient will be defined from this map and only afterwards recovered through class-admissible character channels.

Throughout this subsection we work with one fixed finite cover and one fixed finite nerve presentation. All cochain groups are computed on that presentation. The only comparison maps used are the explicit identification of \v{C}ech cochains with simplicial cochains on the finite nerve and the resulting universal coefficient sequence for that simplicial model. Whenever the branch obstruction class is viewed as a derived $H^2$-class, that comparison is part of the chosen finite nerve presentation rather than an unstated background identification.

The algebraic ingredients themselves are standard: the universal coefficient exact sequence, finite Pontryagin duality, and elementary quotient theory. The point of the subsection is therefore not to present these facts as new algebra, but to isolate the semantic quotient detected by $\mathrm{ev}_{\omega}$ and the pure $\operatorname{Ext}$-type regime in which a nonzero obstruction remains character-blind.

We also make one deliberate restriction of scope. The cover $\mathcal U$ is required to trivialize the band, so the coefficients may be treated as a constant finite abelian group $A$ on the nerve. If the band has nontrivial monodromy on the chosen cover, the correct replacement is cohomology with local coefficients on $N(\mathcal U)$; that case is not treated here.

\begin{definition}[Finite nerve presentation of the obstruction class]\label{def:finite-nerve-presentation}
Fix an admitted reference $r:\mathsf{Ref}_s$ at a state $p$, and assume conditions \textup{(i)}--\textup{(iv)} of \Cref{def:finite-obstruction-presentation}. A \emph{finite nerve presentation} of the branch obstruction class consists of:
\begin{enumerate}[label=(\roman*),nosep]
\item a finite covering family
\[
\mathcal U=\{a_i\to a\}_{i\in I},
\qquad
a=[r]_p,
\]
whose \v{C}ech nerve $N(\mathcal U)$ is finite;
\item a trivialization of the band on $\mathcal U$, so that the band may be viewed as a constant coefficient finite abelian group $A$ on $N(\mathcal U)$;
\item the induced obstruction class
\[
\omega\in H^2(N(\mathcal U),A),
\]
obtained by identifying \v{C}ech cochains on $\mathcal U$ with simplicial cochains on the finite nerve, transporting the branch obstruction cocycle through that identification, and writing
\[
\omega:=\omega_v
\]
for the chosen visible branch. When the class is also viewed as a derived $H^2$-class, this identification is part of the chosen presentation.
\end{enumerate}
Write
\[
\eta_A:H^2(N(\mathcal U),A)\longrightarrow \mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),A)
\]
for the map in the universal coefficient exact sequence
\[
0\to \operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),A)\to H^2(N(\mathcal U),A)\xrightarrow{\eta_A}\mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),A)\to 0
\]
\cite[Sec.~3.1]{Hatcher2002}, and define the \emph{homological evaluation map}
\[
\mathrm{ev}_{\omega}:=\eta_A(\omega).
\]
\end{definition}

\begin{proposition}[Finite nerve comparison isomorphism]\label{prop:finite-nerve-comparison}
Let $\mathcal U=\{a_i\to a\}_{i\in I}$ and $N(\mathcal U)$ be as above. For each $n\ge 0$, each nonempty $(n{+}1)$-fold overlap $a_{i_0\cdots i_n}$ determines a unique $n$-simplex $(i_0,\dots,i_n)$ of $N(\mathcal U)$, and the assignment
\[
\Theta_{\mathcal U}^n:\check C^n(\mathcal U,A)\to C^n(N(\mathcal U),A),\qquad
(\alpha_{i_0\cdots i_n})\longmapsto\bigl((i_0,\dots,i_n)\mapsto \alpha_{i_0\cdots i_n}\bigr)
\]
is an isomorphism of cochain groups. Under this correspondence,
\[
\delta\leftrightarrow d,\qquad
B^n\leftrightarrow B^n,\qquad
Z^n\leftrightarrow Z^n.
\]
Hence cohomology is identified canonically.
\end{proposition}
\begin{proof}
The $n$-cochains on $\mathcal U$ are families of coefficients indexed by nonempty $(n{+}1)$-fold overlaps $a_{i_0\cdots i_n}$, while the simplicial $n$-cochains on $N(\mathcal U)$ are coefficient functions on the $n$-simplices $(i_0,\dots,i_n)$. By definition of the nerve, these two index sets agree, so $\Theta_{\mathcal U}^n$ is a bijection.

The \v{C}ech coboundary of $(\alpha_{i_0\cdots i_n})$ is the alternating sum of restrictions to the faces of each overlap, and the simplicial coboundary of $\Theta_{\mathcal U}^n(\alpha)$ is the alternating sum over the faces of the simplex $(i_0,\dots,i_{n+1})$. Because the face maps of $N(\mathcal U)$ are induced by omitting one index, the two alternating sums coincide term by term. Hence
\[
\Theta_{\mathcal U}^{n+1}(\delta\alpha)=d(\Theta_{\mathcal U}^n(\alpha)).
\]
Therefore $\Theta_{\mathcal U}^\bullet$ is an isomorphism of cochain complexes, and it induces the stated identifications on cocycles, coboundaries, and cohomology.
\end{proof}
\leanverified{paper_conservative_extension_maximal_collapse}

\begin{proposition}[Strict versus intrinsic hidden subgroups]\label{prop:strict-vs-intrinsic-hidden-subgroups}
In the setting of \Cref{def:finite-nerve-presentation}, let $\alpha\in Z^2(N(\mathcal U),A)$ be a simplicial cocycle representing $\omega$, and let
\[
\widetilde\alpha:C_2(N(\mathcal U),\mathbb Z)\to A
\]
be the chain-level evaluation map obtained by linear extension from the oriented $2$-simplices. Then
\[
H_\alpha=\operatorname{Im}(\widetilde\alpha)
\]
and
\[
K_{\omega}:=\mathrm{Im}(\mathrm{ev}_{\omega}).
\]
Thus $H_\alpha$ records presentation-level chain visibility: it is generated by the values of the chosen cocycle on all $2$-chains. By contrast, $K_{\omega}$ records cycle-level class visibility: it is the image that survives after passage to $2$-homology. In particular,
\[
K_{\omega}\subseteq H_\alpha.
\]
\end{proposition}
\begin{proof}
The image of $\widetilde\alpha$ is the subgroup generated by the values of $\alpha$ on the oriented $2$-simplices, equivalently by the values of the original \v{C}ech cocycle on the nonempty triple overlaps of $\mathcal U$. This is exactly the strict subgroup attached to the chosen presentation. The second equality is the definition of $K_\omega$. Since $H_2(N(\mathcal U),\mathbb Z)$ is represented by $2$-cycles inside $C_2(N(\mathcal U),\mathbb Z)$, every homological value counted by $\mathrm{ev}_\omega$ is, a fortiori, a chain-level value of $\widetilde\alpha$, so $K_\omega\subseteq H_\alpha$.
\end{proof}

\begin{definition}[Intrinsic visible quotient]\label{def:intrinsic-visible-quotient}
In the setting of \Cref{def:finite-nerve-presentation}, define the \emph{intrinsic hidden subgroup}
\[
K_{\omega}:=\mathrm{Im}(\mathrm{ev}_{\omega})
\]
and the associated \emph{intrinsic visible quotient}
\[
A_{\mathrm{vis}}^{\omega}:=\operatorname{coker}(\mathrm{ev}_{\omega})=A/K_{\omega}.
\]
\end{definition}

\begin{definition}[Class-admissible characters]\label{def:class-admissible-character}
In the setting of \Cref{def:finite-nerve-presentation}, let
\[
X_{\omega}:=\{\chi\in\widehat A\mid \chi_*(\omega)=0\},
\qquad
\widehat A:=\mathrm{Hom}(A,\mathbb T).
\]
Elements of $X_{\omega}$ are called \emph{class-admissible} for $\omega$.
\end{definition}

\begin{proposition}[Character-theoretic detection of the intrinsic visible quotient]\label{thm:intrinsic-visible-quotient}
In the setting of \Cref{def:class-admissible-character},
\[
X_{\omega}
=
\{\chi\in\widehat A\mid \chi\circ \mathrm{ev}_{\omega}=0\}
=
\mathrm{Ann}(K_{\omega}).
\]
Consequently,
\[
K_{\omega}
=
\bigcap_{\chi\in X_{\omega}}\ker\chi,
\qquad
A_{\mathrm{vis}}^{\omega}\cong A\Big/\!\bigcap_{\chi\in X_{\omega}}\ker\chi.
\]
\leanverified{paper_conservative_extension_intrinsic_visible_quotient}
\end{proposition}

\begin{proof}
For every character $\chi:A\to\mathbb T$, naturality of the universal coefficient sequence gives a commutative square
\[
\begin{CD}
H^2(N(\mathcal U),A) @>\eta_A>> \mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),A) \\
@V\chi_*VV @VV(\chi\circ -)V \\
H^2(N(\mathcal U),\mathbb T) @>\eta_{\mathbb T}>> \mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),\mathbb T).
\end{CD}
\]
Since $\mathbb T$ is divisible and hence injective in $\mathbf{Ab}$, one has
\[
\operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),\mathbb T)=0
\]
\cite[Sec.~2.3]{Weibel1994}. Thus $\eta_{\mathbb T}$ is an isomorphism by \Cref{def:finite-nerve-presentation}. Therefore
\[
\chi_*(\omega)=0
\Longleftrightarrow
\eta_{\mathbb T}(\chi_*(\omega))=0
\Longleftrightarrow
\chi\circ \eta_A(\omega)=0
\Longleftrightarrow
\chi\circ \mathrm{ev}_{\omega}=0.
\]
This proves the first displayed identity, and the second is simply the definition of the annihilator of a subgroup of $A$.

Let $H:=K_{\omega}=\mathrm{Im}(\mathrm{ev}_{\omega})\le A$. By the first part of the proof, $X_{\omega}=\mathrm{Ann}(H)$. Finite abelian duality implies
\[
\bigcap_{\chi\in \mathrm{Ann}(H)}\ker\chi=H
\]
\cite[Part~I, Chaps.~2, 12]{Terras1999}. Hence
\[
K_{\omega}
=
\bigcap_{\chi\in X_{\omega}}\ker\chi.
\]
The final isomorphism follows immediately:
\[
A_{\mathrm{vis}}^{\omega}
=
A/K_{\omega}
\cong
A\Big/\!\bigcap_{\chi\in X_{\omega}}\ker\chi.
\qedhere
\]
\end{proof}

\begin{corollary}[Character detection of the intrinsic visible quotient]\label{cor:intrinsic-character-detection}
The quotient map
\[
\pi_{\omega}:A\twoheadrightarrow A_{\mathrm{vis}}^{\omega}
\]
induces a natural isomorphism
\[
\widehat{A_{\mathrm{vis}}^{\omega}}\cong X_{\omega}.
\]
For elements $x,y\in A$, one has
\[
\pi_{\omega}(x)=\pi_{\omega}(y)
\quad\Longleftrightarrow\quad
\chi(x)=\chi(y)\ \text{for every class-admissible }\chi.
\]
\leanverified{paper_conservative_extension_intrinsic_character_detection}
\end{corollary}

\begin{proof}
By \Cref{thm:intrinsic-visible-quotient},
\[
X_{\omega}=\mathrm{Ann}(K_{\omega})=K_{\omega}^{\perp}.
\]
Characters of $A_{\mathrm{vis}}^{\omega}=A/K_{\omega}$ are exactly the characters of $A$ that kill $K_{\omega}$, so precomposition with $\pi_{\omega}$ identifies $\widehat{A_{\mathrm{vis}}^{\omega}}$ with $X_{\omega}$. The separation statement follows from finite abelian duality applied to the finite group $A_{\mathrm{vis}}^{\omega}$ \cite[Part~I, Chap.~2]{Terras1999}.
\end{proof}

\begin{proposition}[Initiality among pure-extension quotients]\label{thm:intrinsic-pure-ext-initiality}
Let
\[
q:A\twoheadrightarrow B
\]
be a surjective homomorphism onto a finite abelian group $B$. In the setting of \Cref{def:finite-nerve-presentation}, the following are equivalent:
\begin{enumerate}[label=(\roman*),nosep]
\item the pushed-forward class $q_*(\omega)\in H^2(N(\mathcal U),B)$ lies in the subgroup
\[
\operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),B);
\]
\item $\eta_B(q_*(\omega))=0$;
\item $q\circ \mathrm{ev}_{\omega}=0$;
\item $\mathrm{Im}(\mathrm{ev}_{\omega})\subseteq \ker q$;
\item $q$ factors uniquely through
\[
\pi_{\omega}:A\twoheadrightarrow A_{\mathrm{vis}}^{\omega}.
\]
\end{enumerate}
Hence $A_{\mathrm{vis}}^{\omega}$ is the initial coefficient quotient through which the $H_2$-visible part of the obstruction disappears, leaving only a pure $\operatorname{Ext}$-type remnant.
\end{proposition}

\begin{proof}
The equivalence of \textup{(i)} and \textup{(ii)} is exactly the universal coefficient sequence for $B$ \cite[Sec.~3.1]{Hatcher2002}. Naturality gives
\[
\eta_B(q_*(\omega))=q\circ \eta_A(\omega)=q\circ \mathrm{ev}_{\omega},
\]
so \textup{(ii)} and \textup{(iii)} are equivalent. Conditions \textup{(iii)} and \textup{(iv)} are equivalent by definition of the image of a homomorphism. By \Cref{def:intrinsic-visible-quotient}, $\mathrm{Im}(\mathrm{ev}_{\omega})=K_{\omega}$, so \textup{(iv)} is equivalent to
\[
K_{\omega}\subseteq \ker q.
\]
By the universal property of the quotient by $K_{\omega}$, this is equivalent to unique factorization of $q$ through $\pi_{\omega}$, proving \textup{(v)}.
\end{proof}

\begin{corollary}[Class-trivializing initiality when $H_1$ vanishes]\label{cor:h1-vanishing-class-trivializing}
Assume
\[
H_1(N(\mathcal U),\mathbb Z)=0.
\]
Then, for every surjective homomorphism $q:A\twoheadrightarrow B$, the following are equivalent:
\begin{enumerate}[label=(\roman*),nosep]
\item $q_*(\omega)=0$ in $H^2(N(\mathcal U),B)$;
\item $q$ factors uniquely through
\[
\pi_{\omega}:A\twoheadrightarrow A_{\mathrm{vis}}^{\omega}.
\]
\end{enumerate}
Hence, when $H_1(N(\mathcal U),\mathbb Z)=0$, the intrinsic visible quotient is initial among coefficient quotients that annihilate the obstruction class itself.
\leanverified{paper\_conservative\_extension\_h1\_vanishing\_class\_trivializing}
\end{corollary}

\begin{proof}
If $H_1(N(\mathcal U),\mathbb Z)=0$, then
\[
\operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),B)=0
\]
for every abelian group $B$. Thus the pure-extension condition in \Cref{thm:intrinsic-pure-ext-initiality}(i) is equivalent to $q_*(\omega)=0$, and the conclusion follows.
\end{proof}

\begin{remark}[Canonical visibility without a canonical splitting]\label{rem:no-canonical-splitting}
The quotient $A_{\mathrm{vis}}^{\omega}$ and the subgroup $K_{\omega}=\mathrm{Im}(\mathrm{ev}_{\omega})$ are canonical. By contrast, the universal coefficient sequence need not split canonically. Thus the theory isolates a canonical visible quotient and a canonical blind subgroup without requiring any noncanonical decomposition of $\omega$ into visible and invisible summands.
\end{remark}

\begin{corollary}[Comparison with strict visible quotients]\label{cor:strict-to-intrinsic-visible}
Let $\alpha$ be any finite obstruction presentation of the class $\omega$. Then the intrinsic visible quotient
\[
A_{\mathrm{vis}}^{\omega}
\]
admits a canonical surjective homomorphism onto the strict visible quotient
\[
A/H_\alpha.
\]
Equivalently, every strict visible quotient is a quotient of the intrinsic visible quotient.
\leanverified{paper_conservative_extension_strict_to_intrinsic_visible}
\end{corollary}

\begin{proof}
Let $q:A\twoheadrightarrow A/H_\alpha$ be the canonical quotient. By \Cref{thm:initial-visible-quotient}, the quotient kills every cocycle value $\alpha_{ijk}$ pointwise, so the pushed-forward cocycle representative is zero. Hence $q_*(\omega)=0$, and in particular it lies in the $\operatorname{Ext}$-subgroup of $H^2(N(\mathcal U),A/H_\alpha)$. Apply \Cref{thm:intrinsic-pure-ext-initiality}.
\end{proof}

\begin{definition}[Chain-level evaluation of a presentation]\label{def:chain-level-evaluation}
Let $\alpha\in Z^2(N(\mathcal U),A)$ be a finite obstruction presentation of the class $\omega$. Write
\[
C_2(N(\mathcal U),\mathbb Z)
\]
for the simplicial group of integral $2$-chains on the finite nerve. The \emph{chain-level evaluation map} of $\alpha$ is the homomorphism
\[
\widetilde\alpha:C_2(N(\mathcal U),\mathbb Z)\to A
\]
defined by linear extension from the oriented $2$-simplices:
\[
\widetilde\alpha\!\left(\sum_{\sigma} n_{\sigma}\sigma\right)
:=
\sum_{\sigma} n_{\sigma}\alpha(\sigma).
\]
\end{definition}

\begin{proposition}[Strict visibility versus homological visibility]\label{prop:chain-vs-homological-visibility}
In the setting of \Cref{def:chain-level-evaluation}, the following hold.
\begin{enumerate}[label=(\roman*),nosep]
\item $\widetilde\alpha$ vanishes on $2$-boundaries and therefore induces a homomorphism
\[
\overline\alpha:H_2(N(\mathcal U),\mathbb Z)\to A.
\]
\item The induced homomorphism $\overline\alpha$ is exactly the homological evaluation map $\mathrm{ev}_{\omega}$.
\item The strict obstruction subgroup and the intrinsic hidden subgroup satisfy
\[
H_\alpha=\mathrm{Im}(\widetilde\alpha),
\qquad
K_{\omega}=\mathrm{Im}(\overline\alpha)=\widetilde\alpha(Z_2(N(\mathcal U),\mathbb Z)).
\]
\end{enumerate}
Thus the gap between strict visibility and intrinsic visibility is exactly the gap between the image of the cocycle on all $2$-chains and its image on $2$-cycles.
\end{proposition}

\begin{proof}
If $c\in C_3(N(\mathcal U),\mathbb Z)$, then
\[
\widetilde\alpha(\partial c)=(\delta\alpha)(c)=0
\]
because $\alpha$ is a cocycle. Hence $\widetilde\alpha$ vanishes on $B_2(N(\mathcal U),\mathbb Z)$ and induces a homomorphism
\[
\overline\alpha:H_2(N(\mathcal U),\mathbb Z)=Z_2/B_2\to A.
\]
This proves \textup{(i)}.

For any $2$-cycle $z=\sum_{\sigma} n_{\sigma}\sigma\in Z_2(N(\mathcal U),\mathbb Z)$, set
\[
\alpha^{\sharp}(z):=\sum_{\sigma} n_{\sigma}\alpha(\sigma)=\widetilde\alpha(z).
\]
If $z'$ is another $2$-cycle homologous to $z$, then $z-z'=\partial c$ for some $c\in C_3(N(\mathcal U),\mathbb Z)$. By \textup{(i)},
\[
\alpha^{\sharp}(z)-\alpha^{\sharp}(z')
=
\widetilde\alpha(z-z')
=
\widetilde\alpha(\partial c)
=
0.
\]
Therefore $\alpha^{\sharp}$ depends only on the homology class of $z$ and defines a homomorphism
\[
H_2(N(\mathcal U),\mathbb Z)\to A,\qquad
[z]\longmapsto \alpha^{\sharp}(z).
\]
By the definition of the universal coefficient map $\eta_A$, this homomorphism is exactly the image of the cohomology class $\omega=[\alpha]$ under
\[
\eta_A:H^2(N(\mathcal U),A)\to \operatorname{Hom}(H_2(N(\mathcal U),\mathbb Z),A).
\]
Thus
\[
\overline\alpha([z])=\alpha^{\sharp}(z)=\eta_A(\omega)([z])=\mathrm{ev}_{\omega}([z])
\]
for every $[z]\in H_2(N(\mathcal U),\mathbb Z)$, proving \textup{(ii)}.

For \textup{(iii)}, the image of $\widetilde\alpha$ is exactly the subgroup generated by the simplex values $\alpha_{ijk}$, hence it is $H_\alpha$ by \Cref{def:finite-obstruction-presentation}. The equality
\[
\mathrm{Im}(\overline\alpha)=\widetilde\alpha(Z_2(N(\mathcal U),\mathbb Z))
\]
is immediate from \textup{(i)}, and \Cref{def:intrinsic-visible-quotient} identifies this image with $K_{\omega}$.
\end{proof}

\begin{corollary}[Criterion for coincidence of strict and intrinsic visibility]\label{cor:strict-intrinsic-coincidence-criterion}
Let $\alpha$ be a finite obstruction presentation of $\omega$. The following are equivalent:
\begin{enumerate}[label=(\roman*),nosep]
\item $A/H_\alpha\cong A_{\mathrm{vis}}^{\omega}$;
\item $H_\alpha=K_{\omega}$;
\item $\widetilde\alpha(C_2(N(\mathcal U),\mathbb Z))=\widetilde\alpha(Z_2(N(\mathcal U),\mathbb Z))$.
\end{enumerate}
Equivalently, strict visibility agrees with intrinsic visibility exactly when every presentation-level visible value already arises from the evaluation of the cocycle on a genuine $2$-cycle.
\leanverified{paper_conservative_extension_strict_intrinsic_coincidence_criterion}
\end{corollary}

\begin{proof}
The equivalence of \textup{(i)} and \textup{(ii)} is immediate from the quotient descriptions
\[
A/H_\alpha
\qquad\text{and}\qquad
A_{\mathrm{vis}}^{\omega}=A/K_{\omega}.
\]
The equivalence of \textup{(ii)} and \textup{(iii)} is exactly \Cref{prop:chain-vs-homological-visibility}(iii).
\end{proof}

\begin{theorem}[Character-blind pure-extension obstructions]\label{thm:character-blind-obstructions}
In the setting of \Cref{def:finite-nerve-presentation}, the following are equivalent:
\begin{enumerate}[label=(\roman*),nosep]
\item $K_{\omega}=0$;
\item $A_{\mathrm{vis}}^{\omega}=A$;
\item $\mathrm{ev}_{\omega}=0$;
\item $\omega$ lies in the subgroup
\[
\operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),A)\subseteq H^2(N(\mathcal U),A).
\]
\end{enumerate}
If these conditions hold and $\omega\neq 0$, then the chosen branch gerbe $\mathfrak G_v$ is non-neutral, while every character of $A$ is class-admissible for that branch. In particular, if $|\Vis_{p,s}(r)|=1$, then
\[
M,p\Vdash \Null^{\mathrm{glue}}_s(r).
\]
\end{theorem}

\begin{proof}
By \Cref{def:intrinsic-visible-quotient},
\[
K_{\omega}=\mathrm{Im}(\mathrm{ev}_{\omega}),
\]
so \textup{(i)} and \textup{(iii)} are equivalent. Since $A_{\mathrm{vis}}^{\omega}=A/K_{\omega}$, conditions \textup{(i)} and \textup{(ii)} are equivalent. By the universal coefficient sequence, \textup{(iii)} holds exactly when $\omega\in\ker(\eta_A)$, and this kernel is the stated $\operatorname{Ext}$-subgroup. Hence all four conditions are equivalent.

If they hold and $\omega\neq 0$, then \Cref{prop:derived-branch-classification} shows that the branch gerbe $\mathfrak G_v$ is non-neutral.
On the other hand, $\mathrm{ev}_{\omega}=0$ implies
\[
\chi\circ \mathrm{ev}_{\omega}=0
\]
for every $\chi\in\widehat A$, so \Cref{thm:intrinsic-visible-quotient} shows that every character is class-admissible.

If $|\Vis_{p,s}(r)|=1$, then the chosen branch $v$ is the unique visible branch and, in particular, $\Vis_{p,s}(r)\neq\varnothing$. Hence \Cref{prop:compsecs-vs-visible-branches} gives $M,p\Vdash\CompSec_s(r)$. If $M,p\Vdash\Sec_s(r)$, then \Cref{lem:global-section-neutral-branch} would produce a neutral visible branch. Since $v$ is the only visible branch, this would force $\mathfrak G_v$ to be neutral, contradicting the non-neutrality just proved. Therefore $M,p\nVdash\Sec_s(r)$, and \Cref{lem:admitted-gluing-null-reduction} gives
\[
M,p\Vdash \Null^{\mathrm{glue}}_s(r).
\]
\leanverified{paper_conservative_extension_character_blind_obstructions}
\end{proof}

\begin{corollary}[Automatic blindness when $H_2$ vanishes]\label{cor:h2-vanishing-blindness}
Assume
\[
H_2(N(\mathcal U),\mathbb Z)=0.
\]
Then, for every obstruction class $\omega\in H^2(N(\mathcal U),A)$,
\[
\mathrm{ev}_{\omega}=0,
\qquad
K_{\omega}=0,
\qquad
A_{\mathrm{vis}}^{\omega}=A.
\]
Every nonzero branch obstruction class on such a nerve is therefore character-blind.
\leanverified{paper_gluing_failure_h2_vanishing_blindness}
\end{corollary}

\begin{proof}
If $H_2(N(\mathcal U),\mathbb Z)=0$, then
\[
\mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),A)=0,
\]
so $\mathrm{ev}_{\omega}=0$ for every $\omega$. Apply \Cref{thm:character-blind-obstructions}.
\end{proof}

\begin{corollary}[No character-blind obstructions when $H_1$ is free]\label{cor:h1-free-no-blindness}
Assume that $H_1(N(\mathcal U),\mathbb Z)$ is free abelian. Then
\[
K_{\omega}=0
\quad\Longleftrightarrow\quad
\omega=0.
\]
Equivalently, every nonzero branch obstruction class has a nontrivial $H_2$-visible part. In particular, if $|\Vis_{p,s}(r)|=1$, every nontrivial gluing obstruction on such a nerve produces a proper intrinsic visible quotient.
\leanverified{paper_conservative_extension_h1_free_no_blindness}
\end{corollary}

\begin{proof}
If $H_1(N(\mathcal U),\mathbb Z)$ is free abelian, then
\[
\operatorname{Ext}^1(H_1(N(\mathcal U),\mathbb Z),A)=0.
\]
Hence the universal coefficient map $\eta_A$ is injective. By \Cref{thm:character-blind-obstructions},
\[
K_{\omega}=0
\quad\Longleftrightarrow\quad
\mathrm{ev}_{\omega}=0
\quad\Longleftrightarrow\quad
\eta_A(\omega)=0.
\]
Injectivity of $\eta_A$ therefore gives $K_{\omega}=0$ if and only if $\omega=0$. If $\omega\neq 0$, then $K_{\omega}\neq 0$, so $A_{\mathrm{vis}}^{\omega}=A/K_{\omega}$ is a proper quotient of the finite group $A$.
\end{proof}

\begin{lemma}[Nonvacuous stackified connected-components witness]\label{lem:nonvacuous-stackified-components}
Assume $r:\mathsf{Ref}_s$ is admitted at $p$, write
\[
a=[r]_p,\qquad
\mathcal C_a:=\mathcal C_{p,s}/a,
\]
and let $\mathfrak P_r\to\mathcal C_a$ be a realization prestack for $r$ with stackification
\[
\mathfrak L_r:=a(\mathfrak P_r).
\]
If
\[
\pi_0^{\mathrm{pre}}(\mathfrak P_r)(a)=\varnothing
\qquad\text{and}\qquad
\pi_0(\mathfrak L_r)(a)\neq\varnothing,
\]
then
\[
M,p\Vdash \CompSec_s(r)\land M,p\nVdash \Sec_s(r).
\]
\end{lemma}

\begin{proof}
The first condition and \Cref{def:realization-prestack}(i) imply
\[
F_{p,s}(a)=\varnothing.
\]
By \Cref{prop:stackification-components}, there is a canonical isomorphism
\[
a_{p,s}F_{p,s}|_{\mathcal C_a}\cong \pi_0(\mathfrak L_r),
\]
so the second condition gives
\[
a_{p,s}F_{p,s}|_{\mathcal C_a}(a)\neq\varnothing.
\]
By \Cref{thm:sheafification-characterization} and \Cref{def:existence-levels}, this is exactly the conjunction
\[
M,p\Vdash \CompSec_s(r)\land M,p\nVdash \Sec_s(r).
\]
\end{proof}

\begin{example}[A character-blind obstruction on an $\mathbb{R}P^2$-type nerve]\label{ex:rp2-character-blind}
Let $\mathcal U$ be a finite cover whose \v{C}ech nerve is a finite triangulation of $\mathbb{R}P^2$, and take
\[
A=\mathbb Z/2\mathbb Z.
\]
Then
\[
H_2(\mathbb{R}P^2,\mathbb Z)=0,
\qquad
H_1(\mathbb{R}P^2,\mathbb Z)\cong \mathbb Z/2\mathbb Z
\]
\cite[Sec.~2.2]{Hatcher2002}. The universal coefficient sequence therefore yields
\[
H^2(\mathbb{R}P^2,\mathbb Z/2\mathbb Z)
\cong
\operatorname{Ext}^1(\mathbb Z/2\mathbb Z,\mathbb Z/2\mathbb Z)
\cong
\mathbb Z/2\mathbb Z
\]
\cite[Sec.~3.1]{Hatcher2002}. Let $\omega$ be the nonzero class. Then
\[
\mathrm{ev}_{\omega}=0,
\qquad
K_{\omega}=0,
\qquad
A_{\mathrm{vis}}^{\omega}=A,
\]
by \Cref{thm:character-blind-obstructions}. Thus no class-level quotient collapse occurs at all, yet the obstruction class is nonzero. Hence this branch is completely character-blind. If it is the unique visible branch of the admitted reference under consideration, \Cref{lem:nonvacuous-stackified-components} and \Cref{thm:gerbe-null-semantics} yield a genuine gluing-level absence that is invisible to every character channel.
\end{example}

\begin{example}[A single-branch maximal-collapse obstruction on $S^2$]\label{ex:s2-single-branch-maximal-collapse}
Let $\mathcal U=\{U_i\to S^2\}_{i\in I}$ be a finite good cover of $S^2$, let $N:=N(\mathcal U)$ be its nerve, and let $\mathcal C_a$ be the corresponding slice site for one admitted reference $r:\mathsf{Ref}_s$ with $a=[r]_p$. Take the constant band
\[
A=\mathbb Z/2\mathbb Z
\]
on $\mathcal C_a$, and let $\mathfrak G$ be an $A$-banded gerbe representing the nonzero class in
\[
H^2(\mathcal C_a,A)\cong H^2(N,A)\cong \mathbb Z/2\mathbb Z
\]
\cite[Chap.~IV]{Giraud1971}, \cite[Tag 06NY]{StacksProject}. Put
\[
\mathfrak P_r:=\mathfrak G,
\qquad
F_{p,s}|_{\mathcal C_a}:=\pi_0^{\mathrm{pre}}(\mathfrak P_r).
\]
Since $S^2$ is paracompact and the chosen family of finite good covers is cofinal, \Cref{thm:good-cover-gerbe-realization} applies. Hence
\[
\Vis_{p,s}(r)=\{v\},
\qquad
\mathfrak G_v=\mathfrak G,
\qquad
M,p\Vdash \Null^{\mathrm{glue}}_s(r).
\]
Moreover, the restriction of $F_{p,s}$ to the good-cover skeleton is a support presheaf, so this example lies in the natural realization class singled out there.

Since
\[
H_1(N,\mathbb Z) = 0,
\qquad
H_2(N,\mathbb Z) \cong \mathbb Z
\]
\cite[Secs.~2.1, 2.2]{Hatcher2002}, the universal coefficient theorem gives
\[
H^2(N, \mathbb{Z}/2\mathbb{Z})
\cong
\mathrm{Hom}(\mathbb{Z}, \mathbb{Z}/2\mathbb{Z})
\cong
\mathbb{Z}/2\mathbb{Z}.
\]
Let $\omega$ be the nonzero class, and let $[S^2]$ denote the fundamental class in $H_2(N,\mathbb Z)$. Then
\[
\mathrm{ev}_\omega([S^2]) = 1 \in \mathbb{Z}/2\mathbb{Z},
\]
so the homological evaluation map is surjective. By \Cref{cor:maximal-collapse},
\[
K_\omega = A = \mathbb{Z}/2\mathbb{Z},
\qquad
A_{\mathrm{vis}}^\omega = 0.
\]
Hence
\[
M,p\Vdash \Null^{\mathrm{glue}}_s(r),
\qquad
\omega=\omega_v\neq 0,
\qquad
\mathrm{ev}_{\omega}:H_2(N,\mathbb Z)\twoheadrightarrow A,
\qquad
K_\omega=A,
\qquad
A_{\mathrm{vis}}^\omega=0.
\]
This one natural model therefore verifies the full hypothesis chain
\[
(\mathcal C_a,\mathfrak P_r,A,\omega,\mathrm{ev}_\omega,K_\omega,A_{\mathrm{vis}}^\omega)
\]
inside a concrete site-theoretic example, with the realization hypotheses supplied internally by \Cref{thm:good-cover-gerbe-realization}. It is the opposite extreme from the $\mathbb{R}P^2$ blind case of \Cref{ex:rp2-character-blind}: there the obstruction is purely $\operatorname{Ext}$-type with maximal $A_{\mathrm{vis}}^\omega = A$, while here it is purely $H_2$-type with $A_{\mathrm{vis}}^\omega = 0$.
\end{example}

\begin{definition}[Intrinsic strictification code]\label{def:intrinsic-strictification-code}
An \emph{intrinsic strictification code} for the obstruction class $\omega$ consists of a finite set $R$ and an injective map
\[
\iota_{\omega}:A\hookrightarrow A_{\mathrm{vis}}^{\omega}\times R
\]
such that
\[
\mathrm{pr}_1\circ\iota_{\omega}=\pi_{\omega}.
\]
\end{definition}

\begin{corollary}[Intrinsic hidden-state lower bound]\label{cor:intrinsic-hidden-state-lower-bound}
Every intrinsic strictification code satisfies
\[
|R|\ge |K_{\omega}|.
\]
Equivalently,
\[
\log_2|R|\ge \log_2|K_{\omega}|.
\]
This lower bound is sharp.
\leanverified{paper_conservative_extension_intrinsic_hidden_state_lower_bound_publication}
\end{corollary}

\begin{proof}
By \Cref{def:intrinsic-visible-quotient}, the fibers of $\pi_{\omega}$ are exactly the cosets of $K_{\omega}$, so every fiber has cardinality $|K_{\omega}|$. The same fiber-counting argument as in \Cref{thm:minimal-information-cost-of-strictification} therefore shows that any injective code preserving the first projection must satisfy $|R|\ge |K_{\omega}|$.

For sharpness, take $R:=K_{\omega}$ and choose a set-theoretic section $s_{\omega}:A_{\mathrm{vis}}^{\omega}\to A$ of $\pi_{\omega}$. Then
\[
\iota_{\omega}(x):=\bigl(\pi_{\omega}(x),x-s_{\omega}(\pi_{\omega}(x))\bigr)
\]
is injective.
\end{proof}

\begin{proposition}[Prime decomposition of intrinsic visibility and budget]\label{thm:intrinsic-prime-decomposition}
Write the finite abelian group $A$ as a direct sum of its $p$-primary components:
\[
A=\bigoplus_p A_{(p)}.
\]
Let $\omega_p\in H^2(N(\mathcal U),A_{(p)})$ be the image of $\omega$ under the projection onto the $p$-primary summand. Then
\[
\mathrm{ev}_{\omega}=\bigoplus_p \mathrm{ev}_{\omega_p},
\qquad
K_{\omega}=\bigoplus_p K_{\omega,p},
\qquad
A_{\mathrm{vis}}^{\omega}\cong \bigoplus_p A_{(p)}/K_{\omega,p},
\]
where
\[
K_{\omega,p}:=\mathrm{Im}(\mathrm{ev}_{\omega_p}).
\]
Consequently,
\[
|K_{\omega}|=\prod_p |K_{\omega,p}|,
\qquad
\log_2|K_{\omega}|=\sum_p \log_2|K_{\omega,p}|.
\]
\leanverified{paper_conservative_extension_intrinsic_prime_decomposition}
\end{proposition}

\begin{proof}
Both $H^2(N(\mathcal U),-)$ and $\mathrm{Hom}(H_2(N(\mathcal U),\mathbb Z),-)$ commute with finite direct sums, so the universal coefficient map is compatible with the primary decomposition of $A$. Hence
\[
\eta_A(\omega)=\bigoplus_p \eta_{A_{(p)}}(\omega_p),
\]
which is exactly the asserted decomposition of $\mathrm{ev}_{\omega}$.

Taking images gives
\[
K_{\omega}
=
\mathrm{Im}(\mathrm{ev}_{\omega})
=
\bigoplus_p \mathrm{Im}(\mathrm{ev}_{\omega_p})
=
\bigoplus_p K_{\omega,p},
\]
by \Cref{def:intrinsic-visible-quotient}. The quotient decomposition follows immediately:
\[
A_{\mathrm{vis}}^{\omega}
=
A/K_{\omega}
\cong
\bigoplus_p A_{(p)}/K_{\omega,p}.
\]
Since finite direct sums correspond to products of cardinalities, the budget formula is immediate.
\end{proof}

\begin{corollary}[Intrinsic budget versus strict budget]\label{cor:strict-intrinsic-budget-comparison}
For every finite obstruction presentation $\alpha$ of the class $\omega$,
\[
|K_{\omega}|\le |H_\alpha|,
\qquad
\log_2|K_{\omega}|\le \log_2|H_\alpha|.
\]
Hence the intrinsic hidden-state budget never exceeds the strict presentation-level budget. Equality holds if and only if the canonical surjection
\[
A_{\mathrm{vis}}^{\omega}\twoheadrightarrow A/H_\alpha
\]
is an isomorphism.
\leanverified{paper_conservative_extension_strict_intrinsic_budget_comparison}
\end{corollary}

\begin{proof}
By \Cref{cor:strict-to-intrinsic-visible}, there is a surjective homomorphism
\[
A_{\mathrm{vis}}^{\omega}=A/K_{\omega}\twoheadrightarrow A/H_\alpha.
\]
For finite groups this implies
\[
|A|/|K_{\omega}|\ge |A|/|H_\alpha|,
\]
hence $|K_{\omega}|\le |H_\alpha|$. The logarithmic inequality is immediate. Equality of cardinalities holds exactly when the displayed surjection is bijective.
\end{proof}

\begin{corollary}[Maximal collapse criterion]\label{cor:maximal-collapse}
In the setting of \Cref{def:finite-nerve-presentation}, the following are equivalent:
\begin{enumerate}[label=(\roman*),nosep]
\item $\mathrm{ev}_{\omega}:H_2(N(\mathcal U),\mathbb Z)\to A$ is surjective;
\item $K_{\omega}=A$;
\item $A_{\mathrm{vis}}^{\omega}=0$;
\item the intrinsic strictification budget is maximal:
\[
\log_2|K_{\omega}|=\log_2|A|.
\]
\end{enumerate}
\end{corollary}

\begin{proof}
By \Cref{def:intrinsic-visible-quotient}, $K_{\omega}=\mathrm{Im}(\mathrm{ev}_{\omega})$, so \textup{(i)} and \textup{(ii)} are equivalent. Condition \textup{(ii)} is equivalent to \textup{(iii)} by the definition of $A_{\mathrm{vis}}^{\omega}=A/K_{\omega}$. For finite groups, \textup{(ii)} is also equivalent to $|K_{\omega}|=|A|$, which is exactly \textup{(iv)} after taking $\log_2$.
\end{proof}
