\appendix
\section{Finite-State Complexity Upper Bounds}\label{sec:appendix-complexity}

This appendix is supplementary to the main proof chain. It records only routine finite-state upper bounds for the support notions introduced in \Cref{sec:multiaxis-refinement}.

For this appendix, assume that the multi-axis structure is finite and that a distinguished formula $\chi$ is part of the input together with polynomial-time decidable tables for the relations
\[
q\mapsto (M,q\Vdash\chi)
\qquad\text{and}\qquad
q\mapsto (M,q\Vdash\neg\chi).
\]
This is the natural finite encoding once the underlying semantic state graph has already been computed and the support problems are asked over that graph. No matching lower bounds are claimed here.

We consider the decision problems \textup{\textsc{Support}}, \textup{\textsc{Indispensable}}, and \textup{\textsc{MinSupport}} associated with \Cref{def:axis-support,def:axis-indispensability}. For \textup{\textsc{MinSupport}} we use the standard class $\mathrm{D}^P$ in the sense of Papadimitriou and Yannakakis \cite{PapadimitriouYannakakis1984}.

\begin{theorem}[Complexity upper bounds]\label{thm:axis-complexity-upper-bounds}
Under the finite encoding above, the following hold.
\begin{enumerate}[label=(\roman*),nosep]
\item \textup{\textsc{Support}} belongs to coNP.
\item \textup{\textsc{Indispensable}} belongs to NP.
\item \textup{\textsc{MinSupport}} belongs to $\mathrm{D}^P$.
\end{enumerate}
\end{theorem}

\begin{proof}
For \textup{\textsc{Support}}, non-support is witnessed by a pair $(q,r)$ of $K$-reachable states that agree on the relevant $K$-information, both decide $\chi$, and force opposite truth values. Reachability, agreement of axis information, and decision of $\chi$ are all polynomial-time checkable in the finite input, so the complement of \textup{\textsc{Support}} is in NP.

For \textup{\textsc{Indispensable}}, a witness is again a pair $(q,r)$, now only required to agree on all axes except the distinguished axis $j$ and to force opposite truth values of $\chi$. This certificate is polynomially checkable, so \textup{\textsc{Indispensable}} lies in NP.

For \textup{\textsc{MinSupport}}, the condition that $K$ supports $\chi$ is in coNP by the first clause. The condition that every axis in $K$ is indispensable is in NP, because one may guess one witness pair for each axis $j\in K$, and the total certificate size remains polynomial. By \Cref{prop:indispensability-implies-minimality,cor:minimal-support-characterization}, the conjunction of these two conditions is equivalent to minimality, so \textup{\textsc{MinSupport}} lies in $\mathrm{D}^P$.
\end{proof}
