\section{Introduction}\label{sec:introduction}

Standard sheaf-theoretic semantics captures local-to-global passage: compatible local sections glue when the underlying presheaf is a sheaf, and fail to glue when it is not. The obstruction to gluing is recorded by cohomology classes, and the sheaf-theoretic analysis of contextuality \cite{AbramskyBrandenburger2011} treats the absence of a global section as the central semantic defect. Related refinements include stochastic and dynamic presheaf semantics for contextuality \cite{Kishida2014} and Kishida's logic of local inference for the local-consistency side of contextuality \cite{Kishida2016}. The lower-fragment comparison considered in this paper is narrower: within the typed reference-aware $\mathbb L_1$ comparison language, the distinction between compatible local sectionability and global sectionability is invisible under singleton-realization evaluation. A formula in that fragment can register only lower reference-aware data; under the syntactic restrictions made precise later, it does not detect that compatible local data exist but fail to glue.

The present work sits in the forcing and sheaf-theoretic tradition in logic, building on the Kripke semantics for intuitionistic logic \cite{Kripke1965}, Beth's semantic tableaux constructions \cite{Beth1956}, Fitting's model-theoretic forcing \cite{Fitting1969}, Tierney's early topos-theoretic forcing perspective \cite{Tierney1972}, and the later expositions of Goldblatt and Johnstone \cite{Goldblatt2006,Johnstone2002}. In relation to recent work in this journal on embeddings of richer semantic languages into predicate logic \cite{BarberoGalliani2022}, our first theorem proves a deliberately restricted opposite-direction separation: the local-object predicates used below are not recovered inside the typed $\mathbb L_1$ reference-aware comparison fragment. What distinguishes our setting is the explicit stratification of the semantics into conservative extension layers, so that each enrichment---from pointwise truth to reference-aware information-state forcing, from forcing to local objects, from local objects to gluing obstructions---is accompanied by a precise invariance guarantee.

The novelty claimed here is correspondingly limited. The sheafification criterion, gerbe classification input, and universal coefficient algebra are standard tools; the contribution is the way they are separated into typed conservative layers, the restricted lower-fragment undefinability result at the $\mathbb L_1$--$\mathbb L_2$ boundary, the conditional branch-indexed gerbe packaging, and the semantic reading of the homological evaluation quotient. None of the main theorems is presented as a new sheafification theorem, a new Giraud classification theorem, or a new universal coefficient theorem.

The theorem-by-theorem comparison is as follows. Theorem~A should be read against recent APAL work on embeddings of richer team-style languages into predicate logic \cite{BarberoGalliani2022}: it proves only a restricted opposite-direction separation, namely non-recovery of the local-object predicates inside the typed lower fragment $\Frm_1^{\mathrm{ref}}(s)$. Theorem~B packages the standard gerbe classification machinery in a branch-indexed form, but only after the gluing-sensitive realization and comparison hypotheses have been supplied; \Cref{thm:good-cover-gerbe-realization} then verifies that package on a natural in-paper class of banded-gerbe models on paracompact good-cover sites. Theorem~C uses standard universal-coefficient algebra to isolate the semantically visible quotient and the pure $\operatorname{Ext}$ blind case, connecting that analysis back to the cohomological contextuality literature \cite{AbramskyMansfieldSoaresBarbosa2012,AbramskyBarbosaKishidaLalMansfield2015,Caru2017} and to Kishida's dynamic and local-inference refinements \cite{Kishida2014,Kishida2016}.

\begin{center}
\begin{tabular}{p{0.14\linewidth}p{0.14\linewidth}p{0.30\linewidth}p{0.30\linewidth}}
\hline
result & novelty rating & closest background used here & precise increment claimed\\
\hline
Theorem~A & low & APAL embedding and definability work for richer team-style languages \cite{BarberoGalliani2022} & a restricted opposite-direction non-definability result for $\CompSec_s$, $\Sec_s$, and $\Null_s^{\mathrm{glue}}$ inside the typed reference-aware fragment $\Frm_1^{\mathrm{ref}}(s)$\\
Theorem~B & medium, conditional but instantiated on good-cover gerbe models & standard stackification and banded-gerbe classification \cite{Giraud1971,StacksProject} & branch-indexed packaging of visible components, with gluing-level absence requiring nonempty visible branches and non-neutrality of every component gerbe\\
Theorem~C & principal contribution & universal coefficient algebra and finite duality \cite{Hatcher2002,Terras1999,Weibel1994} & semantic identification of the quotient detected by $\mathrm{ev}_\omega$ and of the pure $\operatorname{Ext}$ regime in which class obstructions are character-blind\\
\hline
\end{tabular}
\end{center}

The novelty level is correspondingly uneven. Theorem~A has low novelty but is proof-critical: its sheafification half is standard, and its separation claim is only the fixed-reduct non-definability statement for the typed lower fragment. Theorem~B is a conditional synthesis of standard stack and gerbe tools in a branch-indexed package; it should not be read as an unconditional realization theorem, but the paper now verifies its full hypothesis package on paracompact good-cover gerbe models in \Cref{thm:good-cover-gerbe-realization}. Theorem~C is the main new contribution of the paper: it identifies the intrinsic quotient detected by homological evaluation and the pure $\operatorname{Ext}$ blind regime in the semantic language of visible obstruction data.

This gap motivates the present paper. We construct a five-layer conservative extension chain, equivalently a four-step chain of adjacent conservative extensions,
\[
\mathbb{L}_0 \preceq \mathbb{L}_1 \preceq \mathbb{L}_2 \preceq \mathbb{L}_3 \preceq \mathbb{L}_4,
\]
where $\mathbb{L}_0$ is pointwise semantics, $\mathbb{L}_1$ adds information-state forcing together with reference sorts and admissibility predicates, $\mathbb{L}_2$ adds sites on the existing admitted reference classes, local object functors, and the sectionability predicates, $\mathbb{L}_3$ adds structural absence and gluing-obstruction data, and $\mathbb{L}_4$ adds refinement dynamics. Thus $\mathsf{Ref}_s$ and $\Adm_s$ are not introduced at $\mathbb L_2$; they are already $\Sigma_1$ symbols, and the $\mathbb L_2\to\mathbb L_1$ reduct retains their domains and interpretations. The adjacent embeddings, forgetful maps, and state projections are spelled out in \Cref{prop:chain-instantiation}, and repeated composition gives the layer-preservation statement recorded in \Cref{cor:whole-chain-invariance}. The chain is the minimal framework in which the restricted logical separation theorem, the conditional gerbe semantics, and the homological visibility quotient can all be stated together.

\medskip
\noindent\textbf{Research question.}
Once an admitted reference carries a finite abelian gluing obstruction, what part of that obstruction is semantically visible without enlarging the state space, and how much additional information is required to recover the hidden part?

\medskip
\noindent\textbf{Main results.}
The answer is delivered by three groups of results, each located at a specific boundary of the extension chain. The homological result is the main mathematical payoff; the forcing layer isolates the logical comparison class, and the gerbe layer is retained as a conditional semantic framework making the later hypotheses explicit.

\medskip
\noindent\textbf{Theorem~A} (forcing necessity; $\mathbb{L}_1$--$\mathbb{L}_2$ boundary).
Compatible local sectionability is characterized by sheafification (\Cref{thm:sheafification-characterization}), and the predicates $\CompSec_s$, $\Sec_s$, and $\Null^{\mathrm{glue}}_s$ are not definable by parameter-free formulas of the reference-aware fragment $\Frm_1^{\mathrm{ref}}(s)$ within the syntactic comparison class fixed in \Cref{prop:theorem-a-boundary-typing,cor:layer-correct-gluing-boundary,thm:pointwise-irreducibility}. Here $\Frm_1^{\mathrm{ref}}(s)$ is not the original pointwise language: it is the $\Sigma_1$ fragment with variables of sort $\mathsf{Ref}_s$ and the admissibility predicate $\Adm_s$, but without the $\Sigma_2$ local-object predicates. The admitted reference classes on which the later sites are built are already lower-layer classes in the sense of \Cref{def:lower-reference-data}. The proof keeps the $\Sigma_1$ reduct, state, admitted reference, site, and covers fixed and changes only the local object functor; lower formulas are therefore invariant by reduct preservation, while the local predicates change truth value. This remains true for lower formulas using reference equality, $\Adm_s$, and quantification over $\mathsf{Ref}_s$, because all of that data is held fixed in the common lower reduct. This separation result justifies the enrichment on logical grounds: under those explicit hypotheses, the gluing predicates that drive the obstruction theory are invisible at the typed lower boundary between $\mathbb{L}_1$ and $\mathbb{L}_2$.
Thus the displayed free variable in Theorem~A is well typed before any local-object structure is present; what is added at $\Sigma_2$ is not the reference sort, but the ability to inspect compatible local families and global sectionability.

\medskip
\noindent\textbf{Theorem~B} (conditional branched gerbe semantics; $\mathbb{L}_2$--$\mathbb{L}_3$ boundary).
The gerbe layer is conditional on a gluing-sensitive realization prestack whose stackification recovers the sheafification of isomorphism classes, is globally conservative at the terminal fibre, and is equipped with a cofinal \v{C}ech-to-derived comparison family of covers. Under those hypotheses, visible value classes are the global connected components of the realization stack (\Cref{thm:visible-value-components}), each visible branch $v$ determines a banded component gerbe $\mathfrak{G}_v$ (\Cref{thm:component-gerbe-decomposition}), and gluing-level absence holds exactly when visible branches exist but every component gerbe is non-neutral (\Cref{thm:gerbe-null-semantics}). The conditional package is realized internally for banded gerbes on paracompact good-cover sites by \Cref{thm:good-cover-gerbe-realization}.

\medskip
\noindent\textbf{Theorem~C} (homological visibility; at $\mathbb{L}_3$).
Under explicit finite-nerve and realizability hypotheses, and after restricting to covers on which the finite abelian band is trivialized, the universal coefficient theorem yields a homological evaluation map $\mathrm{ev}_\omega : H_2(N(\mathcal{U}),\mathbb{Z}) \to A$ whose image is the intrinsic hidden subgroup $K_\omega$ and whose cokernel is the intrinsic visible quotient $A_{\mathrm{vis}}^\omega$. The preparatory algebra is standard; the substantive point is the semantic interpretation: the class-admissible characters are exactly the annihilator of that image, and the blind residual---the part of the obstruction invisible to every admissible character---is precisely the pure $\operatorname{Ext}$-type regime singled out in \Cref{thm:character-blind-obstructions}. In the unique-branch support-presheaf setting, the theory yields a formal comparison with the Abramsky--Brandenburger no-global-section picture under the same explicit hypotheses (\Cref{thm:unique-branch-contextuality-comparison}).

\medskip
\noindent\textbf{Secondary consequences.}
The multi-branch theory yields an exact comparison between branch-sensitive and branch-uniform visibility: the canonical surjection
\[
A_{\mathrm{vis}}^{\mathrm{br}}(r) \twoheadrightarrow A_{\mathrm{vis}}^{\mathrm{unif}}(r)
\]
has kernel $(\sum_v K_v)/(\bigcap_v K_v)$, so the hidden budget behind uniform visibility factors into a branch-resolution term and a within-branch residual (\Cref{thm:branch-comparison-exact-sequence,thm:branch-resolution-budget,cor:uniform-hidden-budget-factorization}). Worked finite examples (\Cref{ex:one-bit-hidden-obstruction,ex:rp2-character-blind,ex:s2-single-branch-maximal-collapse,ex:two-branch-sphere-separation}) separate the visible, blind, and multi-branch regimes.

\medskip
\noindent\textbf{Hypothesis stratification.}
The hypotheses used in these results are stratified rather than cumulative. Reference sorts and admissibility predicates belong already to $\Sigma_1$; only the site, cover, local-object, and sectionability symbols enter at $\Sigma_2$. The site-theoretic assumptions enter at that local-object layer: existing admitted reference classes are organized as a thin category equipped with a Grothendieck pretopology and the pullbacks needed for matching families. Branch constancy enters only for the decomposition corollary (\Cref{cor:branch-decomposition}), not for the component-gerbe theorem itself. Global conservativity is used only when stack-global objects are compared back to the original local object functor. Finite nerve presentations appear only in the finite abelian visibility theorems, and those theorems are explicitly restricted to covers on which the band is trivialized; nontrivial monodromy would require local coefficient systems. The logical separation result is a typed lower-fragment non-definability theorem, the gerbe semantics is conditional on gluing-sensitive realization data together with a \v{C}ech-to-derived comparison hypothesis, and the cohomological statements are conditional on finite-nerve realizability together with a fixed cohomological model.

The dependency map is the following. Theorem~A uses only the $\Sigma_1$ placement of references and admissibility, the $\Sigma_2$ site on admitted reference classes, and the parameter-free lower-fragment comparison. Theorem~B additionally uses a gluing-sensitive realization prestack, global conservativity at the terminal fibre, and the cofinal \v{C}ech-to-derived comparison covers; \Cref{thm:good-cover-gerbe-realization} shows that these hypotheses are satisfied on paracompact good-cover gerbe models. Theorem~C uses Theorem~B's branch classes only after a finite nerve presentation and a coverwise trivialization of the finite abelian band have been fixed. The unique-branch contextuality comparison adds the external identification with a support presheaf and does not claim that arbitrary empirical models satisfy the realization hypotheses automatically.

\begin{center}
\begin{tabular}{p{0.42\linewidth}p{0.25\linewidth}p{0.20\linewidth}}
\hline
input & first layer where it appears & used in\\
\hline
$\mathsf{Ref}_s,\ \Adm_s$ & $\Sigma_1$ & Theorem~A\\
$\mathcal C_{p,s},\Cov_{p,s},F_{p,s},\LocSec_s,\CompSec_s,\Sec_s$ & $\Sigma_2$ & Theorems~A--C\\
gluing-sensitive realization and global conservativity & $\Sigma_3$ data & Theorem~B\\
\v{C}ech-to-derived comparison covers & $\Sigma_3$ hypothesis & Theorem~B\\
finite nerve and trivialized finite abelian band & homological presentation & Theorem~C\\
\hline
\end{tabular}
\end{center}

\medskip
\noindent\textbf{Hypothesis ledger.}
\begin{enumerate}[label=(\roman*),nosep]
\item The placement $\mathsf{Ref}_s,\Adm_s\in\Sigma_1$ is used in Theorem~A and fixed in \Cref{prop:theorem-a-boundary-typing,prop:no-reference-increment,def:reference-aware-boundary}.
\item The pretopological site on admitted reference classes with pullbacks is used in Theorems~A--C and fixed in \Cref{def:local-reference-frame}.
\item Gluing-sensitive realization and global conservativity are used in Theorem~B and fixed in \Cref{thm:gluing-sensitive-realization,def:global-conservativity}.
\item Cofinal \v{C}ech-to-derived comparison covers are used in Theorem~B and fixed in \Cref{def:cech-gluing-obstruction}.
\item The finite nerve and trivialized finite abelian band are used in Theorem~C and fixed in \Cref{def:finite-nerve-presentation}.
\end{enumerate}

\medskip
\noindent\textbf{Plan of the paper.}
\Cref{sec:preliminaries} defines logic layers and conservative extensions and introduces the five-layer chain, together with its concrete adjacent interfaces. \Cref{sec:information-states} defines information states, forcing, and lifted pointwise soundness. \Cref{sec:null-decomposition} upgrades the reference-aware state-forcing layer to local geometry on admitted reference classes, proves the sheafification characterization, states the restricted forcing-necessity theorem (Theorem~A), and isolates the additional gluing-sensitive realization hypotheses needed later. The gerbe-obstruction subsection establishes the conditional branched gerbe semantics (Theorem~B) under explicit \v{C}ech-comparison hypotheses and then verifies that package on good-cover gerbe models. The homological-visibility subsection fixes the cohomological conventions for finite nerve presentations and then establishes the intrinsic visible quotient and the character-blind obstruction theorem (Theorem~C), together with the multi-branch comparison sequence and a formal contextuality dictionary. \Cref{sec:multiaxis-refinement} records supplementary refinement monotonicity results at $\mathbb{L}_4$, and a short appendix records finite-state upper bounds.
