\section{Preliminaries}\label{sec:preliminaries}

The notation in this section follows the hierarchy of the paper. We
first fix primitive orbit counts for the base shift, then the
finite-group extension and Adams operations used in the main theorem
package. The reduced determinant, finite part, and cyclic-lift notation
come afterwards because they serve only as scalar normalisation and
canonical cyclic-lift comparison data.

\subsection{Mixing subshifts and primitive orbit counts}

Our standing symbolic system is a finite directed multigraph
\(\Gamma=(V,E_A)\).  We write \(o(e)\) and \(t(e)\) for the initial and
terminal vertices of an edge and
\[
  A_{ij}:=\#\{e\in E_A:o(e)=i,\ t(e)=j\}
\]
for its non-negative integer adjacency matrix.  The matrix \(A\) is
assumed primitive.  The associated one-sided edge shift is
\[
  X_A^+
  =
  \bigl\{(e_n)_{n\ge 0}\in E_A^{\NN}:
    t(e_n)=o(e_{n+1})\ \text{for all }n\ge 0\bigr\},
\]
equipped with the left shift \(\sigma\).  Thus closed length-\(n\) edge
paths are the same as fixed points of \(\sigma^n\), and we write
\[
  a_n(A):=\Tr(A^n)=\#\operatorname{Fix}(\sigma^n)
\]
for the number of points of period dividing \(n\). The number of
primitive orbits of least period \(n\) is
\begin{equation}\label{eq:primitive-mobius}
  p_n(A)=\frac{1}{n}\sum_{d\mid n}\mu(d)\,a_{n/d}(A),
\end{equation}
and the necklace identity reads
\begin{equation}\label{eq:trace-primitive-identity}
  a_n(A)=\sum_{d\mid n}d\,p_d(A).
\end{equation}

This is the convention in the finite-group sections: \(E_A\) is an edge
set and primitive orbits are orbits in the edge shift.  In the scalar
finite-matrix sections and appendices we also allow \(A\) to be an
arbitrary primitive non-negative real matrix.  There \(p_n(A)\) denotes
the algebraic M\"obius coefficient in \eqref{eq:primitive-mobius}; it is
an actual non-negative orbit count only when \(A\) is the adjacency
matrix of a finite directed multigraph.

\begin{proposition}[Trace, primitive orbits, and Euler product]
  \label{prop:prelim-trace-primitive}
  For the adjacency matrix \(A\) of a primitive finite directed
  multigraph, the dynamical \(\zeta\)-function
  satisfies
  \begin{equation}\label{eq:zeta-def}
    \zeta_A(z)
    :=\exp\!\left(\sum_{n\ge 1}\frac{\Tr(A^n)}{n}z^n\right)
    =\frac{1}{\det(I-zA)}
    =\prod_{n\ge 1}(1-z^n)^{-p_n(A)}.
  \end{equation}
\end{proposition}

\begin{proof}
  The last equality follows from \eqref{eq:trace-primitive-identity},
  since
  \[
    \sum_{n\ge 1}\frac{a_n(A)}{n}z^n
    =\sum_{m\ge 1}\sum_{k\ge 1}\frac{p_m(A)}{k}z^{mk}
    =-\sum_{m\ge 1}p_m(A)\log(1-z^m).
  \]
  The determinant identity is standard for finite matrices.
\end{proof}

The finite-group Adams--M\"obius arguments only need the following
uniform Perron--Frobenius trace bound, which remains valid even when
\(\lambda=1\).

\begin{lemma}[Perron--Frobenius trace bound]
  \label{lem:pf-trace-bound}
  Let \(A\) be a primitive non-negative \(d\times d\) matrix with
  Perron root \(\lambda=\rho(A)\). Then
  \[
    \abs{\Tr(A^n)}\le d\,\lambda^n
    \qquad (n\ge 1).
  \]
  In particular, there is a constant \(C_A>0\) such that
  \[
    \abs{\Tr(A^n)}\le C_A\lambda^n
    \qquad (n\ge 1).
  \]
\end{lemma}

\begin{proof}
  Let \(\lambda,\mu_2,\dots,\mu_d\) be the eigenvalues of \(A\), counted
  with algebraic multiplicity. Since \(A\) is primitive,
  Perron--Frobenius theory gives \(\lambda>0\), the eigenvalue
  \(\lambda\) is simple, and \(|\mu_j|<\lambda\) for \(j\ge 2\). Hence
  \[
    \abs{\Tr(A^n)}
    =
    \abs{\lambda^n+\sum_{j=2}^d\mu_j^n}
    \le
    \lambda^n+\sum_{j=2}^d|\mu_j|^n
    \le
    d\,\lambda^n,
  \]
  which proves the claim.
\end{proof}

If \(\lambda=\rho(A)\) is the Perron root and \(\lambda>1\), then
\[
  \Lambda(A):=\max\{|\mu|:\mu\in\spec(A),\ \mu\neq\lambda\},
\]
with the maximum interpreted as \(0\) if there are no non-Perron
eigenvalues,
then the M\"obius coefficients \(p_n(A)\) enjoy the usual square-root
loss from the higher divisors in \eqref{eq:primitive-mobius}.

\begin{lemma}[Primitive M\"obius asymptotic]
  \label{lem:primitive-orbit-asymptotic}
  Let \(A\) be a primitive non-negative matrix with Perron root
  \(\lambda>1\), and let
  \[
    \Lambda(A):=\max\{|\mu|:\mu\in\spec(A),\ \mu\neq\lambda\}.
  \]
  If the set is empty, put \(\Lambda(A)=0\).
  Then
  \[
    \Tr(A^n)=\lambda^n+O(\Lambda(A)^n)
  \]
  and
  \[
    K_{\mathrm{PF}}(A)
    :=
    \begin{cases}
      \displaystyle
      \sup_{n\ge 1}
        \frac{\abs{\Tr(A^n)-\lambda^n}}{\Lambda(A)^n},
      & \Lambda(A)>0, \\[2ex]
      0,
      & \Lambda(A)=0,
    \end{cases}
    \qquad
    K_{\mathrm{tr}}(A)
    :=
    \sup_{\ell\ge 1}\frac{\abs{\Tr(A^\ell)}}{\lambda^\ell}
  \]
  are finite. Moreover,
  \[
    \abs{p_n(A)-\lambda^n/n}
    \le
    \frac{K_{\mathrm{PF}}(A)\Lambda(A)^n}{n}
    +\frac{K_{\mathrm{tr}}(A)}{n}
      \sum_{\substack{m\mid n\\ m>1}}\lambda^{n/m}
  \]
  and therefore
  \[
    \abs{p_n(A)-\lambda^n/n}
    \le
    \frac{K_{\mathrm{PF}}(A)\Lambda(A)^n}{n}
    +\frac{\lambda K_{\mathrm{tr}}(A)}{\lambda-1}
      \frac{\lambda^{n/2}}{n}.
  \]
  Hence
  \[
    p_n(A)
    =
    \frac{\lambda^n}{n}
    +O\!\left(
      \frac{\max\{\Lambda(A)^n,\lambda^{n/2}\}}{n}
    \right).
  \]
  Consequently, with
  \[
    \vartheta_A:=\max\{\Lambda(A)/\lambda,\lambda^{-1/2}\}<1,
  \]
  one has
  \[
    \frac{p_n(A)}{\lambda^n}
    =
    \frac{1}{n}+O\!\left(\frac{\vartheta_A^n}{n}\right),
  \]
  and the series
  \[
    \sum_{n\ge 1}
      \left(\frac{p_n(A)}{\lambda^n}-\frac{1}{n}\right)
  \]
  converges absolutely.
\end{lemma}

\begin{proof}
  List the eigenvalues of \(A\) with algebraic multiplicity as
  \(\lambda,\mu_2,\dots,\mu_d\). Since \(A\) is primitive,
  Perron--Frobenius theory gives that \(\lambda\) is simple and
  \(|\mu_j|<\lambda\) for \(j\ge 2\). For a finite matrix,
  the trace of \(A^n\) is the sum of the \(n\)-th powers of its
  eigenvalues, counted with algebraic multiplicity. Hence
  \[
    \Tr(A^n)
    =
    \lambda^n+\sum_{j=2}^d\mu_j^n
    =
    \lambda^n+O(\Lambda(A)^n).
  \]
  If \(\Lambda(A)=0\), then every non-Perron eigenvalue is \(0\), so
  \(\Tr(A^n)=\lambda^n\) for all \(n\ge 1\), and the convention
  \(K_{\mathrm{PF}}(A)=0\) is exact. If \(\Lambda(A)>0\), then
  \[
    \frac{\abs{\Tr(A^n)-\lambda^n}}{\Lambda(A)^n}
    \le
    \sum_{j=2}^d\left|\frac{\mu_j}{\Lambda(A)}\right|^n
    \le d-1,
  \]
  so \(K_{\mathrm{PF}}(A)<\infty\). The trace bound from
  \Cref{lem:pf-trace-bound} gives \(K_{\mathrm{tr}}(A)\le d\), so
  \(K_{\mathrm{tr}}(A)<\infty\).

  Insert this estimate into the M\"obius formula
  \[
    p_n(A)=\frac{1}{n}\sum_{m\mid n}\mu(m)\Tr(A^{n/m}).
  \]
  The divisor \(m=1\) contributes \(\lambda^n/n\), with error bounded
  by \(K_{\mathrm{PF}}(A)\Lambda(A)^n/n\). For the remaining divisors,
  \(m>1\) implies \(n/m\le n/2\), and the definition of
  \(K_{\mathrm{tr}}(A)\) gives
  \[
    \sum_{\substack{m\mid n\\ m>1}}\abs{\Tr(A^{n/m})}
    \le
    K_{\mathrm{tr}}(A)\sum_{1\le k\le n/2}\lambda^k
    \le
    \frac{\lambda K_{\mathrm{tr}}(A)}{\lambda-1}\lambda^{n/2}.
  \]
  Dividing by \(n\) gives the two explicit displayed bounds and hence
  the stated M\"obius-coefficient asymptotic.
  Since \(\Lambda(A)<\lambda\) and \(\lambda>1\), the number
  \(\vartheta_A\) lies in \((0,1)\), and the final two assertions follow
  by dividing the asymptotic by \(\lambda^n\).
\end{proof}

\begin{definition}[Square-root bound]\label{def:kernel-rh}
We say that \(A\) satisfies the \emph{square-root bound} if
\[
  \Lambda(A)\le \lambda^{1/2}.
\]
\end{definition}

\subsection{Finite-group extensions and Adams operations}

Let \(G\) be a finite group, let \(\tau:E_A\to G\) be an edge cocycle,
and let \(\widetilde\Gamma\) be the skew-product extension on vertex set
\(V\times G\), with one edge
\[
  (i,g)\longrightarrow (j,g\tau(e))
  \qquad \text{for each } e\in E_A \text{ with } o(e)=i,\ t(e)=j.
\]
Its adjacency matrix is denoted by \(\widetilde A\).

For an irreducible unitary representation
\(\varrho:G\to U(V_\varrho)\) with character \(\chi_\varrho\), define
the twisted transition matrix
\begin{equation}\label{eq:twisted-matrix-def}
  (A_\varrho)_{ij}:=\sum_{e:i\to j}\varrho(\tau(e)),
\end{equation}
viewed as an operator on \(\CC^V\otimes V_\varrho\).

If \(\gamma=e_1\cdots e_n\) is a primitive orbit, write
\[
  \tau(\gamma)=\tau(e_1)\cdots\tau(e_n)\in G
\]
for its group label. A different choice of the initial edge conjugates
this element by a prefix product, so the conjugacy class of
\(\tau(\gamma)\) is intrinsic. All orbit-level evaluations below are
therefore made with class functions. For a class function
\(\phi:G\to\CC\), write
\begin{equation}\label{eq:class-function-fourier}
  \phi
  =\sum_{\varrho\in\Irr(G)}\widehat\phi(\varrho)\chi_\varrho,
  \qquad
  \widehat\phi(\varrho)
  :=\frac{1}{|G|}\sum_{g\in G}\phi(g)\overline{\chi_\varrho(g)}.
\end{equation}
With this Hermitian Fourier convention, the coefficient attached to a
conjugacy class \(C\subset G\) is the conjugated character value
\(\overline{\chi_\varrho(C)}=\chi_\varrho(C^{-1})\). Whenever inverse
classes appear below, they encode this conjugation; we do not identify
\(\chi_\varrho(C)\) with \(\chi_\varrho(C^{-1})\) unless
\(\chi_\varrho\) is real-valued.
For \(m\ge 1\), define the Adams operation on class functions by
\begin{equation}\label{eq:adams-class-function}
  (\psi^m\phi)(g):=\phi(g^m).
\end{equation}

\subsection{Reduced determinants and finite parts}

Let \(z_\ast=\lambda^{-1}\). Since \(\lambda\) is a simple eigenvalue,
\(\zeta_A\) has a simple pole at \(z_\ast\), and we define the residue
constant
\begin{equation}\label{eq:residue-constant-def}
  C(A):=\lim_{z\to z_\ast}(1-\lambda z)\zeta_A(z).
\end{equation}
If \(\mu_2,\dots,\mu_d\) are the non-Perron eigenvalues, then
\begin{equation}\label{eq:reduced-det-poly}
  F_A(t):=\frac{\det(I-tA/\lambda)}{1-t}
  =\prod_{j=2}^d\left(1-\frac{\mu_j}{\lambda}t\right).
\end{equation}
We call \(F_A\) the \emph{reduced spectral polynomial}. Moreover,
\begin{equation}\label{eq:reduced-det-constant}
  C(A)=\det\nolimits'(I-A/\lambda)^{-1}=F_A(1)^{-1}.
\end{equation}

\begin{lemma}[Reduced constant at a simple eigenvalue]
  \label{lem:simple-eigenvalue-reduced-constant}
  Let \(M\) be a complex square matrix and let
  \(\alpha\neq 0\) be a simple eigenvalue of \(M\). Define
  \[
    C_\alpha(M)
    :=\lim_{z\to \alpha^{-1}}
      (1-\alpha z)\det(I-zM)^{-1}.
  \]
  Then the limit exists and
  \[
    C_\alpha(M)
    =
    \prod_{\substack{\nu\in\spec(M)\\ \nu\neq\alpha}}
      \left(1-\frac{\nu}{\alpha}\right)^{-1},
  \]
  with algebraic multiplicities in the product. In particular, if \(A\)
  is primitive with Perron root \(\lambda\), then
  \(C(A)=C_\lambda(A)\).
\end{lemma}

\begin{proof}
  Since \(\alpha\) has algebraic multiplicity one,
  \[
    \det(I-zM)
    =(1-\alpha z)
      \prod_{\substack{\nu\in\spec(M)\\ \nu\neq\alpha}}(1-\nu z),
  \]
  where the remaining eigenvalues are listed with algebraic
  multiplicity. Multiplying by \(1-\alpha z\) after inverting and then
  taking \(z\to \alpha^{-1}\) gives the displayed formula. For a
  primitive matrix, the Perron root is simple, so the definition agrees
  with \eqref{eq:residue-constant-def}.
\end{proof}

When \(\lambda>1\), define the primitive Abel generating series
\[
  P_A(r):=\sum_{n\ge 1}p_n(A)\lambda^{-n}r^n,
  \qquad 0\le r<1.
\]
By \Cref{lem:primitive-orbit-asymptotic},
\(p_n(A)\lambda^{-n}=1/n+O(\vartheta^n/n)\) for some
\(\vartheta\in(0,1)\), the renormalised expression
\[
  P_A(r)+\log(1-r)
\]
has a finite limit as \(r\uparrow 1\).

\begin{definition}[Abel finite-part constant]\label{def:abel-finite-part}
For a primitive matrix \(A\) with Perron root \(\lambda>1\), we write
\[
  \FP(A):=\lim_{r\uparrow 1}\bigl(P_A(r)+\log(1-r)\bigr).
\]
The associated orbit-Mertens constant is
\[
  \mathcal M(A):=\gamma+\FP(A).
\]
\end{definition}

\subsection{Cyclic lifts}

For \(q\ge 1\) let \(P_q\) be the \(q\)-cycle permutation matrix and
define the cyclic lift
\begin{equation}\label{eq:cyclic-lift-def}
  A^{\langle q\rangle}:=A\otimes P_q.
\end{equation}
The basic spectral identity
\[
  \spec(A^{\langle q\rangle})
  =\{\mu\,\omega_q^j:\mu\in\spec(A),\ 0\le j\le q-1\}
\]
will be used repeatedly. In particular,
\(\rho(A^{\langle q\rangle})=\rho(A)\). For \(q>1\) this cyclic lift is
typically periodic rather than primitive, but \(\lambda=\rho(A)\) is
still a simple eigenvalue of \(A^{\langle q\rangle}\). We therefore use
the branch-lifted logarithmic record
\[
  \Psi_q(A):=
  -\sum_{j=2}^d
    \Log\!\left(1-\left(\frac{\mu_j}{\lambda}\right)^q\right)
\]
to encode the non-Perron spectrum. Since \(|\mu_j|<\lambda\), every
factor \(1-(\mu_j/\lambda)^q\) lies in the open right half-plane, so the
principal branch \(\Log\) is well defined here. By
\Cref{lem:simple-eigenvalue-reduced-constant},
\[
  \exp(\Psi_q(A))=q\,C_\lambda(A^{\langle q\rangle}),
\]
so the reduced constants determine only the exponentiated values unless
a branch choice is included as part of the data.
When interpreted as a \(\ZZ/q\ZZ\)-extension, this is the special
all-edges-label-\(1\) cocycle; it is not the same object as an arbitrary
cyclic quotient of a later finite-group cocycle.

\begin{definition}[Spectral fingerprints]
  \label{def:spectral-fingerprints}
  The \emph{cyclic spectral fingerprint} of \(A\) is the sequence
  \[
    \{\Psi_q(A)\}_{q\ge 1}.
  \]
  For each \(q\ge 1\), the \emph{\(q\)-torsion spectral fingerprint} is
  the root-of-unity layer
  \[
    \{F_A(\omega_q^j)\}_{0\le j\le q-1}.
  \]
  The associated \emph{exponential fingerprint} is the growth rate
  \[
    \limsup_{q\to\infty}\abs{\Psi_q(A)}^{1/q}.
  \]
\end{definition}

\subsection{Standing notation}

Throughout the paper \(\omega_q=e^{2\pi i/q}\) denotes a primitive
\(q\)th root of unity, and \(\Phi_q\) the \(q\)th cyclotomic
polynomial. We reserve the notation \(\varrho\) for irreducible
representations of finite groups and keep \(\rho(\cdot)\) for spectral
radii. Sections~\ref{sec:chebotarev} and~\ref{sec:shadows} use the
graph and cocycle picture for finite-group extensions. The later scalar
section, Section~\ref{sec:finite-part}, uses only the primitive matrix
formalism and cyclic lifts.
