\section{Introduction}\label{sec:introduction}

The problem addressed in this paper is primitive orbit reconstruction
for finite-group extensions of mixing subshifts of finite type. Let
\((X_A^+,\sigma)\) be a mixing subshift with primitive adjacency matrix
\(A\), let \(\lambda=\rho(A)\), and let \(\tau:E_A\to G\) be a cocycle
into a finite group. For every irreducible representation \(\varrho\)
of \(G\), the twisted transition matrix \(A_\varrho\) records periodic
labels through its traces. These traces are periodic data. The
primitive labels are subtler: when a primitive orbit with label \(g\)
is repeated \(m\) times, its label becomes \(g^m\). Thus primitive
extraction is not ordinary M\"obius inversion in the non-abelian
setting; it must be coupled to the Adams operations
\(\psi^m\phi(g)=\phi(g^m)\) on class functions.

The mathematical centre of the paper is the Adams--M\"obius passage
from twisted determinant data to primitive Peter--Weyl channels. The
word ``main'' is used below only for the finite-group spine built from
this passage. There are not three parallel theorem strands. The spine
has two parts: first, the determinant-to-primitive Adams--M\"obius
mechanism; second, its Perron-point class-Mertens consequence after an
explicit twisted gap is assumed. The final character-orthogonality
formula is included because it identifies what the class constants
encode, but it is a readout of those constants, not a separate source of
primitive information. Scalar determinant computations and cyclic
root-of-unity sampling sit outside this spine. They remain in the paper
because the finite-group theorem needs a trivial-channel normalisation
and because cyclic data give a useful one-dimensional comparison, but
neither part produces non-abelian primitive information. The hierarchy is
therefore intentionally asymmetric.
\begin{enumerate}[label=\textup{(\arabic*)},leftmargin=2.2em]
\item \emph{Main Theorem I.} Twisted determinants are
  linearised in class-function coordinates, and Adams--M\"obius
  inversion converts the resulting periodic Peter--Weyl channels into
  primitive Peter--Weyl series.
\item \emph{Main Theorem II.} After imposing the explicit
  twisted gap \(\eta<\lambda\), the primitive series are evaluated at
  \(z=\lambda^{-1}\), giving conjugacy-class Mertens constants. The
  non-trivial primitive channels can then be read off from these
  constants by ordinary character orthogonality.
\item \emph{Structural finite-group consequences.} Quotient
  functoriality identifies the abelianised primitive profile, and the
  linear span of cyclic quotient observables is the one-dimensional
  Peter--Weyl shadow of the primitive class profile. The orthogonal
  complement is the genuine non-abelian defect. The root-of-unity layers
  used later are obtained only after replacing the cocycle by the
  canonical cyclic lift, where every edge is labelled by the generator of
  \(\ZZ/q\ZZ\); they are not profiles attached to arbitrary cyclic
  quotients.
\item \emph{Auxiliary finite-matrix inputs, not main theorems.} The
  scalar finite part of \(\zeta_A\) supplies only the trivial-channel
  normalisation \(\mathcal M(A)\), and one-layer root-of-unity sampling
  is only the elementary finite Fourier calculation in the separate
  canonical cyclic-lift model.
\end{enumerate}

This hierarchy is also the logical direction of the proofs:
twisted determinants first give periodic Peter--Weyl channels; the
Adams--M\"obius step extracts primitive Peter--Weyl channels; the
explicit gap permits evaluation at \(z=\lambda^{-1}\) and hence class
Mertens constants; and non-abelian character orthogonality reads the
Perron-point primitive channels from those constants. Thus the
dependency graph is
\[
  \begin{array}{c}
    \text{twisted determinants}
    \xrightarrow{\text{trace}}
    \text{periodic Peter--Weyl channels} \\[2pt]
    \xrightarrow{\text{Adams--M\"obius}}
    \text{primitive Peter--Weyl channels} \\[2pt]
    \xrightarrow[\eta<\lambda]{z=\lambda^{-1}}
    \text{class Mertens constants}
    \xrightarrow{\text{orthogonality}}
    \text{primitive class profile}.
  \end{array}
\]

The novelty boundary is deliberately placed at the Adams--M\"obius
arrow. The determinant logarithms are periodic fixed-point data;
ordinary M\"obius inversion would remove repetitions only in the scalar
case. In the finite-group setting the repeated label is \(g^m\), so the
primitive extraction must use the Adams operators on the character ring.
Once the primitive Peter--Weyl values at \(z=\lambda^{-1}\) have been
constructed, the final Fourier formula is the standard orthogonality of
irreducible characters. Its role is diagnostic rather than generative:
it identifies precisely which primitive channels are contained in the
class constants already produced by the preceding Adams--M\"obius and
spectral-gap arguments.

This is also why the scalar and cyclic finite-matrix statements are
placed after the finite-group sections. They remain true without a
finite group cocycle and can be read independently of
\(\tau:E_A\to G\). By contrast, the two main theorem packages depend on
the twisted matrices \(A_\varrho\), the Adams operations on the
character ring, and the primitive conjugacy-class profile. The paper is
therefore organised as a primitive finite-group inverse problem with
scalar normalisation and abelian sampling supplied only as auxiliary
inputs.

Concretely, Main Theorem~I is the package
\Cref{lem:twisted-trace-identity,thm:class-function-linearisation,%
cor:class-function-adams-mobius,cor:primitive-peter-weyl-determinant}.
Main Theorem~II is the package
\Cref{thm:class-mertens-explicit,prop:class-mertens-deviations,%
thm:class-mertens-fourier}. The quotient and cyclic-detection results
of \Cref{sec:shadows} are structural consequences of these packages.
The scalar finite-part and root-of-unity statements are auxiliary
finite-matrix inputs proved later, after the finite-group mechanism.
The latter root-of-unity statements refer to the canonical cyclic-lift
model of
\Cref{rem:abelian-toy-model-cyclic,prop:canonical-cyclic-lift-root-layer},
where every edge is labelled by the generator of \(\ZZ/q\ZZ\); general
cyclic quotients of an
abelianised cocycle give their own quotient profiles, not those
root-of-unity samples.
This terminology is used as a status convention throughout the paper:
a result is called main only when it contains either the
Adams--M\"obius primitive extraction from twisted determinant channels,
or the Perron-point evaluation of the resulting non-trivial
Peter--Weyl channels under the explicit twisted gap. Results proved by
scalar Euler products, Newton identities, or finite Fourier inversion
are not placed on the same level, even when they are useful for
normalisation or comparison.
\begin{center}
\small
\begin{tabular}{L{0.34\linewidth}L{0.36\linewidth}L{0.20\linewidth}}
\toprule
Mechanism & Proof location & Status \\
\midrule
Determinants \(\to\) primitive Peter--Weyl channels &
\Cref{thm:class-function-linearisation,%
cor:class-function-adams-mobius,%
cor:primitive-peter-weyl-determinant} &
Main Theorem~I \\
Primitive channels \(\to\) class constants at \(\lambda^{-1}\) &
\Cref{thm:class-mertens-explicit,%
prop:class-mertens-deviations} &
Main Theorem~II \\
Class constants \(\to\) primitive Perron channels &
\Cref{thm:class-mertens-fourier} &
Main Theorem~II readout \\
Quotients and cyclic shadows &
\Cref{thm:quotient-functoriality,cor:abelian-cyclic-shadow,%
prop:canonical-cyclic-lift-root-layer,%
thm:abelian-shadow-defect,%
thm:cyclic-detection-boundary} &
Structural consequence \\
Scalar finite part of \(\zeta_A\) &
\Cref{thm:finite-part-primitive-closed-form} &
Auxiliary normalisation \\
One root-of-unity layer &
\Cref{thm:finite-part-cyclic-lift-single-q-torsion-reconstruction} &
Auxiliary finite DFT \\
\bottomrule
\end{tabular}
\end{center}

The table is a proof-dependency table, not a list of parallel headline
results. Removing the auxiliary rows would leave Main Theorem~I
unchanged and would affect Main Theorem~II only by leaving its scalar
trivial-channel constant unnamed. Removing Main Theorem~I, by contrast,
would remove the only mechanism in the paper that turns periodic
twisted determinant data into primitive non-abelian data.

The scalar dynamical \(\zeta\)-function
\[
  \zeta_A(z)
  =\exp\!\left(\sum_{n\ge 1}\frac{\Tr(A^n)}{n}z^n\right)
  =\det(I-zA)^{-1}
\]
enters this graph only through the trivial representation channel. Its
rationality and orbit-counting role are classical, going back to
Bowen--Lanford \cite{BowenLanford1970Zeta} and Manning
\cite{Manning1971Axiom}, with further development through Ruelle's
thermodynamic formalism
\cite{Ruelle1976ZetaExpanding,Ruelle1978Thermodynamic} and the
Parry--Pollicott transfer-operator programme
\cite{ParryPollicott1990Zeta}. The finite-part proposition recorded
below, and proved only after the finite-group sections, is therefore a
classical Euler-product/M\"obius computation at the trivial channel. The
one-layer DFT corollary is coefficient recovery for a polynomial sampled
at enough roots of unity. Neither statement generates non-abelian
primitive information; both are kept outside the main theorem spine and
are used only to normalise or test the Adams--M\"obius theorem package.

\subsection{Finite-group main theorem spine}

Let \(\tau:E_A\to G\) be a cocycle into a finite group, and let
\(A_\varrho\) be the twisted transition matrices. For a class function
\(\phi:G\to\CC\) write
\[
  T_{\phi,n}:=\sum_{x\in\operatorname{Fix}(\sigma^n)}
    \phi(\tau_n(x)),
  \qquad
  \mathcal L_\phi(z):=\sum_{n\ge 1}\frac{T_{\phi,n}}{n}z^n,
\]
and
\[
  \Pi_\phi(z):=\sum_{\gamma\in\cP}\phi(\tau(\gamma))z^{|\gamma|}.
\]
For a conjugacy class \(C\subset G\), let \(p_{n,C}\) denote the number
of primitive orbits of length \(n\) whose cocycle label lies in \(C\).
The scalar orbit-Mertens constant \(\mathcal M(A)\) is the
trivial-channel constant defined in \Cref{def:abel-finite-part}; the
finite-matrix formula behind this normalisation is the auxiliary
calculation of \Cref{thm:finite-part-primitive-closed-form}.

\begin{theorem}[Main Theorem I: Adams--M\"obius primitive
  determinant linearisation]
  \label{thm:intro-finite-group}
  \label{thm:intro-adams-mobius}
  For every class function \(\phi:G\to\CC\) and every
  \(|z|<\lambda^{-1}\),
  \[
    \mathcal L_\phi(z)
    =
    \sum_{\varrho\in\Irr(G)}
      \widehat\phi(\varrho)
      \log\det(I-zA_\varrho)^{-1},
  \]
  where \(\widehat\phi(\varrho)\) is the character coefficient of
  \(\phi\). Moreover,
  \[
    \Pi_\phi(z)
    =
    \sum_{m\ge 1}\frac{\mu(m)}{m}\,
      \mathcal L_{\psi^m\phi}(z^m).
  \]
  Equivalently, if
  \(\psi^m\chi_\varrho
    =\sum_{\sigma\in\Irr(G)}a_{\varrho,\sigma}^{(m)}\chi_\sigma\),
  then
  \[
    \Pi_\varrho(z)
    =
    \sum_{m\ge 1}\frac{\mu(m)}{m}
      \sum_{\sigma\in\Irr(G)}
        a_{\varrho,\sigma}^{(m)}
        \log\det(I-z^mA_\sigma)^{-1}.
  \]
\end{theorem}

\begin{theorem}[Main Theorem II: class Mertens constants under
  the twisted gap]
  \label{thm:intro-class-mertens}
  Assume the explicit twisted spectral gap
  \[
    \eta:=\max_{\varrho\neq\mathbf 1}\rho(A_\varrho)<\lambda,
    \qquad
    \Lambda_*:=\max\{\Lambda(A),\eta,\lambda^{1/2}\}.
  \]
  Here \(\Lambda(A)\) denotes the maximum modulus of the non-Perron
  eigenvalues of the base matrix \(A\). The inequality
  \(\eta<\lambda\) is an assumption of this theorem, not a consequence
  proved in the paper.
  Then for every conjugacy class \(C\subset G\),
  \[
    p_{n,C}
    =
    \frac{|C|}{|G|}\frac{\lambda^n}{n}
    +O\!\left(\frac{\Lambda_*^n}{n}\right),
  \]
  and
  \[
    \sum_{n\le N}\frac{p_{n,C}}{\lambda^n}
    =
    \frac{|C|}{|G|}\log N+\mathcal M_C(A)+O(N^{-1}).
  \]
  The deviations
  \[
    \Delta_C(A):=\mathcal M_C(A)-\frac{|C|}{|G|}\mathcal M(A)
  \]
  have the following ordinary character-orthogonality readout:
  \[
    \Pi_\sigma(\lambda^{-1})
    =
    \sum_{C\subset G}\Delta_C(A)\chi_\sigma(C)
    \qquad(\sigma\neq\mathbf 1).
  \]
  The last display is not an additional analytic input. It is ordinary
  class-function orthogonality applied after Main Theorem~I and the gap
  have made the non-trivial primitive values absolutely convergent at
  \(z=\lambda^{-1}\) and identified the class constants.
\end{theorem}

\begin{corollary}[Structural finite-group consequence: quotient
  functoriality and the abelian shadow]
  \label{thm:intro-quotient-shadow}
  Assume the hypotheses of Main Theorem~II, so the class deviations
  \(\Delta_C(A)\) are defined. Quotient maps \(q:G\twoheadrightarrow Q\)
  intertwine class-function logarithms and primitive series:
  \[
    \mathcal L_{q^\ast f}^{G}(z)=\mathcal L_f^{Q}(z),
    \qquad
    \Pi_{q^\ast f}^{G}(z)=\Pi_f^{Q}(z)
    \qquad(|z|<\lambda^{-1}).
  \]
  If
  \[
    \mathcal B_A(C):=\frac{|G|}{|C|}\Delta_C(A),
  \]
  then \(\mathcal B_A\) splits orthogonally into its one-dimensional
  Peter--Weyl part and its higher-dimensional Peter--Weyl part. The span
  of all cyclic quotient observables is exactly the one-dimensional
  sector; hence their aggregate projection is the abelian shadow and is
  orthogonal to the genuine non-abelian defect.
  This is a statement about quotient primitive profiles. The
  root-of-unity layer of the auxiliary finite-part section is obtained
  only after replacing the given cocycle by the canonical cyclic-lift
  cocycle of \Cref{rem:abelian-toy-model-cyclic}, where every edge has
  label \(1\in\ZZ/q\ZZ\)
  (\Cref{prop:canonical-cyclic-lift-root-layer}). In particular, the full
  primitive class profile is generated by cyclic quotient shadows if and
  only if every higher-dimensional primitive channel
  \(\Pi_\varrho(\lambda^{-1})\) vanishes.
\leanverified{paper_etds_intro_finite_group}
\end{corollary}

An explicit \(S_3\)-extension shows that vanishing periodic traces do
not force vanishing primitive bias: the sign-character periodic channel
vanishes identically, yet the corresponding primitive channel is
non-zero.

At the level of the introduction, the two displayed theorem statements
above are the only main theorem statements. The quotient/shadow result
is a finite-group consequence of them. The scalar finite-part formula,
the cyclic-lift inverse-spectral proposition, and the one-layer DFT
formula below are deliberately stated only as propositions or
corollaries, and they are not part of the proof spine for primitive
finite-group reconstruction. The finite part supplies the
trivial-channel normalisation \(\mathcal M(A)\), the cyclic-lift
fingerprint gives an auxiliary inverse-spectral comparison, and the
one-layer DFT formula is exactly coefficient recovery for a polynomial
sampled at enough roots of unity. Thus Main Theorem~I is proved without
any finite-part or root-of-unity input, and Main Theorem~II uses only the
scalar constant \(\mathcal M(A)\) together with the explicit twisted gap.

The structural quotient/shadow consequence is proved in
\Cref{thm:quotient-functoriality,%
cor:abelian-cyclic-shadow,%
prop:canonical-cyclic-lift-root-layer,%
thm:abelian-shadow-defect,%
thm:cyclic-detection-boundary,%
cor:s3-example-profile}.
The following contribution map records dependencies, hypotheses, and
logical status. The first two rows are the main finite-group theorem
package; the quotient/shadow row is its structural finite-group
consequence; the scalar and cyclic rows are auxiliary finite-matrix
inputs.
\begin{center}
\small
\begin{tabular}{L{0.20\linewidth}L{0.30\linewidth}L{0.21\linewidth}L{0.17\linewidth}}
\toprule
Statement & Hypotheses & Role in the proof & Contribution level \\
\midrule
Main Theorem~I &
Primitive base matrix \(A\), finite group \(G\), cocycle
\(\tau:E_A\to G\), and \(|z|<\lambda^{-1}\) &
Adams--M\"obius primitive determinant linearisation &
Main result; new finite-group mechanism; no twisted gap \\
Main Theorem~II &
Main Theorem~I plus the explicit twisted gap
\(\eta=\max_{\varrho\neq\mathbf 1}\rho(A_\varrho)<\lambda\) &
Class Mertens constants, followed by a character-orthogonality readout &
Main result; novelty is Perron-point Adams--M\"obius evaluation under the
gap; the displayed Fourier formula is formal character orthogonality \\
Structural quotient/shadow consequence &
Main Theorem~II plus finite quotient maps \(G\twoheadrightarrow Q\) &
Abelian shadow and non-abelian defect &
Finite-group consequence \\
Scalar finite-part proposition &
Primitive non-negative matrix \(A\) with \(\lambda=\rho(A)>1\) &
Trivial-channel normalisation \(\mathcal M(A)\) &
Auxiliary classical determinant/M\"obius input \\
Cyclic-lift spectrum proposition &
Primitive \(d\times d\) matrix and cyclic-lift constants
\(\{\Psi_q(A)\}_{q\ge 1}\) &
Reduced-spectrum reconstruction from cyclic lifts &
Auxiliary inverse-spectral comparison; not primitive reconstruction \\
One-layer DFT corollary &
One root-of-unity layer with \(q\ge d\) &
Canonical cyclic-lift finite Fourier calculation &
Auxiliary finite Fourier calculation; no primitive finite-group content \\
\bottomrule
\end{tabular}
\end{center}
Thus the determinant identities and primitive inversion of Main
Theorem~I are unconditional inside the open Perron disc, while the
Perron-point class constants and their character-orthogonality readout
in Main Theorem~II use the displayed spectral-gap assumption. The size condition
\(q\ge d\) belongs only to the one-layer discrete Fourier corollary and
plays no role in the finite-group reconstruction.
The section order follows this hierarchy: the finite-group mechanism is
proved first, the quotient/shadow consequences next, and only then are
the scalar finite-part and root-of-unity computations proved.

For reference, the conditional status of the hypotheses used in these
statements is as follows.
\begin{center}
\small
\begin{tabular}{L{0.20\linewidth}L{0.38\linewidth}L{0.30\linewidth}}
\toprule
Statement & Hypothesis & Status in this paper \\
\midrule
Main Theorem~I &
Primitive base matrix, finite group, and fixed cocycle
\(\tau:E_A\to G\) &
Standing data; all identities hold for \(|z|<\lambda^{-1}\) \\
Main Theorem~I &
No twisted spectral gap and no cyclic sampling condition &
The determinant identities and Adams--M\"obius primitive inversion are
formal finite-dimensional identities in the open Perron disc \\
Main Theorem~II &
\(\eta=\max_{\varrho\neq\mathbf 1}\rho(A_\varrho)<\lambda\) &
Explicit assumption; not proved from primitivity of \(A\) \\
Canonical cyclic-lift comparison with \(q>1\) &
\(G=\ZZ/q\ZZ\), \(\tau_q(e)=1\), and
\(A_{\chi_j}^{\tau_q}=\omega_q^jA\) &
Does not satisfy the twisted gap, since
\(\rho(A_{\chi_j}^{\tau_q})=\lambda\) for every \(j\neq0\);
see \Cref{prop:canonical-cyclic-lift-root-layer}; used only as the
auxiliary finite-matrix root-of-unity comparison \\
Main Theorem~II &
\(\Lambda_*=\max\{\Lambda(A),\eta,\lambda^{1/2}\}<\lambda\) &
Consequence of primitivity of \(A\) and the displayed twisted gap \\
Main Theorem~II &
Character-orthogonality readout from all conjugacy-class constants &
Formal readout by irreducible-character orthogonality after the
Perron-point constants have been constructed \\
Auxiliary one-layer DFT corollary &
\(q\ge d\) &
Only a polynomial-sampling condition; not used in the finite-group
theorems \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Auxiliary finite-matrix inputs after the theorem spine}

We next record the scalar finite-matrix components only as a catalogue
of inputs and comparison cases for the finite-group package. This is not
a third theorem block. Its two purposes are narrower: to name the scalar
normalisation \(\mathcal M(A)\) appearing in Main Theorem~II, and to
record the canonical cyclic-lift polynomial-sampling calculation kept
separate from the cyclic quotient shadows in \Cref{sec:shadows}. The proofs are
deferred to \Cref{sec:finite-part}, after the Adams--M\"obius
construction and its quotient/shadow consequences have already been
proved. In particular, the formal statements in this subsection are
orientation statements for notation and comparison; their proofs, and
the places where they are used, are the later auxiliary statements in
\Cref{sec:finite-part}. After removing the Perron factor one obtains the
reduced
determinant
\[
  C(A):=\det\nolimits'(I-A/\lambda)^{-1}
\]
and the reduced spectral polynomial
\[
  F_A(t):=\frac{\det(I-tA/\lambda)}{1-t}
  =\prod_{j=2}^d\left(1-\frac{\mu_j}{\lambda}t\right),
\]
where \(\mu_2,\dots,\mu_d\) are the non-Perron eigenvalues. The
Abel-regularised finite-part constant
\[
  \FP(A):=\lim_{r\uparrow 1}\bigl(P_A(r)+\log(1-r)\bigr),
  \qquad
  P_A(r):=\sum_{n\ge 1}p_n(A)\lambda^{-n}r^n,
\]
measures the deviation of the primitive orbit sum from the harmonic
series.

\begin{proposition}[Auxiliary scalar finite-part computation at the
  Perron pole]
  \label{thm:intro-finite-part}
  Let \(A\) be a primitive non-negative matrix with Perron root
  \(\lambda=\rho(A)>1\), reduced determinant \(C(A)\), and primitive
  orbit counts \(p_n(A)\). Then
  \[
    \FP(A)
    =\log C(A)
      +\sum_{k\ge 2}\frac{\mu(k)}{k}
        \log\det(I-\lambda^{-k}A)^{-1}.
  \]
  Equivalently,
  \[
    \FP(A)
    =\log C(A)
      +\sum_{n\ge 1}p_n(A)
        \bigl(\lambda^{-n}+\log(1-\lambda^{-n})\bigr).
  \]
  In particular, the orbit-Mertens constant
  \(\mathcal M(A)=\gamma+\FP(A)\) satisfies
  \[
    \sum_{n\le N}\frac{p_n(A)}{\lambda^n}
    =\log N+\mathcal M(A)+o(1).
  \]
\leanverified{paper_etds_intro_finite_part}
\end{proposition}

This identifies only the scalar constant that appears in the classwise
deviations above. The argument is a short Euler-product and
M\"obius-inversion computation in the classical SFT determinant
calculus, and it does not interact with the non-trivial Peter--Weyl
channels;
\Cref{thm:intro-finite-part} is proved in
\Cref{thm:finite-part-primitive-closed-form,cor:finite-part-mertens-asymptotic}.

The second scalar component addresses spectral reconstruction from
root-of-unity data, a theme connected to inverse spectral problems for
SFTs \cite{LindMarcus1995}. The useful inverse-spectral part is the
recovery of the reduced spectrum from the full canonical cyclic-lift
fingerprint. The single-layer statement is separated as the elementary
discrete Fourier corollary for that reset model; it is not a quotient of
an arbitrary abelianised cocycle and is not a substitute for the
class-constant readout in Main Theorem~II.

\begin{proposition}[Auxiliary cyclic-lift reconstruction
  of the reduced spectrum]
  \label{thm:intro-cyclic-reconstruction}
  Let \(A\) be a primitive non-negative \(d\times d\) matrix, and set
  \[
    \Psi_q(A):=\log q+\log C_\lambda(A^{\langle q\rangle}),
  \]
  where \(A^{\langle q\rangle}=A\otimes P_q\) is the cyclic lift and
  \(C_\lambda\) denotes the reduced constant at the simple eigenvalue
  \(\lambda=\rho(A)\). Then the sequence
  \(\{\Psi_q(A)\}_{q\ge 1}\) determines \(F_A\), and hence the
  non-Perron spectrum of \(A\) with algebraic multiplicity. Moreover,
  \[
    \limsup_{q\to\infty}\abs{\Psi_q(A)}^{1/q}
    =
    \Lambda(A)/\lambda.
  \]
\end{proposition}

\begin{corollary}[Elementary one-layer DFT corollary]
  \label{cor:intro-one-layer-reconstruction}
  Let \(A\) be a primitive non-negative \(d\times d\) matrix. If
  \(q\ge d\), then the single root-of-unity layer
  \(\{F_A(\omega_q^j)\}_{0\le j\le q-1}\) determines every coefficient
  of the reduced spectral polynomial
  \[
    F_A(t)=\sum_{n=0}^{d-1}a_nt^n,
  \]
  by the finite Fourier formula
  \[
    a_n=\frac{1}{q}\sum_{j=0}^{q-1}
      F_A(\omega_q^j)\omega_q^{-jn}
    \qquad (0\le n\le d-1).
  \]
\leanverified{paper_etds_intro_cyclic_reconstruction_seeds}
\end{corollary}

The proof combines M\"obius inversion on multiples with Newton's
identities for the cyclic-lift spectrum proposition, and finite Fourier
inversion for \Cref{cor:intro-one-layer-reconstruction}; see
\Cref{thm:finite-part-cyclic-lift-spectrum-identifiability,%
thm:finite-part-cyclic-lift-single-q-torsion-reconstruction,%
cor:finite-part-cyclic-lift-zeta-torsion-determines-spectrum}.
The terminology of cyclic, torsion, and exponential spectral
fingerprints is fixed in \Cref{def:spectral-fingerprints}. None of
these auxiliary finite-matrix statements is used to prove the
Adams--M\"obius primitive inversion, the class Mertens theorem, or the
non-abelian reconstruction formula.

\subsection{Relation to existing work}

The dynamical \(\zeta\)-function for SFTs goes back to
Artin--Mazur \cite{ArtinMazur1965PeriodicPoints} and Bowen--Lanford
\cite{BowenLanford1970Zeta,Manning1971Axiom}. The prime-orbit
asymptotic for shifts of finite type is classical: Parry
\cite{Parry1983PrimeOrbit} proved the analogue of the prime number
theorem for closed orbits of SFTs and their suspensions. The existence
of Mertens-type asymptotics for Axiom~A flows was established by Sharp
\cite{Sharp1991Mertens}. For zeta functions attached to broader
group-action orbit data, Lind's \(\ZZ^d\)-action zeta function is the
standard reference \cite{Lind1996ZetaZdActions}. The present paper does
not claim a new prime-orbit theorem. Its primary contribution is the
Adams--M\"obius reconstruction of primitive Peter--Weyl channels from
twisted determinant data. The scalar finite-part identity at the Perron
pole is included as an explicit finite-matrix input for the classwise
constants and as a checkable comparison with the classical Mertens
constant.

The transfer-operator approach is developed systematically in
\cite{Ruelle1976ZetaExpanding,Ruelle1978Thermodynamic,%
ParryPollicott1990Zeta,Baladi2000PositiveTransfer}.
Twisted determinant and \(L\)-function constructions for dynamical
systems and finite-group extensions go back at least to the twisted
Perron--Frobenius framework of Adachi--Sunada
\cite{AdachiSunada1987TwistedPF}. The leading Chebotarev asymptotics for
finite extensions are classical: Parry--Pollicott treat Axiom~A Galois
coverings \cite{ParryPollicott1986Chebotarev}, and Noorani--Parry prove
the corresponding theorem for finite homogeneous extensions of shifts
\cite{NooraniParry1992ChebotarevShifts}; see also
\cite{PollicottSharp2007Chebotarev,Parry1999LivsicNonAbelian}. The
present paper does not reprove those leading density theorems by a new
Tauberian method. Instead it isolates the finite determinant algebra
needed after the leading term: primitive class constants and their
Peter--Weyl coordinates are recovered from twisted determinant data by
Adams--M\"obius inversion.

Finite group actions on SFTs have a separate symbolic-dynamics
literature, and this is the closest context for the finite-group part of
the paper. Adler--Kitchens--Marcus
\cite{AdlerKitchensMarcus1985FiniteGroupActions} initiated a systematic
study of finite group actions on shifts of finite type. Fiebig's work on
periodic points and finite group actions on SFTs
\cite{Fiebig1993PeriodicFiniteGroupActions} is a direct ETDS antecedent:
it attaches computable periodic data to finite actions, studies fixed
subsystems and quotient systems, and compares the corresponding zeta
functions. Silver--Williams
\cite{SilverWilliams2005InvariantFiniteGroupActions} develop the
left- and right-reduced shift invariants for finite group actions and
place Fiebig's quotient-zeta construction in a broader invariant
framework. This is precisely the periodic-data side of the present
setup. The present paper starts from the same kind of finite periodic
data, but asks a different primitive inverse question: after decomposing
the twisted determinant data into Peter--Weyl channels, how does one
remove the repetitions that send a primitive label \(g\) to \(g^m\)?
The answer is the Adams--M\"obius step of Main Theorem~I.
Boyle--Schmieding
\cite{BoyleSchmieding2017FiniteGroupExtensions} analyse finite group
extensions of SFTs through \(K\)-theory, Parry obstructions, and
Liv\v{s}ic obstructions. Boyle--Carlsen--Eilers
\cite{BoyleCarlsenEilers2020FlowEquivalenceGSFT} develop the
classification and equivariant flow-equivalence theory of \(G\)-SFTs
using algebraic matrix structures over group rings. Those works address
existence, classification, equivariant flow equivalence, and
periodic-data obstructions for group actions and extensions. Here the
cocycle \(\tau:E_A\to G\) and its twisted matrices are fixed; no
classification or flow-equivalence invariant is being proposed. The
question is inverse-determinantal and primitive: which primitive
conjugacy-class Mertens constants and non-abelian Fourier coefficients
are forced by the twisted matrices \(A_\varrho\)? The Adams--M\"obius
step is precisely the passage not supplied by ordinary periodic-data
zeta functions: it undoes the repetitions that replace a primitive
label \(g\) by its powers \(g^m\).

\begin{center}
\small
\begin{tabular}{L{0.27\linewidth}L{0.32\linewidth}L{0.31\linewidth}}
\toprule
Closest antecedent & What it supplies & Present use and boundary \\
\midrule
Adler--Kitchens--Marcus
\cite{AdlerKitchensMarcus1985FiniteGroupActions} &
Finite group actions on SFTs as a systematic object of study &
Background for finite actions; no primitive Peter--Weyl inversion is
claimed there or needed from it \\
Fiebig
\cite{Fiebig1993PeriodicFiniteGroupActions} &
Periodic data, fixed subsystems, quotient systems, and zeta functions
for finite actions &
Direct periodic-data antecedent; Main Theorem~I starts from this kind of
periodic information and performs the missing primitive
Adams--M\"obius extraction \\
Silver--Williams
\cite{SilverWilliams2005InvariantFiniteGroupActions} &
Reduced-shift invariants and quotient-zeta context for finite group
actions &
Invariant-theoretic background for quotient data; the present work keeps
the cocycle fixed and extracts primitive Peter--Weyl channels \\
Boyle--Schmieding
\cite{BoyleSchmieding2017FiniteGroupExtensions} &
\(K\)-theoretic, Parry, and Liv\v{s}ic obstructions for finite group
extensions &
Structural context for when twisted channels can have a gap; the gap
itself remains an explicit hypothesis here \\
Boyle--Carlsen--Eilers
\cite{BoyleCarlsenEilers2020FlowEquivalenceGSFT} &
Equivariant flow equivalence and algebraic classification of \(G\)-SFTs &
Classification context only; the present invariant is attached to a
fixed cocycle and fixed twisted matrices \\
Parry--Pollicott and Noorani--Parry
\cite{ParryPollicott1986Chebotarev,NooraniParry1992ChebotarevShifts} &
Leading Chebotarev orbit-counting asymptotics for finite extensions &
The leading density is not new here; the contribution is the finite
determinant algebra for primitive class constants beyond that leading
term \\
\bottomrule
\end{tabular}
\end{center}

This distinction is important for interpreting the theorem labels. The
paper does not promote the classical scalar determinant computation, the
leading Chebotarev density, or finite Fourier inversion on roots of
unity as new contributions. Those ingredients are used as boundary
conditions and comparison tests. The contribution asserted here is the
finite-group primitive inverse problem: starting from the twisted
determinants attached to a fixed cocycle, recover primitive
Peter--Weyl channels and, under the explicit gap, their classwise
Mertens constants.

Tauberian refinements and closed-orbit asymptotics are treated in
\cite{Pollicott1992AgmonTauberian} and the monograph
\cite{ParryPollicott1990Zeta}. For the functorial orbit-counting
perspective underlying Section~\ref{sec:shadows} one should also compare
\cite{PakapongpunWard2009FunctorialOrbit}.
For graph coverings and graph \(L\)-functions one should compare the
Ihara--Bass and Hashimoto determinant formulas, the Stark--Terras theory
of graph coverings and Artin \(L\)-functions, and directed or weighted
variants
\cite{Hashimoto1989ZetaFiniteGraphs,%
StarkTerras1996FiniteGraphsCoverings,%
StarkTerras2000FiniteGraphsCoveringsII,%
StarkTerras2007FiniteGraphsCoveringsIII,%
Bass1992Ihara,%
MizunoSato2001ZetaDigraphs,%
Horton2007IharaDigraphs,%
ChoeKwakParkSato2007Bartholdi,%
Sato2008WeightedBartholdiCoverings}.
For more recent perspectives on dynamical zeta functions and dynamical
determinants beyond the classical symbolic-dynamics sources, compare
Baladi's modern monograph \cite{Baladi2018ZetaDeterminants}, the
parameter-regularity results of J\'ez\'equel
\cite{Jezequel2019ParameterRegularity}, and Shen's work on twisted
Ruelle zeta functions on locally symmetric spaces
\cite{Shen2023ComplexValuedAnalyticTorsion}. Our results stay in the
finite-state symbolic setting but fit naturally into this broader
determinant-based programme.

The relation between the present results and their closest antecedents
is therefore organised with the finite-group mechanism at the centre.
Classical \(\zeta\)-rationality supplies the finite-part input at the
Perron pole, and root-of-unity sampling for canonical cyclic lifts
supplies the separate cyclic-lift comparison case. The Chebotarev theory
of finite-group extensions is where the new ingredient appears: the Adams--M\"obius
passage from twisted determinant data to primitive Peter--Weyl series.
The appendices record secondary consequences of these inputs rather than
additional primary theorem chains.

\subsection{Plan of the paper}

\Cref{sec:preliminaries} fixes notation for primitive orbit counts,
reduced determinants, cyclic lifts, and finite-group extensions.
\Cref{sec:chebotarev} develops the Adams--M\"obius mechanism for
finite-group extensions, derives the primitive Peter--Weyl determinant
series, and establishes the class Mertens theorem under a twisted
spectral gap. \Cref{sec:shadows} proves quotient functoriality,
identifies the abelian shadow and the exact boundary of cyclic
detection, and closes with the \(S_3\)-model demonstrating the Adams
obstruction. \Cref{sec:finite-part} then proves the auxiliary scalar
finite-part formula, the cyclic-lift reconstruction of the reduced
spectrum, and the one-layer discrete Fourier corollary used as the
canonical cyclic-lift comparison case. The appendices collect secondary
consequences:
holomorphic variation of the finite part (\Cref{app:variation}),
arithmetic root-of-unity corollaries including a Dirichlet-character
inversion (\Cref{app:arithmetic}), and rigidity statements for growing
families (\Cref{app:rigidity}).
