\begin{corollary}[Holographic identity under the Lee--Yang unit-circle hypothesis]\label{cor:group-jg-leyang-holographic-identity}
In the setting of Theorem \ref{thm:group-jg-pullback-factorization}, assume that \(K\subseteq\CC\), that every zero \(z_j\) of \(P\) satisfies \(|z_j|=1\), and that \(r\in\RR_{>0}\). Then for every \(z\in\CC^\times\),
\[
Q_r\!\bigl(J_r(z)\bigr)=r^n\,P(z)\,P\!\bigl(r^{-2}z^{-1}\bigr).
\]
Although the algebraic identity itself does not use the unit-circle hypothesis, we retain that assumption here because it is needed in the geometric corollaries below.
\end{corollary}

\begin{proof}
By Theorem \ref{thm:group-jg-pullback-factorization},
\[
r^n z^n\,Q_r\!\bigl(J_r(z)\bigr)=P(z)\,P^{\vee}(r^2 z).
\]
Since \(P^{\vee}(u)=u^nP(1/u)\), substituting \(u=r^2 z\) gives
\[
P^{\vee}(r^2 z)=r^{2n}z^n P(r^{-2}z^{-1}).
\]
Cancelling the common factor \(z^n\) yields the claim.
\end{proof}

\begin{corollary}[Double-resultant elimination and concentric spectral shells]\label{cor:group-jg-leyang-double-resultant}
Retain the hypotheses of Corollary \ref{cor:group-jg-leyang-holographic-identity}, and define
\[
S_r(z):=\operatorname{Res}_w\!\bigl(Q_r(w),\,r z^2-w z+r^{-1}\bigr)\in K(r)[z].
\]
Then
\[
S_r(z)=(-1)^n r^{n}\,P(z)\,\Bigl[z^{n}P\!\bigl(r^{-2}z^{-1}\bigr)\Bigr].
\]
Consequently,
\[
\mathrm{Zero}(S_r)
=\{z_1,\dots,z_n\}\cup\{r^{-2}z_1^{-1},\dots,r^{-2}z_n^{-1}\},
\]
where the two clusters lie on the circles \(|z|=1\) and \(|z|=r^{-2}\), respectively.
\end{corollary}

\begin{proof}
Because \(r z^2-w z+r^{-1}\) is linear in \(w\), the linear specialization formula for the resultant gives
\[
\operatorname{Res}_w\!\bigl(Q_r(w),\,r z^2-w z+r^{-1}\bigr)
=(-z)^n\,Q_r\!\left(\frac{r z^2+r^{-1}}{z}\right)
=(-z)^n Q_r\!\bigl(J_r(z)\bigr).
\]
Substituting Corollary \ref{cor:group-jg-leyang-holographic-identity} yields
\[
S_r(z)=(-z)^n r^n P(z)P(r^{-2}z^{-1})
=(-1)^n r^{n}P(z)\,\Bigl[z^{n}P\!\bigl(r^{-2}z^{-1}\bigr)\Bigr],
\]
which is the required factorization. The description of the zero set follows immediately from the two factors.
\end{proof}

\begin{corollary}[Single-scale holographic reconstruction]\label{cor:group-jg-leyang-single-scale-reconstruction}
In the setting of Corollary \ref{cor:group-jg-leyang-double-resultant}, assume further that \(r>1\) and that \(P\) is separable and monic. Then the transport map
\[
\mathcal T_r:\ P\longmapsto Q_r
\]
is injective. Moreover, \(P\) can be reconstructed uniquely from \((r,Q_r)\): compute \(S_r\) from \(Q_r\), and then \(P\) is the unique monic degree-\(n\) factor of \(S_r\) whose zeros all lie on the unit circle.
\end{corollary}

\begin{proof}
By Corollary \ref{cor:group-jg-leyang-double-resultant},
\[
S_r(z)=(-1)^n r^n\,P(z)\,\Bigl[z^nP\!\bigl(r^{-2}z^{-1}\bigr)\Bigr].
\]
When \(r>1\), every zero of the second factor lies on \(|z|=r^{-2}<1\), whereas the zeros of \(P\) lie on \(|z|=1\). The two spectral shells are therefore disjoint.

Hence \(S_r\) has exactly one degree-\(n\) factor whose zeros lie on the unit circle, namely \(P\) itself up to normalization. Since \(P\) is monic, that normalization is unique. If \(\mathcal T_r(P_1)=\mathcal T_r(P_2)\), then the corresponding polynomials \(S_r\) coincide, so the unit-circle factor is the same in both cases; thus \(P_1=P_2\).
\end{proof}
