\documentclass[11pt,a4paper]{amsart}

\usepackage{amsmath,amssymb,amsthm,mathrsfs}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{enumitem}
\usepackage[protrusion=true,expansion=false]{microtype}
\usepackage{mathtools}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\spec}{spec}
\DeclareMathOperator{\Res}{Res}
\DeclareMathOperator{\Fix}{Fix}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\norm}[1]{\lVert #1 \rVert}
\newcommand{\MM}{\mathcal M}
\newcommand{\dd}{\mathop{}\!\mathrm{d}}
\newcommand{\sone}{\mathcal S_1}

\title[Fredholm Determinants and Spectral Rigidity]{Fredholm determinants, cyclic-block realisations,\\
and spectral rigidity for symbolic \(\zeta\)-functions}

\author{Haobo Ma}
\address{CHRONOAI}

\date{}

\begin{document}

\begin{abstract}
We place symbolic dynamical \(\zeta\)-functions in a trace-class
Fredholm framework. The Fredholm--Witt identity and the associated
spectral rigidity are recorded as connective background. The main new
result is a cyclic-block characterization: an Euler product
\[
  \prod_j (1-\beta_j z^{n_j})^{-m_j}
\]
with
\(
  \sum_j m_j n_j\abs{\beta_j}^{1/n_j}<\infty
\)
admits a trace-class normal realization if and only if, up to an
arbitrary zero summand and unitary equivalence, its non-zero part is a
direct sum of cyclic blocks \(C(n_j,\alpha_j)\otimes I_{m_j}\) with
\(\alpha_j^{n_j}=\beta_j\). In particular, the total non-zero block
size is forced and each factor requires exactly \(m_jn_j\) non-zero
dimensions.

We then record determinant-level applications in symbolic dynamics.
For analytic families of primitive matrices we obtain explicit
group-inverse and Fredholm formulas for the variation of the finite-part
constants and of the Abel renormalization, together with a rapidly
convergent truncation scheme. For locally constant observables, twisted
Fredholm determinants rigidly determine periodic-orbit moment data and
asymptotic cumulants; under the standard finite-state aperiodicity
hypothesis, they also yield uniform off-origin decay of characteristic
functions. These applications are finite-dimensional consequences of the
usual spectral-gap formalism, included here to show how the determinant
package organizes them in one place. Explicit computations are given
for the golden-mean family and for the full two-shift.
\end{abstract}

\subjclass[2020]{37B10, 47B10, 47A55, 60F05}
\keywords{Dynamical \(\zeta\)-function, Fredholm determinant, trace-class operator,
cyclic-block realisation, spectral rigidity}

\maketitle

\input{sec_introduction}
\input{sec_preliminaries}
\input{sec_fredholm}
\input{sec_perturbation}
\input{sec_conclusion}

\bibliographystyle{amsalpha}
\bibliography{references}

\end{document}
