\section{Further remarks}\label{sec:conclusion}

The operator-theoretic statements in this note have different logical
status. The Fredholm--Witt bridge
(\Cref{thm:fredholm-witt-bridge}) is the classical connective identity
that places traces, Witt coordinates, and determinants in one common
language. The sole new contribution is the cyclic-block characterization
theorem (\Cref{thm:cyclic-fredholm-realisation}): an Euler product
admits a normal trace-class block-cyclic realisation if and only if the
summability threshold
\(
  \sum_j m_j n_j\abs{\beta_j}^{1/n_j}<\infty
\)
holds, and in that case the canonical cyclic-block model is already
trace-norm and support optimal within the normal class. Spectral
rigidity (\Cref{thm:fredholm-spectral-rigidity}) then shows that this
extra operator-theoretic freedom is geometric rather than spectral:
once the determinant is fixed, both the non-zero spectrum and the
minimal non-zero size are fixed.

As discussed in \Cref{sec:positioning}, the individual ingredients of
\Cref{thm:cyclic-fredholm-realisation}---Simon's canonical zero
description and the spectral theorem for normal operators---are
classical. The theorem's contribution is the exact cyclic-block
packaging and the support-optimality statement, which are not
consequences of either ingredient alone.

The applications section (\Cref{sec:perturbation}) is included for
context and illustration: in finite dimension, the formulas for
\(C(A)\), \(\MM(A)\), cumulants, and characteristic functions remain
consequences of the usual perturbative and spectral-gap machinery. What
the determinant package adds is a common organizing object: it records
periodic data, reduced determinants, and twisted moment information in
one invariant germ. The explicit golden-mean and full two-shift
computations show that these determinant-level formulas are concrete and
computable.

Several natural extensions merit further investigation. The non-normal
trace-class case is not addressed here: whether a given Euler product
admits a non-normal trace-class realisation with fewer non-zero
dimensions, or under a weaker summability condition, remains open.
Extending the spectral-rigidity statement beyond the Hilbert-space
trace-class case would require a precise determinant theory and a
corresponding canonical zero description. Quantitative refinements of
the symbolic applications, such as Berry--Esseen bounds or moderate
deviations, would likewise require additional spectral input not pursued
here.
