\section{Finite-State \texorpdfstring{$\zeta$}{zeta}-Functions and Finite Parts}
\label{sec:preliminaries}

\subsection{Primitive orbit counts}

Let \(\mathcal A\) be a finite alphabet and let
\(A\in\{0,1\}^{\mathcal A\times\mathcal A}\) be a primitive adjacency
matrix. The associated one-sided subshift of finite type is
\[
  X_A^+
  =\bigl\{(x_n)_{n\ge 0}\in \mathcal A^{\NN}:
    A_{x_n,x_{n+1}}=1\ \text{for all }n\ge 0\bigr\},
\]
equipped with the left shift \(\sigma\). We write
\[
  a_n(A):=\Tr(A^n)
  =\#\Fix(\sigma^n)
\]
for the number of points of period dividing \(n\). The number of
primitive orbits of least period \(n\) is
\begin{equation}\label{eq:primitive-mobius}
  p_n(A)=\frac{1}{n}\sum_{d\mid n}\mu(d)\,a_{n/d}(A),
\end{equation}
and the inverse necklace relation is
\begin{equation}\label{eq:trace-primitive-identity}
  a_n(A)=\sum_{d\mid n}d\,p_d(A).
\end{equation}

\begin{proposition}[Trace, primitive orbits, and Euler product]
  \label{prop:prelim-trace-primitive}
  For every primitive matrix \(A\),
  \begin{equation}\label{eq:zeta-def}
    \zeta_A(z)
    :=\exp\!\left(\sum_{n\ge 1}\frac{\Tr(A^n)}{n}z^n\right)
    =\det(I-zA)^{-1}
    =\prod_{n\ge 1}(1-z^n)^{-p_n(A)}.
  \end{equation}
\end{proposition}

\begin{proof}
  By \eqref{eq:trace-primitive-identity},
  \begin{align*}
    \sum_{n\ge 1}\frac{a_n(A)}{n}z^n
    &=\sum_{n\ge 1}\sum_{d\mid n}\frac{d\,p_d(A)}{n}z^n \\
    &=\sum_{m\ge 1}\sum_{r\ge 1}\frac{p_m(A)}{r}z^{mr}
      =-\sum_{m\ge 1}p_m(A)\log(1-z^m).
  \end{align*}
  Exponentiating gives the Euler product, and the determinant identity
  is standard for finite matrices.
\end{proof}

If \(\lambda=\rho(A)\) is the Perron root and
\[
  \Lambda(A):=\max\{|\mu|:\mu\in\spec(A),\ \mu\neq\lambda\},
\]
then the primitive orbit counts satisfy the usual Perron asymptotic.

\begin{lemma}\label{lem:primitive-orbit-asymptotic}
For a primitive matrix \(A\),
\[
  p_n(A)=\frac{\lambda^n}{n}
  +O\!\left(\frac{\max\{\Lambda(A)^n,\lambda^{n/2}\}}{n}\right).
\]
\end{lemma}

\begin{proof}
  Perron--Frobenius theory gives
  \(
    a_n(A)=\lambda^n+O(\Lambda(A)^n)
  \).
  In \eqref{eq:primitive-mobius}, the divisor \(d=1\) gives the main
  term. For \(d\ge 2\), one has \(n/d\le n/2\), hence
  \[
    \sum_{\substack{d\mid n\\ d\ge 2}}\frac{|a_{n/d}(A)|}{n}
    \ll \frac{\lambda^{n/2}}{n},
  \]
  where the implied constant depends only on the Jordan data of the
  non-Perron spectrum of \(A\).
  This proves the estimate.
\end{proof}
\leanverified{paper_fredholm_witt_primitive_orbit_asymptotic}

\subsection{Finite-part constants}

Let \(z_\ast=\lambda^{-1}\). Since \(\lambda\) is a simple eigenvalue,
\(\zeta_A\) has a simple pole at \(z_\ast\), and we define
\begin{equation}\label{eq:residue-constant-def}
  C(A):=\lim_{z\to z_\ast}(1-\lambda z)\zeta_A(z).
\end{equation}
If \(\mu_2,\dots,\mu_d\) are the non-Perron eigenvalues of \(A\), then
\begin{equation}\label{eq:reduced-det-poly}
  F_A(t):=\frac{\det(I-tA/\lambda)}{1-t}
  =\prod_{j=2}^d\left(1-\frac{\mu_j}{\lambda}t\right)
\end{equation}
and
\begin{equation}\label{eq:reduced-det-constant}
  C(A)=F_A(1)^{-1}.
\end{equation}

Define the primitive Abel generating series
\[
  P_A(r):=\sum_{n\ge 1}p_n(A)\lambda^{-n}r^n,
  \qquad 0\le r<1.
\]
Because
\(
  p_n(A)\lambda^{-n}=1/n+O(\vartheta^n/n)
\)
for some \(\vartheta\in(0,1)\), the renormalised quantity
\[
  P_A(r)+\log(1-r)
\]
has a finite limit as \(r\uparrow 1\).

\begin{definition}[Abel finite-part constant]
  \label{def:abel-finite-part}
  We write
  \[
    \log\MM(A):=\lim_{r\uparrow 1}\bigl(P_A(r)+\log(1-r)\bigr).
  \]
  Indeed,
  \[
    P_A(r)+\log(1-r)
    =\sum_{n\ge 1}\left(p_n(A)\lambda^{-n}-\frac1n\right)r^n,
  \]
  and the estimate
  \(p_n(A)\lambda^{-n}=1/n+O(\vartheta^n/n)\) compares the difference
  termwise with the absolutely convergent series
  \(\sum_{n\ge 1}\vartheta^n r^n/n\).
\end{definition}

\begin{theorem}[Primitive-orbit closed form]
  \label{thm:finite-part-primitive-closed-form}
  Let \(A\) be primitive with Perron root \(\lambda>1\). Then
  \(\log\MM(A)\) exists and satisfies
  \[
    \log\MM(A)
    =\log C(A)+\sum_{k\ge 2}\frac{\mu(k)}{k}
      \log\det(I-\lambda^{-k}A)^{-1}.
  \]
  Equivalently,
  \[
    \log\MM(A)
    =\log C(A)
      +\sum_{n\ge 1}p_n(A)
       \bigl(\lambda^{-n}+\log(1-\lambda^{-n})\bigr).
  \]
  \leanverified{paper_fredholm_witt_finite_part_primitive_closed_form}
\end{theorem}

\begin{proof}
  By \Cref{prop:prelim-trace-primitive}, for \(|z|<\lambda^{-1}\),
  \[
    \log\det(I-zA)^{-1}
    =\sum_{n\ge 1}\frac{\Tr(A^n)}{n}z^n.
  \]
  Hence, for \(0\le r<1\),
  \begin{align*}
    \sum_{k\ge 1}\frac{\mu(k)}{k}
      \log\det(I-\lambda^{-k}r^kA)^{-1}
    &=\sum_{k\ge 1}\frac{\mu(k)}{k}
      \sum_{m\ge 1}\frac{\Tr(A^m)}{m}\lambda^{-km}r^{km} \\
    &=\sum_{n\ge 1}\frac{(\lambda^{-1}r)^n}{n}
      \sum_{k\mid n}\mu(k)\Tr(A^{n/k}) \\
    &=\sum_{n\ge 1}p_n(A)\lambda^{-n}r^n
      =P_A(r).
  \end{align*}
  The term \(k=1\) has asymptotic
  \[
    \log\det(I-\lambda^{-1}rA)^{-1}
    =-\log(1-r)+\log C(A)+o(1)
    \qquad (r\uparrow 1),
  \]
  while every term with \(k\ge 2\) converges absolutely at \(r=1\).
  This proves the first identity.

  For the second identity, use the Euler product in
  \Cref{prop:prelim-trace-primitive} and the identity
  \[
    -\sum_{k\ge 1}\frac{\mu(k)}{k}\log(1-x^k)=x
    \qquad (\abs{x}<1).
  \]
  Indeed,
  \[
    -\sum_{k\ge 1}\frac{\mu(k)}{k}\log(1-x^k)
    =\sum_{k\ge 1}\frac{\mu(k)}{k}
      \sum_{r\ge 1}\frac{x^{kr}}{r}
    =\sum_{n\ge 1}\frac{x^n}{n}\sum_{k\mid n}\mu(k)
    =x.
  \]
  Hence
  \[
    -\sum_{k\ge 2}\frac{\mu(k)}{k}\log(1-x^k)
    =x+\log(1-x).
  \]
  Setting \(x=\lambda^{-n}\) and \(n\) as the cycle-length index in the
  Euler product gives
  \[
    \log\MM(A)=\log C(A)+\sum_{n\ge 1}
    p_n(A)\bigl(\lambda^{-n}+\log(1-\lambda^{-n})\bigr).
  \]
  Rearranging the absolutely convergent sums yields the stated formula.
\end{proof}
\begin{remark}[Finite-state arithmetic]
  \label{rem:finite-state-arithmetic}
  For primitive finite matrices, the numbers \(p_n(A)\) are
  non-negative integers because they count primitive periodic orbits.
  This arithmetic feature belongs to the finite-state category and will
  not survive after passage to general trace-class operators.
\end{remark}
