\documentclass[11pt,reqno]{amsart}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}

\numberwithin{equation}{section}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\NN}{\mathbb{N}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\Xinf}{X_\infty}
\newcommand{\Unif}{\mathsf{Unif}}
\DeclareMathOperator{\Fold}{Fold}
\DeclareMathOperator{\Law}{Law}

\title[Truncation defects in finite-window Zeckendorf normalization]{Truncation defects in finite-window Zeckendorf normalization}
\author{Haobo Ma}
\address{AELF PTE LTD., \#14-02, Marina Bay Financial Centre Tower 1, 8 Marina Blvd, Singapore 018981}
\email{auric@aelf.io}
\author{Wenlin Zhang}
\thanks{Corresponding author.}
\address{National University of Singapore, Singapore}
\email{e1327962@u.nus.edu}
\subjclass{Primary 11B39; Secondary 37B10, 68Q45}
\keywords{Zeckendorf expansion, Fibonacci numeration, finite-window normalization, truncation defect}
\date{}

\begin{document}
\begin{abstract}
For $m\ge 1$, let $\Fold_m$ send a binary word of length $m$ to the first $m$
digits of the Zeckendorf expansion of its Fibonacci value, and let
$D_{n\to m}=\Fold_m\circ r_{n\to m}\oplus \pi_{n\to m}\circ \Fold_n$ be the
defect measuring the failure of finite-window normalization to commute with
truncation.  We prove that the adjacent defect at scale $m+1\to m$ is nonzero
if and only if the appended digit is $1$ and the lower block has Fibonacci
value at least $F_{m+1}$.  This yields a telescoping formula for $D_{n\to m}$
and, more importantly, a one-bit reduction of the whole upper tail: if
$\omega=(p,q)\in\Omega_m\times\Omega_{n-m}$, then
$D_{n\to m}(p,q)=\kappa_{m+1\to m}(p,\beta(q))$ for an explicit bit
$\beta(q)\in\{0,1\}$.  Consequently the uniform defect law is
$\mu_{n,m}=(1-\rho_{n-m})\delta_0+\rho_{n-m}\lambda_m$, where
$\rho_\ell=(2^\ell-(-1)^\ell)/(3\cdot 2^\ell)$ and $\lambda_m$ is the adjacent
law with the top digit fixed to $1$.  We determine $\lambda_m$ completely: its
nonzero support is the chain $v_{m,1}=0^{m-1}1$,
$v_{m,r}=0^{m-r}10 1^{r-2}$ $(2\le r\le m)$, and
$\lambda_m(v_{m,r})=(2^{m-r+1}-(-1)^{m-r+1})/(3\cdot 2^m)$.  Conditioned on a
nonzero defect, the corresponding depth therefore converges to the geometric
law $2^{-r}$.  We also identify $\beta(q)$ with the parity of the initial run
of $1$'s in $q$, prove that this is the minimal exact tail statistic for the
entire defect family, and derive the corresponding stationary one-step Markov
extension.  In the Markov setting the adjacent spectrum can also be determined
exactly: conditioned on a nonzero defect, the finite-window depth law is
independent of the tail length and converges, as the window grows, to the
geometric law $p_{10}p_{11}^{r-1}$.
\end{abstract}

\maketitle

\section{Introduction}

\subsection{Main results}

For $m\ge 1$, let
\[
\Fold_m:\Omega_m=\{0,1\}^m\longrightarrow X_m
\]
be the map that evaluates a binary word in the Fibonacci basis, replaces it by
the Zeckendorf expansion of the same value, and retains the first $m$ digits.
Here $X_m$ denotes the set of admissible words of length $m$, namely those
with no adjacent $1$'s.  If $m\le n$ and $\omega\in\Omega_n$, one may either
truncate first and normalize on the lower window, or normalize at level $n$
and then project to level $m$.  These two procedures need not agree, and we
write
\[
D_{n\to m}(\omega)
:=
\Fold_m(r_{n\to m}(\omega))
\oplus
\pi_{n\to m}(\Fold_n(\omega))
\in \{0,1\}^m
\]
for the resulting defect vector.  The object of the paper is to determine this
observable exactly.

The first theorem isolates the creation of a defect at a single scale.  If
$\eta=(p,b)\in\Omega_{m+1}$ and $s=N(p)$, then
Theorem~\ref{thm:adjacent-threshold} shows that
\[
K_{m+1\to m}(\eta)=b\,\mathbf{1}_{\{s\ge F_{m+1}\}}.
\]
Thus a new defect appears precisely when the appended digit is $1$ and the
lower block has already crossed the threshold $F_{m+1}$.  Once this one-step
criterion is available, the full defect follows from the telescoping formula
of Proposition~\ref{thm:telescoping} and the cocycle identity of
Proposition~\ref{prop:cocycle}.

The main structural point is that the entire upper block retains only one bit
of information.  Theorem~\ref{thm:tail-collapse} constructs a recursively
defined map $\beta:\bigsqcup_{r\ge 0}\Omega_r\to\{0,1\}$ such that for every
decomposition $\omega=(p,q)\in\Omega_m\times\Omega_{n-m}$,
\[
D_{n\to m}(p,q)=\kappa_{m+1\to m}(p,\beta(q)).
\]
Consequently the uniform defect law is a two-component mixture,
\[
\mu_{n,m}=(1-\rho_{n-m})\delta_0+\rho_{n-m}\lambda_m,
\qquad
\rho_\ell=\frac{2^\ell-(-1)^\ell}{3\cdot 2^\ell},
\]
where $\lambda_m$ is the adjacent law with top digit fixed to $1$.  For fixed
$m$, this gives the exact limiting law
\[
\frac23\delta_0+\frac13\lambda_m
\qquad (\ell\to\infty),
\]
with an oscillatory error of order $2^{-\ell}$.

The adjacent law itself admits a complete description.  By
Theorem~\ref{thm:adjacent-spectrum}, its nonzero support is the chain
\[
v_{m,1}=0^{m-1}1,
\qquad
v_{m,r}=0^{m-r}10 1^{r-2}\quad (2\le r\le m),
\]
and the corresponding weights are
\[
\lambda_m(v_{m,r})
=
\frac{2^{m-r+1}-(-1)^{m-r+1}}{3\cdot 2^m}.
\]
Hence the vector-valued law is supported on an explicit carry-depth chain, and
the depth conditioned on $D_{n\to m}\neq 0$ converges to the geometric law
$2^{-r}$ as $m\to\infty$.

The bit $\beta(q)$ has a direct dynamical meaning.  Theorem~\ref{thm:tail-rigidity}
identifies it with the parity of the initial run of $1$'s in the tail and
shows that no strictly coarser tail statistic determines the entire family of
defects exactly.  In that sense $\beta$ is the minimal exact tail factor for
the full defect family, a finite-state feature that may be compared with
Myhill--Nerode minimization \cite{Myhill1957,Nerode1958}.  The same
identification yields an exact stationary one-step Markov extension: after
conditioning on the boundary digit $\omega_m$, the defect law is again a
two-component mixture, now with coefficient
$p_{a1}(1-(-p_{11})^\ell)/(1+p_{11})$ at tail length $\ell$.  The conditional
adjacent laws may also be resolved explicitly: the boundary state $0$
contributes only depth $1$, the boundary state $1$ contributes the entire chain
$\{v_{m,r}\}_{r=1}^m$, and after conditioning on a nonzero defect the resulting
depth law is independent of the tail length and tends to the geometric law
$p_{10}p_{11}^{r-1}$ as $m\to\infty$.

\subsection{Context}

Classical background on Zeckendorf expansions goes back to
\cite{Lekkerkerker1952,Zeckendorf1972,Carlitz1968,CarlitzScovHoggatt1972,%
Fraenkel1985}.

Automata-theoretic normalization in Fibonacci numeration, and
more generally in order-two or confluent linear numeration systems, was
developed in
\cite{Frougny1988,Frougny1991,FrougnyConfluent1992,Frougny1992,%
FrougnySolomyak1996}
and
\cite{FrougnySakarovitch1999,Frougny1999,FrougnyS2010,LabbeLepsova2023,%
ShallitShan2024,MarquesTrojovsky2025}; see also
\cite{BerstelReutenauer2011} for related finite-state and rational-series
methods.

Scalar carry statistics and successor dynamics have been studied in
\cite{Frougny1996Successor,Frougny1997Sequential,%
AhlbachUsatineFrougnyPippenger2013,BertheFrougnyRigoSakarovitch2020}, while
probabilistic and combinatorial questions for Zeckendorf expansions appear in
\cite{KologluKoppMillerWang2011,DrmotaSteiner2002,Steiner2005,MillerWang2012}
and
\cite{Griffiths2010,BaratGrabner2014,Dekking2021,Hosten2024}.  For the
symbolic-dynamical background used here we appeal only to standard references
on the golden-mean shift and inverse limits \cite{LindMarcus1995,Kitchens1998};
for Markov chains we use
\cite{Parry1964,Norris1997,LevinPeresWilmer2009MarkovMixing}.

The defect studied here is of a different nature from the scalar carry
statistics that arise in the successor literature and from the output of a
normalization transducer viewed in isolation.  It records the vector
discrepancy between two finite-window routes to the same admissible block.
Our results are therefore entirely finite-scale and exact: within the
Fibonacci setting we obtain closed formulas for the adjacent threshold, the
telescoping decomposition, the full adjacent spectrum, the one-bit tail
factor, and the resulting uniform and stationary Markov laws.

\subsection{Organization}

Section~\ref{sec:adjacent} develops the finite-window Fold map and proves the
threshold criterion for the adjacent defect.  Section~\ref{sec:telescoping}
derives the telescoping and cocycle formulas.  Section~\ref{sec:uniform}
determines the adjacent law and the exact defect distribution under the
uniform measure.  Section~\ref{sec:markov} identifies the tail factor, extends
the argument to stationary one-step Markov inputs, and computes the exact
Markov depth spectrum together with its conditioned geometric limit.  Small-scale
examples are incorporated into the main text, where they help isolate
phenomena that are invisible at the level of scalar defect indicators.

\section{Finite-window normalization}
\label{sec:adjacent}

We begin by fixing notation and collecting the elementary normalization facts
used throughout.  Digits are indexed from the low-order end: position $k$
carries Fibonacci weight $F_{k+1}$, where $F_1=F_2=1$ and
$F_{j+2}=F_{j+1}+F_j$.  Thus $N(10)=F_2=1$ and $N(01)=F_3=2$, and the prefix
map $r_{n\to m}$ keeps the first $m$ low-order digits.

\subsection{The finite-window Fold map}

For $m\ge 1$ set
\[
\Omega_m:=\{0,1\}^m,
\qquad
X_m:=\{x=(x_1,\dots,x_m)\in\{0,1\}^m:\ x_ix_{i+1}=0\text{ for }1\le i<m\}.
\]
For $\omega=\omega_1\cdots\omega_m\in\Omega_m$, define its Fibonacci value by
\[
N(\omega):=\sum_{k=1}^m\omega_kF_{k+1}.
\]
For $N\ge 0$, let $Z(N)=(z_k)_{k\ge 1}\in\{0,1\}^{\NN}$ be the unique finitely
supported sequence such that
\[
N=\sum_{k\ge 1}z_kF_{k+1}
\qquad\text{and}\qquad
z_kz_{k+1}=0\ \text{ for all }k\ge 1;
\]
we set $Z(0)=0^\infty$.  The Fold map is
\[
\Fold_m(\omega):=\pi_{\infty\to m}(Z(N(\omega)))\in X_m,
\]
where $\pi_{\infty\to m}(x_1,x_2,\dots)=(x_1,\dots,x_m)$.

For $n\ge m$ we use the truncation maps
\[
r_{n\to m}(\omega_1\cdots\omega_n):=\omega_1\cdots\omega_m,
\qquad
\pi_{n\to m}(x_1,\dots,x_n):=(x_1,\dots,x_m),
\]
and, on the ambient cube,
\[
\tau_{n\to m}(u_1,\dots,u_n):=(u_1,\dots,u_m).
\]
We use $\tau_{n\to m}$ only for defect vectors in the ambient cube.

The following elementary facts will be used repeatedly.

\begin{lemma}
\label{lem:max-prefix-value}
For $m\ge 1$,
\[
\max_{x\in X_m}N(x)=F_{m+2}-1.
\]
\end{lemma}

\begin{proof}
Let $M_m:=\max_{x\in X_m}N(x)$ and set $M_0:=0$.  Then $M_1=1=F_3-1$.  For
$m\ge 2$, either $x_m=0$, in which case $N(x)\le M_{m-1}$, or $x_m=1$, in
which case $x_{m-1}=0$ and
\[
N(x)=F_{m+1}+N(x_1,\dots,x_{m-2})\le F_{m+1}+M_{m-2}.
\]
Thus
\[
M_m=\max\{M_{m-1},\,F_{m+1}+M_{m-2}\}.
\]
Using the induction hypothesis,
\[
M_m=\max\{F_{m+1}-1,\,F_{m+1}+F_m-1\}=F_{m+2}-1.
\]
\end{proof}

\begin{proposition}
\label{prop:fold-residue}
For every $m\ge 1$ and every $\omega\in\Omega_m$, write
\[
N(\omega)=qF_{m+2}+r,
\qquad
q\in\{0,1\},
\qquad
0\le r<F_{m+2}.
\]
Then
\[
\Fold_m(\omega)=\pi_{\infty\to m}(Z(r)).
\]
Equivalently, for $x\in X_m$,
\[
\Fold_m(\omega)=x
\quad\Longleftrightarrow\quad
N(\omega)\equiv N(x)\pmod{F_{m+2}}.
\]
\end{proposition}

\begin{proof}
Since
\[
N(\omega)\le \sum_{k=1}^mF_{k+1}=F_{m+3}-2<2F_{m+2},
\]
we have $q\in\{0,1\}$.  If $q=0$, there is nothing to prove.  If $q=1$, then
\[
0\le r=N(\omega)-F_{m+2}\le F_{m+1}-2<F_{m+1}.
\]
Hence $Z(r)$ has no nonzero digit in positions $m,m+1,\dots$, so appending a
single $1$ in position $m+1$ produces the Zeckendorf expansion of
$N(\omega)=F_{m+2}+r$.  Its first $m$ digits are exactly those of $Z(r)$,
which proves the first assertion.

For the equivalence, note that any $x\in X_m$ is already the Zeckendorf
expansion of its own value $N(x)$, and Lemma~\ref{lem:max-prefix-value} gives
$N(x)\le F_{m+2}-1$.  Thus $\pi_{\infty\to m}(Z(r))=x$ if and only if
$r=N(x)$.
\end{proof}

\begin{lemma}
\label{lem:prefix-functoriality}
If $n\ge k\ge m$, then
\[
\pi_{n\to m}=\pi_{k\to m}\circ \pi_{n\to k},
\qquad
r_{n\to m}=r_{k\to m}\circ r_{n\to k},
\qquad
\tau_{n\to m}=\tau_{k\to m}\circ \tau_{n\to k}.
\]
Moreover, for all $u,v\in\{0,1\}^n$,
\[
\tau_{n\to m}(u\oplus v)=\tau_{n\to m}(u)\oplus\tau_{n\to m}(v).
\]
\end{lemma}

\begin{proof}
This is immediate from the definitions of the truncation maps.
\end{proof}

\begin{example}
\label{ex:noncommuting-square}
Take $n=2$, $m=1$, and $\omega=11$.  Then
\[
r_{2\to 1}(11)=1,
\qquad
\Fold_1(r_{2\to 1}(11))=1.
\]
On the other hand, $N(11)=F_2+F_3=3=F_4$, so the Zeckendorf expansion of
$N(11)$ begins with $00$.  Hence
\[
\Fold_2(11)=00,
\qquad
\pi_{2\to 1}(\Fold_2(11))=0.
\]
Thus truncation does not commute with normalization.
\end{example}

\subsection{Adjacent defects}

\noindent
The first nontrivial case is the adjacent step $m+1\to m$.  The point is that
the indicator of the adjacent defect is governed by a single Fibonacci
threshold, while the defect vector itself remains sensitive to the precise
position of the prefix value within the threshold range.

\begin{definition}
\label{def:adjacent-defect}
For $m\ge 1$ and $\eta\in\Omega_{m+1}$, the adjacent defect is
\[
\kappa_{m+1\to m}(\eta)
:=
\Fold_m(r_{m+1\to m}(\eta))
\oplus
\pi_{m+1\to m}(\Fold_{m+1}(\eta))
\in\{0,1\}^m.
\]
Its indicator is
\[
K_{m+1\to m}(\eta):=\mathbf{1}_{\{\kappa_{m+1\to m}(\eta)\ne 0\}}.
\]
\end{definition}

\begin{theorem}[Threshold criterion for the adjacent defect]
\label{thm:adjacent-threshold}
Let $m\ge 1$, let $p\in\Omega_m$, let $b\in\{0,1\}$, and write
\[
\eta=(p,b)\in\Omega_{m+1},
\qquad
s:=N(p).
\]
Then
\[
K_{m+1\to m}(\eta)=b\,\mathbf{1}_{\{s\ge F_{m+1}\}}.
\]
More precisely,
\[
\kappa_{m+1\to m}(p,0)=0,
\]
and
\[
\kappa_{m+1\to m}(p,1)=
\begin{cases}
0,& s<F_{m+1},\\[4pt]
\pi_{\infty\to m}(Z(s\bmod F_{m+2}))
\oplus
\pi_{\infty\to m}(Z(s-F_{m+1})),
& s\ge F_{m+1}.
\end{cases}
\]
When $s\ge F_{m+1}$, the two admissible words in the second line are distinct;
in particular the defect is nonzero.
\end{theorem}

\begin{proof}
If $b=0$, then $N(p,0)=s$, so
\[
\Fold_{m+1}(p,0)=\pi_{\infty\to m+1}(Z(s))
\]
and therefore
\[
\pi_{m+1\to m}(\Fold_{m+1}(p,0))
=
\pi_{\infty\to m}(Z(s))
=
\Fold_m(p).
\]
Thus $\kappa_{m+1\to m}(p,0)=0$.

Assume now that $b=1$.  If $s<F_{m+1}$, then the Zeckendorf expansion of $s$
has no nonzero digit in positions $m,m+1,\dots$.  Consequently
$Z(s+F_{m+2})$ is obtained from $Z(s)$ by inserting a single $1$ in position
$m+1$, and hence
\[
\pi_{m+1\to m}(\Fold_{m+1}(p,1))
=
\pi_{\infty\to m}(Z(s))
=
\Fold_m(p).
\]
So the defect vanishes.

Suppose finally that $s\ge F_{m+1}$.  Write
\[
s=F_{m+1}+t
\qquad\text{with}\qquad
0\le t<F_{m+2}.
\]
By Proposition~\ref{prop:fold-residue},
\[
\Fold_m(p)=\pi_{\infty\to m}(Z(s\bmod F_{m+2})).
\]
Since
\[
N(p,1)=s+F_{m+2}=F_{m+3}+t,
\]
the same proposition at level $m+1$ gives
\[
\Fold_{m+1}(p,1)=\pi_{\infty\to m+1}(Z(t)),
\]
and therefore
\[
\pi_{m+1\to m}(\Fold_{m+1}(p,1))
=
\pi_{\infty\to m}(Z(t))
=
\pi_{\infty\to m}(Z(s-F_{m+1})).
\]
This is the displayed formula.

If the two admissible words in the second line were equal, then
Proposition~\ref{prop:fold-residue} at level $m$ would imply
$s\bmod F_{m+2}=t=s-F_{m+1}$, which is impossible because
$0<F_{m+1}<F_{m+2}$.  Hence the defect is nonzero.
\end{proof}

\begin{example}[Normalization cascade without a truncation defect]
\label{ex:cascade-without-defect}
Take $m=4$, $p=1100$, and append $b=1$.  Normalizing the raw word $(p,1)=11001$
produces $\Fold_5(11001)=00101$, so a nontrivial carry cascade does occur.
Nevertheless
\[
\Fold_4(p)=0010
\qquad\text{and}\qquad
\pi_{5\to 4}(\Fold_5(11001))=0010,
\]
hence $\kappa_{5\to 4}(p,1)=0$.  In particular, the existence of a
normalization cascade does not by itself force a truncation defect.
\end{example}

\section{Telescoping and cocycle formulas}
\label{sec:telescoping}

We now pass from the adjacent defect to the full defect.  The argument in this
section is purely formal: once the one-step defect is understood, the global
defect is recovered by iterating the truncation and projection maps.

\begin{definition}
\label{def:global-defect}
Let $n\ge m$.  For $\omega\in\Omega_n$, define the global defect
\[
D_{n\to m}(\omega)
:=
\Fold_m(r_{n\to m}(\omega))
\oplus
\pi_{n\to m}(\Fold_n(\omega))
\in\{0,1\}^m.
\]
\end{definition}

When $n=m$, the definition gives $D_{m\to m}=0$.  For $n>m$, the defect
measures the accumulated noncommutation between the two routes from $\Omega_n$
to $X_m$.

For $v\in\{0,1\}^m$, write
\[
\lvert v\rvert_0:=\#\{i:\ v_i\ne 0\}.
\]

For fixed $n>m$ and $\omega\in\Omega_n$, set
\[
A_j(\omega):=\Fold_j(r_{n\to j}(\omega)),
\qquad
B_j(\omega):=\pi_{n\to j}(\Fold_n(\omega)),
\]
and
\[
\Delta_j(\omega):=A_j(\omega)\oplus B_j(\omega)
\qquad (m\le j\le n).
\]
Then $\Delta_m(\omega)=D_{n\to m}(\omega)$ and $\Delta_n(\omega)=0$.

\begin{lemma}
\label{lem:adjacent-recursion}
For every $m<j\le n$,
\[
\Delta_{j-1}(\omega)
=
\kappa_{j\to j-1}(r_{n\to j}(\omega))
\oplus
\tau_{j\to j-1}(\Delta_j(\omega)).
\]
\end{lemma}

\begin{proof}
By Lemma~\ref{lem:prefix-functoriality},
\[
B_{j-1}(\omega)=\pi_{j\to j-1}(B_j(\omega)),
\qquad
A_{j-1}(\omega)=\Fold_{j-1}(r_{j\to j-1}(r_{n\to j}(\omega))).
\]
Therefore
\[
\begin{aligned}
\Delta_{j-1}(\omega)
&=A_{j-1}(\omega)\oplus B_{j-1}(\omega) \\
&=
\Bigl(A_{j-1}(\omega)\oplus \pi_{j\to j-1}(A_j(\omega))\Bigr)
\oplus
\Bigl(\pi_{j\to j-1}(A_j(\omega))\oplus \pi_{j\to j-1}(B_j(\omega))\Bigr).
\end{aligned}
\]
The first bracket is $\kappa_{j\to j-1}(r_{n\to j}(\omega))$, while the second
equals $\tau_{j\to j-1}(\Delta_j(\omega))$ because $\tau_{j\to j-1}$ agrees
with $\pi_{j\to j-1}$ on admissible words and is an xor homomorphism.
\end{proof}

\begin{proposition}[Telescoping identity]
\label{thm:telescoping}
For every $n>m$ and every $\omega\in\Omega_n$,
\[
D_{n\to m}(\omega)
=
\bigoplus_{j=m}^{n-1}
\tau_{j\to m}\bigl(\kappa_{j+1\to j}(r_{n\to j+1}(\omega))\bigr).
\]
\end{proposition}

\begin{proof}
We claim that for each $j\in\{m,\dots,n\}$,
\[
\Delta_m(\omega)
=
\tau_{j\to m}(\Delta_j(\omega))
\oplus
\bigoplus_{t=m}^{j-1}
\tau_{t\to m}\bigl(\kappa_{t+1\to t}(r_{n\to t+1}(\omega))\bigr).
\]
For $j=m$ this is tautological.  Assume it holds at some
$j\in\{m,\dots,n-1\}$.  Using Lemma~\ref{lem:adjacent-recursion},
\[
\Delta_j(\omega)
=
\kappa_{j+1\to j}(r_{n\to j+1}(\omega))
\oplus
\tau_{j+1\to j}(\Delta_{j+1}(\omega)).
\]
Applying $\tau_{j\to m}$ and using Lemma~\ref{lem:prefix-functoriality} gives
\[
\tau_{j\to m}(\Delta_j(\omega))
=
\tau_{j\to m}\bigl(\kappa_{j+1\to j}(r_{n\to j+1}(\omega))\bigr)
\oplus
\tau_{j+1\to m}(\Delta_{j+1}(\omega)).
\]
Substituting this into the induction hypothesis proves the claim for $j+1$.
Taking $j=n$ and recalling that $\Delta_n(\omega)=0$ finishes the proof.
\end{proof}

\begin{proposition}[Cocycle identity]
\label{prop:cocycle}
If $m<k<n$, then for every $\omega\in\Omega_n$,
\[
D_{n\to m}(\omega)
=
D_{k\to m}(r_{n\to k}(\omega))
\oplus
\tau_{k\to m}(D_{n\to k}(\omega)).
\]
Consequently,
\[
\lvert D_{n\to m}(\omega)\rvert_0
\le
\lvert D_{k\to m}(r_{n\to k}(\omega))\rvert_0
+
\lvert D_{n\to k}(\omega)\rvert_0.
\]
\end{proposition}

\begin{proof}
Split the sum in Proposition~\ref{thm:telescoping} at the intermediate scale $k$:
\[
\begin{aligned}
D_{n\to m}(\omega)
&=
\bigoplus_{j=m}^{k-1}
\tau_{j\to m}\bigl(\kappa_{j+1\to j}(r_{n\to j+1}(\omega))\bigr) \\
&\quad\oplus
\bigoplus_{j=k}^{n-1}
\tau_{j\to m}\bigl(\kappa_{j+1\to j}(r_{n\to j+1}(\omega))\bigr).
\end{aligned}
\]
The first block is exactly $D_{k\to m}(r_{n\to k}(\omega))$.  In the second
block, Lemma~\ref{lem:prefix-functoriality} gives
$\tau_{j\to m}=\tau_{k\to m}\circ\tau_{j\to k}$, so the second block is
$\tau_{k\to m}(D_{n\to k}(\omega))$.  This proves the cocycle identity.

The Hamming estimate follows from
\[
\lvert u\oplus v\rvert_0\le \lvert u\rvert_0+\lvert v\rvert_0
\]
and the fact that prefix projection cannot increase Hamming weight.
\end{proof}

\begin{example}[A lower-level defect with trivial top indicator]
\label{ex:defect-without-top-indicator}
Let $n=4$, $m=2$, and $\omega=0110$.  Direct computation gives
\[
D_{4\to 2}(\omega)=(0,1)\ne 0,
\]
whereas the top adjacent indicator satisfies
\[
K_{4\to 3}(\omega)=0.
\]
Thus the global defect can be created strictly below the top scale.
\end{example}

\section{The adjacent law and uniform inputs}
\label{sec:uniform}

We next turn to exact distributional questions under the uniform law on raw
words.  The first step is to describe the adjacent law itself.

\subsection{The adjacent law}

\begin{proposition}[Exact one-step defect count]
\label{thm:explicit-count}
For $m\ge 1$, let
\[
C_m:=\#\{\alpha\in\Omega_m:\ N(\alpha)<F_{m+1}\}.
\]
Then
\[
C_1=1,
\qquad
C_2=2,
\qquad
C_m=2^{m-2}+C_{m-2}\quad (m\ge 3).
\]
Consequently,
\[
C_{2r}=\frac{2^{2r}+2}{3},
\qquad
C_{2r+1}=\frac{2^{2r+1}+1}{3}
\qquad (r\ge 1),
\]
and
\[
\PP_{\Unif_{m+1}}[K_{m+1\to m}=1]
=
\frac{2^m-C_m}{2^{m+1}}
=
\frac{\lfloor 2^{m+1}/3\rfloor}{2^{m+1}}.
\]
\end{proposition}

\begin{proof}
Consider the last two digits of $\alpha\in\Omega_m$.

If $\alpha_m=1$, then $N(\alpha)\ge F_{m+1}$, so no such word contributes to
$C_m$.

If $\alpha_m=0$ and $\alpha_{m-1}=0$, then
\[
N(\alpha)=N(\alpha_1,\dots,\alpha_{m-2})
\le
\sum_{k=1}^{m-2}F_{k+1}
=
F_{m+1}-2
<
F_{m+1},
\]
so every choice of the first $m-2$ digits is allowed.  This contributes
$2^{m-2}$ words.

If $\alpha_m=0$ and $\alpha_{m-1}=1$, then
\[
N(\alpha)=F_m+N(\alpha_1,\dots,\alpha_{m-2}),
\]
and the condition $N(\alpha)<F_{m+1}$ becomes
$N(\alpha_1,\dots,\alpha_{m-2})<F_{m-1}$.  This contributes $C_{m-2}$ words.

Hence $C_m=2^{m-2}+C_{m-2}$ for $m\ge 3$.  The initial values are immediate:
$C_1=1$ because only $0$ has value $<F_2=1$, and $C_2=2$ because the words
$00$ and $10$ are exactly those of value $<F_3=2$.

The closed forms follow by unrolling the recurrence.  Setting $C_0:=1$, one
finds
\[
C_{2r}=2^{2r-2}+2^{2r-4}+\cdots+2^0+C_0=\frac{2^{2r}+2}{3},
\]
and
\[
C_{2r+1}=2^{2r-1}+2^{2r-3}+\cdots+2^1+C_1=\frac{2^{2r+1}+1}{3}.
\]
Finally, Theorem~\ref{thm:adjacent-threshold} shows that
$K_{m+1\to m}(\alpha,b)=1$ exactly when $b=1$ and $N(\alpha)\ge F_{m+1}$, so
there are $2^m-C_m$ defect-producing words in $\Omega_{m+1}$.  Dividing by
$2^{m+1}$ gives the probability formula.
\end{proof}

For $v\in\{0,1\}^m$, define
\[
\lambda_m(v):=2^{-m}\#\{p\in\Omega_m:\ \kappa_{m+1\to m}(p,1)=v\}.
\]

\begin{theorem}[Complete adjacent defect spectrum]
\label{thm:adjacent-spectrum}
For $m\ge 1$ and $1\le r\le m$, let
\[
I_{m,r}
:=
[F_{m+3}-F_{m-r+3},\,F_{m+3}-F_{m-r+2}-1].
\]
Let $v_{m,1}=0^{m-1}1$, and for $2\le r\le m$ let
\[
v_{m,r}:=0^{m-r}10 1^{r-2}
\]
in the low-order-first notation of the paper.  If $p\in\Omega_m$ and
$s:=N(p)$, then
\[
\kappa_{m+1\to m}(p,1)=v_{m,r}
\quad\Longleftrightarrow\quad
s\in I_{m,r}.
\]
Consequently
\[
\{\kappa_{m+1\to m}(p,1):\ p\in\Omega_m,\ \kappa_{m+1\to m}(p,1)\ne 0\}
=
\{v_{m,1},\dots,v_{m,m}\},
\]
and
\[
\lambda_m(v_{m,r})
=
\frac{2^{m-r+1}-(-1)^{m-r+1}}{3\cdot 2^m}
\qquad (1\le r\le m).
\]
\end{theorem}

\begin{proof}
The intervals $I_{m,r}$ are disjoint, their lengths are
\[
\lvert I_{m,r}\rvert=F_{m-r+1},
\]
and their union is exactly $[F_{m+1},F_{m+3}-2]$.  Indeed, the right endpoint
of $I_{m,r}$ is one less than the left endpoint of $I_{m,r+1}$, so the
intervals partition the whole threshold range on which
Theorem~\ref{thm:adjacent-threshold} allows a nonzero defect.

If $r=1$, then every $s\in I_{m,1}$ has the form $s=F_{m+1}+u$ with
$0\le u<F_m$.  Since $u<F_m$, the Zeckendorf expansion of $s$ is obtained by
placing a single digit $1$ in position $m$ above $Z(u)$.  Hence
\[
\pi_{\infty\to m}(Z(s))
=
\pi_{\infty\to m}(Z(u))\oplus v_{m,1},
\qquad
\pi_{\infty\to m}(Z(s-F_{m+1}))
=
\pi_{\infty\to m}(Z(u)),
\]
and Theorem~\ref{thm:adjacent-threshold} gives
\[
\kappa_{m+1\to m}(p,1)=v_{m,1}.
\]

Fix $r\ge 2$ and write
\[
s=F_{m+3}-F_{m-r+3}+u,
\qquad
0\le u<F_{m-r+1}.
\]
Set
\[
A_{m,r}:=F_{m+1}-F_{m-r+3},
\qquad
B_{m,r}:=F_{m+2}-F_{m-r+3}.
\]
Then
\[
s\bmod F_{m+2}=A_{m,r}+u,
\qquad
s-F_{m+1}=B_{m,r}+u.
\]

If $r=2q$, repeated use of $F_{j+1}-F_{j-1}=F_j$ gives
\[
A_{m,2q}
=
\sum_{j=0}^{q-2}F_{m-2j},
\qquad
B_{m,2q}
=
\sum_{j=0}^{q-2}F_{m+1-2j}+F_{m-2q+2}.
\]
If $r=2q+1$, the same identity gives
\[
A_{m,2q+1}
=
\sum_{j=0}^{q-2}F_{m-2j}+F_{m-2q+1},
\qquad
B_{m,2q+1}
=
\sum_{j=0}^{q-1}F_{m+1-2j}.
\]
Since $u<F_{m-r+1}$, the support of $Z(u)$ is contained in positions
$1,\dots,m-r$.  Therefore the first $m$ digits of $Z(A_{m,r}+u)$ and
$Z(B_{m,r}+u)$ are obtained by placing the same lower block $Z(u)$ beneath the
fixed Zeckendorf heads encoded by the displayed sums.  If $r=2q$, those heads
occupy the positions
\[
\{m-1,m-3,\dots,m-2q+3\}
\quad\text{and}\quad
\{m,m-2,\dots,m-2q+4,m-2q+1\},
\]
whereas if $r=2q+1$, they occupy the positions
\[
\{m-1,m-3,\dots,m-2q+3,m-2q\}
\quad\text{and}\quad
\{m,m-2,\dots,m-2q+2\}.
\]
In both cases the xor of the two fixed heads is
$v_{m,r}=0^{m-r}10 1^{r-2}$.  Hence
\[
\pi_{\infty\to m}(Z(s\bmod F_{m+2}))
\oplus
\pi_{\infty\to m}(Z(s-F_{m+1}))
=
v_{m,r},
\]
and Theorem~\ref{thm:adjacent-threshold} yields
\[
\kappa_{m+1\to m}(p,1)=v_{m,r}.
\]
Since Theorem~\ref{thm:adjacent-threshold} also shows that the adjacent defect
is nonzero exactly on $[F_{m+1},F_{m+3}-2]$, the classification by the
intervals $I_{m,r}$ is complete.

For $1\le r\le m$, let
\[
M_{m,r}
:=
\#\{p\in\Omega_m:\ \kappa_{m+1\to m}(p,1)=v_{m,r}\}.
\]
If $r\ge 2$, then $I_{m,r}\subset[F_{m+2},F_{m+3}-2]$, so every
$p\in\Omega_m$ with $N(p)\in I_{m,r}$ has top digit $p_m=1$.  Writing
$p=(u,1)$ with $u\in\Omega_{m-1}$ and using
\[
I_{m,r}=F_{m+1}+I_{m-1,r-1},
\]
one obtains a bijection between the words counted by $M_{m,r}$ and those
counted by $M_{m-1,r-1}$.  Thus
\[
M_{m,r}=M_{m-1,r-1}
\qquad (2\le r\le m).
\]
By Proposition~\ref{thm:explicit-count},
\[
\sum_{r=1}^m M_{m,r}
=
\#\{p\in\Omega_m:\ \kappa_{m+1\to m}(p,1)\ne 0\}
=
\left\lfloor\frac{2^{m+1}}{3}\right\rfloor.
\]
Therefore
\[
M_{m,1}
=
\left\lfloor\frac{2^{m+1}}{3}\right\rfloor
-
\left\lfloor\frac{2^m}{3}\right\rfloor
=
\frac{2^m-(-1)^m}{3},
\]
and repeated use of the recursion gives
\[
M_{m,r}=M_{m-r+1,1}
=
\frac{2^{m-r+1}-(-1)^{m-r+1}}{3}.
\]
Dividing by $2^m$ proves the formula for $\lambda_m(v_{m,r})$.
\end{proof}

\begin{example}
\label{ex:adjacent-spectrum-m4}
For $m=4$ the interval decomposition is
\[
I_{4,1}=[5,7],
\qquad
I_{4,2}=[8,9],
\qquad
I_{4,3}=\{10\},
\qquad
I_{4,4}=\{11\}.
\]
Hence the nonzero adjacent defects are
\[
v_{4,1}=0001,
\qquad
v_{4,2}=0010,
\qquad
v_{4,3}=0101,
\qquad
v_{4,4}=1011,
\]
and their weights are
\[
\lambda_4(0001)=\frac{5}{16},
\qquad
\lambda_4(0010)=\frac{3}{16},
\qquad
\lambda_4(0101)=\lambda_4(1011)=\frac{1}{16}.
\]
This is the first scale at which the entire carry-depth chain
$\{v_{m,r}\}_{r=1}^m$ is visible.
\end{example}

\subsection{Tail collapse and exact distribution}

For $n>m$ and $v\in\{0,1\}^m$, define
\[
\mu_{n,m}(v):=2^{-n}\#\{\omega\in\Omega_n:\ D_{n\to m}(\omega)=v\}.
\]
Thus $\mu_{n,m}$ is the exact defect law under the uniform raw-word measure.
Theorem~\ref{thm:tail-collapse} shows that, after the upper block has been
marginalized, a single recursively defined tail bit is sufficient to recover
the full distribution.

\begin{lemma}
\label{lem:zero-barrier}
Let $m\ge 1$, let $p\in\Omega_m$, and let $b\in\{0,1\}$.  Then
\[
\tau_{m+1\to m}\bigl(\kappa_{m+2\to m+1}(p,0,b)\bigr)=0.
\]
\end{lemma}

\begin{proof}
If $b=0$, Theorem~\ref{thm:adjacent-threshold} gives
$\kappa_{m+2\to m+1}(p,0,0)=0$.

If $b=1$, write
\[
s:=N(p,0)=N(p).
\]
Since the $(m+1)$-st digit of $(p,0)$ is zero,
\[
s\le \sum_{k=1}^m F_{k+1}=F_{m+2}-2<F_{m+2}.
\]
Theorem~\ref{thm:adjacent-threshold} at level $m+2\to m+1$ therefore gives
$\kappa_{m+2\to m+1}(p,0,1)=0$.
\end{proof}

\begin{lemma}
\label{lem:double-one-reduction}
Let $m\ge 1$, let $p\in\Omega_m$, and let $b\in\{0,1\}$.  Then
\[
\tau_{m+2\to m}\bigl(\kappa_{m+3\to m+2}(p,1,1,b)\bigr)
=
\kappa_{m+1\to m}(p,b).
\]
\end{lemma}

\begin{proof}
If $b=0$, both sides vanish by Theorem~\ref{thm:adjacent-threshold}.

Assume $b=1$ and write
\[
s:=N(p).
\]
Since $p\in\Omega_m$,
\[
0\le s\le F_{m+3}-2.
\]
At level $m+3\to m+2$, the prefix $(p,1,1)$ has value
\[
N(p,1,1)=s+F_{m+2}+F_{m+3}=s+F_{m+4}.
\]
Theorem~\ref{thm:adjacent-threshold} therefore yields
\[
\tau_{m+2\to m}\bigl(\kappa_{m+3\to m+2}(p,1,1,1)\bigr)
=
\pi_{\infty\to m}(Z(s))
\oplus
\pi_{\infty\to m}(Z(s+F_{m+2})).
\]

If $s<F_{m+1}$, then $s+F_{m+2}<F_{m+3}$ and the Zeckendorf expansion of
$s+F_{m+2}$ is obtained by inserting a single $1$ in position $m+1$ above
$Z(s)$.  Hence
\[
\pi_{\infty\to m}(Z(s+F_{m+2}))=\pi_{\infty\to m}(Z(s)),
\]
so the left-hand side is zero.  Theorem~\ref{thm:adjacent-threshold} gives
$\kappa_{m+1\to m}(p,1)=0$ in the same range.

Assume now that $s\ge F_{m+1}$.  Since $s<F_{m+3}$,
\[
0\le s-F_{m+1}<F_{m+2},
\]
and
\[
s+F_{m+2}=F_{m+3}+(s-F_{m+1}).
\]
Thus the first $m$ digits of $Z(s+F_{m+2})$ are exactly those of
$Z(s-F_{m+1})$, so
\[
\pi_{\infty\to m}(Z(s+F_{m+2}))
=
\pi_{\infty\to m}(Z(s-F_{m+1})).
\]
Also,
\[
\pi_{\infty\to m}(Z(s))
=
\pi_{\infty\to m}(Z(s\bmod F_{m+2}))
\]
because $s<F_{m+3}$.  Theorem~\ref{thm:adjacent-threshold} at level
$m+1\to m$ therefore gives
\[
\pi_{\infty\to m}(Z(s))
\oplus
\pi_{\infty\to m}(Z(s+F_{m+2}))
=
\kappa_{m+1\to m}(p,1).
\]
This proves the claim.
\end{proof}

\begin{theorem}[Tail collapse]
\label{thm:tail-collapse}
Fix $n>m$ and write $\ell:=n-m$.  Define
\[
\beta:\bigsqcup_{r\ge 0}\Omega_r\longrightarrow\{0,1\}
\]
by the recursion
\[
\beta(\varnothing)=0,
\qquad
\beta(1)=1,
\qquad
\beta(0\alpha)=0,
\qquad
\beta(10\alpha)=1,
\qquad
\beta(11\alpha)=\beta(\alpha).
\]
Every nonempty finite binary word has exactly one of the forms $0\alpha$,
$1$, $10\alpha$, or $11\alpha$, so the recursion is unambiguous.
Then for every $p\in\Omega_m$ and every $q\in\Omega_\ell$,
\[
D_{n\to m}(p,q)=\kappa_{m+1\to m}(p,\beta(q)).
\]
For $\ell\ge 0$ set
\[
\rho_\ell:=2^{-\ell}\#\{q\in\Omega_\ell:\ \beta(q)=1\}.
\]
Then
\[
\mu_{n,m}(v)=
\begin{cases}
(1-\rho_\ell)+\rho_\ell\lambda_m(0),& v=0,\\[4pt]
\rho_\ell\lambda_m(v),& v\ne 0.
\end{cases}
\]
Equivalently, if $\delta_0$ denotes the point mass at $0\in\{0,1\}^m$, then
\[
\mu_{n,m}=(1-\rho_\ell)\delta_0+\rho_\ell\lambda_m.
\]
Moreover,
\[
\rho_\ell=\frac{2^\ell-(-1)^\ell}{3\cdot 2^\ell}
\qquad (\ell\ge 0),
\]
and therefore
\[
\PP_{\Unif_n}[D_{n\to m}\ne 0]
=
\rho_{n-m}\,\frac{\lfloor 2^{m+1}/3\rfloor}{2^m}
=
\frac{(2^{n-m}-(-1)^{n-m})\lfloor 2^{m+1}/3\rfloor}{3\cdot 2^n}.
\]
\end{theorem}

\begin{proof}
For $r\ge 0$, for $p\in\Omega_m$, and for $q\in\Omega_r$, write
\[
E_r(p,q):=D_{m+r\to m}(p,q).
\]
The identity
\[
E_r(p,q)=\kappa_{m+1\to m}(p,\beta(q))
\]
is established by induction on $r$.

For $r=0$, one has $q=\varnothing$, $E_0(p,\varnothing)=0$, and
$\beta(\varnothing)=0$.

Assume $r\ge 1$ and that the claim holds for shorter tails.

If $q=1$, then
\[
E_1(p,1)=D_{m+1\to m}(p,1)=\kappa_{m+1\to m}(p,1)
=
\kappa_{m+1\to m}(p,\beta(1)).
\]
This settles the length-one tail.

If $q=0\alpha$, then Proposition~\ref{prop:cocycle} with the intermediate scale
$m+1$ gives
\[
E_r(p,0\alpha)
=
D_{m+1\to m}(p,0)
\oplus
\tau_{m+1\to m}(D_{m+r\to m+1}(p,0,\alpha)).
\]
The first term is zero.  The induction hypothesis at level $m+1$ gives
\[
D_{m+r\to m+1}(p,0,\alpha)
=
\kappa_{m+2\to m+1}(p,0,\beta(\alpha)).
\]
Lemma~\ref{lem:zero-barrier} then implies $E_r(p,0\alpha)=0$, which equals
$\kappa_{m+1\to m}(p,\beta(0\alpha))$.

If $q=10\alpha$, then Proposition~\ref{prop:cocycle} with the intermediate
scale $m+2$ gives
\[
E_r(p,10\alpha)
=
D_{m+2\to m}(p,1,0)
\oplus
\tau_{m+2\to m}(D_{m+r\to m+2}(p,1,0,\alpha)).
\]
Since
\[
N(p,1,0)=N(p,1),
\]
the first term is exactly $\kappa_{m+1\to m}(p,1)$.  The induction hypothesis
at level $m+2$ yields
\[
D_{m+r\to m+2}(p,1,0,\alpha)
=
\kappa_{m+3\to m+2}(p,1,0,\beta(\alpha)),
\]
and Lemma~\ref{lem:zero-barrier}, applied with the $(m+1)$-digit prefix
$(p,1)$, shows that its projection to level $m$ is zero.  Hence
\[
E_r(p,10\alpha)=\kappa_{m+1\to m}(p,1)
=
\kappa_{m+1\to m}(p,\beta(10\alpha)).
\]

If $q=11\alpha$, then Proposition~\ref{prop:cocycle} with the intermediate
scale $m+2$ gives
\[
E_r(p,11\alpha)
=
D_{m+2\to m}(p,1,1)
\oplus
\tau_{m+2\to m}(D_{m+r\to m+2}(p,1,1,\alpha)).
\]
Now
\[
N(p,1,1)=N(p)+F_{m+2}+F_{m+3}=N(p)+F_{m+4}.
\]
Since $N(p)<F_{m+3}$, the Zeckendorf expansion of $N(p,1,1)$ is obtained by
placing an isolated digit $1$ in position $m+3$ above $Z(N(p))$.  Therefore
\[
D_{m+2\to m}(p,1,1)=0.
\]
The induction hypothesis at level $m+2$ gives
\[
D_{m+r\to m+2}(p,1,1,\alpha)
=
\kappa_{m+3\to m+2}(p,1,1,\beta(\alpha)).
\]
Lemma~\ref{lem:double-one-reduction} then yields
\[
E_r(p,11\alpha)
=
\kappa_{m+1\to m}(p,\beta(\alpha))
=
\kappa_{m+1\to m}(p,\beta(11\alpha)).
\]
This completes the induction.

Set
\[
B_\ell:=\#\{q\in\Omega_\ell:\ \beta(q)=1\}.
\]
The defining recursion for $\beta$ gives
\[
B_0=0,
\qquad
B_1=1,
\qquad
B_\ell=2^{\ell-2}+B_{\ell-2}\quad (\ell\ge 2),
\]
because
\[
\Omega_\ell=(0\Omega_{\ell-1})\sqcup(10\Omega_{\ell-2})\sqcup(11\Omega_{\ell-2})
\]
is a disjoint union.  The closed form
\[
B_\ell=\frac{2^\ell-(-1)^\ell}{3}
\]
follows by induction, and therefore
\[
\rho_\ell=\frac{B_\ell}{2^\ell}
=
\frac{2^\ell-(-1)^\ell}{3\cdot 2^\ell}.
\]

If $v\ne 0$, the established identity shows that
$D_{n\to m}(p,q)=v$ holds if and only if $\beta(q)=1$ and
$\kappa_{m+1\to m}(p,1)=v$.  Hence
\[
\mu_{n,m}(v)=\rho_\ell\lambda_m(v)
\qquad (v\ne 0).
\]
The formula at $v=0$ follows from the decomposition into the cases
$\beta(q)=0$ and $\beta(q)=1$.

Finally,
\[
\PP_{\Unif_n}[D_{n\to m}\ne 0]
=
\rho_\ell\sum_{v\ne 0}\lambda_m(v).
\]
By Proposition~\ref{thm:explicit-count},
\[
\sum_{v\ne 0}\lambda_m(v)
=
2^{-m}\#\{p\in\Omega_m:\ \kappa_{m+1\to m}(p,1)\ne 0\}
=
\frac{\lfloor 2^{m+1}/3\rfloor}{2^m}.
\]
Substituting the formula for $\rho_\ell$ proves the last identity.
\end{proof}

\begin{corollary}[Explicit defect-depth law]
\label{cor:explicit-depth-law}
For $n>m$ and $1\le r\le m$,
\[
\mu_{n,m}(v_{m,r})
=
\frac{(2^{n-m}-(-1)^{n-m})(2^{m-r+1}-(-1)^{m-r+1})}{9\cdot 2^n}.
\]
If $R_{n,m}\in\{0,1,\dots,m\}$ is defined by $R_{n,m}=0$ on $\{D_{n\to m}=0\}$
and by $D_{n\to m}=v_{m,R_{n,m}}$ on $\{D_{n\to m}\ne 0\}$, then
\[
\PP_{\Unif_n}[R_{n,m}=r\mid D_{n\to m}\ne 0]
=
\frac{2^{m-r+1}-(-1)^{m-r+1}}{3\lfloor 2^{m+1}/3\rfloor}
\qquad (1\le r\le m).
\]
For each fixed $r\ge 1$,
\[
\PP_{\Unif_n}[R_{n,m}=r\mid D_{n\to m}\ne 0]
\xrightarrow[m\to\infty]{}
2^{-r},
\]
uniformly in the tail length $n-m\ge 1$.
\end{corollary}

\begin{proof}
Theorem~\ref{thm:tail-collapse} gives
\[
\mu_{n,m}(v)=\rho_{n-m}\lambda_m(v)
\qquad (v\ne 0).
\]
Combining this with Theorem~\ref{thm:adjacent-spectrum} yields the explicit
formula for $\mu_{n,m}(v_{m,r})$.  Dividing by
\[
\PP_{\Unif_n}[D_{n\to m}\ne 0]
=
\rho_{n-m}\frac{\lfloor 2^{m+1}/3\rfloor}{2^m}
\]
gives the conditional law of $R_{n,m}$.  For fixed $r$,
\[
\frac{2^{m-r+1}-(-1)^{m-r+1}}{3\lfloor 2^{m+1}/3\rfloor}
\longrightarrow
2^{-r},
\]
and the expression is independent of $n$, so the convergence is uniform in the
tail length.
\end{proof}

\begin{remark}
\label{rem:limit-law}
For fixed $m$, the preceding theorem gives the exact asymptotic law along
increasing tail length:
\[
\mu_{m+\ell,m}
=
\left(\frac23+\frac{(-1)^\ell}{3\cdot 2^\ell}\right)\delta_0
+
\left(\frac13-\frac{(-1)^\ell}{3\cdot 2^\ell}\right)\lambda_m.
\]
Hence
\[
\mu_{m+\ell,m}\xrightarrow[\ell\to\infty]{}\nu_m
:=
\frac23\delta_0+\frac13\lambda_m.
\]
Moreover,
\[
d_{\mathrm{TV}}(\mu_{m+\ell,m},\nu_m)
=
\frac{1-\lambda_m(0)}{3\cdot 2^\ell}
=
\frac{\lfloor 2^{m+1}/3\rfloor}{3\cdot 2^{m+\ell}},
\]
because
\[
1-\lambda_m(0)
=
2^{-m}\#\{p\in\Omega_m:\ \kappa_{m+1\to m}(p,1)\ne 0\}
=
\frac{\lfloor 2^{m+1}/3\rfloor}{2^m}
\]
by Proposition~\ref{thm:explicit-count}.
\end{remark}

\subsection{Small-scale values}

\begin{remark}
\label{rem:small-values}
The exact probabilities $P_{n,m}:=\PP_{\Unif_n}[D_{n\to m}\ne 0]$ obtained by
direct enumeration for $n\le 5$ are listed in Table~\ref{tab:small-values}.
In particular,
\[
P_{2,1}=\frac14,
\qquad
P_{3,1}=\frac18,
\qquad
P_{4,1}=\frac{3}{16},
\]
so the scalar probabilities are not monotone in $n$ at fixed $m$.  Combined
with Example~\ref{ex:defect-without-top-indicator}, this shows that the full
defect law cannot be recovered from the scalar event $\{D_{n\to m}\ne 0\}$
alone.
\end{remark}

\begin{table}[t]
\centering
\small
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{ccc}
\toprule
$(n,m)$ & $\#\{\omega:\ D_{n\to m}\ne 0\}$ & $P_{n,m}$ \\
\midrule
$(2,1)$ & $1$ & $1/4$ \\
$(3,1)$ & $1$ & $1/8$ \\
$(3,2)$ & $2$ & $1/4$ \\
$(4,1)$ & $3$ & $3/16$ \\
$(4,2)$ & $2$ & $1/8$ \\
$(4,3)$ & $5$ & $5/16$ \\
$(5,1)$ & $5$ & $5/32$ \\
$(5,2)$ & $6$ & $3/16$ \\
$(5,3)$ & $5$ & $5/32$ \\
$(5,4)$ & $10$ & $5/16$ \\
\bottomrule
\end{tabular}
\caption{Exact small-scale defect probabilities under the uniform raw-word law.}
\label{tab:small-values}
\end{table}

\section{Tail parity and Markov inputs}
\label{sec:markov}

We now identify the tail bit explicitly and deduce that it is the coarsest
exact tail statistic for the family of defects.  The stationary one-step
Markov law then follows from the same deterministic reduction together with a
direct computation of the parity of the initial run of $1$'s.

\subsection{Parity of the initial run}

\begin{theorem}[Identification and minimality of the tail bit]
\label{thm:tail-rigidity}
For a finite word $q=q_1\cdots q_r$, let
\[
\iota(q):=\max\{s\in\{0,\dots,r\}:\ q_1=\cdots=q_s=1\}
\]
be the length of its initial run of $1$'s, with $\iota(\varnothing):=0$.
Then
\[
\beta(q)\equiv \iota(q)\pmod 2
\qquad
\text{for every finite tail word }q.
\]
Define an equivalence relation on the disjoint union
$\bigsqcup_{r\ge 0}\Omega_r$ by
\[
q\sim q'
\quad\Longleftrightarrow\quad
D_{m+|q|\to m}(p,q)=D_{m+|q'|\to m}(p,q')
\text{ for every }m\ge 1\text{ and }p\in\Omega_m.
\]
Then
\[
q\sim q'
\quad\Longleftrightarrow\quad
\beta(q)=\beta(q').
\]
Consequently, if $S$ is any set, if
$\gamma:\bigsqcup_{r\ge 0}\Omega_r\to S$, and if for each $m\ge 1$ there is a
map $G_m:\Omega_m\times S\to\{0,1\}^m$ such that
\[
D_{m+r\to m}(p,q)=G_m(p,\gamma(q))
\qquad
(r\ge 0,\ p\in\Omega_m,\ q\in\Omega_r),
\]
then there exists a unique map
\[
\psi:\gamma\!\left(\bigsqcup_{r\ge 0}\Omega_r\right)\to\{0,1\}
\]
satisfying
\[
\beta=\psi\circ\gamma.
\]
In particular, every exact tail factor for the full defect has at least two
states, and the factor furnished by Theorem~\ref{thm:tail-collapse} is minimal
up to relabeling of $\{0,1\}$.
\end{theorem}

\begin{proof}
If $q=\varnothing$, then $\beta(q)=0$ and $\iota(q)=0$.  Assume that $q$ is
nonempty.

If $\iota(q)=0$, then $q=0\alpha$ for some $\alpha$, so $\beta(q)=0$ by
definition.  If $\iota(q)=2t+1$ is odd, then either $q=1^{2t+1}$ or
$q=1^{2t+1}0\alpha$.  Repeated use of the rule $\beta(11\alpha)=\beta(\alpha)$
reduces both words to the case $\beta(1)=1$, so $\beta(q)=1$.  If
$\iota(q)=2t$ is positive and even, then either $q=1^{2t}$ or
$q=1^{2t}0\alpha$.  The same reduction gives
$\beta(q)=\beta(\varnothing)=0$.  Hence $\beta(q)\equiv\iota(q)\pmod 2$.

If $\beta(q)=\beta(q')$, Theorem~\ref{thm:tail-collapse} yields
\[
D_{m+|q|\to m}(p,q)
=
\kappa_{m+1\to m}(p,\beta(q))
=
\kappa_{m+1\to m}(p,\beta(q'))
=
D_{m+|q'|\to m}(p,q')
\]
for every $m$ and $p$, so $q\sim q'$.

Conversely, assume $\beta(q)\ne\beta(q')$.  Set $m:=1$ and $p:=1$.  Then
Theorem~\ref{thm:tail-collapse} gives
\[
D_{1+|q|\to 1}(1,q)=\kappa_{2\to 1}(1,\beta(q)),
\qquad
D_{1+|q'|\to 1}(1,q')=\kappa_{2\to 1}(1,\beta(q')).
\]
Also,
\[
\kappa_{2\to 1}(1,0)=0,
\qquad
\kappa_{2\to 1}(1,1)=1,
\]
because $\Fold_1(1)=1$ and $\Fold_2(11)=00$.  Therefore the two defects are
different, so $q\nsim q'$.

For the factorization claim, define
\[
\psi:\gamma\!\left(\bigsqcup_{r\ge 0}\Omega_r\right)\to\{0,1\}
\]
by choosing $q$ with $\gamma(q)=s$ and setting $\psi(s):=\beta(q)$.  This is
well defined because
$\gamma(q)=\gamma(q')$ implies
\[
D_{m+|q|\to m}(p,q)=G_m(p,\gamma(q))=G_m(p,\gamma(q'))=D_{m+|q'|\to m}(p,q')
\]
for every $m$ and $p$, hence $q\sim q'$ and therefore $\beta(q)=\beta(q')$.
The identity $\beta=\psi\circ\gamma$ is immediate from the definition of
$\psi$, and uniqueness holds on the image of $\gamma$.
\end{proof}

\subsection{Stationary Markov inputs}

Theorem~\ref{thm:tail-rigidity} converts the deterministic collapse into an
exact non-uniform law for one-step Markov inputs.  The only tail information
that survives is the parity of the initial run of $1$'s, and that parity has
a closed form under Markov sampling.

\begin{theorem}[Stationary one-step Markov extension]
\label{thm:markov-defect-law}
Let $\PP_P$ be the stationary irreducible one-step Markov measure on
$\{0,1\}^{\NN}$ with transition matrix
\[
P=(p_{ab})_{a,b\in\{0,1\}}
\]
and stationary distribution $\pi=(\pi_0,\pi_1)$.  Let
$\omega^\infty=(\omega_1,\omega_2,\dots)$ denote the coordinate process, and
for $m\ge 1$ write
\[
r_{\infty\to m}(\omega^\infty):=(\omega_1,\dots,\omega_m)\in\Omega_m.
\]
Fix $m\ge 1$ and $\ell\ge 0$, and set $n:=m+\ell$.  For $a\in\{0,1\}$ define
probability measures on $\{0,1\}^m$ by
\[
\lambda_{m,a}(v)
:=
\PP_P\!\left[
\kappa_{m+1\to m}(r_{\infty\to m}(\omega^\infty),1)=v
\ \middle|\
\omega_m=a
\right],
\]
and define
\[
\rho_\ell^{(a)}
:=
\PP_P\!\left[
\beta(\omega_{m+1}\cdots\omega_{m+\ell})=1
\ \middle|\
\omega_m=a
\right].
\]
Then
\[
\Law_{\PP_P}\!\bigl(D_{n\to m}(r_{\infty\to n}(\omega^\infty))\bigr)
=
\sum_{a=0}^1
\pi_a\bigl((1-\rho_\ell^{(a)})\delta_0+\rho_\ell^{(a)}\lambda_{m,a}\bigr),
\]
and
\[
\rho_0^{(a)}=0,
\qquad
\rho_\ell^{(a)}
=
\frac{p_{a1}(1-(-p_{11})^\ell)}{1+p_{11}}
\qquad (\ell\ge 1).
\]
Consequently
\[
\PP_P\!\bigl[D_{n\to m}(r_{\infty\to n}(\omega^\infty))\ne 0\bigr]
=
\sum_{a=0}^1
\pi_a\,\rho_\ell^{(a)}\bigl(1-\lambda_{m,a}(0)\bigr),
\]
the defect law converges as $\ell\to\infty$ to
\[
\nu_{m,P}
:=
\sum_{a=0}^1
\pi_a\left(
\left(1-\frac{p_{a1}}{1+p_{11}}\right)\delta_0
\;+\;
\frac{p_{a1}}{1+p_{11}}\lambda_{m,a}
\right),
\]
and
\[
d_{\mathrm{TV}}
\!\left(
\Law_{\PP_P}\!\bigl(D_{m+\ell\to m}(r_{\infty\to m+\ell}(\omega^\infty))\bigr),
\nu_{m,P}
\right)
\le
\frac{\pi_1}{1+p_{11}}\,p_{11}^{\ell}.
\]
If $p_{01}=p_{11}=u$, then
\[
\rho_\ell^{(0)}=\rho_\ell^{(1)}
=
\frac{u(1-(-u)^\ell)}{1+u},
\]
and the uniform law is recovered at $u=\frac12$.
\end{theorem}

\begin{proof}
Set
\[
P_{m,\ell}
:=
\Law_{\PP_P}\!\bigl(D_{m+\ell\to m}(r_{\infty\to m+\ell}(\omega^\infty))\bigr).
\]
Because the chain is irreducible on two states, $\pi_a>0$ for $a=0,1$, so the
conditional laws $\lambda_{m,a}$ are well defined.
Fix $p\in\Omega_m$ and write $a:=p_m$.  By the Markov property, conditioned on
$r_{\infty\to m}(\omega^\infty)=p$, the future block
$q:=(\omega_{m+1},\dots,\omega_{m+\ell})$ is independent of the past and is
distributed as a length-$\ell$ Markov chain started from state $a$.
Theorem~\ref{thm:tail-collapse} therefore gives
\[
\Law_{\PP_P}\!\bigl(D_{m+\ell\to m}(p,q)\,\big|\,r_{\infty\to m}(\omega^\infty)=p\bigr)
=
(1-\rho_\ell^{(a)})\delta_0+\rho_\ell^{(a)}\delta_{\kappa_{m+1\to m}(p,1)}.
\]
Conditioning further on $\omega_m=a$ and averaging over prefixes with last
digit $a$ yields
\[
\Law_{\PP_P}\!\bigl(
D_{m+\ell\to m}(r_{\infty\to m+\ell}(\omega^\infty))
\ \big|\ \omega_m=a
\bigr)
=
(1-\rho_\ell^{(a)})\delta_0+\rho_\ell^{(a)}\lambda_{m,a}.
\]
Since $\PP_P[\omega_m=a]=\pi_a$ by stationarity, averaging over $a$ proves the
first displayed formula for $P_{m,\ell}$.  The formula for the nonzero defect
probability is immediate.

Theorem~\ref{thm:tail-rigidity} identifies $\beta(q)$ with the parity of the
initial run of $1$'s.  Under the Markov law started from state $a$, the event
$\{\beta(q)=1\}$ is therefore the disjoint union of the cylinder events
corresponding to odd initial run lengths.  For $\ell=0$, this probability is
zero.  For $\ell\ge 1$, write $\ell=2r$ or $\ell=2r+1$.

If $\ell=2r$, then an odd initial run has length $2j+1$ for some
$0\le j\le r-1$, so
\[
\rho_{2r}^{(a)}
=
\sum_{j=0}^{r-1} p_{a1}p_{11}^{2j}(1-p_{11})
=
\frac{p_{a1}(1-p_{11}^{2r})}{1+p_{11}}
=
\frac{p_{a1}(1-(-p_{11})^{2r})}{1+p_{11}}.
\]
If $\ell=2r+1$, then either the odd initial run has length $2j+1$ with
$0\le j\le r-1$ and is followed by $0$, or the entire block is $1^{2r+1}$.
Hence
\[
\rho_{2r+1}^{(a)}
=
\sum_{j=0}^{r-1} p_{a1}p_{11}^{2j}(1-p_{11})
+ p_{a1}p_{11}^{2r}
=
\frac{p_{a1}(1+p_{11}^{2r+1})}{1+p_{11}}
=
\frac{p_{a1}(1-(-p_{11})^{2r+1})}{1+p_{11}}.
\]
This proves the closed form for $\rho_\ell^{(a)}$.

Let
\[
\rho_\infty^{(a)}:=\frac{p_{a1}}{1+p_{11}}.
\]
Since the chain is irreducible, one has $p_{11}<1$.  Therefore
\[
\nu_{m,P}
=
\sum_{a=0}^1
\pi_a\bigl((1-\rho_\infty^{(a)})\delta_0+\rho_\infty^{(a)}\lambda_{m,a}\bigr),
\]
so
\[
P_{m,\ell}-\nu_{m,P}
=
\sum_{a=0}^1
\pi_a\bigl(\rho_\ell^{(a)}-\rho_\infty^{(a)}\bigr)(\lambda_{m,a}-\delta_0).
\]
Since $d_{\mathrm{TV}}(\lambda_{m,a},\delta_0)\le 1$,
\[
d_{\mathrm{TV}}(P_{m,\ell},\nu_{m,P})
\le
\sum_{a=0}^1 \pi_a\bigl|\rho_\ell^{(a)}-\rho_\infty^{(a)}\bigr|
=
\frac{p_{11}^{\ell}}{1+p_{11}}
\sum_{a=0}^1 \pi_a p_{a1}.
\]
Stationarity gives $\pi_0p_{01}+\pi_1p_{11}=\pi_1$, hence
\[
d_{\mathrm{TV}}(P_{m,\ell},\nu_{m,P})
\le
\frac{\pi_1}{1+p_{11}}\,p_{11}^{\ell}.
\]
If $p_{01}=p_{11}=u$, then the chain is i.i.d.\ Bernoulli$(u)$, so
$\rho_\ell^{(0)}=\rho_\ell^{(1)}$.  At $u=\frac12$ one has
$\pi_0=\pi_1=\frac12$, and the stationary average of the conditional adjacent
laws recovers the uniform adjacent law:
\[
\frac12\lambda_{m,0}+\frac12\lambda_{m,1}=\lambda_m.
\]
Hence the first displayed formula for the defect law reduces to
\[
(1-\rho_\ell)\delta_0+\rho_\ell\lambda_m,
\]
which is exactly Theorem~\ref{thm:tail-collapse}.
\end{proof}

\subsection{Exact Markov spectrum and conditioned depth}

The stationary mixture of Theorem~\ref{thm:markov-defect-law} leaves the
adjacent laws $\lambda_{m,a}$ implicit.  The next theorem determines them
exactly and shows that, after conditioning on a nonzero defect, the depth law
is independent of the tail length.

\begin{theorem}[Exact Markov spectrum and conditioned defect depth]\textbf{[New-4 | 2026-04-18]}
\label{thm:markov-spectrum}
Keep the notation of Theorem~\ref{thm:markov-defect-law}.  Define
\[
A_0:=\pi_1,
\qquad
A_{j+1}:=\pi_1p_{10}p_{00}+p_{01}p_{10}A_j
\qquad (j\ge 0).
\]
Then
\[
A_j
=
\frac{\pi_1p_{10}p_{00}}{1-p_{01}p_{10}}
+
\frac{\pi_1p_{11}}{1-p_{01}p_{10}}(p_{01}p_{10})^j
\qquad (j\ge 0).
\]
For $m\ge 1$, $a\in\{0,1\}$, and $1\le r\le m$, set
\[
\Lambda_{m,a}(r):=\lambda_{m,a}(v_{m,r}),
\]
where $\lambda_{m,a}$ is the conditional adjacent law from
Theorem~\ref{thm:markov-defect-law}.  Then
\[
\Lambda_{m,0}(r)=0
\qquad (2\le r\le m),
\]
\[
\Lambda_{m,0}(1)
=
\frac{p_{10}}{\pi_0}
\bigl(
\pi_1-A_{\lfloor (m-1)/2\rfloor}
\bigr),
\]
\[
\Lambda_{m,1}(1)
=
\frac{A_{\lfloor m/2\rfloor}}{\pi_1},
\]
and, for $2\le r\le m$,
\[
\Lambda_{m,1}(r)
=
\frac{p_{11}^{r-2}}{\pi_1}
\Bigl(
p_{01}p_{10}\bigl(\pi_1-A_{\lfloor (m-r)/2\rfloor}\bigr)
+
p_{11}A_{\lfloor (m-r+1)/2\rfloor}
\Bigr).
\]
In particular,
\[
\sum_{r=1}^m\Lambda_{m,1}(r)=1.
\]

Let $R_{n,m}$ be the defect-depth variable from
Corollary~\ref{cor:explicit-depth-law}.  If $\ell\ge 1$ and $n:=m+\ell$, then
\[
\PP_P\!\left[
R_{n,m}=1
\ \middle|\
D_{n\to m}(r_{\infty\to n}(\omega^\infty))\ne 0
\right]
=
\frac{p_{10}\Lambda_{m,0}(1)+p_{11}\Lambda_{m,1}(1)}
{p_{10}\Lambda_{m,0}(1)+p_{11}},
\]
and, for $2\le r\le m$,
\[
\PP_P\!\left[
R_{n,m}=r
\ \middle|\
D_{n\to m}(r_{\infty\to n}(\omega^\infty))\ne 0
\right]
=
\frac{p_{11}\Lambda_{m,1}(r)}
{p_{10}\Lambda_{m,0}(1)+p_{11}}.
\]
Hence the conditioned depth law is independent of the tail length $\ell$.
Moreover, for each fixed $r\ge 1$,
\[
\Lambda_{m,0}(1)\xrightarrow[m\to\infty]{}
\frac{p_{01}p_{11}}{1-p_{01}p_{10}},
\]
\[
\Lambda_{m,1}(1)\xrightarrow[m\to\infty]{}
\frac{p_{10}p_{00}}{1-p_{01}p_{10}},
\qquad
\Lambda_{m,1}(r)\xrightarrow[m\to\infty]{}
\frac{p_{10}p_{11}^{r-1}}{1-p_{01}p_{10}}
\quad (r\ge 2),
\]
and therefore
\[
\PP_P\!\left[
R_{m+\ell,m}=r
\ \middle|\
D_{m+\ell\to m}(r_{\infty\to m+\ell}(\omega^\infty))\ne 0
\right]
\xrightarrow[m\to\infty]{}
p_{10}p_{11}^{r-1}
\]
uniformly in $\ell\ge 1$.
\end{theorem}

\begin{proof}
For $m\ge 1$, let
\[
\mathcal{T}_m
:=
\{\alpha\in\Omega_m:\ N(\alpha)<F_{m+1}\},
\qquad
T_m:=\PP_P[r_{\infty\to m}(\omega^\infty)\in\mathcal{T}_m].
\]
The proof of Proposition~\ref{thm:explicit-count} is purely combinatorial and
therefore yields, for $m\ge 3$,
\[
\mathcal{T}_m=\Omega_{m-2}00\sqcup \mathcal{T}_{m-2}10.
\]
Since $\PP_P[\omega_{m-1}=0,\omega_m=0]=\pi_0p_{00}$ by stationarity and every
word in $\mathcal{T}_{m-2}$ ends with $0$, this gives
\[
T_m=\pi_0p_{00}+p_{01}p_{10}T_{m-2}
\qquad (m\ge 3).
\]
Also,
\[
T_1=T_2=\pi_0,
\]
because $\mathcal{T}_1=\{0\}$ and $\mathcal{T}_2=\{00,10\}$ are exactly the
length-$1$ and length-$2$ words whose top digit is $0$.  It follows by
induction that
\[
T_{2j-1}=T_{2j}
\qquad (j\ge 1).
\]
Write this common value as $\Theta_j$.  Then
\[
\Theta_1=\pi_0,
\qquad
\Theta_{j+1}=\pi_0p_{00}+p_{01}p_{10}\Theta_j
\qquad (j\ge 1).
\]
For $j\ge 1$ define
\[
A_j:=p_{01}\Theta_j.
\]
Using $\pi_0p_{01}=\pi_1p_{10}$, one obtains
\[
A_{j+1}=\pi_1p_{10}p_{00}+p_{01}p_{10}A_j
\qquad (j\ge 1),
\]
and the same identity holds at $j=0$ because
\[
A_1=p_{01}\Theta_1=p_{01}\pi_0=\pi_1p_{10}
=\pi_1p_{10}p_{00}+p_{01}p_{10}A_0.
\]
Hence the displayed affine recursion for $(A_j)_{j\ge 0}$ is valid.  Solving
it gives the closed form
\[
A_j
=
\frac{\pi_1p_{10}p_{00}}{1-p_{01}p_{10}}
+
\left(
\pi_1-\frac{\pi_1p_{10}p_{00}}{1-p_{01}p_{10}}
\right)
(p_{01}p_{10})^j,
\]
which simplifies to the stated formula because
\[
\pi_1-\frac{\pi_1p_{10}p_{00}}{1-p_{01}p_{10}}
=
\frac{\pi_1p_{11}}{1-p_{01}p_{10}}.
\]

For $m\ge 1$ and $1\le r\le m$, let
\[
\mathcal{S}_{m,r}
:=
\{p\in\Omega_m:\ \kappa_{m+1\to m}(p,1)=v_{m,r}\}.
\]
The interval description in Theorem~\ref{thm:adjacent-spectrum} implies
\[
\mathcal{S}_{1,1}=\{1\},
\]
and, for $m\ge 2$,
\[
\mathcal{S}_{m,1}
=
\mathcal{T}_{m-1}1
\sqcup
\mathcal{T}_{m-2}^{\,c}10.
\]
Indeed, the first component corresponds to the case $p_m=1$ and
$N(p_1,\dots,p_{m-1})<F_m$, while the second corresponds to the case $p_m=0$,
$p_{m-1}=1$, and $N(p_1,\dots,p_{m-2})\ge F_{m-1}$.

The two components of $\mathcal{S}_{m,1}$ have distinct top digits.  For the
words with top digit $1$,
\[
\PP_P[r_{\infty\to m}(\omega^\infty)\in \mathcal{T}_{m-1}1]
=
\begin{cases}
\pi_1,& m=1,\\[4pt]
p_{01}T_{m-1},& m\ge 2,
\end{cases}
\]
hence
\[
\PP_P[r_{\infty\to m}(\omega^\infty)\in \mathcal{T}_{m-1}1]
=
A_{\lfloor m/2\rfloor}
\qquad (m\ge 1).
\]
For the words with top digit $0$, note that the total probability of a length
$m-1$ prefix whose top digit is $1$ equals $\pi_1$.  Among those prefixes, the
subset $\mathcal{T}_{m-2}1$ has probability
$A_{\lfloor (m-1)/2\rfloor}$.  Therefore
\[
\PP_P[r_{\infty\to m}(\omega^\infty)\in \mathcal{T}_{m-2}^{\,c}10]
=
p_{10}\bigl(\pi_1-A_{\lfloor (m-1)/2\rfloor}\bigr).
\]
Dividing by $\pi_0$ or $\pi_1$ yields
\[
\Lambda_{m,0}(1)
=
\frac{p_{10}}{\pi_0}
\bigl(
\pi_1-A_{\lfloor (m-1)/2\rfloor}
\bigr),
\qquad
\Lambda_{m,1}(1)
=
\frac{A_{\lfloor m/2\rfloor}}{\pi_1}.
\]
The equality $\Lambda_{m,0}(r)=0$ for $r\ge 2$ follows from
Theorem~\ref{thm:adjacent-spectrum}, because every interval $I_{m,r}$ with
$r\ge 2$ lies in $[F_{m+2},F_{m+3}-2]$ and therefore requires the top digit
$p_m$ to be $1$.

For $2\le r\le m$, repeated use of
\[
I_{j,s}=F_{j+1}+I_{j-1,s-1}
\qquad (2\le s\le j)
\]
from Theorem~\ref{thm:adjacent-spectrum} gives
\[
\mathcal{S}_{m,r}=\mathcal{S}_{m-r+1,1}1^{r-1}.
\]
Set $k:=m-r+1$.  The decomposition of $\mathcal{S}_{k,1}$ above therefore
gives
\[
\mathcal{S}_{m,r}
=
\mathcal{T}_{k-1}1^r
\sqcup
\mathcal{T}_{k-2}^{\,c}10\,1^{r-1}.
\]
The first cylinder has probability
\[
p_{11}^{r-1}A_{\lfloor k/2\rfloor},
\]
and the second has probability
\[
p_{01}p_{10}p_{11}^{r-2}
\bigl(
\pi_1-A_{\lfloor (k-1)/2\rfloor}
\bigr).
\]
Every word in $\mathcal{S}_{m,r}$ ends with the digit $1$, so dividing by
$\pi_1$ yields
\[
\Lambda_{m,1}(r)
=
\frac{p_{11}^{r-2}}{\pi_1}
\Bigl(
p_{01}p_{10}\bigl(\pi_1-A_{\lfloor (m-r)/2\rfloor}\bigr)
+
p_{11}A_{\lfloor (m-r+1)/2\rfloor}
\Bigr),
\]
which is the stated formula.  If $\omega_m=1$, then
$N(r_{\infty\to m}(\omega^\infty))\ge F_{m+1}$, so
Theorem~\ref{thm:adjacent-threshold} implies
$\kappa_{m+1\to m}(r_{\infty\to m}(\omega^\infty),1)\ne 0$.  Hence
\[
\sum_{r=1}^m\Lambda_{m,1}(r)=1.
\]

Fix $\ell\ge 1$ and set $n:=m+\ell$.  Theorem~\ref{thm:markov-defect-law}
gives
\[
\PP_P[D_{n\to m}=v_{m,r}]
=
\pi_0\rho_\ell^{(0)}\Lambda_{m,0}(r)
+
\pi_1\rho_\ell^{(1)}\Lambda_{m,1}(r).
\]
For $\ell\ge 1$,
\[
\rho_\ell^{(0)}
=
\frac{p_{01}(1-(-p_{11})^\ell)}{1+p_{11}},
\qquad
\rho_\ell^{(1)}
=
\frac{p_{11}(1-(-p_{11})^\ell)}{1+p_{11}}.
\]
Using $\pi_0p_{01}=\pi_1p_{10}$ and $\Lambda_{m,0}(r)=0$ for $r\ge 2$, one
obtains
\[
\PP_P[D_{n\to m}=v_{m,1}]
=
\frac{\pi_1(1-(-p_{11})^\ell)}{1+p_{11}}
\bigl(
p_{10}\Lambda_{m,0}(1)+p_{11}\Lambda_{m,1}(1)
\bigr),
\]
\[
\PP_P[D_{n\to m}=v_{m,r}]
=
\frac{\pi_1(1-(-p_{11})^\ell)}{1+p_{11}}
p_{11}\Lambda_{m,1}(r)
\qquad (2\le r\le m).
\]
Summing over $r$ and using $\sum_{r=1}^m\Lambda_{m,1}(r)=1$ gives
\[
\PP_P[D_{n\to m}\ne 0]
=
\frac{\pi_1(1-(-p_{11})^\ell)}{1+p_{11}}
\bigl(
p_{10}\Lambda_{m,0}(1)+p_{11}
\bigr).
\]
Dividing the preceding identities by this common factor proves the displayed
formulas for the conditioned depth law and shows that the result is independent
of $\ell$.

Since $p_{01}p_{10}<1$, the closed form for $A_j$ yields
\[
A_j\xrightarrow[j\to\infty]{}
A_\infty
:=
\frac{\pi_1p_{10}p_{00}}{1-p_{01}p_{10}}.
\]
Hence
\[
\Lambda_{m,1}(1)
=
\frac{A_{\lfloor m/2\rfloor}}{\pi_1}
\xrightarrow[m\to\infty]{}
\frac{p_{10}p_{00}}{1-p_{01}p_{10}}.
\]
Also,
\[
\pi_1-A_\infty
=
\frac{\pi_1p_{11}}{1-p_{01}p_{10}},
\]
so
\[
\Lambda_{m,0}(1)
=
\frac{p_{10}}{\pi_0}
\bigl(
\pi_1-A_{\lfloor (m-1)/2\rfloor}
\bigr)
\xrightarrow[m\to\infty]{}
\frac{p_{10}\pi_1}{\pi_0}\,
\frac{p_{11}}{1-p_{01}p_{10}}
=
\frac{p_{01}p_{11}}{1-p_{01}p_{10}}.
\]
For fixed $r\ge 2$, both indices in the formula for $\Lambda_{m,1}(r)$ tend to
$+\infty$ with $m$, hence
\[
\Lambda_{m,1}(r)
\xrightarrow[m\to\infty]{}
\frac{p_{11}^{r-2}}{\pi_1}
\bigl(
p_{01}p_{10}(\pi_1-A_\infty)+p_{11}A_\infty
\bigr).
\]
Using the two identities above,
\[
p_{01}p_{10}(\pi_1-A_\infty)+p_{11}A_\infty
=
\frac{\pi_1p_{10}p_{11}}{1-p_{01}p_{10}},
\]
and therefore
\[
\Lambda_{m,1}(r)
\xrightarrow[m\to\infty]{}
\frac{p_{10}p_{11}^{r-1}}{1-p_{01}p_{10}}.
\]

Finally,
\[
p_{10}\Lambda_{m,0}(1)+p_{11}
\xrightarrow[m\to\infty]{}
\frac{p_{10}p_{01}p_{11}}{1-p_{01}p_{10}}+p_{11}
=
\frac{p_{11}}{1-p_{01}p_{10}}.
\]
The limit for $r=1$ is therefore
\[
\frac{
\dfrac{p_{10}p_{01}p_{11}}{1-p_{01}p_{10}}
+
\dfrac{p_{11}p_{10}p_{00}}{1-p_{01}p_{10}}
}{
\dfrac{p_{11}}{1-p_{01}p_{10}}
}
=
p_{10},
\]
while for $r\ge 2$ the limit is
\[
\frac{
\dfrac{p_{11}p_{10}p_{11}^{r-1}}{1-p_{01}p_{10}}
}{
\dfrac{p_{11}}{1-p_{01}p_{10}}
}
=
p_{10}p_{11}^{r-1}.
\]
Because the conditioned depth law is exactly independent of $\ell$, the
convergence is uniform in $\ell\ge 1$.
\end{proof}

\bibliographystyle{plain}
\bibliography{references}

\end{document}
