% ------------------------------------------------------------------
%  Finite observation of open symbolic dynamics
% ------------------------------------------------------------------
% ETDS/Cambridge-compatible preprint source: single-column article
% layout, AMS theorem environments, and natbib bibliography.
\documentclass[10pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{array}
\usepackage{enumitem}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{microtype}
\geometry{margin=2.5cm}
\microtypesetup{expansion=false}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem*{lemma*}{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\DeclareMathOperator{\card}{card}
\DeclareMathOperator{\diam}{diam}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\Col}{Col}
\DeclareMathOperator{\spec}{spec}

\newcommand{\NN}{\mathbb{N}}
\newcommand{\NNzero}{\mathbb{N}_0}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\norm}[1]{\lVert #1\rVert}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\ind}{\mathbf{1}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\Leb}{\mathrm{Leb}}

\title{Finite-State Observation of First-Entry Escape Rates\\
and Quasistationary Ambiguity for Symbolic Holes}
\author{Haobo Ma\\
\small CHRONOAI PTE.\ LTD., Singapore\\
\small\texttt{haobo@chronoai.co}}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Let \(Y\) be a topologically mixing one-step subshift of finite type,
\(\nu_\varphi\) the equilibrium Markov measure of a one-step potential,
\(H\) a Markov hole.  This paper is centered on one finite-state
theorem for first-entry residue events into \(H\): the Bayes error of a
depth-\(m\) observation is exactly a killed-matrix path sum with
residue-dependent terminal ambiguity weights.  From this identity we
obtain the degenerate-residue dichotomy and, in the two-terminal
non-degenerate case, recovery of the escape rate
\(P_Y(\varphi)-P_{Y_H}(\varphi)\) together with an explicit
residue-dependent quasistationary amplitude.  The rest of the manuscript
is organized around this theorem.  Two deliberately secondary
finite-dimensional consequences are recorded separately.  One is a
short survivor R\'enyi pressure corollary, which yields the spectrum
\(sP_{Y_H}(\varphi)-P_{Y_H}(s\varphi)\), the conditioned survivor
R\'enyi-\(2\) entropy rate \(h_{2,H}\), and iid Poisson collision
thresholds on the scale
\(\binom{N_m}{2}e^{-mh_{2,H}}\asymp 1\), provided the third-moment
condition \(N_m^3S_3^H(m)\to0\) holds; at the pure exponential scale this
is \(h_{3,H}>\frac32h_{2,H}\).  The other is a finite-dimensional
algebraic resolvent reformulation of the same killed path sum, whose possible
scalar poles lie in a reciprocal cyclotomic lift of the killed matrix
spectrum and whose visible poles are selected by a spectral-projection
non-cancellation criterion.  The H\"older material is kept separate from
this finite-state theorem: an appendix gives an explicit \(C^\alpha\)
open-transfer statement for finite-cylinder holes after block recoding
and separately proves the finite-prefix Gibbs comparison used for escape
exponents.  Neither appendix statement is used in the finite-state
resolvent proposition.
\end{abstract}

{\small
\noindent\textbf{Keywords:}
open symbolic dynamics, escape rates, killed spectra, R\'enyi
pressures, symbolic holes, Poisson approximation.

\medskip
\noindent\textbf{MSC 2020:}
37B10, 37A50, 37D35, 94A17.
}

\bigskip

\input{sec_introduction}
\input{sec_preliminaries}
\input{sec_clarity}
\input{sec_open_system}
\input{sec_open_system_resonance}
\input{sec_double_budget}
\input{sec_conclusion}

\appendix
\input{sec_holder_extension}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
