\section{Introduction}\label{sec:introduction}

\subsection*{Problem}

Let \((X,\mathcal{B},\mu,T)\) be a measure-preserving system and let
\(\mathcal W=\{W_i:i\in\Sigma\}\) be a finite measurable observation
partition.  The coding map
\(\omega_\infty(x)=(s_t(x))_{t\geq 0}\), where
\(T^t x\in W_{s_t(x)}\), defines a measurable symbolic factor over the
finite alphabet \(\Sigma\).  For a visible event \(\mathsf P\), the
depth-\(m\) Bayes error
\[
  \varepsilon_m(\mathsf P;\mu)
\]
is the smallest classification error achievable from the first \(m\)
visible symbols.  This paper studies the sequence
\(m\mapsto \varepsilon_m(\mathsf P;\mu)\) when \(\mathsf P\) is defined
by first entry into a symbolic hole.  The main question is how much of
the open system can be recovered from the finite-prefix Bayes error
rather than from direct access to survivor probabilities or transfer
operators.

\subsection*{Setting}

Fix a topologically mixing one-step subshift of finite type \(Y\), a
one-step potential \(\varphi\), and the equilibrium Markov measure
\(\nu_\varphi\).  Let \(H\) be a Markov hole with mixing survivor
subshift \(Y_H\), and write
\[
  \gamma_H(\varphi):=P_Y(\varphi)-P_{Y_H}(\varphi)
\]
for the escape rate.  For a non-empty proper subset
\(R\subsetneq \ZZ/q\ZZ\), let \(\mathsf P_R\) be the event that the
first entry time into \(H\) is finite and belongs to \(R\) modulo
\(q\).
There is one unavoidable degeneracy: from some safe state and residue,
the future may force the first-hit residue either always into \(R\) or
always outside \(R\).  Proposition~\ref{prop:degenerate-residue-dichotomy}
classifies this case.  Along each residue subsequence the Bayes error is
either eventually zero or has the survivor-mass asymptotic governed by a
quasistationary ambiguity amplitude.  The main amplitude theorem below
treats the two-terminal case, where every safe state and residue can
still hit both sides of the residue partition.

\subsection*{Main results}

The main result is the escape-rate and quasistationary-amplitude theorem.
The survivor R\'enyi and resonance statements are consequences of the
same killed path-sum formalism.  The genuinely finite-observation part is
the conversion of Bayes error into a killed path sum with terminal
ambiguity weights; after that point the asymptotics are finite
Perron--Frobenius calculations.  For clarity, the concrete contributions
are separated as follows:
\begin{enumerate}[label=(\roman*)]
  \item an exact killed-matrix formula and degenerate-residue dichotomy
    for the Bayes error of first-entry residue events of Markov holes,
    identifying the non-degenerate exponential decay with the escape
    rate and refining the leading term by residue-dependent
    quasistationary ambiguity amplitudes;
  \item a supplementary finite-dimensional resolvent reformulation that
    places all possible scalar poles of the Markov error generating
    function in the reciprocal \(q\)-th cyclotomic lift of the killed
    matrix spectrum and characterizes the lifted eigenvalues that are
    actually visible;
  \item a killed-matrix power-sum identity for survivor words, whose
    standard survivor R\'enyi pressure corollary yields the
    R\'enyi-\(2\) entropy rate and the resulting iid Poisson birthday
    thresholds, with the required third-moment condition stated below.
\end{enumerate}
The center of gravity is the escape-rate theorem together with its
quasistationary amplitude refinement; the resonance, survivor-R\'enyi,
and collision statements are subsequent consequences of the same
killed-operator formalism.
Thus the new finite-observation input is the Bayes-error observable,
its exact killed-matrix representation, and the residue-dependent leading
amplitude.  Once that representation is in place, the exponential and
amplitude asymptotics are finite-dimensional Perron--Frobenius
consequences, while the survivor R\'enyi and resolvent statements are
included to record the pressure and finite-matrix spectral bookkeeping
needed later in the paper.
The finite-state assumptions used for the main theorems are finite graph
or matrix hypotheses.  The first-entry residue theorem uses the
two-terminal residue reachability condition; the survivor R\'enyi
corollary uses only the one-step Markov presentation and the mixing
survivor subshift.  These are not spectral assumptions for an open Ruelle
operator, and the main theorems do not contain a H\"older open-operator
extension.  The only cyclotomic resolvent theorem in the paper is the
finite-state theorem below.

\paragraph{Logical scope of the hypotheses.}
The hypotheses are used in four separate layers.
\begin{description}[leftmargin=2.8cm,style=nextline]
  \item[Escape and amplitudes.]
  Theorems~\ref{thm:main-amplitude} and
  \ref{thm:first-entry-escape-rate} assume a finite Markov presentation,
  a mixing survivor subshift, and the two-terminal residue reachability
  condition in a finite graph.  These are checkable graph or matrix
  hypotheses; they do not assert a spectral gap beyond the
  Perron--Frobenius theorem for a finite primitive matrix.
  \item[Survivor R\'enyi spectrum.]
  Corollary~\ref{cor:main-renyi} assumes the same one-step Markov model
  and mixing survivor subshift, but it does not use a residue modulus, a
  residue set, or the two-terminal reachability condition.
  \item[Resonance lift.]
  Theorem~\ref{thm:main-resonance} is only the rational resolvent of the
  finite lifted matrix \(\widehat B_H\).  Its poles are finite-matrix
  poles, not Banach-space Ruelle resonances.
  \item[H\"older appendix.]
  Theorem~\ref{thm:holder-open-transfer} is a standalone
  \(C^\alpha\)-space statement for one-step Markov presentations with
  one-step holes: it defines the open Ruelle operator, proves the
  isolated leading eigenvalue and adjoint eigenfunctional, and identifies
  the survivor pressure from standard Ruelle--Perron--Frobenius theory
  plus a kernel estimate.  Theorem~\ref{thm:holder-finite-cylinder-escape}
  first chooses a block length that dominates both an ambient Markov
  memory and the cylinder length of the hole, and then proves the
  finite-prefix Bayes-error exponent from Gibbs distortion and the finite
  reachability condition.  Neither appendix theorem supplies a hypothesis
  to the finite-state main theorem.
\end{description}

\begin{center}
\scriptsize
\setlength{\tabcolsep}{2pt}
\begin{tabular}{
  >{\raggedright\arraybackslash}p{0.22\textwidth}
  >{\raggedright\arraybackslash}p{0.17\textwidth}
  >{\raggedright\arraybackslash}p{0.20\textwidth}
  >{\raggedright\arraybackslash}p{0.16\textwidth}
  >{\raggedright\arraybackslash}p{0.19\textwidth}}
\hline
Hypothesis or datum & Used in & Status & Where stated & Role \\
\hline
\(Y\) is a topologically mixing one-step SFT
& Theorems~\ref{thm:main-amplitude},
\ref{thm:main-resonance}; Corollary~\ref{cor:main-renyi}
& finite adjacency condition
& theorem statements
& gives a finite primitive Markov presentation \\
\(\varphi\) is one-step and \(\nu_\varphi\) is its equilibrium Markov
measure
& Theorems~\ref{thm:main-amplitude},
\ref{thm:main-resonance}; Corollary~\ref{cor:main-renyi}
& standard finite-state thermodynamic data
& theorem statements
& gives the positive transition matrix \(K\) and cylinder weights \\
\(H=\bigcup_{h\in A_{\mathrm{hole}}}[h]\) with
\(\emptyset\neq A_{\mathrm{hole}}\subsetneq A\)
& Theorems~\ref{thm:main-amplitude},
\ref{thm:main-resonance}; Corollary~\ref{cor:main-renyi}
& finite one-step hole data
& theorem statements
& makes the first-entry and survivor problems non-trivial \\
\(Y_H\) is topologically mixing
& Theorems~\ref{thm:main-amplitude},
\ref{thm:main-resonance}; Corollary~\ref{cor:main-renyi}
& additional finite graph condition, not automatic
& theorem statements
& makes the surviving class primitive and fixes the leading pressure gap \\
\(q\ge2\) and
\(\emptyset\neq R\subsetneq\ZZ/q\ZZ\)
& Theorems~\ref{thm:main-amplitude},
\ref{thm:main-resonance}
& chosen non-trivial residue observable
& theorem statements
& excludes residue events that are empty or certain by definition \\
Two-terminal residue reachability in \(\mathcal G_{H,q}\)
& Theorems~\ref{thm:main-amplitude},
\ref{thm:main-resonance}
& additional finite graph non-degeneracy condition
& theorem statements; Lemma~\ref{lem:residue-reachability}
& gives \(0<\beta_{i,r}\leq1/2\); not used by
Corollary~\ref{cor:main-renyi} \\
Spectral non-cancellation for a lifted eigenvalue
& Theorem~\ref{thm:main-resonance}
& not a standing hypothesis
& conclusion of the theorem
& decides whether a possible finite-matrix pole is visible in the scalar
generating function \\
\hline
\end{tabular}
\end{center}

\begin{center}
\small
\begin{tabular}{
  >{\raggedright\arraybackslash}p{0.24\textwidth}
  >{\raggedright\arraybackslash}p{0.34\textwidth}
  >{\raggedright\arraybackslash}p{0.29\textwidth}}
\hline
Result & Hypotheses used & Status of the conclusion \\
\hline
Theorems~\ref{thm:main-amplitude} and
\ref{thm:first-entry-escape-rate}
& One-step Markov model, mixing survivor subshift, finite
two-terminal reachability
& Exact killed-matrix identity, escape exponent, and amplitude from
finite Perron--Frobenius theory \\
Corollary~\ref{cor:main-renyi}
& One-step Markov model and mixing survivor subshift; no residue
reachability hypothesis
& Killed-matrix power-sum identity, standard survivor pressure
corollary, and iid collision threshold \\
Theorem~\ref{thm:main-resonance}
& The finite matrix \(B_H\) and its residue lift \(\widehat B_H\)
& Finite matrix resolvent; possible poles come from \(\widehat B_H\),
with scalar visibility characterized by non-cancellation \\
Appendix~\ref{app:holder-finite-cylinder}
& Block length dominating ambient memory and hole cylinders,
one-step block hole, \(C^\alpha\) transfer space, Gibbs property,
survivor mixing, finite reachability
& Standalone open-operator spectral gap and finite-prefix escape
exponent; no Banach-space cyclotomic resolvent claim \\
\hline
\end{tabular}
\end{center}

\begin{theorem}[Quasistationary ambiguity amplitudes]
\label{thm:main-amplitude}
Let \(Y\subseteq A^{\NNzero}\) be a topologically mixing one-step
subshift of finite type, let \(\nu_\varphi\) be the fully supported
equilibrium Markov measure of a one-step potential, and write \(K\) for
its transition matrix.  Let
\[
  H=\bigcup_{h\in A_{\mathrm{hole}}}[h],
  \qquad
  \emptyset\neq A_{\mathrm{hole}}\subsetneq A,
  \qquad
  \mathcal S=A\setminus A_{\mathrm{hole}},
\]
and assume that the survivor subshift \(Y_H\) is topologically mixing.
Fix \(q\geq2\) and
\(\emptyset\neq R\subsetneq \ZZ/q\ZZ\).  Let
\[
  \mathsf P_R
  :=\{y:\tau_H(y)<\infty,\ \tau_H(y)\bmod q\in R\}.
\]
Form the \(q\)-cyclic graph \(\mathcal G_{H,q}\) with vertices
\(\mathcal S\times\ZZ/q\ZZ\) and terminal vertices
\(\star_R,\star_{R^c}\), with safe edges
\[
  (i,r)\to(j,r+1)
  \quad\text{when }j\in\mathcal S\text{ and }K_{ij}>0,
\]
and terminal edges \((i,r)\to\star_R\) or
\((i,r)\to\star_{R^c}\) when \(K_{ih}>0\) for some
\(h\in A_{\mathrm{hole}}\) and \(r+1\) lies respectively in \(R\) or in
\(\ZZ/q\ZZ\setminus R\).  Assume the two-terminal residue reachability
hypothesis: for every \((i,r)\in\mathcal S\times\ZZ/q\ZZ\), both
\(\star_R\) and \(\star_{R^c}\) are reachable from \((i,r)\).  Equivalently,
from every safe state \(i\) and starting residue \(r\), there are
admissible futures whose first entry into \(A_{\mathrm{hole}}\) occurs
at residues in both \(R-r\) and \((\ZZ/q\ZZ\setminus R)-r\).

For \(i\in\mathcal S\), let
\(\PP_i\) be the Markov law started at \(i\), set
\[
  \tau_H^+:=\inf\{n\geq1:y_n\in A_{\mathrm{hole}}\},
  \qquad
  p_{i,r}:=\PP_i(r+\tau_H^+\in R\bmod q),
\]
and define \(b_r(i):=\min\{p_{i,r},1-p_{i,r}\}\).  Let \(B_H\) be the
killed transition matrix on \(\mathcal S\),
\[
  B_H(i,j):=K_{ij}\mathbf 1_{\{j\in\mathcal S\}},
\]
let \(\mathcal S_\infty\) be the states appearing in \(Y_H\), and put
\(\mathcal T=\mathcal S\setminus\mathcal S_\infty\).  Order
\(\mathcal S=\mathcal S_\infty\sqcup\mathcal T\) and write
\[
  B_H=
  \begin{pmatrix}
    B_\infty & C\\
    0 & Q
  \end{pmatrix},
\]
where \(B_\infty=B_H|_{\mathcal S_\infty}\).  Let
\(\ell_\infty>0\) be a left Perron--Frobenius eigenvector of
\(B_\infty\) for
\(\rho_H:=\rho(B_\infty)=\rho(B_H)\), define
\[
  \ell_H^{\mathsf T}
  :=
  \bigl(
    \ell_\infty^{\mathsf T},
    \ell_\infty^{\mathsf T}C(\rho_H I-Q)^{-1}
  \bigr),
  \qquad
  \xi_H:=\frac{\ell_H}{\ell_H^{\mathsf T}\mathbf 1}.
\]
Then for every residue class \(r\),
\[
  \lim_{n\to\infty}
  \frac{\varepsilon_{nq+r+1}(\mathsf P_R;\nu_\varphi)}
       {\nu_\varphi(\tau_H\ge nq+r+1)}
  = \xi_H^{\mathsf T}b_r.
\]
Equivalently,
\[
  \varepsilon_m(\mathsf P_R;\nu_\varphi)
  = \bigl(\xi_H^{\mathsf T}b_{m-1\bmod q}+o(1)\bigr)\,
    \nu_\varphi(\tau_H\ge m).
\]
In particular, finite observation records not only the escape exponent
\(\gamma_H(\varphi)\) but also a residue-dependent quasistationary
ambiguity amplitude.
\end{theorem}

\noindent
The next statement is the pressure corollary of the matrix identity in
Lemma~\ref{lem:survivor-renyi-matrix}.  It is included here because its
\(s=2\) and \(s=3\) cases determine the iid collision threshold for
conditioned survivor samples.

\begin{corollary}[Survivor R\'enyi pressure spectrum]
\label{cor:main-renyi}
Assume again that \(Y\) is a mixing one-step subshift of finite type,
\(\nu_\varphi\) is the equilibrium Markov measure of a one-step
potential \(\varphi\), and \(H\) is a Markov hole with mixing survivor
subshift \(Y_H\).  For \(m\geq 1\), let \(\mathcal L_m^H\) be the set of
admissible length-\(m\) words that avoid \(H\) for the first \(m\)
steps, and for \(s>0\) define
\[
  \mathcal Z_{H,m}(s)
  := \sum_{a\in \mathcal L_m^H} \nu_\varphi([a])^s.
\]
Then
\[
  \lim_{m\to\infty}
  -\frac1m\log \mathcal Z_{H,m}(s)
  = sP_Y(\varphi)-P_{Y_H}(s\varphi).
\]

If \(\lambda_m^H\) denotes the law of the survivor word conditioned on
survival to depth \(m\),
\[
  \lambda_m^H(a)
  := \frac{\nu_\varphi([a])}{\nu_\varphi(\tau_H\ge m)},
  \qquad a\in\mathcal L_m^H,
\]
then
\[
  \lim_{m\to\infty}
  -\frac1m\log
  \sum_{a\in\mathcal L_m^H} \lambda_m^H(a)^s
  = sP_{Y_H}(\varphi)-P_{Y_H}(s\varphi).
\]

In particular, at \(s=2\) the conditioned survivor law has R\'enyi-\(2\)
entropy rate
\[
  h_{2,H}:=2P_{Y_H}(\varphi)-P_{Y_H}(2\varphi),
\]
while the ambient two-sample exponent for simultaneous survival and
visible collision is
\[
  \chi_H:=2P_Y(\varphi)-P_{Y_H}(2\varphi)
        = 2\gamma_H(\varphi)+h_{2,H}.
\]
\end{corollary}

This corollary is the survivor analogue of the standard \(L^s\)-spectrum
or R\'enyi pressure identity for Gibbs weights; compare
\citet{Renyi1961,HentschelProcaccia1983,Pesin1997}.  Its role here is
not to introduce a new pressure formalism, but to record the exact
pressure consequence needed for the collision estimates after
conditioning on survival.

The identity \(\chi_H = 2\gamma_H(\varphi)+h_{2,H}\) places the
collision story inside open-system thermodynamic formalism.  Combined
with a Chen--Stein argument (Section~\ref{sec:double-budget}), it yields
iid Poisson birthday thresholds for surviving samples on the scale
\(\binom{N_m}{2}e^{-m h_{2,H}}\asymp 1\), provided
\[
  N_m^3S_3^H(m)\to0,
  \qquad
  S_3^H(m)=e^{-mh_{3,H}+o(m)},
  \qquad
  h_{3,H}:=3P_{Y_H}(\varphi)-P_{Y_H}(3\varphi).
\]
Thus at the pure exponential threshold
\(N_m=e^{(h_{2,H}/2+o(1))m}\), Poisson convergence requires
\[
  h_{3,H}>\frac32h_{2,H}.
\]
These collision statements are iid sampling results for independent
survivor prefixes; no orbit-sampling or weak-dependence Poisson theorem
is asserted here.

\medskip
\noindent
The next theorem is the finite-state resolvent reformulation of the
preceding asymptotics.  The possible-pole set is the ordinary spectrum of
the finite Kronecker lift \(\widehat B_H\); the scalar content is the
non-cancellation criterion deciding which lifted eigenvalues are visible
in the one-variable error generating function.  It has only the finite
Markov hypotheses stated in Theorem~\ref{thm:first-entry-escape-rate};
the H\"older block-recoding comparison in
Appendix~\ref{app:holder-finite-cylinder} is a separate finite-prefix
exponent statement.

\begin{theorem}[Finite-state cyclotomic resolvent]
\label{thm:main-resonance}
For first-entry residue events of one-step Markov holes under the
finite-state hypotheses of
Theorem~\ref{thm:first-entry-escape-rate}, the Bayes error generating
function is the finite-dimensional residue-twisted killed resolvent
\[
  \mathcal E_{\mathsf P_R}(z)
  = z\,\widehat u^{\mathsf T}(I-z\widehat B_H)^{-1}\widehat b,
\]
and every pole belongs to the \(q\)-th cyclotomic lift of the killed
spectrum after taking reciprocals.  The positive real pole of smallest
modulus is
\[
  z_H=e^{\gamma_H(\varphi)}
      =e^{P_Y(\varphi)-P_{Y_H}(\varphi)},
\]
which is \(\rho_H^{-1}\), the reciprocal of the leading killed
eigenvalue \(\rho_H=e^{-\gamma_H(\varphi)}\).
For a non-zero lifted eigenvalue \(\alpha\), let \(\Pi_\alpha\) be its
spectral projection and set
\(N_\alpha=(\widehat B_H-\alpha I)\Pi_\alpha\).  The reciprocal
\(\alpha^{-1}\) is an actual scalar pole precisely when
\(\widehat u^{\mathsf T}N_\alpha^j\Pi_\alpha\widehat b\) is non-zero for
some nilpotent level \(j\); equivalently, for a simple lifted eigenvalue,
the product
\((\widehat u^{\mathsf T}v_\alpha)
(\ell_\alpha^{\mathsf T}\widehat b)\) is non-zero.
This theorem is only a statement about the finite matrix
\(\widehat B_H\).
\end{theorem}

The H\"older material appears in
Appendix~\ref{app:holder-finite-cylinder} as two separate block-recoding
statements for finite unions of cylinders and H\"older potentials.  The
open-transfer theorem is formulated on the explicit Banach space
\(C^\alpha\) and proves its leading spectral data; the Bayes-error
theorem uses only Gibbs distortion, survivor pressure, and finite
reachability to obtain the finite-prefix escape exponent.  Both appendix
statements are independent of the finite-state resolvent theorem.

\subsection*{Context and comparison}

Thermodynamic formalism for symbolic systems and equilibrium states is
classical; we use standard material from
\citet{Bowen2008,Ruelle1978,Walters1975GMeasures,WaltersErgodicTheory}.
The open-system viewpoint---escape through holes, killed transfer
operators, and quasistationary behaviour---goes back to
\citet{PianigianiYorke1979} and was developed for symbolic systems by
\citet{ColletMartinezSchmitt1997}, \citet{DemersYoung2006},
\citet{KellerLiverani2009}, \citet{BunimovichYurchenko2011},
\citet{DemersWrightYoung2012}, and
\citet{FergusonPollicott2012}.  For closely related work on symbolic
holes, Markov holes, symbolic codings of hyperbolic open systems, and
the dependence of escape rates on hole placement, see
\citet{BundfussKrugerTroubetzkoy2011},
\citet{BruinDemersTodd2018,CheriyathAgarwal2023},
\citet{BonannoCristadoroLenci2022}, and
\citet{AgarwalCheriyathTikekar2024}; see also
\citet{DemersTodd2017} for pressure and equilibrium states in the
presence of holes for multimodal maps and
\citet{BruinDemers2022} for recent extensions to dispersing billiards.
For local escape rates and the connection between hitting and escaping
statistics in mixing systems, see
\citet{BruinDemersTodd2018,HaydnYang2020}.
Two recent benchmarks are close in open-system theme but different in
observable: \citet{DemersTodd2021} prove asymptotic escape rates and
limiting conditionally invariant distributions for multimodal maps with
small holes and H\"older/geometric potentials, while
\citet{AtnipFroylandGonzalezTokmanVaienti2023} develop equilibrium
states for random open and closed interval systems without transitivity
assumptions and relate Lyapunov exponents to escape through random
holes.  The present paper stays in the finite-observation symbolic
setting and asks which open-system invariants are visible from finite
Bayes errors and collision statistics.
The observable here is not the survivor set itself.  A finite prefix
that has not yet entered the hole leaves a residue-dependent future
ambiguity, and the Bayes error sums this ambiguity over all safe
prefixes.  This produces the terminal weights \(b_r\), the
degenerate-residue dichotomy, and the quasistationary amplitude
\(\xi_H^{\mathsf T}b_r\).  These quantities do not appear in the usual
escape-rate pressure identity, where survivor mass is summed with
terminal weight \(1\).
The Poisson approximation input is the Chen--Stein method of
\citet{ArratiaGoldsteinGordon1989}.  What is specific here is the finite
observation question: the visible symbolic coding recovers the
escape-rate pressure gap and the residue-dependent quasistationary
amplitudes, while the cyclotomic resolvent and survivor R\'enyi spectrum
are downstream spectral and pressure corollaries.

\subsection*{Theorem hierarchy}

The manuscript is organized around one dynamical theme: finite
observation detects the open-system pressure gap and its
quasistationary refinement.
Section~\ref{sec:open-system} is therefore the center of gravity.  Its
first subsection proves the escape-rate theorem and the amplitude
refinement.  The resolvent discussion in the remainder of
Section~\ref{sec:open-system} is supplementary to that main theorem,
packaging the same killed-operator data in spectral language.
Section~\ref{sec:double-budget} records secondary thermodynamic
consequences of the same calculation.  The H\"older finite-cylinder
appendix is retained only as supporting block-recoding material for
finite-cylinder holes and is not part of the finite-state resolvent
argument.

\subsection*{Organization}

Section~\ref{sec:preliminaries} introduces the visible symbolic factor
and finite observation \(\sigma\)-algebras.
Section~\ref{sec:clarity} develops the exact Bayes error formulas and
the weighted boundary transfer theorem.
Section~\ref{sec:open-system} proves
Theorems~\ref{thm:main-amplitude} and~\ref{thm:main-resonance}.
Section~\ref{sec:double-budget} proves
Corollary~\ref{cor:main-renyi} and derives the collision consequences.
Section~\ref{sec:conclusion} collects brief final remarks.
Appendix~\ref{app:holder-finite-cylinder} gives the supporting H\"older
finite-cylinder open-operator and escape-exponent theorems after
block recoding with block length at least the ambient Markov memory.
