% ------------------------------------------------------------------
%  Holder Gibbs finite-cylinder holes: block recoding and pressure
% ------------------------------------------------------------------

\section{H\"older cylinder holes after block recoding}
\label{app:holder-finite-cylinder}

This appendix is independent of the finite-state resolvent proposition.  It
has two limited purposes.  First, for a one-step Markov presentation with
a one-step cylinder hole it records the actual H\"older
transfer-operator statement on the Banach space
\(C^\alpha\): the open operator is defined explicitly, its leading
eigenvalue and adjoint eigenfunctional are obtained from the survivor
Ruelle--Perron--Frobenius theorem, and the complementary part is shown to
have smaller spectral radius.  Second, after a finite block recoding
whose block length also dominates an ambient Markov memory, it proves the
finite-prefix Bayes-error comparison needed for general Gibbs weights.
No statement in this appendix is a hypothesis for the finite-state
cyclotomic resolvent statement.  At the operator level, this appendix is
closest in spirit to the open-transfer frameworks of
\citet{KellerLiverani2009,FroylandStancevic2010,AtnipFroylandGonzalezTokmanVaienti2023},
but it is used here only as supporting finite-cylinder material rather
than as the source of the finite-state cyclotomic proposition.

The Bayes-error comparison has four finite steps.  First, Gibbs
distortion gives uniform lower bounds for bounded-length continuations.
Second, a finite two-terminal reachability condition gives uniform
conditional ambiguity on every surviving prefix.  Third, a safe-prefix
decomposition identifies the Bayes error with a survivor-cylinder sum.
Fourth, the standard pressure formula for the closed survivor subshift
gives the exponential rate of that sum.

Throughout the one-step part of this appendix, a one-step Markov
presentation means that the ambient shift is given by a finite directed
graph on its symbols and the hole is determined by a subset of symbols.
The safe edges used below are precisely the admissible directed edges of
this one-step presentation whose endpoints are safe.  For a general
finite-cylinder hole, Lemma~\ref{lem:holder-block-reduction} first passes
to a higher-block presentation whose block length is large enough both to
make the ambient SFT one-step and to make the hole a one-step symbol
hole.

We use the following standard form of the Gibbs property.  If \(Z\) is a
topologically mixing one-sided subshift of finite type, \(\psi\) is
H\"older, and \(\nu_\psi\) is its equilibrium state, then there are
constants \(C_G\geq1\) and \(V_\psi<\infty\) such that for every
admissible word \(w=w_0\cdots w_{n-1}\) and every \(x\in[w]\),
\[
  C_G^{-1}e^{S_n\psi(x)-nP_Z(\psi)}
  \leq \nu_\psi([w])
  \leq
  C_G e^{S_n\psi(x)-nP_Z(\psi)},
\]
and \(\abs{S_n\psi(x)-S_n\psi(y)}\leq V_\psi\) whenever
\(x,y\in[w]\).

\begin{lemma}[Uniform Gibbs continuation]
\label{lem:holder-gibbs-continuation}
Let \(Z,\psi,\nu_\psi\) be as above.  For every integer \(L\geq1\) there
is a constant \(c_L>0\) such that, whenever \(w\) and \(v\) are
admissible words, \(\abs{v}\leq L\), and the concatenation \(wv\) is
admissible,
\[
  \frac{\nu_\psi([wv])}{\nu_\psi([w])}\geq c_L .
\]
\end{lemma}

\begin{proof}
Write \(n=\abs w\) and \(k=\abs v\leq L\), and choose
\(x\in[wv]\).  The Gibbs estimate applied to \(wv\), and then to \(w\),
gives
\[
  \frac{\nu_\psi([wv])}{\nu_\psi([w])}
  \geq
  C_G^{-2}
  \exp\bigl(S_{n+k}\psi(x)-(n+k)P_Z(\psi)
            -\sup_{y\in[w]}(S_n\psi(y)-nP_Z(\psi))\bigr).
\]
Since \(x\in[w]\), the Bowen distortion bound gives
\[
  S_n\psi(x)-\sup_{y\in[w]}S_n\psi(y)\geq -V_\psi.
\]
The remaining \(k\) iterates are bounded below by
\(-L\norm{\psi}_\infty\), while \(-kP_Z(\psi)\geq -L\abs{P_Z(\psi)}\).
Thus
\[
  \frac{\nu_\psi([wv])}{\nu_\psi([w])}
  \geq
  C_G^{-2}
  \exp\bigl(-V_\psi-L\norm{\psi}_\infty
              -L\abs{P_Z(\psi)}\bigr),
\]
which is positive and depends only on \(L\), \(Z\), and \(\psi\).
\end{proof}

\begin{lemma}[Uniform finite-prefix ambiguity]
\label{lem:holder-extension-ambiguity}
Let \(Z\) be a topologically mixing one-sided one-step subshift of finite
type, let \(\psi\) be a H\"older potential, and let \(\nu_\psi\) be its
equilibrium state.  Let \(H\) be a non-empty proper one-step Markov hole
with safe alphabet \(\mathcal S\).  Fix \(q\geq2\) and
\(\emptyset\neq R\subsetneq\ZZ/q\ZZ\).  Assume the finite two-terminal
residue reachability condition on the graph
\[
  \mathcal S\times\ZZ/q\ZZ,\qquad \star_R,\quad\star_{R^c},
\]
where safe edges increase the residue by one along admissible safe
transitions, and a terminal edge records the residue of the first
admissible transition into \(H\): both terminal vertices are reachable
from every safe-residue vertex.

Let
\[
  \tau_H(z):=\inf\{n\geq0:z_n\notin\mathcal S\},
  \qquad
  \mathsf P_R:=\{z:\tau_H(z)<\infty,\ \tau_H(z)\in R\bmod q\}.
\]
Then there is \(c_0>0\) such that, for every admissible safe word
\(w=w_0\cdots w_{n-1}\) of length \(n\geq1\),
\[
  c_0
  \leq
  \nu_\psi(\mathsf P_R\mid[w])
  \leq
  1-c_0 .
\]
\end{lemma}

\begin{proof}
For each pair \((i,r)\in\mathcal S\times\ZZ/q\ZZ\), the reachability
condition gives two admissible first-hit continuations from \(i\): one
whose first hit has residue in \(R-r\), and one whose first hit has
residue in \((\ZZ/q\ZZ\setminus R)-r\).  Concretely, these are finite
words \(v^{R}_{i,r}\) and \(v^{R^c}_{i,r}\), appended after the symbol
\(i\), whose intermediate symbols are safe, whose last symbol lies in
the hole, and whose first-hit residues force \(\mathsf P_R\) and
\(\mathsf P_R^c\), respectively.  There are only finitely many
\((i,r)\), so let \(L\) be the maximum of the two continuation lengths.

Let \(w=w_0\cdots w_{n-1}\) be safe, put \(i=w_{n-1}\), and put
\(r=n-1\bmod q\).  The cylinders \([wv^{R}_{i,r}]\) and
\([wv^{R^c}_{i,r}]\) are contained in \([w]\), and they force
\(\mathsf P_R\) and \(\mathsf P_R^c\), respectively.  Lemma
\ref{lem:holder-gibbs-continuation} gives a constant \(c_L>0\) such that
\[
  \nu_\psi([wv^{R}_{i,r}]\mid[w])\geq c_L,
  \qquad
  \nu_\psi([wv^{R^c}_{i,r}]\mid[w])\geq c_L .
\]
Taking \(c_0=\min\{c_L,\tfrac12\}\) gives
\[
  \nu_\psi(\mathsf P_R\mid[w])\geq c_0,
  \qquad
  \nu_\psi(\mathsf P_R^c\mid[w])\geq c_0,
\]
which is the claimed two-sided ambiguity bound.
\end{proof}

\begin{lemma}[Safe-prefix Bayes decomposition]
\label{lem:holder-safe-prefix-bayes-decomposition}
Let \(Z\) be a topologically mixing one-sided one-step subshift of finite
type, let \(\nu\) be a fully supported probability measure on its
cylinders, and let \(H\) be a one-step Markov hole with safe alphabet
\(\mathcal S\).  Fix \(q\geq2\), a non-empty proper
\(R\subsetneq\ZZ/q\ZZ\), and
\[
  \tau_H(z):=\inf\{n\geq0:z_n\notin\mathcal S\},
  \qquad
  \mathsf P_R:=\{z:\tau_H(z)<\infty,\ \tau_H(z)\in R\bmod q\}.
\]
Then for every \(n\geq1\),
\[
  \varepsilon_n(\mathsf P_R;\nu)
  =
  \sum_{\substack{w\in\mathcal L_n(Z)\\
                  w_0,\ldots,w_{n-1}\in\mathcal S}}
    \nu([w])\,
    \min\{
      \nu(\mathsf P_R\mid[w]),
      1-\nu(\mathsf P_R\mid[w])
    \}.
\]
\end{lemma}

\begin{proof}
Apply the cylinder Bayes formula
Proposition~\ref{prop:scan-error-cylinder} to the length-\(n\) cylinder
partition.  If \(w=w_0\cdots w_{n-1}\) contains a symbol outside
\(\mathcal S\), let \(t<n\) be its first such position.  Then
\(\tau_H=t\) on the whole cylinder \([w]\), so the truth value of
\(\mathsf P_R\) is already known from the prefix and the corresponding
Bayes factor is zero.  The remaining cylinders are exactly the safe
prefixes in the displayed sum.
\end{proof}

\begin{lemma}[Safe-word pressure]
\label{lem:holder-safe-word-pressure}
Let \(Z\) be a topologically mixing one-sided one-step subshift of finite
type, let \(\psi\) be a H\"older potential, and let \(\nu_\psi\) be its
equilibrium state.  Let \(H\) be a non-empty proper one-step Markov hole
with safe alphabet \(\mathcal S\), and assume that the survivor subshift
\[
  Z_H:=\{z\in Z:z_n\in\mathcal S\text{ for all }n\geq0\}
\]
is topologically mixing.  Then
\[
  \lim_{n\to\infty}
  -\frac1n\log \nu_\psi(\tau_H\geq n)
  =
  P_Z(\psi)-P_{Z_H}(\psi).
\]
\end{lemma}

\begin{proof}
Let \(\mathcal L_n^{\mathrm{safe}}\) be the admissible length-\(n\)
words all of whose symbols lie in \(\mathcal S\), and set
\[
  Z_n^{\mathrm{safe}}(\psi)
  :=
  \sum_{w\in\mathcal L_n^{\mathrm{safe}}}
    \exp\bigl(\sup_{[w]}S_n\psi\bigr).
\]
The Gibbs estimate gives constants \(C_1,C_2>0\) such that
\[
  C_1e^{-nP_Z(\psi)}Z_n^{\mathrm{safe}}(\psi)
  \leq
  \nu_\psi(\tau_H\geq n)
  \leq
  C_2e^{-nP_Z(\psi)}Z_n^{\mathrm{safe}}(\psi).
\]
It remains to compute the exponential growth of
\(Z_n^{\mathrm{safe}}(\psi)\).

Let \(\mathcal S_\infty\subseteq\mathcal S\) be the symbols that occur
in \(Z_H\), and put \(\mathcal T=\mathcal S\setminus\mathcal S_\infty\).
There is no safe edge from \(\mathcal T\) into \(\mathcal S_\infty\);
otherwise an infinite survivor path beginning in \(\mathcal S_\infty\)
could be appended after that edge, placing the initial transient symbol
in \(Z_H\).  The directed graph on \(\mathcal T\) is acyclic, since a
cycle in \(\mathcal T\) could be repeated indefinitely to produce a
survivor orbit through a transient symbol.  Hence there is a number
\(K\) such that every safe word is a survivor word over
\(\mathcal S_\infty\), possibly followed by a transient suffix of length
at most \(K\).  Words of total length at most \(K\) are irrelevant for
the limit.

Let
\[
  Z_n^{H}(\psi)
  :=
  \sum_{\substack{w\text{ admissible in }Z,\ \abs w=n\\
                  w_0,\ldots,w_{n-1}\in\mathcal S_\infty}}
    \exp\bigl(\sup_{[w]\cap Z_H}S_n\psi\bigr)
\]
be the usual survivor partition sum.  Since survivor words are safe,
\(Z_n^{\mathrm{safe}}(\psi)\geq Z_n^H(\psi)\).  Conversely, if
\(w=uv\) with \(u\) a survivor word and \(v\) a transient suffix of
length \(k\leq K\), then
\[
  \exp\bigl(\sup_{[w]}S_n\psi\bigr)
  \leq
  e^{V_\psi+K\norm{\psi}_\infty}
  \exp\bigl(\sup_{[u]\cap Z_H}S_{n-k}\psi\bigr).
\]
Here the factor \(e^{V_\psi}\) compares the ambient supremum over the
survivor prefix cylinder \([u]\) with the supremum over
\([u]\cap Z_H\); the latter is non-empty because \(u\) is a word in the
mixing survivor subshift.
There are only finitely many possible transient suffixes of length at
most \(K\).  Therefore, for a constant \(C_3<\infty\),
\[
  Z_n^{\mathrm{safe}}(\psi)
  \leq
  C_3\sum_{k=0}^{K} Z_{n-k}^{H}(\psi)
\]
for all large \(n\).  Because \(Z_H\) is a mixing subshift of finite type
and \(\psi|_{Z_H}\) is H\"older,
\[
  \lim_{n\to\infty}\frac1n\log Z_n^H(\psi)
  =
  P_{Z_H}(\psi)
\]
by the standard pressure formula for H\"older potentials on mixing SFTs
\citep{Bowen2008,Ruelle1978}.  The preceding lower and upper bounds show
that \(Z_n^{\mathrm{safe}}(\psi)\) has the same exponential growth rate.
Combining this with the Gibbs comparison for
\(\nu_\psi(\tau_H\geq n)\) proves the displayed escape exponent.
\end{proof}

\begin{lemma}[Lipschitz retraction onto the survivor subshift]
\label{lem:holder-survivor-extension}
Let \(Z\) be a topologically mixing one-sided one-step subshift of finite
type, let \(H\) be a non-empty proper one-step Markov hole with safe
alphabet \(\mathcal S\), and assume that the survivor subshift
\[
  Z_H=\{z\in Z:z_n\in\mathcal S\text{ for all }n\geq0\}
\]
is topologically mixing.  Let \(\mathcal S_\infty\subseteq\mathcal S\)
be the alphabet of symbols that occur in \(Z_H\).  For each
\(i\in\mathcal S_\infty\), choose a point
\(\omega^{(i)}\in Z_H\cap[i]\), and choose one base point
\(\omega^\ast\in Z_H\).

Then there is a \(1\)-Lipschitz retraction
\[
  \pi_H:Z\longrightarrow Z_H
\]
with \(\pi_H|_{Z_H}=\mathrm{id}_{Z_H}\).  Consequently
\[
  \mathcal J_H:\mathcal B_\theta(Z_H)\longrightarrow \mathcal B_\theta(Z),
  \qquad
  \mathcal J_H g:=g\circ\pi_H,
\]
is a bounded linear extension operator satisfying
\[
  \mathcal R\mathcal J_H=\mathrm{id}_{\mathcal B_\theta(Z_H)},
  \qquad
  \norm{\mathcal J_H g}_\infty=\norm{g}_\infty,
  \qquad
  \abs{\mathcal J_H g}_\theta\leq\abs{g}_\theta .
\]
\end{lemma}

\begin{proof}
For \(x\in Z\), let
\[
  n(x):=\inf\{m\geq0:x_m\notin\mathcal S_\infty\},
\]
with the convention \(n(x)=\infty\) if the set is empty.  Because \(Z\)
is one-step and every symbol in \(\mathcal S_\infty\) lies in a safe
infinite path, a point belongs to \(Z_H\) exactly when all of its
symbols lie in \(\mathcal S_\infty\).  Thus \(n(x)=\infty\) if and only
if \(x\in Z_H\).

Define \(\pi_H(x)\) as follows.
\begin{enumerate}[label=(\roman*)]
  \item If \(n(x)=\infty\), set \(\pi_H(x)=x\).
  \item If \(n(x)=0\), set \(\pi_H(x)=\omega^\ast\).
  \item If \(1\leq n(x)<\infty\), write \(n=n(x)\) and
  \(i=x_{n-1}\in\mathcal S_\infty\), and set
  \[
    \pi_H(x):=
    x_0x_1\cdots x_{n-1}\omega^{(i)}_1\omega^{(i)}_2\cdots .
  \]
\end{enumerate}
This point lies in \(Z_H\), and \(\pi_H(x)=x\) for \(x\in Z_H\), so
\(\pi_H\) is a retraction.

To prove the Lipschitz bound, let \(x,y\in Z\) agree in their first
\(N\geq0\) symbols.  If both \(n(x)\) and \(n(y)\) are at least \(N\),
then \(\pi_H(x)\) and \(\pi_H(y)\) share the same first \(N\) symbols,
so \(d_\theta(\pi_H(x),\pi_H(y))\leq\theta^N=d_\theta(x,y)\).  If, say,
\(n(x)<N\), then \(x_{n(x)}\notin\mathcal S_\infty\), hence also
\(y_{n(x)}=x_{n(x)}\notin\mathcal S_\infty\).  Therefore
\(n(y)=n(x)\).  In that case the definitions use the same initial word
and the same reference tail, so \(\pi_H(x)=\pi_H(y)\).  Thus
\(d_\theta(\pi_H(x),\pi_H(y))\leq d_\theta(x,y)\) in all cases.

Now let \(g\in\mathcal B_\theta(Z_H)\).  Linearity of \(\mathcal J_H\)
is immediate.  Since \(\pi_H\) is a retraction,
\(\mathcal R\mathcal J_H g=g\).  The sup-norm is preserved because
\(\pi_H(Z)=Z_H\), and the Lipschitz seminorm estimate follows from the
\(1\)-Lipschitz property of \(\pi_H\):
\[
  \abs{\mathcal J_H g(x)-\mathcal J_H g(y)}
  =
  \abs{g(\pi_Hx)-g(\pi_Hy)}
  \leq
  \abs{g}_\theta\,d_\theta(\pi_Hx,\pi_Hy)
  \leq
  \abs{g}_\theta\,d_\theta(x,y).
\]
\end{proof}

\begin{lemma}[Spectral comparison through a complemented quotient]
\label{lem:holder-quotient-spectrum}
Let \(X,Y\) be Banach spaces, let
\(R:X\to Y\) be a bounded surjection, and let
\(J:Y\to X\) be a bounded right inverse of \(R\), so
\(RJ=\mathrm{id}_Y\).
Set \(E:=\ker R\).  Let \(T\in\mathcal B(X)\) and
\(S\in\mathcal B(Y)\) satisfy
\[
  RT=SR .
\]
Then \(E\) is \(T\)-invariant.  Write \(T_E:=T|_E\).  The following hold.
\begin{enumerate}[label=(\roman*)]
  \item If \(\lambda\notin\spec(T_E)\cup\spec(S)\), then
  \(\lambda\notin\spec(T)\).  More precisely,
  \[
    (T-\lambda)^{-1}u
    =
    J(S-\lambda)^{-1}Ru
    +
    (T_E-\lambda)^{-1}
    \bigl(u-(T-\lambda)J(S-\lambda)^{-1}Ru\bigr).
  \]
  \item If \(\lambda\notin\spec(T_E)\) and \(\lambda\notin\spec(T)\),
  then \(\lambda\notin\spec(S)\).
  \item Consequently
  \[
    \spec(T)\setminus\spec(T_E)
    =
    \spec(S)\setminus\spec(T_E).
  \]
  \item If \(\lambda\notin\spec(T_E)\) and \(\lambda\) is an isolated
  eigenvalue of \(S\) of algebraic multiplicity one, then \(\lambda\) is
  an isolated eigenvalue of \(T\) of algebraic multiplicity one.
  Moreover the restriction map
  \[
    R:\ker(T-\lambda)\longrightarrow\ker(S-\lambda)
  \]
  is an isomorphism, and \(T\) has no non-zero generalized eigenvector
  at \(\lambda\).
\end{enumerate}
\end{lemma}

\begin{proof}
If \(e\in E\), then \(RTe=SRe=0\), so \(Te\in E\).  Thus \(E\) is
\(T\)-invariant.

For (i), fix \(u\in X\) and set \(y:=(S-\lambda)^{-1}Ru\),
\(x_0:=Jy\).  Then
\[
  R\bigl(u-(T-\lambda)x_0\bigr)
  =
  Ru-(S-\lambda)y
  =
  0,
\]
so \(e:=u-(T-\lambda)x_0\) lies in \(E\).  Since
\(\lambda\notin\spec(T_E)\), the element
\[
  x:=x_0+(T_E-\lambda)^{-1}e
\]
is well defined and satisfies \((T-\lambda)x=u\).  This gives the
displayed inverse formula, and every factor in it is bounded.

For (ii), assume \(\lambda\notin\spec(T_E)\) and \(T-\lambda\) is
invertible on \(X\).  If \(e\in E\), then the equation
\((T-\lambda)x=e\) has a unique solution \(x=(T-\lambda)^{-1}e\) in
\(X\).  It also has a unique solution in \(E\), namely
\((T_E-\lambda)^{-1}e\).  By uniqueness these two solutions coincide, so
\((T-\lambda)^{-1}E\subseteq E\).  Therefore \(T-\lambda\) induces a
bounded bijection on the quotient \(X/E\), whose inverse is induced by
\((T-\lambda)^{-1}\).  Since \(R\) identifies \(X/E\) with \(Y\), the
induced quotient map is exactly \(S-\lambda\).  Hence \(S-\lambda\) is
invertible.

Assertion (iii) is immediate from (i) and (ii).

For (iv), assertion (iii) shows that \(\lambda\) is isolated in
\(\spec(T)\) whenever it is isolated in \(\spec(S)\).  If
\(x\in\ker(T-\lambda)\) and \(Rx=0\), then \(x\in E\) and
\((T_E-\lambda)x=0\), hence \(x=0\) because
\(\lambda\notin\spec(T_E)\).  Thus the restriction map \(R\) is
injective on \(\ker(T-\lambda)\).

Now let \(y\in\ker(S-\lambda)\).  Put \(x_0:=Jy\).  Then
\[
  R\bigl((\lambda-T)x_0\bigr)
  =
  (\lambda-S)y
  =
  0,
\]
so \((\lambda-T)x_0\in E\).  Define
\[
  x:=x_0+(T_E-\lambda)^{-1}(\lambda-T)x_0 .
\]
Because \(T_E-\lambda=-(\lambda-T_E)\) on \(E\), this satisfies
\((T-\lambda)x=0\), while \(Rx=y\) because the correction term lies in
\(E\).  Therefore
\[
  R:\ker(T-\lambda)\longrightarrow\ker(S-\lambda)
\]
is surjective, hence an isomorphism.  In particular the eigenspace of
\(T\) at \(\lambda\) is one-dimensional when the eigenspace of \(S\) is.

Finally, let \(v\neq0\) be an eigenvector of \(T\) at \(\lambda\), and
suppose that \((T-\lambda)x=v\) for some \(x\in X\).  Applying \(R\)
gives
\[
  (S-\lambda)Rx=Rv .
\]
Since \(\lambda\) has algebraic multiplicity one for \(S\), the operator
\(S\) has no non-zero generalized eigenvector at \(\lambda\), so the
right-hand side must vanish: \(Rv=0\).  But then \(v\in E\) and
\((T_E-\lambda)v=0\), contradicting \(\lambda\notin\spec(T_E)\).  Thus
\(T\) has no non-zero generalized eigenvector at \(\lambda\).
\end{proof}

\begin{theorem}[Open transfer operator on the H\"older space]
\label{thm:holder-open-transfer}
Let \(Z\) be a topologically mixing one-sided one-step subshift of finite
type, let \(\psi\) be a H\"older potential, and let \(H\) be a
non-empty proper one-step Markov hole with safe alphabet \(\mathcal S\).
Assume that the survivor subshift
\[
  Z_H=\{z\in Z:z_n\in\mathcal S\text{ for all }n\geq0\}
\]
is topologically mixing.  Choose a standard symbolic metric
\(d_\theta\), \(0<\theta<1\), for which \(\psi\) is Lipschitz, and set
\[
  \mathcal B_\theta(Z):=
  \{f:Z\to\mathbb C:\norm{f}_\theta<\infty\},
  \qquad
  \norm{f}_\theta:=\norm{f}_\infty+
  \sup_{x\neq y}\frac{\abs{f(x)-f(y)}}{d_\theta(x,y)} .
\]
Define the open Ruelle operator
\[
  (\mathcal L_{\psi,H}f)(x)
  :=
  \sum_{\substack{\sigma y=x\\ y_0\in\mathcal S}}
    e^{\psi(y)}f(y),
  \qquad f\in\mathcal B_\theta(Z).
\]
Then \(\mathcal L_{\psi,H}\) is a bounded operator on
\(\mathcal B_\theta(Z)\).  With
\[
  \lambda_H:=e^{P_{Z_H}(\psi)},
\]
there exist \(h_H\in\mathcal B_\theta(Z)\), a non-zero positive
functional \(\eta_H\in\mathcal B_\theta(Z)^*\), and a bounded operator
\(\mathcal N_H\) on \(\mathcal B_\theta(Z)\) such that
\[
  \mathcal L_{\psi,H}h_H=\lambda_H h_H,\qquad
  \mathcal L_{\psi,H}^*\eta_H=\lambda_H\eta_H,\qquad
  \eta_H(h_H)=1,
\]
\[
  \mathcal L_{\psi,H}^n f
  =
  \lambda_H^n\eta_H(f)h_H+\mathcal N_H^n f,
  \qquad
  \rho(\mathcal N_H)<\lambda_H .
\]
Equivalently, for every \(f\in\mathcal B_\theta(Z)\) and every
\(\Lambda\in\mathcal B_\theta(Z)^*\),
\[
  \lambda_H^{-n}\Lambda(\mathcal L_{\psi,H}^n f)
  \longrightarrow
  \Lambda(h_H)\eta_H(f).
\]
Thus the leading open spectral data on the H\"older Banach space are
conclusions of the theorem, not hypotheses.
\end{theorem}

\begin{proof}
Write \(\abs{f}_\theta\) for the Lipschitz seminorm associated to
\(d_\theta\), and let \(p\) be a memory for the SFT \(Z\).  Boundedness is
the standard Ruelle estimate, and we recall the argument because the same
pairing is used below.  If \(x,x'\) agree for \(N\geq p\) symbols, then
every admissible safe one-step preimage of \(x\) is paired with the
preimage obtained by prepending the same safe symbol to \(x'\).  The
paired preimages agree for \(N+1\) symbols, and the Lipschitz property of
\(\psi\) gives
\[
  \abs{e^{\psi(y)}-e^{\psi(y')}}\leq C_\psi d_\theta(x,x').
\]
Since the number of one-step preimages is uniformly bounded,
\[
  \norm{\mathcal L_{\psi,H}f}_\theta
  \leq C\norm{f}_\theta
\]
for a constant \(C\) depending only on \(Z,\psi,H\), and \(\theta\).  The
finitely many cases \(N<p\) are absorbed by the sup-norm bound, since
\(d_\theta(x,x')\geq\theta^{p-1}\) there.

Let
\[
  \mathcal R:\mathcal B_\theta(Z)\longrightarrow
  \mathcal B_\theta(Z_H),
  \qquad
  \mathcal R f=f|_{Z_H},
\]
let \(\mathcal E:=\ker\mathcal R\), and let
\(\mathcal J_H:\mathcal B_\theta(Z_H)\to\mathcal B_\theta(Z)\) be the
bounded extension operator from
Lemma~\ref{lem:holder-survivor-extension}.  In particular
\(\mathcal R\mathcal J_H=\mathrm{id}\), so \(\mathcal R\) is onto.  Moreover
\[
  \mathcal R\mathcal L_{\psi,H}
  =
  \mathcal L_{\psi|Z_H}\mathcal R,
\]
where \(\mathcal L_{\psi|Z_H}\) is the ordinary Ruelle operator on the
closed survivor subshift.  Indeed, if \(x\in Z_H\), every safe preimage
of \(x\) is again in \(Z_H\); a preimage from a transient safe symbol
would give that symbol an infinite safe continuation.  The classical
Ruelle--Perron--Frobenius theorem on the mixing SFT \(Z_H\) gives a
simple isolated eigenvalue
\(\lambda_H=e^{P_{Z_H}(\psi)}\) for
\(\mathcal L_{\psi|Z_H}\), with the rest of its spectrum contained in a
disk of radius \(r_H<\lambda_H\)
\citep{Bowen2008,Ruelle1978,Walters1975GMeasures}.

It remains to check that the part of the open operator acting on
\(\mathcal E\) has smaller spectral radius.  Let
\(\mathcal S_\infty\) be the survivor alphabet and
\(\mathcal T=\mathcal S\setminus\mathcal S_\infty\).  As in the proof of
Lemma~\ref{lem:holder-safe-word-pressure}, there is no safe edge from
\(\mathcal T\) to \(\mathcal S_\infty\), and the graph on
\(\mathcal T\) is acyclic.  Hence there is \(K\) such that every
sufficiently long safe word is a word over \(\mathcal S_\infty\) followed
by a transient suffix of length at most \(K\).

For later use set
\[
  W_n(x):=
  \sum_{\substack{\sigma^n y=x\\ y_0,\ldots,y_{n-1}\in\mathcal S}}
  e^{S_n\psi(y)} .
\]
The finite-suffix decomposition just recalled, bounded distortion, and
the Ruelle--Perron--Frobenius bound for the survivor system give a
constant \(C_W<\infty\) such that
\[
  W_n(x)\leq C_W\lambda_H^n
  \qquad (x\in Z,\ n\geq1).
\]
Indeed, for \(n>K\), decompose any safe \(n\)-prefix \(u\) as \(u=sv\),
where \(s\) is a word in the survivor alphabet and \(v\) is a transient
suffix of length \(k\leq K\).  For fixed \(v\), the contribution of the
last \(k\) symbols is bounded by \(e^{K\norm{\psi}_\infty}\).  Since
\([s]\cap Z_H\neq\emptyset\), bounded distortion compares the first
\(n-k\) iterates along \(svx\) with
\(\sup_{[s]\cap Z_H}S_{n-k}\psi\), up to a uniform multiplicative
constant.  Summing over the finitely many possible suffixes \(v\) gives
an upper bound by a constant times
\[
  \sum_{k=0}^{K} Z_{n-k}^H(\psi),
\]
where \(Z_N^H(\psi)\) is the survivor partition sum from
Lemma~\ref{lem:holder-safe-word-pressure}.  The
Ruelle--Perron--Frobenius theorem on the mixing survivor subshift gives
\(Z_N^H(\psi)\leq C\lambda_H^N\).  This proves the displayed bound for
\(n>K\), and increasing \(C_W\) covers the finitely many smaller \(n\).

Take \(f\in\mathcal E\).  If \(y\) occurs in an \(n\)-step summand of
\(\mathcal L_{\psi,H}^n f(x)\), then \(y_0,\ldots,y_{n-1}\) are safe.
For \(n>K\), the first \(n-K\) symbols of \(y\) form a survivor word.
Choose \(\bar y\in Z_H\) with the same first \(n-K\) symbols.  Since
\(f|_{Z_H}=0\),
\[
  \abs{f(y)}
  =
  \abs{f(y)-f(\bar y)}
  \leq
  \norm{f}_\theta\,\theta^{n-K}.
\]
Together with the preceding bound on \(W_n(x)\), this gives
\[
  \norm{\mathcal L_{\psi,H}^n f}_\infty
  \leq
  C_2\theta^n e^{nP_{Z_H}(\psi)}\norm{f}_\theta .
\]

The H\"older seminorm satisfies the same bound.  Suppose first that
\(x,x'\) agree for \(N\geq p\) symbols, so \(d_\theta(x,x')=\theta^N\).
Pair the \(n\)-step safe preimages by the same safe prefix:
\(y=ux\) and \(y'=ux'\), \(\abs u=n\).  Then \(y\) and \(y'\) agree for
\(n+N\) symbols.  Hence
\[
  \abs{f(y)-f(y')}
  \leq
  \abs{f}_\theta\,\theta^{n+N}
  =
  \abs{f}_\theta\,\theta^n d_\theta(x,x').
\]
Also, by the previous comparison with \(Z_H\),
\[
  \abs{f(y')} \leq \norm{f}_\theta\theta^{n-K}.
\]
Finally the Bowen distortion estimate along the paired \(n\)-blocks gives
\[
  \abs{S_n\psi(y)-S_n\psi(y')}
  \leq
  C_\psi d_\theta(x,x'),
\]
and therefore
\[
  \abs{e^{S_n\psi(y)}-e^{S_n\psi(y')}}
  \leq
  C_\psi e^{S_n\psi(y)}d_\theta(x,x')
\]
after increasing \(C_\psi\).  For each paired summand,
\[
\begin{aligned}
 &\abs{e^{S_n\psi(y)}f(y)-e^{S_n\psi(y')}f(y')}        \\
 &\qquad\leq
  C e^{S_n\psi(y)}\norm{f}_\theta
  \theta^n d_\theta(x,x').
\end{aligned}
\]
Summing over safe prefixes and using \(W_n(x)\leq C_W\lambda_H^n\)
therefore yields
\[
  \abs{\mathcal L_{\psi,H}^n f}_\theta
  \leq
  C_3\theta^n e^{nP_{Z_H}(\psi)}\norm{f}_\theta .
\]
If \(N<p\), the same estimate follows from the sup-norm bound divided by
the lower bound \(d_\theta(x,x')\geq\theta^{p-1}\).  Thus the displayed
seminorm estimate holds for all \(x\neq x'\).
Thus
\[
  \rho\bigl(\mathcal L_{\psi,H}|_{\mathcal E}\bigr)
  \leq
  \theta e^{P_{Z_H}(\psi)}
  < \lambda_H .
\]

Now apply Lemma~\ref{lem:holder-quotient-spectrum} with
\[
  X=\mathcal B_\theta(Z),\qquad
  Y=\mathcal B_\theta(Z_H),\qquad
  T=\mathcal L_{\psi,H},\qquad
  S=\mathcal L_{\psi|Z_H},\qquad
  R=\mathcal R,\qquad
  J=\mathcal J_H .
\]
Since \(\rho(T_{\mathcal E})<\lambda_H\), the number \(\lambda_H\) does
not belong to \(\spec(T_{\mathcal E})\).  The survivor
Ruelle--Perron--Frobenius theorem gives that \(\lambda_H\) is a simple
isolated eigenvalue of \(S\), with the rest of \(\spec(S)\) contained in
the disk \(\{z:\abs z\leq r_H\}\).  Therefore
Lemma~\ref{lem:holder-quotient-spectrum}(iii) shows that every spectral
value of \(T\) outside the disk of radius
\(\max\{\rho(T_{\mathcal E}),r_H\}\) comes from \(S\), and
Lemma~\ref{lem:holder-quotient-spectrum}(iv) gives that \(\lambda_H\) is
an isolated eigenvalue of \(T\) of algebraic multiplicity one.

Let \(h_H^0\in\mathcal B_\theta(Z_H)\) be a non-zero eigenfunction of
\(S\) with \(Sh_H^0=\lambda_Hh_H^0\).  The lifting formula in
Lemma~\ref{lem:holder-quotient-spectrum}(iv) gives an eigenfunction
\[
  h_H
  :=
  \mathcal J_Hh_H^0
  +
  (T_{\mathcal E}-\lambda_H)^{-1}
  (\lambda_H-T)\mathcal J_Hh_H^0
\]
for \(T\), and \(\mathcal R h_H=h_H^0\), so \(h_H\neq0\).  If
\(\eta_H^0\) is the positive
Ruelle--Perron--Frobenius eigenfunctional on
\(\mathcal B_\theta(Z_H)\), then
\[
  \eta_H(f):=\eta_H^0(\mathcal Rf)
\]
is a non-zero positive eigenfunctional for \(T^*\).  Normalizing
\(h_H\) so that \(\eta_H(h_H)=1\), the operator
\[
  \Pi_H f:=\eta_H(f)h_H
\]
is the rank-one spectral projection of \(T\) at \(\lambda_H\), and
\(T\Pi_H=\Pi_H T=\lambda_H\Pi_H\).  Setting
\[
  \mathcal N_H:=T(I-\Pi_H),
\]
the spectrum of \(\mathcal N_H\) is the spectrum of \(T\) with the
simple eigenvalue \(\lambda_H\) removed, so
\(\rho(\mathcal N_H)<\lambda_H\), and the displayed spectral
decomposition follows.
\end{proof}

\begin{lemma}[One-step Gibbs error-survival comparison]
\label{lem:holder-one-step-error-comparison}
Let \(Z\) be a topologically mixing one-sided one-step subshift of finite
type, let \(\psi\) be a H\"older potential, and let \(\nu_\psi\) be its
equilibrium state.  Let \(H\) be a non-empty proper one-step Markov hole
with safe alphabet \(\mathcal S\).  Assume that the survivor subshift
\[
  Z_H:=\{z\in Z:z_n\in\mathcal S\text{ for all }n\geq0\}
\]
is topologically mixing, and assume the finite two-terminal residue
reachability condition from
Lemma~\ref{lem:holder-extension-ambiguity}.  For
\[
  \tau_H(z):=\inf\{n\geq0:z_n\notin\mathcal S\},
  \qquad
  \mathsf P_R:=\{z:\tau_H(z)<\infty,\ \tau_H(z)\in R\bmod q\},
\]
there is \(\theta_0>0\) such that, for every \(n\geq1\),
\[
  \theta_0\,\nu_{\psi}(\tau_H\geq n)
  \leq
  \varepsilon_n(\mathsf P_R;\nu_{\psi})
  \leq
  \frac12\,\nu_{\psi}(\tau_H\geq n).
\]
Consequently
\[
  \lim_{n\to\infty}
  -\frac1n\log\varepsilon_n(\mathsf P_R;\nu_\psi)
  =
  P_Z(\psi)-P_{Z_H}(\psi).
\]
\end{lemma}

\begin{proof}
Lemma~\ref{lem:holder-extension-ambiguity} gives a constant \(c_0>0\),
uniform in the safe word \(w\), such that for every admissible safe
length-\(n\) word \(w\),
\[
  c_0
  \leq
  \nu_\psi(\mathsf P_R\mid[w])
  \leq
  1-c_0 .
\]
Hence
\[
  c_0
  \leq
  \min\{
    \nu_\psi(\mathsf P_R\mid[w]),
    1-\nu_\psi(\mathsf P_R\mid[w])
  \}
  \leq \frac12 .
\]
The safe-prefix decomposition
Lemma~\ref{lem:holder-safe-prefix-bayes-decomposition} therefore gives
\[
  \varepsilon_n(\mathsf P_R;\nu_\psi)
  =
  \sum_{\substack{w\in\mathcal L_n(Z)\\
                  w_0,\ldots,w_{n-1}\in\mathcal S}}
    \nu_\psi([w])\,
    \min\{
      \nu_\psi(\mathsf P_R\mid[w]),
      1-\nu_\psi(\mathsf P_R\mid[w])
    \}.
\]
The sum of the cylinder weights over safe words is
\(\nu_\psi(\tau_H\geq n)\).  Taking \(\theta_0=c_0\) proves the
two-sided comparison.

Lemma~\ref{lem:holder-safe-word-pressure} gives
\[
  \lim_{n\to\infty}
  -\frac1n\log\nu_\psi(\tau_H\geq n)
  =
  P_Z(\psi)-P_{Z_H}(\psi),
\]
and the comparison transfers the same exponential rate to
\(\varepsilon_n(\mathsf P_R;\nu_\psi)\).
\end{proof}

\begin{lemma}[Finite-cylinder block reduction]
\label{lem:holder-block-reduction}
Let \(Y\) be a topologically mixing one-sided subshift of finite type,
let \(\varphi\) be H\"older, and let \(\nu_\varphi\) be its equilibrium
state.  Let \(p_Y\geq1\) be a Markov memory for \(Y\), meaning that
any word of length at least \(p_Y+1\) is admissible if and only if all
of its subwords of length \(p_Y+1\) are admissible.  Suppose that
\(H\subset Y\) is a finite union of cylinders and choose
\(\ell\geq p_Y\) such that \(H\) is a union of \(\ell\)-cylinders.
Let
\[
  \Pi_\ell:Y\longrightarrow Y^{(\ell)}
\]
be the \(\ell\)-block conjugacy, let
\(\psi:=\varphi\circ\Pi_\ell^{-1}\), and let \(H^{(\ell)}\) be the
one-step hole in \(Y^{(\ell)}\) whose symbols are the \(\ell\)-blocks
contained in \(H\).  Then \(Y^{(\ell)}\) is a topologically mixing
one-step subshift of finite type, and:
\begin{enumerate}[label=(\roman*)]
  \item \((\Pi_\ell)_\ast\nu_\varphi\) is the equilibrium state
    \(\nu_\psi\) of \(\psi\) on \(Y^{(\ell)}\);
  \item \(P_{Y^{(\ell)}}(\psi)=P_Y(\varphi)\), and the survivor subshift
    of \(H^{(\ell)}\) is conjugate to \(Y_H\), so its pressure is
    \(P_{Y_H}(\varphi)\);
  \item if \(m\geq\ell\) and \(n=m-\ell+1\), then the
    length-\(m\) cylinder \(\sigma\)-algebra on \(Y\) is the pullback of
    the length-\(n\) cylinder \(\sigma\)-algebra on \(Y^{(\ell)}\);
  \item first entry into \(H\) at time \(t\) is exactly first entry into
    \(H^{(\ell)}\) at block time \(t\).  Consequently first-entry residue
    events correspond under \(\Pi_\ell\), and their depth-\(m\) and
    depth-\(n\) Bayes errors are equal.
\end{enumerate}
\end{lemma}

\begin{proof}
The map \(\Pi_\ell\) sends \(y\) to the sequence of overlapping blocks
\((y_t\cdots y_{t+\ell-1})_{t\geq0}\).  This is a topological conjugacy
from \(Y\) to the usual \(\ell\)-block presentation.  Since
\(\ell\geq p_Y\), an allowable sequence of overlapping \(\ell\)-blocks
is admissible exactly when each adjacent pair of block symbols is
admissible in \(Y\).  Hence \(Y^{(\ell)}\) is a one-step SFT, and it is
topologically mixing because it is conjugate to \(Y\).
The variational principle is invariant under conjugacy, so the
pushforward of the equilibrium state of \(\varphi\) is the equilibrium
state of \(\psi=\varphi\circ\Pi_\ell^{-1}\), and the two ambient
pressures agree.
The same conjugacy sends points that avoid \(H\) at every time precisely
to points whose block symbols avoid \(H^{(\ell)}\) at every time, giving
the pressure identity on the survivor subsystem.

A length-\(m\) word \(y_0\cdots y_{m-1}\), with \(m\geq\ell\), determines
exactly the \(n=m-\ell+1\) consecutive \(\ell\)-blocks
\[
  y_0\cdots y_{\ell-1},\;
  y_1\cdots y_\ell,\;\ldots,\;
  y_{m-\ell}\cdots y_{m-1},
\]
and the admissibility constraints in the block presentation recover the
original word from those blocks.  Thus the finite observation
\(\sigma\)-algebras coincide under pullback.  Finally,
\(\sigma^t y\in H\) if and only if the \(t\)-th block of
\(\Pi_\ell y\) belongs to \(H^{(\ell)}\).  The first-entry time and its
residue are therefore preserved, and the Bayes errors computed on the
corresponding finite \(\sigma\)-algebras are identical.
\end{proof}

\begin{theorem}[Block-coded H\"older operator and Gibbs escape comparison]
\label{thm:holder-finite-cylinder-escape}
Let \(Y\) be a topologically mixing one-sided subshift of finite type,
let \(\varphi\) be a H\"older potential, and let \(\nu_\varphi\) be its
equilibrium state.  Let \(H\subset Y\) be a non-empty proper finite union
of cylinders, and define
\[
  Y_H:=\{y\in Y:\sigma^n y\notin H\text{ for every }n\geq0\}.
\]
Fix \(q\geq 2\) and a non-empty proper subset
\(R\subsetneq \ZZ/q\ZZ\), and set
\[
  \tau_H(y):=\inf\{n\geq0:\sigma^n y\in H\},
  \qquad
  \mathsf P_R:=\{y:\tau_H(y)<\infty,\ \tau_H(y)\in R\bmod q\}.
\]
Let \(p_Y\geq1\) be a Markov memory for \(Y\), and choose
\(\ell\geq p_Y\) so that \(H\) is a union of \(\ell\)-cylinders.  Let
\((Y^{(\ell)},\sigma^{(\ell)})\) be the resulting one-step
\(\ell\)-block presentation of \(Y\), let \(\psi\) be the lifted
H\"older potential, and let \(H^{(\ell)}\) be the corresponding one-step
hole.  Let
\(\mathcal S\) be the safe block alphabet.

Assume the following two finite hypotheses.
\begin{enumerate}[label=(G\arabic*)]
  \item The survivor subshift of \(H^{(\ell)}\) is topologically mixing.
  \item In the directed graph with vertices
\[
  \mathcal S\times\ZZ/q\ZZ,\qquad \star_R,\quad\star_{R^c},
\]
with safe edges
\((i,r)\to(j,r+1)\) for admissible safe transitions \(i\to j\), and with
terminal edges \((i,r)\to\star_R\) or \((i,r)\to\star_{R^c}\) according
as an admissible next block enters the hole with residue \(r+1\) in
\(R\) or in \(R^c\), both terminal vertices are reachable from every
\((i,r)\in\mathcal S\times\ZZ/q\ZZ\).
\end{enumerate}

Choose a symbolic metric \(d_\theta\) on \(Y^{(\ell)}\) for which
\(\psi\) is Lipschitz, and let
\[
  (\mathcal L_{\psi,H^{(\ell)}}f)(x)
  :=
  \sum_{\substack{\sigma^{(\ell)} y=x\\ y_0\in\mathcal S}}
    e^{\psi(y)}f(y)
\]
be the open Ruelle operator on
\(\mathcal B_\theta(Y^{(\ell)})=C^\alpha(Y^{(\ell)})\).  Then
\(\mathcal L_{\psi,H^{(\ell)}}\) has the spectral decomposition from
Theorem~\ref{thm:holder-open-transfer}, with leading eigenvalue
\[
  \lambda_H
  =
  e^{P_{(Y^{(\ell)})_{H^{(\ell)}}}(\psi)}
  =
  e^{P_{Y_H}(\varphi)}.
\]

For the first-entry residue event \(\mathsf P_R^{(\ell)}\) in the block
system, let
\(\varepsilon_n^{(\ell)}(\mathsf P_R^{(\ell)};\nu_{\psi})\) denote the
Bayes error computed from block prefixes of length \(n\).  Then, for
every \(m\geq\ell\), with \(n=m-\ell+1\),
\[
  \varepsilon_m(\mathsf P_R;\nu_\varphi)
  =
  \varepsilon_n^{(\ell)}
    (\mathsf P_R^{(\ell)};\nu_{\psi}).
\]
There is \(\theta_0>0\), depending only on \(Y,\varphi,H,q,R\) and the
chosen block presentation, such that, for every \(n\geq1\),
\[
  \theta_0\,\nu_{\psi}(\tau_{H^{(\ell)}}\geq n)
  \leq
  \varepsilon_n^{(\ell)}
    (\mathsf P_R^{(\ell)};\nu_{\psi})
  \leq
  \frac12\,\nu_{\psi}(\tau_{H^{(\ell)}}\geq n).
\]
Consequently
\[
  \lim_{m\to\infty}
  -\frac1m\log
  \varepsilon_m(\mathsf P_R;\nu_\varphi)
  =
  P_Y(\varphi)-P_{Y_H}(\varphi).
\]
The operator assertion is the explicit \(C^\alpha\)-space statement
above applied to the block presentation.  The finite-observation
assertion is only the finite-prefix comparison and escape exponent; it
does not claim a Banach-space cyclotomic resolvent.
\end{theorem}

\begin{proof}
The equality of Bayes errors and the pressure identifications under
block recoding are exactly Lemma~\ref{lem:holder-block-reduction}.  The
operator statement is Theorem~\ref{thm:holder-open-transfer} applied to
the block system \(Y^{(\ell)}\), the H\"older potential \(\psi\), and the
one-step hole \(H^{(\ell)}\); the block pressure identity in
Lemma~\ref{lem:holder-block-reduction} gives the displayed value of
\(\lambda_H\).

Apply Lemma~\ref{lem:holder-one-step-error-comparison} to the block
system \(Y^{(\ell)}\), the potential \(\psi\), and the one-step hole
\(H^{(\ell)}\).  Its two-sided comparison is exactly the displayed
comparison for
\(\varepsilon_n^{(\ell)}(\mathsf P_R^{(\ell)};\nu_\psi)\), and it also
gives
\[
  \lim_{n\to\infty}
  -\frac1n\log
  \varepsilon_n^{(\ell)}(\mathsf P_R^{(\ell)};\nu_\psi)
  =
  P_{Y^{(\ell)}}(\psi)
  -P_{(Y^{(\ell)})_{H^{(\ell)}}}(\psi).
\]
The fixed shift \(n=m-\ell+1\) does not affect the exponential rate.
Block conjugacy preserves pressure and identifies
\((Y^{(\ell)})_{H^{(\ell)}}\) with \(Y_H\).  Hence the last displayed
quantity is \(P_Y(\varphi)-P_{Y_H}(\varphi)\), and the block-recoding
equality gives the claimed limit for
\(\varepsilon_m(\mathsf P_R;\nu_\varphi)\).
\end{proof}
