\documentclass[../../../main.tex]{subfiles}

\begin{document}
\input{sec__spg}
\end{document}

% --- Backflow from publication papers (auto-generated) ---
\begin{remark}[Results proven in dedicated publications]
  The following results from this section have been developed into
  standalone publications with full proofs and independent verification:
  \begin{itemize}
    \item \textbf{Exact cylinder formula} (Proposition \texttt{prop:scan-error-cylinder}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Bayes rule and functional stability} (Proposition \texttt{prop:bayes-optimality}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Tanaka identity} (Theorem \texttt{thm:tanaka-stokes}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{cor:clarity-monotone} (Corollary \texttt{cor:clarity-monotone}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Weighted boundary decomposition} (Proposition \texttt{prop:weighted-boundary-decomposition}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Upper boundary estimate} (Theorem \texttt{thm:clarity-boundary-dimension}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Weighted boundary transfer} (Theorem \texttt{thm:weighted-boundary-transfer}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Weighted boundary pole law} (Corollary \texttt{cor:weighted-boundary-pole}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Pressure gap formula} (Corollary \texttt{cor:pressure-gap}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Survivor R\'enyi pressure spectrum} (Theorem \texttt{thm:survivor-renyi-pressure}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Collision consequences of the survivor spectrum} (Corollary \texttt{cor:survivor-spectrum-collision}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Poisson threshold for collision failure} (Theorem \texttt{thm:double-budget-poisson}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Subcritical capacity region} (Corollary \texttt{cor:capacity-region}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Entropy ledger for finite refinements} (Theorem \texttt{thm:entropy-ledger}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Register lower bound} (Corollary \texttt{thm:register-lower-bound}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Injective refinements restore the microscopic collision profile} (Theorem \texttt{thm:collision-profile-restoration}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Quasistationary ambiguity amplitudes} (Theorem \texttt{thm:main-amplitude}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Survivor R\'enyi pressure spectrum} (Theorem \texttt{thm:main-renyi}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Cyclotomic resonance lift} (Theorem \texttt{thm:main-resonance}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{First-entry residue events and escape rates} (Theorem \texttt{thm:first-entry-escape-rate}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Survivor pressure recovery} (Corollary \texttt{cor:survivor-pressure-recovery}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Quasistationary ambiguity amplitudes} (Theorem \texttt{thm:quasistationary-ambiguity}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Error resolvent and cyclotomic resonance lift} (Theorem \texttt{thm:error-resolvent-cyclotomic}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{H\"older Gibbs cyclotomic lift after block recoding} (Theorem \texttt{thm:holder-open-resonance}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Doubling map and the Fibonacci escape law} (Proposition \texttt{prop:doubling-fibonacci-escape}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Symbolic factor} (Proposition \texttt{prop:symbolic-factor}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Each recursive step is a sliding-block factor} (Proposition \texttt{prop:sliding-block-factor}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{\(\sigma\)-algebra non-expansion} (Corollary \texttt{cor:sigma-nonexpansion}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Entropy monotonicity} (Corollary \texttt{cor:entropy-monotonicity}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Finite-depth collapse theorem} (Theorem \texttt{thm:finite-depth-collapse}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Inverse-limit address space} (Proposition \texttt{prop:inverse-limit-space}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Stable address determinacy on the visible factor} (Theorem \texttt{thm:inverse-limit-determinacy}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Complete address reconstruction} (Theorem \texttt{thm:complete-address-reconstruction}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Clopen decidability} (Proposition \texttt{prop:decidable-clopen}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
    \item \textbf{Sturmian bounds} (Proposition \texttt{prop:sturmian-cylinder-sandwich}) --- Ergodic Theory and Dynamical Systems paper \texttt{scan_projection_address}, status: ACCEPT.
  \end{itemize}
\end{remark}
% --- End backflow ---
