\begin{abstract}
We study the finite-window Fibonacci normalization which sends a binary word of length $m$ to the unique golden-mean admissible word with the same Fibonacci valuation modulo $F_{m+2}$. The map is usually many-to-one on a single window, but its sliding overlap extension has a sharp threshold. We prove that for every $m\ge3$ it is a finite-memory conjugacy from the full two-shift onto a subshift of finite type, with an explicit decoder. For $m=2$ the only nontrivial path-space fiber is the pair of constant sequences; this yields exact conditional-entropy and Kullback--Leibler defect formulas and a mod-zero inverse off the branch. The proof combines a bi-resolving overlap presentation with a residue decoder, so the degree-one conclusion is an arithmetic fact rather than a formal consequence of resolving theory. We also show that the golden case is exceptional among metallic finite-window systems: every metallic parameter $A\ge2$ is already invertible at window length two. Above threshold the overlap presentation is identified with the right Fischer cover. Further residue, Fourier, second-moment and Sturmian discrepancy estimates record the associated finite-window statistics.
\end{abstract}

\medskip
\noindent{\bf 2020 Mathematics Subject Classification.}
Primary: 37B10, 37B05; Secondary: 11B39, 37A35, 37D35

\medskip
\noindent{\bf Key words and phrases.}
Fibonacci stabilization, symbolic dynamics, finite-memory conjugacy, branch locus, Ostrowski numeration, Fischer cover, Zeckendorf representation, sofic shift
