\section{Introduction}\label{sec:intro}

Finite-window recodings arise naturally in symbolic dynamics, numeration, and finite-resolution observation. A local map may be many-to-one on isolated words and yet become invertible once one passes to overlapping windows along a trajectory. The problem studied here is whether this recovery-by-overlap occurs for the canonical Fibonacci stabilization map, and if so at what window size.

For each $m\ge 1$ we consider
\[
\Fold_m:\{0,1\}^m\to X_m,
\]
the map that sends a binary word to the unique Fibonacci-admissible word with the same valuation modulo $F_{m+2}$. On individual words $\Fold_m$ is genuinely many-to-one: different binary blocks can lie in the same residue class and therefore stabilize to the same admissible block. The central question is whether this local folding persists on path space for the associated sliding-block map, or whether overlap restores the lost information.

The answer is sharp. The main contributions of the paper are:
\begin{enumerate}[label=(\roman*),leftmargin=*,itemsep=2pt]
  \item The overlap graph yields a finite-state lift with bounded degree, and the threshold window is exactly $m=3$: Lemma~\ref{lem:m3-decoder} gives an explicit three-window decoder and Theorem~\ref{thm:block-invertibility} proves injectivity for every $m\ge 3$.
  \item The exceptional case $m=2$ is completely rigid rather than generically non-injective. Theorem~\ref{thm:m2-branch-locus} identifies a unique branch orbit, Corollary~\ref{cor:m2-defect-formulas} gives exact entropy and Kullback--Leibler defect formulas, and Corollary~\ref{cor:m2-mod-zero} shows mod-zero invertibility off that branch set.
  \item Above threshold, the stabilized image is a finite-type recoding of the full two-shift. Theorem~\ref{thm:two-sided-conjugacy} supplies a finite-memory inverse and identifies the image shift $Y_m$ as a subshift of finite type.
  \item The conjugacy has exact dynamical consequences. Proposition~\ref{prop:conjugacy-consequences} and Theorem~\ref{thm:continuous-potentials} transport invariant measures, entropy, continuous potentials, pressure, and equilibrium states exactly, while Section~\ref{sec:fp:tv} gives quantitative discrepancy-based corollaries in the Sturmian benchmark.
\end{enumerate}

This places the paper closer to symbolic dynamics and arithmetic coding than to a measurement-theory manuscript. The cut-and-project and detector language serve only to fix the benchmark source law and to explain why the local problem must be posed on the full cube $\{0,1\}^m$. Once the source law and admissible grammar are fixed, the proofs are symbolic and arithmetic.

The relevant literature splits into four strands. Classical work on Sturmian words and Fibonacci numeration explains admissibility, normalization, and automaton structure at the level of words \cite{MorseHedlund1940,Lothaire2002,Queffelec2010,Frougny1992,Berstel2001}; the broader interaction between numeration systems and symbolic dynamics is surveyed in \cite{BertheRigo2010}, and $S$-adic generalizations appear in \cite{BertheSteinerThuswaldner2019}. General sofic-factor theory provides the language of labeled graphs, resolving maps, and finite-to-one factors \cite{LindMarcus1995,Kitchens1998,Nasu1995}; within that framework, resolving factor maps between irreducible shifts of finite type with equal entropy are known to be conjugacies under general conditions \cite{Ashley1991}, and lower-entropy sofic factors have been characterized abstractly \cite{Boyle1983}. The connection between $\beta$-expansions and shift spaces was introduced by Parry~\cite{Parry1960}. Standard thermodynamic formalism explains what must be preserved under topological conjugacy \cite{Walters1982}. What these frameworks do not provide on their own is the arithmetic threshold phenomenon proved here: the explicit fact that $m=2$ fails in exactly one branch locus, whereas every $m\ge 3$ admits a finite-memory inverse. That conclusion depends on the residue arithmetic of Fibonacci stabilization and, in particular, on the decoder of Lemma~\ref{lem:m3-decoder}; it is not a formal consequence of the general sofic setup.

\subsection{Theorem-by-theorem comparison with prior work}

The statements now labelled as supporting propositions are not presented as theorem-level novelties. Proposition~\ref{thm:noise-full-support} is the elementary support consequence of a strictly positive finite channel in the sense of standard information theory \cite{CoverThomas2006}; Propositions~\ref{thm:fold-suite} and~\ref{thm:fiber-residue} unpack Zeckendorf uniqueness and the finite-window normalization viewpoint developed in Fibonacci numeration \cite{Lothaire2002,Frougny1992,Berstel2001}; and Proposition~\ref{thm:kl-ledger} is the entropy chain rule plus the KL projection onto a fiber-uniform lift \cite{CoverThomas2006}. Their role is to fix notation and bookkeeping, not to claim priority.

The new finite-window arithmetic starts with Theorem~\ref{thm:fiber-dft}, which applies the classical roots-of-unity filter \cite{Stanley2012} to the specific modulo-$F_{m+2}$ Fibonacci fibers and records the resulting second-moment formula. Lemma~\ref{lem:cross-multiplication} is the standard d'Ocagne identity \cite{Koshy2001}, while Theorem~\ref{thm:affine-transfer} is the new use of that identity: it gives an explicit affine permutation identifying the canonical and shifted-weight fiber spectra. Theorem~\ref{prop:second-moment-recurrence} is also new relative to prior work on Fibonacci representations \cite{Berstel2001}, because it gives a closed recurrence for the cyclic residue collision statistic of this particular stabilization map.

The central dynamical increment is not supplied by general sofic-factor theory. Hedlund's sliding-block formalism \cite{Hedlund1969} and the standard theory of labeled graph presentations and resolving maps \cite{LindMarcus1995,Kitchens1998,Nasu1995,Ashley1991,Boyle1983} explain the ambient language, but Theorem~\ref{thm:finite-state-lift} adds the specific bi-resolving overlap presentation for $\Fold_m$, Lemma~\ref{lem:m3-decoder} gives the explicit three-window arithmetic decoder, Theorem~\ref{thm:block-invertibility} proves the sharp injectivity threshold $m=3$, and Theorem~\ref{thm:m2-branch-locus} proves that the two-window obstruction is exactly one branch orbit rather than a generic sofic ambiguity.

The finite-memory consequences should be read as conditional on that threshold result. Theorem~\ref{thm:two-sided-conjugacy} upgrades injectivity to a concrete finite-memory conjugacy and finite-type image; Theorem~\ref{thm:exact-sufficiency} is the corresponding block-experiment equivalence in the sense of Blackwell comparison \cite{Blackwell1953}; and Theorem~\ref{thm:continuous-potentials} is the standard transport of entropy, pressure, and equilibrium states through a topological conjugacy \cite{Walters1982}. Corollaries on phase diagrams, finite-range energetics, occupancy, and block-level information geometry similarly add the explicit local formulas for this Fibonacci readout, with the information-geometric invariance resting on classical divergence theory \cite{AliSilvey1966}.

The quantitative Sturmian results are downstream certificates, not new discrepancy theory. Theorem~\ref{thm:tv-certificate} combines standard rotation discrepancy estimates \cite{DrmotaTichy1997} with the deterministic stabilized-block pushforward to obtain the explicit $(m+n)$ total-variation bound, and Theorem~\ref{thm:operational-observable-recovery} combines that bound with the finite-memory inverse to recover every finite-range observable from the stabilized trajectory with explicit discrepancy and boundary-layer error terms.

A second conceptual point is that fixed-window aliasing and path-space information loss are different objects. The entropy ledger of Section~\ref{sec:entropy-ledger} measures the unresolved multiplicity of isolated windows. Above threshold this does not become a new source of dynamical irreversibility, because Section~\ref{sec:measurement} proves that overlap reconstructs the raw trajectory exactly. The local defect at $m=2$ is therefore a sharp exceptional branch, not evidence of a generic extensive loss mechanism.

The Fibonacci benchmark is a natural place to study this question. It is the smallest aperiodic binary system in which the admissible language, residue fibers, threshold decoder, and discrepancy bounds can all be written explicitly. This makes it possible to prove the path-space consequences at theorem level rather than to suggest them heuristically.

The original motivation nevertheless comes from finite-resolution observation of quasiperiodic order \cite{Shechtman1984,Jagannathan2021,BaakeGrimm2013}. In the Sturmian benchmark, any strictly positive local detector noise makes every binary word of length $m$ appear with positive probability, so the ambient cube is physically forced at the readout layer. That application explains why the local folding problem is worth studying, but it does not determine the mathematical identity of the paper.

The paper is organized as follows. Section~\ref{sec:readout} records the symbolic setup, the Sturmian benchmark, and the full-support mechanism. Section~\ref{sec:stabilization} defines Fibonacci stabilization on words, gives the residue description of the fibers, and includes a concrete worked example. Section~\ref{sec:stab:conjugacy} develops the overlap graph, the sharp three-window threshold, and the two-window branch locus. Section~\ref{sec:entropy-ledger} isolates the fixed-window aliasing ledger, and Section~\ref{sec:measurement} turns the threshold theorem into finite-memory conjugacy, path-space equivalence, and transport of thermodynamic formalism. Sections~\ref{sec:fiber-spectrum} and~\ref{sec:fp:tv} record arithmetic fallout and quantitative discrepancy bounds in the Sturmian benchmark. Appendix~\ref{sec:superspace} contains the source-layer model-set input.
