\section{Consequences of the finite-memory conjugacy}\label{sec:measurement}

Theorem~\ref{thm:two-sided-conjugacy} turns the threshold theorem into a conjugacy statement.  This section records the standard consequences for finite blocks, path-space laws and thermodynamic formalism, with the hypotheses stated explicitly.

\subsection{Block presentations above threshold}\label{sec:measurement:sufficiency}

For $m\ge 3$, the finite-block map is a bijection onto its image once the output length is at least $m$.  We record this in a form which is useful for comparing statistical experiments.

Fix $m\ge 3$ and $n\ge m$. Write
\[
\mathcal Y_{m,n}:=\Phi_{m,n}\bigl(\{0,1\}^{n+m-1}\bigr)\subset X_m^n
\]
for the legal length-$n$ stabilized output blocks, and let
\[
\Psi_{m,n}:\mathcal Y_{m,n}\to\{0,1\}^{n+m-1}
\]
denote the inverse of $\Phi_{m,n}$ from Theorem~\ref{thm:block-invertibility}.

\begin{theorem}[Equivalence of block presentations above threshold]\label{thm:exact-sufficiency}
Fix $m\ge 3$ and $n\ge m$. Let $\{P_u:u\in\mathcal U\}$ be any family of probability laws on the raw block space $\{0,1\}^{n+m-1}$, and let
\[
Q_u:=(\Phi_{m,n})_*P_u
\]
be the induced family on $\mathcal Y_{m,n}$. Then:
\begin{enumerate}[label=(\roman*),leftmargin=*,itemsep=2pt]
  \item For every $u\in\mathcal U$, if $A^{(n)}\sim P_u$ and $Y^{(n)}:=\Phi_{m,n}(A^{(n)})$, then
  \[
  A^{(n)}=\Psi_{m,n}(Y^{(n)})\qquad P_u\text{-a.s.}
  \]
  Consequently,
  \[
  H_{P_u}(A^{(n)}\mid Y^{(n)})=0.
  \]
  \item For any random parameter $U$ with conditional law $A^{(n)}\mid\{U=u\}\sim P_u$,
  \[
  I(U;Y^{(n)})=I(U;A^{(n)}).
  \]
  \item The raw-block and stabilized-block experiments are deterministically equivalent. In particular, for every $u,v\in\mathcal U$,
  \[
  D_{\KL}(P_u\|P_v)=D_{\KL}(Q_u\|Q_v),
  \qquad
  D_{\TV}(P_u,P_v)=D_{\TV}(Q_u,Q_v),
  \]
  the Chernoff information agrees, and every Bayesian or Neyman--Pearson decision problem has the same optimal risk under the two experiments.
\end{enumerate}
\end{theorem}

\begin{proof}
Theorem~\ref{thm:block-invertibility} states that $\Phi_{m,n}$ is injective, so $\Psi_{m,n}$ is well-defined on $\mathcal Y_{m,n}$ and satisfies
\[
\Psi_{m,n}(\Phi_{m,n}(a))=a
\qquad\text{for every }a\in\{0,1\}^{n+m-1}.
\]
Hence $A^{(n)}=\Psi_{m,n}(Y^{(n)})$ almost surely under every $P_u$, proving (i).

For (ii), $Y^{(n)}$ is a deterministic function of $A^{(n)}$, so by data processing
\[
I(U;Y^{(n)})\le I(U;A^{(n)}).
\]
Conversely, $A^{(n)}$ is a deterministic function of $Y^{(n)}$ by (i), so the reverse inequality also holds.

For (iii), each $Q_u$ is the pushforward of $P_u$ under a bijection between $\{0,1\}^{n+m-1}$ and $\mathcal Y_{m,n}$. Therefore any divergence or decision criterion that is invariant under deterministic bijections is identical for the two experiments. The displayed identities for $D_{\KL}$ and $D_{\TV}$ are obtained by rewriting the defining sums on $\mathcal Y_{m,n}$ via the inverse map $\Psi_{m,n}$. Chernoff information is preserved for the same reason, and a decision rule on one experiment transports to the other by composition with $\Phi_{m,n}$ or $\Psi_{m,n}$ without changing its loss distribution.
\end{proof}

\subsection{Path-space equivalence}

The same finite-memory inverse gives a measurable isomorphism of path-space experiments.  Consequently the many-to-one character of $\Fold_m$ on a single window does not change relative-entropy rates after overlap is imposed.

\begin{corollary}[Path-space equivalence and divergence-rate preservation]\label{cor:path-space-divergence}
Assume $m\ge 3$, let $\{\nu_u:u\in\mathcal U\}$ be a family of shift-invariant Borel probability measures on $\{0,1\}^{\mathbb Z}$, and write
\[
\eta_u:=(\Phi_m)_*\nu_u
\]
for their images on $Y_m$. Then $\Phi_m$ is a measurable isomorphism between the path-space experiments $(\{0,1\}^{\mathbb Z},\nu_u)$ and $(Y_m,\eta_u)$. In particular, for every $u,v\in\mathcal U$, the finite-block Kullback--Leibler divergences agree exactly after the block identification by $\Phi_{m,n}$, so the relative entropy rates satisfy
\[
h(\nu_u\|\nu_v)=h(\eta_u\|\eta_v)
\]
whenever these specific relative entropy rates exist.
\end{corollary}

\begin{proof}
The path-space isomorphism is exactly Theorem~\ref{thm:two-sided-conjugacy}(i)--(ii). For every $n\ge m$, the block map $\Phi_{m,n}$ identifies the length-$(n+m-1)$ raw marginals of $\nu_u$ and $\nu_v$ with the length-$n$ stabilized marginals of $\eta_u$ and $\eta_v$. The finite-block Kullback--Leibler divergences therefore coincide term by term. Normalizing by $n$ or by $n+m-1$ gives the same limiting rate, since $(n+m-1)/n\to1$. Letting $n\to\infty$ yields the equality of specific relative entropies whenever these limits exist.
\end{proof}

\begin{proposition}[Fixed-window and path-space information]\label{prop:irreversibility-localization}
Assume $m\ge 3$. Let $\nu$ be a shift-invariant Borel probability measure on $\{0,1\}^{\mathbb Z}$, let $\eta:=(\Phi_m)_*\nu$, and let $\mu_m$ and $\pi_m$ denote the corresponding length-$m$ marginals on $\{0,1\}^m$ and $X_m$. Then:
\begin{enumerate}[label=(\roman*),leftmargin=*,itemsep=2pt]
  \item the finite-window entropy decomposition of Proposition~\ref{thm:kl-ledger},
  \[
  H(\mu_m)=H(\pi_m)+\EE_{\pi_m}[\log d_m(X)]-D_{\KL}(\mu_m\|\mu_m^e),
  \]
  describes the unresolved multiplicity of isolated windows;
  \item the path-space systems $(\{0,1\}^{\mathbb Z},\nu)$ and $(Y_m,\eta)$ are measurably isomorphic.  Equivalently, if a second source law $\nu'$ is added and $\eta':=(\Phi_m)_*\nu'$, then the specific relative entropy rates $h(\nu\|\nu')$ and $h(\eta\|\eta')$ agree whenever both are defined.
\end{enumerate}
Consequently canonical stabilization above threshold does not change path-space relative entropy rates.
\end{proposition}

\begin{proof}
Item~(i) is Proposition~\ref{thm:kl-ledger}, formulated at fixed window length $m$. Item~(ii) is the single-process specialization of Corollary~\ref{cor:path-space-divergence}. The asserted conclusion follows because the path-space laws are measurably isomorphic.
\end{proof}

\subsection{Continuous potentials and finite-range interactions}\label{sec:measurement:renorm}

We next record the thermodynamic consequences of the conjugacy.  The first statement is for arbitrary continuous potentials.  The following finite-range formulas make the transported interaction explicit.

\begin{theorem}[Transport of continuous potentials under the conjugacy]\label{thm:continuous-potentials}
Assume $m\ge 3$, let $\varphi:\{0,1\}^{\mathbb Z}\to\RR$ be continuous, and define
\[
\widetilde\varphi:=\varphi\circ \Phi_m^{-1}:Y_m\to\RR.
\]
Then:
\begin{enumerate}[label=(\roman*),leftmargin=*,itemsep=2pt]
  \item for every shift-invariant Borel probability measure $\mu$ on $\{0,1\}^{\mathbb Z}$,
  \[
  \int_{Y_m}\widetilde\varphi\,d(\Phi_{m*}\mu)=\int \varphi\,d\mu,
  \qquad
  h_{\Phi_{m*}\mu}(\sigma)=h_\mu(\sigma);
  \]
  \item the topological pressures agree:
  \[
  P_{Y_m}(\widetilde\varphi)=P_{\{0,1\}^{\mathbb Z}}(\varphi);
  \]
  \item $\mu$ is an equilibrium state for $\varphi$ if and only if $\Phi_{m*}\mu$ is an equilibrium state for $\widetilde\varphi$.
\end{enumerate}
\end{theorem}

\begin{proof}
Theorem~\ref{thm:two-sided-conjugacy}(i) states that $\Phi_m$ is a topological conjugacy onto $Y_m$, so $\Phi_m^{-1}$ is continuous and $\widetilde\varphi$ is again continuous. The integral identity is immediate from the definition of pushforward. The entropy identity is Proposition~\ref{prop:measure-entropy-preservation}. Therefore for every invariant measure $\mu$,
\[
h_{\Phi_{m*}\mu}(\sigma)+\int \widetilde\varphi\,d(\Phi_{m*}\mu)
=
h_\mu(\sigma)+\int \varphi\,d\mu.
\]
Taking the supremum over invariant measures and using the variational principle on both systems yields the pressure identity. The equilibrium-state correspondence is then immediate from equality of the variational functional under pushforward.
\end{proof}

\begin{corollary}[Parameterized pressure functions]\label{cor:phase-diagram}
Assume $m\ge 3$, and let $\{\varphi_\lambda:\lambda\in\Lambda\}$ be any family of continuous potentials on the raw shift. Then for every parameter value $\lambda$,
\[
P_{Y_m}(\widetilde\varphi_\lambda)=P_{\{0,1\}^{\mathbb Z}}(\varphi_\lambda),
\qquad
\widetilde\varphi_\lambda:=\varphi_\lambda\circ\Phi_m^{-1}.
\]
Hence every property formulated solely in terms of the pressure function, and every equilibrium-state correspondence defined by the variational principle, is preserved pointwise after transporting the potential by the stabilized conjugacy.
\end{corollary}

\begin{proof}
Apply Theorem~\ref{thm:continuous-potentials} to each parameter value separately.
\end{proof}

\begin{proposition}[Finite-range energetics under local inversion]\label{prop:finite-range-renorm}
Assume $m\ge 3$. Let $\phi:\{0,1\}^r\to\mathbb R$ be a finite-range interaction of range $r\ge 1$ on the raw shift. Define the induced interaction on $Y_m$ by
\[
\widetilde{\phi}(y_t,\dots,y_{t+r+m-2})
:=
\phi\bigl(\psi_m(y_t,\dots,y_{t+m-1}),\dots,\psi_m(y_{t+r-1},\dots,y_{t+r+m-2})\bigr).
\]
Then $\widetilde{\phi}$ has range $r+m-1$ on $Y_m$, and for every $a\in\{0,1\}^{\mathbb Z}$, $y:=\Phi_m(a)$, and $n\ge 1$,
\[
\sum_{t=0}^{n-1}\phi(a_t,\dots,a_{t+r-1})
=
\sum_{t=0}^{n-1}\widetilde{\phi}(y_t,\dots,y_{t+r+m-2}).
\]
\end{proposition}

\begin{proof}
For every $t\in\mathbb Z$, Theorem~\ref{thm:two-sided-conjugacy}(ii) gives
\[
a_t=\psi_m(y_t,\dots,y_{t+m-1}).
\]
Substituting this identity into $\phi(a_t,\dots,a_{t+r-1})$ yields the definition of $\widetilde{\phi}$ and proves the equality of the finite-volume energy sums term by term.
\end{proof}

\begin{corollary}[Thermodynamic transport above threshold]\label{cor:thermo-equivalence}
Assume $m\ge 3$. For every finite-range interaction $\phi$ on the raw full shift, the induced interaction $\widetilde{\phi}$ on $Y_m$ satisfies:
\begin{enumerate}[label=(\roman*),leftmargin=*,itemsep=2pt]
  \item finite-volume Hamiltonians agree exactly on matched bi-infinite configurations and therefore differ by at most an $O(1)$ boundary term under any standard finite-volume boundary convention;
  \item the corresponding pressure and free-energy density coincide;
  \item equilibrium states are transported bijectively by $\Phi_m$.
\end{enumerate}
In this restricted sense the stabilized readout transports finite-range interactions exactly above the three-window threshold.
\end{corollary}

\begin{proof}
Item~(i) is Proposition~\ref{prop:finite-range-renorm}. For item~(ii), regard $\phi$ and $\widetilde{\phi}$ as the corresponding continuous potentials on the two-sided shifts. The pressure identity is Theorem~\ref{thm:continuous-potentials}; the free-energy-density statement is the same identity in finite-range thermodynamic notation. Item~(iii) is the equilibrium-state correspondence in Theorem~\ref{thm:continuous-potentials}.
\end{proof}

\begin{corollary}[Occupancy and nearest-neighbour energy under stabilized readout]\label{cor:occupancy-energy}
Assume $m\ge 3$ and interpret $A_t\in\{0,1\}$ as a binary occupancy readout along the scan. Define the finite-volume observables
\[
N_n(A):=\sum_{t=0}^{n-1}A_t,
\qquad
C_n(A):=\sum_{t=0}^{n-1}A_tA_{t+1},
\]
and the nearest-neighbour energy
\[
\mathcal H_n(A):=-J\,C_n(A)-h\,N_n(A).
\]
Then for $Y=\Phi_m(A)$ there exist local observables on the stabilized process,
\[
\widetilde A_t(Y):=\psi_m(Y_t,\dots,Y_{t+m-1}),
\qquad
\widetilde C_t(Y):=\widetilde A_t(Y)\widetilde A_{t+1}(Y),
\]
such that
\[
N_n(A)=\sum_{t=0}^{n-1}\widetilde A_t(Y),
\qquad
C_n(A)=\sum_{t=0}^{n-1}\widetilde C_t(Y),
\qquad
\mathcal H_n(A)=\widetilde{\mathcal H}_n(Y)
\]
with
\[
\widetilde{\mathcal H}_n(Y):=-J\sum_{t=0}^{n-1}\widetilde C_t(Y)-h\sum_{t=0}^{n-1}\widetilde A_t(Y).
\]
Consequently, occupancy density, pair density, and the corresponding free-energy density are preserved exactly by stabilized readout.
\end{corollary}

\begin{proof}
The identities for $N_n$ and $C_n$ follow immediately from Theorem~\ref{thm:two-sided-conjugacy}(ii), which recovers each occupancy bit as a local observable of the stabilized trajectory. The energy identity is the special case of Proposition~\ref{prop:finite-range-renorm} with range $r=2$.
\end{proof}

\subsection{Block-level divergences}\label{sec:measurement:info}

Since the block maps are bijections above threshold, the usual finite-dimensional divergences are unchanged after pushforward.

\begin{corollary}[Block-level divergence identities]\label{cor:info-geometry}
Fix $m\ge 3$ and $n\ge m$.
\begin{enumerate}[label=(\roman*),leftmargin=*,itemsep=2pt]
  \item For any convex function $f:[0,\infty)\to\RR$ with $f(1)=0$, let
  \[
  D_f(P\|Q):=\sum_a Q(a)\,f\!\left(\frac{P(a)}{Q(a)}\right)
  \]
  be the associated Csiszar $f$-divergence on the finite raw block space, with the usual conventions at $Q(a)=0$. Then for any probability laws $P,Q$ on $\{0,1\}^{n+m-1}$,
  \[
  D_f(P\|Q)=D_f\bigl((\Phi_{m,n})_*P\|(\Phi_{m,n})_*Q\bigr).
  \]
  In particular, the Hellinger integrals and the Bhattacharyya coefficient are preserved exactly.
  \item Let $\{P_\theta:\theta\in\Theta\}$ be a family of strictly positive mass functions on $\{0,1\}^{n+m-1}$ such that $p_\theta(a):=P_\theta(\{a\})$ is $C^1$ in $\theta$ for every raw block $a$. Let $Q_\theta:=(\Phi_{m,n})_*P_\theta$ and write
  \[
  q_\theta(y):=Q_\theta(\{y\}),
  \qquad y\in\mathcal Y_{m,n}.
  \]
  Then
  \[
  q_\theta(y)=p_\theta(\Psi_{m,n}(y)),
  \qquad
  \nabla_\theta \log q_\theta(y)=\nabla_\theta \log p_\theta(\Psi_{m,n}(y)).
  \]
  Consequently the Fisher information matrices of the raw and stabilized block experiments are identical:
  \[
  \mathcal I_Y(\theta)=\mathcal I_A(\theta).
  \]
\end{enumerate}
\end{corollary}

\begin{proof}
For (i), $\Phi_{m,n}$ is a bijection between the raw block space and $\mathcal Y_{m,n}$, with inverse $\Psi_{m,n}$. Writing the defining sum for the $f$-divergence on $\mathcal Y_{m,n}$ and changing variables by $a=\Psi_{m,n}(y)$ gives the identity immediately. The Hellinger and Bhattacharyya quantities are special cases obtained by choosing the corresponding functions of the likelihood ratio.

For (ii), the mass-function identity
\[
q_\theta(y)=p_\theta(\Psi_{m,n}(y))
\]
is exactly the pushforward relation under a bijection. Differentiating with respect to $\theta$ gives the score identity. The Fisher information matrix is the expectation of the score outer product, so the same change of variables under $\Psi_{m,n}$ shows that the raw and stabilized expectations coincide.
\end{proof}
