\section{Final remarks}\label{sec:discussion}

We have proved a sharp overlap threshold for a locally many-to-one Fibonacci folding.  The word map $\Fold_m$ collapses residue classes on isolated windows, but Theorems~\ref{thm:block-invertibility} and~\ref{thm:two-sided-conjugacy} show that overlapping windows restore injectivity for every $m\ge 3$ and give a finite-memory conjugacy onto a finite-type image.

The failure at $m=2$ is equally rigid.  Theorem~\ref{thm:m2-branch-locus} identifies the unique exceptional orbit, Corollary~\ref{cor:m2-defect-formulas} computes the entropy and Kullback--Leibler defect concentrated there, and Corollary~\ref{cor:m2-mod-zero} shows that the map is already invertible modulo null sets away from this branch.

The metallic comparison in Theorem~\ref{thm:metallic-two-window-rigidity} shows that this two-window branch is not a generic feature of the metallic finite-window normalization family.  For every metallic parameter $A\ge2$, the corresponding two-window code has a local inverse.  The golden case is therefore the unique member of this family in which the two-window overlap still fails to separate the hidden carry.

The image shift also has an intrinsic finite-state description.  Theorem~\ref{thm:fischer-cover-stabilized-image} proves that the overlap presentation is the right Fischer cover of $Y_m$ for every $m\ge3$.  Thus the $2^{m-1}$ overlap states are not an artifact of the proof; they are forced by the follower structure of the stabilized image shift.

The fixed-window entropy decomposition and the path-space result should therefore be read together.  Section~\ref{sec:entropy-ledger} measures multiplicity inside a single fiber, while Section~\ref{sec:measurement} proves that this multiplicity disappears after overlap above threshold.  In particular, invariant measures, pressure and equilibrium states are transported through the finite-memory conjugacy.

The residue formulas, Fourier representation, second-moment recurrence and discrepancy estimates give quantitative information about the same finite-window map.  In the Sturmian model they provide explicit deterministic control of stabilized block laws and transported local observables.

From the sofic-shift perspective, the threshold theorem can be read as follows: the sliding stabilization map $\Phi_m$ is a right-resolving factor of degree at most $2^{m-1}$ for every $m\ge 2$. General theory guarantees that a right-resolving one-to-one factor between irreducible shifts of finite type is a conjugacy \cite{Ashley1991,LindMarcus1995}. The content specific to this paper is that degree one is achieved at exactly $m=3$, and that the obstruction at $m=2$ is confined to a single orbit pair rather than a diffuse set. Neither conclusion follows from the abstract framework; both require the explicit Fibonacci residue arithmetic of Lemma~\ref{lem:m3-decoder} and Theorem~\ref{thm:m2-branch-locus}.

The Fibonacci setting is used because the necessary arithmetic can be made explicit.  Its Sturmian language has minimal aperiodic binary complexity, and in this case the decoder and branch classification can be proved in closed form; the thermodynamic identities then follow formally from the resulting conjugacy.  Section~\ref{sec:metallic-cover} gives a limited metallic comparison at the two-window boundary; no uniform threshold theorem is asserted for arbitrary numeration systems.

The quasicrystal readout motivation explains why the map is naturally posed on the full cube and why finite-sample discrepancy bounds are relevant.  The main theorem itself, however, is a symbolic-dynamical and arithmetic statement: above the three-window threshold, Fibonacci stabilization is a finite-memory recoding of the full two-shift.
