\begin{thebibliography}{99}

\bibitem{AlloucheShallit2003}
J.-P. Allouche and J. Shallit, \textit{Automatic Sequences: Theory, Applications, Generalizations}, Cambridge University Press, Cambridge, 2003.

\bibitem{Ashley1991}
J. Ashley, Resolving factor maps for shifts of finite type with equal entropy, \textit{Ergod. Theory Dynam. Systems}, \textbf{11} (1991), 219--240.

\bibitem{BaakeGrimm2013}
M. Baake and U. Grimm, \textit{Aperiodic Order. Vol. 1. A Mathematical Invitation}, Cambridge University Press, Cambridge, 2013.

\bibitem{Berstel2001}
J. Berstel, An exercise on Fibonacci representations, \textit{RAIRO Theor. Inform. Appl.}, \textbf{35} (2001), 491--498.

\bibitem{Berthe2001Ostrowski}
V. Berth\'e, Autour du syst\`eme de num\'eration d'Ostrowski, \textit{Bull. Belg. Math. Soc. Simon Stevin}, \textbf{8} (2001), 209--239.

\bibitem{BertheRigo2010}
V. Berth\'e and M. Rigo (eds.), \textit{Combinatorics, Automata and Number Theory}, Encyclopedia of Mathematics and its Applications, Vol. 135, Cambridge University Press, Cambridge, 2010.

\bibitem{BertheSteinerThuswaldner2019}
V. Berth\'e, W. Steiner and J. M. Thuswaldner, Geometry, dynamics, and arithmetic of $S$-adic shifts, \textit{Ann. Inst. Fourier (Grenoble)}, \textbf{69} (2019), 1347--1409.

\bibitem{Boyle1983}
M. Boyle, Lower entropy factors of sofic systems, \textit{Ergod. Theory Dynam. Systems}, \textbf{3} (1983), 541--557.

\bibitem{CoverThomas2006}
T. M. Cover and J. A. Thomas, \textit{Elements of Information Theory}, 2nd edition, Wiley-Interscience, Hoboken, NJ, 2006.

\bibitem{DrmotaTichy1997}
M. Drmota and R. F. Tichy, \textit{Sequences, Discrepancies and Applications}, Lecture Notes in Mathematics, Vol. 1651, Springer, Berlin, 1997.

\bibitem{Frougny1992}
C. Frougny, Representations of numbers and finite automata, \textit{Math. Systems Theory}, \textbf{25} (1992), 37--60.

\bibitem{Hedlund1969}
G. A. Hedlund, Endomorphisms and automorphisms of the shift dynamical system, \textit{Math. Systems Theory}, \textbf{3} (1969), 320--375.

\bibitem{Jagannathan2021}
A. Jagannathan, The Fibonacci quasicrystal: Case study of hidden dimensions and multifractality, \textit{Rev. Mod. Phys.}, \textbf{93} (2021), 045001.

\bibitem{Kechris1995}
A. S. Kechris, \textit{Classical Descriptive Set Theory}, Graduate Texts in Mathematics, Vol. 156, Springer, New York, 1995.

\bibitem{Kitchens1998}
B. P. Kitchens, \textit{Symbolic Dynamics. One-Sided, Two-Sided and Countable State Markov Shifts}, Springer, Berlin, 1998.

\bibitem{Koshy2001}
T. Koshy, \textit{Fibonacci and Lucas Numbers with Applications}, Wiley-Interscience, New York, 2001.

\bibitem{KuipersNiederreiter1974}
L. Kuipers and H. Niederreiter, \textit{Uniform Distribution of Sequences}, Wiley, New York, 1974.

\bibitem{LindMarcus1995}
D. Lind and B. Marcus, \textit{An Introduction to Symbolic Dynamics and Coding}, Cambridge University Press, Cambridge, 1995.

\bibitem{Lothaire2002}
M. Lothaire, \textit{Algebraic Combinatorics on Words}, Cambridge University Press, Cambridge, 2002.

\bibitem{Moody2000}
R. V. Moody, Model sets: A survey, in \textit{From Quasicrystals to More Complex Systems} (eds. F. Axel, F. D\'enoyer and J.-P. Gazeau), Springer, Berlin, 2000, 145--166.

\bibitem{MorseHedlund1940}
M. Morse and G. A. Hedlund, Symbolic dynamics II. Sturmian trajectories, \textit{Amer. J. Math.}, \textbf{62} (1940), 1--42.

\bibitem{Nasu1995}
M. Nasu, Textile systems for endomorphisms and automorphisms of the shift, \textit{Mem. Amer. Math. Soc.}, \textbf{114} (1995), no. 546.

\bibitem{Ostrowski1922}
A. Ostrowski, Bemerkungen zur Theorie der diophantischen Approximationen, \textit{Abh. Math. Sem. Univ. Hamburg}, \textbf{1} (1922), 77--98.

\bibitem{Parry1960}
W. Parry, On the $\beta$-expansions of real numbers, \textit{Acta Math. Acad. Sci. Hungar.}, \textbf{11} (1960), 401--416.

\bibitem{Queffelec2010}
M. Queff\'elec, \textit{Substitution Dynamical Systems--Spectral Analysis}, 2nd edition, Lecture Notes in Mathematics, Vol. 1294, Springer, Berlin, 2010.

\bibitem{Shechtman1984}
D. Shechtman, I. Blech, D. Gratias and J. W. Cahn, Metallic phase with long-range orientational order and no translational symmetry, \textit{Phys. Rev. Lett.}, \textbf{53} (1984), 1951--1953.

\bibitem{Stanley2012}
R. P. Stanley, \textit{Enumerative Combinatorics. Vol. 1}, 2nd edition, Cambridge University Press, Cambridge, 2012.

\bibitem{Walters1982}
P. Walters, \textit{An Introduction to Ergodic Theory}, Graduate Texts in Mathematics, Vol. 79, Springer, New York, 1982.

\bibitem{Zeckendorf1972}
E. Zeckendorf, Repr\'esentation des nombres naturels par une somme de nombres de Fibonacci ou de nombres de Lucas, \textit{Bull. Soc. Roy. Sci. Li\`ege}, \textbf{41} (1972), 179--182.

\end{thebibliography}
