\documentclass[11pt,a4paper]{amsart}

\usepackage{amsmath,amssymb,amsthm,mathrsfs}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage[protrusion=true,expansion=false]{microtype}
\usepackage{mathtools}
\raggedbottom

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\spec}{spec}
\DeclareMathOperator{\Res}{Res}
\DeclareMathOperator{\disc}{Disc}
\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\Rot}{Rot}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\Irr}{Irr}
\DeclareMathOperator{\Li}{Li}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\lcm}{lcm}
\DeclareMathOperator{\FP}{FP}
\DeclareMathOperator{\Log}{Log}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\cP}{\mathcal P}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\norm}[1]{\lVert #1 \rVert}

\title[Finite Parts and Adams--M\"obius Inversion]{Finite parts and Adams--M\"obius inversion for finite group extensions of shifts of finite type}

\author{Haobo Ma}
\address[Haobo Ma]{AELF PTE LTD., \#14-02, Marina Bay Financial Centre Tower 1, 8 Marina Blvd, Singapore 018981}
\email[Haobo Ma]{auric@aelf.io}

\author{Wenlin Zhang}
\address[Wenlin Zhang]{National University of Singapore, Singapore}
\email[Wenlin Zhang]{e1327962@u.nus.edu}

\date{}

\begin{document}

\begin{abstract}
For a mixing subshift of finite type with primitive adjacency matrix
\(A\) and Perron root \(\lambda\), we study the passage from dynamical
determinants to primitive orbit data. For a primitive matrix we obtain
an explicit finite-part formula at \(z=\lambda^{-1}\) in
reduced-determinant coordinates and show that the residue constants of
the cyclic lifts determine the reduced spectral polynomial. For finite
group extensions, the basic device is an Adams--M\"obius inversion
formula: ordinary M\"obius inversion removes repetitions of orbit
length, while Adams operations record repetitions of the group label.
This gives determinant formulas for the primitive Peter--Weyl channels
throughout the disc \(|z|<\lambda^{-1}\). Under the twisted spectral-gap
hypothesis
\(\eta:=\max_{\varrho\neq\mathbf{1}}\rho(A_\varrho)<\lambda\), we deduce
classwise Mertens constants, a non-abelian Fourier inversion formula
for these constants, and, when
\(\max\{\Lambda(A),\eta\}>\lambda^{1/2}\), the first spectral-edge
correction. We also characterize the boundary case
\(\rho(A_\varrho)=\lambda\) by a cocycle-equivariant line field, give a
sufficient criterion for \(\eta<\lambda\), and extend the same formalism
to locally constant finite-group cocycles over real H\"older potentials
on a mixing shift of finite type. In that H\"older setting, every
spectral window above the larger of the common essential radius and
\(e^{P/2}\) yields an explicit weighted correction term governed by the
corresponding twisted Ruelle projectors.
\end{abstract}

\subjclass[2020]{37B10, 37D35, 11M41, 20C15}
\keywords{Dynamical $\zeta$-function, shift of finite type, finite-group extension,
primitive orbit, Adams operation, twisted determinant}

\maketitle

\input{sec_introduction}
\input{sec_preliminaries}
\input{sec_finite_part}
\input{sec_chebotarev}
\input{sec_shadows}
\appendix
\input{sec_appendices}

\begin{thebibliography}{99}

\bibitem{AdachiSunada1987TwistedPerronFrobenius}
T.~Adachi and T.~Sunada,
Twisted {P}erron--{F}robenius theorem and {$L$}-functions,
\textit{J. Funct. Anal.} \textbf{71} (1987), no.~1, 1--46.

\bibitem{Baladi2000PositiveTransfer}
V.~Baladi,
\textit{Positive Transfer Operators and Decay of Correlations},
Advanced Series in Nonlinear Dynamics, vol.~16,
World Scientific, River Edge, NJ, 2000.

\bibitem{Baladi2018ZetaDeterminants}
V.~Baladi,
\textit{Dynamical Zeta Functions and Dynamical Determinants for Hyperbolic Maps:
A Functional Approach},
Ergebnisse der Mathematik und ihrer Grenzgebiete. 3. Folge / A Series of Modern Surveys in Mathematics, vol.~68,
Springer, Cham, 2018.

\bibitem{BowenLanford1970Zeta}
R.~Bowen and O.~E.~Lanford, III,
Zeta functions of restrictions of the shift transformation,
in \textit{Global Analysis},
Proc. Sympos. Pure Math., vol.~14,
Amer. Math. Soc., Providence, RI, 1970, pp.~43--49.

\bibitem{BoyleSchmieding2017FiniteGroupExtensions}
M.~Boyle and S.~Schmieding,
Finite group extensions of shifts of finite type: {$K$}-theory, {P}arry and {L}iv\v{s}ic,
\textit{Ergodic Theory Dynam. Systems} \textbf{37} (2017), no.~4, 1026--1059.

\bibitem{DougallSharp2021AnosovCovers}
R.~Dougall and R.~Sharp,
Anosov flows, growth rates on covers and group extensions of subshifts,
\textit{Invent. Math.} \textbf{223} (2021), no.~2, 445--483.

\bibitem{HasegawaSaito2016GraphMertens}
T.~Hasegawa and S.~Saito,
On graph theory {M}ertens' theorems,
\textit{Graphs Combin.} \textbf{32} (2016), no.~5, 1915--1930.

\bibitem{HasegawaSaitoSato2015AbelianCovering}
T.~Hasegawa, S.~Saito, and I.~Sato,
A generalization of a graph theory {M}ertens' theorem: abelian covering case,
\textit{Q. J. Math.} \textbf{66} (2015), no.~3, 809--836.

\bibitem{HasegawaSaitoSato2018GaloisCovering}
T.~Hasegawa, S.~Saito, and I.~Sato,
A generalization of a graph theory {M}ertens' theorem: {G}alois covering case,
\textit{Forum Math.} \textbf{30} (2018), no.~3, 599--615.

\bibitem{KotaniSunada2000ZetaFiniteGraphs}
M.~Kotani and T.~Sunada,
Zeta functions of finite graphs,
\textit{J. Math. Sci. Univ. Tokyo} \textbf{7} (2000), no.~1, 7--25.

\bibitem{LindMarcus1995}
D.~Lind and B.~Marcus,
\textit{An Introduction to Symbolic Dynamics and Coding},
Cambridge University Press, Cambridge, 1995.

\bibitem{Manning1971Axiom}
A.~Manning,
Axiom {A} diffeomorphisms have rational zeta functions,
\textit{Bull. London Math. Soc.} \textbf{3} (1971), 215--220.

\bibitem{ArtinMazur1965PeriodicPoints}
M.~Artin and B.~Mazur,
On periodic points,
\textit{Ann. of Math. (2)} \textbf{81} (1965), no.~1, 82--99.

\bibitem{Parry1983PrimeOrbit}
W.~Parry,
An analogue of the prime number theorem for closed orbits of shifts of
finite type and their suspensions,
\textit{Amer. J. Math.} \textbf{105} (1983), no.~3, 702--723.

\bibitem{PollicottSharp1998Mertens}
M.~Pollicott and R.~Sharp,
Asymptotic expansions for closed orbits in homology classes,
\textit{Geom. Dedicata} \textbf{87} (2001), no.~1--3, 123--160.

\bibitem{Sharp1991Mertens}
R.~Sharp,
An analogue of {M}ertens' theorem for closed orbits of {A}xiom {A} flows,
\textit{Bol. Soc. Brasil. Mat.} \textbf{21} (1991), no.~2, 205--229.

\bibitem{Sharp1993PrimeOrbitAnnals}
R.~Sharp,
Prime orbit theorems with multi-dimensional constraints for {A}xiom {A} flows,
\textit{Monatsh. Math.} \textbf{114} (1993), no.~3--4, 261--304.

\bibitem{Sharp2020AmenableGroupExtensions}
R.~Sharp,
Convergence of zeta functions for amenable group extensions of shifts,
in \textit{Dynamics: Topology and Numbers},
Contemp. Math., vol.~744,
Amer. Math. Soc., Providence, RI, 2020, pp.~245--257.

\bibitem{StarkTerras1996ZetaGraphCoverings}
H.~M.~Stark and A.~A.~Terras,
Zeta functions of finite graphs and coverings,
\textit{Adv. Math.} \textbf{121} (1996), no.~1, 124--165.

\bibitem{StarkTerras2000ZetaGraphCoveringsII}
H.~M.~Stark and A.~A.~Terras,
Zeta functions of finite graphs and coverings, {II},
\textit{Adv. Math.} \textbf{154} (2000), no.~1, 132--195.

\bibitem{Stadlbauer2013KestenAmenability}
M.~Stadlbauer,
An extension of {K}esten's criterion for amenability to topological
{M}arkov chains,
\textit{Adv. Math.} \textbf{235} (2013), 450--468.

\bibitem{Stadlbauer2019ConformalMeasures}
M.~Stadlbauer,
On conformal measures and harmonic functions for group extensions,
in \textit{New Trends in One-Dimensional Dynamics},
Springer Proc. Math. Stat., vol.~285,
Springer, Cham, 2019, pp.~271--299.

\bibitem{TerrasStark2007ZetaGraphCoveringsIII}
A.~A.~Terras and H.~M.~Stark,
Zeta functions of finite graphs and coverings, {III},
\textit{Adv. Math.} \textbf{208} (2007), no.~1, 467--489.

\bibitem{Sunada1989DynamicalLFunctions}
T.~Sunada,
Dynamical {$L$}-functions and homology of closed orbits,
\textit{Bull. Amer. Math. Soc. (N.S.)} \textbf{20} (1989), no.~1, 73--77.

\bibitem{NooraniParry1992ChebotarevShifts}
M.~S.~M.~Noorani and W.~Parry,
A {C}hebotarev theorem for finite homogeneous extensions of shifts,
\textit{Bol. Soc. Brasil. Mat.} \textbf{23} (1992), no.~1--2, 137--151.

\bibitem{Parry1999LivsicNonAbelian}
W.~Parry,
The {L}iv\v{s}ic periodic point theorem for non-abelian cocycles,
\textit{Ergodic Theory Dynam. Systems} \textbf{19} (1999), no.~3, 687--701.

\bibitem{ParryPollicott1986Chebotarev}
W.~Parry and M.~Pollicott,
The {C}hebotarov theorem for {G}alois coverings of {A}xiom {A} flows,
\textit{Ergodic Theory Dynam. Systems} \textbf{6} (1986), no.~1, 133--148.

\bibitem{ParryPollicott1990Zeta}
W.~Parry and M.~Pollicott,
\textit{Zeta Functions and the Periodic Orbit Structure of Hyperbolic Dynamics},
Ast\'erisque, nos.~187--188,
Soci\'et\'e Math\'ematique de France, 1990.

\bibitem{PollicottSharp2007Chebotarev}
M.~Pollicott and R.~Sharp,
Chebotarev-type theorems in homology classes,
\textit{Proc. Amer. Math. Soc.} \textbf{135} (2007), no.~10, 3323--3332.

\bibitem{PollicottZhang2025RapidMixing}
M.~Pollicott and D.~Zhang,
Rapid mixing for compact group extensions of hyperbolic flows,
\textit{Trans. Amer. Math. Soc.} \textbf{378} (2025), no.~7, 5011--5056.

\bibitem{Ruelle1976ZetaExpanding}
D.~Ruelle,
Zeta-functions for expanding maps and {A}nosov flows,
\textit{Invent. Math.} \textbf{34} (1976), 231--242.

\bibitem{Ruelle1978Thermodynamic}
D.~Ruelle,
\textit{Thermodynamic Formalism},
Addison-Wesley, Reading, MA, 1978.

\bibitem{Serre1977LinearRepresentations}
J.-P.~Serre,
\textit{Linear Representations of Finite Groups},
Graduate Texts in Mathematics, vol.~42,
Springer-Verlag, New York-Heidelberg, 1977.

\end{thebibliography}

\end{document}
