\section{Holomorphic variation and effective truncation}
\label{app:variation}

This appendix records holomorphic variation of the Perron-pole finite
part and an explicit truncation estimate. The results here are routine
consequences of the main-text formulas, are not part of the paper's
novelty claims, and may be omitted on a first reading.

\begin{proposition}[Reduced-determinant gradient formula]
  \label{thm:finite-part-reduced-determinant-group-inverse-gradient}
  Let \(\theta\mapsto A_\theta\) be a holomorphic family of primitive
  matrices near \(\theta_0\), and assume that
  \(\lambda(\theta):=\rho(A_\theta)\) stays simple. Choose left and
  right Perron vectors \(\ell_\theta,r_\theta\) normalised by
  \(\ell_\theta^{\mathsf T}r_\theta=1\), let
  \(P_\theta:=r_\theta\ell_\theta^{\mathsf T}\), and set
  \[
    B_\theta:=I-\frac{A_\theta}{\lambda(\theta)},
    \qquad
    R_\theta:=B_\theta^{\#},
  \]
  where \(B_\theta^{\#}\) is the group inverse, characterised by
  \[
    B_\theta R_\theta=R_\theta B_\theta=I-P_\theta,
    \qquad
    R_\theta P_\theta=P_\theta R_\theta=0.
  \]
  Then
  \[
    \partial_\theta\log C(A_\theta)
    =\Tr\!\left(
      R_\theta\left(
        \frac{\partial_\theta A_\theta}{\lambda(\theta)}
        -\frac{\partial_\theta\lambda(\theta)}{\lambda(\theta)^2}
         A_\theta\right)\right),
  \]
  with
  \[
    \partial_\theta\lambda(\theta)
    =\ell_\theta^{\mathsf T}(\partial_\theta A_\theta)r_\theta.
  \]
\end{proposition}

\begin{proof}
  Since the Perron eigenvalue is simple, \(B_\theta\) has a simple zero
  eigenvalue and all other eigenvalues are non-zero; hence the group
  inverse \(R_\theta\) exists. On the invariant complement
  \((I-P_\theta)\CC^d\), the reduced determinant is the ordinary
  determinant of \(B_\theta\), so the usual logarithmic derivative gives
  \[
    \partial_\theta \log\det\nolimits'(B_\theta)
    =\Tr(R_\theta\,\partial_\theta B_\theta).
  \]
  Using \(\log C(A_\theta)=-\log\det\nolimits'(B_\theta)\) from
  \Cref{prop:finite-part-residue-reduced-determinant} and
  \[
    \partial_\theta B_\theta
    =-\frac{\partial_\theta A_\theta}{\lambda(\theta)}
      +\frac{\partial_\theta\lambda(\theta)}{\lambda(\theta)^2}A_\theta
  \]
  gives the first formula. The derivative of the Perron root is the
  standard perturbation identity
  \[
    \partial_\theta\lambda(\theta)
    =\ell_\theta^{\mathsf T}(\partial_\theta A_\theta)r_\theta.
  \]
  Compare Kato's analytic perturbation theory for simple isolated
  eigenvalues \cite{Kato1995Perturbation}.
\end{proof}

\subsection{Holomorphic variation and effective truncation}

\begin{proposition}[Holomorphic variation formula]
  \label{thm:logM-holomorphic-variation}
  Let \(\theta\mapsto A_\theta\) be a holomorphic family of primitive
  matrices near \(\theta_0\), and assume that
  \(\lambda(\theta):=\rho(A_\theta)\) stays simple. Then
  \(\theta\mapsto\FP(A_\theta)\) is holomorphic near \(\theta_0\) and
  \[
    \frac{\dd}{\dd\theta}\FP(A_\theta)
    =\frac{\dd}{\dd\theta}\log C(A_\theta)
     +\sum_{k\ge 2}\frac{\mu(k)}{k}
      \frac{\dd}{\dd\theta}
      \log\det(I-\lambda(\theta)^{-k}A_\theta)^{-1}.
  \]
  Moreover, for every fixed \(z\) with \(|z|<\lambda(\theta)^{-1}\),
  \[
    \partial_\theta\log\det(I-zA_\theta)^{-1}
    =\Tr\!\bigl((I-zA_\theta)^{-1}z\,\partial_\theta A_\theta\bigr).
  \]
\end{proposition}

\begin{proof}
  Apply \Cref{thm:finite-part-primitive-closed-form} to each \(A_\theta\).
  Because \(|\lambda(\theta)^{-k}|<\lambda(\theta)^{-1}\) uniformly for
  \(k\ge 2\) near \(\theta_0\), the sum of holomorphic terms converges
  locally uniformly. Hence \(\FP(A_\theta)\) is holomorphic and may
  be differentiated term by term. The first derivative formula uses
  \Cref{thm:finite-part-reduced-determinant-group-inverse-gradient} for
  the \(C(A_\theta)\)-term.

  For the resolvent identity, apply Jacobi's formula in finite
  dimensions:
  \[
    \partial_\theta \log\det(I-zA_\theta)
    =
    \Tr\!\bigl((I-zA_\theta)^{-1}(-z\,\partial_\theta A_\theta)\bigr).
  \]
  Multiplying by \(-1\) gives the stated derivative formula for
  \(\log\det(I-zA_\theta)^{-1}\).
\end{proof}

\begin{proposition}[Truncation bound]
  \label{thm:logM-truncation-bound}
  Let \(A\) be a primitive non-negative \(d\times d\) matrix with
  Perron root \(\lambda>1\), and define
  \[
    \FP_K(A)
    :=\log C(A)
      +\sum_{n\le K}p_n(A)
       \bigl(\lambda^{-n}+\log(1-\lambda^{-n})\bigr).
  \]
  Then
  \[
    \abs{\FP(A)-\FP_K(A)}
    \le
    \frac{d\,\lambda^{-(K+1)}}{(K+1)(1-\lambda^{-1})^2}.
  \]
\end{proposition}

\begin{proof}
  By \Cref{thm:finite-part-primitive-closed-form},
  \[
    \abs{\FP(A)-\FP_K(A)}
    \le \sum_{n>K} p_n(A)\psi(\lambda^{-n}).
  \]
  Since
  \[
    \psi(x)=-x-\log(1-x)\le \sum_{m\ge 2}x^m=\frac{x^2}{1-x},
  \]
  and
  \[
    p_n(A)\le \frac{\Tr(A^n)}{n}\le \frac{d\,\lambda^n}{n},
  \]
  one obtains
  \[
    p_n(A)\psi(\lambda^{-n})
    \le \frac{d}{n}\frac{\lambda^{-n}}{1-\lambda^{-n}}
    \le \frac{d}{n}\frac{\lambda^{-n}}{1-\lambda^{-1}}.
  \]
  Summing over \(n>K\) gives
  \[
    \abs{\FP(A)-\FP_K(A)}
    \le \frac{d}{1-\lambda^{-1}}
      \sum_{n>K}\frac{\lambda^{-n}}{n}
    \le
    \frac{d\,\lambda^{-(K+1)}}{(K+1)(1-\lambda^{-1})^2}.
  \]
\end{proof}

