\section{Introduction}\label{sec:introduction}

The dynamical \(\zeta\)-function of a mixing subshift of finite type
\((X_A^+,\sigma)\) with primitive adjacency matrix \(A\) is the
rational function
\[
  \zeta_A(z)
  =\exp\!\left(\sum_{n\ge 1}\frac{\Tr(A^n)}{n}z^n\right)
  =\det(I-zA)^{-1}.
\]
Since Bowen--Lanford \cite{BowenLanford1970Zeta} and Manning
\cite{Manning1971Axiom}, the analytic properties of \(\zeta_A\) have
been central to the orbit-counting theory of symbolic dynamics, with
further development through the thermodynamic formalism of Ruelle
\cite{Ruelle1976ZetaExpanding,Ruelle1978Thermodynamic} and the
transfer-operator programme of Parry--Pollicott
\cite{ParryPollicott1990Zeta}. The spectral theory of the associated
Ruelle operator underlies both the prime orbit theorem and
equidistribution results for periodic orbits in hyperbolic dynamics
\cite{Baladi2000PositiveTransfer,ParryPollicott1990Zeta}.

The Perron root \(\lambda=\rho(A)\) gives \(\zeta_A\) a simple pole at
\(z_\ast=\lambda^{-1}\). The existence of Mertens-type asymptotics at
this pole is classical: an explicit Mertens constant in the
Axiom~A-flow setting was established by Sharp \cite{Sharp1991Mertens},
and the same structural form appears in the orbit-counting framework of
Pollicott--Sharp \cite{PollicottSharp1998Mertens}, in the
graph-theoretic covering-zeta literature of Stark--Terras
\cite{StarkTerras1996ZetaGraphCoverings,StarkTerras2000ZetaGraphCoveringsII,%
TerrasStark2007ZetaGraphCoveringsIII}, and in the graph-theoretic
Mertens refinements of Hasegawa--Saito
\cite{HasegawaSaito2016GraphMertens}. The present paper is centred on a
narrower point. Its new ingredient is a class-function
Adams--M\"obius inversion that recovers primitive Peter--Weyl channels
from twisted determinant logarithms. The reduced-determinant finite-part
formula and the cyclic-lift reconstruction recorded first are
supporting finite-state tools, not separate main theorem lines. The
Perron-point class constants are then conditional corollaries under the
explicit hypothesis \(\eta<\lambda\), and the paper does not claim a
general criterion for that spectral gap.

After removing the Perron factor one obtains the reduced determinant
\[
  C(A):=\det\nolimits'(I-A/\lambda)^{-1}
\]
and the reduced spectral polynomial
\[
  F_A(t):=\frac{\det(I-tA/\lambda)}{1-t}
  =\prod_{j=2}^d\left(1-\frac{\mu_j}{\lambda}t\right),
\]
where \(\mu_2,\dots,\mu_d\) are the non-Perron eigenvalues. The
Abel-regularised finite-part constant
\[
  \FP(A):=\lim_{r\uparrow 1}\bigl(P_A(r)+\log(1-r)\bigr),
  \qquad
  P_A(r):=\sum_{n\ge 1}p_n(A)\lambda^{-n}r^n,
\]
measures the deviation of the primitive orbit sum from the harmonic
series.

\begin{proposition}[Supporting finite-part formula at the Perron pole]
  \label{thm:intro-finite-part}
  Let \(A\) be a primitive non-negative matrix with Perron root
  \(\lambda=\rho(A)>1\), reduced determinant \(C(A)\), and primitive
  orbit counts \(p_n(A)\). Then
  \[
    \FP(A)
    =\log C(A)
      +\sum_{k\ge 2}\frac{\mu(k)}{k}
        \log\det(I-\lambda^{-k}A)^{-1}.
  \]
  Equivalently,
  \[
    \FP(A)
    =\log C(A)
      +\sum_{n\ge 1}p_n(A)
        \bigl(\lambda^{-n}+\log(1-\lambda^{-n})\bigr).
  \]
  In particular,
  \[
    \sum_{n\le N}\frac{p_n(A)}{\lambda^n}
    =\log N+\mathcal M(A)+o(1),
    \qquad
    \mathcal M(A):=\gamma+\FP(A).
  \]
\end{proposition}

\Cref{thm:intro-finite-part} is proved in
\Cref{thm:finite-part-primitive-closed-form,cor:finite-part-mertens-asymptotic}.
It is a reduced-determinant packaging of the classical dynamical
Mertens asymptotic, included because the same determinant coordinates
reappear in the finite-group setting.

The same supporting section also records a cyclic-lift reconstruction
fact that we use as the abelian comparison model.

\begin{proposition}[Supporting cyclic reconstruction from cyclic lifts]
  \label{thm:intro-cyclic-reconstruction}
  Let \(A\) be a primitive non-negative \(d\times d\) matrix, and set
  \[
    \Psi_q(A):=\log q+\log C(A^{\langle q\rangle}),
  \]
  where \(A^{\langle q\rangle}=A\otimes P_q\) is the cyclic lift. Then
  the dimension \(d\) together with the sequence
  \(\{\Psi_q(A)\}_{q\ge 1}\) determines \(F_A\), and hence the
  non-Perron spectrum of \(A\) with algebraic multiplicity.
\end{proposition}

The proof combines M\"obius inversion on multiples with Newton's
identities; see \Cref{thm:finite-part-cyclic-lift-spectrum-identifiability}.
A one-root-of-unity refinement is recorded in the appendix.

For finite-group extensions the primitive problem is genuinely
different. In the Chebotarev theory initiated by Parry--Pollicott
\cite{ParryPollicott1986Chebotarev} and, in the finite homogeneous SFT
setting, carried out by Noorani--Parry
\cite{NooraniParry1992ChebotarevShifts}, the Peter--Weyl decomposition
of the extension cocycle leads to twisted transition matrices
\(A_\varrho\) and their determinants. What is missing in the classical
periodic-level picture is a mechanism that recovers a \emph{primitive}
Peter--Weyl channel directly from those determinants. The main theorem
of this paper supplies that mechanism.

\begin{theorem}[Theorem A: Adams--M\"obius primitive inversion with
  conditional Perron-point corollaries]
  \label{thm:intro-finite-group}
  Let \(\tau:E_A\to G\) be a cocycle into a finite group, let
  \(A_\varrho\) be the twisted transition matrices, and for a class
  function \(\phi:G\to\CC\) write
  \[
    \mathcal L_\phi(z)
    :=\sum_{n\ge 1}\frac{T_{\phi,n}}{n}z^n,
    \qquad
    \Pi_\phi(z):=\sum_{\gamma\in\cP}\phi(\tau(\gamma))z^{|\gamma|}.
  \]
  Then:
  \begin{enumerate}[label=\textup{(\roman*)}]
    \item \textup{(Adams--M\"obius inversion.)} Class-function
      determinant linearisation is standard, and the new step is that
      Adams--M\"obius inversion expresses each primitive Peter--Weyl
      series directly in terms of those logarithms:
      \[
        \Pi_\phi(z)
        =\sum_{m\ge 1}\frac{\mu(m)}{m}\,
          \mathcal L_{\psi^m\phi}(z^m).
      \]
    \item \textup{(Conditional Perron-point package.)}
      Assume
      \[
        \eta:=\max_{\varrho\neq\mathbf 1}\rho(A_\varrho)<\lambda.
      \]
      Then for every conjugacy class \(C\subset G\) one has
      \[
        p_{n,C}\sim \frac{|C|}{|G|}\frac{\lambda^n}{n},
      \]
      and
      \[
        \sum_{n\le N}\frac{p_{n,C}}{\lambda^n}
        =\frac{|C|}{|G|}\log N+\mathcal M_C(A)+O(N^{-1}).
      \]
      Moreover, the deviations
      \(\Delta_C(A):=\mathcal M_C(A)-\frac{|C|}{|G|}\mathcal M(A)\)
      determine \(\Pi_\varrho(\lambda^{-1})\) for every non-trivial
      \(\varrho\) by non-abelian Fourier inversion.
  \end{enumerate}
\end{theorem}

The determinant identities and primitive inversion of
\Cref{thm:intro-finite-group}\textup{(i)} hold unconditionally inside
the open Perron disc. The Perron-point class constants and their
Fourier reconstruction in
\Cref{thm:intro-finite-group}\textup{(ii)} rely on the explicit
hypothesis \(\eta<\lambda\); the paper does not solve this spectral-gap
problem in general. Section~\ref{sec:shadows} then records only the
quotient compatibility needed for the main text and the \(S_3\)-example
showing that vanishing periodic traces need not force vanishing
primitive bias.

\subsection{Relation to existing work}\label{subsec:prior-work}

\subsubsection*{Supporting finite-state preliminaries}
The dynamical \(\zeta\)-function for SFTs goes back to Artin--Mazur
\cite{ArtinMazur1965PeriodicPoints} and Bowen--Lanford
\cite{BowenLanford1970Zeta,Manning1971Axiom}. The prime-orbit
asymptotic for shifts of finite type is classical: Parry
\cite{Parry1983PrimeOrbit} proved the analogue of the prime number
theorem for closed orbits of SFTs and their suspensions. The existence
of Mertens-type asymptotics for Axiom~A flows was established by Sharp
\cite{Sharp1991Mertens}; see also \cite{Sharp1993PrimeOrbitAnnals} for
multi-dimensional refinements. Pollicott--Sharp
\cite{PollicottSharp1998Mertens} give explicit Mertens-constant
formulas in homology classes, and graph-theoretic analogues appear in
\cite{StarkTerras1996ZetaGraphCoverings,StarkTerras2000ZetaGraphCoveringsII,%
TerrasStark2007ZetaGraphCoveringsIII,HasegawaSaito2016GraphMertens}.
\Cref{thm:intro-finite-part} does not claim a new prime-orbit theorem;
it records the corresponding constant in reduced-determinant
coordinates. Likewise, \Cref{thm:intro-cyclic-reconstruction} is
supporting finite-state reconstruction infrastructure rather than a
separate main theorem line.

\subsubsection*{Twisted \(L\)-functions and the missing primitive step}
The twisted Perron--Frobenius theorem and the associated dynamical
\(L\)-functions were introduced by Adachi--Sunada
\cite{AdachiSunada1987TwistedPerronFrobenius} and Sunada
\cite{Sunada1989DynamicalLFunctions}. The Peter--Weyl determinant
factorisation of the extension zeta function and Chebotarev-type
equidistribution for finite-group extensions of Axiom~A systems are due
to Parry--Pollicott \cite{ParryPollicott1986Chebotarev}; the finite
homogeneous SFT case was carried out by Noorani--Parry
\cite{NooraniParry1992ChebotarevShifts}. Further developments include
the Chebotarev results of Pollicott--Sharp
\cite{PollicottSharp2007Chebotarev} in homology classes, Parry's
non-abelian Liv\v{s}ic theorem \cite{Parry1999LivsicNonAbelian}, and
the \(K\)-theory approach of Boyle--Schmieding
\cite{BoyleSchmieding2017FiniteGroupExtensions}.

Known input gives the periodic determinant factorisation and the
class-function linearisation
\[
  \mathcal L_\phi(z)
  =
  \sum_{\varrho\in\Irr(G)}
    \widehat\phi(\varrho)\log\det(I-zA_\varrho)^{-1}.
\]
It also gives the leading Chebotarev asymptotic
\(p_{n,C}\sim(|C|/|G|)\lambda^n/n\). What it does \emph{not} give is a
primitive irreducible channel directly from the twisted determinant
data. Direct M\"obius inversion of a single twisted logarithm leaves
higher-power contamination \(\chi_\varrho(g^r)\) from repeated orbit
labels. The missing step is therefore to intertwine M\"obius inversion
with the Adams operations \(g\mapsto g^m\) on class functions. This is
the content of \Cref{cor:class-function-adams-mobius} and
\Cref{cor:primitive-peter-weyl-determinant}; the precise sense in which
direct trace-level M\"obius inversion is weaker is isolated in
\Cref{prop:naive-mobius-contamination}. The constant-term packaging in
\Cref{thm:class-mertens-explicit,thm:class-mertens-fourier} is then a
non-trivial Perron-point consequence of this primitive calculus under
\(\eta<\lambda\).

\subsubsection*{Group extensions and broader dynamical context}
For the graph-theoretic side of the determinant formalism, see
Kotani--Sunada \cite{KotaniSunada2000ZetaFiniteGraphs} and the
covering-zeta programme of Stark--Terras
\cite{StarkTerras1996ZetaGraphCoverings,StarkTerras2000ZetaGraphCoveringsII,%
TerrasStark2007ZetaGraphCoveringsIII}. Ward
\cite{Ward1999TypicalExtensions} studies typical isometric extensions of
full shifts and shows that the corresponding dynamical \(\zeta\)-functions
are generically far from algebraic; this is directly relevant
background for zeta functions of shift extensions, but it does not
address primitive Peter--Weyl extraction from twisted determinants.
Dougall--Sharp \cite{DougallSharp2021AnosovCovers} analyse growth rates
on covers and group extensions of subshifts and Anosov flows through
entropy, amenability, and counting on covers, while Pollicott--Zhang
\cite{PollicottZhang2025RapidMixing} treat rapid mixing for compact
group extensions of hyperbolic flows. Together with Sharp's work on
amenable group extensions of shifts
\cite{Sharp2020AmenableGroupExtensions}, Stadlbauer's pressure and
amenability results
\cite{Stadlbauer2013KestenAmenability,Stadlbauer2019ConformalMeasures},
and Boyle--Schmieding's analysis of Parry and Liv\v{s}ic obstructions
\cite{BoyleSchmieding2017FiniteGroupExtensions}, these papers form the
broader context for spectral and growth questions in group extensions.
They do not, however, provide the Adams--M\"obius primitive inversion
or the ensuing determinant packaging of the Perron-point constants that
is the focus of the present paper.

\subsubsection*{Structure of the paper's contribution}
The paper is organised around one primary theorem chain:
Adams--M\"obius primitive inversion, the irreducible-channel determinant
formula, and the resulting class Mertens constants under
\(\eta<\lambda\). The finite-part and cyclic-lift formulas are included
as supporting finite-state preliminaries, and the quotient results of
\Cref{sec:shadows} are kept only to record quotient functoriality and to
interpret the \(S_3\)-example. Classical ingredients such as twisted
determinant factorisation, class-function linearisation, and the leading
Chebotarev asymptotic are treated throughout as input rather than as
separate contributions.

\subsection{Theorem-by-theorem comparison with prior work}

\Cref{thm:intro-finite-part} is only an introduction-level summary of the
classical dynamical Mertens theorem of Sharp \cite{Sharp1991Mertens},
recast in reduced-determinant coordinates for primitive matrices.
\Cref{thm:intro-cyclic-reconstruction} is not classical input: it states
the paper's supporting cyclic-lift reconstruction principle, for which we
did not find a numbered prior theorem recovering the whole reduced
spectral polynomial from the scalar data \(q\mapsto \Psi_q(A)\).
\Cref{prop:prelim-trace-primitive} records the standard
Artin--Mazur/Bowen--Lanford/Parry--Pollicott zeta formalism and is used
only as input.
\Cref{lem:primitive-orbit-asymptotic} is likewise a standard prime-orbit
estimate in the shift-of-finite-type setting, included only as a
building block for later error terms.
\Cref{def:kernel-rh} introduces paper-specific shorthand for the
verifiable inequality \(\Lambda(A)\le \lambda^{1/2}\), not a classical
named definition.
\Cref{def:abel-finite-part} isolates the classical Mertens constant as an
Abel finite part at the Perron pole, which is the normalization used in
the rest of the manuscript.
\Cref{thm:finite-part-primitive-closed-form} adds a concrete
reduced-determinant identity for that constant beyond the earlier
orbit-counting formulations of Sharp and Pollicott--Sharp.
\Cref{prop:finite-part-residue-reduced-determinant} packages the pole
residue as the deleted determinant \(\det'(I-A/\lambda)^{-1}\), which is
the form needed for the later cyclic-lift discussion.
\Cref{cor:finite-part-mertens-asymptotic} is classical input and does not
claim more than the usual Mertens asymptotic.
\Cref{thm:finite-part-cyclic-lift-reduced-constant-closed} gives a new
closed formula for cyclic-lift residue constants in terms of the reduced
eigenvalues \(\mu_j/\lambda\).
\Cref{thm:finite-part-cyclic-lift-spectrum-identifiability} then extracts
from those cyclic-lift constants a full reconstruction of the reduced
spectrum, which again is not something we found stated in the cited
covering-zeta literature.

\Cref{thm:intro-finite-group} already signals the paper's main
positioning: the periodic Chebotarev theory of
Parry--Pollicott~\cite{ParryPollicott1986Chebotarev} and
Noorani--Parry~\cite{NooraniParry1992ChebotarevShifts} is treated as
classical input, while the primitive Adams--M\"obius step is the genuine
new ingredient.
\Cref{prop:class-indicator-expansion} is the standard second
orthogonality relation for characters, as in Serre
\cite{Serre1977LinearRepresentations}.
\Cref{prop:kernel-peter-weyl-block-diagonalisation} is the standard
determinant factorisation for finite-group extensions already present in
the classical twisted-\(L\)-function framework.
\Cref{prop:twisted-spectral-radius-bound} is also classical input,
serving only to control the non-trivial twisted channels.
\Cref{thm:class-function-linearisation} is not claimed as new: it is the
standard periodic determinant linearisation for class functions from the
same Chebotarev tradition.
\Cref{cor:class-function-adams-mobius} is the first genuinely new step in
the finite-group part, because it inserts Adams operations into
M\"obius inversion to recover primitive class-function series directly
from periodic determinant logarithms.
\Cref{prop:naive-mobius-contamination} clarifies why that extra Adams step
is needed by isolating the exact higher-power contamination left by
direct trace-level M\"obius inversion.

\Cref{lem:adams-coefficients-bounded} supplies the uniform Adams-coefficient
bound needed for the later channel estimates, and we did not find this
notation-level statement isolated in the cited literature.
\Cref{cor:primitive-peter-weyl-determinant} upgrades the classical
periodic determinant factorisation to a primitive irreducible-channel
determinant formula.
\Cref{cor:primitive-peter-weyl-trace} then turns that determinant formula
into a coefficientwise primitive trace identity.
\Cref{thm:class-primitive-decomposition} refines the classical
class-function decomposition from periodic orbit counts to primitive
generating functions and primitive coefficients.
\Cref{cor:class-artin-mobius-trace} makes that refinement explicit as a
primitive Artin--M\"obius trace formula for each conjugacy class.
\Cref{lem:class-primitive-trivial-split} separates the trivial channel
from the non-trivial Peter--Weyl remainder in exactly the form used for
the Perron-point argument.
\Cref{lem:nontrivial-peter-weyl-channel-bound} adds a concrete primitive
irreducible-channel bound of size
\(\max\{\eta^n,\lambda^{n/2}\}/n\), rather than only a classwise leading
equidistribution statement.
\Cref{prop:eta-gap-coboundary-obstruction} translates the classical
Liv\v{s}ic coboundary obstruction into a concrete spectral equality
\(\rho(A_\chi)=\lambda\) for one-dimensional characters.

\Cref{lem:nontrivial-perron-point-convergence} is conditional on the
explicit hypothesis \(\eta<\lambda\) and converts the channel estimate
into absolute convergence at the Perron point.
\Cref{lem:geometric-harmonic-tail} is an elementary but useful explicit
tail bound that we did not find isolated as a numbered result in the
cited references.
\Cref{lem:perron-channel-contribution} is also conditional on
\(\eta<\lambda\) and sharpens the classical trivial-channel Mertens
asymptotic to an exponentially decaying remainder.
\Cref{thm:class-mertens-explicit} is therefore best read as a conditional
corollary to the new primitive calculus: under \(\eta<\lambda\), it
upgrades the classical leading Chebotarev asymptotic to classwise
Mertens sums with explicit constants and Perron-point evaluations.
\Cref{prop:class-mertens-deviations} isolates the deviation constants
\(\Delta_C(A)\) and expresses the non-trivial Perron-point primitive
channels directly in determinant coordinates.
\Cref{thm:class-mertens-fourier} then shows that those conditional class
constants determine every non-trivial Perron-point primitive channel by
non-abelian Fourier inversion and satisfy a Plancherel identity.
\Cref{def:normalised-primitive-class-profile} is merely paper-specific
notation for packaging those conditional deviation constants as a class
function.

\Cref{thm:quotient-functoriality} records exact compatibility of the
primitive generating functions and class counts under quotient maps, a
functoriality statement we did not find stated in this primitive form in
the earlier literature.
\Cref{thm:quadratic-adams-obstruction} is a new specialisation of the
Adams--M\"obius mechanism showing explicitly that periodic vanishing in a
quadratic channel need not force primitive vanishing.
\Cref{prop:s3-example-channels} gives a fully explicit \(S_3\) example
with complete twisted-channel data and a verified spectral gap
\(\eta=1<2=\lambda\).
\Cref{prop:s3-example-primitive-channels} shows on that example that a
non-trivial primitive bias can survive even when one periodic logarithm
vanishes identically.
\Cref{cor:s3-example-profile} completes the example by computing the
primitive class counts, the class-constant deviations, and the
abelian/non-abelian energy split of the resulting profile.

\subsection{Plan of the paper}

\Cref{sec:preliminaries} fixes notation for primitive orbit counts,
reduced determinants, cyclic lifts, and finite-group extensions.
\Cref{sec:finite-part} records the supporting finite-state formulas:
the finite-part identity at the Perron pole and the cyclic-lift
reconstruction of the reduced spectrum. \Cref{sec:chebotarev}, which
forms the core of the paper, develops the Adams--M\"obius mechanism for
finite-group extensions, proves the irreducible primitive determinant
formula, and packages the conditional Perron-point consequences under a
twisted spectral gap; it also includes a precise comparison with the
classical twisted-\(L\)-function framework and a focused discussion of
why \(\eta<\lambda\) is an extra hypothesis rather than an automatic
fact. \Cref{sec:shadows} keeps only the quotient functoriality needed in
the main text and the \(S_3\)-model. The appendix collects the longer
cyclic-lift reconstructions, the orthogonal cyclic-detection formalism,
and the perturbative variation and truncation formulas.
