\section{Finite parts at the Perron pole and cyclic reconstruction}
\label{sec:finite-part}

Throughout this section, \(A\) denotes a primitive non-negative matrix
with Perron root \(\lambda>1\). No further symbolic-dynamical input is
used: the arguments depend only on the primitive matrix formalism, the
Perron pole, and the non-Perron spectrum. The section is included as
supporting finite-state infrastructure for the finite-group arguments of
\Cref{sec:chebotarev}; it is not intended as a separate main theorem
branch.

\medskip\noindent\textbf{Logarithm convention.}
All logarithms in this section are taken in the principal branch
\(\Log\) on \(\CC\setminus(-\infty,0]\). Since the non-Perron
eigenvalues satisfy \(|\mu_j|<\lambda\), each factor
\(1-(\mu_j/\lambda)^q\) has positive real part for all \(q\ge 1\), so
\(\Log(1-(\mu_j/\lambda)^q)\) is unambiguous. Likewise, the reduced
spectral polynomial satisfies \(F_A(e^{i\theta})\ne 0\) for all
\(\theta\in\RR\) (since \(|\mu_j|<\lambda\) implies \(|\mu_j/\lambda|<1\)),
so \(\Log F_A(e^{i\theta})\) is well defined and
\(-\Log F_A(e^{i\theta})=\sum_{n\ge 1}u_n(A)e^{in\theta}/n\) is the
Fourier series of a continuous function on the torus. All subsequent
power series and determinant logarithms use this branch consistently.

\subsection{Relation with classical dynamical Mertens constants}
\label{subsec:relation-classical-mertens}

The existence of Mertens-type asymptotics for orbit sums in hyperbolic
dynamics is well established. Sharp \cite{Sharp1991Mertens} proved
that for a mixing Axiom~A flow with topological entropy \(h\), the
closed-orbit analogue of Mertens' theorem takes the form
\[
  \sum_{\ell(\gamma)\le T}\frac{1}{e^{h\ell(\gamma)}}
  =\log T+c_0+o(1),
\]
where \(\ell(\gamma)\) is the period of the primitive orbit \(\gamma\)
and \(c_0\) is a dynamical Mertens constant determined by the
non-leading spectral data of the Ruelle operator. Pollicott--Sharp
\cite{PollicottSharp1998Mertens} refined this to homology-class Mertens
constants with explicit leading-order expressions involving the
determinant of the Hessian of the pressure function. In the
graph-theoretic setting, Stark--Terras
\cite{StarkTerras1996ZetaGraphCoverings} obtained analogous Mertens
constants for the Ihara zeta function of a finite graph via the
Bass--Ihara determinant formula, and Hasegawa--Saito
\cite{HasegawaSaito2016GraphMertens} derived graph-theoretic Mertens
refinements explicitly parallel to the Sharp--Pollicott formulas.

The finite-part formula below is a specialisation of these results to
the finite-state symbolic setting. The structural form of the constant
parallels those in the cited works; the specific contribution is the
explicit expression as a reduced-determinant identity in the
primitive-matrix language of finite non-negative matrices:
\begin{enumerate}[label=(\roman*)]
  \item it expresses the classical constant as an explicit
    reduced-determinant identity,
  \item the determinant form connects directly to the cyclic-lift
    reconstruction of the non-Perron spectrum in the next subsection,
    and
  \item it provides the abelian baseline for the Adams--M\"obius
    mechanism in \Cref{sec:chebotarev}.
\end{enumerate}
In particular, \Cref{thm:finite-part-primitive-closed-form} does not
claim a new prime-orbit theorem; it is the translation of the classical
Mertens asymptotic into finite-state reduced-determinant coordinates.

\subsection{Finite parts and reduced determinants}

\begin{proposition}[Primitive-orbit closed form for the finite part]
  \label{thm:finite-part-primitive-closed-form}
  The finite-part constant \(\FP(A)\) exists and satisfies
  \[
    \FP(A)
    =\log C(A)+\sum_{k\ge 2}\frac{\mu(k)}{k}
      \log\det(I-\lambda^{-k}A)^{-1}.
  \]
  Equivalently,
  \[
    \FP(A)
    =\log C(A)
      +\sum_{n\ge 1}p_n(A)
        \bigl(\lambda^{-n}+\log(1-\lambda^{-n})\bigr).
  \]
\end{proposition}

\begin{proof}
  We separate the argument into two steps.

  First, the Euler product and M\"obius inversion identify the Abel
  generating series with a determinant expression. By
  \Cref{prop:prelim-trace-primitive}, for \(|z|<\lambda^{-1}\),
  \[
    \log\det(I-zA)^{-1}
    =\sum_{m\ge 1}\frac{\Tr(A^m)}{m}z^m.
  \]
  Hence, for \(0\le r<1\),
  \begin{align*}
    \sum_{k\ge 1}\frac{\mu(k)}{k}
      \log\det(I-\lambda^{-k}r^kA)^{-1}
    &=
    \sum_{k\ge 1}\frac{\mu(k)}{k}
      \sum_{m\ge 1}\frac{\Tr(A^m)}{m}\lambda^{-km}r^{km} \\
    &=
    \sum_{n\ge 1}\frac{(\lambda^{-1}r)^n}{n}
      \sum_{k\mid n}\mu(k)\Tr(A^{n/k}) \\
    &=
    \sum_{n\ge 1}p_n(A)\lambda^{-n}r^n
    =P_A(r),
  \end{align*}
  where the last identity is \eqref{eq:primitive-mobius}.

  Second, we pass to the finite part at \(r\uparrow 1\). The \(k=1\)
  term has asymptotic
  \[
    \log\det(I-\lambda^{-1}rA)^{-1}
    =-\log(1-r)+\log C(A)+o(1),
    \qquad r\uparrow 1,
  \]
  because the Perron pole is simple. Every term with \(k\ge 2\)
  converges absolutely at \(r=1\), so
  \[
    \FP(A)
    =\lim_{r\uparrow 1}\bigl(P_A(r)+\log(1-r)\bigr)
    =\log C(A)+\sum_{k\ge 2}\frac{\mu(k)}{k}
      \log\det(I-\lambda^{-k}A)^{-1}.
  \]

  For the primitive-orbit form, use again
  \Cref{prop:prelim-trace-primitive}:
  \[
    \log\det(I-zA)^{-1}
    =-\sum_{n\ge 1}p_n(A)\log(1-z^n).
  \]
  Substituting \(z=\lambda^{-k}\), summing over \(k\ge 2\), and
  exchanging the two absolutely convergent sums gives
  \[
    \sum_{k\ge 2}\frac{\mu(k)}{k}\log\det(I-\lambda^{-k}A)^{-1}
    =
    -\sum_{n\ge 1}p_n(A)
      \sum_{k\ge 2}\frac{\mu(k)}{k}\log(1-\lambda^{-nk}).
  \]
  For \(|x|<1\), absolute convergence permits the coefficientwise
  calculation
  \[
    \sum_{k\ge 1}\frac{\mu(k)}{k}\log(1-x^k)
    =-\sum_{k\ge 1}\frac{\mu(k)}{k}\sum_{r\ge 1}\frac{x^{kr}}{r}
    =-\sum_{n\ge 1}\frac{x^n}{n}\sum_{k\mid n}\mu(k)
    =-x,
  \]
  because \(\sum_{k\mid n}\mu(k)=0\) unless \(n=1\). Hence
  \[
    \sum_{k\ge 2}\frac{\mu(k)}{k}\log(1-x^k)=-x-\log(1-x).
  \]
  Taking \(x=\lambda^{-n}\) yields the second formula.
\end{proof}

\begin{proposition}[Residue constant and reduced determinant]
  \label{prop:finite-part-residue-reduced-determinant}
  Let \(M\) be a finite matrix and let \(\lambda\) be a simple
  eigenvalue of \(M\). If \(\nu_2,\dots,\nu_d\) are the remaining
  eigenvalues of \(M\), counted with algebraic multiplicity, then
  \[
    C_\lambda(M)
    =\prod_{j=2}^d\left(1-\frac{\nu_j}{\lambda}\right)^{-1}
    =\det\nolimits'(I-M/\lambda)^{-1}.
  \]
  In particular, if \(M=A\) is primitive with Perron root \(\lambda\),
  then
  \[
    C(A)=C_\lambda(A)
    =\prod_{j=2}^d\left(1-\frac{\mu_j}{\lambda}\right)^{-1}
    =\det\nolimits'(I-A/\lambda)^{-1}.
  \]
\end{proposition}

\begin{proof}
  Factor
  \[
    \det(I-zM)
    =(1-\lambda z)\prod_{j=2}^d(1-\nu_j z).
  \]
  Multiplying by \(1-\lambda z\) and letting \(z\to\lambda^{-1}\) gives
  the claim.
\end{proof}

\begin{corollary}[Orbit-Mertens asymptotic]
  \label{cor:finite-part-mertens-asymptotic}
  For every primitive matrix \(A\),
  \[
    \sum_{n\le N}\frac{p_n(A)}{\lambda^n}
    =\log N+\mathcal M(A)+o(1)
    \qquad (N\to\infty),
  \]
  where \(\mathcal M(A)=\gamma+\FP(A)\).
\end{corollary}

\begin{proof}
  By \Cref{prop:prelim-prime-orbit},
  \[
    \sum_{n\ge 1}\left|\frac{p_n(A)}{\lambda^n}-\frac{1}{n}\right|
    <\infty.
  \]
  Hence
  \[
    \sum_{n\le N}\frac{p_n(A)}{\lambda^n}
    =
    \sum_{n\le N}\frac{1}{n}
    +\sum_{n\ge 1}
      \left(\frac{p_n(A)}{\lambda^n}-\frac{1}{n}\right)
    +o(1).
  \]
  The convergent correction term equals \(\FP(A)\), while
  \(\sum_{n\le N}1/n=\log N+\gamma+o(1)\).
\end{proof}

\subsection{Cyclic lifts and reduced spectral reconstruction}

Before proceeding, we extend the residue-constant notation to matrices
that need not be primitive. For any finite matrix \(M\) with a simple
eigenvalue \(\lambda\) at its spectral radius, define
\begin{equation}\label{eq:residue-constant-general}
  C_\lambda(M):=\lim_{z\to\lambda^{-1}}(1-\lambda z)\det(I-zM)^{-1}.
\end{equation}
When \(M=A\) is primitive, this coincides with the earlier
\(C(A)\). For a cyclic lift \(A^{\langle q\rangle}=A\otimes P_q\)
with \(q>1\), the matrix is imprimitive (it has period \(q\)), but
the Perron root \(\lambda=\rho(A)\) remains a simple eigenvalue
of \(A^{\langle q\rangle}\), so \(C_\lambda(A^{\langle q\rangle})\)
is well defined. In the sequel we write
\(C(A^{\langle q\rangle}):=C_\lambda(A^{\langle q\rangle})\) for
brevity.

\medskip
Let
\[
  \Lambda_{\mathrm{rel}}(A):=\Lambda(A)/\lambda\in[0,1)
\]
and
\[
  u_n(A):=\sum_{j=2}^d\left(\frac{\mu_j}{\lambda}\right)^n
  \qquad (n\ge 1).
\]

\begin{proposition}[Reduced determinants of cyclic lifts]
  \label{thm:finite-part-cyclic-lift-reduced-constant-closed}
  For every \(q\ge 1\),
  \[
    C(A^{\langle q\rangle})
    =\frac{1}{q}\prod_{j=2}^d
      \left(1-\left(\frac{\mu_j}{\lambda}\right)^q\right)^{-1}.
  \]
  Equivalently,
  \[
    \Psi_q(A):=\log q+\log C(A^{\langle q\rangle})
    =-\sum_{j=2}^d
      \log\!\left(1-\left(\frac{\mu_j}{\lambda}\right)^q\right).
  \]
\end{proposition}

\begin{proof}
  The spectrum of \(A^{\langle q\rangle}=A\otimes P_q\) is
  \[
    \{\mu\,\omega_q^k:\mu\in\spec(A),\ 0\le k\le q-1\}.
  \]
  Applying \Cref{prop:finite-part-residue-reduced-determinant} to
  \(M=A^{\langle q\rangle}\), with the simple eigenvalue \(\lambda\),
  yields
  \[
    C(A^{\langle q\rangle})
    =\prod_{k=1}^{q-1}(1-\omega_q^k)^{-1}
      \prod_{j=2}^d\prod_{k=0}^{q-1}
        \left(1-\frac{\mu_j}{\lambda}\omega_q^k\right)^{-1}.
  \]
  The first product equals \(1/q\), while
  \[
    \prod_{k=0}^{q-1}(1-x\omega_q^k)=1-x^q.
  \]
  Substituting \(x=\mu_j/\lambda\) proves the claim.
\end{proof}

The auxiliary multiple-sum expansion, M\"obius inversion on multiples,
and the limsup estimate used in the next proposition are recorded in the
appendix.

\begin{proposition}[Determination of the reduced spectrum from cyclic lifts]
  \label{thm:finite-part-cyclic-lift-spectrum-identifiability}
  Given the matrix dimension \(d\) and the sequence
  \(\{\Psi_q(A)\}_{q\ge 1}\) (equivalently
  \(\{C(A^{\langle q\rangle})\}_{q\ge 1}\)), the reduced
  spectral polynomial \(F_A\) is uniquely determined, and hence the
  non-Perron spectrum of \(A\) with algebraic multiplicity.
  Moreover,
  \[
    \limsup_{q\to\infty}\abs{\Psi_q(A)}^{1/q}
    =\Lambda_{\mathrm{rel}}(A).
  \]
\end{proposition}

\begin{proof}
  By \Cref{prop:finite-part-cyclic-lift-mobius-inversion}, the cyclic
  data determine every power sum \(u_n(A)\). Since
  \[
    F_A(t)=\prod_{j=2}^d\left(1-\frac{\mu_j}{\lambda}t\right)
  \]
  has degree \(d-1\) (which is part of the input), Newton's identities
  recover its coefficients from the first \(d-1\) power sums, and hence
  from the full sequence \(\{u_n(A)\}_{n\ge 1}\). This reconstructs the
  non-Perron spectrum with algebraic multiplicity. (The degree \(d-1\)
  is needed as input: the power-sum sequence alone does not determine
  the degree of a monic polynomial without further information, since a
  trailing zero coefficient is invisible to Newton's identities.)

  For the exponential growth rate,
  \(u_q(A)=\sum_{j=2}^d(\mu_j/\lambda)^q\), so
  \Cref{lem:finite-part-exponential-sum-limsup} gives
  \[
    \limsup_{q\to\infty}\abs{u_q(A)}^{1/q}
    =\Lambda_{\mathrm{rel}}(A).
  \]
  The explicit remainder bound from
  \Cref{prop:finite-part-cyclic-lift-dirichlet-multiple-sum} shows that
  \[
    \Psi_q(A)-u_q(A)
    =O\!\left(\Lambda_{\mathrm{rel}}(A)^{2q}\right).
  \]
  If \(\Lambda_{\mathrm{rel}}(A)=0\), then both sequences are
  eventually zero. Otherwise the error term has strictly smaller
  exponential rate than \(u_q(A)\), so
  \[
    \limsup_{q\to\infty}\abs{\Psi_q(A)}^{1/q}
    =\Lambda_{\mathrm{rel}}(A).
  \]
\end{proof}

An additional one-root-of-unity reconstruction statement and a toy
example are recorded in the appendix.

