\section{Discussion and open problems}\label{sec:conclusion}

\subsection{Summary of the main contributions}

The central contribution is the Adams--M\"obius primitive inversion
mechanism for finite-group extensions, which expresses each irreducible
Peter--Weyl channel of the primitive orbit series in terms of twisted
determinants and the Adams operations on the character ring
(\Cref{cor:class-function-adams-mobius,%
cor:primitive-peter-weyl-determinant}). Under the twisted spectral-gap
hypothesis \(\eta<\lambda\), this yields explicit class Mertens
constants (\Cref{thm:class-mertens-explicit}) and a non-abelian
Fourier reconstruction (\Cref{thm:class-mertens-fourier}).

As supporting infrastructure, \Cref{sec:finite-part} records a closed
determinant formula for the finite-part constant at the Perron pole
(\Cref{thm:finite-part-primitive-closed-form}), specialising the
classical dynamical Mertens asymptotic
\cite{Sharp1991Mertens,PollicottSharp1998Mertens} to finite-state
reduced-determinant coordinates, and shows that cyclic-lift data
reconstruct the non-Perron spectrum. The quotient-functoriality results
of \Cref{sec:shadows} identify cyclic data as the exact abelian shadow
of the full primitive class profile.

The determinant identities and the Adams--M\"obius inversion hold
throughout the open Perron disc without extra hypotheses. The twisted
spectral-gap condition \(\eta<\lambda\) is used precisely when the
primitive Peter--Weyl series are evaluated at the boundary point
\(z=\lambda^{-1}\) and converted into conjugacy-class Mertens
constants.

\subsection{Open problems}

We close with three problems that arise naturally from the results
above but remain outside the present scope.

\begin{enumerate}
\item \emph{Dynamical criteria for the twisted spectral gap.} The
  condition
  \(\eta=\max_{\varrho\neq\mathbf 1}\rho(A_\varrho)<\lambda\) is
  assumed in \Cref{thm:class-mertens-explicit}. For the abelian
  (cyclic-lift) case this reduces to the classical condition that the
  Perron root is strictly dominant, which holds by primitivity.  For
  non-abelian extensions the analogous condition was also assumed
  without derivation in \cite{ParryPollicott1986Chebotarev}.
  Identifying computable sufficient conditions on the cocycle
  \(\tau:E_A\to G\) that guarantee \(\eta<\lambda\) for broad classes
  of finite-group extensions remains open. The analysis of Parry and
  Liv\v{s}ic obstructions in
  \cite{BoyleSchmieding2017FiniteGroupExtensions}, together with
  Parry's non-abelian periodic-point theorem
  \cite{Parry1999LivsicNonAbelian}, gives a partial structural answer,
  and the \(S_3\)-model of \Cref{prop:s3-example-channels} (where
  \(\rho(A_\rho)=1<2=\lambda\)) provides one explicit verification.

\item \emph{Non-abelian analogues of the arithmetic appendix.} The
  Dirichlet-character inversion of \Cref{app:arithmetic} and the
  growth criterion for the square-root bound are abelian in nature,
  relying on cyclic quotient data. It would be of interest to know
  which of these arithmetic consequences extend to the non-abelian
  setting, for instance whether boundary multiplicities
  (\Cref{thm:finite-part-boundary-multiplicity-qaxis-energy}) admit
  an analogue formulated in terms of the higher-dimensional twisted
  channels.

\item \emph{Transfer-operator extensions.} The present paper works
  entirely at the level of finite matrices and algebraic determinants.
  For countable-state shifts or for the Ruelle transfer operator acting
  on spaces of H\"older-continuous functions, the finite-part constant
  and the Adams--M\"obius mechanism should have natural analogues
  involving the Fredholm determinant of the transfer operator. Relating
  \(\FP(A)\) to the spectral data of the transfer operator on an
  appropriate Banach space would connect the present results to the
  broader programme of Baladi
  \cite{Baladi2000PositiveTransfer,Baladi2018ZetaDeterminants}, the
  regularity results of J\'ez\'equel
  \cite{Jezequel2019ParameterRegularity}, and the
  Ruelle--Pollicott resonance theory.
\end{enumerate}
