\section{Preliminaries}\label{sec:preliminaries}

We collect here the notation and standard facts used throughout the
paper.

\subsection{Mixing subshifts and primitive orbit counts}

Let \(\mathcal A\) be a finite alphabet and let
\(A\in\{0,1\}^{\mathcal A\times\mathcal A}\) be a primitive adjacency
matrix. The associated one-sided subshift of finite type is
\[
  X_A^+
  =\bigl\{(x_n)_{n\ge 0}\in \mathcal A^{\NN}:
    A_{x_n,x_{n+1}}=1\ \text{for all }n\ge 0\bigr\},
\]
equipped with the left shift \(\sigma\); see \cite[Ch.~2]{LindMarcus1995}.
We write
\[
  a_n(A):=\Tr(A^n)=\#\operatorname{Fix}(\sigma^n)
\]
for the number of points of period dividing \(n\). The number of
primitive orbits of least period \(n\) is
\begin{equation}\label{eq:primitive-mobius}
  p_n(A)=\frac{1}{n}\sum_{d\mid n}\mu(d)\,a_{n/d}(A),
\end{equation}
and the necklace identity reads
\begin{equation}\label{eq:trace-primitive-identity}
  a_n(A)=\sum_{d\mid n}d\,p_d(A).
\end{equation}

\begin{proposition}[Trace, primitive orbits, and Euler product]
  \label{prop:prelim-trace-primitive}
  For every primitive matrix \(A\), the dynamical \(\zeta\)-function
  satisfies
  \begin{equation}\label{eq:zeta-def}
    \zeta_A(z)
    :=\exp\!\left(\sum_{n\ge 1}\frac{\Tr(A^n)}{n}z^n\right)
    =\frac{1}{\det(I-zA)}
    =\prod_{n\ge 1}(1-z^n)^{-p_n(A)}.
  \end{equation}
\end{proposition}

\begin{proof}
  The last equality follows from \eqref{eq:trace-primitive-identity},
  since
  \[
    \sum_{n\ge 1}\frac{a_n(A)}{n}z^n
    =\sum_{m\ge 1}\sum_{k\ge 1}\frac{p_m(A)}{k}z^{mk}
    =-\sum_{m\ge 1}p_m(A)\log(1-z^m).
  \]
  The determinant identity is standard for finite matrices.
\end{proof}

If \(\lambda=\rho(A)\) is the Perron root and
\[
  \Lambda(A):=\max\{|\mu|:\mu\in\spec(A),\ \mu\neq\lambda\},
\]
then the divisor sum in \eqref{eq:primitive-mobius} gives the usual
square-root loss in primitive orbit counting.

\begin{lemma}\label{lem:primitive-orbit-asymptotic}\label{prop:prelim-prime-orbit}
For a primitive matrix \(A\),
\[
  p_n(A)=\frac{\lambda^n}{n}
  +O\!\left(\frac{\max\{\Lambda(A)^n,\lambda^{n/2}\}}{n}\right).
\]
\end{lemma}

\begin{proof}
By Perron--Frobenius theory,
\(
  a_n(A)=\lambda^n+O(\Lambda(A)^n)
\).
In \eqref{eq:primitive-mobius}, the divisor \(d=1\) contributes
\(\lambda^n/n+O(\Lambda(A)^n/n)\). For \(d\ge 2\) one has
\(n/d\le n/2\), hence
\[
  \sum_{\substack{d\mid n\\ d\ge 2}}\frac{|a_{n/d}(A)|}{n}
  \ll \frac{1}{n}\sum_{\substack{d\mid n\\ d\ge 2}}\lambda^{n/d}
  \ll \frac{\lambda^{n/2}}{n}.
\]
Combining the two contributions proves the claim.
\end{proof}

\subsection{Reduced determinants and finite parts}

Let \(z_\ast=\lambda^{-1}\). Since \(\lambda\) is a simple eigenvalue,
\(\zeta_A\) has a simple pole at \(z_\ast\), and we define the residue
constant
\begin{equation}\label{eq:residue-constant-def}
  C(A):=\lim_{z\to z_\ast}(1-\lambda z)\zeta_A(z).
\end{equation}
If \(\mu_2,\dots,\mu_d\) are the non-Perron eigenvalues, then
\begin{equation}\label{eq:reduced-det-poly}
  F_A(t):=\frac{\det(I-tA/\lambda)}{1-t}
  =\prod_{j=2}^d\left(1-\frac{\mu_j}{\lambda}t\right)
\end{equation}
and we call \(F_A\) the \emph{reduced spectral polynomial}. Moreover,
\begin{equation}\label{eq:reduced-det-constant}
  C(A)=\det\nolimits'(I-A/\lambda)^{-1}=F_A(1)^{-1}.
\end{equation}

Define the primitive Abel generating series
\[
  P_A(r):=\sum_{n\ge 1}p_n(A)\lambda^{-n}r^n,
  \qquad 0\le r<1.
\]
Since \(p_n(A)\lambda^{-n}=1/n+O(\vartheta^n/n)\) for some
\(\vartheta\in(0,1)\), the renormalised expression
\[
  P_A(r)+\log(1-r)
\]
has a finite limit as \(r\uparrow 1\).

\begin{definition}[Abel finite-part constant]\label{def:abel-finite-part}
For a primitive matrix \(A\) we write
\[
  \FP(A):=\lim_{r\uparrow 1}\bigl(P_A(r)+\log(1-r)\bigr).
\]
The associated orbit-Mertens constant is
\[
  \mathcal M(A):=\gamma+\FP(A).
\]
\end{definition}

\subsection{Cyclic lifts}

For \(q\ge 1\) let \(P_q\) be the \(q\)-cycle permutation matrix and
define the cyclic lift
\begin{equation}\label{eq:cyclic-lift-def}
  A^{\langle q\rangle}:=A\otimes P_q.
\end{equation}
The spectral identity
\[
  \spec(A^{\langle q\rangle})
  =\{\mu\,\omega_q^j:\mu\in\spec(A),\ 0\le j\le q-1\}
\]
will be used repeatedly. In particular,
\(\rho(A^{\langle q\rangle})=\rho(A)\), and the constants
\[
  \Psi_q(A):=\log q+\log C(A^{\langle q\rangle})
\]
encode the non-Perron spectrum. For each \(q\ge 1\), the corresponding
root-of-unity layer is
\[
  \{F_A(\omega_q^j)\}_{0\le j\le q-1},
\]
and the exponential rate
\[
  \limsup_{q\to\infty}\abs{\Psi_q(A)}^{1/q}
\]
detects the largest non-Perron modulus relative to \(\lambda\); see
\Cref{thm:finite-part-cyclic-lift-spectrum-identifiability}.

\subsection{Finite-group extensions}

Let \(G\) be a finite group, let \(\tau:E(\Gamma)\to G\) be an edge
cocycle, and let \(\widetilde\Gamma\) be the skew-product extension on
vertex set \(V\times G\), with edges
\[
  (i,g)\longrightarrow (j,g\tau(e))
  \qquad \text{whenever } e:i\to j.
\]
Its adjacency matrix is denoted by \(\widetilde A\).

For an irreducible unitary representation
\(\varrho:G\to U(V_\varrho)\) with character \(\chi_\varrho\), define
the twisted transition matrix
\begin{equation}\label{eq:twisted-matrix-def}
  (A_\varrho)_{ij}:=\sum_{e:i\to j}\varrho(\tau(e)),
\end{equation}
viewed as an operator on \(\CC^V\otimes V_\varrho\).

If \(\gamma=e_1\cdots e_n\) is a primitive orbit, write
\[
  \tau(\gamma)=\tau(e_1)\cdots\tau(e_n)\in G
\]
for its group label. For a class function \(\phi:G\to\CC\), write
\begin{equation}\label{eq:class-function-fourier}
  \phi
  =\sum_{\varrho\in\Irr(G)}\widehat\phi(\varrho)\chi_\varrho,
  \qquad
  \widehat\phi(\varrho)
  :=\frac{1}{|G|}\sum_{g\in G}\phi(g)\overline{\chi_\varrho(g)}.
\end{equation}
For \(m\ge 1\), define the Adams operation on class functions by
\begin{equation}\label{eq:adams-class-function}
  (\psi^m\phi)(g):=\phi(g^m).
\end{equation}
These notions underlie the class-function calculus of
\Cref{sec:chebotarev}.

\begin{remark}[Inversion and complex conjugation for characters]
  \label{rem:character-inversion-conjugation}
  If \(\varrho\in\Irr(G)\) is unitary, then for every \(g\in G\),
  \[
    \chi_\varrho(g^{-1})
    =\overline{\chi_\varrho(g)}.
  \]
  Hence for every conjugacy class \(C\subset G\),
  \[
    \chi_\varrho(C^{-1})
    =\overline{\chi_\varrho(C)}.
  \]
\end{remark}

\subsection{Standing notation}

Throughout the paper \(\omega_q=e^{2\pi i/q}\) denotes a primitive
\(q\)th root of unity, and \(\Phi_q\) the \(q\)th cyclotomic
polynomial. We reserve the notation \(\varrho\) for irreducible
representations of finite groups and keep \(\rho(\cdot)\) for spectral
radii. In \Cref{sec:finite-part} we work only with the primitive matrix
formalism, whereas \Cref{sec:chebotarev,sec:shadows} return to the
graph and cocycle description of finite-group extensions.
