\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{hyperref}
\usepackage{microtype}

\numberwithin{equation}{section}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\RR}{\mathbb{R}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\norm}[1]{\lVert #1 \rVert}
\newcommand{\Haus}{\mathcal{H}}
\newcommand{\coeff}{\mathrm{coeff}}
\newcommand{\cell}{\mathrm{cell}}
\newcommand{\coord}{\mathrm{coord}}
\DeclareMathOperator{\Vol}{Vol}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\Div}{div}
\DeclareMathOperator{\Tr}{Tr}

\title{Sharp coefficient bounds for the centered homotopy operator and quantitative boundary readout on boxes}
\author{
  Haobo Ma\thanks{AELF PTE LTD., \#14-02, Marina Bay Financial Centre Tower 1, 8 Marina Blvd, Singapore 018981. Email: \texttt{auric@aelf.io}}
  \and
  Wenlin Zhang\thanks{Corresponding author. National University of Singapore, Singapore. Email: \texttt{e1327962@u.nus.edu}}
}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We study a variational inverse problem for primitives of constant
differential forms on axis-aligned boxes. Given a smooth $(k-1)$-form $\eta$
with $\dd\eta=\dd x_1\wedge\cdots\wedge\dd x_k$, we ask how much of the
boundary trace of $\eta$ can be recovered from sharp, or nearly sharp,
saturation of the coefficient-$L^\infty$ norm in the interior. The analysis
starts from the midpoint-centered radial homotopy operator. On the class of
coordinate-monomial forms $f\,\dd x_I$ we prove the sharp estimate
$\norm{K_k\omega}_{\coeff,\infty}\le (2k)^{-1}\norm{\omega}_{\coeff,\infty}$
and identify the precise incidence obstruction to extending this bound to
larger orientation classes. For a box $R=\prod_{j=1}^k[0,L_j]$ we compute the
extremal value $m(R)=(2\sum_{j=1}^kL_j^{-1})^{-1}$ for primitives of
$\dd x_1\wedge\cdots\wedge\dd x_k$, construct an explicit affine minimizer
$\eta_R$, and establish a linear stability estimate which controls the
boundary trace by the interior norm gap
$\norm{\eta}_{\coeff,\infty}-m(R)$. In particular, every minimizer has the
canonical trace on each codimension-one face, while near-minimizers determine
that trace quantitatively. A slicing argument extends the readout mechanism to
constant coordinate-monomial forms on higher-dimensional boxes, and a one-cube
Whitney construction yields the corresponding cubical analogue.
\end{abstract}

\bigskip
\noindent\textbf{2020 Mathematics Subject Classification.} 58A12, 35R30, 49Q20.

\noindent\textbf{Keywords.} inverse problem, stability estimate,
differential forms, homotopy operator, Whitney forms.

\section{Introduction}

Inverse problems are often formulated through a measurement operator: one
observes boundary or exterior data and seeks to recover an inaccessible
coefficient, source, or geometry. The present paper studies a different, but
still genuinely inverse, mechanism. The forward constraint is an interior
exactness relation, and the inverse information is encoded by sharp saturation
of an optimal coefficient-$L^\infty$ bound for a primitive. More precisely, we
ask whether a sharp interior bound determines the boundary trace of the
primitive. For constant differential forms on axis-aligned boxes the answer is
yes, and the determination is quantitative.

The question is motivated by local reconstruction problems in which a
primitive is controlled in an interior norm while the quantity exchanged
between neighboring cells or charts is its boundary trace. In such settings it
is natural to ask whether an optimal interior estimate already fixes the
boundary data. The box geometry considered here is simple enough for the sharp
constants to be computed explicitly, but rich enough to display a clear
inverse phenomenon: a small deficit from the extremal coefficient norm forces
a proportionally small boundary-trace error. In this sense, the paper provides
a local cubical model in which sharp interior estimates carry explicit inverse
information at the boundary.

Let
\[
I^k=[0,1]^k,
\qquad
\omega_0:=\dd x_1\wedge\cdots\wedge \dd x_k,
\]
and consider
\begin{equation}
\label{eq:intro-extremal-problem}
m_k:=
\inf\left\{
\norm{\eta}_{\coeff,\infty}:\ \eta\in\Omega^{k-1}_{\mathrm{sm}}(I^k),\
\dd\eta=\omega_0
\right\}.
\end{equation}
Here $\norm{\cdot}_{\coeff,\infty}$ denotes the supremum norm of the coordinate
coefficients. If the classical radial homotopy operator is centered at
$c=(1/2,\dots,1/2)$, one obtains the explicit primitive
\begin{equation}
\label{eq:intro-explicit-minimizer}
\eta_\ast
=
\frac{1}{k}\sum_{j=1}^k (-1)^{j-1}\left(x_j-\frac12\right)
\dd x_1\wedge\cdots\wedge\widehat{\dd x_j}\wedge\cdots\wedge \dd x_k,
\end{equation}
and therefore $m_k\le 1/(2k)$. The main point of the paper is that this upper
bound is exact and that the corresponding equality and near-equality regimes
determine the boundary trace.

\medskip
\noindent\textbf{Main results.}
The analysis has three components. First, we isolate the class on which the
midpoint-centered homotopy operator carries the sharp coefficient bound
\[
\norm{K_k\omega}_{\coeff,\infty}\le \frac{1}{2k}\norm{\omega}_{\coeff,\infty}.
\]
This holds for coordinate-monomial forms $\omega=f\,\dd x_I$, and
Proposition~\ref{prop:incidence-criterion} identifies the precise
combinatorial obstruction which prevents the same factor from extending to
larger orientation classes.

Second, we turn the sharp constant into an inverse statement. On every
axis-aligned box $R=\prod_{j=1}^k[0,L_j]$ we compute the exact extremal value
\[
m(R)=\frac{1}{2\sum_{j=1}^k L_j^{-1}}
\]
and construct an affine primitive $\eta_R$ attaining it. If $\eta$ is any
smooth primitive of $\dd x_1\wedge\cdots\wedge\dd x_k$ on $R$, then
\[
\sum_{F\subset\partial R}
\norm{\Tr_F(\eta-\eta_R)}_{L^1(F;\coeff)}
\le
2P_1(R)\bigl(\norm{\eta}_{\coeff,\infty}-m(R)\bigr).
\]
Here the sum extends over the codimension-one faces of $R$, $\Tr_F$ denotes
the pullback to the face $F$, and $P_1(R)$ is the anisotropic $\ell_1$
perimeter. In particular, every minimizer has the canonical trace on each
face, while near-minimizers determine that trace with linear stability. On the
unit cube this gives the explicit boundary formula
\[
\iota_{j,\varepsilon}^\ast\eta
=
(-1)^{j-1}\frac{\varepsilon-\frac12}{k}\,
\dd x_1\wedge\cdots\wedge\widehat{\dd x_j}\wedge\cdots\wedge \dd x_k,
\qquad
\varepsilon\in\{0,1\}.
\]
Third, by slicing we extend the same readout mechanism to constant
coordinate-monomial data on higher-dimensional boxes. Under the weaker
hypothesis that $\dd\omega$ is coordinate-monomial, the homotopy identity also
gives the sharp decomposition
\[
\omega=\dd\eta+E,
\qquad
\norm{E}_{\coeff,\infty}\le \frac{1}{2(k+1)}\norm{\dd\omega}_{\coeff,\infty},
\]
again with sharp constant. We conclude with a one-cube Whitney analogue, which
shows that the same structure persists at the level of cubical cochains.

\medskip
\noindent\textbf{Relation with earlier work.}
The homotopy formula used here is classical; see, for instance,
\cite[section~3]{BottTu1982} and
\cite{Whitney1957,Dodziuk1976,Dupont1976,DupontBook1978,Albert2020,
ArnoldFalkWinther2010,ArnoldAwanou2014}. For bounded right inverses on more
general domains we refer to
\cite{Bogovskii1979,IwaniecLutoborski1993,CostabelMcIntosh2010,
GuzmanSalgado2021,ChaumontFreletLichtVohralik2025}. On the anisotropic side,
related sharp constants for scalar Poincar\'e inequalities on convex domains
were studied in \cite{PayneWeinberger1960,AcostaDuran2004}. Our argument
belongs to the circle of ideas surrounding crystalline isoperimetry, Wulff
shapes, and calibration methods; see
\cite{Taylor1978,Fonseca1991,FonsecaMuller1991,DacorognaPfister1992,
FigalliZhang2022,DeMason2024,Strang1983,Nozawa1990,Grieser2006,Strang2010,
CasellesFaccioloMeinhardt2009,CasellesChambolleMollNovaga2008,Leonardi2015}.
The quotient $\Vol(E)/P_1(E)$ appearing below is the crystalline analogue of
the Cheeger ratio; compare
\cite{Cheeger1970,Leonardi2015,Parini2011}. For rigidity and uniqueness
phenomena in related Cheeger-type problems we also refer to
\cite{CasellesChambolleNovaga2007,CasellesChambolleNovaga2011}.
Our contribution is not a new homotopy formula or a new calibration
principle. Rather, we extract from these classical ingredients a sharp
coefficient-$L^\infty$ estimate on the restricted class $\omega=f\,\dd x_I$,
identify the combinatorial obstruction beyond that class, and show that the
resulting extremal problem yields quantitative boundary determination on
boxes. From the viewpoint of geometric discretization, the box theorem may be
read as a local statement on when a sharp interior certificate fixes the
boundary data.

From the inverse-problems viewpoint, the present problem is complementary to
the usual coefficient-recovery setting of Calder\'on type; compare
\cite{Calderon1980,Uhlmann2009}. Instead of recovering an interior coefficient
from a boundary measurement map, we recover boundary traces from sharp or
near-sharp interior norm saturation.

\subsection{Theorem-by-theorem comparison with prior work}

Definition~\ref{def:decomposable-form} introduces the paper's
coordinate-monomial class and is best viewed as a manuscript-specific
restriction of the ambient de~Rham complex rather than a standard named object
from the literature. Definition~\ref{def:homotopy-operator}, by contrast, is
classical: it is the usual homotopy operator on a star-shaped domain in the
form used by Dupont and Bott--Tu \cite{DupontBook1978,BottTu1982}. The new
coefficient-level content in Section~2 is carried by
Lemma~\ref{lem:contraction-estimate} and
Proposition~\ref{prop:incidence-criterion}, which isolate the exact $1/2$
contraction factor on a single coordinate-monomial term and the exact
incidence obstruction when several orientations are allowed. The main
continuous theorem, Theorem~\ref{thm:optimal-continuous-inverse}, should
therefore be read as a sharp coefficient-$L^\infty$ refinement of the
classical homotopy identity, and
Proposition~\ref{thm:homotopy-decomposition} together with
Corollary~\ref{cor:curl-reconstruction} as derived consequences with explicit
optimal defect constants.

The anisotropic variational input starts from classical weak duality.
Lemma~\ref{lem:boundary-calibration} is the continuous max-flow/min-cut
estimate in the $\ell_\infty/\ell_1$ pair, in the spirit of
\cite{Strang2010}, and Lemma~\ref{lem:box-self-cheeger} records the classical
fact that boxes are extremal objects in this crystalline geometry; compare the
same max-flow/min-cut framework and the calibrable-set literature
\cite{Strang2010,CasellesChambolleMollNovaga2008}. The genuinely new inverse
step is Proposition~\ref{prop:quantitative-boundary-saturation}, which turns
boundary saturation on an optimal domain into an exact deficit identity, and
its decisive specialization is Theorem~\ref{thm:box-boundary-stability}, where
that identity becomes an explicit affine minimizer and a linear stability
estimate recovering all codimension-one traces from the interior norm gap.

Definition~\ref{def:face-trace} is again manuscript-specific notation, added
because the information recovered here lives on codimension-one faces rather
than in the interior. Corollary~\ref{thm:boundary-rigidity} is the unit-cube
specialization of the box theorem and yields the explicit trace formula
\eqref{eq:boundary-rigidity-formula}, while
Corollary~\ref{cor:explicit-minimizer} identifies the classical centered radial
primitive as an actual extremizer for the coefficient norm. Theorem~\ref{thm:sliced-active-readout}
appears to be new at the level stated: by a
slice-by-slice reduction, it extends the readout mechanism from full volume
forms on a $k$-box to arbitrary constant coordinate-monomial data on an
$n$-box and isolates exactly which active boundary coefficients are rigid.

In the discrete section, Proposition~\ref{prop:cubical-whitney} records
standard Whitney-form identities on one cell, now written in a cubical rather
than simplicial notation and consistent with the classical Whitney--Dupont
picture and later cubical FEEC constructions
\cite{Whitney1957,Dupont1976,ArnoldAwanou2014}. The paper-specific discrete
definitions are Definition~\ref{def:whitney-decomposable} and
Proposition~\ref{prop:whitney-characterization}, which single out the Whitney
coordinate-monomial class on which the sharp estimate survives. With that class
in place, Proposition~\ref{thm:discrete-stokes-poincare} is the discrete
counterpart of the continuous sharp decomposition theorem, retaining the same
optimal factor $(2(k+1))^{-1}$, and Corollary~\ref{cor:hypercube-gradient} is
its one-cochain specialization with the optimal $1/4$ remainder bound.

\medskip
\noindent\textbf{Outline of the paper.}
Section~2 introduces the notation and recalls the centered homotopy operator.
Section~3 proves the sharp coefficient estimate on the coordinate-monomial
class together with the decomposition theorem. Section~4 contains the inverse
part of the paper: the extremal primitive problem on boxes, the quantitative
boundary readout theorem, and the sliced extension to higher-dimensional
boxes. Section~5 gives the one-cube Whitney counterpart, and Appendix~A
collects the auxiliary Whitney identities used there.

\section{Preliminaries and the centered homotopy operator}

We begin by fixing notation and recalling the midpoint-centered radial
homotopy operator in the form needed below.

Fix $n\ge 1$ and write
\[
I^n:=[0,1]^n,
\qquad
c:=\left(\frac12,\dots,\frac12\right).
\]
For $1\le k\le n$, let $\Omega^k_{\mathrm{sm}}(I^n)$ denote the smooth
$k$-forms on a neighborhood of $I^n$, restricted to the cube. If
$R\subset\RR^n$ is any axis-aligned box, we use $\Omega^k_{\mathrm{sm}}(R)$ for
the analogous space of smooth $k$-forms on a neighborhood of $R$, restricted
to $R$. If $F$ is a codimension-one face of such a box and
$\iota_F:F\hookrightarrow R$ denotes the inclusion, we write
\[
\Tr_F(\eta):=\iota_F^\ast\eta
\]
for the face trace of a differential form $\eta$ on $R$. If
\[
\omega=\sum_{\abs{I}=k} a_I(x)\,\dd x_I,
\]
we set
\[
\norm{\omega}_{\coeff,\infty}:=
\max_{\abs{I}=k}\sup_{x\in I^n}\abs{a_I(x)}.
\]
All $L^\infty$ estimates below are taken with respect to this coordinate norm.
For a form evaluated at a fixed point $y$, we write
\[
\abs{\omega(y)}_{\coeff,\infty}:=
\max_{\abs{I}=k}\abs{a_I(y)}.
\]
If
\[
\beta=\sum_{\abs{J}=m} b_J(x)\,\dd x_J,
\]
we also write $\coeff_J(\beta):=b_J$ for the coefficient of $\dd x_J$.

\begin{remark}
Throughout, $\dd$ denotes the exterior derivative on smooth forms, whereas $d$
is reserved for the cubical coboundary used in Section~5.
\end{remark}

\begin{definition}[Coordinate-monomial form]
\label{def:decomposable-form}
A smooth $k$-form on $I^n$ is coordinate-monomial if
\[
\omega=f(x)\,\dd x_{i_1}\wedge\cdots\wedge \dd x_{i_k}
\]
for some smooth function $f$ and some ordered index set
$1\le i_1<\cdots<i_k\le n$.
\end{definition}

\begin{remark}
This requirement is stronger than decomposability in
$\Lambda^kT^\ast I^n$. For instance,
$(\dd x_1+\dd x_2)\wedge \dd x_3$ is decomposable in the exterior algebra, but
it is not coordinate-monomial.
\end{remark}

Let $X(x):=x-c$ be the radial vector field centered at $c$, and let
\[
H:[0,1]\times I^n\to I^n,
\qquad
H(t,x):=c+t(x-c).
\]
For each $t\in[0,1]$ we write $H_t(x):=H(t,x)$.

\begin{definition}[Radial homotopy operator]
\label{def:homotopy-operator}
For $1\le k\le n$ define
\[
K_k:\Omega^k_{\mathrm{sm}}(I^n)\to \Omega^{k-1}_{\mathrm{sm}}(I^n)
\]
by
\begin{equation}
\label{eq:homotopy-operator}
(K_k\omega)(x)
:=
\int_0^1 t^{k-1}\,\bigl(\iota_{x-c}\omega\bigr)\!\bigl(c+t(x-c)\bigr)\,\dd t.
\end{equation}
Equivalently, if
$\omega=\sum_{\abs{I}=k}a_I(y)\,\dd y_I$, then
\begin{equation}
\label{eq:homotopy-operator-explicit}
(K_k\omega)(x)
=
\int_0^1 t^{k-1}
\sum_{\abs{I}=k}\sum_{r=1}^k (-1)^{r-1}(x_{i_r}-c_{i_r})\,
a_I\bigl(c+t(x-c)\bigr)\,
\dd x_{I\setminus\{i_r\}}\,\dd t.
\end{equation}
\end{definition}

This is the standard radial homotopy formula centered at the midpoint of the
cube; see \cite[Sec.~3]{BottTu1982} and
\cite{Dupont1976,DupontBook1978,Albert2020}. In particular, for
$\omega_0=\dd x_1\wedge\cdots\wedge \dd x_k$ one obtains
\begin{equation}
\label{eq:explicit-minimizer}
K_k\omega_0
=
\frac{1}{k}\sum_{j=1}^k (-1)^{j-1}\left(x_j-\frac12\right)
\dd x_1\wedge\cdots\wedge\widehat{\dd x_j}\wedge\cdots\wedge \dd x_k.
\end{equation}

\begin{lemma}[Contraction estimate]
\label{lem:contraction-estimate}
For every coordinate-monomial form
\[
\omega=f(y)\,\dd x_{i_1}\wedge\cdots\wedge \dd x_{i_k}
\in \Omega^k_{\mathrm{sm}}(I^n)
\]
and all $x,y\in I^n$ one has
\[
\abs{\iota_{x-c}\omega(y)}_{\coeff,\infty}\le
\frac12\norm{\omega}_{\coeff,\infty}.
\]
In particular,
\[
\norm{\iota_X\omega}_{\coeff,\infty}\le \frac12\norm{\omega}_{\coeff,\infty}.
\]
\end{lemma}

\begin{proof}
For such a form,
\[
\iota_{x-c}\omega(y)=
\sum_{r=1}^k (-1)^{r-1} f(y)\bigl(x_{i_r}-\tfrac12\bigr)\,
\dd x_{i_1}\wedge\cdots\wedge\widehat{\dd x_{i_r}}\wedge\cdots\wedge
\dd x_{i_k}.
\]
Each component of $x-c$ has absolute value at most $1/2$ on $I^n$, and the
coefficient function of $\iota_{x-c}\omega(y)$ is therefore bounded in absolute
value by $\frac12\abs{f(y)}\le \frac12\norm{\omega}_{\coeff,\infty}$.
\end{proof}

\begin{proposition}[Incidence criterion]
\label{prop:incidence-criterion}
Let
\[
\omega=\sum_{\abs{I}=k} a_I(x)\,\dd x_I
\in\Omega^k_{\mathrm{sm}}(I^n).
\]
For each $(k-1)$-index $J$ define
\[
m_J(\omega):=
\#\{i\notin J:\ a_{J\cup\{i\}}\not\equiv 0\},
\qquad
m(\omega):=\max_{\abs{J}=k-1}m_J(\omega).
\]
Here $J\cup\{i\}$ is always written in increasing order.
Then
\[
\norm{\iota_X\omega}_{\coeff,\infty}
\le
\frac{m(\omega)}{2}\norm{\omega}_{\coeff,\infty}.
\]
In particular, let $\mathcal M$ be a family of $k$-indices and let
$\Omega^k_{\mathcal M}(I^n)$ denote the subspace of forms
\[
\omega=\sum_{I\in\mathcal M} a_I(x)\,\dd x_I.
\]
Then the universal bound
\[
\norm{\iota_X\omega}_{\coeff,\infty}\le \frac12\norm{\omega}_{\coeff,\infty}
\qquad
\text{for every }\omega\in\Omega^k_{\mathcal M}(I^n)
\]
holds if and only if, for each $(k-1)$-index $J$, there is at most one
$i\notin J$ such that $J\cup\{i\}\in\mathcal M$.
\end{proposition}

\begin{proof}
For a fixed $(k-1)$-index $J$, the coefficient of $\dd x_J$ in $\iota_X\omega$
is
\[
\sum_{i\notin J}\sigma(J,i)\bigl(x_i-\tfrac12\bigr)a_{J\cup\{i\}}(x),
\]
where $\sigma(J,i)\in\{\pm1\}$ is the sign determined by the insertion of $i$
into the ordered set $J$. At each point $x$, at most $m_J(\omega)$ summands are
present, and each has absolute value bounded by
$\frac12\norm{\omega}_{\coeff,\infty}$. This gives the estimate.

For the final statement, the incidence condition immediately implies
$m(\omega)\le 1$ for every $\omega\in\Omega^k_{\mathcal M}(I^n)$, hence the
bound with factor $1/2$ follows from the estimate already proved.

Conversely, suppose that the incidence condition fails. Then there exist a
$(k-1)$-index $J$ and distinct indices $i,\ell\notin J$ such that
$J\cup\{i\}$ and $J\cup\{\ell\}$ both belong to $\mathcal M$. Consider the two
allowed monomials $\dd x_{J\cup\{i\}}$ and $\dd x_{J\cup\{\ell\}}$, which
share the same $(k-1)$-factor $\dd x_J$. Choosing the sign in
\[
\omega_0:=\dd x_{J\cup\{i\}}\pm \dd x_{J\cup\{\ell\}}
\]
so that the corresponding contributions to the $\dd x_J$-coefficient add, one
has $\norm{\omega_0}_{\coeff,\infty}=1$. At the corner of $I^n$ with
$x_i=x_\ell=1$ if the common sign is positive and $x_i=x_\ell=0$ if the common
sign is negative, the coefficient of $\dd x_J$ in $\iota_X\omega_0$ has
absolute value
\[
\abs{x_i-\tfrac12}+\abs{x_\ell-\tfrac12}=1.
\]
Hence $\norm{\iota_X\omega_0}_{\coeff,\infty}=1$, so the factor $1/2$ cannot
hold on $\Omega^k_{\mathcal M}(I^n)$.
\end{proof}

\begin{remark}
\label{rem:non-coord-monomial}
The restriction to coordinate-monomial forms is essential for the coefficient
norm. On $I^3$ consider
\[
\omega:=\dd x_1\wedge \dd x_3+\dd x_2\wedge \dd x_3.
\]
Then $\norm{\omega}_{\coeff,\infty}=1$, whereas
\[
\iota_X\omega
=
\bigl(x_1-\tfrac12+x_2-\tfrac12\bigr)\dd x_3
-\bigl(x_3-\tfrac12\bigr)(\dd x_1+\dd x_2),
\]
so $\norm{\iota_X\omega}_{\coeff,\infty}=1$. A direct computation gives
\[
K_2\omega
=
\frac{x_1+x_2-1}{2}\,\dd x_3
-\frac{x_3-\tfrac12}{2}\,(\dd x_1+\dd x_2),
\]
hence $\norm{K_2\omega}_{\coeff,\infty}=\frac12>\frac14$. Thus the estimate
proved in Section~3 is genuinely tied to the coordinate-monomial setting.
\end{remark}

\section{Sharp coefficient bounds for coordinate-monomial data}

We now isolate the regime in which the centered homotopy operator carries the
factor $1/(2k)$ in the coefficient norm. For general forms, contraction by the
radial field may combine several active coefficients into a single output
coefficient. Proposition~\ref{prop:incidence-criterion} makes this obstruction
explicit. On the coordinate-monomial class no such interaction occurs, and the
optimal estimate follows directly from the homotopy formula.

\begin{theorem}[Sharp coefficient bound for the centered homotopy operator]
\label{thm:optimal-continuous-inverse}
Let $1\le k\le n$, and set $K_{n+1}:=0$. Then
\begin{equation}
\label{eq:homotopy-identity}
\dd(K_k\omega)+K_{k+1}(\dd\omega)=\omega
\end{equation}
for every $\omega\in\Omega^k_{\mathrm{sm}}(I^n)$. If, in addition, $\omega$ is
coordinate-monomial, then
\begin{equation}
\label{eq:continuous-supnorm-bound}
\norm{K_k\omega}_{\coeff,\infty}\le \frac{1}{2k}\norm{\omega}_{\coeff,\infty}.
\end{equation}
The constant $1/(2k)$ is sharp on the class of closed coordinate-monomial
$k$-forms. In fact, if
$R:\Omega^k_{\mathrm{sm}}(I^n)\to\Omega^{k-1}_{\mathrm{sm}}(I^n)$ is linear and
$\dd(R\omega)=\omega$ for every closed coordinate-monomial $k$-form $\omega$,
then necessarily
\[
\sup\left\{
\frac{\norm{R\omega}_{\coeff,\infty}}{\norm{\omega}_{\coeff,\infty}}:\ 
\omega\neq 0,\ \omega \text{ closed coordinate-monomial}
\right\}
\ge \frac{1}{2k}.
\]
\end{theorem}

\begin{proof}
Write $\omega=\sum_{\abs{I}=k}a_I(y)\,\dd y_I$ in a neighborhood of $I^n$ and
set $H(t,x)=c+t(x-c)$. Since
\[
H^\ast(\dd y_j)=t\,\dd x_j+(x_j-c_j)\,\dd t,
\]
each wedge product contains at most one $\dd t$ factor. Expanding $H^\ast\omega$
therefore gives
\[
H^\ast\omega=A+\dd t\wedge B,
\]
where
\[
A(t,x)=t^k\sum_{\abs{I}=k} a_I(H(t,x))\,\dd x_I
\]
and
\[
B(t,x)=
t^{k-1}\sum_{\abs{I}=k}\sum_{r=1}^k (-1)^{r-1}
\bigl(x_{i_r}-c_{i_r}\bigr)a_I(H(t,x))\,
\dd x_{I\setminus\{i_r\}}.
\]
Thus
\[
K_k\omega(x)=\int_0^1 B(t,x)\,\dd t.
\]

Applying the same expansion to $\dd\omega$ yields
\[
H^\ast(\dd\omega)=C+\dd t\wedge D
\]
for a suitable pair $(C,D)$, with
\[
K_{k+1}(\dd\omega)(x)=\int_0^1 D(t,x)\,\dd t.
\]
Since $\dd(H^\ast\omega)=H^\ast(\dd\omega)$ and
$\dd=\dd_x+\dd t\wedge\partial_t$ on $[0,1]\times I^n$, one has
\[
\dd(A+\dd t\wedge B)=\dd_xA+\dd t\wedge(\partial_tA-\dd_xB).
\]
Comparing the $\dd t$-components gives
\[
\partial_tA-\dd_xB=D.
\]
Integrating in $t$ from $0$ to $1$ yields
\[
A(1,x)-A(0,x)=
\dd_x\!\left(\int_0^1 B(t,x)\,\dd t\right)+\int_0^1 D(t,x)\,\dd t.
\]
Since $A(1,x)=\omega(x)$ and $A(0,x)=0$ for $k\ge 1$, this is exactly
\eqref{eq:homotopy-identity}.

Assume now that $\omega=f(y)\,\dd x_I$ is coordinate-monomial. By
\eqref{eq:homotopy-operator}, each coefficient of $K_k\omega$ at a point $x$ is
bounded by
\[
\int_0^1 t^{k-1}\cdot
\frac12\abs{f(c+t(x-c))}\,\dd t
\]
by Lemma~\ref{lem:contraction-estimate}. Hence
\[
\norm{K_k\omega}_{\coeff,\infty}
\le
\int_0^1 t^{k-1}\,\dd t\cdot \frac12\norm{\omega}_{\coeff,\infty}
=
\frac{1}{2k}\norm{\omega}_{\coeff,\infty}.
\]

For sharpness, consider the constant form
\[
\omega_0:=\dd x_1\wedge\cdots\wedge \dd x_k
\]
and the $k$-face
\[
F:=[0,1]^k\times\{0\}^{n-k}\subset I^n.
\]
We endow $F$ with the orientation induced by
$\dd x_1\wedge\cdots\wedge \dd x_k$. If
$\eta\in\Omega^{k-1}_{\mathrm{sm}}(I^n)$ satisfies $\dd\eta=\omega_0$, then
Stokes' theorem gives
\[
1
=
\int_F \iota_F^\ast\omega_0
=
\int_{\partial F}\iota_{\partial F}^\ast\eta.
\]
Since $\partial F$ has $2k$ faces of unit $(k-1)$-volume,
\[
1
\le
2k\,\norm{\eta}_{\coeff,\infty}.
\]
Hence every primitive of $\omega_0$ has coefficient norm at least $1/(2k)$.
In particular, if $R$ is linear and satisfies $\dd(R\omega)=\omega$ for every
closed coordinate-monomial $k$-form, then $R\omega_0$ is such a primitive and
\[
\norm{R\omega_0}_{\coeff,\infty}\ge \frac{1}{2k}
=
\frac{1}{2k}\norm{\omega_0}_{\coeff,\infty}.
\]
On the other hand, \eqref{eq:explicit-minimizer} gives
\[
K_k\omega_0=
\frac{1}{k}\sum_{j=1}^k (-1)^{j-1}\left(x_j-\frac12\right)
\dd x_1\wedge\cdots\wedge\widehat{\dd x_j}\wedge\cdots\wedge \dd x_k,
\]
and thus
\[
\norm{K_k\omega_0}_{\coeff,\infty}=\frac{1}{2k}.
\]
This completes the proof.
\end{proof}

\begin{proposition}[A sharp decomposition with controlled defect]
\label{thm:homotopy-decomposition}
Let $1\le k\le n-1$ and let $\omega\in\Omega^k_{\mathrm{sm}}(I^n)$. Assume
that $\dd\omega$ is coordinate-monomial. If
\[
\eta:=K_k\omega,
\qquad
E:=K_{k+1}(\dd\omega),
\]
then
\begin{equation}
\label{eq:exact-plus-defect}
\omega=\dd\eta+E,
\qquad
\dd E=\dd\omega,
\end{equation}
and
\begin{equation}
\label{eq:optimal-defect-bound}
\norm{E}_{\coeff,\infty}\le
\frac{1}{2(k+1)}\norm{\dd\omega}_{\coeff,\infty}.
\end{equation}
The constant $1/(2(k+1))$ cannot be improved in any estimate of this form that
is required to hold uniformly for all $\omega$ with coordinate-monomial
exterior derivative.
\end{proposition}

\begin{proof}
The identity \eqref{eq:exact-plus-defect} is just
\eqref{eq:homotopy-identity} with $\eta=K_k\omega$ and
$E=K_{k+1}(\dd\omega)$. Since $\dd\omega$ is coordinate-monomial,
\eqref{eq:continuous-supnorm-bound} applied to $K_{k+1}$ yields
\eqref{eq:optimal-defect-bound}.

To prove sharpness, it is convenient to formulate the competing estimate as a
uniform property. Take
\[
\nu_0:=\dd x_1\wedge\cdots\wedge \dd x_{k+1}
\]
and set $\omega_0:=K_{k+1}\nu_0$. Then $\dd\omega_0=\nu_0$ and
$\norm{\dd\omega_0}_{\coeff,\infty}=1$. Suppose that a constant $C$ had the
following uniform property: for every $\omega$ with coordinate-monomial
exterior derivative there exist $\phi$ and $\widetilde E$ such that
\[
\omega=\dd\phi+\widetilde E,
\qquad
\dd\widetilde E=\dd\omega,
\qquad
\norm{\widetilde E}_{\coeff,\infty}\le C\norm{\dd\omega}_{\coeff,\infty}
\]
Applying this property to $\omega_0$ yields
\[
\omega_0=\dd\phi+\widetilde E,
\qquad
\dd\widetilde E=\nu_0,
\qquad
\norm{\widetilde E}_{\coeff,\infty}\le C.
\]
Thus $\widetilde E$ is a primitive of $\nu_0$ whose norm is controlled by $C$.
By Stokes' theorem on $I^{k+1}$,
\[
1
=
\int_{I^{k+1}}\nu_0
=
\int_{\partial I^{k+1}}\iota_{\partial I^{k+1}}^\ast\widetilde E.
\]
Since $\partial I^{k+1}$ has $2(k+1)$ faces of unit $k$-volume,
\[
1\le 2(k+1)\norm{\widetilde E}_{\coeff,\infty}.
\]
Hence $C\ge 1/(2(k+1))$. Equality is attained by the present construction,
because
\[
E=K_{k+1}\nu_0=\omega_0
\]
and \eqref{eq:explicit-minimizer} with $k$ replaced by $k+1$ gives
$\norm{E}_{\coeff,\infty}=1/(2(k+1))$.
\end{proof}

\begin{corollary}[Potential reconstruction]
\label{cor:curl-reconstruction}
Let $n\ge 2$, let $v$ be a smooth vector field on a neighborhood of $I^n$, and
set
\[
\omega:=\sum_{i=1}^n v_i\,\dd x_i.
\]
Assume that $\dd\omega$ is coordinate-monomial. Then
$U:=K_1\omega$ satisfies
\begin{equation}
\label{eq:curl-potential-bound}
\norm{v-\nabla U}_{\coeff,\infty}\le
\frac14\norm{\dd\omega}_{\coeff,\infty}.
\end{equation}
The constant $1/4$ is optimal in the sense that no smaller uniform constant can
hold in \eqref{eq:curl-potential-bound}.
\end{corollary}

\begin{proof}
By \eqref{eq:homotopy-identity} with $k=1$,
\[
\omega=\dd U+K_2(\dd\omega).
\]
The coefficient vector of $\omega-\dd U$ is exactly $v-\nabla U$, so
\eqref{eq:continuous-supnorm-bound} yields \eqref{eq:curl-potential-bound}.

For sharpness, let
\[
\nu_0:=\dd x_1\wedge \dd x_2,
\qquad
\alpha:=K_2\nu_0.
\]
Then $\dd\alpha=\nu_0$ and $\norm{\nu_0}_{\coeff,\infty}=1$. For every smooth
function $U$,
\[
\dd(\alpha-\dd U)=\nu_0.
\]
Let
\[
Q:=[0,1]^2\times\{0\}^{n-2}\subset I^n.
\]
Equip $Q$ with the orientation induced by $\dd x_1\wedge\dd x_2$.
Stokes' theorem on $Q$ gives
\[
1
=
\int_Q \iota_Q^\ast\nu_0
=
\int_{\partial Q}\iota_{\partial Q}^\ast(\alpha-\dd U).
\]
Since $\partial Q$ has four edges of unit length,
\[
1\le 4\norm{\alpha-\dd U}_{\coeff,\infty}.
\]
Hence no smaller uniform constant can hold.
\end{proof}

\section{Extremal primitives and boundary readout on boxes}

We now turn to the inverse statement. The sharp lower bound in
\eqref{eq:intro-extremal-problem} is governed by the crystalline perimeter
\[
P_1(E):=\int_{\partial E}\abs{\nu_E}_1\,\dd\Haus^{n-1},
\qquad
\abs{\nu_E}_1:=\sum_{i=1}^n \abs{(\nu_E)_i}.
\]
For background on crystalline isoperimetry, the Wulff theorem, and calibrable
sets we refer to
\cite{Taylor1978,Fonseca1991,FonsecaMuller1991,DacorognaPfister1992,
FigalliZhang2022,DeMason2024} and
\cite{Strang1983,Nozawa1990,Grieser2006,Strang2010,
CasellesFaccioloMeinhardt2009,CasellesChambolleMollNovaga2008,
Leonardi2015,Parini2011}. The argument proceeds in three steps. We first prove
a boundary saturation identity for domains that attain the optimal ratio
$\Vol(\Omega)/P_1(\Omega)$. We then verify that every axis-aligned box enjoys
this property, which leads to a linear stability estimate for the boundary
trace of near-minimizers. Finally, a slicing argument transfers the same
mechanism to constant coordinate-monomial forms on higher-dimensional boxes.

\begin{lemma}[Anisotropic calibration bound]
\label{lem:boundary-calibration}
Let $\Omega\subset\RR^n$ be a bounded Lipschitz domain, and set
\[
\mathrm{vol}:=\dd x_1\wedge\cdots\wedge \dd x_n.
\]
Define
\[
\mathcal C_\infty(\Omega):=
\inf\Bigl\{
\norm{V}_{L^\infty(\Omega;\ell_\infty)}:\ 
V\in C^\infty(U;\RR^n),\ U\supset\overline{\Omega}\text{ open},
\ \Div V=1\text{ on }\Omega
\Bigr\}.
\]
Equivalently, $\mathcal C_\infty(\Omega)$ is the infimum of
$\norm{\eta}_{\coeff,\infty}$ over smooth $(n-1)$-forms $\eta$ defined on a
neighborhood of $\overline{\Omega}$ and satisfying $\dd\eta=\mathrm{vol}$ on
$\Omega$. Then every admissible vector field $V$ and every Lipschitz subdomain
$E\subset\Omega$ satisfy
\[
\Vol(E)\le \norm{V}_{L^\infty(\Omega;\ell_\infty)}\,P_1(E).
\]
Consequently,
\[
\mathcal C_\infty(\Omega)\ge
\sup\left\{
\frac{\Vol(E)}{P_1(E)}:\ E\subset\Omega,\ E\text{ Lipschitz}
\right\}.
\]
\end{lemma}

\begin{proof}
For a smooth vector field $V=\sum_{j=1}^n V_j\,\partial_{x_j}$, the form
\[
\eta:=\iota_V\mathrm{vol}
=
\sum_{j=1}^n (-1)^{j-1}V_j\,
\dd x_1\wedge\cdots\wedge\widehat{\dd x_j}\wedge\cdots\wedge \dd x_n
\]
satisfies
$\dd\eta=(\Div V)\mathrm{vol}$, and
\[
\norm{\eta}_{\coeff,\infty}=
\norm{V}_{L^\infty(\Omega;\ell_\infty)}.
\]
Conversely, every smooth $(n-1)$-form on a neighborhood of $\overline{\Omega}$
admits a unique representation of this type, so the two definitions of
$\mathcal C_\infty(\Omega)$ agree. The divergence theorem on Lipschitz domains
used below is standard; see \cite{Federer1969,EvansGariepy2015,Maggi2012}.

Now let $E\subset\Omega$ be a Lipschitz subdomain and let $V$ be admissible.
Then
\[
\Vol(E)
=
\int_E \Div V\,\dd x
=
\int_{\partial E}V\cdot \nu_E\,\dd\Haus^{n-1}
\le
\norm{V}_{L^\infty(\Omega;\ell_\infty)}
\int_{\partial E}\abs{\nu_E}_1\,\dd\Haus^{n-1}.
\]
The inequality uses the duality bound
$\abs{V\cdot \nu_E}\le \norm{V}_{\ell_\infty}\abs{\nu_E}_1$ pointwise on
$\partial E$, and therefore gives the required estimate.
\end{proof}

\begin{proposition}[Quantitative boundary saturation on optimal domains]
\label{prop:quantitative-boundary-saturation}
Let $\Omega\subset\RR^n$ be a bounded Lipschitz domain and assume that
\[
\mathcal C_\infty(\Omega)=\frac{\Vol(\Omega)}{P_1(\Omega)}.
\]
If $V\in C^\infty(U;\RR^n)$ is admissible for $\mathcal C_\infty(\Omega)$,
that is, $\Div V=1$ on $\Omega$, and if
\[
M:=\norm{V}_{L^\infty(\Omega;\ell_\infty)},
\]
then
\begin{equation}
\label{eq:general-boundary-defect}
\int_{\partial\Omega}
\left(
M\abs{\nu_\Omega}_1-V\cdot \nu_\Omega
\right)\,\dd\Haus^{n-1}
=
P_1(\Omega)\left(M-\mathcal C_\infty(\Omega)\right).
\end{equation}
In particular, if $M=\mathcal C_\infty(\Omega)+\delta$, then
\[
\left\|
M\abs{\nu_\Omega}_1-V\cdot \nu_\Omega
\right\|_{L^1(\partial\Omega)}
=
\delta\,P_1(\Omega).
\]
Every minimizer $V$ for $\mathcal C_\infty(\Omega)$ therefore satisfies
\[
V\cdot \nu_\Omega
=
\mathcal C_\infty(\Omega)\abs{\nu_\Omega}_1
\qquad
\text{for $\Haus^{n-1}$-a.e.\ }x\in\partial\Omega.
\]
\end{proposition}

\begin{proof}
The pointwise duality estimate
\[
V\cdot \nu_\Omega\le \norm{V}_{\ell_\infty}\abs{\nu_\Omega}_1\le
M\abs{\nu_\Omega}_1
\]
shows that the integrand in \eqref{eq:general-boundary-defect} is
nonnegative. Since $\Div V=1$ on $\Omega$, the divergence theorem gives
\[
\int_{\partial\Omega}V\cdot \nu_\Omega\,\dd\Haus^{n-1}
=
\int_\Omega \Div V\,\dd x
=
\Vol(\Omega).
\]
Therefore
\[
\int_{\partial\Omega}
\left(
M\abs{\nu_\Omega}_1-V\cdot \nu_\Omega
\right)\,\dd\Haus^{n-1}
=
M P_1(\Omega)-\Vol(\Omega),
\]
which equals the right-hand side of \eqref{eq:general-boundary-defect} by the
assumption on $\mathcal C_\infty(\Omega)$. The remaining assertions follow at
once.
\end{proof}

\begin{lemma}[Axis-aligned boxes attain the crystalline ratio]
\label{lem:box-self-cheeger}
Let
\[
R:=\prod_{j=1}^k [0,L_j]\subset\RR^k,
\qquad
m_R:=\frac{1}{2\sum_{j=1}^k L_j^{-1}}.
\]
Then every Lipschitz subdomain $E\subset R$ satisfies
\begin{equation}
\label{eq:box-cheeger-inequality}
2\left(\sum_{j=1}^k \frac{1}{L_j}\right)\Vol(E)\le P_1(E).
\end{equation}
Consequently,
\[
\sup\left\{
\frac{\Vol(E)}{P_1(E)}:\ E\subset R,\ E\text{ Lipschitz}
\right\}
=
\frac{\Vol(R)}{P_1(R)}
=
m_R,
\]
and
\[
\mathcal C_\infty(R)=m_R.
\]
Moreover, the affine vector field
\begin{equation}
\label{eq:box-calibrator}
V_R(x):=
2m_R\sum_{j=1}^k \frac{x_j-\frac{L_j}{2}}{L_j}\,\partial_{x_j}
\end{equation}
is admissible and satisfies
\[
\norm{V_R}_{L^\infty(R;\ell_\infty)}=m_R.
\]
\end{lemma}

\begin{proof}
Fix $j\in\{1,\dots,k\}$ and consider the vector field
\[
W_j(x):=\left(x_j-\frac{L_j}{2}\right)\partial_{x_j}.
\]
Since $\Div W_j=1$, the divergence theorem on $E$ gives
\[
\Vol(E)
=
\int_E \Div W_j\,\dd x
=
\int_{\partial E}\left(x_j-\frac{L_j}{2}\right)(\nu_E)_j\,\dd\Haus^{k-1}.
\]
Because $\abs{x_j-\frac{L_j}{2}}\le \frac{L_j}{2}$ on $R$, it follows that
\[
\Vol(E)\le \frac{L_j}{2}\int_{\partial E}\abs{(\nu_E)_j}\,\dd\Haus^{k-1}.
\]
Multiplying by $2/L_j$ and summing over $j$ yields
\eqref{eq:box-cheeger-inequality}. Applying this inequality to $E=R$ shows
that $\Vol(R)/P_1(R)=m_R$, hence the displayed supremum is exactly $m_R$.
Lemma~\ref{lem:boundary-calibration} now gives
\[
\mathcal C_\infty(R)\ge m_R.
\]
On the other hand, the field $V_R$ in \eqref{eq:box-calibrator} satisfies
\[
\Div V_R
=
2m_R\sum_{j=1}^k \frac{1}{L_j}
=
1,
\]
while each component of $V_R$ has absolute value at most $m_R$ on $R$. Thus
$V_R$ is admissible and
\[
\mathcal C_\infty(R)\le
\norm{V_R}_{L^\infty(R;\ell_\infty)}
=
m_R.
\]
Hence $\mathcal C_\infty(R)=m_R$.
\end{proof}

\begin{theorem}[Quantitative boundary readout on axis-aligned boxes]
\label{thm:box-boundary-stability}
Let
\[
R:=\prod_{j=1}^k [0,L_j]\subset\RR^k,
\qquad
\omega_R:=\dd x_1\wedge\cdots\wedge \dd x_k,
\qquad
m_R:=\frac{1}{2\sum_{j=1}^k L_j^{-1}},
\]
and define
\begin{equation}
\label{eq:box-explicit-minimizer}
\eta_R
:=
2m_R\sum_{j=1}^k (-1)^{j-1}
\frac{x_j-\frac{L_j}{2}}{L_j}\,
\dd x_1\wedge\cdots\wedge\widehat{\dd x_j}\wedge\cdots\wedge \dd x_k.
\end{equation}
Then $\dd\eta_R=\omega_R$ and
$\norm{\eta_R}_{\coeff,\infty}=m_R$. Let
$\eta\in\Omega^{k-1}_{\mathrm{sm}}(R)$ satisfy $\dd\eta=\omega_R$, and set
\[
M:=\norm{\eta}_{\coeff,\infty}.
\]
For
\[
F_{j,\varepsilon}:=\{x_j=\varepsilon\},
\qquad
\varepsilon\in\{0,L_j\},
\]
let $\iota_{j,\varepsilon}:F_{j,\varepsilon}\hookrightarrow R$ denote the
inclusion. Writing
\[
\iota_{j,\varepsilon}^\ast\eta
=
f_{j,\varepsilon}\,
\dd x_1\wedge\cdots\wedge\widehat{\dd x_j}\wedge\cdots\wedge \dd x_k
\]
and setting
\[
s_{j,L_j}:=(-1)^{j-1},
\qquad
s_{j,0}:=-(-1)^{j-1},
\]
one has
\begin{equation}
\label{eq:box-deficit-identity}
\sum_{j=1}^k\sum_{\varepsilon\in\{0,L_j\}}
\int_{F_{j,\varepsilon}}
\left(
M-s_{j,\varepsilon}f_{j,\varepsilon}
\right)\,\dd\Haus^{k-1}
=
P_1(R)(M-m_R).
\end{equation}
If, for a $(k-1)$-form $\beta$ on a face $F$, we set
\[
\norm{\beta}_{L^1(F;\coeff)}:=
\int_F \abs{\beta(x)}_{\coeff,\infty}\,\dd\Haus^{k-1}(x),
\]
then
\begin{equation}
\label{eq:box-trace-stability}
\sum_{j=1}^k\sum_{\varepsilon\in\{0,L_j\}}
\norm{\iota_{j,\varepsilon}^\ast(\eta-\eta_R)}_{L^1(F_{j,\varepsilon};\coeff)}
\le
2P_1(R)(M-m_R).
\end{equation}
In particular, $M\ge m_R$, the value $m_R$ is sharp, and every minimizer has
the canonical face trace
\[
\iota_{j,\varepsilon}^\ast\eta
=
\iota_{j,\varepsilon}^\ast\eta_R
\qquad
\text{on every }F_{j,\varepsilon}.
\]
\end{theorem}

\begin{proof}
Every smooth $(k-1)$-form on $R$ can be written uniquely as
\[
\eta=
\sum_{j=1}^k (-1)^{j-1}u_j\,
\dd x_1\wedge\cdots\wedge\widehat{\dd x_j}\wedge\cdots\wedge \dd x_k,
\]
and then
\[
\dd\eta=(\Div V)\,\dd x_1\wedge\cdots\wedge \dd x_k,
\qquad
V:=\sum_{j=1}^k u_j\,\partial_{x_j}.
\]
Moreover,
\[
\norm{\eta}_{\coeff,\infty}=
\norm{V}_{L^\infty(R;\ell_\infty)}=M.
\]
The explicit formula \eqref{eq:box-explicit-minimizer} is exactly
$\eta_R=\iota_{V_R}\omega_R$, where $V_R$ is the field in
\eqref{eq:box-calibrator}. In particular $\dd\eta_R=\omega_R$ and
$\norm{\eta_R}_{\coeff,\infty}=m_R$.

Since $\dd\eta=\omega_R$, the associated vector field $V$ is admissible for
$\mathcal C_\infty(R)$. By Lemma~\ref{lem:box-self-cheeger} and
Proposition~\ref{prop:quantitative-boundary-saturation},
\[
\int_{\partial R}
\left(
M\abs{\nu_R}_1-V\cdot \nu_R
\right)\,\dd\Haus^{k-1}
=
P_1(R)(M-m_R).
\]
On each face $F_{j,\varepsilon}$ one has $\abs{\nu_R}_1=1$ and
\[
V\cdot \nu_R=s_{j,\varepsilon}f_{j,\varepsilon},
\]
which is exactly \eqref{eq:box-deficit-identity}.

The trace of $\eta_R$ on $F_{j,\varepsilon}$ is
\[
\iota_{j,\varepsilon}^\ast\eta_R
=
s_{j,\varepsilon}m_R\,
\dd x_1\wedge\cdots\wedge\widehat{\dd x_j}\wedge\cdots\wedge \dd x_k.
\]
Therefore
\[
\norm{\iota_{j,\varepsilon}^\ast(\eta-\eta_R)}_{L^1(F_{j,\varepsilon};\coeff)}
=
\int_{F_{j,\varepsilon}}
\abs{s_{j,\varepsilon}f_{j,\varepsilon}-m_R}\,\dd\Haus^{k-1}.
\]
Since $s_{j,\varepsilon}f_{j,\varepsilon}\le M$ pointwise on each face, we
have
\[
\abs{s_{j,\varepsilon}f_{j,\varepsilon}-m_R}
\le
\left(M-s_{j,\varepsilon}f_{j,\varepsilon}\right)+(M-m_R).
\]
Summing over all faces and using \eqref{eq:box-deficit-identity} gives
\eqref{eq:box-trace-stability}. Because the left-hand side of
\eqref{eq:box-trace-stability} is nonnegative, one must have $M\ge m_R$.
Finally, if $M=m_R$, then \eqref{eq:box-deficit-identity} shows that the
nonnegative integrand $m_R-s_{j,\varepsilon}f_{j,\varepsilon}$ has zero
integral on every face. Since $f_{j,\varepsilon}$ is continuous on
$F_{j,\varepsilon}$, it follows that
$m_R-s_{j,\varepsilon}f_{j,\varepsilon}\equiv 0$ on that face. Hence every
minimizer has the canonical face trace.
\end{proof}

\begin{definition}[Face trace]
\label{def:face-trace}
For $\eta\in\Omega^{k-1}_{\mathrm{sm}}(I^k)$ define
\[
\Tr_{\partial I^k}(\eta):=
\left\{\iota_{j,\varepsilon}^\ast\eta\right\}_{1\le j\le k,\ \varepsilon\in\{0,1\}},
\]
where
\[
F_{j,\varepsilon}:=\{x_j=\varepsilon\},
\qquad
\iota_{j,\varepsilon}:F_{j,\varepsilon}\hookrightarrow I^k.
\]
\end{definition}

\begin{corollary}[Boundary determination on the unit cube]
\label{thm:boundary-rigidity}
Let $\eta\in\Omega^{k-1}_{\mathrm{sm}}(I^k)$ satisfy
\[
\dd\eta=\omega_0.
\]
Then
\[
\norm{\eta}_{\coeff,\infty}\ge \frac{1}{2k}.
\]
If
\[
\norm{\eta}_{\coeff,\infty}\le \frac{1}{2k}+\delta,
\]
then
\begin{equation}
\label{eq:cube-boundary-stability}
\sum_{j=1}^k\sum_{\varepsilon\in\{0,1\}}
\norm{\iota_{j,\varepsilon}^\ast(\eta-\eta_\ast)}_{L^1(F_{j,\varepsilon};\coeff)}
\le
4k\,\delta.
\end{equation}
In particular, if $\norm{\eta}_{\coeff,\infty}=1/(2k)$, then
\[
\Tr_{\partial I^k}(\eta)=\Tr_{\partial I^k}(\eta_\ast).
\]
Equivalently, for every $j\in\{1,\dots,k\}$ and $\varepsilon\in\{0,1\}$,
\begin{equation}
\label{eq:boundary-rigidity-formula}
\iota_{j,\varepsilon}^\ast\eta
=
(-1)^{j-1}\frac{\varepsilon-\tfrac12}{k}\,
\dd x_1\wedge\cdots\wedge\widehat{\dd x_j}\wedge\cdots\wedge \dd x_k
\end{equation}
on the face $F_{j,\varepsilon}$.
\end{corollary}

\begin{proof}
\eqref{eq:cube-boundary-stability} is \eqref{eq:box-trace-stability} with
$L_j=1$ for every $j$. If $\norm{\eta}_{\coeff,\infty}=1/(2k)$, then the
left-hand side of \eqref{eq:cube-boundary-stability} vanishes, which gives the
claimed face trace formula.
\end{proof}

\begin{corollary}[The centered radial primitive is extremal]
\label{cor:explicit-minimizer}
The form $\eta_\ast$ is a minimizer of $\dd\eta=\omega_0$ on $I^k$, and its
boundary trace is the canonical trace described in
Corollary~\ref{thm:boundary-rigidity}.
\end{corollary}

\begin{proof}
\eqref{eq:explicit-minimizer} shows that $\dd\eta_\ast=\omega_0$ and
$\norm{\eta_\ast}_{\coeff,\infty}=1/(2k)$. Therefore
Corollary~\ref{thm:boundary-rigidity} applies. Restricting $\eta_\ast$ to the
faces $x_j=0$ and $x_j=1$ gives
\eqref{eq:boundary-rigidity-formula} directly.
\end{proof}

\begin{theorem}[Sliced boundary readout for constant coordinate-monomial data]
\label{thm:sliced-active-readout}
Let
\[
B:=\prod_{j=1}^n [0,L_j]\subset\RR^n,
\qquad
I=\{i_1<\cdots<i_k\}\subset\{1,\dots,n\},
\qquad
\omega_I:=\dd x_I.
\]
Set
\[
m_{B,I}:=\frac{1}{2\sum_{r=1}^k L_{i_r}^{-1}},
\qquad
P_I(B):=2\Vol(B)\sum_{r=1}^k L_{i_r}^{-1},
\]
and define
\begin{equation}
\label{eq:sliced-explicit-minimizer}
\eta_{B,I}
:=
2m_{B,I}\sum_{r=1}^k (-1)^{r-1}
\frac{x_{i_r}-\frac{L_{i_r}}{2}}{L_{i_r}}\,
\dd x_{I\setminus\{i_r\}}.
\end{equation}
Then $\dd\eta_{B,I}=\omega_I$ and
$\norm{\eta_{B,I}}_{\coeff,\infty}=m_{B,I}$. Let
$\eta\in\Omega^{k-1}_{\mathrm{sm}}(B)$ satisfy $\dd\eta=\omega_I$, and set
\[
M:=\norm{\eta}_{\coeff,\infty}.
\]
For each $r\in\{1,\dots,k\}$ and $\varepsilon\in\{0,L_{i_r}\}$, let
\[
G_{r,\varepsilon}:=\{x_{i_r}=\varepsilon\}
\]
and let $\iota_{r,\varepsilon}:G_{r,\varepsilon}\hookrightarrow B$ denote the
inclusion. Then
\begin{equation}
\label{eq:sliced-boundary-stability}
\sum_{r=1}^k\sum_{\varepsilon\in\{0,L_{i_r}\}}
\left\|
\coeff_{I\setminus\{i_r\}}
\bigl(
\iota_{r,\varepsilon}^\ast(\eta-\eta_{B,I})
\bigr)
\right\|_{L^1(G_{r,\varepsilon})}
\le
2P_I(B)(M-m_{B,I}).
\end{equation}
In particular, $M\ge m_{B,I}$, and every minimizer satisfies
\[
\coeff_{I\setminus\{i_r\}}
\bigl(
\iota_{r,\varepsilon}^\ast\eta
\bigr)
=
\coeff_{I\setminus\{i_r\}}
\bigl(
\iota_{r,\varepsilon}^\ast\eta_{B,I}
\bigr)
\qquad
\text{on every }G_{r,\varepsilon}.
\]
\end{theorem}

\begin{proof}
Write
\[
R_I:=\prod_{r=1}^k [0,L_{i_r}],
\qquad
R_{I^c}:=\prod_{j\notin I}[0,L_j].
\]
For $z\in R_{I^c}$ let
\[
\iota_z:R_I\to B
\]
be the slice embedding obtained by fixing the inactive coordinates equal to
$z$. Then
\[
\dd(\iota_z^\ast\eta)=\dd x_{i_1}\wedge\cdots\wedge \dd x_{i_k}
\]
on the $k$-box $R_I$, and
\[
\norm{\iota_z^\ast\eta}_{\coeff,\infty}\le M.
\]
The restriction $\iota_z^\ast\eta_{B,I}$ is the box minimizer
\eqref{eq:box-explicit-minimizer} for $R_I$, because it depends only on the
active coordinates. Applying Theorem~\ref{thm:box-boundary-stability} on each
slice gives
\[
\sum_{r=1}^k\sum_{\varepsilon\in\{0,L_{i_r}\}}
\left\|
\coeff_{I\setminus\{i_r\}}
\bigl(
\iota_{z,r,\varepsilon}^\ast(\iota_z^\ast\eta-\iota_z^\ast\eta_{B,I})
\bigr)
\right\|_{L^1(\partial R_I\cap\{x_{i_r}=\varepsilon\})}
\le
2P_1(R_I)(M-m_{B,I}),
\]
where $\iota_{z,r,\varepsilon}$ denotes the inclusion of the corresponding
slice face. Integrating this estimate over $z\in R_{I^c}$ and using Fubini's
theorem yields \eqref{eq:sliced-boundary-stability}. Indeed, the left-hand
side becomes exactly the sum of the $L^1$ norms of the active coefficients on
the full faces $G_{r,\varepsilon}$, while
\[
\Vol(R_{I^c})\,P_1(R_I)
=
2\Vol(B)\sum_{r=1}^k L_{i_r}^{-1}
=
P_I(B).
\]
Since the left-hand side of \eqref{eq:sliced-boundary-stability} is
nonnegative, one has $M\ge m_{B,I}$. If $M=m_{B,I}$, then no slice can have
coefficient norm smaller than the sharp box value, so
\[
\norm{\iota_z^\ast\eta}_{\coeff,\infty}=m_{B,I}
\qquad
\text{for every }z\in R_{I^c}.
\]
The slice-wise equality case in Theorem~\ref{thm:box-boundary-stability}
therefore forces the active boundary coefficient on every slice to agree with
that of $\eta_{B,I}$, and hence the same conclusion holds pointwise on each
full face $G_{r,\varepsilon}$.
\end{proof}

\begin{remark}
When $n>k$, Theorem~\ref{thm:sliced-active-readout} determines only the active
coefficient $\coeff_{I\setminus\{i_r\}}$ on the active faces
$G_{r,\varepsilon}$. Components involving inactive differentials are invisible
to the $k$-dimensional slice problem. The theorem therefore identifies the
maximal boundary information that is rigid under the present inverse data.
\end{remark}

\begin{remark}
For $k\ge 2$, Corollary~\ref{thm:boundary-rigidity} yields rigidity of the
boundary trace, but not uniqueness in the interior. If
$\psi$ is a smooth $(k-2)$-form compactly supported in the interior of $I^k$
and $\eta_\varepsilon:=\eta_\ast+\varepsilon\,\dd\psi$, then
$\dd\eta_\varepsilon=\omega_0$ for every $\varepsilon$,
$\Tr_{\partial I^k}(\eta_\varepsilon)=\Tr_{\partial I^k}(\eta_\ast)$, and
$\norm{\eta_\varepsilon}_{\coeff,\infty}=1/(2k)$ for all sufficiently small
$|\varepsilon|$. Indeed, $\dd\psi$ vanishes on a neighborhood of
$\partial I^k$, where $\eta_\ast$ attains the value $1/(2k)$ in the
coefficient norm, whereas $\abs{\eta_\ast}_{\coeff,\infty}<1/(2k)$ on the
compact support of $\dd\psi$.
\end{remark}

\section{A cubical Whitney counterpart}

We conclude with the cubical counterpart on a single cell. The construction is
obtained by composing the continuous homotopy operator with the classical
Whitney map and the integration map. This produces a discrete version of the
same decomposition and boundary-determination mechanism; compare
\cite{Whitney1957,Dodziuk1976,Dupont1976,DupontBook1978,Albert2020,
ArnoldFalkWinther2010,ArnoldAwanou2014,KaczynskiMischaikowMrozek2004}.

Let $C^k(I^n;\RR)$ be the space of cubical $k$-cochains on the standard cell
decomposition of $I^n$, with coboundary $d$. For $\theta\in C^k(I^n;\RR)$ we
set
\[
\norm{\theta}_{\cell,\infty}:=
\max_{\sigma^{(k)}}\abs{\theta(\sigma)},
\]
where the maximum is taken over all oriented $k$-cells.

For $I=\{i_1<\cdots<i_k\}\subset\{1,\dots,n\}$ and
$\varepsilon\in\{0,1\}^{I^c}$, let
\[
\sigma_{I,\varepsilon}
:=
\{x\in I^n:\ x_j=\varepsilon_j \text{ for every } j\notin I\}
\]
be the oriented $k$-cell parallel to
$\dd x_I=\dd x_{i_1}\wedge\cdots\wedge \dd x_{i_k}$, and let
$\delta_{I,\varepsilon}\in C^k(I^n;\RR)$ be the dual basis cochain. Define
\[
\phi_0(t):=1-t,
\qquad
\phi_1(t):=t,
\qquad
\lambda_{I,\varepsilon}(x):=
\prod_{j\notin I}\phi_{\varepsilon_j}(x_j).
\]
The cubical Whitney map is defined by
\[
W(\delta_{I,\varepsilon}):=
\lambda_{I,\varepsilon}(x)\,\dd x_I
\]
and extended linearly. For a smooth $k$-form $\alpha$ on a neighborhood of
$I^n$, the integration map is
\[
\mathcal I(\alpha)(\sigma):=
\int_\sigma \iota_\sigma^\ast\alpha.
\]
Since each $\lambda_{I,\varepsilon}$ is a multilinear polynomial on the whole
cube, $W\theta$ is a globally smooth differential form on $I^n$ for every
cochain $\theta$. Thus the continuous homotopy operator $K_j$ applies without
any regularity issue in the present one-cube setting. The next two
propositions collect the elementary identities needed for the discrete
argument.

\begin{proposition}[Whitney formulas and norm-one bounds]
\label{prop:cubical-whitney}
For every $k$ and every $\theta\in C^k(I^n;\RR)$, the maps $W$ and
$\mathcal I$ satisfy
\[
W\theta=
\sum_{\abs{I}=k}
\left(
\sum_{\varepsilon\in\{0,1\}^{I^c}}
\theta_{I,\varepsilon}\lambda_{I,\varepsilon}(x)
\right)\dd x_I,
\qquad
\theta=\sum_{\abs{I}=k}\sum_{\varepsilon}\theta_{I,\varepsilon}\delta_{I,\varepsilon},
\]
\[
\mathcal I\circ W=\mathrm{Id},
\qquad
\dd\circ W=W\circ d,
\qquad
\mathcal I\circ \dd=d\circ \mathcal I,
\]
and
\[
\norm{W}_{\infty\to\infty}=1,
\qquad
\norm{\mathcal I}_{\infty\to\infty}=1.
\]
\end{proposition}

\begin{proof}
The proof is elementary and is deferred to Appendix~A.
\end{proof}

\begin{definition}[Whitney coordinate-monomial cochain]
\label{def:whitney-decomposable}
A cubical $k$-cochain $\theta$ is called Whitney coordinate-monomial if
$W\theta$ is coordinate-monomial in the sense of
Definition~\ref{def:decomposable-form}.
\end{definition}

\begin{proposition}[Characterization of the Whitney coordinate-monomial class]
\label{prop:whitney-characterization}
A cubical $k$-cochain $\theta$ on $I^n$ is Whitney coordinate-monomial if and
only if there exists a single coordinate set
$I\subset\{1,\dots,n\}$ with $\abs{I}=k$ such that $\theta$ vanishes on every
$k$-cell not parallel to $\dd x_I$. Equivalently,
\[
W\theta=f(x_{I^c})\,\dd x_I
\]
for a multilinear polynomial $f$ in the complementary variables.
\end{proposition}

\begin{proof}
The proof is given in Appendix~A.
\end{proof}

\begin{proposition}[One-cube Whitney decomposition]
\label{thm:discrete-stokes-poincare}
Let $1\le k\le n-1$ and let $\omega\in C^k(I^n;\RR)$. Assume that $d\omega$ is
Whitney coordinate-monomial. Then there exist
\[
\eta\in C^{k-1}(I^n;\RR),
\qquad
E\in C^k(I^n;\RR)
\]
such that
\begin{equation}
\label{eq:discrete-decomposition}
\omega=d\eta+E,
\qquad
dE=d\omega,
\end{equation}
and
\begin{equation}
\label{eq:discrete-defect-bound}
\norm{E}_{\cell,\infty}\le
\frac{1}{2(k+1)}\norm{d\omega}_{\cell,\infty}.
\end{equation}
The constant $1/(2(k+1))$ cannot be improved in any estimate of this form that
is required to hold uniformly for all $\omega$ with Whitney
coordinate-monomial coboundary.
\end{proposition}

\begin{proof}
For each $1\le j\le n$ define
\[
\mathcal H_j:=\mathcal I\circ K_j\circ W:
C^j(I^n;\RR)\to C^{j-1}(I^n;\RR).
\]
We also set $\mathcal H_{n+1}:=0$. Because every Whitney form on a single cube is
a globally smooth multilinear form, the continuous homotopy identity
\eqref{eq:homotopy-identity} applies directly to $W\theta$ for each cochain
$\theta$. By Proposition~\ref{prop:cubical-whitney} and
\eqref{eq:homotopy-identity},
for every $1\le j\le n$ one has
\[
d\mathcal H_j+\mathcal H_{j+1}d=\mathrm{Id}
\qquad
\text{on }C^j(I^n;\RR).
\]
Indeed, if $\theta\in C^j(I^n;\RR)$, then
\[
\begin{aligned}
d\mathcal H_j\theta+\mathcal H_{j+1}d\theta
&=
d(\mathcal I K_jW\theta)+\mathcal I K_{j+1}W(d\theta) \\
&=
\mathcal I(\dd K_jW\theta)+\mathcal I K_{j+1}(\dd W\theta) \\
&=
\mathcal I(\dd K_j+K_{j+1}\dd)W\theta \\
&=
\mathcal I W\theta
=
\theta.
\end{aligned}
\]

Now set
\[
\eta:=\mathcal H_k\omega,
\qquad
E:=\mathcal H_{k+1}(d\omega).
\]
The identity above yields \eqref{eq:discrete-decomposition}, namely
\[
\omega=d(\mathcal H_k\omega)+\mathcal H_{k+1}(d\omega),
\]
and moreover
\[
\norm{E}_{\cell,\infty}
\le
\norm{\mathcal I}_{\infty\to\infty}\,
\norm{K_{k+1}(W(d\omega))}_{\coeff,\infty}
\le
\frac{1}{2(k+1)}\norm{W(d\omega)}_{\coeff,\infty}
\le
\frac{1}{2(k+1)}\norm{d\omega}_{\cell,\infty},
\]
because $W(d\omega)$ is coordinate-monomial by definition and both $W$ and
$\mathcal I$ have norm $1$.

For sharpness it is enough to treat the case $n=k+1$. Let
$\theta_{\mathrm{top}}\in C^{k+1}(I^{k+1};\RR)$ be the cochain taking the value
$1$ on the unique oriented top cell. Since
$W\theta_{\mathrm{top}}=\dd x_1\wedge\cdots\wedge \dd x_{k+1}$,
$\theta_{\mathrm{top}}$ is Whitney coordinate-monomial. Set
\[
\omega_0:=\mathcal H_{k+1}\theta_{\mathrm{top}}.
\]
Because $d\theta_{\mathrm{top}}=0$, the homotopy identity gives
$d\omega_0=\theta_{\mathrm{top}}$. If
\[
\omega_0=d\phi+\widetilde E,
\qquad
d\widetilde E=\theta_{\mathrm{top}},
\]
then
\[
1
=
\theta_{\mathrm{top}}(I^{k+1})
=
d\widetilde E(I^{k+1})
=
\sum_{F\subset\partial I^{k+1}}\pm \widetilde E(F).
\]
There are $2(k+1)$ codimension-one faces, and therefore
\[
1\le 2(k+1)\norm{\widetilde E}_{\cell,\infty}.
\]
This proves the optimality statement.
\end{proof}

\begin{remark}
The construction is local to a single cube. On a general cubical complex one
needs additional compatibility arguments across adjacent cubes in order to
produce a global bounded cochain homotopy; compare the algebraic framework in
\cite{Dupont1976,DupontBook1978,Albert2020}.
\end{remark}

\begin{corollary}[Hypercube gradient reconstruction]
\label{cor:hypercube-gradient}
Let $m\ge 2$, let $\omega\in C^1(I^m;\RR)$, and assume that $d\omega$ is Whitney
coordinate-monomial. Set $\varepsilon:=\norm{d\omega}_{\cell,\infty}$. Then
there exist
\[
f\in C^0(I^m;\RR),
\qquad
r\in C^1(I^m;\RR)
\]
such that
\[
\omega=df+r,
\qquad
dr=d\omega,
\qquad
\norm{r}_{\cell,\infty}\le \frac{\varepsilon}{4}.
\]
The constant $1/4$ is optimal in the sense that no smaller uniform constant can
hold in this estimate.
\end{corollary}

\begin{proof}
Apply Proposition~\ref{thm:discrete-stokes-poincare} with $k=1$.
\end{proof}

\appendix

\section{Whitney identities on a single cube}

For completeness we collect the elementary Whitney identities used in
Section~5.

\begin{proof}[Proof of Proposition~\ref{prop:cubical-whitney}]
The explicit formula for $W\theta$ follows directly from linearity. To prove
$\mathcal I\circ W=\mathrm{Id}$, it is enough to test on the basis cochains.
Let $\sigma_{J,\eta}$ be an oriented $k$-cell. Then
\[
\mathcal I(W\delta_{I,\varepsilon})(\sigma_{J,\eta})=
\int_{\sigma_{J,\eta}}
\iota_{\sigma_{J,\eta}}^\ast\!\bigl(\lambda_{I,\varepsilon}\,\dd x_I\bigr).
\]
If $J\neq I$, the pullback of $\dd x_I$ vanishes. If $J=I$ but
$\eta\neq\varepsilon$, then $\lambda_{I,\varepsilon}$ vanishes identically on
$\sigma_{I,\eta}$. Finally, if $(J,\eta)=(I,\varepsilon)$, then
$\lambda_{I,\varepsilon}\equiv 1$ on $\sigma_{I,\varepsilon}$ and the integral
is equal to $1$. Hence $\mathcal I(W\delta_{I,\varepsilon})=\delta_{I,\varepsilon}$.

To prove $\dd W=Wd$, we again work on a basis cochain. For $\ell\notin I$, let
\[
p(I,\ell):=1+\#\{i\in I:\ i<\ell\},
\]
the position of $\ell$ in the ordered set $I\cup\{\ell\}$. The only
$(k+1)$-cell containing $\sigma_{I,\varepsilon}$ as a face with free
coordinate $\ell$ is
$\sigma_{I\cup\{\ell\},\,\varepsilon|_{(I\cup\{\ell\})^c}}$, and its boundary
orientation yields the incidence sign $(-1)^{p(I,\ell)-\varepsilon_\ell}$.
Therefore
\[
d\delta_{I,\varepsilon}=
\sum_{\ell\notin I}
(-1)^{p(I,\ell)-\varepsilon_\ell}
\delta_{I\cup\{\ell\},\,\varepsilon|_{(I\cup\{\ell\})^c}}.
\]
On the other hand,
\[
\dd\bigl(W(\delta_{I,\varepsilon})\bigr)=
\sum_{\ell\notin I}
\phi_{\varepsilon_\ell}'(x_\ell)
\prod_{\substack{j\notin I\\ j\neq \ell}}
\phi_{\varepsilon_j}(x_j)\,
\dd x_\ell\wedge \dd x_I.
\]
Since $\phi_0'=-1$, $\phi_1'=1$, and
\[
\dd x_\ell\wedge \dd x_I=(-1)^{p(I,\ell)-1}\dd x_{I\cup\{\ell\}},
\]
this is exactly $W(d\delta_{I,\varepsilon})$.

The identity $\mathcal I\circ \dd=d\circ \mathcal I$ is simply Stokes' theorem
on each cell. For the operator norm of $W$, write
\[
W\theta=
\sum_{\abs{I}=k} b_I(x)\,\dd x_I,
\qquad
b_I(x):=
\sum_{\varepsilon\in\{0,1\}^{I^c}}
\theta_{I,\varepsilon}\lambda_{I,\varepsilon}(x).
\]
For fixed $I$, the functions $\lambda_{I,\varepsilon}$ are nonnegative and sum
to $1$, so $b_I(x)$ is a convex combination of the numbers
$\theta_{I,\varepsilon}$. Hence
\[
\norm{W\theta}_{\coeff,\infty}\le\norm{\theta}_{\cell,\infty},
\]
and equality is attained on each basis cochain $\delta_{I,\varepsilon}$.

Finally, if
\[
\alpha=\sum_{\abs{I}=k} a_I(x)\,\dd x_I,
\]
then for every oriented cell $\sigma_{I,\varepsilon}$,
\[
\abs{\mathcal I(\alpha)(\sigma_{I,\varepsilon})}
\le
\sup_{x\in \sigma_{I,\varepsilon}}\abs{a_I(x)}
\le
\norm{\alpha}_{\coeff,\infty}.
\]
Equality is attained on the constant form $\dd x_I$, so
$\norm{\mathcal I}_{\infty\to\infty}=1$ as well.
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:whitney-characterization}]
If $\theta$ vanishes on every $k$-cell outside one orientation class $I$, then
Proposition~\ref{prop:cubical-whitney} gives
\[
W\theta=
\left(
\sum_{\varepsilon\in\{0,1\}^{I^c}}
\theta_{I,\varepsilon}\lambda_{I,\varepsilon}(x)
\right)\dd x_I,
\]
which is of the required form.

Conversely, suppose that $W\theta=f(x_{I^c})\,\dd x_I$. For each
$J\neq I$ with $\abs{J}=k$, the coefficient of $\dd x_J$ in $W\theta$ is
\[
g_J(x):=
\sum_{\varepsilon\in\{0,1\}^{J^c}}
\theta_{J,\varepsilon}\lambda_{J,\varepsilon}(x),
\]
and $g_J$ vanishes identically. Evaluating $g_J$ at a point with
$x_j=\varepsilon_j$ for all $j\notin J$ gives $\lambda_{J,\varepsilon}(x)=1$
and $\lambda_{J,\varepsilon'}(x)=0$ for every $\varepsilon'\neq\varepsilon$.
Hence $\theta_{J,\varepsilon}=0$ for all $\varepsilon$.
\end{proof}

\section*{Data availability statement}

No new data were created or analysed in this study.

\bibliographystyle{unsrt}
\bibliography{references_local}

\end{document}
