\section{The Completed Determinant and Cyclotomic Twists}
\label{sec:completion-twists}

\subsection*{Base field and smooth projective models}

Throughout the geometric discussion we take \(k=\QQ\). Since
irreducibility, smoothness, and genus are unchanged after extending the
constant field to \(\overline{\QQ}\), we pass freely between \(k\) and
\(\overline{\QQ}\) when convenient. For an affine curve
\(C_{\mathrm{aff}}\subset\mathbb{A}^2_k\) defined by an irreducible
polynomial equation, we write \(\overline{X}\) (resp.\ \(\overline{Y}\))
for the smooth projective curve whose function field is
\(k(C_{\mathrm{aff}})\) (resp.\ its degree-two subfield under the
relevant involution). All genus and Hurwitz statements are made for
these complete nonsingular models.

\subsection{The unweighted spectrum}

\begin{proposition}[Explicit determinant]\label{prop:kernel-delta-explicit}
One has
\[
  \Delta(z,u)
  =1-(1+u)z-5uz^2+3u(1+u)z^3-u(u^2-3u+1)z^4
\]
\[
  \qquad
  +u(u^3-3u^2-3u+1)z^5+u^2(u^2+u+1)z^6.
\]
\end{proposition}

\begin{proof}
A symbolic determinant computation from the explicit \(10\times 10\)
matrix \(B(u)\) of \Cref{def:sync-kernel} gives the displayed formula;
the exact computation is recorded in
\Cref{sec:appendix-computation}\,\textup{(a)}.
As a consistency check, the closed form satisfies the self-duality
identity of \Cref{prop:sync-self-duality}; in particular, comparing
\(\Delta(z,u)\) with \(\Delta(uz,u^{-1})\) pairs the coefficients of
\(z^k u^\ell\) and \(z^k u^{k-\ell}\) exactly as required.
\end{proof}

Although \(B(u)\) is a \(10\times 10\) matrix, the determinant has
degree \(6\) in \(z\) because four rows are redundant for every \(u\):
rows \(5,6,8,10\) are \(u\)-multiples of rows \(1,2,7,9\),
respectively. Hence \(\operatorname{rank} B(u)\le 6\), so \(0\) is a
persistent eigenvalue of multiplicity at least four.

\begin{proposition}[Characteristic polynomial and residue constant]
  \label{prop:sync-kernel-spectrum}
  The matrix \(B=B(1)\) is primitive and satisfies
  \[
    \det(I-zB)=(z-1)(z+1)(3z-1)(z^3-z^2+z+1).
  \]
  In particular \(\rho(B)=3\), and the Perron residue constant
  \(C_B\) of \Cref{def:perron-residue-constant} satisfies
  \[
    C_B=\frac{243}{272}.
  \]
\end{proposition}

\begin{proof}
Every state reaches \(\st{000}\) in at most three steps by the paths
listed in \Cref{def:sync-kernel}, and \(\st{000}\) reaches every state
by iterating the transitions through \(\st{001}\), \(\st{002}\),
\(\st{010}\), and \(\st{0-12}\); hence the digraph of \(B\) is strongly
connected. Since there is also the self-loop
\(\st{000}\to\st{000}\), the period is \(1\), so \(B\) is primitive.
Set \(u=1\) in \Cref{prop:kernel-delta-explicit}. The factorisation is
then immediate. Since \((3z-1)\) is the unique factor whose reciprocal
lies outside the unit circle, \(\rho(B)=3\). Evaluating the cofactor at
\(z=1/3\) gives the residue constant. This proves part~\textup{(i)} of
\Cref{thm:intro-kernel-geometry}.
\end{proof}

Together with \Cref{lem:primitive-orbit-asymptotic}, this yields
\[
  p_n(B)=\frac{3^n}{n}
  +O\!\left(\frac{\max\{\Lambda(B)^n,3^{n/2}\}}{n}\right).
\]

\subsection{Completion and algebraic geometry}

Introducing
\[
  u=r^2,\qquad s=r+r^{-1},\qquad w=zr,
\]
the self-duality of \(\Delta(z,u)\) implies that
\(\Delta(w/r,r^2)\) is a polynomial in \(w\) and \(s\), which we denote
by \(\widehat\Delta(w,s)\).

\begin{proposition}[Explicit completed determinant]
  \label{prop:sync-hatdelta-completion}
  The completed determinant is
  \[
    \widehat\Delta(w,s)
    =1-sw-5w^2+3sw^3+(5-s^2)w^4+(s^3-6s)w^5+(s^2-1)w^6.
  \]
  It satisfies
  \[
    \widehat\Delta(w,-s)=\widehat\Delta(-w,s).
  \]
\end{proposition}

\begin{proof}
Substitute \(u=r^2\) and \(z=w/r\) into the explicit formula of
\Cref{prop:kernel-delta-explicit}. Terms in \(r\) and \(r^{-1}\)
combine to give a polynomial in \(s=r+r^{-1}\), and a direct expansion
produces the displayed formula. The parity relation is immediate. This
proves part~\textup{(ii)} of \Cref{thm:intro-kernel-geometry}.
\end{proof}

\begin{remark}
The phrase ``completed determinant'' is non-standard. In this note it
simply denotes the polynomial obtained by rewriting \(\Delta(w/r,r^2)\)
in the self-dual invariants \(w\) and \(s=r+r^{-1}\).
\end{remark}



\begin{proposition}[Generic Galois group]
  \label{prop:sync-hatdelta-galois-s6-generic}
  Viewed as a polynomial in \(w\), the completed determinant
  \(\widehat\Delta(w,s)\) has generic Galois group \(S_6\) over
  \(\QQ(s)\).
\end{proposition}

\begin{proof}
Let \(G\le S_6\) be the generic Galois group of
\(\widehat\Delta(w,s)\) over \(\QQ(s)\). We identify \(G\) in four
steps.

\medskip
\noindent\textit{Step 1 (discriminant, \(G\not\subset A_6\)).}\quad
Write \(D(s):=\disc_w\widehat\Delta(w,s)\in\QQ[s]\). The explicit
computation recorded in \Cref{sec:appendix-computation}\,(d) shows
that \(D(s)\) has degree~\(20\) with leading coefficient~\(-256\).
Since the leading coefficient is negative, \(D(s)\to-\infty\) as
\(s\to+\infty\), so \(D(s)\) takes negative real values and is
therefore not a square in~\(\QQ(s)\). Since the discriminant of \(\widehat\Delta(w,s)\) as a polynomial in \(w\) changes sign under odd permutations, \(D(s)\notin \QQ(s)^2\) implies \(G\not\subset A_6\).

\medskip
\noindent\textit{Step 2 (specialisation input).}\quad
We use the following standard fact: if \(s_0\in\QQ\) is such that
\(\widehat\Delta(w,s_0)\) is separable over \(\QQ\) (equivalently,
\(D(s_0)\neq 0\)), then the Galois group \(G_{s_0}\) of
\(\widehat\Delta(w,s_0)\) over \(\QQ\) embeds into \(G\); see
\cite[\S 4.4]{Serre2008Topics}. The condition \(D(s_0)\neq 0\) is
verified at each chosen \(s_0\) by computing
\(\gcd(\widehat\Delta(w,s_0),\partial_w\widehat\Delta(w,s_0))\bmod p\)
for a convenient prime \(p\).

To read off cycle types inside \(G_{s_0}\), we apply Dedekind's
theorem: if \(f(w)\in\ZZ[w]\) is monic, \(p\nmid\text{disc}(f)\),
and \(f\equiv f_1^{e_1}\cdots f_r^{e_r}\pmod p\) with the \(f_i\)
distinct and irreducible over \(\mathbb{F}_p\), then the Frobenius
at \(p\) in \(\Gal(f/\QQ)\) has cycle type
\((\deg f_1,\dots,\deg f_r)\); see \cite[Ch.~4,
\S{}4.3]{Serre2008Topics}. The exact mod-\(p\) factorisation data are
recorded in \Cref{sec:appendix-computation}\,(d).

\medskip
\noindent\textit{Step 3 (three Frobenius specialisations).}

\smallskip
\noindent\textit{Transitivity (\(s=3\), mod \(7\)).}\quad
One checks that
\(\gcd(\widehat\Delta(w,3),\partial_w\widehat\Delta(w,3))\equiv 1\pmod 7\),
so \(D(3)\neq 0\) and the specialisation is separable.
Modulo \(7\),
\[
  \widehat\Delta(w,3)
  \equiv w^6+2w^5+3w^4+2w^3+2w^2+4w+1\pmod 7,
\]
which is irreducible over \(\mathbb{F}_7\). Hence the Frobenius at \(7\)
in \(G_3\subset G\) is a \(6\)-cycle, so \(G\) acts transitively
on the six roots.

\smallskip
\noindent\textit{Transposition (\(s=2\), mod \(7\)).}\quad
One checks that
\(\gcd(\widehat\Delta(w,2),\partial_w\widehat\Delta(w,2))\equiv 1\pmod 7\),
so \(D(2)\neq 0\). Modulo \(7\),
\[
  \widehat\Delta(w,2)
  \equiv (w+6)(w+5)(w+2)(w+1)(3w^2+3w+2)\pmod 7,
\]
where \(3w^2+3w+2\) is irreducible over \(\mathbb{F}_7\). The
factorisation type is \((1)(1)(1)(1)(2)\), so the Frobenius at \(7\)
in \(G_2\subset G\) is a transposition. Hence \(G\) contains a
transposition.

\smallskip
\noindent\textit{5-cycle (\(s=3\), mod \(19\)).}\quad
One checks that
\(\gcd(\widehat\Delta(w,3),\partial_w\widehat\Delta(w,3))\equiv 1\pmod{19}\).
Modulo \(19\),
\[
  \widehat\Delta(w,3)
  \equiv (w+16)\bigl(8w^5+14w^4+9w^2+3w+6\bigr)\pmod{19},
\]
with the quintic factor irreducible over \(\mathbb{F}_{19}\). The
factorisation type is \((1)(5)\), so the Frobenius at \(19\) in
\(G_3\subset G\) is a \(5\)-cycle. Hence \(G\) contains a
\(5\)-cycle.

\medskip
\noindent\textit{Step 4 (subgroup classification).}\quad
We now have: \(G\le S_6\) is transitive, contains a transposition, and
contains a \(5\)-cycle. A transitive subgroup of \(S_6\) that contains
a \(5\)-cycle is primitive: if \(\sigma\in G\) is a \(5\)-cycle and
\(\mathcal{B}\) is a non-trivial block, then \(|\mathcal{B}|\) divides
\(6\). If \(\mathcal{B}\) meets the support of \(\sigma\), then
\(\sigma(\mathcal{B})\cap\mathcal{B}\neq\varnothing\), which forces
\(\sigma(\mathcal{B})=\mathcal{B}\), so \(\mathcal{B}\) must contain all
five moved points of \(\sigma\), giving \(|\mathcal{B}|\in\{5,6\}\). No
proper block of size \(5\) or \(6\) can exist in a transitive group on
six points, so \(G\) is primitive \cite[Ch.~3, \S{}3.3]{Dixon1996}.
A primitive subgroup of \(S_6\) containing a transposition is the full
symmetric group \cite[Ch.~3, Theorem~3.3A]{Dixon1996}. Hence \(G=S_6\),
completing part~\textup{(iii)} of \Cref{thm:intro-kernel-geometry}.
\end{proof}

\begin{corollary}\label{lem:hatdelta-irreducible}
The polynomial \(\widehat\Delta(w,s)\) is irreducible in \(\QQ[w,s]\).
\end{corollary}

\begin{proof}
Since \(\Gal(\widehat\Delta/\QQ(s))=S_6\), the action on the six roots is
transitive, so \(\widehat\Delta(w,s)\) is irreducible over \(\QQ(s)\).
Gauss's lemma then implies irreducibility in \(\QQ[w,s]\).
\end{proof}
We first construct the quotient curve \(\overline{Y}\) and establish its
smoothness; then we record the integrality of \(C_{\mathrm{aff}}\) and
the non-squareness of \(x\) in \(k(\overline{Y})\), which together
justify the definition of \(\overline{X}\) as a well-defined smooth
projective curve.

\begin{proposition}[The quotient is a smooth plane quartic]
  \label{prop:sync-hatdelta-quotient}
  The invariants
  \[
    x:=w^2,\qquad y:=sw
  \]
  identify \(\overline{Y}\) birationally with the smooth projective closure of
  the affine curve
  \[
    F(x,y):=
    1-y-5x+3xy+5x^2-xy^2+xy^3-6x^2y+x^2y^2-x^3=0.
  \]
  Consequently,
  \[
    g(\overline{Y})=3.
  \]
\end{proposition}

\begin{proof}
By \Cref{lem:self-dual-compression}, the coefficient of \(w^j\) in
\(\widehat\Delta(w,s)\) has degree at most \(j\) in \(s\) and the
parity of \(j\). Hence
\Cref{lem:sign-quotient-normal-form} applies to the involution
\(\iota:(w,s)\mapsto(-w,-s)\): the invariants
\[
  x=w^2,\qquad y=sw
\]
generate the invariant subring, and there is a unique polynomial
\(F(x,y)\in\QQ[x,y]\) with
\(\widehat\Delta(w,s)=F(w^2,sw)\).
A direct substitution gives the displayed formula for \(F\). Conversely,
adjoining \(w=\sqrt{x}\) recovers the function field of \(\overline{X}\)
from that of the curve \(F=0\), showing that \(\overline{Y}\) is
birationally the smooth projective closure of \(\{F=0\}\).

We verify that this projective closure is smooth.  Homogenise by setting
\(X=x\), \(Y=y\), \(Z=1\) and clearing denominators to obtain
\[
\begin{aligned}
  F_h(X,Y,Z)
  &= Z^4 - YZ^3 - 5XZ^2 + 3XYZ + 5X^2Z^2 - XY^2Z \\
  &\quad + XY^3 - 6X^2YZ + X^2Y^2 - X^3Z .
\end{aligned}
\]
On the affine chart \(Z=1\),
\[
  \partial_x F
  =-5+3y+10x-y^2+y^3-12xy+2xy^2-3x^2,
\]
\[
  \partial_y F
  =-1+3x-2xy+3xy^2-6x^2+2x^2y.
\]
A Gr\"obner-basis computation of the ideal
\(\langle F,\,\partial_x F,\,\partial_y F\rangle\subset\QQ[x,y]\)
gives \(\langle 1\rangle\), so there is no affine singular point.

The line at infinity \(Z=0\) meets the quartic where the leading form
\(XY^2(X+Y)=0\), giving the three points
\[
  [0:1:0],\qquad [1:0:0],\qquad [1:-1:0].
\]
At \([0:1:0]\) (chart \(Y=1\)): one computes
\[
\begin{aligned}
  \partial_X F_h\big|_{(0,1,0)}
  &=
  \bigl[-5Z^2+3YZ-Y^2Z+Y^3 \\
  &\qquad -12XYZ+2XY^2-3X^2Z\bigr]_{(0,1,0)}
  = 1\neq 0 .
\end{aligned}
\]
At the other two points at infinity,
\[
  \partial_Z F_h\big|_{(1,0,0)}=-1\neq 0,
  \qquad
  \partial_Z F_h\big|_{(1,-1,0)}=1\neq 0 .
\]

Hence the projective curve is smooth everywhere, so by the
genus formula for smooth plane curves \cite{Hartshorne1977},
\(g(\overline{Y})=(4-1)(4-2)/2=3\).
\end{proof}

\begin{lemma}[Integrality and function-field setup]
  \label{lem:integral-function-field}
  Let \(k=\QQ\) and let \(\overline{Y}\) be the smooth projective quartic
  of \Cref{prop:sync-hatdelta-quotient}. Then:
  \begin{enumerate}[label=\textup{(\roman*)}]
    \item The affine curve \(C_{\mathrm{aff}}:=\{\widehat\Delta(w,s)=0\}
          \subset\mathbb{A}^2_k\) is integral.
    \item The element \(x=w^2\in k(\overline{Y})\) is not a square
          in \(k(\overline{Y})\).
    \item The extension \(k(\overline{Y})(\sqrt{x})/k(\overline{Y})\)
          has degree exactly two, and its function field is \(k(C_{\mathrm{aff}})\).
  \end{enumerate}
\end{lemma}

\begin{proof}
Part~(i): \(\widehat\Delta(w,s)\) is irreducible in \(\QQ[w,s]\) by
\Cref{lem:hatdelta-irreducible}, so its zero locus is an irreducible
variety. Since \(\widehat\Delta\) is prime in the UFD \(\QQ[w,s]\), the
coordinate ring \(\QQ[w,s]/(\widehat\Delta)\) is a domain, so the
variety is also reduced. Hence \(C_{\mathrm{aff}}\) is integral.

Part~(ii): From \Cref{prop:sync-hatdelta-quotient}, the affine quartic
satisfies \(F(0,y)=1-y\), so \(x=0\) meets \(\overline{Y}\) only at
the point \(P=(0,1)\), with \(\partial_y F(0,1)=-1\neq 0\), giving
\(\operatorname{ord}_P(x)=1\). A square in \(k(\overline{Y})\) would
have even valuation at every place; the odd order at \(P\) shows
\(x\notin k(\overline{Y})^2\).

Part~(iii): Since \(x\notin k(\overline{Y})^2\), the extension
\(k(\overline{Y})(\sqrt{x})/k(\overline{Y})\) has degree two. The
identifications \(w=\sqrt{x}\) and \(s=y/\sqrt{x}\) show that this
extension is the function field of \(C_{\mathrm{aff}}\).
\end{proof}

\begin{proposition}[Two-sheeted cover of genus six]
  \label{prop:sync-hatdelta-double-cover}
  The covering map \(\overline{X}\to\overline{Y}\) is a degree-two map
  branched at exactly two points of \(\overline{Y}\). Hence
  \[
    g(\overline{X})=6.
  \]
\end{proposition}

\begin{proof}
By \Cref{lem:integral-function-field}, the extension of function fields
\(k(\overline{X})/k(\overline{Y})\) is \(k(\overline{Y})(\sqrt{x})\),
which has degree exactly two.

We use the standard criterion for ramification in quadratic extensions
of function fields: let \(K\) be a function field of characteristic
\(\neq 2\), and let \(L=K(\sqrt{f}\,)\) with \(f\in K^\times\) not a
square. Then a place \(P\) of \(K\) ramifies in \(L/K\) if and only if
\(\operatorname{ord}_P(f)\) is odd; see \cite[III.7.3]{Stichtenoth2009}.
Here \(f=x\), and we determine the divisor of \(x\) on \(\overline{Y}\).

On the affine chart, \(F(0,y)=1-y\), so the only affine point with
\(x=0\) is \(P=(0,1)\). Since \(\partial_y F(0,1)=-1\neq 0\), the line
\(x=0\) meets \(\overline{Y}\) transversely there, and therefore
\[
  \operatorname{ord}_P(x)=1.
\]

For the points at infinity, work on the chart \(X=1\) with
\(y=Y/X\) and \(z=Z/X\). The equation becomes
\[
  G(y,z)=z^4-yz^3-y^2z+y^3-3yz+y^2-z.
\]
At \(Q_0=[1:0:0]\), one has \(\partial_z G(0,0)=-1\), so \(z\) is an
analytic function of \(y\), and solving \(G(y,z)=0\) gives
\[
  z=y^2-2y^3+O(y^4).
\]
Hence \(x=X/Z=1/z\) has pole order \(2\) at \(Q_0\), so
\[
  \operatorname{ord}_{Q_0}(x)=-2.
\]
At
\(Q_\infty=[1:-1:0]\), writing \(y=-1+t\) gives
\[
  G(-1+t,z)=t-2t^2+t^3+z-tz-t^2z+z^3-tz^3+z^4,
\]
so \(\partial_z G(-1,0)=1\neq 0\), and
\[
  z=-t+t^2+O(t^3).
\]
Thus \(x=1/z\) has a simple pole at \(Q_\infty\), so
\[
  \operatorname{ord}_{Q_\infty}(x)=-1.
\]

At the remaining point \(Q_1=[0:1:0]\), use the chart \(Y=1\) with
\(u=X/Y\) and \(v=Z/Y\). The homogenised equation begins as
\[
  u-v^3+O(uv,v^4,u^2)=0,
\]
so \(u=v^3+O(v^4)\), and therefore \(x=X/Z=u/v\) has a zero of order
\(2\) at \(Q_1\), so
\[
  \operatorname{ord}_{Q_1}(x)=2.
\]

Thus
\[
  \operatorname{div}(x)=P+2Q_1-2Q_0-Q_\infty.
\]
The odd valuations occur exactly at \(P\) and \(Q_\infty\). By the
quadratic ramification criterion, these are exactly the two ramified
places of \(k(\overline{X})/k(\overline{Y})\), and each contributes one
simple branch point. Hurwitz's formula \cite{Hartshorne1977} then gives
\[
  2g(\overline{X})-2=2(2g(\overline{Y})-2)+2=2(2\cdot 3-2)+2,
\]
hence \(g(\overline{X})=6\).
\end{proof}





\subsection{Cyclic lifts and primitive real factors}

For \(q\ge 1\), let \(P_q\) be the cyclic permutation matrix on
\(\ZZ/q\ZZ\), acting on the standard basis by \(P_qe_t=e_{t+1}\). The
natural \(q\)-cyclic skew-product lift attached to the binary output
labelling is
\[
  \widetilde B_q:=B_0\otimes I_q+B_1\otimes P_q.
\]
Its state set is \(\mathcal S\times\ZZ/q\ZZ\), and an edge
\(x\to y\) of output \(e\in\{0,1\}\) in the original kernel becomes
\((x,t)\to(y,t+e)\) in the lift. This is the standard cyclic extension
attached to the output cocycle
\cite{AdlerKitchensMarcus1985,Fiebig1993,BoyleSchmieding2017}.

\begin{proposition}[Fourier decomposition of the cyclic lift]
  \label{prop:sync-cyclic-lift-factorisation}
  Let \(\omega_q=e^{2\pi i/q}\). Then
  \[
    \det(I-z\widetilde B_q)
    =\prod_{j=0}^{q-1}\Delta(z,\omega_q^j).
  \]
\end{proposition}

\begin{proof}
Let \(F_q=(\omega_q^{ab})_{0\le a,b\le q-1}\) be the discrete Fourier
matrix. Since
\[
  F_q^{-1}P_qF_q=\operatorname{diag}(1,\omega_q,\dots,\omega_q^{q-1}),
\]
conjugating \(I-z\widetilde B_q\) by \(I_{10}\otimes F_q\) gives
\[
  \bigoplus_{j=0}^{q-1}\bigl(I-z(B_0+\omega_q^jB_1)\bigr)
  =\bigoplus_{j=0}^{q-1}\bigl(I-zB(\omega_q^j)\bigr).
\]
Taking determinants yields the stated factorisation.
\end{proof}

\begin{proposition}[Primitive real factors of the cyclic lifts]
  \label{prop:sync-primitive-real-factor}
  Fix \(m\ge 2\). For
  \(a\in(\ZZ/2m\ZZ)^\times/\{\pm1\}\), set
  \[
    r_a:=e^{\pi ia/m},
    \qquad
    s_a:=r_a+r_a^{-1}=2\cos(\pi a/m).
  \]
  Then
  \[
    \Delta(z,r_a^2)=\widehat\Delta(zr_a,s_a).
  \]
  Consequently, if
  \[
    A_m:=\Bigl\{2\cos\!\tfrac{\pi a}{m}:
      a\in(\ZZ/2m\ZZ)^\times/\{\pm1\}\Bigr\},
  \]
  then the primitive Fourier blocks of the \(m\)-cyclic lift, after
  pairing conjugate characters and changing variables \(w=zr_a\), give
  the real polynomial
  \[
    \prod_{\alpha\in A_m}\widehat\Delta(w,\alpha)
    =\Res_s\bigl(\widehat\Delta(w,s),\Phi_m^+(s)\bigr).
  \]
\end{proposition}

\begin{proof}
The identity
\(\Delta(z,r_a^2)=\widehat\Delta(zr_a,s_a)\) is exactly the defining
change of variables \(u=r^2\), \(w=zr\), \(s=r+r^{-1}\) applied at
\(r=r_a\). If \(a\) is coprime to \(2m\), then \(r_a\) is a primitive
\((2m)\)-th root of unity and \(r_a^2\) is a primitive \(m\)-th root of
unity, so the corresponding factor is one of the primitive Fourier
blocks in \Cref{prop:sync-cyclic-lift-factorisation}. Replacing \(a\) by
\(-a\) conjugates \(r_a\) and leaves \(s_a=r_a+r_a^{-1}\) unchanged, so
conjugate primitive characters collapse to the same real slice
\(\widehat\Delta(w,s_a)\).

By definition,
\[
  \Phi_m^+(s)=\prod_{\alpha\in A_m}(s-\alpha),
\]
so the product formula for resultants gives
\[
  \Res_s\bigl(\widehat\Delta(w,s),\Phi_m^+(s)\bigr)
  =\prod_{\alpha\in A_m}\widehat\Delta(w,\alpha),
\]
which is exactly the real primitive factor obtained from the paired
primitive Fourier blocks.
\end{proof}

\subsection{Cyclotomic resultants and twisted Perron roots}

For \(m\ge 2\), let
\[
  s_m:=2\cos(\pi/m),
\]
let \(\Phi_m^+(s)\) be the minimal polynomial of \(s_m\) over \(\QQ\), and
define
\[
  R_m(w):=\Res_s\bigl(\widehat\Delta(w,s),\Phi_m^+(s)\bigr)\in\ZZ[w].
\]
The notation \(\Phi_m^+\) is non-standard; it denotes the real
cyclotomic polynomial of \(2\cos(\pi/m)\), namely the minimal
polynomial of the real part of a primitive \(2m\)-th root of unity.
Concretely, the conjugates of \(s_m\) under \(\Gal(\QQ(s_m)/\QQ)\) are
\[
  A_m:=\Bigl\{2\cos\!\tfrac{\pi a}{m}:a\in(\ZZ/2m\ZZ)^\times/\{\pm 1\}\Bigr\},
\]
and \(\Phi_m^+(s)=\prod_{\alpha\in A_m}(s-\alpha)\). The degree is
\(\deg\Phi_m^+=|A_m|=\varphi(2m)/2\). By
\Cref{prop:sync-primitive-real-factor}, the polynomial \(R_m(w)\) is
precisely the primitive real factor extracted from the \(m\)-cyclic
lift.

For orientation, the first few real cyclotomic polynomials are
\[
  \Phi_2^+(s)=s,\qquad
  \Phi_3^+(s)=s-1,\qquad
  \Phi_4^+(s)=s^2-2,\qquad
  \Phi_5^+(s)=s^2-s-1.
\]
The cases \(m=2\) and \(m=3\) are exceptional for different reasons:
\(R_2(w)=\widehat\Delta(w,0)\) is handled separately in
\Cref{cor:rho-m2-closed-form}, while for \(m=3\) one has
\(\Phi_3^+(s)=s-1\) and the leading coefficient \(s^2-1\) of
\(\widehat\Delta(w,s)\) vanishes at \(s=1\), so \(\deg R_3=5\) rather
than \(6\). This is why the degree law begins at \(m\ge 4\).

The corresponding twisted Perron root is
\[
  \rho_m:=\frac{1}{\min\{\abs{w}:R_m(w)=0\}}.
\]

\begin{theorem}[Cyclotomic resultant degree law]
  \label{thm:sync-cyclotomic-degree-law}
  For every \(m\ge 4\),
  \[
    \deg_w R_m=3\varphi(2m).
  \]
  Moreover, if \(m\) is even then \(R_m(w)\) is an even polynomial.
\end{theorem}

\begin{proof}
Let \(d_m:=\deg\Phi_m^+=\varphi(2m)/2\), and let \(A_m\) be the
conjugate set
\[
  A_m:=\Bigl\{2\cos\!\tfrac{\pi a}{m}:a\in(\ZZ/2m\ZZ)^\times/\{\pm 1\}\Bigr\}.
\]
Then \(\Phi_m^+(s)=\prod_{\alpha\in A_m}(s-\alpha)\), and the
product formula for resultants \cite{CoxLittleOShea2005} gives
\[
  R_m(w)
  =\Res_s\bigl(\widehat\Delta(w,s),\Phi_m^+(s)\bigr)
  =\prod_{\alpha\in A_m}\widehat\Delta(w,\alpha).
\]
For \(m\ge 4\), the leading coefficient of \(\widehat\Delta(w,\alpha)\)
as a polynomial in \(w\) is \(\alpha^2-1\). Since \(\alpha=2\cos(\pi a/m)\)
for \(a\in(\ZZ/2m\ZZ)^\times\), we have \(\alpha=\pm 1\) only if
\(\zeta+\zeta^{-1}=\pm 1\) for a primitive \(2m\)-th root of unity
\(\zeta\), which occurs exactly when \(2m=6\), i.e.\ \(m=3\). Hence
for \(m\ge 4\) one has \(\alpha^2-1\neq 0\) for every \(\alpha\in A_m\),
so each factor \(\widehat\Delta(w,\alpha)\) has degree exactly \(6\) in
\(w\). Therefore
\[
  \deg_w R_m=6\,|A_m|=6d_m=3\varphi(2m).
\]

For evenness when \(m\) is even: the set \(A_m\) is stable under
\(\alpha\mapsto-\alpha\), because \(\cos(\pi a/m)\) and
\(-\cos(\pi a/m)=\cos(\pi(m-a)/m)\) belong to the same conjugate set.
The symmetry \(\widehat\Delta(w,-s)=\widehat\Delta(-w,s)\) then gives
\[
  R_m(-w)
  =\prod_{\alpha\in A_m}\widehat\Delta(-w,\alpha)
  =\prod_{\alpha\in A_m}\widehat\Delta(w,-\alpha)
  =\prod_{\alpha\in A_m}\widehat\Delta(w,\alpha)
  =R_m(w).\qedhere
\]
\end{proof}

\begin{corollary}[The twist \(m=2\)]
  \label{cor:rho-m2-closed-form}
  One has
  \[
    R_2(w)=\widehat\Delta(w,0)
    =1-5w^2+5w^4-w^6
    =(1-w^2)(w^4-4w^2+1),
  \]
  and therefore
  \[
    \rho_2=\sqrt{2+\sqrt3}.
  \]
\end{corollary}

\begin{proof}
Since \(s_2=0\), the resultant is obtained by direct substitution.
A minimal-modulus zero of \(R_2\) is \(\sqrt{2-\sqrt3}\), whose
reciprocal is the stated value. Together with
\Cref{thm:sync-cyclotomic-degree-law}, this proves
parts~\textup{(i)} and~\textup{(ii)} of
\Cref{thm:intro-kernel-cyclotomic}.
\end{proof}

\subsection{Endpoint jets and large-\texorpdfstring{$m$}{m} asymptotics}

Let \(w(s)\) denote the simple branch of \(\widehat\Delta(w,s)=0\) near
\((w,s)=(1/3,2)\), and set
\[
  \rho(s):=\frac{1}{w(s)}.
\]

\begin{proposition}[Endpoint jet]
  \label{prop:sync-endpoint-jet}
  The branch \(w(s)\) is analytic near \(s=2\), with
  \[
    w(2)=\frac13,\qquad
    w'(2)=-\frac{11}{153},\qquad
    \rho'(2)=\frac{11}{17},\qquad
    \rho''(2)=\frac{1351}{9826}.
  \]
  Moreover, if \(\delta:=2-s\), then
  \begin{align*}
    \rho(2-\delta)
    &=3-\frac{11}{17}\delta+\frac{1351}{19652}\delta^2
      -\frac{17133}{45435424}\delta^3 \\
    &\quad
      -\frac{350037409}{105046700288}\delta^4
      -\frac{47614595705}{242867971065856}\delta^5 \\
    &\quad
      +\frac{141889658001627}{561510749104259072}\delta^6
      +O(\delta^7).
  \end{align*}
\end{proposition}

\begin{proof}
One has
\[
  \widehat\Delta(1/3,2)=0,\qquad
  \partial_w\widehat\Delta(1/3,2)\neq 0.
\]
Hence the implicit function theorem gives a unique analytic branch
\(w(s)\) near \(s=2\). Differentiating the identity
\(\widehat\Delta(w(s),s)=0\) gives
\[
  \partial_w\widehat\Delta(w(s),s)\,w'(s)
  +\partial_s\widehat\Delta(w(s),s)=0.
\]
At \((w,s)=(1/3,2)\) one computes
\[
  \partial_w\widehat\Delta(1/3,2)=-\frac{272}{81},
  \qquad
  \partial_s\widehat\Delta(1/3,2)=-\frac{176}{729},
\]
hence
\[
  w'(2)
  =-\frac{\partial_s\widehat\Delta(1/3,2)}
          {\partial_w\widehat\Delta(1/3,2)}
  =-\frac{11}{153}.
\]
Since \(\rho(s)=w(s)^{-1}\), this immediately gives
\[
  \rho'(2)=-\frac{w'(2)}{w(2)^2}=\frac{11}{17}.
\]
Differentiating a second time yields
\[
  \partial_{ww}\widehat\Delta\,(w')^2
  +2\partial_{ws}\widehat\Delta\,w'
  +\partial_{ss}\widehat\Delta
  +\partial_w\widehat\Delta\,w''=0,
\]
which determines \(w''(2)\), and therefore
\[
  \rho''(2)=\frac{2w'(2)^2}{w(2)^3}-\frac{w''(2)}{w(2)^2}
  =\frac{1351}{9826}.
\]
Continuing the same recursion through order six yields the displayed
jet of \(\rho(2-\delta)\). The exact recursive identities and the
resulting coefficients are recorded in
\Cref{sec:appendix-computation}\,\textup{(e)}.
\end{proof}

\begin{lemma}[Root separation]
  \label{lem:root-separation}
  For all sufficiently large \(m\), the minimal-modulus roots of
  \(R_m(w)\) are precisely the roots contributed by
  \(\widehat\Delta(w,s_m)\), together with the symmetric factor
  \(\widehat\Delta(w,-s_m)\) when \(-s_m\in A_m\) (equivalently, when
  \(m\) is even). Specifically, letting \(w^*(s_m)\) denote the unique
  root of \(\widehat\Delta(w,s_m)=0\) near \(w=1/3\), one has
  \(\rho_m=\rho(s_m)=1/|w^*(s_m)|\).
\end{lemma}

\begin{proof}
Define
\[
  \lambda(s):=\min\bigl\{|w|:\widehat\Delta(w,s)=0\bigr\}.
\]
By the symmetry \(\widehat\Delta(w,-s)=\widehat\Delta(-w,s)\), one has
\(\lambda(-s)=\lambda(s)\) for all \(s\). By the product formula,
\[
  R_m(w)=\prod_{\alpha\in A_m}\widehat\Delta(w,\alpha),
\]
and \(\rho_m=1/\min\{|w|:R_m(w)=0\}=1/\min_{\alpha\in A_m}\lambda(\alpha)\).
We must show that \(\lambda(s_m)\) is strictly smaller than
\(\lambda(\alpha)\) for all other \(\alpha\in A_m\), provided \(m\)
is large enough.

\medskip
\noindent\textit{Step 1 (Endpoint branch near \(s=2\)).}\quad
At \(s=2\), the polynomial \(\widehat\Delta(w,2)\) factorises as
\[
  \widehat\Delta(w,2)=(w-1)(w+1)(3w-1)(w^3-w^2+w+1).
\]
The minimal-modulus root is \(w=1/3\), and it is simple. The cubic factor has a unique real root between \(-11/20\) and \(-27/50\) because it changes sign there. Moreover,
\[
  |w^3-w^2+w|\le |w|^3+|w|^2+|w|<0.54^3+0.54^2+0.54<1
\]
whenever \(|w|\le 27/50\), so \(w^3-w^2+w+1\neq 0\) on the closed disc of radius \(27/50\). Hence every root other than \(w=1/3\) has modulus at least \(27/50>1/2\).
By the implicit function theorem, there exist \(\eta>0\) and a unique real-analytic branch
\(w^*:[2-\eta,2]\to\RR\) with \(w^*(2)=1/3\) and
\(\widehat\Delta(w^*(s),s)=0\). Shrinking \(\eta\) if necessary, we may
assume that
\[
  0<w^*(s)<\frac12
  \qquad (2-\eta\le s\le 2),
\]
while every other root of \(\widehat\Delta(\cdot,s)\) has modulus at
least \(1/2\). Hence
\[
  \lambda(s)=w^*(s)
  \qquad (2-\eta\le s\le 2).
\]
This is the same endpoint branch as in \Cref{prop:sync-endpoint-jet},
so \(\rho(s)=1/w^*(s)\) on \([2-\eta,2]\).
Since \(w^{*\prime}(2)=-11/153<0\) by \Cref{prop:sync-endpoint-jet},
continuity of \(w^{*\prime}\) allows us to shrink \(\eta\) further so
that \(w^{*\prime}(s)<0\) on \([2-\eta,2]\). Thus \(w^*(s)>1/3\) for
\(s<2\), and
\[
  \rho(s)=\frac{1}{w^*(s)}
\]
is strictly increasing on \([2-\eta,2]\).

\medskip
\noindent\textit{Step 2 (Strict spectral gap away from the untwisted
endpoints).}\quad
Let \(s\in(-2,2)\), choose \(\vartheta\in(0,\pi)\) with
\(s=2\cos\vartheta\), and set \(r=e^{i\vartheta}\), \(u=r^2=e^{2i\vartheta}\).
Then \(\widehat\Delta(w,s)=\Delta(w/r,r^2)\), so
\(\lambda(s)^{-1}\) is the spectral radius of \(B(u)\). Since every row
of \(B(1)\) has sum \(3\), one has \(\rho(B(u))\le 3\).

Suppose for contradiction that \(\rho(B(u))=3\). Choose an eigenvalue
\(\mu\) of \(B(u)\) with \(|\mu|=3\) and a corresponding eigenvector
\(\mathbf{x}=(x_1,\dots,x_{10})\), normalised by
\(\max_j |x_j|=1\). Pick \(i\) with \(|x_i|=1\). The \(i\)-th row of
\(B(u)\) contains exactly three non-zero entries, each of modulus \(1\),
so
\[
  3=|\mu x_i|
   =\Bigl|\sum_j b_{ij}(u)x_j\Bigr|
  \le \sum_j |b_{ij}(u)x_j|
  \le 3.
\]
Hence equality holds throughout: the three target coordinates in that
row all have modulus \(1\), and the corresponding summands have a
common phase. Because the digraph is strongly connected, iterating this
argument along directed paths shows that \(|x_j|=1\) for every
\(j\).

Now inspect the specific rows of \(B(u)\). From the row of
\(\st{000}\),
\[
  3=|\mu x_1|=|x_1+x_2+x_3|,
\]
so \(x_1=x_2=x_3\) and \(\mu=3\). Then the rows of
\(\st{001}\), \(\st{0-12}\), and \(\st{01-1}\) successively give
\[
  x_4=x_5=x_6=x_7=x_8=x_9=x_1.
\]
Finally, the row of \(\st{002}\) reads
\[
  3x_3=ux_1+ux_2+x_{10}.
\]
Again each summand has modulus \(1\), so equality forces
\(ux_1=ux_2=x_{10}=x_3=x_1\), hence \(u=1\), contradicting
\(\vartheta\in(0,\pi)\). Therefore
\[
  \lambda(s)>\frac13
  \qquad (-2<s<2).
\]

\medskip
\noindent\textit{Step 3 (Compactness away from the endpoints).}\quad
The function \(\lambda:[-2,2]\to\RR_{>0}\) is continuous. On the compact
set
\[
  K:=[-2+\eta,\,2-\eta],
\]
Step~2 gives \(\lambda(s)>1/3\) for every \(s\in K\). Hence
\[
  c_0:=\min_{s\in K}\lambda(s)
\]
is well defined and satisfies \(c_0>1/3\).

\medskip
\noindent\textit{Step 4 (Comparison of endpoint-near conjugates).}\quad
Let \(\alpha\in A_m\). If \(|\alpha|>2-\eta\), then for large enough
\(m\) one has \(|\alpha|\in[2-\eta,2]\). Among the conjugates
\(\alpha=2\cos(\pi a/m)\) with \(1\le a<m\), the quantity \(|\alpha|\)
is maximal at \(a=1\), giving \(s_m=2\cos(\pi/m)\); if \(m\) is even,
the same maximal absolute value also occurs at \(a=m-1\), giving
\(-s_m\). Thus every \(\alpha\in A_m\) with \(\alpha\neq \pm s_m\)
satisfies
\[
  |\alpha|<s_m.
\]
Using \(\lambda(-s)=\lambda(s)\) and the strict monotonicity of
\(\rho(s)=1/\lambda(s)\) on \([2-\eta,2]\), we obtain
\[
  \lambda(\alpha)=\lambda(|\alpha|)>\lambda(s_m)
  \qquad
  (\alpha\in A_m,\ \alpha\neq \pm s_m,\ |\alpha|>2-\eta).
\]
If \(|\alpha|\le 2-\eta\), then \(\alpha\in K\) and Step~3 gives
\[
  \lambda(\alpha)\ge c_0.
\]
Since \(\lambda(s_m)=w^*(s_m)\to 1/3\) as \(m\to\infty\), we have
\(\lambda(s_m)<c_0\) for all sufficiently large \(m\). Therefore, for
all sufficiently large \(m\),
\[
  \lambda(s_m)\le \lambda(\alpha)
  \qquad
  (\alpha\in A_m),
\]
with strict inequality for \(\alpha\neq \pm s_m\) and equality only for
\(\alpha=-s_m\) when \(-s_m\in A_m\).
Consequently the minimal-modulus roots of \(R_m(w)\) come from the
factor \(\widehat\Delta(w,s_m)\), together with
\(\widehat\Delta(w,-s_m)\) when \(m\) is even, and
\[
  \rho_m=\frac{1}{\lambda(s_m)}=\rho(s_m). \qedhere
\]
\end{proof}

\begin{theorem}[Asymptotic expansion of the twisted Perron roots]
  \label{thm:rho-gap-m12}
  As \(m\to\infty\),
  \begin{align*}
    \rho_m
    &=3-\frac{11\pi^2}{17m^2}
      +\frac{1808\pi^4}{14739m^4}
      -\frac{27872249\pi^6}{2044594080m^6} \\
    &\quad
      -\frac{77645433997\pi^8}{33089710590720m^8}
      +\frac{2976802482983509\pi^{10}}
             {3442653489858508800m^{10}} \\
    &\quad
      +\frac{3832209193654464509\pi^{12}}
             {23878244605658617036800m^{12}}
      +O(m^{-14}).
  \end{align*}
  In particular,
  \[
    3-\rho_m\sim \frac{11\pi^2}{17m^2}.
  \]
\end{theorem}

\begin{proof}
By \Cref{lem:root-separation}, \(\rho_m=\rho(s_m)\) for all sufficiently
large \(m\). The Taylor series of \(2-s_m=2-2\cos(\pi/m)\) at infinity is
\[
\begin{aligned}
  2-s_m
  &=\frac{\pi^2}{m^2}-\frac{\pi^4}{12m^4}
    +\frac{\pi^6}{360m^6}-\frac{\pi^8}{20160m^8} \\
  &\quad
    +\frac{\pi^{10}}{1814400m^{10}}
    -\frac{\pi^{12}}{239500800m^{12}}+O(m^{-14}).
\end{aligned}
\]
Setting \(\delta=2-s_m\) and substituting into the sixth-order jet of
\Cref{prop:sync-endpoint-jet} yields the displayed asymptotic series,
where the coefficient of \(m^{-2k}\) is determined by the \(\delta^k\)
term of the jet together with the lower-order terms of the Taylor series
of \(\delta\). This proves part~\textup{(iii)} of
\Cref{thm:intro-kernel-cyclotomic}.
\end{proof}

\begin{remark}
The asymptotic expansion is already numerically accurate for moderate
values of \(m\). For the displayed values \(m=4,6,8,10\), the quantity
\(\rho_m\) is verified directly from the definition
\[
  \rho_m^{-1}=\min_{\alpha\in A_m}\min\{\abs{w}:\widehat\Delta(w,\alpha)=0\},
\]
using the finite SageMath checks recorded in
\Cref{sec:appendix-computation}\,\textup{(e)}. Thus the table below does
not use \Cref{lem:root-separation}, which is only invoked for
sufficiently large \(m\). Solving the specialised equations for these
small values and comparing with the truncation of
\Cref{thm:rho-gap-m12} gives
\[
  \begin{array}{c|c|c|c}
    m & \rho_m & \text{12th-order approximation} & \text{absolute error} \\
    \hline
    4  & 2.644080451866 & 2.644083856227 & 3.40\times 10^{-6} \\
    6  & 2.831532258235 & 2.831532269934 & 1.17\times 10^{-8} \\
    8  & 2.903081377378 & 2.903081377585 & 2.07\times 10^{-10} \\
    10 & 2.937319429068 & 2.937319429077 & 9.09\times 10^{-12}
  \end{array}
\]
for \(m=4,6,8,10\).
\end{remark}










