\section{Further remarks}\label{sec:conclusion}

This note isolates one explicit weighted self-dual kernel in which
self-duality collapses all unit-circle twists to a single completed
determinant. The point is not a general theory but a complete
calculation: in this example the same degree-\(6\) polynomial controls
the quotient geometry, the generic Galois group, the cyclic-lift
factors, and the cyclotomic asymptotics. The cyclotomic resultants
therefore reduce the unit-circle twists to a one-parameter elimination
problem and yield twisted Perron roots with an effective large-\(m\)
expansion.

Two directions seem natural. The first is to identify structural
conditions on a general self-dual weighted kernel that force a
low-degree completed determinant or, more concretely, a quotient curve
of genus at most \(3\). The second is to understand to what extent the
large-\(m\) asymptotics of \(\rho_m\) depend only on the local jet of
\(\widehat\Delta\) at the endpoint \(s=2\).

\appendix

\section{Explicit matrix and computational verification}
\label{sec:appendix-computation}

\subsection*{The transition matrix \(B(u)\)}

In the state order
\((\mathtt{000},\mathtt{001},\mathtt{002},\mathtt{010},\mathtt{100},
\mathtt{101},\mathtt{0{-}12},\mathtt{1{-}12},\mathtt{01{-}1},\mathtt{11{-}1})\),
the weighted adjacency matrix \(B(u)=B_0+uB_1\) of
\Cref{def:sync-kernel} is
\[
B(u)=
\begin{pmatrix}
1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 1 & 1 & 1 & 0 & 0 & 0 & 0 \\
u & u & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 \\
0 & 0 & 0 & 0 & 1 & 1 & u & 0 & 0 & 0 \\
u & u & u & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & u & u & u & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 1 & 0 \\
0 & 0 & 0 & u & u & 0 & 0 & 0 & u & 0 \\
0 & 1 & 1 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
0 & u & u & 0 & 0 & 0 & 0 & u & 0 & 0
\end{pmatrix}.
\]
Here row \(i\) lists the outgoing edge weights from state \(i\): each
output-\(0\) edge contributes weight \(1\) and each output-\(1\) edge
contributes weight \(u\), as in \Cref{def:sync-kernel}. The matrix \(B(1)\)
is the unweighted adjacency matrix, and \(\det(I-zB(u))\) is the formula
of \Cref{prop:kernel-delta-explicit}.

\subsection*{Reproducibility guide}

The cyclic-lift factorisation and the primitive real-factor
interpretation are proved in the main text in
\Cref{prop:sync-cyclic-lift-factorisation,prop:sync-primitive-real-factor}.
The computer-assisted parts of the paper are organised here as a
theorem-to-command map rather than as a single executable script,
because the verification splits naturally into SageMath and Mathematica
blocks.

Items \textup{(a)}--\textup{(e)} were run in SageMath~10.2, and item
\textup{(f)} was run in Mathematica~14.0. Their roles in the body are:
\begin{itemize}[leftmargin=2.5em]
  \item \textup{(a)} certifies \Cref{prop:kernel-delta-explicit};
  \item \textup{(b)} certifies
        \Cref{prop:sync-hatdelta-completion,prop:sync-hatdelta-quotient};
  \item \textup{(c)} certifies the affine smoothness step in
        \Cref{prop:sync-hatdelta-quotient};
  \item \textup{(d)} certifies the discriminant and finite-field
        factorisations used in
        \Cref{prop:sync-hatdelta-galois-s6-generic};
  \item \textup{(e)} certifies the finite small-\(m\) checks used in
        the numerical remark after \Cref{thm:rho-gap-m12};
  \item \textup{(f)} certifies
        \Cref{prop:sync-endpoint-jet,thm:rho-gap-m12}.
\end{itemize}
Each item records the exact output used in the argument together with
the corresponding command block.

\subsection*{Exact verification data}

\begin{enumerate}[label=(\alph*)]
  \item \textit{Determinant expansion and self-duality check.}\quad
    Expanding \(\det(I-zB(u))\) from the displayed matrix gives
    \[
      \Delta(z,u)
      =1-(1+u)z-5uz^2+3u(1+u)z^3-u(u^2-3u+1)z^4
    \]
    \[
      \qquad
      +u(u^3-3u^2-3u+1)z^5+u^2(u^2+u+1)z^6.
    \]
    Substituting \(u=1\) yields
    \[
      \Delta(z,1)=1-2z-5z^2+6z^3+z^4-4z^5+3z^6
      =(z-1)(z+1)(3z-1)(z^3-z^2+z+1).
    \]
    The same expansion satisfies
    \(\Delta(uz,u^{-1})=\Delta(z,u)\) coefficientwise, which is the
    explicit consistency check used in
    \Cref{prop:kernel-delta-explicit}.

    \medskip
    \noindent\textit{Sage commands.}
\begin{verbatim}
R.<u,z> = QQ[]
B = Matrix(R, [
    [1,1,1,0,0,0,0,0,0,0],
    [0,0,0,1,1,1,0,0,0,0],
    [u,u,0,0,0,0,0,0,0,1],
    [0,0,0,0,1,1,u,0,0,0],
    [u,u,u,0,0,0,0,0,0,0],
    [0,0,0,u,u,u,0,0,0,0],
    [0,0,0,1,1,0,0,0,1,0],
    [0,0,0,u,u,0,0,0,u,0],
    [0,1,1,0,0,0,0,1,0,0],
    [0,u,u,0,0,0,0,u,0,0]
])
Delta = (identity_matrix(R, 10) - z*B).det().expand()
factor(Delta.subs(u=1))
\end{verbatim}

  \item \textit{Completion and quotient polynomial.}\quad
    After the substitution \(u=r^2\), \(z=w/r\), one obtains
    \[
      \widehat\Delta(w,s)
      =1-sw-5w^2+3sw^3+(5-s^2)w^4+(s^3-6s)w^5+(s^2-1)w^6,
    \]
    with \(s=r+r^{-1}\). Passing to the invariants \(x=w^2\) and
    \(y=sw\) gives the quotient equation
    \[
      F(x,y)=1-y-5x+3xy+5x^2-xy^2+xy^3-6x^2y+x^2y^2-x^3.
    \]

    \medskip
    \noindent\textit{Sage commands.}
\begin{verbatim}
R.<r,w> = LaurentPolynomialRing(QQ)
Delta_rw = 1-(1+r^2)*(w/r)-5*r^2*(w/r)^2+3*r^2*(1+r^2)*(w/r)^3 \
    - r^2*(r^4-3*r^2+1)*(w/r)^4 \
    + r^2*(r^6-3*r^4-3*r^2+1)*(w/r)^5 \
    + r^4*(r^4+r^2+1)*(w/r)^6
Delta_rw.expand()

S.<s,w> = PolynomialRing(QQ)
H = 1 - s*w - 5*w^2 + 3*s*w^3 + (5 - s^2)*w^4 \
    + (s^3 - 6*s)*w^5 + (s^2 - 1)*w^6

T.<x,y> = PolynomialRing(QQ)
F = 1 - y - 5*x + 3*x*y + 5*x^2 - x*y^2 + x*y^3 \
    - 6*x^2*y + x^2*y^2 - x^3
H, F
\end{verbatim}

  \item \textit{Affine smoothness certificate.}\quad
    For the quartic \(F(x,y)\) above,
    \[
      \partial_xF
      =-5+3y+10x-y^2+y^3-12xy+2xy^2-3x^2,
    \]
    \[
      \partial_yF
      =-1+3x-2xy+3xy^2-6x^2+2x^2y.
    \]
    The reduced Gr\"obner basis of the ideal
    \(\langle F,\,\partial_xF,\,\partial_yF\rangle\subset\QQ[x,y]\)
    with respect to graded reverse lexicographic order is \(\{1\}\).
    This is the exact affine smoothness certificate used in
    \Cref{prop:sync-hatdelta-quotient}.

    \medskip
    \noindent\textit{Sage commands.}
\begin{verbatim}
R.<x,y> = QQ[]
F = 1 - y - 5*x + 3*x*y + 5*x^2 - x*y^2 + x*y^3 \
    - 6*x^2*y + x^2*y^2 - x^3
ideal(F, diff(F, x), diff(F, y)).groebner_basis()
\end{verbatim}

  \item \textit{Discriminant and mod-\(p\) factorisations.}\quad
    The discriminant \(D(s)=\disc_w\widehat\Delta(w,s)\) has degree
    \(20\) in \(s\) with leading coefficient \(-256\). The command below
    computes the full polynomial \(D(s)\), confirming both the degree
    and the leading coefficient. The specialisations
    used in \Cref{prop:sync-hatdelta-galois-s6-generic} are
    \[
      \widehat\Delta(w,3)=8w^6+9w^5-4w^4+9w^3-5w^2-3w+1
    \]
    and
    \[
      \widehat\Delta(w,2)=3w^6-4w^5+w^4+6w^3-5w^2-2w+1.
    \]
    Their exact finite-field outputs are
    \[
      \gcd\!\bigl(\widehat\Delta(w,3),\,\partial_w\widehat\Delta(w,3)\bigr)
      \equiv 1 \pmod 7,
    \]
    \[
      \widehat\Delta(w,3)\equiv
      w^6+2w^5+3w^4+2w^3+2w^2+4w+1 \pmod 7,
    \]
    with the degree-\(6\) factor irreducible over \(\mathbb{F}_7\),
    \[
      \gcd\!\bigl(\widehat\Delta(w,2),\,\partial_w\widehat\Delta(w,2)\bigr)
      \equiv 1 \pmod 7,
    \]
    \[
      \widehat\Delta(w,2)\equiv
      (w+6)(w+5)(w+2)(w+1)(3w^2+3w+2)\pmod 7,
    \]
    where \(3w^2+3w+2\) is irreducible over \(\mathbb{F}_7\), and
    \[
      \gcd\!\bigl(\widehat\Delta(w,3),\,\partial_w\widehat\Delta(w,3)\bigr)
      \equiv 1 \pmod{19},
    \]
    \[
      \widehat\Delta(w,3)\equiv
      (w+16)\bigl(8w^5+14w^4+9w^2+3w+6\bigr)\pmod{19},
    \]
    with the quintic factor irreducible over \(\mathbb{F}_{19}\).

    \medskip
    \noindent\textit{Sage commands.}
\begin{verbatim}
R.<w,s> = QQ[]
H = 1 - s*w - 5*w^2 + 3*s*w^3 + (5 - s^2)*w^4 \
    + (s^3 - 6*s)*w^5 + (s^2 - 1)*w^6
D = H.discriminant(w)
D.degree(), D.leading_coefficient()

R7.<w> = PolynomialRing(GF(7))
f37 = R7(H(s=3))
gcd(f37, f37.derivative()), factor(f37)

f27 = R7(H(s=2))
gcd(f27, f27.derivative()), factor(f27)

R19.<w> = PolynomialRing(GF(19))
f319 = R19(H(s=3))
gcd(f319, f319.derivative()), factor(f319)
\end{verbatim}

  \item \textit{Finite verification for the displayed small values.}\quad
    For the table following \Cref{thm:rho-gap-m12}, one computes
    \[
      \lambda(\alpha):=\min\{\abs{w}:\widehat\Delta(w,\alpha)=0\}
    \]
    for every conjugate \(\alpha\in A_m\) with \(m=4,6,8,10\). For
    \(m=4\) and \(m=6\), the conjugate sets are
    \(A_m=\{\pm s_m\}\), so there is nothing further to check. For
    \(m=8\) and \(m=10\), the remaining positive conjugates are
    \(2\cos(3\pi/8)\) and \(2\cos(3\pi/10)\), and the direct output is
    \[
      \lambda\!\bigl(2\cos(\pi/8)\bigr)
      < \lambda\!\bigl(2\cos(3\pi/8)\bigr),
    \]
    \[
      \lambda\!\bigl(2\cos(\pi/10)\bigr)
      < \lambda\!\bigl(2\cos(3\pi/10)\bigr).
    \]
    Together with the symmetry \(\lambda(-\alpha)=\lambda(\alpha)\),
    this verifies that the displayed values \(\rho_m\) are obtained from
    the slices \(\alpha=\pm s_m\).

    \medskip
    \noindent\textit{Sage commands.}
\begin{verbatim}
R.<w> = PolynomialRing(QQ)
def H(alpha):
    return 1 - alpha*w - 5*w^2 + 3*alpha*w^3 + (5 - alpha^2)*w^4 \
        + (alpha^3 - 6*alpha)*w^5 + (alpha^2 - 1)*w^6

def lambda_alpha(alpha):
    roots = H(alpha).change_ring(CC).roots(multiplicities=False)
    return min(abs(root) for root in roots)

for m in [4, 6, 8, 10]:
    reps = []
    for a in range(1, 2*m):
        if gcd(a, 2*m) == 1 and a <= m:
            reps.append(2*cos(pi*a/m))
    print(m, [(alpha.n(), lambda_alpha(alpha).n()) for alpha in reps])
\end{verbatim}

  \item \textit{Endpoint jet recursion.}\quad
    At the endpoint \((w,s)=(1/3,2)\), one has
    \[
      \partial_w\widehat\Delta(1/3,2)=-\frac{272}{81},
      \qquad
      \partial_s\widehat\Delta(1/3,2)=-\frac{176}{729}.
    \]
    Writing \(w(s)=\sum_{j\ge 0}a_j(s-2)^j\) with \(a_0=1/3\) and
    substituting into \(\widehat\Delta(w(s),s)=0\) gives a triangular
    recursion for the coefficients \(a_j\). After inversion
    \(\rho(s)=1/w(s)\), the exact coefficients of
    \(\rho(2-\delta)=3+\sum_{j=1}^6 c_j\delta^j+O(\delta^7)\) are
    \[
      c_1=-\frac{11}{17},\qquad
      c_2=\frac{1351}{19652},\qquad
      c_3=-\frac{17133}{45435424},
    \]
    \[
      c_4=-\frac{350037409}{105046700288},\qquad
      c_5=-\frac{47614595705}{242867971065856},
    \]
    \[
      c_6=\frac{141889658001627}{561510749104259072}.
    \]
    These are exactly the jet coefficients used in
    \Cref{prop:sync-endpoint-jet}.

    \medskip
    \noindent\textit{Mathematica commands.}
\begin{verbatim}
Clear[w, s, \[Delta], a1, a2, a3, a4, a5, a6];
H[w_, s_] := 1 - s w - 5 w^2 + 3 s w^3 + (5 - s^2) w^4 +
    (s^3 - 6 s) w^5 + (s^2 - 1) w^6;
ws = 1/3 + a1 \[Delta] + a2 \[Delta]^2 + a3 \[Delta]^3 +
    a4 \[Delta]^4 + a5 \[Delta]^5 + a6 \[Delta]^6;
seriesH = Normal[
    Series[H[ws, 2 - \[Delta]], {\[Delta], 0, 6}]
];
eqns = Thread[
    CoefficientList[seriesH, \[Delta]] == 0
];
sol = First[Solve[eqns, {a1, a2, a3, a4, a5, a6}]];
Normal[Series[1/(ws /. sol), {\[Delta], 0, 6}]]
\end{verbatim}
\end{enumerate}
