\section{Introduction}\label{sec:introduction}

Let \(B(u)=B_0+uB_1\) be the weighted adjacency matrix of the explicit
ten-state digraph displayed in \Cref{def:sync-kernel} and
Figure~\ref{fig:sync-kernel}. The symbol ``synchronisation kernel'' is
purely terminological: the state names \(\st{000},\ldots,\st{11-1}\)
are nothing more than labels for the ten vertices of that digraph, and
all combinatorial data appear verbatim in the transition table. Thus the
manuscript analyzes a single concrete self-dual kernel rather than a
general construction. Even this small finite-state object already
produces a completed determinant whose smooth projective spectral curve
has genus \(6\), whose quotient by the sign involution is a smooth
quartic of genus \(3\), and whose generic Galois group is the full
symmetric group~\(S_6\).

The permutation implementing the self-duality exchanges output-zero and
output-one edges. \Cref{lem:self-dual-compression} records the formal
consequences for any such kernel: the determinant satisfies the
functional equation \(\Delta(z,u)=\Delta(uz,u^{-1})\) and compresses to a
polynomial \(\widehat\Delta(w,s)\in\ZZ[s][w]\) of degree six after the
change of variables \(u=r^2,\ s=r+r^{-1},\ w=zr\). Every other statement in
the paper is specific to the ten-state kernel of \Cref{def:sync-kernel}.
For each fixed \(u\), the matrix \(B(u)\) defines a finite-state
dynamical \(\zeta\)-function through the determinant
\[
  \Delta(z,u):=\det(I-zB(u)).
\]
The same weighted matrix also controls the natural cyclic
skew-product lifts
\[
  \widetilde B_q:=B_0\otimes I_q+B_1\otimes P_q,
\]
where \(P_q\) is the cyclic permutation matrix on \(\ZZ/q\ZZ\). Their
Fourier blocks are \(B(\omega_q^j)\), so the unit-circle twists are the
natural character factors of the labelled cyclic lifts. The purpose of
this note is to record the additional structure coming from the
explicit self-duality
\[
  \Delta(z,u)=\Delta(uz,u^{-1}),
\]
which compresses the unit-circle twists \(u=e^{2i\theta}\) to a single
real parameter after the standard change of variables
\[
  r=e^{i\theta},\qquad u=r^2,\qquad s=r+r^{-1}=2\cos\theta,\qquad w=zr.
\]
The resulting completed determinant \(\widehat\Delta(w,s)\) provides a
convenient algebraic model for the twisted spectrum, and the
cyclotomic resultants considered below are precisely the primitive real
factors extracted from the Fourier decomposition of the cyclic lifts.


\paragraph{Terminology.}
We call \((w,s)\mapsto \widehat\Delta(w,s)\) the completed determinant in analogy with completed \(L\)-functions: it is simply the polynomial obtained from \(\Delta(w/r,r^2)\) after rewriting in the self-dual invariants \(w\) and \(s=r+r^{-1}\). Likewise \(\Phi_m^+\) denotes the real cyclotomic polynomial of \(2\cos(\pi/m)\); concretely
\[
  \Phi_m^+(s)
  =\prod_{a\in(\ZZ/2m\ZZ)^\times/\{\pm1\}}
    \bigl(s-2\cos(\pi a/m)\bigr),
\]
so \(\Phi_m^+\) is the standard real Chebyshev factor attached to the primitive \((2m)\)-th roots of unity.



This note has two strands. The first is geometric: the affine
curve \(\widehat\Delta(w,s)=0\) admits a natural involution, and its
normalisation and quotient can be described explicitly. The second is
arithmetic: eliminating the parameter \(s\) along cyclotomic slices
produces one-variable resultants whose minimal-modulus roots control
the twisted Perron radii. This yields explicit asymptotics for the
cyclotomic twists of the kernel.

\paragraph{Computer-assisted verification.}
All determinant expansions, discriminant computations, finite-field
factorisations, and endpoint jet recursions leading to
Propositions~\ref{prop:kernel-delta-explicit},
\ref{prop:sync-hatdelta-quotient},
\ref{prop:sync-hatdelta-galois-s6-generic},
\ref{prop:sync-endpoint-jet}, and
Theorem~\ref{thm:rho-gap-m12} were carried out in SageMath~10.2 and
Mathematica~14.0. The appendix is organised as a theorem-to-command map
rather than as a single executable script: items
\textup{(a)}--\textup{(e)} record the SageMath certificates for the
determinant, completion, smoothness, Galois-theoretic specialisations,
and the finite small-\(m\) slice checks, while item \textup{(f)}
records the Mathematica recursion for the endpoint jet. Each item
states the exact output used in the corresponding proof.

What makes this particular example worth isolating is a structural
coincidence that does not hold for generic self-dual kernels. The
self-duality \(\Delta(z,u)=\Delta(uz,u^{-1})\) compresses the full
circle of unitary twists \(u=e^{2i\theta}\) to a single real variable
\(s=2\cos\theta\), so that the twist data for all cyclic lifts of all
orders is encoded in one polynomial \(\widehat\Delta(w,s)\in\ZZ[s][w]\)
of degree six. The significance of the example is therefore concrete
rather than general: once everything is packaged into this one small
polynomial, the quotient geometry, the generic Galois group, and the
large-\(m\) asymptotics can all be computed completely and audited
term by term. Beyond the formal consequences of
\Cref{lem:self-dual-compression} and the cyclic-lift factorisation of
\Cref{prop:sync-cyclic-lift-factorisation}, every claim in the paper is
deliberately confined to the ten-state kernel.


\subsection{Main results}

Our first theorem records the explicit determinant and geometric
information. All curves in the main results are considered over
\(\overline{\QQ}\), and \(\overline{X}\) denotes the normalisation of
the projective closure of the irreducible affine curve
\(\widehat\Delta(w,s)=0\).

\begin{theorem}\label{thm:intro-kernel-geometry}
Let \(B(u)\) be the weighted synchronisation kernel of
\Cref{def:sync-kernel}. Then:
\begin{enumerate}[label=\textup{(\roman*)}]
  \item The determinant is
  \[
    \Delta(z,u)
    =1-(1+u)z-5uz^2+3u(1+u)z^3-u(u^2-3u+1)z^4
  \]
  \[
    \qquad
    +u(u^3-3u^2-3u+1)z^5+u^2(u^2+u+1)z^6.
  \]
  In particular,
  \[
    \Delta(z,1)=(z-1)(z+1)(3z-1)(z^3-z^2+z+1),
    \qquad
    C_B=\frac{243}{272}.
  \]

  \item The completed determinant
  \[
    \widehat\Delta(w,s):=\Delta(w/r,r^2)
    \qquad (u=r^2,\ s=r+r^{-1})
  \]
  is given by
  \[
    \widehat\Delta(w,s)
    =1-sw-5w^2+3sw^3+(5-s^2)w^4+(s^3-6s)w^5+(s^2-1)w^6.
  \]
  It satisfies
  \[
    \widehat\Delta(w,-s)=\widehat\Delta(-w,s).
  \]

  \item The smooth projective curve \(\overline{X}\) has genus \(6\).
  Its quotient by the involution \((w,s)\mapsto(-w,-s)\) is a smooth
  plane quartic of genus \(3\), and the generic Galois group of
  \(\widehat\Delta(w,s)\) over \(\QQ(s)\) is \(S_6\).
\end{enumerate}
\end{theorem}

Our second theorem concerns cyclotomic twists.

\begin{theorem}\label{thm:intro-kernel-cyclotomic}
For \(m\ge 2\), let \(s_m=2\cos(\pi/m)\), let \(\Phi_m^+\) be the
minimal polynomial of \(s_m\), and define
\[
  R_m(w):=\Res_s\bigl(\widehat\Delta(w,s),\Phi_m^+(s)\bigr).
\]
If
\[
  \rho_m:=\frac{1}{\min\{\abs{w}:R_m(w)=0\}},
\]
then:
\begin{enumerate}[label=\textup{(\roman*)}]
  \item For every \(m\ge 4\),
  \[
    \deg_w R_m=3\varphi(2m).
  \]
  If \(m\) is even, then \(R_m(w)\) is an even polynomial.

  \item One has
  \[
    \rho_2=\sqrt{2+\sqrt3}.
  \]

  \item As \(m\to\infty\),
  \[
    \rho_m
    =3-\frac{11\pi^2}{17m^2}+O(m^{-4}).
  \]
  Moreover, \Cref{thm:rho-gap-m12} gives the expansion through order
  \(m^{-12}\). In particular,
  \[
    3-\rho_m\sim \frac{11\pi^2}{17m^2}.
  \]
\end{enumerate}
\end{theorem}

We also record the primitive-orbit asymptotics of the unweighted kernel.

\subsection{Previous work}

The rationality of dynamical \(\zeta\)-functions for shifts of finite
type goes back to Bowen--Lanford, Manning, Parry--Pollicott, and
Bass~\cite{BowenLanford1970Zeta,Manning1971Axiom,ParryPollicott1983,Bass1992};
General background on symbolic dynamics may be found in
\cite{LindMarcus2021,Kitchens1998}.

The closest literature falls into three layers.

Classical background on graph zeta functions, coverings, and
Artin-style \(L\)-factorisations goes through Hashimoto
\cite{Hashimoto1990,Hashimoto1992}, Stark and Terras
\cite{StarkTerras1996,StarkTerras2000,StarkTerras2007}, and
Kotani--Sunada \cite{KotaniSunada2000}; see also Terras's monograph
\cite{Terras2011}. The voltage-graph and covering language for cyclic
lifts is classical \cite{GrossTucker2001}, while the twisted-operator
viewpoint relevant to the present Fourier decomposition connects to the
covering theory of Cimasoni--Kassel \cite{CimasoniKassel2023}. Spectral
properties of arbitrary lifts of graphs, directly relevant to
\Cref{prop:sync-cyclic-lift-factorisation}, are studied systematically
in \cite{DalfoFiolGarriga2019}.

Direct precedents for determinant expressions and covering
factorisations on digraphs and weighted graphs are due to Mizuno and
Sato \cite{MizunoSato2000,MizunoSato2001,MizunoSato2002,MizunoSato2004},
to Sato \cite{Sato2006}, and, on the dynamical/determinantal side, to
Morita and collaborators
\cite{Morita2020,IdeIshikawaMoritaSatoSegawa2021}. These works are the
closest antecedents to the cyclic-lift/resultant discussion in the
present note.

Finite-group extensions of shifts of finite type provide the
symbolic-dynamical framework for tracking output labels in a skew
product. This viewpoint goes back to Adler--Kitchens--Marcus
\cite{AdlerKitchensMarcus1985} and Fiebig \cite{Fiebig1993}; periodic
data and recurrence in \(\ZZ_q\)-extensions were studied by
Pollicott--Sharp \cite{PollicottSharp1994}, and the
matrix-over-group-ring framework was developed further by
Boyle--Schmieding \cite{BoyleSchmieding2017}. The underlying two-track
comparison may be seen directly as the fibre product of labelled graphs
in the sense of Nasu's textile systems \cite{Nasu1995}. The present
paper uses only the simplest cyclic extension attached to the binary
output labels. Its novelty is correspondingly narrower: not a new
general theory, but the complete analysis of one explicit self-dual
kernel for which the completed determinant remains degree \(6\) while
still carrying nontrivial quotient geometry, generic \(S_6\) Galois
group, and explicit large-\(m\) cyclotomic asymptotics.

\subsection{Organisation}

\Cref{sec:preliminaries} recalls primitive orbit counts and defines the
weighted synchronisation kernel together with its self-duality.
\Cref{sec:completion-twists} contains the explicit determinant, the
cyclic-lift Fourier decomposition, the irreducibility and integrality
of the completed curve, its geometry, the resultant degree law, and the
asymptotics of the twisted Perron roots.
\Cref{sec:appendix-computation} records the explicit transition matrix
and the exact theorem-to-command map for the computer-assisted checks.







