\input{preamble}

\title{Finite-window Zeckendorf normalization, compatible restriction, and the pair shift}
\author{
  Haobo Ma\thanks{AELF PTE LTD., \#14-02, Marina Bay Financial Centre Tower 1, 8 Marina Blvd, Singapore 018981. Email: \texttt{auric@aelf.io}}
  \and
  Wenlin Zhang\thanks{Corresponding author. National University of Singapore, Singapore. Email: \texttt{e1327962@u.nus.edu}}
}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We study finite-window Zeckendorf normalization for binary words in Fibonacci weights. Given a word of length \(m\), its Fibonacci value is replaced by the Zeckendorf expansion and then truncated back to the original window. This defines an idempotent map onto the length-\(m\) language of the golden-mean shift, but it does not commute with naive restriction when the window size varies. There is nevertheless a unique compatible restriction on microstates, and the discrepancy from naive truncation may occupy the whole visible window; in particular, no uniformly bounded tail correction can restore commutation.

A structural theorem is proved for zero-synchronized local normalizers of rewrite span \(r\): the associated interior pair language of raw and normalized digits is always a synchronized sofic shift whose follower memory is at most \(r-1\). The Fibonacci rule \(011\to100\) is the span-\(3\) case, and its follower graph collapses canonically to the four-state pair shift studied in the paper. A second structural theorem reduces local invertibility of window codes to finitely many ambiguous cores of length at most \(r+1\); for Fibonacci this yields the exact conjugacy threshold \(m\ge 3\).

From the explicit four-state presentation one obtains the exact discrepancy statistics, including density \(4/9\), asymptotic variance \(118/243\), closed covariance formulas, and a rational power spectrum. The same graph also provides a transfer-matrix description of weighted fibers, from which the free energy is recovered as the Legendre transform of a strictly concave entropy profile.
\end{abstract}

\paragraph{Keywords.}
Zeckendorf normalization; Fibonacci numeration; sofic shifts; sliding block codes; transfer matrices; thermodynamic formalism.

\paragraph{2020 Mathematics Subject Classification.}
Primary 37B10, 11B39; Secondary 37A50, 68Q45.

\input{sections/01-introduction}
\input{sections/02-preliminaries}
\input{sections/03-multiscale}
\input{sections/04-gauge-anomaly}
\input{sections/05-sequence-level}
\input{sections/06-partition-function}
\input{sections/09-discussion}

\appendix
\input{appendices/appA-rewriting}
\input{appendices/appB-pressure-clt}
\input{appendices/appC-sofic-conjugacy}
\input{appendices/appD-recurrence-discriminant}

\bibliographystyle{plain}
\bibliography{references}

\end{document}
