\section{Model and basic objects}\label{sec:model}

This section records the minimal data needed later: the rational generating function, the companion transfer matrix, and the spectral polynomial.
All of these constructions are standard for rational sequences \cite{FlajoletSedgewick2009,Stanley2012}, and the proofs are omitted unless the argument interacts with the balanced-family structure.

\subsection{Generating function}
\label{subsec:ogf}

Let \(y\in\CC\) be a complex parameter. Define the ordinary generating function
\[
Z(t,y)=\sum_{m\ge 0} Z_m(y)t^m,
\qquad
Z(t,y)=\frac{N(t,y)}{D(t,y)},
\]
where
\begin{align*}
N(t,y) &= 1+yt+(y^2-y-1)t^2+(y^3-2y)t^3,\\
D(t,y) &= 1-t-(2y+1)t^2+t^3+y(y+1)t^4.
\end{align*}
The coefficients \(Z_m(y)\) are uniquely determined by \(D(t,y)Z(t,y)=N(t,y)\) and satisfy \(Z_m(y)\in\ZZ[y]\).
At \(y=1\) we have \(Z(t,1)=1/(1-2t)\), hence \(Z_m(1)=2^m\).

\paragraph{Initial values.}
Comparing coefficients in \(D(t,y)Z(t,y)=N(t,y)\) gives
\[
Z_0(y)=1,\qquad Z_1(y)=y+1,\qquad Z_2(y)=(y+1)^2,\qquad Z_3(y)=(y+1)^3.
\]

\subsection{Order-\texorpdfstring{\(4\)}{4} recurrence and the spectral polynomial}
\label{subsec:recurrence}

From \(D(t,y)Z(t,y)=N(t,y)\) we obtain, for all \(m\ge 4\), the order-\(4\) linear recurrence
\begin{equation}\label{eq:recurrence}
Z_m(y)-Z_{m-1}(y)-(2y+1)Z_{m-2}(y)+Z_{m-3}(y)+y(y+1)Z_{m-4}(y)=0.
\end{equation}

The associated characteristic (spectral) polynomial is
\begin{equation}\label{eq:spectral-poly}
\Pi(\lambda,y)=\lambda^4-\lambda^3-(2y+1)\lambda^2+\lambda+y(y+1),
\end{equation}
which satisfies \(\Pi(\lambda,y)=\lambda^4 D(1/\lambda,y)\).
Let \(\lambda_1(y),\dots,\lambda_4(y)\) be the roots of \(\Pi(\lambda,y)=0\), counted with multiplicity.

\subsection{Companion transfer matrix}
\label{subsec:transfer-matrix}

The recurrence \eqref{eq:recurrence} admits a canonical $4\times 4$ transfer-matrix realization.
Define
\begin{equation}\label{eq:companion-matrix}
T(y)=
\begin{pmatrix}
1 & 2y+1 & -1 & -y(y+1)\\
1 & 0 & 0 & 0\\
0 & 1 & 0 & 0\\
0 & 0 & 1 & 0
\end{pmatrix}.
\end{equation}
Then
\[
\begin{pmatrix}
Z_m\\ Z_{m-1}\\ Z_{m-2}\\ Z_{m-3}
\end{pmatrix}
=
T(y)
\begin{pmatrix}
Z_{m-1}\\ Z_{m-2}\\ Z_{m-3}\\ Z_{m-4}
\end{pmatrix}
\qquad (m\ge 4),
\]
so the polynomial family is generated by a finite-dimensional non-Hermitian companion problem.
The characteristic polynomial of \(T(y)\) is exactly \(\Pi(\lambda,y)\):
\[
\det(\lambda I-T(y))=\Pi(\lambda,y).
\]
All subsequent references to dominant eigenvalues, branch points, and the dominant branch over $y>0$ are therefore statements about the spectral geometry of the companion family \eqref{eq:companion-matrix}.

\begin{lemma}[Cyclicity of the companion family]\label{lem:cyclic-companion}
For every \(y\in\CC\), the minimal polynomial of \(T(y)\) equals its characteristic polynomial \(\Pi(\lambda,y)\).
In particular, if \(\Pi(\cdot,y_c)\) has a double root and no higher multiplicity, then \(T(y_c)\) has a size-\(2\) Jordan block at that eigenvalue.
\end{lemma}

\begin{proof}
$T(y)$ is a companion matrix, so its rational canonical form is a single block and the minimal and characteristic polynomials coincide \cite{HornJohnson2012}.
\end{proof}

In this paper the induced quartic transfer family is the primary object.
No separate microscopic realization is assumed within the present analysis.

