\section{Classification of real branch points and dominant-modulus roles}\label{sec:ep-classification}

Among the five finite branch values identified in Theorem~\ref{thm:elliptic-reduction}, exactly three are real.
By Lemma~\ref{lem:cyclic-companion}, every repeated root of \(\Pi(\cdot,y)\) corresponds to a defective eigenvalue of the companion family \(T(y)\).
In this section we determine their spectral role by analyzing the eigenvalue collision structure at each branch point and, crucially, by comparing the modulus of the colliding eigenvalues against the remaining roots.

The following elementary comparison is the only general local fact needed below:
a branch that leaves a repeated root at order \(t^{1/\nu}\) dominates any non-bifurcating simple branch whose modulus changes only at order \(t\).

\begin{lemma}[Local dominance from a double-root splitting]\label{lem:local-dominance}
Let \(P(\lambda,t)=\prod_{j=1}^n(\lambda-\mu_j(t))\) be a monic degree-\(n\) polynomial whose coefficients are analytic in \(t\) near \(t=0\).
Suppose that at \(t=0\) exactly two roots collide:
\[
\mu_1(0)=\mu_2(0)=:\mu_*,
\qquad
\mu_3(0),\dots,\mu_n(0) \text{ simple and distinct from } \mu_*.
\]
Assume that for all \(j\ge 3\),
\begin{equation}\label{eq:local-dominance-hyp}
|\mu_j(0)|\le |\mu_*|.
\end{equation}
Assume that one of the two bifurcating roots satisfies
\begin{equation}\label{eq:bifurc-form}
\abs{\mu_1(t)} \ge \abs{\mu_*} + a t^{1/2} - B t
\end{equation}
for some constants \(a,B>0\) and all real \(t>0\) sufficiently small.
Then there exists \(\delta>0\) such that
\begin{equation}\label{eq:local-dominance-concl}
\abs{\mu_1(t)}>\abs{\mu_j(t)}
\qquad\text{for every } j\ge 3 \text{ and } 0<t<\delta.
\end{equation}

\begin{proof}
Since the non-colliding roots \(\mu_3(0),\dots,\mu_n(0)\) are simple, the implicit function theorem gives analytic continuations \(\mu_j(t)\) and constants \(L_j>0\) such that
\[
\abs{\mu_j(t)-\mu_j(0)}\le L_j t
\qquad (j\ge 3,\ 0\le t\le t_0)
\]
for some \(t_0>0\).
\[
\abs{\mu_j(t)}
\le
\abs{\mu_j(0)}+L_j t
\le
\abs{\mu_*}+L_j t
=
\abs{\mu_*}+O(t)
\qquad (j\ge 3).
\]
On the other hand, the bifurcating branch satisfies
\[
\abs{\mu_1(t)}\ge \abs{\mu_*}+a t^{1/2}-Bt.
\]
Let \(L:=\max_{j\ge 3}L_j\).
Since \(t^{1/2}\gg t\) as \(t\to0^+\), choose \(0<\delta\le t_0\) so that
\[
Bt\le \frac{a}{2}t^{1/2},
\qquad
Lt\le \frac{a}{2}t^{1/2}
\qquad (0<t<\delta).
\]
Then for \(0<t<\delta\),
\[
\abs{\mu_1(t)}
\ge
\abs{\mu_*}+a t^{1/2}-Bt
\ge
\abs{\mu_*}+\frac{a}{2}t^{1/2}
>
\abs{\mu_*}+Lt
\ge
\abs{\mu_j(t)}
\]
for every \(j\ge 3\), proving \eqref{eq:local-dominance-concl}.
\end{proof}
\end{lemma}

\begin{theorem}[Real branch-point classification]\label{thm:ep-classification}
The three real branch values are
\[
y_{\mathrm{sub}} < 0, \qquad y_0 = 0, \qquad y_1 = 1,
\]
where $y_{\mathrm{sub}}$ is the unique real root of $c(y) = 256y^3 + 411y^2 + 165y + 32$.
Their spectral roles are as follows.
\begin{enumerate}
\item At $y = 0$, the spectral polynomial factors as
\[
\Pi(\lambda, 0) = \lambda(\lambda + 1)(\lambda - 1)^2.
\]
The double root $\lambda=1$ ties in modulus with the simple root $\lambda=-1$.
Nevertheless, for $y>0$ sufficiently small, the pair of branches issuing from $\lambda=1$ contains the unique dominant root.
Thus $y=0$ is the endpoint of the dominant branch over $y>0$.

\item At $y = 1$, the spectral polynomial factors as
\[
\Pi(\lambda, 1) = (\lambda - 2)(\lambda - 1)(\lambda + 1)^2.
\]
The double root $\lambda = -1$ is not the dominant eigenvalue; the positive real root $\lambda = 2$ remains simple.
Hence $y = 1$ is not a branch point of the dominant branch over $y>0$.

\item At $y = y_{\mathrm{sub}} \approx -1.13445$, the colliding double root is \emph{strictly subdominant}: it is exceeded in modulus by a pair of complex-conjugate roots.
Therefore $y_{\mathrm{sub}}$ is a subdominant branch point, not on the dominant-modulus envelope.
\end{enumerate}
\end{theorem}

\begin{proof}
We begin with the local dominance claim in part~(1).

\medskip
\noindent\textbf{Continuity of roots at $y=0$.}
At $y=0$ the factorization is $\Pi(\lambda,0)=\lambda(\lambda+1)(\lambda-1)^2$.
The three simple roots $\lambda\in\{0,-1\}$ extend analytically in $y$ by the implicit function theorem, since $\partial_\lambda\Pi$ is nonzero at each of them.
The double root at $\lambda=1$ produces a square-root ramification as described next.

\medskip
\noindent\textbf{Self-contained local expansion.}
Write $y=s^2$ for real $s>0$ small.
Set $\lambda=1+h$ directly in $\Pi(\lambda,y)=\lambda^4-\lambda^3-(2y+1)\lambda^2+\lambda+y(y+1)$ and collect terms:
\begin{align*}
\Pi(1+h,y)
&= (1+4h+6h^2+4h^3+h^4) - (1+3h+3h^2+h^3)\\
&\quad - (1+2y)(1+2h+h^2) + (1+h) + y^2+y\\
&= 2h^2 - y - 4yh + O(y^2, h^3, yh^2).
\end{align*}
Setting $y=s^2$ and seeking $h=\sigma s+O(s^2)$, the leading terms give $2\sigma^2 s^2 - s^2 = 0$, hence $\sigma^2=\tfrac12$, i.e.\ $\sigma=\pm\tfrac{1}{\sqrt{2}}$.
The correct square-root opening is confirmed by evaluating the quadratic discriminant:
\[
\frac{\partial^2\Pi}{\partial\lambda^2}\bigg|_{(1,0)} = 2\cdot\frac{\partial}{\partial\lambda}(4\lambda^3-3\lambda^2-(2y+1)\cdot 2\lambda+1)\bigg|_{(1,0)}=2(12-6-4)=4,
\]
\[
\frac{\partial\Pi}{\partial y}\bigg|_{(1,0)} = (-2\lambda^2+2y+1)\bigg|_{(1,0)}=-2+1=-1.
\]
By the Puiseux implicit function theorem for a double root (see, e.g., \cite{BrieskornKnorrer1986}), the two branches issuing from $(1,0)$ satisfy
\begin{equation}\label{eq:local-bifurc}
\lambda_\pm(y)
= 1 \pm \sqrt{\frac{-\partial_y\Pi}{\tfrac12\partial_\lambda^2\Pi}}\bigg|_{(1,0)} y^{1/2} + O(y)
= 1 \pm \frac{1}{\sqrt{2}}\,y^{1/2} + O(y).
\end{equation}
In particular, setting $y=s^2$,
\begin{equation}\label{eq:lambda-pm-local}
\lambda_+(s) = 1 + \frac{1}{\sqrt{2}}\,s + O(s^2),
\qquad
\lambda_-(s) = 1 - \frac{1}{\sqrt{2}}\,s + O(s^2).
\end{equation}

\medskip
\noindent\textbf{Spectral-radius comparison.}
For the remaining two roots near $(-1,0)$ and $(0,0)$, the implicit function theorem gives simple analytic extensions:
\begin{equation}\label{eq:other-roots-local}
\lambda_{-1}(y) = -1 + O(y),
\qquad
\lambda_0(y) = O(y).
\end{equation}
More precisely, differentiating $\Pi(\lambda,y)=\lambda^4-\lambda^3-(2y+1)\lambda^2+\lambda+y(y+1)$ with respect to $\lambda$:
\[
\partial_\lambda\Pi(\lambda,y) = 4\lambda^3-3\lambda^2-2(2y+1)\lambda+1.
\]
At $(\lambda,y)=(-1,0)$: $\partial_\lambda\Pi(-1,0) = -4-3+2+1=-4\neq 0$, so the implicit function theorem applies.
For the $y$-derivative: $\partial_y\Pi(\lambda,y)=-2\lambda^2+2y+1$, giving at $(-1,0)$: $\partial_y\Pi(-1,0)=-2+1=-1$.
Hence
\[
\frac{d\lambda_{-1}}{dy}\bigg|_{y=0} = -\frac{\partial_y\Pi(-1,0)}{\partial_\lambda\Pi(-1,0)} = -\frac{-1}{-4} = -\frac{1}{4},
\]
and
\[
\lambda_{-1}(y) = -1 - \frac{1}{4}y + O(y^2).
\]
For the root near $\lambda=0$: $\partial_\lambda\Pi(0,0)=1$ and $\partial_y\Pi(0,0)=1$, so
\[
\lambda_0(y) = -y+O(y^2).
\]
Therefore $|\lambda_{-1}(s^2)|=1+\tfrac{1}{4}s^2+O(s^4)>1$ for $s>0$ small, while $|\lambda_+(s)|=1+\tfrac{1}{\sqrt{2}}s+O(s^2)>1$, and $|\lambda_0(s^2)|=s^2+O(s^4)\to 0$ for $s\to 0^+$.

The critical comparison is $|\lambda_+(s)|$ vs $|\lambda_{-1}(s^2)|$: the bifurcating root $\lambda_+$ departs from $1$ at rate $\tfrac{1}{\sqrt{2}}s$ (first order in $s$), while the non-bifurcating root $\lambda_{-1}$ departs from $-1$ (hence $|\lambda_{-1}|$ departs from $1$) at rate $\tfrac{1}{4}s^2$ (second order in $s$):
\[
|\lambda_+(s)| - |\lambda_{-1}(s^2)| = \frac{1}{\sqrt{2}}s - \frac{1}{4}s^2 + O(s^3) > 0
\]
for $0<s<4/\sqrt{2}=2\sqrt{2}$.
For $|\lambda_-|$ and $|\lambda_0|$: $|\lambda_-(s)|=1-\tfrac{1}{\sqrt{2}}s+O(s^2)<1$ and $|\lambda_0(s^2)|=s^2+O(s^4)\to 0$.
Hence, for $s>0$ sufficiently small,
\begin{equation}\label{eq:dominance-ineqs}
\abs{\lambda_+(s)}
>
\abs{\lambda_{-1}(s^2)},
\qquad
\abs{\lambda_+(s)}
>
\abs{\lambda_-(s)},
\qquad
\abs{\lambda_+(s)}
>
\abs{\lambda_0(s^2)}.
\end{equation}
Thus the branch $\lambda_+(y^{1/2})$ is the unique dominant root for $y>0$ small, proving part~(1).

Part~(2) is immediate from
\[
\Pi(\lambda,1)=(\lambda-2)(\lambda-1)(\lambda+1)^2,
\]
since the double root has modulus \(1\) while the simple root \(\lambda=2\) has modulus \(2\).

We give the proof of part~(3) in detail.

\medskip
\noindent\textbf{Localization of the real root.}
Let
\begin{equation}\label{eq:cubic-lambda}
p(\lambda) = 16\lambda^3 - 9\lambda^2 + 1.
\end{equation}
Since $p(-1/2) = -13/4 < 0$ and $p(0) = 1 > 0$, the intermediate value theorem places a real root $\alpha \in (-1/2, 0)$.
To confirm uniqueness in $\RR$, note first that
\[
p'(\lambda) = 48\lambda^2 - 18\lambda = 6\lambda(8\lambda - 3).
\]
For $\lambda < 0$ both factors on the right-hand side are negative, so $p'(\lambda) > 0$ on $(-\infty,0)$.
Hence $p$ is strictly increasing on $(-\infty,0)$ and therefore has exactly one negative root.
On $[0,\infty)$, the only critical points are $\lambda = 0$ (local max, $p(0) = 1$) and $\lambda = 3/8$ (local min).
At the local minimum,
\[
p(3/8) = 16 \cdot \frac{27}{512} - 9 \cdot \frac{9}{64} + 1
= \frac{27}{32} - \frac{81}{64} + 1 = \frac{37}{64} > 0,
\]
so $p$ has no zero on $[0,\infty)$.
Therefore $\alpha$ is the unique real root of $p$.
Numerically, $\alpha \approx -0.273434810652454$.

\medskip
\noindent\textbf{Branch value.}
The corresponding branch value is
\begin{equation}\label{eq:ysub-formula}
y_{\mathrm{sub}} = \frac{4\alpha^3 - 3\alpha^2 - 2\alpha + 1}{4\alpha}
\approx -1.134452003008312.
\end{equation}

\medskip
\noindent\textbf{Factorization at $y = y_{\mathrm{sub}}$.}
At this branch value, $\lambda = \alpha$ is a double root.
Write
\begin{equation}\label{eq:Pi-ysub-factor}
\Pi(\lambda, y_{\mathrm{sub}}) = (\lambda - \alpha)^2 \bigl(\lambda^2 + b\lambda + q\bigr).
\end{equation}
Expanding and comparing the $\lambda^3$ coefficient gives
\[
-2\alpha + b = -1, \qquad \text{hence } b = 2\alpha - 1.
\]
Comparing the $\lambda$ coefficient yields
\[
-2\alpha q + \alpha^2(2\alpha - 1) = 1.
\]
Using $16\alpha^3 - 9\alpha^2 + 1 = 0$ to eliminate $\alpha^3 = (9\alpha^2 - 1)/16$, one solves
\begin{equation}\label{eq:q-formula}
q = \frac{\alpha^2 - 9}{16\alpha}.
\end{equation}
Since $\alpha \in (-1/2, 0)$, the numerator $\alpha^2 - 9 < 0$ and the denominator $16\alpha < 0$, so $q > 0$.

\medskip
\noindent\textbf{The remaining roots are complex.}
The discriminant of the residual quadratic is
\[
b^2 - 4q = (2\alpha - 1)^2 - \frac{\alpha^2 - 9}{4\alpha}
= \frac{-2\alpha^2 + \alpha + 2}{\alpha}.
\]
On $(-1/2, 0)$, the numerator $-2\alpha^2 + \alpha + 2$ is positive (its roots are $\alpha = (1 \pm \sqrt{17})/4$, both outside $(-1/2, 0)$), and the denominator $\alpha < 0$, so $b^2 - 4q < 0$.
The remaining two roots are therefore complex conjugates with modulus $\sqrt{q}$.

\medskip
\noindent\textbf{Subdominance of the double root.}
It remains to show $\sqrt{q} > |\alpha|$, i.e., $q > \alpha^2$.
We compute
\begin{equation}\label{eq:q-minus-alpha-sq}
q - \alpha^2 = \frac{\alpha^2 - 9 - 16\alpha^3}{16\alpha}.
\end{equation}
Using $16\alpha^3 = 9\alpha^2 - 1$, the numerator becomes
\[
\alpha^2 - 9 - (9\alpha^2 - 1) = -8\alpha^2 - 8 = -8(\alpha^2 + 1).
\]
Hence
\[
q - \alpha^2 = \frac{-8(\alpha^2 + 1)}{16\alpha} = -\frac{\alpha^2 + 1}{2\alpha} > 0,
\]
where positivity holds because $\alpha < 0$ and $\alpha^2 + 1 > 0$.
Therefore $\sqrt{q} > |\alpha|$, confirming that the double root at $\lambda = \alpha$ is strictly subdominant.
\end{proof}

\begin{remark}\label{rem:ysub-local-gap}
The strict inequality $\sqrt{q}>|\alpha|$ gives a positive modulus gap at \(y=y_{\mathrm{sub}}\).
By continuity of the roots of \(\Pi(\lambda,y)\) with respect to \(y\), there exists a neighborhood \(U\) of \(y_{\mathrm{sub}}\) such that, for \(y\in U\setminus\{y_{\mathrm{sub}}\}\), the two branches colliding at \(y_{\mathrm{sub}}\) remain strictly below the dominant modulus.
Thus \(y_{\mathrm{sub}}\) does not enter the dominant-modulus envelope even locally.
\end{remark}

\begin{remark}\label{rem:ep-physical}
In the spectral decomposition of the polynomial sequence, the limiting zero-accumulation curve is governed by equimodular competition among dominant eigenvalues, not by the discriminant locus alone.
A branch point belongs to the dominant spectral boundary only when the colliding eigenvalue is dominant.
Theorem~\ref{thm:ep-classification} shows that $y_{\mathrm{sub}}$ fails this dominance test: the colliding eigenvalue pair at $\lambda = \alpha$ is exceeded in modulus by the complex-conjugate pair of modulus $\sqrt{q}$.
Thus $y_{\mathrm{sub}}$ is a branch point of the discriminant curve that does not enter the dominant-modulus envelope.
\end{remark}

\begin{remark}\label{rem:yang-lee-distinction}
The Yang--Lee and Fisher papers \cite{YangLee1952I,LeeYang1952II,Fisher1978,KortmanGriffiths1971} concern zeros arising from specific statistical-mechanical models.
The present paper does not derive such a model, so the comparison is only terminological: the point $y=0$ found here is a square-root endpoint of the dominant-modulus boundary for this quartic family, not a derived Yang--Lee edge in the original physical sense.
\end{remark}

\subsection{Real ovals and the exact real-spectral phase diagram}

The elliptic normalization also determines the complete real-root count of the quartic transfer polynomial.

\begin{theorem}[Real-spectral phase diagram]\label{thm:real-phase-diagram}
Let $\beta_1 < \beta_2 < \beta_3$ be the three real roots of
\[
4\lambda^3 - 4\lambda + 1 = 0.
\]
Then the real locus of the elliptic curve
\[
E:\quad Y^2 = 4\lambda^3 - 4\lambda + 1
\]
decomposes as
\[
E(\RR)=E_{\mathrm{c}}(\RR)\sqcup E_{\mathrm{n}}(\RR),
\]
where
\[
E_{\mathrm{c}}(\RR)
=\{(\lambda,Y)\in\RR^2:\beta_1\le \lambda\le \beta_2,\ Y^2=4\lambda^3-4\lambda+1\}
\]
is a compact oval and
\[
E_{\mathrm{n}}(\RR)
=\{(\lambda,Y)\in\RR^2:\lambda\ge \beta_3,\ Y^2=4\lambda^3-4\lambda+1\}\cup\{O\}
\]
is the noncompact oval completed by the point at infinity $O$.
Under the degree-$4$ \(y\)-projection
\[
y=\lambda^2-\frac12+\frac12Y,
\]
their images are exactly
\begin{equation}\label{eq:real-oval-images}
y\bigl(E_{\mathrm{c}}(\RR)\bigr)=[y_{\mathrm{sub}},1],
\qquad
y\bigl(E_{\mathrm{n}}(\RR)\bigr)=[0,\infty).
\end{equation}
Consequently, for real $y$ away from the branch values $y_{\mathrm{sub}},0,1$, the quartic equation
\[
\Pi(\lambda,y)=0
\]
has exactly
\begin{equation}\label{eq:real-root-counts}
\#\{\lambda\in\RR:\Pi(\lambda,y)=0\}
=
\begin{cases}
0, & y<y_{\mathrm{sub}},\\
2, & y_{\mathrm{sub}}<y<0,\\
4, & 0<y<1,\\
2, & y>1.
\end{cases}
\end{equation}
In particular, the transfer family is completely real-rooted if and only if $y\in[0,1]$, agreeing with Proposition~\ref{prop:real-fibers-specialization}.
\end{theorem}

\begin{proof}
The cubic $4\lambda^3-4\lambda+1$ has three distinct real roots because its derivative
\[
12\lambda^2-4=4(3\lambda^2-1)
\]
vanishes at $\lambda=\pm 1/\sqrt{3}$, with
\[
4\Bigl(-\frac{1}{\sqrt{3}}\Bigr)^3-4\Bigl(-\frac{1}{\sqrt{3}}\Bigr)+1
=1+\frac{8}{3\sqrt{3}}>0,
\]
while
\[
4\Bigl(\frac{1}{\sqrt{3}}\Bigr)^3-4\Bigl(\frac{1}{\sqrt{3}}\Bigr)+1
=1-\frac{8}{3\sqrt{3}}<0.
\]
Hence the real locus of $E$ has two connected components, one lying above the bounded interval $[\beta_1,\beta_2]$ and one lying above $[\beta_3,\infty)$ together with $O$; this is the standard topology of a real elliptic curve with three real branch points \cite{Silverman2009,Miranda1995}.

The critical points of the \(y\)-projection on $E$ are determined by $\dd y=0$, hence by
\[
(\lambda-1)(\lambda+1)(16\lambda^3-9\lambda^2+1)=0
\]
from \eqref{eq:ramification-poly}.
By Theorem~\ref{thm:ep-classification}, the cubic factor has a unique real root $\alpha\in(-1/2,0)$, and its image is $y_{\mathrm{sub}}$.
Moreover,
\[
4(-1)^3-4(-1)+1=1>0,
\qquad
4\cdot 0^3-4\cdot 0+1=1>0,
\qquad
4\cdot 1+(-4)+1=1>0,
\]
while the cubic is negative at $\lambda=1/\sqrt3$.
Hence $\beta_1<-1$, $\beta_2\in(0,1/\sqrt3)$, and $\beta_3\in(1/\sqrt3,1)$, so
\[
\beta_1<-1<\alpha<\beta_2<\beta_3<1.
\]
the real ramification points split as follows:
the compact oval contains the two critical points with $\lambda=-1$ and $\lambda=\alpha$, while the noncompact oval contains the critical point with $\lambda=1$ and the pole at infinity.

Restrict $y$ to the compact oval $E_{\mathrm{c}}(\RR)$.
This restriction is a continuous real-valued function on a compact connected one-manifold, and its only critical values are $1$ and $y_{\mathrm{sub}}$.
Therefore the image is the closed interval $[y_{\mathrm{sub}},1]$.
Similarly, on the noncompact oval the only finite critical point is $y=0$, while $y\to+\infty$ along either branch approaching $O$ because $y$ has a pole of order $4$ at infinity.
Hence the noncompact image is $[0,\infty)$.
This proves \eqref{eq:real-oval-images}.

For real $y$ away from the branch locus, each point of $y^{-1}(y)\cap E(\RR)$ is simple and contributes a distinct real root $\lambda$ of $\Pi(\lambda,y)=0$.
Moreover, each regular value in the interior of an oval image has exactly two preimages on that oval.
Therefore:
for $y<y_{\mathrm{sub}}$ neither oval contributes real points;
for $y_{\mathrm{sub}}<y<0$ only the compact oval contributes, giving two real roots;
for $0<y<1$ both ovals contribute, giving four real roots;
and for $y>1$ only the noncompact oval contributes, giving two real roots.
The endpoint statement at $y=0$ and $y=1$ follows from Theorem~\ref{thm:ep-classification}, so complete real-rootedness occurs exactly on $[0,1]$.
\end{proof}
