\section{Discussion and scope}\label{sec:discussion-new}

The paper should be read as an explicit quartic case study.
Its main outputs are all tied to the specialization \((a,c,d,g)=(-1,-1,1,0)\): the elliptic normalization of the projective quartic, the exact finite branch values, the real-spectral phase diagram, the proof that the positive branch is globally dominant on \(y>0\), the endpoint cosine-law asymptotics at \(y=0\), the quarter-power splitting at projective infinity, and the degree-growth law for the polynomial sequence.
Section~\ref{sec:genus-one} also records three auxiliary balanced-family observations that organize the computation: the uniform Weierstrass reduction of Proposition~\ref{prop:balanced-family}, the componentwise real-projection description of Theorem~\ref{thm:balanced-real-projection}, and the normalized double-root determination of Theorem~\ref{thm:balanced-classification}.
Those statements are useful bookkeeping devices, but the paper does not claim a general classification of balanced quartics.

Relative to Biggs's equimodular-curve picture, the Beraha--Kahane--Weiss framework, and the recurrence literature of Tran, Forg\'acs--Tran, Tran--Zumba, Ndikubwayo, and related authors, the gain here comes from specializing completely to one quartic.
That specialization makes it possible to write the normalization, the branch values, the real component images, and the dominant positive branch in closed form.
The novelty is therefore not new general machinery, but a full exact analysis in a family where the relevant geometry is still tractable.

One conceptual point emerging from the calculation is the separation between the full ramification locus and the dominant-modulus envelope.
At \(y=0\), the branch collision occurs on the dominant branch and governs the first edge zeros.
At \(y=1\) and \(y=y_{\mathrm{sub}}\), the collisions are real but subdominant and therefore do not produce the same boundary behavior.
The real-spectral phase diagram makes this distinction visible, while the localized cosine law of Section~\ref{sec:physical-sheet} converts it into the odd-integer endpoint quantization.

At the opposite end of the parameter range, the large-\(y\) expansion is controlled by a double tangency at projective infinity.
This produces the quarter-power splitting of the dominant root.
Together with the degree law of Section~\ref{sec:support-law}, it shows that the top-degree sector is a rigid algebraic boundary rather than an exponentially large multiplicity regime.

Two extensions appear natural.
One is to derive the quartic family from a concrete combinatorial or statistical-mechanical model and thereby attach an interpretation to the coefficients.
The other is to study how much of Section~\ref{sec:genus-one} survives for other balanced parameters: the real branch-locus quintic \eqref{eq:balanced-quintic} and the interval comparison of Theorem~\ref{thm:balanced-real-projection} give the basic data, but the present paper does not attempt a general dominant-branch classification or a family-level zero-asymptotic theory.
