% ══════════════════════════════════════════════════════════
\section{Introduction}\label{sec:intro}
% ══════════════════════════════════════════════════════════

The idea that a single primitive can generate a rich family of
objects is a recurring structural motif in mathematics and logic.
In Boolean algebra the Sheffer stroke (NAND) suffices to express
every truth function~\cite{Sheffer1913}.
In the $\lambda$-calculus a single variable-binding construct
generates all computable functions~\cite{Church1936}.
In combinatory logic two combinators $S$ and~$K$ are enough;
with Sch\"onfinkel's encoding even a single combinator
suffices~\cite{Schonfinkel1924}.

A recent and striking addition to this lineage is
Odrzywołek's demonstration that \emph{every} elementary function
available on a standard scientific calculator---the 36-element
basis comprising arithmetic, algebraic, trigonometric, hyperbolic,
and logarithmic operations---can be reconstructed from a single
binary operator
\begin{equation}\label{eq:eml-def-intro}
  \eml(x,y) \;=\; e^{x} - \ln y
\end{equation}
together with the constant~$1$, using the minimal context-free
grammar $S \to 1 \mid \eml(S,S)$~\cite{Odrzywołek2026}.
This collapses the \emph{expression library} to a uniform binary
tree syntax---we call this \textbf{expression universality}.

The present note identifies expression universality as the bottom
rung of a three-level hierarchy whose upper levels are qualitatively
stronger.

\begin{enumerate}[label=\textbf{L\arabic*},leftmargin=2.5em]
\item \textbf{Expression universality} (Odrzywołek):
  a diverse function library is collapsed to a single operator and
  a tree grammar.  The operations of composition and evaluation
  remain external to the formalism.
\item \textbf{Protocol universality} (this paper, building on
  \cite{OmegaSPG2026}):
  integer arithmetic operations---addition and multiplication
  themselves---are not primitive but \emph{emerge} from a single
  scan--projection protocol.  The unique primitive is function
  composition~$\circ$, and $+$~and~$\times$ are its level-0 and
  level-1 readouts.  Canonical forms are computable, and the word
  problem for projection-word equivalence is decidable.
\item \textbf{Certificate universality} (this paper, building on
  \cite{OmegaSPG2026}):
  the fold fibers carry collision-moment sequences
  $\{S_q(m)\}_{q \ge 0}$ whose Hankel matrices are positive
  semi-definite.  A finite Prony--Hankel reconstruction recovers
  the full multiplicity spectrum from $2r$ moments, yielding an
  auditable, finite-dimensional spectral certificate at each
  resolution~$m$.
\end{enumerate}

\begin{table}[ht]
\centering\small
\caption{Three levels of single-primitive universality.}
\label{tab:hierarchy}
\begin{tabular}{@{}llll@{}}
\toprule
Level & What is unified & Primitive & Canonical form \\
\midrule
L0: Expression  & function library  & $\eml + 1$  & none intrinsic \\
L1: Protocol    & arithmetic ops    & $\circ$ (composition) & Zeckendorf NF \\
L2: Certificate & audit structure   & finite-state kernel & spectral fingerprint \\
\bottomrule
\end{tabular}
\end{table}

The hierarchy is strict (\Cref{sec:strictness}):
L1 subsumes~L0 because any expression-universal system can be
compiled into the protocol framework, but not conversely---the
obstruction is the undecidability of expression-tree equivalence
versus the decidability of fold normal forms.
L2 extends~L1 with an intrinsic positive-definite kernel that
is absent at the protocol level alone.

\paragraph{Relation to the main work.}
The protocol and certificate layers are developed in full
generality in~\cite{OmegaSPG2026}.  The purpose of this note is
narrower: to make the hierarchy explicit, to state the strictness
results cleanly, and to anchor the comparison with Odrzywołek's
EML operator as a concrete external reference point.
