\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{array}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{enumitem}

\numberwithin{equation}{section}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\EE}{\mathbb{E}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ind}{\mathbf{1}}
\DeclareMathOperator{\sgn}{sgn}

% Notation: \mathsf{P} denotes a measurable event (subset of the probability
% space X); italic P denotes a subset of the symbolic shift space
% \Omega_\infty=\{0,1\}^\mathbb{N}. When X=\Omega_\infty both coincide and we
% use \mathsf{P} throughout.

% target: Electronic Communications in Probability

\title{A Note on Prefix-Partition Scan Error and Conditional Boundary Rate Laws}
\author{
  Haobo Ma\thanks{AELF PTE LTD., \#14-02, Marina Bay Financial Centre Tower 1, 8 Marina Blvd, Singapore 018981. Email: \texttt{auric@aelf.io}}
  \and
  Wenlin Zhang\thanks{Corresponding author. National University of Singapore, Singapore. Email: \texttt{e1327962@u.nus.edu}}
}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We study oracle classification error along a fixed prefix filtration and keep
the scope deliberately conditional: once the filtration and reference measure
are fixed, the Bayes risk $\varepsilon_m(\mathsf{P};\mu)$ decays exactly as
the weighted boundary mass.  The structural input is the discrete Tanaka
identity $\varepsilon_m=\varepsilon_0-\EE[\mathcal{T}_m]$ applied to the posterior
martingale, together with a weighted boundary law expressing
$\varepsilon_m$ as the boundary-mass sum of local ambiguities.  Standard
count-based upper and lower bounds follow whenever cylinders of the same
length are uniformly comparable, and the same decomposition enforces lower
growth of mixed cylinders from any prescribed lower bound on $\varepsilon_m$.
The genuinely new content is an intrinsic hidden-Markov/sofic theorem:
for primitive absorbing prefix filters on mixing shifts of finite type with the
Parry measure, the mixed posterior states form a finite strongly connected
graph whose spectral radius simultaneously yields automatic thickness and
determines the exponential growth of boundary cylinders.  This places the
conditional rate laws inside the classical hidden-Markov literature while
remaining honest about scope.  The worked examples include a
positive-measure run-length hidden-Markov example, the parity and self-similar
cases, and a boundary-growth failure illustrating the necessity of
thickness.
A polynomial-distortion variant is recorded separately as an auxiliary
remark rather than as a Gibbs theorem.
\end{abstract}

\tableofcontents

\section{Introduction}
\label{sec:intro}

This note studies a fixed-filtration oracle question at the level appropriate
for a short communication: how much does refining prefix information improve
the best possible decision error for a target event?  Let
$(X,\mathcal{B},\mu)$ be a probability space with an increasing filtration
$(\mathcal{F}_m)_{m\ge 0}$ generated by $\{0,1\}$-valued functions
$s_0,s_1,\ldots$  The scan error profile
\[
m\longmapsto\varepsilon_m(\mathsf{P};\mu)
:=\inf\bigl\{\mu(\mathsf{P}\triangle\mathsf{Q}):\mathsf{Q}\in\mathcal{F}_m\bigr\}
\]
records the smallest classification error at each prefix depth.

The Bayes cylinder decomposition and the optimality of the threshold classifier
$p_m\ge 1/2$ are classical
\cite{Tsybakov2009Nonparametric,DevroyeGyorfiLugosi1996}.  We recall them in
Section~\ref{sec:scan} only in the form needed later.  The only source of
genuinely new mathematics is the hidden-Markov/sofic theorem stated below;
all remaining statements are reorganizations of standard identities inside a
fixed filtration.  The main points are:
\begin{enumerate}[label=(\roman*)]
\item A \emph{discrete Tanaka identity}
  (Proposition~\ref{prop:scan-tanaka}) connecting error decay to the accumulation
  of threshold crossings of the martingale $(p_m)$.  This is a structural
  observation that follows from the standard discrete-time Tanaka--Meyer
  formula \cite{LochowskiOblojPromelSiopaes2021}.
\item A \emph{weighted boundary law}
  (Proposition~\ref{prop:weighted-boundary-law}) expressing $\varepsilon_m$ as
  boundary mass times averaged local ambiguity, together with count-based
  consequences and a converse-style lower-growth statement
  (Propositions~\ref{prop:boundary-dimension}--\ref{prop:volume-boundary-scaling}
  and Proposition~\ref{prop:boundary-growth-necessary}).
\item An \emph{intrinsic hidden-Markov/sofic class}
  (Definition~\ref{def:hidden-markov-filter} and
  Theorem~\ref{thm:hidden-markov-thickness}) whose mixed posterior states form
  a finite strongly connected graph.  Primitive absorbing prefix filters thereby
  deliver automatic thickness and a spectral formula for the boundary growth,
  placing the conditional rate laws squarely inside the standard literature on
  sofic factors and probabilistic automata
  \cite{Blackwell1957,Heller1965,BaumPetrie1966,BoylePetersenHMP,PollicottKempton2016,Verbitskiy2011,Rabin1963,Paz1971,PerrinPin2004}.
\item \emph{Worked examples and failure modes}: a benchmark hidden-Markov
  factor example, the first-hit parity event, the self-similar boundary exponent
  $\frac12$, and failure examples showing the necessity of thickness, the effect
  of filtration choice, and the detailed computation for
  Example~\ref{ex:thickness-failure}.
\end{enumerate}
The novelty claim is therefore precise: the hidden-Markov/sofic theorem is
the only new statement, whereas the remaining material is recorded so that the
conditional rate laws can be cited cleanly.  The structural observations are
nonetheless useful because they isolate what is genuinely filtration-driven
and what must be assumed from outside (thickness, cylinder comparability).
Section~\ref{sub:related} provides pointers to the closest partition, hidden
Markov, and Gibbs literatures.
To keep the focus on Theorem~\ref{thm:hidden-markov-thickness}, the
preliminary Sections~\ref{sec:prefix}--\ref{sec:tanaka} and the auxiliary
log-Gibbs discussion in Section~\ref{sub:nonuniform} have been shortened to
bare statements with explicit citations; routine proofs are omitted.

\subsection{Heuristic comparison with margin assumptions in classification}
\label{sub:margin}

The thickness hypothesis in Proposition~\ref{prop:volume-boundary-scaling}
should be read only as a heuristic comparison with margin and noise conditions
in nonparametric classification.
Mammen and Tsybakov \cite{MammenTsybakov1999} introduce a margin condition
with exponent $\alpha\ge 0$: for the regression function
$p_\infty(x)=\mu(\mathsf{P}\mid\mathcal{F}_\infty)(x)$,
\[
\mu\!\left(\left|p_\infty(x)-\tfrac12\right|\le t\right)\le C_0\,t^\alpha
\qquad\text{for all }t>0.
\]
Tsybakov \cite{Tsybakov2004} extends this to a general noise-exponent
framework, showing that margin conditions control the rate of optimal
plug-in classifiers.

Our uniform thickness condition
$\min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\}\ge\theta>0$ for all
boundary cylinders $C\in\partial_m\mathsf{P}$ is a \emph{uniform two-sided
nondegeneracy} assumption on individual mixed cylinders: it demands that
every boundary cylinder carries a bounded-away-from-zero fraction of
both $\mathsf{P}$ and $\mathsf{P}^c$.  The Mammen--Tsybakov condition is,
by contrast, a \emph{global distributional} statement about the regression
function $p_\infty$ near $\frac12$; the exponent $\alpha$ is not a fixed
threshold but a rate parameter.

The two conditions are related but structurally different:
\begin{enumerate}[label=(\arabic*)]
\item Uniform thickness prevents one-sided concentration on individual
  cylinders, while the margin condition integrates over the whole space.
  A measure satisfying the margin condition with large $\alpha$
  can still have individual boundary cylinders with $\mu(\mathsf{P}\mid C)\to 0$,
  violating uniform thickness.
\item Thickness controls \emph{oracle} error (best decision with known measure),
  whereas the margin exponent controls \emph{estimation} error (rate of
  learning from data).  These are different quantities, and the fixed
  constant $\theta$ does not play the same role as $\alpha$.
\end{enumerate}
For a more direct point of comparison in the dyadic-partition context, Scott
and Nowak \cite{ScottNowak2006} combine a dyadic box-counting boundary class
with noise assumptions that are closer in spirit to our setup; see
Section~\ref{sub:related} for a detailed comparison.

\subsection{Related work}
\label{sub:related}

Bayes risk on partitions and optimality of threshold classifiers are classical;
see \cite{DevroyeGyorfiLugosi1996} and \cite{AudibertTsybakov2007} for refined
plug-in rates.  The underlying martingale convergence is Doob's upward theorem
\cite{Doob1953,Williams1991,Durrett2019}.  The discrete Tanaka formula is a
lattice analogue of Tanaka's formula for Brownian local time
\cite{Tanaka1963,RevuzYor1999}; Proposition~\ref{prop:scan-tanaka} is used here
only as a structural identity, not as a new Tanaka theorem.

\paragraph{Classical hidden-Markov references.}
The probabilistic-function program initiated by
\cite{Blackwell1957,Heller1965,BaumPetrie1966} already describes finite-state
filters of Markov chains, while \cite{Thomas1990,PerrinPin2004,Kitchens1998}
develop the automata-on-infinite-words and symbolic-dynamics background that we
invoke.  Section~\ref{sec:rates} uses this machinery only to the extent needed
to identify when the fixed-filtration oracle quantities inherit automatic
thickness and Perron--Frobenius growth.

Within that classical landscape, Theorem~\ref{thm:hidden-markov-thickness}
chooses the absorbing prefix-filter viewpoint: we keep the reference filtration
and measure fixed, treat the finite-state filter as a transducer producing the
oracle posterior, and ask only for quantitative consequences (automatic
thickness and explicit boundary growth) for that filtration.  Thus the result
does not reprove the existence of hidden-Markov factors in the sense of
\cite{Blackwell1957,BaumPetrie1966} or of omega-regular acceptance conditions
\`a la \cite{Thomas1990,PerrinPin2004}; instead it pinpoints how those classical
structures feed into the scan-error decay studied here.

Cylinder partition structure in symbolic dynamics is treated in
\cite{LindMarcus1995}; boundary combinatorics connect to
\cite{Fischer1975,Krieger1984,Boyle1983}.  Exponential cylinder estimates for
the Parry measure (measure of maximal entropy) are in \cite{Parry1964}; for
general Gibbs measures the corresponding cylinder bounds are
potential-dependent rather than uniform over all cylinders of a fixed length
\cite{Bowen1975,WaltersErgodicTheory}.
The Shannon--McMillan--Breiman theorem \cite{Shannon1948,McMillan1953,Breiman1957}
gives almost-sure asymptotics for the cylinder \emph{containing} a given
sample point; this is an almost-sure pointwise statement, weaker than the
uniform all-cylinder bounds assumed in our rate theorems (see
Section~\ref{sec:discussion}).  Background on dynamics and information theory
is in \cite{WaltersErgodicTheory,LindMarcus1995,CoverThomas2006,MorseHedlund1940}.

\paragraph{Closer neighboring statistical literature.}
A closer comparison point than nonlinear approximation alone is the literature
on adaptive partition classification and level-set estimation.  Binev et al.\
\cite{BinevCohenDahmenDeVore2014} study adaptive partition classifiers as set
approximators to the Bayes set.  Willett and Nowak
\cite{WillettNowak2007}, Singh et al.\ \cite{SinghScottNowak2009}, and Scott
and Davenport \cite{ScottDavenport2007} analyze level-set estimation with
hierarchical partitions, Hausdorff loss, or cost-sensitive reductions.
Historically, CART \cite{BreimanFriedmanOlshenStone1984} is the baseline tree
partition framework behind much of this literature.  The distinction from the
present paper is threefold: the measure is known rather than estimated, the
filtration is fixed rather than adaptively selected, and the quantity of
interest is the exact oracle Bayes risk along that single filtration rather
than minimax estimation risk.  In that sense the present results lie on the
decision-theoretic side of partition refinement; compare Blackwell's
comparison-of-experiments viewpoint \cite{Blackwell1951}.

\paragraph{Comparison with dyadic decision-tree and approximation literature.}
The boundary-count viewpoint is closest to dyadic decision-tree work in
statistical learning, especially Scott and Nowak \cite{ScottNowak2006} and
Blanchard et al.~\cite{BlanchardSchaferRozenholcMuller2007}.  Scott and
Nowak study \emph{sample-based} classification with dyadic decision trees
under a dyadic box-counting boundary class together with noise assumptions.
Blanchard et al.\ analyze optimal dyadic decision trees and oracle
inequalities for data-driven tree selection.  A second comparison point is
the adaptive-partition approximation literature of Donoho
\cite{Donoho1997}, DeVore \cite{DeVore1998}, and Cohen et al.\
\cite{CohenDahmenDaubechiesDeVore2001}, where dyadic trees encode functions
through best-basis or best-tree approximants.  Our setting is narrower: the
measure is known, the filtration is fixed in advance, and the object of study
is the \emph{oracle prefix-Bayes error} $\varepsilon_m(\mathsf{P};\mu)$ for a
single prescribed chain of prefix partitions.  The shared ingredient is
therefore the count of mixed dyadic cells, but the conclusions differ:
here we obtain direct oracle identities and rate bounds along that fixed
filtration, rather than minimax learning rates, empirical oracle
inequalities, or best-tree approximation theorems.

\begin{table}[ht]
\centering
{\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{1.00in}>{\raggedright\arraybackslash}p{0.70in}>{\raggedright\arraybackslash}p{0.76in}>{\raggedright\arraybackslash}p{0.82in}>{\raggedright\arraybackslash}p{1.18in}>{\raggedright\arraybackslash}p{1.00in}@{}}
\toprule
reference & setting & object & information & complexity / noise & conclusion \\
\midrule
Scott--Nowak \cite{ScottNowak2006} &
dyadic classification &
classifier / boundary &
sampled data &
box-counting boundary class plus noise assumptions &
minimax classification rates \\
Blanchard et al.\ \cite{BlanchardSchaferRozenholcMuller2007} &
empirical dyadic trees &
tree classifier &
sampled data &
tree penalty and oracle model selection &
oracle inequalities; optimal trees \\
Donoho \cite{Donoho1997}; DeVore \cite{DeVore1998}; Cohen et al.\ \cite{CohenDahmenDaubechiesDeVore2001} &
adaptive approximation &
function / encoding &
target known exactly &
best-basis or tree budget; no noise layer &
best-tree approximation and optimal encoding \\
this paper &
fixed prefix filtration &
event indicator / Bayes decision &
known measure &
mixed-cylinder count plus thickness on mixed cells &
oracle decay of $\varepsilon_m$ \\
\bottomrule
\end{tabular}
\par}
\caption{Side-by-side positioning against the closest dyadic-tree and adaptive-approximation references.}
\label{tab:dyadic-comparison}
\end{table}

\paragraph{Hidden-Markov and Gibbs viewpoints.}
Hidden-Markov and sofic measures are treated extensively in the symbolic
dynamics literature \cite{BoylePetersenHMP,Verbitskiy2011}, building on the
classical probabilistic-function program of
\cite{Blackwell1957,Heller1965,BaumPetrie1966}.  Automata on infinite objects
and symbolic-coding references such as
\cite{Thomas1990,PerrinPin2004,Kitchens1998} provide the complementary
background for the automaton formulation used here.  Factors of Gibbs measures
give the dynamical context for potential-weighted cylinder estimates
\cite{PollicottKempton2016}, and the probabilistic-automata models of Rabin and
Paz \cite{Rabin1963,Paz1971} already encode the finite-state filters that
appear in Section~\ref{sec:rates}.  The present paper does not attempt a full
thermodynamic treatment of hidden-Markov factors; our contribution is instead to
identify exactly where such classical structures provide automatic thickness and
spectral growth inside the fixed-filtration oracle picture.

\section{The Prefix Filtration Setting}
\label{sec:prefix}

The results hold in the following general setting; $T$-invariance and dynamics
are not used in any proof.

\begin{definition}[Prefix filtration]
\label{def:prefix-filtration}
Let $(X,\mathcal{B},\mu)$ be a probability space and $(s_t)_{t\ge 0}$ a
sequence of $\{0,1\}$-valued measurable functions.  Set $\mathcal{F}_0:=\{\varnothing,X\}$
and $\mathcal{F}_m:=\sigma(s_0,\ldots,s_{m-1})$ for $m\ge 1$.  The family
$(\mathcal{F}_m)_{m\ge 0}$ is the \emph{prefix filtration}; set
$\mathcal{F}_\infty:=\bigvee_m\mathcal{F}_m$.
\end{definition}

The dynamical specialization is $s_t(x):=\ind_W(T^tx)$ for a measurable set
$W$; the symbolic specialization is $X=\Omega_\infty:=\{0,1\}^{\mathbb{N}}$
and $s_t(\xi):=\xi_t$.  All results below hold for any prefix filtration.

For $a=a_0\cdots a_{m-1}\in\{0,1\}^m$, the length-$m$ cylinder in
$\Omega_\infty$ is $[a]:=\{\xi:\xi_0\cdots\xi_{m-1}=a\}$.  The readout map
$\omega_\infty(x):=s_0(x)s_1(x)\cdots$ pulls back sigma-algebras:
$\mathcal{F}_m=\omega_\infty^{-1}(\sigma([a]:a\in\{0,1\}^m))$.

\begin{proposition}[Prefix-decidable events are clopen cylinder unions]
\label{prop:decidable-clopen}
A set $P\subseteq\Omega_\infty$ is in $\mathcal{A}_m:=\sigma([a]:a\in\{0,1\}^m)$
if and only if it is a union of length-$m$ cylinders $[a]$, equivalently a
finite union of clopen balls of radius $2^{-m}$, hence a clopen set in the
prefix ultrametric.  Pulling back along
$\omega_\infty$, if $\mathsf{P}\in\mathcal{F}_m$ then the truth of
$\mathsf{P}$ is determined by the first $m$ symbols of the readout.
\end{proposition}

This classical observation is immediate because the atoms of
$\mathcal{A}_m$ are precisely the length-$m$ cylinders, which are clopen balls
in the prefix ultrametric; pulling back along $\omega_\infty$ therefore
transfers the statement to $X$, so we suppress the formal proof.

\subsection{Standing assumptions and scope}
\label{sub:standing}

Electronic Communications in Probability favors short notes, so we spell out
the only hypotheses used repeatedly:
\begin{enumerate}[label=(\alph*)]
\item \emph{Events.}  All events lie in $\mathcal{F}_\infty$ so that the
  martingale limit $p_\infty$ and $\varepsilon_m\to 0$ are well defined.
\item \emph{Cylinder comparability.}  Whenever a theorem invokes uniform (or
  polynomial) cylinder estimates we state it explicitly; otherwise the identity
  $\varepsilon_m=\mathcal{W}_m\overline{\theta}_m$ is the only conclusion.
\item \emph{Thickness.}  Lower bounds require a uniform two-sided ambiguity on
  mixed cylinders.  This is treated as an assumption except in
  Theorem~\ref{thm:hidden-markov-thickness}, which derives it for primitive
  filters.
\end{enumerate}
Referencing these conditions directly keeps later diagnostics simple (for
instance, Example~\ref{ex:thickness-failure} violates only thickness).

\section{The Scan Error Profile}
\label{sec:scan}

The next two propositions are standard partition-Bayes facts.  They are
included only to fix notation and to isolate the exact decomposition reused in
Section~\ref{sec:rates}.

\begin{proposition}[Exact cylinder decomposition]
\label{prop:scan-error-cylinder}
Let $\mathcal{C}_m$ be the atoms of $\mathcal{F}_m$ and $p_m:=\mu(\mathsf{P}\mid\mathcal{F}_m)$.
For every $\mathsf{P}\in\mathcal{F}_\infty$ and $m\ge 0$,
\[
\varepsilon_m(\mathsf{P};\mu)
=\sum_{C\in\mathcal{C}_m}\min\{\mu(\mathsf{P}\cap C),\mu(C\setminus\mathsf{P})\}
=\EE[\min\{p_m,1-p_m\}].
\]
\end{proposition}

The statement is the standard Bayes partition identity
\cite[Chapter~6]{DevroyeGyorfiLugosi1996}; on each atom $C\in\mathcal{C}_m$ the
optimal action chooses the smaller of $\mu(\mathsf{P}\cap C)$ and
$\mu(C\setminus\mathsf{P})$, and summing those minima yields the
expectation formula above.

\begin{proposition}[Bayes-optimality]
\label{prop:bayes-optimality}
The threshold classifier $\mathsf{P}_m^\star:=\{x:p_m(x)\ge 1/2\}$ achieves
$\mu(\mathsf{P}\triangle\mathsf{P}_m^\star)=\varepsilon_m(\mathsf{P};\mu)$, and
$|\int\varphi\ind_{\mathsf{P}}\,d\mu-\int\varphi\ind_{\mathsf{P}_m^\star}\,d\mu|\le\|\varphi\|_\infty\varepsilon_m$
for every bounded $\mathcal{F}_\infty$-measurable $\varphi$.
\end{proposition}

The proof follows by applying Proposition~\ref{prop:scan-error-cylinder} to the
threshold decision on each atom; see again \cite{DevroyeGyorfiLugosi1996}.

\begin{corollary}[Monotonicity and decidability]
\label{cor:basic-properties}
The profile $m\mapsto\varepsilon_m(\mathsf{P};\mu)$ is nonincreasing, and
$\varepsilon_m(\mathsf{P};\mu)=0$ iff $\mathsf{P}$ coincides $\mu$-a.s.\ with
an event in $\mathcal{F}_m$.
\end{corollary}

\section{Discrete Tanaka Representation}
\label{sec:tanaka}

We record only the specialized Tanaka representation used later.  The point is
not to establish a new stochastic-calculus formula, but to rewrite the exact
Bayes error in a form adapted to the posterior martingale.

\begin{proposition}[Discrete Tanaka identity for the scan error]
\label{prop:scan-tanaka}
The following identity is a direct consequence of the discrete-time
Tanaka--Meyer formula applied to the posterior martingale $(p_m)$; see
\cite{LochowskiOblojPromelSiopaes2021} for the general c\`adl\`ag setting.
Let $p_m:=\mu(\mathsf{P}\mid\mathcal{F}_m)$.  For $m\ge 1$ define the
threshold-crossing compensator
\[
\mathcal{T}_m
:=\sum_{k=0}^{m-1}
\left(
\left|p_{k+1}-\tfrac12\right|
-\left|p_k-\tfrac12\right|
-\sgn\!\left(p_k-\tfrac12\right)(p_{k+1}-p_k)
\right),
\]
with $\sgn(0):=0$.  The quantity $\mathcal{T}_m$ measures how often $(p_m)$
crosses the decision threshold $\frac12$, so no new local-time statement is
needed.  Moreover, $\mathcal{T}_m\ge 0$ and $m\mapsto \mathcal{T}_m$ is
nondecreasing, and
\[
\varepsilon_m(\mathsf{P};\mu)=\varepsilon_0(\mathsf{P};\mu)-\EE[\mathcal{T}_m],
\qquad
\varepsilon_0(\mathsf{P};\mu)=\min\{\mu(\mathsf{P}),1-\mu(\mathsf{P})\}.
\]
Since $\mathsf{P}\in\mathcal{F}_\infty$, we have $\varepsilon_m(\mathsf{P};\mu)\to 0$
and $\EE[\mathcal{T}_\infty]=\varepsilon_0(\mathsf{P};\mu)$.
\end{proposition}

\begin{proof}
Apply the discrete Tanaka--Meyer formula from
\cite{Tanaka1963,LochowskiOblojPromelSiopaes2021} to $p_m$ at the threshold
$\tfrac12$, which produces the compensator $\mathcal{T}_m$ above and the
identity $\varepsilon_m=\varepsilon_0-\EE[\mathcal{T}_m]$.  Monotone
convergence for $(\mathcal{T}_m)$ together with Doob's martingale convergence
theorem then yields $\varepsilon_m\downarrow 0$ and
$\EE[\mathcal{T}_\infty]=\varepsilon_0$.
\end{proof}

\section{Boundary-Controlled Rates}
\label{sec:rates}

\subsection{Boundary cylinders, boundary mass, and the growth scale}

\begin{definition}[Boundary cylinders, boundary mass, and filtration boundary exponent]
\label{def:boundary-cylinder}
For $\mathsf{P}\in\mathcal{F}_\infty$, the boundary-cylinder family at
resolution $m$ is
\[
\partial_m\mathsf{P}
:=\{C\in\mathcal{C}_m:\mu(\mathsf{P}\cap C)>0\text{ and }\mu(C\setminus\mathsf{P})>0\},
\]
with cardinality $N_m(\partial\mathsf{P}):=\#(\partial_m\mathsf{P})$.  Its
total boundary mass is
\[
\mathcal{W}_m(\mathsf{P};\mu):=\sum_{C\in\partial_m\mathsf{P}}\mu(C).
\]

Let $M_m:=\#\{C\in\mathcal{C}_m:\mu(C)>0\}$.  When the limit
$\lambda_{\mathrm{cyl}}:=\lim_{m\to\infty}M_m^{1/m}$ exists and exceeds $1$,
it is the \emph{cylinder growth base}.  For a mixing SFT,
$\lambda_{\mathrm{cyl}}$ is the spectral radius of the transition matrix
\cite{LindMarcus1995}.  The \emph{upper filtration boundary exponent} is
\[
\overline{\beta}_{\mathcal{F}}(\mathsf{P})
:=\limsup_{m\to\infty}
\frac{\log N_m(\partial\mathsf{P})}{m\log\lambda_{\mathrm{cyl}}}.
\]
\end{definition}

The quantity $\beta=\overline{\beta}_{\mathcal{F}}(\mathsf{P})\in[0,1]$ is an
exponential growth rate attached to the chosen filtration, not an intrinsic
geometric dimension of the underlying measurable set.  Fixed block recodings
preserve $\beta$ because both prefix length and cylinder-growth base are
rescaled by the same constant, but general filtration changes need not; see
Example~\ref{ex:filtration-dependence}.

The count $N_m(\partial\mathsf{P})$ is measure-theoretic rather than purely
topological: a cylinder is counted only when both $\mu(\mathsf{P}\cap C)$ and
$\mu(C\setminus\mathsf{P})$ are positive.  This is close in spirit to the
dyadic box-counting boundary classes of Scott and Nowak
\cite{ScottNowak2006}, but the role is different.  Their boundary counts enter
sample-based classification rates under noise assumptions, whereas here the
same style of count controls the oracle scan error for a fixed known measure.
Likewise, Blanchard et al. \cite{BlanchardSchaferRozenholcMuller2007} optimize
over dyadic trees chosen from data, while our filtration is fixed in advance
and only the best decision rule on each prefix level changes.

\subsection{Weighted and count-based rate laws}

\begin{proposition}[Weighted boundary law]
\label{prop:weighted-boundary-law}
For each $m\ge 0$, define the averaged boundary ambiguity
\[
\overline{\theta}_m(\mathsf{P};\mu)
:=
\begin{cases}
\dfrac{\sum_{C\in\partial_m\mathsf{P}}
\mu(C)\min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\}}
{\mathcal{W}_m(\mathsf{P};\mu)}, & \mathcal{W}_m(\mathsf{P};\mu)>0,\\[2ex]
0, & \mathcal{W}_m(\mathsf{P};\mu)=0.
\end{cases}
\]
Then
\[
\varepsilon_m(\mathsf{P};\mu)
=\mathcal{W}_m(\mathsf{P};\mu)\,\overline{\theta}_m(\mathsf{P};\mu).
\]
Moreover, $0\le \overline{\theta}_m(\mathsf{P};\mu)\le \frac12$ for all $m$.
If $\mathcal{W}_m(\mathsf{P};\mu)>0$ eventually and
$\liminf_{m\to\infty}\overline{\theta}_m(\mathsf{P};\mu)>0$, then
$\varepsilon_m(\mathsf{P};\mu)\asymp \mathcal{W}_m(\mathsf{P};\mu)$.
\end{proposition}

\begin{proof}
By Proposition~\ref{prop:scan-error-cylinder},
\[
\varepsilon_m(\mathsf{P};\mu)
=\sum_{C\in\partial_m\mathsf{P}}
\mu(C)\min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\},
\]
because non-boundary cylinders contribute zero.  If $\mathcal{W}_m(\mathsf{P};\mu)=0$,
then both sides vanish.  Otherwise dividing by $\mathcal{W}_m(\mathsf{P};\mu)$ gives the
formula.  The bound $\overline{\theta}_m\le\frac12$ is immediate from
$\min\{p,1-p\}\le\frac12$.  The final comparability statement follows by
bounding $\overline{\theta}_m$ above by $\frac12$ and below by a positive
constant for large $m$.
\end{proof}

The next two statements are count-based corollaries of
Proposition~\ref{prop:weighted-boundary-law}: same-length cylinder estimates
control $\mathcal{W}_m(\mathsf{P};\mu)$ by comparing each boundary cylinder to
$\lambda_{\mathrm{cyl}}^{-m}$.

\begin{proposition}[Count-based upper bound from cylinder masses]
\label{prop:boundary-dimension}
Suppose $\mu(C)\le K\lambda_{\mathrm{cyl}}^{-m}$ for all $m\ge 1$ and all
$C\in\mathcal{C}_m$.  Then
\[
\varepsilon_m(\mathsf{P};\mu)\le K\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P}).
\]
If $\overline{\beta}_{\mathcal{F}}(\mathsf{P})=\beta<1$, then
\[
\varepsilon_m(\mathsf{P};\mu)\le\lambda_{\mathrm{cyl}}^{-(1-\beta)m+o(m)}.
\]
\end{proposition}

\begin{proof}
By Proposition~\ref{prop:weighted-boundary-law},
\[
\varepsilon_m(\mathsf{P};\mu)\le \mathcal{W}_m(\mathsf{P};\mu)
\le K\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P}).
\]
If $\overline{\beta}_{\mathcal{F}}(\mathsf{P})=\beta<1$, fix $\eta>0$.
For all large $m$,
$N_m(\partial\mathsf{P})\le\lambda_{\mathrm{cyl}}^{(\beta+\eta)m}$, so
\[
\varepsilon_m(\mathsf{P};\mu)\le
K\lambda_{\mathrm{cyl}}^{-(1-\beta-\eta)m}.
\]
Since $\eta$ and the constant $K$ contribute only $o(m)$ on the logarithmic
scale, the claimed exponent follows.
\end{proof}

\begin{proposition}[Two-sided count law under boundary thickness]
\label{prop:volume-boundary-scaling}
Suppose $c_-\lambda_{\mathrm{cyl}}^{-m}\le\mu(C)\le c_+\lambda_{\mathrm{cyl}}^{-m}$
for all $C\in\mathcal{C}_m$ and all $m\ge 1$, and suppose there exist
$\theta\in(0,\frac12]$ and $m_0$ such that for all $m\ge m_0$ and all
$C\in\partial_m\mathsf{P}$,
\[
\min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\}\ge\theta.
\]
Then for all $m\ge m_0$,
\[
\theta c_-\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})
\le\varepsilon_m(\mathsf{P};\mu)
\le\frac{c_+}{2}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P}).
\]
If $N_m(\partial\mathsf{P})=\lambda_{\mathrm{cyl}}^{\beta m+o(m)}$, then
\[
\lim_{m\to\infty}\left(-\frac{1}{m}
\log_{\lambda_{\mathrm{cyl}}}\varepsilon_m(\mathsf{P};\mu)\right)=1-\beta.
\]
\end{proposition}

\begin{proof}
By Proposition~\ref{prop:weighted-boundary-law},
\[
\varepsilon_m(\mathsf{P};\mu)
=\sum_{C\in\partial_m\mathsf{P}}
\mu(C)\min\{\mu(\mathsf{P}\mid C),1-\mu(\mathsf{P}\mid C)\}.
\]
The lower bound follows from the thickness hypothesis together with
$\mu(C)\ge c_-\lambda_{\mathrm{cyl}}^{-m}$ on every boundary cylinder.  The
upper bound uses $\min\{p,1-p\}\le\frac12$ and
$\mu(C)\le c_+\lambda_{\mathrm{cyl}}^{-m}$.  If
$N_m(\partial\mathsf{P})=\lambda_{\mathrm{cyl}}^{\beta m+o(m)}$, inserting this
into the two-sided estimate and dividing by $m$ on the logarithmic scale gives
$1-\beta$.
\end{proof}

\begin{proposition}[Lower error forces boundary growth]
\label{prop:boundary-growth-necessary}
Assume the upper cylinder bound from Proposition~\ref{prop:boundary-dimension}.  If
there exist $\beta\in[0,1]$, $c>0$, and an infinite sequence $m_j\to\infty$
such that
\[
\varepsilon_{m_j}(\mathsf{P};\mu)\ge c\lambda_{\mathrm{cyl}}^{-(1-\beta)m_j}
\qquad\text{for all }j,
\]
then
\[
N_{m_j}(\partial\mathsf{P})\ge
K^{-1}c\,\lambda_{\mathrm{cyl}}^{\beta m_j}
\qquad\text{for all }j.
\]
In particular, any lower exponential rate for $\varepsilon_m$ forces
comparable lower growth of boundary cylinders along the same subsequence.
\end{proposition}

\begin{proof}
Proposition~\ref{prop:boundary-dimension} gives
\[
c\lambda_{\mathrm{cyl}}^{-(1-\beta)m_j}
\le \varepsilon_{m_j}(\mathsf{P};\mu)
\le K\lambda_{\mathrm{cyl}}^{-m_j}N_{m_j}(\partial\mathsf{P}),
\]
and rearranging yields the claim.
\end{proof}

\subsection{Primitive absorbing prefix filters}

For this subsection we keep the symbolic specialization
$X=\Sigma_A$ of a mixing shift of finite type with alphabet $\mathcal{A}$ and
adjacency matrix $A$, endowed with its Parry (maximal-entropy) measure.  The
class we isolate is intrinsic in the sense of the hidden-Markov and sofic
literature \cite{BoylePetersenHMP,PollicottKempton2016,Verbitskiy2011} and
matches the probabilistic automata viewpoint of
\cite{Rabin1963,Paz1971}.  It is exactly the first-hit deterministic transducer
class familiar from probabilistic functions of Markov chains
\cite{Blackwell1957,Heller1965,BaumPetrie1966}, specialized to the prefix
filtration used throughout.  A word $w=a_0\cdots a_{m-1}$ corresponds to a path
$v_0\to v_1\to\cdots\to v_m$ in the defining graph for $\Sigma_A$, and the
Parry measure assigns uniformly comparable mass to every admissible cylinder
$[w]$ \cite{Parry1964}.

Throughout we use the vertex model, so $\delta(r,w)$ abbreviates the iterated
application $\delta(\delta(\cdots\delta(r,a_0),\ldots),a_{m-1})$.  The pair
$(v_m,\delta(r_\ast,w))$ is the terminal state of the product graph obtained by
running the automaton $\mathcal{M}$ along the admissible path determined by
$w$.

\begin{definition}[Primitive absorbing prefix filter]
\label{def:hidden-markov-filter}
Let $\Sigma_A$ be mixing and let
$\mathcal{M}=(R,\delta,r_\ast,R_+,R_-)$ be a deterministic partial automaton on
$\mathcal{A}$.  The decision states $R_+$ and $R_-$ are disjoint absorbing
classes and $\delta(r,a)$ is defined whenever $r\notin R_+\cup R_-$ and
$a\in\mathcal{A}$ labels an admissible edge of $\Sigma_A$.  A finite word
$w=a_0\cdots a_{m-1}$ declares ``$+$'' if $\delta(r_\ast,w)\in R_+$ and every
proper prefix stays outside $R_+\cup R_-$; analogously for ``$-$''.  The event
\[
\mathsf{P}:=\bigcup_{\substack{w\in\mathcal{A}^\ast\\
    \delta(r_\ast,w)\in R_+\\
    \delta(r_\ast,w')\notin R_+\cup R_- \text{ for }w'\prec w}}[w]
\]
is therefore decided the first time the automaton hits $R_+$ (before $R_-$),
so $\mathsf{P}$ is an open prefix event as required by
Definition~\ref{def:boundary-cylinder}.

Write $V(A)$ for the state set of $A$.  The ambiguous product states are
\[
\mathcal{H}:=\{(v,r)\in V(A)\times (R\setminus(R_+\cup R_-)):
 \text{$(v,r)$ is reachable without decisions and both $R_\pm$ remain reachable}\}.
\]
The induced transition matrix $B_\partial$ on $\mathcal{H}$ is primitive when the
strictly ambiguous component is strongly connected.  Here
$B_\partial\bigl((v,r),(v',r')\bigr)=1$ whenever there exists
$a\in\mathcal{A}$ with $v\to v'$ in $A$ and $\delta(r,a)=r'\notin R_+\cup R_-$;
otherwise the entry is $0$.  Let $\mathbf{u}$ be the row vector indexed by
$\mathcal{H}$ whose $(v,r)$ component equals $1$ when the empty word keeps the
product state $(v,r_\ast)$ ambiguous and admits a first symbol taking it to
$(v,r)$, and equals $0$ otherwise.  We call $\mathsf{P}$ a
\emph{primitive absorbing prefix filter} if $\mathcal{H}\ne\varnothing$ and
$B_\partial$ is primitive.
\end{definition}

\paragraph{Absorption hypothesis.}
Throughout this subsection we impose a one-sided absorption assumption:
every reachable $(v,r)\in V(A)\times (R\setminus(R_+\cup R_-))$ hits
$R_+\cup R_-$ almost surely under the Parry measure, and whenever
$(v,r)\notin\mathcal{H}$ exactly one of $R_+$ or $R_-$ remains reachable
without encountering the other.  This prevents nondecision traps and makes
the boundary condition equivalent to membership in $\mathcal{H}$.

\begin{remark}
We deliberately adopt the absorbing prefix-filter viewpoint here: an event is
declared as soon as $\mathcal{M}$ reaches $R_+$ or $R_-$.  This is the
``narrow path'' singled out by the referee and it matches the concrete models
used later, especially Examples~\ref{ex:first-hit-parity}
and~\ref{ex:competing-runs}.  Infinite-word safety properties such as the even
shift therefore sit outside Theorem~\ref{thm:hidden-markov-thickness}, keeping
the statement honest about its scope.
\end{remark}

The automaton $\mathcal{M}$ records the finite-state filter that observes the
SFT through a deterministic transducer; the states in $\mathcal{H}$ are exactly
those product states where both decisions remain possible.  Primitivity of
$B_\partial$ expresses that the ambiguous component is strongly connected and
aperiodic, matching the probabilistic-function viewpoint of
\cite{Blackwell1957,Heller1965,BaumPetrie1966,BoylePetersenHMP,Verbitskiy2011}.

\begin{lemma}[Uniform decision windows]
\label{lem:uniform-windows}
Let $\mathsf{P}$ be a primitive absorbing prefix filter on $\Sigma_A$ with Parry
measure $\mu$.  There exist $L\ge 1$ and $\delta\in(0,\frac12]$ such that for
each $(v,r)\in\mathcal{H}$ one can choose admissible words
$u_{(v,r)}^+$ and $u_{(v,r)}^-$ of lengths at most $L$ satisfying:
\begin{enumerate}[label=(\alph*)]
\item $\delta(r,u_{(v,r)}^+)\in R_+$ and $\delta(r,u_{(v,r)}^-)\in R_-$, and no
  proper prefix of $u_{(v,r)}^\pm$ hits $R_+\cup R_-$;
\item for every boundary cylinder $[w]$ whose terminal product state is
  $(v,r)$, the conditional probabilities satisfy
  $\mu([wu_{(v,r)}^+]\mid[w])\ge\delta$ and
  $\mu([wu_{(v,r)}^-]\mid[w])\ge\delta$.
\end{enumerate}
\end{lemma}

\begin{proof}
Because $(v,r)\in\mathcal{H}$, there exist words steering $r$ to both $R_+$
and $R_-$ without intermediate decisions.  Taking the longest such witness
across the finite set $\mathcal{H}$ yields an upper bound $L_0$ on the word
lengths.  Primitivity of $B_\partial$ supplies an integer $L_1$ such that from
any ambiguous state there is a word of length at most $L_1$ staying inside
$\mathcal{H}$ and ending at a prescribed ambiguous state.  Concatenating these
finite witnesses shows that from each $(v,r)$ there exist ``yes'' and ``no''
words of length at most $L:=L_0+L_1$ satisfying (a).  The Parry measure obeys
the uniform same-length bounds of Proposition~\ref{prop:volume-boundary-scaling},
so any admissible length-$\ell$ cylinder has mass between
$c_- \lambda_{\mathrm{cyl}}^{-\ell}$ and $c_+ \lambda_{\mathrm{cyl}}^{-\ell}$.
For $[w]$ ending at $(v,r)$,
\[
\frac{\mu([wu_{(v,r)}^\pm])}{\mu([w])}
\ge \frac{c_- \lambda_{\mathrm{cyl}}^{-|u_{(v,r)}^\pm|}}
           {c_+}
\ge \frac{c_-}{c_+}\lambda_{\mathrm{cyl}}^{-L}.
\]
Taking $\delta:=\frac12\frac{c_-}{c_+}\lambda_{\mathrm{cyl}}^{-L}$ establishes
(b).
\end{proof}

By (a), the subcylinders $[wu_{(v,r)}^+]$ and $[wu_{(v,r)}^-]$ lie entirely in
$\mathsf{P}$ and $\mathsf{P}^c$, respectively, so (b) also yields
$\mu(\mathsf{P}\mid[w])\ge\delta$ and $\mu(\mathsf{P}^c\mid[w])\ge\delta$ for
every boundary cylinder $[w]$.

\begin{theorem}[Primitive absorbing prefix filters give automatic thickness and spectral counts]
\label{thm:hidden-markov-thickness}
Let $\mathsf{P}$ be a primitive absorbing prefix filter on a mixing
$\Sigma_A$ with Parry measure $\mu$ that satisfies the absorption hypothesis,
and let $B_\partial$ be the matrix from
Definition~\ref{def:hidden-markov-filter}.  Then:
\begin{enumerate}[label=(\arabic*)]
\item For each finite word $w$, the conditional mass satisfies
  $0<\mu(\mathsf{P}\mid[w])<1$ if and only if the terminal product state
  $(v_m,\delta(r_\ast,w))$ belongs to $\mathcal{H}$.  Equivalently,
  $[w]\in\partial_m\mathsf{P}$ iff that terminal state is ambiguous.
\item There exists $\theta\in(0,\frac12]$, depending only on $A$ and the
  automaton, such that every boundary cylinder $C$ satisfies
  $\theta\le\mu(\mathsf{P}\mid C)\le 1-\theta$.  Hence the uniform thickness
  hypothesis of Proposition~\ref{prop:volume-boundary-scaling} holds automatically.
\item For every $m\ge 1$,
  $N_m(\partial\mathsf{P})=\mathbf{u}^\top B_\partial^{m-1} \mathbf{1}$, where
  $\mathbf{u}$ is the accessibility vector from
  Definition~\ref{def:hidden-markov-filter}.  Consequently
  $N_m(\partial\mathsf{P})=\rho_\partial^{\,m-1}(\kappa+o(1))$ with
  $\rho_\partial:=\rho(B_\partial)$ the Perron eigenvalue and some constant
  $\kappa>0$.  The filtration boundary exponent is therefore
  $\overline{\beta}_{\mathcal{F}}(\mathsf{P})=\log_{\lambda_{\mathrm{cyl}}}\rho_\partial$,
  and the scan error obeys the two-sided bounds
  \[
  c_1\lambda_{\mathrm{cyl}}^{-m}\rho_\partial^{\,m-1}
  \le \varepsilon_m(\mathsf{P};\mu)
  \le c_2\lambda_{\mathrm{cyl}}^{-m}\rho_\partial^{\,m-1}
  \qquad\text{for all }m\ge 1
  \]
  for suitable constants $c_1,c_2>0$.
\end{enumerate}
\end{theorem}

\begin{proof}
Every word determines an admissible path in the product graph.  The conditional
probability $\mu(\mathsf{P}\mid[w])$ depends only on the terminal state
$(v_m,\delta(r_\ast,w))$.  If that state lies outside $\mathcal{H}$ then the
absorption hypothesis guarantees that exactly one decision state remains
reachable and is hit almost surely, hence the conditional probability is $0$
or $1$.  Conversely, if $(v_m,\delta(r_\ast,w))\in\mathcal{H}$, both
continuations occur with positive conditional probability by
Lemma~\ref{lem:uniform-windows}, so $(1)$ follows.

For part (2), Lemma~\ref{lem:uniform-windows} yields a uniform $\delta>0$ such
that both $\mu(\mathsf{P}\mid[w])$ and $\mu(\mathsf{P}^c\mid[w])$ exceed
$\delta$ on every boundary cylinder $[w]$.  Taking $\theta:=\delta$ provides
the claimed thickness constant because the Parry measure grants uniform
comparability of cylinders.

For (3), a length-$m$ word contributes to $N_m(\partial\mathsf{P})$ precisely
when its path stays inside $\mathcal{H}$ for the first $m-1$ steps after the
initial symbol.  The number of such paths equals the matrix coefficient
$\mathbf{u}^\top B_\partial^{\,m-1} \mathbf{1}$ by the same automaton enumerations as
in \cite{Rabin1963,BoylePetersenHMP}.  The Perron--Frobenius theorem yields
$N_m(\partial\mathsf{P})=\rho_\partial^{\,m-1}(\kappa+o(1))$ with $\kappa>0$ because
$B_\partial$ is primitive.  Combining the automatic thickness from (2) with the
Parry cylinder bounds allows
Proposition~\ref{prop:volume-boundary-scaling} to be applied, giving the
two-sided bounds for $\varepsilon_m(\mathsf{P};\mu)$ and the value of the
boundary exponent.
\end{proof}

\paragraph{Positioning relative to classical results.}
Items (1) and (3) rest on the standard probabilistic-function analysis of
hidden-Markov filters via the product automaton and Perron--Frobenius theory
\cite{Blackwell1957,Heller1965,BaumPetrie1966,BoylePetersenHMP,Verbitskiy2011}.
The novelty here is that these intrinsic statements are welded to the
fixed-filtration boundary law from Section~\ref{sec:rates}: automatic thickness
and the spectral count now feed directly into the oracle scan-error decay
through Proposition~\ref{prop:volume-boundary-scaling}.

\subsection{Auxiliary polynomial-distortion variant}
\label{sub:nonuniform}

The uniform bounds $\mu(C)\asymp\lambda_{\mathrm{cyl}}^{-m}$ hold for the
Parry measure (measure of maximal entropy) on a mixing SFT
\cite{Parry1964}.  By contrast, a general Gibbs measure has
potential-dependent cylinder masses of the form
$e^{S_m\phi(x)-mP(\phi)}$ rather than a same-length bound that depends only on
$m$ \cite{Bowen1975,WaltersErgodicTheory}.  For that reason the next statement
is recorded only as an auxiliary polynomial-distortion variant; it is not a
genuine Gibbs theorem.

\begin{definition}[Weak-Gibbs distortion]
\label{def:log-regular}
A measure $\mu$ has \emph{weak-Gibbs (quasi-Bernoulli) distortion with exponent $\alpha>0$} if there exist
$0<c_1\le c_2<\infty$ such that for all $m\ge 1$ and all $C\in\mathcal{C}_m$
with $\mu(C)>0$,
\[
c_1 m^{-\alpha}\lambda_{\mathrm{cyl}}^{-m}\le\mu(C)\le c_2 m^{\alpha}\lambda_{\mathrm{cyl}}^{-m}.
\]
\end{definition}

This polynomial distortion bound is standard for weak Gibbs measures and their
factors; see \cite{Bowen1975,WaltersErgodicTheory}.

\begin{proposition}[Weak-Gibbs polynomial-distortion variant]
\label{prop:nonuniform-cylinder}
Let $\mu$ have weak-Gibbs distortion with exponent $\alpha>0$.  Assume the thickness
condition of Proposition~\ref{prop:volume-boundary-scaling} holds with some
$\theta>0$ and $m_0$.  Then for every $m\ge m_0$,
\[
\theta c_1 m^{-\alpha}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})
\le\varepsilon_m(\mathsf{P};\mu)
\le \frac{c_2}{2}m^{\alpha}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P}).
\]
If $N_m(\partial\mathsf{P})=\lambda_{\mathrm{cyl}}^{\beta m+o(m)}$, then
\[
\varepsilon_m(\mathsf{P};\mu)=\lambda_{\mathrm{cyl}}^{-(1-\beta)m+o(m)},
\]
equivalently,
\[
\lim_{m\to\infty}\left(-\frac{1}{m}\log_{\lambda_{\mathrm{cyl}}}\varepsilon_m(\mathsf{P};\mu)\right)=1-\beta.
\]
In particular, polynomial distortion in cylinder masses changes only the
subexponential factor, not the exponential rate.
\end{proposition}

\begin{proof}
The same decomposition argument as in Proposition~\ref{prop:volume-boundary-scaling}
applies with $c_1 m^{-\alpha}$ and $c_2 m^\alpha$ in place of $c_-$ and
$c_+/2$.  Hence for every $m\ge m_0$,
\[
\theta c_1 m^{-\alpha}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})
\le\varepsilon_m(\mathsf{P};\mu)
\le \frac{c_2}{2}m^{\alpha}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P}).
\]
If $N_m(\partial\mathsf{P})=\lambda_{\mathrm{cyl}}^{\beta m+o(m)}$, substitution
gives
\[
\theta c_1 m^{-\alpha}\lambda_{\mathrm{cyl}}^{-(1-\beta)m+o(m)}
\le\varepsilon_m(\mathsf{P};\mu)
\le \frac{c_2}{2}m^{\alpha}\lambda_{\mathrm{cyl}}^{-(1-\beta)m+o(m)}.
\]
Since $\log_{\lambda_{\mathrm{cyl}}} m=o(m)$, the polynomial factors are
absorbed into the $o(m)$ term, yielding
$\varepsilon_m(\mathsf{P};\mu)=\lambda_{\mathrm{cyl}}^{-(1-\beta)m+o(m)}$ and the
equivalent logarithmic limit.
\end{proof}

\begin{table}[ht]
\centering\small
\begin{tabular}{p{2.8in}p{3.1in}}
\toprule
assumption & conclusion for $\varepsilon_m(\mathsf{P};\mu)$ \\
\midrule
$\mathcal{W}_m(\mathsf{P};\mu)$ and averaged ambiguity $\overline{\theta}_m$ &
$\varepsilon_m=\mathcal{W}_m(\mathsf{P};\mu)\,\overline{\theta}_m(\mathsf{P};\mu)$ \\
$\mu(C)\le K\lambda_{\mathrm{cyl}}^{-m}$ on all $m$-cylinders &
$\varepsilon_m\le K\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})$ \\
uniform $c_\pm\lambda_{\mathrm{cyl}}^{-m}$ bounds plus thickness $\theta$ &
$\theta c_-\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})\le\varepsilon_m
\le \frac{c_+}{2}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})$;
if $N_m=\lambda_{\mathrm{cyl}}^{\beta m+o(m)}$, then
$\varepsilon_m=\lambda_{\mathrm{cyl}}^{-(1-\beta)m+o(m)}$ \\
weak-Gibbs distortion ($c_i m^{\pm\alpha}\lambda_{\mathrm{cyl}}^{-m}$) plus thickness $\theta$ (auxiliary) &
$\theta c_1 m^{-\alpha}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})\le\varepsilon_m\le \frac{c_2}{2}m^{\alpha}\lambda_{\mathrm{cyl}}^{-m}N_m(\partial\mathsf{P})$;
if $N_m=\lambda_{\mathrm{cyl}}^{\beta m+o(m)}$, then $\varepsilon_m=\lambda_{\mathrm{cyl}}^{-(1-\beta)m+o(m)}$ \\
\bottomrule
\end{tabular}
\caption{Summary of the exact weighted law, the main count-based corollaries,
and the auxiliary polynomial-distortion variant.}
\label{tab:rate-laws}
\end{table}

\section{Examples and Failure Modes}
\label{sec:examples}

The measure-theoretic boundary from Definition~\ref{def:boundary-cylinder} will
be used verbatim: a cylinder $C$ lies in $\partial_m\mathsf{P}$ if and only if
$\mu(\mathsf{P}\cap C)>0$ and $\mu(C\setminus\mathsf{P})>0$.  Events of measure
$0$ or $1$ therefore contribute no boundary cylinders, and every example below
has been checked under this convention.

\begin{example}[Finite-horizon events]
\label{ex:finite-horizon-event}
If $\mathsf{P}\in\mathcal{F}_r$ for some finite $r$, then membership is decided
once the first $r$ symbols are read, so $\partial_m\mathsf{P}=\varnothing$ and
$\varepsilon_m(\mathsf{P};\mu)=0$ for all $m\ge r$.
\end{example}

\begin{example}[A natural first-hit parity event]
\label{ex:first-hit-parity}
Let $X=\Omega_\infty=\{0,1\}^{\mathbb{N}}$ with the fair Bernoulli measure and
define
\[
\tau(\xi):=\inf\{t\ge 0:\xi_t=1\},
\qquad
\mathsf{P}_{\mathrm{even}}:=\{\tau<\infty\text{ and }\tau\text{ is even}\}.
\]
The null sequence $0^\infty$ is ignored by $\tau<\infty$, and the automaton from
Definition~\ref{def:hidden-markov-filter} simply tracks the parity of the
current zero run until the first $1$ appears.  At depth $m$ the only unresolved
cylinder is $C_m:=[0^m]$, so $N_m(\partial\mathsf{P}_{\mathrm{even}})=1$ and
$\overline{\beta}_{\mathcal{F}}(\mathsf{P}_{\mathrm{even}})=0$.  A geometric
calculation shows that
$\mu(\mathsf{P}_{\mathrm{even}}\mid[0^m])=\frac23$ for even $m$ and
$\frac13$ for odd $m$, whence every boundary cylinder has local ambiguity
$\frac13$ and the thickness constant is automatic.  Because
$\mu([0^m])=2^{-m}$, the exact cylinder decomposition yields
$\varepsilon_m(\mathsf{P}_{\mathrm{even}};\mu)=\frac13\,2^{-m}$, so this example
verifies Proposition~\ref{prop:volume-boundary-scaling} with a minimal
primitive absorbing prefix filter.
\end{example}

\begin{example}[Competing runs trigger spectral boundary growth]
\label{ex:competing-runs}
Let $X=\Omega_\infty$ with the fair Bernoulli measure.
Stop scanning once either the pattern $11$ or the pattern $000$ has appeared,
and declare $\mathsf{P}_{\mathrm{run}}$ if $11$ occurs first.  The elementary
Markov calculation
\[
p_H=\tfrac12+\tfrac12p_Z,\quad
p_Z=\tfrac12p_{ZZ}+\tfrac12p_H,\quad
p_{ZZ}=\tfrac12p_H
\]
shows that the ``first-run'' event has positive measure
$\mu(\mathsf{P}_{\mathrm{run}})=\tfrac12(p_H+p_Z)=0.7$.

The absorbing automaton has decision states $R_+=\{\text{``11''}\}$ and
$R_-=\{\text{``000''}\}$ together with the ambiguous states
``$H$'' (most recent symbol $1$), ``$Z$'' (a single trailing $0$), and
``$ZZ$'' (two trailing zeros).  The transitions are
\[
S\stackrel{1}{\longrightarrow}H,\quad
S\stackrel{0}{\longrightarrow}Z,\quad
H\stackrel{0}{\longrightarrow}Z,\quad
Z\stackrel{1}{\longrightarrow}H,\quad
Z\stackrel{0}{\longrightarrow}ZZ,\quad
ZZ\stackrel{1}{\longrightarrow}H,
\]
while $H\stackrel{1}{\longrightarrow}R_+$ and $ZZ\stackrel{0}{\longrightarrow}R_-$.
Thus the event fits Definition~\ref{def:hidden-markov-filter} with a primitive
ambiguous component.  Ordering the states $(H,Z,ZZ)$ gives
\[
B_\partial
=\begin{pmatrix}
0&1&0\\[0.5ex]
1&0&1\\[0.5ex]
1&0&0
\end{pmatrix},
\qquad
\mathbf{u}=(1,1,0)^\top,
\]
because the first symbol places the run either in $H$ or $Z$.  The characteristic
polynomial of $B_\partial$ is $x^3-x-1$, so the spectral radius is the plastic
constant $\rho_\partial\approx 1.3247$.  Consequently
\[
N_m(\partial\mathsf{P}_{\mathrm{run}})
=\mathbf{u}^\top B_\partial^{\,m-1}\mathbf{1}
=\rho_\partial^{\,m-1}(\kappa+o(1)),
\]
and the filtration boundary exponent is $\log_2\rho_\partial$.  By
Lemma~\ref{lem:uniform-windows} the thickness constant is uniform, so
Proposition~\ref{prop:volume-boundary-scaling} yields
\[
c_1\,2^{-m}\rho_\partial^{\,m-1}
\le \varepsilon_m(\mathsf{P}_{\mathrm{run}};\mu)
\le c_2\,2^{-m}\rho_\partial^{\,m-1}.
\]
The corrected counts begin with
\[
\begin{array}{c|ccccc}
m & 1 & 2 & 3 & 4 & 5\\ \hline
N_m(\partial\mathsf{P}_{\mathrm{run}}) & 2 & 3 & 4 & 5 & 7
\end{array}
\]
showing the $m\mapsto m-1$ shift explicitly.  This example keeps all
hypotheses of Theorem~\ref{thm:hidden-markov-thickness} while producing a
genuinely spectral boundary law with positive measure.
\end{example}

\begin{example}[Full two-shift: an explicit positive-measure event with filtration boundary exponent $1/2$]
\label{ex:self-similar-bernoulli}
Let $X=\Omega_\infty=\{0,1\}^{\mathbb{N}}$ with the fair Bernoulli measure
$\mu=(\tfrac12,\tfrac12)^{\otimes\mathbb{N}}$, which is the Parry measure for
the full two-shift.  Then $\lambda_{\mathrm{cyl}}=2$ and every length-$m$
cylinder has mass exactly $2^{-m}$.

\paragraph{Recursive construction of $\mathsf{E}$.}
Set $B_0:=\{\varnothing\}$.  For each $k\ge 0$ and each $w\in B_k$, split the
four grandchildren of $[w]$ at depth $2k+2$ into
\[
B_{k+1}:=\{w00,w11:w\in B_k\},\qquad
I_{k+1}:=\{w01:w\in B_k\},\qquad
O_{k+1}:=\{w10:w\in B_k\}.
\]
Thus every unresolved cylinder produces one cylinder declared
``inside'' ($01$), one declared ``outside'' ($10$), and two cylinders
($00$ and $11$) on which the same construction repeats.  Define
\[
K:=\bigcap_{k\ge 0}\bigcup_{w\in B_k}[w],
\qquad
\mathsf{E}:=K\cup\bigcup_{k\ge 1}\bigcup_{u\in I_k}[u].
\]
The families $\{[u]:u\in I_k,\ k\ge 1\}$, $\{[v]:v\in O_k,\ k\ge 1\}$, and $K$
are pairwise disjoint and partition $X$, so $\mathsf{E}\in\mathcal{F}_\infty$.
Conditioning on the first two symbols gives
\[
\mu(\mathsf{E})
=\mu([01])+\mu([00])\mu(\mathsf{E}\mid[00])+\mu([11])\mu(\mathsf{E}\mid[11])
=\frac14+\frac14\mu(\mathsf{E})+\frac14\mu(\mathsf{E}),
\]
because the conditional law on $[00]$ or $[11]$ is a scaled copy of the whole
construction.  Hence $\mu(\mathsf{E})=\frac12$, so the example has positive
measure.

\paragraph{Boundary count.}
At even depth $2k$, the boundary cylinders are exactly $\{[w]:w\in B_k\}$:
any cylinder declared inside or outside at an earlier stage is already
decided, while each $[w]$ with $w\in B_k$ contains both an inside descendant
$[w01]$ and an outside descendant $[w10]$.  Since $\#B_k=2^k$,
\[
N_{2k}(\partial\mathsf{E})=2^k.
\]
At odd depth $2k+1$, the boundary cylinders are exactly
$\{[w0],[w1]:w\in B_k\}$, so
\[
N_{2k+1}(\partial\mathsf{E})=2^{k+1}.
\]
Therefore
\[
\overline{\beta}_{\mathcal{F}}(\mathsf{E})
=\limsup_{m\to\infty}\frac{\log N_m(\partial\mathsf{E})}{m\log 2}
=\frac12.
\]

\paragraph{Thickness verification.}
For $w\in B_k$, let $p_w:=\mu(\mathsf{E}\mid[w])$.  The same self-similarity
as above gives
\[
p_w=\frac14+\frac12 p_w,
\]
hence $p_w=\frac12$.  Thus every even-depth boundary cylinder carries exactly
half of $\mathsf{E}$ and half of $\mathsf{E}^c$.  For odd-depth boundary
cylinders,
\[
\mu(\mathsf{E}\mid[w0])
=\frac12\mu(\mathsf{E}\mid[w00])+\frac12\mu(\mathsf{E}\mid[w01])
=\frac12\cdot\frac12+\frac12\cdot 1
=\frac34,
\]
while
\[
\mu(\mathsf{E}\mid[w1])
=\frac12\mu(\mathsf{E}\mid[w10])+\frac12\mu(\mathsf{E}\mid[w11])
=\frac12\cdot 0+\frac12\cdot\frac12
=\frac14.
\]
Hence every boundary cylinder satisfies
\[
\min\{\mu(\mathsf{E}\mid C),1-\mu(\mathsf{E}\mid C)\}\ge\frac14,
\]
so the thickness condition holds with $\theta=\frac14$.

\paragraph{Rate conclusion.}
Proposition~\ref{prop:volume-boundary-scaling} applies with
$\lambda_{\mathrm{cyl}}=2$, $c_-=c_+=1$, and $\theta=\frac14$, yielding
$\varepsilon_m(\mathsf{E};\mu)\asymp 2^{-m/2}$.  In fact the exact cylinder
decomposition gives
\[
\varepsilon_{2k}(\mathsf{E};\mu)=2^k\cdot 2^{-2k}\cdot\frac12=2^{-k-1},
\qquad
\varepsilon_{2k+1}(\mathsf{E};\mu)=2^{k+1}\cdot 2^{-(2k+1)}\cdot\frac14=2^{-k-2},
\]
so $\varepsilon_m(\mathsf{E};\mu)\asymp 2^{-m/2}$ with explicit constants.
\end{example}

\begin{remark}[Subsequence oscillation of boundary counts]
\label{rem:oscillation}
Example~\ref{ex:self-similar-bernoulli} already shows that exact boundary
counts can oscillate by subsequence:
\[
N_{2k}(\partial\mathsf{E})=2^k,
\qquad
N_{2k+1}(\partial\mathsf{E})=2^{k+1}=2\,N_{2k}(\partial\mathsf{E}).
\]
The filtration boundary exponent remains $\frac12$, but no single asymptotic
constant governs all depths.  This is why the rate statements in
Section~\ref{sec:rates} are phrased with $\limsup$ or $o(m)$ rather than with
full asymptotic limits.
\end{remark}

\begin{example}[Boundary growth without thickness]
\label{ex:thickness-failure}
Let $X=\Omega_\infty$ with the fair Bernoulli measure and consider
\[
\mathsf{U}:=\bigcup_{n\ge 1}[0^n1^n].
\]
At depth $m$, a cylinder is mixed exactly when it belongs to one of the two
families
\[
C_m^{(0)}:=[0^m]
\qquad\text{or}\qquad
C_{n,r}:=[0^n1^r]\text{ with }n+r=m,\ 1\le r<n.
\]
Such prefixes can still be completed to $0^n1^n$, but they are not yet
decided.  Hence
\[
N_m(\partial\mathsf{U})=\#\{n:\lfloor m/2\rfloor+1\le n\le m\}
=\left\lceil\frac{m}{2}\right\rceil
=2^{o(m)},
\]
so the filtration boundary exponent is $0$.

For the pure zero cylinder $C_m^{(0)}$ there are infinitely many admissible
continuations inside $\mathsf{U}$, one for each $n\ge m$, whence
\[
\mu(\mathsf{U}\mid C_m^{(0)})=\sum_{k\ge 0}2^{-(m+2k)}
=\frac{4}{3}\,2^{-m},
\qquad
\mu(C_m^{(0)})=2^{-m}.
\]
For $C_{n,r}$ with $1\le r<n$, the only continuation inside $\mathsf{U}$ is to
finish the remaining block of $n-r$ ones, so
\[
\mu(\mathsf{U}\mid C_{n,r})=2^{-(n-r)}.
\]
Thus the local ambiguities on boundary cylinders go to $0$, and the uniform
thickness hypothesis fails.  The exact decomposition now reads
\[
\varepsilon_m(\mathsf{U};\mu)
=2^{-m}\min\!\left\{\frac{4}{3}\,2^{-m},1-\frac{4}{3}\,2^{-m}\right\}
+\sum_{n=\lfloor m/2\rfloor+1}^{m}2^{-m}\,2^{-(2n-m)}.
\]
The first term is $O(2^{-2m})$, whereas the sum equals
$\sum_{n=\lfloor m/2\rfloor+1}^{m}4^{-n}$ and hence remains $\asymp 2^{-m}$.
By contrast, the lower bound in
Proposition~\ref{prop:volume-boundary-scaling} would have order
$2^{-m}N_m(\partial\mathsf{U})\asymp m2^{-m}$ if thickness were available.
The missing factor $m$ is exactly the cost of the collapsing local ambiguity.
\end{example}

\begin{example}[Dependence on the chosen filtration]
\label{ex:filtration-dependence}
Let $X=\Omega_\infty$ with the fair Bernoulli measure and consider the event
\[
\mathsf{P}:=\{\xi_1=1\}.
\]
With the usual prefix filtration
$\mathcal{F}_m=\sigma(\xi_0,\ldots,\xi_{m-1})$, the event is decided once the
second coordinate is read, so $\varepsilon_m(\mathsf{P};\mu)=0$ for all
$m\ge 2$ and $\overline{\beta}_{\mathcal{F}}(\mathsf{P})=0$.

Now replace this by the even-coordinate filtration
\[
\mathcal{G}_m:=\sigma(\xi_0,\xi_2,\ldots,\xi_{2m-2}).
\]
The event $\mathsf{P}$ is independent of every $\mathcal{G}_m$, so
$\varepsilon_m(\mathsf{P};\mu)=\frac12$ for all $m$.  Every positive-mass atom
of $\mathcal{G}_m$ is therefore boundary, and there are exactly $2^m$ such
atoms.  Thus
\[
N_m^{(\mathcal{G})}(\partial\mathsf{P})=2^m,
\qquad
\overline{\beta}_{\mathcal{G}}(\mathsf{P})=1.
\]
The same measurable event therefore has trivial boundary growth for one
filtration and maximal boundary growth for another.  The filtration boundary
exponent is a property of the representation, not of the set alone.
\end{example}

\section{Discussion and Limitations}
\label{sec:discussion}

The fixed-filtration picture established here is intentionally modest.  The
scan error profile admits an exact partition decomposition
(Proposition~\ref{prop:scan-error-cylinder}), a Bayes-optimal threshold rule
(Proposition~\ref{prop:bayes-optimality}), and a discrete Tanaka
representation (Proposition~\ref{prop:scan-tanaka}).  The new organizational
point of Section~\ref{sec:rates} is that the decisive quantity is the weighted
boundary mass $\mathcal{W}_m(\mathsf{P};\mu)$: count-based rates are corollaries once
one compares $\mathcal{W}_m$ to same-length cylinder masses, while
Proposition~\ref{prop:boundary-growth-necessary} shows that lower error bounds
force corresponding lower growth of mixed cylinders.  The primitive
hidden-Markov class of Theorem~\ref{thm:hidden-markov-thickness} provides one
setting in which thickness is verified intrinsically.
Examples~\ref{ex:first-hit-parity}
and~\ref{ex:self-similar-bernoulli} illustrate this mechanism, while
Examples~\ref{ex:thickness-failure} and~\ref{ex:filtration-dependence} show
that thickness and filtration choice are genuinely structural assumptions, not
merely cosmetic ones.

\paragraph{Shannon--McMillan--Breiman and uniform bounds.}
The SMB theorem \cite{Shannon1948,McMillan1953,Breiman1957} states that for
$\mu$-a.e.\ $x$,
\[
-\frac{1}{m}\log\mu([x]_m)\to h(\mu)
\quad\text{as }m\to\infty,
\]
where $[x]_m$ is the length-$m$ cylinder containing $x$.  This is a pointwise
almost-sure statement about a single cylinder along a single orbit; it does
\emph{not} give a uniform bound over all admissible cylinders simultaneously.
The uniform bounds $\mu(C)\asymp\lambda_{\mathrm{cyl}}^{-m}$ in
Propositions~\ref{prop:boundary-dimension} and~\ref{prop:volume-boundary-scaling}
are strictly stronger.  For mixing SFTs with the Parry measure, these
uniform bounds are established by Parry \cite{Parry1964}; for general Gibbs
measures Bowen's theory gives potential-dependent cylinder weights rather
than the uniform same-length estimates used here \cite{Bowen1975}.  Thus the
main count theorems apply directly to the Parry / maximal-entropy case, while
Proposition~\ref{prop:nonuniform-cylinder} is only a polynomial-distortion
variant.  A genuine Gibbs theorem would require a weighted boundary partition
function that keeps track of the potential rather than of simple cylinder
counts.

\paragraph{Heuristic relation to margin conditions.}
The comparison with \cite{MammenTsybakov1999,Tsybakov2004} is only heuristic.
Thickness is a uniform lower bound on the two class masses inside each mixed
cylinder, whereas the Mammen--Tsybakov condition controls the global
distribution of $p_\infty$ near $\frac12$.  The former concerns oracle
approximation on a fixed filtration; the latter concerns estimation from data.
Scott and Nowak's dyadic boundary/noise framework
\cite{ScottNowak2006} is therefore the closer statistical-learning analogue.

\paragraph{What the counterexamples show.}
Example~\ref{ex:thickness-failure} keeps the boundary count at the
subexponential scale $2^{o(m)}$ but lets the local ambiguity on each mixed
cylinder collapse.  The lower bound from
Proposition~\ref{prop:volume-boundary-scaling} then fails by an unbounded factor.
Example~\ref{ex:filtration-dependence} shows that the filtration boundary
exponent $\overline{\beta}_{\mathcal{F}}(\mathsf{P})$ can change from $0$ to
$1$ for the same measurable event when the observed coordinate sequence is
changed.  Together with the parity oscillation in
Remark~\ref{rem:oscillation}, these examples explain why the paper works with
weighted boundary mass, thickness hypotheses, and limsup exponents rather than
claiming an intrinsic geometric boundary theory.

\paragraph{Extensions.}
Natural generalizations include: finite alphabets of size $q$ (the discrete
Tanaka structure extends with $q-1$ threshold parameters); the continuous-time
limit to Tanaka--Meyer formula \cite{RevuzYor1999}; weakening the thickness
hypothesis to an $L^p$ condition on boundary-cylinder probabilities; and
replacing the simple count $N_m(\partial\mathsf{P})$ by a genuinely
potential-weighted partition function in the Gibbs setting.  It is also
natural to ask when the filtration boundary exponent
$\overline{\beta}_{\mathcal{F}}(\mathsf{P})$ can be expressed in dynamical
terms such as entropy, pressure, or automaton growth rates for larger symbolic
classes.

\paragraph{Limitations.}
The rate laws assume exponential cylinder complexity; subexponential systems
(Sturmian codings, substitution shifts) require separate analysis.  The proofs
do not use $T$-invariance: all results are purely filtration-theoretic, and the
dynamical framing motivates the cylinder structure but is not invoked in any
argument.  Beyond the hidden-Markov filter class, the paper does not derive
thickness or boundary growth from a broad intrinsic dynamical hypothesis.  A
substantially stronger paper would need such a derivation, or an equally
robust converse theory, for natural symbolic or Gibbs classes.

\begin{thebibliography}{99}

\bibitem{AudibertTsybakov2007}
Jean-Yves Audibert and Alexandre~B. Tsybakov.
\newblock Fast learning rates for plug-in classifiers.
\newblock {\em The Annals of Statistics}, 35(2):608--633, 2007.

\bibitem{BinevCohenDahmenDeVore2014}
Peter Binev, Albert Cohen, Wolfgang Dahmen, and Ronald DeVore.
\newblock Classification algorithms using adaptive partitioning.
\newblock {\em The Annals of Statistics}, 42(6):2141--2163, 2014.

\bibitem{Blackwell1951}
David Blackwell.
\newblock Comparison of experiments.
\newblock In Jerzy Neyman, editor, {\em Proceedings of the Second Berkeley
  Symposium on Mathematical Statistics and Probability}, pages 93--102.
  University of California Press, Berkeley, CA, 1951.

\bibitem{BlanchardSchaferRozenholcMuller2007}
Gilles Blanchard, Christian Sch\"afer, Yves Rozenholc, and Klaus-Robert
  M\"uller.
\newblock Optimal dyadic decision trees.
\newblock {\em Machine Learning}, 66(2--3):209--241, 2007.

\bibitem{Bowen1975}
Rufus Bowen.
\newblock {\em Equilibrium States and the Ergodic Theory of Anosov
  Diffeomorphisms}, volume 470 of {\em Lecture Notes in Mathematics}.
\newblock Springer, Berlin, 1975.

\bibitem{Boyle1983}
Mike Boyle.
\newblock Lower entropy factors of sofic systems.
\newblock {\em Ergodic Theory and Dynamical Systems}, 3(4):541--557, 1983.

\bibitem{Breiman1957}
Leo Breiman.
\newblock The individual ergodic theorem of information theory.
\newblock {\em The Annals of Mathematical Statistics}, 28(3):809--811, 1957.

\bibitem{BreimanFriedmanOlshenStone1984}
Leo Breiman, Jerome~H. Friedman, Richard~A. Olshen, and Charles~J. Stone.
\newblock {\em Classification and Regression Trees}.
\newblock Chapman and Hall, New York, 1984.

\bibitem{CohenDahmenDaubechiesDeVore2001}
Albert Cohen, Wolfgang Dahmen, Ingrid Daubechies, and Ronald DeVore.
\newblock Tree approximation and optimal encoding.
\newblock {\em Applied and Computational Harmonic Analysis}, 11(2):192--226,
  2001.

\bibitem{CoverThomas2006}
Thomas~M. Cover and Joy~A. Thomas.
\newblock {\em Elements of Information Theory}.
\newblock Wiley-Interscience, Hoboken, NJ, second edition, 2006.

\bibitem{DeVore1998}
Ronald~A. DeVore.
\newblock Nonlinear approximation.
\newblock {\em Acta Numerica}, 7:51--150, 1998.

\bibitem{DevroyeGyorfiLugosi1996}
Luc Devroye, L\'aszl\'o Gy\"orfi, and G\'abor Lugosi.
\newblock {\em A Probabilistic Theory of Pattern Recognition}, volume~31 of
  {\em Applications of Mathematics}.
\newblock Springer, New York, 1996.

\bibitem{Donoho1997}
David~L. Donoho.
\newblock {CART} and best-ortho-basis: a connection.
\newblock {\em The Annals of Statistics}, 25(5):1870--1911, 1997.

\bibitem{Doob1953}
J.~L. Doob.
\newblock {\em Stochastic Processes}.
\newblock Wiley, New York, 1953.

\bibitem{Durrett2019}
Rick Durrett.
\newblock {\em Probability: Theory and Examples}.
\newblock Cambridge Series in Statistical and Probabilistic Mathematics.
  Cambridge University Press, Cambridge, fifth edition, 2019.

\bibitem{Fischer1975}
Roland Fischer.
\newblock Sofic systems and graphs.
\newblock {\em Monatshefte f\"ur Mathematik}, 80(3):179--186, 1975.

\bibitem{Krieger1984}
Wolfgang Krieger.
\newblock On sofic systems {I}.
\newblock {\em Israel Journal of Mathematics}, 48(2--3):305--330, 1984.

\bibitem{LindMarcus1995}
Douglas Lind and Brian Marcus.
\newblock {\em An Introduction to Symbolic Dynamics and Coding}.
\newblock Cambridge University Press, Cambridge, 1995.

\bibitem{LochowskiOblojPromelSiopaes2021}
Rafa{\l}~M. {\L}ochowski, Jan Ob{\l}\'oj, David~J. Pr\"omel, and Pietro
  Siorpaes.
\newblock Local times and {T}anaka--{M}eyer formulae for c\`adl\`ag paths.
\newblock {\em Bernoulli}, 27(3):1597--1633, 2021.

\bibitem{MammenTsybakov1999}
Enno Mammen and Alexandre~B. Tsybakov.
\newblock Smooth discrimination analysis.
\newblock {\em The Annals of Statistics}, 27(6):1808--1829, 1999.

\bibitem{McMillan1953}
Brockway McMillan.
\newblock The basic theorems of information theory.
\newblock {\em The Annals of Mathematical Statistics}, 24(2):196--219, 1953.

\bibitem{MorseHedlund1940}
Marston Morse and Gustav~A. Hedlund.
\newblock Symbolic dynamics {II}. {S}turmian trajectories.
\newblock {\em American Journal of Mathematics}, 62(1):1--42, 1940.

\bibitem{Parry1964}
William Parry.
\newblock Intrinsic {M}arkov chains.
\newblock {\em Transactions of the American Mathematical Society},
  112(1):55--66, 1964.

\bibitem{RevuzYor1999}
Daniel Revuz and Marc Yor.
\newblock {\em Continuous Martingales and {B}rownian Motion}, volume 293 of
  {\em Grundlehren der mathematischen Wissenschaften}.
\newblock Springer, Berlin, third edition, 1999.

\bibitem{ScottDavenport2007}
Clayton Scott and Mark Davenport.
\newblock Regression level set estimation via cost-sensitive classification.
\newblock {\em IEEE Transactions on Signal Processing}, 55(6):2752--2760,
  2007.

\bibitem{ScottNowak2006}
Clayton Scott and Robert~D. Nowak.
\newblock Minimax-optimal classification with dyadic decision trees.
\newblock {\em IEEE Transactions on Information Theory}, 52(4):1335--1353,
  2006.

\bibitem{Shannon1948}
Claude~E. Shannon.
\newblock A mathematical theory of communication.
\newblock {\em Bell System Technical Journal}, 27(3):379--423, 1948.

\bibitem{SinghScottNowak2009}
Aarti Singh, Clayton Scott, and Robert Nowak.
\newblock Adaptive {H}ausdorff estimation of density level sets.
\newblock {\em The Annals of Statistics}, 37(5B):2760--2782, 2009.

\bibitem{Tanaka1963}
Hiroshi Tanaka.
\newblock Note on continuous additive functionals of the 1-dimensional
  {B}rownian path.
\newblock {\em Zeitschrift f\"ur Wahrscheinlichkeitstheorie und verwandte
  Gebiete}, 1:251--257, 1963.

\bibitem{Tsybakov2004}
Alexandre~B. Tsybakov.
\newblock Optimal aggregation of classifiers in statistical learning.
\newblock {\em The Annals of Statistics}, 32(1):135--166, 2004.

\bibitem{Tsybakov2009Nonparametric}
Alexandre~B. Tsybakov.
\newblock {\em Introduction to Nonparametric Estimation}.
\newblock Springer Series in Statistics. Springer, New York, 2009.

\bibitem{WaltersErgodicTheory}
Peter Walters.
\newblock {\em An Introduction to Ergodic Theory}, volume~79 of {\em Graduate
  Texts in Mathematics}.
\newblock Springer, New York, 1982.

\bibitem{WillettNowak2007}
Rebecca~M. Willett and Robert~D. Nowak.
\newblock Minimax optimal level-set estimation.
\newblock {\em IEEE Transactions on Image Processing}, 16(12):2965--2979,
  2007.

\bibitem{Williams1991}
David Williams.
\newblock {\em Probability with Martingales}.
\newblock Cambridge University Press, Cambridge, 1991.

\bibitem{BoylePetersenHMP}
Mike Boyle and Karl Petersen.
\newblock Hidden Markov processes in the context of symbolic dynamics.
\newblock In B. Marcus, K. Petersen, and T. Weissman, editors, {\em Entropy of
  Hidden Markov Processes and Connections to Dynamical Systems}, London
  Mathematical Society Lecture Note Series, vol.~385, pages 1--38. Cambridge
  University Press, 2011.

\bibitem{PollicottKempton2016}
Mark Pollicott and Thomas Kempton.
\newblock Factors of Gibbs measures for full shifts.
\newblock {\em Proceedings of the American Mathematical Society},
  144(5):2161--2173, 2016.

\bibitem{Verbitskiy2011}
Evgeny Verbitskiy.
\newblock Thermodynamics of hidden Markov processes.
\newblock {\em Letters in Mathematical Physics}, 97(3):341--361, 2011.

\bibitem{Rabin1963}
Michael O. Rabin.
\newblock Probabilistic automata.
\newblock {\em Information and Control}, 6(3):230--245, 1963.

\bibitem{Paz1971}
Azaria Paz.
\newblock {\em Introduction to Probabilistic Automata}.
\newblock Academic Press, New York, 1971.

\bibitem{Torgersen1991}
Erik Torgersen.
\newblock {\em Comparison of Statistical Experiments}, volume 36 of {\em
  Encyclopedia of Mathematics and Its Applications}.
\newblock Cambridge University Press, Cambridge, 1991.

\bibitem{Blackwell1957}
David Blackwell.
\newblock The identifiability problem for functions of finite {M}arkov chains.
\newblock {\em The Annals of Mathematical Statistics}, 28(4):1011--1015, 1957.

\bibitem{Heller1965}
Alex Heller.
\newblock On stochastic processes derived from {M}arkov chains.
\newblock {\em The Annals of Mathematical Statistics}, 36(4):1286--1291, 1965.

\bibitem{BaumPetrie1966}
Leonard~E. Baum and Ted Petrie.
\newblock Statistical inference for probabilistic functions of finite state
  {M}arkov chains.
\newblock {\em The Annals of Mathematical Statistics}, 37(6):1554--1563, 1966.

\bibitem{Thomas1990}
Wolfgang Thomas.
\newblock Automata on infinite objects.
\newblock In Jan van Leeuwen, editor, {\em Handbook of Theoretical Computer
  Science}, volume~B, pages 133--191. Elsevier, Amsterdam, 1990.

\bibitem{PerrinPin2004}
Dominique Perrin and Jean-{\'E}ric Pin.
\newblock {\em Infinite Words: Automata, Semigroups, Logic and Games}.
\newblock Academic Press, London, 2004.

\bibitem{Kitchens1998}
Bruce~P. Kitchens.
\newblock {\em Symbolic Dynamics: One-Sided, Two-Sided and Countable State
  Markov Shifts}.
\newblock Springer, Berlin, 1998.

\end{thebibliography}

\end{document}
