\section{Conclusion}\label{sec:conclusion}

\subsection*{Summary of results}

The paper answers a single question: given an admitted reference carrying a finite abelian gluing obstruction, what part of that obstruction is semantically visible without enlarging the state space?

Theorem~A establishes that the question cannot even be stated at the information-state forcing layer. Compatible local sectionability and gluing-level absence are undefinable in the lower forcing language, so the local-object enrichment is logically necessary rather than a notational convenience.

Theorem~B provides the semantic framework for the answer. Visible value classes are the global connected components of the realization stack; each visible branch determines a banded component gerbe; and gluing-level absence holds exactly when every component gerbe is non-neutral.

Theorem~C delivers the quantitative answer. The intrinsic visible quotient is the cokernel of the homological evaluation map. The $H_2$-visible part of the obstruction is captured exactly by this quotient, while the blind residual---the part invisible to every admissible character---is a pure $\operatorname{Ext}$-type class. In the unique-branch support-presheaf setting, this recovers the Abramsky--Brandenburger no-global-section picture and identifies the precise mechanism behind the cohomological blind cases highlighted by Car\`u \cite{Caru2017}.

\subsection*{External significance}

The connection to contextuality gives the results an anchor outside pure sheaf theory. The forcing-necessity theorem shows that the passage from pointwise semantics to local objects is a logical enrichment in the precise sense that gluing predicates are undefinable at the lower layer. The branched gerbe semantics replaces the binary local-to-global defect of standard contextuality analysis with a graded, branch-indexed account. The homological visibility theorem turns the universal coefficient sequence into a tool for computing the exact boundary between what is semantically accessible and what remains hidden.

\subsection*{Relation to topos-theoretic forcing}

The four-layer chain developed here shares its conceptual ancestry with topos-theoretic forcing as systematized by Tierney \cite{MacLaneMoerdijk1994} and elaborated in Johnstone's comprehensive treatment \cite{Johnstone2002}. In the topos-theoretic tradition, forcing is modeled by the internal logic of a sheaf topos, and conservativity results take the form of geometric morphism adjunctions that preserve logical structure. The present framework is narrower in one respect---it works with a fixed site rather than varying the ambient topos---but broader in another: the explicit stratification into numbered layers with explicit forgetful functors and state projections (\Cref{def:conservative-extension}) makes the conservativity claims syntactically verifiable at each boundary, rather than relying on the general preservation properties of geometric morphisms. The gerbe-theoretic layer, in particular, uses the Giraud classification in a site-specific rather than topos-internal way, which permits the finite combinatorial analysis of visibility that drives Theorem~C. Whether the full four-layer chain admits a natural topos-internal reformulation---with each layer corresponding to a subterminal object or a Lawvere--Tierney topology on a single ambient topos---remains an open question that would connect the present results to the broader categorical semantics programme.

\subsection*{Open problems}

Two directions remain open. First, the finite-state complexity upper bounds in the appendix have no matching lower bounds; determining the exact complexity of the support and minimal-support problems is an open combinatorial question. Second, the present paper treats a single admitted reference at one state. The multi-observer extension---indexed families of references with coupled states and observer symmetry---is deferred to a sequel. The conservativity proposition for observer-indexed extensions is straightforward, but the interaction between branch visibility and observer perspectives introduces new structure not addressed here.
