\section{Compressed abstract calibrated toy-model corollaries for response loci and homogeneous shells}\label{app:abstract-homogeneous-main}

The paper proves one exact homogeneous detector-field realization in
\Cref{prop:symmetry-homogeneous-main}, and that realization has constant lapse.
Everything collected here is therefore only appendix-level algebra for one
fixed calibrated response function. No separate exact nonconstant-lapse UDW
model with
\[
\mathcal S_x(\Omega)\equiv\mathcal S(\Omega)
\]
is constructed here.

\subsection{Threshold-selected response loci for one fixed calibrated response function}

This subsection only inserts the two already selected record thresholds into
one prescribed calibrated response profile and then invokes
\Cref{thm:response-locus-main} for range and regularity.

\begin{corollary}[Abstract calibrated toy-model corollary: response loci for the classical sampled-counter thresholds]\label{thm:shells-general-main}
\ResponseStandingClause the two classical record thresholds
from \Cref{sec:records} lift to the following response loci for one fixed
stationary detector response model. Fix a detector design
$(\Omega,\lambda_c,\Delta\tau,\kappa_{\mathrm r})$ in a static KMS background,
so that \Cref{ass:response-standing-main} supplies the gap-regular absolutely
continuous spectrum near $\pm\Omega$, the absence of atoms there, the
positivity of $\widehat W_x(\Omega)$, and the $C^1$ regularity of
$x\mapsto\mathcal S_x(\Omega)$ and $R_\Omega$ on the region of interest. For
that fixed response function $R_\Omega$, the two record benchmarks are
equivalent to the following response loci. Their
nonemptiness is exactly the range check
$\gamma_\varphi,\gamma_{\mathrm{mem}}\in\operatorname{range}(R_\Omega)$, and
shell language applies only at points where $dR_\Omega\neq0$.
\begin{enumerate}[label=\textup{(\roman*)},leftmargin=2.2em,itemsep=0.25em]
\item The ideal one-step blanking counter mode reaches the Parry benchmark
exactly on the response locus
\begin{equation}\label{eq:shell-phi-general-main}
\Sigma_{\varphi}=
\left\{x:
\frac{\lambda_{\mathrm{sw}}^2\mathcal S_x(\Omega)\Delta\tau}{e^{\beta_\infty N(x)\Omega}-1}=\log\varphi
\right\}.
\end{equation}
Equivalently, this is the fixed-response locus $R_\Omega^{-1}(\gamma_\varphi)$.
Wherever $dR_\Omega\neq0$ on this locus, \Cref{thm:response-locus-main}
upgrades it to a smooth codimension-one shell; otherwise it remains only a
response locus inside the chosen calibrated model.

\item The sampled one-recovery-opportunity boundary-reset counter mode reaches
the memory-transition point exactly on the response locus
\begin{equation}\label{eq:shell-mem-general-main}
\Sigma_{\mathrm{mem}}=
\left\{x:
\frac{\lambda_{\mathrm{sw}}^2\mathcal S_x(\Omega)}{e^{\beta_\infty N(x)\Omega}-1}
=\Gamma_c(\kappa_{\mathrm r},\Delta\tau)
\right\}.
\end{equation}
Equivalently, this is the fixed-response locus
$R_\Omega^{-1}(\gamma_{\mathrm{mem}})$. Wherever $dR_\Omega\neq0$ on this
locus, it is a smooth codimension-one shell. On the locus itself, the sampled
record is exact $1$-dependent.
\end{enumerate}
\end{corollary}

\begin{proof}
For the ideal counter mode, Proposition~\ref{prop:ideal-main} says that the
Parry benchmark is equivalent to $\Gamma\Delta\tau=\log\varphi$. Under the
Model~A assumption of \Cref{ass:udw-rate-calibration-main}, this becomes
$R_\Omega(x)\Delta\tau=\log\varphi$, i.e. $R_\Omega(x)=\gamma_\varphi$ with
the standing shorthand from \eqref{eq:target-rates-main}. This condition is
exactly \eqref{eq:shell-phi-general-main}, so the locus is
$R_\Omega^{-1}(\gamma_\varphi)$. The range and regularity statements are then
exactly the content of \Cref{thm:response-locus-main}.

For the sampled one-recovery-opportunity boundary-reset model,
\Cref{prop:recovery-main,lem:one-dep-criterion-main} say that the memory
transition is equivalent to $\Gamma=\Gamma_c(\kappa_{\mathrm r},\Delta\tau)$.
Under the same Model~A assumption, this becomes
$R_\Omega(x)=\Gamma_c(\kappa_{\mathrm r},\Delta\tau)$, i.e.
$R_\Omega(x)=\gamma_{\mathrm{mem}}$. With the same fixed calibrated response
function, this is exactly \eqref{eq:shell-mem-general-main}, so the locus is
$R_\Omega^{-1}(\gamma_{\mathrm{mem}})$. Every point satisfying that equation
has an exact $1$-dependent sampled record in that model. The
\Cref{lem:one-gap-kms-underdetermination-main} explains why this corollary is
only conditional on one fixed calibrated response function: changing
$\mathcal S_x(\Omega)$ changes the loci while preserving one-gap detailed
balance.
\end{proof}

\subsection{Homogeneous specialization of the fixed calibrated response theorem}

No new detector-field input is added below. The remaining statements start from
\Cref{thm:response-locus-main,thm:shells-general-main} and then impose extra
calibration or background hypotheses in the one-way order
\[
\text{fixed }R_\Omega
\Longrightarrow
\text{response loci}
\Longrightarrow
\text{pair transport}
\Longrightarrow
\text{homogeneous toy-model specializations}.
\]
Any nonconstant lapse is inserted only at the last stage of that chain.

\begin{lemma}[Abstract calibrated toy-model corollary: homogeneous specialization]\label{lem:abstract-homogeneous-specialization-main}
\ResponseStandingClause specialize the fixed calibrated response function $R_\Omega$ to the homogeneous form
below. Fix a static region $U$, a gap $\Omega>0$, a Tolman parameter
$\beta_\infty>0$, a positive constant $A$, and a $C^1$ lapse
$N:U\to(0,\infty)$. Define the calibrated response function
\[
R_\Omega^{\mathrm{hom}}(x)=\frac{A}{e^{\beta_\infty N(x)\Omega}-1}.
\]
Then for any target rate $\gamma>0$ the response locus
\[
L_\gamma^{\mathrm{hom}}
=\{x\in U:R_\Omega^{\mathrm{hom}}(x)=\gamma\}
\]
is exactly the lapse level set
\[
\beta_\infty N(x)=\frac{1}{\Omega}\log\!\left(1+\frac{A}{\gamma}\right).
\]
If $dN(x_0)\neq0$ at $x_0\in L_\gamma^{\mathrm{hom}}$, then
$L_\gamma^{\mathrm{hom}}$ is near $x_0$ a $C^1$ codimension-one shell. This is
an abstract calibrated toy-model specialization of a fixed response function.
The only exact detector-field realization proved in this paper is
\Cref{prop:symmetry-homogeneous-main}, where the lapse is constant; whenever
later appendices use homogeneous shell geometry with nonconstant lapse, they do
so through this level-set specialization rather than through a separate exact
UDW construction.
\end{lemma}

\begin{proof}
Solving $R_\Omega^{\mathrm{hom}}(x)=\gamma$ gives
\[
e^{\beta_\infty N(x)\Omega}=1+\frac{A}{\gamma},
\]
hence the displayed lapse level set. Differentiating yields
\[
dR_\Omega^{\mathrm{hom}}
=-A\frac{\beta_\infty\Omega e^{\beta_\infty N\Omega}}
{(e^{\beta_\infty N\Omega}-1)^2}\,dN,
\]
whose scalar prefactor is strictly negative. Therefore
$dR_\Omega^{\mathrm{hom}}(x_0)\neq0$ exactly when $dN(x_0)\neq0$, and the
regular level-set theorem gives the shell statement.
\end{proof}

\begin{corollary}[Abstract calibrated toy-model corollary: homogeneous lapse-level sets]\label{cor:homogeneous-main}
\ResponseStandingClause assume in addition the calibrated homogeneous spectral-weight regime.
Assume in addition that $\mathcal S_x(\Omega)\equiv\mathcal S(\Omega)$ on the
region under study. Then both response loci are pure lapse level sets, with local inverse
temperatures
\begin{equation}\label{eq:betaphi-betamem-main}
\beta_{\varphi}=
\frac{1}{\Omega}
\log\!\left(1+\frac{\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)\Delta\tau}{\log\varphi}\right),
\qquad
\beta_{\mathrm{mem}}=
\frac{1}{\Omega}
\log\!\left(1+\frac{\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)}
{\Gamma_c(\kappa_{\mathrm r},\Delta\tau)}\right),
\end{equation}
so that
\begin{equation}\label{eq:lapse-level-main}
\Sigma_{\varphi}=\{x:\beta_{\mathrm{loc}}(x)=\beta_{\varphi}\},
\qquad
\Sigma_{\mathrm{mem}}=\{x:\beta_{\mathrm{loc}}(x)=\beta_{\mathrm{mem}}\}.
\end{equation}
If $dN\neq0$ on one of these level sets, then
\Cref{thm:response-locus-main,lem:abstract-homogeneous-specialization-main}
identify it as the corresponding shell; for nonconstant lapse this remains an
abstract calibrated toy-model specialization.
\end{corollary}

\begin{proof}
With $\mathcal S_x(\Omega)\equiv\mathcal S(\Omega)$, the two response-locus
equations contain $x$ only through $\beta_\infty N(x)$. Solving the
ideal-benchmark equation for $\beta_\infty N(x)$ gives $\beta_\varphi$, and
solving the memory equation gives $\beta_{\mathrm{mem}}$. Hence the loci are
the lapse level sets in~\eqref{eq:lapse-level-main}.
\end{proof}

For later pair comparisons write $r_\varphi$ and
$r_{\mathrm{mem}}$ for the homogeneous ideal-benchmark and
memory-transition radii, and set
\[
D_\varphi=e^{\beta_\infty\Omega N(r_\varphi)}-1,
\qquad
D_{\mathrm{mem}}=e^{\beta_\infty\Omega N(r_{\mathrm{mem}})}-1,
\qquad
Q_{\varphi/\mathrm{mem}}=\frac{D_\varphi}{D_{\mathrm{mem}}}.
\]
The cancellation identity used later is proved in \Cref{thm:ratio-main}; here
we use only the level-set formulas above.

\subsection{Monotone spherical lapse bookkeeping in the abstract calibrated toy model}

Everything here stays at the level of
\Cref{lem:abstract-homogeneous-specialization-main}. No new exact
detector-field realization with
$\mathcal S_x(\Omega)\equiv\mathcal S(\Omega)$ is being claimed here.
Any nonconstant-lapse insertion is only bookkeeping for a fixed calibrated
response function.

\begin{proposition}[Abstract calibrated toy-model corollary: monotone spherical existence]\label{prop:existence-main}
\ResponseStandingClause assume in addition the homogeneous
specialization on a monotone spherical region.
Assume the homogeneous specialization of \Cref{cor:homogeneous-main} in a static
spherically symmetric region $r\in I$. Suppose
$N:I\to N(I)$ is $C^1$ and strictly monotone. For either threshold
$\star\in\{\varphi,\mathrm{mem}\}$, the corresponding response locus exists in $I$ if
and only if
\[
\frac{\beta_\star}{\beta_\infty}\in N(I),
\]
where $\beta_\varphi$ and $\beta_{\mathrm{mem}}$ are defined in~\eqref{eq:betaphi-betamem-main}. When it exists, its radius is unique and equals
\[
r_\star=N^{-1}\!\left(\frac{\beta_\star}{\beta_\infty}\right).
\]
If $N'(r_\star)\neq0$, this response locus is a regular spherical shell inside
the abstract calibrated toy model.
\end{proposition}

\begin{proof}
By \Cref{cor:homogeneous-main}, the shell equation is exactly
$\beta_\infty N(r)=\beta_\star$. Since $N$ is strictly monotone on $I$, it is
one-to-one onto its range. Hence the equation has a solution precisely when
$\beta_\star/\beta_\infty$ lies in $N(I)$, and the solution is then the
displayed inverse image. Thus shell existence on a monotone spherical region is
completely equivalent to the calibrated target lapse lying in the available
range of $N$. The nonzero derivative condition is the one-dimensional form of
the regular level-set condition.
\end{proof}

Under this homogeneous specialization, both response loci are lapse level sets
selected by the two record thresholds and become shells only after the regular
level-set check. The default comparison remains the transport identity
\Cref{cor:spectral-transport-main}; Appendix~\ref{app:schwarzschild-toy}
records a further Schwarzschild-lapse toy specialization.

\begin{proposition}[Abstract calibrated toy-model corollary: detector-controlled ordering]\label{prop:order-main}
\ResponseStandingClause assume in addition the homogeneous
specialization. Assume the calibrated
homogeneous spectral-weight regime, and define
\[
\Gamma_{\varphi}=\frac{\log\varphi}{\Delta\tau}.
\]
Let $N_{\varphi}$ and $N_{\mathrm{mem}}$ denote the lapse values of the two
response loci $\Sigma_{\varphi}$ and $\Sigma_{\mathrm{mem}}$. Then
\begin{equation}\label{eq:order-main}
N_{\varphi}<N_{\mathrm{mem}}
\iff
\Gamma_{\varphi}>\Gamma_c(\kappa_{\mathrm r},\Delta\tau),
\end{equation}
with equality if and only if $\Gamma_{\varphi}=\Gamma_c(\kappa_{\mathrm r},\Delta\tau)$. Thus the ordering of the two homogeneous response loci is detector controlled; when the regularity condition holds, this is the shell ordering inside the abstract calibrated toy model.
\end{proposition}

\begin{proof}
Both response loci are lapse level sets in the calibrated homogeneous
spectral-weight regime. Writing $A=\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)>0$,
the local excitation rate as a function of the lapse is
\[
\Gamma_+(N)=\frac{A}{e^{\beta_\infty\Omega N}-1}.
\]
For $\beta_\infty,\Omega,N>0$,
\[
\frac{d\Gamma_+}{dN}
=
-A\,\frac{\beta_\infty\Omega e^{\beta_\infty\Omega N}}
{(e^{\beta_\infty\Omega N}-1)^2}<0 .
\]
Hence the higher threshold rate corresponds exactly to the smaller lapse value, giving~\eqref{eq:order-main} and the equality statement.
\end{proof}
