
\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{hyperref}
\setstretch{1.08}
\setlength{\parindent}{1.2em}
\setlength{\parskip}{0.45em}
\begin{document}

\begin{flushright}
Haobo Ma and Wenlin Zhang\\
CHRONOAI PTE. LTD., Singapore; National University of Singapore, Singapore\\
\href{mailto:auric@aelf.io}{auric@aelf.io}; \href{mailto:e1327962@u.nus.edu}{e1327962@u.nus.edu}\\
\today
\end{flushright}

Editor-in-Chief\\
\emph{J. Math. Phys.}

Dear Editor,

Please consider our manuscript, \emph{``A conditional note on response-model shells from a sampled classical counter driven by the stationary Unruh--DeWitt excitation rate in static KMS spacetimes''}, for publication as a Research Article in \emph{J. Math. Phys.}.

The paper shows that the full stationary binary record of a sampled classical counter, whose attempt intensity is calibrated by the stationary Unruh--DeWitt excitation rate, selects geometric response level sets that rate thermometry alone would not single out without the readout model. Dead time and finite recovery are intrinsic features of counting detectors, and exploiting their full stochastic output goes beyond summarizing the detector by its mean transition rate. The mathematical core of the work---symbolic dynamics on subshifts of finite type, renewal-process theory, spectral analysis, and the implicit function theorem for local ratio-level data inversion---is presented as a self-contained theorem chain suited to \emph{J. Math. Phys.}.

The present manuscript is a self-contained treatment, with results on universal antibunching, soft dead time, shell-pair inverse formulas, and a complete Schwarzschild worked example.

The principal results are the following.
\begin{enumerate}
\item In the ideal dead-time model, the sampled record is a Markov chain on the golden-mean shift, and the Parry benchmark is realized at a unique detector threshold.
\item In the finite-recovery model, the sampled record is a full-support renewal process with strictly increasing hazard, so the Parry benchmark is excluded. The relevant selector is instead the memory-transition shell at exact $1$-dependence.
\item Finite recovery approaches the ideal grammar only softly: consecutive-click violations are quadratically suppressed in the sampling interval, yet the hazard remains genuinely history dependent at every positive sampling interval.
\item The finite-recovery record is universally antibunched, with strictly increasing spectral density on $[0,\pi]$ and an asymptotic Fano factor below one across the whole parameter range.
\item In the homogeneous spectral-weight regime, the shell pair obeys an exact amplitude-cancellation law. For shell pairs tending to a nonextremal Hartle--Hawking horizon this yields a universal leading-order ratio law for shell distances.
\item Within the same response model, a calibrated Schwarzschild shell pair gives explicit inverse formulas for the redshift scale and mass parameter, and two operating modes sharing the same detector core provide a local two-ratio inverse test on static spherical two-parameter families, with full shell-data compatibility checked separately.
\end{enumerate}

The main article contains the complete theorem proofs. The Supplementary Material records auxiliary calculations, finite-counting checks, coordinate expansions, and numerical bookkeeping supporting the displayed formulas. The scripts generating the figures and tables are available upon request.

The work is original, has not been published previously, and is not under consideration elsewhere. All authors have approved the submission. We have no competing interests to declare.

Thank you for your consideration.

Sincerely,

Haobo Ma and Wenlin Zhang

\end{document}
