\section{Conditional shell-pair observables and ratio identifiability}\label{sec:inverse}

We now use the shell loci of \Cref{thm:shells-general-main}. Table~\ref{tab:hypothesis-ledger}
continues to be the standing map, and the boxed quotient convention below fixes
the notation once for the whole section. The theorem chain has three layers:
\Cref{cor:spectral-transport-main} gives the default shell-pair comparison for
a fixed calibrated response function, \Cref{thm:ratio-main} specializes it to
the homogeneous case, and the remaining subsections extract near-horizon and
fixed-radius inverse consequences of that homogeneous specialization. As in
\Cref{sec:kms-shells}, the theorem statements retain the explicit sampled-counter
closure in their first sentence, so the prose below records only the
additional input specific to each layer.

\subsection{Physical spectral transport and abstract homogeneous specialization}

The shell-pair ratio used below is a quotient of the two Planck denominators,
not a quotient of the two target rates. Its orientation is fixed once and then
kept throughout the section. For shell radii $r_\varphi$ and
$r_{\mathrm{mem}}$ define
\begin{equation}\label{eq:denominator-convention-main}
D_\varphi=e^{\beta_\infty\Omega N(r_\varphi)}-1,\qquad
D_{\mathrm{mem}}=e^{\beta_\infty\Omega N(r_{\mathrm{mem}})}-1,\qquad
Q_{\varphi/\mathrm{mem}}=\frac{D_\varphi}{D_{\mathrm{mem}}}.
\end{equation}
Thus $Q_{\varphi/\mathrm{mem}}$ always means ideal-benchmark denominator divided by
memory-transition denominator. The exact-clock condition contains the sampling
factor $\Delta\tau$, whereas the memory-transition condition is already a rate
condition. Keeping these two source equations in denominator form is what fixes
the quotient below.

The two members of the pair are obtained from the same field-facing detector
core: the gap, coupling scale, detector profile, and field state are fixed,
while the classical counter mode is changed from ideal one-step blanking to
the sampled one-recovery-opportunity boundary-reset mode. In the calibrated homogeneous spectral-weight regime,
this shared core contributes one common positive factor
	\[
	A=\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)
	\]
to both source shell equations.

Under the calibrated homogeneous spectral-weight regime and a common
field-facing detector core, the two source equations in the fixed denominator
notation are
\begin{equation}\label{eq:denominator-shells-main}
\frac{A\Delta\tau}{D_\varphi}=\log\varphi,
\qquad
\frac{A}{D_{\mathrm{mem}}}
=\Gamma_c(\kappa_{\mathrm r},\Delta\tau),
\qquad
	A=\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)>0 .
\end{equation}
Here the ideal-benchmark member uses ideal one-step blanking, while the
memory-transition member uses the sampled one-recovery-opportunity
boundary-reset mode and the same sampling
interval and recovery rate. Since all factors are positive, these two source
equations are equivalently
\begin{equation}\label{eq:denominator-rearranged-main}
D_\varphi=\frac{A\Delta\tau}{\log\varphi},
\qquad
D_{\mathrm{mem}}=\frac{A}{\Gamma_c(\kappa_{\mathrm r},\Delta\tau)} .
\end{equation}
The factor $\Delta\tau$ appears only in the ideal-benchmark source equation because
the ideal record threshold is $\Gamma_+\Delta\tau=\log\varphi$, whereas the
memory threshold is already a rate equation,
$\Gamma_+=\Gamma_c(\kappa_{\mathrm r},\Delta\tau)$.

The physically robust comparison for one fixed calibrated response function
keeps the two shell points distinct and therefore produces the
spectral-transport factor of \Cref{cor:spectral-transport-main}. The
homogeneous quotient law used later in the horizon and inverse subsections is
an additional abstract calibrated specialization obtained by imposing
$\mathcal S_x(\Omega)\equiv\mathcal S(\Omega)$. Because the later algebra uses
its denominator bookkeeping, we record that specialization first and then state
the general physical comparison.

\medskip
\noindent
\fbox{%
\parbox{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}{%
\textbf{Convention for shell-pair quotients.}
The slash notation records numerator over denominator in the written order:
\[
Q_{\varphi/\mathrm{mem}}:=\frac{D_\varphi}{D_{\mathrm{mem}}},
\qquad
Q_{\mathrm{mem}/\varphi}:=\frac{D_{\mathrm{mem}}}{D_\varphi}
=Q_{\varphi/\mathrm{mem}}^{-1}.
\]
We abbreviate the detector-side constant by
\[
C_D:=\frac{\Delta\tau\,\Gamma_c(\kappa_{\mathrm r},\Delta\tau)}{\log\varphi}.
\]
Thus the homogeneous denominator quotient proved below is
$Q_{\varphi/\mathrm{mem}}=C_D$, while the reciprocal target-rate quotient is
\[
\frac{\gamma_\varphi}{\gamma_{\mathrm{mem}}}
=Q_{\mathrm{mem}/\varphi}
=C_D^{-1}.
\]
}}%
\medskip

\begin{corollary}[Abstract homogeneous denominator quotient for a fixed calibrated $R_\Omega$]\label{thm:ratio-main}
Assume the sampled classical interval instrument of
\Cref{prop:instrument-kernels-main}, with independent Poisson attempts and
fresh interval ancillas, and impose only the modelling closure
$\Gamma=\Gamma_+(x,\Omega)$. Assume the calibrated homogeneous spectral-weight
regime. Thus the fixed calibrated response function $R_\Omega$ has homogeneous
spectral weight. Let the ideal-benchmark and memory-transition shell radii be
produced by one field-facing detector core
with fixed gap, coupling scale, detector profile, and field state. The
ideal-benchmark member uses ideal one-step blanking with sampling interval
$\Delta\tau$, and the memory member uses the sampled one-recovery-opportunity
boundary-reset mode and the same
sampling interval and recovery rate $\kappa_{\mathrm r}$. Using the fixed
denominator convention in \eqref{eq:denominator-convention-main}, the
shell-pair denominator quotient is
\begin{equation}\label{eq:ratio-main}
Q_{\varphi/\mathrm{mem}}
=\frac{D_\varphi}{D_{\mathrm{mem}}}
=
\frac{e^{\beta_\infty\Omega N(r_\varphi)}-1}
     {e^{\beta_\infty\Omega N(r_{\mathrm{mem}})}-1}
=
\frac{\Delta\tau\,\Gamma_c(\kappa_{\mathrm r},\Delta\tau)}
{\log\varphi}.
\end{equation}
The reciprocal constant
\[
C_D^{-1}
=\frac{\log\varphi}
{\Delta\tau\,\Gamma_c(\kappa_{\mathrm r},\Delta\tau)}
\]
belongs to the reciprocal denominator quotient
$Q_{\mathrm{mem}/\varphi}$ and not to
$Q_{\varphi/\mathrm{mem}}$. In particular, for the two record-selected target
rates
\[
\gamma_\varphi=\frac{\log\varphi}{\Delta\tau},
\qquad
\gamma_{\mathrm{mem}}=\Gamma_c(\kappa_{\mathrm r},\Delta\tau),
\]
one has
\begin{equation}\label{eq:target-quotient-main}
\frac{\gamma_\varphi}{\gamma_{\mathrm{mem}}}
=
Q_{\mathrm{mem}/\varphi}
=C_D^{-1}
=Q_{\varphi/\mathrm{mem}}^{-1}.
\end{equation}
This is a cancellation identity inside the calibrated homogeneous response
model; it does not derive the common amplitude $A$ or the homogeneity of
$\mathcal S_x(\Omega)$ from the KMS condition. For a general fixed calibrated
response function without the homogeneous specialization, the default physical
statement is instead \Cref{cor:spectral-transport-main}.
\end{corollary}

\begin{proof}
The ideal-benchmark and memory-transition thresholds are supplied by
Proposition~\ref{prop:ideal-main} and
\Cref{lem:one-dep-criterion-main}. In the calibrated homogeneous specialization of
\Cref{thm:shells-general-main,cor:homogeneous-main},
\[
\Gamma_+(r,\Omega)=\frac{A}{e^{\beta_\infty\Omega N(r)}-1},
\qquad A=\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)>0 .
\]
Substitution gives the two source equations
\eqref{eq:denominator-shells-main}, hence the two positive denominator
identities
\[
D_\varphi=\frac{A\Delta\tau}{\log\varphi},
\qquad
D_{\mathrm{mem}}=\frac{A}{\Gamma_c(\kappa_{\mathrm r},\Delta\tau)},
\qquad
A=\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)>0.
\]
Dividing the first identity by the second cancels exactly the common response
amplitude $A$ and gives \eqref{eq:ratio-main}. Dividing in the opposite
orientation gives the reciprocal denominator quotient. Since the ideal-benchmark
and memory target rates are
$\gamma_\varphi=\log\varphi/\Delta\tau$ and
$\gamma_{\mathrm{mem}}=\Gamma_c(\kappa_{\mathrm r},\Delta\tau)$, their quotient
is precisely the reciprocal denominator quotient, proving
\eqref{eq:target-quotient-main}. No further geometric hypothesis is used beyond
the stated response-model and common-core inputs.
\end{proof}

This cancellation law uses one common field-facing core and homogeneous
spectral weight; without the second input the transport factor in
\Cref{cor:spectral-transport-main} remains. The sampling and recovery
parameters still enter through the readout constant
\[
C_D=\frac{\Delta\tau\,\Gamma_c(\kappa_{\mathrm r},\Delta\tau)}{\log\varphi},
\]
so the ratio law is amplitude-free, not detector-free. By the boxed
convention, all later ratio maps keep the orientation
$Q_{\varphi/\mathrm{mem}}=C_D$ unless the reciprocal is named explicitly. The
next corollary is the general physical identity of which
\Cref{thm:ratio-main} is the homogeneous specialization.

\begin{corollary}[Physical shell-pair spectral-transport identity for a fixed calibrated $R_\Omega$]\label{cor:spectral-transport-main}
Assume the sampled classical interval instrument of
\Cref{prop:instrument-kernels-main}, with independent Poisson attempts and
fresh interval ancillas, and impose only the modelling closure
$\Gamma=\Gamma_+(x,\Omega)$. Assume the static response regularity and
calibration of \Cref{ass:kms-response-main}, so that $R_\Omega$ is fixed. Let
$x_\varphi$ and $x_{\mathrm{mem}}$ be points on the ideal-benchmark and
memory-transition loci for one common field-facing detector core with the same
gap, coupling scale, detector profile, and field state, but with the two
counter modes specified above. Thus
\[
\frac{\lambda_{\mathrm{sw}}^2\mathcal S_{x_\varphi}(\Omega)\Delta\tau}
     {e^{\beta_\infty\Omega N(x_\varphi)}-1}
=\log\varphi,
\qquad
\frac{\lambda_{\mathrm{sw}}^2\mathcal S_{x_{\mathrm{mem}}}(\Omega)}
     {e^{\beta_\infty\Omega N(x_{\mathrm{mem}})}-1}
=\Gamma_c(\kappa_{\mathrm r},\Delta\tau).
\]
Then
\begin{equation}\label{eq:spectral-transport-main}
\frac{e^{\beta_\infty\Omega N(x_{\varphi})}-1}
     {e^{\beta_\infty\Omega N(x_{\mathrm{mem}})}-1}
=
\frac{\Delta\tau\,\Gamma_c(\kappa_{\mathrm r},\Delta\tau)}
{\log\varphi}\,
\frac{\mathcal S_{x_{\varphi}}(\Omega)}
     {\mathcal S_{x_{\mathrm{mem}}}(\Omega)} .
\end{equation}
Thus every departure from the homogeneous denominator-ratio law is exactly the
displayed spectral-weight transport factor. If the lapse model is
independently known, the same identity can be read as a measurement of that
spectral-weight ratio.
\end{corollary}

\begin{proof}
Set
\[
D_\varphi(x_\varphi)=e^{\beta_\infty\Omega N(x_\varphi)}-1,\qquad
D_{\mathrm{mem}}(x_{\mathrm{mem}})
=e^{\beta_\infty\Omega N(x_{\mathrm{mem}})}-1 .
\]
The two displayed shell equations are positive and therefore equivalent to
\[
D_\varphi(x_\varphi)
=\frac{\lambda_{\mathrm{sw}}^2\mathcal S_{x_\varphi}(\Omega)\Delta\tau}{\log\varphi},
\qquad
D_{\mathrm{mem}}(x_{\mathrm{mem}})
=
\frac{\lambda_{\mathrm{sw}}^2\mathcal S_{x_{\mathrm{mem}}}(\Omega)}
{\Gamma_c(\kappa_{\mathrm r},\Delta\tau)} .
\]
Taking their quotient in the fixed orientation
$D_\varphi/D_{\mathrm{mem}}$ cancels the common coupling scale
$\lambda_{\mathrm{sw}}^2$ and leaves exactly \eqref{eq:spectral-transport-main}. In the
homogeneous specialization
$\mathcal S_x(\Omega)\equiv\mathcal S(\Omega)$ the final factor is one, so
\Cref{thm:ratio-main} is recovered.
\end{proof}

\begin{proposition}[Quantitative local homogeneity criterion for a fixed calibrated $R_\Omega$]\label{prop:spectral-stability-main}
Assume the sampled classical interval instrument of
\Cref{prop:instrument-kernels-main}, with independent Poisson attempts and
fresh interval ancillas, and impose only the modelling closure
$\Gamma=\Gamma_+(x,\Omega)$. Assume the hypotheses of
Corollary~\ref{cor:spectral-transport-main}. Let $U$ be a geodesically convex
subset of the static slice containing $x_\varphi$ and $x_{\mathrm{mem}}$, and
suppose $\log \mathcal S_x(\Omega)$ is $C^1$ on $U$. Fix any auxiliary
Riemannian norm $|\cdot|$ on the slice and write
\[
L_U(\Omega)=\sup_{x\in U}|d\log \mathcal S_x(\Omega)|,
\qquad
d_U=d_U(x_\varphi,x_{\mathrm{mem}})
\]
for the corresponding geodesic distance between the two shell points. Let
\[
C_D=\frac{\Delta\tau\,\Gamma_c(\kappa_{\mathrm r},\Delta\tau)}{\log\varphi}
\]
and
\[
R=
\frac{e^{\beta_\infty\Omega N(x_{\varphi})}-1}
     {e^{\beta_\infty\Omega N(x_{\mathrm{mem}})}-1}.
\]
Then
\begin{equation}\label{eq:spectral-stability-main}
e^{-L_U(\Omega)d_U}
\le
\frac{\mathcal S_{x_\varphi}(\Omega)}
     {\mathcal S_{x_{\mathrm{mem}}}(\Omega)}
\le
e^{L_U(\Omega)d_U},
\end{equation}
and therefore
\begin{equation}\label{eq:spectral-stability-ratio-main}
C_D e^{-L_U(\Omega)d_U}
\le R\le
C_D e^{L_U(\Omega)d_U},
\qquad
\left|\frac{R}{C_D}-1\right|
\le e^{L_U(\Omega)d_U}-1.
\end{equation}
In particular, for any family with $L_{U_t}(\Omega(t))\,d_t\to0$ one has
\[
R(t)=C_D(t)\bigl(1+O(L_{U_t}(\Omega(t))\,d_t)\bigr).
\]
Thus the homogeneous ratio law is recovered asymptotically whenever the shell separation is small compared with the spatial variation scale of $\log\mathcal S_x(\Omega)$.
\end{proposition}

\begin{proof}
Let $\gamma:[0,d_U]\to U$ be a unit-speed minimizing geodesic from $x_{\mathrm{mem}}$ to $x_\varphi$. Then
\[
\log\frac{\mathcal S_{x_\varphi}(\Omega)}
         {\mathcal S_{x_{\mathrm{mem}}}(\Omega)}
=
\int_0^{d_U}
d\log\mathcal S_{\gamma(u)}(\Omega)\bigl[\dot\gamma(u)\bigr]\,du.
\]
Hence
\[
\left|
\log\frac{\mathcal S_{x_\varphi}(\Omega)}
         {\mathcal S_{x_{\mathrm{mem}}}(\Omega)}
\right|
\le
\int_0^{d_U}L_U(\Omega)\,du
=L_U(\Omega)d_U,
\]
which gives \eqref{eq:spectral-stability-main} after exponentiation. Multiplying the latter by the exact spectral-transport identity \eqref{eq:spectral-transport-main} yields \eqref{eq:spectral-stability-ratio-main}. If $L_{U_t}(\Omega(t))d_t\to0$, then $e^{L_{U_t}(\Omega(t))d_t}-1=O(L_{U_t}(\Omega(t))d_t)$, giving the final estimate.
\end{proof}

\begin{corollary}[Uniform perturbative form]\label{cor:spectral-stability-perturbative-main}
Assume the sampled classical interval instrument of
\Cref{prop:instrument-kernels-main}, with independent Poisson attempts and
fresh interval ancillas, and impose only the modelling closure
$\Gamma=\Gamma_+(x,\Omega)$. Assume the hypotheses of
Proposition~\ref{prop:spectral-stability-main}. If, on the same region $U$,
\[
\mathcal S_x(\Omega)=\mathcal S(\Omega)(1+\varepsilon(x)),
\qquad
|\varepsilon(x)|\le\varepsilon_0<1,
\]
then
\[
C_D\,\frac{1-\varepsilon_0}{1+\varepsilon_0}
\le R\le
C_D\,\frac{1+\varepsilon_0}{1-\varepsilon_0},
\qquad
\left|\frac{R}{C_D}-1\right|
\le
\frac{2\varepsilon_0}{1-\varepsilon_0}.
\]
\end{corollary}

\begin{proof}
By \Cref{cor:spectral-transport-main},
\[
\frac{R}{C_D}
=
\frac{1+\varepsilon(x_\varphi)}{1+\varepsilon(x_{\mathrm{mem}})}.
\]
The two-sided estimate follows by maximizing and minimizing the numerator and denominator under $|\varepsilon|\le\varepsilon_0$. The absolute bound follows from
\[
\left|
\frac{1+\varepsilon_\varphi}{1+\varepsilon_{\mathrm{mem}}}-1
\right|
=
\left|\frac{\varepsilon_\varphi-\varepsilon_{\mathrm{mem}}}{1+\varepsilon_{\mathrm{mem}}}\right|
\le
\frac{2\varepsilon_0}{1-\varepsilon_0}.
\]
\end{proof}

\subsection{Abstract calibrated small-shell near-horizon hierarchy}

Near a nonextremal Hartle--Hawking horizon, the primary small-shell statement in
this subsection is an exact collar identity for the homogeneous shell-pair
denominator ratio inside the abstract calibrated homogeneous specialization.
That statement requires only that the shell pair enter a Gaussian normal
collar. The further proper-distance and area-excess hierarchy appears only
after an additional linearization step, and that step requires the
dimensionless small-gap condition $\Omega(t)\epsilon(t)\to0$. Without that
extra condition the valid near-horizon conclusion remains the exact nonlinear
collar ratio; no proper-distance ratio follows in general.

\begin{assumption}[Abstract calibrated small-shell horizon family]\label{ass:small-shell-main}
A family of shell pairs in the abstract calibrated homogeneous specialization is called \emph{small-shell} if it is indexed by a parameter $t$, has detector gap $\Omega(t)>0$, and has proper distances
\[
\rho_\varphi(t)>0,\qquad \rho_{\mathrm{mem}}(t)>0
\]
from the horizon, satisfies the exact homogeneous ratio law
\[
\frac{e^{\beta_\infty\Omega(t) N(\rho_\varphi(t))}-1}
     {e^{\beta_\infty\Omega(t) N(\rho_{\mathrm{mem}}(t))}-1}
=C_D(t),
\qquad
C_D(t)=
\frac{\Delta\tau(t)\,\Gamma_c(\kappa_{\mathrm r}(t),\Delta\tau(t))}
{\log\varphi},
\]
and obeys
\[
\epsilon(t)=\max\{\rho_\varphi(t),\rho_{\mathrm{mem}}(t)\}\downarrow0 .
\]
The detector parameters may be fixed or may vary with $t$. The hypothesis is part of the statement, not a consequence of the exact ratio law: outside such a family the horizon expansion below is not being invoked. This assumption by itself gives only an exact Planck-ratio relation in a normal collar. The linear proper-distance and area-excess laws below are an additional consequence obtained only after the dimensionless small-gap condition $\Omega(t)\epsilon(t)\to0$ is imposed. \Cref{lem:abstract-homogeneous-specialization-main} supplies the response-level setting in which this family is interpreted.
\end{assumption}

\begin{lemma}[Near-horizon Planck denominator]\label{lem:varying-gap-denominator-main}
Assume the nonextremal horizon expansion
\[
N(\rho)=\kappa_{\mathrm H}\rho+O(\rho^3),
\qquad
\beta_\infty=\frac{2\pi}{\kappa_{\mathrm H}}.
\]
Let $\Omega(t)>0$, let $0<\rho(t)\le\epsilon(t)$, and suppose
$\epsilon(t)\downarrow0$. Then, uniformly for such $\rho(t)$,
\begin{equation}\label{eq:exact-near-horizon-denominator-main}
e^{\beta_\infty\Omega(t)N(\rho(t))}-1
=
e^{2\pi\Omega(t)\rho(t)(1+O(\epsilon(t)^2))}-1 .
\end{equation}
If, in addition, $\Omega(t)\epsilon(t)\to0$, then the denominator linearizes
uniformly:
\begin{equation}\label{eq:varying-gap-denominator-main}
e^{\beta_\infty\Omega(t)N(\rho(t))}-1
=2\pi\Omega(t)\rho(t)
\Bigl(1+O(\Omega(t)\rho(t))+O(\rho(t)^2)\Bigr).
\end{equation}
In particular, if $\Omega(t)$ is bounded, the relative error is $O(\rho(t))$.
\end{lemma}

\begin{proof}
The horizon expansion and the Hartle--Hawking relation give
\[
\beta_\infty\Omega(t)N(\rho)
=2\pi\Omega(t)\rho\bigl(1+O(\rho^2)\bigr),
\]
uniformly in a fixed normal collar. Since $\rho\le\epsilon(t)$ and
$\epsilon(t)\downarrow0$, this proves
\eqref{eq:exact-near-horizon-denominator-main}. If also
$\Omega(t)\epsilon(t)\to0$, the exponent is $o(1)$ uniformly along the family.
Using $e^y-1=y(1+O(y))$ for $y=o(1)$ then gives
\[
e^{\beta_\infty\Omega(t)N(\rho)}-1
=2\pi\Omega(t)\rho
\bigl(1+O(\rho^2)\bigr)
\bigl(1+O(\Omega(t)\rho)\bigr),
\]
which is \eqref{eq:varying-gap-denominator-main}. For bounded $\Omega(t)$ the
term $O(\Omega(t)\rho(t))$ is $O(\rho(t))$, while
$O(\rho(t)^2)$ is smaller.
\end{proof}

\begin{lemma}[Exact abstract near-horizon collar ratio]\label{lem:exact-collar-ratio-main}
Assume the sampled classical interval instrument of
\Cref{prop:instrument-kernels-main}, with independent Poisson attempts and
fresh interval ancillas, and impose only the modelling closure
$\Gamma=\Gamma_+(x,\Omega)$. Assume the small-shell horizon-family hypothesis in
\Cref{ass:small-shell-main} and the nonextremal Hartle--Hawking expansion
\[
N(\rho)=\kappa_{\mathrm H}\rho+O(\rho^3),
\qquad
\beta_\infty=\frac{2\pi}{\kappa_{\mathrm H}}.
\]
Then the exact homogeneous ratio law is equivalent to
\begin{equation}\label{eq:exact-collar-ratio-main}
C_D(t)=
\frac{
e^{2\pi\Omega(t)\rho_\varphi(t)(1+O(\epsilon(t)^2))}-1}
{
e^{2\pi\Omega(t)\rho_{\mathrm{mem}}(t)(1+O(\epsilon(t)^2))}-1},
\end{equation}
with uniform $O(\epsilon(t)^2)$ terms along the family. This is the primary
near-horizon statement for small-shell families: it remains valid even when
$\Omega(t)\epsilon(t)$ does not tend to zero, and by itself it does not imply a
linear proper-distance or area-excess ratio.
\end{lemma}

\begin{proof}
Apply \Cref{lem:varying-gap-denominator-main} without the small-gap
linearization at $\rho_\varphi(t)$ and $\rho_{\mathrm{mem}}(t)$, then substitute
the exact homogeneous ratio law in \Cref{ass:small-shell-main}.
\end{proof}

\begin{corollary}[Abstract small-gap linearization of the near-horizon collar ratio]\label{thm:horizon-main}
Assume the sampled classical interval instrument of
\Cref{prop:instrument-kernels-main}, with independent Poisson attempts and
fresh interval ancillas, and impose only the modelling closure
$\Gamma=\Gamma_+(x,\Omega)$. Assume the small-shell horizon-family hypothesis in
\Cref{ass:small-shell-main}, so that the homogeneous shell pair comes from a
fixed calibrated response function $R_\Omega$, and the dimensionless
small-gap condition
\[
\eta(t):=\Omega(t)\epsilon(t)\longrightarrow0 .
\]
Suppose that the static region ends at a nonextremal Hartle--Hawking horizon, so that $\beta_\infty=2\pi/\kappa_{\mathrm H}$ and, in proper distance $\rho$ from the horizon,
\begin{equation}\label{eq:horizon-main}
N(\rho)=\kappa_{\mathrm H}\rho+O(\rho^3)
\qquad(\rho\downarrow0).
\end{equation}
Then \Cref{lem:exact-collar-ratio-main} linearizes, and as
$\epsilon(t)\downarrow0$,
\begin{equation}\label{eq:rho-ratio-main}
\frac{\rho_{\varphi}(t)}{\rho_{\mathrm{mem}}(t)}
=
C_D(t)\Bigl(1+O(\eta(t))+O(\epsilon(t)^2)\Bigr).
\end{equation}
In Schwarzschild, if $A_H=16\pi M^2$ is the horizon area and $A_{\varphi}(t)$ and $A_{\mathrm{mem}}(t)$ are the areas of the shell spheres, then
\begin{equation}\label{eq:area-ratio-main}
\frac{A_{\varphi}(t)-A_H}{A_{\mathrm{mem}}(t)-A_H}
=
C_D(t)^2\Bigl(1+O(\eta(t))+O(\epsilon(t)^2)\Bigr).
\end{equation}
If the detector gap is fixed, or more generally uniformly bounded, these two
estimates reduce to the simpler $O(\epsilon(t))$ relative-error form.
This is the small-gap linearization of
\Cref{lem:exact-collar-ratio-main}; without $\eta(t)\to0$ only the exact
nonlinear collar identity remains. \Cref{lem:high-gap-obstruction-main} gives
an explicit regime where the linear ratio fails.
\end{corollary}

The preceding warning is part of the logical content of the subsection: the
proper-distance ratio \eqref{eq:rho-ratio-main} is not a consequence of the
shell law alone and not a consequence of the exact denominator ratio alone.
Without the extra condition $\eta(t)=\Omega(t)\epsilon(t)\to0$, one has only
\[
C_D(t)=
\frac{
e^{2\pi\Omega(t)\rho_\varphi(t)(1+O(\epsilon(t)^2))}-1}
{
e^{2\pi\Omega(t)\rho_{\mathrm{mem}}(t)(1+O(\epsilon(t)^2))}-1},
\]
which is precisely \Cref{lem:exact-collar-ratio-main}. That formula is already
the full near-horizon information available in the high-gap regime.

\begin{proof}
All estimates below are along the small-shell family, and the constants in the
$O(\cdot)$ terms are uniform after restricting to a sufficiently small normal
collar. Applying \Cref{lem:varying-gap-denominator-main} at
$\rho_\varphi(t)$ and $\rho_{\mathrm{mem}}(t)$ gives
\[
\begin{aligned}
\frac{e^{\beta_\infty\Omega(t) N(\rho_\varphi(t))}-1}
     {e^{\beta_\infty\Omega(t) N(\rho_{\mathrm{mem}}(t))}-1}
&=
\frac{\rho_\varphi(t)}{\rho_{\mathrm{mem}}(t)}
\frac{1+O(\Omega(t)\rho_\varphi(t))+O(\rho_\varphi(t)^2)}
     {1+O(\Omega(t)\rho_{\mathrm{mem}}(t))+O(\rho_{\mathrm{mem}}(t)^2)}  \\
&=
\frac{\rho_\varphi(t)}{\rho_{\mathrm{mem}}(t)}
\Bigl(1+O(\eta(t))+O(\epsilon(t)^2)\Bigr).
\end{aligned}
\]
The exact homogeneous ratio law says that the left-hand side is $C_D(t)$.
Hence
\[
C_D(t)=
\frac{\rho_\varphi(t)}{\rho_{\mathrm{mem}}(t)}
\bigl(1+\delta(t)\bigr),
\qquad
\delta(t)=O(\eta(t))+O(\epsilon(t)^2).
\]
After restricting to $\epsilon(t)$ so small that $|\delta(t)|<1/2$,
$(1+\delta(t))^{-1}=1+O(\eta(t))+O(\epsilon(t)^2)$, which proves
\eqref{eq:rho-ratio-main}.

In Schwarzschild, $\rho$ is not the areal-radius increment. With $r_H=2M$,
\[
\rho=2\sqrt{r_H(r-r_H)}+O((r-r_H)^{3/2}),
\qquad
r-r_H=\frac{\rho^2}{4r_H}+O(\rho^4).
\]
Therefore
\[
A(r)-A_H=4\pi(r^2-r_H^2)=2\pi\rho^2+O(\rho^4)
=2\pi\rho^2\bigl(1+O(\rho^2)\bigr).
\]
Applying this at the two shell radii gives
\[
\frac{A_{\varphi}(t)-A_H}{A_{\mathrm{mem}}(t)-A_H}
=
\left(\frac{\rho_\varphi(t)}{\rho_{\mathrm{mem}}(t)}\right)^2
\frac{1+O(\rho_\varphi(t)^2)}
     {1+O(\rho_{\mathrm{mem}}(t)^2)}
=
\left(\frac{\rho_\varphi(t)}{\rho_{\mathrm{mem}}(t)}\right)^2
\bigl(1+O(\epsilon(t)^2)\bigr).
\]
Combining this with~\eqref{eq:rho-ratio-main} gives
\eqref{eq:area-ratio-main}.
\end{proof}

\begin{corollary}[Asymptotically homogeneous small-gap linearization]\label{cor:asymptotic-horizon-main}
Assume the sampled classical interval instrument of
\Cref{prop:instrument-kernels-main}, with independent Poisson attempts and
fresh interval ancillas, and impose only the modelling closure
$\Gamma=\Gamma_+(x,\Omega)$. Assume the same calibrated horizon hypotheses as in
\Cref{thm:horizon-main}, but replace exact homogeneity by the
spectral-transport setting of \Cref{cor:spectral-transport-main}. Suppose the
shell points lie on the same static radial segment, define
\[
L_\epsilon(t)=\sup_{0\le \rho\le \epsilon(t)}
\left|\partial_\rho \log \mathcal S_\rho(\Omega(t))\right|,
\]
and assume in addition that
\[
L_\epsilon(t)\,\epsilon(t)\longrightarrow0 .
\]
Then
\begin{equation}\label{eq:asymptotic-rho-ratio-main}
\frac{\rho_{\varphi}(t)}{\rho_{\mathrm{mem}}(t)}
=
C_D(t)\Bigl(1+O(\eta(t))+O(\epsilon(t)^2)+O(L_\epsilon(t)\epsilon(t))\Bigr).
\end{equation}
Thus the homogeneous small-gap linearization remains the leading term whenever
the shell separation shrinks faster than the spatial variation scale of
$\log \mathcal S_\rho(\Omega)$.
\end{corollary}

\begin{proof}
Because both shell points lie in $0\le\rho\le\epsilon(t)$ on the same radial segment, the one-dimensional mean-value estimate gives
\[
\left|
\log\frac{\mathcal S_{\rho_\varphi(t)}(\Omega(t))}
         {\mathcal S_{\rho_{\mathrm{mem}}(t)}(\Omega(t))}
\right|
\le
L_\epsilon(t)\,|\rho_\varphi(t)-\rho_{\mathrm{mem}}(t)|
\le
L_\epsilon(t)\,\epsilon(t).
\]
Hence
\[
\frac{\mathcal S_{\rho_\varphi(t)}(\Omega(t))}
     {\mathcal S_{\rho_{\mathrm{mem}}(t)}(\Omega(t))}
=
1+O(L_\epsilon(t)\epsilon(t)).
\]
Combining this with the spectral-transport identity \eqref{eq:spectral-transport-main} yields
\[
\frac{e^{\beta_\infty\Omega(t) N(\rho_\varphi(t))}-1}
     {e^{\beta_\infty\Omega(t) N(\rho_{\mathrm{mem}}(t))}-1}
=
C_D(t)\bigl(1+O(L_\epsilon(t)\epsilon(t))\bigr).
\]
Applying the proof of \Cref{thm:horizon-main} with this modified exact ratio replaces the left-hand side constant $C_D(t)$ by
$C_D(t)\bigl(1+O(L_\epsilon(t)\epsilon(t))\bigr)$, which gives
\eqref{eq:asymptotic-rho-ratio-main}.
\end{proof}

\begin{lemma}[High-gap obstruction to the linear ratio]\label{lem:high-gap-obstruction-main}
Assume the sampled classical interval instrument of
\Cref{prop:instrument-kernels-main}, with independent Poisson attempts and
fresh interval ancillas, and impose only the modelling closure
$\Gamma=\Gamma_+(x,\Omega)$. The condition $\Omega(t)\epsilon(t)\to0$ in
\Cref{thm:horizon-main} cannot be dropped. Suppose, along a small-shell
family, that
\[
\Omega(t)\epsilon(t)\to\Lambda\in(0,\infty),\qquad
\frac{\rho_\varphi(t)}{\epsilon(t)}\to a,\qquad
\frac{\rho_{\mathrm{mem}}(t)}{\epsilon(t)}\to b
\]
with $a,b\in(0,1]$. Then the exact collar ratio gives
\begin{equation}\label{eq:high-gap-limit-main}
C_D(t)\longrightarrow
\frac{e^{2\pi\Lambda a}-1}{e^{2\pi\Lambda b}-1},
\end{equation}
whereas the proper-distance ratio tends to $a/b$. These two limits are not equal
in general.
\end{lemma}

\begin{proof}
Under the displayed hypotheses the correction
$O(\epsilon(t)^2)$ in \eqref{eq:exact-collar-ratio-main} contributes
$o(1)$ to the exponent, because
$\Omega(t)\rho_\star(t)O(\epsilon(t)^2)=O(\epsilon(t)^2)$ for
$\star\in\{\varphi,\mathrm{mem}\}$. Taking the limit in
\eqref{eq:exact-collar-ratio-main} gives \eqref{eq:high-gap-limit-main}. The
last assertion follows, for example, by taking $a\ne b$ and any fixed
$\Lambda>0$: the function $(e^{2\pi\Lambda x}-1)/x$ is not constant on
$(0,\infty)$.
\end{proof}

\Cref{lem:exact-collar-ratio-main} is the valid near-horizon statement before
linearization. \Cref{thm:horizon-main} is only the additional small-gap
linearization of that lemma, obtained in the dimensionless regime
$\Omega(t)\epsilon(t)\to0$, and \Cref{lem:high-gap-obstruction-main} shows why
that hypothesis is essential. A fixed shell pair away from the horizon still
satisfies the exact ratio law proved in \Cref{thm:ratio-main}, but the
near-horizon expansion does not by itself supply proper-distance or area-excess
asymptotics; in the high-gap regime it supplies only the exact collar ratio.
Within the dimensionless small-gap small-shell regime, the leading ratio is
independent of the common response amplitude and depends on the detector only
through the known readout constant $C_D(t)$.

A particularly informative regime is the fast-sampling limit.
\begin{lemma}[Fast-sampling critical rate]\label{lem:fast-gammac-main}
For fixed $\kappa_{\mathrm r}>0$ and $\Delta\tau\downarrow0$, equivalently $s=\kappa_{\mathrm r}\Delta\tau\downarrow0$,
\begin{equation}\label{eq:fast-sampling-main}
\Gamma_c(\kappa_{\mathrm r},\Delta\tau)
=
\frac{L+\log L+o(1)}{\Delta\tau},
\qquad
L=\log\frac{1}{s}.
\end{equation}
\end{lemma}

\begin{proof}
Write $x=\Gamma_c/\kappa_{\mathrm r}$. The critical equation~\eqref{eq:gamma-c-main} is
\[
s=\frac{\log x}{x-1}.
\]
For $s\downarrow0$ the non-removable solution has $x\to\infty$. Put $x=(L+u)/s$, with $L=\log(1/s)$. Then
\[
\log x
=\log(L+u)-\log s
=L+\log(L+u),
\]
while
\[
s(x-1)=L+u-s.
\]
The equation $\log x=s(x-1)$ therefore becomes
\[
u=\log(L+u)+s .
\]
Since the right-hand side is $O(\log L)$, this identity first gives $u=O(\log L)$, and substituting back gives
$u=\log L+o(1)$. Hence
\[
\Gamma_c=\kappa_{\mathrm r}x
=\frac{s}{\Delta\tau}\frac{L+\log L+o(1)}{s}
=\frac{L+\log L+o(1)}{\Delta\tau}.
\]
\end{proof}

In the calibrated homogeneous near-horizon setting with fixed detector gap and response amplitude, both target rates diverge as $\Delta\tau\downarrow0$, so the corresponding lapse levels tend to zero. Thus the fast-sampling scaling produces a concrete algebraic small-shell family inside the homogeneous specialization of \Cref{lem:abstract-homogeneous-specialization-main}, and it satisfies the dimensionless small-gap condition in Corollary~\ref{thm:horizon-main}. Combining that small-gap linearization with Lemma~\ref{lem:fast-gammac-main} gives
\begin{equation}\label{eq:fast-ratio-main}
\frac{\rho_{\mathrm{mem}}}{\rho_{\varphi}}
\sim
\frac{\log\varphi}{L+\log L}
\longrightarrow0.
\end{equation}
Thus the memory-transition shell collapses toward the horizon parametrically faster than the ideal-benchmark shell. Within this calibrated specialization, the shell pair therefore resolves a genuine hierarchy of stretched-horizon scales rather than two nearby realizations of the same thermal profile.

\subsection{Static spherical ratio identifiability}

The shell pair also carries fixed-radius ratio inverse information in static spherical settings,
but there are two distinct uses. A single shell pair in a fixed Schwarzschild
response model gives explicit algebra once the detector constants
$\beta_\varphi(D)$ and $\beta_{\mathrm{mem}}(D)$ have been calibrated; that
calculation uses the absolute response scale. The two-mode construction below
is different: it first divides within each mode and solves only the resulting
ratio equations. The common response amplitude is absent from those two
equations, but it remains an admissibility variable that must be checked after
a candidate background parameter has been found.

Thus the main local theorem in this subsection uses the conservative inverse
that the ratio equations actually justify: it fixes one radius record, inverts
the two ratio values, and keeps the full common-core admissibility conditions in
the same theorem statement. If the four radii are also varied as algebraic
parameters, the resulting ambient residual graph is recorded separately in the
appendix; that enlarged bookkeeping is not an inverse theorem from nearby shell
records. Physical common-core shell data still form a generally smaller subset,
cut out by the undivided shell equations, positivity of one common hardware
prefactor, and the exterior range of the chosen background family.

For the Schwarzschild statement, ``model-admissible'' means that the displayed radii are generated by the calibrated homogeneous shell equations with a common positive response amplitude, with $r_\varphi>2M$, $r_{\mathrm{mem}}>2M$, and positive detector constants. Equivalently, the algebraic right-hand sides below must give $\beta_\infty^2>0$, $M>0$, and lapse values in $(0,1)$ at both radii. This condition is not cosmetic: arbitrary pairs of positive radii and detector constants need not describe two exterior Schwarzschild shell locations.

\begin{proposition}[Schwarzschild algebra for a fixed calibrated homogeneous $R_\Omega$]\label{prop:schw-main}
Assume the sampled classical interval instrument of
\Cref{prop:instrument-kernels-main}, with independent Poisson attempts and
fresh interval ancillas, and impose only the modelling closure
$\Gamma=\Gamma_+(x,\Omega)$. Assume the calibrated homogeneous spectral-weight
regime, so that the fixed calibrated response function $R_\Omega$ is
homogeneous, and define the detector constants
\begin{equation}\label{eq:beta-constants-main}
\beta_{\varphi}(D)=
\frac{1}{\Omega}
\log\!\left(1+\frac{\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)\Delta\tau}{\log\varphi}\right),
\qquad
\beta_{\mathrm{mem}}(D)=
\frac{1}{\Omega}
\log\!\left(1+\frac{\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)}
{\Gamma_c(\kappa_{\mathrm r},\Delta\tau)}\right).
\end{equation}
If model-admissible shell data generated by the same response amplitude occur at distinct radii $r_{\varphi}\neq r_{\mathrm{mem}}$ in Schwarzschild, then
\begin{equation}\label{eq:beta-infty-main}
\beta_\infty^2=
\frac{\beta_{\varphi}(D)^2r_{\varphi}-\beta_{\mathrm{mem}}(D)^2r_{\mathrm{mem}}}
{r_{\varphi}-r_{\mathrm{mem}}},
\end{equation}
and
\begin{equation}\label{eq:mass-main}
M=
\frac{r_{\varphi}r_{\mathrm{mem}}\bigl(\beta_{\varphi}(D)^2-\beta_{\mathrm{mem}}(D)^2\bigr)}
{2\bigl(\beta_{\varphi}(D)^2r_{\varphi}-\beta_{\mathrm{mem}}(D)^2r_{\mathrm{mem}}\bigr)}.
\end{equation}
\end{proposition}

\begin{proof}
In Schwarzschild,
\[
N(r)^2=1-\frac{2M}{r}.
\]
The two homogeneous shell equations are
\[
\frac{\beta_{\varphi}(D)^2}{\beta_\infty^2}=1-\frac{2M}{r_\varphi},
\qquad
\frac{\beta_{\mathrm{mem}}(D)^2}{\beta_\infty^2}=1-\frac{2M}{r_{\mathrm{mem}}}.
\]
Equivalently,
\[
\beta_\infty^2-\frac{2M\beta_\infty^2}{r_\varphi}=\beta_{\varphi}(D)^2,
\qquad
\beta_\infty^2-\frac{2M\beta_\infty^2}{r_{\mathrm{mem}}}=\beta_{\mathrm{mem}}(D)^2.
\]
This is a nonsingular linear system for $\beta_\infty^2$ and $2M\beta_\infty^2$ when $r_\varphi\neq r_{\mathrm{mem}}$. Solving it gives~\eqref{eq:beta-infty-main} and~\eqref{eq:mass-main}. The model-admissibility hypothesis is what guarantees that the displayed right-hand sides are positive and correspond to a shell outside the horizon; arbitrary pairs of radii and detector constants need not do so.
\end{proof}

Proposition~\ref{prop:schw-main} is exact algebra within the calibrated
response model. It is not amplitude-free, because
$\beta_\varphi(D)$ and $\beta_{\mathrm{mem}}(D)$ still contain the common
response amplitude.

A related construction extends beyond Schwarzschild in a local ratio form. Each
detector operating mode supplies one shell pair and hence one scalar ratio
equation. Two genuinely distinct modes can provide two equations for a
two-parameter background family. The data objects used in those equations are
defined explicitly as follows. There are two different maps. The first is the
abstract fixed-radius ratio map
\[
\mathcal R_\eta:\Theta\to(0,\infty)^2,
\]
obtained after dividing within each detector mode. The second is the full
fixed-setting data map
\[
\mathcal F_\eta:\Theta\times(0,\infty)\to(0,\infty)^4,
\]
which keeps the common unknown amplitude and the four undivided shell
equations. The inverse theorem below concerns only the abstract ratio map and
its local sensitivity to nearby ratio data in $(0,\infty)^2$; a physical shell
record at the same fixed radii is encoded by the separate four-equation map
$\mathcal F_\eta$, which for fixed detector settings forces the single target
vector selected by those settings.

For later reference, the three data objects are:
\[
\begin{aligned}
\eta &\quad \text{fixed radius record},\\
y &\in (0,\infty)^2 \quad \text{ratio datum at those fixed radii},\\
b &\in (0,\infty)^4 \quad \text{full fixed-setting shell target}.
\end{aligned}
\]
In the theorem below, the detector settings are held fixed, so $C$ and $b$ are
held fixed as well. Accordingly, a nearby point $y$ means nearby measured or
estimated denominator quotients for those same settings and that same radius
record $\eta_*$. It does not mean that the detector settings have been varied;
changing the settings would replace the constants $C_m$ and the target vector
$b$, hence would define a different inverse problem.

A two-mode radius record is written
\[
\eta=(\eta_{\varphi,1},\eta_{{\mathrm{mem}},1},
      \eta_{\varphi,2},\eta_{{\mathrm{mem}},2}).
\]
Here $J$ denotes the radial-coordinate interval for the chosen static spherical
coordinate convention. When radius records are varied only as ambient algebraic
variables, they are taken in an open set $\mathcal D\subset J^4$ on which the
response factors below are well defined, and the corresponding algebraic ratio
data lie in an open set $\mathcal E\subset\mathcal D\times(0,\infty)^2$.
For each $m$, the ideal-benchmark radius is obtained from ideal one-step
blanking with sampling interval $\Delta\tau_m$, while the memory radius is
obtained from the sampled one-recovery-opportunity boundary-reset mode and the
same sampling interval and recovery rate $\kappa_{{\mathrm r},m}$. The detector
settings supply the known physical ratio targets
\begin{equation}\label{eq:two-mode-ratio-constants-main}
C_m=\frac{\Delta\tau_m\,\Gamma_c(\kappa_{{\mathrm r},m},\Delta\tau_m)}{\log\varphi},
\qquad m=1,2.
\end{equation}

For a candidate background parameter $\theta$, define the two radius-dependent ratio observables
\begin{equation}\label{eq:two-mode-ratio-functions-main}
R_m(\theta;\eta)=
\frac{e^{\beta_\infty(\theta)\Omega N(\eta_{\varphi,m};\theta)}-1}
     {e^{\beta_\infty(\theta)\Omega N(\eta_{{\mathrm{mem}},m};\theta)}-1},
\qquad
\mathcal R_\eta(\theta)=(R_1(\theta;\eta),R_2(\theta;\eta)).
\end{equation}
The physical residuals are
\[
H_m(\theta,\eta)=R_m(\theta;\eta)-C_m,\qquad
H(\theta,\eta)=\mathcal R_\eta(\theta)-C,
\]
where $C=(C_1,C_2)$.
If one physically estimates the two denominator quotients at the fixed radii
and detector settings, the resulting datum is a point
$\widehat y\in(0,\infty)^2$ approximating $\mathcal R_\eta(\theta)$. Thus a
nearby point of ratio space means a pair of nearby measured denominator
quotients for those same fixed radii and readout settings, not a perturbed
radius quadruple. Equivalently, the local inverse theorem below takes as input
estimated ratio data
\[
\widehat y\approx\mathcal R_{\eta_*}(\theta_*)
\]
at one fixed radius record $\eta_*$ and one fixed pair of detector settings.
When those settings are kept fixed, exact full shell data are not a nearby
two-parameter family in ratio space: they still correspond only to the single
physical target $C$ in the ratio projection of the four-equation map
$\mathcal F_{\eta_*}$.

The full shell bookkeeping is separate from the two ratio equations. Let
\[
\mathcal A=\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)
\]
denote the unknown homogeneous hardware prefactor. For a candidate background parameter $\theta$, radius record $\eta$, and mode $m$, the undivided shell equations infer
\[
\mathcal A_{\varphi,m}(\theta;\eta)=
\frac{\log\varphi}{\Delta\tau_m}
\bigl(e^{\beta_\infty(\theta)\Omega N(\eta_{\varphi,m};\theta)}-1\bigr),
\]
and
\[
\mathcal A_{{\mathrm{mem}},m}(\theta;\eta)=
\Gamma_c(\kappa_{{\mathrm r},m},\Delta\tau_m)
\bigl(e^{\beta_\infty(\theta)\Omega N(\eta_{{\mathrm{mem}},m};\theta)}-1\bigr).
\]
Equivalently, for fixed $\eta$ the full undivided shell data are encoded by
\begin{equation}\label{eq:full-fixed-radius-map-main}
\mathcal F_\eta(\theta,\mathcal A)=
\begin{pmatrix}
\dfrac{\mathcal A\Delta\tau_1}
      {e^{\beta_\infty(\theta)\Omega N(\eta_{\varphi,1};\theta)}-1} \\
\dfrac{\mathcal A}
      {e^{\beta_\infty(\theta)\Omega N(\eta_{{\mathrm{mem}},1};\theta)}-1} \\
\dfrac{\mathcal A\Delta\tau_2}
      {e^{\beta_\infty(\theta)\Omega N(\eta_{\varphi,2};\theta)}-1} \\
\dfrac{\mathcal A}
      {e^{\beta_\infty(\theta)\Omega N(\eta_{{\mathrm{mem}},2};\theta)}-1}
\end{pmatrix},
\end{equation}
with target vector
\begin{equation}\label{eq:full-fixed-radius-target-main}
b=
\bigl(
\log\varphi,\,
\Gamma_c(\kappa_{{\mathrm r},1},\Delta\tau_1),\,
\log\varphi,\,
\Gamma_c(\kappa_{{\mathrm r},2},\Delta\tau_2)
\bigr).
\end{equation}
Thus $\eta$ is a full common-core shell record exactly when there exist
$(\theta,\mathcal A)$ with $\mathcal A>0$ such that
$\mathcal F_\eta(\theta,\mathcal A)=b$. Dividing the first pair and the second
pair of entries in \eqref{eq:full-fixed-radius-map-main} eliminates
$\mathcal A$ and leaves the two ratio equations
$\mathcal R_\eta(\theta)=C$.

\begin{lemma}[Ratio residuals and full common-core data]\label{lem:ratio-admissibility-main}
Assume the sampled classical interval instrument of
\Cref{prop:instrument-kernels-main}, with independent Poisson attempts and
fresh interval ancillas, and impose only the modelling closure
$\Gamma=\Gamma_+(x,\Omega)$. For a fixed physical mode $m$, candidate parameter $\theta$, and radius record
$\eta$, the equation $H_m(\theta,\eta)=0$ is equivalent to the within-mode
amplitude equality
\[
\mathcal A_{\varphi,m}(\theta;\eta)=\mathcal A_{{\mathrm{mem}},m}(\theta;\eta).
\]
The pair of ratio equations $H(\theta,\eta)=0$ contains no unknown
$\mathcal A$. A radius record is an exact full common-core shell record at the
same $\theta$ only if, in addition, the four estimates
\[
\mathcal A_{\varphi,1},\quad \mathcal A_{{\mathrm{mem}},1},\quad
\mathcal A_{\varphi,2},\quad \mathcal A_{{\mathrm{mem}},2}
\]
are all positive and equal to one common value $\mathcal A>0$, and the four
radii lie in the exterior coordinate ranges of the background family.
Equivalently, the abstract ratio condition $H(\theta,\eta)=0$ becomes a
physical fixed-radius shell solution only after the full map in
\eqref{eq:full-fixed-radius-map-main} hits the target vector
\eqref{eq:full-fixed-radius-target-main}. These extra conditions are not
consequences of the ratio equations. In particular, once the detector settings
and hence the target vector $b$ are fixed, every exact full common-core shell
solution satisfies
\[
\mathcal R_\eta(\theta)=C.
\]
Therefore a nearby point $y\neq C$ has no exact interpretation as full
fixed-setting shell data unless one explicitly changes the detector targets
themselves or treats $y$ only as an estimated ratio datum in the abstract ratio
space.
\end{lemma}

\begin{proof}
By the definitions above,
\[
\frac{\mathcal A_{\varphi,m}(\theta;\eta)}
{\mathcal A_{{\mathrm{mem}},m}(\theta;\eta)}
=
\frac{e^{\beta_\infty(\theta)\Omega N(\eta_{\varphi,m};\theta)}-1}
     {e^{\beta_\infty(\theta)\Omega N(\eta_{{\mathrm{mem}},m};\theta)}-1}
\frac{\log\varphi}
{\Delta\tau_m\,\Gamma_c(\kappa_{{\mathrm r},m},\Delta\tau_m)}
=\frac{H_m(\theta,\eta)+C_m}{C_m}.
\]
Thus $H_m(\theta,\eta)=0$ is exactly the equality of the two within-mode
amplitude estimates. The formula for $H_m$ contains only the radius record, the
background family, the gap, and the readout constants; the common prefactor
$\mathcal A$ has cancelled. Equality across the two modes, positivity, and the
exterior-range restrictions are separate requirements. They cannot be inferred
from $H(\theta,\eta)=0$, because the ratio operation has deliberately discarded
the absolute scale and does not encode the allowed radial range. In
\eqref{eq:full-fixed-radius-map-main}, that discarded scale is exactly the
unknown common amplitude $\mathcal A$. Conversely, if
$\mathcal F_\eta(\theta,\mathcal A)=b$ for some $\mathcal A>0$, then dividing
the first and second entries and also the third and fourth entries recovers
$R_m(\theta;\eta)=C_m$ for $m=1,2$, hence $\mathcal R_\eta(\theta)=C$.
\end{proof}

\begin{theorem}[Local estimator branch for the abstract fixed-radius ratio map of a fixed calibrated $R_\Omega$]\label{thm:two-mode-main}
Assume the sampled classical interval instrument of
\Cref{prop:instrument-kernels-main}, with independent Poisson attempts and
fresh interval ancillas, and impose only the modelling closure
$\Gamma=\Gamma_+(x,\Omega)$. Let $\Theta\subset\RR^2$ be open, let $J$ be the radial interval used by a fixed
static spherical coordinate chart, and suppose
\[
N:J\times\Theta\to(0,\infty),
\qquad
\beta_\infty:\Theta\to(0,\infty)
\]
are $C^1$. Choose an exact radius quadruple
\[
\eta_*=(r_{\varphi,1},r_{{\mathrm{mem}},1},
        r_{\varphi,2},r_{{\mathrm{mem}},2})\in J^4
\]
on which the ratio functions in \eqref{eq:two-mode-ratio-functions-main} are
defined. The detector settings are two readout settings of one field-facing
core at gap $\Omega$, with a positive homogeneous prefactor
		$\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)$ shared across both settings but not assumed
known; equivalently, the ratio map is built from a fixed calibrated
homogeneous response function $R_\Omega$.

Let $C=(C_1,C_2)$ be the ratio targets from
\eqref{eq:two-mode-ratio-constants-main}, and set
\[
\mathcal R_*=\mathcal R_{\eta_*}.
\]
Assume that the exact radius record is produced at
$\theta_*\in\Theta$ by admissible full common-core shell data with one positive
shared amplitude $\mathcal A_*>0$, so that
\[
\mathcal R_*(\theta_*)=C
\]
and
\[
\mathcal A_{\varphi,1}(\theta_*;\eta_*)
=\mathcal A_{{\mathrm{mem}},1}(\theta_*;\eta_*)
=\mathcal A_{\varphi,2}(\theta_*;\eta_*)
=\mathcal A_{{\mathrm{mem}},2}(\theta_*;\eta_*)
=\mathcal A_*>0,
\]
with all four radii lying in the exterior coordinate range of the background
family at $\theta_*$.
If
\begin{equation}\label{eq:two-mode-ratio-jacobian-main}
\det D_\theta\mathcal R_{\eta_*}(\theta_*)\neq0,
\end{equation}
then there are neighbourhoods $U\subset\Theta$ of $\theta_*$ and
$Y\subset(0,\infty)^2$ of $C$, together with a unique $C^1$ branch
\[
\mathcal I_{\eta_*}:Y\to U,\qquad \mathcal I_{\eta_*}(C)=\theta_*,
\]
such that
\[
\mathcal R_*(\mathcal I_{\eta_*}(y))=y\qquad(y\in Y).
\]
Here $y$ is a point of ratio space $Y\subset(0,\infty)^2$, namely a nearby
value of the abstract fixed-radius ratio map
$\mathcal R_{\eta_*}:\Theta\to(0,\infty)^2$; it is not an arbitrary nearby
radius quadruple. Physically, $y$ represents the pair of denominator-quotient
values estimated from the two detector modes while the same four radii in
$\eta_*$ and the same readout settings are held fixed. For those fixed detector
settings, the ratio target $C=(C_1,C_2)$ and the full undivided target vector
$b$ are fixed as well. The branch $\mathcal I_{\eta_*}$ therefore solves only
the abstract ratio equations and should be read only as an estimator or
sensitivity map from nearby observed denominator quotients $y$ to background
parameters. It is not, by itself, a theorem that nearby $y\neq C$ are exact
fixed-setting shell records. If one instead changes detector settings, then the
physical target constants $C_m$ and the target vector $b$ change, so one is no
longer in the same fixed-setting estimator problem.

Moreover, shrink $U$ if necessary so that the fixed radii in $\eta_*$ remain
exterior-admissible for every $\theta\in U$. Then define
\[
\widetilde{\mathcal A}_{\varphi,m}(y)
=\mathcal A_{\varphi,m}(\mathcal I_{\eta_*}(y);\eta_*),
\qquad
\widetilde{\mathcal A}_{{\mathrm{mem}},m}(y)
=\mathcal A_{{\mathrm{mem}},m}(\mathcal I_{\eta_*}(y);\eta_*),
\qquad m=1,2.
\]
For any ratio datum $y\in Y$, the recovered parameter
$\theta=\mathcal I_{\eta_*}(y)$ is the unique local background parameter whose
abstract ratio map agrees with $y$, and the four pulled-back amplitudes above
record how that inferred parameter enters the undivided shell equations at the
fixed radius record. At the physical point $y=C$, they coincide and recover the
shared response amplitude $\mathcal A_*$. Moreover, if $\theta\in U$ and
$\mathcal R_*(\theta)=y\in Y$, then $\theta=\mathcal I_{\eta_*}(y)$. Thus the
concrete increment is a local inverse for the abstract fixed-radius ratio map,
to be used as an estimator or sensitivity statement for nearby estimated ratio
data at fixed $\eta_*$.
\end{theorem}

\begin{proof}
With $\eta_*$ fixed, the stated regularity makes
$\mathcal R_*:\Theta\to\RR^2$ a $C^1$ map. Exact admissible full common-core shell data
satisfy the two homogeneous ratio equations by Corollary~\ref{thm:ratio-main},
hence $\mathcal R_*(\theta_*)=C$.

To make the independence condition explicit, write
\[
B_{\varphi,m}(\theta)=\beta_\infty(\theta)\Omega N(r_{\varphi,m};\theta),
\qquad
B_{\mathrm{mem},m}(\theta)=\beta_\infty(\theta)\Omega N(r_{{\mathrm{mem}},m};\theta).
\]
Then
\[
d_\theta R_m=
\frac{
e^{B_{\varphi,m}}\,d_\theta B_{\varphi,m}\,(e^{B_{\mathrm{mem},m}}-1)
-(e^{B_{\varphi,m}}-1)e^{B_{\mathrm{mem},m}}\,d_\theta B_{\mathrm{mem},m}
}
{(e^{B_{\mathrm{mem},m}}-1)^2}.
\]
Thus the determinant condition is the concrete requirement that the two row
covectors supplied by the two fixed-radius ratio equations are not collinear at
$\theta_*$. It can fail because the detector modes are redundant, because the
two shell pairs sample the background family redundantly, or both.

Apply the ordinary inverse function theorem to the abstract fixed-radius ratio
map $\mathcal R_*:\Theta\to\RR^2$ at $\theta_*$. The determinant in
\eqref{eq:two-mode-ratio-jacobian-main} is exactly the Jacobian determinant of
this map. The inverse function theorem gives the inverse branch
$\mathcal I_{\eta_*}$ and the stated uniqueness. The amplitude estimates
\[
\theta\longmapsto \mathcal A_{\varphi,m}(\theta;\eta_*),\qquad
\theta\longmapsto \mathcal A_{{\mathrm{mem}},m}(\theta;\eta_*),
\]
are $C^1$ because they are compositions of the $C^1$ maps
$N,\beta_\infty$ with the smooth exponential function. Hence the pulled-back
fixed-record amplitudes
$\widetilde{\mathcal A}_{\varphi,m}$ and
$\widetilde{\mathcal A}_{{\mathrm{mem}},m}$ are $C^1$ on $Y$.

Because the given record is generated by admissible full common-core shell
data at $\theta_*$, the four amplitudes agree there and equal one positive
value $\mathcal A_*$. Exterior admissibility at the fixed radii is an open
condition in $\theta$, so $U$ can be shrunk so that the fixed record remains in
the allowed coordinate range throughout $U$.
\end{proof}

\begin{corollary}[Delta-method covariance for estimated ratio data]\label{cor:two-mode-delta-main}
Assume the hypotheses of \Cref{thm:two-mode-main}. Let $\widehat y_N\in Y$ be
estimated ratio data such that
\[
\sqrt{N}\,(\widehat y_N-C)\Longrightarrow \mathcal N(0,\Sigma_y)
\]
for some positive semidefinite $2\times2$ matrix $\Sigma_y$. Set
\[
\widehat\theta_N=\mathcal I_{\eta_*}(\widehat y_N).
\]
Then
\[
\sqrt{N}\,(\widehat\theta_N-\theta_*)
\Longrightarrow
\mathcal N(0,\Sigma_\theta),
\]
with
\begin{equation}\label{eq:delta-covariance-main}
\Sigma_\theta
=D\mathcal I_{\eta_*}(C)\,\Sigma_y\,D\mathcal I_{\eta_*}(C)^\top
=\bigl[D_\theta\mathcal R_{\eta_*}(\theta_*)\bigr]^{-1}
\Sigma_y
\bigl[D_\theta\mathcal R_{\eta_*}(\theta_*)\bigr]^{-{\top}}.
\end{equation}
If $\widehat\Sigma_{y,N}\to\Sigma_y$ in probability, then the plug-in estimator
\[
\widehat\Sigma_{\theta,N}
=D\mathcal I_{\eta_*}(\widehat y_N)\,
\widehat\Sigma_{y,N}\,
D\mathcal I_{\eta_*}(\widehat y_N)^\top
\]
converges in probability to $\Sigma_\theta$.
\end{corollary}

\begin{proof}
By \Cref{thm:two-mode-main}, the inverse branch $\mathcal I_{\eta_*}$ is $C^1$
on $Y$. Therefore the multivariate delta method applied at $C$ gives
\[
\sqrt{N}\,(\mathcal I_{\eta_*}(\widehat y_N)-\mathcal I_{\eta_*}(C))
\Longrightarrow
\mathcal N\!\bigl(0,
D\mathcal I_{\eta_*}(C)\Sigma_y D\mathcal I_{\eta_*}(C)^\top\bigr).
\]
Since $\mathcal I_{\eta_*}(C)=\theta_*$, this is the first claim. The identity
\[
D\mathcal I_{\eta_*}(C)=\bigl[D_\theta\mathcal R_{\eta_*}(\theta_*)\bigr]^{-1}
\]
is the chain rule for the local inverse branch. The plug-in consistency follows
from the continuity of $D\mathcal I_{\eta_*}$ on $Y$ and Slutsky's theorem.
\end{proof}

\begin{theorem}[Fixed-setting admissibility for the undivided shell equations]\label{thm:fixed-radius-admissibility-main}
Assume the hypotheses of \Cref{thm:two-mode-main}, and keep the same fixed
radius record $\eta_*$, ratio target $C$, and undivided target vector $b$ from
\eqref{eq:full-fixed-radius-target-main}. Then exact full common-core shell
data at those fixed detector settings occur only at the physical ratio target
\[
y=C.
\]
More precisely, if $y\in Y$ and there exists $\mathcal A>0$ such that
\[
\mathcal F_{\eta_*}(\mathcal I_{\eta_*}(y),\mathcal A)=b,
\]
then necessarily $y=C$. Conversely, at $y=C$ one has
\[
\mathcal F_{\eta_*}(\theta_*,\mathcal A_*)=b.
\]
Thus nearby values $y\neq C$ are not exact fixed-setting shell records; they
are admissible only as estimated ratio data for the estimator branch of
\Cref{thm:two-mode-main}.

To test fixed-setting admissibility, equip $\mathbb R^4$ with the Euclidean
norm and define
\begin{equation}\label{eq:full-shell-misfit-main}
\mathcal E(y)=
\inf_{\mathcal A>0}
\bigl\|\mathcal F_{\eta_*}(\mathcal I_{\eta_*}(y),\mathcal A)-b\bigr\|^2,
\qquad y\in Y.
\end{equation}
If
\[
v(y)=\mathcal F_{\eta_*}(\mathcal I_{\eta_*}(y),1)\in(0,\infty)^4,
\]
then the unique minimizing amplitude is
\[
\mathcal A_{\mathrm{LS}}(y)=\frac{\langle v(y),b\rangle}{\|v(y)\|^2}>0,
\]
and
\[
\mathcal E(y)=
\bigl\|\mathcal A_{\mathrm{LS}}(y)\,v(y)-b\bigr\|^2.
\]
In particular, $\mathcal E(C)=0$, and on the neighbourhood $Y$ one has
\[
\mathcal E(y)=0\quad\Longleftrightarrow\quad y=C.
\]
If one wants exact shell records for nearby ratio values, the detector targets
must vary with $y$, so the exact equation becomes
$\mathcal F_{\eta_*}(\mathcal I_{\eta_*}(y),\mathcal A(y))=b(y)$ rather than
the fixed-target equation with $b$.
\end{theorem}

\begin{proof}
For any $y\in Y$, set $\theta=\mathcal I_{\eta_*}(y)$. Then
$\mathcal R_{\eta_*}(\theta)=y$ by \Cref{thm:two-mode-main}. If
$\mathcal F_{\eta_*}(\theta,\mathcal A)=b$ for some $\mathcal A>0$, then
\Cref{lem:ratio-admissibility-main} implies
$\mathcal R_{\eta_*}(\theta)=C$, hence $y=C$. This proves the necessity of the
physical target value. The converse at $y=C$ is part of the exact shell-data
hypothesis in \Cref{thm:two-mode-main}, namely
$\mathcal F_{\eta_*}(\theta_*,\mathcal A_*)=b$.

Now fix $y\in Y$ and write $v(y)=\mathcal F_{\eta_*}(\mathcal I_{\eta_*}(y),1)$.
Since every entry of $v(y)$ is positive,
\[
\mathcal F_{\eta_*}(\mathcal I_{\eta_*}(y),\mathcal A)=\mathcal A\,v(y),
\]
so minimizing over $\mathcal A>0$ in \eqref{eq:full-shell-misfit-main} is the
one-dimensional least-squares problem for the positive ray generated by $v(y)$.
Its unique minimizer is the displayed coefficient
$\mathcal A_{\mathrm{LS}}(y)=\langle v(y),b\rangle/\|v(y)\|^2$, which is
strictly positive because $v(y)$ and $b$ have positive entries. Therefore
\[
\mathcal E(y)=
\bigl\|\mathcal A_{\mathrm{LS}}(y)\,v(y)-b\bigr\|^2.
\]
By construction, $\mathcal E(y)=0$ if and only if there exists
$\mathcal A>0$ such that
$\mathcal F_{\eta_*}(\mathcal I_{\eta_*}(y),\mathcal A)=b$, and the first part
of the theorem shows that this happens exactly at $y=C$ inside $Y$.
\end{proof}

\begin{remark}[Estimator meaning versus retuned exact targets]\label{rem:fixed-target-versus-retuned-main}
The interpretation of \Cref{thm:fixed-radius-admissibility-main} is strictly
fixed-setting. Once the detector settings and the radius record $\eta_*$ are
frozen, the ratio target is the single point $C=(C_1,C_2)$ and the undivided
shell target is the single vector $b$. The inverse branch
$\mathcal I_{\eta_*}(y)$ should therefore be read as an estimator or
sensitivity map for nearby observed denominator quotients $y$, while the
residual $\mathcal E(y)$ measures how far the inferred parameter is from
satisfying the exact fixed-target shell equations with one common positive
amplitude.

If one instead wants nearby $y$ to represent exact shell data, the detector
targets must vary with $y$. In the present notation the corresponding
$y$-dependent target vector is
\[
b(y)=
\left(
\log\varphi,\,
\frac{y_1\log\varphi}{\Delta\tau_1},\,
\log\varphi,\,
\frac{y_2\log\varphi}{\Delta\tau_2}
\right),
\]
because the exact within-mode quotient relation is
$y_m=\Delta\tau_m\Gamma_{{\mathrm c},m}/\log\varphi$. The exact equation in
that retuned problem is then
\[
\mathcal F_{\eta_*}(\mathcal I_{\eta_*}(y),\mathcal A(y))=b(y),
\]
which is a different inverse problem from the fixed-setting estimator problem
treated in \Cref{thm:two-mode-main,thm:fixed-radius-admissibility-main}.
\end{remark}

If one instead varies the radius quadruple as an ambient algebraic parameter,
the corresponding residual graph and the restriction to full common-core shell
data are recorded separately in
\Cref{cor:ambient-ratio-graph-app-main,thm:full-shell-subset-main,cor:full-shell-restriction-main}. Those appendix statements describe the local geometry of the ratio residual equations in the enlarged data space; they are not inverse theorems from nearby shell records to background geometry.

\begin{remark}\label{rem:jacobian-main}
The Jacobian condition~\eqref{eq:two-mode-ratio-jacobian-main} is a local identifiability condition on both the detector modes and the background family, evaluated at the exact radius record and physical ratio target. It can fail because the two modes produce redundant ratio functions, or because the chosen background parametrization affects the two recorded radii in linearly dependent ways. Equality $C_1=C_2$ is therefore not a complete degeneracy criterion; it is a warning sign that the two modes may not provide independent constraints. Choosing distinct sampling and recovery parameters is necessary in practice, but the determinant in~\eqref{eq:two-mode-ratio-jacobian-main} is the actual test.
\end{remark}

\subsection{Schwarzschild worked example}\label{subsec:worked-example}

As a concrete verification of the preceding formulas, \cref{tab:worked} presents a worked example in the Schwarzschild background with $r_{\mathrm s}=2M=1$ (so $\beta_\infty=4\pi$), $\Omega=1$, and $\lambda_{\mathrm{sw}}^2\mathcal S(\Omega)=2$. Two detector modes $D_1=(\Omega,\lambda_{\mathrm{sw}},0.5,\,8.0)$ and $D_2=(\Omega,\lambda_{\mathrm{sw}},1.2,\,3.0)$ are tabulated; these modes have distinct ratio constants $C_1=\Delta\tau_1\Gamma_{c,1}/\log\varphi=0.164812$ and $C_2=\Delta\tau_2\Gamma_{c,2}/\log\varphi=0.228130$, but the actual local identifiability test is the Jacobian condition in \cref{thm:two-mode-main}. For each mode, the critical rate $\Gamma_c$ is obtained from~\eqref{eq:gamma-c-main} and the shell radii from the lapse conditions in \cref{cor:homogeneous-main}. The mode-specific ratio $C_m$ is verified to agree with the denominator quotient in~\eqref{eq:ratio-main}. The inverse formulas of \cref{prop:schw-main} are then applied to the $D_1$ shell pair alone to recover $\beta_\infty$ and $M$ to full numerical precision in the same normalization.

For the fixed-radius ratio map, the nonsingularity check in \cref{thm:two-mode-main} is first evaluated at the fixed radius record $\eta_*$ in the ambient coordinate chart $\theta=(\beta_\infty,M)$, giving
\[
\det D_\theta\mathcal R_{\eta_*}(\theta_*)\approx-2.97\times10^{-1}.
\]
This confirms the estimator branch of \cref{thm:two-mode-main} for nearby ratio estimates at fixed $\eta_*$. The corresponding asymptotic covariance is then given by \cref{cor:two-mode-delta-main}. For exact fixed detector settings, however, \cref{thm:fixed-radius-admissibility-main} shows that the full shell equations still single out the physical target value $C$, so nearby $y\neq C$ must be read as estimated ratio data rather than exact shell records. For physical Schwarzschild records the parameters are constrained by the Hartle--Hawking relation $\beta_\infty=8\pi M$ (so the family is one-dimensional), so the independent test is the corresponding constrained one-parameter sensitivity check along that curve rather than a two-parameter ambient Jacobian.

\input{generated/tab_worked_example}
