\begin{thebibliography}{47}

% === Algebraic and quantum field theory ===

\bibitem{Takesaki2003}
Takesaki, M.:
Theory of Operator Algebras II.
Springer, Berlin (2003)

\bibitem{Davies1976}
Davies, E.B.:
Quantum Theory of Open Systems.
Academic Press, London (1976)

\bibitem{Holevo2001}
Holevo, A.S.:
Statistical Structure of Quantum Theory.
Lecture Notes in Physics, vol.~461.
Springer, Berlin (2001)

\bibitem{FewsterVerch2012}
Fewster, C.J., Verch, R.:
Dynamical locality and covariance: what makes a physical theory the same in all spacetimes?
Ann.\ Henri Poincar\'e \textbf{13}, 1613--1674 (2012)

\bibitem{Fewster2020}
Fewster, C.J.:
A generally covariant measurement scheme for quantum field theory in curved spacetimes.
In: Finster, F., Giulini, D., Kleiner, J., Tolksdorf, J. (eds.)
Progress and Visions in Quantum Theory in View of Gravity, pp.~253--268.
Birkh\"auser, Cham (2020)

\bibitem{PercheMartinMartinez2022}
Perche, T.R., Mart\'in-Mart\'inez, E.:
Geometry of spacetime from quantum measurements.
Phys.\ Rev.\ D \textbf{105}(6), 066011 (2022)

% === Unruh--DeWitt detectors ===

\bibitem{Unruh1976}
Unruh, W.G.:
Notes on black-hole evaporation.
Phys.\ Rev.\ D \textbf{14}(4), 870--892 (1976)

\bibitem{DeWitt1979}
DeWitt, B.S.:
Quantum gravity: the new synthesis.
In: Hawking, S.W., Israel, W. (eds.)
General Relativity: An Einstein Centenary Survey, pp.~680--745.
Cambridge University Press, Cambridge (1979)

\bibitem{BirDav1982}
Birrell, N.D., Davies, P.C.W.:
Quantum Fields in Curved Space.
Cambridge University Press, Cambridge (1982)

\bibitem{Takagi1986}
Takagi, S.:
Vacuum noise and stress induced by uniform acceleration: Hawking--Unruh effect in Rindler manifold of arbitrary dimension.
Prog.\ Theor.\ Phys.\ Suppl.\ \textbf{88}, 1--142 (1986)

\bibitem{LoukoSatz2006}
Louko, J., Satz, A.:
How often does the Unruh--DeWitt detector click? Regularisation by a spatial profile.
Class.\ Quantum Grav.\ \textbf{23}(22), 6321--6343 (2006)

\bibitem{LoukoSatz2008}
Louko, J., Satz, A.:
Transition rate of the Unruh--DeWitt detector in curved spacetime.
Class.\ Quantum Grav.\ \textbf{25}(5), 055012 (2008)

\bibitem{CrispinoHiguchiMatsas2008}
Crispino, L.C.B., Higuchi, A., Matsas, G.E.A.:
The Unruh effect and its applications.
Rev.\ Mod.\ Phys.\ \textbf{80}(3), 787--838 (2008)

\bibitem{HodgkinsonLouko2012}
Hodgkinson, L., Louko, J.:
How often does the Unruh--DeWitt detector click beyond four dimensions?
J.\ Math.\ Phys.\ \textbf{53}(8), 082301 (2012)

\bibitem{HendersonEtAl2018}
Henderson, L.J., Hennigar, R.A., Mann, R.B., Smith, A.R.H., Zhang, J.:
Harvesting entanglement from the black hole vacuum.
Class.\ Quantum Grav.\ \textbf{35}(21), 21LT02 (2018)

\bibitem{TjoaMann2020}
Tjoa, E., Mann, R.B.:
Harvesting correlations in Schwarzschild and collapsing shell spacetimes.
J.\ High Energy Phys.\ \textbf{2020}, 155 (2020)

\bibitem{PranziniEtAl2024}
Pranzini, N., Garc\'ia-P\'erez, G., Keski-Vakkuri, E., Maniscalco, S.:
Repeated measurements on non-replicable systems and their consequences for Unruh--DeWitt detectors.
Quantum \textbf{8}, 1490 (2024)

\bibitem{LimaEtAl2023Qudit}
Lima, C., Patterson, E., Tjoa, E., Mann, R.B.:
Unruh phenomena and thermalization for qudit detectors.
Phys.\ Rev.\ D \textbf{108}(10), 105020 (2023)

% === KMS states, thermodynamics of spacetime ===

\bibitem{KayWald1991}
Kay, B.S., Wald, R.M.:
Theorems on the uniqueness and thermal properties of stationary, nonsingular, quasifree states on spacetimes with a bifurcate Killing horizon.
Phys.\ Rep.\ \textbf{207}(2), 49--136 (1991)

\bibitem{BuchholzSchlemmer2007}
Buchholz, D., Schlemmer, J.:
Local temperature in curved spacetime.
Class.\ Quantum Grav.\ \textbf{24}(7), F25--F31 (2007)

\bibitem{Solveen2012}
Solveen, C.:
Local thermal equilibrium and KMS states in curved spacetime.
Class.\ Quantum Grav.\ \textbf{29}(24), 245015 (2012)

\bibitem{BuchholzSolveen2013}
Buchholz, D., Solveen, C.:
Unruh effect and the concept of temperature.
Class.\ Quantum Grav.\ \textbf{30}(8), 085011 (2013)

\bibitem{GranseePinamontiVerch2017}
Gransee, M., Pinamonti, N., Verch, R.:
KMS-like properties of local equilibrium states in quantum field theory.
J.\ Geom.\ Phys.\ \textbf{117}, 15--35 (2017)

\bibitem{Jacobson1995}
Jacobson, T.:
Thermodynamics of spacetime: the Einstein equation of state.
Phys.\ Rev.\ Lett.\ \textbf{75}(7), 1260--1263 (1995)

% === Quantum clocks ===

\bibitem{SaleckerWigner1958}
Salecker, H., Wigner, E.P.:
Quantum limitations of the measurement of space-time distances.
Phys.\ Rev.\ \textbf{109}(2), 571--577 (1958)

\bibitem{BraunsteinCaves1994}
Braunstein, S.L., Caves, C.M.:
Statistical distance and the geometry of quantum states.
Phys.\ Rev.\ Lett.\ \textbf{72}(22), 3439--3443 (1994)

\bibitem{WangEtAl2014Metrology}
Wang, J., Tian, Z., Jing, J., Fan, H.:
Quantum metrology and estimation of Unruh effect.
Sci.\ Rep.\ \textbf{4}, 7195 (2014)

\bibitem{HaoWu2016Unruh}
Hao, Y., Wu, S.:
Quantum parameter estimation in the Unruh--DeWitt detector model.
Phys.\ Rev.\ D \textbf{94}(2), 024058 (2016)

\bibitem{LockFuentes2017}
Lock, M.P.E., Fuentes, I.:
Relativistic quantum clocks.
In: Renner, R., Stupar, S. (eds.)
Time in Physics, pp.~51--68.
Birkh\"auser, Cham (2017)

\bibitem{ErkerEtAl2017}
Erker, P., Mitchison, M.T., Silva, R., Woods, M.P., Brunner, N., Huber, M.:
Autonomous Quantum Clocks: Does Thermodynamics Limit Our Ability to Measure Time?
Phys.\ Rev.\ X \textbf{7}(3), 031022 (2017)

\bibitem{CostaEtAl2020Ramsey}
Costa, H.A.S., da Paz, I.G., Carvalho, P.R.S., Sampaio, M.:
Ramsey interferometry as a witness of acceleration radiation.
Ann.\ Phys.\ \textbf{416}, 168158 (2020)

\bibitem{DuMann2021AdS}
Du, H., Mann, R.B.:
Fisher information as a probe of spacetime structure: relativistic quantum metrology in (A)dS.
J.\ High Energy Phys.\ \textbf{2021}, 112 (2021)

\bibitem{FengZhang2022QFI}
Feng, J., Zhang, J.-J.:
Quantum Fisher information as a probe for Unruh thermality.
Phys.\ Lett.\ B \textbf{827}, 136992 (2022)

\bibitem{BassoMazieroCeleri2025}
Basso, M.L.W., Maziero, J., C\'eleri, L.C.:
Quantum detailed fluctuation theorem in curved spacetimes: the observer dependent nature of entropy production.
Phys.\ Rev.\ Lett.\ \textbf{134}(5), 050406 (2025)

% === Dead-time effects ===

\bibitem{Muller1973}
M\"uller, J.W.:
Dead-time problems.
Nucl.\ Instrum.\ Methods \textbf{112}, 47--57 (1973)

\bibitem{CantorTeich1975}
Cantor, B.I., Teich, M.C.:
Dead-time-corrected photocounting distributions for laser radiation.
J.\ Opt.\ Soc.\ Am.\ \textbf{65}(7), 786--791 (1975)

% === Symbolic dynamics ===

\bibitem{LindMarcus1995}
Lind, D., Marcus, B.:
An Introduction to Symbolic Dynamics and Coding.
Cambridge University Press, Cambridge (1995)

\bibitem{Parry1964}
Parry, W.:
Intrinsic Markov chains.
Trans.\ Amer.\ Math.\ Soc.\ \textbf{112}(1), 55--66 (1964)

% === Probability and information theory ===

\bibitem{Billingsley1995}
Billingsley, P.:
Probability and Measure, 3rd edn.
Wiley, New York (1995)

\bibitem{CoverThomas2006}
Cover, T.M., Thomas, J.A.:
Elements of Information Theory, 2nd edn.
Wiley-Interscience, Hoboken (2006)

\bibitem{DaleyVereJones2008}
Daley, D.J., Vere-Jones, D.:
An Introduction to the Theory of Point Processes, Vol.~II, 2nd edn.
Springer, New York (2008)

\bibitem{KipnisVaradhan1986}
Kipnis, C., Varadhan, S.R.S.:
Central limit theorem for additive functionals of reversible Markov chains and applications to simple exclusions.
Comm.\ Math.\ Phys.\ \textbf{104}(1), 1--19 (1986)

\bibitem{Gordin1969}
Gordin, M.I.:
The central limit theorem for stationary processes.
Soviet Math.\ Doklady \textbf{10}, 1174--1176 (1969)

% === Open quantum systems ===

\bibitem{BreuerPetruccione2002}
Breuer, H.-P., Petruccione, F.:
The Theory of Open Quantum Systems.
Oxford University Press, Oxford (2002)

\bibitem{Carmichael1993Book}
Carmichael, H.:
An Open Systems Approach to Quantum Optics.
Lecture Notes in Physics Monographs, vol.~18.
Springer, Berlin (1993)

\bibitem{WisemanMilburn2009}
Wiseman, H.M., Milburn, G.J.:
Quantum Measurement and Control.
Cambridge University Press, Cambridge (2009)

\bibitem{HanEtAl2025}
Han, S.-W., Chen, W., Chen, L., Ouyang, Z., Feng, J.:
Reliable quantum master equation of the Unruh--DeWitt detector.
Phys.\ Rev.\ D \textbf{112}(8), 085014 (2025)

\bibitem{ShallueCarroll2025}
Shallue, C.J., Carroll, S.M.:
What Hawking radiation looks like as you fall into a black hole.
Phys.\ Rev.\ D \textbf{112}(8), 085013 (2025)

\bibitem{WilkinsonEtAl2025Backaction}
Wilkinson, A.S., Parry, L.J.A., Louko, J., Unruh, W.G.:
Back-action from inertial and non-inertial Unruh--DeWitt detectors revisited in covariant perturbation theory.
arXiv:2512.16217 [gr-qc] (2025)

\end{thebibliography}
