\section{Discussion and scope}\label{sec:discussion-new}

The main tools used here are classical: companion matrices, rational generating functions, finite Hankel rank, Weierstrass reduction, Riemann--Hurwitz, local Puiseux expansions, and the Beraha--Kahane--Weiss description of limiting zeros.
What is specific to this paper is an exact worked example for the concrete quartic family \(\Pi(\lambda,y)\) together with the balanced-family reduction plus the double-root normalization (Theorem~\ref{thm:balanced-classification}) that singles this quartic out once the colliding points are fixed at $(1,0)$ and $(-1,1)$.
For this family we obtain: a projective normalization by an elliptic curve, the factorization of the ramification polynomial, the real-branch classification, the real-spectral phase diagram, the dominant-branch identification on \(y>0\), the endpoint zero quantization at \(y=0\), and the degree law.
The structural contributions that go beyond the single specialization are Proposition~\ref{prop:balanced-family} and Theorem~\ref{thm:balanced-real-projection}: the former identifies a balanced quartic family for which the substitution \(y-\lambda^2\) removes the quartic obstruction uniformly, while the latter gives a general criterion for whether the real $y$-projection of any balanced-family member consists of one or two disjoint intervals.

Relative to Biggs's equimodular-curve viewpoint \cite{Biggs2002}, to the general Beraha--Kahane--Weiss/Sokal framework, and to the rational-generating-function zero literature of Tran and Forg\'acs--Tran, the present paper is narrower and should be read as a worked example.
Those works describe broad accumulation mechanisms or large parameter families; the four-term recurrence papers of Tran--Zumba \cite{TranZumba2018,TranZumba2021}, Adams \cite{Adams2020}, and Ndikubwayo \cite{Ndikubwayo2022} address real-zero conditions from different angles but do not produce the elliptic normalization or the dominant-branch geometry that the balanced-family reduction makes accessible.
Here the gain from specializing to one quartic is that the elliptic normalization, the real branch values, and the dominant positive branch can all be written and analyzed explicitly.
The paper is therefore best read as a detailed quartic case study: Proposition~\ref{prop:balanced-family} and Theorem~\ref{thm:balanced-real-projection} provide the reusable balanced-family statements, while the subsequent geometry is specific to the normalization \((a,c,d,g)=(-1,-1,1,0)\).

The scope of the present analysis is limited but precise.
All results concern the quartic companion family defined in Section~\ref{sec:model}.
No microscopic derivation of this family is claimed here, nor is one needed for the spectral statements proved above.
Accordingly, terms such as dominant branch, branch point, and \(f_{\mathrm{dom}}(y)=\log r(y)\) are used purely in the spectral sense internal to the quartic family itself.

From this viewpoint, the main structural point is the separation between the full ramification locus and the dominant-modulus envelope.
At \(y=0\), the branch collision occurs on the dominant branch over \(y>0\) and therefore controls the first edge zeros.
At \(y=1\) and \(y=y_{\mathrm{sub}}\), by contrast, the collisions are subdominant and do not create the same boundary behavior.
The real-spectral phase diagram makes this distinction visible, while the local cosine law of Section~\ref{sec:physical-sheet} explains the resulting odd-integer edge quantization.

At the opposite end of the parameter range, the large-\(y\) expansion is governed by a double tangency at projective infinity.
This gives a second closed-form mechanism, distinct from the finite branch-point analysis, for the quarter-power splitting of the dominant root.
Together with the degree law of Section~\ref{sec:support-law}, it shows that the top-degree sector is rigid in a precise algebraic sense.

Two extensions appear natural.
One is to derive the quartic family from an explicit combinatorial or statistical-mechanical model and thereby recover an interpretation of the coefficients.
The other is to push the balanced-family viewpoint beyond the normalization and real projection criterion.
Theorem~\ref{thm:balanced-real-projection} gives the coarse topological answer (one versus two intervals); the next level of precision would determine, for general $(a,c,d,g)$, how the dominant branch point depends on the parameters and when the endpoint zero asymptotics take the cosine-law form.
That broader program lies beyond the scope of the present paper, but the real branch-locus quintic \eqref{eq:balanced-quintic} and the criterion of Theorem~\ref{thm:balanced-real-projection} provide the starting data.
