\section{Appendix A: Termination and confluence of the fold normalization}
\label{app:rewriting}

We briefly record a verifiable normalization procedure realizing Zeckendorf form. This appendix uses MSD-first convention only. Appendix~\ref{app:pressure} does \emph{not} use this left-to-right scan; it works in LH convention and reconstructs the bulk pair shift from synchronized reversed pair blocks. Consider a binary Fibonacci digit string written in MSD-first convention. After prefixing a leading zero, repeatedly apply the local rewrite
\[
011 \to 100
\]
whenever it occurs, scanning left-to-right and continuing until no occurrence remains. In the low-to-high window convention used in the body of the paper, this is the equivalent rule \(110\to001\).

\begin{proposition}[Termination]\label{prop:termination}
Every rewrite sequence terminates after finitely many steps.
\end{proposition}

\begin{proof}
Each application of \(011\to 100\) strictly decreases the number of \(1\)s by one, hence termination follows.
\end{proof}

\begin{proposition}[Confluence and uniqueness of normal form]\label{prop:confluence}
The rewriting system is confluent: every input string rewrites to a unique normal form. This normal form contains no block \(11\) and equals the Zeckendorf representation of the same Fibonacci value.
\end{proposition}

\begin{proof}
Each rewrite preserves Fibonacci value. Indeed, if the block \(011\) occupies positions \(k,k+1,k+2\) in MSD-first convention, then
\[
\Fib_{k+2}+\Fib_{k+1}=\Fib_{k+3},
\]
so replacing \(011\) by \(100\) leaves the represented integer unchanged.

To prove confluence, it is enough to prove local confluence because termination is already known from Proposition~\ref{prop:termination}. The rule \(011\to 100\) has no nontrivial critical overlaps: two occurrences of \(011\) are either disjoint or identical. Hence if two rewrites start from the same word, they commute when they are distinct, and there is nothing to check when they coincide. Therefore the system is locally confluent, and Newman's lemma gives global confluence.

Let \(w^\sharp\) be the unique normal form. Since it is irreducible, the prefixed word \(0w^\sharp\) contains no occurrence of \(011\). If \(w^\sharp\) contained a block \(11\), take the leftmost occurrence and examine the symbol immediately to its left inside \(0w^\sharp\). That symbol must be \(0\) (it is either the prefixed \(0\) or the digit that separates this run of \(1\)s from the previous digit), so \(0w^\sharp\) would contain \(011\), contradicting irreducibility. Hence \(w^\sharp\) contains no block \(11\).

Finally, \(w^\sharp\) has the same Fibonacci value as the initial word, and it has no adjacent \(1\)s. By Zeckendorf uniqueness \cite{Lekkerkerker1952,Zeckendorf1972}, there is exactly one binary Fibonacci expansion with that property. Therefore \(w^\sharp\) is precisely the Zeckendorf representation of the input value.
\end{proof}

Together, Propositions~\ref{prop:termination}--\ref{prop:confluence} imply that the window-level fold \(\Fold_m\) is well-defined. Idempotence follows since any already-normal form is fixed by the rewrite, and surjectivity follows because every stable word \(x\in X_m\) is a fixed point: \(\Fold_m(x)=x\).

