\section{Appendix C: Sofic graph models and the conjugacy threshold}
\label{app:sofic}

\subsection{The labeled graph presentation}
Proposition~\ref{prop:sofic} constructs a labeled directed graph
\[
G_m=(V_m,E_m,\lambda_m),
\qquad
V_m=\{0,1\}^{m-1},
\]
whose edges are the length-\(m\) words \(vb\) (\(v\in\{0,1\}^{m-1}\), \(b\in\{0,1\}\)), with
\[
v\longrightarrow v'=\mathrm{suffix}_{m-1}(vb),
\qquad
\lambda_m(vb)=\Fold_m(vb)\in X_m.
\]
The edge shift of \(G_m\) is canonically conjugate to the full two-sided shift \(\{0,1\}^{\ZZ}\): the underlying microstate sequence is recovered by reading off the appended bit on each edge.

\subsection{Right-resolving property}
\begin{proposition}\label{prop:right-resolving}
For every \(m\ge 1\), the graph \(G_m\) is right-resolving: for each fixed vertex \(v\in V_m\), the two outgoing edges \(v0\) and \(v1\) have distinct labels.
\end{proposition}

\begin{proof}
The two words \(v0\) and \(v1\) have values differing by \(\Fib_{m+1}\). On the other hand, a length-\(m\) word has value at most
\[
\sum_{k=1}^m \Fib_{k+1}=\Fib_{m+3}-2<2\Fib_{m+2},
\]
so the full Zeckendorf normalization of any length-\(m\) word can have at most one nonzero digit beyond the first \(m\) coordinates, namely a possible \(1\) in position \(m+1\). Consequently, if two length-\(m\) words have the same fold label, then their values differ by a multiple of \(\Fib_{m+2}\): the first \(m\) Zeckendorf digits agree, and only the possible carry digit at position \(m+1\) can differ. Since
\[
0<\Fib_{m+1}<\Fib_{m+2},
\]
the words \(v0\) and \(v1\) cannot have the same fold label. Hence \(\Fold_m(v0)\neq \Fold_m(v1)\), as required.
\end{proof}

\subsection{Fiber products and injectivity for \texorpdfstring{$m\ge 3$}{m >= 3}}
Let \(F_m\) be the fiber product of \(G_m\) with itself. Thus the vertices of \(F_m\) are pairs \((u,v)\in V_m\times V_m\), and there is an edge
\[
(u,v)\longrightarrow (u',v')
\]
whenever \(u\to u'\) and \(v\to v'\) are edges of \(G_m\) carrying the same label. The diagonal copy of \(V_m\) always gives a subgraph of \(F_m\); off the diagonal, \(F_m\) records label collisions between distinct vertex paths.

We prove the absence of off-diagonal cycles through a sequence of lemmas that characterize same-label pairs, bound the elementary moves connecting them, and establish a strict potential decrease.

\begin{lemma}[Characterization of same-label pairs]\label{lem:same-label}
Two distinct words \(p,q\in\{0,1\}^m\) satisfy \(\Fold_m(p)=\Fold_m(q)\) if and only if \(p\) can be transformed into \(q\) by a finite sequence of the following \emph{elementary boundary moves} applied within the window \(\{1,\dots,m\}\):
\begin{enumerate}
\item[(E1)] \emph{Interior reverse rewrite:} replace \(001\) at positions \((k,k+1,k+2)\) with \(k+2\le m\) by \(110\), or vice versa. This preserves the Fibonacci value on positions \(1,\dots,m\).
\item[(E2)] \emph{Right-boundary carry absorption:} replace \((0,0)\) at positions \((m-1,m)\) by \((1,1)\), or vice versa, when the Fibonacci value modulo \(\Fib_{m+2}\) accounts for a carry at position \(m+1\). Concretely, \(\Fold_m\) retains only the first \(m\) Zeckendorf digits, so two words whose values differ by exactly \(\Fib_{m+2}\) receive the same fold label.
\end{enumerate}
No other type of move is needed.
\end{lemma}

\begin{proof}
If \(\Fold_m(p)=\Fold_m(q)\), then the Zeckendorf representations of \(N(p)\) and \(N(q)\) agree on positions \(1,\dots,m\). Hence \(N(p)\) and \(N(q)\) differ by a multiple of \(\Fib_{m+2}\). By the value bound in Proposition~\ref{prop:right-resolving}, the difference is either \(0\) or \(\pm\Fib_{m+2}\).

Case 1: \(N(p)=N(q)\). Then \(p\) and \(q\) are both Fibonacci representations of the same integer. By Appendix~\ref{app:rewriting}, they are connected by interior rewrites \(110\leftrightarrow 001\), which are exactly moves of type (E1).

Case 2: \(|N(p)-N(q)|=\Fib_{m+2}=\Fib_{m+1}+\Fib_m\). Without loss of generality, \(N(q)=N(p)+\Fib_{m+2}\). Then \(q\) is obtained from some intermediate word \(p'\) (connected to \(p\) by (E1) moves) by adding \(\Fib_{m+1}+\Fib_m\) at positions \(m\) and \(m-1\). This is exactly a type (E2) boundary move: flipping positions \((m-1,m)\) from \((0,0)\) to \((1,1)\) adds \(\Fib_m+\Fib_{m+1}=\Fib_{m+2}\). Combining (E1) and (E2) moves connects \(p\) to \(q\).
\end{proof}

\begin{lemma}[Minimal connecting chains]\label{lem:minimal-chain}
Let \(p\neq q\) with \(\Fold_m(p)=\Fold_m(q)\), and let \(\mathcal{C}\) be a connecting chain of elementary moves (E1)--(E2) from \(p\) to \(q\) with the minimum number of moves. Let \(r(\mathcal{C})\) be the number of moves in \(\mathcal{C}\), and let \(a(\mathcal{C})\in\{1,\dots,m\}\) be the leftmost position touched by any move in \(\mathcal{C}\). Then:
\begin{enumerate}
\item[(i)] \(r(\mathcal{C})\ge 1\).
\item[(ii)] Each (E1) move touches positions in a contiguous triple \((k,k+1,k+2)\); each (E2) move touches positions \((m-1,m)\).
\item[(iii)] For \(m\ge 3\), the support of \(\mathcal{C}\) --- the set of all positions touched by some move --- is contained in \(\{a(\mathcal{C}),\dots,m\}\) and has size at most \(2r(\mathcal{C})+1\).
\end{enumerate}
\end{lemma}

\begin{proof}
Part (i) is immediate since \(p\neq q\). Part (ii) follows from the definitions of (E1) and (E2). For part (iii), each move touches at most \(3\) consecutive positions, and moves in a minimal chain cannot have disjoint supports that leave a gap (otherwise the chain decomposes into independent subchains, contradicting that \(p\neq q\) requires them all). The leftmost position is \(a(\mathcal{C})\) and the rightmost is at most \(m\).
\end{proof}

\begin{lemma}[Potential decrease under shift]\label{lem:potential-decrease}
Let \(m\ge 3\). Consider an off-diagonal edge \((u,v)\to(u',v')\) in \(F_m\), corresponding to two distinct words \(p=ub_1\) and \(q=vb_2\) with \(\Fold_m(p)=\Fold_m(q)\), where \(b_1,b_2\in\{0,1\}\) are the appended bits and \(u'=\mathrm{suffix}_{m-1}(p)\), \(v'=\mathrm{suffix}_{m-1}(q)\). Let \(\mathcal{C}\) be a minimal connecting chain for \((p,q)\) with parameters \(r=r(\mathcal{C})\) and \(a=a(\mathcal{C})\). Define the integer-valued potential
\[
\mathcal{P}(u,v):=(m+1)\,r+a.
\]
Then \(u'\neq v'\) implies that any minimal connecting chain \(\mathcal{C}'\) for the successor pair \((p',q')\) --- where \(p'=u'b_1'\) and \(q'=v'b_2'\) are the words on the next edge with the same label --- satisfies
\[
\mathcal{P}(u',v')<\mathcal{P}(u,v).
\]
\end{lemma}

\begin{proof}
The shift from \(p\) to \(p'\) deletes position \(1\) and appends a new digit at position \(m\). We consider two cases.

\emph{Case 1: \(a>1\).} Every move in \(\mathcal{C}\) touches only positions \(\{a,\dots,m\}\), so deleting position \(1\) shifts all moves one position to the left. The number of moves \(r'=r\) is unchanged, but the leftmost position decreases: \(a'=a-1\). Hence \(\mathcal{P}'=(m+1)r+(a-1)=\mathcal{P}-1\).

\emph{Case 2: \(a=1\).} The leftmost move in \(\mathcal{C}\) touches position \(1\). After the shift deletes position \(1\), this move either exits the window entirely (if it was an (E1) move at positions \((1,2,3)\) and the shift removes position \(1\)) or its support shrinks. In either case, the connecting chain for the successor pair loses at least one move that involved position \(1\). Since the right-resolving property (Proposition~\ref{prop:right-resolving}) ensures that the appended bits \(b_1'\) and \(b_2'\) are uniquely determined by the common label, the new pair \((p',q')\) (if off-diagonal) has a minimal chain \(\mathcal{C}'\) with \(r'\le r-1\). Even if the new leftmost position \(a'\) is as large as \(m\), we have
\[
\mathcal{P}'=(m+1)(r-1)+a'\le (m+1)(r-1)+m=(m+1)r-1<(m+1)r+1\le \mathcal{P}.
\]
In both cases, \(\mathcal{P}\) strictly decreases.
\end{proof}

\begin{proposition}\label{prop:sync}
If \(m\ge 3\), the off-diagonal subgraph of \(F_m\) has no directed cycle. Equivalently, two distinct bi-infinite paths in \(G_m\) cannot carry the same bi-infinite label sequence.
\end{proposition}

\begin{proof}
By Lemma~\ref{lem:potential-decrease}, the potential \(\mathcal{P}\) strictly decreases along every off-diagonal edge of \(F_m\). Since \(\mathcal{P}\) is a positive integer, a directed cycle is impossible: any cycle would require \(\mathcal{P}\) to return to its starting value after finitely many strict decreases.

For the bi-infinite formulation: suppose two distinct bi-infinite paths carry the same label sequence. Restricting to any sufficiently long finite window gives an off-diagonal path segment in \(F_m\) of arbitrary length. But the potential \(\mathcal{P}\) is bounded below by \(1\) and decreases by at least \(1\) at each step, so no off-diagonal path can have length exceeding the initial potential value. This contradicts the arbitrary length.
\end{proof}

\subsection{An explicit decoder for \(m=3\)}
For \(m=3\) the decoder can be written down directly. We work in the low-to-high (LH) convention throughout: position \(k\) carries weight \(F_{k+1}\), so positions \(1,2,3\) carry weights \(1,2,3\) respectively (with \(F_2=1\), \(F_3=2\), \(F_4=3\)).

Given vertex \(v=(v_1,v_2)\) and appended bit \(b\), the three-digit LH word is \((v_1,v_2,b)\) with value \(N = v_1+2v_2+3b\). The successor vertex is \((v_2,b)\), and the edge label is \(\Fold_3(v_1 v_2 b)\), obtained by Zeckendorf-normalizing \(N\) and reading off the first three digits. The case \((v_1,v_2,b)=(0,1,1)\) deserves explicit comment: \(N=5=F_5\), and the Zeckendorf representation of \(F_5\) has its only nonzero digit at position \(k=4\) (where the weight is \(F_5=5\)), which lies outside the window; the first three Zeckendorf digits are therefore all zero. The complete edge-label table is:
\[
00\xrightarrow{000}00,\quad 00\xrightarrow{001}01,\quad
01\xrightarrow{010}10,\quad 01\xrightarrow{000}11,
\]
\[
10\xrightarrow{100}00,\quad 10\xrightarrow{101}01,\quad
11\xrightarrow{001}10,\quad 11\xrightarrow{100}11.
\]

We use the past-window convention: the vertex represents the two-digit suffix of the microstate sequence up to the current time. After reading edge \(b=\eta_t\) from vertex \(v=(\eta_{t-2},\eta_{t-1})\), the terminal vertex is \(v'=(\eta_{t-1},\eta_t)\). The last (rightmost) bit of \(v'\) is \(\eta_t\), the microstate bit to be decoded.

Among label pairs \((y_{t-1},y_t)\), the only two that admit more than one possible terminal vertex are
\[
(000,001):\text{ terminal vertex is }01\text{ or }10,
\qquad
(100,001):\text{ terminal vertex is }01\text{ or }10.
\]
In both cases the two candidate vertices have disjoint successor outgoing label sets:
\[
\mathcal{L}^+(01)=\{000,010\},\qquad \mathcal{L}^+(10)=\{100,101\}.
\]
The next label \(y_{t+1}\) therefore resolves the ambiguity: \(y_{t+1}\in\{000,010\}\) forces terminal vertex \(01\) (last bit \(1\)), and \(y_{t+1}\in\{100,101\}\) forces terminal vertex \(10\) (last bit \(0\)). Every other two-label block already determines the terminal vertex from \((y_{t-1},y_t)\) alone.

Since the two exceptional pairs each require one look-ahead label, the inverse code \(\Psi_3\) has anticipation \(1\): the decoded bit \((\Psi_3(y))_t\) is determined by \((y_{t-2},y_{t-1},y_t,y_{t+1})\). The last bit of the resolved terminal vertex gives \((\Psi_3(y))_t\).

\begin{proof}[Proof of Theorem~\ref{thm:conjugacy}]
The edge shift of \(G_m\) is conjugate to \(\{0,1\}^{\ZZ}\) by reading the appended bit on each edge. Under this identification, the label map of \(G_m\) is exactly \(\Phi_m\). Proposition~\ref{prop:right-resolving} shows that the presentation is deterministic to the future, and Proposition~\ref{prop:sync} shows that the fiber product has no off-diagonal bi-infinite path. Therefore the label map is injective on bi-infinite edge paths; see \cite[Chapter~8]{LindMarcus1995}. Since \(Y_m\) is its image by definition, the map is bijective onto \(Y_m\). Compactness of the domain and Hausdorffness of the codomain imply that the inverse is continuous, and the Curtis--Hedlund--Lyndon theorem identifies that inverse with a sliding block code \(\Psi_m\). The two displayed identities are then immediate.
\end{proof}

\subsection{The degenerate case \texorpdfstring{$m=2$}{m = 2}}
For \(m=2\), the label table is particularly simple:
\[
00\mapsto 00,\qquad
01\mapsto 01,\qquad
10\mapsto 10,\qquad
11\mapsto 00.
\]
Thus the only ambiguity is the collision \(00\leftarrow\{00,11\}\).

\begin{proof}[Proof of Theorem~\ref{thm:m2}]
The fiber product description makes the degeneracy transparent. Off the diagonal there are only the two ordered vertex pairs \((0,1)\) and \((1,0)\), and each has a single outgoing edge labeled \(00\) back to itself. These are exactly the two lifts of the constant label sequence \((00)^\ZZ\), namely the constant microstate sequences \(0^\ZZ\) and \(1^\ZZ\).

Every other label sequence contains either \(01\) or \(10\). Those labels have unique preimages \(01\) and \(10\), so once one such block appears the overlapping-window structure propagates a unique lift to the entire bi-infinite sequence. Hence \((00)^\ZZ\) is the only non-injective fiber, and it has precisely the two preimages stated in Theorem~\ref{thm:m2}.
\end{proof}
