\section{The partition-function recurrence and discriminant}
\label{app:recurrence}

This appendix collects the transfer-matrix calculations used in Section~\ref{sec:partition}. Once the four-state presentation from Appendix~\ref{app:pressure} is fixed, the remaining arguments are linear algebra and symbolic elimination, so we separate them from the main narrative.

\subsection{Weighted transfer matrix and boundary vectors}
Weight each edge of the four-state graph from Appendix~\ref{app:pressure} by \(y^{Y}\), where \(Y\in\{0,1\}\) is the second component of the label \((X,Y)\). In the state order \((00,01,10,11)\) this gives the weighted adjacency matrix
\[
M(y)=
\begin{pmatrix}
1&y&0&y\\
0&0&1&0\\
1&1+y&0&0\\
1&0&0&0
\end{pmatrix}.
\]
Its characteristic polynomial is
\[
\det(\lambda I-M(y))
=
\lambda^4-\lambda^3-(2y+1)\lambda^2+\lambda+y(y+1)
=
\Pi(\lambda,y).
\]
Replacing \(\lambda\) by \(t^{-1}\) gives
\[
\det(I-tM(y))
=
1-t-(2y+1)t^2+t^3+y(y+1)t^4
=
D(t;y).
\]
Because the synchronized graph is finite, the weighted language counted by \(Z_m(y)\) is rational and admits a four-dimensional linear representation; compare the standard transfer-matrix formalism in \cite[Chapter~6]{AlloucheShallit2003} and \cite{GrabnerKirschenhoferTichy2002}. With \(\alpha\) fixed by the synchronized initial state \(00\), a convenient terminal vector is obtained by solving
\[
\alpha^{\mathsf T}M(y)^m\beta(y)=Z_m(y)\qquad (m=0,1,2,3),
\]
using the directly enumerated initial polynomials
\[
Z_0(y)=1,\qquad
Z_1(y)=1+y,\qquad
Z_2(y)=2+2y,\qquad
Z_3(y)=2+5y+y^2.
\]
The resulting representation is
\[
Z_m(y)=\alpha^{\mathsf T}M(y)^m\beta(y),
\qquad
\alpha=
\begin{pmatrix}
1\\ 0\\ 0\\ 0
\end{pmatrix},
\qquad
\beta(y)=
\begin{pmatrix}
1\\[1mm]
\dfrac{1}{1+y}\\[2mm]
\dfrac{1}{y}\\[2mm]
\dfrac{y}{1+y}
\end{pmatrix}.
\]
The rational entries of \(\beta(y)\) are an artifact of this compact \(4\times4\) representation; after multiplication by \(M(y)^m\) all poles cancel and one recovers the polynomial \(Z_m(y)\).

\subsection{Numerator computation}
The generating function \(Z(t;y)=\sum_{m\ge 0}Z_m(y)t^m\) satisfies
\[
Z(t;y)=\frac{N(t;y)}{D(t;y)},
\]
where \(N(t;y)=Z(t;y)\cdot D(t;y)\) is determined by the initial values. Setting
\[
\begin{aligned}
N(t;y)
&=\sum_{m=0}^{3}
\Bigl(
Z_m(y)-\sum_{k=1}^{\min(m,4)}c_k(y)Z_{m-k}(y)
\Bigr)t^m,
\end{aligned}
\]
where \(c_1=1\), \(c_2=2y+1\), \(c_3=-1\), \(c_4=-y(y+1)\) are the recurrence coefficients, a direct calculation using \(Z_0=1\), \(Z_1=1+y\), \(Z_2=2+2y\), \(Z_3=2+5y+y^2\) gives
\[
N(t;y)=1+yt-yt^2-y^2t^3=(1+ty)(1-yt^2).
\]
The factorization is verified by expanding the product.

\subsection{Recurrence and Hankel rank}
Because \(Z_m(y)\) is obtained by summing the weights \(y^{\abs{\Fold_m(\omega)}_1}\) over all length-\(m\) microstate words, it is a linear functional of the weighted transfer-matrix dynamics. Consequently \(Z_m(y)\) satisfies the characteristic recurrence of \(M(y)\),
\[
Z_m(y)=Z_{m-1}(y)+(2y+1)Z_{m-2}(y)-Z_{m-3}(y)-y(y+1)Z_{m-4}(y)
\qquad (m\ge 4).
\]
The initial values
\[
Z_0(y)=1,\qquad
Z_1(y)=1+y,\qquad
Z_2(y)=2+2y,\qquad
Z_3(y)=2+5y+y^2
\]
come from direct enumeration. Therefore
\[
Z(t;y)=\sum_{m\ge0}Z_m(y)t^m
\]
is a rational function with denominator \(D(t;y)\) and numerator \(N(t;y)=(1+ty)(1-yt^2)\).

\subsection{Discriminant factorization}
Let
\[
\Pi(\lambda,y)=\lambda^4-\lambda^3-(2y+1)\lambda^2+\lambda+y(y+1).
\]
A direct symbolic computation gives
\[
\Disc_\lambda \Pi(\lambda,y)
=-y(y-1)\bigl(256y^3+411y^2+165y+32\bigr).
\]
The cubic factor has exactly one negative real root,
\[
y_{\mathrm{c}}\approx -1.13445.
\]
This is the unique negative real value of \(y\) at which \(\Pi(\lambda,y)\) has a repeated root, apart from the zeros at \(y=0\) and \(y=1\).

\subsection{Root analysis at the discriminant zeros}
At each discriminant zero, we compute the roots of \(\Pi(\lambda,y)\) explicitly.

\emph{At \(y=0\):} \(\Pi(\lambda,0)=\lambda^4-\lambda^3-\lambda^2+\lambda=\lambda(\lambda-1)^2(\lambda+1)\). The repeated root is \(\lambda=1\), which is dominant. Since \(Z(t;0)=1/((1-t)^2(1+t))\), the coefficients \(Z_m(0)=\abs{\Fold_m^{-1}(0^m)}\) grow linearly in \(m\), exactly as expected from the double dominant root.

\emph{At \(y=1\):} \(\Pi(\lambda,1)=\lambda^4-\lambda^3-3\lambda^2+\lambda+2=(\lambda-2)(\lambda+1)^2(\lambda-1)\). The repeated root is \(\lambda=-1\), which is subdominant (dominated by \(\lambda=2\)).

\emph{At \(y=y_{\mathrm{c}}\approx -1.13445\):} the quartic \(\Pi(\lambda,y_{\mathrm{c}})\) evaluates to approximately \(\lambda^4-\lambda^3+1.2689\lambda^2+\lambda+0.1525\). Numerical root-finding gives a repeated real root near \(\lambda\approx -0.2734\) and a complex-conjugate pair near \(\lambda\approx 0.7734\pm 1.2008\,i\) with modulus approximately \(1.428\). Since \(|-0.2734|<1.428\), the repeated root is strictly subdominant. Hence the discriminant zero at \(y_{\mathrm{c}}\) does not produce a singularity of the dominant eigenvalue.

\begin{remark}
For \(y>0\), the Perron--Frobenius theorem guarantees that the dominant root of \(M(y)\) is simple, real, and positive. The free energy \(f(y)=\log\lambda_\star(y)\) is therefore real-analytic on \((0,\infty)\). On the negative real axis, \(Z_m(y)\) involves sign-alternating terms and a separate dominant-sheet analysis would be needed; we do not pursue this.
\end{remark}

\subsection{Output-weight cumulants}
Set \(y=e^\theta\). The scaled cumulant generating function for \(H_m\) is
\[
\Lambda_Y(\theta)
=
\log \rho\!\bigl(M(e^\theta)\bigr)-\log 2.
\]
Write
\[
F(\lambda,z):=\lambda^4-\lambda^3-(2z+1)\lambda^2+\lambda+z(z+1).
\]
Since \(F(\rho(\theta),e^\theta)=0\) and \(F(2,1)=0\), implicit differentiation at \(\theta=0\) gives
\[
F_\lambda(2,1)\rho'(0)+F_z(2,1)=0.
\]
Here
\[
F_\lambda(2,1)=9,\qquad F_z(2,1)=-5,
\]
so \(\rho'(0)=5/9\). A second differentiation yields
\[
F_{\lambda\lambda}(2,1)\rho'(0)^2+2F_{\lambda z}(2,1)\rho'(0)+F_{zz}(2,1)+F_z(2,1)+F_\lambda(2,1)\rho''(0)=0.
\]
Using
\[
F_{\lambda\lambda}(2,1)=30,\qquad
F_{\lambda z}(2,1)=-8,\qquad
F_{zz}(2,1)=2,
\]
one obtains \(\rho''(0)=71/243\). Consequently,
\[
\Lambda_Y'(0)=\frac{5}{18},
\qquad
\Lambda_Y''(0)=\frac{67}{972}.
\]
These are the mean density and asymptotic variance appearing in Corollary~\ref{thm:weight-stats}. Since \(\theta\mapsto \rho(M(e^\theta))\) is real-analytic on \(\RR\), the G\"artner--Ellis rate function is
\[
I(a)=\sup_{\theta\in\RR}\bigl(\theta a-\Lambda_Y(\theta)\bigr).
\]
