\documentclass[10pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{array}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[expansion=false]{microtype}
\usepackage{enumitem}
\usepackage{mathrsfs}
\usepackage{tikz}
\usepackage[margin=2.5cm]{geometry}

\numberwithin{equation}{section}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\NN}{\mathbb{N}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\ind}{\mathbf{1}}
\newcommand{\Aadm}{\mathcal{A}_{\mathrm{adm}}}
\newcommand{\Ainf}{\mathcal{A}_{\infty}}
\newcommand{\Avis}{A_{\mathrm{vis}}(\mathfrak{U},[\alpha])}
\newcommand{\Xadm}{X_{\mathrm{adm}}}
\newcommand{\Rpre}{\mathscr{R}}
\newcommand{\Comp}{\mathsf{Comp}}
\DeclareMathOperator{\diam}{diam}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\Isom}{Isom}
\DeclareMathOperator{\Ob}{Ob}
\DeclareMathOperator{\Aut}{Aut}

\title{Recursive Addressing via Prefix Sites, Local Sections, and Inverse Limits: An Expository Note}
\author{
  Haobo Ma\thanks{AELF PTE LTD., \#14-02, Marina Bay Financial Centre Tower 1, 8 Marina Blvd, Singapore 018981. Email: \texttt{auric@aelf.io}}
  \and
  Wenlin Zhang\thanks{Corresponding author. National University of Singapore, Singapore. Email: \texttt{e1327962@u.nus.edu}}
}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We present an expository note on recursive addressing.  Starting from a
measurable dynamical system and a finite family of observable concepts, we
record the standard fact that recursively generated address cylinders refine
visible information without enlarging the visible $\sigma$-algebra.  Fixing a
time origin, the induced symbolic coding produces admissible words and cylinder
objects for a prefix site.  Within this language, address failure splits into
the standard presheaf-theoretic phenomena of empty local fibers, descent
failure, and non-separatedness.  For a finite cover together with local
readout objects carrying an abelian band, we define the associated intrinsic
\v{C}ech defect class and the visible quotient obtained by killing that class.
This quotient yields
a register lower bound.  We also give a concrete symbolic computation: a
four-chart cylinder-algebra cover of the full binary shift whose ordered nerve is the
boundary of a tetrahedron and whose prescribed $\ZZ/2\ZZ$-valued defect class
is nonzero; a companion $\ZZ/4\ZZ$ example exhibits partial visibility.
Finally, for prefix inverse systems we formulate single-valued readout at an
inverse-limit point via the local site of cylinders containing that point,
together with stalks and germs.  The emphasis throughout is on organizing
standard sheaf-theoretic language and on isolating the additional semantic
input needed to generate the relevant local objects and bands functorially.
\end{abstract}

\section{Introduction}

The slogan behind recursive addressing is simple: addresses precede values.
Used carelessly, however, that slogan invites too much secondary terminology.
This article is intentionally expository.  We fix a single vocabulary for the
measurable and symbolic faces of recursive addressing, spell out the minimal
formalism needed to discuss local sections, and package the few universal
constructions that arise in practice (notably the visible quotient in
Section~\ref{sec:visible-quotient}).  Every numbered statement either records a
standard fact or isolates the additional hypotheses required for a routine
construction.  The emphasis is therefore on readability and on exposing exactly
where genuinely new semantic input would have to enter.  In particular,
Sections~\ref{sec:measurable}--\ref{sec:visible-quotient} are driven by quoted
background, while Section~\ref{sec:semantic-example} presents a worked example
extracted directly from a measurable model to illustrate how the abstract
constructions behave in practice.

Once the measurable statement is isolated, the rest of the paper can be stated
in standard site language.  Proposition~\ref{prop:measurable-coding-prefix-bridge}
spells out the bridge: fixing a time origin turns the measurable readout
process into a symbolic coding map, and the measurable cylinder events are
exactly pullbacks of symbolic cylinders.  Admissible addresses then give rise
to a prefix site of cylinders and finite unions of cylinders, with covers given
by finite cylinder refinements.  Over that site we place a groupoid-valued
prestack $\Rpre$ of admissible local readout objects (introduced formally in
Section~\ref{sec:descent-groupoid}), so comparisons and torsors are intrinsic.
What informal recursive-addressing language treated as address failure is then
rewritten internally:
\begin{itemize}
\item there may be no local object over a given address;
\item compatible local objects may fail to descend to a global one;
\item global objects may fail to be separated by their restrictions.
\end{itemize}
The three cases correspond respectively to empty local fibers, descent
obstructions, and non-separatedness.  In particular, most of the paper records
standard sheaf-theoretic mechanisms in recursive-addressing notation.  The
point of the note is not to claim new general theorems about arbitrary sites or
presheaves, but to isolate the places where additional comparison data become
necessary and to work out one concrete symbolic model in detail.

\medskip
\noindent\textbf{Structure of the paper and scope.}
Section~\ref{sec:measurable} is background: it records the measurable
non-expansion statement that underlies recursive refinement
(Proposition~\ref{prop:recursive-addressing-refinement} and
Corollary~\ref{cor:recursive-addressing-visible-sigma-nonexpansion}) and the
coding bridge to the symbolic setting
(Proposition~\ref{prop:measurable-coding-prefix-bridge});
see~\cite{WaltersErgodicTheory} for standard measurable-dynamics
prerequisites.  Section~\ref{sec:prefix-sites} reformulates admissible
addresses as a site of trace cylinders inside $\Xadm$, using standard material
on sites and Grothendieck topologies from
\cite{MacLaneMoerdijk1992,Johnstone2002Elephant,SGA4I}; the associated readout
prestack and failure-mode trichotomy
(Corollary~\ref{cor:null-trichotomy-local-section}) are direct specializations
of the separated-presheaf formalism.  For modern local-to-global viewpoints in
the same spirit, see also
\cite{AbramskyBrandenburger2011,AbramskyMansfieldBarbosa2012,Curry2014Sheaves,DowlingKaliesVandervorst2023}.

Section~\ref{sec:descent} contains the local-section dictionary and the
normalized \v{C}ech obstruction theory.  We first give the formal cover nerve
and alternating cochain complex
(Lemma~\ref{lem:cech-complex-formal}), following the general framework of
\cite{StacksProject,Giraud1971CohomologieNonAbelienne,Bredon1997SheafTheory,HatcherAlgebraicTopology}.
We then define the intrinsic $A$-banded comparison torsors and the associated
\v{C}ech defect class
(Definition~\ref{def:prefix-site-h2-gluing-obstruction}) once local objects
with an abelian band have been fixed.

Section~\ref{sec:visible-quotient} packages a routine universal-construction
observation: for a fixed finite cover $\mathfrak{U}$ with defect class
$[\alpha]$, the visible quotient $\Avis$ is the initial quotient of the band on
which $[\alpha]$ vanishes
(Proposition~\ref{thm:intrinsic-visible-quotient}).  The pushed-forward class
vanishes on $\Avis$, and every homomorphism killing $[\alpha]$ factors uniquely
through $\Avis$.  The register bound
(Lemma~\ref{thm:prefix-site-register-lower-bound}) and spectral criterion
(Proposition~\ref{thm:prefix-site-spectral-definability-avis}) then follow.

Section~\ref{sec:forced-obstruction} is a worked symbolic model rather than a
general forcing theorem.  We construct a four-chart cylinder-algebra cover of the full
binary shift whose nerve is the boundary of a tetrahedron, prescribe
comparison data directly on the ordered nerve, and compute the resulting
nonzero defect class (Proposition~\ref{thm:forced-obstruction},
Example~\ref{ex:tetrahedral-cover}, and
Example~\ref{ex:tetrahedral-z4}).

Section~\ref{sec:inverse-limits} is background for prefix inverse-limit towers
\cite{RibesZalesskii2010ProfiniteGroups,LindMarcus1995SymbolicDynamics,Kitchens1998SymbolicDynamics},
with one local reformulation: single-valued readout at an inverse-limit point
is restated in stalk/germ language
(Theorem~\ref{thm:recursive-addressing-readout-separatedness-null}), using the
local site of all cylinders containing a point and the stalk as a filtered
colimit.

\medskip
\noindent\textbf{What is standard and what is specific here.}
\begin{enumerate}[label=\textup{(\arabic*)}]
\item Proposition~\ref{prop:recursive-addressing-refinement} and
Corollary~\ref{cor:recursive-addressing-visible-sigma-nonexpansion}: the
measurable non-expansion statement (background).
\item Proposition~\ref{prop:prefix-site-grothendieck-topology},
Definition~\ref{def:prefix-sheaf-readout}, and
Corollary~\ref{cor:null-trichotomy-local-section}: the prefix site,
readout prestack, and failure-mode trichotomy (reformulation of standard
material).
\item Lemma~\ref{lem:cech-complex-formal},
Definition~\ref{def:prefix-site-h2-gluing-obstruction}, and
Proposition~\ref{prop:null-as-h2-obstruction}: the formal \v{C}ech complex and
the defect obstruction attached to banded local objects on a fixed cover.
\item Proposition~\ref{thm:intrinsic-visible-quotient}: the visible quotient
attached to a defect class via universal property, and its consequences
(Lemma~\ref{thm:prefix-site-register-lower-bound},
Proposition~\ref{thm:prefix-site-spectral-definability-avis}).
\item Proposition~\ref{thm:forced-obstruction},
Example~\ref{ex:tetrahedral-cover}, and
Example~\ref{ex:tetrahedral-z4}: explicit symbolic computations on a
tetrahedral cover.
\item Theorem~\ref{thm:recursive-addressing-readout-separatedness-null}:
single-valued readout at an inverse-limit point reformulated via local stalks
and germs.
\end{enumerate}

To keep the vocabulary fixed, we will repeatedly translate three recurring
phrases into standard sheaf language: ``addresses precede values'' means that
sections only exist after an admissible cylinder has been chosen, ``semantic
null'' refers to an empty local fiber $\Rpre(a)=\varnothing$, ``protocol
null'' to failure of compatible local data to glue, and ``collision null'' to
non-separatedness of global threads under restriction.  These translations are
used verbatim in every section, so no additional terminology tables are needed.

\medskip
\noindent\textbf{Related literature.}
The general local-to-global obstruction program originates in the contextuality
framework of Abramsky and collaborators
\cite{AbramskyBrandenburger2011,Abramsky2014ContextualSemantics,AbramskyMansfieldBarbosa2012,AbramskyEtAl2015ContextualityCohomologyParadox},
where measurement contexts canonically produce the comparison torsors and bands
whose failure to glue obstructs global sections.  Staton and Uijlen
\cite{StatonUijlen2018EffectAlgebras} provide a semantics-facing refinement in
terms of effect algebras and presheaves, and the continuation-sheaf story of
\cite{DowlingKaliesVandervorst2023} supplies a dynamical complement.  The
present note sits upstream of those works: we freeze the ``addresses precede
values'' vocabulary, isolate the precise hypotheses needed for basic descent
arguments, and treat the local objects together with their abelian band as the
input datum.  Once that datum is fixed, the comparison torsors are intrinsic;
Section~\ref{sec:semantic-example} records one example where the semantics
supplies the datum directly.  The references cited above remain the canonical
sources for genuinely new obstruction theorems.

\section{Recursive Addressing in Measurable Systems}
\label{sec:measurable}

Let $(X,\mathcal{B},T)$ be a measurable dynamical system
\cite{WaltersErgodicTheory}.  Fix a finite
family of measurable sets
\[
\mathcal{C}^{(L)} = \{C_1^{(L)},\dots,C_{r_L}^{(L)}\}\subset \mathcal{B},
\]
thought of as the observable concepts available at level $L$.  Their binary
readout process is
\[
\mathbf{b}^{(L)}_t(x)
:=
\bigl(
\ind_{C_1^{(L)}}(T^t x),\dots,\ind_{C_{r_L}^{(L)}}(T^t x)
\bigr)\in\{0,1\}^{r_L}.
\]
For a word
\[
a=(a_0,\dots,a_{\ell-1})\in \bigl(\{0,1\}^{r_L}\bigr)^\ell
\]
and an integer $t$, define the derived cylinder event
\[
C_{a,t}^{(L+1)}
:=
\{x\in X:\mathbf{b}^{(L)}_t(x)=a_0,\dots,\mathbf{b}^{(L)}_{t+\ell-1}(x)=a_{\ell-1}\}.
\]
Thus the next level is produced by finite windows in the previous one.  The
statements recorded below are classical consequences of the definitions; we
quote them for completeness because they anchor the later symbolic language.

\begin{proposition}[Recursive addressing is measurable refinement]
\label{prop:recursive-addressing-refinement}
Let
\[
\mathcal{G}^{(L)}
:=
\sigma\bigl(T^{-t}C_i^{(L)}:\ 1\le i\le r_L,\ t\in\ZZ\bigr)
\]
be the visible $\sigma$-algebra generated by the level-$L$ concepts and their
time shifts.  Then every derived cylinder event $C_{a,t}^{(L+1)}$ belongs to
$\mathcal{G}^{(L)}$.
\end{proposition}

\begin{proof}
For fixed $i$ and $t$, the map
\[
x\longmapsto \ind_{C_i^{(L)}}(T^t x)=\ind_{T^{-t}C_i^{(L)}}(x)
\]
is $\mathcal{G}^{(L)}$-measurable.  Hence the event
\[
\{x:\mathbf{b}^{(L)}_t(x)=a_0,\dots,\mathbf{b}^{(L)}_{t+\ell-1}(x)=a_{\ell-1}\}
\]
is a finite intersection of sets of the form $T^{-s}C_i^{(L)}$ and their
complements, and therefore lies in $\mathcal{G}^{(L)}$.
\end{proof}

\begin{corollary}[Visible $\sigma$-algebras do not expand]
\label{cor:recursive-addressing-visible-sigma-nonexpansion}
Suppose $\mathcal{C}^{(L+1)}$ is chosen from the derived cylinder events
$C_{a,t}^{(L+1)}$, and define
\[
\mathcal{G}^{(L+1)}
:=
\sigma\bigl(T^{-t}C:\ C\in \mathcal{C}^{(L+1)},\ t\in\ZZ\bigr).
\]
Then
\[
\mathcal{G}^{(L+1)}\subseteq \mathcal{G}^{(L)}.
\]
\end{corollary}

\begin{proof}
Each set $C\in\mathcal{C}^{(L+1)}$ is one of the derived cylinder events
$C_{a,t_0}^{(L+1)}$, so Proposition~\ref{prop:recursive-addressing-refinement}
gives $C\in\mathcal{G}^{(L)}$.  Since $\mathcal{G}^{(L)}$ is generated by all
time shifts $T^{-s}C_i^{(L)}$, it is stable under every shift $T^{-t}$:
\[
T^{-t}\bigl(T^{-s}C_i^{(L)}\bigr)=T^{-(s+t)}C_i^{(L)}\in \mathcal{G}^{(L)}.
\]
Hence $T^{-t}C\in\mathcal{G}^{(L)}$ for every $C\in\mathcal{C}^{(L+1)}$ and
every $t\in\ZZ$.  Therefore every generator of $\mathcal{G}^{(L+1)}$ lies in
$\mathcal{G}^{(L)}$, so
\[
\mathcal{G}^{(L+1)}\subseteq \mathcal{G}^{(L)}.
\]
\end{proof}

\begin{remark}
This is the measurable content of recursive addressing.  New addresses are
produced by reorganizing visible cylinder events; they do not enlarge the
observable $\sigma$-algebra.  The rest of the paper repackages this fact in
site-theoretic language.
\end{remark}

To connect the measurable setup with the symbolic site used below, fix the
alphabet
\[
\Sigma_L:=\{0,1\}^{r_L}
\]
and define the coding map
\[
\kappa_L\colon X\to \Sigma_L^{\ZZ},
\qquad
(\kappa_L(x))_t:=\mathbf{b}^{(L)}_t(x).
\]
For a finite word $a=(a_0,\dots,a_{\ell-1})\in \Sigma_L^\ell$ and an integer
$t$, write
\[
[a]_t
:=
\{y\in \Sigma_L^{\ZZ}: y_t=a_0,\dots,y_{t+\ell-1}=a_{\ell-1}\}.
\]
Let $\kappa_L^+\colon X\to \Sigma_L^{\NN}$ be the forward coding map
\[
(\kappa_L^+(x))_t:=\mathbf{b}^{(L)}_t(x)
\qquad (t\ge 0),
\]
and for $a\in \Sigma_L^{<\infty}$ let $C_a\subseteq \Sigma_L^{\NN}$ denote the
usual one-sided cylinder of forward sequences beginning with $a$.  Define the
admissible forward words by
\[
\Aadm^{(L)}
:=
\{a\in \Sigma_L^{<\infty}: (\kappa_L^+)^{-1}(C_a)\neq\varnothing\}.
\]

\begin{proposition}[Coding bridge from measurable events to prefix cylinders]
\label{prop:measurable-coding-prefix-bridge}
For every finite word $a=(a_0,\dots,a_{\ell-1})\in \Sigma_L^\ell$ and every
$t\in\ZZ$,
\[
C_{a,t}^{(L+1)}=\kappa_L^{-1}([a]_t).
\]
In particular, for every $a\in \Aadm^{(L)}$,
\[
(\kappa_L^+)^{-1}(C_a)=C_{a,0}^{(L+1)},
\]
so the forward symbolic cylinders are exactly the measurable derived cylinder
events at time $0$.  More generally,
\[
\kappa_L^{-1}(\sigma^{-t}[a]_0)=C_{a,t}^{(L+1)},
\]
where $\sigma$ is the shift on $\Sigma_L^{\ZZ}$.
\end{proposition}

\begin{proof}
By definition, $x\in C_{a,t}^{(L+1)}$ if and only if
\[
\mathbf{b}^{(L)}_t(x)=a_0,\dots,\mathbf{b}^{(L)}_{t+\ell-1}(x)=a_{\ell-1},
\]
which is equivalent to $\kappa_L(x)\in [a]_t$.  This proves the first
identity.  The statement for $t=0$ is exactly the definition of the one-sided
cylinder $C_a$, and the shifted form is the same identity written as a pullback
of a translated cylinder in $\Sigma_L^{\ZZ}$.
\end{proof}

\section{Prefix Sites and Readout Presheaves}
\label{sec:prefix-sites}

From now on we stay inside the admissible image shift
\[
\Xadm:=\im(\kappa_L^+)\subseteq \Sigma_L^{\NN}
\]
produced by the measurable coding map, so every admissible word is literally a
trace that occurs in $\Xadm$.
Proposition~\ref{prop:measurable-coding-prefix-bridge} shows that, after fixing
a time origin, the symbolic site is obtained from the measurable readout
process by pulling back cylinders inside $\Xadm$, not inside the full shift,
so every object considered below is a subset of $\Xadm$.

\subsection{The Cylinder Category}

Fix a finite alphabet $\Sigma$ and let $\Xadm\subseteq \Sigma^{\NN}$ be any
subshift that arises as an admissible image $\im(\kappa_L^+)$ of a measurable
coding map.  Let $\Sigma^{<\infty}$ be the set of finite words, write
$a\preceq b$ when $a$ is a prefix of $b$, and set
\[
\Aadm:=\{a\in\Sigma^{<\infty}:\ C_a^{\mathrm{full}}\cap \Xadm\neq\varnothing\},
\]
the set of words that actually occur inside $\Xadm$.  For such a word $a$, let
$C_a\subseteq \Xadm$ be the trace cylinder consisting of all sequences of
$\Xadm$ that begin with $a$
\cite{LindMarcus1995SymbolicDynamics,Kitchens1998SymbolicDynamics}; equivalently,
$C_a=\Xadm\cap C_a^{\mathrm{full}}$, where $C_a^{\mathrm{full}}$ is the usual
cylinder in the full shift $\Sigma^{\NN}$.

\begin{definition}[Cylinder algebra]
\label{def:cylinder-algebra}
The \emph{cylinder algebra} $\mathfrak{C}(\Aadm)$ is the Boolean subalgebra of
$\mathcal{P}(\Xadm)$ generated by the trace cylinders $\{C_a:a\in\Aadm\}$.  Its
elements are finite unions of admissible cylinders.  Because $\Xadm$ consists
only of actual traces, the following properties hold automatically.
\begin{enumerate}[label=\textup{(CA\arabic*)}]
\item\label{ax:nesting} if $b\succeq a$, then $C_b\subseteq C_a$;
\item\label{ax:intersection} if $a,b\in\Aadm$ and $C_a\cap C_b\neq\varnothing$,
then there exists $d\in\Aadm$ with $d\succeq a$, $d\succeq b$, and
$C_d = C_a\cap C_b$ (take $d$ to be the longer of the two words);
\item\label{ax:refinement} every nonempty element of $\mathfrak{C}(\Aadm)$ can
be written as a finite disjoint union of cylinders $C_{a_i}$ with
$a_i\in\Aadm$.
\end{enumerate}
\end{definition}

Axiom~\ref{ax:intersection} simply records that if a sequence in $\Xadm$
begins with both $a$ and $b$, then one word extends the other, so the overlap
agrees with the cylinder of the longer word.  The refinement axiom follows
because every finite union of basic trace cylinders can be refined by
splitting into disjoint cylinders of a common depth.  When $\Xadm$ is the full
shift, these statements reduce to the usual combinatorics of prefixes in
$\Sigma^{\NN}$.

\begin{definition}[Prefix site]
\label{def:prefix-site}
The \emph{prefix site} $\mathsf{Pref}(\Aadm)$ is the category whose objects
are the elements of the cylinder algebra $\mathfrak{C}(\Aadm)$ and whose
morphisms are inclusions.  A finite family $\{U_i\hookrightarrow U\}_{i=1}^n$
is declared to be a \emph{cover} if
\[
U = \bigcup_{i=1}^n U_i.
\]
\end{definition}

When convenient, we identify the cylinder $C_a$ with its address $a$ and write
$b\to a$ for the inclusion $C_b\hookrightarrow C_a$ when $b\succeq a$.  In
this notation, a \emph{cylinder-algebra cover} of $a$ is a finite family
$\{b_i\to a\}_{i=1}^n$ such that $C_a=\bigcup_i C_{b_i}$.

\begin{proposition}[Cylinder-algebra covers generate a Grothendieck topology]
\label{prop:prefix-site-grothendieck-topology}
The cylinder-algebra covers satisfy the identity, base-change, and transitivity axioms
of a finitary pretopology on $\mathsf{Pref}(\Aadm)$
\cite[Chapter~III]{MacLaneMoerdijk1992},
\cite[\S C2.1]{Johnstone2002Elephant}.  Hence they generate a
Grothendieck topology $J_{\mathrm{cyl}}$, and
\[
(\mathsf{Pref}(\Aadm),J_{\mathrm{cyl}})
\]
is a well-defined site.
\end{proposition}

\begin{proof}
The singleton family $\{U\hookrightarrow U\}$ is a cover because $U=U$.

For base change, let $\{U_i\hookrightarrow U\}_{i=1}^n$ be a cover and let
$V\hookrightarrow U$ be a morphism (inclusion).  Then
\[
V
=
V\cap U
=
\bigcup_{i=1}^n (V\cap U_i).
\]
Each $V\cap U_i$ belongs to $\mathfrak{C}(\Aadm)$ by
axiom~\ref{ax:intersection} (applied iteratively to the cylinder
decompositions of $V$ and $U_i$) and axiom~\ref{ax:refinement}.  Writing each
nonempty $V\cap U_i$ as a finite union of admissible cylinders gives a cover of
$V$ that refines the original cover of $U$.

For transitivity, if $\{U_i\hookrightarrow U\}_{i=1}^n$ is a cover and each
$\{V_{ij}\hookrightarrow U_i\}_{j=1}^{m_i}$ is a cover of $U_i$, then
\[
U
=
\bigcup_{i=1}^n U_i
=
\bigcup_{i=1}^n \bigcup_{j=1}^{m_i} V_{ij},
\]
so the composite family is again a cover.
Therefore the covers form a finitary pretopology and hence generate a
Grothendieck topology \cite[Proposition~III.2.2]{MacLaneMoerdijk1992}.
\end{proof}

\subsection{The Readout Prestack}

The paper keeps the readout object deliberately abstract.  A local readout over
$a$ is any object that can be evaluated using only the information visible on
$C_a$ and that remains admissible under further refinement.

\begin{definition}[Readout prestack]
\label{def:prefix-sheaf-readout}
A \emph{readout prestack} on $\mathsf{Pref}(\Aadm)$ is a contravariant pseudofunctor
\[
\Rpre:\mathsf{Pref}(\Aadm)^{\mathrm{op}}\to \mathbf{Gpd}.
\]
For $a\in\Aadm$, the groupoid $\Rpre(a)$ records the admissible local readout
objects over $C_a$ and their internal morphisms.  For a refinement $b\succeq a$,
the restriction functor
\[
\rho_{a\leftarrow b}\colon \Rpre(a)\longrightarrow \Rpre(b)
\]
is functorial and encodes admissibility under further refinement.  We write
$\Ob\Rpre(a)$ for its object set and $\Aut_{\Rpre}(s)$ for the automorphism
group of $s\in\Ob\Rpre(a)$.  Whenever we state that $\Rpre(a)=\varnothing$, we
mean that $\Ob\Rpre(a)$ has no objects, and all previous set-valued terminology
applies to $\Ob\Rpre$ via the forgetful functor $\mathbf{Gpd}\to\mathbf{Set}$.
We say that $\Rpre$ is \emph{separated} with respect to a cover
$\{b_i\to a\}_{i=1}^n$ if the canonical map
$\Ob\Rpre(a)\to\prod_i\Ob\Rpre(b_i)$ is injective, equivalently if the
underlying set-valued presheaf of objects is separated on that cover
\cite[\S II.1]{MacLaneMoerdijk1992}.
\end{definition}

\section{Local Sections and Descent}
\label{sec:descent}
\label{sec:descent-groupoid}

\subsection{Local Sections}

\begin{proposition}[Addressability as existence of a local section]
\label{prop:null-as-local-section-obstruction}
For an admissible address $a\in\Aadm$, readout is internally defined at $a$ if
and only if
\[
\Rpre(a)\neq\varnothing.
\]
\end{proposition}

\begin{proof}
By Definition~\ref{def:prefix-sheaf-readout}, the object set
$\Ob\Rpre(a)$ consists exactly of the admissible local readout objects over
the cylinder $C_a$.  Therefore $\Rpre(a)\neq\varnothing$ (meaning
$\Ob\Rpre(a)\neq\varnothing$) if and only if at least one such local readout
exists.
\end{proof}

\begin{corollary}[Three standard failure modes]
\label{cor:null-trichotomy-local-section}
For an admissible finite-level address $a\in\Aadm$, failure of a well-defined
readout at $a$ has three internal
forms:
\begin{enumerate}[label=\textup{(\arabic*)}]
\item \emph{empty local fiber}: $\Rpre(a)=\varnothing$ (i.e., $\Ob\Rpre(a)$ has
no objects);
\item \emph{descent failure}: local sections exist on a cylinder-algebra cover of $a$
and are pairwise compatible, but no object of $\Rpre(a)$ restricts to them;
\item \emph{non-separatedness}: two distinct objects of $\Rpre(a)$ have the
same restrictions to a cylinder-algebra cover of $a$.
\end{enumerate}
These modes are not mutually exclusive, and the latter two are relative to a
chosen cylinder-algebra cover of $a$.
\end{corollary}

\begin{proof}
The first case is Proposition~\ref{prop:null-as-local-section-obstruction}.  If
$\Rpre(a)\neq\varnothing$ but a compatible family on a cylinder-algebra cover does not
come from any object of $\Rpre(a)$, then the failure is one of effective
descent.  If distinct objects of $\Rpre(a)$ induce the same compatible family
on a cover, then the failure is one of separatedness.  These are the three
standard internal failure modes for the existence and uniqueness of readout on
a presheaf over a fixed admissible object and a fixed cover
\cite[\S II.1--2]{MacLaneMoerdijk1992}.
\end{proof}

\begin{example}[Binary prefix site and the three failure modes]
\label{ex:binary-prefix-site-failure-modes}
Let $X=\{0,1\}^{\NN}$, let $\Sigma=\{0,1\}$, let $\Aadm=\Sigma^{<\infty}$, and
let $C_a\subseteq X$ be the usual cylinder of sequences beginning with the word
$a$.  The root address $\varepsilon$ is covered by the two depth-one cylinders
$0\to \varepsilon$ and $1\to \varepsilon$.

Three explicit readout presheaves on this same binary prefix site realize the
three failure modes from Corollary~\ref{cor:null-trichotomy-local-section}.
\begin{enumerate}[label=\textup{(\arabic*)}]
\item \emph{Empty fiber}: let $\Rpre_{\mathrm{emp}}(\varepsilon)=\varnothing$
and $\Rpre_{\mathrm{emp}}(b)=\{*\}$ for every nonempty word $b$, with the
unique possible restriction maps.  Readout fails immediately at the root.
\item \emph{Descent failure}: let $\Rpre_{\mathrm{desc}}(\varepsilon)=\{g\}$,
let $\Rpre_{\mathrm{desc}}(0)=\{x_0,y_0\}$ and
$\Rpre_{\mathrm{desc}}(1)=\{x_1,y_1\}$, and let every deeper cylinder below
$0$ or $1$ inherit the corresponding two-point set with identity restrictions.
The two restriction maps from $\varepsilon$ send $g$ to $x_0$ and $x_1$.
The pair overlap $C_0\cap C_1$ is empty, so every pair of local sections on the
cover is automatically compatible.  In particular, the family $(y_0,y_1)$ is
compatible on the cover
$\{0,1\to\varepsilon\}$ but does not descend to a global section.
\item \emph{Non-separatedness}: let
$\Rpre_{\mathrm{ns}}(\varepsilon)=\{u,v\}$ and
$\Rpre_{\mathrm{ns}}(b)=\{*\}$ for every nonempty word $b$, with both
restrictions from $\varepsilon$ constant.  Then $u\neq v$, but their
restrictions to the cover $\{0,1\to\varepsilon\}$ agree.
\end{enumerate}
This concrete binary site keeps the geometry fixed while isolating the three
internal presheaf-theoretic mechanisms.
\end{example}

\subsection{Intrinsic Comparison Torsors}
\label{subsec:comparison-torsors}

Let $\mathfrak{U}=\{b_i\to a\}_{i=1}^n$ be a finite cylinder-algebra cover.  Because
the cylinder algebra is closed under finite intersections
(Definition~\ref{def:cylinder-algebra}), every nonempty overlap
$C_{b_{i_0}}\cap\cdots\cap C_{b_{i_q}}$ is itself a cylinder
$C_{b_{i_0\dots i_q}}$ refining each $b_{i_r}$.  For an object
$s_i\in\Ob\Rpre(b_i)$ and such an overlap we abbreviate
\[
s_{i|i_0\dots i_q}
:=
\rho_{b_i\leftarrow b_{i_0\dots i_q}}(s_i)\in\Ob\Rpre(b_{i_0\dots i_q}).
\]
In particular, for a pair overlap $b_{ij}$ we obtain the comparison groupoid
\[
\Isom_{ij}:=
\Isom_{\Rpre(b_{ij})}\bigl(s_{j|ij},s_{i|ij}\bigr),
\]
and when this set is nonempty it is acted on simply transitively by the
automorphism group of the target object $s_{i|ij}$ via postcomposition.

\begin{definition}[Abelian band along a cover]\label{def:abelian-band}
Fix local objects $\{s_i\in\Ob\Rpre(b_i)\}_{i=1}^n$.  An \emph{abelian band}
for $\{s_i\}$ consists of an abelian group $A$ together with isomorphisms
\[
\vartheta_{i,V}\colon \Aut_{\Rpre}(s_{i|V})\xrightarrow{\;\cong\;} A
\]
for every nonempty refinement $V\to b_i$ obtained from a finite intersection
of members of $\mathfrak{U}$, such that $\vartheta_{i,V}$ is natural under
further restriction: if $V'\to V$ is induced by an additional intersection,
then $\vartheta_{i,V'}\circ \rho_{V\leftarrow V'}=\vartheta_{i,V}$.  Equivalently,
the family $\{\Aut_{\Rpre}(s_{i|V})\}$ is identified with the constant band $A$
along the cover.
\end{definition}

Under this hypothesis every comparison set inherits a canonical $A$-torsor
structure: for $\xi\in \Isom_{ij}$ and $a\in A$ we set
\[
a\cdot \xi := \vartheta_{i,b_{ij}}^{-1}(a)\circ \xi.
\]
This action is free and transitive, and it is compatible with composition on
triple overlaps because the identifications $\vartheta_{i,V}$ are natural.
Consequently the torsors and their composition maps required for the defect
construction are intrinsic to $\Rpre$ as soon as an abelian band has been
specified; no extra comparison datum has to be prescribed externally.
Moreover, for all $a,b\in A$ and compatible comparison elements we have
\begin{equation}\label{eq:torsor-additivity}
m_{ijk}(a\cdot \xi,b\cdot \eta) = (a+b)\cdot m_{ijk}(\xi,\eta),
\end{equation}
since the action is given by postcomposition with automorphisms of the target.

\subsection{The \v{C}ech Complex of a Cylinder-Algebra Cover}
\label{subsec:cech-complex}

The groupoid-valued prestack $\Rpre$ therefore produces canonical torsors of
comparisons on every overlap once an abelian band is fixed.  To speak about a
\v{C}ech defect, we now review the underlying cochain complex.  The general
framework is classical; see \cite[Tag~01FG]{StacksProject} for the abstract
\v{C}ech complex, \cite[Chapter~V]{Bredon1997SheafTheory} for
sheaf-cohomological background, and
\cite[\S IV.3]{Giraud1971CohomologieNonAbelienne} for the banded (possibly
non-abelian) refinement.

\begin{lemma}[Normalized \v{C}ech complex of a finite cylinder-algebra cover]
\label{lem:cech-complex-formal}
Let $\mathfrak{U}=\{b_i\to a\}_{i=1}^n$ be a finite cylinder-algebra cover of
$a\in\Aadm$, and let $A$ be an abelian group.  Define the \emph{ordered cover
nerve} $N(\mathfrak{U})$ as the abstract simplicial complex on
$\{1,\dots,n\}$ whose $q$-simplices are the strictly increasing tuples
\[
N_q(\mathfrak{U})
:=
\left\{
(i_0,\dots,i_q): 1\le i_0<\cdots<i_q\le n,\ 
C_{b_{i_0}}\cap\cdots\cap C_{b_{i_q}}\neq\varnothing
\right\}.
\]
Define the normalized \v{C}ech cochain groups
\[
\check{C}^q(\mathfrak{U},A)
:=
\prod_{(i_0,\dots,i_q)\in N_q(\mathfrak{U})} A.
\]
The coboundary $\delta^q\colon \check{C}^q\to \check{C}^{q+1}$ is given by
\[
(\delta^q f)_{i_0\dots i_{q+1}}
:=
\sum_{s=0}^{q+1} (-1)^s\, f_{i_0\dots\widehat{i_s}\dots i_{q+1}},
\]
for $i_0<\cdots<i_{q+1}$, where $\widehat{i_s}$ denotes omission.  Then
$\delta^{q+1}\circ\delta^q=0$, and
\[
\check{H}^q(\mathfrak{U},A)
:=
\ker\delta^q / \im\delta^{q-1}
\]
is the $q$-th normalized \v{C}ech cohomology group of $\mathfrak{U}$ with
coefficients in $A$.  In particular, a $2$-cocycle is a family
$(\alpha_{ijk})_{i<j<k}\in\check{C}^2(\mathfrak{U},A)$ satisfying
\begin{equation}\label{eq:cocycle-condition}
\alpha_{jkl} - \alpha_{ikl} + \alpha_{ijl} - \alpha_{ijk} = 0
\end{equation}
for every ordered quadruple $i<j<k<l$ with
$C_{b_i}\cap C_{b_j}\cap C_{b_k}\cap C_{b_l}\neq\varnothing$, and two
$2$-cocycles $\alpha,\alpha'$ are
cohomologous when $\alpha-\alpha'=\delta^1 h$ for some
$1$-cochain $(h_{ij})_{i<j}\in\check{C}^1(\mathfrak{U},A)$.
\end{lemma}

\begin{proof}
This is the standard simplicial or normalized \v{C}ech complex of an ordered
finite nerve.  The identity $\delta^{q+1}\circ\delta^q=0$ is the usual
alternating-sum cancellation, and the remaining assertions are definitions.
See \cite{HatcherAlgebraicTopology} for simplicial cochains and
\cite[Tag~01FG]{StacksProject} for the \v{C}ech formulation.
\end{proof}

\subsection{\v{C}ech-Type Gluing Obstructions}

\begin{definition}[Intrinsic $A$-banded comparison datum and \v{C}ech defect]
\label{def:prefix-site-h2-gluing-obstruction}
Fix an admissible address $a\in\Aadm$, a cylinder-algebra cover
$\mathfrak{U}=\{b_i\to a\}_{i=1}^n$, and an abelian group $A$.  Suppose
$\{s_i\in\Ob\Rpre(b_i)\}$ is a family of local sections equipped with an
abelian band in the sense of Definition~\ref{def:abelian-band}.  For every
nonempty pair overlap $b_{ij}$ we form the comparison torsor
\[
\Comp_{ij}
:=
\Isom_{\Rpre(b_{ij})}(s_{j|ij},s_{i|ij}),
\]
viewed as an $A$-torsor via the action $a\cdot\xi
:=\vartheta_{i,b_{ij}}^{-1}(a)\circ\xi$.  On each nonempty ordered triple
overlap the composition map
\[
m_{ijk}\colon \Comp_{ij}\times\Comp_{jk}\longrightarrow \Comp_{ik},
\qquad
m_{ijk}(\xi,\eta):=\xi\circ\eta,
\]
is $A$-equivariant and associative by construction
\cite[\S IV.3]{Giraud1971CohomologieNonAbelienne}.  Choosing elements
$\gamma_{ij}\in \Comp_{ij}$ for all nonempty ordered pair overlaps therefore
produces the data previously listed.

Because $\Comp_{ik}$ is an $A$-torsor, on each nonempty ordered triple overlap
there is a unique element $\alpha_{ijk}\in A$ such that
\begin{equation}\label{eq:defect-via-composition}
m_{ijk}(\gamma_{ij},\gamma_{jk})=\alpha_{ijk}\cdot \gamma_{ik}.
\end{equation}
The family $(\alpha_{ijk})$ is the associated \emph{\v{C}ech defect}.  The
associativity of the composition maps guarantees that $(\alpha_{ijk})$
satisfies the cocycle condition~\eqref{eq:cocycle-condition} on every
$4$-fold overlap, so it defines a class
$[\alpha]\in\check{H}^2(\mathfrak{U},A)$.
\end{definition}

The next lemma records the effect of changing representatives.

\begin{lemma}[Change of trivialization]\label{lem:change-trivialization}
Let $\beta=(\beta_{ij})_{i<j}\in\check{C}^1(\mathfrak{U},A)$ and set
$\gamma'_{ij}:=\beta_{ij}\cdot \gamma_{ij}$.  Then the defect computed from the
family $\{\gamma'_{ij}\}$ is the cohomologous cocycle
$\alpha'=\alpha+\delta^1\beta$.
\end{lemma}

\begin{proof}
Equation~\eqref{eq:torsor-additivity} gives
\[
m_{ijk}(\gamma'_{ij},\gamma'_{jk})
=
m_{ijk}(\beta_{ij}\cdot\gamma_{ij},\beta_{jk}\cdot \gamma_{jk})
=
(\beta_{ij}+\beta_{jk})\cdot m_{ijk}(\gamma_{ij},\gamma_{jk}).
\]
Substituting~\eqref{eq:defect-via-composition} produces
\[
m_{ijk}(\gamma'_{ij},\gamma'_{jk})
=
(\beta_{ij}+\beta_{jk}+\alpha_{ijk})\cdot \gamma_{ik}
=
(\beta_{ij}+\beta_{jk}+\alpha_{ijk}-\beta_{ik})\cdot \gamma'_{ik},
\]
which is precisely the statement that $\alpha'=\alpha+\delta^1\beta$.  Hence
the class $[\alpha]$ depends only on the banded comparison datum.
\end{proof}

\begin{proposition}[Nontrivial \v{C}ech defect prevents coherent global readout]
\label{prop:null-as-h2-obstruction}
In the setting of Definition~\ref{def:prefix-site-h2-gluing-obstruction},
suppose further that there exists a global section $s\in\Ob\Rpre(a)$ whose
restrictions to the cover are the sections $\{s_i\}$ and whose induced
comparisons on every pair overlap $(i,j)$ produce distinguished elements
$\gamma_{ij}^s\in\Comp_{ij}$ that are compatible with the composition maps:
\[
m_{ijk}(\gamma_{ij}^s,\gamma_{jk}^s)=\gamma_{ik}^s
\]
on every triple overlap.  Then the defect class $[\alpha]$ computed from the
induced elements $\gamma^s_{ij}$ is trivial.  In particular, if
$[\alpha]\neq 0$, then no such global section $s$ exists.
\end{proposition}

\begin{proof}
The global section $s$ restricts to $s_i$ on each $b_i$, and on every pair
overlap the induced comparison between $s_j$ and $s_i$ via $s$ gives a
distinguished element $\gamma_{ij}^s$.  On every triple overlap these compose
coherently:
\[
m_{ijk}(\gamma_{ij}^s,\gamma_{jk}^s)=\gamma_{ik}^s.
\]
By Definition~\ref{def:prefix-site-h2-gluing-obstruction}, this means
$\alpha_{ijk}=0$ on every triple overlap.  Hence the resulting defect cocycle
is zero, so $[\alpha]=0$.
\end{proof}

\section{Visible Quotients and Quantitative Recovery}
\label{sec:visible-quotient}

Let $A$ be a finite abelian group, $\mathfrak{U}$ a finite cylinder-algebra cover, and
$[\alpha]\in\check{H}^2(\mathfrak{U},A)$ a defect class as in
Section~\ref{sec:descent}.  We always regard $A$ as the automorphism band of
the chosen local sections on $\mathfrak{U}$ furnished by
Definition~\ref{def:abelian-band}.

\smallskip

\noindent\textbf{Standing hypothesis.}
Throughout this section we assume that the ordered nerve
$N(\mathfrak{U})$ has torsion-free first homology
$H_1\bigl(N(\mathfrak{U});\ZZ\bigr)$.  This holds, for example, when the
nerve is a tree or a finite contractible complex, and it removes the potential
$\operatorname{Ext}\bigl(H_1(N(\mathfrak{U})),{-}\bigr)$ contribution in
degree~$2$ cohomology.  Because $N(\mathfrak{U})$ is a finite simplicial
complex, normalized \v{C}ech cohomology with constant coefficients agrees with
simplicial cohomology, and the universal coefficient theorem
\cite[Theorem~3.2]{HatcherAlgebraicTopology} identifies
$\check{H}^2(\mathfrak{U},B)$ with $\Hom(H_2(N(\mathfrak{U})),B)$ for every
abelian group $B$, so injections on coefficients induce injections on
$\check{H}^2$.

\begin{remark}[Role of the torsion-free hypothesis]\label{rem:torsion-free-role}
Part~\textup{(2)} of Proposition~\ref{thm:intrinsic-visible-quotient} uses two
facts supplied by this hypothesis: (i) normalized \v{C}ech cohomology agrees
with simplicial cohomology for constant coefficients on $N(\mathfrak{U})$, and
(ii) the universal coefficient theorem expresses $\check{H}^2(\mathfrak{U},B)$
as $\Hom(H_2(N(\mathfrak{U})),B)$ without an
$\operatorname{Ext}^1(H_1(N(\mathfrak{U})),B)$ term.  Consequently a
coefficient inclusion $j\colon C\hookrightarrow D$ induces an injective map
$j_*\colon \check{H}^2(\mathfrak{U},C)\to \check{H}^2(\mathfrak{U},D)$.  If
$H_1(N(\mathfrak{U}))$ carried torsion, the $\operatorname{Ext}^1$ term could
survive and be annihilated by $j_*$, so the factorization argument below would
require additional work; we therefore leave the torsionful case to future
extensions.
\end{remark}

The next proposition records the routine universal construction obtained by
quotienting $A$ by the intersection of all kernels through which the class
vanishes.

\begin{proposition}[Visible quotient associated to a defect class]
\label{thm:intrinsic-visible-quotient}
For each subgroup $N\le A$, let $q_N\colon A\twoheadrightarrow A/N$ be the
quotient map, and define
\[
\mathcal{N}([\alpha])
:=
\bigl\{
N\le A:\ (q_N)_*[\alpha]=0\ \text{in}\ \check{H}^2(\mathfrak{U},A/N)
\bigr\}.
\]
Because $A$ is finite, the collection $\mathcal{N}([\alpha])$ is finite; write
\[
\mathcal{N}([\alpha])=\{N_1,\dots,N_m\}.
\]
Define
\[
\Phi
:=
\bigl(q_{N_1},\dots,q_{N_m}\bigr)\colon
A\longrightarrow \prod_{r=1}^m A/N_r,
\qquad
K:=\ker\Phi=\bigcap_{r=1}^m N_r,
\qquad
\Avis:=\im\Phi.
\]
Then, under the standing torsion-free hypothesis on $N(\mathfrak{U})$:
\begin{enumerate}[label=\textup{(\arabic*)}]
\item The pushed-forward class vanishes in $\Avis$: the canonical surjection
$q\colon A\twoheadrightarrow \Avis$ satisfies $q_*[\alpha]=0$ in
$\check{H}^2(\mathfrak{U},\Avis)$.
\item \textup{(Universal property.)} If $\varphi\colon A\to B$ is any
homomorphism with $\varphi_*[\alpha]=0$, then $\varphi$ factors uniquely
through $q$: there exists a unique $\bar\varphi\colon \Avis\to B$ with
$\varphi=\bar\varphi\circ q$.
\item The quotient map $q$ identifies $\Avis$ with $A/K$.
\end{enumerate}
\end{proposition}

\begin{proof}
Because $A$ is finite, it has only finitely many subgroups, so the set
$\mathcal{N}([\alpha])$ is finite.  For each $r$, the class
$(q_{N_r})_*[\alpha]$ vanishes by definition.  Since
\[
\check{C}^q\!\Bigl(\mathfrak{U},\prod_{r=1}^m A/N_r\Bigr)
\cong
\prod_{r=1}^m \check{C}^q(\mathfrak{U},A/N_r)
\]
and the \v{C}ech coboundaries respect this product decomposition, the same is
true in cohomology:
\[
\check{H}^2\!\Bigl(\mathfrak{U},\prod_{r=1}^m A/N_r\Bigr)
\cong
\prod_{r=1}^m \check{H}^2(\mathfrak{U},A/N_r).
\]
Therefore $\Phi_*[\alpha]=0$.  Let
$i\colon \Avis\hookrightarrow \prod_{r=1}^m A/N_r$ be the inclusion.  Then
$i\circ q=\Phi$, so
\[
i_*\bigl(q_*[\alpha]\bigr)=\Phi_*[\alpha]=0.
\]
By the universal coefficient theorem and the standing torsion-free hypothesis,
$\check{H}^2(\mathfrak{U},B)\cong \Hom(H_2(N(\mathfrak{U})),B)$ for every
coefficient group $B$, hence $i_\ast$ identifies with the injective map
$\Hom(H_2(N(\mathfrak{U})),\Avis)\hookrightarrow \Hom(H_2(N(\mathfrak{U})),
\prod_r A/N_r)$.  Consequently $q_*[\alpha]=0$, proving~\textup{(1)}.

For~\textup{(2)}, let $\varphi\colon A\to B$ satisfy $\varphi_*[\alpha]=0$ and
factor it as
\[
A\xrightarrow{\ \bar\varphi\ }\im(\varphi)\xrightarrow{\ j\ } B,
\]
where $\bar\varphi$ is the canonical surjection and $j$ is the inclusion.
The induced maps on cohomology fit into the commutative diagram
\[
\begin{array}{ccc}
\check{H}^2(\mathfrak{U},A) & \xrightarrow{\ \bar\varphi_*\ } &
\check{H}^2(\mathfrak{U},\im(\varphi))\\
& \searrow_{\varphi_*} & \downarrow{j_*}\\
&& \check{H}^2(\mathfrak{U},B).
\end{array}
\]
Write $N:=\ker(\varphi)=\ker(\bar\varphi)$.  The torsion-free hypothesis on
$N(\mathfrak{U})$ identifies $\check{H}^2(\mathfrak{U},-)$ with the
Hom-functor $\Hom(H_2(N(\mathfrak{U})),{-})$, so $j_*$ is injective for every
injective coefficient map.  Since $\varphi_*[\alpha]=0$, the diagram gives
\[
0=\varphi_*[\alpha]=j_*\bigl(\bar\varphi_*[\alpha]\bigr)
\qquad\Longrightarrow\qquad
\bar\varphi_*[\alpha]=0.
\]
Identifying $\im(\varphi)$ with $A/N$ by the first isomorphism theorem, the
map $\bar\varphi$ becomes the quotient map $q_N\colon A\twoheadrightarrow A/N$,
so
$N\in\mathcal{N}([\alpha])$.  Because $K$ is the intersection of all such
subgroups, $K\subseteq N=\ker\varphi$.  By the universal property of the
quotient $A\twoheadrightarrow A/K$, there exists a unique homomorphism
$\widetilde{\varphi}\colon A/K\to B$ with $\varphi=\widetilde{\varphi}\circ
\pi_K$, where $\pi_K$ denotes the quotient map.  The first isomorphism theorem
identifies $A/K$ canonically with $\Avis=\im\Phi$, and under this
identification $\pi_K=q$.  Hence $\varphi$ factors uniquely through $q$.

Part~\textup{(3)} is exactly the first isomorphism theorem applied to
$q\colon A\twoheadrightarrow\Avis$.
\end{proof}

\begin{remark}\label{rem:cocycle-vs-pointwise}
The universal property in Proposition~\ref{thm:intrinsic-visible-quotient}
is intrinsic to the class $[\alpha]$.  In particular, it does not require a
choice of cocycle representative or a subgroup generated by cocycle values.
The subgroup $K$ is defined directly from the quotients on which the class
vanishes, and $\Avis$ is the corresponding initial quotient.  The notation
already records the dependence on the fixed cover and class.
\end{remark}

\begin{lemma}[Register lower bound for full-resolution recovery]
\label{thm:prefix-site-register-lower-bound}
Assume $A$ is finite and let $\Avis$ be the visible
quotient attached to the fixed cover and class.  Suppose one wishes to recover
the full $A$-resolution from $\Avis$ by adjoining a finite register set $R$ in
such a way that there exists an injection
\[
A\hookrightarrow \Avis\times R.
\]
Then
\[
\lvert R\rvert \ge \lvert K\rvert.
\]
This bound is sharp: equality can be attained with $R=K$.
\end{lemma}

\begin{proof}
Every fiber of the quotient map $A\twoheadrightarrow \Avis$ is a coset of
$K$, hence has cardinality $\lvert K\rvert$.  If
$A\hookrightarrow \Avis\times R$, then restricting the injection to any one of
these fibers yields an injection into $R$.  Therefore
\[
\lvert R\rvert\ge \lvert K\rvert.
\]

For sharpness, take $R:=K$.  Choose a set-theoretic section
$s:\Avis\to A$ of the quotient map.  Then every $a\in A$ can be written
uniquely as
\[
a = s(\bar a)+h
\qquad
(\bar a\in\Avis,\ h\in K),
\]
and the map $a\mapsto (\bar a,h)$ is injective.
\end{proof}

\begin{definition}[Character-based spectral readout]
\label{def:character-based-spectral-readout}
Let $\chi\in \widehat A:=\Hom(A,\TT)$.  Pushing an $A$-banded comparison datum
forward along $\chi$ produces a $\TT$-banded comparison datum, and hence a
pushed-forward defect class $\chi_*[\alpha]$.  The resulting $\TT$-valued
descent problem is called the \emph{character-based spectral readout}
associated to $\chi$.  It is said to be \emph{spectrally definable} if
$\chi_*[\alpha]=0$ in
$\check{H}^2(\mathfrak{U},\TT)$.
\end{definition}

\begin{proposition}[Spectral definability occurs only on the visible quotient]
\label{thm:prefix-site-spectral-definability-avis}
Let $\widehat A:=\Hom(A,\TT)$ be the Pontryagin dual, and define the
annihilator
\[
K^\perp
:=
\{\chi\in \widehat A:\ \chi(k)=1\ \text{for all }k\in K\}.
\]
Then $K^\perp\cong \widehat{\Avis}$, and the character-based spectral
readout of Definition~\ref{def:character-based-spectral-readout} is spectrally
definable if and only if the chosen character lies in $K^\perp$.
\end{proposition}

\begin{proof}
Because $A$ is finite abelian, a character of $A$ factors through the quotient
$A\twoheadrightarrow \Avis\cong A/K$ if and only if it is trivial on $K$.
This identifies $\widehat{\Avis}$ with the subgroup
$K^\perp\le \widehat A$.

Now let $\chi\in\widehat A$.  By
Proposition~\ref{thm:intrinsic-visible-quotient}, the pushed-forward class
$\chi_*[\alpha]\in\check{H}^2(\mathfrak{U},\TT)$ vanishes if and only if $\chi$
factors through $q\colon A\twoheadrightarrow \Avis$.  By the first paragraph,
that happens exactly when $\chi$ is trivial on $K$.  Hence the spectral
readout is definable if and only if $\chi\in K^\perp$.
\end{proof}

\section{Appendix: A Symbolic Cover with Nontrivial Defect}
\label{sec:forced-obstruction}

The constructions of Sections~\ref{sec:descent}--\ref{sec:visible-quotient}
apply once one has fixed local objects and an abelian band on a finite
cylinder-algebra cover.  In applications, the main difficulty is to generate
that datum canonically from an actual readout semantics.  As a concrete model
computation, we now work out a symbolic example in which both the cover nerve
and a nontrivial defect class can be written down explicitly.  The present
section is deliberately positioned as a combinatorial appendix: it specializes
the general formalism to a prescribed banded datum on the ordered nerve, and it
is included as a sanity check rather than as a semantically canonical readout
model.

\subsection{A Tetrahedral Cylinder-Algebra Cover}

Work on the one-sided full binary shift
\[
X:=\{0,1\}^{\NN},
\]
with $\Aadm=\{0,1\}^{<\infty}$ and the usual cylinder sets
$C_w\subseteq X$ of sequences beginning with the word $w$
\cite[\S 1.2]{LindMarcus1995SymbolicDynamics}.  Let
\[
D_1:=C_{00},\qquad D_2:=C_{01},\qquad D_3:=C_{10},\qquad D_4:=C_{11}.
\]
These four depth-$2$ cylinders are pairwise disjoint and cover $X$.

Define four objects of the cylinder algebra by omitting one depth-$2$ cylinder
at a time:
\[
\begin{aligned}
U_1&:=D_2\cup D_3\cup D_4, &
U_2&:=D_1\cup D_3\cup D_4,\\
U_3&:=D_1\cup D_2\cup D_4, &
U_4&:=D_1\cup D_2\cup D_3.
\end{aligned}
\]

\begin{proposition}[Tetrahedral cover of the full shift]
\label{prop:tetrahedral-cover}
Let $\mathfrak{U}=\{U_1,U_2,U_3,U_4\}\to \varepsilon$.  Then:
\begin{enumerate}[label=\textup{(\arabic*)}]
\item $\mathfrak{U}$ is a cylinder-algebra cover of the root object $C_\varepsilon=X$;
\item for $i\neq j$, the pair overlap $U_i\cap U_j$ is the union of exactly two
depth-$2$ cylinders;
\item for distinct $i,j,k$, the triple overlap $U_i\cap U_j\cap U_k$ is the
single depth-$2$ cylinder $D_\ell$, where $\ell$ is the unique index not in
$\{i,j,k\}$;
\item the quadruple overlap $U_1\cap U_2\cap U_3\cap U_4$ is empty.
\end{enumerate}
Consequently the nerve $N(\mathfrak{U})$ is the boundary of a tetrahedron.
\end{proposition}

\begin{proof}
Each $U_i$ is a finite union of cylinders, hence an object of
$\mathsf{Pref}(\Aadm)$.  Since every $D_r$ lies in exactly three of the
$U_i$'s, the union $U_1\cup U_2\cup U_3\cup U_4$ is all of $X$, proving~(1).

For $i\neq j$, the set $U_i\cap U_j$ is obtained by removing the two omitted
depth-$2$ cylinders, so it is the union of the remaining two, proving~(2).  If
$i,j,k$ are distinct, then exactly one depth-$2$ cylinder survives in all three
sets, giving~(3).  Finally, every $D_r$ is omitted by one of the $U_i$'s, so
no point lies in all four sets; this is~(4).  The description of the nerve is
therefore immediate.
\end{proof}

\subsection{A Prescribed \texorpdfstring{$\ZZ/2\ZZ$}{Z/2Z}-Valued 2-Cocycle}


\begin{figure}[ht]
\centering
\begin{tikzpicture}[scale=1.15,every node/.style={circle,fill=black!70,inner sep=1.7pt}]
\node[label=above:$U_1$] (u1) at (0,1.9) {};
\node[label=below left:$U_2$] (u2) at (-1.5,0) {};
\node[label=below right:$U_3$] (u3) at (1.5,0) {};
\node[label=below:$U_4$] (u4) at (0,-1.4) {};
\draw (u1)--(u2)--(u3)--(u1);
\draw (u2)--(u4)--(u3);
\draw[dashed] (u1)--(u4);
\draw[dashed] (u2)--(u3);
\end{tikzpicture}
\caption{Planar projection of the tetrahedral nerve $N(\mathfrak{U})$ from Proposition~\ref{prop:tetrahedral-cover}.  Dashed edges denote faces hidden in this drawing.}
\label{fig:tetrahedral-nerve}
\end{figure}

Figure~\ref{fig:tetrahedral-nerve} depicts the ordered nerve that drives the
combinatorial cocycle computation below.

The comparison data in this subsection are prescribed directly on the ordered
nerve.  No claim is made that they are canonically induced by a readout
prestack; the purpose is only to supply a complete finite-cover computation in
the formalism of Section~\ref{sec:descent}.  Concretely, we choose compatible
local objects whose automorphism band identifies with the constant group
$A=\ZZ/2\ZZ$ so that every comparison torsor is literally a copy of $A$.

\begin{definition}[Prescribed comparison datum on the tetrahedral cover]
\label{def:tetrahedral-prescribed-datum}
Let $A=\ZZ/2\ZZ$ and let $\mathfrak{U}$ be the tetrahedral cover from
Proposition~\ref{prop:tetrahedral-cover}.  For each ordered edge
$1\le i<j\le 4$, set
\[
\Comp_{ij}:=A
\]
viewed as an $A$-torsor by translation, and choose the comparison element
$\gamma_{ij}:=0$.  On the four ordered faces define
\[
m_{ijk}(x,y):=x+y+\alpha_{ijk}
\qquad
(1\le i<j<k\le 4),
\]
where
\[
\alpha_{123}=1,
\qquad
\alpha_{124}=\alpha_{134}=\alpha_{234}=0.
\]
\end{definition}

\begin{proposition}[A nontrivial prescribed defect on the tetrahedral cover]
\label{thm:forced-obstruction}
Let $A=\ZZ/2\ZZ$, let $\mathfrak{U}$ be the tetrahedral cover from
Proposition~\ref{prop:tetrahedral-cover}, and consider the prescribed
comparison datum of Definition~\ref{def:tetrahedral-prescribed-datum}.
Then:
\begin{enumerate}[label=\textup{(\arabic*)}]
\item these data form an $A$-banded comparison datum on $\mathfrak{U}$;
\item the resulting defect class
$[\alpha]\in\check{H}^2(\mathfrak{U},\ZZ/2\ZZ)$ is nonzero;
\item the visible quotient is $\Avis=0$, and any register set
$R$ with an injection $A\hookrightarrow \Avis\times R$ satisfies
$|R|\ge 2$.
\end{enumerate}
\end{proposition}

\begin{proof}
Because the quadruple overlap is empty by
Proposition~\ref{prop:tetrahedral-cover}, the associativity condition on
quadruple overlaps is vacuous.  Hence the stated torsors, composition maps, and
comparison elements define an $A$-banded comparison datum, proving~\textup{(1)}.

By Proposition~\ref{prop:tetrahedral-cover}, the ordered nerve
$N(\mathfrak{U})$ is the simplicial boundary $\partial\Delta^3$ of a tetrahedron
\cite{HatcherAlgebraicTopology}.  Its ordered $2$-simplices are
\[
(123),\ (124),\ (134),\ (234),
\]
and there is no $3$-simplex, so every $2$-cochain is automatically a cocycle.
Let
\[
z:=(234)-(134)+(124)-(123)
\]
be the usual fundamental $2$-cycle of $\partial\Delta^3$.  For any $1$-cochain
$h$,
\[
\langle \delta^1 h,z\rangle = \langle h,\partial z\rangle = 0,
\]
because $\partial z=0$.  On the other hand, the present defect cochain satisfies
\[
\langle \alpha,z\rangle
=
\alpha_{123}+\alpha_{124}+\alpha_{134}+\alpha_{234}
=
1.
\]
Therefore $\alpha$ is not a coboundary, so $[\alpha]\neq 0$.  This proves
\textup{(2)}.

For \textup{(3)}, note that $A=\ZZ/2\ZZ$ has only two subgroups: $0$ and $A$.
The identity quotient $A\twoheadrightarrow A/0=A$ does not kill the class
because $[\alpha]\neq 0$, whereas the zero quotient $A\twoheadrightarrow A/A=0$
does.  Hence $\mathcal{N}([\alpha])=\{A\}$ in the notation of
Proposition~\ref{thm:intrinsic-visible-quotient}.  Therefore
$K=A$, so $\Avis\cong A/K=0$.  The register bound from
Lemma~\ref{thm:prefix-site-register-lower-bound} then gives
$|R|\ge |K|=2$.
\end{proof}

\begin{example}[Interpretation of the tetrahedral model]
\label{ex:tetrahedral-cover}
The cover in Proposition~\ref{prop:tetrahedral-cover} is obtained by deleting
one depth-$2$ cylinder at a time.  Pair overlaps remember two depth-$2$
cylinders, while triple overlaps isolate a single depth-$2$ cylinder.  The
nonzero face value $\alpha_{123}=1$ may therefore be read as a phase mismatch
localized on the cylinder $D_4=C_{11}$.  Because the cover nerve is a
tetrahedral boundary rather than a simplex, this local mismatch survives in
cohomology and forces the visible quotient to collapse to $0$.  The example
provides a compact symbolic test case on which the obstruction and register
bound can be computed exactly.  It should be read as a standalone combinatorial
obstruction calculation on a symbolic cover rather than as a semantically
canonical readout model.
\end{example}

\begin{example}[A partially visible \texorpdfstring{$\ZZ/4\ZZ$}{Z/4Z} computation]
\label{ex:tetrahedral-z4}
Keep the same ordered tetrahedral nerve but take $A=\ZZ/4\ZZ$ and prescribe
the $2$-cocycle
\[
\alpha_{123}=2,
\qquad
\alpha_{124}=\alpha_{134}=\alpha_{234}=0.
\]
Evaluating on the fundamental cycle
$z=(234)-(134)+(124)-(123)$ gives
\[
\langle \alpha,z\rangle = 2\in \ZZ/4\ZZ,
\]
so $[\alpha]\neq 0$ in $\check{H}^2(\mathfrak{U},\ZZ/4\ZZ)$.  The subgroups of
$\ZZ/4\ZZ$ are $0$, $2\ZZ/4\ZZ$, and $\ZZ/4\ZZ$.  The identity quotient does
not kill the class, while the quotient
\[
q\colon \ZZ/4\ZZ\twoheadrightarrow (\ZZ/4\ZZ)/(2\ZZ/4\ZZ)\cong \ZZ/2\ZZ
\]
does because $q(\alpha_{123})=0$.  Hence the kernels killing the class are
exactly $2\ZZ/4\ZZ$ and $\ZZ/4\ZZ$, so
\[
K=2\ZZ/4\ZZ,
\qquad
\Avis\cong (\ZZ/4\ZZ)/(2\ZZ/4\ZZ)\cong \ZZ/2\ZZ.
\]
This is the basic case in which the visible quotient and the hidden kernel are
both nontrivial.  Lemma~\ref{thm:prefix-site-register-lower-bound} again gives
the lower bound $|R|\ge 2$ for full $\ZZ/4\ZZ$-recovery from the visible
quotient.
\end{example}

\section{A Semantically Grounded GHZ Example}
\label{sec:semantic-example}

We now tie Sections~\ref{sec:measurable}--\ref{sec:visible-quotient} together
via a worked example drawn directly from the GHZ contextuality scenario of
\cite{AbramskyBrandenburger2011,AbramskyEtAl2015ContextualityCohomologyParadox}.
The measurable GHZ experiment supplies canonical comparison torsors, so the
resulting defect is intrinsic rather than prescribed.

\subsection{Coding the GHZ experiment as a prefix cover}

Each GHZ run chooses measurements $s_i\in\{X,Y\}$ for parties
$i=1,2,3$.  Encode $X$ as $0$ and $Y$ as $1$, and push this alphabet through
the coding map of Proposition~\ref{prop:measurable-coding-prefix-bridge}.
The four GHZ contexts used in the parity argument are
\[
w_1=001\ (XXY),\quad
w_2=010\ (XYX),\quad
w_3=100\ (YXX),\quad
w_4=111\ (YYY).
\]
As in Section~\ref{sec:forced-obstruction}, define the context cover
\[
\mathfrak{U}_{\mathrm{GHZ}}
:=\bigl\{U_i\to \varepsilon\bigr\}_{i=1}^4,
\qquad
U_i:=\bigcup_{j\ne i} C_{w_j}.
\]
Triple overlaps isolate the single omitted context, and the quadruple overlap
is empty, so the ordered nerve is the boundary of a tetrahedron.

\subsection{Local readouts from GHZ parity constraints}

Let $\sigma_i\in\{\pm 1\}$ denote the outcome on party $i$.  The GHZ state
predicts
\[
\sigma_1\sigma_2\sigma_3=
\begin{cases}
 +1,& \text{for }XXY,XYX,YXX,\\
 -1,& \text{for }YYY.
\end{cases}
\]
Transport $\{\pm1\}$ to $\ZZ/2$ via $+1\mapsto 0$, $-1\mapsto 1$, and write
$v_i\in(\ZZ/2)^3$ for the bit triple of a context.  The solution sets
\[
S_i
:=
\Bigl\{v\in(\ZZ/2)^3:\ v_1+v_2+v_3\equiv \delta_i \pmod 2\Bigr\},
\qquad
\delta_i=
\begin{cases}
0,& i=1,2,3,\\
1,& i=4,
\end{cases}
\]
form torsors for the even-weight subgroup
\[
A:=
\Bigl\{a\in(\ZZ/2)^3:\ a_1+a_2+a_3\equiv 0\Bigr\}\cong(\ZZ/2)^2,
\]
acting by componentwise addition.  The readout prestack $\Rpre$ assigns to each
$U_i$ the groupoid of functions that pick an element of $S_j$ for every context
$C_{w_j}\subseteq U_i$.  Restricting to a triple overlap simply evaluates the
chosen assignment on the unique surviving context, so the comparison torsor on
that overlap is canonically identified with $A$.

Fix the lexicographically smallest GHZ assignment on each context:
\[
\begin{aligned}
s^{(1)}(XXY)&=(0,0,0),&
s^{(2)}(XYX)&=(0,0,0),\\
s^{(3)}(YXX)&=(0,0,0),&
s^{(4)}(YYY)&=(0,0,1).
\end{aligned}
\]
Evaluating
Definition~\ref{def:prefix-site-h2-gluing-obstruction} on these choices produces
the defect cocycle
\[
\alpha_{123}=(1,1,0)\in A,
\qquad
\alpha_{124}=\alpha_{134}=\alpha_{234}=0,
\]
because the three even contexts enforce parity $0$ while the omitted $YYY$
context enforces parity $1$.  Any other family of GHZ assignments differs by an
$A$-valued $1$-cochain, so $\alpha$ is well defined up to coboundary.

\subsection{Visible quotient and interpretation}

The ordered nerve is the tetrahedral boundary, so the fundamental $2$-cycle
$z=(234)-(134)+(124)-(123)$ evaluates to
\[
\langle \alpha,z\rangle = (1,1,0)\neq 0\in A,
\]
hence $[\alpha]\neq 0$ in $\check{H}^2(\mathfrak{U}_{\mathrm{GHZ}},A)$.  The
visible quotient from Proposition~\ref{thm:intrinsic-visible-quotient} is
therefore trivial: every homomorphism that kills $[\alpha]$ factors through
$A/A=0$.  Lemma~\ref{thm:prefix-site-register-lower-bound} shows that at least
four register states are required to recover the full $(\ZZ/2)^2$ band, and
Proposition~\ref{thm:prefix-site-spectral-definability-avis} implies that only
those characters trivial on $A$ admit spectrally definable readouts.  In other
words, the canonical GHZ parity mismatch manifests directly as the intrinsic
defect of $\mathfrak{U}_{\mathrm{GHZ}}$, providing a semantically grounded
example that threads together the measurable, symbolic, and obstruction-theoretic
layers.

\section{Prefix Inverse Limits and Separatedness}
\label{sec:inverse-limits}

\subsection{Stable Objects in a Prefix Tower}

We now restrict to inverse systems that already arise from admissible prefixes.
This is the rooted-tree or path-space situation naturally attached to the
prefix site of Section~\ref{sec:prefix-sites}.  The next proposition records
the standard compactness and uniqueness facts in that setting; see
\cite{RibesZalesskii2010ProfiniteGroups,
Wilson1998ProfiniteGroups,Munkres2000Topology,Bourbaki1989GeneralTopology}.

\begin{proposition}[Standard inverse-limit facts for prefix towers]
\label{thm:recursive-addressing-dyadic-inverse-limit-stable-object}
Let $\Sigma$ be a finite alphabet, and for each $b\ge 0$ let
$\mathcal{A}_b\subseteq \Sigma^b$ be a nonempty finite set such that the
truncation maps
\[
\pi_{b'\to b}\colon \mathcal{A}_{b'}\twoheadrightarrow \mathcal{A}_b
\]
sending a word to its prefix of length $b$ are surjective for all $b'\ge b$.
Put
\[
\Aadm:=\bigcup_{b\ge 0}\mathcal{A}_b,
\qquad
\Ainf
:=
\varprojlim_b \mathcal{A}_b
=
\left\{
(a_b)_{b\ge 0}\in \prod_{b\ge 0}\mathcal{A}_b:\
\pi_{b'\to b}(a_{b'})=a_b\ \text{for all }b'\ge b
\right\}.
\]
Equivalently,
\[
\Ainf
=
\{x\in \Sigma^{\NN}: x|_b\in \mathcal{A}_b\ \text{for all }b\ge 0\},
\]
the branch space of the rooted prefix tree determined by the sets
$\mathcal{A}_b$.
Then:
\begin{enumerate}[label=\textup{(\arabic*)}]
\item $\Ainf\neq\varnothing$;
\item the dyadic metric
\[
d_{\mathrm{dy}}(a,a')
:=
\begin{cases}
0,& a=a',\\[1mm]
2^{-b_\ast(a,a')},& a\neq a',
\end{cases}
\qquad
b_\ast(a,a')
:=
\max\{b\ge 0:\ a_b=a'_b\}
\]
makes $\Ainf$ into a compact, complete, totally disconnected ultrametric
space;
\item if $Y$ is a complete metric space and for each $a_b\in \mathcal{A}_b$ one
is given a nonempty closed set $F_b(a_b)\subseteq Y$ such that
\[
F_{b'}(a_{b'})\subseteq F_b(a_b)
\quad\text{whenever}\quad
\pi_{b'\to b}(a_{b'})=a_b
\]
and
\[
\diam\bigl(F_b(a_b)\bigr)\le C\,2^{-b}
\qquad
(\forall b,\ \forall a_b\in\mathcal{A}_b)
\]
for some constant $C>0$, then every $a=(a_b)_{b\ge 0}\in \Ainf$ determines a
unique stable object
\[
x(a)\in \bigcap_{b\ge 0} F_b(a_b).
\]
\end{enumerate}
\end{proposition}

\begin{proof}[Proof sketch]
For each $b\ge 0$, let
\[
E_b
:=
\bigcup_{w\in \mathcal{A}_b} C_w
\subseteq \Sigma^{\NN}.
\]
Each $E_b$ is a nonempty compact subset of $\Sigma^{\NN}$, and surjectivity of
the truncation maps implies $E_{b+1}\subseteq E_b$.  By compactness,
$\bigcap_{b\ge 0}E_b$ is nonempty.  A sequence lies in that intersection if and
only if all of its length-$b$ prefixes belong to $\mathcal{A}_b$, which is
equivalent to being a compatible thread in the inverse limit.  Thus
$\Ainf=\bigcap_{b\ge 0}E_b\neq\varnothing$.

The dyadic metric is the standard prefix ultrametric
\cite[\S 2.1]{RibesZalesskii2010ProfiniteGroups}.
Compactness of $\Ainf$ follows because it is a closed subset of the compact
space $\Sigma^{\NN}$.  Completeness is obtained by coordinatewise stabilization
of Cauchy sequences, and total disconnectedness follows because the basic
cylinder neighborhoods
\[
U_b(a):=\{a'\in\Ainf:\ a'_b=a_b\}
\]
are clopen and form a basis.

For the final statement, fix $a=(a_b)\in\Ainf$.  The family
$\{F_b(a_b)\}_{b\ge 0}$ is a nested sequence of nonempty closed subsets of $Y$
with diameters tending to zero.  Choose $x_b\in F_b(a_b)$.  The diameter bound
makes $(x_b)$ a Cauchy sequence, hence it converges to some $x\in Y$ because
$Y$ is complete.  Since each $F_b(a_b)$ is closed and contains all sufficiently
deep $x_{b'}$, it follows that $x\in F_b(a_b)$ for every $b$.  Uniqueness
follows from
\[
d_Y(x,y)\le \diam(F_b(a_b))\le C2^{-b}
\qquad
(\forall b),
\]
which forces $x=y$ as $b\to\infty$.
\end{proof}

\begin{example}[Dyadic intervals]
\label{ex:dyadic-intervals}
Take $\mathcal{A}_b=\{0,1\}^b$ with the bonding maps that delete the last
digit.  Let $Y=[0,1]$, and for
$a_b=(\varepsilon_1,\dots,\varepsilon_b)\in\{0,1\}^b$ define
\[
F_b(a_b):=
\left[
\sum_{j=1}^b \varepsilon_j2^{-j},
\sum_{j=1}^b \varepsilon_j2^{-j}+2^{-b}
\right].
\]
Then $F_{b'}(a_{b'})\subseteq F_b(a_b)$ whenever
$\pi_{b'\to b}(a_{b'})=a_b$, and
$\diam(F_b(a_b))=2^{-b}$.
Proposition~\ref{thm:recursive-addressing-dyadic-inverse-limit-stable-object}
recovers the fact that every compatible binary thread determines a unique point
of $[0,1]$.
\end{example}

\subsection{Stalks, Germs, and Single-Valued Readout}

Fix $a=(a_b)_{b\ge 0}\in \Ainf$.  Because each $a_b\in \Aadm$, the cylinders
$C_{a_b}$ form a cofinal neighborhood basis of the branch $a$ inside
$\mathsf{Pref}(\Aadm)$.  We define the \emph{local site at $a$} to be the full
subcategory of $\mathsf{Pref}(\Aadm)$ on all objects
$C\in\mathfrak{C}(\Aadm)$ that contain $a$.  Concretely, every such object
contains some basic cylinder $C_{a_b}$, so the directed system may be indexed
by the chain
\[
\{a_b\}_{b\ge 0},
\]
ordered by refinement (larger $b$ = finer information).

For the stalk discussion only the object-level separatedness question matters.
Accordingly we pass from the groupoid-valued prestack $\Rpre$ to its underlying
set-valued presheaf of objects $U\mapsto \Ob\Rpre(U)$, and by abuse of notation
continue to denote that presheaf by $\Rpre$.

\begin{definition}[Stalk and germ]
\label{def:stalk-germ}
The \emph{stalk} of the underlying object presheaf $\Rpre$ at $a\in\Ainf$ is the
filtered colimit
\[
\Rpre_a := \varinjlim_{b\ge 0} \Ob\Rpre(a_b),
\]
where the colimit is taken over the refinement chain
$a_0\preceq a_1\preceq a_2\preceq \cdots$.  Two elements
$s\in\Ob\Rpre(a_b)$ and $s'\in\Ob\Rpre(a_{b'})$ define the same \emph{germ} at $a$
if there exists $b''\ge b,b'$ such that
$\rho_{a_b\leftarrow a_{b''}}(s)=\rho_{a_{b'}\leftarrow a_{b''}}(s')$.  The
germ of $s$ at $a$ is denoted $[s]_a\in\Rpre_a$.
\end{definition}

For each $b\ge 0$, write
\[
U_b(a):=\{a'\in\Ainf:\ a'_b=a_b\}
\]
for the basic clopen neighborhood of $a$ at level~$b$.  If
$c\in\mathcal{A}_n$ satisfies $\pi_{n\to b}(c)=a_b$, write
$U(c):=\{a'\in\Ainf:\ a'_n=c\}\subseteq U_b(a)$.
A \emph{finite cylinder-algebra cover of $U_b(a)$} is a finite family
$\{U(c_i)\}_{i=1}^r$ with $U_b(a)=\bigcup_{i=1}^r U(c_i)$.
We write $\Rpre(U_b(a)):=\Ob\Rpre(a_b)$ and $\Rpre(U(c)):=\Ob\Rpre(c)$.

\begin{theorem}[Stalk criterion for single-valued readout]
\label{thm:recursive-addressing-readout-separatedness-null}
For a stable address $a\in\Ainf$ and each $b\ge 0$, let
\[
\lambda_b\colon \Rpre(U_b(a))\longrightarrow \Rpre_a
\]
be the natural map from sections on the basic neighborhood $U_b(a)$ to the
stalk at $a$.  Then:
\begin{enumerate}[label=\textup{(\arabic*)}]
\item For every $b\ge 0$ and every pair of sections
$s,t\in \Rpre(U_b(a))$, one has $\lambda_b(s)=\lambda_b(t)$ if and only if
there exists $b'\ge b$ such that
\[
\rho_{a_b\leftarrow a_{b'}}(s)=\rho_{a_b\leftarrow a_{b'}}(t).
\]
\item The following are equivalent.
\begin{enumerate}[label=\textup{(\alph*)}]
\item \textup{(Single-valued readout at $a$.)} For every $b\ge 0$, the map
$\lambda_b$ is injective.
\item For every $b\ge 0$ and every pair $s,t\in\Rpre(U_b(a))$, equality of
germs at $a$ forces equality of sections: if $\lambda_b(s)=\lambda_b(t)$, then
$s=t$.
\end{enumerate}
\item Assume single-valued readout at $a$.  Let
$\{U(c_i)\to U_b(a)\}_{i=1}^r$ be any finite cylinder-algebra cover and let
$s,t\in \Rpre(U_b(a))$.  If
\[
\rho_{a_b\leftarrow c_i}(s)=\rho_{a_b\leftarrow c_i}(t)
\qquad
(1\le i\le r),
\]
then $s=t$.
\end{enumerate}

Moreover, failure of a well-defined readout at $a$ has three internal forms:
\begin{enumerate}[label=\textup{(\roman*)}]
\item \emph{empty local fiber}: $\Rpre(U_b(a))=\varnothing$ for some
sufficiently deep neighborhood;
\item \emph{descent failure}: compatible local sections on a finite
cylinder-algebra cover of some $U_b(a)$ do not glue to a section over $U_b(a)$;
\item \emph{non-separatedness at $a$}: for some $b$, the map
$\lambda_b$ is not injective, equivalently two distinct sections over
$U_b(a)$ define the same germ at $a$.
\end{enumerate}
\end{theorem}

\begin{proof}
\textup{(1)} Every object of the local site at $a$ contains some basic
neighborhood $U_{b'}(a)$: if $V\in\mathfrak{C}(\Aadm)$ contains $a$, then one of
the cylinder summands in a disjoint cylinder decomposition of $V$ contains $a$,
and a sufficiently deep basic neighborhood $U_{b'}(a)$ lies inside that
summand.  Thus the chain $\{U_{b'}(a)\}_{b'\ge b}$ is cofinal among all
neighborhoods of $a$ contained in $U_b(a)$.  By the definition of the stalk as
a filtered colimit, two sections over $U_b(a)$ have the same germ at $a$ if and
only if they agree after restriction to some deeper basic neighborhood
$U_{b'}(a)$.  This is exactly the stated criterion.

\textup{(2)} This is immediate from~\textup{(1)}: injectivity of $\lambda_b$ is
precisely the assertion that equality of germs at $a$ forces equality of
sections over $U_b(a)$.

\textup{(3)} Because $\{U(c_i)\}_{i=1}^r$ covers $U_b(a)$ and $a\in U_b(a)$,
there exists an index $i_0$ with $a\in U(c_{i_0})$.  By the cofinality argument
from~\textup{(1)}, some deeper basic neighborhood $U_{b'}(a)$ is contained in
$U(c_{i_0})$.  The assumed equality on the cover therefore implies
\[
\rho_{a_b\leftarrow a_{b'}}(s)=\rho_{a_b\leftarrow a_{b'}}(t),
\]
so $\lambda_b(s)=\lambda_b(t)$ by~\textup{(1)}.  Injectivity of $\lambda_b$
then gives $s=t$.

The trichotomy follows from the same local-site dictionary as in
Corollary~\ref{cor:null-trichotomy-local-section}, but now applied at the fixed
point $a$.  If some sufficiently deep neighborhood has empty fiber, readout
fails by absence of local data.  If compatible local sections on a
cylinder-algebra cover of some $U_b(a)$ do not glue, the failure is one of
descent.  If
$\lambda_b$ fails to be injective for some $b$, then two distinct sections over
$U_b(a)$ define the same germ at $a$, so uniqueness fails.
\end{proof}

\section{Discussion}

The paper isolates one clean categorical chain:
\[
\begin{aligned}
\text{recursive refinement of visible events}
&\Longrightarrow \text{prefix site of cylinder sets} \\
&\Longrightarrow \text{local-section, descent, and separatedness obstructions} \\
&\Longrightarrow \text{intrinsic visible quotient and register bound} \\
&\Longrightarrow \text{explicit symbolic defect on a finite cover nerve} \\
&\Longrightarrow \text{inverse-limit local stalk and germ criteria.}
\end{aligned}
\]

Section~\ref{sec:measurable} records the standard measurable non-expansion
statement; Section~\ref{sec:prefix-sites} reformulates it as a site of cylinder
sets closed under finite intersections, following
\cite{MacLaneMoerdijk1992,Johnstone2002Elephant,SGA4I}; Section~\ref{sec:descent}
is the \v{C}ech obstruction theory based on
\cite{StacksProject,Giraud1971CohomologieNonAbelienne,Bredon1997SheafTheory};
Section~\ref{sec:inverse-limits} uses standard inverse-limit machinery from
\cite{RibesZalesskii2010ProfiniteGroups,Wilson1998ProfiniteGroups}.

The paper-specific content lies in Sections~\ref{sec:visible-quotient}
and~\ref{sec:forced-obstruction}, but even there the emphasis is on organizing
the formalism rather than claiming a new general theory.  In
Section~\ref{sec:visible-quotient}, Proposition~\ref{thm:intrinsic-visible-quotient}
defines the visible quotient $\Avis$ intrinsically via the family of quotients
on which the class $[\alpha]$ vanishes.  This removes the dependence on a
chosen cocycle representative and makes the subsequent register and spectral
statements class-level consequences.

Section~\ref{sec:forced-obstruction} supplies a fully explicit symbolic model.
Proposition~\ref{thm:forced-obstruction} computes a nontrivial $\ZZ/2\ZZ$-valued
defect on a four-chart cylinder-algebra cover of the full shift whose nerve is the
boundary of a tetrahedron, and Example~\ref{ex:tetrahedral-cover} interprets
that class as a phase mismatch localized on a single depth-$2$ cylinder.

The stalk formulation of single-valued readout in
Theorem~\ref{thm:recursive-addressing-readout-separatedness-null} says that
sections over a neighborhood of an inverse-limit address are determined by
their germs at that address exactly when the maps to the stalk are injective.
Equality on any actual cylinder-algebra cover containing the point then forces equality
of sections.

Two directions for further development are visible.  The first is to construct
the local objects and their abelian bands functorially from honest symbolic or
measurable readout semantics, rather than selecting them by hand on a chosen
cover.  The second is to
extend the comparison data from the abelian banded setting to a full
stack-theoretic or gerbe-theoretic framework
\cite{Giraud1971CohomologieNonAbelienne}, which would allow non-abelian
symmetry groups and produce higher descent obstructions.

\bibliographystyle{alpha}
\bibliography{references_local}

\end{document}



